### 2021-04-27 main.tex
first draft

\documentclass[a4paper,11pt]{article}
%\setcounter{page}{1}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}

\usepackage{hyperref}
%\usepackage[bookmarks,bookmarksnumbered,naturalnames,plainpages=false]{hyperref}
%usepackage{url}

% for footnote ref
\usepackage{refcount}

% symbols
\usepackage{amsmath} 
\usepackage{amssymb} 
\usepackage{amsbsy}
\usepackage{bbold}
\usepackage{latexsym}
%\usepackage{amsfonts}
\usepackage{stmaryrd}
%\usepackage{mathabx}
%\usepackage{MnSymbol}
\usepackage{harmony} % simple music fonts
\usepackage{mathtools} % for arrows
%\usepackage{mathptmx}

%% theorem envs
\usepackage{theorem}
\newtheorem{theorem}{Theorem} %[section]
\newtheorem{definition}[theorem]{Definition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newenvironment{proof}{\vspace{-2ex}{\it Proof. }}{\hspace*{\fill} $\Box$\smallskip }
\theorembodyfont{\slshape}
\newtheorem{example}[theorem]{Example}
\newtheorem{remark}[theorem]{Remark}

% extension of enumerate env. (style for displaying counters)
% \usepackage{enumerate}

%% pictures
% \usepackage{graphicx} 
% \DeclareGraphicsExtensions{.pdf,.png,.jpg}
% \graphicspath{fig/}

%% PGF, Tikz
%% \usepackage{pgfplots}
%% \usepgfplotslibrary{dateplot}
%% \usepackage{pgf,pgfarrows,pgfnodes, pgfautomata}
% \usepackage{tikz}
%% \usetikzlibrary{arrows}
%% \usetikzlibrary{calc}
%% \usetikzlibrary{snakes}
%% \usetikzlibrary{backgrounds}
% \usetikzlibrary{trees}
%% \usetikzlibrary{automata}
%% \usetikzlibrary{positioning}
%% \usetikzlibrary{matrix}
%% \usetikzlibrary{patterns}
%% \usetikzlibrary{shapes}

%% for new macros
\usepackage{xspace}

%% arrows etc
%\input{rewriting}

%% Misc macros

\def\ie{\textit{i.e.}\xspace}
\def\eg{\textit{e.g.}\xspace}
\def\wrt{\textit{wrt}\xspace}
%\def\wlog{\textit{wlog}\xspace}
\def\etc{\textit{etc}\xspace}

\def\<#1>{\langle #1 \rangle}
\newcommand{\pair}[2]{\langle{#1}, {#2}\rangle}
\newcommand{\A}{\mathcal{A}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\D}{\mathbb{D}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\W}{\mathbb{W}}

\newcommand{\Semiring}{\mathbb{S}}
\newcommand{\zero}{\mathbb{0}}
\newcommand{\one}{\mathbb{1}}
\newcommand{\dom}{\ensuremath{\mathit{dom}}}

\newcommand{\weight}{\mathsf{weight}}
\newcommand{\init}{\mathsf{in}}
\newcommand{\final}{\mathsf{out}}

%\sloppy

\title{Weighted Visibly Pushdown Automata and Automated Music Transcription}
%\author{Florent Jacquemard}
%\institute{INRIA \& CNAM, Paris, France\\
%\email{florent.jacquemard@inria.fr}}

%\titlerunning{WVPA \& AMT}
%\authorrunning{Florent Jacquemard}

\date{\today}

\begin{document}
\thispagestyle{empty}
\maketitle
\abstract{\ldots}

Symbolic Weighted (SW) extension of symbolic automata where...

\paragraph{Semirings.}
We shall consider semiring domains for weight values.
%
A \emph{semiring} $\< \Semiring, \oplus, \zero, \otimes, \one>$ 
is a structure with a domain~$\Semiring$,
equipped with two associative
binary operators~$\oplus$ and $\otimes$
with respective neutral elements $\zero$ and $\one$ and such that:
%$\< \mathbb{S}, \oplus, \zero>$ is a commutative monoid
%$\< \mathbb{S}, \otimes, \one>$ is a monoid
$\oplus$ is commutative, 
$\otimes$ distributes over~$\oplus$:  $\forall x, y, z \in \mathbb{S}$,
$x \otimes (y \oplus z) = (x \otimes y) \oplus (x \otimes z)$, 
%  and $(x \oplus y) \otimes z = (x \otimes z) \oplus (y \otimes z)$;
and $\zero$ is absorbing for~$\otimes$: 
$\forall x\in \mathbb{S}$, $\zero \otimes x = x \otimes \zero = \zero$.
%Components of a semiring~$\Semiring$ may be subscripted by~$\Semiring$ when needed.
%We simply write $x \in \Semiring$ to mean $x \in \mathbb{S}$.
%
In the application presented in this paper, intuitively,
$\oplus$ selects an optimal value amongst two values and 
$\otimes$ combines two values into a single value.

%\noindent %properties:
A semiring $\Semiring$ is \emph{commutative} if $\otimes$ is commutative.
%\francesco{How we deal with that since it is non commutative?}
It \emph{bounded}~\cite{Mohri02semiring} if $\forall x \in \dom(\Semiring), \one \oplus x = \one$,
and \emph{idempotent} if for all $x \in \Semiring$, $x \oplus x = x$.
%We restrict in the following of \emph{idempotent} semirings.
Note that every bounded semiring is idempotent:
by boundedness, 
$\one \oplus \one = \one$, and idempotency follows by multiplying
both sides by $x$ and distributing. %

%\noindent
A semiring $\Semiring$ 
is \emph{monotonic} \wrt a partial ordering~$\leq$ 
iff for all $x, y, z  \in \Semiring$,  $x \leq y$ implies
$x \oplus z \leq y \oplus z$,
$x \otimes z \leq y \otimes z$
and $z \otimes x \leq z \otimes y$,
%
and it is \emph{superior} %\wrt a partial ordering~$\leq$
\wrt~$\leq$ iff for all $x, y \in \Semiring$,  
$x \leq x \otimes y$ and 
$y \leq x \otimes y$~\cite{Huang08advanceddynamic}.
The latter property corresponds to the 
\emph{non-negative weights} condition in shortest-path algorithms~\cite{Dijkstra59anote}.
Intuitively, it means that combining elements always increase their weight. 
% always get worse in term of weight
Note that when $\Semiring$ is superior \wrt~$\leq$, then $\one \leq \zero$
and moreover, for all $x \in \Semiring$, $\one \leq x \leq \zero$.

Every idempotent semiring~$\Semiring$ induces 
a partial ordering~$\leq_\Semiring$ 
called the \emph{natural ordering} of~$\Semiring$
and defined by: 
%implicitly defined by the semiring $\Semiring$ 
for all $x$ and $y$,
$x \leq_\Semiring y \;\mbox{iff}\; x \oplus y = x$.
This ordering is sometimes defined in the opposite direction~\cite{DrosteKuich09semirings};
The above definition follows \cite{Mohri02semiring}, 
and coincides than the usual ordering on the Tropical semiring (\emph{min-plus}).
%
It holds that $\Semiring$ is {monotonic} \wrt~$\leq_\Semiring$.
An idempotent Semiring $\Semiring$~is called \emph{total} if
it~$\leq_\Semiring$ is total
\ie when for all $x, y \in \Semiring$, either $x \oplus y = x$ or $x \oplus y = y$.

We shall consider below infinite sums with~$\oplus$.
A semiring~$\Semiring$ is called \emph{complete} 
if for every family
$(x_i)_{i \in I}$ %$\{ x_i \mid i \in I \}$
of elements of $\dom(\Semiring)$ over an index set $I \subset \mathbb{N}$,
the infinite sum $\bigoplus_{i \in I} x_i$
is well-defined and in $\dom(\Semiring)$,
and the following properties hold:
\begin{description}
\item[$i.$]
\emph{infinite sums extend finite sums:}
$\displaystyle\bigoplus_{i \in \emptyset} x_i = \zero$,\quad 
      $\forall j\in \mathbb{N}, \displaystyle\bigoplus_{i \in \{ j \}} x_i = x_j$,\quad
      $\forall j, k\in \mathbb{N}, j\neq k, 
      \displaystyle\bigoplus_{i \in \{ j, k \}} x_i = x_j \oplus x_k$,
%
\item[$ii.$]
\emph{associativity and commutativity:}
for all $I \subseteq \mathbb{N}$
and all partition $(I_{j})_{j \in J}$ of $I$, %\subseteq \mathbb{N}$, 
\(
\displaystyle
\bigoplus_{j \in J}\bigoplus_{i \in I_j} x_i = 
\bigoplus_{i \in I} x_i
\),
%
\item[$iii.$] 
\emph{distributivity of product over infinite sum:}\\
for all $I \subseteq \mathbb{N}$,
\(
\displaystyle
\bigoplus_{i \in I} (x \otimes y_i) = x \otimes \bigoplus_{i\in I} y_i\), and
\(
\displaystyle
\bigoplus_{i \in I} (x_i \otimes y) = (\bigoplus_{i \in I} x_i ) \otimes y\).
\end{description}

\section{SW Visibly Pushdown Automata}
We follow the approach of~\cite{Mohri03EDWA} for the computation of distances...

\subsection{SW Automata and Transducers}
The following definition of weighted transducers over infinite alphabets 
generalizes weighted transducers over finite alphabets, see  e.g.~\cite{Mohri03EDWA}, 
by considering weight functions generalizing the guards of symbolic automata

Let $\Sigma$ and $\Delta$ be respectively an input and output \emph{alphabets}, 
which are finite or infinite sets of symbols, and let $\Semiring$ be a semiring.
%
A \emph{label theory} 
is a 4-uplet of recursively enumerable sets:
%$\Phi_0$, \Phi_\Sigma \uplus \Phi_\Delta \uplus \Phi_{\Sigma, \Delta}$ 
$\Phi_0$ containing constant functions valued in $\Semiring$, 
$\Phi_\Sigma$ and $\Phi_\Delta$, 
containing unary functions in $\Sigma \to \Semiring$, resp. $\Delta \to \Semiring$, 
and $\Phi_{\Sigma, \Delta}$  containing binary functions in $\Sigma \times \Delta \to \Semiring$.
Moreover, we assume that each of these sets is closed under $\oplus$ and $\otimes$, 
and all partial applications of functions $\Phi_{\Sigma, \Delta}$, 
resp. 
$f_a: y \mapsto f(a, y)$ for $a \in \Sigma$ and $y \in \Delta$
and 
$f_b: x \mapsto f(x, b)$ for $b \in \Delta$ and $x \in \Sigma$, 
belong resp. to $\Phi_\Sigma$ and $\Phi_\Delta$.

\begin{definition}
A weighted transducer $T$ over the input and output alphabet $\Sigma$ and $\Delta$ 
and the semiring $\Semiring$ is a tuple
$T = \< Q, \init, \weight, \final >$,
where $Q$ is a finite set of states, 
$\mathsf{in} : Q \to \Semiring$, 
respectively $\mathsf{out} : Q \to \Semiring,$
is a function defining the weight for entering, 
respectively leaving, a state, 
and $\weight$ is a transition %weight 
function of $Q \times Q$ into %$\Phi$.
$\< \Phi_0, \Phi_\Sigma, \Phi_\Delta, \Phi_{\Sigma, \Delta}>$.
\end{definition}
We extend the above transition function into a function from
$Q \times (\Sigma \cup \{ \epsilon \}) \times (\Delta \cup \{ \epsilon \}) \times Q$
into $\Semiring$, also called $\weight$ for simplicity, such that 
for all $q, q' \in Q$, $a \in \Sigma$, $b \in \Delta$, 
and with 
$\< \phi_\epsilon, \phi_\Sigma, \phi_\Delta, \phi_{\Sigma, \Delta}> = \weight(q, q')$, 
\[
\begin{array}{rcl}
\weight(q, \epsilon, \epsilon, q') & = & \phi_\epsilon\\
\weight(q, a, \epsilon, q') & = & \phi_\Sigma(a)\\
\weight(q, \epsilon, b, q') & = & \phi_\Delta(b)\\
\weight(q, a, b, q') & = & \phi_{\Sigma, \Delta}(a, b)
\end{array}      
\]
%More precisely, $Q \times Q$, 
%resp. $Q \times \Sigma \times Q$,
%$Q \times \Delta \times Q$,
%$Q \times \Sigma \times \Delta \times Q$,
%into resp. $\Phi_0$, $\Phi_\Sigma$, $\Phi_\Delta$ $\Phi_{\Sigma, \Delta}$.

These functions $\phi$ act as guards for the transducer's transitions, 
preventing a transition when they return the absorbing $\zero$ of $\Semiring$.

\subsection{Distance between words or languages}
distance $d$:
defined over $\Sigma^* \times \Sigma^*$ 
into a semiring  $\Semiring = ( \Semiring, \oplus, \zero, \otimes, \one)$.

\paragraph{Edit-Distance.}
...algebraic definition of edit-distance of Mohri, in \cite{Mohri03EDWA}
% Mehryar Mohri 
% Edit-distance of weighted automata: General definitions and algorithms
% International Journal of Foundations of Computer Science 14.06 (2003): 957-982.
%
\noindent
Let $\Omega = \Sigma \cup \{ \epsilon \} \times \Sigma \cup \{ \epsilon \} \setminus \{ (\epsilon, \epsilon) \}$,
and let $h$ be the morphism from $\Omega^*$ into $\Sigma^* \times \Sigma^*$  
defined over the concatenation of strings of $\Sigma^*$ (that removes the $\epsilon$'s).
%
\noindent
An \emph{alignment} between 2 strings  $s, t \in \Sigma^*$ is an element $\omega \in \Omega^*$ 
such that $h(\omega) = (s, t)$.
%
\noindent
We assume a base cost function $\Omega$ : $\delta: \Omega \to S$, extended to $\Omega^*$ as follows  
(for $\omega \in \Omega^*$): 
\(
\displaystyle\delta(\omega) = \bigotimes_{0 \leq i < |\omega|} \delta(\omega_i)
\).

\noindent
\begin{definition}
For  $s, t \in \Sigma^*$, the edit-distance between $s$ and $t$ is  
\( 
d(s, t) = \displaystyle\bigoplus_{\omega \in \Omega^*\, h(\omega) = (s, t)} \delta(\omega)
\).
\end{definition}

e.g. Levenstein edit-distance: $S$ is min-plus and $\delta(a, b) = 1$ for all $(a, b) \in \Omega$.

%\paragraph{Distance between a word and a regular language}

\subsection{SW Visibly Pushdown Automata}

\section{Application}
Symbolic Automated Music Transcription

\subsection{Representations}

\paragraph{Performance.}

\paragraph{Score.}

\subsection{Transducer for Distance Computation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% BIBLIO                                                                     %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\bibliographystyle{plain}
%\bibliographystyle{plainurl} 
\bibliographystyle{abbrv}
%\bibliographystyle{splncs04}
\bibliography{references}

\end{document}

### 2021-04-27 main.tex
def. transducers and automata

@@
 \begin{definition}
-A weighted transducer $T$ over the input and output alphabet $\Sigma$ and $\Delta$ 
+A \emph{weighted transducer} $T$ over the input and output alphabet $\Sigma$ and $\Delta$ 
 and the semiring $\Semiring$ is a tuple
@@
 preventing a transition when they return the absorbing $\zero$ of $\Semiring$.
+
+The weighted transducer $T$ defines a mapping 
+from the pairs of strings of $\Sigma^* \times \Delta^*$ 
+into the weights of~$\Semiring$,
+based on the following intermediate function $\weight_\A$
+defined recursively for every $q, q' \in Q$, 
+%$a \in \Sigma$, $b \in \Delta$ 
+for every strings of $s \in \Sigma^*$, $t \in \Delta^*$:
+\[
+\begin{array}{rccl}
+\weight_\A(q, s, t, q') & = & & \weight(q, \epsilon, \epsilon, q')\\
+ & & \oplus \displaystyle\bigoplus_{\begin{array}{c}
+                                      \scriptstyle q'' \in Q\\[-2pt]
+                                      \scriptstyle s = au, a \in \Sigma
+                                      \end{array}} &
+     \weight(q, a, \epsilon, q'') \otimes \weight_\A(q'', u, t, q')\\
+ & & \oplus \displaystyle\bigoplus_{\begin{array}{c}
+                                    \scriptstyle q'' \in Q\\[-2pt]
+                                    \scriptstyle t = bv, b \in \Delta
+                                    \end{array}} &
+      \weight(q, \epsilon, b, q'') \otimes \weight_\A(q'', s, v, q')\\ 
+      & & \oplus \displaystyle\bigoplus_{\begin{array}{c}
+                                         \scriptstyle q'' \in Q\\[-2pt]
+                                         \scriptstyle s = au, a \in \Sigma\\[-2pt]
+                                         \scriptstyle t = bv, b \in \Delta
+                                         \end{array}} &
+      \weight(q, a, b, q'') \otimes \weight_\A(q'', u, v, q')\\ 
+\end{array}
+\]
+Recall that by convention, an empty sum with $\oplus$ is $\zero$.
+
+\noindent
+The weight associated by $T$ to  $\< s, t> \in \Sigma^* \times \Delta^*$
+is then defined as follows: 
+\[
+T(s, t)  = 
+\displaystyle\bigoplus_{q, q' \in Q} \mathsf{in}(q) 
+\mathop{\otimes} \weight_\A(q, s, t, q') \mathop{\otimes} \mathsf{out}(q').
+\]
+
+A \emph{weighted automata} $T = \< Q, \init, \weight, \final >$
+over $\Sigma$ and $\Semiring$ 
+is defined in a similar way by simply omitting the output symbols,
+\ie $\weight$ is a function of~$Q \times Q$ into %$\Phi$.
+$\< \Phi_0, \Phi_\Sigma >$, 
+or equivalently from~$Q \times (\Sigma \cup \{ \epsilon \}) \times Q$ into~$\Semiring$.
 

### 2021-04-28 main.tex
def. SWVPA

--- 2021-04-27 main.tex
@@
 
-\newcommand{\weight}{\mathsf{weight}}
-\newcommand{\init}{\mathsf{in}}
-\newcommand{\final}{\mathsf{out}}
+\def\SWT{\textsf{SWT}\xspace}
+\def\SWA{\textsf{SWA}\xspace}
+\def\SWVPA{\textsf{SWVPA}\xspace}
+\def\weight{\mathsf{weight}}
+\def\wei{\mathsf{w}}
+\def\init{\mathsf{in}}
+\def\final{\mathsf{out}}
+\newcommand{\call}[1]{\ensuremath \langle_{#1}}
+\newcommand{\return}[1]{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
+\def\Sigmai{{\Sigma_\mathsf{i}}}
+\def\Sigmac{{\Sigma_\mathsf{c}}}
+\def\Sigmar{{\Sigma_\mathsf{r}}}
+\def\Phii{{\Phi_\mathsf{i}}}
+\def\Phic{{\Phi_\mathsf{c}}}
+\def\Phir{{\Phi_\mathsf{r}}}
+\def\Phicr{{\Phi_\mathsf{cr}}}
+\def\weii{{\wei_\mathsf{i}}}
+\def\weic{{\wei_\mathsf{c}}}
+\def\weir{{\wei_\mathsf{r}}}
+\def\weie{{\wei_\mathsf{e}}}
+\newcommand{\config}[2]{\ensuremath \genfrac{[}{]}{0pt}{}{#1}{#2}}
 
@@
 
-%\noindent %properties:
-A semiring $\Semiring$ is \emph{commutative} if $\otimes$ is commutative.
-%\francesco{How we deal with that since it is non commutative?}
-It \emph{bounded}~\cite{Mohri02semiring} if $\forall x \in \dom(\Semiring), \one \oplus x = \one$,
-and \emph{idempotent} if for all $x \in \Semiring$, $x \oplus x = x$.
-%We restrict in the following of \emph{idempotent} semirings.
-Note that every bounded semiring is idempotent:
-by boundedness, 
-$\one \oplus \one = \one$, and idempotency follows by multiplying
-both sides by $x$ and distributing. %
+and let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a {semiring},
 
@@
 
-\section{SW Visibly Pushdown Automata}
-We follow the approach of~\cite{Mohri03EDWA} for the computation of distances...
+\section{SW Automata and Transducers}
+\label{section:transducer}
+We follow the approach of~\cite{Mohri03EDWA} for the computation of distances
+between words with transducers.
 
-\subsection{SW Automata and Transducers}
 The following definition of weighted transducers over infinite alphabets 
@@
 
+%(SWT)
 \begin{definition}
-A weighted transducer $T$ over the input and output alphabet $\Sigma$ and $\Delta$ 
+A \emph{symbolic-weighted transducer} $T$ over the input and output alphabet $\Sigma$ and $\Delta$ 
 and the semiring $\Semiring$ is a tuple
-$T = \< Q, \init, \weight, \final >$,
+$T = \< Q, \init, \wei, \final >$,
 where $Q$ is a finite set of states, 
@@
 respectively $\mathsf{out} : Q \to \Semiring,$
-is a function defining the weight for entering, 
+are functions defining the weight for entering, 
 respectively leaving, a state, 
-and $\weight$ is a transition %weight 
-function of $Q \times Q$ into %$\Phi$.
+and $\wei$ is a transition %weight 
+function from $Q \times Q$ into %$\Phi$.
 $\< \Phi_0, \Phi_\Sigma, \Phi_\Delta, \Phi_{\Sigma, \Delta}>$.
 \end{definition}
+%
 We extend the above transition function into a function from
 $Q \times (\Sigma \cup \{ \epsilon \}) \times (\Delta \cup \{ \epsilon \}) \times Q$
-into $\Semiring$, also called $\weight$ for simplicity, such that 
+into $\Semiring$, also called $\wei$ for simplicity, such that 
 for all $q, q' \in Q$, $a \in \Sigma$, $b \in \Delta$, 
 and with 
-$\< \phi_\epsilon, \phi_\Sigma, \phi_\Delta, \phi_{\Sigma, \Delta}> = \weight(q, q')$, 
+$\< \phi_\epsilon, \phi_\Sigma, \phi_\Delta, \phi_{\Sigma, \Delta}> = \wei(q, q')$, 
 \[
 \begin{array}{rcl}
-\weight(q, \epsilon, \epsilon, q') & = & \phi_\epsilon\\
-\weight(q, a, \epsilon, q') & = & \phi_\Sigma(a)\\
-\weight(q, \epsilon, b, q') & = & \phi_\Delta(b)\\
-\weight(q, a, b, q') & = & \phi_{\Sigma, \Delta}(a, b)
+\wei(q, \epsilon, \epsilon, q') & = & \phi_\epsilon\\
+\wei(q, a, \epsilon, q') & = & \phi_\Sigma(a)\\
+\wei(q, \epsilon, b, q') & = & \phi_\Delta(b)\\
+\wei(q, a, b, q') & = & \phi_{\Sigma, \Delta}(a, b)
 \end{array}      
@@
 
-\subsection{Distance between words or languages}
-distance $d$:
-defined over $\Sigma^* \times \Sigma^*$ 
-into a semiring  $\Semiring = ( \Semiring, \oplus, \zero, \otimes, \one)$.
+The symbolic-weighted transducer $T$ defines a mapping 
+from the pairs of strings of $\Sigma^* \times \Delta^*$ 
+into the weights of~$\Semiring$,
+based on the following intermediate function $\weight_T$
+defined recursively for every $q, q' \in Q$, 
+%$a \in \Sigma$, $b \in \Delta$ 
+for every strings of $s \in \Sigma^*$, $t \in \Delta^*$:
+\[
+\begin{array}{rccl}
+\weight_T(q, s, t, q') & = & & \wei(q, \epsilon, \epsilon, q')\\
+ & & \oplus \displaystyle\bigoplus_{\begin{array}{c}
+                                      \scriptstyle q'' \in Q\\[-2pt]
+                                      \scriptstyle s = au, a \in \Sigma
+                                      \end{array}} &
+     \wei(q, a, \epsilon, q'') \otimes \weight_T(q'', u, t, q')\\
+ & & \oplus \displaystyle\bigoplus_{\begin{array}{c}
+                                    \scriptstyle q'' \in Q\\[-2pt]
+                                    \scriptstyle t = bv, b \in \Delta
+                                    \end{array}} &
+      \wei(q, \epsilon, b, q'') \otimes \weight_T(q'', s, v, q')\\ 
+      & & \oplus \displaystyle\bigoplus_{\begin{array}{c}
+                                         \scriptstyle q'' \in Q\\[-2pt]
+                                         \scriptstyle s = au, a \in \Sigma\\[-2pt]
+                                         \scriptstyle t = bv, b \in \Delta
+                                         \end{array}} &
+      \wei(q, a, b, q'') \otimes \weight_\A(q'', u, v, q')\\ 
+\end{array}
+\]
+Recall that by convention, an empty sum with $\oplus$ is $\zero$. 
+%
+\noindent
+The weight associated by $T$ to  $\< s, t> \in \Sigma^* \times \Delta^*$
+is then defined as follows: 
+\[
+T(s, t)  = 
+\displaystyle\bigoplus_{q, q' \in Q} \mathsf{in}(q) 
+\mathop{\otimes} \weight_T(q, s, t, q') \mathop{\otimes} \mathsf{out}(q').
+\]
 
-\paragraph{Edit-Distance.}
+A \emph{symbolic weighted automata} (\SWA) $A = \< Q, \init, \weight, \final >$
+over $\Sigma$ and $\Semiring$ 
+is defined in a similar way by simply omitting the output symbols,
+\ie $\wei$ is a function of~$Q \times Q$ into %$\Phi$.
+$\< \Phi_0, \Phi_\Sigma >$, 
+or equivalently from~$Q \times (\Sigma \cup \{ \epsilon \}) \times Q$ into~$\Semiring$.
+
+\begin{proposition}
+Given a \SWT $T$ 
+over $\Sigma$, $\Delta$ and $\Semiring$, 
+and a word $s \in \Sigma^*$, 
+one can construct a \SWA 
+$A_{s, T}$ such that for all $t \in \Delta^*$, 
+$A_{s, T}(t) = T(s, t)$.
+\end{proposition}
+The construction time and size of $A_{s, T}$ are $O(| s | . \| T \|)$.
+
+\section{SW Visibly Pushdown Automata}
+The following model generalizes Symbolic VPA~\cite{DAntonyAlur14SVPDA}
+from Boolean semirings to arbitrary semiring weight domains.
+
+Let $\Sigma$ be an input alphabet, finite (large) or infinite,
+that we assume partitioned into :
+\begin{itemize}
+\item a set $\Sigmai$ of \emph{internal symbols} denoted $a$,
+\item a set $\Sigmac$ of \emph{call symbols} denoted $\call{a}$,
+\item a set $\Sigmar$ of \emph{return symbols} denoted $\return{a}$.
+\end{itemize}
+In order to simplify notations, and following the definition of Section~\ref{section:transducer}, 
+we shall write respectively~$\Phii$, $\Phic$, $\Phir$ and~$\Phicr$
+for~$\Phi_\Sigmai$, $\Phi_\Sigmac$, $\Phi_\Sigmar$ and~$\Phi_{\Sigmac, \Sigmar}$,
+
+\begin{definition}
+A \emph{Symbolic Weighted Visibly Pushdown Automata} (\SWVPA) $A$ 
+over the input $\Sigma = \Sigmai \uplus \Sigmac \uplus \Sigmar$
+and the semiring $\Semiring$ is a tuple
+$T = \< Q, P, \init, \weii, \weic, \weir, \weie, \final >$,
+where $Q$ is a finite set of states, 
+$P$ is a finite set of stack symbols, 
+$\mathsf{in} : Q \to \Semiring$, 
+respectively $\mathsf{out} : Q \to \Semiring,$
+are functions defining the weight for entering, 
+respectively leaving, a state, 
+and 
+$\weii : Q \times Q \to \Phii$,  
+$\weic : Q \times Q \times P \to \Phic$,  
+$\weir : Q \times P \times Q \to \Phicr$,  
+$\weie : Q \times Q \to \Phir$,  
+are transition functions.
+\end{definition}
+%
+Similarly as in Section~\ref{section:transducer}, 
+we extend the above transition functions as follows
+for all $q, q' \in Q$, $p \in P$, 
+$a \in \Sigmai$, 
+$\call{c} \in \Sigmac$, 
+$\return{r} \in \Sigmar$, 
+overloading the names for simplicity:
+\[
+\begin{array}{lll}
+\weii: Q \times \Sigmai \times Q \to \Semiring & 
+\weii(q, a, q') = \phi_\mathsf{i}(a) & 
+\mathrm{where~} \phi_\mathsf{i} = \weii(q, q')\\
+%
+\weic: Q \times \Sigmac \times Q \times P \to \Semiring & 
+\weic(q, \call{c}, q', p) = \phi_\mathsf{c}(\call{c}) & 
+\mathrm{where~} \phi_\mathsf{c} = \weic(q, q', p)\\
+%
+\weir: Q \times \Sigmac \times P \times \Sigmar \times Q \to \Semiring & 
+\weir(q, {\call{c}},  p, {\return{r}}, q') = \phi_\mathsf{r}({\call{c}},  {\return{r}}) & 
+\mathrm{where~} \phi_\mathsf{r} = \weir(q, p, q')\\
+%
+\weie: Q \times \Sigmar \times Q \to \Semiring & 
+\weie(q, {\return{r}}, q') = \phi_\mathsf{e}({\return{r}}) &
+\mathrm{where~} \phi_\mathsf{e} = \weie(q, q')\\
+\end{array}      
+\]
+
+\noindent
+The intuition is the following for the above transitions.
+\begin{description}
+\item $\weii$ : read the input internal symbol $a$, change state to $q'$.
+\item $\weic$ : read the input symbol $\call{c}$, push it to the stack along with $p$, change state to $q'$.
+\item $\weir$ : when the stack is not empty, 
+      read and pop from stack a pair made of $\call{c}$ and $p$, 
+      read the input symbol $\return{r}$, change state to $q'$.
+      In this case, the weight function $\phi_\mathsf{r}$ checks a matching between the call and return symbols.
+\item $\weie$ : when the stack is empty, 
+      read the input symbol $\call{r}$, change state to $q'$.
+\end{description}
+
+We give now a formal definition of these transitions of the automaton $A$
+in term of a weight value
+computed by an intermediate function $\weight_A$.
+In the case of a pushdown automaton, a configuration is composed 
+of a state $q \in Q$ and a stack content $\theta \in \Theta^*$, where $\Theta = \Sigmac \times P$.
+Therefore, $\weight_A$ is a function from 
+$Q \times \Theta^* \times \Sigma^* \times Q \times \Theta^*$ into $\Semiring$.
+\[
+\begin{array}{rcl}
+\weight_A\bigl(\config{q}{\theta}, a u, \config{q'}{\theta'}\bigr) & = & 
+ {\displaystyle\bigoplus_{q'' \in Q}} \weii(q, a, q'') 
+ \otimes \weight_A\bigl(\config{q''}{\theta}, u, \config{q'}{\theta'}\bigr)\\
+%
+\weight_A\bigl(\config{q}{\theta}, {\call{c}} u, \config{q'}{\theta'}\bigr) & = & 
+ {\displaystyle\bigoplus_{\begin{array}{c}
+                          \scriptstyle q'' \in Q\\[-2pt]
+                          \scriptstyle p \in P
+                          \end{array}}}
+ \weic\bigl(q, {\call{c}}, q'', p\bigr) 
+ \otimes \weight_A\bigl(\config{q''}{{\call{c}}\, p\cdot \theta}, u, \config{q'}{\theta'}\bigr)\\[1mm]
+%
+\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot \theta}, {\return{r}} u, \config{q'}{\theta'}\bigr) & = & 
+ {\displaystyle\bigoplus_{q'' \in Q}} 
+ \weir\bigl(q, {\call{c}}, p, {\return{r}}, q''\bigr) 
+ \otimes \weight_A\bigl(\config{q''}{\theta}, u, \config{q'}{\theta'}\bigr)\\
+%
+\weight_A\bigl(\config{q}{\bot}, {\return{r}} u, \config{q'}{\theta'}\bigr) & = & 
+ {\displaystyle\bigoplus_{q'' \in Q}} \weie(q, {\return{r}}, q'') 
+ \otimes \weight_A\bigl(\config{q''}{\bot}, u, \config{q'}{\theta'}\bigr)\\
+\end{array}
+\]
+where $\bot$ denotes the empty stack and ${\call{c}}\, p\cdot \theta$ 
+denotes a stack with the pair made of ${\call{c}}$ and $p$ on its top and $\theta$
+as the rest of stack.
+
+\noindent
+The weight associated by $A$ to $s \in \Sigma^*$
+is then defined as follows by empty stack computation: 
+\[
+A(s)  = 
+{\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
+\mathsf{in}(q) \mathop{\otimes} 
+\weight_A\bigl(\config{q}{\bot}, s, \config{q'}{\bot}\bigr) 
+\mathop{\otimes} \mathsf{out}(q').
+\]
+
+% The weight of a transition acts as a guard: 
+% a transition is activated for a symbol $`a`$ iff its weight $`\phi(a) \neq 0`$ (or $`\psi(a, b) \neq 0`$), 
+% the absorbing element.
+
+\section{Application}
+Symbolic Automated Music Transcription
+
+\subsection{Representations}
+
+\paragraph{Performance.}
+
+\paragraph{Score.}
+
+\subsection{Transducer for Distance Computation}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% BIBLIO                                                                     %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\bibliographystyle{plain}
+%\bibliographystyle{plainurl} 
+\bibliographystyle{abbrv}
+%\bibliographystyle{splncs04}
+\bibliography{references}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% APPENDIX                                                                   %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newpage
+\appendix
+
+\section{Edit-Distance}
+
+%\subsection{Distance between words or languages}
 ...algebraic definition of edit-distance of Mohri, in \cite{Mohri03EDWA}
@@
 % International Journal of Foundations of Computer Science 14.06 (2003): 957-982.
-%
-\noindent
+distance $d$ over $\Sigma^* \times \Sigma^*$ 
+into a semiring  $\Semiring = ( \Semiring, \oplus, \zero, \otimes, \one)$.
+
+%\noindent
 Let $\Omega = \Sigma \cup \{ \epsilon \} \times \Sigma \cup \{ \epsilon \} \setminus \{ (\epsilon, \epsilon) \}$,
@@
 
-\subsection{SW Visibly Pushdown Automata}
-
-\section{Application}
-Symbolic Automated Music Transcription
-
-\subsection{Representations}
-
-\paragraph{Performance.}
-
-\paragraph{Score.}
-
-\subsection{Transducer for Distance Computation}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% BIBLIO                                                                     %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\bibliographystyle{plain}
-%\bibliographystyle{plainurl} 
-\bibliographystyle{abbrv}
-%\bibliographystyle{splncs04}
-\bibliography{references}
-
 \end{document}

### 2021-04-29 main.tex
representations score, performances

@@
 \def\Sigmar{{\Sigma_\mathsf{r}}}
+\def\Deltai{{\Delta_\mathsf{i}}}
+\def\Deltac{{\Delta_\mathsf{c}}}
+\def\Deltar{{\Delta_\mathsf{r}}}
 \def\Phii{{\Phi_\mathsf{i}}}
@@
 \section{SW Visibly Pushdown Automata}
+\label{section:SWVPA}
 The following model generalizes Symbolic VPA~\cite{DAntonyAlur14SVPDA}
@@
 Symbolic Automated Music Transcription
+and analysis of music performances
 
-\subsection{Representations}
+\subsection{Time Scales}
+Real-Time Unit (RTU) = seconds
 
-\paragraph{Performance.}
+\noindent 
+Musical-Time Unit (MTU) = number of measures
 
-\paragraph{Score.}
+\noindent 
+conversion via tempo value
 
-\subsection{Transducer for Distance Computation}
+\subsection{Representation of Music Performances}
+A musical performance is represented symbolically as a finite sequence of timestamped events.
+similar to piano roll representation~\cite{Muller15fundamentals} chap.1
+
+\noindent
+We consider an infinite alphabet $\Sigma$ of MIDI-like 
+events made of:
+\begin{itemize}
+\item a timestamp in RTU\\ %(real or int = nb samples ?)
+      or IOI in RTU
+\item a pitch value in 0..127
+\item ON | OFF flag
+\item a velocity value in 0..127
+\end{itemize}
+
+\subsection{Representation of Music Scores}
+we consider here the case of monodies for the sake of simplificity.
+
+A music score ris epresented as a structured word
+= sequence of quantified events + markups,
+see nested words~\cite{AlurMadhusudan09nested}.
+
+We consider an alphabet $\Delta$, every symbol of which is 
+composed of a tag, in a finite set $\Xi$, 
+and an IOI duration value.
+Moreover, $\Delta$ is partitioned into 
+$\Delta = \Deltai \uplus \Deltac \uplus \Deltar$, 
+like in Section~\ref{section:SWVPA}.
+
+\noindent
+The elements of $\Deltai$ represent events:
+% (infinite alphabet of internal symbols) made of:
+whose tags are one of:
+\begin{itemize}
+\item continuation (0) : tie or dot
+\item note, grace note (pitch) or chord (pitch+)
+\item rest
+\item ...
+\end{itemize}
+
+\noindent
+The elements of $\Deltac$ and $\Deltar$ are 
+markups for the representation of groups of events
+(linearization of rhythm trees \cite{jacquemard:hal-01138642}...)
+- parentheses for time divisions : tuplets, bars...
+tag %labels: 
+contain info such as tuple number, beaming policy...
+
+\noindent
+The date or duration of events, in MTU (rational), 
+can be computed with the markups and tags (e.g. grace note has duration 0).
+
+\noindent
+There are simultaneous events, since grace notes has duration 0. They are ordered.
+
+\noindent
+Finite bound on the number of duration ratio. ?
+
+\subsection{Performance/Score Distance Computation}
+with a transducer
 

### 2021-04-29 main.tex
revise definition label theory

@@
 \title{Weighted Visibly Pushdown Automata and Automated Music Transcription}
-%\author{Florent Jacquemard}
+\author{Florent Jacquemard}
 %\institute{INRIA \& CNAM, Paris, France\\
@@
 \maketitle
-\abstract{\ldots}
 
-Symbolic Weighted (SW) extension of symbolic automata where...
+%\begin{abstract}
+%Symbolic Weighted (SW) extension of symbolic automata...
+%\end{abstract}
 
-\paragraph{Semirings.}
+\section{SW Automata and Transducers}
+\label{section:transducer}
+
+We follow the approach of~\cite{Mohri03EDWA} for the computation of distances
+between words with weighted transducers, and propose models of weighted automata 
+and weighted transducers over infinite alphabets.
+
+These models generalize weighted automata and transducers over finite alphabets, see  e.g.~\cite{Mohri03EDWA}, 
+by labeling each transition with a weight functions that takes the 
+input and output symbols as parameters, instead of a simple weight value.
+These functions are similar to the guards of symbolic automata~\cite{Veanes13ciaa}, 
+but they can return values in an arbitrary semiring, 
+where the latter guards are restricted to the Boolean semiring.
+
+\subsection{Semirings}
+\label{section:semiring}
 We shall consider semiring domains for weight values.
@@
 %
-In the application presented in this paper, intuitively,
-$\oplus$ selects an optimal value amongst two values and 
-$\otimes$ combines two values into a single value.
+% In the application presented in this paper, intuitively,
+% $\oplus$ selects an optimal value amongst two values and 
+% $\otimes$ combines two values into a single value.
+%and let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a {semiring},
 
-and let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a {semiring},
+\noindent
+A semiring $\Semiring$ is \emph{commutative} if $\otimes$ is commutative.
+It is \emph{idempotent} if for each $x \in \dom(\Semiring)$, $x \oplus x = x$.
+%
+Following the terminology of~\cite{Mohri02semiring},
+when $\forall x \in \dom(\Semiring), \one \oplus x = \one$,
+the semiring $\Semiring$ is is called \emph{bounded}.
+Note that every bounded semiring is idempotent:
+by boundedness, 
+$\one \oplus \one = \one$, and idempotency follows by multiplying
+both sides by $x$ and distributing.
 
-%\noindent
+% \noindent
 A semiring $\Semiring$ 
@@
 It holds that $\Semiring$ is {monotonic} \wrt~$\leq_\Semiring$.
-An idempotent Semiring $\Semiring$~is called \emph{total} if
+An idempotent semiring $\Semiring$~is called \emph{total} if
 it~$\leq_\Semiring$ is total
@@
 
+\medskip
 We shall consider below infinite sums with~$\oplus$.
@@
 
-\section{SW Automata and Transducers}
-\label{section:transducer}
-We follow the approach of~\cite{Mohri03EDWA} for the computation of distances
-between words with transducers.
-
-The following definition of weighted transducers over infinite alphabets 
-generalizes weighted transducers over finite alphabets, see  e.g.~\cite{Mohri03EDWA}, 
-by considering weight functions generalizing the guards of symbolic automata
+\subsection{Label Theory}
+\label{section:symbols}
 
 Let $\Sigma$ and $\Delta$ be respectively an input and output \emph{alphabets}, 
-which are finite or infinite sets of symbols, and let $\Semiring$ be a semiring.
-%
-A \emph{label theory} 
-is a 4-uplet of recursively enumerable sets:
+which are countable (finite or infinite) sets of symbols, 
+and let $\Semiring$ be a commutative semiring.
+
+\noindent 
+A \emph{label theory} is made of 4 recursively enumerable sets:
 %$\Phi_0$, \Phi_\Sigma \uplus \Phi_\Delta \uplus \Phi_{\Sigma, \Delta}$ 
-$\Phi_0$ containing constant functions valued in $\Semiring$, 
+$\Phi_\epsilon \subseteq \Semiring$, % containing constant functions valued in $\Semiring$, 
 $\Phi_\Sigma$ and $\Phi_\Delta$, 
@@
 and $\Phi_{\Sigma, \Delta}$  containing binary functions in $\Sigma \times \Delta \to \Semiring$.
-Moreover, we assume that each of these sets is closed under $\oplus$ and $\otimes$, 
+Moreover, we assume that these sets are closed under the operators 
+$\oplus$ and $\otimes$ of~$\Semiring$.
+More precisely, for all $\phi, \phi' \in \Phi_\Sigma$ all 
+$\psi, \psi' \in \Phi_\Delta$, 
+and $\eta, \eta' \in \Phi_{\Sigma, \Delta}$, the function
+\begin{description}
+\item $\phi \otimes \phi' : x \mapsto \phi(x) \otimes \phi'(x)$ belongs to $\Phi_\Sigma$, 
+\item $\psi \otimes \psi' : y \mapsto \psi(y) \otimes \psi'(y)$ belongs to $\Phi_\Delta$,
+\item $\phi \otimes \eta : x, y \mapsto \phi(x) \otimes \eta(x, y)$ belongs to $\Phi_{\Sigma, \Delta}$,
+\item $\eta \otimes \psi : x, y \mapsto \eta(x, y) \otimes \psi(y)$ belongs to $\Phi_{\Sigma, \Delta}$,
+\item $\eta \otimes \eta' : x, y \mapsto \eta(x, y) \otimes \eta'(x, y)$ belongs to $\Phi_{\Sigma, \Delta}$.
+\end{description}
+The same also holds for the binary sum operator $\oplus$.
+
+\noindent
+\marginpar{maybe not necessary}
+Finally, it is assumed that the codomain of every function of $\Phi_\Sigma$ and $\Phi_\Delta$
+is a subset of $\Phi_\epsilon$.
 and all partial applications of functions $\Phi_{\Sigma, \Delta}$, 
-resp. 
-$f_a: y \mapsto f(a, y)$ for $a \in \Sigma$ and $y \in \Delta$
-and 
-$f_b: x \mapsto f(x, b)$ for $b \in \Delta$ and $x \in \Sigma$, 
+resp.  $f_a: y \mapsto f(a, y)$ for $a \in \Sigma$ and $y \in \Delta$
+and  $f_b: x \mapsto f(x, b)$ for $b \in \Delta$ and $x \in \Sigma$, 
 belong resp. to $\Phi_\Sigma$ and $\Phi_\Delta$.
@@
 
-Let $\Sigma$ be an input alphabet, finite (large) or infinite,
+Let $\Sigma$ be a countable alphabet 
+%finite (large) or infinite,
 that we assume partitioned into :
@@
 \end{itemize}
-In order to simplify notations, and following the definition of Section~\ref{section:transducer}, 
-we shall write respectively~$\Phii$, $\Phic$, $\Phir$ and~$\Phicr$
-for~$\Phi_\Sigmai$, $\Phi_\Sigmac$, $\Phi_\Sigmar$ and~$\Phi_{\Sigmac, \Sigmar}$,
+Let $\Semiring$ be a commutative semiring and  
+let  $(\Phi_\epsilon, \Phic, \Phir, \Phicr)$ be a label theory over $\Semiring$
+%In order to simplify notations, %and following the definition of Section~\ref{section:transducer}, 
+%we shall write respectively 
+where $\Phic$, $\Phir$ and~$\Phicr$ stand respectively 
+for~$\Phi_\Sigmac$, $\Phi_\Sigmar$ and~$\Phi_{\Sigmac, \Sigmar}$.
+%
+Moreover, we extend this theory with a set $\Phii$ 
+of unary functions in $\Sigmai \to \Semiring$,
+closed under $\oplus$ and $\otimes$.
 
 \begin{definition}
-A \emph{Symbolic Weighted Visibly Pushdown Automata} (\SWVPA) $A$ 
-over the input $\Sigma = \Sigmai \uplus \Sigmac \uplus \Sigmar$
-and the semiring $\Semiring$ is a tuple
+A \emph{Symbolic Weighted Visibly Pushdown Automata} (\SWVPA)~$A$ 
+over  $\Sigma = \Sigmai \uplus \Sigmac \uplus \Sigmar$ and $\Semiring$ is a tuple
 $T = \< Q, P, \init, \weii, \weic, \weir, \weie, \final >$,
@@
 \weii(q, a, q') = \phi_\mathsf{i}(a) & 
-\mathrm{where~} \phi_\mathsf{i} = \weii(q, q')\\
+\mathrm{where~} \phi_\mathsf{i} = \weii(q, q'),\\
 %
@@
 \weic(q, \call{c}, q', p) = \phi_\mathsf{c}(\call{c}) & 
-\mathrm{where~} \phi_\mathsf{c} = \weic(q, q', p)\\
+\mathrm{where~} \phi_\mathsf{c} = \weic(q, q', p),\\
 %
@@
 \weir(q, {\call{c}},  p, {\return{r}}, q') = \phi_\mathsf{r}({\call{c}},  {\return{r}}) & 
-\mathrm{where~} \phi_\mathsf{r} = \weir(q, p, q')\\
+\mathrm{where~} \phi_\mathsf{r} = \weir(q, p, q'),\\
 %
@@
 \weie(q, {\return{r}}, q') = \phi_\mathsf{e}({\return{r}}) &
-\mathrm{where~} \phi_\mathsf{e} = \weie(q, q')\\
+\mathrm{where~} \phi_\mathsf{e} = \weie(q, q').\\
 \end{array}      
@@
 \item $\weii$ : read the input internal symbol $a$, change state to $q'$.
-\item $\weic$ : read the input symbol $\call{c}$, push it to the stack along with $p$, change state to $q'$.
+\item $\weic$ : read the input call symbol $\call{c}$, push it to the stack along with $p$, change state to $q'$.
 \item $\weir$ : when the stack is not empty, 
       read and pop from stack a pair made of $\call{c}$ and $p$, 
-      read the input symbol $\return{r}$, change state to $q'$.
-      In this case, the weight function $\phi_\mathsf{r}$ checks a matching between the call and return symbols.
+      read the input return symbol $\return{r}$, change state to $q'$.
+      In this case, the weight function $\phi_\mathsf{r}$ 
+      computes the value of matching between the call and return symbols.
+      It might be $\zero$ to express that the symbols do not match.
 \item $\weie$ : when the stack is empty, 
@@
 The weight associated by $A$ to $s \in \Sigma^*$
-is then defined as follows by empty stack computation: 
+is then defined as follows,
+following empty stack semantics: 
 \[
@@
 
+\subsection{1-best Computation}
+
+Regarding the infinite sum operator, it must hold that
+$\bigoplus_{x \in \Phi_\Sigma} \phi(x)$, 
+$\bigoplus_{y \in \Phi_\Delta} \psi(y)$, and
+
 % The weight of a transition acts as a guard: 
@@
 % the absorbing element.
+
+property of bounded convexity of weight functions:
 

### 2021-05-01 main.tex
1-best algo

@@
 \usepackage{refcount}
+
+% extension of enumerate env. (style for displaying counters)
+% \usepackage{enumerate}
+
+%% pictures
+% \usepackage{graphicx} 
+% \DeclareGraphicsExtensions{.pdf,.png,.jpg}
+% \graphicspath{fig/}
+
+%% PGF, Tikz
+%% \usepackage{pgfplots}
+%% \usepgfplotslibrary{dateplot}
+%% \usepackage{pgf,pgfarrows,pgfnodes, pgfautomata}
+% \usepackage{tikz}
+%% \usetikzlibrary{arrows}
+%% \usetikzlibrary{calc}
+%% \usetikzlibrary{snakes}
+%% \usetikzlibrary{backgrounds}
+% \usetikzlibrary{trees}
+%% \usetikzlibrary{automata}
+%% \usetikzlibrary{positioning}
+%% \usetikzlibrary{matrix}
+%% \usetikzlibrary{patterns}
+%% \usetikzlibrary{shapes}
 
@@
 \newtheorem{remark}[theorem]{Remark}
+{\theorembodyfont{\rmfamily} \theoremstyle{break} \newtheorem{algo}{Algorithm}}
 
-% extension of enumerate env. (style for displaying counters)
-% \usepackage{enumerate}
-
-%% pictures
-% \usepackage{graphicx} 
-% \DeclareGraphicsExtensions{.pdf,.png,.jpg}
-% \graphicspath{fig/}
-
-%% PGF, Tikz
-%% \usepackage{pgfplots}
-%% \usepgfplotslibrary{dateplot}
-%% \usepackage{pgf,pgfarrows,pgfnodes, pgfautomata}
-% \usepackage{tikz}
-%% \usetikzlibrary{arrows}
-%% \usetikzlibrary{calc}
-%% \usetikzlibrary{snakes}
-%% \usetikzlibrary{backgrounds}
-% \usetikzlibrary{trees}
-%% \usetikzlibrary{automata}
-%% \usetikzlibrary{positioning}
-%% \usetikzlibrary{matrix}
-%% \usetikzlibrary{patterns}
-%% \usetikzlibrary{shapes}
+%% algorithms
+%\usepackage{algorithm} 
+%\usepackage{program}
 
@@
 \newcommand{\pair}[2]{\langle{#1}, {#2}\rangle}
-\newcommand{\A}{\mathcal{A}}
-\newcommand{\B}{\mathcal{B}}
+%\newcommand{\A}{\mathcal{A}}
+%\newcommand{\B}{\mathcal{B}}
 \newcommand{\D}{\mathbb{D}}
@@
 \def\final{\mathsf{out}}
-\newcommand{\call}[1]{\ensuremath \langle_{#1}}
-\newcommand{\return}[1]{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
+\newcommand{\call}[1]{\ensuremath #1} %{\ensuremath \langle_{#1}}
+\newcommand{\return}[1]{\ensuremath #1} %{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
 \def\Sigmai{{\Sigma_\mathsf{i}}}
@@
 \newcommand{\config}[2]{\ensuremath \genfrac{[}{]}{0pt}{}{#1}{#2}} 
+\newcommand{\opluseq}{\ensuremath\mathrel{\oplus}=}
 
@@
 %\end{abstract}
+
+weighted parsing
+
+principle of quantitative parsing / parsing of words over infinite alphabet:
 
@@
 
-\noindent
-A semiring $\Semiring$ is \emph{commutative} if $\otimes$ is commutative.
-It is \emph{idempotent} if for each $x \in \dom(\Semiring)$, $x \oplus x = x$.
-%
-Following the terminology of~\cite{Mohri02semiring},
-when $\forall x \in \dom(\Semiring), \one \oplus x = \one$,
-the semiring $\Semiring$ is is called \emph{bounded}.
-Note that every bounded semiring is idempotent:
-by boundedness, 
-$\one \oplus \one = \one$, and idempotency follows by multiplying
-both sides by $x$ and distributing.
-
 % \noindent
@@
 
+%\noindent
+A semiring $\Semiring$ is \emph{commutative} if $\otimes$ is commutative.
+It is \emph{idempotent} if for each $x \in \dom(\Semiring)$, $x \oplus x = x$.
+%
+%Following the terminology of~\cite{Mohri02semiring},
+%when $\forall x \in \dom(\Semiring), \one \oplus x = \one$,
+%the semiring $\Semiring$ is is called \emph{bounded}.
+%Note that every bounded semiring is idempotent:
+%by boundedness, 
+%$\one \oplus \one = \one$, and idempotency follows by multiplying
+%both sides by $x$ and distributing. 
+%
 Every idempotent semiring~$\Semiring$ induces 
-a partial ordering~$\leq_\Semiring$ 
+a partial ordering~$\leq_\oplus$ 
 called the \emph{natural ordering} of~$\Semiring$
@@
 for all $x$ and $y$,
-$x \leq_\Semiring y \;\mbox{iff}\; x \oplus y = x$.
+$x \leq_\oplus y \;\mbox{iff}\; x \oplus y = x$.
 This ordering is sometimes defined in the opposite direction~\cite{DrosteKuich09semirings};
@@
 %
-It holds that $\Semiring$ is {monotonic} \wrt~$\leq_\Semiring$.
+It holds that $\Semiring$ is {monotonic} \wrt~$\leq_\oplus$.
 An idempotent semiring $\Semiring$~is called \emph{total} if
-it~$\leq_\Semiring$ is total
+it~$\leq_\oplus$ is total
 \ie when for all $x, y \in \Semiring$, either $x \oplus y = x$ or $x \oplus y = y$.
@@
 
+\begin{example}
+semirings      
+\end{example}
+
 \subsection{Label Theory}
@@
 \noindent 
-A \emph{label theory} is made of 4 recursively enumerable sets:
+A \emph{label theory} over $\Sigma$ and $\Delta$
+is made of 4 recursively enumerable sets:
 %$\Phi_0$, \Phi_\Sigma \uplus \Phi_\Delta \uplus \Phi_{\Sigma, \Delta}$ 
@@
 \noindent
-\marginpar{maybe not necessary}
+\marginpar{necessary?}
 Finally, it is assumed that the codomain of every function of $\Phi_\Sigma$ and $\Phi_\Delta$
@@
 
+\subsection{Definitions}
 %(SWT)
@@
 function from $Q \times Q$ into %$\Phi$.
-$\< \Phi_0, \Phi_\Sigma, \Phi_\Delta, \Phi_{\Sigma, \Delta}>$.
+$\< \Phi_\epsilon, \Phi_\Sigma, \Phi_\Delta, \Phi_{\Sigma, \Delta}>$.
 \end{definition}
@@
                                          \end{array}} &
-      \wei(q, a, b, q'') \otimes \weight_\A(q'', u, v, q')\\ 
+      \wei(q, a, b, q'') \otimes \weight_T(q'', u, v, q')\\ 
 \end{array}
@@
 
+\begin{example}
+comparison of two sequences of timestamped events      
+\end{example}
+
+\noindent
 A \emph{symbolic weighted automata} (\SWA) $A = \< Q, \init, \weight, \final >$
@@
 is defined in a similar way by simply omitting the output symbols,
-\ie $\wei$ is a function of~$Q \times Q$ into %$\Phi$.
-$\< \Phi_0, \Phi_\Sigma >$, 
+\ie $\wei$ is a function from~$Q \times Q$ into %$\Phi$.
+$\< \Phi_\epsilon, \Phi_\Sigma >$, 
 or equivalently from~$Q \times (\Sigma \cup \{ \epsilon \}) \times Q$ into~$\Semiring$.
+
+\subsection{Properties}
 
@@
 \end{proposition}
-The construction time and size of $A_{s, T}$ are $O(| s | . \| T \|)$.
+The construction time and size for $A_{s, T}$ are $O(| s | . \| T \|)$.
+
+\begin{proposition}
+      ...removal of $\epsilon$ transitions for \SWA...
+      \end{proposition}
 
@@
 
+\subsection{Definition}
 Let $\Sigma$ be a countable alphabet 
 %finite (large) or infinite,
-that we assume partitioned into :
-\begin{itemize}
-\item a set $\Sigmai$ of \emph{internal symbols} denoted $a$,
-\item a set $\Sigmac$ of \emph{call symbols} denoted $\call{a}$,
-\item a set $\Sigmar$ of \emph{return symbols} denoted $\return{a}$.
-\end{itemize}
+that we assume partitioned into three subsets 
+$\Sigmai \uplus \Sigmac \uplus \Sigmac$ 
+respectively called of \emph{internal}, \emph{call} and \emph{return}
+symbols.
+% \begin{itemize}
+% \item a set $\Sigmai$ of \emph{internal symbols} denoted $a$,
+% \item a set $\Sigmac$ of \emph{call symbols} denoted $\call{a}$,
+% \item a set $\Sigmar$ of \emph{return symbols} denoted $\return{a}$.
+% \end{itemize}
 Let $\Semiring$ be a commutative semiring and  
@@
 $\return{r} \in \Sigmar$, 
-overloading the names for simplicity:
+overloading their names: % for simplicity:
 \[
@@
       In this case, the weight function $\phi_\mathsf{r}$ 
-      computes the value of matching between the call and return symbols.
-      It might be $\zero$ to express that the symbols do not match.
+      computes a value of matching between the call and return symbols.
+      This value might be $\zero$ in order to express that the symbols do not match.
 \item $\weie$ : when the stack is empty, 
@@
 \begin{array}{rcl}
-\weight_A\bigl(\config{q}{\theta}, a u, \config{q'}{\theta'}\bigr) & = & 
+\weight_A\bigl(\config{q}{\theta}, a\, u, \config{q'}{\theta'}\bigr) & = & 
  {\displaystyle\bigoplus_{q'' \in Q}} \weii(q, a, q'') 
@@
 %
-\weight_A\bigl(\config{q}{\theta}, {\call{c}} u, \config{q'}{\theta'}\bigr) & = & 
+\weight_A\bigl(\config{q}{\theta}, {\call{c}}\, u, \config{q'}{\theta'}\bigr) & = & 
  {\displaystyle\bigoplus_{\begin{array}{c}
@@
 %
-\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot \theta}, {\return{r}} u, \config{q'}{\theta'}\bigr) & = & 
+\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot \theta}, {\return{r}}\, u, \config{q'}{\theta'}\bigr) & = & 
  {\displaystyle\bigoplus_{q'' \in Q}} 
@@
 %
-\weight_A\bigl(\config{q}{\bot}, {\return{r}} u, \config{q'}{\theta'}\bigr) & = & 
+\weight_A\bigl(\config{q}{\bot}, {\return{r}}\, u, \config{q'}{\theta'}\bigr) & = & 
  {\displaystyle\bigoplus_{q'' \in Q}} \weie(q, {\return{r}}, q'') 
@@
 
+\begin{example}
+structured words...
+intro language of music notation ?
+\end{example}
+
+\subsection{Properties}
+SWVPA are closed under the binary operators of the underlying semiring.
+%
+\begin{proposition}
+Let $A_1$ and $A_2$ be two (\SWVPA)
+over the same $\Sigma$ and $\Semiring$.
+There exists two $\SWVPA$ $A_1 \oplus A_2$ and $A_1 \otimes A_2$, 
+effectively constructible, 
+such that for all $s \in \Sigma^*$, 
+$(A_1 \oplus A_2)(s) = A_1(s) \oplus A_2(s)$ and 
+$(A_1 \otimes A_2)(s) = A_1(s) \otimes A_2(s)$.
+\end{proposition}
+The construction is essentially the same 
+as the one of~\cite{DAntonyAlur14SVPDA}, in the case of the Boolean semiring.
+
 \subsection{1-best Computation}
+Assume that the semiring $\Semiring$ is commutative, idempotent, and complete.
+% total ? 
+% monotonic and superior writ natural ordering
+...
+%Regarding the infinite sum operator, note that
+%$\bigoplus_{x \in \Phi_\Sigma} \phi(x)$, 
+%$\bigoplus_{y \in \Phi_\Delta} \psi(y)$, and  
+%... exist and in $\Semiring$.
 
-Regarding the infinite sum operator, it must hold that
-$\bigoplus_{x \in \Phi_\Sigma} \phi(x)$, 
-$\bigoplus_{y \in \Phi_\Delta} \psi(y)$, and
+We propose a Dijkstra algorithm computing the minimal weight, 
+by $A$ for a word in $\Semiring^*$.
+It runs by labeling iteratively all pair $\<q, q'>$ of states of $A$ 
+by weight values in $\Semiring$, 
+in order to converge to 
+$\bigoplus_{s \in \Sigma^*} \weight_A\bigl(\config{q}{\bot}, s, \config{q'}{\bot}\bigr)$.
+%
+More precisely, we compute a marking $d : Q \times \{ \bot, \top \} \times Q$
+such that eventually:
+$d(q, \sigma ,q') = \bigoplus_{s \in \Sigma^*} \weight_A\bigl(\config{q}{\sigma}, s, \config{q'}{\sigma}\bigr)$
+where $\sigma \in \{ \bot, \top \}$, and $\top$ is a fresh stack symbol which does not belong to $\Theta$.
+Note that having a stack reduced to such a symbol makes impossible the application of the 
+two last cases in the definition of $\weight_A$ (return symbol and empty stack). 
+However, it is possible to apply the two first cases 
+(internal symbol or call symbol, with a push on the top of $\top$).
+Algorithm~\ref{algo:Dijkstra} uses a priority queue $P$.
+
+\begin{algo}[1-best for SWVPA] \label{algo:Dijkstra}
+\textbf{initially} $P$ contains all %triplets 
+$\< q_1, \bot, q_2>$ and $\<q_1, \top, q_2>$
+for $q_1, q_2 \in Q$, with 
+$d(q_1, \bot, q_2) = d(q_1, \top, q_2) = \one$ if $q_1 = q_2$ and
+$d(q_1, \bot, q_2) = d(q_1, \top, q_2) = \zero$ otherwise.
+
+\smallskip\noindent
+\textbf{while} $P$ is not empty
+
+\noindent
+\quad\textbf{extract} from $P$ $\< q_1, \sigma, q_2>$ with a minimal $d(q_1, \sigma, q_2)$ wrt $\leq_\oplus$.
+
+\noindent\quad \textbf{for all} $q_0, q_3 \in Q$ and $p \in P$ do
+
+\noindent
+\qquad
+\(\begin{array}{rrcl}
+ & d(q_1, \sigma, q_3) & \opluseq &
+   d(q_1, \sigma, q_2) \otimes \displaystyle\bigoplus_{a \in \Sigmai} \weii(q_2, a, q_3)\\
+%
+%  & d(q_0, \sigma, q_2) & \opluseq & 
+%    \displaystyle\bigoplus_{a \in \Sigmai} \weii(q_0, a, q_1) \otimes d(q_1, \sigma, q_2)\\
+%
+\mathrm{if~} \sigma = \top & d(q_0, \top, q_3) & \opluseq &
+d(q_1, \sigma, q_2) \otimes
+%\displaystyle\bigoplus_{p \in P} 
+\displaystyle\bigoplus_{{\call{c}} \in \Sigmac} 
+\displaystyle\bigoplus_{{\return{r}} \in \Sigmar} \phi({\call{c}}, {\return{r}})\\
+\mathrm{and} &  d(q_0, \bot, q_3) & \opluseq &
+d(q_1, \sigma, q_2) \otimes
+%\displaystyle\bigoplus_{p \in P} 
+\displaystyle\bigoplus_{{\call{c}} \in \Sigmac} 
+\displaystyle\bigoplus_{{\return{r}} \in \Sigmar} \phi({\call{c}}, {\return{r}})\\
+& & & \mathrm{where~} \phi = \weic(q_0, q_1, p) \otimes \weir(q_2, p, q_3)\\
+%
+\mathrm{if~} \sigma = \bot & d(q_1, \bot, q_3) & \opluseq &
+ d(q_1, \sigma, q_2) \otimes \displaystyle\bigoplus_{\return{r} \in \Sigmar} \weie(q_2, {\return{r}}, q_3)\\
+%
+%  & d(q_1, \sigma, q_3) & \opluseq &
+%  d(q_1, \sigma, q_2) \otimes d(q_2, \sigma, q_3)
+\end{array}
+\) 
+\end{algo}
+
+\noindent
+The infinite sums in the updates of $d$ in Algorithm~\ref{algo:Dijkstra} are well defined
+since~$\Semiring$ is complete.
+
+% \begin{algorithm}
+% \caption{<your caption for this algorithm>}
+% \label{<your label for references later in your document>}
+% \begin{program}
+% \seq{|line| := `` '', i := 1};
+% \WHILE i \neq n+1 \DO
+%       |line| := |item|[i] \concat `` '' \concat |number|[i]; 
+%       i := i+1; 
+%       \WHILE i \neq n+1 \AND |item|[i] = |item|[i-1] \DO 
+%       |line| := |line| \concat ``, '' \concat |number|[i]);
+%       i := i+1 \OD ; 
+%       |write|(|line|) \OD 
+% \end{program}
+% \end{algorithm}
 
@@
 
+\medskip
+In order to obtain effectively a word of $\Semiring^*$ of minimal weight, 
+we require an additional 
 property of bounded convexity of weight functions:
@@
 \item a pitch value in 0..127
-\item ON | OFF flag
+\item \textsf{on} | \textsf{off} flag
 \item a velocity value in 0..127
@@
 
-A music score ris epresented as a structured word
+A music score is represented as a structured word
 = sequence of quantified events + markups,

### 2021-05-01 main.tex
1-best algo update

@@
 \newcommand{\opluseq}{\ensuremath\mathrel{\oplus}=}
+\newcommand{\ioi}[1]{\mathsf{ioi}({#1})}
 
@@
 
-\title{Weighted Visibly Pushdown Automata and Automated Music Transcription}
+% Parsing over infinite alphabet as optimal alignment computation 
+% as edit-distance between string and language
+% Symbolic Parsing
+\title{Weighted Visibly Pushdown Automata\\ Application to Automated Music Transcription}
 \author{Florent Jacquemard}
@@
 
-weighted parsing
+problem
+quantitative parsing 
+or symbolic parsing
+parsing of words over infinite input alphabet.
 
-principle of quantitative parsing / parsing of words over infinite alphabet:
+generalizes weighted parsing: 
+finding the best derivation for a weighted grammar.
+
+model of weighted CFG over infinite set of terminal symbol
+(but finite set non-terminals and production rules).
+in production rules, terminal symbols appear as variables (parameters)
+and the weight associated to the production rule is a function of these variables.
+It is the approach of Symbolic Automata, except that the domain of weight values 
+is not restricted to be Boolean, like  for the guards in the rules of SA, 
+but can be an arbitrary commutative semiring (assuming some restrictions).
+
+wrt an edit distance
+
+we consider a strictly restricted model of weighted grammars : VPA 
+and general edit-distances that can be defined by a weighted word transducer
 
@@
 \noindent
-\quad\textbf{extract} from $P$ $\< q_1, \sigma, q_2>$ with a minimal $d(q_1, \sigma, q_2)$ wrt $\leq_\oplus$.
+\quad\textbf{extract} $\< q_1, \sigma, q_2>$ from $P$ 
+ such that $d(q_1, \sigma, q_2)$ is minimal wrt $\leq_\oplus$.
 
@@
 \displaystyle\bigoplus_{{\call{c}} \in \Sigmac} 
-\displaystyle\bigoplus_{{\return{r}} \in \Sigmar} \phi({\call{c}}, {\return{r}})\\
+\displaystyle\bigoplus_{{\return{r}} \in \Sigmar} \eta({\call{c}}, {\return{r}})\\
 \mathrm{and} &  d(q_0, \bot, q_3) & \opluseq &
@@
 \displaystyle\bigoplus_{{\call{c}} \in \Sigmac} 
-\displaystyle\bigoplus_{{\return{r}} \in \Sigmar} \phi({\call{c}}, {\return{r}})\\
-& & & \mathrm{where~} \phi = \weic(q_0, q_1, p) \otimes \weir(q_2, p, q_3)\\
+\displaystyle\bigoplus_{{\return{r}} \in \Sigmar} \eta({\call{c}}, {\return{r}})\\
+& & & \mathrm{where~} \eta = \weic(q_0, q_1, p) \otimes \weir(q_2, p, q_3)\\
 %
@@
 %
-%  & d(q_1, \sigma, q_3) & \opluseq &
-%  d(q_1, \sigma, q_2) \otimes d(q_2, \sigma, q_3)
+& d(q_0, \bot, q_2) & \opluseq & 
+d(q_0, \bot, q_1) \otimes d(q_1, \sigma, q_2), \mathrm{if~} \< q_0, \bot, q_1> \notin P \\
+%
+\mathrm{if~} \sigma = \top & d(q_0, \top, q_2) & \opluseq & 
+d(q_0, \top, q_1) \otimes d(q_1, \sigma, q_2), \mathrm{if~} \< q_0, \top, q_1> \notin P \\
+%
+& d(q_1, \bot, q_3) & \opluseq & 
+   d(q_1, \sigma, q_2) \otimes d(q_2, \bot, q_3), \mathrm{if~} \< q_2, \bot, q_3> \notin P \\
+%
+\mathrm{if~} \sigma = \top & d(q_1, \top, q_3) & \opluseq & 
+    d(q_1, \sigma, q_2) \otimes d(q_2, \top, q_3), \mathrm{if~} \< q_2, \top, q_3> \notin P \\
 \end{array}
@@
 \subsection{Representation of Music Performances}
-A musical performance is represented symbolically as a finite sequence of timestamped events.
-similar to piano roll representation~\cite{Muller15fundamentals} chap.1
+We consider symbolic representation of musical performances, as finite sequences of events.
+It corresponds to the concrete case of a MIDI file~\cite{SMF} 
+recorded  from an electronic keyboard, 
+or the output of a transcription from audio files~\cite{Benetos18AMTsurvey}.
+%
+For the sake of simplicity, we shall only consider here the case of monophonic performances, 
+where at most one note is sounding at a time -- the approach however extends to the polyphonic case.
 
-\noindent
-We consider an infinite alphabet $\Sigma$ of MIDI-like 
-events made of:
-\begin{itemize}
-\item a timestamp in RTU\\ %(real or int = nb samples ?)
-      or IOI in RTU
-\item a pitch value in 0..127
-\item \textsf{on} | \textsf{off} flag
-\item a velocity value in 0..127
-\end{itemize}
+A music performance is a finite sequence of events in a set $\Sigma$.
+Every event $e \in \Sigma$ has attributes such from a finite domain, 
+like a number of key for a note 
+or a flag indicating that it is a rest 
+(\textsf{ON} and \textsf{OFF} messages in~\cite{SMF})
+and a velocity value (0..127 in~\cite{SMF})).
+%This representation is similar to the piano roll ~\cite{Muller15fundamentals} chap.1. 
+Moreover, it contains a RTU value $\ioi{e}$ (real number) 
+representing the time distance to the next event, or the to the end of performance 
+(also called \emph{inter-onset interval}).
 
 \subsection{Representation of Music Scores}
-we consider here the case of monodies for the sake of simplificity.
-
-A music score is represented as a structured word
-= sequence of quantified events + markups,
-see nested words~\cite{AlurMadhusudan09nested}.
+Music score are represented as structured words
+made of quantified events and parenthesized markups,
+akin of nested words~\cite{AlurMadhusudan09nested}.
 
@@
 composed of a tag, in a finite set $\Xi$, 
-and an IOI duration value.
-Moreover, $\Delta$ is partitioned into 
+and an MTU (rational) IOI duration value.
+The alphabet $\Delta$ is partitioned into 
 $\Delta = \Deltai \uplus \Deltac \uplus \Deltar$, 
 like in Section~\ref{section:SWVPA}.
-
+%
 \noindent
-The elements of $\Deltai$ represent events:
+The symbols of $\Deltai$ represent events:
 % (infinite alphabet of internal symbols) made of:
-whose tags are one of:
-\begin{itemize}
-\item continuation (0) : tie or dot
-\item note, grace note (pitch) or chord (pitch+)
-\item rest
-\item ...
-\end{itemize}
+with tags indicate a new note or grace-note (with null IOI), 
+a rest or the continuation of the previous note (tie or dot).
+The elements of $\Deltac \uplus \Deltar$ are matched
+and are the markups for describing the structure of the score.
 
@@
 with a transducer
+
+tempo value in a finite domain (e.g. 30..300 bpm)
+can be fixed 
+or recomputed by the transducer when reading each event, 
+according to a perceptive/cognitive model of tempo 
+such as~\cite{LargeJones99tempo}
+(also used in the context of score following~\cite{Cont10TPAMI}).
+% we wont detail here.
 

### 2021-05-02 main.tex
lemmas n-best, application, debut intro

@@
 
-\def\SWT{\textsf{SWT}\xspace}
-\def\SWA{\textsf{SWA}\xspace}
-\def\SWVPA{\textsf{SWVPA}\xspace}
+\def\SWT{\textsf{swT}\xspace}
+\def\SWA{\textsf{swA}\xspace}
+\def\SWVPA{\textsf{sw-VPA}\xspace}
 \def\weight{\mathsf{weight}}
@@
 
-problem
-quantitative parsing 
-or symbolic parsing
-parsing of words over infinite input alphabet.
+the problem
+%quantitative parsing or symbolic parsing
+of parsing of words over infinite input alphabet.
 
-generalizes weighted parsing: 
-finding the best derivation for a weighted grammar.
+A weighted CFG associates a weight value to each of its derivation trees.
+There exists some specialized parsing algorithms~\cite{nbest} 
+to compute the $n$ best derivations of a given input word.
 
-model of weighted CFG over infinite set of terminal symbol
-(but finite set non-terminals and production rules).
-in production rules, terminal symbols appear as variables (parameters)
-and the weight associated to the production rule is a function of these variables.
-It is the approach of Symbolic Automata, except that the domain of weight values 
+Such models and algorithm only deal with finite input alphabets.
+Considering infinite alphabets can however be of practical interest, 
+for instance when dealing with sequences of timestamped events.
+...example of music transcription\dots
+
+%generalizes weighted parsing: 
+%finding the best derivation for a weighted grammar.
+
+We propose here models of weighted recognizers 
+model of weighted CFG over infinite input alphabets %sets of terminal symbol
+(but with finite sets of states and transitions rules). 
+% non-terminals and production rules).
+In the production rules, input symbols appear as variables (parameters)
+and the weight associated to the transition rule is a function of these variables.
+%
+It is the approach of Symbolic Automata~\cite{VeanesdAntoniJACM}, 
+except that the domain of weight values 
 is not restricted to be Boolean, like  for the guards in the rules of SA, 
@@
 
-wrt an edit distance
+models: automata, transducer and pushdown automata with visibly restriction.
+%wrt an edit distance
+The latter corresponds in expressiveness
+to a strict restricted model of weighted grammars. %VPA 
+That permits to define a general approach for parsing 
+over infinite input alphabets.
 
-we consider a strictly restricted model of weighted grammars : VPA 
-and general edit-distances that can be defined by a weighted word transducer
+First one general edit-distance is defined by a weighted word 
+transducer~\cite{Mohri}
+%Symbolic automata transducers are extended models~\cite{VeanesdAntoniJACM}
+%dealing with infinite set of input symbols...
+value in a semiring...
 
@@
 which are countable (finite or infinite) sets of symbols, 
-and let $\Semiring$ be a commutative semiring.
+and let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative semiring.
 
@@
 \begin{definition}
-A \emph{symbolic-weighted transducer} $T$ over the input and output alphabet $\Sigma$ and $\Delta$ 
+A \emph{symbolic-weighted transducer} (\SWT) 
+$T$ over the input and output alphabet $\Sigma$ and $\Delta$ 
 and the semiring $\Semiring$ is a tuple
@@
 is then defined as follows: 
-\[
+\begin{equation}
 T(s, t)  = 
@@
 \mathop{\otimes} \weight_T(q, s, t, q') \mathop{\otimes} \mathsf{out}(q').
-\]
+\label{eq:weightT}
+\end{equation}
 
@@
 % \end{itemize}
-Let $\Semiring$ be a commutative semiring and  
+Let~$\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative semiring and  
 let  $(\Phi_\epsilon, \Phic, \Phir, \Phicr)$ be a label theory over $\Semiring$
@@
 following empty stack semantics: 
-\[
+\begin{equation}
 A(s)  = 
@@
 \mathop{\otimes} \mathsf{out}(q').
-\]
+\label{eq:weightA}
+\end{equation}
 
@@
 \subsection{Properties}
-SWVPA are closed under the binary operators of the underlying semiring.
+The class of \SWVPA is closed under the binary operators of the underlying semiring.
 %
@@
 
-\subsection{1-best Computation}
-Assume that the semiring $\Semiring$ is commutative, idempotent, and complete.
+\subsection{Best-first Search}
+Assume that the semiring $\Semiring$ is 
+commutative, idempotent, superior, total and complete.
 % total ? 
 % monotonic and superior writ natural ordering
-...
 %Regarding the infinite sum operator, note that
@@
 %... exist and in $\Semiring$.
+We propose a Dijkstra algorithm computing the minimal weight by $A$, 
+\wrt $\leq_\oplus$, for a word in $\Semiring^*$.
 
-We propose a Dijkstra algorithm computing the minimal weight, 
-by $A$ for a word in $\Semiring^*$.
-It runs by labeling iteratively all pair $\<q, q'>$ of states of $A$ 
-by weight values in $\Semiring$, 
-in order to converge to 
-$\bigoplus_{s \in \Sigma^*} \weight_A\bigl(\config{q}{\bot}, s, \config{q'}{\bot}\bigr)$.
+\noindent
+More precisely, 
+let $\top$ be a fresh stack symbol which does not belong to $\Theta$,
+and for every two states $q, q' \in Q$ and $\sigma \in \{ \bot, \top \}$, 
+let 
+\[ 
+  d_0(q, \sigma, q') = \bigoplus_{s\in \Sigma^*} 
+  \textstyle
+  \weight_A\bigl(\config{q}{\sigma}, s, \config{q'}{\sigma}\bigr). 
+\]
+Since $\Semiring$ is complete, this infinite sum is well defined, 
+and since $\leq_\oplus$ is assumed total, it is the minimum in $\Sigma^*$
+of $s \mapsto \weight_A(\config{q}{\sigma}, s, \config{q'}{\sigma})$ \wrt this ordering.
 %
-More precisely, we compute a marking $d : Q \times \{ \bot, \top \} \times Q$
-such that eventually:
-$d(q, \sigma ,q') = \bigoplus_{s \in \Sigma^*} \weight_A\bigl(\config{q}{\sigma}, s, \config{q'}{\sigma}\bigr)$
-where $\sigma \in \{ \bot, \top \}$, and $\top$ is a fresh stack symbol which does not belong to $\Theta$.
-Note that having a stack reduced to such a symbol makes impossible the application of the 
-two last cases in the definition of $\weight_A$ (return symbol and empty stack). 
-However, it is possible to apply the two first cases 
-(internal symbol or call symbol, with a push on the top of $\top$).
-Algorithm~\ref{algo:Dijkstra} uses a priority queue $P$.
+When $\sigma = \bot$, $d_0(q, \sigma, q')$ is the central 
+expression in a term of the definition (\ref{eq:weightA}) of $A(s_0)$ for the minimum $s_0$
+(for the above function).
+When $\sigma = \bot$, intuitively, it is the minimum weight of a computation of $A$
+starting in state $q$ with a stack $\theta \in \Theta^*$ (possibly empty), 
+and ending in state $q$ with the same stack $\theta$, such that moreover 
+the computation does not touch a symbol of $\theta$. 
+That means that during the computation, 
+$A$ may apply the first case of in the definition of $\weight_A$ (internal symbol),
+as well as the second case, to can push call symbols on the top of $\theta$, 
+and may pop these symbols with the third case (return symbol). 
+However, it cannot apply one of the two last cases (return symbol and empty stack)
+when the current stack is $\theta$.
+%pop symbols in $\theta$.
+% Note that having a stack reduced to such a symbol makes impossible the application of the 
+% two last cases in the definition of $\weight_A$ (return symbol and empty stack). 
+% However, it is possible to apply the two first cases 
+% (internal symbol or call symbol, with a push on the top of $\top$).
 
-\begin{algo}[1-best for SWVPA] \label{algo:Dijkstra}
-\textbf{initially} $P$ contains all %triplets 
-$\< q_1, \bot, q_2>$ and $\<q_1, \top, q_2>$
-for $q_1, q_2 \in Q$, with 
+The algorithm~\ref{algo:Dijkstra}
+constructs iteratively a a marking $d : Q \times \{ \bot, \top \} \times Q \to \Semiring$
+%of the triplets $\<q, \sigma, q'>$ 
+%of states of $A$ by weight values in $\Semiring$, 
+that converges eventually to $d_0(q, \sigma, q')$. 
+%It uses for that purpose a priority queue $P$ containing triplets of 
+%$Q \times \{ \bot, \top \} \times Q$.
+
+\begin{algo}[1-best for \SWVPA] \label{algo:Dijkstra}
+\textbf{initially} let $P = Q \times \{ \bot, \top \} \times Q$, %contains all %triplets 
+%$\< q_1, \bot, q_2>$ and $\<q_1, \top, q_2>$ for $q_1, q_2 \in Q$, with 
+and 
 $d(q_1, \bot, q_2) = d(q_1, \top, q_2) = \one$ if $q_1 = q_2$ and
@@
 %
-& d(q_0, \bot, q_2) & \opluseq & 
-d(q_0, \bot, q_1) \otimes d(q_1, \sigma, q_2), \mathrm{if~} \< q_0, \bot, q_1> \notin P \\
+%& d(q_0, \bot, q_2) & \opluseq & 
+%d(q_0, \bot, q_1) \otimes d(q_1, \sigma, q_2), \mathrm{if~} \< q_0, \bot, q_1> \notin P \\
 %
-\mathrm{if~} \sigma = \top & d(q_0, \top, q_2) & \opluseq & 
-d(q_0, \top, q_1) \otimes d(q_1, \sigma, q_2), \mathrm{if~} \< q_0, \top, q_1> \notin P \\
+%\mathrm{if~} \sigma = \top & d(q_0, \top, q_2) & \opluseq & 
+%d(q_0, \top, q_1) \otimes d(q_1, \sigma, q_2), \mathrm{if~} \< q_0, \top, q_1> \notin P \\
 %
@@
 since~$\Semiring$ is complete.
+The algorithm performs $2.|Q|^2$ iterations until $P$ is empty, 
+and each iteration has a time complexity $O(|Q|^2 . |P|)$.
+This gives a time complexity $O(|Q|^4 . |P|)$. 
+It can be reduced by implementing $P$ as a priority queue, 
+prioritized by the value returned by $d$
+***complete***. %$|Q|^3.\log(|Q|^2)$
+
+The correctness of Algorithm~\ref{algo:Dijkstra} 
+is ensured by the invariant expressed in the following lemma.
+\begin{lemma}
+For all $\< q_1, \sigma, q_2> \notin P$, $d(q_1, \sigma, q_2) =  d_0(q_1, \sigma, q_2)$.
+\end{lemma}
+The proof is by contradiction, 
+assuming a counter-example minimal in the length of the witness word.
 
@@
 
-\medskip
-In order to obtain effectively a word of $\Semiring^*$ of minimal weight, 
-we require an additional 
-property of bounded convexity of weight functions:
+%\begin{lemma}
+\noindent
+For computing the minimal weight of a computation of $A$, we use the fact that,
+at the termination of Algorithm~\ref{algo:Dijkstra}, %it holds that,
+%There exist $q_1, q_2 \in Q$
+\[
+  {\displaystyle \bigoplus_{s \in \Sigma^*} A(s)} = 
+  {\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
+  \mathsf{in}(q) \mathop{\otimes} d(q, \bot, q') \mathop{\otimes} \mathsf{out}(q').
+\]
+%\end{lemma}
+
+%\medskip
+\noindent
+In order to obtain effectively a witness 
+(word of $\Semiring^*$ with computation of $A$ of minimal weight), 
+we require an additional property the of weight functions.
+\begin{definition}
+Let $\Sigma$ be an alphabet and $\Semiring$ a complete semiring.
+A function $\phi$ from $\Sigma^n$ into $\Semiring$
+is called $k$-\emph{convex} for a natural number $k$ iff 
+$\mathit{card}\{ \mathbf{a} \in \Sigma^n \mid \phi(\mathbf{a}) = 
+                 \displaystyle\bigoplus_{\mathbf{p} \in \Sigma^n} \phi(\mathbf{p}) \} \leq k$.
+\end{definition}
+A label theory is $k$-convex if all its functions are $k$-convex.
+
+\begin{proposition}
+For a \SWVPA $A$ over a 
+commutative, idempotent, superior, total and complete semiring 
+and an alphabet $\Sigma$ with a $k$-convex label theory, 
+one can construct in PTIME a word $s \in \Sigma^*$ 
+such that $A(s)$ is minimal \wrt the natural ordering for $\Semiring$. 
+%$\leg_\oplus$.
+\end{proposition}
 
@@
 \subsection{Representation of Music Performances}
-We consider symbolic representation of musical performances, as finite sequences of events.
+We consider symbolic representations of musical performances, as finite sequences of events.
 It corresponds to the concrete case of a MIDI file~\cite{SMF} 
@@
 %
-For the sake of simplicity, we shall only consider here the case of monophonic performances, 
-where at most one note is sounding at a time -- the approach however extends to the polyphonic case.
+For the sake of simplicity, 
+we shall only consider here the case of monophonic performances, 
+where at most one note is sounding at a time. 
+The approach however extends to the polyphonic case.
 
-A music performance is a finite sequence of events in a set $\Sigma$.
+A music performance is a finite sequence of events in a set~$\Sigma$.
 Every event $e \in \Sigma$ has attributes such from a finite domain, 
@@
 Moreover, it contains a RTU value $\ioi{e}$ (real number) 
-representing the time distance to the next event, or the to the end of performance 
-(also called \emph{inter-onset interval}).
+representing the time distance to the next event, 
+or to the end of performance for the last event,
+also called \emph{inter-onset interval}.
 
@@
 Music score are represented as structured words
-made of quantified events and parenthesized markups,
+made of timed %quantified 
+events and parenthesized markups,
 akin of nested words~\cite{AlurMadhusudan09nested}.
@@
 and an MTU (rational) IOI duration value.
-The alphabet $\Delta$ is partitioned into 
+%The alphabet $\Delta$ 
+It is partitioned into 
 $\Delta = \Deltai \uplus \Deltac \uplus \Deltar$, 
@@
 % (infinite alphabet of internal symbols) made of:
-with tags indicate a new note or grace-note (with null IOI), 
+with tags indicating a new note or grace-note (with null IOI), 
 a rest or the continuation of the previous note (tie or dot).
+%
 The elements of $\Deltac \uplus \Deltar$ are matched
-and are the markups for describing the structure of the score.
+markups for describing the structure of the score, 
+\ie the hierarchical grouping of events, and also, 
+importantly the division of time in measures, tuplets...
+%- parentheses for time divisions : tuplets, bars...
+(linearization of rhythm trees \cite{jacquemard:hal-01138642}...).
+They contain additional info such as tuple number, beaming policy...
 
 \noindent
-The elements of $\Deltac$ and $\Deltar$ are 
-markups for the representation of groups of events
-(linearization of rhythm trees \cite{jacquemard:hal-01138642}...)
-- parentheses for time divisions : tuplets, bars...
-tag %labels: 
-contain info such as tuple number, beaming policy...
+The duration values of letters of $\Delta$, in MTU (rational), 
+can be computed with the markups and tags (\eg grace note has duration 0).
 
-\noindent
-The date or duration of events, in MTU (rational), 
-can be computed with the markups and tags (e.g. grace note has duration 0).
+%\noindent
+%There are simultaneous events, since grace notes has duration 0. They are ordered.
+%
+%\noindent
+%Finite bound on the number of duration ratio. ?
 
-\noindent
-There are simultaneous events, since grace notes has duration 0. They are ordered.
-
-\noindent
-Finite bound on the number of duration ratio. ?
+\begin{example}
+...      
+\end{example}
 
 \subsection{Performance/Score Distance Computation}
-with a transducer
+\label{app:distance}
+We define a distance between performance and score representations
+by a $\SWT$ $T = \< Q, \init, \wei, \final >$, over a semiring $\Semiring$.
+** detail the elements of $\Semiring$ ....**
+%are quadruplets of the form
+%$\< t, s, \delta_t, \delta_s>$
 
-tempo value in a finite domain (e.g. 30..300 bpm)
-can be fixed 
-or recomputed by the transducer when reading each event, 
+Every state of $Q$ contains a 
+tempo value in a finite domain (e.g. 30..300 bpm).
+This value can be fixed 
+or recomputed by the $T$ %transducer 
+after reading each event, 
 according to a perceptive/cognitive model of tempo 
@@
 % we wont detail here.
+
+\subsection{Transcription by Best-first Search}
+We assume a score language defined by a \SWVPA over the semiring 
+$\Semiring$ of Section~\ref{app:distance}.
 

### 2021-05-02 polyphony.tex
lemmas n-best, application, debut intro

\section{Application}
Symbolic Automated Music Transcription
and analysis of music performances

\subsection{Time Scales}
Real-Time Unit (RTU) = seconds

\noindent 
Musical-Time Unit (MTU) = number of measures

\noindent 
conversion via tempo value

\subsection{Representation of Music Performances}
A musical performance is represented symbolically as a finite sequence of timestamped events.
This representation is similar to the piano roll ~\cite{Muller15fundamentals} chap.1.

\noindent
We consider an infinite alphabet $\Sigma$ of MIDI-like 
events made of:
\begin{itemize}
\item a timestamp in RTU\\ %(real or int = nb samples ?)
      or IOI in RTU
\item a pitch value in 0..127
\item \textsf{on} | \textsf{off} flag
\item a velocity value in 0..127
\end{itemize}

### 2021-05-03 main.tex
intro (debut)

@@
 % as edit-distance between string and language
-% Symbolic Parsing
-\title{Weighted Visibly Pushdown Automata\\ Application to Automated Music Transcription}
+% 
+\title{Symbolic Parsing and\\ Automated Music Transcription}
+%\title{Weighted Visibly Pushdown Automata\\ Automated Music Transcription}
 \author{Florent Jacquemard}
@@
 
-the problem
-%quantitative parsing or symbolic parsing
-of parsing of words over infinite input alphabet.
+%the problem: quantitative parsing or symbolic parsing
+%of parsing of words over infinite input alphabet.
 
-A weighted CFG associates a weight value to each of its derivation trees.
-There exists some specialized parsing algorithms~\cite{nbest} 
+Weighted parsing~\cite{Goodman99SemiringParsing,Nederhof03weightedParsing,MorbitzVogler19weighted-parsing} 
+consists in analyzing a word in input \wrt a weighted language model, \eg a probabilistic context-free grammar (CFG).
+In such a model, a weight value is associated to each transition rule, % production rule
+and the rule's weights can be combined with a associative product operator $\otimes$ into 
+the weight of a derivation (or \emph{abstract syntax tree}, \ie the result of an analyze).
+A second operator $\oplus$, associative and commutative, 
+is used to handle ambiguity of the model, 
+by summing the weights of the possibly several (in general exponentially many) 
+syntax trees associated to a given input word.
+Intuitively, $\oplus$ aims at selecting, or ranking, the syntax trees.
+The weight domain, equipped with these two operators is assumed, at minima, 
+to form a \emph{semiring} $\Semiring$. %\cite{Eilenberg74automata}
+%by ranking 
+%making the weight domain a semiring.
+Some efficient specialized parsing algorithms~\cite{Huang05kbest} have been proposed in this context 
+in order to compute the $n$ best syntax trees of a given input word without having to enumerate them all.
+Generally based on dynamic programming, these algorithms rely on 
+additional algebraic properties of $\Semiring$.
+-- see \eg~\cite{Huang05kbest} for some NLP applications.
+%The extraction of $n$ best list is useful
+
+** DEL ** 
+%Parsing produces an analyze of an input word in the form of a parse tree 
+%representing the derivation of a grammar.
+In case of ambiguity, several analyses, in general an exponential number, 
+are possible for the same input word. 
+Weighted CFG associate a weight value to each of its derivation trees, 
+based on weighted values associated to each production rules, 
+and are therefore useful for the compact representation and the ranking of large sets of parse trees.
+Moreover, there exist some efficient specialized parsing algorithms~\cite{Huang05kbest}
 to compute the $n$ best derivations of a given input word.
 
-Such models and algorithm only deal with finite input alphabets.
+The models and algorithms mentioned above deal with finite input alphabets.
 Considering infinite alphabets can however be of practical interest, 
 for instance when dealing with sequences of timestamped events.
-...example of music transcription\dots
+(approach of music transcription by parsing~\cite{MCM19})
 
@@
 
-We propose here models of weighted recognizers 
-model of weighted CFG over infinite input alphabets %sets of terminal symbol
-(but with finite sets of states and transitions rules). 
+We propose here some models of weighted recognizers 
+%model of weighted CFG 
+computing on words over infinite input alphabets %sets of terminal symbol
+- but with finite sets of states and transitions rules. 
 % non-terminals and production rules).
-In the production rules, input symbols appear as variables (parameters)
+In their transition rules, input symbols appear as variables (parameters)
 and the weight associated to the transition rule is a function of these variables.
 %
-It is the approach of Symbolic Automata~\cite{VeanesdAntoniJACM}, 
+This approach is close to the one of Symbolic Automata (SA)~\cite{VeanesdAntoniJACM}, 
 except that the domain of weight values 
@@
 
-models: automata, transducer and pushdown automata with visibly restriction.
+The models defined here are finite automata, transducers and pushdown automata with visibly restriction.
 %wrt an edit distance
+The latter model of these automata computes on nested words~\cite{AlurMadhusudan09nested}, 
+a structured format of words parenthesized with markup symbols, 
+which can correspond to a linearization of trees.
+
 The latter corresponds in expressiveness
-to a strict restricted model of weighted grammars. %VPA 
-That permits to define a general approach for parsing 
-over infinite input alphabets.
-
+to a strict restricted model of weighted grammars
+and can be used in a general approach for parsing over infinite input alphabets.
+\dots
 First one general edit-distance is defined by a weighted word 
@@
 between words with weighted transducers, and propose models of weighted automata 
-and weighted transducers over infinite alphabets.
+and weighted transducers handling infinite alphabets.
 

### 2021-05-03 main.tex
intro (cleanup)

@@
 
-** DEL ** 
-%Parsing produces an analyze of an input word in the form of a parse tree 
-%representing the derivation of a grammar.
-In case of ambiguity, several analyses, in general an exponential number, 
-are possible for the same input word. 
-Weighted CFG associate a weight value to each of its derivation trees, 
-based on weighted values associated to each production rules, 
-and are therefore useful for the compact representation and the ranking of large sets of parse trees.
-Moreover, there exist some efficient specialized parsing algorithms~\cite{Huang05kbest}
-to compute the $n$ best derivations of a given input word.
-
-The models and algorithms mentioned above deal with finite input alphabets.
-Considering infinite alphabets can however be of practical interest, 
-for instance when dealing with sequences of timestamped events.
-(approach of music transcription by parsing~\cite{MCM19})
-
+The various approaches to weighted parsing usually assume a finite input alphabet. %models and algorithms 
+Considering infinite or large alphabets can however be of practical interest, 
+for instance when processing strings in UTF encodings or for vulnerability detection in Web-applications 
+(see \eg~\cite{dAntoni21CACM}), for model checking, 
+or when dealing with sequences of timestamped events
+(like in the approach of~\cite{MCM19} of automated music transcription by parsing).
 %generalizes weighted parsing: 
@@
 %
-This approach is close to the one of Symbolic Automata (SA)~\cite{VeanesdAntoniJACM}, 
-except that the domain of weight values 
-is not restricted to be Boolean, like for the guards in the rules of SA, 
+This approach is close to the case of Symbolic Automata (SA)~\cite{VeanesdAntoniJACM}, 
+except that the domain for weight values is not restricted to be Boolean, 
+like for the guards in the rules of SA, 
 but can be an arbitrary commutative semiring (assuming some restrictions).
-
+%
 The models defined here are finite automata, transducers and pushdown automata with visibly restriction.

### 2021-05-03 main.tex
intro

@@
 % 
-\title{Symbolic Parsing and\\ Automated Music Transcription}
+\title{Symbolic Weighted Parsing and\\ Automated Music Transcription}
 %\title{Weighted Visibly Pushdown Automata\\ Automated Music Transcription}
@@
 The weight domain, equipped with these two operators is assumed, at minima, 
-to form a \emph{semiring} $\Semiring$. %\cite{Eilenberg74automata}
+to form a \emph{semiring} $\Semiring$, %\cite{Eilenberg74automata}
+such as a min-plus algebra of domain $\mathbb{R} \cup \{ +\infty\}$, 
+where $\oplus$ is min and $\otimes$ is plus.
 %by ranking 
@@
 Generally based on dynamic programming, these algorithms rely on 
-additional algebraic properties of $\Semiring$.
+additional algebraic properties of~$\Semiring$.
 -- see \eg~\cite{Huang05kbest} for some NLP applications.
@@
 Considering infinite or large alphabets can however be of practical interest, 
-for instance when processing strings in UTF encodings or for vulnerability detection in Web-applications 
-(see \eg~\cite{dAntoni21CACM}), for model checking, 
+for instance in the context of processing strings in UTF16 encodings, 
+for vulnerability detection in Web-applications 
+(see \eg~\cite{dAntoni21CACM}), 
+for the representation of execution traces in model checking, 
 or when dealing with sequences of timestamped events
-(like in the approach of~\cite{MCM19} of automated music transcription by parsing).
+-- we describe briefly at the end of the paper a parsing based approach to 
+of automated music transcription, that is the conversion of a symbolic 
+representation of a music performance into a structured music score.
+%\cite{MCM19} 
 %generalizes weighted parsing: 
@@
 
-We propose here some models of weighted recognizers 
+We present here some weighted language models
 %model of weighted CFG 
@@
 % non-terminals and production rules).
-In their transition rules, input symbols appear as variables (parameters)
-and the weight associated to the transition rule is a function of these variables.
+In their transition rules, input symbols appear as variables %(parameters)
+and the weight associated to a transition rule is a function of these variables.
 % 
-This approach is close to the case of Symbolic Automata (SA)~\cite{VeanesdAntoniJACM}, 
+This approach is close to the case of Symbolic Automata (SA)~\cite{dAntoni21CACM}, 
 except that the domain for weight values is not restricted to be Boolean, 
@@
 %
-The models defined here are finite automata, transducers and pushdown automata with visibly restriction.
+The models defined here are finite automata (\SWA), 
+transducers (\SWT) and pushdown automata
+with a visibly restriction~\cite{AlurMadhusudan09nested} (\SWVPA).
 %wrt an edit distance
-The latter model of these automata computes on nested words~\cite{AlurMadhusudan09nested}, 
-a structured format of words parenthesized with markup symbols, 
-which can correspond to a linearization of trees.
+The latter model of automata computes on \emph{nested words}~\cite{AlurMadhusudan09nested}, 
+a structured form of words parenthesized with markup symbols, 
+corresponding to a linearization of trees.
+In expressiveness, they correspond to a strict restriction
+of weighted CFG. 
+%and can be used in a general approach for parsing over infinite input alphabets.
+Let $A$ be a \SWVPA, associating $A(w) \in \Semiring$ 
+to a given a nested word $w$ (representing a parse tree),
+and let a \SWT compute a distance $d$, in $\Semiring$, 
+between 2 strings over respectively an infinite input alphabet and the 
+same (infinite) alphabet of $A$.
+Then, the problem of Symbolic Weighted Parsing is, 
+given an input string $s$, to find a nested word $w$ minimizing 
+(according to the ranking defined by $\oplus$)
+the distance $d(s, w) \otimes A(w)$ between $s$ and $A$, 
+as defined in~\cite{Mohri03EDWA}.
+%following
 
-The latter corresponds in expressiveness
-to a strict restricted model of weighted grammars
-and can be used in a general approach for parsing over infinite input alphabets.
-\dots
-First one general edit-distance is defined by a weighted word 
-transducer~\cite{Mohri}
-%Symbolic automata transducers are extended models~\cite{VeanesdAntoniJACM}
-%dealing with infinite set of input symbols...
-value in a semiring...
+% First one general edit-distance is defined by a weighted word 
+% transducer~\cite{Mohri}
+% %Symbolic automata transducers are extended models~\cite{VeanesdAntoniJACM}
+% %dealing with infinite set of input symbols...
+% value in a semiring...
+
+In Section~\ref{section:SWA} we introduce \SWA and \SWT, 
+then \SWVPA are defined in Section~\ref{section:SWPDA}, 
+where a polynomial 1-best algorithm is described that can be use to solve
+Symbolic Weighted Parsing.
+Finally, \label{sec:transcription}, we present an application 
+of this approach to automated music transcription that has been implemented.
 
@@
 \label{section:transducer}
+\label{section:SWA}
+\label{section:SWT}
 
@@
 \section{SW Visibly Pushdown Automata}
-\label{section:SWVPA}
+\label{section:SWVPA}\label{sec:SWVPA}
 The following model generalizes Symbolic VPA~\cite{DAntonyAlur14SVPDA}
@@
 \section{Application}
+\label{sec:transcription}
 Symbolic Automated Music Transcription

### 2021-05-04 main.tex
intro
abstract

@@
-\documentclass[a4paper,11pt]{article}
+\documentclass[runningheads]{llncs}
+%\documentclass[a4paper,11pt]{article}
 %\setcounter{page}{1}
@@
 \usepackage{theorem}
-\newtheorem{theorem}{Theorem} %[section]
-\newtheorem{definition}[theorem]{Definition}
-\newtheorem{lemma}[theorem]{Lemma}
-\newtheorem{corollary}[theorem]{Corollary}
-\newtheorem{proposition}[theorem]{Proposition}
-\newenvironment{proof}{\vspace{-2ex}{\it Proof. }}{\hspace*{\fill} $\Box$\smallskip }
-\theorembodyfont{\slshape}
-\newtheorem{example}[theorem]{Example}
-\newtheorem{remark}[theorem]{Remark}
+% \newtheorem{theorem}{Theorem} %[section]
+% \newtheorem{definition}[theorem]{Definition}
+% \newtheorem{lemma}[theorem]{Lemma}
+% \newtheorem{corollary}[theorem]{Corollary}
+% \newtheorem{proposition}[theorem]{Proposition}
+% \newenvironment{proof}{\vspace{-2ex}{\it Proof. }}{\hspace*{\fill} $\Box$\smallskip }
+% \theorembodyfont{\slshape}
+% \newtheorem{example}[theorem]{Example}
+% \newtheorem{remark}[theorem]{Remark}
 {\theorembodyfont{\rmfamily} \theoremstyle{break} \newtheorem{algo}{Algorithm}}
@@
 % 
-\title{Symbolic Weighted Parsing and\\ Automated Music Transcription}
+%\title{Symbolic Weighted Parsing and Automated Music Transcription}
+\title{Symbolic Weighted Language Models and Automated Music Transcription}
 %\title{Weighted Visibly Pushdown Automata\\ Automated Music Transcription}
-\author{Florent Jacquemard}
+%\author{Florent Jacquemard}
 %\institute{INRIA \& CNAM, Paris, France\\
@@
 %\titlerunning{WVPA \& AMT}
+\titlerunning{Symbolic Weighted Parsing and Automated Music Transcription}
 %\authorrunning{Florent Jacquemard}
@@
 
-%\begin{abstract}
+\begin{abstract}
 %Symbolic Weighted (SW) extension of symbolic automata...
-%\end{abstract}
+
+We propose language models called Symbolic Weighted Automata (SWA) 
+at the joint %intersection 
+between Symbolic Automata (SA) and Weighted Automata (WA), 
+as well as Transducers (SWT) and Visibly Pushdown (SWPDA) variants.
+%
+Like SA, SWA deal with large or infinite input alphabets, 
+and like WA, they output a weight value in a semiring domain.
+The transitions of WA are labeled by functions from an infinite alphabet into the weight domain. 
+This is unlike WA whose transitions are guarded by boolean predicates 
+overs symbols in an infinite alphabet 
+and also unlike WA whose transitions are labeled by constant weight values, 
+and who deal only with finite automata.
+%
+We present some properties of WA, WT and SWPDA models, 
+and show how they can be used to define and solve a variant of parsing 
+over infinite alphabets.
+%
+We also briefly describe the application that motivated the introduction of these models: 
+an parse-based approach to automated music transcription.
+\end{abstract}
 
@@
 %of parsing of words over infinite input alphabet.
-
 Weighted parsing~\cite{Goodman99SemiringParsing,Nederhof03weightedParsing,MorbitzVogler19weighted-parsing} 
 consists in analyzing a word in input \wrt a weighted language model, \eg a probabilistic context-free grammar (CFG).
-In such a model, a weight value is associated to each transition rule, % production rule
+In such models, a weight value is associated to each transition rule, % production rule
 and the rule's weights can be combined with a associative product operator $\otimes$ into 
@@
 A second operator $\oplus$, associative and commutative, 
-is used to handle ambiguity of the model, 
+is moreover used to handle ambiguity of the model, 
 by summing the weights of the possibly several (in general exponentially many) 
@@
 % 
-This approach is close to the case of Symbolic Automata (SA)~\cite{dAntoni21CACM}, 
+This approach is close to the case of 
+Symbolic Automata (SA)~\cite{dAntoniVeanes17CAV,dAntoni21CACM}, 
 except that the domain for weight values is not restricted to be Boolean, 
@@
 In Section~\ref{section:SWA} we introduce \SWA and \SWT, 
-then \SWVPA are defined in Section~\ref{section:SWPDA}, 
+then \SWVPA are defined in Section~\ref{sec:SWVPA}, 
 where a polynomial 1-best algorithm is described that can be use to solve
@@
 We follow the approach of~\cite{Mohri03EDWA} for the computation of distances
-between words with weighted transducers, and propose models of weighted automata 
-and weighted transducers handling infinite alphabets.
-
-These models generalize weighted automata and transducers over finite alphabets, see  e.g.~\cite{Mohri03EDWA}, 
+between words and languages, with weighted transducers, 
+extending it to infinite alphabet with models of symbolic weighted automata 
+and transducers. % handling infinite alphabets.
+
+These models generalize weighted automata and transducers over finite alphabets, 
+see  e.g.~\cite{Mohri03EDWA}, 
 by labeling each transition with a weight functions that takes the 
 input and output symbols as parameters, instead of a simple weight value.
-These functions are similar to the guards of symbolic automata~\cite{Veanes13ciaa}, 
+These functions are similar to the guards of symbolic automata~\cite{dAntoniVeanes17CAV,dAntoni21CACM},
 but they can return values in an arbitrary semiring, 
@@
 \label{section:SWVPA}\label{sec:SWVPA}
-The following model generalizes Symbolic VPA~\cite{DAntonyAlur14SVPDA}
+The following model generalizes Symbolic VPA~\cite{dAntonyAlur14SVPDA}
 from Boolean semirings to arbitrary semiring weight domains.
@@
 The construction is essentially the same 
-as the one of~\cite{DAntonyAlur14SVPDA}, in the case of the Boolean semiring.
+as the one of~\cite{dAntonyAlur14SVPDA}, in the case of the Boolean semiring.
 

### 2021-05-05 main.tex
conflict
1 passe sec. 1 (debut)

@@
 %\title{Symbolic Weighted Parsing and Automated Music Transcription}
-\title{Symbolic Weighted Language Models and Automated Music Transcription}
-%\title{Weighted Visibly Pushdown Automata\\ Automated Music Transcription}
-%\author{Florent Jacquemard}
-%\institute{INRIA \& CNAM, Paris, France\\
-%\email{florent.jacquemard@inria.fr}}
+%\title{Symbolic Weighted Language Models and Automated Music Transcription}
+\title{Symbolic Weighted Language Models and Parsing over Infinite Alphabets}
+%\title{Weighted Visibly Pushdown Automata and Automated Music Transcription}
+\author{Florent Jacquemard}
+\institute{INRIA \& CNAM, Paris, France\\
+\email{florent.jacquemard@inria.fr}}
 
@@
 We present some properties of WA, WT and SWPDA models, 
-and show how they can be used to define and solve a variant of parsing 
+%and show how they can be used 
+that we use to define and solve a variant of parsing 
 over infinite alphabets.
@@
 
-The various approaches to weighted parsing usually assume a finite input alphabet. %models and algorithms 
+The approaches to weighted parsing usually assume a finite input alphabet. %models and algorithms 
 Considering infinite or large alphabets can however be of practical interest, 
@@
 between words and languages, with weighted transducers, 
-extending it to infinite alphabet with models of symbolic weighted automata 
-and transducers. % handling infinite alphabets.
-
-These models generalize weighted automata and transducers over finite alphabets, 
-see  e.g.~\cite{Mohri03EDWA}, 
-by labeling each transition with a weight functions that takes the 
+and extending it to infinite alphabets.
+% with models of symbolic weighted automata and transducers. 
+%
+The models introduced in this section generalize 
+weighted automata and transducers~\cite{Droste09handbook} 
+%over finite alphabets, see  e.g.~\cite{Mohri03EDWA}, 
+by labeling each transition with a weight function that takes the 
 input and output symbols as parameters, instead of a simple weight value.
@@
 \label{section:semiring}
-We shall consider semiring domains for weight values.
+We shall consider semiring domains for the weight values of our language models.
 %
@@
 %$\< \mathbb{S}, \otimes, \one>$ is a monoid
-$\oplus$ is commutative, 
-$\otimes$ distributes over~$\oplus$:  $\forall x, y, z \in \mathbb{S}$,
+\begin{itemize}
+\item $\oplus$ is commutative 
+-- therefore $\< \Semiring, \oplus, \zero>$ is a commutative monoid 
+   and $\< \Semiring, \otimes, \one>$ is a monoid,
+\item $\otimes$ distributes over~$\oplus$:  $\forall x, y, z \in \mathbb{S}$,
 $x \otimes (y \oplus z) = (x \otimes y) \oplus (x \otimes z)$, 
 %  and $(x \oplus y) \otimes z = (x \otimes z) \oplus (y \otimes z)$;
-and $\zero$ is absorbing for~$\otimes$: 
+\item $\zero$ is absorbing for~$\otimes$: 
 $\forall x\in \mathbb{S}$, $\zero \otimes x = x \otimes \zero = \zero$.
+\end{itemize}
 %Components of a semiring~$\Semiring$ may be subscripted by~$\Semiring$ when needed.
@@
 \label{section:symbols}
-
-Let $\Sigma$ and $\Delta$ be respectively an input and output \emph{alphabets}, 
-which are countable (finite or infinite) sets of symbols, 
+ We shall define the functions labeling the transitions of Sw automata and transducers.
+Let $\Sigma$ and $\Delta$ be two countable sets of symbols
+called respectively input and output \emph{alphabets}, 
 and let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative semiring.
@@
 and $\Phi_{\Sigma, \Delta}$  containing binary functions in $\Sigma \times \Delta \to \Semiring$.
-Moreover, we assume that these sets are closed under the operators 
+Moreover, these sets are required to be closed under the operators 
 $\oplus$ and $\otimes$ of~$\Semiring$.
@@
 and  $f_b: x \mapsto f(x, b)$ for $b \in \Delta$ and $x \in \Sigma$, 
-belong resp. to $\Phi_\Sigma$ and $\Phi_\Delta$.
+belong resp. to~$\Phi_\Sigma$ and~$\Phi_\Delta$.
 
@@
 Recall that by convention, an empty sum with $\oplus$ is $\zero$.
-%
-\noindent
+
+The above expression of $\weight_T$ can be seen as a stateful definition of 
+an edit-distance between a word $s \in \Sigma^*$ and a word $t \in \Delta^*$,
+see also~\cite{Mohri03ijfcs}.
+Intuitively, 
+$\wei(q, a, \epsilon, q'')$, in the second line, is the cost of 
+the deletion of a symbol $a \in \Sigma$ in $s$, 
+$\wei(q, \epsilon, b, q'')$, in the third line, is the cost 
+of the deletion of insertion of  $b \in \Delta$ in $s$, 
+$\wei(q, a, b, q'')$, in the last line, is the cost 
+of the substitution of  $a \in \Sigma$ by $b \in \Delta$.
+The cost of a sequence of such operations transforming $s$ into $t$ is 
+the product with $\otimes$ of the costs of individual operations involved,
+and the distance between $s$ and $t$ is the sum with $\oplus$
+of all such cost products.
+
+\medskip\noindent
 The weight associated by $T$ to  $\< s, t> \in \Sigma^* \times \Delta^*$
@@
 \subsection{Best-first Search}
-Assume that the semiring $\Semiring$ is 
-commutative, idempotent, superior, total and complete.
+Assume a semiring $\Semiring$
+commutative, complete, idempotent, superior, and total.
 % total ? 
@@
 
-\section{Application}
+\section{Symbolic Weighted Parsing}
+
+\subsection{Application to Amtomated Music Transcription}
 \label{sec:transcription}
@@
 
-\subsection{Time Scales}
+\subsubsection{Time Scales}
 Real-Time Unit (RTU) = seconds
@@
 
-\subsection{Representation of Music Performances}
+\subsubsection{Representation of Music Performances}
 We consider symbolic representations of musical performances, as finite sequences of events.
@@
 
-\subsection{Representation of Music Scores}
+\subsubsection{Representation of Music Scores}
 Music score are represented as structured words
@@
 
-\subsection{Performance/Score Distance Computation}
+\subsubsection{Performance/Score Distance Computation}
 \label{app:distance}

### 2021-05-05 main.tex
1 passe def. transducers

@@
 \begin{definition}
-A \emph{symbolic-weighted transducer} (\SWT) 
-$T$ over the input and output alphabet $\Sigma$ and $\Delta$ 
+A \emph{symbolic-weighted transducer} (\SWT)~$T$ 
+over the input and output alphabets~$\Sigma$ and $\Delta$ 
 and the semiring $\Semiring$ is a tuple
-$T = \< Q, \init, \wei, \final >$,
+$T = \< Q, \init, \vec{\wei}, \final >$,
 where $Q$ is a finite set of states, 
@@
 respectively leaving, a state, 
-and $\wei$ is a transition %weight 
-function from $Q \times Q$ into %$\Phi$.
-$\< \Phi_\epsilon, \Phi_\Sigma, \Phi_\Delta, \Phi_{\Sigma, \Delta}>$.
+and $\vec{\wei}$ is a tuplet made of the 4 of transition functions 
+$\wei_\epsilon$, $\wei_\Sigma$, $\wei_\Delta$, and $\wei_{\Sigma, \Delta}$
+from $Q \times Q$ into respectively
+$\Phi_\epsilon$, $\Phi_\Sigma$, $\Phi_\Delta$, and $\Phi_{\Sigma, \Delta}$.
 \end{definition}
 %
-We extend the above transition function into a function from
+We synthesize the above 4-uplet of transition functions into a unique 
+function~$\wei$ from
 $Q \times (\Sigma \cup \{ \epsilon \}) \times (\Delta \cup \{ \epsilon \}) \times Q$
-into $\Semiring$, also called $\wei$ for simplicity, such that 
-for all $q, q' \in Q$, $a \in \Sigma$, $b \in \Delta$, 
-and with 
-$\< \phi_\epsilon, \phi_\Sigma, \phi_\Delta, \phi_{\Sigma, \Delta}> = \wei(q, q')$, 
+into $\Semiring$, %also called $\wei$ for simplicity, 
+such that, for all $q, q' \in Q$, $a \in \Sigma$, and $b \in \Delta$, 
+%and with 
+%$\< \phi_\epsilon, \phi_\Sigma, \phi_\Delta, \phi_{\Sigma, \Delta}> = \wei(q, q')$, 
 \[
-\begin{array}{rcl}
-\wei(q, \epsilon, \epsilon, q') & = & \phi_\epsilon\\
-\wei(q, a, \epsilon, q') & = & \phi_\Sigma(a)\\
-\wei(q, \epsilon, b, q') & = & \phi_\Delta(b)\\
-\wei(q, a, b, q') & = & \phi_{\Sigma, \Delta}(a, b)
+\begin{array}{rcll}
+\wei(q, \epsilon, \epsilon, q') & = &  \wei_\epsilon(q, q'),\\ %\phi_\epsilon\\
+\wei(q, a, \epsilon, q') & = & \phi_\Sigma(a) & %\wei_\Sigma(q, q')(a)
+\quad\mathrm{where~} \phi_\Sigma = \wei_\Sigma(q, q'),\\
+\wei(q, \epsilon, b, q') & = & \phi_\Delta(b) & 
+\quad\mathrm{where~} \phi_\Delta = \wei_\Delta(q, q'),\\
+\wei(q, a, b, q') & = & \phi_{\Sigma, \Delta}(a, b) & 
+\quad\mathrm{where~} \phi_{\Sigma, \Delta} = \wei_{\Sigma, \Delta}(q, q').\\
 \end{array}      
@@
 %into resp. $\Phi_0$, $\Phi_\Sigma$, $\Phi_\Delta$ $\Phi_{\Sigma, \Delta}$.
-
-These functions $\phi$ act as guards for the transducer's transitions, 
-preventing a transition when they return the absorbing $\zero$ of $\Semiring$.
 
@@
                                          \end{array}} &
-      \wei(q, a, b, q'') \otimes \weight_T(q'', u, v, q')\\ 
+      \wei(q, a, b, q'') \otimes \weight_T(q'', u, v, q').\\ 
 \end{array}
 \]
-Recall that by convention, an empty sum with $\oplus$ is $\zero$.
+We recall that, by convention, an empty sum with $\oplus$ is $\zero$. 
+%
+Since $\zero$ is absorbing for~$\otimes$ in~$\Semiring$,
+one term $\wei(q, a, b, q'')$ equal to $\zero$ in the above expression 
+will be ignored in the sum, meaning that there is no possible transition
+from state $q$ into state $q'$ while reading $a$ and writing $b$ 
+(including the case $a = \epsilon$ and $a = \epsilon$).
+%These functions $\phi$ act as guards for the transducer's transitions, 
+%preventing a transition when they return the absorbing $\zero$ of $\Semiring$.
+This is analogous to a transition's guard not satisfied by $\<a, b>$ in 
+the case of symbolic transducers.
 
@@
 of the substitution of  $a \in \Sigma$ by $b \in \Delta$.
-The cost of a sequence of such operations transforming $s$ into $t$ is 
-the product with $\otimes$ of the costs of individual operations involved,
-and the distance between $s$ and $t$ is the sum with $\oplus$
-of all such cost products.
+The cost of a sequence of such operations transforming $s$ into $t$, 
+is the product, with $\otimes$, of the individual costs of the operations involved;
+And the distance between $s$ and $t$ is the sum, with $\oplus$,
+of all such product of costs.
 

### 2021-05-05 main.tex
rev. def automata

@@
 %
-% In the application presented in this paper, intuitively,
-% $\oplus$ selects an optimal value amongst two values and 
-% $\otimes$ combines two values into a single value.
+Intuitively, in the models presented in this paper, 
+$\oplus$ selects an optimal value from two given values, 
+in order to handle non-determinism, 
+and $\otimes$ combines two values into a single value, 
+in order to describe the chaining of several transitions.
 %and let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a {semiring},
@@
 \marginpar{necessary?}
-Finally, it is assumed that the codomain of every function of $\Phi_\Sigma$ and $\Phi_\Delta$
-is a subset of $\Phi_\epsilon$.
-and all partial applications of functions $\Phi_{\Sigma, \Delta}$, 
+Finally, it is also required that the codomain of every function of $\Phi_\Sigma$ and $\Phi_\Delta$
+is a subset of $\Phi_\epsilon$,
+and that all partial applications of functions $\Phi_{\Sigma, \Delta}$, 
 resp.  $f_a: y \mapsto f(a, y)$ for $a \in \Sigma$ and $y \in \Delta$
@@
 \begin{definition}
-A \emph{symbolic-weighted transducer} (\SWT)~$T$ 
+\label{def:transducer} \label{def:SWT}
+A \emph{symbolic-weighted transducer} (\SWT)
 over the input and output alphabets~$\Sigma$ and $\Delta$ 
@@
 \noindent
-A \emph{symbolic weighted automata} (\SWA) $A = \< Q, \init, \weight, \final >$
+The \emph{Symbolic Weighted Automata} %$A = \< Q, \init, \weight, \final >$
 over $\Sigma$ and $\Semiring$ 
-is defined in a similar way by simply omitting the output symbols,
-\ie $\wei$ is a function from~$Q \times Q$ into %$\Phi$.
-$\< \Phi_\epsilon, \Phi_\Sigma >$, 
-or equivalently from~$Q \times (\Sigma \cup \{ \epsilon \}) \times Q$ into~$\Semiring$.
+are defined similarly as the transducers of Definition~\ref{def:SWT}, 
+by simply omitting the output symbols.
+%
+\begin{definition}
+A \emph{symbolic-weighted automaton} (\SWA)
+over the input alphabet~$\Sigma$
+and the semiring $\Semiring$ is a tuple
+$A = \< Q, \init, \vec{\wei}, \final >$,
+where $Q$ is a finite set of states, 
+$\mathsf{in} : Q \to \Semiring$, 
+respectively $\mathsf{out} : Q \to \Semiring,$
+are functions defining the weight for entering, 
+respectively leaving, a state, 
+and $\vec{\wei}$ is a tuplet made of the 2 of transition functions 
+$\wei_\epsilon$, and $\wei_\Sigma$ from $Q \times Q$ into respectively
+$\Phi_\epsilon$, and $\Phi_\Sigma$.
+\end{definition}
+
+\noindent
+The 2 above transition functions are synthesized into 
+a function~$\wei$ 
+from~$Q \times (\Sigma \cup \{ \epsilon \}) \times Q$ into~$\Semiring$.
+%
+When $\wei_\epsilon(q, q') = \zero$ for all $q, q' \in Q$, 
+the automaton $A$ is called \emph{without $\epsilon$-transitions}.
 

### 2021-05-05 main.tex
suppr. epsilon

@@
 $\oplus$ and $\otimes$ of~$\Semiring$.
-More precisely, for all $\phi, \phi' \in \Phi_\Sigma$ all 
-$\psi, \psi' \in \Phi_\Delta$, 
+More precisely, for all 
+$\alpha, \alpha' \in \Phi_\epsilon$,
+all $\phi, \phi' \in \Phi_\Sigma$,
+all $\psi, \psi' \in \Phi_\Delta$, 
 and $\eta, \eta' \in \Phi_{\Sigma, \Delta}$, the function
 \begin{description}
+\item $\phi \otimes \alpha : x \mapsto \phi(x) \otimes \alpha$, 
+      and $\alpha \otimes \phi' : x \mapsto \alpha \otimes \phi'(x)$ belong to $\Phi_\Sigma$, 
 \item $\phi \otimes \phi' : x \mapsto \phi(x) \otimes \phi'(x)$ belongs to $\Phi_\Sigma$, 
@@
 and $\vec{\wei}$ is a tuplet made of the 2 of transition functions 
-$\wei_\epsilon$, and $\wei_\Sigma$ from $Q \times Q$ into respectively
-$\Phi_\epsilon$, and $\Phi_\Sigma$.
+$\wei_\epsilon$ and $\wei_\Sigma$ from $Q \times Q$ into respectively
+$\Phi_\epsilon$ and $\Phi_\Sigma$.
 \end{definition}
@@
 \noindent
-The 2 above transition functions are synthesized into 
-a function~$\wei$ 
+The above transition functions $\wei_\epsilon$ and $\wei_\Sigma$
+are synthesized into a function~$\wei$ 
 from~$Q \times (\Sigma \cup \{ \epsilon \}) \times Q$ into~$\Semiring$.
@@
 When $\wei_\epsilon(q, q') = \zero$ for all $q, q' \in Q$, 
-the automaton $A$ is called \emph{without $\epsilon$-transitions}.
+the automaton~$A$ is called \emph{without $\epsilon$-transitions}.
 
@@
 and a word $s \in \Sigma^*$, 
-one can construct a \SWA 
-$A_{s, T}$ such that for all $t \in \Delta^*$, 
-$A_{s, T}(t) = T(s, t)$.
+there exists a \SWA 
+$A_{T, s}$ such that for all $t \in \Delta^*$, 
+$A_{T, s}(t) = T(s, t)$.
 \end{proposition}
@@
 
+% gen. to transducers
 \begin{proposition}
-...removal of $\epsilon$ transitions for \SWA...
+Given a \SWA $A$ over $\Sigma$ and $\Semiring$, 
+there exists a \SWA $A'$ without $\epsilon$-transitions 
+such that for all $s \in \Sigma^*$, $A'(s) = A(s)$.
 \end{proposition}
+%
+\begin{proof}     
+Let $A = \< Q, \init, \<\wei_\epsilon, \wei_\Sigma>, \final >$.
+The automaton $A'$ is $\< Q, \init, \<\wei'_\epsilon, \wei'_\Sigma>, \final >$,
+where, for all $q, q' \in Q$,
+$\wei'_\epsilon(q, q') = \zero$  
+and 
+% \[
+% \begin{array}{rcccl}
+% \wei'_\Sigma(q, q') & = &  & \displaystyle\bigoplus_{q_1 \in Q} &
+%                \wei_\Sigma(q, q_1) \otimes  \wei_\epsilon(q_1, q')\\
+%                     &   & \oplus  & \displaystyle\bigoplus_{q_2 \in Q} &
+%                \wei_\epsilon(q, q_2) \otimes \wei_\Sigma(q_2, q')\\
+% \end{array}
+% \]
+\[
+\wei'_\Sigma(q, q') = 
+\displaystyle\bigoplus_{q_1 \in Q} \wei_\Sigma(q, q_1) \otimes \wei_\epsilon(q_1, q')
+\oplus  \displaystyle\bigoplus_{q_2 \in Q} \wei_\epsilon(q, q_2) \otimes \wei_\Sigma(q_2, q')
+\]
+In the above definition of $\wei'_\Sigma$ we use the operators 
+$\oplus$ and $\otimes$ on $\Phi_\Sigma$ described in Section~\ref{section:symbols}.
+\qed
+\end{proof}
 

### 2021-05-06 main.tex
revise section semiring

@@
 $x \otimes (y \oplus z) = (x \otimes y) \oplus (x \otimes z)$, 
-%  and $(x \oplus y) \otimes z = (x \otimes z) \oplus (y \otimes z)$;
+and $(x \oplus y) \otimes z = (x \otimes z) \oplus (y \otimes z)$,
 \item $\zero$ is absorbing for~$\otimes$: 
@@
 
-% \noindent
-A semiring $\Semiring$ 
-is \emph{monotonic} \wrt a partial ordering~$\leq$ 
-iff for all $x, y, z  \in \Semiring$,  $x \leq y$ implies
-$x \oplus z \leq y \oplus z$,
-$x \otimes z \leq y \otimes z$
-and $z \otimes x \leq z \otimes y$,
-%
-and it is \emph{superior} %\wrt a partial ordering~$\leq$
-\wrt~$\leq$ iff for all $x, y \in \Semiring$,  
-$x \leq x \otimes y$ and 
-$y \leq x \otimes y$~\cite{Huang08advanceddynamic}.
-The latter property corresponds to the 
-\emph{non-negative weights} condition in shortest-path algorithms~\cite{Dijkstra59anote}.
-Intuitively, it means that combining elements always increase their weight. 
-% always get worse in term of weight
-Note that when $\Semiring$ is superior \wrt~$\leq$, then $\one \leq \zero$
-and moreover, for all $x \in \Semiring$, $\one \leq x \leq \zero$.
-
-%\noindent
+\medskip%\noindent
 A semiring $\Semiring$ is \emph{commutative} if $\otimes$ is commutative.
@@
 %
+Every idempotent semiring~$\Semiring$ induces 
+a partial ordering~$\leq_\oplus$ 
+called the \emph{natural ordering} of~$\Semiring$~\cite{Mohri02semiring} 
+and defined,  by: 
+%implicitly defined by the semiring $\Semiring$ 
+for all $x$ and $y$,
+$x \leq_\oplus y \;\mbox{iff}\; x \oplus y = x$.
+%(see~\cite{Mohri02semiring} for the proof that it is an ordering).
+%
+The natural ordering is sometimes defined in the opposite direction~\cite{DrosteKuich09semirings};
+We follow here the direction  %follows \cite{Mohri02semiring}, and 
+that coincides than the usual ordering on the Tropical semiring (\emph{min-plus}).
+
+\begin{lemma}[Monotony, \cite{Mohri02semiring}]
+Let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be an idempotent semiring.
+For all $x, y, z  \in \Semiring$,  
+if $x \leq_\oplus y$ then
+$x \oplus z \leq_\oplus y \oplus z$,
+$x \otimes z \leq_\oplus y \otimes z$
+and $z \otimes x \leq_\oplus z \otimes y$.
+\end{lemma}      
+We say in this case that $\Semiring$ is \emph{monotonic}. % \wrt $\leq_\oplus$.
+%A semiring $\Semiring$ 
+%is \emph{monotonic} \wrt a partial ordering~$\leq$ 
+%iff for all $x, y, z  \in \Semiring$,  $x \leq y$ implies
+%$x \oplus z \leq y \oplus z$,
+%$x \otimes z \leq y \otimes z$,
+%and $z \otimes x \leq z \otimes y$.
+%
+Another important semiring property in the context of optimization
+is {superiority}~\cite{Huang08advanceddynamic}, 
+which corresponds to the 
+\emph{non-negative weights} condition in shortest-path algorithms~\cite{Dijkstra59anote}.
+Intuitively, it means that combining elements with $\otimes$ always increase their weight. 
+Formally, it is defined as the property of the following lemma.
+
+\begin{lemma}[Superiority, Boundedness]
+\label{lem:superior}\label{lem:bounded}
+Let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be an idempotent semiring.
+The two following statements are equivalent:
+\begin{itemize}
+\item [$i.$] for all $x, y \in \Semiring$,  
+$x \leq_\oplus x \otimes y$ and 
+$y \leq_\oplus x \otimes y$
+\item[$ii.$] for all $x \in \Semiring$,  $\one \oplus x = \one$.
+\end{itemize}
+\end{lemma}
+%
+In~\cite{Huang08advanceddynamic}, the property~$i.$ is called 
+$\Semiring$ is \emph{superior} \wrt the ordering~$\leq_\oplus$.
+It implies that 
+%It holds that If $\Semiring$ is superior \wrt~$\leq$, then %$\one \leq \zero$
+for all $x \in \Semiring$, $\one \leq_\oplus x \leq_\oplus \zero$.
+Indeed, by the second inequality of $(i.)$ with $y = \one$, $\one \leq_\oplus \one \otimes x = x$, 
+and, by the first inequality of $(i.)$ with $y = \zero$,  
+$x \leq_\oplus x \otimes \zero = \zero$.
+%
+Intuitively, from an optimization point of view,
+it means that $\one$ is the best value, and $\zero$ the worst.
+%** superior implies $\Semiring$ bounded~\cite{Mohri02semiring} see **
+%
+In~\cite{Mohri02semiring}, the property of Lemma~\ref{lem:bounded}(in the form $ii.$) 
+is called \emph{boundedness} of $\Semiring$ -- we shall use this term in the rest of the paper. 
+% \emph{negative boundedness} 
+It implies that, when looking for a best path in a graph whose edges
+are weighted by values of $\Semiring$, the loops can be safely avoided.
+
+\begin{proof}(Lemma~\ref{lem:bounded})
+For $(ii.) \Rightarrow (i.)$ : 
+$x \oplus (x \otimes y) = x \otimes (\one \oplus y) = x$, 
+by distributivity of $\otimes$ over $\oplus$. 
+Hence $x \leq_\oplus x \otimes y$.
+Similarly, $y \oplus (x \otimes y) = (\one \oplus x) \otimes y = y$, 
+hence $y \leq_\oplus x \otimes y$.
+%
+For $(i.) \Rightarrow (ii.)$ : as we have seen, $(i.)$ implies that 
+$\one \leq_\oplus x$ for all $x \in \Semiring$, 
+\ie, by definition of $\leq_\oplus$, $\one \oplus x = \one$.
+\qed
+\end{proof}
+
 %Following the terminology of~\cite{Mohri02semiring},
@@
 %the semiring $\Semiring$ is is called \emph{bounded}.
-%Note that every bounded semiring is idempotent:
-%by boundedness, 
-%$\one \oplus \one = \one$, and idempotency follows by multiplying
-%both sides by $x$ and distributing. 
-%
-Every idempotent semiring~$\Semiring$ induces 
-a partial ordering~$\leq_\oplus$ 
-called the \emph{natural ordering} of~$\Semiring$
-and defined by: 
-%implicitly defined by the semiring $\Semiring$ 
-for all $x$ and $y$,
-$x \leq_\oplus y \;\mbox{iff}\; x \oplus y = x$.
-This ordering is sometimes defined in the opposite direction~\cite{DrosteKuich09semirings};
-The above definition follows \cite{Mohri02semiring}, 
-and coincides than the usual ordering on the Tropical semiring (\emph{min-plus}).
-%
-It holds that $\Semiring$ is {monotonic} \wrt~$\leq_\oplus$.
+
+\begin{lemma}
+Every bounded semiring is idempotent.
+\end{lemma}
+\begin{proof}
+By boundedness, $\one \oplus \one = \one$, 
+and idempotency follows by multiplying
+both sides by $x$ and distributing. 
+\qed
+\end{proof}
+
+\noindent
 An idempotent semiring $\Semiring$~is called \emph{total} if
@@
 
-\medskip
-We shall consider below infinite sums with~$\oplus$.
+%\medskip
+We shall need below infinite sums with~$\oplus$.
 A semiring~$\Semiring$ is called \emph{complete} 
@@
 \begin{example}
+** classical examples **
 semirings      
@@
 A \emph{label theory} over $\Sigma$ and $\Delta$
-is made of 4 recursively enumerable sets:
-%$\Phi_0$, \Phi_\Sigma \uplus \Phi_\Delta \uplus \Phi_{\Sigma, \Delta}$ 
-$\Phi_\epsilon \subseteq \Semiring$, % containing constant functions valued in $\Semiring$, 
+is made of 3 recursively enumerable sets:
+%$\Phi_\epsilon \subseteq \Semiring$, % containing constant functions valued in $\Semiring$, 
 $\Phi_\Sigma$ and $\Phi_\Delta$, 
 containing unary functions in $\Sigma \to \Semiring$, resp. $\Delta \to \Semiring$, 
-and $\Phi_{\Sigma, \Delta}$  containing binary functions in $\Sigma \times \Delta \to \Semiring$.
-Moreover, these sets are required to be closed under the operators 
-$\oplus$ and $\otimes$ of~$\Semiring$.
-More precisely, for all 
-$\alpha, \alpha' \in \Phi_\epsilon$,
-all $\phi, \phi' \in \Phi_\Sigma$,
+and $\Phi_{\Sigma, \Delta}$  containing binary functions in $\Sigma \times \Delta \to \Semiring$, 
+such that 
+for all $\alpha \in \Semiring$, and $\phi \in \Phi_\Sigma$,
+$\alpha \otimes \phi : x \mapsto \alpha \otimes \phi(x)$, 
+and $\phi \otimes \alpha : x \mapsto \phi(x) \otimes \alpha$
+belong to $\Phi_\Sigma$, and similarly for $\oplus$ 
+and for $\Phi_\Delta$ and $\Phi_{\Sigma, \Delta}$.
+%
+Moreover, these sets are required to be closed under the following 
+operators~$\oplus$ and $\otimes$ of~$\Semiring$:
+for all $\phi, \phi' \in \Phi_\Sigma$,
 all $\psi, \psi' \in \Phi_\Delta$, 
@@
 \begin{description}
-\item $\phi \otimes \alpha : x \mapsto \phi(x) \otimes \alpha$, 
-      and $\alpha \otimes \phi' : x \mapsto \alpha \otimes \phi'(x)$ belong to $\Phi_\Sigma$, 
 \item $\phi \otimes \phi' : x \mapsto \phi(x) \otimes \phi'(x)$ belongs to $\Phi_\Sigma$, 
@@
 \end{description}
-The same also holds for the binary sum operator $\oplus$.
+And the same also holds for the binary sum operator $\oplus$.
 
@@
 \marginpar{necessary?}
-Finally, it is also required that the codomain of every function of $\Phi_\Sigma$ and $\Phi_\Delta$
-is a subset of $\Phi_\epsilon$,
-and that all partial applications of functions $\Phi_{\Sigma, \Delta}$, 
+Finally, it is also required 
+% that the codomain of every function of $\Phi_\Sigma$ and $\Phi_\Delta$ 
+% is a subset of $\Phi_\epsilon$, and
+that all partial applications of functions $\Phi_{\Sigma, \Delta}$, 
 resp.  $f_a: y \mapsto f(a, y)$ for $a \in \Sigma$ and $y \in \Delta$
@@
 $\wei_\epsilon$, $\wei_\Sigma$, $\wei_\Delta$, and $\wei_{\Sigma, \Delta}$
-from $Q \times Q$ into respectively
-$\Phi_\epsilon$, $\Phi_\Sigma$, $\Phi_\Delta$, and $\Phi_{\Sigma, \Delta}$.
+from $Q \times Q$ into respectively~$\Semiring$, %$\Phi_\epsilon$, 
+$\Phi_\Sigma$, $\Phi_\Delta$, and $\Phi_{\Sigma, \Delta}$.
 \end{definition}
@@
 \begin{array}{rccl}
-\weight_T(q, s, t, q') & = & & \wei(q, \epsilon, \epsilon, q')\\
- & & \oplus \displaystyle\bigoplus_{\begin{array}{c}
+\weight_T(q, s, t, q') & = & 
+   \displaystyle\bigoplus_{q'' \in Q} &
+   \wei(q, \epsilon, \epsilon, q'') \otimes \weight_T(q'', s, t, q')\\
+   & \oplus & \displaystyle\bigoplus_{q'' \in Q} &
+              \wei(q'', \epsilon, \epsilon, q') \otimes \weight_T(q, s, t, q'')\\
+ & \oplus & \displaystyle\bigoplus_{\begin{array}{c}
                                       \scriptstyle q'' \in Q\\[-2pt]
@@
    \wei(q, a, \epsilon, q'') \otimes \weight_T(q'', u, t, q')\\
- & & \oplus \displaystyle\bigoplus_{\begin{array}{c}
+ & \oplus & \displaystyle\bigoplus_{\begin{array}{c}
                                     \scriptstyle q'' \in Q\\[-2pt]
@@
     \wei(q, \epsilon, b, q'') \otimes \weight_T(q'', s, v, q')\\ 
-      & & \oplus \displaystyle\bigoplus_{\begin{array}{c}
+ & \oplus & \displaystyle\bigoplus_{\begin{array}{c}
                                     \scriptstyle q'' \in Q\\[-2pt]
@@
 Intuitively, 
-$\wei(q, a, \epsilon, q'')$, in the second line, is the cost of 
+$\wei(q, a, \epsilon, q'')$, in the third line, is the cost of 
 the deletion of a symbol $a \in \Sigma$ in $s$, 
-$\wei(q, \epsilon, b, q'')$, in the third line, is the cost 
+$\wei(q, \epsilon, b, q'')$, in the fourth line, is the cost 
 of the deletion of insertion of  $b \in \Delta$ in $s$, 
@@
 and $\vec{\wei}$ is a tuplet made of the 2 of transition functions 
-$\wei_\epsilon$ and $\wei_\Sigma$ from $Q \times Q$ into respectively
-$\Phi_\epsilon$ and $\Phi_\Sigma$.
+$\wei_\epsilon$ and $\wei_\Sigma$ from $Q \times Q$ into 
+respectively $\Semiring$ %$\Phi_\epsilon$ 
+and $\Phi_\Sigma$.
 \end{definition}
@@
 Given a \SWT $T$ 
-over $\Sigma$, $\Delta$ and $\Semiring$, 
-and a word $s \in \Sigma^*$, 
-there exists a \SWA 
+over alphabet $\Sigma$, $\Delta$ and a commutative semiring~$\Semiring$, 
+and $s \in \Sigma^*$, 
+there exists an effectively constructible \SWA 
 $A_{T, s}$ such that for all $t \in \Delta^*$, 
@@
 \end{proposition}
+%
+\begin{proof}
+Let $T = \< Q, \init, \vec{\wei}, \final >$ and let $n = |s|$, the length of $s$.
+
+\qed
+\end{proof}
+
 The construction time and size for $A_{s, T}$ are $O(| s | . \| T \|)$.
@@
 \begin{proposition}
-Given a \SWA $A$ over $\Sigma$ and $\Semiring$, 
-there exists a \SWA $A'$ without $\epsilon$-transitions 
+Given a \SWA $A$ over alphabet $\Sigma$ and a commutative and bounded semiring $\Semiring$, 
+there exists an effectively constructible \SWA $A'$ 
+without $\epsilon$-transitions 
 such that for all $s \in \Sigma^*$, $A'(s) = A(s)$.
@@
 %
-\begin{proof}     
+\begin{proof}  ** Tb revised **   
 Let $A = \< Q, \init, \<\wei_\epsilon, \wei_\Sigma>, \final >$.
@@
 \wei'_\Sigma(q, q') = 
-\displaystyle\bigoplus_{q_1 \in Q} \wei_\Sigma(q, q_1) \otimes \wei_\epsilon(q_1, q')
-\oplus  \displaystyle\bigoplus_{q_2 \in Q} \wei_\epsilon(q, q_2) \otimes \wei_\Sigma(q_2, q')
+\displaystyle\bigoplus_{q'' \in Q} \wei_\epsilon(q, q'') \otimes \wei_\Sigma(q'', q')
+\oplus \displaystyle\bigoplus_{q_1 \in Q} \wei_\Sigma(q, q'') \otimes \wei_\epsilon(q'', q')
 \]
-In the above definition of $\wei'_\Sigma$ we use the operators 
-$\oplus$ and $\otimes$ on $\Phi_\Sigma$ described in Section~\ref{section:symbols}.
+In the above definition of $\wei'_\Sigma$ we use the operator
+of product of function of $\Phi_\Sigma$ by $\Semiring$ 
+described in Section~\ref{section:symbols}.
+%
+By definition of $\weight_{A}$ and distributivity of $\otimes$ on $\oplus$, 
+** NO. TBC see \cite{Mohri02ijfcs} **
+it holds that $\weight_{A}(q, s, q') = \weight_{A'}(q, s, q')$.
 \qed

### 2021-05-06 main.tex
construction image transducer

@@
 \usepackage{refcount}
+
+% array and tabular
+\usepackage{array}
+\newcolumntype{C}[1]{>{\centering\arraybackslash\hspace{0pt}}p{#1}}
 
@@
 \begin{example}
-** classical examples **
-semirings      
+Figure~\ref{gif:semirings} presented examples of semirings interesting in practice 
+and enjoying the above properties.
 \end{example}
+
+\begin{figure}[t]
+\begin{center}
+\begin{tabular}{|c|c|C{2em}|C{2em}|C{2em}|C{2em}|}
+\hline
+        & domain & $\oplus$ & $\otimes$ & $\zero$  & $\one$\\ %& nat. ordering\\
+\hline\hline
+Boolean & $\{\bot, \top\}$ & $\vee$ & $\wedge$ & $\bot$ & $\top$\\ %& $\top \leq_\oplus \bot$  \\
+\hline
+Counting & $\mathbb{N}$ & $+$ & $\times$ & 0 & 1 \\
+\hline
+Viterbi & $[0, 1] \subset \mathbb{R}$ & $\mathit{max}$ & $\times$ & 0 & 1\\ % & $x \leq_\oplus y \iff x \ge y$  \\
+\hline
+Tropical & $\mathbb{R}_+ \cup \{ +\infty\}$ & $\mathit{min}$ & $+$ & $+\infty$ & $0$\\ % & $x \leq_\oplus y \iff x \leq y$   \\
+\hline
+%MaxPlus & $\mathbb{Q} \cup \{ -\infty\}$ & $\mathsf{max}$ & $+$ & $-\infty$ & $0$ \\
+%\hline
+%Word lang. & $2^{\Sigma^*}$ & $\cup$ & $\cdot$ & $\emptyset$ & $\{ \varepsilon \}$ \\
+%\hline
+\end{tabular}
+\end{center}
+\vskip-1em
+\caption{Some commutative, bounded, total and complete semirings.}
+\label{fig:semirings}
+\end{figure}
 
@@
 \label{section:symbols}
- We shall define the functions labeling the transitions of Sw automata and transducers.
+ We shall now define the functions labeling the transitions of SW-automata and -transducers.
 Let $\Sigma$ and $\Delta$ be two countable sets of symbols
@@
 are synthesized into a function~$\wei$ 
-from~$Q \times (\Sigma \cup \{ \epsilon \}) \times Q$ into~$\Semiring$.
+from~$Q \times (\Sigma \cup \{ \epsilon \}) \times Q$ into~$\Semiring$ 
+as above.
 %
@@
 \subsection{Properties}
+The two following properties will be useful to our approach on 
+symbolic weighted parsing in Section~\ref{sec:parsing}.
 
@@
 \begin{proof}
-Let $T = \< Q, \init, \vec{\wei}, \final >$ and let $n = |s|$, the length of $s$.
-
+Let $T = \< Q, \init, \vec{\wei}, \final >$ 
+where $\vec{\wei} \< \wei_\epsilon, \wei_\Sigma, \wei_\Delta, \wei_{\Sigma, \Delta}>$
+has an associated synthesized form $\wei$, 
+and let $s = s_1 \ldots s_n$ with $s_1, \ldots, s_n \in \Sigma$.
+%$n = |s|$ be the length of $s$.
+%
+Let $Q' = [1..n] \times Q$ be the state set of $A_{T, s}$,
+and let, for all $q \in Q$, 
+$\init'(1, q) = \init(q)$ and $\init'(i, q) = \zero$ for all $1 < i \leq n$, 
+and $\final'(n, q) = \final(q)$ and $\final'(i, q) = \zero$ for all $1 \leq i <  n$.
+
+\noindent
+We define the functions $\wei'_\epsilon$ and $\wei'_\Sigma$ for $A_{T, s}$
+in their synthesized form: for all $1 \leq i, j \leq n$, $q, q' \in Q$ and $a \in \Sigma$, 
+and $b \in \Delta$, 
+\[
+\begin{array}{lclcl}
+\wei'(\< i, q>, \epsilon, \< i, q'>)   & = & \wei(q, \epsilon, \epsilon, q'), & \quad & \\ 
+\wei'(\< i, q>, \epsilon, \< i+1, q'>) & = & \wei(q, s_i, \epsilon, q') & & \mathrm{if~} i < n,\\ 
+\wei'(\< i, q>, \epsilon, \< j, q'>)   & = & \zero & & \mathrm{if~} j \neq i, j \neq i+1,\\ 
+\wei'(\< i, q>, b, \< i, q'>)          & = & \wei(q, \epsilon, b, q'),\\ 
+\wei'(\< i, q>, b, \< i+1, q'>)        & = & \wei(q, s_i, b, q') & & \mathrm{if~} i < n,\\ 
+\wei'(\< i, q>, b, \< j, q'>)          & = & \zero & & \mathrm{if~} j \neq i, j \neq i+1. 
+\end{array}      
+\]
+Let $A_{T, s} = \< Q', \init', \vec{\wei'}, \final' >$.
 \qed
@@
 
-The construction time and size for $A_{s, T}$ are $O(| s | . \| T \|)$.
+The construction time and size for $A_{T, s}$ are $O(\| T \| . | s |)$,
+where the size $\| T \|$ of $T$ is its number of states $|Q|$.
 
@@
 \section{Symbolic Weighted Parsing}
+\label{sec:parsing}
 

### 2021-05-06 main.tex
details section VPA

@@
 \newcommand{\return}[1]{\ensuremath #1} %{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
+\def\Omegai{{\Omega_\mathsf{i}}}
+\def\Omegac{{\Omega_\mathsf{c}}}
+\def\Omegar{{\Omega_\mathsf{r}}}
 \def\Sigmai{{\Sigma_\mathsf{i}}}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% SWT & SWA
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{SW Automata and Transducers}
@@
 Let $\Sigma$ and $\Delta$ be two countable sets of symbols
-called respectively input and output \emph{alphabets}, 
+called respectively \emph{input} and \emph{output} \emph{alphabets}, 
 and let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative semiring.
@@
 \noindent
-\marginpar{necessary?}
+%\marginpar{necessary?}
 Finally, it is also required 
@@
 \]
+%** The application of $\oplus$ can be seen as a ND choice between transitions **
 We recall that, by convention, an empty sum with $\oplus$ is $\zero$. 
@@
 $\wei(q, a, \epsilon, q'')$, in the third line, is the cost of 
-the deletion of a symbol $a \in \Sigma$ in $s$, 
+the deletion of a symbol~$a \in \Sigma$ in~$s$, 
 $\wei(q, \epsilon, b, q'')$, in the fourth line, is the cost 
-of the deletion of insertion of  $b \in \Delta$ in $s$, 
-$\wei(q, a, b, q'')$, in the last line, is the cost 
-of the substitution of  $a \in \Sigma$ by $b \in \Delta$.
+of the deletion of insertion of~$b \in \Delta$ in~$s$, 
+and $\wei(q, a, b, q'')$, in the last line, is the cost 
+of the substitution of  $a \in \Sigma$ by~$b \in \Delta$.
 The cost of a sequence of such operations transforming $s$ into $t$, 
@@
 \begin{example}
-comparison of two sequences of timestamped events      
+DTW for sequences of timestamped events **     
+\end{example}
+
+\begin{example}
+(simpler) poinwise distance between two sequences of timestamped events **
 \end{example}
@@
 there exists an effectively constructible \SWA 
-$A_{T, s}$ such that for all $t \in \Delta^*$, 
-$A_{T, s}(t) = T(s, t)$.
+$A_{T, s}$ over $\Delta$ and $\Semiring$ 
+such that for all $t \in \Delta^*$, $A_{T, s}(t) = T(s, t)$.
 \end{proposition}
@@
 \]
-Let $A_{T, s} = \< Q', \init', \vec{\wei'}, \final' >$.
+Note that the functions defined above belong to the set $\Phi_\Delta$ 
+of the underlying label theory, 
+by the hypothesis of closure under partial application.
+%
+The $\SWA$ wanted is $A_{T, s} = \< Q', \init', \vec{\wei'}, \final' >$.
 \qed
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% SW-VPA
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{SW Visibly Pushdown Automata}
@@
 \subsection{Definition}
-Let $\Sigma$ be a countable alphabet 
+Let $\Omega$ be a countable alphabet 
 %finite (large) or infinite,
 that we assume partitioned into three subsets 
-$\Sigmai \uplus \Sigmac \uplus \Sigmac$ 
+$\Omegai$, $\Omegac$, $\Omegac$,
 respectively called of \emph{internal}, \emph{call} and \emph{return}
@@
 % \begin{itemize}
-% \item a set $\Sigmai$ of \emph{internal symbols} denoted $a$,
-% \item a set $\Sigmac$ of \emph{call symbols} denoted $\call{a}$,
-% \item a set $\Sigmar$ of \emph{return symbols} denoted $\return{a}$.
+% \item a set $\Omegai$ of \emph{internal symbols} denoted $a$,
+% \item a set $\Omegac$ of \emph{call symbols} denoted $\call{a}$,
+% \item a set $\Omegar$ of \emph{return symbols} denoted $\return{a}$.
 % \end{itemize}
@@
 where $\Phic$, $\Phir$ and~$\Phicr$ stand respectively 
-for~$\Phi_\Sigmac$, $\Phi_\Sigmar$ and~$\Phi_{\Sigmac, \Sigmar}$.
+for~$\Phi_\Omegac$, $\Phi_\Omegar$ and~$\Phi_{\Omegac, \Omegar}$.
 %
 Moreover, we extend this theory with a set $\Phii$ 
-of unary functions in $\Sigmai \to \Semiring$,
+of unary functions in $\Omegai \to \Semiring$,
 closed under $\oplus$ and $\otimes$.
@@
 A \emph{Symbolic Weighted Visibly Pushdown Automata} (\SWVPA)~$A$ 
-over  $\Sigma = \Sigmai \uplus \Sigmac \uplus \Sigmar$ and $\Semiring$ is a tuple
+over  $\Omega = \Omegai \uplus \Omegac \uplus \Omegar$ and $\Semiring$ is a tuple
 $T = \< Q, P, \init, \weii, \weic, \weir, \weie, \final >$,
@@
 for all $q, q' \in Q$, $p \in P$, 
-$a \in \Sigmai$, 
-$\call{c} \in \Sigmac$, 
-$\return{r} \in \Sigmar$, 
+$a \in \Omegai$, 
+$\call{c} \in \Omegac$, 
+$\return{r} \in \Omegar$, 
 overloading their names: % for simplicity:
@@
 \begin{array}{lll}
-\weii: Q \times \Sigmai \times Q \to \Semiring & 
+\weii: Q \times \Omegai \times Q \to \Semiring & 
 \weii(q, a, q') = \phi_\mathsf{i}(a) & 
@@
 %
-\weic: Q \times \Sigmac \times Q \times P \to \Semiring & 
+\weic: Q \times \Omegac \times Q \times P \to \Semiring & 
 \weic(q, \call{c}, q', p) = \phi_\mathsf{c}(\call{c}) & 
@@
 %
-\weir: Q \times \Sigmac \times P \times \Sigmar \times Q \to \Semiring & 
+\weir: Q \times \Omegac \times P \times \Sigmar \times Q \to \Semiring & 
 \weir(q, {\call{c}},  p, {\return{r}}, q') = \phi_\mathsf{r}({\call{c}},  {\return{r}}) & 
@@
 %
-\weie: Q \times \Sigmar \times Q \to \Semiring & 
+\weie: Q \times \Omegar \times Q \to \Semiring & 
 \weie(q, {\return{r}}, q') = \phi_\mathsf{e}({\return{r}}) &
@@
 In the case of a pushdown automaton, a configuration is composed 
-of a state $q \in Q$ and a stack content $\theta \in \Theta^*$, where $\Theta = \Sigmac \times P$.
+of a state $q \in Q$ and a stack content $\theta \in \Theta^*$, where $\Theta = \Omegac \times P$.
 Therefore, $\weight_A$ is a function from 
-$Q \times \Theta^* \times \Sigma^* \times Q \times \Theta^*$ into $\Semiring$.
+$Q \times \Theta^* \times \Omega^* \times Q \times \Theta^*$ into $\Semiring$.
 \[
@@
 where $\bot$ denotes the empty stack and ${\call{c}}\, p\cdot \theta$ 
-denotes a stack with the pair made of ${\call{c}}$ and $p$ on its top and $\theta$
-as the rest of stack.
-
-\noindent
-The weight associated by $A$ to $s \in \Sigma^*$
+denotes a stack where the pair made of ${\call{c}}$ and $p$ is the top symbol and $\theta$
+is the rest of stack.
+
+\noindent
+The weight associated by $A$ to $s \in \Omega^*$
 is then defined as follows,
@@
 Let $A_1$ and $A_2$ be two (\SWVPA)
-over the same $\Sigma$ and $\Semiring$.
+over the same $\Omega$ and $\Semiring$.
 There exists two $\SWVPA$ $A_1 \oplus A_2$ and $A_1 \otimes A_2$, 
 effectively constructible, 
-such that for all $s \in \Sigma^*$, 
+such that for all $s \in \Omega^*$, 
 $(A_1 \oplus A_2)(s) = A_1(s) \oplus A_2(s)$ and 
@@
 %Regarding the infinite sum operator, note that
-%$\bigoplus_{x \in \Phi_\Sigma} \phi(x)$, 
+%$\bigoplus_{x \in \Phi_\Omega} \phi(x)$, 
 %$\bigoplus_{y \in \Phi_\Delta} \psi(y)$, and  
@@
 \[ 
-  d_0(q, \sigma, q') = \bigoplus_{s\in \Sigma^*} 
+  d_0(q, \sigma, q') = \bigoplus_{s\in \Omega^*} 
   \textstyle
@@
 Since $\Semiring$ is complete, this infinite sum is well defined, 
-and since $\leq_\oplus$ is assumed total, it is the minimum in $\Sigma^*$
+and since $\leq_\oplus$ is assumed total, it is the minimum in $\Omega^*$
 of $s \mapsto \weight_A(\config{q}{\sigma}, s, \config{q'}{\sigma})$ \wrt this ordering.
@@
  & d(q_1, \sigma, q_3) & \opluseq &
-   d(q_1, \sigma, q_2) \otimes \displaystyle\bigoplus_{a \in \Sigmai} \weii(q_2, a, q_3)\\
+   d(q_1, \sigma, q_2) \otimes \displaystyle\bigoplus_{a \in \Omegai} \weii(q_2, a, q_3)\\
 %
 %  & d(q_0, \sigma, q_2) & \opluseq & 
-%    \displaystyle\bigoplus_{a \in \Sigmai} \weii(q_0, a, q_1) \otimes d(q_1, \sigma, q_2)\\
+%    \displaystyle\bigoplus_{a \in \Omegai} \weii(q_0, a, q_1) \otimes d(q_1, \sigma, q_2)\\
 %
@@
 %\displaystyle\bigoplus_{p \in P} 
-\displaystyle\bigoplus_{{\call{c}} \in \Sigmac} 
-\displaystyle\bigoplus_{{\return{r}} \in \Sigmar} \eta({\call{c}}, {\return{r}})\\
+\displaystyle\bigoplus_{{\call{c}} \in \Omegac} 
+\displaystyle\bigoplus_{{\return{r}} \in \Omegar} \eta({\call{c}}, {\return{r}})\\
 \mathrm{and} &  d(q_0, \bot, q_3) & \opluseq &
@@
 %\displaystyle\bigoplus_{p \in P} 
-\displaystyle\bigoplus_{{\call{c}} \in \Sigmac} 
-\displaystyle\bigoplus_{{\return{r}} \in \Sigmar} \eta({\call{c}}, {\return{r}})\\
+\displaystyle\bigoplus_{{\call{c}} \in \Omegac} 
+\displaystyle\bigoplus_{{\return{r}} \in \Omegar} \eta({\call{c}}, {\return{r}})\\
 & & & \mathrm{where~} \eta = \weic(q_0, q_1, p) \otimes \weir(q_2, p, q_3)\\
@@
 \mathrm{if~} \sigma = \bot & d(q_1, \bot, q_3) & \opluseq &
-  d(q_1, \sigma, q_2) \otimes \displaystyle\bigoplus_{\return{r} \in \Sigmar} \weie(q_2, {\return{r}}, q_3)\\
+  d(q_1, \sigma, q_2) \otimes \displaystyle\bigoplus_{\return{r} \in \Omegar} \weie(q_2, {\return{r}}, q_3)\\
 %
@@
 \[
-  {\displaystyle \bigoplus_{s \in \Sigma^*} A(s)} = 
+  {\displaystyle \bigoplus_{s \in \Omega^*} A(s)} = 
   {\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
@@
 \begin{definition}
-Let $\Sigma$ be an alphabet and $\Semiring$ a complete semiring.
-A function $\phi$ from $\Sigma^n$ into $\Semiring$
+Let $\Omega$ be an alphabet and $\Semiring$ a complete semiring.
+A function $\phi$ from $\Omega^n$ into $\Semiring$
 is called $k$-\emph{convex} for a natural number $k$ iff 
-$\mathit{card}\{ \mathbf{a} \in \Sigma^n \mid \phi(\mathbf{a}) = 
-                 \displaystyle\bigoplus_{\mathbf{p} \in \Sigma^n} \phi(\mathbf{p}) \} \leq k$.
+$\mathit{card}\{ \mathbf{a} \in \Omega^n \mid \phi(\mathbf{a}) = 
+                 \displaystyle\bigoplus_{\mathbf{p} \in \Omega^n} \phi(\mathbf{p}) \} \leq k$.
 \end{definition}
@@
 commutative, idempotent, superior, total and complete semiring 
-and an alphabet $\Sigma$ with a $k$-convex label theory, 
-one can construct in PTIME a word $s \in \Sigma^*$ 
+and an alphabet $\Omega$ with a $k$-convex label theory, 
+one can construct in PTIME a word $s \in \Omega^*$ 
 such that $A(s)$ is minimal \wrt the natural ordering for $\Semiring$. 

### 2021-05-06 main.tex
details section VPA

@@
 \section{SW Automata and Transducers}
-\label{section:transducer}
-\label{section:SWA}
-\label{section:SWT}
+\label{section:transducer}\label{sec:transducer}
+\label{section:SWA}\label{sec:SWA}
+\label{section:SWT}\label{sec:SWT}
 
@@
 in term of a weight value
-computed by an intermediate function $\weight_A$.
+computed by an intermediate function $\weight_A$, like in Section~\ref{sec:SWT}.
 In the case of a pushdown automaton, a configuration is composed 
-of a state $q \in Q$ and a stack content $\theta \in \Theta^*$, where $\Theta = \Omegac \times P$.
+of a state $q \in Q$ and a stack content $\gamma \in \Gamma^*$, where $\Gamma = \Omegac \times P$.
 Therefore, $\weight_A$ is a function from 
-$Q \times \Theta^* \times \Omega^* \times Q \times \Theta^*$ into $\Semiring$.
+$Q \times \Gamma^* \times \Omega^* \times Q \times \Gamma^*$ into $\Semiring$.
 \[
 \begin{array}{rcl}
-\weight_A\bigl(\config{q}{\theta}, a\, u, \config{q'}{\theta'}\bigr) & = & 
+\weight_A\bigl(\config{q}{\gamma}, a\, u, \config{q'}{\gamma'}\bigr) & = & 
  {\displaystyle\bigoplus_{q'' \in Q}} \weii(q, a, q'') 
- \otimes \weight_A\bigl(\config{q''}{\theta}, u, \config{q'}{\theta'}\bigr)\\
-%
-\weight_A\bigl(\config{q}{\theta}, {\call{c}}\, u, \config{q'}{\theta'}\bigr) & = & 
+ \otimes \weight_A\bigl(\config{q''}{\gamma}, u, \config{q'}{\gamma'}\bigr)\\
+%
+\weight_A\bigl(\config{q}{\gamma}, {\call{c}}\, u, \config{q'}{\gamma'}\bigr) & = & 
  {\displaystyle\bigoplus_{\begin{array}{c}
@@
  \weic\bigl(q, {\call{c}}, q'', p\bigr) 
- \otimes \weight_A\bigl(\config{q''}{{\call{c}}\, p\cdot \theta}, u, \config{q'}{\theta'}\bigr)\\[1mm]
-%
-\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot \theta}, {\return{r}}\, u, \config{q'}{\theta'}\bigr) & = & 
+ \otimes \weight_A\bigl(\config{q''}{{\call{c}}\, p\cdot \gamma}, u, \config{q'}{\gamma'}\bigr)\\[1mm]
+%
+\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot \gamma}, {\return{r}}\, u, \config{q'}{\gamma'}\bigr) & = & 
  {\displaystyle\bigoplus_{q'' \in Q}} 
  \weir\bigl(q, {\call{c}}, p, {\return{r}}, q''\bigr) 
- \otimes \weight_A\bigl(\config{q''}{\theta}, u, \config{q'}{\theta'}\bigr)\\
-%
-\weight_A\bigl(\config{q}{\bot}, {\return{r}}\, u, \config{q'}{\theta'}\bigr) & = & 
+ \otimes \weight_A\bigl(\config{q''}{\gamma}, u, \config{q'}{\gamma'}\bigr)\\
+%
+\weight_A\bigl(\config{q}{\bot}, {\return{r}}\, u, \config{q'}{\gamma'}\bigr) & = & 
  {\displaystyle\bigoplus_{q'' \in Q}} \weie(q, {\return{r}}, q'') 
- \otimes \weight_A\bigl(\config{q''}{\bot}, u, \config{q'}{\theta'}\bigr)\\
+ \otimes \weight_A\bigl(\config{q''}{\bot}, u, \config{q'}{\gamma'}\bigr)\\
 \end{array}
 \]
-where $\bot$ denotes the empty stack and ${\call{c}}\, p\cdot \theta$ 
-denotes a stack where the pair made of ${\call{c}}$ and $p$ is the top symbol and $\theta$
-is the rest of stack.
+where $\bot$ denotes the empty stack and ${\call{c}}\, p\cdot \gamma$ 
+denotes a stack where the pair made of ${\call{c}} \in \Omegac$ and $p \in P$ is the top symbol 
+and $\gamma$ is the rest of stack.
 
@@
 The construction is essentially the same 
-as the one of~\cite{dAntonyAlur14SVPDA}, in the case of the Boolean semiring.
+as in the case of the Boolean semiring~\cite{dAntonyAlur14SVPDA}.
 
@@
 More precisely, 
-let $\top$ be a fresh stack symbol which does not belong to $\Theta$,
+let $\top$ be a fresh stack symbol which does not belong to $\Gamma$,
 and for every two states $q, q' \in Q$ and $\sigma \in \{ \bot, \top \}$, 
@@
 When $\sigma = \bot$, intuitively, it is the minimum weight of a computation of $A$
-starting in state $q$ with a stack $\theta \in \Theta^*$ (possibly empty), 
-and ending in state $q$ with the same stack $\theta$, such that moreover 
-the computation does not touch a symbol of $\theta$. 
+starting in state $q$ with a stack $\gamma \in \Gamma^*$ (possibly empty), 
+and ending in state $q$ with the same stack $\gamma$, such that moreover 
+the computation does not touch a symbol of $\gamma$. 
 That means that during the computation, 
 $A$ may apply the first case of in the definition of $\weight_A$ (internal symbol),
-as well as the second case, to can push call symbols on the top of $\theta$, 
+as well as the second case, to can push call symbols on the top of $\gamma$, 
 and may pop these symbols with the third case (return symbol). 
 However, it cannot apply one of the two last cases (return symbol and empty stack)
-when the current stack is $\theta$.
-%pop symbols in $\theta$.
+when the current stack is $\gamma$.
+%pop symbols in $\gamma$.
 % Note that having a stack reduced to such a symbol makes impossible the application of the 

### 2021-05-07 main.tex
corresp. swTA - swVPA

@@
 %% PGF, Tikz
+\usepackage{tikz-cd}
 %% \usepackage{pgfplots}
@@
 % \usepackage{tikz}
+% \usetikzlibrary{cd}
 %% \usetikzlibrary{arrows}
@@
 \newcommand{\W}{\mathbb{W}}
+\newcommand{\T}{\mathcal{T}}
 
@@
 \def\SWA{\textsf{swA}\xspace}
+\def\SWTA{\textsf{swTA}\xspace}
 \def\SWVPA{\textsf{sw-VPA}\xspace}
@@
 \newcommand{\ioi}[1]{\mathsf{ioi}({#1})}
+\newcommand{\rank}{\mathsf{rk}}
+\newcommand{\lin}{\mathsf{lin}}
 
@@
 
+\begin{figure}
+\centering
+\begin{tikzpicture}
+\node (NFA) at (0,4) {%
+  \(
+  \begin{array}{c} 
+  \mathsf{NFA} : \Sigma_{\mathsf{fin}}^* \to \mathbb{B}\\
+  q \xrightarrow{a} q' \quad  a \in \Sigma_\mathsf{fin}
+  \end{array} 
+  \)
+};
+%
+\node (WA) at (-2,2) {%
+  \(
+  \begin{array}{c} 
+  \mathsf{WA} : \Sigma_{\mathsf{fin}}^* \to \mathbb{S}\\
+  q \xrightarrow{a, w} q' \quad  a \in \Sigma_\mathsf{fin}, w \in \mathbb{S}
+  \end{array} 
+  \)
+};
+%
+\node (SA) at (2,2) {%
+  \(
+  \begin{array}{c} 
+  \mathsf{SA} : \Sigma_{\mathsf{inf}}^* \to \mathbb{B}\\
+  q \xrightarrow{\phi} q' \quad \phi : \Sigma_\mathsf{inf} \to \mathbb{B}
+  \end{array} 
+  \)
+};
+%
+\node (SWA) at (0,0) {%
+  \(
+  \begin{array}{c} 
+  \mathsf{SWA} : \Sigma_{\mathsf{inf}}^* \to \mathbb{S}\\
+  q \xrightarrow{\phi} q' \quad \phi : \Sigma_\mathsf{inf} \to \mathbb{S}
+  \end{array} 
+  \)
+};
+\draw[->] (NFA)--(WA);
+\draw[->] (NFA)--(SA);
+\draw[->] (WA)--(SWA);
+\draw[->] (SA)--(SWA);
+%\begin{array}{c} \mathsf{NFA} : \Sigma^* \to \mathbb{B} \end{array} 
+\end{tikzpicture}
+\caption{Classes of Symbolic/Weighted Automata. 
+$\Sigma_\mathsf{fin}$ is a finite alphabet, 
+$\Sigma_\mathsf{inf}$ is a countable alphabet,
+$\mathbb{B}$ is the Boolean algebra, 
+$\mathbb{S}$ is an arbitrary commutative semiring, 
+$q \xrightarrow{\dots} q'$ represents the form of a transition between states $q$ and $q'$.}
+\label{fig:hierarchy}  
+\end{figure}
+
 % First one general edit-distance is defined by a weighted word 
@@
 
-In Section~\ref{section:SWA} we introduce \SWA and \SWT, 
-then \SWVPA are defined in Section~\ref{sec:SWVPA}, 
+In Section~\ref{section:SWA} we introduce \SWA and \SWT.
+Then \SWVPA are defined in Section~\ref{sec:SWVPA}, 
 where a polynomial 1-best algorithm is described that can be use to solve
-Symbolic Weighted Parsing.
-Finally, \label{sec:transcription}, we present an application 
+Symbolic Weighted Parsing, as described Section~\ref{sec:parsing}.
+Finally, in Section~\ref{sec:transcription}, we present an application 
 of this approach to automated music transcription that has been implemented.
@@
 
-\subsection{Definitions}
+\subsection{Definitions} \label{sec:SWTdef}\label{sec:SWAdef}
 %(SWT)
@@
 and the semiring $\Semiring$ is a tuple
-$T = \< Q, \init, \vec{\wei}, \final >$,
+$T = \< Q, \init, \bar{\wei}, \final >$,
 where $Q$ is a finite set of states, 
@@
 respectively leaving, a state, 
-and $\vec{\wei}$ is a tuplet made of the 4 of transition functions 
+and $\bar{\wei}$ is a tuplet made of the 4 of transition functions 
 $\wei_\epsilon$, $\wei_\Sigma$, $\wei_\Delta$, and $\wei_{\Sigma, \Delta}$
@@
 \begin{array}{rccl}
-\weight_T(q, s, t, q') & = & 
+xc(q, s, t, q') & = & 
    \displaystyle\bigoplus_{q'' \in Q} &
@@
 \]
+The sum above can be infinite because of the expression in the first line, 
+however, $\T(\Omega)$ is well defined since $\Semiring$ is complete.
+%
+As we shall see, epsilon-transitions can be removed and hence the sum made finite.
+%
 %** The application of $\oplus$ can be seen as a ND choice between transitions **
@@
 %
-\begin{definition}
+In this case, a label theory over $\Sigma$ is reduced to
+a set $\Phi_\Sigma$ closed under~$\oplus$ and~$\otimes$.
+%
+\begin{definition} \label{def:SWA}
 A \emph{symbolic-weighted automaton} (\SWA)
 over the input alphabet~$\Sigma$
-and the semiring $\Semiring$ is a tuple
-$A = \< Q, \init, \vec{\wei}, \final >$,
+and the commutative semiring $\Semiring$ is a tuple
+$A = \< Q, \init, \bar{\wei}, \final >$,
 where $Q$ is a finite set of states, 
@@
 respectively leaving, a state, 
-and $\vec{\wei}$ is a tuplet made of the 2 of transition functions 
+and $\bar{\wei}$ is a pair of transition functions 
 $\wei_\epsilon$ and $\wei_\Sigma$ from $Q \times Q$ into 
@@
 \begin{proof}
-Let $T = \< Q, \init, \vec{\wei}, \final >$ 
-where $\vec{\wei} \< \wei_\epsilon, \wei_\Sigma, \wei_\Delta, \wei_{\Sigma, \Delta}>$
+Let $T = \< Q, \init, \bar{\wei}, \final >$ 
+where $\bar{\wei} \< \wei_\epsilon, \wei_\Sigma, \wei_\Delta, \wei_{\Sigma, \Delta}>$
 has an associated synthesized form $\wei$, 
@@
 %
-The $\SWA$ wanted is $A_{T, s} = \< Q', \init', \vec{\wei'}, \final' >$.
+The $\SWA$ wanted is $A_{T, s} = \< Q', \init', \bar{\wei'}, \final' >$.
 \qed
@@
 \begin{proposition}
-Given a \SWA $A$ over alphabet $\Sigma$ and a commutative and bounded semiring $\Semiring$, 
+Given a \SWA $A$ over alphabet $\Sigma$ 
+and a commutative and bounded semiring $\Semiring$, 
 there exists an effectively constructible \SWA $A'$ 
@@
 $\Omegai$, $\Omegac$, $\Omegac$,
-respectively called of \emph{internal}, \emph{call} and \emph{return}
-symbols.
+whose elements are respectively called 
+\emph{internal}, \emph{call} and \emph{return} symbols.
 % \begin{itemize}
@@
 \begin{description}
-\item $\weii$ : read the input internal symbol $a$, change state to $q'$.
+\item $\weii$ : read the input internal symbol $a$, change state to $q'$ (stack is untouched).
 \item $\weic$ : read the input call symbol $\call{c}$, push it to the stack along with $p$, change state to $q'$.
@@
       In this case, the weight function $\phi_\mathsf{r}$ 
-      computes a value of matching between the call and return symbols.
-      This value might be $\zero$ in order to express that the symbols do not match.
+      computes a value of matching between the call and return symbols $c$ and $r$.
+      This value might be set to $\zero$ in order to express that the symbols do not match.
 \item $\weie$ : when the stack is empty, 
@@
 
-We give now a formal definition of these transitions of the automaton $A$
+We give now a formal definition of these transitions of the automaton~$A$
 in term of a weight value
@@
 \subsection{Properties}
-The class of \SWVPA is closed under the binary operators of the underlying semiring.
+Like VPA and symbolic VPA, 
+the class of \SWVPA is closed under the binary operators of the underlying semiring.
 %
@@
 \subsection{Best-first Search}
-Assume a semiring $\Semiring$
-commutative, complete, idempotent, superior, and total.
+Let us assume that the semiring $\Semiring$ is
+commutative, bounded, complete, and total.
 % total ? 
@@
 The algorithm~\ref{algo:Dijkstra}
-constructs iteratively a a marking $d : Q \times \{ \bot, \top \} \times Q \to \Semiring$
+constructs iteratively a marking $d : Q \times \{ \bot, \top \} \times Q \to \Semiring$
 %of the triplets $\<q, \sigma, q'>$ 
@@
 \label{sec:parsing}
-
-\subsection{Application to Amtomated Music Transcription}
+The models and results of the former sections allow us 
+to define formally and solve the problem of parsing over infinite alphabet.
+
+\subsection{Definition}
+
+\subsection{Results}
+
+\subsection{Application to Automated Music Transcription}
 \label{sec:transcription}
@@
 
+\section{Trees and Nested-Words}
+see also \cite{Caralp12VPAmult} for a discussion on the correspondence 
+between tree automata and VPA.
+
+Let $\Omega$ be a countable ranked alphabet, such that 
+every symbol $a \in \Omega$ has a rank 
+$\rank(a) \in [0..M]$ where $M$ is a fixed natural number.
+We write $\Omega_k$ the subset of all symbols $a$ of $\Omega$
+with $\rank(a) = k$, where $0 \leq k \leq M$.
+%$\cup_{i = 0}^{i = M} \Omega_i$ 
+%
+$\T(\Omega)$ denotes the free $\Omega$-algebra of finite, ordered, 
+$\Omega$-labeled trees, 
+which is the smallest set such that  $\Omega_0 \subset \T(\Omega)$
+and for all $1 \leq k \leq M$, all $a \in \Omega_k$, 
+and all $t_1, \ldots, t_k \in \T(\Omega)$, $a(t_1, \ldots, t_k) \in \T(\Omega)$.
+%
+% tree = single node (leave) labeled with a symbol of $a \in \Omegai$
+% (such a tree is simply denoted by $a$)
+% or the composition, denoted by $b(t_1,\ldots, t_n$) of a node labeled with $b$
+% and $n$ subtrees $t_1$,\ldots, $t_n$.
+
+\renewcommand{\call}[1]{\ensuremath \langle_{#1}}
+\renewcommand{\return}[1]{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
+
+Let $\hat\Omega$ be the countable (unranked) alphabet obtained from 
+$\Omega$ as follows: 
+$\hat\Omega = \< \Omegai, \Omegac, \Omegar >$, with
+$\Omegai = \Omega_0$, 
+$\Omegac = \{ \; \call{a} \mid a \in \Omega_{>0} \}$,
+$\Omegac = \{ \; \return{a} \mid a \in \Omega_{>0} \}$,
+where $\Omega_{>0}$ denotes $\bigcup_{k = 1}^{M} \Omega_k$.
+
+\noindent
+We define a linearization of trees of $\T(\Omega)$ into 
+words of $\hat{\Omega}^*$ as follows:
+\begin{description}
+\item $\lin(a) = a$ for all $a \in \Omega_0$, 
+\item $\lin\bigl( b(t_1, \ldots, t_k)\bigr) = 
+       \call{b} \; \lin(t_1) \ldots \lin(t_k) \; \return{b}$ 
+       when $b \in \Omega_k$, $1 \leq k \leq M$.
+\end{description}
+
+% A \emph{regular tree grammar} over $\Omega$ 
+% is a triplet $G = \< N, q_\mathsf{i}, R>$ where
+% $N$ is a finite set of non-terminal symbols denoted $q$..., 
+% $q_\mathsf{i} \in N$ is the starting non-terminal, 
+% $R$ is a finite set of production rules of the form
+% $q_0 \to a(q_1\ldots q_k)$ where 
+% $q_0, q_1, \ldots, q_k \in N$
+% $a \in \Omega_k$.
+% A tree $t \in \T(\Omega)$ is in the language of $G$ 
+% if it can be generated from $q_\mathsf{i}$ by 
+% non terminal replacement following the rules of $R$.
+
+Let us assume a label theory $\Phi_{\Omega_k}$ 
+for each $k \in [0..M]$.
+%
+\begin{definition}  \label{def:SWTA}
+A \emph{symbolic-weighted tree automaton} (\SWTA)
+over the ranked input alphabet~$\Omega$
+and the commutative semiring $\Semiring$ 
+is a triplet $A = \< Q, \init, \bar{\wei} >$ where
+$Q$ is a finite set of states, 
+$\mathsf{in} : Q \to \Semiring$ is the starting weight function, 
+and $\bar{\wei}$ is a $M+2$-uplet of transition functions made of:
+$\wei_\epsilon$ from $Q \times Q$ into $\Semiring$, and, 
+for each $k \in [0..M]$, $\wei_{\Omega, k}$
+from $Q \times Q^{k}$ into $\Phi_{\Omega_k}$.
+\end{definition}
+%
+Like in Section~\ref{sec:SWAdef}, we define a function~$\wei$ %from $\bar\wei$
+from 
+$Q \times (\Sigma \cup \{ \epsilon \}) \times \bigcup_{k=0}^{M} Q^k$
+into~$\Semiring$: %also called $\wei$ for simplicity, 
+%such that, for all $q, q' \in Q$, $a \in \Sigma$, and $b \in \Delta$, 
+\[
+\begin{array}{rcll}
+\wei(q_0, \epsilon, q_1) & = &  \wei_\epsilon(q_0, q_1),\\ %\phi_\epsilon\\
+\wei(q_0, a, q_1 \ldots q_k) & = & \phi_{\Omega, k}(a) &
+\quad\mathrm{where~} \phi_{\Omega, k} = \wei_{\Omega, k}(q_0, q_1\ldots q_k).\\
+\end{array}      
+\]
+Intuitively, $\wei(q_0, a, q_1 \ldots q_k)$ can be seen as
+the weight of a production $q_0 \to a(q_1, \ldots, q_k)$ 
+of a regular tree grammar~\cite{tata}, 
+that replaces the non-terminal symbol $q_0$, by $a(q_1, \ldots, q_k)$.
+
+\noindent
+The \SWTA of Definition~\ref{def:SWTA} defines a mapping 
+from trees of $\T(\Omega)$ into the weights of~$\Semiring$,
+based on the intermediate function $\weight_A$
+defined as follows for $q_0 \in Q$ and 
+$t = b(t_1,\ldots, t_k) \in \T(\Omega)$,
+åwith $0 \leq k \leq M$:
+\[
+\begin{array}{rccl}
+\weight_A(q_0, t) & = & 
+   \displaystyle\bigoplus_{q_1 \in Q} &
+   \wei(q, \epsilon, q_1) \otimes \weight_A(q_1, t)\\
+ & \oplus & \displaystyle\bigoplus_{q_1 \ldots q_k \in Q^k} &
+              \wei(q_0, b, q_1 \ldots q_k ) 
+   \otimes \displaystyle\bigotimes_{i=1}^{k}
+           \weight_A(q_i, t_i).\\
+\end{array}
+\]
+
+\medskip\noindent
+The weight associated by $A$ to  $t \in \T(\Omega)$ is then
+\begin{equation}
+A(t)  = 
+\displaystyle\bigoplus_{q \in Q} \mathsf{in}(q) \mathop{\otimes} \weight_A(q, t).
+\label{eq:weightTA}
+\end{equation}
+
+\begin{lemma}  \label{lem:SWTA}
+For all \SWTA $A$ over~$\Omega$ and $\Semiring$, 
+there exists an effectively constructible \SWVPA $A'$ 
+such that for all $t \in \T(\Omega)$, $A'\bigl(\lin(t)\bigr) = A(t)$.
+\end{lemma}  
+\begin{proof}
+\dots
+\qed\end{proof}
+
 \section{Edit-Distance}

### 2021-05-08 main.tex
swTA - swVPA

@@
 %Symbolic Weighted (SW) extension of symbolic automata...
-
-We propose language models called Symbolic Weighted Automata (SWA) 
+We propose a framework for weighted parsing over infinite alphabets.
+
+It is based on language models called Symbolic Weighted Automata (SWA) 
 at the joint %intersection 
@@
 %
-We synthesize the above 4-uplet of transition functions into a unique 
+We summerize the above 4-uplet of transition functions into a unique 
 function~$\wei$ from
@@
 \begin{definition}
-A \emph{Symbolic Weighted Visibly Pushdown Automata} (\SWVPA)~$A$ 
+A \emph{Symbolic Weighted Visibly Pushdown Automata} (\SWVPA) 
 over  $\Omega = \Omegai \uplus \Omegac \uplus \Omegar$ and $\Semiring$ is a tuple
-$T = \< Q, P, \init, \weii, \weic, \weir, \weie, \final >$,
+$A = \< Q, P, \init, \weii, \weic, \weir, \weie, \final >$,
 where $Q$ is a finite set of states, 
@@
 %
-\weir: Q \times \Omegac \times P \times \Sigmar \times Q \to \Semiring & 
+\weir: Q \times \Omegac \times P \times \Omegar \times Q \to \Semiring & 
 \weir(q, {\call{c}},  p, {\return{r}}, q') = \phi_\mathsf{r}({\call{c}},  {\return{r}}) & 
@@
 %
-Like in Section~\ref{sec:SWAdef}, we define a function~$\wei$ %from $\bar\wei$
+Like in Section~\ref{sec:SWAdef}, we define from $\bar{\wei}$ 
+a transition function~$\wei$, %from $\bar\wei$
 from 
@@
 Intuitively, $\wei(q_0, a, q_1 \ldots q_k)$ can be seen as
-the weight of a production $q_0 \to a(q_1, \ldots, q_k)$ 
+the weight of a production rule $q_0 \to a(q_1, \ldots, q_k)$ 
 of a regular tree grammar~\cite{tata}, 
 that replaces the non-terminal symbol $q_0$, by $a(q_1, \ldots, q_k)$. 
+%in a step of tree building.
+%
+Such a grammar computes the weights of the derivation trees 
+of the Context-Free grammar obtained by forgetting the labeling symbols of $\Omega_{>0}$.
 
@@
 \begin{lemma}  \label{lem:SWTA}
-For all \SWTA $A$ over~$\Omega$ and $\Semiring$, 
-there exists an effectively constructible \SWVPA $A'$ 
+For all \SWTA $A$ over~$\Omega$ and $\Semiring$, without $\epsilon$-transitions,
+there exists an effectively constructible \SWVPA $A'$ over 
+$\hat\Omega$ and $\Semiring$
 such that for all $t \in \T(\Omega)$, $A'\bigl(\lin(t)\bigr) = A(t)$.
 \end{lemma} 
+% 
 \begin{proof}
-\dots
+Let $A = \< Q, \init, \bar{\wei} >$ where $\bar{\wei}$ is summarized as above by a function
+$\wei: Q \times (\Sigma \cup \{ \epsilon \}) \times \bigcup_{k=0}^{M} Q^k \to \Semiring$.
+
+We build 
+$A' = \< Q', P', \init', \weii, \weic, \weir, \weie, \final' >$,
+computing over $\hat\Omega = \< \Omegai, \Omegac, \Omegar >$,
+%
+%$\Omega = \Omegai \uplus \Omegac \uplus \Omegar$ and $\Semiring$ 
+where $Q' = \bigcup_{k=0}^{M} Q^k$ be  the set of sequences of state symbols of $A$, 
+of length at most $M$, including the empty sequence denoted by $\epsilon$, 
+and where $P' = Q'$.
+
+\[
+\begin{array}{ll}
+%\weii: Q \times \Omegai \times Q \to \Semiring & 
+\weii(\bar{q}, a, \bar{q} q') = \wei(q', a, \epsilon) & 
+\mathrm{for~all~} a \in \Omega_0,\\
+%
+%\weic: Q \times \Omegac \times Q \times P \to \Semiring & 
+\weic(\bar{q}, \call{b}, \epsilon, \bar{q}) = \one & %\wei(p', b, \bar{p}) & 
+\mathrm{for~all~} b \in \Omega_{>0},\\
+%
+%\weir: Q \times \Omegac \times P \times \Sigmar \times Q \to \Semiring & 
+\weir(\bar{q}, {\call{b}}, \bar{p}, {\return{b}}, \bar{p} q') = \wei(q', b, \bar{q}) & 
+\mathrm{for~all~}  b \in \Omega_{>0},\\
+%
+%\weie: Q \times \Omegar \times Q \to \Semiring & 
+\weie(\bar{p}, {\return{b}}, \bar{q}) = \zero &
+\mathrm{for~all~}  b \in \Omega_{>0}.\\
+\end{array}      
+\]
+
+In practice, it is sufficient to consider in $Q'$ only the prefixes of 
+sequences
+
 \qed\end{proof}

### 2021-05-10 main.tex
SW parsing

@@
 We shall need below infinite sums with~$\oplus$.
-A semiring~$\Semiring$ is called \emph{complete} 
+A semiring~$\Semiring$ is called \emph{complete}~\cite{Kuich97semirings}
 if for every family
@@
 
-\section{Symbolic Weighted Parsing}
-\label{sec:parsing}
-The models and results of the former sections allow us 
-to define formally and solve the problem of parsing over infinite alphabet.
-
-\subsection{Definition}
-
-\subsection{Results}
-
-\subsection{Application to Automated Music Transcription}
-\label{sec:transcription}
-Symbolic Automated Music Transcription
-and analysis of music performances
-
-\subsubsection{Time Scales}
-Real-Time Unit (RTU) = seconds
-
-\noindent 
-Musical-Time Unit (MTU) = number of measures
-
-\noindent 
-conversion via tempo value
-
-\subsubsection{Representation of Music Performances}
-We consider symbolic representations of musical performances, as finite sequences of events.
-It corresponds to the concrete case of a MIDI file~\cite{SMF} 
-recorded  from an electronic keyboard, 
-or the output of a transcription from audio files~\cite{Benetos18AMTsurvey}.
-%
-For the sake of simplicity, 
-we shall only consider here the case of monophonic performances, 
-where at most one note is sounding at a time. 
-The approach however extends to the polyphonic case.
-
-A music performance is a finite sequence of events in a set~$\Sigma$.
-Every event $e \in \Sigma$ has attributes such from a finite domain, 
-like a number of key for a note 
-or a flag indicating that it is a rest 
-(\textsf{ON} and \textsf{OFF} messages in~\cite{SMF})
-and a velocity value (0..127 in~\cite{SMF})).
-%This representation is similar to the piano roll ~\cite{Muller15fundamentals} chap.1. 
-Moreover, it contains a RTU value $\ioi{e}$ (real number) 
-representing the time distance to the next event, 
-or to the end of performance for the last event,
-also called \emph{inter-onset interval}.
-
-\subsubsection{Representation of Music Scores}
-Music score are represented as structured words
-made of timed %quantified 
-events and parenthesized markups,
-akin of nested words~\cite{AlurMadhusudan09nested}.
-
-We consider an alphabet $\Delta$, every symbol of which is 
-composed of a tag, in a finite set $\Xi$, 
-and an MTU (rational) IOI duration value.
-%The alphabet $\Delta$ 
-It is partitioned into 
-$\Delta = \Deltai \uplus \Deltac \uplus \Deltar$, 
-like in Section~\ref{section:SWVPA}.
-%
-\noindent
-The symbols of $\Deltai$ represent events:
-% (infinite alphabet of internal symbols) made of:
-with tags indicating a new note or grace-note (with null IOI), 
-a rest or the continuation of the previous note (tie or dot).
-%
-The elements of $\Deltac \uplus \Deltar$ are matched
-markups for describing the structure of the score, 
-\ie the hierarchical grouping of events, and also, 
-importantly the division of time in measures, tuplets...
-%- parentheses for time divisions : tuplets, bars...
-(linearization of rhythm trees \cite{jacquemard:hal-01138642}...).
-They contain additional info such as tuple number, beaming policy...
-
-\noindent
-The duration values of letters of $\Delta$, in MTU (rational), 
-can be computed with the markups and tags (\eg grace note has duration 0).
-
-%\noindent
-%There are simultaneous events, since grace notes has duration 0. They are ordered.
-%
-%\noindent
-%Finite bound on the number of duration ratio. ?
-
-\begin{example}
-...      
-\end{example}
-
-\subsubsection{Performance/Score Distance Computation}
-\label{app:distance}
-We define a distance between performance and score representations
-by a $\SWT$ $T = \< Q, \init, \wei, \final >$, over a semiring $\Semiring$.
-** detail the elements of $\Semiring$ ....**
-%are quadruplets of the form
-%$\< t, s, \delta_t, \delta_s>$
-
-Every state of $Q$ contains a 
-tempo value in a finite domain (e.g. 30..300 bpm).
-This value can be fixed 
-or recomputed by the $T$ %transducer 
-after reading each event, 
-according to a perceptive/cognitive model of tempo 
-such as~\cite{LargeJones99tempo}
-(also used in the context of score following~\cite{Cont10TPAMI}).
-% we wont detail here.
-
-\subsection{Transcription by Best-first Search}
-We assume a score language defined by a \SWVPA over the semiring 
-$\Semiring$ of Section~\ref{app:distance}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% BIBLIO                                                                     %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\bibliographystyle{plain}
-%\bibliographystyle{plainurl} 
-\bibliographystyle{abbrv}
-%\bibliographystyle{splncs04}
-\bibliography{references}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% APPENDIX                                                                   %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\appendix
-
-\section{Trees and Nested-Words}
-see also \cite{Caralp12VPAmult} for a discussion on the correspondence 
-between tree automata and VPA.
+\subsection{Trees and Nested-Words}
+The hierarchical structure of nested-words, defined the \emph{call} and \emph{return} markup symbols of 
+suggest a correspondence between these word and trees. 
+The lifting of this correspondence to languages, respective of tree automata and VPA have
+been discussed in~\cite{AlurMadhusudan09nested}, 
+see also~\cite{Caralp12VPAmult} for the weighted case.
+In this section, we discuss the correspondence between the symbolic-weighted extensions
+of tree automata and VPA.
 
@@
 In practice, it is sufficient to consider in $Q'$ only the prefixes of 
-sequences
-
+sequences.
 \qed\end{proof}
+
+\section{Symbolic Weighted Parsing}
+\label{sec:parsing}
+Let us now use the models and results of the former sections in %order to define 
+an approach to the problem of parsing over infinite alphabet. %appropriate
+%
+Besides considering infinitely many possible of input symbols, 
+handled with suitable language formalisms, 
+this approach extends conventional parsing 
+by computing a derivation tree modulo 
+a generic distance between words,  
+defined by a SW transducer given in input.
+This enables considering finer word relationships than strict equality 
+as in the conventional parsing approach, 
+opening possibilities of quantitative analysis via this method.
+
+\subsection{Definition}
+
+Let $\Sigma$ be a countable input alphabet 
+and $\Omega = \Omegai \uplus \Omegac \uplus \Omegar$ an output countable alphabet,
+let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative, bounded, complete 
+and total semiring 
+and let  $(\Phi_\epsilon, \Phic, \Phir, \Phicr)$ be a label theory over $\Semiring$, 
+assumed computable and $k$-convex for some fixed $k$.
+%
+\noindent
+Assuming given in input:
+\begin{itemize}
+\item a \SWT $T$ over $\Sigma$,  $\Omega$, and $\Semiring$,
+defining a measure %between words 
+$T: \Sigma^* \times \Omega^* \to \Semiring$,
+\item a \SWVPA $A$ over $\Omega$, and $\Semiring$, defining a series of nested words
+      $A : \Omega^* \to \Semiring$,
+\item an input word $s \in \Sigma^*$,
+\end{itemize}
+the problem of \emph{symbolic weighted parsing} is to find 
+a nested word $t \in \Omega$ minimizing
+\( T(s, t) \otimes A(t)\)
+\wrt $\leq_\oplus$, 
+\ie such that %Hence, it is the problem of finding 
+\( T(s, t) \otimes A(t) = 
+  \displaystyle\bigoplus_{t' \in \Omega^*} T(s, t') \otimes A(t') \).
+
+\noindent 
+Therefore, it is the problem of optimizing a measure 
+called the \emph{edit-distance between $s$ and $A$} in~\cite{Mohri03EDWA}.
+%
+The input language can also be expressed as a \SWTA, or, 
+as a particular case, as a weighted context-free grammar, 
+converted in turn into a \SWVPA following Lemma~\ref{lem:SWTA}.
+%
+In the case of finite alphabets, 
+the problem of searching, in a WTA language, 
+for the best parse tree for a given,  
+sometimes referred as~\emph{weighted parsing}
+(see \cite{Goodman99semiring,MorbitzVogler19weighted-parsing} **more general problems**)
+is a particular case of SW parsing.
+%
+Indeed, it corresponds to the case where $T$
+accepts only the pairs $\<s, t>$ such that 
+$s$ is the projection of $t$ on $\Omegai$. 
+This can be done with a single state $q$ and 
+with transition rules of the form:
+\begin{description}
+\item[] $\wei(q, \epsilon, a, q) = \one$ for all $a \in \Omegac \cup \Omegar$,
+\item[] $\wei(q, a, a, q) = \one$ for all $a \in \Omegai$,
+\item[] $\wei(q, a, b, q) = \zero$ for all $a, b \in \Omegai$, $a \neq b$.
+\end{description}
+
+\subsection{Computation}
+
+\begin{proposition}
+The problem of Symbolic Weighted  parsing 
+can be solved in PTIME in the size of the input \SWT, \SWVPA (or \SWTA) and input word, 
+and the computation time of the functions of the label theory.
+\end{proposition}
+%
+\begin{proof}
+Bar-Hillel construction
+\qed  
+\end{proof}
+
+\subsection{Application to Automated Music Transcription}
+\label{sec:transcription}
+Symbolic Automated Music Transcription
+and analysis of music performances
+
+\subsubsection{Time Scales}
+Real-Time Unit (RTU) = seconds
+
+\noindent 
+Musical-Time Unit (MTU) = number of measures
+
+\noindent 
+conversion via tempo value
+
+\subsubsection{Representation of Music Performances}
+We consider symbolic representations of musical performances, as finite sequences of events.
+It corresponds to the concrete case of a MIDI file~\cite{SMF} 
+recorded  from an electronic keyboard, 
+or the output of a transcription from audio files~\cite{Benetos18AMTsurvey}.
+%
+For the sake of simplicity, 
+we shall only consider here the case of monophonic performances, 
+where at most one note is sounding at a time. 
+The approach however extends to the polyphonic case.
+
+A music performance is a finite sequence of events in a set~$\Sigma$.
+Every event $e \in \Sigma$ has attributes such from a finite domain, 
+like a number of key for a note 
+or a flag indicating that it is a rest 
+(\textsf{ON} and \textsf{OFF} messages in~\cite{SMF})
+and a velocity value (0..127 in~\cite{SMF})).
+%This representation is similar to the piano roll ~\cite{Muller15fundamentals} chap.1. 
+Moreover, it contains a RTU value $\ioi{e}$ (real number) 
+representing the time distance to the next event, 
+or to the end of performance for the last event,
+also called \emph{inter-onset interval}.
+
+\subsubsection{Representation of Music Scores}
+Music score are represented as structured words
+made of timed %quantified 
+events and parenthesized markups,
+akin of nested words~\cite{AlurMadhusudan09nested}.
+
+We consider an alphabet $\Delta$, every symbol of which is 
+composed of a tag, in a finite set $\Xi$, 
+and an MTU (rational) IOI duration value.
+%The alphabet $\Delta$ 
+It is partitioned into 
+$\Delta = \Deltai \uplus \Deltac \uplus \Deltar$, 
+like in Section~\ref{section:SWVPA}.
+%
+\noindent
+The symbols of $\Deltai$ represent events:
+% (infinite alphabet of internal symbols) made of:
+with tags indicating a new note or grace-note (with null IOI), 
+a rest or the continuation of the previous note (tie or dot).
+%
+The elements of $\Deltac \uplus \Deltar$ are matched
+markups for describing the structure of the score, 
+\ie the hierarchical grouping of events, and also, 
+importantly the division of time in measures, tuplets...
+%- parentheses for time divisions : tuplets, bars...
+(linearization of rhythm trees \cite{jacquemard:hal-01138642}...).
+They contain additional info such as tuple number, beaming policy...
+
+\noindent
+The duration values of letters of $\Delta$, in MTU (rational), 
+can be computed with the markups and tags (\eg grace note has duration 0).
+
+%\noindent
+%There are simultaneous events, since grace notes has duration 0. They are ordered.
+%
+%\noindent
+%Finite bound on the number of duration ratio. ?
+
+\begin{example}
+...      
+\end{example}
+
+\subsubsection{Performance/Score Distance Computation}
+\label{app:distance}
+We define a distance between performance and score representations
+by a $\SWT$ $T = \< Q, \init, \wei, \final >$, over a semiring $\Semiring$.
+** detail the elements of $\Semiring$ ....**
+%are quadruplets of the form
+%$\< t, s, \delta_t, \delta_s>$
+
+Every state of $Q$ contains a 
+tempo value in a finite domain (e.g. 30..300 bpm).
+This value can be fixed 
+or recomputed by the $T$ %transducer 
+after reading each event, 
+according to a perceptive/cognitive model of tempo 
+such as~\cite{LargeJones99tempo}
+(also used in the context of score following~\cite{Cont10TPAMI}).
+% we wont detail here.
+
+\subsection{Transcription by SW Parsing} %Best-first Search}
+We assume a score language defined by a \SWVPA over the semiring 
+$\Semiring$ of Section~\ref{app:distance}.
+
+\section*{Conclusion}
+% discussion
+
+\noindent
+-- summary
+
+\noindent
+-- other theoretical properties of SW models
+
+\noindent
+-- room to improve complexity for best-search algorithm
+... modular approach with oracles ...
+
+and extention to $n$-best
+
+\noindent
+-- offline algorithm, 
+semi-online implementation for AMT (bar-by-bar approach)
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% BIBLIO                                                                     %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\bibliographystyle{plain}
+%\bibliographystyle{plainurl} 
+\bibliographystyle{abbrv}
+%\bibliographystyle{splncs04}
+\bibliography{references}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% APPENDIX                                                                   %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newpage
+\appendix
 

### 2021-05-11 main.tex
passe intro

@@
 
-%% theorem envs
+%% theorem environments
 \usepackage{theorem}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% intro
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Introduction} \label{sec:intro}
+Parsing %variant of membership problem for formal languages
+%- given a language model and an input word $s$, compute a derivation of the model that yields $s$.
+is the problem %process 
+of structuring an input linear representation %typically
+(a finite word over an alphabet) according to a language model (a formal grammar).
+% natural language, programming language, 
+%
+Most of the context-free parsing approaches~\cite{GruneJacobs08parsing}  
+assume a finite and reasonably small input alphabet. %models and algorithms 
+Such a restriction makes perfectly sense in the context of NLP tasks 
+like constituency parsing or of programming languages compilers or interpreters.
+Considering large or infinite alphabets can however be of practical interest
+in other cases.
+%
+For instance, when dealing with large characters encodings such as UTF-16, % processing strings in 
+\eg for vulnerability detection in Web-applications~\cite{dAntoni21CACM}, 
+%
+for the analyse of data streams or serialization of structured documents~\cite{Segoufin06csl}, 
+or processing of timed execution traces.
+Regarding the latter case, we describe briefly at the end of the paper 
+a parse-based approach to automated music transcription,
+where a symbolic music performance (sequence of timed musical events) % timestamped
+% a symbolic representation of a music performance 
+is converted into a structured score in Common Western music notation~\cite{foscarin:hal-01988990}.
+%generalizes weighted parsing: 
+%finding the best derivation for a weighted grammar.
+
+Various extensions of automata have been studied for handling infinite alphabets.
+%words carrying data values in an infinite domain (e.g. integers) e.g. data processing 
+To cite a few, some memory extensions allow restricted storage and comparison of input symbols, 
+%and correspond with logics, 
+(see~\cite{Segoufin06csl} for a survey)
+like pebbles for marking positions~\cite{NevenSchwentickVianu04FSMinfinite}, 
+registers~\cite{KaminskiFrancez94}, 
+computations in 2 steps~\cite{Bojanczyk11FO2}... % data words automata.
+%
+%for the and verification of infinite state systems 
+%(model checkers: alphabet = long bit-vectors)
+%...For the representation of  in model checking, verification and 
+similar to representation of automata in model checkers, 
+following a similar idea, 
+Symbolic automata, 
+transitions guarded by predicates over alphabet theories.
+
+another quantitative issue in parsing is non-determinism : 
+with an ambiguous grammar, there may exists several derivations yielding one input word.
 %the problem: quantitative parsing or symbolic parsing
 %of parsing of words over infinite input alphabet.
+selection by ranking derivation according to an associated weight value...
 Weighted parsing~\cite{Goodman99SemiringParsing,Nederhof03weightedParsing,MorbitzVogler19weighted-parsing} 
-consists in analyzing a word in input \wrt a weighted language model, \eg a probabilistic context-free grammar (CFG).
+consists in analysing a word in input \wrt a weighted language model, \eg a probabilistic context-free grammar (CFG).
 In such models, a weight value is associated to each transition rule, % production rule
@@
 
-The approaches to weighted parsing usually assume a finite input alphabet. %models and algorithms 
-Considering infinite or large alphabets can however be of practical interest, 
-for instance in the context of processing strings in UTF16 encodings, 
-for vulnerability detection in Web-applications 
-(see \eg~\cite{dAntoni21CACM}), 
-for the representation of execution traces in model checking, 
-or when dealing with sequences of timestamped events
--- we describe briefly at the end of the paper a parsing based approach to 
-of automated music transcription, that is the conversion of a symbolic 
-representation of a music performance into a structured music score.
-%\cite{MCM19} 
-%generalizes weighted parsing: 
-%finding the best derivation for a weighted grammar.
-
 We present here some weighted language models
@@
 with a visibly restriction~\cite{AlurMadhusudan09nested} (\SWVPA).
-%wrt an edit distance
+%\wrt an edit distance
 The latter model of automata computes on \emph{nested words}~\cite{AlurMadhusudan09nested}, 
@@
 %
-We summerize the above 4-uplet of transition functions into a unique 
+We summarize the above 4-uplet of transition functions into a unique 
 function~$\wei$ from
@@
 \begin{example}
-(simpler) poinwise distance between two sequences of timestamped events **
+(simpler) pointwise distance between two sequences of timestamped events **
 \end{example}
@@
 assumed computable and $k$-convex for some fixed $k$.
-%
+
 \noindent
@@
 sometimes referred as~\emph{weighted parsing}
-(see \cite{Goodman99semiring,MorbitzVogler19weighted-parsing} **more general problems**)
+(see \cite{Goodman99SemiringParsing,MorbitzVogler19weighted-parsing} **more general problems**)
 is a particular case of SW parsing.

### 2021-05-12 main.tex
revise intro

@@
 is the problem %process 
-of structuring an input linear representation %typically
-(a finite word over an alphabet) according to a language model (a formal grammar).
+of structuring a linear representation in input %typically
+(a finite word over an alphabet) according to a language model. % (a formal grammar).
 % natural language, programming language, 
@@
 %
-for the analyse of data streams or serialization of structured documents~\cite{Segoufin06csl}, 
-or processing of timed execution traces.
-Regarding the latter case, we describe briefly at the end of the paper 
+for the analyse (\eg validation or filtering) 
+of data streams or serialization of structured documents 
+(which may contain textual or numerical attributes)~\cite{Segoufin06csl}, 
+or for processing timed execution traces~\cite{Bouyer03algebraic}.
+% algebraic definition of a class of data languages 
+% (notion of monoid recognizability, based on registers, comparable to Bojancszik et al. data words)
+Regarding the latter case, we describe %briefly 
+at the end of the paper 
 a parse-based approach to automated music transcription,
-where a symbolic music performance (sequence of timed musical events) % timestamped
+where a symbolic music performance, 
+presented as a sequence of timed musical events, % timestamped
 % a symbolic representation of a music performance 
-is converted into a structured score in Common Western music notation~\cite{foscarin:hal-01988990}.
+is converted into a structured score in Common Western Music Notation~\cite{foscarin:hal-01988990}.
 %generalizes weighted parsing: 
@@
 
-Various extensions of automata have been studied for handling infinite alphabets.
+Various extensions of language models for handling infinite alphabets have been studied.
 %words carrying data values in an infinite domain (e.g. integers) e.g. data processing 
-To cite a few, some memory extensions allow restricted storage and comparison of input symbols, 
+For instance, some automata with memory extensions 
+allow restricted storage and comparison of input symbols, 
 %and correspond with logics, 
-(see~\cite{Segoufin06csl} for a survey)
-like pebbles for marking positions~\cite{NevenSchwentickVianu04FSMinfinite}, 
+(see~\cite{Segoufin06csl} for a survey),
+with pebbles for marking positions~\cite{NevenSchwentickVianu04FSMinfinite}, 
 registers~\cite{KaminskiFrancez94}, 
-computations in 2 steps~\cite{Bojanczyk11FO2}... % data words automata.
+or %computations in 2 steps
+the possibility to compute on subsequences 
+with the same attribute values~\cite{Bojanczyk11FO2}. % data words automata.
 %
@@
 %...For the representation of  in model checking, verification and 
-similar to representation of automata in model checkers, 
-following a similar idea, 
-Symbolic automata, 
-transitions guarded by predicates over alphabet theories.
-
-another quantitative issue in parsing is non-determinism : 
-with an ambiguous grammar, there may exists several derivations yielding one input word.
-%the problem: quantitative parsing or symbolic parsing
-%of parsing of words over infinite input alphabet.
-selection by ranking derivation according to an associated weight value...
-Weighted parsing~\cite{Goodman99SemiringParsing,Nederhof03weightedParsing,MorbitzVogler19weighted-parsing} 
-consists in analysing a word in input \wrt a weighted language model, \eg a probabilistic context-free grammar (CFG).
-In such models, a weight value is associated to each transition rule, % production rule
-and the rule's weights can be combined with a associative product operator $\otimes$ into 
-the weight of a derivation (or \emph{abstract syntax tree}, \ie the result of an analyze).
-A second operator $\oplus$, associative and commutative, 
+The automata at the core of model checkers
+compute on input symbols represented by large bitvectors~\cite{Vardi07ciaa} %\cite{BaierKatoen08MC}
+(sets of assignments of Boolean variables) %propositional variables)
+and in practice,  %implementation
+each transition accepts a set of such symbols (instead of an individual symbol), 
+represented by Boolean formula or Binary Decision Diagrams.
+%
+Following a similar idea, % and generalizing,
+in Symbolic automata (\textsf{SA})~\cite{dAntoniVeanes17CAV,dAntoni21CACM}, 
+the transitions are guarded by predicates over infinite alphabet domains.
+With closure conditions on the sets of such predicates, % (alphabet theories), 
+all the good properties enjoyed by automata over finite alphabets are preserved.
+
+Other extensions of language models  %(automata and grammars) 
+help in dealing with non-determinism, by the computation of weight values. 
+With an ambiguous grammar, there may exists several derivations 
+(\emph{abstract syntax trees} -- AST), % \ie the result of an analyze.
+yielding one input word. % structuring a word, 
+The association of one weight value %in semiring domain
+to each derivation permits to select a best one (or $n$ bests). % a fixed number of bests 
+% = ...a ranking of derivations 
+This is roughly the principle of \emph{weighted parsing}
+approaches~\cite{Goodman99SemiringParsing,Nederhof03weightedParsing,MorbitzVogler19weighted-parsing}.
+In \emph{weighted language models}, 
+like \eg probabilistic context-free grammars, % (CFG),
+a weight value is associated to each transition rule, % production rule
+and the rule's weights can be combined with a associative product operator~$\otimes$ into 
+the weight of a derivation.
+A second operator~$\oplus$, associative and commutative, 
 is moreover used to handle ambiguity of the model, 
-by summing the weights of the possibly several (in general exponentially many) 
-syntax trees associated to a given input word.
-Intuitively, $\oplus$ aims at selecting, or ranking, the syntax trees.
+by summing the weights of the possibly several (in general exponentially many) AST %syntax trees 
+associated to a given input word.
+Typically, $\oplus$ can select the minimum of two weights.
+%Intuitively,~$\oplus$ selects, or ranking, the syntax trees.
 The weight domain, equipped with these two operators is assumed, at minima, 
-to form a \emph{semiring} $\Semiring$, %\cite{Eilenberg74automata}
-such as a min-plus algebra of domain $\mathbb{R} \cup \{ +\infty\}$, 
+to form a \emph{semiring} $\Semiring$ handling infinite sums, 
+%\cite{Eilenberg74automata}
+such as a min-plus algebra of domain $\mathbb{R}^+ \cup \{ +\infty\}$, 
 where $\oplus$ is min and $\otimes$ is plus.
@@
 %making the weight domain a semiring.
-Some efficient specialized parsing algorithms~\cite{Huang05kbest} have been proposed in this context 
-in order to compute the $n$ best syntax trees of a given input word without having to enumerate them all.
-Generally based on dynamic programming, these algorithms rely on 
-additional algebraic properties of~$\Semiring$.
--- see \eg~\cite{Huang05kbest} for some NLP applications.
+%Some efficient specialized parsing algorithms~\cite{Huang05kbest} have been proposed in this context 
+%% models represented as hypergraphs \cite{Huang05kbest}
+%in order to compute the $n$ best syntax trees of a given input word without having to enumerate them all.
+%Generally based on dynamic programming, these algorithms rely on 
+%additional algebraic properties of~$\Semiring$.
+%-- see \eg~\cite{Huang05kbest} for some NLP applications.
 %The extraction of $n$ best list is useful 
-
-We present here some weighted language models
+%the problem: quantitative parsing or symbolic parsing
+%of parsing of words over infinite input alphabet.
+
+In this paper, we present a uniform framework for weighted parsing over infinite input alphabets.
+It is based on some weighted %finite states 
+language models generalizing 
+both SA, with functions in an arbitrary semiring instead of Boolean guards, 
+and Weighted Automata (\textsf{WA}), by handling infinite alphabets -- Figure~\ref{fig:hierarchy}.
+%***
+%some weighted language models
 %model of weighted CFG 
-computing on words over infinite input alphabets %sets of terminal symbol
-- but with finite sets of states and transitions rules. 
+%computing on words over infinite input alphabets %sets of terminal symbol
+%- but with finite sets of states and transitions rules. 
 % non-terminals and production rules).
@@
 %
-This approach is close to the case of 
-Symbolic Automata (SA)~\cite{dAntoniVeanes17CAV,dAntoni21CACM}, 
-except that the domain for weight values is not restricted to be Boolean, 
-like for the guards in the rules of SA, 
-but can be an arbitrary commutative semiring (assuming some restrictions).
-%
-The models defined here are finite automata (\SWA),
-transducers (\SWT) and pushdown automata
-with a visibly restriction~\cite{AlurMadhusudan09nested} (\SWVPA).
-%\wrt an edit distance
-The latter model of automata computes on \emph{nested words}~\cite{AlurMadhusudan09nested}, 
-a structured form of words parenthesized with markup symbols, 
-corresponding to a linearization of trees.
-In expressiveness, they correspond to a strict restriction
-of weighted CFG. 
-%and can be used in a general approach for parsing over infinite input alphabets.
-Let $A$ be a \SWVPA, associating $A(w) \in \Semiring$ 
-to a given a nested word $w$ (representing a parse tree),
-and let a \SWT compute a distance $d$, in $\Semiring$, 
-between 2 strings over respectively an infinite input alphabet and the 
-same (infinite) alphabet of $A$.
-Then, the problem of Symbolic Weighted Parsing is, 
-given an input string $s$, to find a nested word $w$ minimizing 
-(according to the ranking defined by $\oplus$)
-the distance $d(s, w) \otimes A(w)$ between $s$ and $A$, 
-as defined in~\cite{Mohri03EDWA}.
-%following
-
 \begin{figure}
@@
 \end{figure}
+%
+%This approach is close to the case of 
+%Symbolic Automata (SA)~\cite{dAntoniVeanes17CAV,dAntoni21CACM}, 
+%except that the domain for weight values is not restricted to be Boolean, 
+%like for the guards in the rules of SA, 
+%but can be an arbitrary commutative semiring (assuming some restrictions).
+%
+The models presented here are finite automata called Symbolic Weighted (\SWA),
+transducers (\SWT) and pushdown automata,
+with a visibly restriction~\cite{AlurMadhusudan09nested} (\SWVPA).
+%\wrt an edit distance
+The latter model of automata computes on \emph{nested words}~\cite{AlurMadhusudan09nested}, 
+a structured form of words parenthesized with markup symbols, 
+corresponding to a linearization of trees.
+In the context of parsing, there are used here to represent (weighted) AST of CF grammars.
+More precisely, a \SWVPA $A$ associates a weight value $A(t)$ % \in \Semiring$ 
+to a given a nested word $t$ which is the linearization of an AST. %representing a parse tree.
+%
+On the other hand, 
+a \SWT is used to define a distance $d(s, t)$ between two finite words $s$ and $t$
+over an infinite alphabet, following~\cite{Mohri03EDWA}.
+Then, the \emph{SW-parsing} problem aims at %computing 
+finding $t$ minimizing (\wrt the ranking defined by $\oplus$)
+$d(s, t) \otimes A(w)$ -- 
+this value is called the distance between $s$ and $A$ in~\cite{Mohri03EDWA}.
+
+... like above weighted-parsing methods, 
+our approach to parsing proceeds in 2 steps, 
+Bar-Hillel construction... and best path search \cite{Mohri,Huang05}...
+** revise **
+%In expressiveness, they are equivalent to weighted CFG. 
+%and can be used in a general approach for parsing over infinite input alphabets.
+%
+%Let $A$ be a \SWVPA, associating $A(w) \in \Semiring$ 
+%to a given a nested word $w$ (representing a parse tree),
+%and let a \SWT compute a distance $d$, in $\Semiring$, 
+%between 2 strings over respectively an infinite input alphabet and the 
+%same (infinite) alphabet of $A$.
+%Then, the problem of Symbolic Weighted Parsing is, 
+%given an input string $s$, to find a nested word $w$ minimizing 
+%(according to the ranking defined by $\oplus$)
+%the distance $d(s, w) \otimes A(w)$ between $s$ and $A$, 
+%as defined in~\cite{Mohri03EDWA}.
 
@@
 % value in a semiring...
+
+The main contributions of the paper are: ...
+%generalization of 2 classes
+%framework for parsing over infinite alphabet, based on generic def. of distances
+%use of VPA for parsing, word instead of trees, use comparison input-output
 

### 2021-05-12 main.tex
intro (fin)

@@
 We propose a framework for weighted parsing over infinite alphabets.
-
+%
 It is based on language models called Symbolic Weighted Automata (SWA) 
@@
 %\cite{Eilenberg74automata}
-such as a min-plus algebra of domain $\mathbb{R}^+ \cup \{ +\infty\}$, 
-where $\oplus$ is min and $\otimes$ is plus.
+such as the Viterbi semiring and the tropical min-plus algebra, see Figure~\ref{fig:semirings}. 
+%of domain $\mathbb{R}^+ \cup \{ +\infty\}$, 
+%where $\oplus$ is min and $\otimes$ is plus .
 %by ranking 
@@
 On the other hand, 
-a \SWT is used to define a distance $d(s, t)$ between two finite words $s$ and $t$
+a \SWT is used to define a distance $T(s, t)$ between two finite words $s$ and $t$
 over an infinite alphabet, following~\cite{Mohri03EDWA}.
 Then, the \emph{SW-parsing} problem aims at %computing 
-finding $t$ minimizing (\wrt the ranking defined by $\oplus$)
-$d(s, t) \otimes A(w)$ -- 
-this value is called the distance between $s$ and $A$ in~\cite{Mohri03EDWA}.
-
-... like above weighted-parsing methods, 
-our approach to parsing proceeds in 2 steps, 
-Bar-Hillel construction... and best path search \cite{Mohri,Huang05}...
-** revise **
+finding $t$ minimizing 
+$T(s, t) \otimes A(t)$ (\wrt the ranking defined by $\oplus$)
+-- this value is called the distance between $s$ and $A$ in~\cite{Mohri03EDWA}.
+%
+Similarly to weighted-parsing 
+methods~\cite{Goodman99SemiringParsing,Nederhof03weightedParsing,MorbitzVogler19weighted-parsing}, 
+our approach proceeds in two steps, 
+based on properties of the SW models. 
+The first step is a Bar-Hillel construction where, 
+given a \SWT $T$, a \SWVPA $A$, and an input word $s$, 
+a \SWVPA $A_{T, s}$ is built, such that for all $t$, $A_{T, s}(t) = T(s, t) \otimes A(t)$.
+In the second step, a best AST $t$ is found by applying best search
+similar to shortest distance in graphs~\cite{Mohri02semiring,Huang05kbest} to $A_{T, s}$.
 %In expressiveness, they are equivalent to weighted CFG. 
@@
 
-The main contributions of the paper are: ...
-%generalization of 2 classes
-%framework for parsing over infinite alphabet, based on generic def. of distances
-%use of VPA for parsing, word instead of trees, use comparison input-output
-
-In Section~\ref{section:SWA} we introduce \SWA and \SWT.
-Then \SWVPA are defined in Section~\ref{sec:SWVPA}, 
-where a polynomial 1-best algorithm is described that can be use to solve
-Symbolic Weighted Parsing, as described Section~\ref{sec:parsing}.
-Finally, in Section~\ref{sec:transcription}, we present an application 
-of this approach to automated music transcription that has been implemented.
+The main contributions of the paper are: 
+($i$)~the introduction of automata, \SWA, transducers, \SWT (Section~\ref{section:SWA}),
+and visibly pushdown automata \SWVPA (Section~\ref{sec:SWVPA}),
+generalizing the corresponding classes of symbolic and weighted models, 
+%for the weighted computation on (nested) words over infinite alphabets;  
+($ii$)~a polynomial best-search algorithm for \SWVPA, %(Section~\ref{sec:best})
+% a framework for parsing over infinite alphabets, 
+and ($iii$)~a uniform framework (Section~\ref{sec:parsing}) for parsing over infinite alphabets, 
+the keys to which are 
+($iii.a$)~the \SWT-based definition of generic edit distances between input and output words,
+and ($iii.b$)~the use of nested words, and \SWVPA, 
+instead of syntax trees and grammars. %(\S~\ref{sec:trees}).
+%
+Finally, Section~\ref{sec:transcription} describes
+the implemented application %of this approach 
+to automated music transcription
+that motivated this study.
+
+%based these models and on generic def. of distances;
+%the use of VPA for parsing, word instead of trees, use comparison input-output.
+%and best-search
+
+%In Section~\ref{section:SWA} we introduce \SWA and \SWT.
+%Then \SWVPA are defined in Section~\ref{sec:SWVPA}, 
+%where a polynomial 1-best algorithm is described that can be use to solve
+%Symbolic Weighted Parsing, as described Section~\ref{sec:parsing}.
+%Finally, in Section~\ref{sec:transcription}, we present an application 
+%of this approach to automated music transcription that has been implemented.
 
@@
 \hline
-Tropical & $\mathbb{R}_+ \cup \{ +\infty\}$ & $\mathit{min}$ & $+$ & $+\infty$ & $0$\\ % & $x \leq_\oplus y \iff x \leq y$   \\
+Tropical min-plus & $\mathbb{R}_+ \cup \{ +\infty\}$ & $\mathit{min}$ & $+$ & $+\infty$ & $0$\\ % & $x \leq_\oplus y \iff x \leq y$   \\
 \hline
@@
 \subsection{Best-first Search} 
+\label{sec:best}\label{sec:search}
 Let us assume that the semiring $\Semiring$ is
@@
 \subsection{Trees and Nested-Words}
+\label{sec:trees}
 The hierarchical structure of nested-words, defined the \emph{call} and \emph{return} markup symbols of 

### 2021-05-12 main.tex
intro (fin)

@@
 
+\def\SA{\textsf{SA}\xspace}
+\def\WA{\textsf{WA}\xspace}
 \def\SWT{\textsf{swT}\xspace}
@@
 Following a similar idea, % and generalizing,
-in Symbolic automata (\textsf{SA})~\cite{dAntoniVeanes17CAV,dAntoni21CACM}, 
+in symbolic automata (\SA)~\cite{dAntoniVeanes17CAV,dAntoni21CACM}, 
 the transitions are guarded by predicates over infinite alphabet domains.
@@
 In \emph{weighted language models}, 
-like \eg probabilistic context-free grammars, % (CFG),
+like \eg probabilistic context-free grammars % (CFG),
+and weighted automata (\WA) \cite{Droste09handbook},
 a weight value is associated to each transition rule, % production rule
@@
 associated to a given input word.
-Typically, $\oplus$ can select the minimum of two weights.
+Typically, $\oplus$ will select the best of two weight values.
 %Intuitively,~$\oplus$ selects, or ranking, the syntax trees.
 The weight domain, equipped with these two operators is assumed, at minima, 
-to form a \emph{semiring} $\Semiring$ handling infinite sums, 
+to form a \emph{semiring} %$\Semiring$ 
+where $\oplus$ can be extended to infinite sums, 
 %\cite{Eilenberg74automata}
@@
 In this paper, we present a uniform framework for weighted parsing over infinite input alphabets.
-It is based on some weighted %finite states 
+It is based on weighted %finite states 
 language models generalizing 
-both SA, with functions in an arbitrary semiring instead of Boolean guards, 
-and Weighted Automata (\textsf{WA}), by handling infinite alphabets -- Figure~\ref{fig:hierarchy}.
+both~\SA, with functions in an arbitrary semiring instead of Boolean guards, 
+and~\WA, by handling infinite alphabets -- Figure~\ref{fig:hierarchy}.
 %***
@@
 %
-The models presented here are finite automata called Symbolic Weighted (\SWA),
+The models presented here are finite automata called symbolic-weighted (\SWA),
 transducers (\SWT) and pushdown automata,
@@
 corresponding to a linearization of trees.
-In the context of parsing, there are used here to represent (weighted) AST of CF grammars.
+In the context of parsing, they are used here to represent (weighted) AST of CF grammars.
 More precisely, a \SWVPA $A$ associates a weight value $A(t)$ % \in \Semiring$ 
-to a given a nested word $t$ which is the linearization of an AST. %representing a parse tree.
+to a given a nested word $t$, which is the linearization of an AST. %representing a parse tree.
 %
@@
 a \SWVPA $A_{T, s}$ is built, such that for all $t$, $A_{T, s}(t) = T(s, t) \otimes A(t)$.
-In the second step, a best AST $t$ is found by applying best search
-similar to shortest distance in graphs~\cite{Mohri02semiring,Huang05kbest} to $A_{T, s}$.
+In the second step, a best AST $t$ is found by applying to $A_{T, s}$ 
+a best search algorithm similar to shortest distance in graphs~\cite{Mohri02semiring,Huang05kbest}.
 %In expressiveness, they are equivalent to weighted CFG. 

### 2021-05-12 main.tex
1 pass semirings

@@
 between words and languages, with weighted transducers, 
-and extending it to infinite alphabets.
+extending it to infinite alphabets.
 % with models of symbolic weighted automata and transducers. 
@@
 \label{section:semiring}
-We shall consider semiring domains for the weight values of our language models.
+We shall consider semiring for the weight values of our language models.
 %
@@
 \begin{itemize}
-\item $\oplus$ is commutative 
--- therefore $\< \Semiring, \oplus, \zero>$ is a commutative monoid 
-   and $\< \Semiring, \otimes, \one>$ is a monoid,
+\item $\oplus$ is commutative;
+ $\< \Semiring, \oplus, \zero>$ is a commutative monoid 
+   and $\< \Semiring, \otimes, \one>$ a monoid,
 \item $\otimes$ distributes over~$\oplus$:  $\forall x, y, z \in \mathbb{S}$,
@@
 and $\otimes$ combines two values into a single value, 
-in order to describe the chaining of several transitions.
+in a chaining of transitions.
 %and let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a {semiring},
@@
 We follow here the direction  %follows \cite{Mohri02semiring}, and 
-that coincides than the usual ordering on the Tropical semiring (\emph{min-plus}).
-
-\begin{lemma}[Monotony, \cite{Mohri02semiring}]
+that coincides with the usual ordering on the Tropical semiring \emph{min-plus} 
+(Figure~\ref{fig:semirings}).
+
+\begin{lemma}[Monotony, \cite{Mohri02semiring}] \label{lem:monotonic}
 Let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be an idempotent semiring.
@@
 \end{lemma}   
-We say in this case that $\Semiring$ is \emph{monotonic}. % \wrt $\leq_\oplus$.
+When the property of Lemma~\ref{lem:monotonic} holds, 
+%We say in this case that 
+$\Semiring$ is called  \emph{monotonic}. % \wrt $\leq_\oplus$.
 %A semiring $\Semiring$ 
@@
 Intuitively, it means that combining elements with $\otimes$ always increase their weight. 
-Formally, it is defined as the property of the following lemma.
+Formally, it is defined as the property ($i$) below. %of the following lemma.
 
@@
 %
-In~\cite{Huang08advanceddynamic}, the property~$i.$ is called 
-$\Semiring$ is \emph{superior} \wrt the ordering~$\leq_\oplus$.
-It implies that 
-%It holds that If $\Semiring$ is superior \wrt~$\leq$, then %$\one \leq \zero$
+\begin{proof} %(Lemma~\ref{lem:bounded})
+For $(ii) \Rightarrow (i)$ : 
+$x \oplus (x \otimes y) = x \otimes (\one \oplus y) = x$, 
+by distributivity of~$\otimes$ over~$\oplus$. 
+Hence $x \leq_\oplus x \otimes y$.
+Similarly, $y \oplus (x \otimes y) = (\one \oplus x) \otimes y = y$, 
+hence $y \leq_\oplus x \otimes y$.
+%
+For $(i) \Rightarrow (ii)$ :
+%$(i)$ implies that $\one \leq_\oplus x$ for all $x \in \Semiring$, 
+by the second inequality of ($i$), with $y = \one$, 
+$\one \leq_\oplus x \otimes \one = x$
+\ie, by definition of $\leq_\oplus$, $\one \oplus x = \one$.
+\qed
+\end{proof}
+
+In~\cite{Huang08advanceddynamic}, $\Semiring$ with the property~$(i)$ 
+is called \emph{superior} \wrt the ordering~$\leq_\oplus$.
+We have seen in the proof of Lemma~\ref{lem:bounded} that it implies that 
+$\one \leq_\oplus x$ for all $x \in \Semiring$.
+Similarly, by the first inequality of ($i$) with $y = \zero$,  
+$x \leq_\oplus x \otimes \zero = \zero$.
+%
+Hence, in a superior semiring, % \wrt~$\leq$, 
+it holds that %$\one \leq \zero$
 for all $x \in \Semiring$, $\one \leq_\oplus x \leq_\oplus \zero$.
-Indeed, by the second inequality of $(i.)$ with $y = \one$, $\one \leq_\oplus \one \otimes x = x$, 
-and, by the first inequality of $(i.)$ with $y = \zero$,  
-$x \leq_\oplus x \otimes \zero = \zero$.
 %
@@
 %
-In~\cite{Mohri02semiring}, the property of Lemma~\ref{lem:bounded}(in the form $ii.$) 
+In~\cite{Mohri02semiring}, the property ($ii$) of Lemma~\ref{lem:bounded}  
 is called \emph{boundedness} of $\Semiring$ -- we shall use this term in the rest of the paper. 
@@
 are weighted by values of $\Semiring$, the loops can be safely avoided.
-
-\begin{proof}(Lemma~\ref{lem:bounded})
-For $(ii.) \Rightarrow (i.)$ : 
-$x \oplus (x \otimes y) = x \otimes (\one \oplus y) = x$, 
-by distributivity of $\otimes$ over $\oplus$. 
-Hence $x \leq_\oplus x \otimes y$.
-Similarly, $y \oplus (x \otimes y) = (\one \oplus x) \otimes y = y$, 
-hence $y \leq_\oplus x \otimes y$.
-%
-For $(i.) \Rightarrow (ii.)$ : as we have seen, $(i.)$ implies that 
-$\one \leq_\oplus x$ for all $x \in \Semiring$, 
-\ie, by definition of $\leq_\oplus$, $\one \oplus x = \one$.
-\qed
-\end{proof}
 
@@
 is well-defined and in $\dom(\Semiring)$,
-and the following properties hold:
+and the following holds: %properties
 \begin{description}
@@
 \begin{example}
-Figure~\ref{gif:semirings} presented examples of semirings interesting in practice 
+Figure~\ref{fig:semirings} presented examples of semirings interesting in practice 
 and enjoying the above properties.

### 2021-05-14 main.tex
pass preuve proj. SWT

@@
 \def\wei{\mathsf{w}}
+\def\mei{\mathsf{m}}
 \def\init{\mathsf{in}}
@@
 equipped with two associative
-binary operators~$\oplus$ and $\otimes$
-with respective neutral elements $\zero$ and $\one$ and such that:
+binary operators~$\oplus$ and $\otimes$,
+with respective neutral elements $\zero$ and $\one$, and such that:
 %$\< \mathbb{S}, \oplus, \zero>$ is a commutative monoid
@@
 \begin{proof} %(Lemma~\ref{lem:bounded})
-For $(ii) \Rightarrow (i)$ : 
+$(ii) \Rightarrow (i)$ : 
 $x \oplus (x \otimes y) = x \otimes (\one \oplus y) = x$, 
@@
 %
-For $(i) \Rightarrow (ii)$ :
+$(i) \Rightarrow (ii)$ :
 %$(i)$ implies that $\one \leq_\oplus x$ for all $x \in \Semiring$, 
 by the second inequality of ($i$), with $y = \one$, 
-$\one \leq_\oplus x \otimes \one = x$
-\ie, by definition of $\leq_\oplus$, $\one \oplus x = \one$.
+$\one \leq_\oplus x \otimes \one = x$, \ie, 
+by definition of $\leq_\oplus$, $\one \oplus x = \one$.
 \qed
@@
 
-In~\cite{Huang08advanceddynamic}, $\Semiring$ with the property~$(i)$ 
-is called \emph{superior} \wrt the ordering~$\leq_\oplus$.
+In~\cite{Huang08advanceddynamic}, the property~$(i)$ 
+is called $\Semiring$ \emph{superior} \wrt the ordering~$\leq_\oplus$.
 We have seen in the proof of Lemma~\ref{lem:bounded} that it implies that 
@@
 We shall need below infinite sums with~$\oplus$.
-A semiring~$\Semiring$ is called \emph{complete}~\cite{Kuich97semirings}
-if for every family
+A semiring~$\Semiring$ is called \emph{complete}~\cite{Droste09handbook} 
+%(\cite{Droste09handbook} chapter 1) %\cite{Kuich97semirings}
+if it has an %infinite sum 
+operation $\bigoplus_{i \in I} x_i$
+for every family
 $(x_i)_{i \in I}$ %$\{ x_i \mid i \in I \}$
-of elements of $\dom(\Semiring)$ over an index set $I \subset \mathbb{N}$,
-the infinite sum $\bigoplus_{i \in I} x_i$
-is well-defined and in $\dom(\Semiring)$,
-and the following holds: %properties
+of elements of $\dom(\Semiring)$ over an index set $I \subset \mathbb{N}$, such that
+%is well-defined and in $\dom(\Semiring)$,
+the following holds: %properties
 \begin{description}
 \item[$i.$]
-\emph{infinite sums extend finite sums:}
+\emph{infinite sums extend finite sums:}\\
 $\displaystyle\bigoplus_{i \in \emptyset} x_i = \zero$,\quad 
-      $\forall j\in \mathbb{N}, \displaystyle\bigoplus_{i \in \{ j \}} x_i = x_j$,\quad
+      $\forall j\in \mathbb{N}, \displaystyle\bigoplus_{i \in \{ j \}} x_i = x_j$,
       $\forall j, k\in \mathbb{N}, j\neq k, 
@@
 \item[$ii.$]
-\emph{associativity and commutativity:}
+\emph{associativity and commutativity:}\\
 for all $I \subseteq \mathbb{N}$
@@
 \hline
-Tropical min-plus & $\mathbb{R}_+ \cup \{ +\infty\}$ & $\mathit{min}$ & $+$ & $+\infty$ & $0$\\ % & $x \leq_\oplus y \iff x \leq y$   \\
+Tropical min-plus & $\mathbb{R}_+ \cup \{ \infty\}$ & $\mathit{min}$ & $+$ & $\infty$ & $0$\\ % & $x \leq_\oplus y \iff x \leq y$   \\
 \hline
@@
 \label{section:symbols}
-We shall now define the functions labeling the transitions of SW-automata and -transducers.
-Let $\Sigma$ and $\Delta$ be two countable sets of symbols
-called respectively \emph{input} and \emph{output} \emph{alphabets}, 
-and let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative semiring.
-
-\noindent 
-A \emph{label theory} over $\Sigma$ and $\Delta$
-is made of 3 recursively enumerable sets:
+We shall now define the functions labeling the transitions of SW automata and transducers,
+generalizing the Boolean algebras of~\cite{dAntoniVeanes17CAV} 
+from Boolean to other semiring domains.
+%
+We consider \emph{alphabets}, which are countable sets of symbols 
+denoted $\Sigma$, $\Delta$,...
+%Let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative, complete semiring.
+%
+\noindent 
+Given a semiring $\< \Semiring, \oplus, \zero, \otimes, \one>$, 
+a $\Semiring$-\emph{label theory}
+is a tuplet $\bar\Phi$ of recursively enumerable sets of the form
 %$\Phi_\epsilon \subseteq \Semiring$, % containing constant functions valued in $\Semiring$, 
-$\Phi_\Sigma$ and $\Phi_\Delta$, 
-containing unary functions in $\Sigma \to \Semiring$, resp. $\Delta \to \Semiring$, 
-and $\Phi_{\Sigma, \Delta}$  containing binary functions in $\Sigma \times \Delta \to \Semiring$, 
-such that 
-for all $\alpha \in \Semiring$, and $\phi \in \Phi_\Sigma$,
+$\Phi_\Sigma$, %and $\Phi_\Delta$, 
+containing unary functions of type $\Sigma \to \Semiring$, %resp. $\Delta \to \Semiring$, 
+and $\Phi_{\Sigma, \Delta}$, containing binary functions $\Sigma \times \Delta \to \Semiring$, 
+such that:
+
+\noindent -- 
+for all $\Phi_{\Sigma, \Delta} \in \bar\Phi$, there exists 
+$\Phi_{\Sigma} \in \bar\Phi$ and $\Phi_{\Delta} \in \bar\Phi$
+
+\noindent -- 
+for all $\alpha \in \Semiring$ and $\phi \in \Phi_\Sigma$,
       $\alpha \otimes \phi : x \mapsto \alpha \otimes \phi(x)$, 
-and $\phi \otimes \alpha : x \mapsto \phi(x) \otimes \alpha$
-belong to $\Phi_\Sigma$, and similarly for $\oplus$ 
-and for $\Phi_\Delta$ and $\Phi_{\Sigma, \Delta}$.
-%
-Moreover, these sets are required to be closed under the following 
-operators~$\oplus$ and $\otimes$ of~$\Semiring$:
+      and $\phi \otimes \alpha : x \mapsto \phi(x) \otimes \alpha$\\
+\phantom{--} belong to $\Phi_\Sigma$, and similarly for $\oplus$ 
+      and for $\Phi_{\Sigma, \Delta}$
+
+\noindent -- 
 for all $\phi, \phi' \in \Phi_\Sigma$,
-all $\psi, \psi' \in \Phi_\Delta$, 
-and $\eta, \eta' \in \Phi_{\Sigma, \Delta}$, the function
-\begin{description}
-\item $\phi \otimes \phi' : x \mapsto \phi(x) \otimes \phi'(x)$ belongs to $\Phi_\Sigma$, 
-\item $\psi \otimes \psi' : y \mapsto \psi(y) \otimes \psi'(y)$ belongs to $\Phi_\Delta$,
-\item $\phi \otimes \eta : x, y \mapsto \phi(x) \otimes \eta(x, y)$ belongs to $\Phi_{\Sigma, \Delta}$,
-\item $\eta \otimes \psi : x, y \mapsto \eta(x, y) \otimes \psi(y)$ belongs to $\Phi_{\Sigma, \Delta}$,
-\item $\eta \otimes \eta' : x, y \mapsto \eta(x, y) \otimes \eta'(x, y)$ belongs to $\Phi_{\Sigma, \Delta}$.
-\end{description}
-And the same also holds for the binary sum operator $\oplus$.
-
-\noindent
-%\marginpar{necessary?}
-Finally, it is also required 
-% that the codomain of every function of $\Phi_\Sigma$ and $\Phi_\Delta$ 
-% is a subset of $\Phi_\epsilon$, and
-that all partial applications of functions $\Phi_{\Sigma, \Delta}$, 
-resp.  $f_a: y \mapsto f(a, y)$ for $a \in \Sigma$ and $y \in \Delta$
-and  $f_b: x \mapsto f(x, b)$ for $b \in \Delta$ and $x \in \Sigma$, 
+$\phi \otimes \phi': x \mapsto \phi(x) \otimes \phi'(x)$ belongs to $\Phi_\Sigma$
+
+\noindent -- 
+for all $\eta, \eta' \in \Phi_{\Sigma, \Delta}$
+$\eta \otimes \eta': x, y \mapsto \eta(x, y) \otimes \eta'(x, y)$ belongs to $\Phi_{\Sigma, \Delta}$
+
+\noindent -- 
+for all $\phi \in \Phi_\Sigma$ and $\eta \in \Phi_{\Sigma, \Delta}$,
+$\phi \otimes \eta: x, y \mapsto \phi(x) \otimes \eta(x, y)$ belongs to $\Phi_{\Sigma, \Delta}$
+
+\noindent -- 
+for all $\psi \in \Phi_\Delta$ and $\eta \in \Phi_{\Sigma, \Delta}$,
+$\eta \otimes \psi: x, y \mapsto \eta(x, y) \otimes \psi(y)$ belongs to $\Phi_{\Sigma, \Delta}$
+
+\noindent -- 
+similar closures holds for $\oplus$
+
+\noindent -- 
+partial applications $\eta \in \Phi_{\Sigma, \Delta}$, 
+resp. $\eta_a: y \mapsto \eta(a, y)$ for $a \in \Sigma$ %and $y \in \Delta$
+and\\
+\phantom{--} $\eta_b: x \mapsto \eta(x, b)$ for $b \in \Delta$ %and $x \in \Sigma$, 
 belong resp. to~$\Phi_\Sigma$ and~$\Phi_\Delta$.
+
+%Moreover, these sets are required to be closed under the 
+%operators~$\oplus$ and~$\otimes$ of~$\Semiring$:
+%for all $\phi, \phi' \in \Phi_\Sigma$,
+%$\psi, \psi' \in \Phi_\Delta$, 
+%and $\eta, \eta' \in \Phi_{\Sigma, \Delta}$, %the function
+%%
+%\begin{center}
+%\begin{tabular}{cclll}
+%$\phi \otimes \phi'$ & : & $x \mapsto \phi(x) \otimes \phi'(x)$ & belongs to $\Phi_\Sigma$,\\
+%$\psi \otimes \psi'$ & : & $y \mapsto \psi(y) \otimes \psi'(y)$ & belongs to $\Phi_\Delta$,\\
+%$\phi \otimes \eta$\;  & : & $x, y \mapsto \phi(x) \otimes \eta(x, y)$ & belongs to $\Phi_{\Sigma, \Delta}$,\\
+%$\eta \otimes \psi$  & : & $x, y \mapsto \eta(x, y) \otimes \psi(y)$ & belongs to $\Phi_{\Sigma, \Delta}$,\\
+%$\eta \otimes \eta'$ & : & $x, y \mapsto \eta(x, y) \otimes \eta'(x, y)$ & belongs to $\Phi_{\Sigma, \Delta}$, &
+%\multicolumn{1}{r}{and similarly for $\oplus$.}\\ %the same also holds for the binary sum operator $\oplus$.
+%\end{tabular}
+%\end{center}
+%
+%Finally, it is also required 
+%% that the codomain of every function of $\Phi_\Sigma$ and $\Phi_\Delta$ 
+%% is a subset of $\Phi_\epsilon$, and
+%that the partial applications of a function $\eta \in \Phi_{\Sigma, \Delta}$, 
+%resp.  $\eta_a: y \mapsto f(a, y)$ for $a \in \Sigma$ and $y \in \Delta$
+%and  $\eta_b: x \mapsto f(x, b)$ for $b \in \Delta$ and $x \in \Sigma$, 
+%belong resp. to~$\Phi_\Sigma$ and~$\Phi_\Delta$.
+
+%For a complete semiring $\Semiring$, 
+%a function $\phi: \Sigma \to \Semiring$, 
+%resp. $\eta: \Sigma \times \Delta \to \Semiring$,
+%is called
 
@@
 %(SWT)
+
+Let $\Semiring$ be a commutative and complete semiring, 
+$\Sigma$ and $\Delta$ be alphabets called respectively \emph{input} and \emph{output}, %{alphabets}, 
+and $\bar\Phi = \< \Phi_\Sigma, \Phi_\Delta, \Phi_{\Sigma, \Delta}>$ be an $\Semiring$-label theory.
+
 \begin{definition}
@@
 A \emph{symbolic-weighted transducer} (\SWT)
-over the input and output alphabets~$\Sigma$ and $\Delta$ 
-and the semiring $\Semiring$ is a tuple
+over $\Sigma$, $\Delta$, $\Semiring$ and $\bar\Phi$
+%the input and output alphabets~$\Sigma$ and $\Delta$ with label theory $\bar\Phi$, and the semiring $\Semiring$ 
+is a tuple
 $T = \< Q, \init, \bar{\wei}, \final >$,
@@
 %
-We summarize the above 4-uplet of transition functions into a unique 
-function~$\wei$ from
-$Q \times (\Sigma \cup \{ \epsilon \}) \times (\Delta \cup \{ \epsilon \}) \times Q$
-into $\Semiring$, %also called $\wei$ for simplicity, 
-such that, for all $q, q' \in Q$, $a \in \Sigma$, and $b \in \Delta$, 
+A pair of states $\< q, q'>$ such that $\wei_\epsilon(q, q') \neq \zero$
+is called an \emph{$\epsilon$-transition}.
+
+\noindent We synthesize the 4-uplet of transition functions into a unique 
+function~$\wei:
+ Q \times (\Sigma \cup \{ \epsilon \}) \times (\Delta \cup \{ \epsilon \}) \times Q 
+ \to \Semiring$, %also called $\wei$ for simplicity, 
+such that, for all $q, q' \in Q$, $a \in \Sigma$,  $b \in \Delta$, 
 %and with 
@@
 \wei(q, \epsilon, \epsilon, q') & = &  \wei_\epsilon(q, q'),\\ %\phi_\epsilon\\
-\wei(q, a, \epsilon, q') & = & \phi_\Sigma(a) & %\wei_\Sigma(q, q')(a)
-\quad\mathrm{where~} \phi_\Sigma = \wei_\Sigma(q, q'),\\
-\wei(q, \epsilon, b, q') & = & \phi_\Delta(b) & 
-\quad\mathrm{where~} \phi_\Delta = \wei_\Delta(q, q'),\\
-\wei(q, a, b, q') & = & \phi_{\Sigma, \Delta}(a, b) & 
-\quad\mathrm{where~} \phi_{\Sigma, \Delta} = \wei_{\Sigma, \Delta}(q, q').\\
+\wei(q, a, \epsilon, q') & = & \phi(a) & %\wei_\Sigma(q, q')(a)
+\quad\mathrm{where~} \phi = \wei_\Sigma(q, q') \in \Phi_\Sigma,\\
+\wei(q, \epsilon, b, q') & = & \psi(b) & 
+\quad\mathrm{where~} \psi = \wei_\Delta(q, q') \in \Phi_\Delta,\\
+\wei(q, a, b, q') & = & \eta(a, b) & 
+\quad\mathrm{where~} \eta = \wei_{\Sigma, \Delta}(q, q') \in \Phi_{\Sigma, \Delta}.\\
 \end{array}      
@@
 
-The symbolic-weighted transducer $T$ defines a mapping 
+\noindent
+The symbolic-weighted transducer~$T$ defines a mapping 
 from the pairs of strings of $\Sigma^* \times \Delta^*$ 
-into the weights of~$\Semiring$,
+into~$\Semiring$,
 based on the following intermediate function $\weight_T$
@@
 \begin{array}{rccl}
-xc(q, s, t, q') & = & 
+\weight_T(q, s, t, q') & = & 
    \displaystyle\bigoplus_{q'' \in Q} &
@@
 \]
-The sum above can be infinite because of the expression in the first line, 
-however, $\T(\Omega)$ is well defined since $\Semiring$ is complete.
-%
-As we shall see, epsilon-transitions can be removed and hence the sum made finite.
+
+The above sum may be infinite because of the $\epsilon$-transitions in the first and second lines.
+However, $\weight_T$ is well defined since $\Semiring$ is complete.
+%
+As we shall see, under conditions, the $\epsilon$-transitions can be removed and hence the sum made finite.
 %
@@
 from state $q$ into state $q'$ while reading $a$ and writing $b$ 
-(including the case $a = \epsilon$ and $a = \epsilon$).
+(including the case $a = \epsilon$ and $b = \epsilon$).
 %These functions $\phi$ act as guards for the transducer's transitions, 
@@
 
-The above expression of $\weight_T$ can be seen as a stateful definition of 
+The expression of $\weight_T$ can be seen as a stateful definition of 
 an edit-distance between a word $s \in \Sigma^*$ and a word $t \in \Delta^*$,
@@
 $\wei(q, a, \epsilon, q'')$, in the third line, is the cost of 
-the deletion of a symbol~$a \in \Sigma$ in~$s$, 
+the deletion of the symbol~$a \in \Sigma$ in~$s$, 
 $\wei(q, \epsilon, b, q'')$, in the fourth line, is the cost 
-of the deletion of insertion of~$b \in \Delta$ in~$s$, 
+of the insertion of~$b \in \Delta$ in $t$, 
 and $\wei(q, a, b, q'')$, in the last line, is the cost 
@@
 The weight associated by $T$ to  $\< s, t> \in \Sigma^* \times \Delta^*$
-is then defined as follows: 
+is defined as follows: 
 \begin{equation}
@@
 The \emph{Symbolic Weighted Automata} %$A = \< Q, \init, \weight, \final >$
-over $\Sigma$ and $\Semiring$ 
+%over $\Sigma$,  $\Semiring$ and $\bar\Phi$
 are defined similarly as the transducers of Definition~\ref{def:SWT}, 
@@
 %
-In this case, a label theory over $\Sigma$ is reduced to
-a set $\Phi_\Sigma$ closed under~$\oplus$ and~$\otimes$.
+In this case, the label theory $\bar\Phi$ is reduced to a singleton $\< \Phi_\Sigma>$.
+%over $\Sigma$ is reduced to
+%a set $\Phi_\Sigma$ closed under~$\oplus$ and~$\otimes$.
 %
@@
 A \emph{symbolic-weighted automaton} (\SWA)
-over the input alphabet~$\Sigma$
-and the commutative semiring $\Semiring$ is a tuple
+over $\Sigma$, $\Semiring$ and $\bar\Phi$
+%the input alphabet~$\Sigma$ and the commutative semiring $\Semiring$ 
+is a tuple
 $A = \< Q, \init, \bar{\wei}, \final >$,
@@
 The above transition functions $\wei_\epsilon$ and $\wei_\Sigma$
-are synthesized into a function~$\wei$ 
-from~$Q \times (\Sigma \cup \{ \epsilon \}) \times Q$ into~$\Semiring$ 
-as above.
-%
-When $\wei_\epsilon(q, q') = \zero$ for all $q, q' \in Q$, 
-the automaton~$A$ is called \emph{without $\epsilon$-transitions}.
+can be synthesized as above into a function
+$\wei: Q \times (\Sigma \cup \{ \epsilon \}) \times Q \to \Semiring$.
+
+%
+%When $\wei_\epsilon(q, q') = \zero$ for all $q, q' \in Q$, 
+%the automaton~$A$ is called \emph{without $\epsilon$-transitions}.
 
@@
 \begin{proof}
-Let $T = \< Q, \init, \bar{\wei}, \final >$ 
-where $\bar{\wei} \< \wei_\epsilon, \wei_\Sigma, \wei_\Delta, \wei_{\Sigma, \Delta}>$
-has an associated synthesized form $\wei$, 
+Let $T = \< Q, \init, \bar{\wei}, \final >$,
+let 
+$\wei: Q \times (\Sigma \cup \{ \epsilon \}) \times (\Delta \cup \{ \epsilon \}) \times Q 
+ \to \Semiring$
+be the synthesized form of 
+$\bar{\wei} = \< \wei_\epsilon, \wei_\Sigma, \wei_\Delta, \wei_{\Sigma, \Delta}>$,
 and let $s = s_1 \ldots s_n$ with $s_1, \ldots, s_n \in \Sigma$.
@@
 %
-Let $Q' = [1..n] \times Q$ be the state set of $A_{T, s}$,
-and let, for all $q \in Q$, 
-$\init'(1, q) = \init(q)$ and $\init'(i, q) = \zero$ for all $1 < i \leq n$, 
-and $\final'(n, q) = \final(q)$ and $\final'(i, q) = \zero$ for all $1 \leq i <  n$.
-
-\noindent
-We define the functions $\wei'_\epsilon$ and $\wei'_\Sigma$ for $A_{T, s}$
-in their synthesized form: for all $1 \leq i, j \leq n$, $q, q' \in Q$ and $a \in \Sigma$, 
+The state set of $A_{T, s}$ will be $Q' = [1..n] \times Q$.
+Its state entering weight function is defined by 
+$\init'(1, q) = \init(q)$ for all $q \in Q$, 
+and $\init'(i, q) = \zero$ for all $1 < i \leq n$, 
+and its state leaving weight function by 
+$\final'(n, q) = \final(q)$ or all $q \in Q$, 
+and $\final'(i, q) = \zero$ for all $1 \leq i <  n$.
+
+\noindent
+We define transition functions $\wei'_\epsilon$ and $\wei'_\Sigma$ for $A_{T, s}$, 
+together with their synthesized form $\wei'$ as follows: 
+for all $1 \leq i, j \leq n$, $q, q' \in Q$, %and $a \in \Sigma$, 
 and $b \in \Delta$, 
 \[
-\begin{array}{lclcl}
-\wei'(\< i, q>, \epsilon, \< i, q'>)   & = & \wei(q, \epsilon, \epsilon, q'), & \quad & \\ 
-\wei'(\< i, q>, \epsilon, \< i+1, q'>) & = & \wei(q, s_i, \epsilon, q') & & \mathrm{if~} i < n,\\ 
-\wei'(\< i, q>, \epsilon, \< j, q'>)   & = & \zero & & \mathrm{if~} j \neq i, j \neq i+1,\\ 
-\wei'(\< i, q>, b, \< i, q'>)          & = & \wei(q, \epsilon, b, q'),\\ 
-\wei'(\< i, q>, b, \< i+1, q'>)        & = & \wei(q, s_i, b, q') & & \mathrm{if~} i < n,\\ 
-\wei'(\< i, q>, b, \< j, q'>)          & = & \zero & & \mathrm{if~} j \neq i, j \neq i+1. 
+\begin{array}{lclclcll}
+\wei'_\epsilon\bigl(\< i, q>, \< i, q'>\bigr) & = & \wei_\epsilon(q, q') & \mathrm{\ie} &
+\wei'\bigl(\< i, q>, \epsilon, \< i, q'>\bigr)   & = & \wei(q, \epsilon, \epsilon, q'), & \\ 
+%
+\wei'_\epsilon\bigl(\< i, q>, \< i+1, q'>\bigr) & = & \phi(s_i) & 
+\multicolumn{5}{l}{\mathrm{where~} \phi = \wei_\Sigma(q, q'),}\\
+ & & & \mathrm{\ie} &
+\wei'\bigl(\< i, q>, \epsilon, \< i+1, q'>\bigr) & = & \wei(q, s_i, \epsilon, q') & 
+\mathrm{if~} i < n,\\ 
+%
+\wei'_\epsilon\bigl(\< i, q>, \< j, q'>\bigr) & = & \zero & 
+%\mathrm{\ie} & \wei'(\< i, q>, \epsilon, \< j, q'>)   & = & \zero & & 
+\multicolumn{5}{l}{\mathrm{if~} j \neq i, i+1,}\\ 
+%
+\wei'_\Sigma\bigl(\< i, q>, \< i, q'>\bigr) & = & \wei_\Delta(q, q') & \mathrm{\ie} &
+\wei'\bigl(\< i, q>, b, \< i, q'>\bigr)          & = & \wei(q, \epsilon, b, q'),\\ 
+%
+\wei'_\Sigma\bigl(\< i, q>, \< i+1, q'>\bigr) & = & \eta_{s_i} &
+\multicolumn{5}{l}{\mathrm{where~} \eta = \wei_{\Sigma, \Delta}(q, q'),}\\
+ & & & \mathrm{\ie} & \wei'\bigl(\< i, q>, b, \< i+1, q'>\bigr) & = & \wei(q, s_i, b, q') & 
+ \mathrm{if~} i < n,\\ 
+%
+\wei'_\Sigma\bigl(\< i, q>, \< j, q'>\bigr) & = & \zero & \mathrm{\ie} &
+\wei'\bigl(\< i, q>, b, \< j, q'>\bigr)     & = & \zero & \mathrm{if~} j \neq i, i+1. 
 \end{array}      
 \]
-Note that the functions defined above belong to the set $\Phi_\Delta$ 
-of the underlying label theory, 
-by the hypothesis of closure under partial application.
-%
+We use the assumption that the underlying label theory of $T$ is 
+is closed under partial applications (Section~\ref{section:symbols})
+in the above construction of $\wei'_\Sigma$.
+
+\noindent
 The $\SWA$ wanted is $A_{T, s} = \< Q', \init', \bar{\wei'}, \final' >$.
@@
 where the size $\| T \|$ of $T$ is its number of states $|Q|$.
+
+\noindent 
+When $\Semiring$ is complete, we call \emph{summary} of a 
+to a $\SWT$ $T = \< Q, \init, \bar{\wei}, \final >$
+a triplet of functions 
+$\mei_\Sigma$, $\mei_\Delta$, and $\mei_{\Sigma, \Delta}$
+from $Q \times Q$ into~$\Semiring$, defined by 
+\[
+\begin{array}{rcll}
+\mei_\Sigma(q, q') & = & \bigoplus_{a \in \Sigma} \phi(a) & 
+\quad\mathrm{where~} \phi = \wei_\Sigma(q, q') \in \Phi_\Sigma,\\
+\mei_\Delta(q, q') & = & \bigoplus_{b \in \Delta} \psi(b) & 
+\quad\mathrm{where~} \psi = \wei_\Delta(q, q') \in \Phi_\Delta,\\
+\mei_{\Sigma, \Delta}(q, q') & = & 
+ \bigoplus_{a \in \Sigma} \bigoplus_{b \in \Delta} \eta(a, b) & 
+\quad\mathrm{where~} \eta = \wei_{\Sigma, \Delta}(q, q') \in \Phi_{\Sigma, \Delta}.\\
+\end{array}      
+\]
+Note that these functions are well defined under the assumption that $\Semiring$ is complete.
+%A $\SWT$ is called \emph{presented} when these functions, with finite domains, 
+%are given explicitly.
+For a \SWA, the summary is only defined by $\mei_\Sigma$.
 
@@
 We give now a formal definition of these transitions of the automaton~$A$
-in term of a weight value
-computed by an intermediate function $\weight_A$, like in Section~\ref{sec:SWT}.
+in term of %a weight value computed by 
+an intermediate function $\weight_A$, like in Section~\ref{sec:SWT}.
 In the case of a pushdown automaton, a configuration is composed 
@@
 \label{sec:best}\label{sec:search}
-Let us assume that the semiring $\Semiring$ is
+**hypotheses**
+Let us assume that the semiring~$\Semiring$ is
 commutative, bounded, complete, and total.
+and 
+** assume label theory is s.t. 
+for all $\phi \in \Phii$, 
+$\psi \in \Phir$, 
+and $\eta \in \Phicr$, 
+$\displaystyle\bigoplus_{a \in \Omegai} \phi(a)$ 
+$\displaystyle\bigoplus_{r \in \Omegar} \phi(r)$ and 
+$\displaystyle\bigoplus_{{\call{c}} \in \Omegac} 
+\displaystyle\bigoplus_{{\return{r}} \in \Omegar} \eta({\call{c}}, {\return{r}})$
+are computable...
+
 % total ? 
@@
 %... exist and in $\Semiring$.
+
 We propose a Dijkstra algorithm computing the minimal weight by $A$, 
-\wrt $\leq_\oplus$, for a word in $\Semiring^*$.
+\wrt~$\leq_\oplus$, for a word in~$\Omega^*$.
 
@@
 we require an additional property the of weight functions.
+%
 \begin{definition}

### 2021-05-16 abstract.tex
split files

--- 2021-05-12 main.tex
@@
-\documentclass[runningheads]{llncs}
-%\documentclass[a4paper,11pt]{article}
-%\setcounter{page}{1}
-
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-
-\usepackage{hyperref}
-%\usepackage[bookmarks,bookmarksnumbered,naturalnames,plainpages=false]{hyperref}
-%usepackage{url}
-
-% for footnote ref
-\usepackage{refcount}
-
-% array and tabular
-\usepackage{array}
-\newcolumntype{C}[1]{>{\centering\arraybackslash\hspace{0pt}}p{#1}}
-
-% extension of enumerate env. (style for displaying counters)
-% \usepackage{enumerate}
-
-%% pictures
-% \usepackage{graphicx} 
-% \DeclareGraphicsExtensions{.pdf,.png,.jpg}
-% \graphicspath{fig/}
-
-%% PGF, Tikz
-\usepackage{tikz-cd}
-%% \usepackage{pgfplots}
-%% \usepgfplotslibrary{dateplot}
-%% \usepackage{pgf,pgfarrows,pgfnodes, pgfautomata}
-% \usepackage{tikz}
-% \usetikzlibrary{cd}
-%% \usetikzlibrary{arrows}
-%% \usetikzlibrary{calc}
-%% \usetikzlibrary{snakes}
-%% \usetikzlibrary{backgrounds}
-% \usetikzlibrary{trees}
-%% \usetikzlibrary{automata}
-%% \usetikzlibrary{positioning}
-%% \usetikzlibrary{matrix}
-%% \usetikzlibrary{patterns}
-%% \usetikzlibrary{shapes}
-
-% symbols
-\usepackage{amsmath} 
-\usepackage{amssymb} 
-\usepackage{amsbsy}
-\usepackage{bbold}
-\usepackage{latexsym}
-%\usepackage{amsfonts}
-\usepackage{stmaryrd}
-%\usepackage{mathabx}
-%\usepackage{MnSymbol}
-\usepackage{harmony} % simple music fonts
-\usepackage{mathtools} % for arrows
-%\usepackage{mathptmx}
-
-%% theorem environments
-\usepackage{theorem}
-% \newtheorem{theorem}{Theorem} %[section]
-% \newtheorem{definition}[theorem]{Definition}
-% \newtheorem{lemma}[theorem]{Lemma}
-% \newtheorem{corollary}[theorem]{Corollary}
-% \newtheorem{proposition}[theorem]{Proposition}
-% \newenvironment{proof}{\vspace{-2ex}{\it Proof. }}{\hspace*{\fill} $\Box$\smallskip }
-% \theorembodyfont{\slshape}
-% \newtheorem{example}[theorem]{Example}
-% \newtheorem{remark}[theorem]{Remark}
-{\theorembodyfont{\rmfamily} \theoremstyle{break} \newtheorem{algo}{Algorithm}}
-
-%% algorithms
-%\usepackage{algorithm} 
-%\usepackage{program}
-
-%% for new macros
-\usepackage{xspace}
-
-%% arrows etc
-%\input{rewriting}
-
-%% Misc macros
-
-\def\ie{\textit{i.e.}\xspace}
-\def\eg{\textit{e.g.}\xspace}
-\def\wrt{\textit{wrt}\xspace}
-%\def\wlog{\textit{wlog}\xspace}
-\def\etc{\textit{etc}\xspace}
-
-\def\<#1>{\langle #1 \rangle}
-\newcommand{\pair}[2]{\langle{#1}, {#2}\rangle}
-%\newcommand{\A}{\mathcal{A}}
-%\newcommand{\B}{\mathcal{B}}
-\newcommand{\D}{\mathbb{D}}
-\newcommand{\E}{\mathbb{E}}
-\newcommand{\W}{\mathbb{W}}
-\newcommand{\T}{\mathcal{T}}
-
-\newcommand{\Semiring}{\mathbb{S}}
-\newcommand{\zero}{\mathbb{0}}
-\newcommand{\one}{\mathbb{1}}
-\newcommand{\dom}{\ensuremath{\mathit{dom}}}
-
-\def\SA{\textsf{SA}\xspace}
-\def\WA{\textsf{WA}\xspace}
-\def\SWT{\textsf{swT}\xspace}
-\def\SWA{\textsf{swA}\xspace}
-\def\SWTA{\textsf{swTA}\xspace}
-\def\SWVPA{\textsf{sw-VPA}\xspace}
-\def\weight{\mathsf{weight}}
-\def\wei{\mathsf{w}}
-\def\init{\mathsf{in}}
-\def\final{\mathsf{out}}
-\newcommand{\call}[1]{\ensuremath #1} %{\ensuremath \langle_{#1}}
-\newcommand{\return}[1]{\ensuremath #1} %{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
-\def\Omegai{{\Omega_\mathsf{i}}}
-\def\Omegac{{\Omega_\mathsf{c}}}
-\def\Omegar{{\Omega_\mathsf{r}}}
-\def\Sigmai{{\Sigma_\mathsf{i}}}
-\def\Sigmac{{\Sigma_\mathsf{c}}}
-\def\Sigmar{{\Sigma_\mathsf{r}}}
-\def\Deltai{{\Delta_\mathsf{i}}}
-\def\Deltac{{\Delta_\mathsf{c}}}
-\def\Deltar{{\Delta_\mathsf{r}}}
-\def\Phii{{\Phi_\mathsf{i}}}
-\def\Phic{{\Phi_\mathsf{c}}}
-\def\Phir{{\Phi_\mathsf{r}}}
-\def\Phicr{{\Phi_\mathsf{cr}}}
-\def\weii{{\wei_\mathsf{i}}}
-\def\weic{{\wei_\mathsf{c}}}
-\def\weir{{\wei_\mathsf{r}}}
-\def\weie{{\wei_\mathsf{e}}}
-\newcommand{\config}[2]{\ensuremath \genfrac{[}{]}{0pt}{}{#1}{#2}} 
-\newcommand{\opluseq}{\ensuremath\mathrel{\oplus}=}
-\newcommand{\ioi}[1]{\mathsf{ioi}({#1})}
-\newcommand{\rank}{\mathsf{rk}}
-\newcommand{\lin}{\mathsf{lin}}
-
-%\sloppy
-
-% Parsing over infinite alphabet as optimal alignment computation 
-% as edit-distance between string and language
+% !TEX root = main.tex
 %
-%\title{Symbolic Weighted Parsing and Automated Music Transcription}
-%\title{Symbolic Weighted Language Models and Automated Music Transcription}
-\title{Symbolic Weighted Language Models and Parsing over Infinite Alphabets}
-%\title{Weighted Visibly Pushdown Automata and Automated Music Transcription}
-\author{Florent Jacquemard}
-\institute{INRIA \& CNAM, Paris, France\\
-\email{florent.jacquemard@inria.fr}}
-
-%\titlerunning{WVPA \& AMT}
-\titlerunning{Symbolic Weighted Parsing and Automated Music Transcription}
-%\authorrunning{Florent Jacquemard}
-
-\date{\today}
-
-\begin{document}
-\thispagestyle{empty}
-\maketitle
-
-\begin{abstract}
-%Symbolic Weighted (SW) extension of symbolic automata...
+% Symbolic Weighted (SW) extension of symbolic automata...
+%
 We propose a framework for weighted parsing over infinite alphabets.
@@
 We also briefly describe the application that motivated the introduction of these models: 
-an parse-based approach to automated music transcription.
-\end{abstract}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% intro
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Introduction} \label{sec:intro}
-Parsing %variant of membership problem for formal languages
-%- given a language model and an input word $s$, compute a derivation of the model that yields $s$.
-is the problem %process 
-of structuring a linear representation in input %typically
-(a finite word over an alphabet) according to a language model. % (a formal grammar).
-% natural language, programming language, 
-%
-Most of the context-free parsing approaches~\cite{GruneJacobs08parsing}  
-assume a finite and reasonably small input alphabet. %models and algorithms 
-Such a restriction makes perfectly sense in the context of NLP tasks 
-like constituency parsing or of programming languages compilers or interpreters.
-Considering large or infinite alphabets can however be of practical interest
-in other cases.
-%
-For instance, when dealing with large characters encodings such as UTF-16, % processing strings in 
-\eg for vulnerability detection in Web-applications~\cite{dAntoni21CACM}, 
-%
-for the analyse (\eg validation or filtering) 
-of data streams or serialization of structured documents 
-(which may contain textual or numerical attributes)~\cite{Segoufin06csl}, 
-or for processing timed execution traces~\cite{Bouyer03algebraic}.
-% algebraic definition of a class of data languages 
-% (notion of monoid recognizability, based on registers, comparable to Bojancszik et al. data words)
-Regarding the latter case, we describe %briefly 
-at the end of the paper 
-a parse-based approach to automated music transcription,
-where a symbolic music performance, 
-presented as a sequence of timed musical events, % timestamped
-% a symbolic representation of a music performance 
-is converted into a structured score in Common Western Music Notation~\cite{foscarin:hal-01988990}.
-%generalizes weighted parsing: 
-%finding the best derivation for a weighted grammar.
-
-Various extensions of language models for handling infinite alphabets have been studied.
-%words carrying data values in an infinite domain (e.g. integers) e.g. data processing 
-For instance, some automata with memory extensions 
-allow restricted storage and comparison of input symbols, 
-%and correspond with logics, 
-(see~\cite{Segoufin06csl} for a survey),
-with pebbles for marking positions~\cite{NevenSchwentickVianu04FSMinfinite}, 
-registers~\cite{KaminskiFrancez94}, 
-or %computations in 2 steps
-the possibility to compute on subsequences 
-with the same attribute values~\cite{Bojanczyk11FO2}. % data words automata.
-%
-%for the and verification of infinite state systems 
-%(model checkers: alphabet = long bit-vectors)
-%...For the representation of  in model checking, verification and 
-The automata at the core of model checkers
-compute on input symbols represented by large bitvectors~\cite{Vardi07ciaa} %\cite{BaierKatoen08MC}
-(sets of assignments of Boolean variables) %propositional variables)
-and in practice,  %implementation
-each transition accepts a set of such symbols (instead of an individual symbol), 
-represented by Boolean formula or Binary Decision Diagrams.
-%
-Following a similar idea, % and generalizing,
-in symbolic automata (\SA)~\cite{dAntoniVeanes17CAV,dAntoni21CACM}, 
-the transitions are guarded by predicates over infinite alphabet domains.
-With closure conditions on the sets of such predicates, % (alphabet theories), 
-all the good properties enjoyed by automata over finite alphabets are preserved.
-
-Other extensions of language models  %(automata and grammars) 
-help in dealing with non-determinism, by the computation of weight values. 
-With an ambiguous grammar, there may exists several derivations 
-(\emph{abstract syntax trees} -- AST), % \ie the result of an analyze.
-yielding one input word. % structuring a word, 
-The association of one weight value %in semiring domain
-to each derivation permits to select a best one (or $n$ bests). % a fixed number of bests 
-% = ...a ranking of derivations 
-This is roughly the principle of \emph{weighted parsing}
-approaches~\cite{Goodman99SemiringParsing,Nederhof03weightedParsing,MorbitzVogler19weighted-parsing}.
-In \emph{weighted language models}, 
-like \eg probabilistic context-free grammars % (CFG),
-and weighted automata (\WA) \cite{Droste09handbook},
-a weight value is associated to each transition rule, % production rule
-and the rule's weights can be combined with a associative product operator~$\otimes$ into 
-the weight of a derivation.
-A second operator~$\oplus$, associative and commutative, 
-is moreover used to handle ambiguity of the model, 
-by summing the weights of the possibly several (in general exponentially many) AST %syntax trees 
-associated to a given input word.
-Typically, $\oplus$ will select the best of two weight values.
-%Intuitively,~$\oplus$ selects, or ranking, the syntax trees.
-The weight domain, equipped with these two operators is assumed, at minima, 
-to form a \emph{semiring} %$\Semiring$ 
-where $\oplus$ can be extended to infinite sums, 
-%\cite{Eilenberg74automata}
-such as the Viterbi semiring and the tropical min-plus algebra, see Figure~\ref{fig:semirings}. 
-%of domain $\mathbb{R}^+ \cup \{ +\infty\}$, 
-%where $\oplus$ is min and $\otimes$ is plus .
-%by ranking 
-%making the weight domain a semiring.
-%Some efficient specialized parsing algorithms~\cite{Huang05kbest} have been proposed in this context 
-%% models represented as hypergraphs \cite{Huang05kbest}
-%in order to compute the $n$ best syntax trees of a given input word without having to enumerate them all.
-%Generally based on dynamic programming, these algorithms rely on 
-%additional algebraic properties of~$\Semiring$.
-%-- see \eg~\cite{Huang05kbest} for some NLP applications.
-%The extraction of $n$ best list is useful 
-%the problem: quantitative parsing or symbolic parsing
-%of parsing of words over infinite input alphabet.
-
-In this paper, we present a uniform framework for weighted parsing over infinite input alphabets.
-It is based on weighted %finite states 
-language models generalizing 
-both~\SA, with functions in an arbitrary semiring instead of Boolean guards, 
-and~\WA, by handling infinite alphabets -- Figure~\ref{fig:hierarchy}.
-%***
-%some weighted language models
-%model of weighted CFG 
-%computing on words over infinite input alphabets %sets of terminal symbol
-%- but with finite sets of states and transitions rules. 
-% non-terminals and production rules).
-In their transition rules, input symbols appear as variables %(parameters)
-and the weight associated to a transition rule is a function of these variables.
-%
-\begin{figure}
-\centering
-\begin{tikzpicture}
-\node (NFA) at (0,4) {%
-  \(
-  \begin{array}{c} 
-  \mathsf{NFA} : \Sigma_{\mathsf{fin}}^* \to \mathbb{B}\\
-  q \xrightarrow{a} q' \quad  a \in \Sigma_\mathsf{fin}
-  \end{array} 
-  \)
-};
-%
-\node (WA) at (-2,2) {%
-  \(
-  \begin{array}{c} 
-  \mathsf{WA} : \Sigma_{\mathsf{fin}}^* \to \mathbb{S}\\
-  q \xrightarrow{a, w} q' \quad  a \in \Sigma_\mathsf{fin}, w \in \mathbb{S}
-  \end{array} 
-  \)
-};
-%
-\node (SA) at (2,2) {%
-  \(
-  \begin{array}{c} 
-  \mathsf{SA} : \Sigma_{\mathsf{inf}}^* \to \mathbb{B}\\
-  q \xrightarrow{\phi} q' \quad \phi : \Sigma_\mathsf{inf} \to \mathbb{B}
-  \end{array} 
-  \)
-};
-%
-\node (SWA) at (0,0) {%
-  \(
-  \begin{array}{c} 
-  \mathsf{SWA} : \Sigma_{\mathsf{inf}}^* \to \mathbb{S}\\
-  q \xrightarrow{\phi} q' \quad \phi : \Sigma_\mathsf{inf} \to \mathbb{S}
-  \end{array} 
-  \)
-};
-\draw[->] (NFA)--(WA);
-\draw[->] (NFA)--(SA);
-\draw[->] (WA)--(SWA);
-\draw[->] (SA)--(SWA);
-%\begin{array}{c} \mathsf{NFA} : \Sigma^* \to \mathbb{B} \end{array} 
-\end{tikzpicture}
-\caption{Classes of Symbolic/Weighted Automata. 
-$\Sigma_\mathsf{fin}$ is a finite alphabet, 
-$\Sigma_\mathsf{inf}$ is a countable alphabet,
-$\mathbb{B}$ is the Boolean algebra, 
-$\mathbb{S}$ is an arbitrary commutative semiring, 
-$q \xrightarrow{\dots} q'$ represents the form of a transition between states $q$ and $q'$.}
-\label{fig:hierarchy}  
-\end{figure}
-%
-%This approach is close to the case of 
-%Symbolic Automata (SA)~\cite{dAntoniVeanes17CAV,dAntoni21CACM}, 
-%except that the domain for weight values is not restricted to be Boolean, 
-%like for the guards in the rules of SA, 
-%but can be an arbitrary commutative semiring (assuming some restrictions).
-%
-The models presented here are finite automata called symbolic-weighted (\SWA),
-transducers (\SWT) and pushdown automata,
-with a visibly restriction~\cite{AlurMadhusudan09nested} (\SWVPA).
-%\wrt an edit distance
-The latter model of automata computes on \emph{nested words}~\cite{AlurMadhusudan09nested}, 
-a structured form of words parenthesized with markup symbols, 
-corresponding to a linearization of trees.
-In the context of parsing, they are used here to represent (weighted) AST of CF grammars.
-More precisely, a \SWVPA $A$ associates a weight value $A(t)$ % \in \Semiring$ 
-to a given a nested word $t$, which is the linearization of an AST. %representing a parse tree.
-%
-On the other hand, 
-a \SWT is used to define a distance $T(s, t)$ between two finite words $s$ and $t$
-over an infinite alphabet, following~\cite{Mohri03EDWA}.
-Then, the \emph{SW-parsing} problem aims at %computing 
-finding $t$ minimizing 
-$T(s, t) \otimes A(t)$ (\wrt the ranking defined by $\oplus$)
--- this value is called the distance between $s$ and $A$ in~\cite{Mohri03EDWA}.
-%
-Similarly to weighted-parsing 
-methods~\cite{Goodman99SemiringParsing,Nederhof03weightedParsing,MorbitzVogler19weighted-parsing}, 
-our approach proceeds in two steps, 
-based on properties of the SW models. 
-The first step is a Bar-Hillel construction where, 
-given a \SWT $T$, a \SWVPA $A$, and an input word $s$, 
-a \SWVPA $A_{T, s}$ is built, such that for all $t$, $A_{T, s}(t) = T(s, t) \otimes A(t)$.
-In the second step, a best AST $t$ is found by applying to $A_{T, s}$ 
-a best search algorithm similar to shortest distance in graphs~\cite{Mohri02semiring,Huang05kbest}.
-%In expressiveness, they are equivalent to weighted CFG. 
-%and can be used in a general approach for parsing over infinite input alphabets.
-%
-%Let $A$ be a \SWVPA, associating $A(w) \in \Semiring$ 
-%to a given a nested word $w$ (representing a parse tree),
-%and let a \SWT compute a distance $d$, in $\Semiring$, 
-%between 2 strings over respectively an infinite input alphabet and the 
-%same (infinite) alphabet of $A$.
-%Then, the problem of Symbolic Weighted Parsing is, 
-%given an input string $s$, to find a nested word $w$ minimizing 
-%(according to the ranking defined by $\oplus$)
-%the distance $d(s, w) \otimes A(w)$ between $s$ and $A$, 
-%as defined in~\cite{Mohri03EDWA}.
-
-% First one general edit-distance is defined by a weighted word 
-% transducer~\cite{Mohri}
-% %Symbolic automata transducers are extended models~\cite{VeanesdAntoniJACM}
-% %dealing with infinite set of input symbols...
-% value in a semiring...
-
-The main contributions of the paper are: 
-($i$)~the introduction of automata, \SWA, transducers, \SWT (Section~\ref{section:SWA}),
-and visibly pushdown automata \SWVPA (Section~\ref{sec:SWVPA}),
-generalizing the corresponding classes of symbolic and weighted models, 
-%for the weighted computation on (nested) words over infinite alphabets;  
-($ii$)~a polynomial best-search algorithm for \SWVPA, %(Section~\ref{sec:best})
-% a framework for parsing over infinite alphabets, 
-and ($iii$)~a uniform framework (Section~\ref{sec:parsing}) for parsing over infinite alphabets, 
-the keys to which are 
-($iii.a$)~the \SWT-based definition of generic edit distances between input and output words,
-and ($iii.b$)~the use of nested words, and \SWVPA, 
-instead of syntax trees and grammars. %(\S~\ref{sec:trees}).
-%
-Finally, Section~\ref{sec:transcription} describes
-the implemented application %of this approach 
-to automated music transcription
-that motivated this study.
-
-%based these models and on generic def. of distances;
-%the use of VPA for parsing, word instead of trees, use comparison input-output.
-%and best-search
-
-%In Section~\ref{section:SWA} we introduce \SWA and \SWT.
-%Then \SWVPA are defined in Section~\ref{sec:SWVPA}, 
-%where a polynomial 1-best algorithm is described that can be use to solve
-%Symbolic Weighted Parsing, as described Section~\ref{sec:parsing}.
-%Finally, in Section~\ref{sec:transcription}, we present an application 
-%of this approach to automated music transcription that has been implemented.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SWT & SWA
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{SW Automata and Transducers}
-\label{section:transducer}\label{sec:transducer}
-\label{section:SWA}\label{sec:SWA}
-\label{section:SWT}\label{sec:SWT}
-
-We follow the approach of~\cite{Mohri03EDWA} for the computation of distances
-between words and languages, with weighted transducers, 
-and extending it to infinite alphabets.
-% with models of symbolic weighted automata and transducers. 
-%
-The models introduced in this section generalize 
-weighted automata and transducers~\cite{Droste09handbook} 
-%over finite alphabets, see  e.g.~\cite{Mohri03EDWA}, 
-by labeling each transition with a weight function that takes the 
-input and output symbols as parameters, instead of a simple weight value.
-These functions are similar to the guards of symbolic automata~\cite{dAntoniVeanes17CAV,dAntoni21CACM},
-but they can return values in an arbitrary semiring, 
-where the latter guards are restricted to the Boolean semiring.
-
-\subsection{Semirings}
-\label{section:semiring}
-We shall consider semiring domains for the weight values of our language models.
-%
-A \emph{semiring} $\< \Semiring, \oplus, \zero, \otimes, \one>$ 
-is a structure with a domain~$\Semiring$,
-equipped with two associative
-binary operators~$\oplus$ and $\otimes$
-with respective neutral elements $\zero$ and $\one$ and such that:
-%$\< \mathbb{S}, \oplus, \zero>$ is a commutative monoid
-%$\< \mathbb{S}, \otimes, \one>$ is a monoid
-\begin{itemize}
-\item $\oplus$ is commutative 
--- therefore $\< \Semiring, \oplus, \zero>$ is a commutative monoid 
-   and $\< \Semiring, \otimes, \one>$ is a monoid,
-\item $\otimes$ distributes over~$\oplus$:  $\forall x, y, z \in \mathbb{S}$,
-$x \otimes (y \oplus z) = (x \otimes y) \oplus (x \otimes z)$, 
-and $(x \oplus y) \otimes z = (x \otimes z) \oplus (y \otimes z)$,
-\item $\zero$ is absorbing for~$\otimes$: 
-$\forall x\in \mathbb{S}$, $\zero \otimes x = x \otimes \zero = \zero$.
-\end{itemize}
-%Components of a semiring~$\Semiring$ may be subscripted by~$\Semiring$ when needed.
-%We simply write $x \in \Semiring$ to mean $x \in \mathbb{S}$.
-%
-Intuitively, in the models presented in this paper, 
-$\oplus$ selects an optimal value from two given values, 
-in order to handle non-determinism, 
-and $\otimes$ combines two values into a single value, 
-in order to describe the chaining of several transitions.
-%and let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a {semiring},
-
-\medskip%\noindent
-A semiring $\Semiring$ is \emph{commutative} if $\otimes$ is commutative.
-It is \emph{idempotent} if for each $x \in \dom(\Semiring)$, $x \oplus x = x$.
-%
-Every idempotent semiring~$\Semiring$ induces 
-a partial ordering~$\leq_\oplus$ 
-called the \emph{natural ordering} of~$\Semiring$~\cite{Mohri02semiring} 
-and defined,  by: 
-%implicitly defined by the semiring $\Semiring$ 
-for all $x$ and $y$,
-$x \leq_\oplus y \;\mbox{iff}\; x \oplus y = x$.
-%(see~\cite{Mohri02semiring} for the proof that it is an ordering).
-%
-The natural ordering is sometimes defined in the opposite direction~\cite{DrosteKuich09semirings};
-We follow here the direction  %follows \cite{Mohri02semiring}, and 
-that coincides than the usual ordering on the Tropical semiring (\emph{min-plus}).
-
-\begin{lemma}[Monotony, \cite{Mohri02semiring}]
-Let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be an idempotent semiring.
-For all $x, y, z  \in \Semiring$,  
-if $x \leq_\oplus y$ then
-$x \oplus z \leq_\oplus y \oplus z$,
-$x \otimes z \leq_\oplus y \otimes z$
-and $z \otimes x \leq_\oplus z \otimes y$.
-\end{lemma}      
-We say in this case that $\Semiring$ is \emph{monotonic}. % \wrt $\leq_\oplus$.
-%A semiring $\Semiring$ 
-%is \emph{monotonic} \wrt a partial ordering~$\leq$ 
-%iff for all $x, y, z  \in \Semiring$,  $x \leq y$ implies
-%$x \oplus z \leq y \oplus z$,
-%$x \otimes z \leq y \otimes z$,
-%and $z \otimes x \leq z \otimes y$.
-%
-Another important semiring property in the context of optimization
-is {superiority}~\cite{Huang08advanceddynamic}, 
-which corresponds to the 
-\emph{non-negative weights} condition in shortest-path algorithms~\cite{Dijkstra59anote}.
-Intuitively, it means that combining elements with $\otimes$ always increase their weight. 
-Formally, it is defined as the property of the following lemma.
-
-\begin{lemma}[Superiority, Boundedness]
-\label{lem:superior}\label{lem:bounded}
-Let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be an idempotent semiring.
-The two following statements are equivalent:
-\begin{itemize}
-\item [$i.$] for all $x, y \in \Semiring$,  
-$x \leq_\oplus x \otimes y$ and 
-$y \leq_\oplus x \otimes y$
-\item[$ii.$] for all $x \in \Semiring$,  $\one \oplus x = \one$.
-\end{itemize}
-\end{lemma}
-%
-In~\cite{Huang08advanceddynamic}, the property~$i.$ is called 
-$\Semiring$ is \emph{superior} \wrt the ordering~$\leq_\oplus$.
-It implies that 
-%It holds that If $\Semiring$ is superior \wrt~$\leq$, then %$\one \leq \zero$
-for all $x \in \Semiring$, $\one \leq_\oplus x \leq_\oplus \zero$.
-Indeed, by the second inequality of $(i.)$ with $y = \one$, $\one \leq_\oplus \one \otimes x = x$, 
-and, by the first inequality of $(i.)$ with $y = \zero$,  
-$x \leq_\oplus x \otimes \zero = \zero$.
-%
-Intuitively, from an optimization point of view,
-it means that $\one$ is the best value, and $\zero$ the worst.
-%** superior implies $\Semiring$ bounded~\cite{Mohri02semiring} see **
-%
-In~\cite{Mohri02semiring}, the property of Lemma~\ref{lem:bounded}(in the form $ii.$) 
-is called \emph{boundedness} of $\Semiring$ -- we shall use this term in the rest of the paper. 
-% \emph{negative boundedness} 
-It implies that, when looking for a best path in a graph whose edges
-are weighted by values of $\Semiring$, the loops can be safely avoided.
-
-\begin{proof}(Lemma~\ref{lem:bounded})
-For $(ii.) \Rightarrow (i.)$ : 
-$x \oplus (x \otimes y) = x \otimes (\one \oplus y) = x$, 
-by distributivity of $\otimes$ over $\oplus$. 
-Hence $x \leq_\oplus x \otimes y$.
-Similarly, $y \oplus (x \otimes y) = (\one \oplus x) \otimes y = y$, 
-hence $y \leq_\oplus x \otimes y$.
-%
-For $(i.) \Rightarrow (ii.)$ : as we have seen, $(i.)$ implies that 
-$\one \leq_\oplus x$ for all $x \in \Semiring$, 
-\ie, by definition of $\leq_\oplus$, $\one \oplus x = \one$.
-\qed
-\end{proof}
-
-%Following the terminology of~\cite{Mohri02semiring},
-%when $\forall x \in \dom(\Semiring), \one \oplus x = \one$,
-%the semiring $\Semiring$ is is called \emph{bounded}.
-
-\begin{lemma}
-Every bounded semiring is idempotent.
-\end{lemma}
-\begin{proof}
-By boundedness, $\one \oplus \one = \one$, 
-and idempotency follows by multiplying
-both sides by $x$ and distributing. 
-\qed
-\end{proof}
-
-\noindent
-An idempotent semiring $\Semiring$~is called \emph{total} if
-it~$\leq_\oplus$ is total
-\ie when for all $x, y \in \Semiring$, either $x \oplus y = x$ or $x \oplus y = y$.
-
-%\medskip
-We shall need below infinite sums with~$\oplus$.
-A semiring~$\Semiring$ is called \emph{complete}~\cite{Kuich97semirings}
-if for every family
-$(x_i)_{i \in I}$ %$\{ x_i \mid i \in I \}$
-of elements of $\dom(\Semiring)$ over an index set $I \subset \mathbb{N}$,
-the infinite sum $\bigoplus_{i \in I} x_i$
-is well-defined and in $\dom(\Semiring)$,
-and the following properties hold:
-\begin{description}
-\item[$i.$]
-\emph{infinite sums extend finite sums:}
-$\displaystyle\bigoplus_{i \in \emptyset} x_i = \zero$,\quad 
-      $\forall j\in \mathbb{N}, \displaystyle\bigoplus_{i \in \{ j \}} x_i = x_j$,\quad
-      $\forall j, k\in \mathbb{N}, j\neq k, 
-      \displaystyle\bigoplus_{i \in \{ j, k \}} x_i = x_j \oplus x_k$,
-%
-\item[$ii.$]
-\emph{associativity and commutativity:}
-for all $I \subseteq \mathbb{N}$
-and all partition $(I_{j})_{j \in J}$ of $I$, %\subseteq \mathbb{N}$, 
-\(
-\displaystyle
-\bigoplus_{j \in J}\bigoplus_{i \in I_j} x_i = 
-\bigoplus_{i \in I} x_i
-\),
-%
-\item[$iii.$] 
-\emph{distributivity of product over infinite sum:}\\
-for all $I \subseteq \mathbb{N}$,
-\(
-\displaystyle
-\bigoplus_{i \in I} (x \otimes y_i) = x \otimes \bigoplus_{i\in I} y_i\), and
-\(
-\displaystyle
-\bigoplus_{i \in I} (x_i \otimes y) = (\bigoplus_{i \in I} x_i ) \otimes y\).
-\end{description}
-
-\begin{example}
-Figure~\ref{gif:semirings} presented examples of semirings interesting in practice 
-and enjoying the above properties.
-\end{example}
-
-\begin{figure}[t]
-\begin{center}
-\begin{tabular}{|c|c|C{2em}|C{2em}|C{2em}|C{2em}|}
-\hline
-        & domain & $\oplus$ & $\otimes$ & $\zero$  & $\one$\\ %& nat. ordering\\
-\hline\hline
-Boolean & $\{\bot, \top\}$ & $\vee$ & $\wedge$ & $\bot$ & $\top$\\ %& $\top \leq_\oplus \bot$  \\
-\hline
-Counting & $\mathbb{N}$ & $+$ & $\times$ & 0 & 1 \\
-\hline
-Viterbi & $[0, 1] \subset \mathbb{R}$ & $\mathit{max}$ & $\times$ & 0 & 1\\ % & $x \leq_\oplus y \iff x \ge y$  \\
-\hline
-Tropical min-plus & $\mathbb{R}_+ \cup \{ +\infty\}$ & $\mathit{min}$ & $+$ & $+\infty$ & $0$\\ % & $x \leq_\oplus y \iff x \leq y$   \\
-\hline
-%MaxPlus & $\mathbb{Q} \cup \{ -\infty\}$ & $\mathsf{max}$ & $+$ & $-\infty$ & $0$ \\
-%\hline
-%Word lang. & $2^{\Sigma^*}$ & $\cup$ & $\cdot$ & $\emptyset$ & $\{ \varepsilon \}$ \\
-%\hline
-\end{tabular}
-\end{center}
-\vskip-1em
-\caption{Some commutative, bounded, total and complete semirings.}
-\label{fig:semirings}
-\end{figure}
-
-\subsection{Label Theory}
-\label{section:symbols}
-We shall now define the functions labeling the transitions of SW-automata and -transducers.
-Let $\Sigma$ and $\Delta$ be two countable sets of symbols
-called respectively \emph{input} and \emph{output} \emph{alphabets}, 
-and let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative semiring.
-
-\noindent 
-A \emph{label theory} over $\Sigma$ and $\Delta$
-is made of 3 recursively enumerable sets:
-%$\Phi_\epsilon \subseteq \Semiring$, % containing constant functions valued in $\Semiring$, 
-$\Phi_\Sigma$ and $\Phi_\Delta$, 
-containing unary functions in $\Sigma \to \Semiring$, resp. $\Delta \to \Semiring$, 
-and $\Phi_{\Sigma, \Delta}$  containing binary functions in $\Sigma \times \Delta \to \Semiring$, 
-such that 
-for all $\alpha \in \Semiring$, and $\phi \in \Phi_\Sigma$,
-$\alpha \otimes \phi : x \mapsto \alpha \otimes \phi(x)$, 
-and $\phi \otimes \alpha : x \mapsto \phi(x) \otimes \alpha$
-belong to $\Phi_\Sigma$, and similarly for $\oplus$ 
-and for $\Phi_\Delta$ and $\Phi_{\Sigma, \Delta}$.
-%
-Moreover, these sets are required to be closed under the following 
-operators~$\oplus$ and $\otimes$ of~$\Semiring$:
-for all $\phi, \phi' \in \Phi_\Sigma$,
-all $\psi, \psi' \in \Phi_\Delta$, 
-and $\eta, \eta' \in \Phi_{\Sigma, \Delta}$, the function
-\begin{description}
-\item $\phi \otimes \phi' : x \mapsto \phi(x) \otimes \phi'(x)$ belongs to $\Phi_\Sigma$, 
-\item $\psi \otimes \psi' : y \mapsto \psi(y) \otimes \psi'(y)$ belongs to $\Phi_\Delta$,
-\item $\phi \otimes \eta : x, y \mapsto \phi(x) \otimes \eta(x, y)$ belongs to $\Phi_{\Sigma, \Delta}$,
-\item $\eta \otimes \psi : x, y \mapsto \eta(x, y) \otimes \psi(y)$ belongs to $\Phi_{\Sigma, \Delta}$,
-\item $\eta \otimes \eta' : x, y \mapsto \eta(x, y) \otimes \eta'(x, y)$ belongs to $\Phi_{\Sigma, \Delta}$.
-\end{description}
-And the same also holds for the binary sum operator $\oplus$.
-
-\noindent
-%\marginpar{necessary?}
-Finally, it is also required 
-% that the codomain of every function of $\Phi_\Sigma$ and $\Phi_\Delta$ 
-% is a subset of $\Phi_\epsilon$, and
-that all partial applications of functions $\Phi_{\Sigma, \Delta}$, 
-resp.  $f_a: y \mapsto f(a, y)$ for $a \in \Sigma$ and $y \in \Delta$
-and  $f_b: x \mapsto f(x, b)$ for $b \in \Delta$ and $x \in \Sigma$, 
-belong resp. to~$\Phi_\Sigma$ and~$\Phi_\Delta$.
-
-\subsection{Definitions} \label{sec:SWTdef}\label{sec:SWAdef}
-%(SWT)
-\begin{definition}
-\label{def:transducer} \label{def:SWT}
-A \emph{symbolic-weighted transducer} (\SWT)
-over the input and output alphabets~$\Sigma$ and $\Delta$ 
-and the semiring $\Semiring$ is a tuple
-$T = \< Q, \init, \bar{\wei}, \final >$,
-where $Q$ is a finite set of states, 
-$\mathsf{in} : Q \to \Semiring$, 
-respectively $\mathsf{out} : Q \to \Semiring,$
-are functions defining the weight for entering, 
-respectively leaving, a state, 
-and $\bar{\wei}$ is a tuplet made of the 4 of transition functions 
-$\wei_\epsilon$, $\wei_\Sigma$, $\wei_\Delta$, and $\wei_{\Sigma, \Delta}$
-from $Q \times Q$ into respectively~$\Semiring$, %$\Phi_\epsilon$, 
-$\Phi_\Sigma$, $\Phi_\Delta$, and $\Phi_{\Sigma, \Delta}$.
-\end{definition}
-%
-We summarize the above 4-uplet of transition functions into a unique 
-function~$\wei$ from
-$Q \times (\Sigma \cup \{ \epsilon \}) \times (\Delta \cup \{ \epsilon \}) \times Q$
-into $\Semiring$, %also called $\wei$ for simplicity, 
-such that, for all $q, q' \in Q$, $a \in \Sigma$, and $b \in \Delta$, 
-%and with 
-%$\< \phi_\epsilon, \phi_\Sigma, \phi_\Delta, \phi_{\Sigma, \Delta}> = \wei(q, q')$, 
-\[
-\begin{array}{rcll}
-\wei(q, \epsilon, \epsilon, q') & = &  \wei_\epsilon(q, q'),\\ %\phi_\epsilon\\
-\wei(q, a, \epsilon, q') & = & \phi_\Sigma(a) & %\wei_\Sigma(q, q')(a)
-\quad\mathrm{where~} \phi_\Sigma = \wei_\Sigma(q, q'),\\
-\wei(q, \epsilon, b, q') & = & \phi_\Delta(b) & 
-\quad\mathrm{where~} \phi_\Delta = \wei_\Delta(q, q'),\\
-\wei(q, a, b, q') & = & \phi_{\Sigma, \Delta}(a, b) & 
-\quad\mathrm{where~} \phi_{\Sigma, \Delta} = \wei_{\Sigma, \Delta}(q, q').\\
-\end{array}      
-\]
-%More precisely, $Q \times Q$, 
-%resp. $Q \times \Sigma \times Q$,
-%$Q \times \Delta \times Q$,
-%$Q \times \Sigma \times \Delta \times Q$,
-%into resp. $\Phi_0$, $\Phi_\Sigma$, $\Phi_\Delta$ $\Phi_{\Sigma, \Delta}$.
-
-The symbolic-weighted transducer $T$ defines a mapping 
-from the pairs of strings of $\Sigma^* \times \Delta^*$ 
-into the weights of~$\Semiring$,
-based on the following intermediate function $\weight_T$
-defined recursively for every $q, q' \in Q$, 
-%$a \in \Sigma$, $b \in \Delta$ 
-for every strings of $s \in \Sigma^*$, $t \in \Delta^*$:
-\[
-\begin{array}{rccl}
-xc(q, s, t, q') & = & 
-   \displaystyle\bigoplus_{q'' \in Q} &
-   \wei(q, \epsilon, \epsilon, q'') \otimes \weight_T(q'', s, t, q')\\
-   & \oplus & \displaystyle\bigoplus_{q'' \in Q} &
-              \wei(q'', \epsilon, \epsilon, q') \otimes \weight_T(q, s, t, q'')\\
- & \oplus & \displaystyle\bigoplus_{\begin{array}{c}
-                                      \scriptstyle q'' \in Q\\[-2pt]
-                                      \scriptstyle s = au, a \in \Sigma
-                                      \end{array}} &
-   \wei(q, a, \epsilon, q'') \otimes \weight_T(q'', u, t, q')\\
- & \oplus & \displaystyle\bigoplus_{\begin{array}{c}
-                                    \scriptstyle q'' \in Q\\[-2pt]
-                                    \scriptstyle t = bv, b \in \Delta
-                                    \end{array}} &
-    \wei(q, \epsilon, b, q'') \otimes \weight_T(q'', s, v, q')\\ 
- & \oplus & \displaystyle\bigoplus_{\begin{array}{c}
-                                    \scriptstyle q'' \in Q\\[-2pt]
-                                    \scriptstyle s = au, a \in \Sigma\\[-2pt]
-                                    \scriptstyle t = bv, b \in \Delta
-                                    \end{array}} &
-   \wei(q, a, b, q'') \otimes \weight_T(q'', u, v, q').\\ 
-\end{array}
-\]
-The sum above can be infinite because of the expression in the first line, 
-however, $\T(\Omega)$ is well defined since $\Semiring$ is complete.
-%
-As we shall see, epsilon-transitions can be removed and hence the sum made finite.
-%
-%** The application of $\oplus$ can be seen as a ND choice between transitions **
-We recall that, by convention, an empty sum with $\oplus$ is $\zero$. 
-%
-Since $\zero$ is absorbing for~$\otimes$ in~$\Semiring$,
-one term $\wei(q, a, b, q'')$ equal to $\zero$ in the above expression 
-will be ignored in the sum, meaning that there is no possible transition
-from state $q$ into state $q'$ while reading $a$ and writing $b$ 
-(including the case $a = \epsilon$ and $a = \epsilon$).
-%These functions $\phi$ act as guards for the transducer's transitions, 
-%preventing a transition when they return the absorbing $\zero$ of $\Semiring$.
-This is analogous to a transition's guard not satisfied by $\<a, b>$ in 
-the case of symbolic transducers.
-
-The above expression of $\weight_T$ can be seen as a stateful definition of 
-an edit-distance between a word $s \in \Sigma^*$ and a word $t \in \Delta^*$,
-see also~\cite{Mohri03ijfcs}.
-Intuitively, 
-$\wei(q, a, \epsilon, q'')$, in the third line, is the cost of 
-the deletion of a symbol~$a \in \Sigma$ in~$s$, 
-$\wei(q, \epsilon, b, q'')$, in the fourth line, is the cost 
-of the deletion of insertion of~$b \in \Delta$ in~$s$, 
-and $\wei(q, a, b, q'')$, in the last line, is the cost 
-of the substitution of  $a \in \Sigma$ by~$b \in \Delta$.
-The cost of a sequence of such operations transforming $s$ into $t$, 
-is the product, with $\otimes$, of the individual costs of the operations involved;
-And the distance between $s$ and $t$ is the sum, with $\oplus$,
-of all such product of costs.
-
-\medskip\noindent
-The weight associated by $T$ to  $\< s, t> \in \Sigma^* \times \Delta^*$
-is then defined as follows: 
-\begin{equation}
-T(s, t)  = 
-\displaystyle\bigoplus_{q, q' \in Q} \mathsf{in}(q) 
-\mathop{\otimes} \weight_T(q, s, t, q') \mathop{\otimes} \mathsf{out}(q').
-\label{eq:weightT}
-\end{equation}
-
-\begin{example}
-DTW for sequences of timestamped events **     
-\end{example}
-
-\begin{example}
-(simpler) pointwise distance between two sequences of timestamped events **
-\end{example}
-
-\noindent
-The \emph{Symbolic Weighted Automata} %$A = \< Q, \init, \weight, \final >$
-over $\Sigma$ and $\Semiring$ 
-are defined similarly as the transducers of Definition~\ref{def:SWT}, 
-by simply omitting the output symbols.
-%
-In this case, a label theory over $\Sigma$ is reduced to
-a set $\Phi_\Sigma$ closed under~$\oplus$ and~$\otimes$.
-%
-\begin{definition} \label{def:SWA}
-A \emph{symbolic-weighted automaton} (\SWA)
-over the input alphabet~$\Sigma$
-and the commutative semiring $\Semiring$ is a tuple
-$A = \< Q, \init, \bar{\wei}, \final >$,
-where $Q$ is a finite set of states, 
-$\mathsf{in} : Q \to \Semiring$, 
-respectively $\mathsf{out} : Q \to \Semiring,$
-are functions defining the weight for entering, 
-respectively leaving, a state, 
-and $\bar{\wei}$ is a pair of transition functions 
-$\wei_\epsilon$ and $\wei_\Sigma$ from $Q \times Q$ into 
-respectively $\Semiring$ %$\Phi_\epsilon$ 
-and $\Phi_\Sigma$.
-\end{definition}
-
-\noindent
-The above transition functions $\wei_\epsilon$ and $\wei_\Sigma$
-are synthesized into a function~$\wei$ 
-from~$Q \times (\Sigma \cup \{ \epsilon \}) \times Q$ into~$\Semiring$ 
-as above.
-%
-When $\wei_\epsilon(q, q') = \zero$ for all $q, q' \in Q$, 
-the automaton~$A$ is called \emph{without $\epsilon$-transitions}.
-
-\subsection{Properties}
-The two following properties will be useful to our approach on 
-symbolic weighted parsing in Section~\ref{sec:parsing}.
-
-\begin{proposition}
-Given a \SWT $T$ 
-over alphabet $\Sigma$, $\Delta$ and a commutative semiring~$\Semiring$, 
-and $s \in \Sigma^*$, 
-there exists an effectively constructible \SWA 
-$A_{T, s}$ over $\Delta$ and $\Semiring$ 
-such that for all $t \in \Delta^*$, $A_{T, s}(t) = T(s, t)$.
-\end{proposition}
-%
-\begin{proof}
-Let $T = \< Q, \init, \bar{\wei}, \final >$ 
-where $\bar{\wei} \< \wei_\epsilon, \wei_\Sigma, \wei_\Delta, \wei_{\Sigma, \Delta}>$
-has an associated synthesized form $\wei$, 
-and let $s = s_1 \ldots s_n$ with $s_1, \ldots, s_n \in \Sigma$.
-%$n = |s|$ be the length of $s$.
-%
-Let $Q' = [1..n] \times Q$ be the state set of $A_{T, s}$,
-and let, for all $q \in Q$, 
-$\init'(1, q) = \init(q)$ and $\init'(i, q) = \zero$ for all $1 < i \leq n$, 
-and $\final'(n, q) = \final(q)$ and $\final'(i, q) = \zero$ for all $1 \leq i <  n$.
-
-\noindent
-We define the functions $\wei'_\epsilon$ and $\wei'_\Sigma$ for $A_{T, s}$
-in their synthesized form: for all $1 \leq i, j \leq n$, $q, q' \in Q$ and $a \in \Sigma$, 
-and $b \in \Delta$, 
-\[
-\begin{array}{lclcl}
-\wei'(\< i, q>, \epsilon, \< i, q'>)   & = & \wei(q, \epsilon, \epsilon, q'), & \quad & \\ 
-\wei'(\< i, q>, \epsilon, \< i+1, q'>) & = & \wei(q, s_i, \epsilon, q') & & \mathrm{if~} i < n,\\ 
-\wei'(\< i, q>, \epsilon, \< j, q'>)   & = & \zero & & \mathrm{if~} j \neq i, j \neq i+1,\\ 
-\wei'(\< i, q>, b, \< i, q'>)          & = & \wei(q, \epsilon, b, q'),\\ 
-\wei'(\< i, q>, b, \< i+1, q'>)        & = & \wei(q, s_i, b, q') & & \mathrm{if~} i < n,\\ 
-\wei'(\< i, q>, b, \< j, q'>)          & = & \zero & & \mathrm{if~} j \neq i, j \neq i+1. 
-\end{array}      
-\]
-Note that the functions defined above belong to the set $\Phi_\Delta$ 
-of the underlying label theory, 
-by the hypothesis of closure under partial application.
-%
-The $\SWA$ wanted is $A_{T, s} = \< Q', \init', \bar{\wei'}, \final' >$.
-\qed
-\end{proof}
-
-The construction time and size for $A_{T, s}$ are $O(\| T \| . | s |)$,
-where the size $\| T \|$ of $T$ is its number of states $|Q|$.
-
-% gen. to transducers
-\begin{proposition}
-Given a \SWA $A$ over alphabet $\Sigma$ 
-and a commutative and bounded semiring $\Semiring$, 
-there exists an effectively constructible \SWA $A'$ 
-without $\epsilon$-transitions 
-such that for all $s \in \Sigma^*$, $A'(s) = A(s)$.
-\end{proposition}
-%
-\begin{proof}  ** Tb revised **   
-Let $A = \< Q, \init, \<\wei_\epsilon, \wei_\Sigma>, \final >$.
-The automaton $A'$ is $\< Q, \init, \<\wei'_\epsilon, \wei'_\Sigma>, \final >$,
-where, for all $q, q' \in Q$,
-$\wei'_\epsilon(q, q') = \zero$  
-and 
-% \[
-% \begin{array}{rcccl}
-% \wei'_\Sigma(q, q') & = &  & \displaystyle\bigoplus_{q_1 \in Q} &
-%                \wei_\Sigma(q, q_1) \otimes  \wei_\epsilon(q_1, q')\\
-%                     &   & \oplus  & \displaystyle\bigoplus_{q_2 \in Q} &
-%                \wei_\epsilon(q, q_2) \otimes \wei_\Sigma(q_2, q')\\
-% \end{array}
-% \]
-\[
-\wei'_\Sigma(q, q') = 
-\displaystyle\bigoplus_{q'' \in Q} \wei_\epsilon(q, q'') \otimes \wei_\Sigma(q'', q')
-\oplus \displaystyle\bigoplus_{q_1 \in Q} \wei_\Sigma(q, q'') \otimes \wei_\epsilon(q'', q')
-\]
-In the above definition of $\wei'_\Sigma$ we use the operator
-of product of function of $\Phi_\Sigma$ by $\Semiring$ 
-described in Section~\ref{section:symbols}.
-%
-By definition of $\weight_{A}$ and distributivity of $\otimes$ on $\oplus$, 
-** NO. TBC see \cite{Mohri02ijfcs} **
-it holds that $\weight_{A}(q, s, q') = \weight_{A'}(q, s, q')$.
-\qed
-\end{proof}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SW-VPA
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{SW Visibly Pushdown Automata}
-\label{section:SWVPA}\label{sec:SWVPA}
-The following model generalizes Symbolic VPA~\cite{dAntonyAlur14SVPDA}
-from Boolean semirings to arbitrary semiring weight domains.
-
-\subsection{Definition}
-Let $\Omega$ be a countable alphabet 
-%finite (large) or infinite,
-that we assume partitioned into three subsets 
-$\Omegai$, $\Omegac$, $\Omegac$,
-whose elements are respectively called 
-\emph{internal}, \emph{call} and \emph{return} symbols.
-% \begin{itemize}
-% \item a set $\Omegai$ of \emph{internal symbols} denoted $a$,
-% \item a set $\Omegac$ of \emph{call symbols} denoted $\call{a}$,
-% \item a set $\Omegar$ of \emph{return symbols} denoted $\return{a}$.
-% \end{itemize}
-Let~$\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative semiring and  
-let  $(\Phi_\epsilon, \Phic, \Phir, \Phicr)$ be a label theory over $\Semiring$
-%In order to simplify notations, %and following the definition of Section~\ref{section:transducer}, 
-%we shall write respectively 
-where $\Phic$, $\Phir$ and~$\Phicr$ stand respectively 
-for~$\Phi_\Omegac$, $\Phi_\Omegar$ and~$\Phi_{\Omegac, \Omegar}$.
-%
-Moreover, we extend this theory with a set $\Phii$ 
-of unary functions in $\Omegai \to \Semiring$,
-closed under $\oplus$ and $\otimes$.
-
-\begin{definition}
-A \emph{Symbolic Weighted Visibly Pushdown Automata} (\SWVPA) 
-over  $\Omega = \Omegai \uplus \Omegac \uplus \Omegar$ and $\Semiring$ is a tuple
-$A = \< Q, P, \init, \weii, \weic, \weir, \weie, \final >$,
-where $Q$ is a finite set of states, 
-$P$ is a finite set of stack symbols, 
-$\mathsf{in} : Q \to \Semiring$, 
-respectively $\mathsf{out} : Q \to \Semiring,$
-are functions defining the weight for entering, 
-respectively leaving, a state, 
-and 
-$\weii : Q \times Q \to \Phii$,  
-$\weic : Q \times Q \times P \to \Phic$,  
-$\weir : Q \times P \times Q \to \Phicr$,  
-$\weie : Q \times Q \to \Phir$,  
-are transition functions.
-\end{definition}
-%
-Similarly as in Section~\ref{section:transducer}, 
-we extend the above transition functions as follows
-for all $q, q' \in Q$, $p \in P$, 
-$a \in \Omegai$, 
-$\call{c} \in \Omegac$, 
-$\return{r} \in \Omegar$, 
-overloading their names: % for simplicity:
-\[
-\begin{array}{lll}
-\weii: Q \times \Omegai \times Q \to \Semiring & 
-\weii(q, a, q') = \phi_\mathsf{i}(a) & 
-\mathrm{where~} \phi_\mathsf{i} = \weii(q, q'),\\
-%
-\weic: Q \times \Omegac \times Q \times P \to \Semiring & 
-\weic(q, \call{c}, q', p) = \phi_\mathsf{c}(\call{c}) & 
-\mathrm{where~} \phi_\mathsf{c} = \weic(q, q', p),\\
-%
-\weir: Q \times \Omegac \times P \times \Omegar \times Q \to \Semiring & 
-\weir(q, {\call{c}},  p, {\return{r}}, q') = \phi_\mathsf{r}({\call{c}},  {\return{r}}) & 
-\mathrm{where~} \phi_\mathsf{r} = \weir(q, p, q'),\\
-%
-\weie: Q \times \Omegar \times Q \to \Semiring & 
-\weie(q, {\return{r}}, q') = \phi_\mathsf{e}({\return{r}}) &
-\mathrm{where~} \phi_\mathsf{e} = \weie(q, q').\\
-\end{array}      
-\]
-
-\noindent
-The intuition is the following for the above transitions.
-\begin{description}
-\item $\weii$ : read the input internal symbol $a$, change state to $q'$ (stack is untouched).
-\item $\weic$ : read the input call symbol $\call{c}$, push it to the stack along with $p$, change state to $q'$.
-\item $\weir$ : when the stack is not empty, 
-      read and pop from stack a pair made of $\call{c}$ and $p$, 
-      read the input return symbol $\return{r}$, change state to $q'$.
-      In this case, the weight function $\phi_\mathsf{r}$ 
-      computes a value of matching between the call and return symbols $c$ and $r$.
-      This value might be set to $\zero$ in order to express that the symbols do not match.
-\item $\weie$ : when the stack is empty, 
-      read the input symbol $\call{r}$, change state to $q'$.
-\end{description}
-
-We give now a formal definition of these transitions of the automaton~$A$
-in term of a weight value
-computed by an intermediate function $\weight_A$, like in Section~\ref{sec:SWT}.
-In the case of a pushdown automaton, a configuration is composed 
-of a state $q \in Q$ and a stack content $\gamma \in \Gamma^*$, where $\Gamma = \Omegac \times P$.
-Therefore, $\weight_A$ is a function from 
-$Q \times \Gamma^* \times \Omega^* \times Q \times \Gamma^*$ into $\Semiring$.
-\[
-\begin{array}{rcl}
-\weight_A\bigl(\config{q}{\gamma}, a\, u, \config{q'}{\gamma'}\bigr) & = & 
- {\displaystyle\bigoplus_{q'' \in Q}} \weii(q, a, q'') 
- \otimes \weight_A\bigl(\config{q''}{\gamma}, u, \config{q'}{\gamma'}\bigr)\\
-%
-\weight_A\bigl(\config{q}{\gamma}, {\call{c}}\, u, \config{q'}{\gamma'}\bigr) & = & 
- {\displaystyle\bigoplus_{\begin{array}{c}
-                          \scriptstyle q'' \in Q\\[-2pt]
-                          \scriptstyle p \in P
-                          \end{array}}}
- \weic\bigl(q, {\call{c}}, q'', p\bigr) 
- \otimes \weight_A\bigl(\config{q''}{{\call{c}}\, p\cdot \gamma}, u, \config{q'}{\gamma'}\bigr)\\[1mm]
-%
-\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot \gamma}, {\return{r}}\, u, \config{q'}{\gamma'}\bigr) & = & 
- {\displaystyle\bigoplus_{q'' \in Q}} 
- \weir\bigl(q, {\call{c}}, p, {\return{r}}, q''\bigr) 
- \otimes \weight_A\bigl(\config{q''}{\gamma}, u, \config{q'}{\gamma'}\bigr)\\
-%
-\weight_A\bigl(\config{q}{\bot}, {\return{r}}\, u, \config{q'}{\gamma'}\bigr) & = & 
- {\displaystyle\bigoplus_{q'' \in Q}} \weie(q, {\return{r}}, q'') 
- \otimes \weight_A\bigl(\config{q''}{\bot}, u, \config{q'}{\gamma'}\bigr)\\
-\end{array}
-\]
-where $\bot$ denotes the empty stack and ${\call{c}}\, p\cdot \gamma$ 
-denotes a stack where the pair made of ${\call{c}} \in \Omegac$ and $p \in P$ is the top symbol 
-and $\gamma$ is the rest of stack.
-
-\noindent
-The weight associated by $A$ to $s \in \Omega^*$
-is then defined as follows,
-following empty stack semantics: 
-\begin{equation}
-A(s)  = 
-{\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
-\mathsf{in}(q) \mathop{\otimes} 
-\weight_A\bigl(\config{q}{\bot}, s, \config{q'}{\bot}\bigr) 
-\mathop{\otimes} \mathsf{out}(q').
-\label{eq:weightA}
-\end{equation}
-
-\begin{example}
-structured words...
-intro language of music notation ?
-\end{example}
-
-\subsection{Properties}
-Like VPA and symbolic VPA, 
-the class of \SWVPA is closed under the binary operators of the underlying semiring.
-%
-\begin{proposition}
-Let $A_1$ and $A_2$ be two (\SWVPA)
-over the same $\Omega$ and $\Semiring$.
-There exists two $\SWVPA$ $A_1 \oplus A_2$ and $A_1 \otimes A_2$, 
-effectively constructible, 
-such that for all $s \in \Omega^*$, 
-$(A_1 \oplus A_2)(s) = A_1(s) \oplus A_2(s)$ and 
-$(A_1 \otimes A_2)(s) = A_1(s) \otimes A_2(s)$.
-\end{proposition}
-The construction is essentially the same 
-as in the case of the Boolean semiring~\cite{dAntonyAlur14SVPDA}.
-
-\subsection{Best-first Search} 
-\label{sec:best}\label{sec:search}
-Let us assume that the semiring $\Semiring$ is
-commutative, bounded, complete, and total.
-% total ? 
-% monotonic and superior writ natural ordering
-%Regarding the infinite sum operator, note that
-%$\bigoplus_{x \in \Phi_\Omega} \phi(x)$, 
-%$\bigoplus_{y \in \Phi_\Delta} \psi(y)$, and  
-%... exist and in $\Semiring$.
-We propose a Dijkstra algorithm computing the minimal weight by $A$, 
-\wrt $\leq_\oplus$, for a word in $\Semiring^*$.
-
-\noindent
-More precisely, 
-let $\top$ be a fresh stack symbol which does not belong to $\Gamma$,
-and for every two states $q, q' \in Q$ and $\sigma \in \{ \bot, \top \}$, 
-let 
-\[ 
-  d_0(q, \sigma, q') = \bigoplus_{s\in \Omega^*} 
-  \textstyle
-  \weight_A\bigl(\config{q}{\sigma}, s, \config{q'}{\sigma}\bigr). 
-\]
-Since $\Semiring$ is complete, this infinite sum is well defined, 
-and since $\leq_\oplus$ is assumed total, it is the minimum in $\Omega^*$
-of $s \mapsto \weight_A(\config{q}{\sigma}, s, \config{q'}{\sigma})$ \wrt this ordering.
-%
-When $\sigma = \bot$, $d_0(q, \sigma, q')$ is the central 
-expression in a term of the definition (\ref{eq:weightA}) of $A(s_0)$ for the minimum $s_0$
-(for the above function).
-When $\sigma = \bot$, intuitively, it is the minimum weight of a computation of $A$
-starting in state $q$ with a stack $\gamma \in \Gamma^*$ (possibly empty), 
-and ending in state $q$ with the same stack $\gamma$, such that moreover 
-the computation does not touch a symbol of $\gamma$. 
-That means that during the computation, 
-$A$ may apply the first case of in the definition of $\weight_A$ (internal symbol),
-as well as the second case, to can push call symbols on the top of $\gamma$, 
-and may pop these symbols with the third case (return symbol). 
-However, it cannot apply one of the two last cases (return symbol and empty stack)
-when the current stack is $\gamma$.
-%pop symbols in $\gamma$.
-% Note that having a stack reduced to such a symbol makes impossible the application of the 
-% two last cases in the definition of $\weight_A$ (return symbol and empty stack). 
-% However, it is possible to apply the two first cases 
-% (internal symbol or call symbol, with a push on the top of $\top$).
-
-The algorithm~\ref{algo:Dijkstra}
-constructs iteratively a marking $d : Q \times \{ \bot, \top \} \times Q \to \Semiring$
-%of the triplets $\<q, \sigma, q'>$ 
-%of states of $A$ by weight values in $\Semiring$, 
-that converges eventually to $d_0(q, \sigma, q')$. 
-%It uses for that purpose a priority queue $P$ containing triplets of 
-%$Q \times \{ \bot, \top \} \times Q$.
-
-\begin{algo}[1-best for \SWVPA] \label{algo:Dijkstra}
-\textbf{initially} let $P = Q \times \{ \bot, \top \} \times Q$, %contains all %triplets 
-%$\< q_1, \bot, q_2>$ and $\<q_1, \top, q_2>$ for $q_1, q_2 \in Q$, with 
-and 
-$d(q_1, \bot, q_2) = d(q_1, \top, q_2) = \one$ if $q_1 = q_2$ and
-$d(q_1, \bot, q_2) = d(q_1, \top, q_2) = \zero$ otherwise.
-
-\smallskip\noindent
-\textbf{while} $P$ is not empty
-
-\noindent
-\quad\textbf{extract} $\< q_1, \sigma, q_2>$ from $P$ 
- such that $d(q_1, \sigma, q_2)$ is minimal wrt $\leq_\oplus$.
-
-\noindent\quad \textbf{for all} $q_0, q_3 \in Q$ and $p \in P$ do
-
-\noindent
-\qquad
-\(\begin{array}{rrcl}
- & d(q_1, \sigma, q_3) & \opluseq &
-   d(q_1, \sigma, q_2) \otimes \displaystyle\bigoplus_{a \in \Omegai} \weii(q_2, a, q_3)\\
-%
-%  & d(q_0, \sigma, q_2) & \opluseq & 
-%    \displaystyle\bigoplus_{a \in \Omegai} \weii(q_0, a, q_1) \otimes d(q_1, \sigma, q_2)\\
-%
-\mathrm{if~} \sigma = \top & d(q_0, \top, q_3) & \opluseq &
-d(q_1, \sigma, q_2) \otimes
-%\displaystyle\bigoplus_{p \in P} 
-\displaystyle\bigoplus_{{\call{c}} \in \Omegac} 
-\displaystyle\bigoplus_{{\return{r}} \in \Omegar} \eta({\call{c}}, {\return{r}})\\
-\mathrm{and} &  d(q_0, \bot, q_3) & \opluseq &
-d(q_1, \sigma, q_2) \otimes
-%\displaystyle\bigoplus_{p \in P} 
-\displaystyle\bigoplus_{{\call{c}} \in \Omegac} 
-\displaystyle\bigoplus_{{\return{r}} \in \Omegar} \eta({\call{c}}, {\return{r}})\\
-& & & \mathrm{where~} \eta = \weic(q_0, q_1, p) \otimes \weir(q_2, p, q_3)\\
-%
-\mathrm{if~} \sigma = \bot & d(q_1, \bot, q_3) & \opluseq &
-  d(q_1, \sigma, q_2) \otimes \displaystyle\bigoplus_{\return{r} \in \Omegar} \weie(q_2, {\return{r}}, q_3)\\
-%
-%& d(q_0, \bot, q_2) & \opluseq & 
-%d(q_0, \bot, q_1) \otimes d(q_1, \sigma, q_2), \mathrm{if~} \< q_0, \bot, q_1> \notin P \\
-%
-%\mathrm{if~} \sigma = \top & d(q_0, \top, q_2) & \opluseq & 
-%d(q_0, \top, q_1) \otimes d(q_1, \sigma, q_2), \mathrm{if~} \< q_0, \top, q_1> \notin P \\
-%
-& d(q_1, \bot, q_3) & \opluseq & 
-   d(q_1, \sigma, q_2) \otimes d(q_2, \bot, q_3), \mathrm{if~} \< q_2, \bot, q_3> \notin P \\
-%
-\mathrm{if~} \sigma = \top & d(q_1, \top, q_3) & \opluseq & 
-    d(q_1, \sigma, q_2) \otimes d(q_2, \top, q_3), \mathrm{if~} \< q_2, \top, q_3> \notin P \\
-\end{array}
-\) 
-\end{algo}
-
-\noindent
-The infinite sums in the updates of $d$ in Algorithm~\ref{algo:Dijkstra} are well defined
-since~$\Semiring$ is complete.
-The algorithm performs $2.|Q|^2$ iterations until $P$ is empty, 
-and each iteration has a time complexity $O(|Q|^2 . |P|)$.
-This gives a time complexity $O(|Q|^4 . |P|)$. 
-It can be reduced by implementing $P$ as a priority queue, 
-prioritized by the value returned by $d$
-***complete***. %$|Q|^3.\log(|Q|^2)$
-
-The correctness of Algorithm~\ref{algo:Dijkstra} 
-is ensured by the invariant expressed in the following lemma.
-\begin{lemma}
-For all $\< q_1, \sigma, q_2> \notin P$, $d(q_1, \sigma, q_2) =  d_0(q_1, \sigma, q_2)$.
-\end{lemma}
-The proof is by contradiction, 
-assuming a counter-example minimal in the length of the witness word.
-
-% \begin{algorithm}
-% \caption{<your caption for this algorithm>}
-% \label{<your label for references later in your document>}
-% \begin{program}
-% \seq{|line| := `` '', i := 1};
-% \WHILE i \neq n+1 \DO
-%       |line| := |item|[i] \concat `` '' \concat |number|[i]; 
-%       i := i+1; 
-%       \WHILE i \neq n+1 \AND |item|[i] = |item|[i-1] \DO 
-%       |line| := |line| \concat ``, '' \concat |number|[i]);
-%       i := i+1 \OD ; 
-%       |write|(|line|) \OD 
-% \end{program}
-% \end{algorithm}
-
-% The weight of a transition acts as a guard: 
-% a transition is activated for a symbol $`a`$ iff its weight $`\phi(a) \neq 0`$ (or $`\psi(a, b) \neq 0`$), 
-% the absorbing element.
-
-%\begin{lemma}
-\noindent
-For computing the minimal weight of a computation of $A$, we use the fact that,
-at the termination of Algorithm~\ref{algo:Dijkstra}, %it holds that,
-%There exist $q_1, q_2 \in Q$
-\[
-  {\displaystyle \bigoplus_{s \in \Omega^*} A(s)} = 
-  {\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
-  \mathsf{in}(q) \mathop{\otimes} d(q, \bot, q') \mathop{\otimes} \mathsf{out}(q').
-\]
-%\end{lemma}
-
-%\medskip
-\noindent
-In order to obtain effectively a witness 
-(word of $\Semiring^*$ with computation of $A$ of minimal weight), 
-we require an additional property the of weight functions.
-\begin{definition}
-Let $\Omega$ be an alphabet and $\Semiring$ a complete semiring.
-A function $\phi$ from $\Omega^n$ into $\Semiring$
-is called $k$-\emph{convex} for a natural number $k$ iff 
-$\mathit{card}\{ \mathbf{a} \in \Omega^n \mid \phi(\mathbf{a}) = 
-                 \displaystyle\bigoplus_{\mathbf{p} \in \Omega^n} \phi(\mathbf{p}) \} \leq k$.
-\end{definition}
-A label theory is $k$-convex if all its functions are $k$-convex.
-
-\begin{proposition}
-For a \SWVPA $A$ over a 
-commutative, idempotent, superior, total and complete semiring 
-and an alphabet $\Omega$ with a $k$-convex label theory,
-one can construct in PTIME a word $s \in \Omega^*$ 
-such that $A(s)$ is minimal \wrt the natural ordering for $\Semiring$. 
-%$\leg_\oplus$.
-\end{proposition}
-
-\subsection{Trees and Nested-Words}
-\label{sec:trees}
-The hierarchical structure of nested-words, defined the \emph{call} and \emph{return} markup symbols of 
-suggest a correspondence between these word and trees. 
-The lifting of this correspondence to languages, respective of tree automata and VPA have
-been discussed in~\cite{AlurMadhusudan09nested}, 
-see also~\cite{Caralp12VPAmult} for the weighted case.
-In this section, we discuss the correspondence between the symbolic-weighted extensions
-of tree automata and VPA.
-
-Let $\Omega$ be a countable ranked alphabet, such that 
-every symbol $a \in \Omega$ has a rank 
-$\rank(a) \in [0..M]$ where $M$ is a fixed natural number.
-We write $\Omega_k$ the subset of all symbols $a$ of $\Omega$
-with $\rank(a) = k$, where $0 \leq k \leq M$.
-%$\cup_{i = 0}^{i = M} \Omega_i$ 
-%
-$\T(\Omega)$ denotes the free $\Omega$-algebra of finite, ordered, 
-$\Omega$-labeled trees, 
-which is the smallest set such that  $\Omega_0 \subset \T(\Omega)$
-and for all $1 \leq k \leq M$, all $a \in \Omega_k$, 
-and all $t_1, \ldots, t_k \in \T(\Omega)$, $a(t_1, \ldots, t_k) \in \T(\Omega)$.
-%
-% tree = single node (leave) labeled with a symbol of $a \in \Omegai$
-% (such a tree is simply denoted by $a$)
-% or the composition, denoted by $b(t_1,\ldots, t_n$) of a node labeled with $b$
-% and $n$ subtrees $t_1$,\ldots, $t_n$.
-
-\renewcommand{\call}[1]{\ensuremath \langle_{#1}}
-\renewcommand{\return}[1]{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
-
-Let $\hat\Omega$ be the countable (unranked) alphabet obtained from 
-$\Omega$ as follows: 
-$\hat\Omega = \< \Omegai, \Omegac, \Omegar >$, with
-$\Omegai = \Omega_0$, 
-$\Omegac = \{ \; \call{a} \mid a \in \Omega_{>0} \}$,
-$\Omegac = \{ \; \return{a} \mid a \in \Omega_{>0} \}$,
-where $\Omega_{>0}$ denotes $\bigcup_{k = 1}^{M} \Omega_k$.
-
-\noindent
-We define a linearization of trees of $\T(\Omega)$ into 
-words of $\hat{\Omega}^*$ as follows:
-\begin{description}
-\item $\lin(a) = a$ for all $a \in \Omega_0$, 
-\item $\lin\bigl( b(t_1, \ldots, t_k)\bigr) = 
-       \call{b} \; \lin(t_1) \ldots \lin(t_k) \; \return{b}$ 
-       when $b \in \Omega_k$, $1 \leq k \leq M$.
-\end{description}
-
-% A \emph{regular tree grammar} over $\Omega$ 
-% is a triplet $G = \< N, q_\mathsf{i}, R>$ where
-% $N$ is a finite set of non-terminal symbols denoted $q$..., 
-% $q_\mathsf{i} \in N$ is the starting non-terminal, 
-% $R$ is a finite set of production rules of the form
-% $q_0 \to a(q_1\ldots q_k)$ where 
-% $q_0, q_1, \ldots, q_k \in N$
-% $a \in \Omega_k$.
-% A tree $t \in \T(\Omega)$ is in the language of $G$ 
-% if it can be generated from $q_\mathsf{i}$ by 
-% non terminal replacement following the rules of $R$.
-
-Let us assume a label theory $\Phi_{\Omega_k}$ 
-for each $k \in [0..M]$.
-%
-\begin{definition}  \label{def:SWTA}
-A \emph{symbolic-weighted tree automaton} (\SWTA)
-over the ranked input alphabet~$\Omega$
-and the commutative semiring $\Semiring$ 
-is a triplet $A = \< Q, \init, \bar{\wei} >$ where
-$Q$ is a finite set of states, 
-$\mathsf{in} : Q \to \Semiring$ is the starting weight function, 
-and $\bar{\wei}$ is a $M+2$-uplet of transition functions made of:
-$\wei_\epsilon$ from $Q \times Q$ into $\Semiring$, and, 
-for each $k \in [0..M]$, $\wei_{\Omega, k}$
-from $Q \times Q^{k}$ into $\Phi_{\Omega_k}$.
-\end{definition}
-%
-Like in Section~\ref{sec:SWAdef}, we define from $\bar{\wei}$ 
-a transition function~$\wei$, %from $\bar\wei$
-from 
-$Q \times (\Sigma \cup \{ \epsilon \}) \times \bigcup_{k=0}^{M} Q^k$
-into~$\Semiring$: %also called $\wei$ for simplicity, 
-%such that, for all $q, q' \in Q$, $a \in \Sigma$, and $b \in \Delta$, 
-\[
-\begin{array}{rcll}
-\wei(q_0, \epsilon, q_1) & = &  \wei_\epsilon(q_0, q_1),\\ %\phi_\epsilon\\
-\wei(q_0, a, q_1 \ldots q_k) & = & \phi_{\Omega, k}(a) &
-\quad\mathrm{where~} \phi_{\Omega, k} = \wei_{\Omega, k}(q_0, q_1\ldots q_k).\\
-\end{array}      
-\]
-Intuitively, $\wei(q_0, a, q_1 \ldots q_k)$ can be seen as
-the weight of a production rule $q_0 \to a(q_1, \ldots, q_k)$ 
-of a regular tree grammar~\cite{tata}, 
-that replaces the non-terminal symbol $q_0$, by $a(q_1, \ldots, q_k)$. 
-%in a step of tree building.
-%
-Such a grammar computes the weights of the derivation trees 
-of the Context-Free grammar obtained by forgetting the labeling symbols of $\Omega_{>0}$.
-
-\noindent
-The \SWTA of Definition~\ref{def:SWTA} defines a mapping 
-from trees of $\T(\Omega)$ into the weights of~$\Semiring$,
-based on the intermediate function $\weight_A$
-defined as follows for $q_0 \in Q$ and 
-$t = b(t_1,\ldots, t_k) \in \T(\Omega)$,
-åwith $0 \leq k \leq M$:
-\[
-\begin{array}{rccl}
-\weight_A(q_0, t) & = & 
-   \displaystyle\bigoplus_{q_1 \in Q} &
-   \wei(q, \epsilon, q_1) \otimes \weight_A(q_1, t)\\
- & \oplus & \displaystyle\bigoplus_{q_1 \ldots q_k \in Q^k} &
-              \wei(q_0, b, q_1 \ldots q_k ) 
-   \otimes \displaystyle\bigotimes_{i=1}^{k}
-           \weight_A(q_i, t_i).\\
-\end{array}
-\]
-
-\medskip\noindent
-The weight associated by $A$ to  $t \in \T(\Omega)$ is then
-\begin{equation}
-A(t)  = 
-\displaystyle\bigoplus_{q \in Q} \mathsf{in}(q) \mathop{\otimes} \weight_A(q, t).
-\label{eq:weightTA}
-\end{equation}
-
-\begin{lemma}  \label{lem:SWTA}
-For all \SWTA $A$ over~$\Omega$ and $\Semiring$, without $\epsilon$-transitions,
-there exists an effectively constructible \SWVPA $A'$ over 
-$\hat\Omega$ and $\Semiring$
-such that for all $t \in \T(\Omega)$, $A'\bigl(\lin(t)\bigr) = A(t)$.
-\end{lemma} 
-% 
-\begin{proof}
-Let $A = \< Q, \init, \bar{\wei} >$ where $\bar{\wei}$ is summarized as above by a function
-$\wei: Q \times (\Sigma \cup \{ \epsilon \}) \times \bigcup_{k=0}^{M} Q^k \to \Semiring$.
-
-We build 
-$A' = \< Q', P', \init', \weii, \weic, \weir, \weie, \final' >$,
-computing over $\hat\Omega = \< \Omegai, \Omegac, \Omegar >$,
-%
-%$\Omega = \Omegai \uplus \Omegac \uplus \Omegar$ and $\Semiring$ 
-where $Q' = \bigcup_{k=0}^{M} Q^k$ be  the set of sequences of state symbols of $A$, 
-of length at most $M$, including the empty sequence denoted by $\epsilon$, 
-and where $P' = Q'$.
-
-\[
-\begin{array}{ll}
-%\weii: Q \times \Omegai \times Q \to \Semiring & 
-\weii(\bar{q}, a, \bar{q} q') = \wei(q', a, \epsilon) & 
-\mathrm{for~all~} a \in \Omega_0,\\
-%
-%\weic: Q \times \Omegac \times Q \times P \to \Semiring & 
-\weic(\bar{q}, \call{b}, \epsilon, \bar{q}) = \one & %\wei(p', b, \bar{p}) & 
-\mathrm{for~all~} b \in \Omega_{>0},\\
-%
-%\weir: Q \times \Omegac \times P \times \Sigmar \times Q \to \Semiring & 
-\weir(\bar{q}, {\call{b}}, \bar{p}, {\return{b}}, \bar{p} q') = \wei(q', b, \bar{q}) & 
-\mathrm{for~all~}  b \in \Omega_{>0},\\
-%
-%\weie: Q \times \Omegar \times Q \to \Semiring & 
-\weie(\bar{p}, {\return{b}}, \bar{q}) = \zero &
-\mathrm{for~all~}  b \in \Omega_{>0}.\\
-\end{array}      
-\]
-
-In practice, it is sufficient to consider in $Q'$ only the prefixes of 
-sequences.
-\qed\end{proof}
-
-\section{Symbolic Weighted Parsing}
-\label{sec:parsing}
-Let us now use the models and results of the former sections in %order to define 
-an approach to the problem of parsing over infinite alphabet. %appropriate
-%
-Besides considering infinitely many possible of input symbols, 
-handled with suitable language formalisms, 
-this approach extends conventional parsing 
-by computing a derivation tree modulo 
-a generic distance between words,  
-defined by a SW transducer given in input.
-This enables considering finer word relationships than strict equality 
-as in the conventional parsing approach, 
-opening possibilities of quantitative analysis via this method.
-
-\subsection{Definition}
-
-Let $\Sigma$ be a countable input alphabet 
-and $\Omega = \Omegai \uplus \Omegac \uplus \Omegar$ an output countable alphabet,
-let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative, bounded, complete 
-and total semiring 
-and let  $(\Phi_\epsilon, \Phic, \Phir, \Phicr)$ be a label theory over $\Semiring$, 
-assumed computable and $k$-convex for some fixed $k$.
-
-\noindent
-Assuming given in input:
-\begin{itemize}
-\item a \SWT $T$ over $\Sigma$,  $\Omega$, and $\Semiring$,
-defining a measure %between words 
-$T: \Sigma^* \times \Omega^* \to \Semiring$,
-\item a \SWVPA $A$ over $\Omega$, and $\Semiring$, defining a series of nested words
-      $A : \Omega^* \to \Semiring$,
-\item an input word $s \in \Sigma^*$,
-\end{itemize}
-the problem of \emph{symbolic weighted parsing} is to find 
-a nested word $t \in \Omega$ minimizing
-\( T(s, t) \otimes A(t)\)
-\wrt $\leq_\oplus$, 
-\ie such that %Hence, it is the problem of finding 
-\( T(s, t) \otimes A(t) = 
-  \displaystyle\bigoplus_{t' \in \Omega^*} T(s, t') \otimes A(t') \).
-
-\noindent 
-Therefore, it is the problem of optimizing a measure 
-called the \emph{edit-distance between $s$ and $A$} in~\cite{Mohri03EDWA}.
-%
-The input language can also be expressed as a \SWTA, or, 
-as a particular case, as a weighted context-free grammar, 
-converted in turn into a \SWVPA following Lemma~\ref{lem:SWTA}.
-%
-In the case of finite alphabets, 
-the problem of searching, in a WTA language, 
-for the best parse tree for a given,  
-sometimes referred as~\emph{weighted parsing}
-(see \cite{Goodman99SemiringParsing,MorbitzVogler19weighted-parsing} **more general problems**)
-is a particular case of SW parsing.
-%
-Indeed, it corresponds to the case where $T$
-accepts only the pairs $\<s, t>$ such that 
-$s$ is the projection of $t$ on $\Omegai$. 
-This can be done with a single state $q$ and 
-with transition rules of the form:
-\begin{description}
-\item[] $\wei(q, \epsilon, a, q) = \one$ for all $a \in \Omegac \cup \Omegar$,
-\item[] $\wei(q, a, a, q) = \one$ for all $a \in \Omegai$,
-\item[] $\wei(q, a, b, q) = \zero$ for all $a, b \in \Omegai$, $a \neq b$.
-\end{description}
-
-\subsection{Computation}
-
-\begin{proposition}
-The problem of Symbolic Weighted  parsing 
-can be solved in PTIME in the size of the input \SWT, \SWVPA (or \SWTA) and input word, 
-and the computation time of the functions of the label theory.
-\end{proposition}
-%
-\begin{proof}
-Bar-Hillel construction
-\qed  
-\end{proof}
-
-\subsection{Application to Automated Music Transcription}
-\label{sec:transcription}
-Symbolic Automated Music Transcription
-and analysis of music performances
-
-\subsubsection{Time Scales}
-Real-Time Unit (RTU) = seconds
-
-\noindent 
-Musical-Time Unit (MTU) = number of measures
-
-\noindent 
-conversion via tempo value
-
-\subsubsection{Representation of Music Performances}
-We consider symbolic representations of musical performances, as finite sequences of events.
-It corresponds to the concrete case of a MIDI file~\cite{SMF} 
-recorded  from an electronic keyboard, 
-or the output of a transcription from audio files~\cite{Benetos18AMTsurvey}.
-%
-For the sake of simplicity, 
-we shall only consider here the case of monophonic performances, 
-where at most one note is sounding at a time. 
-The approach however extends to the polyphonic case.
-
-A music performance is a finite sequence of events in a set~$\Sigma$.
-Every event $e \in \Sigma$ has attributes such from a finite domain, 
-like a number of key for a note 
-or a flag indicating that it is a rest 
-(\textsf{ON} and \textsf{OFF} messages in~\cite{SMF})
-and a velocity value (0..127 in~\cite{SMF})).
-%This representation is similar to the piano roll ~\cite{Muller15fundamentals} chap.1. 
-Moreover, it contains a RTU value $\ioi{e}$ (real number) 
-representing the time distance to the next event, 
-or to the end of performance for the last event,
-also called \emph{inter-onset interval}.
-
-\subsubsection{Representation of Music Scores}
-Music score are represented as structured words
-made of timed %quantified 
-events and parenthesized markups,
-akin of nested words~\cite{AlurMadhusudan09nested}.
-
-We consider an alphabet $\Delta$, every symbol of which is 
-composed of a tag, in a finite set $\Xi$, 
-and an MTU (rational) IOI duration value.
-%The alphabet $\Delta$ 
-It is partitioned into 
-$\Delta = \Deltai \uplus \Deltac \uplus \Deltar$, 
-like in Section~\ref{section:SWVPA}.
-%
-\noindent
-The symbols of $\Deltai$ represent events:
-% (infinite alphabet of internal symbols) made of:
-with tags indicating a new note or grace-note (with null IOI), 
-a rest or the continuation of the previous note (tie or dot).
-%
-The elements of $\Deltac \uplus \Deltar$ are matched
-markups for describing the structure of the score, 
-\ie the hierarchical grouping of events, and also, 
-importantly the division of time in measures, tuplets...
-%- parentheses for time divisions : tuplets, bars...
-(linearization of rhythm trees \cite{jacquemard:hal-01138642}...).
-They contain additional info such as tuple number, beaming policy...
-
-\noindent
-The duration values of letters of $\Delta$, in MTU (rational), 
-can be computed with the markups and tags (\eg grace note has duration 0).
-
-%\noindent
-%There are simultaneous events, since grace notes has duration 0. They are ordered.
-%
-%\noindent
-%Finite bound on the number of duration ratio. ?
-
-\begin{example}
-...      
-\end{example}
-
-\subsubsection{Performance/Score Distance Computation}
-\label{app:distance}
-We define a distance between performance and score representations
-by a $\SWT$ $T = \< Q, \init, \wei, \final >$, over a semiring $\Semiring$.
-** detail the elements of $\Semiring$ ....**
-%are quadruplets of the form
-%$\< t, s, \delta_t, \delta_s>$
-
-Every state of $Q$ contains a 
-tempo value in a finite domain (e.g. 30..300 bpm).
-This value can be fixed 
-or recomputed by the $T$ %transducer 
-after reading each event, 
-according to a perceptive/cognitive model of tempo 
-such as~\cite{LargeJones99tempo}
-(also used in the context of score following~\cite{Cont10TPAMI}).
-% we wont detail here.
-
-\subsection{Transcription by SW Parsing} %Best-first Search}
-We assume a score language defined by a \SWVPA over the semiring 
-$\Semiring$ of Section~\ref{app:distance}.
-
-\section*{Conclusion}
-% discussion
-
-\noindent
--- summary
-
-\noindent
--- other theoretical properties of SW models
-
-\noindent
--- room to improve complexity for best-search algorithm
-... modular approach with oracles ...
-
-and extention to $n$-best
-
-\noindent
--- offline algorithm, 
-semi-online implementation for AMT (bar-by-bar approach)
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% BIBLIO                                                                     %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\bibliographystyle{plain}
-%\bibliographystyle{plainurl} 
-\bibliographystyle{abbrv}
-%\bibliographystyle{splncs04}
-\bibliography{references}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% APPENDIX                                                                   %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\appendix
-
-\section{Edit-Distance}
-
-%\subsection{Distance between words or languages}
-...algebraic definition of edit-distance of Mohri, in \cite{Mohri03EDWA}
-% Mehryar Mohri 
-% Edit-distance of weighted automata: General definitions and algorithms
-% International Journal of Foundations of Computer Science 14.06 (2003): 957-982.
-distance $d$ over $\Sigma^* \times \Sigma^*$ 
-into a semiring  $\Semiring = ( \Semiring, \oplus, \zero, \otimes, \one)$.
-
-%\noindent
-Let $\Omega = \Sigma \cup \{ \epsilon \} \times \Sigma \cup \{ \epsilon \} \setminus \{ (\epsilon, \epsilon) \}$,
-and let $h$ be the morphism from $\Omega^*$ into $\Sigma^* \times \Sigma^*$  
-defined over the concatenation of strings of $\Sigma^*$ (that removes the $\epsilon$'s).
-%
-\noindent
-An \emph{alignment} between 2 strings  $s, t \in \Sigma^*$ is an element $\omega \in \Omega^*$ 
-such that $h(\omega) = (s, t)$.
-%
-\noindent
-We assume a base cost function $\Omega$ : $\delta: \Omega \to S$, extended to $\Omega^*$ as follows  
-(for $\omega \in \Omega^*$): 
-\(
-\displaystyle\delta(\omega) = \bigotimes_{0 \leq i < |\omega|} \delta(\omega_i)
-\).
-
-\noindent
-\begin{definition}
-For  $s, t \in \Sigma^*$, the edit-distance between $s$ and $t$ is  
-\( 
-d(s, t) = \displaystyle\bigoplus_{\omega \in \Omega^*\, h(\omega) = (s, t)} \delta(\omega)
-\).
-\end{definition}
-
-e.g. Levenstein edit-distance: $S$ is min-plus and $\delta(a, b) = 1$ for all $(a, b) \in \Omega$.
-
-%\paragraph{Distance between a word and a regular language}
-
-\end{document}
+a parse-based approach to automated music transcription.

### 2021-05-16 introduction.tex
split files

--- 2021-05-12 main.tex
@@
-\documentclass[runningheads]{llncs}
-%\documentclass[a4paper,11pt]{article}
-%\setcounter{page}{1}
-
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-
-\usepackage{hyperref}
-%\usepackage[bookmarks,bookmarksnumbered,naturalnames,plainpages=false]{hyperref}
-%usepackage{url}
-
-% for footnote ref
-\usepackage{refcount}
-
-% array and tabular
-\usepackage{array}
-\newcolumntype{C}[1]{>{\centering\arraybackslash\hspace{0pt}}p{#1}}
-
-% extension of enumerate env. (style for displaying counters)
-% \usepackage{enumerate}
-
-%% pictures
-% \usepackage{graphicx} 
-% \DeclareGraphicsExtensions{.pdf,.png,.jpg}
-% \graphicspath{fig/}
-
-%% PGF, Tikz
-\usepackage{tikz-cd}
-%% \usepackage{pgfplots}
-%% \usepgfplotslibrary{dateplot}
-%% \usepackage{pgf,pgfarrows,pgfnodes, pgfautomata}
-% \usepackage{tikz}
-% \usetikzlibrary{cd}
-%% \usetikzlibrary{arrows}
-%% \usetikzlibrary{calc}
-%% \usetikzlibrary{snakes}
-%% \usetikzlibrary{backgrounds}
-% \usetikzlibrary{trees}
-%% \usetikzlibrary{automata}
-%% \usetikzlibrary{positioning}
-%% \usetikzlibrary{matrix}
-%% \usetikzlibrary{patterns}
-%% \usetikzlibrary{shapes}
-
-% symbols
-\usepackage{amsmath} 
-\usepackage{amssymb} 
-\usepackage{amsbsy}
-\usepackage{bbold}
-\usepackage{latexsym}
-%\usepackage{amsfonts}
-\usepackage{stmaryrd}
-%\usepackage{mathabx}
-%\usepackage{MnSymbol}
-\usepackage{harmony} % simple music fonts
-\usepackage{mathtools} % for arrows
-%\usepackage{mathptmx}
-
-%% theorem environments
-\usepackage{theorem}
-% \newtheorem{theorem}{Theorem} %[section]
-% \newtheorem{definition}[theorem]{Definition}
-% \newtheorem{lemma}[theorem]{Lemma}
-% \newtheorem{corollary}[theorem]{Corollary}
-% \newtheorem{proposition}[theorem]{Proposition}
-% \newenvironment{proof}{\vspace{-2ex}{\it Proof. }}{\hspace*{\fill} $\Box$\smallskip }
-% \theorembodyfont{\slshape}
-% \newtheorem{example}[theorem]{Example}
-% \newtheorem{remark}[theorem]{Remark}
-{\theorembodyfont{\rmfamily} \theoremstyle{break} \newtheorem{algo}{Algorithm}}
-
-%% algorithms
-%\usepackage{algorithm} 
-%\usepackage{program}
-
-%% for new macros
-\usepackage{xspace}
-
-%% arrows etc
-%\input{rewriting}
-
-%% Misc macros
-
-\def\ie{\textit{i.e.}\xspace}
-\def\eg{\textit{e.g.}\xspace}
-\def\wrt{\textit{wrt}\xspace}
-%\def\wlog{\textit{wlog}\xspace}
-\def\etc{\textit{etc}\xspace}
-
-\def\<#1>{\langle #1 \rangle}
-\newcommand{\pair}[2]{\langle{#1}, {#2}\rangle}
-%\newcommand{\A}{\mathcal{A}}
-%\newcommand{\B}{\mathcal{B}}
-\newcommand{\D}{\mathbb{D}}
-\newcommand{\E}{\mathbb{E}}
-\newcommand{\W}{\mathbb{W}}
-\newcommand{\T}{\mathcal{T}}
-
-\newcommand{\Semiring}{\mathbb{S}}
-\newcommand{\zero}{\mathbb{0}}
-\newcommand{\one}{\mathbb{1}}
-\newcommand{\dom}{\ensuremath{\mathit{dom}}}
-
-\def\SA{\textsf{SA}\xspace}
-\def\WA{\textsf{WA}\xspace}
-\def\SWT{\textsf{swT}\xspace}
-\def\SWA{\textsf{swA}\xspace}
-\def\SWTA{\textsf{swTA}\xspace}
-\def\SWVPA{\textsf{sw-VPA}\xspace}
-\def\weight{\mathsf{weight}}
-\def\wei{\mathsf{w}}
-\def\init{\mathsf{in}}
-\def\final{\mathsf{out}}
-\newcommand{\call}[1]{\ensuremath #1} %{\ensuremath \langle_{#1}}
-\newcommand{\return}[1]{\ensuremath #1} %{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
-\def\Omegai{{\Omega_\mathsf{i}}}
-\def\Omegac{{\Omega_\mathsf{c}}}
-\def\Omegar{{\Omega_\mathsf{r}}}
-\def\Sigmai{{\Sigma_\mathsf{i}}}
-\def\Sigmac{{\Sigma_\mathsf{c}}}
-\def\Sigmar{{\Sigma_\mathsf{r}}}
-\def\Deltai{{\Delta_\mathsf{i}}}
-\def\Deltac{{\Delta_\mathsf{c}}}
-\def\Deltar{{\Delta_\mathsf{r}}}
-\def\Phii{{\Phi_\mathsf{i}}}
-\def\Phic{{\Phi_\mathsf{c}}}
-\def\Phir{{\Phi_\mathsf{r}}}
-\def\Phicr{{\Phi_\mathsf{cr}}}
-\def\weii{{\wei_\mathsf{i}}}
-\def\weic{{\wei_\mathsf{c}}}
-\def\weir{{\wei_\mathsf{r}}}
-\def\weie{{\wei_\mathsf{e}}}
-\newcommand{\config}[2]{\ensuremath \genfrac{[}{]}{0pt}{}{#1}{#2}} 
-\newcommand{\opluseq}{\ensuremath\mathrel{\oplus}=}
-\newcommand{\ioi}[1]{\mathsf{ioi}({#1})}
-\newcommand{\rank}{\mathsf{rk}}
-\newcommand{\lin}{\mathsf{lin}}
-
-%\sloppy
-
-% Parsing over infinite alphabet as optimal alignment computation 
-% as edit-distance between string and language
+% !TEX root = main.tex
 %
-%\title{Symbolic Weighted Parsing and Automated Music Transcription}
-%\title{Symbolic Weighted Language Models and Automated Music Transcription}
-\title{Symbolic Weighted Language Models and Parsing over Infinite Alphabets}
-%\title{Weighted Visibly Pushdown Automata and Automated Music Transcription}
-\author{Florent Jacquemard}
-\institute{INRIA \& CNAM, Paris, France\\
-\email{florent.jacquemard@inria.fr}}
-
-%\titlerunning{WVPA \& AMT}
-\titlerunning{Symbolic Weighted Parsing and Automated Music Transcription}
-%\authorrunning{Florent Jacquemard}
-
-\date{\today}
-
-\begin{document}
-\thispagestyle{empty}
-\maketitle
-
-\begin{abstract}
-%Symbolic Weighted (SW) extension of symbolic automata...
-We propose a framework for weighted parsing over infinite alphabets.
+% Introduction
 %
-It is based on language models called Symbolic Weighted Automata (SWA) 
-at the joint %intersection 
-between Symbolic Automata (SA) and Weighted Automata (WA), 
-as well as Transducers (SWT) and Visibly Pushdown (SWPDA) variants.
-%
-Like SA, SWA deal with large or infinite input alphabets, 
-and like WA, they output a weight value in a semiring domain.
-The transitions of WA are labeled by functions from an infinite alphabet into the weight domain. 
-This is unlike WA whose transitions are guarded by boolean predicates 
-overs symbols in an infinite alphabet 
-and also unlike WA whose transitions are labeled by constant weight values, 
-and who deal only with finite automata.
-%
-We present some properties of WA, WT and SWPDA models, 
-%and show how they can be used 
-that we use to define and solve a variant of parsing 
-over infinite alphabets.
-%
-We also briefly describe the application that motivated the introduction of these models: 
-an parse-based approach to automated music transcription.
-\end{abstract}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% intro
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Introduction} \label{sec:intro}
 Parsing %variant of membership problem for formal languages
@@
 %of this approach to automated music transcription that has been implemented.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SWT & SWA
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{SW Automata and Transducers}
-\label{section:transducer}\label{sec:transducer}
-\label{section:SWA}\label{sec:SWA}
-\label{section:SWT}\label{sec:SWT}
-
-We follow the approach of~\cite{Mohri03EDWA} for the computation of distances
-between words and languages, with weighted transducers, 
-and extending it to infinite alphabets.
-% with models of symbolic weighted automata and transducers. 
-%
-The models introduced in this section generalize 
-weighted automata and transducers~\cite{Droste09handbook} 
-%over finite alphabets, see  e.g.~\cite{Mohri03EDWA}, 
-by labeling each transition with a weight function that takes the 
-input and output symbols as parameters, instead of a simple weight value.
-These functions are similar to the guards of symbolic automata~\cite{dAntoniVeanes17CAV,dAntoni21CACM},
-but they can return values in an arbitrary semiring, 
-where the latter guards are restricted to the Boolean semiring.
-
-\subsection{Semirings}
-\label{section:semiring}
-We shall consider semiring domains for the weight values of our language models.
-%
-A \emph{semiring} $\< \Semiring, \oplus, \zero, \otimes, \one>$ 
-is a structure with a domain~$\Semiring$,
-equipped with two associative
-binary operators~$\oplus$ and $\otimes$
-with respective neutral elements $\zero$ and $\one$ and such that:
-%$\< \mathbb{S}, \oplus, \zero>$ is a commutative monoid
-%$\< \mathbb{S}, \otimes, \one>$ is a monoid
-\begin{itemize}
-\item $\oplus$ is commutative 
--- therefore $\< \Semiring, \oplus, \zero>$ is a commutative monoid 
-   and $\< \Semiring, \otimes, \one>$ is a monoid,
-\item $\otimes$ distributes over~$\oplus$:  $\forall x, y, z \in \mathbb{S}$,
-$x \otimes (y \oplus z) = (x \otimes y) \oplus (x \otimes z)$, 
-and $(x \oplus y) \otimes z = (x \otimes z) \oplus (y \otimes z)$,
-\item $\zero$ is absorbing for~$\otimes$: 
-$\forall x\in \mathbb{S}$, $\zero \otimes x = x \otimes \zero = \zero$.
-\end{itemize}
-%Components of a semiring~$\Semiring$ may be subscripted by~$\Semiring$ when needed.
-%We simply write $x \in \Semiring$ to mean $x \in \mathbb{S}$.
-%
-Intuitively, in the models presented in this paper, 
-$\oplus$ selects an optimal value from two given values, 
-in order to handle non-determinism, 
-and $\otimes$ combines two values into a single value, 
-in order to describe the chaining of several transitions.
-%and let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a {semiring},
-
-\medskip%\noindent
-A semiring $\Semiring$ is \emph{commutative} if $\otimes$ is commutative.
-It is \emph{idempotent} if for each $x \in \dom(\Semiring)$, $x \oplus x = x$.
-%
-Every idempotent semiring~$\Semiring$ induces 
-a partial ordering~$\leq_\oplus$ 
-called the \emph{natural ordering} of~$\Semiring$~\cite{Mohri02semiring} 
-and defined,  by: 
-%implicitly defined by the semiring $\Semiring$ 
-for all $x$ and $y$,
-$x \leq_\oplus y \;\mbox{iff}\; x \oplus y = x$.
-%(see~\cite{Mohri02semiring} for the proof that it is an ordering).
-%
-The natural ordering is sometimes defined in the opposite direction~\cite{DrosteKuich09semirings};
-We follow here the direction  %follows \cite{Mohri02semiring}, and 
-that coincides than the usual ordering on the Tropical semiring (\emph{min-plus}).
-
-\begin{lemma}[Monotony, \cite{Mohri02semiring}]
-Let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be an idempotent semiring.
-For all $x, y, z  \in \Semiring$,  
-if $x \leq_\oplus y$ then
-$x \oplus z \leq_\oplus y \oplus z$,
-$x \otimes z \leq_\oplus y \otimes z$
-and $z \otimes x \leq_\oplus z \otimes y$.
-\end{lemma}      
-We say in this case that $\Semiring$ is \emph{monotonic}. % \wrt $\leq_\oplus$.
-%A semiring $\Semiring$ 
-%is \emph{monotonic} \wrt a partial ordering~$\leq$ 
-%iff for all $x, y, z  \in \Semiring$,  $x \leq y$ implies
-%$x \oplus z \leq y \oplus z$,
-%$x \otimes z \leq y \otimes z$,
-%and $z \otimes x \leq z \otimes y$.
-%
-Another important semiring property in the context of optimization
-is {superiority}~\cite{Huang08advanceddynamic}, 
-which corresponds to the 
-\emph{non-negative weights} condition in shortest-path algorithms~\cite{Dijkstra59anote}.
-Intuitively, it means that combining elements with $\otimes$ always increase their weight. 
-Formally, it is defined as the property of the following lemma.
-
-\begin{lemma}[Superiority, Boundedness]
-\label{lem:superior}\label{lem:bounded}
-Let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be an idempotent semiring.
-The two following statements are equivalent:
-\begin{itemize}
-\item [$i.$] for all $x, y \in \Semiring$,  
-$x \leq_\oplus x \otimes y$ and 
-$y \leq_\oplus x \otimes y$
-\item[$ii.$] for all $x \in \Semiring$,  $\one \oplus x = \one$.
-\end{itemize}
-\end{lemma}
-%
-In~\cite{Huang08advanceddynamic}, the property~$i.$ is called 
-$\Semiring$ is \emph{superior} \wrt the ordering~$\leq_\oplus$.
-It implies that 
-%It holds that If $\Semiring$ is superior \wrt~$\leq$, then %$\one \leq \zero$
-for all $x \in \Semiring$, $\one \leq_\oplus x \leq_\oplus \zero$.
-Indeed, by the second inequality of $(i.)$ with $y = \one$, $\one \leq_\oplus \one \otimes x = x$, 
-and, by the first inequality of $(i.)$ with $y = \zero$,  
-$x \leq_\oplus x \otimes \zero = \zero$.
-%
-Intuitively, from an optimization point of view,
-it means that $\one$ is the best value, and $\zero$ the worst.
-%** superior implies $\Semiring$ bounded~\cite{Mohri02semiring} see **
-%
-In~\cite{Mohri02semiring}, the property of Lemma~\ref{lem:bounded}(in the form $ii.$) 
-is called \emph{boundedness} of $\Semiring$ -- we shall use this term in the rest of the paper. 
-% \emph{negative boundedness} 
-It implies that, when looking for a best path in a graph whose edges
-are weighted by values of $\Semiring$, the loops can be safely avoided.
-
-\begin{proof}(Lemma~\ref{lem:bounded})
-For $(ii.) \Rightarrow (i.)$ : 
-$x \oplus (x \otimes y) = x \otimes (\one \oplus y) = x$, 
-by distributivity of $\otimes$ over $\oplus$. 
-Hence $x \leq_\oplus x \otimes y$.
-Similarly, $y \oplus (x \otimes y) = (\one \oplus x) \otimes y = y$, 
-hence $y \leq_\oplus x \otimes y$.
-%
-For $(i.) \Rightarrow (ii.)$ : as we have seen, $(i.)$ implies that 
-$\one \leq_\oplus x$ for all $x \in \Semiring$, 
-\ie, by definition of $\leq_\oplus$, $\one \oplus x = \one$.
-\qed
-\end{proof}
-
-%Following the terminology of~\cite{Mohri02semiring},
-%when $\forall x \in \dom(\Semiring), \one \oplus x = \one$,
-%the semiring $\Semiring$ is is called \emph{bounded}.
-
-\begin{lemma}
-Every bounded semiring is idempotent.
-\end{lemma}
-\begin{proof}
-By boundedness, $\one \oplus \one = \one$, 
-and idempotency follows by multiplying
-both sides by $x$ and distributing. 
-\qed
-\end{proof}
-
-\noindent
-An idempotent semiring $\Semiring$~is called \emph{total} if
-it~$\leq_\oplus$ is total
-\ie when for all $x, y \in \Semiring$, either $x \oplus y = x$ or $x \oplus y = y$.
-
-%\medskip
-We shall need below infinite sums with~$\oplus$.
-A semiring~$\Semiring$ is called \emph{complete}~\cite{Kuich97semirings}
-if for every family
-$(x_i)_{i \in I}$ %$\{ x_i \mid i \in I \}$
-of elements of $\dom(\Semiring)$ over an index set $I \subset \mathbb{N}$,
-the infinite sum $\bigoplus_{i \in I} x_i$
-is well-defined and in $\dom(\Semiring)$,
-and the following properties hold:
-\begin{description}
-\item[$i.$]
-\emph{infinite sums extend finite sums:}
-$\displaystyle\bigoplus_{i \in \emptyset} x_i = \zero$,\quad 
-      $\forall j\in \mathbb{N}, \displaystyle\bigoplus_{i \in \{ j \}} x_i = x_j$,\quad
-      $\forall j, k\in \mathbb{N}, j\neq k, 
-      \displaystyle\bigoplus_{i \in \{ j, k \}} x_i = x_j \oplus x_k$,
-%
-\item[$ii.$]
-\emph{associativity and commutativity:}
-for all $I \subseteq \mathbb{N}$
-and all partition $(I_{j})_{j \in J}$ of $I$, %\subseteq \mathbb{N}$, 
-\(
-\displaystyle
-\bigoplus_{j \in J}\bigoplus_{i \in I_j} x_i = 
-\bigoplus_{i \in I} x_i
-\),
-%
-\item[$iii.$] 
-\emph{distributivity of product over infinite sum:}\\
-for all $I \subseteq \mathbb{N}$,
-\(
-\displaystyle
-\bigoplus_{i \in I} (x \otimes y_i) = x \otimes \bigoplus_{i\in I} y_i\), and
-\(
-\displaystyle
-\bigoplus_{i \in I} (x_i \otimes y) = (\bigoplus_{i \in I} x_i ) \otimes y\).
-\end{description}
-
-\begin{example}
-Figure~\ref{gif:semirings} presented examples of semirings interesting in practice 
-and enjoying the above properties.
-\end{example}
-
-\begin{figure}[t]
-\begin{center}
-\begin{tabular}{|c|c|C{2em}|C{2em}|C{2em}|C{2em}|}
-\hline
-        & domain & $\oplus$ & $\otimes$ & $\zero$  & $\one$\\ %& nat. ordering\\
-\hline\hline
-Boolean & $\{\bot, \top\}$ & $\vee$ & $\wedge$ & $\bot$ & $\top$\\ %& $\top \leq_\oplus \bot$  \\
-\hline
-Counting & $\mathbb{N}$ & $+$ & $\times$ & 0 & 1 \\
-\hline
-Viterbi & $[0, 1] \subset \mathbb{R}$ & $\mathit{max}$ & $\times$ & 0 & 1\\ % & $x \leq_\oplus y \iff x \ge y$  \\
-\hline
-Tropical min-plus & $\mathbb{R}_+ \cup \{ +\infty\}$ & $\mathit{min}$ & $+$ & $+\infty$ & $0$\\ % & $x \leq_\oplus y \iff x \leq y$   \\
-\hline
-%MaxPlus & $\mathbb{Q} \cup \{ -\infty\}$ & $\mathsf{max}$ & $+$ & $-\infty$ & $0$ \\
-%\hline
-%Word lang. & $2^{\Sigma^*}$ & $\cup$ & $\cdot$ & $\emptyset$ & $\{ \varepsilon \}$ \\
-%\hline
-\end{tabular}
-\end{center}
-\vskip-1em
-\caption{Some commutative, bounded, total and complete semirings.}
-\label{fig:semirings}
-\end{figure}
-
-\subsection{Label Theory}
-\label{section:symbols}
-We shall now define the functions labeling the transitions of SW-automata and -transducers.
-Let $\Sigma$ and $\Delta$ be two countable sets of symbols
-called respectively \emph{input} and \emph{output} \emph{alphabets}, 
-and let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative semiring.
-
-\noindent 
-A \emph{label theory} over $\Sigma$ and $\Delta$
-is made of 3 recursively enumerable sets:
-%$\Phi_\epsilon \subseteq \Semiring$, % containing constant functions valued in $\Semiring$, 
-$\Phi_\Sigma$ and $\Phi_\Delta$, 
-containing unary functions in $\Sigma \to \Semiring$, resp. $\Delta \to \Semiring$, 
-and $\Phi_{\Sigma, \Delta}$  containing binary functions in $\Sigma \times \Delta \to \Semiring$, 
-such that 
-for all $\alpha \in \Semiring$, and $\phi \in \Phi_\Sigma$,
-$\alpha \otimes \phi : x \mapsto \alpha \otimes \phi(x)$, 
-and $\phi \otimes \alpha : x \mapsto \phi(x) \otimes \alpha$
-belong to $\Phi_\Sigma$, and similarly for $\oplus$ 
-and for $\Phi_\Delta$ and $\Phi_{\Sigma, \Delta}$.
-%
-Moreover, these sets are required to be closed under the following 
-operators~$\oplus$ and $\otimes$ of~$\Semiring$:
-for all $\phi, \phi' \in \Phi_\Sigma$,
-all $\psi, \psi' \in \Phi_\Delta$, 
-and $\eta, \eta' \in \Phi_{\Sigma, \Delta}$, the function
-\begin{description}
-\item $\phi \otimes \phi' : x \mapsto \phi(x) \otimes \phi'(x)$ belongs to $\Phi_\Sigma$, 
-\item $\psi \otimes \psi' : y \mapsto \psi(y) \otimes \psi'(y)$ belongs to $\Phi_\Delta$,
-\item $\phi \otimes \eta : x, y \mapsto \phi(x) \otimes \eta(x, y)$ belongs to $\Phi_{\Sigma, \Delta}$,
-\item $\eta \otimes \psi : x, y \mapsto \eta(x, y) \otimes \psi(y)$ belongs to $\Phi_{\Sigma, \Delta}$,
-\item $\eta \otimes \eta' : x, y \mapsto \eta(x, y) \otimes \eta'(x, y)$ belongs to $\Phi_{\Sigma, \Delta}$.
-\end{description}
-And the same also holds for the binary sum operator $\oplus$.
-
-\noindent
-%\marginpar{necessary?}
-Finally, it is also required 
-% that the codomain of every function of $\Phi_\Sigma$ and $\Phi_\Delta$ 
-% is a subset of $\Phi_\epsilon$, and
-that all partial applications of functions $\Phi_{\Sigma, \Delta}$, 
-resp.  $f_a: y \mapsto f(a, y)$ for $a \in \Sigma$ and $y \in \Delta$
-and  $f_b: x \mapsto f(x, b)$ for $b \in \Delta$ and $x \in \Sigma$, 
-belong resp. to~$\Phi_\Sigma$ and~$\Phi_\Delta$.
-
-\subsection{Definitions} \label{sec:SWTdef}\label{sec:SWAdef}
-%(SWT)
-\begin{definition}
-\label{def:transducer} \label{def:SWT}
-A \emph{symbolic-weighted transducer} (\SWT)
-over the input and output alphabets~$\Sigma$ and $\Delta$ 
-and the semiring $\Semiring$ is a tuple
-$T = \< Q, \init, \bar{\wei}, \final >$,
-where $Q$ is a finite set of states, 
-$\mathsf{in} : Q \to \Semiring$, 
-respectively $\mathsf{out} : Q \to \Semiring,$
-are functions defining the weight for entering, 
-respectively leaving, a state, 
-and $\bar{\wei}$ is a tuplet made of the 4 of transition functions 
-$\wei_\epsilon$, $\wei_\Sigma$, $\wei_\Delta$, and $\wei_{\Sigma, \Delta}$
-from $Q \times Q$ into respectively~$\Semiring$, %$\Phi_\epsilon$, 
-$\Phi_\Sigma$, $\Phi_\Delta$, and $\Phi_{\Sigma, \Delta}$.
-\end{definition}
-%
-We summarize the above 4-uplet of transition functions into a unique 
-function~$\wei$ from
-$Q \times (\Sigma \cup \{ \epsilon \}) \times (\Delta \cup \{ \epsilon \}) \times Q$
-into $\Semiring$, %also called $\wei$ for simplicity, 
-such that, for all $q, q' \in Q$, $a \in \Sigma$, and $b \in \Delta$, 
-%and with 
-%$\< \phi_\epsilon, \phi_\Sigma, \phi_\Delta, \phi_{\Sigma, \Delta}> = \wei(q, q')$, 
-\[
-\begin{array}{rcll}
-\wei(q, \epsilon, \epsilon, q') & = &  \wei_\epsilon(q, q'),\\ %\phi_\epsilon\\
-\wei(q, a, \epsilon, q') & = & \phi_\Sigma(a) & %\wei_\Sigma(q, q')(a)
-\quad\mathrm{where~} \phi_\Sigma = \wei_\Sigma(q, q'),\\
-\wei(q, \epsilon, b, q') & = & \phi_\Delta(b) & 
-\quad\mathrm{where~} \phi_\Delta = \wei_\Delta(q, q'),\\
-\wei(q, a, b, q') & = & \phi_{\Sigma, \Delta}(a, b) & 
-\quad\mathrm{where~} \phi_{\Sigma, \Delta} = \wei_{\Sigma, \Delta}(q, q').\\
-\end{array}      
-\]
-%More precisely, $Q \times Q$, 
-%resp. $Q \times \Sigma \times Q$,
-%$Q \times \Delta \times Q$,
-%$Q \times \Sigma \times \Delta \times Q$,
-%into resp. $\Phi_0$, $\Phi_\Sigma$, $\Phi_\Delta$ $\Phi_{\Sigma, \Delta}$.
-
-The symbolic-weighted transducer $T$ defines a mapping 
-from the pairs of strings of $\Sigma^* \times \Delta^*$ 
-into the weights of~$\Semiring$,
-based on the following intermediate function $\weight_T$
-defined recursively for every $q, q' \in Q$, 
-%$a \in \Sigma$, $b \in \Delta$ 
-for every strings of $s \in \Sigma^*$, $t \in \Delta^*$:
-\[
-\begin{array}{rccl}
-xc(q, s, t, q') & = & 
-   \displaystyle\bigoplus_{q'' \in Q} &
-   \wei(q, \epsilon, \epsilon, q'') \otimes \weight_T(q'', s, t, q')\\
-   & \oplus & \displaystyle\bigoplus_{q'' \in Q} &
-              \wei(q'', \epsilon, \epsilon, q') \otimes \weight_T(q, s, t, q'')\\
- & \oplus & \displaystyle\bigoplus_{\begin{array}{c}
-                                      \scriptstyle q'' \in Q\\[-2pt]
-                                      \scriptstyle s = au, a \in \Sigma
-                                      \end{array}} &
-   \wei(q, a, \epsilon, q'') \otimes \weight_T(q'', u, t, q')\\
- & \oplus & \displaystyle\bigoplus_{\begin{array}{c}
-                                    \scriptstyle q'' \in Q\\[-2pt]
-                                    \scriptstyle t = bv, b \in \Delta
-                                    \end{array}} &
-    \wei(q, \epsilon, b, q'') \otimes \weight_T(q'', s, v, q')\\ 
- & \oplus & \displaystyle\bigoplus_{\begin{array}{c}
-                                    \scriptstyle q'' \in Q\\[-2pt]
-                                    \scriptstyle s = au, a \in \Sigma\\[-2pt]
-                                    \scriptstyle t = bv, b \in \Delta
-                                    \end{array}} &
-   \wei(q, a, b, q'') \otimes \weight_T(q'', u, v, q').\\ 
-\end{array}
-\]
-The sum above can be infinite because of the expression in the first line, 
-however, $\T(\Omega)$ is well defined since $\Semiring$ is complete.
-%
-As we shall see, epsilon-transitions can be removed and hence the sum made finite.
-%
-%** The application of $\oplus$ can be seen as a ND choice between transitions **
-We recall that, by convention, an empty sum with $\oplus$ is $\zero$. 
-%
-Since $\zero$ is absorbing for~$\otimes$ in~$\Semiring$,
-one term $\wei(q, a, b, q'')$ equal to $\zero$ in the above expression 
-will be ignored in the sum, meaning that there is no possible transition
-from state $q$ into state $q'$ while reading $a$ and writing $b$ 
-(including the case $a = \epsilon$ and $a = \epsilon$).
-%These functions $\phi$ act as guards for the transducer's transitions, 
-%preventing a transition when they return the absorbing $\zero$ of $\Semiring$.
-This is analogous to a transition's guard not satisfied by $\<a, b>$ in 
-the case of symbolic transducers.
-
-The above expression of $\weight_T$ can be seen as a stateful definition of 
-an edit-distance between a word $s \in \Sigma^*$ and a word $t \in \Delta^*$,
-see also~\cite{Mohri03ijfcs}.
-Intuitively, 
-$\wei(q, a, \epsilon, q'')$, in the third line, is the cost of 
-the deletion of a symbol~$a \in \Sigma$ in~$s$, 
-$\wei(q, \epsilon, b, q'')$, in the fourth line, is the cost 
-of the deletion of insertion of~$b \in \Delta$ in~$s$, 
-and $\wei(q, a, b, q'')$, in the last line, is the cost 
-of the substitution of  $a \in \Sigma$ by~$b \in \Delta$.
-The cost of a sequence of such operations transforming $s$ into $t$, 
-is the product, with $\otimes$, of the individual costs of the operations involved;
-And the distance between $s$ and $t$ is the sum, with $\oplus$,
-of all such product of costs.
-
-\medskip\noindent
-The weight associated by $T$ to  $\< s, t> \in \Sigma^* \times \Delta^*$
-is then defined as follows: 
-\begin{equation}
-T(s, t)  = 
-\displaystyle\bigoplus_{q, q' \in Q} \mathsf{in}(q) 
-\mathop{\otimes} \weight_T(q, s, t, q') \mathop{\otimes} \mathsf{out}(q').
-\label{eq:weightT}
-\end{equation}
-
-\begin{example}
-DTW for sequences of timestamped events **     
-\end{example}
-
-\begin{example}
-(simpler) pointwise distance between two sequences of timestamped events **
-\end{example}
-
-\noindent
-The \emph{Symbolic Weighted Automata} %$A = \< Q, \init, \weight, \final >$
-over $\Sigma$ and $\Semiring$ 
-are defined similarly as the transducers of Definition~\ref{def:SWT}, 
-by simply omitting the output symbols.
-%
-In this case, a label theory over $\Sigma$ is reduced to
-a set $\Phi_\Sigma$ closed under~$\oplus$ and~$\otimes$.
-%
-\begin{definition} \label{def:SWA}
-A \emph{symbolic-weighted automaton} (\SWA)
-over the input alphabet~$\Sigma$
-and the commutative semiring $\Semiring$ is a tuple
-$A = \< Q, \init, \bar{\wei}, \final >$,
-where $Q$ is a finite set of states, 
-$\mathsf{in} : Q \to \Semiring$, 
-respectively $\mathsf{out} : Q \to \Semiring,$
-are functions defining the weight for entering, 
-respectively leaving, a state, 
-and $\bar{\wei}$ is a pair of transition functions 
-$\wei_\epsilon$ and $\wei_\Sigma$ from $Q \times Q$ into 
-respectively $\Semiring$ %$\Phi_\epsilon$ 
-and $\Phi_\Sigma$.
-\end{definition}
-
-\noindent
-The above transition functions $\wei_\epsilon$ and $\wei_\Sigma$
-are synthesized into a function~$\wei$ 
-from~$Q \times (\Sigma \cup \{ \epsilon \}) \times Q$ into~$\Semiring$ 
-as above.
-%
-When $\wei_\epsilon(q, q') = \zero$ for all $q, q' \in Q$, 
-the automaton~$A$ is called \emph{without $\epsilon$-transitions}.
-
-\subsection{Properties}
-The two following properties will be useful to our approach on 
-symbolic weighted parsing in Section~\ref{sec:parsing}.
-
-\begin{proposition}
-Given a \SWT $T$ 
-over alphabet $\Sigma$, $\Delta$ and a commutative semiring~$\Semiring$, 
-and $s \in \Sigma^*$, 
-there exists an effectively constructible \SWA 
-$A_{T, s}$ over $\Delta$ and $\Semiring$ 
-such that for all $t \in \Delta^*$, $A_{T, s}(t) = T(s, t)$.
-\end{proposition}
-%
-\begin{proof}
-Let $T = \< Q, \init, \bar{\wei}, \final >$ 
-where $\bar{\wei} \< \wei_\epsilon, \wei_\Sigma, \wei_\Delta, \wei_{\Sigma, \Delta}>$
-has an associated synthesized form $\wei$, 
-and let $s = s_1 \ldots s_n$ with $s_1, \ldots, s_n \in \Sigma$.
-%$n = |s|$ be the length of $s$.
-%
-Let $Q' = [1..n] \times Q$ be the state set of $A_{T, s}$,
-and let, for all $q \in Q$, 
-$\init'(1, q) = \init(q)$ and $\init'(i, q) = \zero$ for all $1 < i \leq n$, 
-and $\final'(n, q) = \final(q)$ and $\final'(i, q) = \zero$ for all $1 \leq i <  n$.
-
-\noindent
-We define the functions $\wei'_\epsilon$ and $\wei'_\Sigma$ for $A_{T, s}$
-in their synthesized form: for all $1 \leq i, j \leq n$, $q, q' \in Q$ and $a \in \Sigma$, 
-and $b \in \Delta$, 
-\[
-\begin{array}{lclcl}
-\wei'(\< i, q>, \epsilon, \< i, q'>)   & = & \wei(q, \epsilon, \epsilon, q'), & \quad & \\ 
-\wei'(\< i, q>, \epsilon, \< i+1, q'>) & = & \wei(q, s_i, \epsilon, q') & & \mathrm{if~} i < n,\\ 
-\wei'(\< i, q>, \epsilon, \< j, q'>)   & = & \zero & & \mathrm{if~} j \neq i, j \neq i+1,\\ 
-\wei'(\< i, q>, b, \< i, q'>)          & = & \wei(q, \epsilon, b, q'),\\ 
-\wei'(\< i, q>, b, \< i+1, q'>)        & = & \wei(q, s_i, b, q') & & \mathrm{if~} i < n,\\ 
-\wei'(\< i, q>, b, \< j, q'>)          & = & \zero & & \mathrm{if~} j \neq i, j \neq i+1. 
-\end{array}      
-\]
-Note that the functions defined above belong to the set $\Phi_\Delta$ 
-of the underlying label theory, 
-by the hypothesis of closure under partial application.
-%
-The $\SWA$ wanted is $A_{T, s} = \< Q', \init', \bar{\wei'}, \final' >$.
-\qed
-\end{proof}
-
-The construction time and size for $A_{T, s}$ are $O(\| T \| . | s |)$,
-where the size $\| T \|$ of $T$ is its number of states $|Q|$.
-
-% gen. to transducers
-\begin{proposition}
-Given a \SWA $A$ over alphabet $\Sigma$ 
-and a commutative and bounded semiring $\Semiring$, 
-there exists an effectively constructible \SWA $A'$ 
-without $\epsilon$-transitions 
-such that for all $s \in \Sigma^*$, $A'(s) = A(s)$.
-\end{proposition}
-%
-\begin{proof}  ** Tb revised **   
-Let $A = \< Q, \init, \<\wei_\epsilon, \wei_\Sigma>, \final >$.
-The automaton $A'$ is $\< Q, \init, \<\wei'_\epsilon, \wei'_\Sigma>, \final >$,
-where, for all $q, q' \in Q$,
-$\wei'_\epsilon(q, q') = \zero$  
-and 
-% \[
-% \begin{array}{rcccl}
-% \wei'_\Sigma(q, q') & = &  & \displaystyle\bigoplus_{q_1 \in Q} &
-%                \wei_\Sigma(q, q_1) \otimes  \wei_\epsilon(q_1, q')\\
-%                     &   & \oplus  & \displaystyle\bigoplus_{q_2 \in Q} &
-%                \wei_\epsilon(q, q_2) \otimes \wei_\Sigma(q_2, q')\\
-% \end{array}
-% \]
-\[
-\wei'_\Sigma(q, q') = 
-\displaystyle\bigoplus_{q'' \in Q} \wei_\epsilon(q, q'') \otimes \wei_\Sigma(q'', q')
-\oplus \displaystyle\bigoplus_{q_1 \in Q} \wei_\Sigma(q, q'') \otimes \wei_\epsilon(q'', q')
-\]
-In the above definition of $\wei'_\Sigma$ we use the operator
-of product of function of $\Phi_\Sigma$ by $\Semiring$ 
-described in Section~\ref{section:symbols}.
-%
-By definition of $\weight_{A}$ and distributivity of $\otimes$ on $\oplus$, 
-** NO. TBC see \cite{Mohri02ijfcs} **
-it holds that $\weight_{A}(q, s, q') = \weight_{A'}(q, s, q')$.
-\qed
-\end{proof}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SW-VPA
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{SW Visibly Pushdown Automata}
-\label{section:SWVPA}\label{sec:SWVPA}
-The following model generalizes Symbolic VPA~\cite{dAntonyAlur14SVPDA}
-from Boolean semirings to arbitrary semiring weight domains.
-
-\subsection{Definition}
-Let $\Omega$ be a countable alphabet 
-%finite (large) or infinite,
-that we assume partitioned into three subsets 
-$\Omegai$, $\Omegac$, $\Omegac$,
-whose elements are respectively called 
-\emph{internal}, \emph{call} and \emph{return} symbols.
-% \begin{itemize}
-% \item a set $\Omegai$ of \emph{internal symbols} denoted $a$,
-% \item a set $\Omegac$ of \emph{call symbols} denoted $\call{a}$,
-% \item a set $\Omegar$ of \emph{return symbols} denoted $\return{a}$.
-% \end{itemize}
-Let~$\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative semiring and  
-let  $(\Phi_\epsilon, \Phic, \Phir, \Phicr)$ be a label theory over $\Semiring$
-%In order to simplify notations, %and following the definition of Section~\ref{section:transducer}, 
-%we shall write respectively 
-where $\Phic$, $\Phir$ and~$\Phicr$ stand respectively 
-for~$\Phi_\Omegac$, $\Phi_\Omegar$ and~$\Phi_{\Omegac, \Omegar}$.
-%
-Moreover, we extend this theory with a set $\Phii$ 
-of unary functions in $\Omegai \to \Semiring$,
-closed under $\oplus$ and $\otimes$.
-
-\begin{definition}
-A \emph{Symbolic Weighted Visibly Pushdown Automata} (\SWVPA) 
-over  $\Omega = \Omegai \uplus \Omegac \uplus \Omegar$ and $\Semiring$ is a tuple
-$A = \< Q, P, \init, \weii, \weic, \weir, \weie, \final >$,
-where $Q$ is a finite set of states, 
-$P$ is a finite set of stack symbols, 
-$\mathsf{in} : Q \to \Semiring$, 
-respectively $\mathsf{out} : Q \to \Semiring,$
-are functions defining the weight for entering, 
-respectively leaving, a state, 
-and 
-$\weii : Q \times Q \to \Phii$,  
-$\weic : Q \times Q \times P \to \Phic$,  
-$\weir : Q \times P \times Q \to \Phicr$,  
-$\weie : Q \times Q \to \Phir$,  
-are transition functions.
-\end{definition}
-%
-Similarly as in Section~\ref{section:transducer}, 
-we extend the above transition functions as follows
-for all $q, q' \in Q$, $p \in P$, 
-$a \in \Omegai$, 
-$\call{c} \in \Omegac$, 
-$\return{r} \in \Omegar$, 
-overloading their names: % for simplicity:
-\[
-\begin{array}{lll}
-\weii: Q \times \Omegai \times Q \to \Semiring & 
-\weii(q, a, q') = \phi_\mathsf{i}(a) & 
-\mathrm{where~} \phi_\mathsf{i} = \weii(q, q'),\\
-%
-\weic: Q \times \Omegac \times Q \times P \to \Semiring & 
-\weic(q, \call{c}, q', p) = \phi_\mathsf{c}(\call{c}) & 
-\mathrm{where~} \phi_\mathsf{c} = \weic(q, q', p),\\
-%
-\weir: Q \times \Omegac \times P \times \Omegar \times Q \to \Semiring & 
-\weir(q, {\call{c}},  p, {\return{r}}, q') = \phi_\mathsf{r}({\call{c}},  {\return{r}}) & 
-\mathrm{where~} \phi_\mathsf{r} = \weir(q, p, q'),\\
-%
-\weie: Q \times \Omegar \times Q \to \Semiring & 
-\weie(q, {\return{r}}, q') = \phi_\mathsf{e}({\return{r}}) &
-\mathrm{where~} \phi_\mathsf{e} = \weie(q, q').\\
-\end{array}      
-\]
-
-\noindent
-The intuition is the following for the above transitions.
-\begin{description}
-\item $\weii$ : read the input internal symbol $a$, change state to $q'$ (stack is untouched).
-\item $\weic$ : read the input call symbol $\call{c}$, push it to the stack along with $p$, change state to $q'$.
-\item $\weir$ : when the stack is not empty, 
-      read and pop from stack a pair made of $\call{c}$ and $p$, 
-      read the input return symbol $\return{r}$, change state to $q'$.
-      In this case, the weight function $\phi_\mathsf{r}$ 
-      computes a value of matching between the call and return symbols $c$ and $r$.
-      This value might be set to $\zero$ in order to express that the symbols do not match.
-\item $\weie$ : when the stack is empty, 
-      read the input symbol $\call{r}$, change state to $q'$.
-\end{description}
-
-We give now a formal definition of these transitions of the automaton~$A$
-in term of a weight value
-computed by an intermediate function $\weight_A$, like in Section~\ref{sec:SWT}.
-In the case of a pushdown automaton, a configuration is composed 
-of a state $q \in Q$ and a stack content $\gamma \in \Gamma^*$, where $\Gamma = \Omegac \times P$.
-Therefore, $\weight_A$ is a function from 
-$Q \times \Gamma^* \times \Omega^* \times Q \times \Gamma^*$ into $\Semiring$.
-\[
-\begin{array}{rcl}
-\weight_A\bigl(\config{q}{\gamma}, a\, u, \config{q'}{\gamma'}\bigr) & = & 
- {\displaystyle\bigoplus_{q'' \in Q}} \weii(q, a, q'') 
- \otimes \weight_A\bigl(\config{q''}{\gamma}, u, \config{q'}{\gamma'}\bigr)\\
-%
-\weight_A\bigl(\config{q}{\gamma}, {\call{c}}\, u, \config{q'}{\gamma'}\bigr) & = & 
- {\displaystyle\bigoplus_{\begin{array}{c}
-                          \scriptstyle q'' \in Q\\[-2pt]
-                          \scriptstyle p \in P
-                          \end{array}}}
- \weic\bigl(q, {\call{c}}, q'', p\bigr) 
- \otimes \weight_A\bigl(\config{q''}{{\call{c}}\, p\cdot \gamma}, u, \config{q'}{\gamma'}\bigr)\\[1mm]
-%
-\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot \gamma}, {\return{r}}\, u, \config{q'}{\gamma'}\bigr) & = & 
- {\displaystyle\bigoplus_{q'' \in Q}} 
- \weir\bigl(q, {\call{c}}, p, {\return{r}}, q''\bigr) 
- \otimes \weight_A\bigl(\config{q''}{\gamma}, u, \config{q'}{\gamma'}\bigr)\\
-%
-\weight_A\bigl(\config{q}{\bot}, {\return{r}}\, u, \config{q'}{\gamma'}\bigr) & = & 
- {\displaystyle\bigoplus_{q'' \in Q}} \weie(q, {\return{r}}, q'') 
- \otimes \weight_A\bigl(\config{q''}{\bot}, u, \config{q'}{\gamma'}\bigr)\\
-\end{array}
-\]
-where $\bot$ denotes the empty stack and ${\call{c}}\, p\cdot \gamma$ 
-denotes a stack where the pair made of ${\call{c}} \in \Omegac$ and $p \in P$ is the top symbol 
-and $\gamma$ is the rest of stack.
-
-\noindent
-The weight associated by $A$ to $s \in \Omega^*$
-is then defined as follows,
-following empty stack semantics: 
-\begin{equation}
-A(s)  = 
-{\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
-\mathsf{in}(q) \mathop{\otimes} 
-\weight_A\bigl(\config{q}{\bot}, s, \config{q'}{\bot}\bigr) 
-\mathop{\otimes} \mathsf{out}(q').
-\label{eq:weightA}
-\end{equation}
-
-\begin{example}
-structured words...
-intro language of music notation ?
-\end{example}
-
-\subsection{Properties}
-Like VPA and symbolic VPA, 
-the class of \SWVPA is closed under the binary operators of the underlying semiring.
-%
-\begin{proposition}
-Let $A_1$ and $A_2$ be two (\SWVPA)
-over the same $\Omega$ and $\Semiring$.
-There exists two $\SWVPA$ $A_1 \oplus A_2$ and $A_1 \otimes A_2$, 
-effectively constructible, 
-such that for all $s \in \Omega^*$, 
-$(A_1 \oplus A_2)(s) = A_1(s) \oplus A_2(s)$ and 
-$(A_1 \otimes A_2)(s) = A_1(s) \otimes A_2(s)$.
-\end{proposition}
-The construction is essentially the same 
-as in the case of the Boolean semiring~\cite{dAntonyAlur14SVPDA}.
-
-\subsection{Best-first Search} 
-\label{sec:best}\label{sec:search}
-Let us assume that the semiring $\Semiring$ is
-commutative, bounded, complete, and total.
-% total ? 
-% monotonic and superior writ natural ordering
-%Regarding the infinite sum operator, note that
-%$\bigoplus_{x \in \Phi_\Omega} \phi(x)$, 
-%$\bigoplus_{y \in \Phi_\Delta} \psi(y)$, and  
-%... exist and in $\Semiring$.
-We propose a Dijkstra algorithm computing the minimal weight by $A$, 
-\wrt $\leq_\oplus$, for a word in $\Semiring^*$.
-
-\noindent
-More precisely, 
-let $\top$ be a fresh stack symbol which does not belong to $\Gamma$,
-and for every two states $q, q' \in Q$ and $\sigma \in \{ \bot, \top \}$, 
-let 
-\[ 
-  d_0(q, \sigma, q') = \bigoplus_{s\in \Omega^*} 
-  \textstyle
-  \weight_A\bigl(\config{q}{\sigma}, s, \config{q'}{\sigma}\bigr). 
-\]
-Since $\Semiring$ is complete, this infinite sum is well defined, 
-and since $\leq_\oplus$ is assumed total, it is the minimum in $\Omega^*$
-of $s \mapsto \weight_A(\config{q}{\sigma}, s, \config{q'}{\sigma})$ \wrt this ordering.
-%
-When $\sigma = \bot$, $d_0(q, \sigma, q')$ is the central 
-expression in a term of the definition (\ref{eq:weightA}) of $A(s_0)$ for the minimum $s_0$
-(for the above function).
-When $\sigma = \bot$, intuitively, it is the minimum weight of a computation of $A$
-starting in state $q$ with a stack $\gamma \in \Gamma^*$ (possibly empty), 
-and ending in state $q$ with the same stack $\gamma$, such that moreover 
-the computation does not touch a symbol of $\gamma$. 
-That means that during the computation, 
-$A$ may apply the first case of in the definition of $\weight_A$ (internal symbol),
-as well as the second case, to can push call symbols on the top of $\gamma$, 
-and may pop these symbols with the third case (return symbol). 
-However, it cannot apply one of the two last cases (return symbol and empty stack)
-when the current stack is $\gamma$.
-%pop symbols in $\gamma$.
-% Note that having a stack reduced to such a symbol makes impossible the application of the 
-% two last cases in the definition of $\weight_A$ (return symbol and empty stack). 
-% However, it is possible to apply the two first cases 
-% (internal symbol or call symbol, with a push on the top of $\top$).
-
-The algorithm~\ref{algo:Dijkstra}
-constructs iteratively a marking $d : Q \times \{ \bot, \top \} \times Q \to \Semiring$
-%of the triplets $\<q, \sigma, q'>$ 
-%of states of $A$ by weight values in $\Semiring$, 
-that converges eventually to $d_0(q, \sigma, q')$. 
-%It uses for that purpose a priority queue $P$ containing triplets of 
-%$Q \times \{ \bot, \top \} \times Q$.
-
-\begin{algo}[1-best for \SWVPA] \label{algo:Dijkstra}
-\textbf{initially} let $P = Q \times \{ \bot, \top \} \times Q$, %contains all %triplets 
-%$\< q_1, \bot, q_2>$ and $\<q_1, \top, q_2>$ for $q_1, q_2 \in Q$, with 
-and 
-$d(q_1, \bot, q_2) = d(q_1, \top, q_2) = \one$ if $q_1 = q_2$ and
-$d(q_1, \bot, q_2) = d(q_1, \top, q_2) = \zero$ otherwise.
-
-\smallskip\noindent
-\textbf{while} $P$ is not empty
-
-\noindent
-\quad\textbf{extract} $\< q_1, \sigma, q_2>$ from $P$ 
- such that $d(q_1, \sigma, q_2)$ is minimal wrt $\leq_\oplus$.
-
-\noindent\quad \textbf{for all} $q_0, q_3 \in Q$ and $p \in P$ do
-
-\noindent
-\qquad
-\(\begin{array}{rrcl}
- & d(q_1, \sigma, q_3) & \opluseq &
-   d(q_1, \sigma, q_2) \otimes \displaystyle\bigoplus_{a \in \Omegai} \weii(q_2, a, q_3)\\
-%
-%  & d(q_0, \sigma, q_2) & \opluseq & 
-%    \displaystyle\bigoplus_{a \in \Omegai} \weii(q_0, a, q_1) \otimes d(q_1, \sigma, q_2)\\
-%
-\mathrm{if~} \sigma = \top & d(q_0, \top, q_3) & \opluseq &
-d(q_1, \sigma, q_2) \otimes
-%\displaystyle\bigoplus_{p \in P} 
-\displaystyle\bigoplus_{{\call{c}} \in \Omegac} 
-\displaystyle\bigoplus_{{\return{r}} \in \Omegar} \eta({\call{c}}, {\return{r}})\\
-\mathrm{and} &  d(q_0, \bot, q_3) & \opluseq &
-d(q_1, \sigma, q_2) \otimes
-%\displaystyle\bigoplus_{p \in P} 
-\displaystyle\bigoplus_{{\call{c}} \in \Omegac} 
-\displaystyle\bigoplus_{{\return{r}} \in \Omegar} \eta({\call{c}}, {\return{r}})\\
-& & & \mathrm{where~} \eta = \weic(q_0, q_1, p) \otimes \weir(q_2, p, q_3)\\
-%
-\mathrm{if~} \sigma = \bot & d(q_1, \bot, q_3) & \opluseq &
-  d(q_1, \sigma, q_2) \otimes \displaystyle\bigoplus_{\return{r} \in \Omegar} \weie(q_2, {\return{r}}, q_3)\\
-%
-%& d(q_0, \bot, q_2) & \opluseq & 
-%d(q_0, \bot, q_1) \otimes d(q_1, \sigma, q_2), \mathrm{if~} \< q_0, \bot, q_1> \notin P \\
-%
-%\mathrm{if~} \sigma = \top & d(q_0, \top, q_2) & \opluseq & 
-%d(q_0, \top, q_1) \otimes d(q_1, \sigma, q_2), \mathrm{if~} \< q_0, \top, q_1> \notin P \\
-%
-& d(q_1, \bot, q_3) & \opluseq & 
-   d(q_1, \sigma, q_2) \otimes d(q_2, \bot, q_3), \mathrm{if~} \< q_2, \bot, q_3> \notin P \\
-%
-\mathrm{if~} \sigma = \top & d(q_1, \top, q_3) & \opluseq & 
-    d(q_1, \sigma, q_2) \otimes d(q_2, \top, q_3), \mathrm{if~} \< q_2, \top, q_3> \notin P \\
-\end{array}
-\) 
-\end{algo}
-
-\noindent
-The infinite sums in the updates of $d$ in Algorithm~\ref{algo:Dijkstra} are well defined
-since~$\Semiring$ is complete.
-The algorithm performs $2.|Q|^2$ iterations until $P$ is empty, 
-and each iteration has a time complexity $O(|Q|^2 . |P|)$.
-This gives a time complexity $O(|Q|^4 . |P|)$. 
-It can be reduced by implementing $P$ as a priority queue, 
-prioritized by the value returned by $d$
-***complete***. %$|Q|^3.\log(|Q|^2)$
-
-The correctness of Algorithm~\ref{algo:Dijkstra} 
-is ensured by the invariant expressed in the following lemma.
-\begin{lemma}
-For all $\< q_1, \sigma, q_2> \notin P$, $d(q_1, \sigma, q_2) =  d_0(q_1, \sigma, q_2)$.
-\end{lemma}
-The proof is by contradiction, 
-assuming a counter-example minimal in the length of the witness word.
-
-% \begin{algorithm}
-% \caption{<your caption for this algorithm>}
-% \label{<your label for references later in your document>}
-% \begin{program}
-% \seq{|line| := `` '', i := 1};
-% \WHILE i \neq n+1 \DO
-%       |line| := |item|[i] \concat `` '' \concat |number|[i]; 
-%       i := i+1; 
-%       \WHILE i \neq n+1 \AND |item|[i] = |item|[i-1] \DO 
-%       |line| := |line| \concat ``, '' \concat |number|[i]);
-%       i := i+1 \OD ; 
-%       |write|(|line|) \OD 
-% \end{program}
-% \end{algorithm}
-
-% The weight of a transition acts as a guard: 
-% a transition is activated for a symbol $`a`$ iff its weight $`\phi(a) \neq 0`$ (or $`\psi(a, b) \neq 0`$), 
-% the absorbing element.
-
-%\begin{lemma}
-\noindent
-For computing the minimal weight of a computation of $A$, we use the fact that,
-at the termination of Algorithm~\ref{algo:Dijkstra}, %it holds that,
-%There exist $q_1, q_2 \in Q$
-\[
-  {\displaystyle \bigoplus_{s \in \Omega^*} A(s)} = 
-  {\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
-  \mathsf{in}(q) \mathop{\otimes} d(q, \bot, q') \mathop{\otimes} \mathsf{out}(q').
-\]
-%\end{lemma}
-
-%\medskip
-\noindent
-In order to obtain effectively a witness 
-(word of $\Semiring^*$ with computation of $A$ of minimal weight), 
-we require an additional property the of weight functions.
-\begin{definition}
-Let $\Omega$ be an alphabet and $\Semiring$ a complete semiring.
-A function $\phi$ from $\Omega^n$ into $\Semiring$
-is called $k$-\emph{convex} for a natural number $k$ iff 
-$\mathit{card}\{ \mathbf{a} \in \Omega^n \mid \phi(\mathbf{a}) = 
-                 \displaystyle\bigoplus_{\mathbf{p} \in \Omega^n} \phi(\mathbf{p}) \} \leq k$.
-\end{definition}
-A label theory is $k$-convex if all its functions are $k$-convex.
-
-\begin{proposition}
-For a \SWVPA $A$ over a 
-commutative, idempotent, superior, total and complete semiring 
-and an alphabet $\Omega$ with a $k$-convex label theory,
-one can construct in PTIME a word $s \in \Omega^*$ 
-such that $A(s)$ is minimal \wrt the natural ordering for $\Semiring$. 
-%$\leg_\oplus$.
-\end{proposition}
-
-\subsection{Trees and Nested-Words}
-\label{sec:trees}
-The hierarchical structure of nested-words, defined the \emph{call} and \emph{return} markup symbols of 
-suggest a correspondence between these word and trees. 
-The lifting of this correspondence to languages, respective of tree automata and VPA have
-been discussed in~\cite{AlurMadhusudan09nested}, 
-see also~\cite{Caralp12VPAmult} for the weighted case.
-In this section, we discuss the correspondence between the symbolic-weighted extensions
-of tree automata and VPA.
-
-Let $\Omega$ be a countable ranked alphabet, such that 
-every symbol $a \in \Omega$ has a rank 
-$\rank(a) \in [0..M]$ where $M$ is a fixed natural number.
-We write $\Omega_k$ the subset of all symbols $a$ of $\Omega$
-with $\rank(a) = k$, where $0 \leq k \leq M$.
-%$\cup_{i = 0}^{i = M} \Omega_i$ 
-%
-$\T(\Omega)$ denotes the free $\Omega$-algebra of finite, ordered, 
-$\Omega$-labeled trees, 
-which is the smallest set such that  $\Omega_0 \subset \T(\Omega)$
-and for all $1 \leq k \leq M$, all $a \in \Omega_k$, 
-and all $t_1, \ldots, t_k \in \T(\Omega)$, $a(t_1, \ldots, t_k) \in \T(\Omega)$.
-%
-% tree = single node (leave) labeled with a symbol of $a \in \Omegai$
-% (such a tree is simply denoted by $a$)
-% or the composition, denoted by $b(t_1,\ldots, t_n$) of a node labeled with $b$
-% and $n$ subtrees $t_1$,\ldots, $t_n$.
-
-\renewcommand{\call}[1]{\ensuremath \langle_{#1}}
-\renewcommand{\return}[1]{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
-
-Let $\hat\Omega$ be the countable (unranked) alphabet obtained from 
-$\Omega$ as follows: 
-$\hat\Omega = \< \Omegai, \Omegac, \Omegar >$, with
-$\Omegai = \Omega_0$, 
-$\Omegac = \{ \; \call{a} \mid a \in \Omega_{>0} \}$,
-$\Omegac = \{ \; \return{a} \mid a \in \Omega_{>0} \}$,
-where $\Omega_{>0}$ denotes $\bigcup_{k = 1}^{M} \Omega_k$.
-
-\noindent
-We define a linearization of trees of $\T(\Omega)$ into 
-words of $\hat{\Omega}^*$ as follows:
-\begin{description}
-\item $\lin(a) = a$ for all $a \in \Omega_0$, 
-\item $\lin\bigl( b(t_1, \ldots, t_k)\bigr) = 
-       \call{b} \; \lin(t_1) \ldots \lin(t_k) \; \return{b}$ 
-       when $b \in \Omega_k$, $1 \leq k \leq M$.
-\end{description}
-
-% A \emph{regular tree grammar} over $\Omega$ 
-% is a triplet $G = \< N, q_\mathsf{i}, R>$ where
-% $N$ is a finite set of non-terminal symbols denoted $q$..., 
-% $q_\mathsf{i} \in N$ is the starting non-terminal, 
-% $R$ is a finite set of production rules of the form
-% $q_0 \to a(q_1\ldots q_k)$ where 
-% $q_0, q_1, \ldots, q_k \in N$
-% $a \in \Omega_k$.
-% A tree $t \in \T(\Omega)$ is in the language of $G$ 
-% if it can be generated from $q_\mathsf{i}$ by 
-% non terminal replacement following the rules of $R$.
-
-Let us assume a label theory $\Phi_{\Omega_k}$ 
-for each $k \in [0..M]$.
-%
-\begin{definition}  \label{def:SWTA}
-A \emph{symbolic-weighted tree automaton} (\SWTA)
-over the ranked input alphabet~$\Omega$
-and the commutative semiring $\Semiring$ 
-is a triplet $A = \< Q, \init, \bar{\wei} >$ where
-$Q$ is a finite set of states, 
-$\mathsf{in} : Q \to \Semiring$ is the starting weight function, 
-and $\bar{\wei}$ is a $M+2$-uplet of transition functions made of:
-$\wei_\epsilon$ from $Q \times Q$ into $\Semiring$, and, 
-for each $k \in [0..M]$, $\wei_{\Omega, k}$
-from $Q \times Q^{k}$ into $\Phi_{\Omega_k}$.
-\end{definition}
-%
-Like in Section~\ref{sec:SWAdef}, we define from $\bar{\wei}$ 
-a transition function~$\wei$, %from $\bar\wei$
-from 
-$Q \times (\Sigma \cup \{ \epsilon \}) \times \bigcup_{k=0}^{M} Q^k$
-into~$\Semiring$: %also called $\wei$ for simplicity, 
-%such that, for all $q, q' \in Q$, $a \in \Sigma$, and $b \in \Delta$, 
-\[
-\begin{array}{rcll}
-\wei(q_0, \epsilon, q_1) & = &  \wei_\epsilon(q_0, q_1),\\ %\phi_\epsilon\\
-\wei(q_0, a, q_1 \ldots q_k) & = & \phi_{\Omega, k}(a) &
-\quad\mathrm{where~} \phi_{\Omega, k} = \wei_{\Omega, k}(q_0, q_1\ldots q_k).\\
-\end{array}      
-\]
-Intuitively, $\wei(q_0, a, q_1 \ldots q_k)$ can be seen as
-the weight of a production rule $q_0 \to a(q_1, \ldots, q_k)$ 
-of a regular tree grammar~\cite{tata}, 
-that replaces the non-terminal symbol $q_0$, by $a(q_1, \ldots, q_k)$. 
-%in a step of tree building.
-%
-Such a grammar computes the weights of the derivation trees 
-of the Context-Free grammar obtained by forgetting the labeling symbols of $\Omega_{>0}$.
-
-\noindent
-The \SWTA of Definition~\ref{def:SWTA} defines a mapping 
-from trees of $\T(\Omega)$ into the weights of~$\Semiring$,
-based on the intermediate function $\weight_A$
-defined as follows for $q_0 \in Q$ and 
-$t = b(t_1,\ldots, t_k) \in \T(\Omega)$,
-åwith $0 \leq k \leq M$:
-\[
-\begin{array}{rccl}
-\weight_A(q_0, t) & = & 
-   \displaystyle\bigoplus_{q_1 \in Q} &
-   \wei(q, \epsilon, q_1) \otimes \weight_A(q_1, t)\\
- & \oplus & \displaystyle\bigoplus_{q_1 \ldots q_k \in Q^k} &
-              \wei(q_0, b, q_1 \ldots q_k ) 
-   \otimes \displaystyle\bigotimes_{i=1}^{k}
-           \weight_A(q_i, t_i).\\
-\end{array}
-\]
-
-\medskip\noindent
-The weight associated by $A$ to  $t \in \T(\Omega)$ is then
-\begin{equation}
-A(t)  = 
-\displaystyle\bigoplus_{q \in Q} \mathsf{in}(q) \mathop{\otimes} \weight_A(q, t).
-\label{eq:weightTA}
-\end{equation}
-
-\begin{lemma}  \label{lem:SWTA}
-For all \SWTA $A$ over~$\Omega$ and $\Semiring$, without $\epsilon$-transitions,
-there exists an effectively constructible \SWVPA $A'$ over 
-$\hat\Omega$ and $\Semiring$
-such that for all $t \in \T(\Omega)$, $A'\bigl(\lin(t)\bigr) = A(t)$.
-\end{lemma} 
-% 
-\begin{proof}
-Let $A = \< Q, \init, \bar{\wei} >$ where $\bar{\wei}$ is summarized as above by a function
-$\wei: Q \times (\Sigma \cup \{ \epsilon \}) \times \bigcup_{k=0}^{M} Q^k \to \Semiring$.
-
-We build 
-$A' = \< Q', P', \init', \weii, \weic, \weir, \weie, \final' >$,
-computing over $\hat\Omega = \< \Omegai, \Omegac, \Omegar >$,
-%
-%$\Omega = \Omegai \uplus \Omegac \uplus \Omegar$ and $\Semiring$ 
-where $Q' = \bigcup_{k=0}^{M} Q^k$ be  the set of sequences of state symbols of $A$, 
-of length at most $M$, including the empty sequence denoted by $\epsilon$, 
-and where $P' = Q'$.
-
-\[
-\begin{array}{ll}
-%\weii: Q \times \Omegai \times Q \to \Semiring & 
-\weii(\bar{q}, a, \bar{q} q') = \wei(q', a, \epsilon) & 
-\mathrm{for~all~} a \in \Omega_0,\\
-%
-%\weic: Q \times \Omegac \times Q \times P \to \Semiring & 
-\weic(\bar{q}, \call{b}, \epsilon, \bar{q}) = \one & %\wei(p', b, \bar{p}) & 
-\mathrm{for~all~} b \in \Omega_{>0},\\
-%
-%\weir: Q \times \Omegac \times P \times \Sigmar \times Q \to \Semiring & 
-\weir(\bar{q}, {\call{b}}, \bar{p}, {\return{b}}, \bar{p} q') = \wei(q', b, \bar{q}) & 
-\mathrm{for~all~}  b \in \Omega_{>0},\\
-%
-%\weie: Q \times \Omegar \times Q \to \Semiring & 
-\weie(\bar{p}, {\return{b}}, \bar{q}) = \zero &
-\mathrm{for~all~}  b \in \Omega_{>0}.\\
-\end{array}      
-\]
-
-In practice, it is sufficient to consider in $Q'$ only the prefixes of 
-sequences.
-\qed\end{proof}
-
-\section{Symbolic Weighted Parsing}
-\label{sec:parsing}
-Let us now use the models and results of the former sections in %order to define 
-an approach to the problem of parsing over infinite alphabet. %appropriate
-%
-Besides considering infinitely many possible of input symbols, 
-handled with suitable language formalisms, 
-this approach extends conventional parsing 
-by computing a derivation tree modulo 
-a generic distance between words,  
-defined by a SW transducer given in input.
-This enables considering finer word relationships than strict equality 
-as in the conventional parsing approach, 
-opening possibilities of quantitative analysis via this method.
-
-\subsection{Definition}
-
-Let $\Sigma$ be a countable input alphabet 
-and $\Omega = \Omegai \uplus \Omegac \uplus \Omegar$ an output countable alphabet,
-let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative, bounded, complete 
-and total semiring 
-and let  $(\Phi_\epsilon, \Phic, \Phir, \Phicr)$ be a label theory over $\Semiring$, 
-assumed computable and $k$-convex for some fixed $k$.
-
-\noindent
-Assuming given in input:
-\begin{itemize}
-\item a \SWT $T$ over $\Sigma$,  $\Omega$, and $\Semiring$,
-defining a measure %between words 
-$T: \Sigma^* \times \Omega^* \to \Semiring$,
-\item a \SWVPA $A$ over $\Omega$, and $\Semiring$, defining a series of nested words
-      $A : \Omega^* \to \Semiring$,
-\item an input word $s \in \Sigma^*$,
-\end{itemize}
-the problem of \emph{symbolic weighted parsing} is to find 
-a nested word $t \in \Omega$ minimizing
-\( T(s, t) \otimes A(t)\)
-\wrt $\leq_\oplus$, 
-\ie such that %Hence, it is the problem of finding 
-\( T(s, t) \otimes A(t) = 
-  \displaystyle\bigoplus_{t' \in \Omega^*} T(s, t') \otimes A(t') \).
-
-\noindent 
-Therefore, it is the problem of optimizing a measure 
-called the \emph{edit-distance between $s$ and $A$} in~\cite{Mohri03EDWA}.
-%
-The input language can also be expressed as a \SWTA, or, 
-as a particular case, as a weighted context-free grammar, 
-converted in turn into a \SWVPA following Lemma~\ref{lem:SWTA}.
-%
-In the case of finite alphabets, 
-the problem of searching, in a WTA language, 
-for the best parse tree for a given,  
-sometimes referred as~\emph{weighted parsing}
-(see \cite{Goodman99SemiringParsing,MorbitzVogler19weighted-parsing} **more general problems**)
-is a particular case of SW parsing.
-%
-Indeed, it corresponds to the case where $T$
-accepts only the pairs $\<s, t>$ such that 
-$s$ is the projection of $t$ on $\Omegai$. 
-This can be done with a single state $q$ and 
-with transition rules of the form:
-\begin{description}
-\item[] $\wei(q, \epsilon, a, q) = \one$ for all $a \in \Omegac \cup \Omegar$,
-\item[] $\wei(q, a, a, q) = \one$ for all $a \in \Omegai$,
-\item[] $\wei(q, a, b, q) = \zero$ for all $a, b \in \Omegai$, $a \neq b$.
-\end{description}
-
-\subsection{Computation}
-
-\begin{proposition}
-The problem of Symbolic Weighted  parsing 
-can be solved in PTIME in the size of the input \SWT, \SWVPA (or \SWTA) and input word, 
-and the computation time of the functions of the label theory.
-\end{proposition}
-%
-\begin{proof}
-Bar-Hillel construction
-\qed  
-\end{proof}
-
-\subsection{Application to Automated Music Transcription}
-\label{sec:transcription}
-Symbolic Automated Music Transcription
-and analysis of music performances
-
-\subsubsection{Time Scales}
-Real-Time Unit (RTU) = seconds
-
-\noindent 
-Musical-Time Unit (MTU) = number of measures
-
-\noindent 
-conversion via tempo value
-
-\subsubsection{Representation of Music Performances}
-We consider symbolic representations of musical performances, as finite sequences of events.
-It corresponds to the concrete case of a MIDI file~\cite{SMF} 
-recorded  from an electronic keyboard, 
-or the output of a transcription from audio files~\cite{Benetos18AMTsurvey}.
-%
-For the sake of simplicity, 
-we shall only consider here the case of monophonic performances, 
-where at most one note is sounding at a time. 
-The approach however extends to the polyphonic case.
-
-A music performance is a finite sequence of events in a set~$\Sigma$.
-Every event $e \in \Sigma$ has attributes such from a finite domain, 
-like a number of key for a note 
-or a flag indicating that it is a rest 
-(\textsf{ON} and \textsf{OFF} messages in~\cite{SMF})
-and a velocity value (0..127 in~\cite{SMF})).
-%This representation is similar to the piano roll ~\cite{Muller15fundamentals} chap.1. 
-Moreover, it contains a RTU value $\ioi{e}$ (real number) 
-representing the time distance to the next event, 
-or to the end of performance for the last event,
-also called \emph{inter-onset interval}.
-
-\subsubsection{Representation of Music Scores}
-Music score are represented as structured words
-made of timed %quantified 
-events and parenthesized markups,
-akin of nested words~\cite{AlurMadhusudan09nested}.
-
-We consider an alphabet $\Delta$, every symbol of which is 
-composed of a tag, in a finite set $\Xi$, 
-and an MTU (rational) IOI duration value.
-%The alphabet $\Delta$ 
-It is partitioned into 
-$\Delta = \Deltai \uplus \Deltac \uplus \Deltar$, 
-like in Section~\ref{section:SWVPA}.
-%
-\noindent
-The symbols of $\Deltai$ represent events:
-% (infinite alphabet of internal symbols) made of:
-with tags indicating a new note or grace-note (with null IOI), 
-a rest or the continuation of the previous note (tie or dot).
-%
-The elements of $\Deltac \uplus \Deltar$ are matched
-markups for describing the structure of the score, 
-\ie the hierarchical grouping of events, and also, 
-importantly the division of time in measures, tuplets...
-%- parentheses for time divisions : tuplets, bars...
-(linearization of rhythm trees \cite{jacquemard:hal-01138642}...).
-They contain additional info such as tuple number, beaming policy...
-
-\noindent
-The duration values of letters of $\Delta$, in MTU (rational), 
-can be computed with the markups and tags (\eg grace note has duration 0).
-
-%\noindent
-%There are simultaneous events, since grace notes has duration 0. They are ordered.
-%
-%\noindent
-%Finite bound on the number of duration ratio. ?
-
-\begin{example}
-...      
-\end{example}
-
-\subsubsection{Performance/Score Distance Computation}
-\label{app:distance}
-We define a distance between performance and score representations
-by a $\SWT$ $T = \< Q, \init, \wei, \final >$, over a semiring $\Semiring$.
-** detail the elements of $\Semiring$ ....**
-%are quadruplets of the form
-%$\< t, s, \delta_t, \delta_s>$
-
-Every state of $Q$ contains a 
-tempo value in a finite domain (e.g. 30..300 bpm).
-This value can be fixed 
-or recomputed by the $T$ %transducer 
-after reading each event, 
-according to a perceptive/cognitive model of tempo 
-such as~\cite{LargeJones99tempo}
-(also used in the context of score following~\cite{Cont10TPAMI}).
-% we wont detail here.
-
-\subsection{Transcription by SW Parsing} %Best-first Search}
-We assume a score language defined by a \SWVPA over the semiring 
-$\Semiring$ of Section~\ref{app:distance}.
-
-\section*{Conclusion}
-% discussion
-
-\noindent
--- summary
-
-\noindent
--- other theoretical properties of SW models
-
-\noindent
--- room to improve complexity for best-search algorithm
-... modular approach with oracles ...
-
-and extention to $n$-best
-
-\noindent
--- offline algorithm, 
-semi-online implementation for AMT (bar-by-bar approach)
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% BIBLIO                                                                     %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\bibliographystyle{plain}
-%\bibliographystyle{plainurl} 
-\bibliographystyle{abbrv}
-%\bibliographystyle{splncs04}
-\bibliography{references}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% APPENDIX                                                                   %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\appendix
-
-\section{Edit-Distance}
-
-%\subsection{Distance between words or languages}
-...algebraic definition of edit-distance of Mohri, in \cite{Mohri03EDWA}
-% Mehryar Mohri 
-% Edit-distance of weighted automata: General definitions and algorithms
-% International Journal of Foundations of Computer Science 14.06 (2003): 957-982.
-distance $d$ over $\Sigma^* \times \Sigma^*$ 
-into a semiring  $\Semiring = ( \Semiring, \oplus, \zero, \otimes, \one)$.
-
-%\noindent
-Let $\Omega = \Sigma \cup \{ \epsilon \} \times \Sigma \cup \{ \epsilon \} \setminus \{ (\epsilon, \epsilon) \}$,
-and let $h$ be the morphism from $\Omega^*$ into $\Sigma^* \times \Sigma^*$  
-defined over the concatenation of strings of $\Sigma^*$ (that removes the $\epsilon$'s).
-%
-\noindent
-An \emph{alignment} between 2 strings  $s, t \in \Sigma^*$ is an element $\omega \in \Omega^*$ 
-such that $h(\omega) = (s, t)$.
-%
-\noindent
-We assume a base cost function $\Omega$ : $\delta: \Omega \to S$, extended to $\Omega^*$ as follows  
-(for $\omega \in \Omega^*$): 
-\(
-\displaystyle\delta(\omega) = \bigotimes_{0 \leq i < |\omega|} \delta(\omega_i)
-\).
-
-\noindent
-\begin{definition}
-For  $s, t \in \Sigma^*$, the edit-distance between $s$ and $t$ is  
-\( 
-d(s, t) = \displaystyle\bigoplus_{\omega \in \Omega^*\, h(\omega) = (s, t)} \delta(\omega)
-\).
-\end{definition}
-
-e.g. Levenstein edit-distance: $S$ is min-plus and $\delta(a, b) = 1$ for all $(a, b) \in \Omega$.
-
-%\paragraph{Distance between a word and a regular language}
-
-\end{document}

### 2021-05-16 main-V0.tex
split files

--- 2021-05-14 main.tex
@@
+% old version
+
 \documentclass[runningheads]{llncs}
@@
 
-%For a complete semiring $\Semiring$, 
-%a function $\phi: \Sigma \to \Semiring$, 
-%resp. $\eta: \Sigma \times \Delta \to \Semiring$,
-%is called
+\medskip\noindent
+When~$\Semiring$ is complete, 
+we call \emph{summary} of a function
+$\phi \in \Phi_\Sigma$,
+resp. $\eta \in \Phi_{\Sigma, \Delta}$,
+the value $\bigoplus_{a \in \Sigma} \phi(a)$, 
+resp. $\bigoplus_{a \in \Sigma} \bigoplus_{b \in \Delta} \eta(a, b)$.
+Note that, by definition of infinite sums in complete semirings, 
+a summary of $\phi \oplus \phi'$ and $\phi \otimes \phi'$ %(and $\alpha \otimes \phi$)
+can be computed from summaries of $\phi$ and $\phi'$ in $\Phi_{\Sigma}$, using the operators of $\Semiring$, 
+and the same holds for $\eta, \eta' \in \Phi_{\Sigma, \Delta}$. 
+%
+A label theory is called \emph{effective} when 
+summaries of $\phi \oplus \eta$, $\phi \otimes \eta$, $\eta \oplus \psi$, $\eta \otimes \psi$, 
+and of the partial applications $\eta_a$ and $\eta_b$, 
+can be computed, using the operators of $\Semiring$, 
+from summaries of 
+$\phi \in \Phi_{\Sigma}$, $\psi \in \Phi_{\Delta}$, $\eta \in \Phi_{\Sigma, \Delta}$.
 
@@
 \begin{example}
+Let $\Sigma$ be an alphabet of symbols with timestamps. 
+The timestamp of $a \in \Sigma$, denoted by $\mathsf{t}(a)$, is expressed as a rational number.
+
 DTW for sequences of timestamped events **     
@@
 $\wei: Q \times (\Sigma \cup \{ \epsilon \}) \times Q \to \Semiring$.
-
 %
@@
 %the automaton~$A$ is called \emph{without $\epsilon$-transitions}.
+
+A $\SWT$ or $\SWA$ is called \emph{summarized}
+when
 
@@
 \[
-\begin{array}{lclclcll}
+\begin{array}{lclcrcll}
 \wei'_\epsilon\bigl(\< i, q>, \< i, q'>\bigr) & = & \wei_\epsilon(q, q') & \mathrm{\ie} &
@@
 \wei'_\Sigma\bigl(\< i, q>, \< j, q'>\bigr) & = & \zero & \mathrm{\ie} &
-\wei'\bigl(\< i, q>, b, \< j, q'>\bigr)     & = & \zero & \mathrm{if~} j \neq i, i+1. 
+\wei'\bigl(\< i, q>, b, \< j, q'>\bigr)     & = & \multicolumn{2}{l}{\zero\ \mathrm{~if~} j \neq i, i+1.}
 \end{array}      
@@
 \noindent
-The $\SWA$ wanted is $A_{T, s} = \< Q', \init', \bar{\wei'}, \final' >$.
+The $\SWA$ wanted is $A_{T, s} = \< Q', \init', {\bar\wei}', \final' >$.
 \qed
@@
 where the size $\| T \|$ of $T$ is its number of states $|Q|$.
-
-\noindent 
-When $\Semiring$ is complete, we call \emph{summary} of a 
-to a $\SWT$ $T = \< Q, \init, \bar{\wei}, \final >$
-a triplet of functions 
-$\mei_\Sigma$, $\mei_\Delta$, and $\mei_{\Sigma, \Delta}$
-from $Q \times Q$ into~$\Semiring$, defined by 
-\[
-\begin{array}{rcll}
-\mei_\Sigma(q, q') & = & \bigoplus_{a \in \Sigma} \phi(a) & 
-\quad\mathrm{where~} \phi = \wei_\Sigma(q, q') \in \Phi_\Sigma,\\
-\mei_\Delta(q, q') & = & \bigoplus_{b \in \Delta} \psi(b) & 
-\quad\mathrm{where~} \psi = \wei_\Delta(q, q') \in \Phi_\Delta,\\
-\mei_{\Sigma, \Delta}(q, q') & = & 
- \bigoplus_{a \in \Sigma} \bigoplus_{b \in \Delta} \eta(a, b) & 
-\quad\mathrm{where~} \eta = \wei_{\Sigma, \Delta}(q, q') \in \Phi_{\Sigma, \Delta}.\\
-\end{array}      
-\]
-Note that these functions are well defined under the assumption that $\Semiring$ is complete.
-%A $\SWT$ is called \emph{presented} when these functions, with finite domains, 
-%are given explicitly.
-For a \SWA, the summary is only defined by $\mei_\Sigma$.
 

### 2021-05-16 semiring.tex
split files

--- 2021-05-12 main.tex
@@
-\documentclass[runningheads]{llncs}
-%\documentclass[a4paper,11pt]{article}
-%\setcounter{page}{1}
-
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-
-\usepackage{hyperref}
-%\usepackage[bookmarks,bookmarksnumbered,naturalnames,plainpages=false]{hyperref}
-%usepackage{url}
-
-% for footnote ref
-\usepackage{refcount}
-
-% array and tabular
-\usepackage{array}
-\newcolumntype{C}[1]{>{\centering\arraybackslash\hspace{0pt}}p{#1}}
-
-% extension of enumerate env. (style for displaying counters)
-% \usepackage{enumerate}
-
-%% pictures
-% \usepackage{graphicx} 
-% \DeclareGraphicsExtensions{.pdf,.png,.jpg}
-% \graphicspath{fig/}
-
-%% PGF, Tikz
-\usepackage{tikz-cd}
-%% \usepackage{pgfplots}
-%% \usepgfplotslibrary{dateplot}
-%% \usepackage{pgf,pgfarrows,pgfnodes, pgfautomata}
-% \usepackage{tikz}
-% \usetikzlibrary{cd}
-%% \usetikzlibrary{arrows}
-%% \usetikzlibrary{calc}
-%% \usetikzlibrary{snakes}
-%% \usetikzlibrary{backgrounds}
-% \usetikzlibrary{trees}
-%% \usetikzlibrary{automata}
-%% \usetikzlibrary{positioning}
-%% \usetikzlibrary{matrix}
-%% \usetikzlibrary{patterns}
-%% \usetikzlibrary{shapes}
-
-% symbols
-\usepackage{amsmath} 
-\usepackage{amssymb} 
-\usepackage{amsbsy}
-\usepackage{bbold}
-\usepackage{latexsym}
-%\usepackage{amsfonts}
-\usepackage{stmaryrd}
-%\usepackage{mathabx}
-%\usepackage{MnSymbol}
-\usepackage{harmony} % simple music fonts
-\usepackage{mathtools} % for arrows
-%\usepackage{mathptmx}
-
-%% theorem environments
-\usepackage{theorem}
-% \newtheorem{theorem}{Theorem} %[section]
-% \newtheorem{definition}[theorem]{Definition}
-% \newtheorem{lemma}[theorem]{Lemma}
-% \newtheorem{corollary}[theorem]{Corollary}
-% \newtheorem{proposition}[theorem]{Proposition}
-% \newenvironment{proof}{\vspace{-2ex}{\it Proof. }}{\hspace*{\fill} $\Box$\smallskip }
-% \theorembodyfont{\slshape}
-% \newtheorem{example}[theorem]{Example}
-% \newtheorem{remark}[theorem]{Remark}
-{\theorembodyfont{\rmfamily} \theoremstyle{break} \newtheorem{algo}{Algorithm}}
-
-%% algorithms
-%\usepackage{algorithm} 
-%\usepackage{program}
-
-%% for new macros
-\usepackage{xspace}
-
-%% arrows etc
-%\input{rewriting}
-
-%% Misc macros
-
-\def\ie{\textit{i.e.}\xspace}
-\def\eg{\textit{e.g.}\xspace}
-\def\wrt{\textit{wrt}\xspace}
-%\def\wlog{\textit{wlog}\xspace}
-\def\etc{\textit{etc}\xspace}
-
-\def\<#1>{\langle #1 \rangle}
-\newcommand{\pair}[2]{\langle{#1}, {#2}\rangle}
-%\newcommand{\A}{\mathcal{A}}
-%\newcommand{\B}{\mathcal{B}}
-\newcommand{\D}{\mathbb{D}}
-\newcommand{\E}{\mathbb{E}}
-\newcommand{\W}{\mathbb{W}}
-\newcommand{\T}{\mathcal{T}}
-
-\newcommand{\Semiring}{\mathbb{S}}
-\newcommand{\zero}{\mathbb{0}}
-\newcommand{\one}{\mathbb{1}}
-\newcommand{\dom}{\ensuremath{\mathit{dom}}}
-
-\def\SA{\textsf{SA}\xspace}
-\def\WA{\textsf{WA}\xspace}
-\def\SWT{\textsf{swT}\xspace}
-\def\SWA{\textsf{swA}\xspace}
-\def\SWTA{\textsf{swTA}\xspace}
-\def\SWVPA{\textsf{sw-VPA}\xspace}
-\def\weight{\mathsf{weight}}
-\def\wei{\mathsf{w}}
-\def\init{\mathsf{in}}
-\def\final{\mathsf{out}}
-\newcommand{\call}[1]{\ensuremath #1} %{\ensuremath \langle_{#1}}
-\newcommand{\return}[1]{\ensuremath #1} %{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
-\def\Omegai{{\Omega_\mathsf{i}}}
-\def\Omegac{{\Omega_\mathsf{c}}}
-\def\Omegar{{\Omega_\mathsf{r}}}
-\def\Sigmai{{\Sigma_\mathsf{i}}}
-\def\Sigmac{{\Sigma_\mathsf{c}}}
-\def\Sigmar{{\Sigma_\mathsf{r}}}
-\def\Deltai{{\Delta_\mathsf{i}}}
-\def\Deltac{{\Delta_\mathsf{c}}}
-\def\Deltar{{\Delta_\mathsf{r}}}
-\def\Phii{{\Phi_\mathsf{i}}}
-\def\Phic{{\Phi_\mathsf{c}}}
-\def\Phir{{\Phi_\mathsf{r}}}
-\def\Phicr{{\Phi_\mathsf{cr}}}
-\def\weii{{\wei_\mathsf{i}}}
-\def\weic{{\wei_\mathsf{c}}}
-\def\weir{{\wei_\mathsf{r}}}
-\def\weie{{\wei_\mathsf{e}}}
-\newcommand{\config}[2]{\ensuremath \genfrac{[}{]}{0pt}{}{#1}{#2}} 
-\newcommand{\opluseq}{\ensuremath\mathrel{\oplus}=}
-\newcommand{\ioi}[1]{\mathsf{ioi}({#1})}
-\newcommand{\rank}{\mathsf{rk}}
-\newcommand{\lin}{\mathsf{lin}}
-
-%\sloppy
-
-% Parsing over infinite alphabet as optimal alignment computation 
-% as edit-distance between string and language
+% !TEX root = main.tex
 %
-%\title{Symbolic Weighted Parsing and Automated Music Transcription}
-%\title{Symbolic Weighted Language Models and Automated Music Transcription}
-\title{Symbolic Weighted Language Models and Parsing over Infinite Alphabets}
-%\title{Weighted Visibly Pushdown Automata and Automated Music Transcription}
-\author{Florent Jacquemard}
-\institute{INRIA \& CNAM, Paris, France\\
-\email{florent.jacquemard@inria.fr}}
-
-%\titlerunning{WVPA \& AMT}
-\titlerunning{Symbolic Weighted Parsing and Automated Music Transcription}
-%\authorrunning{Florent Jacquemard}
-
-\date{\today}
-
-\begin{document}
-\thispagestyle{empty}
-\maketitle
-
-\begin{abstract}
-%Symbolic Weighted (SW) extension of symbolic automata...
-We propose a framework for weighted parsing over infinite alphabets.
+% Semiring basics
 %
-It is based on language models called Symbolic Weighted Automata (SWA) 
-at the joint %intersection 
-between Symbolic Automata (SA) and Weighted Automata (WA), 
-as well as Transducers (SWT) and Visibly Pushdown (SWPDA) variants.
-%
-Like SA, SWA deal with large or infinite input alphabets, 
-and like WA, they output a weight value in a semiring domain.
-The transitions of WA are labeled by functions from an infinite alphabet into the weight domain. 
-This is unlike WA whose transitions are guarded by boolean predicates 
-overs symbols in an infinite alphabet 
-and also unlike WA whose transitions are labeled by constant weight values, 
-and who deal only with finite automata.
-%
-We present some properties of WA, WT and SWPDA models, 
-%and show how they can be used 
-that we use to define and solve a variant of parsing 
-over infinite alphabets.
-%
-We also briefly describe the application that motivated the introduction of these models: 
-an parse-based approach to automated music transcription.
-\end{abstract}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% intro
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Introduction} \label{sec:intro}
-Parsing %variant of membership problem for formal languages
-%- given a language model and an input word $s$, compute a derivation of the model that yields $s$.
-is the problem %process 
-of structuring a linear representation in input %typically
-(a finite word over an alphabet) according to a language model. % (a formal grammar).
-% natural language, programming language, 
-%
-Most of the context-free parsing approaches~\cite{GruneJacobs08parsing}  
-assume a finite and reasonably small input alphabet. %models and algorithms 
-Such a restriction makes perfectly sense in the context of NLP tasks 
-like constituency parsing or of programming languages compilers or interpreters.
-Considering large or infinite alphabets can however be of practical interest
-in other cases.
-%
-For instance, when dealing with large characters encodings such as UTF-16, % processing strings in 
-\eg for vulnerability detection in Web-applications~\cite{dAntoni21CACM}, 
-%
-for the analyse (\eg validation or filtering) 
-of data streams or serialization of structured documents 
-(which may contain textual or numerical attributes)~\cite{Segoufin06csl}, 
-or for processing timed execution traces~\cite{Bouyer03algebraic}.
-% algebraic definition of a class of data languages 
-% (notion of monoid recognizability, based on registers, comparable to Bojancszik et al. data words)
-Regarding the latter case, we describe %briefly 
-at the end of the paper 
-a parse-based approach to automated music transcription,
-where a symbolic music performance, 
-presented as a sequence of timed musical events, % timestamped
-% a symbolic representation of a music performance 
-is converted into a structured score in Common Western Music Notation~\cite{foscarin:hal-01988990}.
-%generalizes weighted parsing: 
-%finding the best derivation for a weighted grammar.
-
-Various extensions of language models for handling infinite alphabets have been studied.
-%words carrying data values in an infinite domain (e.g. integers) e.g. data processing 
-For instance, some automata with memory extensions 
-allow restricted storage and comparison of input symbols, 
-%and correspond with logics, 
-(see~\cite{Segoufin06csl} for a survey),
-with pebbles for marking positions~\cite{NevenSchwentickVianu04FSMinfinite}, 
-registers~\cite{KaminskiFrancez94}, 
-or %computations in 2 steps
-the possibility to compute on subsequences 
-with the same attribute values~\cite{Bojanczyk11FO2}. % data words automata.
-%
-%for the and verification of infinite state systems 
-%(model checkers: alphabet = long bit-vectors)
-%...For the representation of  in model checking, verification and 
-The automata at the core of model checkers
-compute on input symbols represented by large bitvectors~\cite{Vardi07ciaa} %\cite{BaierKatoen08MC}
-(sets of assignments of Boolean variables) %propositional variables)
-and in practice,  %implementation
-each transition accepts a set of such symbols (instead of an individual symbol), 
-represented by Boolean formula or Binary Decision Diagrams.
-%
-Following a similar idea, % and generalizing,
-in symbolic automata (\SA)~\cite{dAntoniVeanes17CAV,dAntoni21CACM}, 
-the transitions are guarded by predicates over infinite alphabet domains.
-With closure conditions on the sets of such predicates, % (alphabet theories), 
-all the good properties enjoyed by automata over finite alphabets are preserved.
-
-Other extensions of language models  %(automata and grammars) 
-help in dealing with non-determinism, by the computation of weight values. 
-With an ambiguous grammar, there may exists several derivations 
-(\emph{abstract syntax trees} -- AST), % \ie the result of an analyze.
-yielding one input word. % structuring a word, 
-The association of one weight value %in semiring domain
-to each derivation permits to select a best one (or $n$ bests). % a fixed number of bests 
-% = ...a ranking of derivations 
-This is roughly the principle of \emph{weighted parsing}
-approaches~\cite{Goodman99SemiringParsing,Nederhof03weightedParsing,MorbitzVogler19weighted-parsing}.
-In \emph{weighted language models}, 
-like \eg probabilistic context-free grammars % (CFG),
-and weighted automata (\WA) \cite{Droste09handbook},
-a weight value is associated to each transition rule, % production rule
-and the rule's weights can be combined with a associative product operator~$\otimes$ into 
-the weight of a derivation.
-A second operator~$\oplus$, associative and commutative, 
-is moreover used to handle ambiguity of the model, 
-by summing the weights of the possibly several (in general exponentially many) AST %syntax trees 
-associated to a given input word.
-Typically, $\oplus$ will select the best of two weight values.
-%Intuitively,~$\oplus$ selects, or ranking, the syntax trees.
-The weight domain, equipped with these two operators is assumed, at minima, 
-to form a \emph{semiring} %$\Semiring$ 
-where $\oplus$ can be extended to infinite sums, 
-%\cite{Eilenberg74automata}
-such as the Viterbi semiring and the tropical min-plus algebra, see Figure~\ref{fig:semirings}. 
-%of domain $\mathbb{R}^+ \cup \{ +\infty\}$, 
-%where $\oplus$ is min and $\otimes$ is plus .
-%by ranking 
-%making the weight domain a semiring.
-%Some efficient specialized parsing algorithms~\cite{Huang05kbest} have been proposed in this context 
-%% models represented as hypergraphs \cite{Huang05kbest}
-%in order to compute the $n$ best syntax trees of a given input word without having to enumerate them all.
-%Generally based on dynamic programming, these algorithms rely on 
-%additional algebraic properties of~$\Semiring$.
-%-- see \eg~\cite{Huang05kbest} for some NLP applications.
-%The extraction of $n$ best list is useful 
-%the problem: quantitative parsing or symbolic parsing
-%of parsing of words over infinite input alphabet.
-
-In this paper, we present a uniform framework for weighted parsing over infinite input alphabets.
-It is based on weighted %finite states 
-language models generalizing 
-both~\SA, with functions in an arbitrary semiring instead of Boolean guards, 
-and~\WA, by handling infinite alphabets -- Figure~\ref{fig:hierarchy}.
-%***
-%some weighted language models
-%model of weighted CFG 
-%computing on words over infinite input alphabets %sets of terminal symbol
-%- but with finite sets of states and transitions rules. 
-% non-terminals and production rules).
-In their transition rules, input symbols appear as variables %(parameters)
-and the weight associated to a transition rule is a function of these variables.
-%
-\begin{figure}
-\centering
-\begin{tikzpicture}
-\node (NFA) at (0,4) {%
-  \(
-  \begin{array}{c} 
-  \mathsf{NFA} : \Sigma_{\mathsf{fin}}^* \to \mathbb{B}\\
-  q \xrightarrow{a} q' \quad  a \in \Sigma_\mathsf{fin}
-  \end{array} 
-  \)
-};
-%
-\node (WA) at (-2,2) {%
-  \(
-  \begin{array}{c} 
-  \mathsf{WA} : \Sigma_{\mathsf{fin}}^* \to \mathbb{S}\\
-  q \xrightarrow{a, w} q' \quad  a \in \Sigma_\mathsf{fin}, w \in \mathbb{S}
-  \end{array} 
-  \)
-};
-%
-\node (SA) at (2,2) {%
-  \(
-  \begin{array}{c} 
-  \mathsf{SA} : \Sigma_{\mathsf{inf}}^* \to \mathbb{B}\\
-  q \xrightarrow{\phi} q' \quad \phi : \Sigma_\mathsf{inf} \to \mathbb{B}
-  \end{array} 
-  \)
-};
-%
-\node (SWA) at (0,0) {%
-  \(
-  \begin{array}{c} 
-  \mathsf{SWA} : \Sigma_{\mathsf{inf}}^* \to \mathbb{S}\\
-  q \xrightarrow{\phi} q' \quad \phi : \Sigma_\mathsf{inf} \to \mathbb{S}
-  \end{array} 
-  \)
-};
-\draw[->] (NFA)--(WA);
-\draw[->] (NFA)--(SA);
-\draw[->] (WA)--(SWA);
-\draw[->] (SA)--(SWA);
-%\begin{array}{c} \mathsf{NFA} : \Sigma^* \to \mathbb{B} \end{array} 
-\end{tikzpicture}
-\caption{Classes of Symbolic/Weighted Automata. 
-$\Sigma_\mathsf{fin}$ is a finite alphabet, 
-$\Sigma_\mathsf{inf}$ is a countable alphabet,
-$\mathbb{B}$ is the Boolean algebra, 
-$\mathbb{S}$ is an arbitrary commutative semiring, 
-$q \xrightarrow{\dots} q'$ represents the form of a transition between states $q$ and $q'$.}
-\label{fig:hierarchy}  
-\end{figure}
-%
-%This approach is close to the case of 
-%Symbolic Automata (SA)~\cite{dAntoniVeanes17CAV,dAntoni21CACM}, 
-%except that the domain for weight values is not restricted to be Boolean, 
-%like for the guards in the rules of SA, 
-%but can be an arbitrary commutative semiring (assuming some restrictions).
-%
-The models presented here are finite automata called symbolic-weighted (\SWA),
-transducers (\SWT) and pushdown automata,
-with a visibly restriction~\cite{AlurMadhusudan09nested} (\SWVPA).
-%\wrt an edit distance
-The latter model of automata computes on \emph{nested words}~\cite{AlurMadhusudan09nested}, 
-a structured form of words parenthesized with markup symbols, 
-corresponding to a linearization of trees.
-In the context of parsing, they are used here to represent (weighted) AST of CF grammars.
-More precisely, a \SWVPA $A$ associates a weight value $A(t)$ % \in \Semiring$ 
-to a given a nested word $t$, which is the linearization of an AST. %representing a parse tree.
-%
-On the other hand, 
-a \SWT is used to define a distance $T(s, t)$ between two finite words $s$ and $t$
-over an infinite alphabet, following~\cite{Mohri03EDWA}.
-Then, the \emph{SW-parsing} problem aims at %computing 
-finding $t$ minimizing 
-$T(s, t) \otimes A(t)$ (\wrt the ranking defined by $\oplus$)
--- this value is called the distance between $s$ and $A$ in~\cite{Mohri03EDWA}.
-%
-Similarly to weighted-parsing 
-methods~\cite{Goodman99SemiringParsing,Nederhof03weightedParsing,MorbitzVogler19weighted-parsing}, 
-our approach proceeds in two steps, 
-based on properties of the SW models. 
-The first step is a Bar-Hillel construction where, 
-given a \SWT $T$, a \SWVPA $A$, and an input word $s$, 
-a \SWVPA $A_{T, s}$ is built, such that for all $t$, $A_{T, s}(t) = T(s, t) \otimes A(t)$.
-In the second step, a best AST $t$ is found by applying to $A_{T, s}$ 
-a best search algorithm similar to shortest distance in graphs~\cite{Mohri02semiring,Huang05kbest}.
-%In expressiveness, they are equivalent to weighted CFG. 
-%and can be used in a general approach for parsing over infinite input alphabets.
-%
-%Let $A$ be a \SWVPA, associating $A(w) \in \Semiring$ 
-%to a given a nested word $w$ (representing a parse tree),
-%and let a \SWT compute a distance $d$, in $\Semiring$, 
-%between 2 strings over respectively an infinite input alphabet and the 
-%same (infinite) alphabet of $A$.
-%Then, the problem of Symbolic Weighted Parsing is, 
-%given an input string $s$, to find a nested word $w$ minimizing 
-%(according to the ranking defined by $\oplus$)
-%the distance $d(s, w) \otimes A(w)$ between $s$ and $A$, 
-%as defined in~\cite{Mohri03EDWA}.
-
-% First one general edit-distance is defined by a weighted word 
-% transducer~\cite{Mohri}
-% %Symbolic automata transducers are extended models~\cite{VeanesdAntoniJACM}
-% %dealing with infinite set of input symbols...
-% value in a semiring...
-
-The main contributions of the paper are: 
-($i$)~the introduction of automata, \SWA, transducers, \SWT (Section~\ref{section:SWA}),
-and visibly pushdown automata \SWVPA (Section~\ref{sec:SWVPA}),
-generalizing the corresponding classes of symbolic and weighted models, 
-%for the weighted computation on (nested) words over infinite alphabets;  
-($ii$)~a polynomial best-search algorithm for \SWVPA, %(Section~\ref{sec:best})
-% a framework for parsing over infinite alphabets, 
-and ($iii$)~a uniform framework (Section~\ref{sec:parsing}) for parsing over infinite alphabets, 
-the keys to which are 
-($iii.a$)~the \SWT-based definition of generic edit distances between input and output words,
-and ($iii.b$)~the use of nested words, and \SWVPA, 
-instead of syntax trees and grammars. %(\S~\ref{sec:trees}).
-%
-Finally, Section~\ref{sec:transcription} describes
-the implemented application %of this approach 
-to automated music transcription
-that motivated this study.
-
-%based these models and on generic def. of distances;
-%the use of VPA for parsing, word instead of trees, use comparison input-output.
-%and best-search
-
-%In Section~\ref{section:SWA} we introduce \SWA and \SWT.
-%Then \SWVPA are defined in Section~\ref{sec:SWVPA}, 
-%where a polynomial 1-best algorithm is described that can be use to solve
-%Symbolic Weighted Parsing, as described Section~\ref{sec:parsing}.
-%Finally, in Section~\ref{sec:transcription}, we present an application 
-%of this approach to automated music transcription that has been implemented.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SWT & SWA
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{SW Automata and Transducers}
-\label{section:transducer}\label{sec:transducer}
-\label{section:SWA}\label{sec:SWA}
-\label{section:SWT}\label{sec:SWT}
-
-We follow the approach of~\cite{Mohri03EDWA} for the computation of distances
-between words and languages, with weighted transducers, 
-extending it to infinite alphabets.
-% with models of symbolic weighted automata and transducers. 
-%
-The models introduced in this section generalize 
-weighted automata and transducers~\cite{Droste09handbook} 
-%over finite alphabets, see  e.g.~\cite{Mohri03EDWA}, 
-by labeling each transition with a weight function that takes the 
-input and output symbols as parameters, instead of a simple weight value.
-These functions are similar to the guards of symbolic automata~\cite{dAntoniVeanes17CAV,dAntoni21CACM},
-but they can return values in an arbitrary semiring, 
-where the latter guards are restricted to the Boolean semiring.
-
-\subsection{Semirings}
-\label{section:semiring}
 We shall consider semiring for the weight values of our language models.
@@
 \end{figure}
-
-\subsection{Label Theory}
-\label{section:symbols}
-We shall now define the functions labeling the transitions of SW-automata and -transducers.
-Let $\Sigma$ and $\Delta$ be two countable sets of symbols
-called respectively \emph{input} and \emph{output} \emph{alphabets}, 
-and let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative semiring.
-
-\noindent 
-A \emph{label theory} over $\Sigma$ and $\Delta$
-is made of 3 recursively enumerable sets:
-%$\Phi_\epsilon \subseteq \Semiring$, % containing constant functions valued in $\Semiring$, 
-$\Phi_\Sigma$ and $\Phi_\Delta$, 
-containing unary functions in $\Sigma \to \Semiring$, resp. $\Delta \to \Semiring$, 
-and $\Phi_{\Sigma, \Delta}$  containing binary functions in $\Sigma \times \Delta \to \Semiring$, 
-such that 
-for all $\alpha \in \Semiring$, and $\phi \in \Phi_\Sigma$,
-$\alpha \otimes \phi : x \mapsto \alpha \otimes \phi(x)$, 
-and $\phi \otimes \alpha : x \mapsto \phi(x) \otimes \alpha$
-belong to $\Phi_\Sigma$, and similarly for $\oplus$ 
-and for $\Phi_\Delta$ and $\Phi_{\Sigma, \Delta}$.
-%
-Moreover, these sets are required to be closed under the following 
-operators~$\oplus$ and $\otimes$ of~$\Semiring$:
-for all $\phi, \phi' \in \Phi_\Sigma$,
-all $\psi, \psi' \in \Phi_\Delta$, 
-and $\eta, \eta' \in \Phi_{\Sigma, \Delta}$, the function
-\begin{description}
-\item $\phi \otimes \phi' : x \mapsto \phi(x) \otimes \phi'(x)$ belongs to $\Phi_\Sigma$, 
-\item $\psi \otimes \psi' : y \mapsto \psi(y) \otimes \psi'(y)$ belongs to $\Phi_\Delta$,
-\item $\phi \otimes \eta : x, y \mapsto \phi(x) \otimes \eta(x, y)$ belongs to $\Phi_{\Sigma, \Delta}$,
-\item $\eta \otimes \psi : x, y \mapsto \eta(x, y) \otimes \psi(y)$ belongs to $\Phi_{\Sigma, \Delta}$,
-\item $\eta \otimes \eta' : x, y \mapsto \eta(x, y) \otimes \eta'(x, y)$ belongs to $\Phi_{\Sigma, \Delta}$.
-\end{description}
-And the same also holds for the binary sum operator $\oplus$.
-
-\noindent
-%\marginpar{necessary?}
-Finally, it is also required 
-% that the codomain of every function of $\Phi_\Sigma$ and $\Phi_\Delta$ 
-% is a subset of $\Phi_\epsilon$, and
-that all partial applications of functions $\Phi_{\Sigma, \Delta}$, 
-resp.  $f_a: y \mapsto f(a, y)$ for $a \in \Sigma$ and $y \in \Delta$
-and  $f_b: x \mapsto f(x, b)$ for $b \in \Delta$ and $x \in \Sigma$, 
-belong resp. to~$\Phi_\Sigma$ and~$\Phi_\Delta$.
-
-\subsection{Definitions} \label{sec:SWTdef}\label{sec:SWAdef}
-%(SWT)
-\begin{definition}
-\label{def:transducer} \label{def:SWT}
-A \emph{symbolic-weighted transducer} (\SWT)
-over the input and output alphabets~$\Sigma$ and $\Delta$ 
-and the semiring $\Semiring$ is a tuple
-$T = \< Q, \init, \bar{\wei}, \final >$,
-where $Q$ is a finite set of states, 
-$\mathsf{in} : Q \to \Semiring$, 
-respectively $\mathsf{out} : Q \to \Semiring,$
-are functions defining the weight for entering, 
-respectively leaving, a state, 
-and $\bar{\wei}$ is a tuplet made of the 4 of transition functions 
-$\wei_\epsilon$, $\wei_\Sigma$, $\wei_\Delta$, and $\wei_{\Sigma, \Delta}$
-from $Q \times Q$ into respectively~$\Semiring$, %$\Phi_\epsilon$, 
-$\Phi_\Sigma$, $\Phi_\Delta$, and $\Phi_{\Sigma, \Delta}$.
-\end{definition}
-%
-We summarize the above 4-uplet of transition functions into a unique 
-function~$\wei$ from
-$Q \times (\Sigma \cup \{ \epsilon \}) \times (\Delta \cup \{ \epsilon \}) \times Q$
-into $\Semiring$, %also called $\wei$ for simplicity, 
-such that, for all $q, q' \in Q$, $a \in \Sigma$, and $b \in \Delta$, 
-%and with 
-%$\< \phi_\epsilon, \phi_\Sigma, \phi_\Delta, \phi_{\Sigma, \Delta}> = \wei(q, q')$, 
-\[
-\begin{array}{rcll}
-\wei(q, \epsilon, \epsilon, q') & = &  \wei_\epsilon(q, q'),\\ %\phi_\epsilon\\
-\wei(q, a, \epsilon, q') & = & \phi_\Sigma(a) & %\wei_\Sigma(q, q')(a)
-\quad\mathrm{where~} \phi_\Sigma = \wei_\Sigma(q, q'),\\
-\wei(q, \epsilon, b, q') & = & \phi_\Delta(b) & 
-\quad\mathrm{where~} \phi_\Delta = \wei_\Delta(q, q'),\\
-\wei(q, a, b, q') & = & \phi_{\Sigma, \Delta}(a, b) & 
-\quad\mathrm{where~} \phi_{\Sigma, \Delta} = \wei_{\Sigma, \Delta}(q, q').\\
-\end{array}      
-\]
-%More precisely, $Q \times Q$, 
-%resp. $Q \times \Sigma \times Q$,
-%$Q \times \Delta \times Q$,
-%$Q \times \Sigma \times \Delta \times Q$,
-%into resp. $\Phi_0$, $\Phi_\Sigma$, $\Phi_\Delta$ $\Phi_{\Sigma, \Delta}$.
-
-The symbolic-weighted transducer $T$ defines a mapping 
-from the pairs of strings of $\Sigma^* \times \Delta^*$ 
-into the weights of~$\Semiring$,
-based on the following intermediate function $\weight_T$
-defined recursively for every $q, q' \in Q$, 
-%$a \in \Sigma$, $b \in \Delta$ 
-for every strings of $s \in \Sigma^*$, $t \in \Delta^*$:
-\[
-\begin{array}{rccl}
-xc(q, s, t, q') & = & 
-   \displaystyle\bigoplus_{q'' \in Q} &
-   \wei(q, \epsilon, \epsilon, q'') \otimes \weight_T(q'', s, t, q')\\
-   & \oplus & \displaystyle\bigoplus_{q'' \in Q} &
-              \wei(q'', \epsilon, \epsilon, q') \otimes \weight_T(q, s, t, q'')\\
- & \oplus & \displaystyle\bigoplus_{\begin{array}{c}
-                                      \scriptstyle q'' \in Q\\[-2pt]
-                                      \scriptstyle s = au, a \in \Sigma
-                                      \end{array}} &
-   \wei(q, a, \epsilon, q'') \otimes \weight_T(q'', u, t, q')\\
- & \oplus & \displaystyle\bigoplus_{\begin{array}{c}
-                                    \scriptstyle q'' \in Q\\[-2pt]
-                                    \scriptstyle t = bv, b \in \Delta
-                                    \end{array}} &
-    \wei(q, \epsilon, b, q'') \otimes \weight_T(q'', s, v, q')\\ 
- & \oplus & \displaystyle\bigoplus_{\begin{array}{c}
-                                    \scriptstyle q'' \in Q\\[-2pt]
-                                    \scriptstyle s = au, a \in \Sigma\\[-2pt]
-                                    \scriptstyle t = bv, b \in \Delta
-                                    \end{array}} &
-   \wei(q, a, b, q'') \otimes \weight_T(q'', u, v, q').\\ 
-\end{array}
-\]
-The sum above can be infinite because of the expression in the first line, 
-however, $\T(\Omega)$ is well defined since $\Semiring$ is complete.
-%
-As we shall see, epsilon-transitions can be removed and hence the sum made finite.
-%
-%** The application of $\oplus$ can be seen as a ND choice between transitions **
-We recall that, by convention, an empty sum with $\oplus$ is $\zero$. 
-%
-Since $\zero$ is absorbing for~$\otimes$ in~$\Semiring$,
-one term $\wei(q, a, b, q'')$ equal to $\zero$ in the above expression 
-will be ignored in the sum, meaning that there is no possible transition
-from state $q$ into state $q'$ while reading $a$ and writing $b$ 
-(including the case $a = \epsilon$ and $a = \epsilon$).
-%These functions $\phi$ act as guards for the transducer's transitions, 
-%preventing a transition when they return the absorbing $\zero$ of $\Semiring$.
-This is analogous to a transition's guard not satisfied by $\<a, b>$ in 
-the case of symbolic transducers.
-
-The above expression of $\weight_T$ can be seen as a stateful definition of 
-an edit-distance between a word $s \in \Sigma^*$ and a word $t \in \Delta^*$,
-see also~\cite{Mohri03ijfcs}.
-Intuitively, 
-$\wei(q, a, \epsilon, q'')$, in the third line, is the cost of 
-the deletion of a symbol~$a \in \Sigma$ in~$s$, 
-$\wei(q, \epsilon, b, q'')$, in the fourth line, is the cost 
-of the deletion of insertion of~$b \in \Delta$ in~$s$, 
-and $\wei(q, a, b, q'')$, in the last line, is the cost 
-of the substitution of  $a \in \Sigma$ by~$b \in \Delta$.
-The cost of a sequence of such operations transforming $s$ into $t$, 
-is the product, with $\otimes$, of the individual costs of the operations involved;
-And the distance between $s$ and $t$ is the sum, with $\oplus$,
-of all such product of costs.
-
-\medskip\noindent
-The weight associated by $T$ to  $\< s, t> \in \Sigma^* \times \Delta^*$
-is then defined as follows: 
-\begin{equation}
-T(s, t)  = 
-\displaystyle\bigoplus_{q, q' \in Q} \mathsf{in}(q) 
-\mathop{\otimes} \weight_T(q, s, t, q') \mathop{\otimes} \mathsf{out}(q').
-\label{eq:weightT}
-\end{equation}
-
-\begin{example}
-DTW for sequences of timestamped events **     
-\end{example}
-
-\begin{example}
-(simpler) pointwise distance between two sequences of timestamped events **
-\end{example}
-
-\noindent
-The \emph{Symbolic Weighted Automata} %$A = \< Q, \init, \weight, \final >$
-over $\Sigma$ and $\Semiring$ 
-are defined similarly as the transducers of Definition~\ref{def:SWT}, 
-by simply omitting the output symbols.
-%
-In this case, a label theory over $\Sigma$ is reduced to
-a set $\Phi_\Sigma$ closed under~$\oplus$ and~$\otimes$.
-%
-\begin{definition} \label{def:SWA}
-A \emph{symbolic-weighted automaton} (\SWA)
-over the input alphabet~$\Sigma$
-and the commutative semiring $\Semiring$ is a tuple
-$A = \< Q, \init, \bar{\wei}, \final >$,
-where $Q$ is a finite set of states, 
-$\mathsf{in} : Q \to \Semiring$, 
-respectively $\mathsf{out} : Q \to \Semiring,$
-are functions defining the weight for entering, 
-respectively leaving, a state, 
-and $\bar{\wei}$ is a pair of transition functions 
-$\wei_\epsilon$ and $\wei_\Sigma$ from $Q \times Q$ into 
-respectively $\Semiring$ %$\Phi_\epsilon$ 
-and $\Phi_\Sigma$.
-\end{definition}
-
-\noindent
-The above transition functions $\wei_\epsilon$ and $\wei_\Sigma$
-are synthesized into a function~$\wei$ 
-from~$Q \times (\Sigma \cup \{ \epsilon \}) \times Q$ into~$\Semiring$ 
-as above.
-%
-When $\wei_\epsilon(q, q') = \zero$ for all $q, q' \in Q$, 
-the automaton~$A$ is called \emph{without $\epsilon$-transitions}.
-
-\subsection{Properties}
-The two following properties will be useful to our approach on 
-symbolic weighted parsing in Section~\ref{sec:parsing}.
-
-\begin{proposition}
-Given a \SWT $T$ 
-over alphabet $\Sigma$, $\Delta$ and a commutative semiring~$\Semiring$, 
-and $s \in \Sigma^*$, 
-there exists an effectively constructible \SWA 
-$A_{T, s}$ over $\Delta$ and $\Semiring$ 
-such that for all $t \in \Delta^*$, $A_{T, s}(t) = T(s, t)$.
-\end{proposition}
-%
-\begin{proof}
-Let $T = \< Q, \init, \bar{\wei}, \final >$ 
-where $\bar{\wei} \< \wei_\epsilon, \wei_\Sigma, \wei_\Delta, \wei_{\Sigma, \Delta}>$
-has an associated synthesized form $\wei$, 
-and let $s = s_1 \ldots s_n$ with $s_1, \ldots, s_n \in \Sigma$.
-%$n = |s|$ be the length of $s$.
-%
-Let $Q' = [1..n] \times Q$ be the state set of $A_{T, s}$,
-and let, for all $q \in Q$, 
-$\init'(1, q) = \init(q)$ and $\init'(i, q) = \zero$ for all $1 < i \leq n$, 
-and $\final'(n, q) = \final(q)$ and $\final'(i, q) = \zero$ for all $1 \leq i <  n$.
-
-\noindent
-We define the functions $\wei'_\epsilon$ and $\wei'_\Sigma$ for $A_{T, s}$
-in their synthesized form: for all $1 \leq i, j \leq n$, $q, q' \in Q$ and $a \in \Sigma$, 
-and $b \in \Delta$, 
-\[
-\begin{array}{lclcl}
-\wei'(\< i, q>, \epsilon, \< i, q'>)   & = & \wei(q, \epsilon, \epsilon, q'), & \quad & \\ 
-\wei'(\< i, q>, \epsilon, \< i+1, q'>) & = & \wei(q, s_i, \epsilon, q') & & \mathrm{if~} i < n,\\ 
-\wei'(\< i, q>, \epsilon, \< j, q'>)   & = & \zero & & \mathrm{if~} j \neq i, j \neq i+1,\\ 
-\wei'(\< i, q>, b, \< i, q'>)          & = & \wei(q, \epsilon, b, q'),\\ 
-\wei'(\< i, q>, b, \< i+1, q'>)        & = & \wei(q, s_i, b, q') & & \mathrm{if~} i < n,\\ 
-\wei'(\< i, q>, b, \< j, q'>)          & = & \zero & & \mathrm{if~} j \neq i, j \neq i+1. 
-\end{array}      
-\]
-Note that the functions defined above belong to the set $\Phi_\Delta$ 
-of the underlying label theory, 
-by the hypothesis of closure under partial application.
-%
-The $\SWA$ wanted is $A_{T, s} = \< Q', \init', \bar{\wei'}, \final' >$.
-\qed
-\end{proof}
-
-The construction time and size for $A_{T, s}$ are $O(\| T \| . | s |)$,
-where the size $\| T \|$ of $T$ is its number of states $|Q|$.
-
-% gen. to transducers
-\begin{proposition}
-Given a \SWA $A$ over alphabet $\Sigma$ 
-and a commutative and bounded semiring $\Semiring$, 
-there exists an effectively constructible \SWA $A'$ 
-without $\epsilon$-transitions 
-such that for all $s \in \Sigma^*$, $A'(s) = A(s)$.
-\end{proposition}
-%
-\begin{proof}  ** Tb revised **   
-Let $A = \< Q, \init, \<\wei_\epsilon, \wei_\Sigma>, \final >$.
-The automaton $A'$ is $\< Q, \init, \<\wei'_\epsilon, \wei'_\Sigma>, \final >$,
-where, for all $q, q' \in Q$,
-$\wei'_\epsilon(q, q') = \zero$  
-and 
-% \[
-% \begin{array}{rcccl}
-% \wei'_\Sigma(q, q') & = &  & \displaystyle\bigoplus_{q_1 \in Q} &
-%                \wei_\Sigma(q, q_1) \otimes  \wei_\epsilon(q_1, q')\\
-%                     &   & \oplus  & \displaystyle\bigoplus_{q_2 \in Q} &
-%                \wei_\epsilon(q, q_2) \otimes \wei_\Sigma(q_2, q')\\
-% \end{array}
-% \]
-\[
-\wei'_\Sigma(q, q') = 
-\displaystyle\bigoplus_{q'' \in Q} \wei_\epsilon(q, q'') \otimes \wei_\Sigma(q'', q')
-\oplus \displaystyle\bigoplus_{q_1 \in Q} \wei_\Sigma(q, q'') \otimes \wei_\epsilon(q'', q')
-\]
-In the above definition of $\wei'_\Sigma$ we use the operator
-of product of function of $\Phi_\Sigma$ by $\Semiring$ 
-described in Section~\ref{section:symbols}.
-%
-By definition of $\weight_{A}$ and distributivity of $\otimes$ on $\oplus$, 
-** NO. TBC see \cite{Mohri02ijfcs} **
-it holds that $\weight_{A}(q, s, q') = \weight_{A'}(q, s, q')$.
-\qed
-\end{proof}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SW-VPA
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{SW Visibly Pushdown Automata}
-\label{section:SWVPA}\label{sec:SWVPA}
-The following model generalizes Symbolic VPA~\cite{dAntonyAlur14SVPDA}
-from Boolean semirings to arbitrary semiring weight domains.
-
-\subsection{Definition}
-Let $\Omega$ be a countable alphabet 
-%finite (large) or infinite,
-that we assume partitioned into three subsets 
-$\Omegai$, $\Omegac$, $\Omegac$,
-whose elements are respectively called 
-\emph{internal}, \emph{call} and \emph{return} symbols.
-% \begin{itemize}
-% \item a set $\Omegai$ of \emph{internal symbols} denoted $a$,
-% \item a set $\Omegac$ of \emph{call symbols} denoted $\call{a}$,
-% \item a set $\Omegar$ of \emph{return symbols} denoted $\return{a}$.
-% \end{itemize}
-Let~$\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative semiring and  
-let  $(\Phi_\epsilon, \Phic, \Phir, \Phicr)$ be a label theory over $\Semiring$
-%In order to simplify notations, %and following the definition of Section~\ref{section:transducer}, 
-%we shall write respectively 
-where $\Phic$, $\Phir$ and~$\Phicr$ stand respectively 
-for~$\Phi_\Omegac$, $\Phi_\Omegar$ and~$\Phi_{\Omegac, \Omegar}$.
-%
-Moreover, we extend this theory with a set $\Phii$ 
-of unary functions in $\Omegai \to \Semiring$,
-closed under $\oplus$ and $\otimes$.
-
-\begin{definition}
-A \emph{Symbolic Weighted Visibly Pushdown Automata} (\SWVPA) 
-over  $\Omega = \Omegai \uplus \Omegac \uplus \Omegar$ and $\Semiring$ is a tuple
-$A = \< Q, P, \init, \weii, \weic, \weir, \weie, \final >$,
-where $Q$ is a finite set of states, 
-$P$ is a finite set of stack symbols, 
-$\mathsf{in} : Q \to \Semiring$, 
-respectively $\mathsf{out} : Q \to \Semiring,$
-are functions defining the weight for entering, 
-respectively leaving, a state, 
-and 
-$\weii : Q \times Q \to \Phii$,  
-$\weic : Q \times Q \times P \to \Phic$,  
-$\weir : Q \times P \times Q \to \Phicr$,  
-$\weie : Q \times Q \to \Phir$,  
-are transition functions.
-\end{definition}
-%
-Similarly as in Section~\ref{section:transducer}, 
-we extend the above transition functions as follows
-for all $q, q' \in Q$, $p \in P$, 
-$a \in \Omegai$, 
-$\call{c} \in \Omegac$, 
-$\return{r} \in \Omegar$, 
-overloading their names: % for simplicity:
-\[
-\begin{array}{lll}
-\weii: Q \times \Omegai \times Q \to \Semiring & 
-\weii(q, a, q') = \phi_\mathsf{i}(a) & 
-\mathrm{where~} \phi_\mathsf{i} = \weii(q, q'),\\
-%
-\weic: Q \times \Omegac \times Q \times P \to \Semiring & 
-\weic(q, \call{c}, q', p) = \phi_\mathsf{c}(\call{c}) & 
-\mathrm{where~} \phi_\mathsf{c} = \weic(q, q', p),\\
-%
-\weir: Q \times \Omegac \times P \times \Omegar \times Q \to \Semiring & 
-\weir(q, {\call{c}},  p, {\return{r}}, q') = \phi_\mathsf{r}({\call{c}},  {\return{r}}) & 
-\mathrm{where~} \phi_\mathsf{r} = \weir(q, p, q'),\\
-%
-\weie: Q \times \Omegar \times Q \to \Semiring & 
-\weie(q, {\return{r}}, q') = \phi_\mathsf{e}({\return{r}}) &
-\mathrm{where~} \phi_\mathsf{e} = \weie(q, q').\\
-\end{array}      
-\]
-
-\noindent
-The intuition is the following for the above transitions.
-\begin{description}
-\item $\weii$ : read the input internal symbol $a$, change state to $q'$ (stack is untouched).
-\item $\weic$ : read the input call symbol $\call{c}$, push it to the stack along with $p$, change state to $q'$.
-\item $\weir$ : when the stack is not empty, 
-      read and pop from stack a pair made of $\call{c}$ and $p$, 
-      read the input return symbol $\return{r}$, change state to $q'$.
-      In this case, the weight function $\phi_\mathsf{r}$ 
-      computes a value of matching between the call and return symbols $c$ and $r$.
-      This value might be set to $\zero$ in order to express that the symbols do not match.
-\item $\weie$ : when the stack is empty, 
-      read the input symbol $\call{r}$, change state to $q'$.
-\end{description}
-
-We give now a formal definition of these transitions of the automaton~$A$
-in term of a weight value
-computed by an intermediate function $\weight_A$, like in Section~\ref{sec:SWT}.
-In the case of a pushdown automaton, a configuration is composed 
-of a state $q \in Q$ and a stack content $\gamma \in \Gamma^*$, where $\Gamma = \Omegac \times P$.
-Therefore, $\weight_A$ is a function from 
-$Q \times \Gamma^* \times \Omega^* \times Q \times \Gamma^*$ into $\Semiring$.
-\[
-\begin{array}{rcl}
-\weight_A\bigl(\config{q}{\gamma}, a\, u, \config{q'}{\gamma'}\bigr) & = & 
- {\displaystyle\bigoplus_{q'' \in Q}} \weii(q, a, q'') 
- \otimes \weight_A\bigl(\config{q''}{\gamma}, u, \config{q'}{\gamma'}\bigr)\\
-%
-\weight_A\bigl(\config{q}{\gamma}, {\call{c}}\, u, \config{q'}{\gamma'}\bigr) & = & 
- {\displaystyle\bigoplus_{\begin{array}{c}
-                          \scriptstyle q'' \in Q\\[-2pt]
-                          \scriptstyle p \in P
-                          \end{array}}}
- \weic\bigl(q, {\call{c}}, q'', p\bigr) 
- \otimes \weight_A\bigl(\config{q''}{{\call{c}}\, p\cdot \gamma}, u, \config{q'}{\gamma'}\bigr)\\[1mm]
-%
-\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot \gamma}, {\return{r}}\, u, \config{q'}{\gamma'}\bigr) & = & 
- {\displaystyle\bigoplus_{q'' \in Q}} 
- \weir\bigl(q, {\call{c}}, p, {\return{r}}, q''\bigr) 
- \otimes \weight_A\bigl(\config{q''}{\gamma}, u, \config{q'}{\gamma'}\bigr)\\
-%
-\weight_A\bigl(\config{q}{\bot}, {\return{r}}\, u, \config{q'}{\gamma'}\bigr) & = & 
- {\displaystyle\bigoplus_{q'' \in Q}} \weie(q, {\return{r}}, q'') 
- \otimes \weight_A\bigl(\config{q''}{\bot}, u, \config{q'}{\gamma'}\bigr)\\
-\end{array}
-\]
-where $\bot$ denotes the empty stack and ${\call{c}}\, p\cdot \gamma$ 
-denotes a stack where the pair made of ${\call{c}} \in \Omegac$ and $p \in P$ is the top symbol 
-and $\gamma$ is the rest of stack.
-
-\noindent
-The weight associated by $A$ to $s \in \Omega^*$
-is then defined as follows,
-following empty stack semantics: 
-\begin{equation}
-A(s)  = 
-{\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
-\mathsf{in}(q) \mathop{\otimes} 
-\weight_A\bigl(\config{q}{\bot}, s, \config{q'}{\bot}\bigr) 
-\mathop{\otimes} \mathsf{out}(q').
-\label{eq:weightA}
-\end{equation}
-
-\begin{example}
-structured words...
-intro language of music notation ?
-\end{example}
-
-\subsection{Properties}
-Like VPA and symbolic VPA, 
-the class of \SWVPA is closed under the binary operators of the underlying semiring.
-%
-\begin{proposition}
-Let $A_1$ and $A_2$ be two (\SWVPA)
-over the same $\Omega$ and $\Semiring$.
-There exists two $\SWVPA$ $A_1 \oplus A_2$ and $A_1 \otimes A_2$, 
-effectively constructible, 
-such that for all $s \in \Omega^*$, 
-$(A_1 \oplus A_2)(s) = A_1(s) \oplus A_2(s)$ and 
-$(A_1 \otimes A_2)(s) = A_1(s) \otimes A_2(s)$.
-\end{proposition}
-The construction is essentially the same 
-as in the case of the Boolean semiring~\cite{dAntonyAlur14SVPDA}.
-
-\subsection{Best-first Search} 
-\label{sec:best}\label{sec:search}
-Let us assume that the semiring $\Semiring$ is
-commutative, bounded, complete, and total.
-% total ? 
-% monotonic and superior writ natural ordering
-%Regarding the infinite sum operator, note that
-%$\bigoplus_{x \in \Phi_\Omega} \phi(x)$, 
-%$\bigoplus_{y \in \Phi_\Delta} \psi(y)$, and  
-%... exist and in $\Semiring$.
-We propose a Dijkstra algorithm computing the minimal weight by $A$, 
-\wrt $\leq_\oplus$, for a word in $\Semiring^*$.
-
-\noindent
-More precisely, 
-let $\top$ be a fresh stack symbol which does not belong to $\Gamma$,
-and for every two states $q, q' \in Q$ and $\sigma \in \{ \bot, \top \}$, 
-let 
-\[ 
-  d_0(q, \sigma, q') = \bigoplus_{s\in \Omega^*} 
-  \textstyle
-  \weight_A\bigl(\config{q}{\sigma}, s, \config{q'}{\sigma}\bigr). 
-\]
-Since $\Semiring$ is complete, this infinite sum is well defined, 
-and since $\leq_\oplus$ is assumed total, it is the minimum in $\Omega^*$
-of $s \mapsto \weight_A(\config{q}{\sigma}, s, \config{q'}{\sigma})$ \wrt this ordering.
-%
-When $\sigma = \bot$, $d_0(q, \sigma, q')$ is the central 
-expression in a term of the definition (\ref{eq:weightA}) of $A(s_0)$ for the minimum $s_0$
-(for the above function).
-When $\sigma = \bot$, intuitively, it is the minimum weight of a computation of $A$
-starting in state $q$ with a stack $\gamma \in \Gamma^*$ (possibly empty), 
-and ending in state $q$ with the same stack $\gamma$, such that moreover 
-the computation does not touch a symbol of $\gamma$. 
-That means that during the computation, 
-$A$ may apply the first case of in the definition of $\weight_A$ (internal symbol),
-as well as the second case, to can push call symbols on the top of $\gamma$, 
-and may pop these symbols with the third case (return symbol). 
-However, it cannot apply one of the two last cases (return symbol and empty stack)
-when the current stack is $\gamma$.
-%pop symbols in $\gamma$.
-% Note that having a stack reduced to such a symbol makes impossible the application of the 
-% two last cases in the definition of $\weight_A$ (return symbol and empty stack). 
-% However, it is possible to apply the two first cases 
-% (internal symbol or call symbol, with a push on the top of $\top$).
-
-The algorithm~\ref{algo:Dijkstra}
-constructs iteratively a marking $d : Q \times \{ \bot, \top \} \times Q \to \Semiring$
-%of the triplets $\<q, \sigma, q'>$ 
-%of states of $A$ by weight values in $\Semiring$, 
-that converges eventually to $d_0(q, \sigma, q')$. 
-%It uses for that purpose a priority queue $P$ containing triplets of 
-%$Q \times \{ \bot, \top \} \times Q$.
-
-\begin{algo}[1-best for \SWVPA] \label{algo:Dijkstra}
-\textbf{initially} let $P = Q \times \{ \bot, \top \} \times Q$, %contains all %triplets 
-%$\< q_1, \bot, q_2>$ and $\<q_1, \top, q_2>$ for $q_1, q_2 \in Q$, with 
-and 
-$d(q_1, \bot, q_2) = d(q_1, \top, q_2) = \one$ if $q_1 = q_2$ and
-$d(q_1, \bot, q_2) = d(q_1, \top, q_2) = \zero$ otherwise.
-
-\smallskip\noindent
-\textbf{while} $P$ is not empty
-
-\noindent
-\quad\textbf{extract} $\< q_1, \sigma, q_2>$ from $P$ 
- such that $d(q_1, \sigma, q_2)$ is minimal wrt $\leq_\oplus$.
-
-\noindent\quad \textbf{for all} $q_0, q_3 \in Q$ and $p \in P$ do
-
-\noindent
-\qquad
-\(\begin{array}{rrcl}
- & d(q_1, \sigma, q_3) & \opluseq &
-   d(q_1, \sigma, q_2) \otimes \displaystyle\bigoplus_{a \in \Omegai} \weii(q_2, a, q_3)\\
-%
-%  & d(q_0, \sigma, q_2) & \opluseq & 
-%    \displaystyle\bigoplus_{a \in \Omegai} \weii(q_0, a, q_1) \otimes d(q_1, \sigma, q_2)\\
-%
-\mathrm{if~} \sigma = \top & d(q_0, \top, q_3) & \opluseq &
-d(q_1, \sigma, q_2) \otimes
-%\displaystyle\bigoplus_{p \in P} 
-\displaystyle\bigoplus_{{\call{c}} \in \Omegac} 
-\displaystyle\bigoplus_{{\return{r}} \in \Omegar} \eta({\call{c}}, {\return{r}})\\
-\mathrm{and} &  d(q_0, \bot, q_3) & \opluseq &
-d(q_1, \sigma, q_2) \otimes
-%\displaystyle\bigoplus_{p \in P} 
-\displaystyle\bigoplus_{{\call{c}} \in \Omegac} 
-\displaystyle\bigoplus_{{\return{r}} \in \Omegar} \eta({\call{c}}, {\return{r}})\\
-& & & \mathrm{where~} \eta = \weic(q_0, q_1, p) \otimes \weir(q_2, p, q_3)\\
-%
-\mathrm{if~} \sigma = \bot & d(q_1, \bot, q_3) & \opluseq &
-  d(q_1, \sigma, q_2) \otimes \displaystyle\bigoplus_{\return{r} \in \Omegar} \weie(q_2, {\return{r}}, q_3)\\
-%
-%& d(q_0, \bot, q_2) & \opluseq & 
-%d(q_0, \bot, q_1) \otimes d(q_1, \sigma, q_2), \mathrm{if~} \< q_0, \bot, q_1> \notin P \\
-%
-%\mathrm{if~} \sigma = \top & d(q_0, \top, q_2) & \opluseq & 
-%d(q_0, \top, q_1) \otimes d(q_1, \sigma, q_2), \mathrm{if~} \< q_0, \top, q_1> \notin P \\
-%
-& d(q_1, \bot, q_3) & \opluseq & 
-   d(q_1, \sigma, q_2) \otimes d(q_2, \bot, q_3), \mathrm{if~} \< q_2, \bot, q_3> \notin P \\
-%
-\mathrm{if~} \sigma = \top & d(q_1, \top, q_3) & \opluseq & 
-    d(q_1, \sigma, q_2) \otimes d(q_2, \top, q_3), \mathrm{if~} \< q_2, \top, q_3> \notin P \\
-\end{array}
-\) 
-\end{algo}
-
-\noindent
-The infinite sums in the updates of $d$ in Algorithm~\ref{algo:Dijkstra} are well defined
-since~$\Semiring$ is complete.
-The algorithm performs $2.|Q|^2$ iterations until $P$ is empty, 
-and each iteration has a time complexity $O(|Q|^2 . |P|)$.
-This gives a time complexity $O(|Q|^4 . |P|)$. 
-It can be reduced by implementing $P$ as a priority queue, 
-prioritized by the value returned by $d$
-***complete***. %$|Q|^3.\log(|Q|^2)$
-
-The correctness of Algorithm~\ref{algo:Dijkstra} 
-is ensured by the invariant expressed in the following lemma.
-\begin{lemma}
-For all $\< q_1, \sigma, q_2> \notin P$, $d(q_1, \sigma, q_2) =  d_0(q_1, \sigma, q_2)$.
-\end{lemma}
-The proof is by contradiction, 
-assuming a counter-example minimal in the length of the witness word.
-
-% \begin{algorithm}
-% \caption{<your caption for this algorithm>}
-% \label{<your label for references later in your document>}
-% \begin{program}
-% \seq{|line| := `` '', i := 1};
-% \WHILE i \neq n+1 \DO
-%       |line| := |item|[i] \concat `` '' \concat |number|[i]; 
-%       i := i+1; 
-%       \WHILE i \neq n+1 \AND |item|[i] = |item|[i-1] \DO 
-%       |line| := |line| \concat ``, '' \concat |number|[i]);
-%       i := i+1 \OD ; 
-%       |write|(|line|) \OD 
-% \end{program}
-% \end{algorithm}
-
-% The weight of a transition acts as a guard: 
-% a transition is activated for a symbol $`a`$ iff its weight $`\phi(a) \neq 0`$ (or $`\psi(a, b) \neq 0`$), 
-% the absorbing element.
-
-%\begin{lemma}
-\noindent
-For computing the minimal weight of a computation of $A$, we use the fact that,
-at the termination of Algorithm~\ref{algo:Dijkstra}, %it holds that,
-%There exist $q_1, q_2 \in Q$
-\[
-  {\displaystyle \bigoplus_{s \in \Omega^*} A(s)} = 
-  {\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
-  \mathsf{in}(q) \mathop{\otimes} d(q, \bot, q') \mathop{\otimes} \mathsf{out}(q').
-\]
-%\end{lemma}
-
-%\medskip
-\noindent
-In order to obtain effectively a witness 
-(word of $\Semiring^*$ with computation of $A$ of minimal weight), 
-we require an additional property the of weight functions.
-\begin{definition}
-Let $\Omega$ be an alphabet and $\Semiring$ a complete semiring.
-A function $\phi$ from $\Omega^n$ into $\Semiring$
-is called $k$-\emph{convex} for a natural number $k$ iff 
-$\mathit{card}\{ \mathbf{a} \in \Omega^n \mid \phi(\mathbf{a}) = 
-                 \displaystyle\bigoplus_{\mathbf{p} \in \Omega^n} \phi(\mathbf{p}) \} \leq k$.
-\end{definition}
-A label theory is $k$-convex if all its functions are $k$-convex.
-
-\begin{proposition}
-For a \SWVPA $A$ over a 
-commutative, idempotent, superior, total and complete semiring 
-and an alphabet $\Omega$ with a $k$-convex label theory,
-one can construct in PTIME a word $s \in \Omega^*$ 
-such that $A(s)$ is minimal \wrt the natural ordering for $\Semiring$. 
-%$\leg_\oplus$.
-\end{proposition}
-
-\subsection{Trees and Nested-Words}
-\label{sec:trees}
-The hierarchical structure of nested-words, defined the \emph{call} and \emph{return} markup symbols of 
-suggest a correspondence between these word and trees. 
-The lifting of this correspondence to languages, respective of tree automata and VPA have
-been discussed in~\cite{AlurMadhusudan09nested}, 
-see also~\cite{Caralp12VPAmult} for the weighted case.
-In this section, we discuss the correspondence between the symbolic-weighted extensions
-of tree automata and VPA.
-
-Let $\Omega$ be a countable ranked alphabet, such that 
-every symbol $a \in \Omega$ has a rank 
-$\rank(a) \in [0..M]$ where $M$ is a fixed natural number.
-We write $\Omega_k$ the subset of all symbols $a$ of $\Omega$
-with $\rank(a) = k$, where $0 \leq k \leq M$.
-%$\cup_{i = 0}^{i = M} \Omega_i$ 
-%
-$\T(\Omega)$ denotes the free $\Omega$-algebra of finite, ordered, 
-$\Omega$-labeled trees, 
-which is the smallest set such that  $\Omega_0 \subset \T(\Omega)$
-and for all $1 \leq k \leq M$, all $a \in \Omega_k$, 
-and all $t_1, \ldots, t_k \in \T(\Omega)$, $a(t_1, \ldots, t_k) \in \T(\Omega)$.
-%
-% tree = single node (leave) labeled with a symbol of $a \in \Omegai$
-% (such a tree is simply denoted by $a$)
-% or the composition, denoted by $b(t_1,\ldots, t_n$) of a node labeled with $b$
-% and $n$ subtrees $t_1$,\ldots, $t_n$.
-
-\renewcommand{\call}[1]{\ensuremath \langle_{#1}}
-\renewcommand{\return}[1]{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
-
-Let $\hat\Omega$ be the countable (unranked) alphabet obtained from 
-$\Omega$ as follows: 
-$\hat\Omega = \< \Omegai, \Omegac, \Omegar >$, with
-$\Omegai = \Omega_0$, 
-$\Omegac = \{ \; \call{a} \mid a \in \Omega_{>0} \}$,
-$\Omegac = \{ \; \return{a} \mid a \in \Omega_{>0} \}$,
-where $\Omega_{>0}$ denotes $\bigcup_{k = 1}^{M} \Omega_k$.
-
-\noindent
-We define a linearization of trees of $\T(\Omega)$ into 
-words of $\hat{\Omega}^*$ as follows:
-\begin{description}
-\item $\lin(a) = a$ for all $a \in \Omega_0$, 
-\item $\lin\bigl( b(t_1, \ldots, t_k)\bigr) = 
-       \call{b} \; \lin(t_1) \ldots \lin(t_k) \; \return{b}$ 
-       when $b \in \Omega_k$, $1 \leq k \leq M$.
-\end{description}
-
-% A \emph{regular tree grammar} over $\Omega$ 
-% is a triplet $G = \< N, q_\mathsf{i}, R>$ where
-% $N$ is a finite set of non-terminal symbols denoted $q$..., 
-% $q_\mathsf{i} \in N$ is the starting non-terminal, 
-% $R$ is a finite set of production rules of the form
-% $q_0 \to a(q_1\ldots q_k)$ where 
-% $q_0, q_1, \ldots, q_k \in N$
-% $a \in \Omega_k$.
-% A tree $t \in \T(\Omega)$ is in the language of $G$ 
-% if it can be generated from $q_\mathsf{i}$ by 
-% non terminal replacement following the rules of $R$.
-
-Let us assume a label theory $\Phi_{\Omega_k}$ 
-for each $k \in [0..M]$.
-%
-\begin{definition}  \label{def:SWTA}
-A \emph{symbolic-weighted tree automaton} (\SWTA)
-over the ranked input alphabet~$\Omega$
-and the commutative semiring $\Semiring$ 
-is a triplet $A = \< Q, \init, \bar{\wei} >$ where
-$Q$ is a finite set of states, 
-$\mathsf{in} : Q \to \Semiring$ is the starting weight function, 
-and $\bar{\wei}$ is a $M+2$-uplet of transition functions made of:
-$\wei_\epsilon$ from $Q \times Q$ into $\Semiring$, and, 
-for each $k \in [0..M]$, $\wei_{\Omega, k}$
-from $Q \times Q^{k}$ into $\Phi_{\Omega_k}$.
-\end{definition}
-%
-Like in Section~\ref{sec:SWAdef}, we define from $\bar{\wei}$ 
-a transition function~$\wei$, %from $\bar\wei$
-from 
-$Q \times (\Sigma \cup \{ \epsilon \}) \times \bigcup_{k=0}^{M} Q^k$
-into~$\Semiring$: %also called $\wei$ for simplicity, 
-%such that, for all $q, q' \in Q$, $a \in \Sigma$, and $b \in \Delta$, 
-\[
-\begin{array}{rcll}
-\wei(q_0, \epsilon, q_1) & = &  \wei_\epsilon(q_0, q_1),\\ %\phi_\epsilon\\
-\wei(q_0, a, q_1 \ldots q_k) & = & \phi_{\Omega, k}(a) &
-\quad\mathrm{where~} \phi_{\Omega, k} = \wei_{\Omega, k}(q_0, q_1\ldots q_k).\\
-\end{array}      
-\]
-Intuitively, $\wei(q_0, a, q_1 \ldots q_k)$ can be seen as
-the weight of a production rule $q_0 \to a(q_1, \ldots, q_k)$ 
-of a regular tree grammar~\cite{tata}, 
-that replaces the non-terminal symbol $q_0$, by $a(q_1, \ldots, q_k)$. 
-%in a step of tree building.
-%
-Such a grammar computes the weights of the derivation trees 
-of the Context-Free grammar obtained by forgetting the labeling symbols of $\Omega_{>0}$.
-
-\noindent
-The \SWTA of Definition~\ref{def:SWTA} defines a mapping 
-from trees of $\T(\Omega)$ into the weights of~$\Semiring$,
-based on the intermediate function $\weight_A$
-defined as follows for $q_0 \in Q$ and 
-$t = b(t_1,\ldots, t_k) \in \T(\Omega)$,
-åwith $0 \leq k \leq M$:
-\[
-\begin{array}{rccl}
-\weight_A(q_0, t) & = & 
-   \displaystyle\bigoplus_{q_1 \in Q} &
-   \wei(q, \epsilon, q_1) \otimes \weight_A(q_1, t)\\
- & \oplus & \displaystyle\bigoplus_{q_1 \ldots q_k \in Q^k} &
-              \wei(q_0, b, q_1 \ldots q_k ) 
-   \otimes \displaystyle\bigotimes_{i=1}^{k}
-           \weight_A(q_i, t_i).\\
-\end{array}
-\]
-
-\medskip\noindent
-The weight associated by $A$ to  $t \in \T(\Omega)$ is then
-\begin{equation}
-A(t)  = 
-\displaystyle\bigoplus_{q \in Q} \mathsf{in}(q) \mathop{\otimes} \weight_A(q, t).
-\label{eq:weightTA}
-\end{equation}
-
-\begin{lemma}  \label{lem:SWTA}
-For all \SWTA $A$ over~$\Omega$ and $\Semiring$, without $\epsilon$-transitions,
-there exists an effectively constructible \SWVPA $A'$ over 
-$\hat\Omega$ and $\Semiring$
-such that for all $t \in \T(\Omega)$, $A'\bigl(\lin(t)\bigr) = A(t)$.
-\end{lemma} 
-% 
-\begin{proof}
-Let $A = \< Q, \init, \bar{\wei} >$ where $\bar{\wei}$ is summarized as above by a function
-$\wei: Q \times (\Sigma \cup \{ \epsilon \}) \times \bigcup_{k=0}^{M} Q^k \to \Semiring$.
-
-We build 
-$A' = \< Q', P', \init', \weii, \weic, \weir, \weie, \final' >$,
-computing over $\hat\Omega = \< \Omegai, \Omegac, \Omegar >$,
-%
-%$\Omega = \Omegai \uplus \Omegac \uplus \Omegar$ and $\Semiring$ 
-where $Q' = \bigcup_{k=0}^{M} Q^k$ be  the set of sequences of state symbols of $A$, 
-of length at most $M$, including the empty sequence denoted by $\epsilon$, 
-and where $P' = Q'$.
-
-\[
-\begin{array}{ll}
-%\weii: Q \times \Omegai \times Q \to \Semiring & 
-\weii(\bar{q}, a, \bar{q} q') = \wei(q', a, \epsilon) & 
-\mathrm{for~all~} a \in \Omega_0,\\
-%
-%\weic: Q \times \Omegac \times Q \times P \to \Semiring & 
-\weic(\bar{q}, \call{b}, \epsilon, \bar{q}) = \one & %\wei(p', b, \bar{p}) & 
-\mathrm{for~all~} b \in \Omega_{>0},\\
-%
-%\weir: Q \times \Omegac \times P \times \Sigmar \times Q \to \Semiring & 
-\weir(\bar{q}, {\call{b}}, \bar{p}, {\return{b}}, \bar{p} q') = \wei(q', b, \bar{q}) & 
-\mathrm{for~all~}  b \in \Omega_{>0},\\
-%
-%\weie: Q \times \Omegar \times Q \to \Semiring & 
-\weie(\bar{p}, {\return{b}}, \bar{q}) = \zero &
-\mathrm{for~all~}  b \in \Omega_{>0}.\\
-\end{array}      
-\]
-
-In practice, it is sufficient to consider in $Q'$ only the prefixes of 
-sequences.
-\qed\end{proof}
-
-\section{Symbolic Weighted Parsing}
-\label{sec:parsing}
-Let us now use the models and results of the former sections in %order to define 
-an approach to the problem of parsing over infinite alphabet. %appropriate
-%
-Besides considering infinitely many possible of input symbols, 
-handled with suitable language formalisms, 
-this approach extends conventional parsing 
-by computing a derivation tree modulo 
-a generic distance between words,  
-defined by a SW transducer given in input.
-This enables considering finer word relationships than strict equality 
-as in the conventional parsing approach, 
-opening possibilities of quantitative analysis via this method.
-
-\subsection{Definition}
-
-Let $\Sigma$ be a countable input alphabet 
-and $\Omega = \Omegai \uplus \Omegac \uplus \Omegar$ an output countable alphabet,
-let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative, bounded, complete 
-and total semiring 
-and let  $(\Phi_\epsilon, \Phic, \Phir, \Phicr)$ be a label theory over $\Semiring$, 
-assumed computable and $k$-convex for some fixed $k$.
-
-\noindent
-Assuming given in input:
-\begin{itemize}
-\item a \SWT $T$ over $\Sigma$,  $\Omega$, and $\Semiring$,
-defining a measure %between words 
-$T: \Sigma^* \times \Omega^* \to \Semiring$,
-\item a \SWVPA $A$ over $\Omega$, and $\Semiring$, defining a series of nested words
-      $A : \Omega^* \to \Semiring$,
-\item an input word $s \in \Sigma^*$,
-\end{itemize}
-the problem of \emph{symbolic weighted parsing} is to find 
-a nested word $t \in \Omega$ minimizing
-\( T(s, t) \otimes A(t)\)
-\wrt $\leq_\oplus$, 
-\ie such that %Hence, it is the problem of finding 
-\( T(s, t) \otimes A(t) = 
-  \displaystyle\bigoplus_{t' \in \Omega^*} T(s, t') \otimes A(t') \).
-
-\noindent 
-Therefore, it is the problem of optimizing a measure 
-called the \emph{edit-distance between $s$ and $A$} in~\cite{Mohri03EDWA}.
-%
-The input language can also be expressed as a \SWTA, or, 
-as a particular case, as a weighted context-free grammar, 
-converted in turn into a \SWVPA following Lemma~\ref{lem:SWTA}.
-%
-In the case of finite alphabets, 
-the problem of searching, in a WTA language, 
-for the best parse tree for a given,  
-sometimes referred as~\emph{weighted parsing}
-(see \cite{Goodman99SemiringParsing,MorbitzVogler19weighted-parsing} **more general problems**)
-is a particular case of SW parsing.
-%
-Indeed, it corresponds to the case where $T$
-accepts only the pairs $\<s, t>$ such that 
-$s$ is the projection of $t$ on $\Omegai$. 
-This can be done with a single state $q$ and 
-with transition rules of the form:
-\begin{description}
-\item[] $\wei(q, \epsilon, a, q) = \one$ for all $a \in \Omegac \cup \Omegar$,
-\item[] $\wei(q, a, a, q) = \one$ for all $a \in \Omegai$,
-\item[] $\wei(q, a, b, q) = \zero$ for all $a, b \in \Omegai$, $a \neq b$.
-\end{description}
-
-\subsection{Computation}
-
-\begin{proposition}
-The problem of Symbolic Weighted  parsing 
-can be solved in PTIME in the size of the input \SWT, \SWVPA (or \SWTA) and input word, 
-and the computation time of the functions of the label theory.
-\end{proposition}
-%
-\begin{proof}
-Bar-Hillel construction
-\qed  
-\end{proof}
-
-\subsection{Application to Automated Music Transcription}
-\label{sec:transcription}
-Symbolic Automated Music Transcription
-and analysis of music performances
-
-\subsubsection{Time Scales}
-Real-Time Unit (RTU) = seconds
-
-\noindent 
-Musical-Time Unit (MTU) = number of measures
-
-\noindent 
-conversion via tempo value
-
-\subsubsection{Representation of Music Performances}
-We consider symbolic representations of musical performances, as finite sequences of events.
-It corresponds to the concrete case of a MIDI file~\cite{SMF} 
-recorded  from an electronic keyboard, 
-or the output of a transcription from audio files~\cite{Benetos18AMTsurvey}.
-%
-For the sake of simplicity, 
-we shall only consider here the case of monophonic performances, 
-where at most one note is sounding at a time. 
-The approach however extends to the polyphonic case.
-
-A music performance is a finite sequence of events in a set~$\Sigma$.
-Every event $e \in \Sigma$ has attributes such from a finite domain, 
-like a number of key for a note 
-or a flag indicating that it is a rest 
-(\textsf{ON} and \textsf{OFF} messages in~\cite{SMF})
-and a velocity value (0..127 in~\cite{SMF})).
-%This representation is similar to the piano roll ~\cite{Muller15fundamentals} chap.1. 
-Moreover, it contains a RTU value $\ioi{e}$ (real number) 
-representing the time distance to the next event, 
-or to the end of performance for the last event,
-also called \emph{inter-onset interval}.
-
-\subsubsection{Representation of Music Scores}
-Music score are represented as structured words
-made of timed %quantified 
-events and parenthesized markups,
-akin of nested words~\cite{AlurMadhusudan09nested}.
-
-We consider an alphabet $\Delta$, every symbol of which is 
-composed of a tag, in a finite set $\Xi$, 
-and an MTU (rational) IOI duration value.
-%The alphabet $\Delta$ 
-It is partitioned into 
-$\Delta = \Deltai \uplus \Deltac \uplus \Deltar$, 
-like in Section~\ref{section:SWVPA}.
-%
-\noindent
-The symbols of $\Deltai$ represent events:
-% (infinite alphabet of internal symbols) made of:
-with tags indicating a new note or grace-note (with null IOI), 
-a rest or the continuation of the previous note (tie or dot).
-%
-The elements of $\Deltac \uplus \Deltar$ are matched
-markups for describing the structure of the score, 
-\ie the hierarchical grouping of events, and also, 
-importantly the division of time in measures, tuplets...
-%- parentheses for time divisions : tuplets, bars...
-(linearization of rhythm trees \cite{jacquemard:hal-01138642}...).
-They contain additional info such as tuple number, beaming policy...
-
-\noindent
-The duration values of letters of $\Delta$, in MTU (rational), 
-can be computed with the markups and tags (\eg grace note has duration 0).
-
-%\noindent
-%There are simultaneous events, since grace notes has duration 0. They are ordered.
-%
-%\noindent
-%Finite bound on the number of duration ratio. ?
-
-\begin{example}
-...      
-\end{example}
-
-\subsubsection{Performance/Score Distance Computation}
-\label{app:distance}
-We define a distance between performance and score representations
-by a $\SWT$ $T = \< Q, \init, \wei, \final >$, over a semiring $\Semiring$.
-** detail the elements of $\Semiring$ ....**
-%are quadruplets of the form
-%$\< t, s, \delta_t, \delta_s>$
-
-Every state of $Q$ contains a 
-tempo value in a finite domain (e.g. 30..300 bpm).
-This value can be fixed 
-or recomputed by the $T$ %transducer 
-after reading each event, 
-according to a perceptive/cognitive model of tempo 
-such as~\cite{LargeJones99tempo}
-(also used in the context of score following~\cite{Cont10TPAMI}).
-% we wont detail here.
-
-\subsection{Transcription by SW Parsing} %Best-first Search}
-We assume a score language defined by a \SWVPA over the semiring 
-$\Semiring$ of Section~\ref{app:distance}.
-
-\section*{Conclusion}
-% discussion
-
-\noindent
--- summary
-
-\noindent
--- other theoretical properties of SW models
-
-\noindent
--- room to improve complexity for best-search algorithm
-... modular approach with oracles ...
-
-and extention to $n$-best
-
-\noindent
--- offline algorithm, 
-semi-online implementation for AMT (bar-by-bar approach)
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% BIBLIO                                                                     %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\bibliographystyle{plain}
-%\bibliographystyle{plainurl} 
-\bibliographystyle{abbrv}
-%\bibliographystyle{splncs04}
-\bibliography{references}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% APPENDIX                                                                   %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\appendix
-
-\section{Edit-Distance}
-
-%\subsection{Distance between words or languages}
-...algebraic definition of edit-distance of Mohri, in \cite{Mohri03EDWA}
-% Mehryar Mohri 
-% Edit-distance of weighted automata: General definitions and algorithms
-% International Journal of Foundations of Computer Science 14.06 (2003): 957-982.
-distance $d$ over $\Sigma^* \times \Sigma^*$ 
-into a semiring  $\Semiring = ( \Semiring, \oplus, \zero, \otimes, \one)$.
-
-%\noindent
-Let $\Omega = \Sigma \cup \{ \epsilon \} \times \Sigma \cup \{ \epsilon \} \setminus \{ (\epsilon, \epsilon) \}$,
-and let $h$ be the morphism from $\Omega^*$ into $\Sigma^* \times \Sigma^*$  
-defined over the concatenation of strings of $\Sigma^*$ (that removes the $\epsilon$'s).
-%
-\noindent
-An \emph{alignment} between 2 strings  $s, t \in \Sigma^*$ is an element $\omega \in \Omega^*$ 
-such that $h(\omega) = (s, t)$.
-%
-\noindent
-We assume a base cost function $\Omega$ : $\delta: \Omega \to S$, extended to $\Omega^*$ as follows  
-(for $\omega \in \Omega^*$): 
-\(
-\displaystyle\delta(\omega) = \bigotimes_{0 \leq i < |\omega|} \delta(\omega_i)
-\).
-
-\noindent
-\begin{definition}
-For  $s, t \in \Sigma^*$, the edit-distance between $s$ and $t$ is  
-\( 
-d(s, t) = \displaystyle\bigoplus_{\omega \in \Omega^*\, h(\omega) = (s, t)} \delta(\omega)
-\).
-\end{definition}
-
-e.g. Levenstein edit-distance: $S$ is min-plus and $\delta(a, b) = 1$ for all $(a, b) \in \Omega$.
-
-%\paragraph{Distance between a word and a regular language}
-
-\end{document}

### 2021-05-17 introduction.tex
example

@@
 % (notion of monoid recognizability, based on registers, comparable to Bojancszik et al. data words)
-Regarding the latter case, we describe %briefly 
+Regarding the latter case, we briefly mention
 at the end of the paper 
-a parse-based approach to automated music transcription,
-where a symbolic music performance, 
+a parse-based approach to automated music transcription~\cite{foscarin:hal-01988990}
+that motivated the present work. 
+In this problem, a symbolic music performance, 
 presented as a sequence of timed musical events, % timestamped
 % a symbolic representation of a music performance 
-is converted into a structured score in Common Western Music Notation~\cite{foscarin:hal-01988990}.
+is converted into a structured score in Common Western Music Notation.
 %generalizes weighted parsing: 
@@
 %
-Finally, Section~\ref{sec:transcription} describes
-the implemented application %of this approach 
-to automated music transcription
-that motivated this study.
+%Finally, Section~\ref{sec:transcription} describes
+%the implemented application %of this approach 
+%to automated music transcription
+%that motivated this study.
 

### 2021-05-17 main.tex
example

@@
 % \newtheorem{remark}[theorem]{Remark}
+\def\endex{\hspace*{\fill} $\Diamond$\smallskip }
 {\theorembodyfont{\rmfamily} \theoremstyle{break} \newtheorem{algo}{Algorithm}}
@@
 %\input{rewriting}
+
+%% music symbols
+% see http://tug.ctan.org/info/latex4musicians/latex4musicians.pdf
+\usepackage{musicography}
 
@@
 
-\def\SA{\textsf{SA}\xspace}
-\def\WA{\textsf{WA}\xspace}
+\def\SA{\textsf{sA}\xspace}
+\def\WA{\textsf{wA}\xspace}
 \def\SWT{\textsf{swT}\xspace}
@@
 %\title{Symbolic Weighted Language Models and Automated Music Transcription}
-\title{Symbolic Weighted Language Models and Parsing over Infinite Alphabets}
+\title{Symbolic Weighted Language Models and\\ Parsing over Infinite Alphabets}
 %\title{Weighted Visibly Pushdown Automata and Automated Music Transcription}
@@
 %\titlerunning{WVPA \& AMT}
-\titlerunning{Symbolic Weighted Parsing and Automated Music Transcription}
+%\titlerunning{Symbolic Weighted Parsing} % and Automated Music Transcription
 %\authorrunning{Florent Jacquemard}
@@
 \begin{abstract}
-%Symbolic Weighted (SW) extension of symbolic automata...
-We propose a framework for weighted parsing over infinite alphabets.
-%
-It is based on language models called Symbolic Weighted Automata (SWA) 
-at the joint %intersection 
-between Symbolic Automata (SA) and Weighted Automata (WA), 
-as well as Transducers (SWT) and Visibly Pushdown (SWPDA) variants.
-%
-Like SA, SWA deal with large or infinite input alphabets, 
-and like WA, they output a weight value in a semiring domain.
-The transitions of WA are labeled by functions from an infinite alphabet into the weight domain. 
-This is unlike WA whose transitions are guarded by boolean predicates 
-overs symbols in an infinite alphabet 
-and also unlike WA whose transitions are labeled by constant weight values, 
-and who deal only with finite automata.
-%
-We present some properties of WA, WT and SWPDA models, 
-%and show how they can be used 
-that we use to define and solve a variant of parsing 
-over infinite alphabets.
-%
-We also briefly describe the application that motivated the introduction of these models: 
-an parse-based approach to automated music transcription.
+\input{abstract}
 \end{abstract}
@@
 \section{Introduction} \label{sec:intro}
-Parsing %variant of membership problem for formal languages
-%- given a language model and an input word $s$, compute a derivation of the model that yields $s$.
-is the problem %process 
-of structuring a linear representation in input %typically
-(a finite word over an alphabet) according to a language model. % (a formal grammar).
-% natural language, programming language, 
-%
-Most of the context-free parsing approaches~\cite{GruneJacobs08parsing}  
-assume a finite and reasonably small input alphabet. %models and algorithms 
-Such a restriction makes perfectly sense in the context of NLP tasks 
-like constituency parsing or of programming languages compilers or interpreters.
-Considering large or infinite alphabets can however be of practical interest
-in other cases.
-%
-For instance, when dealing with large characters encodings such as UTF-16, % processing strings in 
-\eg for vulnerability detection in Web-applications~\cite{dAntoni21CACM}, 
-%
-for the analyse (\eg validation or filtering) 
-of data streams or serialization of structured documents 
-(which may contain textual or numerical attributes)~\cite{Segoufin06csl}, 
-or for processing timed execution traces~\cite{Bouyer03algebraic}.
-% algebraic definition of a class of data languages 
-% (notion of monoid recognizability, based on registers, comparable to Bojancszik et al. data words)
-Regarding the latter case, we describe %briefly 
-at the end of the paper 
-a parse-based approach to automated music transcription,
-where a symbolic music performance, 
-presented as a sequence of timed musical events, % timestamped
-% a symbolic representation of a music performance 
-is converted into a structured score in Common Western Music Notation~\cite{foscarin:hal-01988990}.
-%generalizes weighted parsing: 
-%finding the best derivation for a weighted grammar.
-
-Various extensions of language models for handling infinite alphabets have been studied.
-%words carrying data values in an infinite domain (e.g. integers) e.g. data processing 
-For instance, some automata with memory extensions 
-allow restricted storage and comparison of input symbols, 
-%and correspond with logics, 
-(see~\cite{Segoufin06csl} for a survey),
-with pebbles for marking positions~\cite{NevenSchwentickVianu04FSMinfinite}, 
-registers~\cite{KaminskiFrancez94}, 
-or %computations in 2 steps
-the possibility to compute on subsequences 
-with the same attribute values~\cite{Bojanczyk11FO2}. % data words automata.
-%
-%for the and verification of infinite state systems 
-%(model checkers: alphabet = long bit-vectors)
-%...For the representation of  in model checking, verification and 
-The automata at the core of model checkers
-compute on input symbols represented by large bitvectors~\cite{Vardi07ciaa} %\cite{BaierKatoen08MC}
-(sets of assignments of Boolean variables) %propositional variables)
-and in practice,  %implementation
-each transition accepts a set of such symbols (instead of an individual symbol), 
-represented by Boolean formula or Binary Decision Diagrams.
-%
-Following a similar idea, % and generalizing,
-in symbolic automata (\SA)~\cite{dAntoniVeanes17CAV,dAntoni21CACM}, 
-the transitions are guarded by predicates over infinite alphabet domains.
-With closure conditions on the sets of such predicates, % (alphabet theories), 
-all the good properties enjoyed by automata over finite alphabets are preserved.
-
-Other extensions of language models  %(automata and grammars) 
-help in dealing with non-determinism, by the computation of weight values. 
-With an ambiguous grammar, there may exists several derivations 
-(\emph{abstract syntax trees} -- AST), % \ie the result of an analyze.
-yielding one input word. % structuring a word, 
-The association of one weight value %in semiring domain
-to each derivation permits to select a best one (or $n$ bests). % a fixed number of bests 
-% = ...a ranking of derivations 
-This is roughly the principle of \emph{weighted parsing}
-approaches~\cite{Goodman99SemiringParsing,Nederhof03weightedParsing,MorbitzVogler19weighted-parsing}.
-In \emph{weighted language models}, 
-like \eg probabilistic context-free grammars % (CFG),
-and weighted automata (\WA) \cite{Droste09handbook},
-a weight value is associated to each transition rule, % production rule
-and the rule's weights can be combined with a associative product operator~$\otimes$ into 
-the weight of a derivation.
-A second operator~$\oplus$, associative and commutative, 
-is moreover used to handle ambiguity of the model, 
-by summing the weights of the possibly several (in general exponentially many) AST %syntax trees 
-associated to a given input word.
-Typically, $\oplus$ will select the best of two weight values.
-%Intuitively,~$\oplus$ selects, or ranking, the syntax trees.
-The weight domain, equipped with these two operators is assumed, at minima, 
-to form a \emph{semiring} %$\Semiring$ 
-where $\oplus$ can be extended to infinite sums, 
-%\cite{Eilenberg74automata}
-such as the Viterbi semiring and the tropical min-plus algebra, see Figure~\ref{fig:semirings}. 
-%of domain $\mathbb{R}^+ \cup \{ +\infty\}$, 
-%where $\oplus$ is min and $\otimes$ is plus .
-%by ranking 
-%making the weight domain a semiring.
-%Some efficient specialized parsing algorithms~\cite{Huang05kbest} have been proposed in this context 
-%% models represented as hypergraphs \cite{Huang05kbest}
-%in order to compute the $n$ best syntax trees of a given input word without having to enumerate them all.
-%Generally based on dynamic programming, these algorithms rely on 
-%additional algebraic properties of~$\Semiring$.
-%-- see \eg~\cite{Huang05kbest} for some NLP applications.
-%The extraction of $n$ best list is useful 
-%the problem: quantitative parsing or symbolic parsing
-%of parsing of words over infinite input alphabet.
-
-In this paper, we present a uniform framework for weighted parsing over infinite input alphabets.
-It is based on weighted %finite states 
-language models generalizing 
-both~\SA, with functions in an arbitrary semiring instead of Boolean guards, 
-and~\WA, by handling infinite alphabets -- Figure~\ref{fig:hierarchy}.
-%***
-%some weighted language models
-%model of weighted CFG 
-%computing on words over infinite input alphabets %sets of terminal symbol
-%- but with finite sets of states and transitions rules. 
-% non-terminals and production rules).
-In their transition rules, input symbols appear as variables %(parameters)
-and the weight associated to a transition rule is a function of these variables.
-%
-\begin{figure}
-\centering
-\begin{tikzpicture}
-\node (NFA) at (0,4) {%
-  \(
-  \begin{array}{c} 
-  \mathsf{NFA} : \Sigma_{\mathsf{fin}}^* \to \mathbb{B}\\
-  q \xrightarrow{a} q' \quad  a \in \Sigma_\mathsf{fin}
-  \end{array} 
-  \)
-};
-%
-\node (WA) at (-2,2) {%
-  \(
-  \begin{array}{c} 
-  \mathsf{WA} : \Sigma_{\mathsf{fin}}^* \to \mathbb{S}\\
-  q \xrightarrow{a, w} q' \quad  a \in \Sigma_\mathsf{fin}, w \in \mathbb{S}
-  \end{array} 
-  \)
-};
-%
-\node (SA) at (2,2) {%
-  \(
-  \begin{array}{c} 
-  \mathsf{SA} : \Sigma_{\mathsf{inf}}^* \to \mathbb{B}\\
-  q \xrightarrow{\phi} q' \quad \phi : \Sigma_\mathsf{inf} \to \mathbb{B}
-  \end{array} 
-  \)
-};
-%
-\node (SWA) at (0,0) {%
-  \(
-  \begin{array}{c} 
-  \mathsf{SWA} : \Sigma_{\mathsf{inf}}^* \to \mathbb{S}\\
-  q \xrightarrow{\phi} q' \quad \phi : \Sigma_\mathsf{inf} \to \mathbb{S}
-  \end{array} 
-  \)
-};
-\draw[->] (NFA)--(WA);
-\draw[->] (NFA)--(SA);
-\draw[->] (WA)--(SWA);
-\draw[->] (SA)--(SWA);
-%\begin{array}{c} \mathsf{NFA} : \Sigma^* \to \mathbb{B} \end{array} 
-\end{tikzpicture}
-\caption{Classes of Symbolic/Weighted Automata. 
-$\Sigma_\mathsf{fin}$ is a finite alphabet, 
-$\Sigma_\mathsf{inf}$ is a countable alphabet,
-$\mathbb{B}$ is the Boolean algebra, 
-$\mathbb{S}$ is an arbitrary commutative semiring, 
-$q \xrightarrow{\dots} q'$ represents the form of a transition between states $q$ and $q'$.}
-\label{fig:hierarchy}  
-\end{figure}
-%
-%This approach is close to the case of 
-%Symbolic Automata (SA)~\cite{dAntoniVeanes17CAV,dAntoni21CACM}, 
-%except that the domain for weight values is not restricted to be Boolean, 
-%like for the guards in the rules of SA, 
-%but can be an arbitrary commutative semiring (assuming some restrictions).
-%
-The models presented here are finite automata called symbolic-weighted (\SWA),
-transducers (\SWT) and pushdown automata,
-with a visibly restriction~\cite{AlurMadhusudan09nested} (\SWVPA).
-%\wrt an edit distance
-The latter model of automata computes on \emph{nested words}~\cite{AlurMadhusudan09nested}, 
-a structured form of words parenthesized with markup symbols, 
-corresponding to a linearization of trees.
-In the context of parsing, they are used here to represent (weighted) AST of CF grammars.
-More precisely, a \SWVPA $A$ associates a weight value $A(t)$ % \in \Semiring$ 
-to a given a nested word $t$, which is the linearization of an AST. %representing a parse tree.
-%
-On the other hand, 
-a \SWT is used to define a distance $T(s, t)$ between two finite words $s$ and $t$
-over an infinite alphabet, following~\cite{Mohri03EDWA}.
-Then, the \emph{SW-parsing} problem aims at %computing 
-finding $t$ minimizing 
-$T(s, t) \otimes A(t)$ (\wrt the ranking defined by $\oplus$)
--- this value is called the distance between $s$ and $A$ in~\cite{Mohri03EDWA}.
-%
-Similarly to weighted-parsing 
-methods~\cite{Goodman99SemiringParsing,Nederhof03weightedParsing,MorbitzVogler19weighted-parsing}, 
-our approach proceeds in two steps, 
-based on properties of the SW models. 
-The first step is a Bar-Hillel construction where, 
-given a \SWT $T$, a \SWVPA $A$, and an input word $s$, 
-a \SWVPA $A_{T, s}$ is built, such that for all $t$, $A_{T, s}(t) = T(s, t) \otimes A(t)$.
-In the second step, a best AST $t$ is found by applying to $A_{T, s}$ 
-a best search algorithm similar to shortest distance in graphs~\cite{Mohri02semiring,Huang05kbest}.
-%In expressiveness, they are equivalent to weighted CFG. 
-%and can be used in a general approach for parsing over infinite input alphabets.
-%
-%Let $A$ be a \SWVPA, associating $A(w) \in \Semiring$ 
-%to a given a nested word $w$ (representing a parse tree),
-%and let a \SWT compute a distance $d$, in $\Semiring$, 
-%between 2 strings over respectively an infinite input alphabet and the 
-%same (infinite) alphabet of $A$.
-%Then, the problem of Symbolic Weighted Parsing is, 
-%given an input string $s$, to find a nested word $w$ minimizing 
-%(according to the ranking defined by $\oplus$)
-%the distance $d(s, w) \otimes A(w)$ between $s$ and $A$, 
-%as defined in~\cite{Mohri03EDWA}.
-
-% First one general edit-distance is defined by a weighted word 
-% transducer~\cite{Mohri}
-% %Symbolic automata transducers are extended models~\cite{VeanesdAntoniJACM}
-% %dealing with infinite set of input symbols...
-% value in a semiring...
-
-The main contributions of the paper are: 
-($i$)~the introduction of automata, \SWA, transducers, \SWT (Section~\ref{section:SWA}),
-and visibly pushdown automata \SWVPA (Section~\ref{sec:SWVPA}),
-generalizing the corresponding classes of symbolic and weighted models, 
-%for the weighted computation on (nested) words over infinite alphabets;  
-($ii$)~a polynomial best-search algorithm for \SWVPA, %(Section~\ref{sec:best})
-% a framework for parsing over infinite alphabets, 
-and ($iii$)~a uniform framework (Section~\ref{sec:parsing}) for parsing over infinite alphabets, 
-the keys to which are 
-($iii.a$)~the \SWT-based definition of generic edit distances between input and output words,
-and ($iii.b$)~the use of nested words, and \SWVPA, 
-instead of syntax trees and grammars. %(\S~\ref{sec:trees}).
-%
-Finally, Section~\ref{sec:transcription} describes
-the implemented application %of this approach 
-to automated music transcription
-that motivated this study.
-
-%based these models and on generic def. of distances;
-%the use of VPA for parsing, word instead of trees, use comparison input-output.
-%and best-search
-
-%In Section~\ref{section:SWA} we introduce \SWA and \SWT.
-%Then \SWVPA are defined in Section~\ref{sec:SWVPA}, 
-%where a polynomial 1-best algorithm is described that can be use to solve
-%Symbolic Weighted Parsing, as described Section~\ref{sec:parsing}.
-%Finally, in Section~\ref{sec:transcription}, we present an application 
-%of this approach to automated music transcription that has been implemented.
+\input{introduction}
 
@@
 
-We follow the approach of~\cite{Mohri03EDWA} for the computation of distances
-between words and languages, with weighted transducers, 
-extending it to infinite alphabets.
+We follow the approach of~\cite{Mohri03EDWA} for the computation of distances,
+between words and languages, using weighted transducers, 
+and extend it to infinite alphabets.
 % with models of symbolic weighted automata and transducers. 
@@
 %over finite alphabets, see  e.g.~\cite{Mohri03EDWA}, 
-by labeling each transition with a weight function that takes the 
-input and output symbols as parameters, instead of a simple weight value.
+by labelling each transition with a weight function (instead of a simple weight value), 
+that takes the input and output symbols as parameters. 
 These functions are similar to the guards of symbolic automata~\cite{dAntoniVeanes17CAV,dAntoni21CACM},
-but they can return values in an arbitrary semiring, 
-where the latter guards are restricted to the Boolean semiring.
+but they can return values in an generic semiring, 
+whereas the latter guards are restricted to the Boolean semiring.
 
@@
 \label{section:semiring}
-We shall consider semiring for the weight values of our language models.
-%
-A \emph{semiring} $\< \Semiring, \oplus, \zero, \otimes, \one>$ 
-is a structure with a domain~$\Semiring$,
-equipped with two associative
-binary operators~$\oplus$ and $\otimes$,
-with respective neutral elements $\zero$ and $\one$, and such that:
-%$\< \mathbb{S}, \oplus, \zero>$ is a commutative monoid
-%$\< \mathbb{S}, \otimes, \one>$ is a monoid
-\begin{itemize}
-\item $\oplus$ is commutative;
- $\< \Semiring, \oplus, \zero>$ is a commutative monoid 
-   and $\< \Semiring, \otimes, \one>$ a monoid,
-\item $\otimes$ distributes over~$\oplus$:  $\forall x, y, z \in \mathbb{S}$,
-$x \otimes (y \oplus z) = (x \otimes y) \oplus (x \otimes z)$, 
-and $(x \oplus y) \otimes z = (x \otimes z) \oplus (y \otimes z)$,
-\item $\zero$ is absorbing for~$\otimes$: 
-$\forall x\in \mathbb{S}$, $\zero \otimes x = x \otimes \zero = \zero$.
-\end{itemize}
-%Components of a semiring~$\Semiring$ may be subscripted by~$\Semiring$ when needed.
-%We simply write $x \in \Semiring$ to mean $x \in \mathbb{S}$.
-%
-Intuitively, in the models presented in this paper, 
-$\oplus$ selects an optimal value from two given values, 
-in order to handle non-determinism, 
-and $\otimes$ combines two values into a single value, 
-in a chaining of transitions.
-%and let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a {semiring},
-
-\medskip%\noindent
-A semiring $\Semiring$ is \emph{commutative} if $\otimes$ is commutative.
-It is \emph{idempotent} if for each $x \in \dom(\Semiring)$, $x \oplus x = x$.
-%
-Every idempotent semiring~$\Semiring$ induces 
-a partial ordering~$\leq_\oplus$ 
-called the \emph{natural ordering} of~$\Semiring$~\cite{Mohri02semiring} 
-and defined,  by: 
-%implicitly defined by the semiring $\Semiring$ 
-for all $x$ and $y$,
-$x \leq_\oplus y \;\mbox{iff}\; x \oplus y = x$.
-%(see~\cite{Mohri02semiring} for the proof that it is an ordering).
-%
-The natural ordering is sometimes defined in the opposite direction~\cite{DrosteKuich09semirings};
-We follow here the direction  %follows \cite{Mohri02semiring}, and 
-that coincides with the usual ordering on the Tropical semiring \emph{min-plus} 
-(Figure~\ref{fig:semirings}).
-
-\begin{lemma}[Monotony, \cite{Mohri02semiring}] \label{lem:monotonic}
-Let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be an idempotent semiring.
-For all $x, y, z  \in \Semiring$,  
-if $x \leq_\oplus y$ then
-$x \oplus z \leq_\oplus y \oplus z$,
-$x \otimes z \leq_\oplus y \otimes z$
-and $z \otimes x \leq_\oplus z \otimes y$.
-\end{lemma}   
-When the property of Lemma~\ref{lem:monotonic} holds, 
-%We say in this case that 
-$\Semiring$ is called  \emph{monotonic}. % \wrt $\leq_\oplus$.
-%A semiring $\Semiring$ 
-%is \emph{monotonic} \wrt a partial ordering~$\leq$ 
-%iff for all $x, y, z  \in \Semiring$,  $x \leq y$ implies
-%$x \oplus z \leq y \oplus z$,
-%$x \otimes z \leq y \otimes z$,
-%and $z \otimes x \leq z \otimes y$.
-%
-Another important semiring property in the context of optimization
-is {superiority}~\cite{Huang08advanceddynamic}, 
-which corresponds to the 
-\emph{non-negative weights} condition in shortest-path algorithms~\cite{Dijkstra59anote}.
-Intuitively, it means that combining elements with $\otimes$ always increase their weight. 
-Formally, it is defined as the property ($i$) below. %of the following lemma.
-
-\begin{lemma}[Superiority, Boundedness]
-\label{lem:superior}\label{lem:bounded}
-Let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be an idempotent semiring.
-The two following statements are equivalent:
-\begin{itemize}
-\item [$i.$] for all $x, y \in \Semiring$,  
-$x \leq_\oplus x \otimes y$ and 
-$y \leq_\oplus x \otimes y$
-\item[$ii.$] for all $x \in \Semiring$,  $\one \oplus x = \one$.
-\end{itemize}
-\end{lemma}
-%
-\begin{proof} %(Lemma~\ref{lem:bounded})
-$(ii) \Rightarrow (i)$ : 
-$x \oplus (x \otimes y) = x \otimes (\one \oplus y) = x$, 
-by distributivity of~$\otimes$ over~$\oplus$. 
-Hence $x \leq_\oplus x \otimes y$.
-Similarly, $y \oplus (x \otimes y) = (\one \oplus x) \otimes y = y$, 
-hence $y \leq_\oplus x \otimes y$.
-%
-$(i) \Rightarrow (ii)$ :
-%$(i)$ implies that $\one \leq_\oplus x$ for all $x \in \Semiring$, 
-by the second inequality of ($i$), with $y = \one$, 
-$\one \leq_\oplus x \otimes \one = x$, \ie, 
-by definition of $\leq_\oplus$, $\one \oplus x = \one$.
-\qed
-\end{proof}
-
-In~\cite{Huang08advanceddynamic}, the property~$(i)$ 
-is called $\Semiring$ \emph{superior} \wrt the ordering~$\leq_\oplus$.
-We have seen in the proof of Lemma~\ref{lem:bounded} that it implies that 
-$\one \leq_\oplus x$ for all $x \in \Semiring$.
-Similarly, by the first inequality of ($i$) with $y = \zero$,  
-$x \leq_\oplus x \otimes \zero = \zero$.
-%
-Hence, in a superior semiring, % \wrt~$\leq$, 
-it holds that %$\one \leq \zero$
-for all $x \in \Semiring$, $\one \leq_\oplus x \leq_\oplus \zero$.
-%
-Intuitively, from an optimization point of view,
-it means that $\one$ is the best value, and $\zero$ the worst.
-%** superior implies $\Semiring$ bounded~\cite{Mohri02semiring} see **
-%
-In~\cite{Mohri02semiring}, the property ($ii$) of Lemma~\ref{lem:bounded}  
-is called \emph{boundedness} of $\Semiring$ -- we shall use this term in the rest of the paper. 
-% \emph{negative boundedness} 
-It implies that, when looking for a best path in a graph whose edges
-are weighted by values of $\Semiring$, the loops can be safely avoided.
-
-%Following the terminology of~\cite{Mohri02semiring},
-%when $\forall x \in \dom(\Semiring), \one \oplus x = \one$,
-%the semiring $\Semiring$ is is called \emph{bounded}.
-
-\begin{lemma}
-Every bounded semiring is idempotent.
-\end{lemma}
-\begin{proof}
-By boundedness, $\one \oplus \one = \one$, 
-and idempotency follows by multiplying
-both sides by $x$ and distributing. 
-\qed
-\end{proof}
-
-\noindent
-An idempotent semiring $\Semiring$~is called \emph{total} if
-it~$\leq_\oplus$ is total
-\ie when for all $x, y \in \Semiring$, either $x \oplus y = x$ or $x \oplus y = y$.
-
-%\medskip
-We shall need below infinite sums with~$\oplus$.
-A semiring~$\Semiring$ is called \emph{complete}~\cite{Droste09handbook} 
-%(\cite{Droste09handbook} chapter 1) %\cite{Kuich97semirings}
-if it has an %infinite sum 
-operation $\bigoplus_{i \in I} x_i$
-for every family
-$(x_i)_{i \in I}$ %$\{ x_i \mid i \in I \}$
-of elements of $\dom(\Semiring)$ over an index set $I \subset \mathbb{N}$, such that
-%is well-defined and in $\dom(\Semiring)$,
-the following holds: %properties
-\begin{description}
-\item[$i.$]
-\emph{infinite sums extend finite sums:}\\
-$\displaystyle\bigoplus_{i \in \emptyset} x_i = \zero$,\quad 
-      $\forall j\in \mathbb{N}, \displaystyle\bigoplus_{i \in \{ j \}} x_i = x_j$,
-      $\forall j, k\in \mathbb{N}, j\neq k, 
-      \displaystyle\bigoplus_{i \in \{ j, k \}} x_i = x_j \oplus x_k$,
-%
-\item[$ii.$]
-\emph{associativity and commutativity:}\\
-for all $I \subseteq \mathbb{N}$
-and all partition $(I_{j})_{j \in J}$ of $I$, %\subseteq \mathbb{N}$, 
-\(
-\displaystyle
-\bigoplus_{j \in J}\bigoplus_{i \in I_j} x_i = 
-\bigoplus_{i \in I} x_i
-\),
-%
-\item[$iii.$] 
-\emph{distributivity of product over infinite sum:}\\
-for all $I \subseteq \mathbb{N}$,
-\(
-\displaystyle
-\bigoplus_{i \in I} (x \otimes y_i) = x \otimes \bigoplus_{i\in I} y_i\), and
-\(
-\displaystyle
-\bigoplus_{i \in I} (x_i \otimes y) = (\bigoplus_{i \in I} x_i ) \otimes y\).
-\end{description}
-
-\begin{example}
-Figure~\ref{fig:semirings} presented examples of semirings interesting in practice 
-and enjoying the above properties.
-\end{example}
-
-\begin{figure}[t]
-\begin{center}
-\begin{tabular}{|c|c|C{2em}|C{2em}|C{2em}|C{2em}|}
-\hline
-        & domain & $\oplus$ & $\otimes$ & $\zero$  & $\one$\\ %& nat. ordering\\
-\hline\hline
-Boolean & $\{\bot, \top\}$ & $\vee$ & $\wedge$ & $\bot$ & $\top$\\ %& $\top \leq_\oplus \bot$  \\
-\hline
-Counting & $\mathbb{N}$ & $+$ & $\times$ & 0 & 1 \\
-\hline
-Viterbi & $[0, 1] \subset \mathbb{R}$ & $\mathit{max}$ & $\times$ & 0 & 1\\ % & $x \leq_\oplus y \iff x \ge y$  \\
-\hline
-Tropical min-plus & $\mathbb{R}_+ \cup \{ \infty\}$ & $\mathit{min}$ & $+$ & $\infty$ & $0$\\ % & $x \leq_\oplus y \iff x \leq y$   \\
-\hline
-%MaxPlus & $\mathbb{Q} \cup \{ -\infty\}$ & $\mathsf{max}$ & $+$ & $-\infty$ & $0$ \\
-%\hline
-%Word lang. & $2^{\Sigma^*}$ & $\cup$ & $\cdot$ & $\emptyset$ & $\{ \varepsilon \}$ \\
-%\hline
-\end{tabular}
-\end{center}
-\vskip-1em
-\caption{Some commutative, bounded, total and complete semirings.}
-\label{fig:semirings}
-\end{figure}
+\input{semiring}
 
@@
 \label{section:symbols}
-We shall now define the functions labeling the transitions of SW automata and transducers,
-generalizing the Boolean algebras of~\cite{dAntoniVeanes17CAV} 
-from Boolean to other semiring domains.
-%
-We consider \emph{alphabets}, which are countable sets of symbols 
-denoted $\Sigma$, $\Delta$,...
-%Let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative, complete semiring.
-%
-\noindent 
-Given a semiring $\< \Semiring, \oplus, \zero, \otimes, \one>$, 
-a $\Semiring$-\emph{label theory}
-is a tuplet $\bar\Phi$ of recursively enumerable sets of the form
-%$\Phi_\epsilon \subseteq \Semiring$, % containing constant functions valued in $\Semiring$, 
-$\Phi_\Sigma$, %and $\Phi_\Delta$, 
-containing unary functions of type $\Sigma \to \Semiring$, %resp. $\Delta \to \Semiring$, 
-and $\Phi_{\Sigma, \Delta}$, containing binary functions $\Sigma \times \Delta \to \Semiring$, 
-such that:
-
-\noindent -- 
-for all $\Phi_{\Sigma, \Delta} \in \bar\Phi$, there exists 
-$\Phi_{\Sigma} \in \bar\Phi$ and $\Phi_{\Delta} \in \bar\Phi$
-
-\noindent -- 
-for all $\alpha \in \Semiring$ and $\phi \in \Phi_\Sigma$,
-      $\alpha \otimes \phi : x \mapsto \alpha \otimes \phi(x)$, 
-      and $\phi \otimes \alpha : x \mapsto \phi(x) \otimes \alpha$\\
-\phantom{--} belong to $\Phi_\Sigma$, and similarly for $\oplus$ 
-      and for $\Phi_{\Sigma, \Delta}$
-
-\noindent -- 
-for all $\phi, \phi' \in \Phi_\Sigma$,
-$\phi \otimes \phi': x \mapsto \phi(x) \otimes \phi'(x)$ belongs to $\Phi_\Sigma$
-
-\noindent -- 
-for all $\eta, \eta' \in \Phi_{\Sigma, \Delta}$
-$\eta \otimes \eta': x, y \mapsto \eta(x, y) \otimes \eta'(x, y)$ belongs to $\Phi_{\Sigma, \Delta}$
-
-\noindent -- 
-for all $\phi \in \Phi_\Sigma$ and $\eta \in \Phi_{\Sigma, \Delta}$,
-$\phi \otimes \eta: x, y \mapsto \phi(x) \otimes \eta(x, y)$ belongs to $\Phi_{\Sigma, \Delta}$
-
-\noindent -- 
-for all $\psi \in \Phi_\Delta$ and $\eta \in \Phi_{\Sigma, \Delta}$,
-$\eta \otimes \psi: x, y \mapsto \eta(x, y) \otimes \psi(y)$ belongs to $\Phi_{\Sigma, \Delta}$
-
-\noindent -- 
-similar closures holds for $\oplus$
-
-\noindent -- 
-partial applications $\eta \in \Phi_{\Sigma, \Delta}$, 
-resp. $\eta_a: y \mapsto \eta(a, y)$ for $a \in \Sigma$ %and $y \in \Delta$
-and\\
-\phantom{--} $\eta_b: x \mapsto \eta(x, b)$ for $b \in \Delta$ %and $x \in \Sigma$, 
-belong resp. to~$\Phi_\Sigma$ and~$\Phi_\Delta$.
-
-%Moreover, these sets are required to be closed under the 
-%operators~$\oplus$ and~$\otimes$ of~$\Semiring$:
-%for all $\phi, \phi' \in \Phi_\Sigma$,
-%$\psi, \psi' \in \Phi_\Delta$, 
-%and $\eta, \eta' \in \Phi_{\Sigma, \Delta}$, %the function
-%%
-%\begin{center}
-%\begin{tabular}{cclll}
-%$\phi \otimes \phi'$ & : & $x \mapsto \phi(x) \otimes \phi'(x)$ & belongs to $\Phi_\Sigma$,\\
-%$\psi \otimes \psi'$ & : & $y \mapsto \psi(y) \otimes \psi'(y)$ & belongs to $\Phi_\Delta$,\\
-%$\phi \otimes \eta$\;  & : & $x, y \mapsto \phi(x) \otimes \eta(x, y)$ & belongs to $\Phi_{\Sigma, \Delta}$,\\
-%$\eta \otimes \psi$  & : & $x, y \mapsto \eta(x, y) \otimes \psi(y)$ & belongs to $\Phi_{\Sigma, \Delta}$,\\
-%$\eta \otimes \eta'$ & : & $x, y \mapsto \eta(x, y) \otimes \eta'(x, y)$ & belongs to $\Phi_{\Sigma, \Delta}$, &
-%\multicolumn{1}{r}{and similarly for $\oplus$.}\\ %the same also holds for the binary sum operator $\oplus$.
-%\end{tabular}
-%\end{center}
-%
-%Finally, it is also required 
-%% that the codomain of every function of $\Phi_\Sigma$ and $\Phi_\Delta$ 
-%% is a subset of $\Phi_\epsilon$, and
-%that the partial applications of a function $\eta \in \Phi_{\Sigma, \Delta}$, 
-%resp.  $\eta_a: y \mapsto f(a, y)$ for $a \in \Sigma$ and $y \in \Delta$
-%and  $\eta_b: x \mapsto f(x, b)$ for $b \in \Delta$ and $x \in \Sigma$, 
-%belong resp. to~$\Phi_\Sigma$ and~$\Phi_\Delta$.
-
-%For a complete semiring $\Semiring$, 
-%a function $\phi: \Sigma \to \Semiring$, 
-%resp. $\eta: \Sigma \times \Delta \to \Semiring$,
-%is called
+\input{labels}
 
@@
 \begin{array}{rccl}
+\weight_T(q, \epsilon, \epsilon, q') & = &              \wei(q, \epsilon, \epsilon, q')\\
 \weight_T(q, s, t, q') & = & 
@@
 \]
-
-The above sum may be infinite because of the $\epsilon$-transitions in the first and second lines.
-However, $\weight_T$ is well defined since $\Semiring$ is complete.
-%
+%
+We recall that, by convention, an empty sum with $\bigoplus$ is $\zero$. 
+%
+The above sum may be infinite because of the $\epsilon$-transitions, % in the first and second lines.
+in this case, the infinite sum operators applies since $\Semiring$ is assumed complete.
 As we shall see, under conditions, the $\epsilon$-transitions can be removed and hence the sum made finite.
@@
 %** The application of $\oplus$ can be seen as a ND choice between transitions **
-We recall that, by convention, an empty sum with $\oplus$ is $\zero$. 
-%
+
 Since $\zero$ is absorbing for~$\otimes$ in~$\Semiring$,
@@
 
+% DTW for sequences of timestamped events 
+% impossible, needs reading input/output symbols even by epsilon-transitions.
+
 \begin{example}
-DTW for sequences of timestamped events **     
-\end{example}
-
-\begin{example}
-(simpler) pointwise distance between two sequences of timestamped events **
+% simple pointwise distance between two sequences of timestamped events **
+Let us consider the tropical (\emph{min-plus}) semiring~$\Semiring$ 
+of Figure~\ref{fig:semirings} and 
+let $\Sigma = \mathbb{R}_+$ be an input alphabet of event dates
+and $\Delta = \{ \mathsf{e}, \mathsf{-} \} \times \mathbb{R}_+$ 
+be an output alphabet of symbols with timestamps. 
+A symbol $\< \mathsf{e}, d > \in \Delta$ represents an event starting at date $d$, 
+and $\< \mathsf{-}, d >$ is a continuation of the previous event.
+%More precisely, we let 
+%$\Sigma = \{ a, - \} \times \R_+$ and 
+This example of $\Delta$ is motivated by the case of music notation, 
+where several notated events (notes) can be tied together, 
+with a \emph{tie} or a \emph{dot}
+(like in $\musQuarter\!\!\mathrel{\raisebox{-1.5mm}{$\smile$}}\!\!\musEighth$ 
+ or equivalently~\musQuarterDotted),
+meaning that they will be played as a unique sounding event.
+%The timestamp of $a \in \Sigma$, denoted by $\mathsf{t}(a)$, is expressed as a rational number.
+
+We consider a \SWT with two states $q_0$ and $q_1$ whose purpose 
+is to compare a recorded performance $s \in \Sigma^*$
+with notated music sheet $t \in \Delta^*$.
+One timestamp $d_i \in \Sigma$ may corresponds 
+to one notated event $\<\mathsf{e}, d'_i> \in \Sigma$, in which case 
+the weight value computed by the \SWT is the time distance between both
+(see transitions $\wei_{11}$ below).
+%
+Moreover, if $\<\mathsf{e}, d'_i>$ is followed by continuations 
+$\<\mathsf{-}, d'_{i+1}>$..., they are just skip with no cost (transitions $\wei_{01}$ or weight $\one$).
+%transitions
+\[
+\begin{array}{rclcrcl}
+\wei_{11}(q_0, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d| & \quad &
+\wei_{11}(q_1, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d|\\
+\wei_{01}(q_0, \epsilon, \< \mathsf{-}, d'>, q_0) & = & \one & &
+\wei_{01}(q_1, \epsilon, \< \mathsf{-}, d'>, q_0) & = & \one\\
+\wei_{10}(q_0, d, \epsilon, q_1) & = & \alpha\\
+\end{array}
+\]
+%
+Finally, it may happen that the performers plays an extra note accidentally, but only once in a row. 
+This is modelled by the transition $\wei_{10}$ with an arbitrary weight value $\alpha \in \Semiring$, 
+switch from state $q_0$ (normal) to $q_1$ (error).
+The transitions in the second column below switch back to the normal state $q_0$.
+\endex
 \end{example}
@@
 %the automaton~$A$ is called \emph{without $\epsilon$-transitions}.
+
+The \emph{summary} of a $\SWT$, resp. a $\SWA$, is ***
 

### 2021-05-17 main2.tex
V2 with extended expressiveness

--- 2021-05-16 main-V0.tex
@@
-% old version
+% fork of main.tex with more expressive models
+% SWT read both tapes (under 2 heads) at every transition (even epsilon or semi-epsilon)
+% SWA read tape (under head) at every transition (even epsilon)
+% SW-VPA read top of stack (when not empty) at every transition
 
@@
 \label{section:symbols}
-We shall now define the functions labeling the transitions of SW automata and transducers,
-generalizing the Boolean algebras of~\cite{dAntoniVeanes17CAV} 
-from Boolean to other semiring domains.
-%
-We consider \emph{alphabets}, which are countable sets of symbols 
-denoted $\Sigma$, $\Delta$,...
-%Let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative, complete semiring.
-%
-\noindent 
-Given a semiring $\< \Semiring, \oplus, \zero, \otimes, \one>$, 
-a $\Semiring$-\emph{label theory}
-is a tuplet $\bar\Phi$ of recursively enumerable sets of the form
-%$\Phi_\epsilon \subseteq \Semiring$, % containing constant functions valued in $\Semiring$, 
-$\Phi_\Sigma$, %and $\Phi_\Delta$, 
-containing unary functions of type $\Sigma \to \Semiring$, %resp. $\Delta \to \Semiring$, 
-and $\Phi_{\Sigma, \Delta}$, containing binary functions $\Sigma \times \Delta \to \Semiring$, 
-such that:
-
-\noindent -- 
-for all $\Phi_{\Sigma, \Delta} \in \bar\Phi$, there exists 
-$\Phi_{\Sigma} \in \bar\Phi$ and $\Phi_{\Delta} \in \bar\Phi$
-
-\noindent -- 
-for all $\alpha \in \Semiring$ and $\phi \in \Phi_\Sigma$,
-      $\alpha \otimes \phi : x \mapsto \alpha \otimes \phi(x)$, 
-      and $\phi \otimes \alpha : x \mapsto \phi(x) \otimes \alpha$\\
-\phantom{--} belong to $\Phi_\Sigma$, and similarly for $\oplus$ 
-      and for $\Phi_{\Sigma, \Delta}$
-
-\noindent -- 
-for all $\phi, \phi' \in \Phi_\Sigma$,
-$\phi \otimes \phi': x \mapsto \phi(x) \otimes \phi'(x)$ belongs to $\Phi_\Sigma$
-
-\noindent -- 
-for all $\eta, \eta' \in \Phi_{\Sigma, \Delta}$
-$\eta \otimes \eta': x, y \mapsto \eta(x, y) \otimes \eta'(x, y)$ belongs to $\Phi_{\Sigma, \Delta}$
-
-\noindent -- 
-for all $\phi \in \Phi_\Sigma$ and $\eta \in \Phi_{\Sigma, \Delta}$,
-$\phi \otimes \eta: x, y \mapsto \phi(x) \otimes \eta(x, y)$ belongs to $\Phi_{\Sigma, \Delta}$
-
-\noindent -- 
-for all $\psi \in \Phi_\Delta$ and $\eta \in \Phi_{\Sigma, \Delta}$,
-$\eta \otimes \psi: x, y \mapsto \eta(x, y) \otimes \psi(y)$ belongs to $\Phi_{\Sigma, \Delta}$
-
-\noindent -- 
-similar closures holds for $\oplus$
-
-\noindent -- 
-partial applications $\eta \in \Phi_{\Sigma, \Delta}$, 
-resp. $\eta_a: y \mapsto \eta(a, y)$ for $a \in \Sigma$ %and $y \in \Delta$
-and\\
-\phantom{--} $\eta_b: x \mapsto \eta(x, b)$ for $b \in \Delta$ %and $x \in \Sigma$, 
-belong resp. to~$\Phi_\Sigma$ and~$\Phi_\Delta$.
-
-%Moreover, these sets are required to be closed under the 
-%operators~$\oplus$ and~$\otimes$ of~$\Semiring$:
-%for all $\phi, \phi' \in \Phi_\Sigma$,
-%$\psi, \psi' \in \Phi_\Delta$, 
-%and $\eta, \eta' \in \Phi_{\Sigma, \Delta}$, %the function
-%%
-%\begin{center}
-%\begin{tabular}{cclll}
-%$\phi \otimes \phi'$ & : & $x \mapsto \phi(x) \otimes \phi'(x)$ & belongs to $\Phi_\Sigma$,\\
-%$\psi \otimes \psi'$ & : & $y \mapsto \psi(y) \otimes \psi'(y)$ & belongs to $\Phi_\Delta$,\\
-%$\phi \otimes \eta$\;  & : & $x, y \mapsto \phi(x) \otimes \eta(x, y)$ & belongs to $\Phi_{\Sigma, \Delta}$,\\
-%$\eta \otimes \psi$  & : & $x, y \mapsto \eta(x, y) \otimes \psi(y)$ & belongs to $\Phi_{\Sigma, \Delta}$,\\
-%$\eta \otimes \eta'$ & : & $x, y \mapsto \eta(x, y) \otimes \eta'(x, y)$ & belongs to $\Phi_{\Sigma, \Delta}$, &
-%\multicolumn{1}{r}{and similarly for $\oplus$.}\\ %the same also holds for the binary sum operator $\oplus$.
-%\end{tabular}
-%\end{center}
-%
-%Finally, it is also required 
-%% that the codomain of every function of $\Phi_\Sigma$ and $\Phi_\Delta$ 
-%% is a subset of $\Phi_\epsilon$, and
-%that the partial applications of a function $\eta \in \Phi_{\Sigma, \Delta}$, 
-%resp.  $\eta_a: y \mapsto f(a, y)$ for $a \in \Sigma$ and $y \in \Delta$
-%and  $\eta_b: x \mapsto f(x, b)$ for $b \in \Delta$ and $x \in \Sigma$, 
-%belong resp. to~$\Phi_\Sigma$ and~$\Phi_\Delta$.
-
-\medskip\noindent
-When~$\Semiring$ is complete, 
-we call \emph{summary} of a function
-$\phi \in \Phi_\Sigma$,
-resp. $\eta \in \Phi_{\Sigma, \Delta}$,
-the value $\bigoplus_{a \in \Sigma} \phi(a)$, 
-resp. $\bigoplus_{a \in \Sigma} \bigoplus_{b \in \Delta} \eta(a, b)$.
-Note that, by definition of infinite sums in complete semirings, 
-a summary of $\phi \oplus \phi'$ and $\phi \otimes \phi'$ %(and $\alpha \otimes \phi$)
-can be computed from summaries of $\phi$ and $\phi'$ in $\Phi_{\Sigma}$, using the operators of $\Semiring$, 
-and the same holds for $\eta, \eta' \in \Phi_{\Sigma, \Delta}$. 
-%
-A label theory is called \emph{effective} when 
-summaries of $\phi \oplus \eta$, $\phi \otimes \eta$, $\eta \oplus \psi$, $\eta \otimes \psi$, 
-and of the partial applications $\eta_a$ and $\eta_b$, 
-can be computed, using the operators of $\Semiring$, 
-from summaries of 
-$\phi \in \Phi_{\Sigma}$, $\psi \in \Phi_{\Delta}$, $\eta \in \Phi_{\Sigma, \Delta}$.
+\input{labels}
 
@@
 Let $\Semiring$ be a commutative and complete semiring, 
-$\Sigma$ and $\Delta$ be alphabets called respectively \emph{input} and \emph{output}, %{alphabets}, 
  and $\bar\Phi = \< \Phi_\Sigma, \Phi_\Delta, \Phi_{\Sigma, \Delta}>$ be an $\Semiring$-label theory.
@@
 respectively leaving, a state, 
-and $\bar{\wei}$ is a tuplet made of the 4 of transition functions 
-$\wei_\epsilon$, $\wei_\Sigma$, $\wei_\Delta$, and $\wei_{\Sigma, \Delta}$
-from $Q \times Q$ into respectively~$\Semiring$, %$\Phi_\epsilon$, 
-$\Phi_\Sigma$, $\Phi_\Delta$, and $\Phi_{\Sigma, \Delta}$.
+and $\bar{\wei}$ is a 4-uplet of transition functions 
+$\wei_{00}$, $\wei_{10}$, $\wei_{01}$, and $\wei_{11}$
+%$\wei_{i, j}: Q \times Q \to \Phi_{\Sigma, \Delta}$, with $i, j \in \{ 0, 1\}$.
+from $Q \times Q$ into $\Phi_{\Sigma, \Delta}$.
 \end{definition}
 %
-A pair of states $\< q, q'>$ such that $\wei_\epsilon(q, q') \neq \zero$
-is called an \emph{$\epsilon$-transition}.
-
-\noindent We synthesize the 4-uplet of transition functions into a unique 
-function~$\wei:
- Q \times (\Sigma \cup \{ \epsilon \}) \times (\Delta \cup \{ \epsilon \}) \times Q 
- \to \Semiring$, %also called $\wei$ for simplicity, 
-such that, for all $q, q' \in Q$, $a \in \Sigma$,  $b \in \Delta$, 
-%and with 
-%$\< \phi_\epsilon, \phi_\Sigma, \phi_\Delta, \phi_{\Sigma, \Delta}> = \wei(q, q')$, 
-\[
-\begin{array}{rcll}
-\wei(q, \epsilon, \epsilon, q') & = &  \wei_\epsilon(q, q'),\\ %\phi_\epsilon\\
-\wei(q, a, \epsilon, q') & = & \phi(a) & %\wei_\Sigma(q, q')(a)
-\quad\mathrm{where~} \phi = \wei_\Sigma(q, q') \in \Phi_\Sigma,\\
-\wei(q, \epsilon, b, q') & = & \psi(b) & 
-\quad\mathrm{where~} \psi = \wei_\Delta(q, q') \in \Phi_\Delta,\\
-\wei(q, a, b, q') & = & \eta(a, b) & 
-\quad\mathrm{where~} \eta = \wei_{\Sigma, \Delta}(q, q') \in \Phi_{\Sigma, \Delta}.\\
-\end{array}      
-\]
-%More precisely, $Q \times Q$, 
-%resp. $Q \times \Sigma \times Q$,
-%$Q \times \Delta \times Q$,
-%$Q \times \Sigma \times \Delta \times Q$,
-%into resp. $\Phi_0$, $\Phi_\Sigma$, $\Phi_\Delta$ $\Phi_{\Sigma, \Delta}$.
-
-\noindent
-The symbolic-weighted transducer~$T$ defines a mapping 
-from the pairs of strings of $\Sigma^* \times \Delta^*$ 
-into~$\Semiring$,
+%A pair of states $\< q, q'>$ such that $\wei_\epsilon(q, q') \neq \zero$
+%is called an \emph{$\epsilon$-transition}.
+
+The \SWT $T$ computes on pairs of words $\< s, t> \in \Sigma^* \times \Delta^*$,
+$s$, resp. $t$, being called input word, resp. output word.
+%
+For convenience, when $\wei_{ij}(q, q') = \eta \in \Phi_{\Sigma, \Delta}$, 
+for $q, q' \in Q$ and $i, j \in \{ 0, 1\}$,
+we shall sometimes write, overloading the function names:
+$\wei_{ij}(q, a, b, q') = \eta(a, b)$ when  $a \in \Sigma$,  $b \in \Delta$.
+%
+Intuitively, using a transition $\wei_{ij}(q, a, b, q')$ means for $T$:
+read $a$, resp. $b$, at the current position in the input word, resp. output word;
+if $i = 1$, resp. $j = 1$,
+increment the current position in the input word (otherwise, do not change it);
+change state from $q$ to $q'$.
+
+\noindent
+More precisely, the \SWT~$T$ defines a mapping 
+from the pairs of words of $\Sigma^* \times \Delta^*$ into~$\Semiring$,
 based on the following intermediate function $\weight_T$
@@
 %$a \in \Sigma$, $b \in \Delta$ 
-for every strings of $s \in \Sigma^*$, $t \in \Delta^*$:
+%for every strings of $s \in \Sigma^*$, $t \in \Delta^*$:
+$a \in \Sigma$, $u \in \Sigma^*$, $b\in \Delta$, $v\in \Delta^*$, 
 \[
-\begin{array}{rccl}
-\weight_T(q, s, t, q') & = & 
+\begin{array}{lccl}
+\weight_T(q, \epsilon, \epsilon, q)  & = & \multicolumn{2}{l}{\final(q)}\\
+\weight_T(q, \epsilon, \epsilon, q') & = & \multicolumn{2}{l}{\zero \mathrm{~if~} q \neq q'}\\
+\weight_T(q, au, bv, q') & = & 
    \displaystyle\bigoplus_{q'' \in Q} &
-   \wei(q, \epsilon, \epsilon, q'') \otimes \weight_T(q'', s, t, q')\\
+   \wei_{00}(q, a, b, q'') \otimes \weight_T(q'', au, bv, q')\\
   & \oplus & \displaystyle\bigoplus_{q'' \in Q} &
-              \wei(q'', \epsilon, \epsilon, q') \otimes \weight_T(q, s, t, q'')\\
- & \oplus & \displaystyle\bigoplus_{\begin{array}{c}
-                                      \scriptstyle q'' \in Q\\[-2pt]
-                                      \scriptstyle s = au, a \in \Sigma
-                                      \end{array}} &
-   \wei(q, a, \epsilon, q'') \otimes \weight_T(q'', u, t, q')\\
- & \oplus & \displaystyle\bigoplus_{\begin{array}{c}
-                                    \scriptstyle q'' \in Q\\[-2pt]
-                                    \scriptstyle t = bv, b \in \Delta
-                                    \end{array}} &
-    \wei(q, \epsilon, b, q'') \otimes \weight_T(q'', s, v, q')\\ 
- & \oplus & \displaystyle\bigoplus_{\begin{array}{c}
-                                    \scriptstyle q'' \in Q\\[-2pt]
-                                    \scriptstyle s = au, a \in \Sigma\\[-2pt]
-                                    \scriptstyle t = bv, b \in \Delta
-                                    \end{array}} &
-   \wei(q, a, b, q'') \otimes \weight_T(q'', u, v, q').\\ 
+    \wei_{10}(q'', a, b, q') \otimes \weight_T(q, u, bv, q'')\\
+  & \oplus & \displaystyle\bigoplus_{q'' \in Q} &
+    \wei_{01}(q'', a, b, q') \otimes \weight_T(q, au, v, q'')\\
+  & \oplus & \displaystyle\bigoplus_{q'' \in Q} &
+    \wei_{11}(q'', a, b, q') \otimes \weight_T(q, u, v, q'')\\
 \end{array}
@@
 
-The above sum may be infinite because of the $\epsilon$-transitions in the first and second lines.
+The above sum may be infinite because of the third line.
 However, $\weight_T$ is well defined since $\Semiring$ is complete.
@@
 As we shall see, under conditions, the $\epsilon$-transitions can be removed and hence the sum made finite.
-%
+** SUPPR. $\epsilon$-transition (with $\wei_{00}$) first. **
+
+*** rem. there is no case for $a, \epsilon$ or $\epsilon, b$.
+   It can be simulated with $\wei_{10}$ and $\wei_{01}$ on the last letter. ***
+
 %** The application of $\oplus$ can be seen as a ND choice between transitions **
@@
 Since $\zero$ is absorbing for~$\otimes$ in~$\Semiring$,
-one term $\wei(q, a, b, q'')$ equal to $\zero$ in the above expression 
+one term $\wei_{ij}(q, a, b, q'')$ equal to $\zero$ in the above expression 
 will be ignored in the sum, meaning that there is no possible transition
@@
 Intuitively, 
-$\wei(q, a, \epsilon, q'')$, in the third line, is the cost of 
+$\wei_{10}(q, a, b, q'')$, in the fourth line, is the cost of 
 the deletion of the symbol~$a \in \Sigma$ in~$s$, 
-$\wei(q, \epsilon, b, q'')$, in the fourth line, is the cost 
+$\wei{01}(q, a, b, q'')$, in the fifth line, is the cost 
 of the insertion of~$b \in \Delta$ in $t$, 
-and $\wei(q, a, b, q'')$, in the last line, is the cost 
+and $\wei_{11}(q, a, b, q'')$, in the last line, is the cost 
 of the substitution of  $a \in \Sigma$ by~$b \in \Delta$.
+%
 The cost of a sequence of such operations transforming $s$ into $t$, 
@@
 
-A $\SWT$ or $\SWA$ is called \emph{summarized}
-when
+The \emph{summary} of a $\SWT$, resp. a $\SWA$, is ***
 

### 2021-05-17 main.tex
cleanup

@@
 
-The \emph{summary} of a $\SWT$, resp. a $\SWA$, is ***
+%The \emph{summary} of a $\SWT$, resp. a $\SWA$, is ***
 
@@
 %
-\begin{proof}  ** Tb revised **   
-Let $A = \< Q, \init, \<\wei_\epsilon, \wei_\Sigma>, \final >$.
-The automaton $A'$ is $\< Q, \init, \<\wei'_\epsilon, \wei'_\Sigma>, \final >$,
-where, for all $q, q' \in Q$,
-$\wei'_\epsilon(q, q') = \zero$  
-and 
-% \[
-% \begin{array}{rcccl}
-% \wei'_\Sigma(q, q') & = &  & \displaystyle\bigoplus_{q_1 \in Q} &
-%                \wei_\Sigma(q, q_1) \otimes  \wei_\epsilon(q_1, q')\\
-%                     &   & \oplus  & \displaystyle\bigoplus_{q_2 \in Q} &
-%                \wei_\epsilon(q, q_2) \otimes \wei_\Sigma(q_2, q')\\
-% \end{array}
-% \]
-\[
-\wei'_\Sigma(q, q') = 
-\displaystyle\bigoplus_{q'' \in Q} \wei_\epsilon(q, q'') \otimes \wei_\Sigma(q'', q')
-\oplus \displaystyle\bigoplus_{q_1 \in Q} \wei_\Sigma(q, q'') \otimes \wei_\epsilon(q'', q')
-\]
-In the above definition of $\wei'_\Sigma$ we use the operator
-of product of function of $\Phi_\Sigma$ by $\Semiring$ 
-described in Section~\ref{section:symbols}.
-%
-By definition of $\weight_{A}$ and distributivity of $\otimes$ on $\oplus$, 
-** NO. TBC see \cite{Mohri02ijfcs} **
-it holds that $\weight_{A}(q, s, q') = \weight_{A'}(q, s, q')$.
-\qed
-\end{proof}
+%\begin{proof}  ** Tb revised **   
+%Let $A = \< Q, \init, \<\wei_\epsilon, \wei_\Sigma>, \final >$.
+%The automaton $A'$ is $\< Q, \init, \<\wei'_\epsilon, \wei'_\Sigma>, \final >$,
+%where, for all $q, q' \in Q$,
+%$\wei'_\epsilon(q, q') = \zero$  
+%and 
+%% \[
+%% \begin{array}{rcccl}
+%% \wei'_\Sigma(q, q') & = &  & \displaystyle\bigoplus_{q_1 \in Q} &
+%%                \wei_\Sigma(q, q_1) \otimes  \wei_\epsilon(q_1, q')\\
+%%                     &   & \oplus  & \displaystyle\bigoplus_{q_2 \in Q} &
+%%                \wei_\epsilon(q, q_2) \otimes \wei_\Sigma(q_2, q')\\
+%% \end{array}
+%% \]
+%\[
+%\wei'_\Sigma(q, q') = 
+%\displaystyle\bigoplus_{q'' \in Q} \wei_\epsilon(q, q'') \otimes \wei_\Sigma(q'', q')
+%\oplus \displaystyle\bigoplus_{q_1 \in Q} \wei_\Sigma(q, q'') \otimes \wei_\epsilon(q'', q')
+%\]
+%In the above definition of $\wei'_\Sigma$ we use the operator
+%of product of function of $\Phi_\Sigma$ by $\Semiring$ 
+%described in Section~\ref{section:symbols}.
+%%
+%By definition of $\weight_{A}$ and distributivity of $\otimes$ on $\oplus$, 
+%** NO. TBC see \cite{Mohri02ijfcs} **
+%it holds that $\weight_{A}(q, s, q') = \weight_{A'}(q, s, q')$.
+%\qed
+%\end{proof}
 
@@
 
-\begin{example}
-structured words...
-intro language of music notation ?
-\end{example}
+%\begin{example}
+%structured words...
+%intro language of music notation ?
+%\end{example}
 
@@
 \label{sec:best}\label{sec:search}
-**hypotheses**
+%**hypotheses**
 Let us assume that the semiring~$\Semiring$ is
 commutative, bounded, complete, and total.
-and 
-** assume label theory is s.t. 
-for all $\phi \in \Phii$, 
-$\psi \in \Phir$, 
-and $\eta \in \Phicr$, 
-$\displaystyle\bigoplus_{a \in \Omegai} \phi(a)$ 
-$\displaystyle\bigoplus_{r \in \Omegar} \phi(r)$ and 
-$\displaystyle\bigoplus_{{\call{c}} \in \Omegac} 
-\displaystyle\bigoplus_{{\return{r}} \in \Omegar} \eta({\call{c}}, {\return{r}})$
-are computable...
+and assume an effective label theory.
+%s.t. for all $\phi \in \Phii$, 
+%$\psi \in \Phir$, 
+%and $\eta \in \Phicr$, 
+%$\displaystyle\bigoplus_{a \in \Omegai} \phi(a)$ 
+%$\displaystyle\bigoplus_{r \in \Omegar} \phi(r)$ and 
+%$\displaystyle\bigoplus_{{\call{c}} \in \Omegac} 
+%\displaystyle\bigoplus_{{\return{r}} \in \Omegar} \eta({\call{c}}, {\return{r}})$
+%are computable...
 
@@
 This gives a time complexity $O(|Q|^4 . |P|)$. 
-It can be reduced by implementing $P$ as a priority queue, 
-prioritized by the value returned by $d$
-***complete***. %$|Q|^3.\log(|Q|^2)$
+%
+%It can be reduced by implementing $P$ as a priority queue, 
+%prioritized by the value returned by $d$
+%***complete***. %$|Q|^3.\log(|Q|^2)$
 
@@
 %
-\begin{proof}
-Bar-Hillel construction
-\qed  
-\end{proof}
-
-\subsection{Application to Automated Music Transcription}
-\label{sec:transcription}
-Symbolic Automated Music Transcription
-and analysis of music performances
-
-\subsubsection{Time Scales}
-Real-Time Unit (RTU) = seconds
-
-\noindent 
-Musical-Time Unit (MTU) = number of measures
-
-\noindent 
-conversion via tempo value
-
-\subsubsection{Representation of Music Performances}
-We consider symbolic representations of musical performances, as finite sequences of events.
-It corresponds to the concrete case of a MIDI file~\cite{SMF} 
-recorded  from an electronic keyboard, 
-or the output of a transcription from audio files~\cite{Benetos18AMTsurvey}.
-%
-For the sake of simplicity, 
-we shall only consider here the case of monophonic performances, 
-where at most one note is sounding at a time. 
-The approach however extends to the polyphonic case.
-
-A music performance is a finite sequence of events in a set~$\Sigma$.
-Every event $e \in \Sigma$ has attributes such from a finite domain, 
-like a number of key for a note 
-or a flag indicating that it is a rest 
-(\textsf{ON} and \textsf{OFF} messages in~\cite{SMF})
-and a velocity value (0..127 in~\cite{SMF})).
-%This representation is similar to the piano roll ~\cite{Muller15fundamentals} chap.1. 
-Moreover, it contains a RTU value $\ioi{e}$ (real number) 
-representing the time distance to the next event, 
-or to the end of performance for the last event,
-also called \emph{inter-onset interval}.
-
-\subsubsection{Representation of Music Scores}
-Music score are represented as structured words
-made of timed %quantified 
-events and parenthesized markups,
-akin of nested words~\cite{AlurMadhusudan09nested}.
-
-We consider an alphabet $\Delta$, every symbol of which is 
-composed of a tag, in a finite set $\Xi$, 
-and an MTU (rational) IOI duration value.
-%The alphabet $\Delta$ 
-It is partitioned into 
-$\Delta = \Deltai \uplus \Deltac \uplus \Deltar$, 
-like in Section~\ref{section:SWVPA}.
-%
-\noindent
-The symbols of $\Deltai$ represent events:
-% (infinite alphabet of internal symbols) made of:
-with tags indicating a new note or grace-note (with null IOI), 
-a rest or the continuation of the previous note (tie or dot).
-%
-The elements of $\Deltac \uplus \Deltar$ are matched
-markups for describing the structure of the score, 
-\ie the hierarchical grouping of events, and also, 
-importantly the division of time in measures, tuplets...
-%- parentheses for time divisions : tuplets, bars...
-(linearization of rhythm trees \cite{jacquemard:hal-01138642}...).
-They contain additional info such as tuple number, beaming policy...
-
-\noindent
-The duration values of letters of $\Delta$, in MTU (rational), 
-can be computed with the markups and tags (\eg grace note has duration 0).
-
-%\noindent
-%There are simultaneous events, since grace notes has duration 0. They are ordered.
-%
-%\noindent
-%Finite bound on the number of duration ratio. ?
-
-\begin{example}
-...      
-\end{example}
-
-\subsubsection{Performance/Score Distance Computation}
-\label{app:distance}
-We define a distance between performance and score representations
-by a $\SWT$ $T = \< Q, \init, \wei, \final >$, over a semiring $\Semiring$.
-** detail the elements of $\Semiring$ ....**
-%are quadruplets of the form
-%$\< t, s, \delta_t, \delta_s>$
-
-Every state of $Q$ contains a 
-tempo value in a finite domain (e.g. 30..300 bpm).
-This value can be fixed 
-or recomputed by the $T$ %transducer 
-after reading each event, 
-according to a perceptive/cognitive model of tempo 
-such as~\cite{LargeJones99tempo}
-(also used in the context of score following~\cite{Cont10TPAMI}).
-% we wont detail here.
-
-\subsection{Transcription by SW Parsing} %Best-first Search}
-We assume a score language defined by a \SWVPA over the semiring 
-$\Semiring$ of Section~\ref{app:distance}.
+%\begin{proof}
+%Bar-Hillel construction
+%\qed  
+%\end{proof}
 
@@
 
+\end{document}
+
 \section{Edit-Distance}
@@
 %\paragraph{Distance between a word and a regular language}
-
-\end{document}

### 2021-05-17 transcription.tex
cleanup

--- 2021-05-17 main2.tex
@@
-% fork of main.tex with more expressive models
-% SWT read both tapes (under 2 heads) at every transition (even epsilon or semi-epsilon)
-% SWA read tape (under head) at every transition (even epsilon)
-% SW-VPA read top of stack (when not empty) at every transition
-
-\documentclass[runningheads]{llncs}
-%\documentclass[a4paper,11pt]{article}
-%\setcounter{page}{1}
-
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-
-\usepackage{hyperref}
-%\usepackage[bookmarks,bookmarksnumbered,naturalnames,plainpages=false]{hyperref}
-%usepackage{url}
-
-% for footnote ref
-\usepackage{refcount}
-
-% array and tabular
-\usepackage{array}
-\newcolumntype{C}[1]{>{\centering\arraybackslash\hspace{0pt}}p{#1}}
-
-% extension of enumerate env. (style for displaying counters)
-% \usepackage{enumerate}
-
-%% pictures
-% \usepackage{graphicx} 
-% \DeclareGraphicsExtensions{.pdf,.png,.jpg}
-% \graphicspath{fig/}
-
-%% PGF, Tikz
-\usepackage{tikz-cd}
-%% \usepackage{pgfplots}
-%% \usepgfplotslibrary{dateplot}
-%% \usepackage{pgf,pgfarrows,pgfnodes, pgfautomata}
-% \usepackage{tikz}
-% \usetikzlibrary{cd}
-%% \usetikzlibrary{arrows}
-%% \usetikzlibrary{calc}
-%% \usetikzlibrary{snakes}
-%% \usetikzlibrary{backgrounds}
-% \usetikzlibrary{trees}
-%% \usetikzlibrary{automata}
-%% \usetikzlibrary{positioning}
-%% \usetikzlibrary{matrix}
-%% \usetikzlibrary{patterns}
-%% \usetikzlibrary{shapes}
-
-% symbols
-\usepackage{amsmath} 
-\usepackage{amssymb} 
-\usepackage{amsbsy}
-\usepackage{bbold}
-\usepackage{latexsym}
-%\usepackage{amsfonts}
-\usepackage{stmaryrd}
-%\usepackage{mathabx}
-%\usepackage{MnSymbol}
-\usepackage{harmony} % simple music fonts
-\usepackage{mathtools} % for arrows
-%\usepackage{mathptmx}
-
-%% theorem environments
-\usepackage{theorem}
-% \newtheorem{theorem}{Theorem} %[section]
-% \newtheorem{definition}[theorem]{Definition}
-% \newtheorem{lemma}[theorem]{Lemma}
-% \newtheorem{corollary}[theorem]{Corollary}
-% \newtheorem{proposition}[theorem]{Proposition}
-% \newenvironment{proof}{\vspace{-2ex}{\it Proof. }}{\hspace*{\fill} $\Box$\smallskip }
-% \theorembodyfont{\slshape}
-% \newtheorem{example}[theorem]{Example}
-% \newtheorem{remark}[theorem]{Remark}
-{\theorembodyfont{\rmfamily} \theoremstyle{break} \newtheorem{algo}{Algorithm}}
-
-%% algorithms
-%\usepackage{algorithm} 
-%\usepackage{program}
-
-%% for new macros
-\usepackage{xspace}
-
-%% arrows etc
-%\input{rewriting}
-
-%% Misc macros
-
-\def\ie{\textit{i.e.}\xspace}
-\def\eg{\textit{e.g.}\xspace}
-\def\wrt{\textit{wrt}\xspace}
-%\def\wlog{\textit{wlog}\xspace}
-\def\etc{\textit{etc}\xspace}
-
-\def\<#1>{\langle #1 \rangle}
-\newcommand{\pair}[2]{\langle{#1}, {#2}\rangle}
-%\newcommand{\A}{\mathcal{A}}
-%\newcommand{\B}{\mathcal{B}}
-\newcommand{\D}{\mathbb{D}}
-\newcommand{\E}{\mathbb{E}}
-\newcommand{\W}{\mathbb{W}}
-\newcommand{\T}{\mathcal{T}}
-
-\newcommand{\Semiring}{\mathbb{S}}
-\newcommand{\zero}{\mathbb{0}}
-\newcommand{\one}{\mathbb{1}}
-\newcommand{\dom}{\ensuremath{\mathit{dom}}}
-
-\def\SA{\textsf{SA}\xspace}
-\def\WA{\textsf{WA}\xspace}
-\def\SWT{\textsf{swT}\xspace}
-\def\SWA{\textsf{swA}\xspace}
-\def\SWTA{\textsf{swTA}\xspace}
-\def\SWVPA{\textsf{sw-VPA}\xspace}
-\def\weight{\mathsf{weight}}
-\def\wei{\mathsf{w}}
-\def\mei{\mathsf{m}}
-\def\init{\mathsf{in}}
-\def\final{\mathsf{out}}
-\newcommand{\call}[1]{\ensuremath #1} %{\ensuremath \langle_{#1}}
-\newcommand{\return}[1]{\ensuremath #1} %{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
-\def\Omegai{{\Omega_\mathsf{i}}}
-\def\Omegac{{\Omega_\mathsf{c}}}
-\def\Omegar{{\Omega_\mathsf{r}}}
-\def\Sigmai{{\Sigma_\mathsf{i}}}
-\def\Sigmac{{\Sigma_\mathsf{c}}}
-\def\Sigmar{{\Sigma_\mathsf{r}}}
-\def\Deltai{{\Delta_\mathsf{i}}}
-\def\Deltac{{\Delta_\mathsf{c}}}
-\def\Deltar{{\Delta_\mathsf{r}}}
-\def\Phii{{\Phi_\mathsf{i}}}
-\def\Phic{{\Phi_\mathsf{c}}}
-\def\Phir{{\Phi_\mathsf{r}}}
-\def\Phicr{{\Phi_\mathsf{cr}}}
-\def\weii{{\wei_\mathsf{i}}}
-\def\weic{{\wei_\mathsf{c}}}
-\def\weir{{\wei_\mathsf{r}}}
-\def\weie{{\wei_\mathsf{e}}}
-\newcommand{\config}[2]{\ensuremath \genfrac{[}{]}{0pt}{}{#1}{#2}} 
-\newcommand{\opluseq}{\ensuremath\mathrel{\oplus}=}
-\newcommand{\ioi}[1]{\mathsf{ioi}({#1})}
-\newcommand{\rank}{\mathsf{rk}}
-\newcommand{\lin}{\mathsf{lin}}
-
-%\sloppy
-
-% Parsing over infinite alphabet as optimal alignment computation 
-% as edit-distance between string and language
-% 
-%\title{Symbolic Weighted Parsing and Automated Music Transcription}
-%\title{Symbolic Weighted Language Models and Automated Music Transcription}
-\title{Symbolic Weighted Language Models and Parsing over Infinite Alphabets}
-%\title{Weighted Visibly Pushdown Automata and Automated Music Transcription}
-\author{Florent Jacquemard}
-\institute{INRIA \& CNAM, Paris, France\\
-\email{florent.jacquemard@inria.fr}}
-
-%\titlerunning{WVPA \& AMT}
-\titlerunning{Symbolic Weighted Parsing and Automated Music Transcription}
-%\authorrunning{Florent Jacquemard}
-
-\date{\today}
-
-\begin{document}
-\thispagestyle{empty}
-\maketitle
-
-\begin{abstract}
-\input{abstract}
-\end{abstract}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% intro
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Introduction} \label{sec:intro}
-\input{introduction}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SWT & SWA
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{SW Automata and Transducers}
-\label{section:transducer}\label{sec:transducer}
-\label{section:SWA}\label{sec:SWA}
-\label{section:SWT}\label{sec:SWT}
-
-We follow the approach of~\cite{Mohri03EDWA} for the computation of distances
-between words and languages, with weighted transducers, 
-extending it to infinite alphabets.
-% with models of symbolic weighted automata and transducers. 
-%
-The models introduced in this section generalize 
-weighted automata and transducers~\cite{Droste09handbook} 
-%over finite alphabets, see  e.g.~\cite{Mohri03EDWA}, 
-by labeling each transition with a weight function that takes the 
-input and output symbols as parameters, instead of a simple weight value.
-These functions are similar to the guards of symbolic automata~\cite{dAntoniVeanes17CAV,dAntoni21CACM},
-but they can return values in an arbitrary semiring, 
-where the latter guards are restricted to the Boolean semiring.
-
-\subsection{Semirings} 
-\label{section:semiring}
-\input{semiring}
-
-\subsection{Label Theory}
-\label{section:symbols}
-\input{labels}
-
-\subsection{Definitions} \label{sec:SWTdef}\label{sec:SWAdef}
-%(SWT)
-
-Let $\Semiring$ be a commutative and complete semiring, 
- and $\bar\Phi = \< \Phi_\Sigma, \Phi_\Delta, \Phi_{\Sigma, \Delta}>$ be an $\Semiring$-label theory.
-
-\begin{definition}
-\label{def:transducer} \label{def:SWT}
-A \emph{symbolic-weighted transducer} (\SWT)
-over $\Sigma$, $\Delta$, $\Semiring$ and $\bar\Phi$
-%the input and output alphabets~$\Sigma$ and $\Delta$ with label theory $\bar\Phi$, and the semiring $\Semiring$ 
-is a tuple
-$T = \< Q, \init, \bar{\wei}, \final >$,
-where $Q$ is a finite set of states, 
-$\mathsf{in} : Q \to \Semiring$, 
-respectively $\mathsf{out} : Q \to \Semiring,$
-are functions defining the weight for entering, 
-respectively leaving, a state, 
-and $\bar{\wei}$ is a 4-uplet of transition functions 
-$\wei_{00}$, $\wei_{10}$, $\wei_{01}$, and $\wei_{11}$
-%$\wei_{i, j}: Q \times Q \to \Phi_{\Sigma, \Delta}$, with $i, j \in \{ 0, 1\}$.
-from $Q \times Q$ into $\Phi_{\Sigma, \Delta}$.
-\end{definition}
-%
-%A pair of states $\< q, q'>$ such that $\wei_\epsilon(q, q') \neq \zero$
-%is called an \emph{$\epsilon$-transition}.
-
-The \SWT $T$ computes on pairs of words $\< s, t> \in \Sigma^* \times \Delta^*$,
-$s$, resp. $t$, being called input word, resp. output word.
-%
-For convenience, when $\wei_{ij}(q, q') = \eta \in \Phi_{\Sigma, \Delta}$, 
-for $q, q' \in Q$ and $i, j \in \{ 0, 1\}$,
-we shall sometimes write, overloading the function names:
-$\wei_{ij}(q, a, b, q') = \eta(a, b)$ when  $a \in \Sigma$,  $b \in \Delta$.
-%
-Intuitively, using a transition $\wei_{ij}(q, a, b, q')$ means for $T$:
-read $a$, resp. $b$, at the current position in the input word, resp. output word;
-if $i = 1$, resp. $j = 1$,
-increment the current position in the input word (otherwise, do not change it);
-change state from $q$ to $q'$.
-
-\noindent
-More precisely, the \SWT~$T$ defines a mapping 
-from the pairs of words of $\Sigma^* \times \Delta^*$ into~$\Semiring$,
-based on the following intermediate function $\weight_T$
-defined recursively for every $q, q' \in Q$, 
-%$a \in \Sigma$, $b \in \Delta$ 
-%for every strings of $s \in \Sigma^*$, $t \in \Delta^*$:
-$a \in \Sigma$, $u \in \Sigma^*$, $b\in \Delta$, $v\in \Delta^*$, 
-\[
-\begin{array}{lccl}
-\weight_T(q, \epsilon, \epsilon, q)  & = & \multicolumn{2}{l}{\final(q)}\\
-\weight_T(q, \epsilon, \epsilon, q') & = & \multicolumn{2}{l}{\zero \mathrm{~if~} q \neq q'}\\
-\weight_T(q, au, bv, q') & = & 
-   \displaystyle\bigoplus_{q'' \in Q} &
-   \wei_{00}(q, a, b, q'') \otimes \weight_T(q'', au, bv, q')\\
-  & \oplus & \displaystyle\bigoplus_{q'' \in Q} &
-    \wei_{10}(q'', a, b, q') \otimes \weight_T(q, u, bv, q'')\\
-  & \oplus & \displaystyle\bigoplus_{q'' \in Q} &
-    \wei_{01}(q'', a, b, q') \otimes \weight_T(q, au, v, q'')\\
-  & \oplus & \displaystyle\bigoplus_{q'' \in Q} &
-    \wei_{11}(q'', a, b, q') \otimes \weight_T(q, u, v, q'')\\
-\end{array}
-\]
-
-The above sum may be infinite because of the third line.
-However, $\weight_T$ is well defined since $\Semiring$ is complete.
-%
-As we shall see, under conditions, the $\epsilon$-transitions can be removed and hence the sum made finite.
-** SUPPR. $\epsilon$-transition (with $\wei_{00}$) first. **
-
-*** rem. there is no case for $a, \epsilon$ or $\epsilon, b$.
-   It can be simulated with $\wei_{10}$ and $\wei_{01}$ on the last letter. ***
-
-%** The application of $\oplus$ can be seen as a ND choice between transitions **
-We recall that, by convention, an empty sum with $\oplus$ is $\zero$. 
-%
-Since $\zero$ is absorbing for~$\otimes$ in~$\Semiring$,
-one term $\wei_{ij}(q, a, b, q'')$ equal to $\zero$ in the above expression 
-will be ignored in the sum, meaning that there is no possible transition
-from state $q$ into state $q'$ while reading $a$ and writing $b$ 
-(including the case $a = \epsilon$ and $b = \epsilon$).
-%These functions $\phi$ act as guards for the transducer's transitions, 
-%preventing a transition when they return the absorbing $\zero$ of $\Semiring$.
-This is analogous to a transition's guard not satisfied by $\<a, b>$ in 
-the case of symbolic transducers.
-
-The expression of $\weight_T$ can be seen as a stateful definition of 
-an edit-distance between a word $s \in \Sigma^*$ and a word $t \in \Delta^*$,
-see also~\cite{Mohri03ijfcs}.
-Intuitively, 
-$\wei_{10}(q, a, b, q'')$, in the fourth line, is the cost of 
-the deletion of the symbol~$a \in \Sigma$ in~$s$, 
-$\wei{01}(q, a, b, q'')$, in the fifth line, is the cost 
-of the insertion of~$b \in \Delta$ in $t$, 
-and $\wei_{11}(q, a, b, q'')$, in the last line, is the cost 
-of the substitution of  $a \in \Sigma$ by~$b \in \Delta$.
-%
-The cost of a sequence of such operations transforming $s$ into $t$, 
-is the product, with $\otimes$, of the individual costs of the operations involved;
-And the distance between $s$ and $t$ is the sum, with $\oplus$,
-of all such product of costs.
-
-\medskip\noindent
-The weight associated by $T$ to  $\< s, t> \in \Sigma^* \times \Delta^*$
-is defined as follows: 
-\begin{equation}
-T(s, t)  = 
-\displaystyle\bigoplus_{q, q' \in Q} \mathsf{in}(q) 
-\mathop{\otimes} \weight_T(q, s, t, q') \mathop{\otimes} \mathsf{out}(q').
-\label{eq:weightT}
-\end{equation}
-
-\begin{example}
-Let $\Sigma$ be an alphabet of symbols with timestamps. 
-The timestamp of $a \in \Sigma$, denoted by $\mathsf{t}(a)$, is expressed as a rational number.
-
-DTW for sequences of timestamped events **     
-\end{example}
-
-\begin{example}
-(simpler) pointwise distance between two sequences of timestamped events **
-\end{example}
-
-\noindent
-The \emph{Symbolic Weighted Automata} %$A = \< Q, \init, \weight, \final >$
-%over $\Sigma$,  $\Semiring$ and $\bar\Phi$
-are defined similarly as the transducers of Definition~\ref{def:SWT}, 
-by simply omitting the output symbols.
-%
-In this case, the label theory $\bar\Phi$ is reduced to a singleton $\< \Phi_\Sigma>$.
-%over $\Sigma$ is reduced to
-%a set $\Phi_\Sigma$ closed under~$\oplus$ and~$\otimes$.
-%
-\begin{definition} \label{def:SWA}
-A \emph{symbolic-weighted automaton} (\SWA)
-over $\Sigma$, $\Semiring$ and $\bar\Phi$
-%the input alphabet~$\Sigma$ and the commutative semiring $\Semiring$ 
-is a tuple
-$A = \< Q, \init, \bar{\wei}, \final >$,
-where $Q$ is a finite set of states, 
-$\mathsf{in} : Q \to \Semiring$, 
-respectively $\mathsf{out} : Q \to \Semiring,$
-are functions defining the weight for entering, 
-respectively leaving, a state, 
-and $\bar{\wei}$ is a pair of transition functions 
-$\wei_\epsilon$ and $\wei_\Sigma$ from $Q \times Q$ into 
-respectively $\Semiring$ %$\Phi_\epsilon$ 
-and $\Phi_\Sigma$.
-\end{definition}
-
-\noindent
-The above transition functions $\wei_\epsilon$ and $\wei_\Sigma$
-can be synthesized as above into a function
-$\wei: Q \times (\Sigma \cup \{ \epsilon \}) \times Q \to \Semiring$.
-%
-%When $\wei_\epsilon(q, q') = \zero$ for all $q, q' \in Q$, 
-%the automaton~$A$ is called \emph{without $\epsilon$-transitions}.
-
-The \emph{summary} of a $\SWT$, resp. a $\SWA$, is ***
-
-\subsection{Properties}
-The two following properties will be useful to our approach on 
-symbolic weighted parsing in Section~\ref{sec:parsing}.
-
-\begin{proposition}
-Given a \SWT $T$ 
-over alphabet $\Sigma$, $\Delta$ and a commutative semiring~$\Semiring$, 
-and $s \in \Sigma^*$, 
-there exists an effectively constructible \SWA 
-$A_{T, s}$ over $\Delta$ and $\Semiring$ 
-such that for all $t \in \Delta^*$, $A_{T, s}(t) = T(s, t)$.
-\end{proposition}
-%
-\begin{proof}
-Let $T = \< Q, \init, \bar{\wei}, \final >$,
-let 
-$\wei: Q \times (\Sigma \cup \{ \epsilon \}) \times (\Delta \cup \{ \epsilon \}) \times Q 
- \to \Semiring$
-be the synthesized form of 
-$\bar{\wei} = \< \wei_\epsilon, \wei_\Sigma, \wei_\Delta, \wei_{\Sigma, \Delta}>$,
-and let $s = s_1 \ldots s_n$ with $s_1, \ldots, s_n \in \Sigma$.
-%$n = |s|$ be the length of $s$.
-%
-The state set of $A_{T, s}$ will be $Q' = [1..n] \times Q$.
-Its state entering weight function is defined by 
-$\init'(1, q) = \init(q)$ for all $q \in Q$, 
-and $\init'(i, q) = \zero$ for all $1 < i \leq n$, 
-and its state leaving weight function by 
-$\final'(n, q) = \final(q)$ or all $q \in Q$, 
-and $\final'(i, q) = \zero$ for all $1 \leq i <  n$.
-
-\noindent
-We define transition functions $\wei'_\epsilon$ and $\wei'_\Sigma$ for $A_{T, s}$, 
-together with their synthesized form $\wei'$ as follows: 
-for all $1 \leq i, j \leq n$, $q, q' \in Q$, %and $a \in \Sigma$, 
-and $b \in \Delta$, 
-\[
-\begin{array}{lclcrcll}
-\wei'_\epsilon\bigl(\< i, q>, \< i, q'>\bigr) & = & \wei_\epsilon(q, q') & \mathrm{\ie} &
-\wei'\bigl(\< i, q>, \epsilon, \< i, q'>\bigr)   & = & \wei(q, \epsilon, \epsilon, q'), & \\ 
-%
-\wei'_\epsilon\bigl(\< i, q>, \< i+1, q'>\bigr) & = & \phi(s_i) & 
-\multicolumn{5}{l}{\mathrm{where~} \phi = \wei_\Sigma(q, q'),}\\
- & & & \mathrm{\ie} &
-\wei'\bigl(\< i, q>, \epsilon, \< i+1, q'>\bigr) & = & \wei(q, s_i, \epsilon, q') & 
-\mathrm{if~} i < n,\\ 
-%
-\wei'_\epsilon\bigl(\< i, q>, \< j, q'>\bigr) & = & \zero & 
-%\mathrm{\ie} & \wei'(\< i, q>, \epsilon, \< j, q'>)   & = & \zero & & 
-\multicolumn{5}{l}{\mathrm{if~} j \neq i, i+1,}\\ 
-%
-\wei'_\Sigma\bigl(\< i, q>, \< i, q'>\bigr) & = & \wei_\Delta(q, q') & \mathrm{\ie} &
-\wei'\bigl(\< i, q>, b, \< i, q'>\bigr)          & = & \wei(q, \epsilon, b, q'),\\ 
-%
-\wei'_\Sigma\bigl(\< i, q>, \< i+1, q'>\bigr) & = & \eta_{s_i} &
-\multicolumn{5}{l}{\mathrm{where~} \eta = \wei_{\Sigma, \Delta}(q, q'),}\\
- & & & \mathrm{\ie} & \wei'\bigl(\< i, q>, b, \< i+1, q'>\bigr) & = & \wei(q, s_i, b, q') & 
- \mathrm{if~} i < n,\\ 
-%
-\wei'_\Sigma\bigl(\< i, q>, \< j, q'>\bigr) & = & \zero & \mathrm{\ie} &
-\wei'\bigl(\< i, q>, b, \< j, q'>\bigr)     & = & \multicolumn{2}{l}{\zero\ \mathrm{~if~} j \neq i, i+1.}
-\end{array}      
-\]
-We use the assumption that the underlying label theory of $T$ is 
-is closed under partial applications (Section~\ref{section:symbols})
-in the above construction of $\wei'_\Sigma$.
-
-\noindent
-The $\SWA$ wanted is $A_{T, s} = \< Q', \init', {\bar\wei}', \final' >$.
-\qed
-\end{proof}
-
-The construction time and size for $A_{T, s}$ are $O(\| T \| . | s |)$,
-where the size $\| T \|$ of $T$ is its number of states $|Q|$.
-
-% gen. to transducers
-\begin{proposition}
-Given a \SWA $A$ over alphabet $\Sigma$ 
-and a commutative and bounded semiring $\Semiring$, 
-there exists an effectively constructible \SWA $A'$ 
-without $\epsilon$-transitions 
-such that for all $s \in \Sigma^*$, $A'(s) = A(s)$.
-\end{proposition}
-%
-\begin{proof}  ** Tb revised **   
-Let $A = \< Q, \init, \<\wei_\epsilon, \wei_\Sigma>, \final >$.
-The automaton $A'$ is $\< Q, \init, \<\wei'_\epsilon, \wei'_\Sigma>, \final >$,
-where, for all $q, q' \in Q$,
-$\wei'_\epsilon(q, q') = \zero$  
-and 
-% \[
-% \begin{array}{rcccl}
-% \wei'_\Sigma(q, q') & = &  & \displaystyle\bigoplus_{q_1 \in Q} &
-%                \wei_\Sigma(q, q_1) \otimes  \wei_\epsilon(q_1, q')\\
-%                     &   & \oplus  & \displaystyle\bigoplus_{q_2 \in Q} &
-%                \wei_\epsilon(q, q_2) \otimes \wei_\Sigma(q_2, q')\\
-% \end{array}
-% \]
-\[
-\wei'_\Sigma(q, q') = 
-\displaystyle\bigoplus_{q'' \in Q} \wei_\epsilon(q, q'') \otimes \wei_\Sigma(q'', q')
-\oplus \displaystyle\bigoplus_{q_1 \in Q} \wei_\Sigma(q, q'') \otimes \wei_\epsilon(q'', q')
-\]
-In the above definition of $\wei'_\Sigma$ we use the operator
-of product of function of $\Phi_\Sigma$ by $\Semiring$ 
-described in Section~\ref{section:symbols}.
-%
-By definition of $\weight_{A}$ and distributivity of $\otimes$ on $\oplus$, 
-** NO. TBC see \cite{Mohri02ijfcs} **
-it holds that $\weight_{A}(q, s, q') = \weight_{A'}(q, s, q')$.
-\qed
-\end{proof}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SW-VPA
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{SW Visibly Pushdown Automata}
-\label{section:SWVPA}\label{sec:SWVPA}
-The following model generalizes Symbolic VPA~\cite{dAntonyAlur14SVPDA}
-from Boolean semirings to arbitrary semiring weight domains.
-
-\subsection{Definition}
-Let $\Omega$ be a countable alphabet 
-%finite (large) or infinite,
-that we assume partitioned into three subsets 
-$\Omegai$, $\Omegac$, $\Omegac$,
-whose elements are respectively called 
-\emph{internal}, \emph{call} and \emph{return} symbols.
-% \begin{itemize}
-% \item a set $\Omegai$ of \emph{internal symbols} denoted $a$,
-% \item a set $\Omegac$ of \emph{call symbols} denoted $\call{a}$,
-% \item a set $\Omegar$ of \emph{return symbols} denoted $\return{a}$.
-% \end{itemize}
-Let~$\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative semiring and  
-let  $(\Phi_\epsilon, \Phic, \Phir, \Phicr)$ be a label theory over $\Semiring$
-%In order to simplify notations, %and following the definition of Section~\ref{section:transducer}, 
-%we shall write respectively 
-where $\Phic$, $\Phir$ and~$\Phicr$ stand respectively 
-for~$\Phi_\Omegac$, $\Phi_\Omegar$ and~$\Phi_{\Omegac, \Omegar}$.
-%
-Moreover, we extend this theory with a set $\Phii$ 
-of unary functions in $\Omegai \to \Semiring$,
-closed under $\oplus$ and $\otimes$.
-
-\begin{definition}
-A \emph{Symbolic Weighted Visibly Pushdown Automata} (\SWVPA) 
-over  $\Omega = \Omegai \uplus \Omegac \uplus \Omegar$ and $\Semiring$ is a tuple
-$A = \< Q, P, \init, \weii, \weic, \weir, \weie, \final >$,
-where $Q$ is a finite set of states, 
-$P$ is a finite set of stack symbols, 
-$\mathsf{in} : Q \to \Semiring$, 
-respectively $\mathsf{out} : Q \to \Semiring,$
-are functions defining the weight for entering, 
-respectively leaving, a state, 
-and 
-$\weii : Q \times Q \to \Phii$,  
-$\weic : Q \times Q \times P \to \Phic$,  
-$\weir : Q \times P \times Q \to \Phicr$,  
-$\weie : Q \times Q \to \Phir$,  
-are transition functions.
-\end{definition}
-%
-Similarly as in Section~\ref{section:transducer}, 
-we extend the above transition functions as follows
-for all $q, q' \in Q$, $p \in P$, 
-$a \in \Omegai$, 
-$\call{c} \in \Omegac$, 
-$\return{r} \in \Omegar$, 
-overloading their names: % for simplicity:
-\[
-\begin{array}{lll}
-\weii: Q \times \Omegai \times Q \to \Semiring & 
-\weii(q, a, q') = \phi_\mathsf{i}(a) & 
-\mathrm{where~} \phi_\mathsf{i} = \weii(q, q'),\\
-%
-\weic: Q \times \Omegac \times Q \times P \to \Semiring & 
-\weic(q, \call{c}, q', p) = \phi_\mathsf{c}(\call{c}) & 
-\mathrm{where~} \phi_\mathsf{c} = \weic(q, q', p),\\
-%
-\weir: Q \times \Omegac \times P \times \Omegar \times Q \to \Semiring & 
-\weir(q, {\call{c}},  p, {\return{r}}, q') = \phi_\mathsf{r}({\call{c}},  {\return{r}}) & 
-\mathrm{where~} \phi_\mathsf{r} = \weir(q, p, q'),\\
-%
-\weie: Q \times \Omegar \times Q \to \Semiring & 
-\weie(q, {\return{r}}, q') = \phi_\mathsf{e}({\return{r}}) &
-\mathrm{where~} \phi_\mathsf{e} = \weie(q, q').\\
-\end{array}      
-\]
-
-\noindent
-The intuition is the following for the above transitions.
-\begin{description}
-\item $\weii$ : read the input internal symbol $a$, change state to $q'$ (stack is untouched).
-\item $\weic$ : read the input call symbol $\call{c}$, push it to the stack along with $p$, change state to $q'$.
-\item $\weir$ : when the stack is not empty, 
-      read and pop from stack a pair made of $\call{c}$ and $p$, 
-      read the input return symbol $\return{r}$, change state to $q'$.
-      In this case, the weight function $\phi_\mathsf{r}$ 
-      computes a value of matching between the call and return symbols $c$ and $r$.
-      This value might be set to $\zero$ in order to express that the symbols do not match.
-\item $\weie$ : when the stack is empty, 
-      read the input symbol $\call{r}$, change state to $q'$.
-\end{description}
-
-We give now a formal definition of these transitions of the automaton~$A$
-in term of %a weight value computed by 
-an intermediate function $\weight_A$, like in Section~\ref{sec:SWT}.
-In the case of a pushdown automaton, a configuration is composed 
-of a state $q \in Q$ and a stack content $\gamma \in \Gamma^*$, where $\Gamma = \Omegac \times P$.
-Therefore, $\weight_A$ is a function from 
-$Q \times \Gamma^* \times \Omega^* \times Q \times \Gamma^*$ into $\Semiring$.
-\[
-\begin{array}{rcl}
-\weight_A\bigl(\config{q}{\gamma}, a\, u, \config{q'}{\gamma'}\bigr) & = & 
- {\displaystyle\bigoplus_{q'' \in Q}} \weii(q, a, q'') 
- \otimes \weight_A\bigl(\config{q''}{\gamma}, u, \config{q'}{\gamma'}\bigr)\\
-%
-\weight_A\bigl(\config{q}{\gamma}, {\call{c}}\, u, \config{q'}{\gamma'}\bigr) & = & 
- {\displaystyle\bigoplus_{\begin{array}{c}
-                          \scriptstyle q'' \in Q\\[-2pt]
-                          \scriptstyle p \in P
-                          \end{array}}}
- \weic\bigl(q, {\call{c}}, q'', p\bigr) 
- \otimes \weight_A\bigl(\config{q''}{{\call{c}}\, p\cdot \gamma}, u, \config{q'}{\gamma'}\bigr)\\[1mm]
-%
-\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot \gamma}, {\return{r}}\, u, \config{q'}{\gamma'}\bigr) & = & 
- {\displaystyle\bigoplus_{q'' \in Q}} 
- \weir\bigl(q, {\call{c}}, p, {\return{r}}, q''\bigr) 
- \otimes \weight_A\bigl(\config{q''}{\gamma}, u, \config{q'}{\gamma'}\bigr)\\
-%
-\weight_A\bigl(\config{q}{\bot}, {\return{r}}\, u, \config{q'}{\gamma'}\bigr) & = & 
- {\displaystyle\bigoplus_{q'' \in Q}} \weie(q, {\return{r}}, q'') 
- \otimes \weight_A\bigl(\config{q''}{\bot}, u, \config{q'}{\gamma'}\bigr)\\
-\end{array}
-\]
-where $\bot$ denotes the empty stack and ${\call{c}}\, p\cdot \gamma$ 
-denotes a stack where the pair made of ${\call{c}} \in \Omegac$ and $p \in P$ is the top symbol 
-and $\gamma$ is the rest of stack.
-
-\noindent
-The weight associated by $A$ to $s \in \Omega^*$
-is then defined as follows,
-following empty stack semantics: 
-\begin{equation}
-A(s)  = 
-{\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
-\mathsf{in}(q) \mathop{\otimes} 
-\weight_A\bigl(\config{q}{\bot}, s, \config{q'}{\bot}\bigr) 
-\mathop{\otimes} \mathsf{out}(q').
-\label{eq:weightA}
-\end{equation}
-
-\begin{example}
-structured words...
-intro language of music notation ?
-\end{example}
-
-\subsection{Properties}
-Like VPA and symbolic VPA, 
-the class of \SWVPA is closed under the binary operators of the underlying semiring.
-%
-\begin{proposition}
-Let $A_1$ and $A_2$ be two (\SWVPA)
-over the same $\Omega$ and $\Semiring$.
-There exists two $\SWVPA$ $A_1 \oplus A_2$ and $A_1 \otimes A_2$, 
-effectively constructible, 
-such that for all $s \in \Omega^*$, 
-$(A_1 \oplus A_2)(s) = A_1(s) \oplus A_2(s)$ and 
-$(A_1 \otimes A_2)(s) = A_1(s) \otimes A_2(s)$.
-\end{proposition}
-The construction is essentially the same 
-as in the case of the Boolean semiring~\cite{dAntonyAlur14SVPDA}.
-
-\subsection{Best-first Search} 
-\label{sec:best}\label{sec:search}
-**hypotheses**
-Let us assume that the semiring~$\Semiring$ is
-commutative, bounded, complete, and total.
-and 
-** assume label theory is s.t. 
-for all $\phi \in \Phii$, 
-$\psi \in \Phir$, 
-and $\eta \in \Phicr$, 
-$\displaystyle\bigoplus_{a \in \Omegai} \phi(a)$ 
-$\displaystyle\bigoplus_{r \in \Omegar} \phi(r)$ and 
-$\displaystyle\bigoplus_{{\call{c}} \in \Omegac} 
-\displaystyle\bigoplus_{{\return{r}} \in \Omegar} \eta({\call{c}}, {\return{r}})$
-are computable...
-
-% total ? 
-% monotonic and superior writ natural ordering
-%Regarding the infinite sum operator, note that
-%$\bigoplus_{x \in \Phi_\Omega} \phi(x)$, 
-%$\bigoplus_{y \in \Phi_\Delta} \psi(y)$, and  
-%... exist and in $\Semiring$.
-
-We propose a Dijkstra algorithm computing the minimal weight by $A$, 
-\wrt~$\leq_\oplus$, for a word in~$\Omega^*$.
-
-\noindent
-More precisely, 
-let $\top$ be a fresh stack symbol which does not belong to $\Gamma$,
-and for every two states $q, q' \in Q$ and $\sigma \in \{ \bot, \top \}$, 
-let 
-\[ 
-  d_0(q, \sigma, q') = \bigoplus_{s\in \Omega^*} 
-  \textstyle
-  \weight_A\bigl(\config{q}{\sigma}, s, \config{q'}{\sigma}\bigr). 
-\]
-Since $\Semiring$ is complete, this infinite sum is well defined, 
-and since $\leq_\oplus$ is assumed total, it is the minimum in $\Omega^*$
-of $s \mapsto \weight_A(\config{q}{\sigma}, s, \config{q'}{\sigma})$ \wrt this ordering.
-%
-When $\sigma = \bot$, $d_0(q, \sigma, q')$ is the central 
-expression in a term of the definition (\ref{eq:weightA}) of $A(s_0)$ for the minimum $s_0$
-(for the above function).
-When $\sigma = \bot$, intuitively, it is the minimum weight of a computation of $A$
-starting in state $q$ with a stack $\gamma \in \Gamma^*$ (possibly empty), 
-and ending in state $q$ with the same stack $\gamma$, such that moreover 
-the computation does not touch a symbol of $\gamma$. 
-That means that during the computation, 
-$A$ may apply the first case of in the definition of $\weight_A$ (internal symbol),
-as well as the second case, to can push call symbols on the top of $\gamma$, 
-and may pop these symbols with the third case (return symbol). 
-However, it cannot apply one of the two last cases (return symbol and empty stack)
-when the current stack is $\gamma$.
-%pop symbols in $\gamma$.
-% Note that having a stack reduced to such a symbol makes impossible the application of the 
-% two last cases in the definition of $\weight_A$ (return symbol and empty stack). 
-% However, it is possible to apply the two first cases 
-% (internal symbol or call symbol, with a push on the top of $\top$).
-
-The algorithm~\ref{algo:Dijkstra}
-constructs iteratively a marking $d : Q \times \{ \bot, \top \} \times Q \to \Semiring$
-%of the triplets $\<q, \sigma, q'>$ 
-%of states of $A$ by weight values in $\Semiring$, 
-that converges eventually to $d_0(q, \sigma, q')$. 
-%It uses for that purpose a priority queue $P$ containing triplets of 
-%$Q \times \{ \bot, \top \} \times Q$.
-
-\begin{algo}[1-best for \SWVPA] \label{algo:Dijkstra}
-\textbf{initially} let $P = Q \times \{ \bot, \top \} \times Q$, %contains all %triplets 
-%$\< q_1, \bot, q_2>$ and $\<q_1, \top, q_2>$ for $q_1, q_2 \in Q$, with 
-and 
-$d(q_1, \bot, q_2) = d(q_1, \top, q_2) = \one$ if $q_1 = q_2$ and
-$d(q_1, \bot, q_2) = d(q_1, \top, q_2) = \zero$ otherwise.
-
-\smallskip\noindent
-\textbf{while} $P$ is not empty
-
-\noindent
-\quad\textbf{extract} $\< q_1, \sigma, q_2>$ from $P$ 
- such that $d(q_1, \sigma, q_2)$ is minimal wrt $\leq_\oplus$.
-
-\noindent\quad \textbf{for all} $q_0, q_3 \in Q$ and $p \in P$ do
-
-\noindent
-\qquad
-\(\begin{array}{rrcl}
- & d(q_1, \sigma, q_3) & \opluseq &
-   d(q_1, \sigma, q_2) \otimes \displaystyle\bigoplus_{a \in \Omegai} \weii(q_2, a, q_3)\\
-%
-%  & d(q_0, \sigma, q_2) & \opluseq & 
-%    \displaystyle\bigoplus_{a \in \Omegai} \weii(q_0, a, q_1) \otimes d(q_1, \sigma, q_2)\\
-%
-\mathrm{if~} \sigma = \top & d(q_0, \top, q_3) & \opluseq &
-d(q_1, \sigma, q_2) \otimes
-%\displaystyle\bigoplus_{p \in P} 
-\displaystyle\bigoplus_{{\call{c}} \in \Omegac} 
-\displaystyle\bigoplus_{{\return{r}} \in \Omegar} \eta({\call{c}}, {\return{r}})\\
-\mathrm{and} &  d(q_0, \bot, q_3) & \opluseq &
-d(q_1, \sigma, q_2) \otimes
-%\displaystyle\bigoplus_{p \in P} 
-\displaystyle\bigoplus_{{\call{c}} \in \Omegac} 
-\displaystyle\bigoplus_{{\return{r}} \in \Omegar} \eta({\call{c}}, {\return{r}})\\
-& & & \mathrm{where~} \eta = \weic(q_0, q_1, p) \otimes \weir(q_2, p, q_3)\\
-%
-\mathrm{if~} \sigma = \bot & d(q_1, \bot, q_3) & \opluseq &
-  d(q_1, \sigma, q_2) \otimes \displaystyle\bigoplus_{\return{r} \in \Omegar} \weie(q_2, {\return{r}}, q_3)\\
-%
-%& d(q_0, \bot, q_2) & \opluseq & 
-%d(q_0, \bot, q_1) \otimes d(q_1, \sigma, q_2), \mathrm{if~} \< q_0, \bot, q_1> \notin P \\
-%
-%\mathrm{if~} \sigma = \top & d(q_0, \top, q_2) & \opluseq & 
-%d(q_0, \top, q_1) \otimes d(q_1, \sigma, q_2), \mathrm{if~} \< q_0, \top, q_1> \notin P \\
-%
-& d(q_1, \bot, q_3) & \opluseq & 
-   d(q_1, \sigma, q_2) \otimes d(q_2, \bot, q_3), \mathrm{if~} \< q_2, \bot, q_3> \notin P \\
-%
-\mathrm{if~} \sigma = \top & d(q_1, \top, q_3) & \opluseq & 
-    d(q_1, \sigma, q_2) \otimes d(q_2, \top, q_3), \mathrm{if~} \< q_2, \top, q_3> \notin P \\
-\end{array}
-\) 
-\end{algo}
-
-\noindent
-The infinite sums in the updates of $d$ in Algorithm~\ref{algo:Dijkstra} are well defined
-since~$\Semiring$ is complete.
-The algorithm performs $2.|Q|^2$ iterations until $P$ is empty, 
-and each iteration has a time complexity $O(|Q|^2 . |P|)$.
-This gives a time complexity $O(|Q|^4 . |P|)$. 
-It can be reduced by implementing $P$ as a priority queue, 
-prioritized by the value returned by $d$
-***complete***. %$|Q|^3.\log(|Q|^2)$
-
-The correctness of Algorithm~\ref{algo:Dijkstra} 
-is ensured by the invariant expressed in the following lemma.
-\begin{lemma}
-For all $\< q_1, \sigma, q_2> \notin P$, $d(q_1, \sigma, q_2) =  d_0(q_1, \sigma, q_2)$.
-\end{lemma}
-The proof is by contradiction, 
-assuming a counter-example minimal in the length of the witness word.
-
-% \begin{algorithm}
-% \caption{<your caption for this algorithm>}
-% \label{<your label for references later in your document>}
-% \begin{program}
-% \seq{|line| := `` '', i := 1};
-% \WHILE i \neq n+1 \DO
-%       |line| := |item|[i] \concat `` '' \concat |number|[i]; 
-%       i := i+1; 
-%       \WHILE i \neq n+1 \AND |item|[i] = |item|[i-1] \DO 
-%       |line| := |line| \concat ``, '' \concat |number|[i]);
-%       i := i+1 \OD ; 
-%       |write|(|line|) \OD 
-% \end{program}
-% \end{algorithm}
-
-% The weight of a transition acts as a guard: 
-% a transition is activated for a symbol $`a`$ iff its weight $`\phi(a) \neq 0`$ (or $`\psi(a, b) \neq 0`$), 
-% the absorbing element.
-
-%\begin{lemma}
-\noindent
-For computing the minimal weight of a computation of $A$, we use the fact that,
-at the termination of Algorithm~\ref{algo:Dijkstra}, %it holds that,
-%There exist $q_1, q_2 \in Q$
-\[
-  {\displaystyle \bigoplus_{s \in \Omega^*} A(s)} = 
-  {\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
-  \mathsf{in}(q) \mathop{\otimes} d(q, \bot, q') \mathop{\otimes} \mathsf{out}(q').
-\]
-%\end{lemma}
-
-%\medskip
-\noindent
-In order to obtain effectively a witness 
-(word of $\Semiring^*$ with computation of $A$ of minimal weight), 
-we require an additional property the of weight functions.
-%
-\begin{definition}
-Let $\Omega$ be an alphabet and $\Semiring$ a complete semiring.
-A function $\phi$ from $\Omega^n$ into $\Semiring$
-is called $k$-\emph{convex} for a natural number $k$ iff 
-$\mathit{card}\{ \mathbf{a} \in \Omega^n \mid \phi(\mathbf{a}) = 
-                 \displaystyle\bigoplus_{\mathbf{p} \in \Omega^n} \phi(\mathbf{p}) \} \leq k$.
-\end{definition}
-A label theory is $k$-convex if all its functions are $k$-convex.
-
-\begin{proposition}
-For a \SWVPA $A$ over a 
-commutative, idempotent, superior, total and complete semiring 
-and an alphabet $\Omega$ with a $k$-convex label theory,
-one can construct in PTIME a word $s \in \Omega^*$ 
-such that $A(s)$ is minimal \wrt the natural ordering for $\Semiring$. 
-%$\leg_\oplus$.
-\end{proposition}
-
-\subsection{Trees and Nested-Words}
-\label{sec:trees}
-The hierarchical structure of nested-words, defined the \emph{call} and \emph{return} markup symbols of 
-suggest a correspondence between these word and trees. 
-The lifting of this correspondence to languages, respective of tree automata and VPA have
-been discussed in~\cite{AlurMadhusudan09nested}, 
-see also~\cite{Caralp12VPAmult} for the weighted case.
-In this section, we discuss the correspondence between the symbolic-weighted extensions
-of tree automata and VPA.
-
-Let $\Omega$ be a countable ranked alphabet, such that 
-every symbol $a \in \Omega$ has a rank 
-$\rank(a) \in [0..M]$ where $M$ is a fixed natural number.
-We write $\Omega_k$ the subset of all symbols $a$ of $\Omega$
-with $\rank(a) = k$, where $0 \leq k \leq M$.
-%$\cup_{i = 0}^{i = M} \Omega_i$ 
-%
-$\T(\Omega)$ denotes the free $\Omega$-algebra of finite, ordered, 
-$\Omega$-labeled trees, 
-which is the smallest set such that  $\Omega_0 \subset \T(\Omega)$
-and for all $1 \leq k \leq M$, all $a \in \Omega_k$, 
-and all $t_1, \ldots, t_k \in \T(\Omega)$, $a(t_1, \ldots, t_k) \in \T(\Omega)$.
-%
-% tree = single node (leave) labeled with a symbol of $a \in \Omegai$
-% (such a tree is simply denoted by $a$)
-% or the composition, denoted by $b(t_1,\ldots, t_n$) of a node labeled with $b$
-% and $n$ subtrees $t_1$,\ldots, $t_n$.
-
-\renewcommand{\call}[1]{\ensuremath \langle_{#1}}
-\renewcommand{\return}[1]{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
-
-Let $\hat\Omega$ be the countable (unranked) alphabet obtained from 
-$\Omega$ as follows: 
-$\hat\Omega = \< \Omegai, \Omegac, \Omegar >$, with
-$\Omegai = \Omega_0$, 
-$\Omegac = \{ \; \call{a} \mid a \in \Omega_{>0} \}$,
-$\Omegac = \{ \; \return{a} \mid a \in \Omega_{>0} \}$,
-where $\Omega_{>0}$ denotes $\bigcup_{k = 1}^{M} \Omega_k$.
-
-\noindent
-We define a linearization of trees of $\T(\Omega)$ into 
-words of $\hat{\Omega}^*$ as follows:
-\begin{description}
-\item $\lin(a) = a$ for all $a \in \Omega_0$, 
-\item $\lin\bigl( b(t_1, \ldots, t_k)\bigr) = 
-       \call{b} \; \lin(t_1) \ldots \lin(t_k) \; \return{b}$ 
-       when $b \in \Omega_k$, $1 \leq k \leq M$.
-\end{description}
-
-% A \emph{regular tree grammar} over $\Omega$ 
-% is a triplet $G = \< N, q_\mathsf{i}, R>$ where
-% $N$ is a finite set of non-terminal symbols denoted $q$..., 
-% $q_\mathsf{i} \in N$ is the starting non-terminal, 
-% $R$ is a finite set of production rules of the form
-% $q_0 \to a(q_1\ldots q_k)$ where 
-% $q_0, q_1, \ldots, q_k \in N$
-% $a \in \Omega_k$.
-% A tree $t \in \T(\Omega)$ is in the language of $G$ 
-% if it can be generated from $q_\mathsf{i}$ by 
-% non terminal replacement following the rules of $R$.
-
-Let us assume a label theory $\Phi_{\Omega_k}$ 
-for each $k \in [0..M]$.
-%
-\begin{definition}  \label{def:SWTA}
-A \emph{symbolic-weighted tree automaton} (\SWTA)
-over the ranked input alphabet~$\Omega$
-and the commutative semiring $\Semiring$ 
-is a triplet $A = \< Q, \init, \bar{\wei} >$ where
-$Q$ is a finite set of states, 
-$\mathsf{in} : Q \to \Semiring$ is the starting weight function, 
-and $\bar{\wei}$ is a $M+2$-uplet of transition functions made of:
-$\wei_\epsilon$ from $Q \times Q$ into $\Semiring$, and, 
-for each $k \in [0..M]$, $\wei_{\Omega, k}$
-from $Q \times Q^{k}$ into $\Phi_{\Omega_k}$.
-\end{definition}
-%
-Like in Section~\ref{sec:SWAdef}, we define from $\bar{\wei}$ 
-a transition function~$\wei$, %from $\bar\wei$
-from 
-$Q \times (\Sigma \cup \{ \epsilon \}) \times \bigcup_{k=0}^{M} Q^k$
-into~$\Semiring$: %also called $\wei$ for simplicity, 
-%such that, for all $q, q' \in Q$, $a \in \Sigma$, and $b \in \Delta$, 
-\[
-\begin{array}{rcll}
-\wei(q_0, \epsilon, q_1) & = &  \wei_\epsilon(q_0, q_1),\\ %\phi_\epsilon\\
-\wei(q_0, a, q_1 \ldots q_k) & = & \phi_{\Omega, k}(a) &
-\quad\mathrm{where~} \phi_{\Omega, k} = \wei_{\Omega, k}(q_0, q_1\ldots q_k).\\
-\end{array}      
-\]
-Intuitively, $\wei(q_0, a, q_1 \ldots q_k)$ can be seen as
-the weight of a production rule $q_0 \to a(q_1, \ldots, q_k)$ 
-of a regular tree grammar~\cite{tata}, 
-that replaces the non-terminal symbol $q_0$, by $a(q_1, \ldots, q_k)$. 
-%in a step of tree building.
-%
-Such a grammar computes the weights of the derivation trees 
-of the Context-Free grammar obtained by forgetting the labeling symbols of $\Omega_{>0}$.
-
-\noindent
-The \SWTA of Definition~\ref{def:SWTA} defines a mapping 
-from trees of $\T(\Omega)$ into the weights of~$\Semiring$,
-based on the intermediate function $\weight_A$
-defined as follows for $q_0 \in Q$ and 
-$t = b(t_1,\ldots, t_k) \in \T(\Omega)$,
-åwith $0 \leq k \leq M$:
-\[
-\begin{array}{rccl}
-\weight_A(q_0, t) & = & 
-   \displaystyle\bigoplus_{q_1 \in Q} &
-   \wei(q, \epsilon, q_1) \otimes \weight_A(q_1, t)\\
- & \oplus & \displaystyle\bigoplus_{q_1 \ldots q_k \in Q^k} &
-              \wei(q_0, b, q_1 \ldots q_k ) 
-   \otimes \displaystyle\bigotimes_{i=1}^{k}
-           \weight_A(q_i, t_i).\\
-\end{array}
-\]
-
-\medskip\noindent
-The weight associated by $A$ to  $t \in \T(\Omega)$ is then
-\begin{equation}
-A(t)  = 
-\displaystyle\bigoplus_{q \in Q} \mathsf{in}(q) \mathop{\otimes} \weight_A(q, t).
-\label{eq:weightTA}
-\end{equation}
-
-\begin{lemma}  \label{lem:SWTA}
-For all \SWTA $A$ over~$\Omega$ and $\Semiring$, without $\epsilon$-transitions,
-there exists an effectively constructible \SWVPA $A'$ over 
-$\hat\Omega$ and $\Semiring$
-such that for all $t \in \T(\Omega)$, $A'\bigl(\lin(t)\bigr) = A(t)$.
-\end{lemma} 
-% 
-\begin{proof}
-Let $A = \< Q, \init, \bar{\wei} >$ where $\bar{\wei}$ is summarized as above by a function
-$\wei: Q \times (\Sigma \cup \{ \epsilon \}) \times \bigcup_{k=0}^{M} Q^k \to \Semiring$.
-
-We build 
-$A' = \< Q', P', \init', \weii, \weic, \weir, \weie, \final' >$,
-computing over $\hat\Omega = \< \Omegai, \Omegac, \Omegar >$,
-%
-%$\Omega = \Omegai \uplus \Omegac \uplus \Omegar$ and $\Semiring$ 
-where $Q' = \bigcup_{k=0}^{M} Q^k$ be  the set of sequences of state symbols of $A$, 
-of length at most $M$, including the empty sequence denoted by $\epsilon$, 
-and where $P' = Q'$.
-
-\[
-\begin{array}{ll}
-%\weii: Q \times \Omegai \times Q \to \Semiring & 
-\weii(\bar{q}, a, \bar{q} q') = \wei(q', a, \epsilon) & 
-\mathrm{for~all~} a \in \Omega_0,\\
-%
-%\weic: Q \times \Omegac \times Q \times P \to \Semiring & 
-\weic(\bar{q}, \call{b}, \epsilon, \bar{q}) = \one & %\wei(p', b, \bar{p}) & 
-\mathrm{for~all~} b \in \Omega_{>0},\\
-%
-%\weir: Q \times \Omegac \times P \times \Sigmar \times Q \to \Semiring & 
-\weir(\bar{q}, {\call{b}}, \bar{p}, {\return{b}}, \bar{p} q') = \wei(q', b, \bar{q}) & 
-\mathrm{for~all~}  b \in \Omega_{>0},\\
-%
-%\weie: Q \times \Omegar \times Q \to \Semiring & 
-\weie(\bar{p}, {\return{b}}, \bar{q}) = \zero &
-\mathrm{for~all~}  b \in \Omega_{>0}.\\
-\end{array}      
-\]
-
-In practice, it is sufficient to consider in $Q'$ only the prefixes of 
-sequences.
-\qed\end{proof}
-
-\section{Symbolic Weighted Parsing}
-\label{sec:parsing}
-Let us now use the models and results of the former sections in %order to define 
-an approach to the problem of parsing over infinite alphabet. %appropriate
-%
-Besides considering infinitely many possible of input symbols, 
-handled with suitable language formalisms, 
-this approach extends conventional parsing 
-by computing a derivation tree modulo 
-a generic distance between words,  
-defined by a SW transducer given in input.
-This enables considering finer word relationships than strict equality 
-as in the conventional parsing approach, 
-opening possibilities of quantitative analysis via this method.
-
-\subsection{Definition}
-
-Let $\Sigma$ be a countable input alphabet 
-and $\Omega = \Omegai \uplus \Omegac \uplus \Omegar$ an output countable alphabet,
-let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative, bounded, complete 
-and total semiring 
-and let  $(\Phi_\epsilon, \Phic, \Phir, \Phicr)$ be a label theory over $\Semiring$, 
-assumed computable and $k$-convex for some fixed $k$.
-
-\noindent
-Assuming given in input:
-\begin{itemize}
-\item a \SWT $T$ over $\Sigma$,  $\Omega$, and $\Semiring$,
-defining a measure %between words 
-$T: \Sigma^* \times \Omega^* \to \Semiring$,
-\item a \SWVPA $A$ over $\Omega$, and $\Semiring$, defining a series of nested words
-      $A : \Omega^* \to \Semiring$,
-\item an input word $s \in \Sigma^*$,
-\end{itemize}
-the problem of \emph{symbolic weighted parsing} is to find 
-a nested word $t \in \Omega$ minimizing
-\( T(s, t) \otimes A(t)\)
-\wrt $\leq_\oplus$, 
-\ie such that %Hence, it is the problem of finding 
-\( T(s, t) \otimes A(t) = 
-  \displaystyle\bigoplus_{t' \in \Omega^*} T(s, t') \otimes A(t') \).
-
-\noindent 
-Therefore, it is the problem of optimizing a measure 
-called the \emph{edit-distance between $s$ and $A$} in~\cite{Mohri03EDWA}.
-%
-The input language can also be expressed as a \SWTA, or, 
-as a particular case, as a weighted context-free grammar, 
-converted in turn into a \SWVPA following Lemma~\ref{lem:SWTA}.
-%
-In the case of finite alphabets, 
-the problem of searching, in a WTA language, 
-for the best parse tree for a given,  
-sometimes referred as~\emph{weighted parsing}
-(see \cite{Goodman99SemiringParsing,MorbitzVogler19weighted-parsing} **more general problems**)
-is a particular case of SW parsing.
-%
-Indeed, it corresponds to the case where $T$
-accepts only the pairs $\<s, t>$ such that 
-$s$ is the projection of $t$ on $\Omegai$. 
-This can be done with a single state $q$ and 
-with transition rules of the form:
-\begin{description}
-\item[] $\wei(q, \epsilon, a, q) = \one$ for all $a \in \Omegac \cup \Omegar$,
-\item[] $\wei(q, a, a, q) = \one$ for all $a \in \Omegai$,
-\item[] $\wei(q, a, b, q) = \zero$ for all $a, b \in \Omegai$, $a \neq b$.
-\end{description}
-
-\subsection{Computation}
-
-\begin{proposition}
-The problem of Symbolic Weighted  parsing 
-can be solved in PTIME in the size of the input \SWT, \SWVPA (or \SWTA) and input word, 
-and the computation time of the functions of the label theory.
-\end{proposition}
-%
-\begin{proof}
-Bar-Hillel construction
-\qed  
-\end{proof}
-
 \subsection{Application to Automated Music Transcription}
@@
 $\Semiring$ of Section~\ref{app:distance}.
-
-\section*{Conclusion}
-% discussion
-
-\noindent
--- summary
-
-\noindent
--- other theoretical properties of SW models
-
-\noindent
--- room to improve complexity for best-search algorithm
-... modular approach with oracles ...
-
-and extention to $n$-best
-
-\noindent
--- offline algorithm, 
-semi-online implementation for AMT (bar-by-bar approach)
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% BIBLIO                                                                     %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\bibliographystyle{plain}
-%\bibliographystyle{plainurl} 
-\bibliographystyle{abbrv}
-%\bibliographystyle{splncs04}
-\bibliography{references}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% APPENDIX                                                                   %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\appendix
-
-\section{Edit-Distance}
-
-%\subsection{Distance between words or languages}
-...algebraic definition of edit-distance of Mohri, in \cite{Mohri03EDWA}
-% Mehryar Mohri 
-% Edit-distance of weighted automata: General definitions and algorithms
-% International Journal of Foundations of Computer Science 14.06 (2003): 957-982.
-distance $d$ over $\Sigma^* \times \Sigma^*$ 
-into a semiring  $\Semiring = ( \Semiring, \oplus, \zero, \otimes, \one)$.
-
-%\noindent
-Let $\Omega = \Sigma \cup \{ \epsilon \} \times \Sigma \cup \{ \epsilon \} \setminus \{ (\epsilon, \epsilon) \}$,
-and let $h$ be the morphism from $\Omega^*$ into $\Sigma^* \times \Sigma^*$  
-defined over the concatenation of strings of $\Sigma^*$ (that removes the $\epsilon$'s).
-%
-\noindent
-An \emph{alignment} between 2 strings  $s, t \in \Sigma^*$ is an element $\omega \in \Omega^*$ 
-such that $h(\omega) = (s, t)$.
-%
-\noindent
-We assume a base cost function $\Omega$ : $\delta: \Omega \to S$, extended to $\Omega^*$ as follows  
-(for $\omega \in \Omega^*$): 
-\(
-\displaystyle\delta(\omega) = \bigotimes_{0 \leq i < |\omega|} \delta(\omega_i)
-\).
-
-\noindent
-\begin{definition}
-For  $s, t \in \Sigma^*$, the edit-distance between $s$ and $t$ is  
-\( 
-d(s, t) = \displaystyle\bigoplus_{\omega \in \Omega^*\, h(\omega) = (s, t)} \delta(\omega)
-\).
-\end{definition}
-
-e.g. Levenstein edit-distance: $S$ is min-plus and $\delta(a, b) = 1$ for all $(a, b) \in \Omega$.
-
-%\paragraph{Distance between a word and a regular language}
-
-\end{document}

### 2021-05-17 main.tex
cleanup

@@
 %** The application of $\oplus$ can be seen as a ND choice between transitions **
+% choose to advance head in input, in output or both (or none)
 
@@
 The transitions in the second column below switch back to the normal state $q_0$.
+% the metric computed by the \SWT is the smallest sum of point wise distances 
+% between dates of input and output events.
 \endex
@@
 \noindent
-The above transition functions $\wei_\epsilon$ and $\wei_\Sigma$
-can be synthesized as above into a function
+As above in the case of \SWT, 
+the transition functions $\wei_\epsilon$ and $\wei_\Sigma$
+can be synthesized into a single function
 $\wei: Q \times (\Sigma \cup \{ \epsilon \}) \times Q \to \Semiring$.
@@
 
+%\input{transcription}
+
 \section*{Conclusion}

### 2021-05-17 transcription.tex
cleanup

@@
+% !TEX root = main.tex
+%
+% Application to AMT
+%
 \subsection{Application to Automated Music Transcription}
@@
 %This representation is similar to the piano roll ~\cite{Muller15fundamentals} chap.1. 
-Moreover, it contains a RTU value $\ioi{e}$ (real number) 
+Moreover, it contains a time value $\ioi{e}$ (real number) in RTU (seconds)
 representing the time distance to the next event, 
@@
 composed of a tag, in a finite set $\Xi$, 
-and an MTU (rational) IOI duration value.
+and an IOI duration value expressed in MTU (rational number).
 %The alphabet $\Delta$ 
@@
 $\Semiring$ of Section~\ref{app:distance}.
+
+\subsection{Implementation}
+
+\textsf{qparse} based on a similar approach.
+
+slightly different models of symbolic weighted tree automata
+...WTA  specification of a prior score language.
+
+% we hope to gain flexibility with the present framework and extend application field
+% from transicription to other problems related to music notation and symbolic recorded performances.

### 2021-05-18 main2.tex
revise V2:  extended expressiveness

@@
 Let $\Semiring$ be a commutative and complete semiring, 
+$\Sigma$ and $\Delta$ be alphabets called respectively \emph{input} and \emph{output}, %{alphabets}, 
 and $\bar\Phi = \< \Phi_\Sigma, \Phi_\Delta, \Phi_{\Sigma, \Delta}>$ be an $\Semiring$-label theory.
@@
 \end{definition}
-%
-%A pair of states $\< q, q'>$ such that $\wei_\epsilon(q, q') \neq \zero$
-%is called an \emph{$\epsilon$-transition}.
-
-The \SWT $T$ computes on pairs of words $\< s, t> \in \Sigma^* \times \Delta^*$,
-$s$, resp. $t$, being called input word, resp. output word.
+
+A pair of states $\< q, q'>$ such that $\wei_{00}(q, q') \neq \zero$
+is called an \emph{$\epsilon$-transition}.
 %
@@
 $\wei_{ij}(q, a, b, q') = \eta(a, b)$ when  $a \in \Sigma$,  $b \in \Delta$.
-%
-Intuitively, using a transition $\wei_{ij}(q, a, b, q')$ means for $T$:
-read $a$, resp. $b$, at the current position in the input word, resp. output word;
-if $i = 1$, resp. $j = 1$,
-increment the current position in the input word (otherwise, do not change it);
-change state from $q$ to $q'$.
-
-\noindent
-More precisely, the \SWT~$T$ defines a mapping 
-from the pairs of words of $\Sigma^* \times \Delta^*$ into~$\Semiring$,
+
+\noindent 
+The \SWT $T$ computes on pairs of words $\< s, t> \in \Sigma^* \times \Delta^*$,
+$s$, resp. $t$, being called input word, resp. output word.
+%
+% \noindent
+More precisely, the symbolic-weighted transducer~$T$ defines a mapping 
+from the pairs of strings of $\Sigma^* \times \Delta^*$ 
+into~$\Semiring$,
 based on the following intermediate function $\weight_T$
@@
 \begin{array}{lccl}
-\weight_T(q, \epsilon, \epsilon, q)  & = & \multicolumn{2}{l}{\final(q)}\\
+\weight_T(q, \epsilon, \epsilon, q)  & = & \multicolumn{2}{l}{\one}\\ % or \final(q)
 \weight_T(q, \epsilon, \epsilon, q') & = & \multicolumn{2}{l}{\zero \mathrm{~if~} q \neq q'}\\
@@
   & \oplus & \displaystyle\bigoplus_{q'' \in Q} &
-    \wei_{10}(q'', a, b, q') \otimes \weight_T(q, u, bv, q'')\\
+    \wei_{10}(q, a, b, q'') \otimes \weight_T(q'', u, bv, q')\\
   & \oplus & \displaystyle\bigoplus_{q'' \in Q} &
-    \wei_{01}(q'', a, b, q') \otimes \weight_T(q, au, v, q'')\\
+    \wei_{01}(q, a, b, q'') \otimes \weight_T(q'', au, v, q')\\
   & \oplus & \displaystyle\bigoplus_{q'' \in Q} &
-    \wei_{11}(q'', a, b, q') \otimes \weight_T(q, u, v, q'')\\
+    \wei_{11}(q, a, b, q'') \otimes \weight_T(q'', u, v, q')\\
 \end{array}
 \]
-
-The above sum may be infinite because of the third line.
-However, $\weight_T$ is well defined since $\Semiring$ is complete.
-%
-As we shall see, under conditions, the $\epsilon$-transitions can be removed and hence the sum made finite.
-** SUPPR. $\epsilon$-transition (with $\wei_{00}$) first. **
-
-*** rem. there is no case for $a, \epsilon$ or $\epsilon, b$.
-   It can be simulated with $\wei_{10}$ and $\wei_{01}$ on the last letter. ***
+%
+We recall that, by convention, an empty sum with $\bigoplus$ is $\zero$. 
+%
+The above sum may be infinite because of the $\epsilon$-transitions, % in the first and second lines.
+in this case, the infinite sum operators applies since $\Semiring$ is assumed complete.
+%As we shall see, under conditions, the $\epsilon$-transitions can be removed and hence the sum made finite.
+%** SUPPR. $\epsilon$-transition (with $\wei_{00}$) first. **
+%
+%*** rem. there is no case for $a, \epsilon$ or $\epsilon, b$.
+%   It can be simulated with $\wei_{10}$ and $\wei_{01}$ on the last letter. ***
 
 %** The application of $\oplus$ can be seen as a ND choice between transitions **
-We recall that, by convention, an empty sum with $\oplus$ is $\zero$. 
-%
+% choose to advance head in input, in output or both (or none)
+Intuitively, using a transition $\wei_{ij}(q, a, b, q')$ means for $T$:
+read $a$, resp. $b$, at the current position in the input word, resp. output word;
+if $i = 1$, resp. $j = 1$,
+increment the current position in the input word (otherwise, do not change it);
+change state from $q$ to $q'$.
+The input and output symbols are always read at current positions, 
+even during transition that do not increment the head's position.
+This is in contrast with the models of weighted transducers over finite alphabets~\cite{Mohri03ijfcs},
+which ignore the symbol at current position in this case.
+In our case, symbols in a infinite alphabet cannot be stored in the finite memory of the transducer, 
+whereas considering the current symbol may be useful to compute a transition weight .
+%(even when it does not change the head's position, like with $\epsilon$-transitions).
+%
+The cases $\weight_T(q, au, \epsilon, q')$ and $\weight_T(q, \epsilon, bv, q')$
+are missing in the definition of $\weight_T$.
+It means that $T$ must avoid configurations where it reached the end of 
+the output word and not of the input one, or vice-versa.
+This can be done by using $\wei_{10}$ and $\wei_{01}$
+before reaching the end of word, and using a special state for this purpose.
+
 Since $\zero$ is absorbing for~$\otimes$ in~$\Semiring$,
@@
 \begin{example}
-Let $\Sigma$ be an alphabet of symbols with timestamps. 
-The timestamp of $a \in \Sigma$, denoted by $\mathsf{t}(a)$, is expressed as a rational number.
-
-DTW for sequences of timestamped events **     
+We can define with a \SWT the computation of a similarity measure 
+between timed sequences
+similar to dynamic time warping (DTW).
+
+Let $\Semiring$  be the tropical (\emph{min-plus}) semiring of Figure~\ref{fig:semirings} and 
+let $\Sigma = \Delta = \mathbb{R}_+$ be sets of timestamps.
+We consider a \SWT with one state $q$ and transitions
+$\wei_{11}(q, d, d', q) = 
+ \wei_{01}(q, d, d', q) = 
+ \wei_{10}(q, d, d', q) = |d' - d|$,
+and $\wei_{00}(q, d, d', q) = \zero$, for all $d, d' \in \mathbb{R}_+$.
+% needs reading input/output symbols even by epsilon-transitions.
+The recursive definition of $\weight_T$ correspond to the dynamic programming equations of DTW
+for the computation of an optimal match between words, 
+the matching cost for two symbols being the 
+the time distance between them.
+\endex
 \end{example}
@@
 \begin{example}
-(simpler) pointwise distance between two sequences of timestamped events **
+% simple pointwise distance between two sequences of timestamped events **
+Let us consider the tropical (\emph{min-plus}) semiring~$\Semiring$ 
+of Figure~\ref{fig:semirings} and 
+let $\Sigma = \mathbb{R}_+$ be an input alphabet of event dates
+and $\Delta = \{ \mathsf{e}, \mathsf{-} \} \times \mathbb{R}_+$ 
+be an output alphabet of symbols with timestamps. 
+A symbol $\< \mathsf{e}, d > \in \Delta$ represents an event starting at date $d$, 
+and $\< \mathsf{-}, d >$ is a continuation of the previous event.
+%More precisely, we let 
+%$\Sigma = \{ a, - \} \times \R_+$ and 
+This example of $\Delta$ is motivated by the case of music notation, 
+where several notated events (notes) can be tied together, 
+with a \emph{tie} or a \emph{dot}
+(like in $\musQuarter\!\!\mathrel{\raisebox{-1.5mm}{$\smile$}}\!\!\musEighth$ 
+ or equivalently~\musQuarterDotted),
+meaning that they will be played as a unique sounding event.
+%The timestamp of $a \in \Sigma$, denoted by $\mathsf{t}(a)$, is expressed as a rational number.
+
+We consider a \SWT with two states $q_0$ and $q_1$ whose purpose 
+is to compare a recorded performance $s \in \Sigma^*$
+with notated music sheet $t \in \Delta^*$.
+One timestamp $d_i \in \Sigma$ may corresponds 
+to one notated event $\<\mathsf{e}, d'_i> \in \Sigma$, in which case 
+the weight value computed by the \SWT is the time distance between both
+(see transitions $\wei_{11}$ below).
+%
+Moreover, if $\<\mathsf{e}, d'_i>$ is followed by continuations 
+$\<\mathsf{-}, d'_{i+1}>$..., they are just skip with no cost (transitions $\wei_{01}$ or weight $\one$).
+%transitions
+\[
+\begin{array}{rclcrcl}
+\wei_{11}(q_0, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d| & \quad &
+\wei_{11}(q_1, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d|\\
+\wei_{01}(q_0, d, \< \mathsf{-}, d'>, q_0) & = & \one & &
+\wei_{01}(q_1, d, \< \mathsf{-}, d'>, q_0) & = & \one\\
+\wei_{10}(q_0, d, b, q_1) & = & \alpha & & \multicolumn{3}{l}{\mathrm{for~all~} b \in \Delta}
+\end{array}
+\]
+%
+Moreover, it may happen that the performers plays an extra note accidentally, but only once in a row. 
+This is modelled by the transition $\wei_{10}$ with an arbitrary weight value $\alpha \in \Semiring$, 
+switching from state $q_0$ (normal) to $q_1$ (error).
+The transitions in the second column below switch back to the normal state $q_0$.
+% the metric computed by the \SWT is the smallest sum of point wise distances 
+% between dates of input and output events.
+At last, we let $q_0$ be the only initial and final state, with
+$\init(q_0) = \one$, $\init(q_1) = \zero$, 
+$\final(q_0) = \one$, $\final(q_1) = \zero$, 
+\endex
 \end{example}
@@
 %
-In this case, the label theory $\bar\Phi$ is reduced to a singleton $\< \Phi_\Sigma>$.
+In this case, the label theory $\bar\Phi$ can be reduced to a singleton $\< \Phi_\Sigma>$.
 %over $\Sigma$ is reduced to
@@
 and $\bar{\wei}$ is a pair of transition functions 
-$\wei_\epsilon$ and $\wei_\Sigma$ from $Q \times Q$ into 
-respectively $\Semiring$ %$\Phi_\epsilon$ 
-and $\Phi_\Sigma$.
+$\wei_{0}$, $\wei_{1}$
+%$\wei_{i, j}: Q \times Q \to \Phi_{\Sigma, \Delta}$, with $i, j \in \{ 0, 1\}$.
+from $Q \times Q$ into $\Phi_{\Sigma}$.
 \end{definition}
-
-\noindent
-The above transition functions $\wei_\epsilon$ and $\wei_\Sigma$
-can be synthesized as above into a function
-$\wei: Q \times (\Sigma \cup \{ \epsilon \}) \times Q \to \Semiring$.
+A pair of states $\< q, q'>$ such that $\wei_{0}(q, q') \neq \zero$
+is called an \emph{$\epsilon$-transition}.
+
+\noindent
+As above in the case of \SWT, 
+the transition functions $\wei_0$ and $\wei_1$
+can be synthesized into a single function
+$\wei: Q \times \Sigma \times Q \to \Semiring$.
 %
@@
 \begin{proposition}
-Given a \SWT $T$ 
-over alphabet $\Sigma$, $\Delta$ and a commutative semiring~$\Semiring$, 
+Given a \SWT $T$ over $\Sigma$, $\Delta$, $\Semiring$ and $\bar\Phi$,
 and $s \in \Sigma^*$, 
 there exists an effectively constructible \SWA 
-$A_{T, s}$ over $\Delta$ and $\Semiring$ 
+$A_{T, s}$ over $\Delta$, $\Semiring$ and $\bar\Phi$,
 such that for all $t \in \Delta^*$, $A_{T, s}(t) = T(s, t)$.
@@
 let 
-$\wei: Q \times (\Sigma \cup \{ \epsilon \}) \times (\Delta \cup \{ \epsilon \}) \times Q 
- \to \Semiring$
+$\wei: Q \times \Sigma \times \Delta \times Q \to \Semiring$
 be the synthesized form of 
-$\bar{\wei} = \< \wei_\epsilon, \wei_\Sigma, \wei_\Delta, \wei_{\Sigma, \Delta}>$,
+$\bar{\wei} = \< \wei_{00}, \wei_{10}, \wei_{01}, \wei_{11}>$,
 and let $s = s_1 \ldots s_n$ with $s_1, \ldots, s_n \in \Sigma$.
@@
 \noindent
-We define transition functions $\wei'_\epsilon$ and $\wei'_\Sigma$ for $A_{T, s}$, 
+We define transition functions $\wei'_0$ and $\wei'_1$ for $A_{T, s}$, 
 together with their synthesized form $\wei'$ as follows: 
@@
 \[
-\begin{array}{lclcrcll}
-\wei'_\epsilon\bigl(\< i, q>, \< i, q'>\bigr) & = & \wei_\epsilon(q, q') & \mathrm{\ie} &
-\wei'\bigl(\< i, q>, \epsilon, \< i, q'>\bigr)   & = & \wei(q, \epsilon, \epsilon, q'), & \\ 
-%
-\wei'_\epsilon\bigl(\< i, q>, \< i+1, q'>\bigr) & = & \phi(s_i) & 
-\multicolumn{5}{l}{\mathrm{where~} \phi = \wei_\Sigma(q, q'),}\\
- & & & \mathrm{\ie} &
-\wei'\bigl(\< i, q>, \epsilon, \< i+1, q'>\bigr) & = & \wei(q, s_i, \epsilon, q') & 
-\mathrm{if~} i < n,\\ 
-%
-\wei'_\epsilon\bigl(\< i, q>, \< j, q'>\bigr) & = & \zero & 
-%\mathrm{\ie} & \wei'(\< i, q>, \epsilon, \< j, q'>)   & = & \zero & & 
-\multicolumn{5}{l}{\mathrm{if~} j \neq i, i+1,}\\ 
-%
-\wei'_\Sigma\bigl(\< i, q>, \< i, q'>\bigr) & = & \wei_\Delta(q, q') & \mathrm{\ie} &
-\wei'\bigl(\< i, q>, b, \< i, q'>\bigr)          & = & \wei(q, \epsilon, b, q'),\\ 
-%
-\wei'_\Sigma\bigl(\< i, q>, \< i+1, q'>\bigr) & = & \eta_{s_i} &
-\multicolumn{5}{l}{\mathrm{where~} \eta = \wei_{\Sigma, \Delta}(q, q'),}\\
- & & & \mathrm{\ie} & \wei'\bigl(\< i, q>, b, \< i+1, q'>\bigr) & = & \wei(q, s_i, b, q') & 
- \mathrm{if~} i < n,\\ 
-%
-\wei'_\Sigma\bigl(\< i, q>, \< j, q'>\bigr) & = & \zero & \mathrm{\ie} &
-\wei'\bigl(\< i, q>, b, \< j, q'>\bigr)     & = & \multicolumn{2}{l}{\zero\ \mathrm{~if~} j \neq i, i+1.}
+\begin{array}{lclcr}
+\wei'_0\bigl(\< i, q>, \< i, q'>\bigr) & = & \wei_{00}(q, q')\\ 
+%
+\wei'_0\bigl(\< i, q>, \< i+1, q'>\bigr) & = & \phi(s_i) & 
+\mathrm{if} & i < n, \mathrm{where~} \phi = \wei_{10}(q, q'),\\
+%
+\wei'_0\bigl(\< i, q>, \< j, q'>\bigr) & = & \zero & 
+\mathrm{if} & j \neq i, i+1,\\ 
+%
+\wei'_1\bigl(\< i, q>, \< i, q'>\bigr) & = & \wei_{01}(q, q')\\
+%
+\wei'_1\bigl(\< i, q>, \< i+1, q'>\bigr) & = & \eta_{s_i} &
+\mathrm{if} & i < n, \mathrm{where~} \eta = \wei_{11}(q, q'),\\
+%
+\wei'_1\bigl(\< i, q>, \< j, q'>\bigr) & = & \zero & 
+\mathrm{if} & j \neq i, i+1.
 \end{array}      
@@
 is closed under partial applications (Section~\ref{section:symbols})
-in the above construction of $\wei'_\Sigma$.
+in the above construction of $\wei'_1$.
 
@@
 \begin{proposition}
-Given a \SWA $A$ over alphabet $\Sigma$ 
-and a commutative and bounded semiring $\Semiring$, 
+When $\Semiring$ is commutative, bounded and complete, 
+given a \SWA $A$ 
+over $\Sigma$, $\Semiring$ and $\bar\Phi$
 there exists an effectively constructible \SWA $A'$ 
 without $\epsilon$-transitions 
+over $\Sigma$, $\Semiring$ and $\bar\Phi$,
 such that for all $s \in \Sigma^*$, $A'(s) = A(s)$.
@@
 %
-\begin{proof}  ** Tb revised **   
-Let $A = \< Q, \init, \<\wei_\epsilon, \wei_\Sigma>, \final >$.
-The automaton $A'$ is $\< Q, \init, \<\wei'_\epsilon, \wei'_\Sigma>, \final >$,
+\begin{proof} Let $A = \< Q, \init, \<\wei_0, \wei_1>, \final >$.
+The automaton $A'$ is $\< Q, \init, \<\wei'_0, \wei'_1>, \final >$,
 where, for all $q, q' \in Q$,
-$\wei'_\epsilon(q, q') = \zero$  
+$\wei'_0(q, q') = \zero$  
 and 
+** Tb revised **   
 % \[
@@
 % \end{itemize}
-Let~$\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative semiring and  
-let  $(\Phi_\epsilon, \Phic, \Phir, \Phicr)$ be a label theory over $\Semiring$
+Let~$\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative and complete semiring and  
+let  $(\Phi_\epsilon, \Phii, \Phic, \Phir, \Phicr)$ be a label theory over $\Semiring$
 %In order to simplify notations, %and following the definition of Section~\ref{section:transducer}, 
 %we shall write respectively 
-where $\Phic$, $\Phir$ and~$\Phicr$ stand respectively 
-for~$\Phi_\Omegac$, $\Phi_\Omegar$ and~$\Phi_{\Omegac, \Omegar}$.
-%
-Moreover, we extend this theory with a set $\Phii$ 
-of unary functions in $\Omegai \to \Semiring$,
-closed under $\oplus$ and $\otimes$.
+where $\Phii$, $\Phic$, $\Phir$ and~$\Phicr$ stand respectively 
+for~$\Phi_\Omegai$, $\Phi_\Omegac$, $\Phi_\Omegar$ and~$\Phi_{\Omegac, \Omegar}$.
+%
+%Moreover, we extend this theory with a set $\Phii$ 
+%of unary functions in $\Omegai \to \Semiring$,
+%closed under $\oplus$ and $\otimes$.
 
@@
 commutative, bounded, complete, and total.
-and 
-** assume label theory is s.t. 
-for all $\phi \in \Phii$, 
-$\psi \in \Phir$, 
-and $\eta \in \Phicr$, 
-$\displaystyle\bigoplus_{a \in \Omegai} \phi(a)$ 
-$\displaystyle\bigoplus_{r \in \Omegar} \phi(r)$ and 
-$\displaystyle\bigoplus_{{\call{c}} \in \Omegac} 
-\displaystyle\bigoplus_{{\return{r}} \in \Omegar} \eta({\call{c}}, {\return{r}})$
-are computable...
+and assume an effective label theory.
+%s.t. for all $\phi \in \Phii$, 
+%$\psi \in \Phir$, 
+%and $\eta \in \Phicr$, 
+%$\displaystyle\bigoplus_{a \in \Omegai} \phi(a)$ 
+%$\displaystyle\bigoplus_{r \in \Omegar} \phi(r)$ and 
+%$\displaystyle\bigoplus_{{\call{c}} \in \Omegac} 
+%\displaystyle\bigoplus_{{\return{r}} \in \Omegar} \eta({\call{c}}, {\return{r}})$
+%are computable...
 
@@
 %
-\begin{proof}
-Bar-Hillel construction
-\qed  
-\end{proof}
-
-\subsection{Application to Automated Music Transcription}
-\label{sec:transcription}
-Symbolic Automated Music Transcription
-and analysis of music performances
-
-\subsubsection{Time Scales}
-Real-Time Unit (RTU) = seconds
-
-\noindent 
-Musical-Time Unit (MTU) = number of measures
-
-\noindent 
-conversion via tempo value
-
-\subsubsection{Representation of Music Performances}
-We consider symbolic representations of musical performances, as finite sequences of events.
-It corresponds to the concrete case of a MIDI file~\cite{SMF} 
-recorded  from an electronic keyboard, 
-or the output of a transcription from audio files~\cite{Benetos18AMTsurvey}.
-%
-For the sake of simplicity, 
-we shall only consider here the case of monophonic performances, 
-where at most one note is sounding at a time. 
-The approach however extends to the polyphonic case.
-
-A music performance is a finite sequence of events in a set~$\Sigma$.
-Every event $e \in \Sigma$ has attributes such from a finite domain, 
-like a number of key for a note 
-or a flag indicating that it is a rest 
-(\textsf{ON} and \textsf{OFF} messages in~\cite{SMF})
-and a velocity value (0..127 in~\cite{SMF})).
-%This representation is similar to the piano roll ~\cite{Muller15fundamentals} chap.1. 
-Moreover, it contains a RTU value $\ioi{e}$ (real number) 
-representing the time distance to the next event, 
-or to the end of performance for the last event,
-also called \emph{inter-onset interval}.
-
-\subsubsection{Representation of Music Scores}
-Music score are represented as structured words
-made of timed %quantified 
-events and parenthesized markups,
-akin of nested words~\cite{AlurMadhusudan09nested}.
-
-We consider an alphabet $\Delta$, every symbol of which is 
-composed of a tag, in a finite set $\Xi$, 
-and an MTU (rational) IOI duration value.
-%The alphabet $\Delta$ 
-It is partitioned into 
-$\Delta = \Deltai \uplus \Deltac \uplus \Deltar$, 
-like in Section~\ref{section:SWVPA}.
-%
-\noindent
-The symbols of $\Deltai$ represent events:
-% (infinite alphabet of internal symbols) made of:
-with tags indicating a new note or grace-note (with null IOI), 
-a rest or the continuation of the previous note (tie or dot).
-%
-The elements of $\Deltac \uplus \Deltar$ are matched
-markups for describing the structure of the score, 
-\ie the hierarchical grouping of events, and also, 
-importantly the division of time in measures, tuplets...
-%- parentheses for time divisions : tuplets, bars...
-(linearization of rhythm trees \cite{jacquemard:hal-01138642}...).
-They contain additional info such as tuple number, beaming policy...
-
-\noindent
-The duration values of letters of $\Delta$, in MTU (rational), 
-can be computed with the markups and tags (\eg grace note has duration 0).
-
-%\noindent
-%There are simultaneous events, since grace notes has duration 0. They are ordered.
-%
-%\noindent
-%Finite bound on the number of duration ratio. ?
-
-\begin{example}
-...      
-\end{example}
-
-\subsubsection{Performance/Score Distance Computation}
-\label{app:distance}
-We define a distance between performance and score representations
-by a $\SWT$ $T = \< Q, \init, \wei, \final >$, over a semiring $\Semiring$.
-** detail the elements of $\Semiring$ ....**
-%are quadruplets of the form
-%$\< t, s, \delta_t, \delta_s>$
-
-Every state of $Q$ contains a 
-tempo value in a finite domain (e.g. 30..300 bpm).
-This value can be fixed 
-or recomputed by the $T$ %transducer 
-after reading each event, 
-according to a perceptive/cognitive model of tempo 
-such as~\cite{LargeJones99tempo}
-(also used in the context of score following~\cite{Cont10TPAMI}).
-% we wont detail here.
-
-\subsection{Transcription by SW Parsing} %Best-first Search}
-We assume a score language defined by a \SWVPA over the semiring 
-$\Semiring$ of Section~\ref{app:distance}.
+%\begin{proof}
+%Bar-Hillel construction
+%\qed  
+%\end{proof}
+
+%\input{transcription}
 

### 2021-05-20 main2-epsilon.tex
revise V2: last version with epsilon-transitions before fork (diff. in defs, assumption on semiring needed)

--- 2021-05-18 main2.tex
@@
 \def\Phir{{\Phi_\mathsf{r}}}
+\def\Phici{{\Phi_\mathsf{ci}}}
+\def\Phicc{{\Phi_\mathsf{cc}}}
 \def\Phicr{{\Phi_\mathsf{cr}}}
+\def\Phicx{{\Phi_\mathsf{cx}}}
 \def\weii{{\wei_\mathsf{i}}}
@@
 \subsection{Semirings} 
-\label{section:semiring}
+\label{section:semiring}\label{sec:semiring}
 \input{semiring}
@@
 where $Q$ is a finite set of states, 
-$\mathsf{in} : Q \to \Semiring$, 
-respectively $\mathsf{out} : Q \to \Semiring,$
+$\mathsf{in} : Q \to \Phi_{\Sigma, \Delta}$, 
+respectively $\mathsf{out} : Q \to \Phi_{\Sigma, \Delta}$,
 are functions defining the weight for entering, 
@@
 
-A pair of states $\< q, q'>$ such that $\wei_{00}(q, q') \neq \zero$
-is called an \emph{$\epsilon$-transition}.
+%A pair of states $\< q, q'>$ such that $\wei_{00}(q, q') \neq \zero$
+%is called an \emph{$\epsilon$-transition}.
 %
@@
 for $q, q' \in Q$ and $i, j \in \{ 0, 1\}$,
+respectively $\mathsf{in}(q) = \eta$, $\mathsf{out}(q') = \eta$,
 we shall sometimes write, overloading the function names:
-$\wei_{ij}(q, a, b, q') = \eta(a, b)$ when  $a \in \Sigma$,  $b \in \Delta$.
-
-\noindent 
-The \SWT $T$ computes on pairs of words $\< s, t> \in \Sigma^* \times \Delta^*$,
-$s$, resp. $t$, being called input word, resp. output word.
+$\wei_{ij}(q, a, b, q') = \eta(a, b)$ for $a \in \Sigma$,  $b \in \Delta$,
+respectively $\mathsf{in}(q, a, b) = \eta(a, b)$,
+$\mathsf{out}(q', a, b) = \eta(a, b)$.
+
+\noindent 
+The \SWT $T$ computes on pairs of words $\< s, t> \in \Sigma^+ \times \Delta^+$,
+$s$ and $t$, being respectively called input and output word.
 %
@@
 More precisely, the symbolic-weighted transducer~$T$ defines a mapping 
-from the pairs of strings of $\Sigma^* \times \Delta^*$ 
+from the pairs of strings of $\Sigma^+ \times \Delta^+$ 
 into~$\Semiring$,
@@
 \begin{array}{lccl}
-\weight_T(q, \epsilon, \epsilon, q)  & = & \multicolumn{2}{l}{\one}\\ % or \final(q)
-\weight_T(q, \epsilon, \epsilon, q') & = & \multicolumn{2}{l}{\zero \mathrm{~if~} q \neq q'}\\
+%\weight_T(q, \epsilon, \epsilon, q)  & = & \multicolumn{2}{l}{\one}\\ % or \final(q)
+%\weight_T(q, \epsilon, \epsilon, q') & = & \multicolumn{2}{l}{\zero \mathrm{~if~} q \neq q'}\\
 \weight_T(q, au, bv, q') & = & 
-   \displaystyle\bigoplus_{q'' \in Q} &
-   \wei_{00}(q, a, b, q'') \otimes \weight_T(q'', au, bv, q')\\
-  & \oplus & \displaystyle\bigoplus_{q'' \in Q} &
-    \wei_{10}(q, a, b, q'') \otimes \weight_T(q'', u, bv, q')\\
-  & \oplus & \displaystyle\bigoplus_{q'' \in Q} &
-    \wei_{01}(q, a, b, q'') \otimes \weight_T(q'', au, v, q')\\
-  & \oplus & \displaystyle\bigoplus_{q'' \in Q} &
-    \wei_{11}(q, a, b, q'') \otimes \weight_T(q'', u, v, q')\\
+   \displaystyle\bigoplus_{r \in Q} &
+   \wei_{00}(q, a, b, r) \otimes \weight_T(r, au, bv, q')\\
+  & \oplus & \displaystyle\bigoplus_{r \in Q, u \neq \epsilon} &
+    \wei_{10}(q, a, b, r) \otimes \weight_T(r, u, bv, q')\\
+  & \oplus & \displaystyle\bigoplus_{r \in Q, v \neq \epsilon} &
+    \wei_{01}(q, a, b, r) \otimes \weight_T(r, au, v, q')\\
+  & \oplus & \displaystyle\bigoplus_{r \in Q, u,v \neq \epsilon} &
+    \wei_{11}(q, a, b, r) \otimes \weight_T(r, u, v, q')\\
+  & \oplus & \displaystyle\bigoplus_{u,v = \epsilon} &
+    \wei_{11}(q, a, b, q')
 \end{array}
@@
 %
-We recall that, by convention, an empty sum with $\bigoplus$ is $\zero$. 
-%
-The above sum may be infinite because of the $\epsilon$-transitions, % in the first and second lines.
-in this case, the infinite sum operators applies since $\Semiring$ is assumed complete.
+We recall that, by convention (Section~\ref{sec:semiring}), 
+an empty sum with $\bigoplus$ is equal to~$\zero$. 
+Since $\zero$ is absorbing for~$\otimes$ in~$\Semiring$,
+one term $\wei_{ij}(q, a, b, q'')$ equal to $\zero$ in the above expression 
+will be ignored in the sum, meaning that there is no possible transition
+from state $q$ into state $q'$ while reading $a$ and $b$.
+%(including the case $a = \epsilon$ and $b = \epsilon$).
+%These functions $\phi$ act as guards for the transducer's transitions, 
+%preventing a transition when they return the absorbing $\zero$ of $\Semiring$.
+This is analogous to the case of a transition's guard not satisfied by $\<a, b>$ for 
+symbolic transducers.
+% 
+The above sum may be infinite because of the transitions $\wei_{00}$. % in the first and second lines.
+In this case, the infinite sum operators applies since $\Semiring$ is assumed complete.
 %As we shall see, under conditions, the $\epsilon$-transitions can be removed and hence the sum made finite.
@@
 Intuitively, using a transition $\wei_{ij}(q, a, b, q')$ means for $T$:
-read $a$, resp. $b$, at the current position in the input word, resp. output word;
-if $i = 1$, resp. $j = 1$,
-increment the current position in the input word (otherwise, do not change it);
-change state from $q$ to $q'$.
-The input and output symbols are always read at current positions, 
-even during transition that do not increment the head's position.
-This is in contrast with the models of weighted transducers over finite alphabets~\cite{Mohri03ijfcs},
-which ignore the symbol at current position in this case.
-In our case, symbols in a infinite alphabet cannot be stored in the finite memory of the transducer, 
-whereas considering the current symbol may be useful to compute a transition weight .
+when reading respectively $a$ and $b$ at the current positions in the input and output words, 
+increment the current position in the input word if if $i = 1$ (otherwise, do not change it),
+increment the current position in the output word if if $j = 1$, 
+and change state from $q$ to $q'$.
+%
+In contrast with the models of weighted transducers over finite alphabets~\cite{Mohri03ijfcs},
+the input and output symbols at current positions are read by every transitions, 
+even transitions that do not increment the head's position.
+%This is in contrast with the models of weighted transducers over finite alphabets~\cite{Mohri03ijfcs},
+%which ignore the symbol at current position in this case.
+This is an important feature in the case of an infinite alphabet, 
+since symbols cannot be stored in the finite memory of the transducer in this case.
+%whereas considering the current symbol may be useful to compute a transition weight.
 %(even when it does not change the head's position, like with $\epsilon$-transitions).
 %
-The cases $\weight_T(q, au, \epsilon, q')$ and $\weight_T(q, \epsilon, bv, q')$
-are missing in the definition of $\weight_T$.
-It means that $T$ must avoid configurations where it reached the end of 
-the output word and not of the input one, or vice-versa.
-This can be done by using $\wei_{10}$ and $\wei_{01}$
-before reaching the end of word, and using a special state for this purpose.
-
-Since $\zero$ is absorbing for~$\otimes$ in~$\Semiring$,
-one term $\wei_{ij}(q, a, b, q'')$ equal to $\zero$ in the above expression 
-will be ignored in the sum, meaning that there is no possible transition
-from state $q$ into state $q'$ while reading $a$ and writing $b$ 
-(including the case $a = \epsilon$ and $b = \epsilon$).
-%These functions $\phi$ act as guards for the transducer's transitions, 
-%preventing a transition when they return the absorbing $\zero$ of $\Semiring$.
-This is analogous to a transition's guard not satisfied by $\<a, b>$ in 
-the case of symbolic transducers.
+%The cases $\weight_T(q, au, \epsilon, q')$ and $\weight_T(q, \epsilon, bv, q')$
+%are missing in the definition of $\weight_T$.
+%It means that $T$ must avoid configurations where it reached the end of 
+%the output word and not of the input one, or vice-versa.
+%This can be done by using $\wei_{10}$ and $\wei_{01}$
+%before reaching the end of word, and using a special state for this purpose.
 
@@
 Intuitively, 
-$\wei_{10}(q, a, b, q'')$, in the fourth line, is the cost of 
+$\wei_{10}(q, a, b, r)$ is the cost of 
 the deletion of the symbol~$a \in \Sigma$ in~$s$, 
-$\wei{01}(q, a, b, q'')$, in the fifth line, is the cost 
+$\wei{01}(q, a, b, r)$ is the cost 
 of the insertion of~$b \in \Delta$ in $t$, 
-and $\wei_{11}(q, a, b, q'')$, in the last line, is the cost 
+and $\wei_{11}(q, a, b, r)$ is the cost 
 of the substitution of  $a \in \Sigma$ by~$b \in \Delta$.
@@
 \medskip\noindent
-The weight associated by $T$ to  $\< s, t> \in \Sigma^* \times \Delta^*$
+Let $\< s, t> \in \Sigma^+ \times \Delta^+$, with $s = s_0\ldots s_m$, $t = t_0\ldots t_n$). 
+The weight associated by $T$ to $\< s, t>$
 is defined as follows:
@@
 T(s, t)  = 
-\displaystyle\bigoplus_{q, q' \in Q} \mathsf{in}(q) 
-\mathop{\otimes} \weight_T(q, s, t, q') \mathop{\otimes} \mathsf{out}(q').
-\label{eq:weightT}
+\displaystyle\bigoplus_{q, q' \in Q} \mathsf{in}(q, s_0, t_0) 
+\mathop{\otimes} \weight_T(q, s, t, q') \mathop{\otimes} \mathsf{out}(q', s_m, t_n)
+\label{eq:SWT-value}
 \end{equation}
@@
 where $Q$ is a finite set of states, 
-$\mathsf{in} : Q \to \Semiring$, 
-respectively $\mathsf{out} : Q \to \Semiring,$
+$\mathsf{in} : Q \to \Phi_\Sigma$, 
+respectively $\mathsf{out} : Q \to \Phi_\Sigma,$
 are functions defining the weight for entering, 
@@
 can be synthesized into a single function
-$\wei: Q \times \Sigma \times Q \to \Semiring$.
+$\wei: Q \times \Sigma \times Q \to \Semiring$, 
+and the computation of $A$ on words $s \in \Sigma^+$
+is defined with an intermediate function $\weight_A$, 
+%$a \in \Sigma$, $b \in \Delta$ 
+%for every strings of $s \in \Sigma^*$, $t \in \Delta^*$:
+defined as follows for $q, q' \in Q$, $a \in \Sigma$, $u \in \Sigma^*$,
+\[
+\begin{array}{lccl}
+%\weight_T(q, \epsilon, \epsilon, q)  & = & \multicolumn{2}{l}{\one}\\ % or \final(q)
+%\weight_T(q, \epsilon, \epsilon, q') & = & \multicolumn{2}{l}{\zero \mathrm{~if~} q \neq q'}\\
+\weight_A(q, au, q') & = & 
+   \displaystyle\bigoplus_{r \in Q} &
+   \wei_{0}(q, a, r) \otimes \weight_A(r, au, q')\\
+  & \oplus & \displaystyle\bigoplus_{r \in Q, u \neq \epsilon} &
+    \wei_{1}(q, a, r) \otimes \weight_A(r, u, q')\\
+  & \oplus & \displaystyle\bigoplus_{u = \epsilon} & 
+    \wei_{1}(q, a, q')
+\end{array}
+\]
+%
+\noindent
+and the weight value associated by $A$ to 
+$s = s_0\ldots s_m \in \Sigma^+$ is: % defined as follows:
+\begin{equation}
+A(s)  = 
+\displaystyle\bigoplus_{q, q' \in Q} \mathsf{in}(q, s_0) 
+\mathop{\otimes} \weight_A(q, s, q') \mathop{\otimes} \mathsf{out}(q', s_m)
+\label{eq:SWA-value}
+\end{equation}
+
 %
@@
 
-\begin{proposition}
+\begin{proposition} \label{prop:epsilon}
 Given a \SWT $T$ over $\Sigma$, $\Delta$, $\Semiring$ and $\bar\Phi$,
@@
 \[
-\begin{array}{lclcr}
+\begin{array}{lclcl}
 \wei'_0\bigl(\< i, q>, \< i, q'>\bigr) & = & \wei_{00}(q, q')\\ 
@@
 \end{proof}
-
+%
+\noindent
 The construction time and size for $A_{T, s}$ are $O(\| T \| . | s |)$,
@@
 % gen. to transducers
+\medskip
+The following proposition shows that the transitions of type $\wei_{00}$
+(analogous of $\epsilon$-transitions of automata over finite alphabets)
+are not necessary to the expressiveness of $\SWA$, 
+and can safely nullified, 
+with an appropriate update of the other transitions.
+%
 \begin{proposition}
@@
 there exists an effectively constructible \SWA $A'$ 
-without $\epsilon$-transitions 
 over $\Sigma$, $\Semiring$ and $\bar\Phi$,
+without $\epsilon$-transitions,
 such that for all $s \in \Sigma^*$, $A'(s) = A(s)$.
@@
 %
-\begin{proof} Let $A = \< Q, \init, \<\wei_0, \wei_1>, \final >$.
-The automaton $A'$ is $\< Q, \init, \<\wei'_0, \wei'_1>, \final >$,
-where, for all $q, q' \in Q$,
-$\wei'_0(q, q') = \zero$  
+\begin{proof} 
+Let $A = \< Q, \init, \<\wei_0, \wei_1>, \final >$.
+%
+We associate to every sequence in $Q^+$ called \emph{$\epsilon$-path}, 
+and every $a \in \Sigma$, a weight value 
+defined recursively as follows (for $q_1, q_2 \in Q$ and $\sigma \in Q^*$):
+\[
+\begin{array}{rcl}
+\weight_\epsilon(q_1, a) & = & \one\\
+\weight_\epsilon(q_1 q_2 \sigma, a) & = & \wei_0(q_1, a, q_2) \otimes \weight_\epsilon(q_2 \sigma, a)\\
+\end{array}
+\]
+%
+We shall construct a \SWA $A' = \< Q, \init, \<\wei'_0, \wei'_1>, \final' >$,
+where, for all $q, q' \in Q$, and $a \in \Sigma$, 
 and 
-** Tb revised **   
-% \[
-% \begin{array}{rcccl}
-% \wei'_\Sigma(q, q') & = &  & \displaystyle\bigoplus_{q_1 \in Q} &
-%                \wei_\Sigma(q, q_1) \otimes  \wei_\epsilon(q_1, q')\\
-%                     &   & \oplus  & \displaystyle\bigoplus_{q_2 \in Q} &
-%                \wei_\epsilon(q, q_2) \otimes \wei_\Sigma(q_2, q')\\
-% \end{array}
-% \]
-\[
-\wei'_\Sigma(q, q') = 
-\displaystyle\bigoplus_{q'' \in Q} \wei_\epsilon(q, q'') \otimes \wei_\Sigma(q'', q')
-\oplus \displaystyle\bigoplus_{q_1 \in Q} \wei_\Sigma(q, q'') \otimes \wei_\epsilon(q'', q')
-\]
-In the above definition of $\wei'_\Sigma$ we use the operator
+\[
+\begin{array}{rcl}
+\wei'_0(q, a, q') & = & \zero\\
+\wei'_1(q, a, q') & = & 
+\displaystyle\bigoplus_{r \in Q} 
+\displaystyle\bigoplus_{\sigma \in Q^*} 
+\weight_\epsilon(q\sigma r, a) \otimes  \wei_1(r, a, q')\\
+\final'(q, a) & = & 
+\final(a, q) \oplus
+\displaystyle\bigoplus_{r \in Q} 
+\displaystyle\bigoplus_{\sigma \in Q^*} 
+\weight_\epsilon(q\sigma r, a) \otimes \final(r, a)\\
+\end{array}
+\]
+Hence $\wei'_0(q, q')$ is the constant function $\zero$ of $\Phi_\Sigma$.
+Following Equation~(\ref{eq:SWA-value}),
+\[
+\begin{array}{lccl}
+\weight_{A'}(q, au, q') & = & 
+    \displaystyle\bigoplus_{r \in Q, u \neq \epsilon} &
+    \wei'_{1}(q, a, r) \otimes \weight_{A'}(r, u, q')
+    \oplus \displaystyle\bigoplus_{u = \epsilon} \wei'_{1}(q, a, q') \\
+ & = & \displaystyle\bigoplus_{r \in Q, u \neq \epsilon} & 
+  \displaystyle\bigoplus_{\sigma \in Q^*} 
+  \weight_\epsilon(q\sigma r, a) \otimes  \wei_1(r, a, q') \otimes \weight_{A'}(r, u, q')\\
+ & \oplus & \displaystyle\bigoplus_{r \in Q, u = \epsilon} &
+   \displaystyle\bigoplus_{\sigma \in Q^*} \weight_\epsilon(q\sigma r, a) \otimes \wei_{1}(r, a, q')  
+\end{array}
+\]
+Moreover, we can show that
+\[
+\begin{array}{lccl}
+\weight_{A}(q, au, q') & = & 
+    \displaystyle\bigoplus_{r \in Q, u \neq \epsilon} &
+    \wei'_{1}(q, a, r) \otimes \weight_{A'}(r, u, q')\\
+ & = & \displaystyle\bigoplus_{r \in Q, u \neq \epsilon} & 
+  \displaystyle\bigoplus_{\sigma \in Q^*} 
+  \weight_\epsilon(q\sigma r, a) \otimes  \wei_1(r, a, q')
+\end{array}
+\]
+
+%By definition of $A(s)$ and $A'(s)$ (Equation~\ref{eq:SWA-value}), based on 
+%$\weight_A(q, s, q')$ and $\weight_{A'}(q, s, q')$, 
+%and by distributy of $\otimes$ over $\bigoplus$.
+
+In the above definition of $\wei'_1(q, a, q')$ and $\final'(q, a)$, 
+the first sum is finite and the second is not. 
+%
+We will show how to compute effectively the sum 
+$\bigoplus_{\sigma \in Q^*} \weight_\epsilon(q\sigma r, a)$
+for given $q$, $r$ and $a$, 
+using a shortest distance algorithm.
+
+** In the above definition of $\wei'_\Sigma$ we use the operator
 of product of function of $\Phi_\Sigma$ by $\Semiring$ 
@@
 \end{proof}
+
+The construction of Proposition~\ref{prop:epsilon} 
+can be straigthforwardly generalized to \SWT.
+
+not bounded: 
+procedure of removal of $\epsilon$-transitions
+... essentially the same as for \WA~\cite{Mohri02ijfcs}.
 
@@
 % \end{itemize}
-Let~$\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative and complete semiring and  
-let  $(\Phi_\epsilon, \Phii, \Phic, \Phir, \Phicr)$ be a label theory over $\Semiring$
+Let~$\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative and complete semiring and let  
+$\bar\Phi = \< \Phii, \Phic, \Phir, \Phici, \Phicc, \Phicr>$ 
+be a label theory over $\Semiring$
 %In order to simplify notations, %and following the definition of Section~\ref{section:transducer}, 
 %we shall write respectively 
-where $\Phii$, $\Phic$, $\Phir$ and~$\Phicr$ stand respectively 
-for~$\Phi_\Omegai$, $\Phi_\Omegac$, $\Phi_\Omegar$ and~$\Phi_{\Omegac, \Omegar}$.
+where $\Phii$, $\Phic$, $\Phir$ and~$\Phicx$ (with $\mathsf{x} \in \{ \mathsf{i}, \mathsf{c}, \mathsf{r}\}$) 
+stand respectively 
+for~$\Phi_\Omegai$, $\Phi_\Omegac$, $\Phi_\Omegar$ and~$\Phi_{\Omegac, \Omega_\mathsf{x}}$.
 %
@@
 A \emph{Symbolic Weighted Visibly Pushdown Automata} (\SWVPA) 
-over  $\Omega = \Omegai \uplus \Omegac \uplus \Omegar$ and $\Semiring$ is a tuple
-$A = \< Q, P, \init, \weii, \weic, \weir, \weie, \final >$,
+over  $\Omega = \Omegai \uplus \Omegac \uplus \Omegar$, $\Semiring$ and $\bar\Phi$ 
+is a tuple $A = \< Q, P, \init, \weii, \weic, \weir, \weie, \final >$,
 where $Q$ is a finite set of states, 
@@
 and 
-$\weii : Q \times Q \to \Phii$,  
-$\weic : Q \times Q \times P \to \Phic$,  
+$\weii : Q \times P \times Q \to \Phici$,  
+$\weic : Q \times P \times Q \times P \to \Phicc$,  
 $\weir : Q \times P \times Q \to \Phicr$,  
@@
 \begin{array}{lll}
-\weii: Q \times \Omegai \times Q \to \Semiring & 
-\weii(q, a, q') = \phi_\mathsf{i}(a) & 
-\mathrm{where~} \phi_\mathsf{i} = \weii(q, q'),\\
-%
-\weic: Q \times \Omegac \times Q \times P \to \Semiring & 
-\weic(q, \call{c}, q', p) = \phi_\mathsf{c}(\call{c}) & 
-\mathrm{where~} \phi_\mathsf{c} = \weic(q, q', p),\\
+\weii: Q \times \Omegac \times P \times \Omegai \times Q \to \Semiring & 
+\weii(q, c, p, a, q') = \eta_\mathsf{ci}(c, a) & 
+\mathrm{where~} \eta_\mathsf{ci} = \weii(q, p, q'),\\
+%
+\weic: Q \times \Omegac \times P \times  \Omegac \times Q \times P \to \Semiring & 
+\weic(q, c, p, \call{c'}, q', p') = \eta_\mathsf{cc}(c, \call{c'}) & 
+\mathrm{where~} \eta_\mathsf{cc} = \weic(q, p, q', p'),\\
 %
 \weir: Q \times \Omegac \times P \times \Omegar \times Q \to \Semiring & 
-\weir(q, {\call{c}},  p, {\return{r}}, q') = \phi_\mathsf{r}({\call{c}},  {\return{r}}) & 
-\mathrm{where~} \phi_\mathsf{r} = \weir(q, p, q'),\\
+\weir(q, {\call{c}},  p, {\return{r}}, q') = \eta_\mathsf{cr}({\call{c}},  {\return{r}}) & 
+\mathrm{where~} \eta_\mathsf{cr} = \weir(q, p, q'),\\
 %

### 2021-05-20 semiring.tex
example
revise V2: last version with epsilon-transitions before fork (diff. in defs, assumption on semiring needed)

@@
 %
-We shall consider semiring for the weight values of our language models.
+We shall consider semirings for the weight values of our language models.
 %
@@
 \begin{example}
-Figure~\ref{fig:semirings} presented examples of semirings interesting in practice 
+Figure~\ref{fig:semirings} presents examples of semirings interesting in practice 
 and enjoying the above properties.
@@
 \end{figure}
+
+%%% semiring properties used in paper
+% - commutative
+% - bounded (implies idempotent and superior)
+% - total
+% - complete

### 2021-05-21 epsilon-removal.tex
applicaion partielle transducteur

--- 2021-05-20 main2-epsilon.tex
@@
-% fork of main.tex with more expressive models
-% SWT read both tapes (under 2 heads) at every transition (even epsilon or semi-epsilon)
-% SWA read tape (under head) at every transition (even epsilon)
-% SW-VPA read top of stack (when not empty) at every transition
-
-\documentclass[runningheads]{llncs}
-%\documentclass[a4paper,11pt]{article}
-%\setcounter{page}{1}
-
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-
-\usepackage{hyperref}
-%\usepackage[bookmarks,bookmarksnumbered,naturalnames,plainpages=false]{hyperref}
-%usepackage{url}
-
-% for footnote ref
-\usepackage{refcount}
-
-% array and tabular
-\usepackage{array}
-\newcolumntype{C}[1]{>{\centering\arraybackslash\hspace{0pt}}p{#1}}
-
-% extension of enumerate env. (style for displaying counters)
-% \usepackage{enumerate}
-
-%% pictures
-% \usepackage{graphicx} 
-% \DeclareGraphicsExtensions{.pdf,.png,.jpg}
-% \graphicspath{fig/}
-
-%% PGF, Tikz
-\usepackage{tikz-cd}
-%% \usepackage{pgfplots}
-%% \usepgfplotslibrary{dateplot}
-%% \usepackage{pgf,pgfarrows,pgfnodes, pgfautomata}
-% \usepackage{tikz}
-% \usetikzlibrary{cd}
-%% \usetikzlibrary{arrows}
-%% \usetikzlibrary{calc}
-%% \usetikzlibrary{snakes}
-%% \usetikzlibrary{backgrounds}
-% \usetikzlibrary{trees}
-%% \usetikzlibrary{automata}
-%% \usetikzlibrary{positioning}
-%% \usetikzlibrary{matrix}
-%% \usetikzlibrary{patterns}
-%% \usetikzlibrary{shapes}
-
-% symbols
-\usepackage{amsmath} 
-\usepackage{amssymb} 
-\usepackage{amsbsy}
-\usepackage{bbold}
-\usepackage{latexsym}
-%\usepackage{amsfonts}
-\usepackage{stmaryrd}
-%\usepackage{mathabx}
-%\usepackage{MnSymbol}
-\usepackage{harmony} % simple music fonts
-\usepackage{mathtools} % for arrows
-%\usepackage{mathptmx}
-
-%% theorem environments
-\usepackage{theorem}
-% \newtheorem{theorem}{Theorem} %[section]
-% \newtheorem{definition}[theorem]{Definition}
-% \newtheorem{lemma}[theorem]{Lemma}
-% \newtheorem{corollary}[theorem]{Corollary}
-% \newtheorem{proposition}[theorem]{Proposition}
-% \newenvironment{proof}{\vspace{-2ex}{\it Proof. }}{\hspace*{\fill} $\Box$\smallskip }
-% \theorembodyfont{\slshape}
-% \newtheorem{example}[theorem]{Example}
-% \newtheorem{remark}[theorem]{Remark}
-\def\endex{\hspace*{\fill} $\Diamond$\smallskip }
-{\theorembodyfont{\rmfamily} \theoremstyle{break} \newtheorem{algo}{Algorithm}}
-
-%% algorithms
-%\usepackage{algorithm} 
-%\usepackage{program}
-
-%% for new macros
-\usepackage{xspace}
-
-%% arrows etc
-%\input{rewriting}
-
-%% music symbols
-% see http://tug.ctan.org/info/latex4musicians/latex4musicians.pdf
-\usepackage{musicography}
-
-%% Misc macros
-
-\def\ie{\textit{i.e.}\xspace}
-\def\eg{\textit{e.g.}\xspace}
-\def\wrt{\textit{wrt}\xspace}
-%\def\wlog{\textit{wlog}\xspace}
-\def\etc{\textit{etc}\xspace}
-
-\def\<#1>{\langle #1 \rangle}
-\newcommand{\pair}[2]{\langle{#1}, {#2}\rangle}
-%\newcommand{\A}{\mathcal{A}}
-%\newcommand{\B}{\mathcal{B}}
-\newcommand{\D}{\mathbb{D}}
-\newcommand{\E}{\mathbb{E}}
-\newcommand{\W}{\mathbb{W}}
-\newcommand{\T}{\mathcal{T}}
-
-\newcommand{\Semiring}{\mathbb{S}}
-\newcommand{\zero}{\mathbb{0}}
-\newcommand{\one}{\mathbb{1}}
-\newcommand{\dom}{\ensuremath{\mathit{dom}}}
-
-\def\SA{\textsf{sA}\xspace}
-\def\WA{\textsf{wA}\xspace}
-\def\SWT{\textsf{swT}\xspace}
-\def\SWA{\textsf{swA}\xspace}
-\def\SWTA{\textsf{swTA}\xspace}
-\def\SWVPA{\textsf{sw-VPA}\xspace}
-\def\weight{\mathsf{weight}}
-\def\wei{\mathsf{w}}
-\def\mei{\mathsf{m}}
-\def\init{\mathsf{in}}
-\def\final{\mathsf{out}}
-\newcommand{\call}[1]{\ensuremath #1} %{\ensuremath \langle_{#1}}
-\newcommand{\return}[1]{\ensuremath #1} %{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
-\def\Omegai{{\Omega_\mathsf{i}}}
-\def\Omegac{{\Omega_\mathsf{c}}}
-\def\Omegar{{\Omega_\mathsf{r}}}
-\def\Sigmai{{\Sigma_\mathsf{i}}}
-\def\Sigmac{{\Sigma_\mathsf{c}}}
-\def\Sigmar{{\Sigma_\mathsf{r}}}
-\def\Deltai{{\Delta_\mathsf{i}}}
-\def\Deltac{{\Delta_\mathsf{c}}}
-\def\Deltar{{\Delta_\mathsf{r}}}
-\def\Phii{{\Phi_\mathsf{i}}}
-\def\Phic{{\Phi_\mathsf{c}}}
-\def\Phir{{\Phi_\mathsf{r}}}
-\def\Phici{{\Phi_\mathsf{ci}}}
-\def\Phicc{{\Phi_\mathsf{cc}}}
-\def\Phicr{{\Phi_\mathsf{cr}}}
-\def\Phicx{{\Phi_\mathsf{cx}}}
-\def\weii{{\wei_\mathsf{i}}}
-\def\weic{{\wei_\mathsf{c}}}
-\def\weir{{\wei_\mathsf{r}}}
-\def\weie{{\wei_\mathsf{e}}}
-\newcommand{\config}[2]{\ensuremath \genfrac{[}{]}{0pt}{}{#1}{#2}} 
-\newcommand{\opluseq}{\ensuremath\mathrel{\oplus}=}
-\newcommand{\ioi}[1]{\mathsf{ioi}({#1})}
-\newcommand{\rank}{\mathsf{rk}}
-\newcommand{\lin}{\mathsf{lin}}
-
-%\sloppy
-
-% Parsing over infinite alphabet as optimal alignment computation 
-% as edit-distance between string and language
-% 
-%\title{Symbolic Weighted Parsing and Automated Music Transcription}
-%\title{Symbolic Weighted Language Models and Automated Music Transcription}
-\title{Symbolic Weighted Language Models and\\ Parsing over Infinite Alphabets}
-%\title{Weighted Visibly Pushdown Automata and Automated Music Transcription}
-\author{Florent Jacquemard}
-\institute{INRIA \& CNAM, Paris, France\\
-\email{florent.jacquemard@inria.fr}}
-
-%\titlerunning{WVPA \& AMT}
-%\titlerunning{Symbolic Weighted Parsing} % and Automated Music Transcription
-%\authorrunning{Florent Jacquemard}
-
-\date{\today}
-
-\begin{document}
-\thispagestyle{empty}
-\maketitle
-
-\begin{abstract}
-\input{abstract}
-\end{abstract}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% intro
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Introduction} \label{sec:intro}
-\input{introduction}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SWT & SWA
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{SW Automata and Transducers}
-\label{section:transducer}\label{sec:transducer}
-\label{section:SWA}\label{sec:SWA}
-\label{section:SWT}\label{sec:SWT}
-
-We follow the approach of~\cite{Mohri03EDWA} for the computation of distances,
-between words and languages, using weighted transducers, 
-and extend it to infinite alphabets.
-% with models of symbolic weighted automata and transducers. 
-%
-The models introduced in this section generalize 
-weighted automata and transducers~\cite{Droste09handbook} 
-%over finite alphabets, see  e.g.~\cite{Mohri03EDWA}, 
-by labelling each transition with a weight function (instead of a simple weight value), 
-that takes the input and output symbols as parameters. 
-These functions are similar to the guards of symbolic automata~\cite{dAntoniVeanes17CAV,dAntoni21CACM},
-but they can return values in an generic semiring, 
-whereas the latter guards are restricted to the Boolean semiring.
-
-\subsection{Semirings} 
-\label{section:semiring}\label{sec:semiring}
-\input{semiring}
-
-\subsection{Label Theory}
-\label{section:symbols}
-\input{labels}
-
-\subsection{Definitions} \label{sec:SWTdef}\label{sec:SWAdef}
-%(SWT)
-
-Let $\Semiring$ be a commutative and complete semiring, 
-$\Sigma$ and $\Delta$ be alphabets called respectively \emph{input} and \emph{output}, %{alphabets}, 
-and $\bar\Phi = \< \Phi_\Sigma, \Phi_\Delta, \Phi_{\Sigma, \Delta}>$ be an $\Semiring$-label theory.
-
-\begin{definition}
-\label{def:transducer} \label{def:SWT}
-A \emph{symbolic-weighted transducer} (\SWT)
-over $\Sigma$, $\Delta$, $\Semiring$ and $\bar\Phi$
-%the input and output alphabets~$\Sigma$ and $\Delta$ with label theory $\bar\Phi$, and the semiring $\Semiring$ 
-is a tuple
-$T = \< Q, \init, \bar{\wei}, \final >$,
-where $Q$ is a finite set of states, 
-$\mathsf{in} : Q \to \Phi_{\Sigma, \Delta}$, 
-respectively $\mathsf{out} : Q \to \Phi_{\Sigma, \Delta}$,
-are functions defining the weight for entering, 
-respectively leaving, a state, 
-and $\bar{\wei}$ is a 4-uplet of transition functions 
-$\wei_{00}$, $\wei_{10}$, $\wei_{01}$, and $\wei_{11}$
-%$\wei_{i, j}: Q \times Q \to \Phi_{\Sigma, \Delta}$, with $i, j \in \{ 0, 1\}$.
-from $Q \times Q$ into $\Phi_{\Sigma, \Delta}$.
-\end{definition}
-
-%A pair of states $\< q, q'>$ such that $\wei_{00}(q, q') \neq \zero$
-%is called an \emph{$\epsilon$-transition}.
-%
-For convenience, when $\wei_{ij}(q, q') = \eta \in \Phi_{\Sigma, \Delta}$, 
-for $q, q' \in Q$ and $i, j \in \{ 0, 1\}$,
-respectively $\mathsf{in}(q) = \eta$, $\mathsf{out}(q') = \eta$,
-we shall sometimes write, overloading the function names:
-$\wei_{ij}(q, a, b, q') = \eta(a, b)$ for $a \in \Sigma$,  $b \in \Delta$,
-respectively $\mathsf{in}(q, a, b) = \eta(a, b)$,
-$\mathsf{out}(q', a, b) = \eta(a, b)$.
-
-\noindent 
-The \SWT $T$ computes on pairs of words $\< s, t> \in \Sigma^+ \times \Delta^+$,
-$s$ and $t$, being respectively called input and output word.
-%
-% \noindent
-More precisely, the symbolic-weighted transducer~$T$ defines a mapping 
-from the pairs of strings of $\Sigma^+ \times \Delta^+$ 
-into~$\Semiring$,
-based on the following intermediate function $\weight_T$
-defined recursively for every $q, q' \in Q$, 
-%$a \in \Sigma$, $b \in \Delta$ 
-%for every strings of $s \in \Sigma^*$, $t \in \Delta^*$:
-$a \in \Sigma$, $u \in \Sigma^*$, $b\in \Delta$, $v\in \Delta^*$, 
-\[
-\begin{array}{lccl}
-%\weight_T(q, \epsilon, \epsilon, q)  & = & \multicolumn{2}{l}{\one}\\ % or \final(q)
-%\weight_T(q, \epsilon, \epsilon, q') & = & \multicolumn{2}{l}{\zero \mathrm{~if~} q \neq q'}\\
-\weight_T(q, au, bv, q') & = & 
-   \displaystyle\bigoplus_{r \in Q} &
-   \wei_{00}(q, a, b, r) \otimes \weight_T(r, au, bv, q')\\
-  & \oplus & \displaystyle\bigoplus_{r \in Q, u \neq \epsilon} &
-    \wei_{10}(q, a, b, r) \otimes \weight_T(r, u, bv, q')\\
-  & \oplus & \displaystyle\bigoplus_{r \in Q, v \neq \epsilon} &
-    \wei_{01}(q, a, b, r) \otimes \weight_T(r, au, v, q')\\
-  & \oplus & \displaystyle\bigoplus_{r \in Q, u,v \neq \epsilon} &
-    \wei_{11}(q, a, b, r) \otimes \weight_T(r, u, v, q')\\
-  & \oplus & \displaystyle\bigoplus_{u,v = \epsilon} &
-    \wei_{11}(q, a, b, q')
-\end{array}
-\]
-%
-We recall that, by convention (Section~\ref{sec:semiring}), 
-an empty sum with $\bigoplus$ is equal to~$\zero$. 
-Since $\zero$ is absorbing for~$\otimes$ in~$\Semiring$,
-one term $\wei_{ij}(q, a, b, q'')$ equal to $\zero$ in the above expression 
-will be ignored in the sum, meaning that there is no possible transition
-from state $q$ into state $q'$ while reading $a$ and $b$.
-%(including the case $a = \epsilon$ and $b = \epsilon$).
-%These functions $\phi$ act as guards for the transducer's transitions, 
-%preventing a transition when they return the absorbing $\zero$ of $\Semiring$.
-This is analogous to the case of a transition's guard not satisfied by $\<a, b>$ for 
-symbolic transducers.
-% 
-The above sum may be infinite because of the transitions $\wei_{00}$. % in the first and second lines.
-In this case, the infinite sum operators applies since $\Semiring$ is assumed complete.
-%As we shall see, under conditions, the $\epsilon$-transitions can be removed and hence the sum made finite.
-%** SUPPR. $\epsilon$-transition (with $\wei_{00}$) first. **
-%
-%*** rem. there is no case for $a, \epsilon$ or $\epsilon, b$.
-%   It can be simulated with $\wei_{10}$ and $\wei_{01}$ on the last letter. ***
-
-%** The application of $\oplus$ can be seen as a ND choice between transitions **
-% choose to advance head in input, in output or both (or none)
-Intuitively, using a transition $\wei_{ij}(q, a, b, q')$ means for $T$:
-when reading respectively $a$ and $b$ at the current positions in the input and output words, 
-increment the current position in the input word if if $i = 1$ (otherwise, do not change it),
-increment the current position in the output word if if $j = 1$, 
-and change state from $q$ to $q'$.
-%
-In contrast with the models of weighted transducers over finite alphabets~\cite{Mohri03ijfcs},
-the input and output symbols at current positions are read by every transitions, 
-even transitions that do not increment the head's position.
-%This is in contrast with the models of weighted transducers over finite alphabets~\cite{Mohri03ijfcs},
-%which ignore the symbol at current position in this case.
-This is an important feature in the case of an infinite alphabet, 
-since symbols cannot be stored in the finite memory of the transducer in this case.
-%whereas considering the current symbol may be useful to compute a transition weight.
-%(even when it does not change the head's position, like with $\epsilon$-transitions).
-%
-%The cases $\weight_T(q, au, \epsilon, q')$ and $\weight_T(q, \epsilon, bv, q')$
-%are missing in the definition of $\weight_T$.
-%It means that $T$ must avoid configurations where it reached the end of 
-%the output word and not of the input one, or vice-versa.
-%This can be done by using $\wei_{10}$ and $\wei_{01}$
-%before reaching the end of word, and using a special state for this purpose.
-
-The expression of $\weight_T$ can be seen as a stateful definition of 
-an edit-distance between a word $s \in \Sigma^*$ and a word $t \in \Delta^*$,
-see also~\cite{Mohri03ijfcs}.
-Intuitively, 
-$\wei_{10}(q, a, b, r)$ is the cost of 
-the deletion of the symbol~$a \in \Sigma$ in~$s$, 
-$\wei{01}(q, a, b, r)$ is the cost 
-of the insertion of~$b \in \Delta$ in $t$, 
-and $\wei_{11}(q, a, b, r)$ is the cost 
-of the substitution of  $a \in \Sigma$ by~$b \in \Delta$.
-%
-The cost of a sequence of such operations transforming $s$ into $t$, 
-is the product, with $\otimes$, of the individual costs of the operations involved;
-And the distance between $s$ and $t$ is the sum, with $\oplus$,
-of all such product of costs.
-
-\medskip\noindent
-Let $\< s, t> \in \Sigma^+ \times \Delta^+$, with $s = s_0\ldots s_m$, $t = t_0\ldots t_n$). 
-The weight associated by $T$ to $\< s, t>$
-is defined as follows:
-\begin{equation}
-T(s, t)  = 
-\displaystyle\bigoplus_{q, q' \in Q} \mathsf{in}(q, s_0, t_0) 
-\mathop{\otimes} \weight_T(q, s, t, q') \mathop{\otimes} \mathsf{out}(q', s_m, t_n)
-\label{eq:SWT-value}
-\end{equation}
-
-\begin{example}
-We can define with a \SWT the computation of a similarity measure 
-between timed sequences
-similar to dynamic time warping (DTW).
-
-Let $\Semiring$  be the tropical (\emph{min-plus}) semiring of Figure~\ref{fig:semirings} and 
-let $\Sigma = \Delta = \mathbb{R}_+$ be sets of timestamps.
-We consider a \SWT with one state $q$ and transitions
-$\wei_{11}(q, d, d', q) = 
- \wei_{01}(q, d, d', q) = 
- \wei_{10}(q, d, d', q) = |d' - d|$,
-and $\wei_{00}(q, d, d', q) = \zero$, for all $d, d' \in \mathbb{R}_+$.
-% needs reading input/output symbols even by epsilon-transitions.
-The recursive definition of $\weight_T$ correspond to the dynamic programming equations of DTW
-for the computation of an optimal match between words, 
-the matching cost for two symbols being the 
-the time distance between them.
-\endex
-\end{example}
-
-\begin{example}
-% simple pointwise distance between two sequences of timestamped events **
-Let us consider the tropical (\emph{min-plus}) semiring~$\Semiring$ 
-of Figure~\ref{fig:semirings} and 
-let $\Sigma = \mathbb{R}_+$ be an input alphabet of event dates
-and $\Delta = \{ \mathsf{e}, \mathsf{-} \} \times \mathbb{R}_+$ 
-be an output alphabet of symbols with timestamps. 
-A symbol $\< \mathsf{e}, d > \in \Delta$ represents an event starting at date $d$, 
-and $\< \mathsf{-}, d >$ is a continuation of the previous event.
-%More precisely, we let 
-%$\Sigma = \{ a, - \} \times \R_+$ and 
-This example of $\Delta$ is motivated by the case of music notation, 
-where several notated events (notes) can be tied together, 
-with a \emph{tie} or a \emph{dot}
-(like in $\musQuarter\!\!\mathrel{\raisebox{-1.5mm}{$\smile$}}\!\!\musEighth$ 
- or equivalently~\musQuarterDotted),
-meaning that they will be played as a unique sounding event.
-%The timestamp of $a \in \Sigma$, denoted by $\mathsf{t}(a)$, is expressed as a rational number.
-
-We consider a \SWT with two states $q_0$ and $q_1$ whose purpose 
-is to compare a recorded performance $s \in \Sigma^*$
-with notated music sheet $t \in \Delta^*$.
-One timestamp $d_i \in \Sigma$ may corresponds 
-to one notated event $\<\mathsf{e}, d'_i> \in \Sigma$, in which case 
-the weight value computed by the \SWT is the time distance between both
-(see transitions $\wei_{11}$ below).
-%
-Moreover, if $\<\mathsf{e}, d'_i>$ is followed by continuations 
-$\<\mathsf{-}, d'_{i+1}>$..., they are just skip with no cost (transitions $\wei_{01}$ or weight $\one$).
-%transitions
-\[
-\begin{array}{rclcrcl}
-\wei_{11}(q_0, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d| & \quad &
-\wei_{11}(q_1, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d|\\
-\wei_{01}(q_0, d, \< \mathsf{-}, d'>, q_0) & = & \one & &
-\wei_{01}(q_1, d, \< \mathsf{-}, d'>, q_0) & = & \one\\
-\wei_{10}(q_0, d, b, q_1) & = & \alpha & & \multicolumn{3}{l}{\mathrm{for~all~} b \in \Delta}
-\end{array}
-\]
-%
-Moreover, it may happen that the performers plays an extra note accidentally, but only once in a row. 
-This is modelled by the transition $\wei_{10}$ with an arbitrary weight value $\alpha \in \Semiring$, 
-switching from state $q_0$ (normal) to $q_1$ (error).
-The transitions in the second column below switch back to the normal state $q_0$.
-% the metric computed by the \SWT is the smallest sum of point wise distances 
-% between dates of input and output events.
-At last, we let $q_0$ be the only initial and final state, with
-$\init(q_0) = \one$, $\init(q_1) = \zero$, 
-$\final(q_0) = \one$, $\final(q_1) = \zero$, 
-\endex
-\end{example}
-
-\noindent
-The \emph{Symbolic Weighted Automata} %$A = \< Q, \init, \weight, \final >$
-%over $\Sigma$,  $\Semiring$ and $\bar\Phi$
-are defined similarly as the transducers of Definition~\ref{def:SWT}, 
-by simply omitting the output symbols.
-%
-In this case, the label theory $\bar\Phi$ can be reduced to a singleton $\< \Phi_\Sigma>$.
-%over $\Sigma$ is reduced to
-%a set $\Phi_\Sigma$ closed under~$\oplus$ and~$\otimes$.
-%
-\begin{definition} \label{def:SWA}
-A \emph{symbolic-weighted automaton} (\SWA)
-over $\Sigma$, $\Semiring$ and $\bar\Phi$
-%the input alphabet~$\Sigma$ and the commutative semiring $\Semiring$ 
-is a tuple
-$A = \< Q, \init, \bar{\wei}, \final >$,
-where $Q$ is a finite set of states, 
-$\mathsf{in} : Q \to \Phi_\Sigma$, 
-respectively $\mathsf{out} : Q \to \Phi_\Sigma,$
-are functions defining the weight for entering, 
-respectively leaving, a state, 
-and $\bar{\wei}$ is a pair of transition functions 
-$\wei_{0}$, $\wei_{1}$
-%$\wei_{i, j}: Q \times Q \to \Phi_{\Sigma, \Delta}$, with $i, j \in \{ 0, 1\}$.
-from $Q \times Q$ into $\Phi_{\Sigma}$.
-\end{definition}
-A pair of states $\< q, q'>$ such that $\wei_{0}(q, q') \neq \zero$
-is called an \emph{$\epsilon$-transition}.
-
-\noindent
-As above in the case of \SWT, 
-the transition functions $\wei_0$ and $\wei_1$
-can be synthesized into a single function
-$\wei: Q \times \Sigma \times Q \to \Semiring$, 
-and the computation of $A$ on words $s \in \Sigma^+$
-is defined with an intermediate function $\weight_A$, 
-%$a \in \Sigma$, $b \in \Delta$ 
-%for every strings of $s \in \Sigma^*$, $t \in \Delta^*$:
-defined as follows for $q, q' \in Q$, $a \in \Sigma$, $u \in \Sigma^*$,
-\[
-\begin{array}{lccl}
-%\weight_T(q, \epsilon, \epsilon, q)  & = & \multicolumn{2}{l}{\one}\\ % or \final(q)
-%\weight_T(q, \epsilon, \epsilon, q') & = & \multicolumn{2}{l}{\zero \mathrm{~if~} q \neq q'}\\
-\weight_A(q, au, q') & = & 
-   \displaystyle\bigoplus_{r \in Q} &
-   \wei_{0}(q, a, r) \otimes \weight_A(r, au, q')\\
-  & \oplus & \displaystyle\bigoplus_{r \in Q, u \neq \epsilon} &
-    \wei_{1}(q, a, r) \otimes \weight_A(r, u, q')\\
-  & \oplus & \displaystyle\bigoplus_{u = \epsilon} & 
-    \wei_{1}(q, a, q')
-\end{array}
-\]
-%
-\noindent
-and the weight value associated by $A$ to 
-$s = s_0\ldots s_m \in \Sigma^+$ is: % defined as follows:
-\begin{equation}
-A(s)  = 
-\displaystyle\bigoplus_{q, q' \in Q} \mathsf{in}(q, s_0) 
-\mathop{\otimes} \weight_A(q, s, q') \mathop{\otimes} \mathsf{out}(q', s_m)
-\label{eq:SWA-value}
-\end{equation}
-
-%
-%When $\wei_\epsilon(q, q') = \zero$ for all $q, q' \in Q$, 
-%the automaton~$A$ is called \emph{without $\epsilon$-transitions}.
-
-%The \emph{summary} of a $\SWT$, resp. a $\SWA$, is ***
-
-\subsection{Properties}
-The two following properties will be useful to our approach on 
-symbolic weighted parsing in Section~\ref{sec:parsing}.
-
-\begin{proposition} \label{prop:epsilon}
-Given a \SWT $T$ over $\Sigma$, $\Delta$, $\Semiring$ and $\bar\Phi$,
-and $s \in \Sigma^*$, 
-there exists an effectively constructible \SWA 
-$A_{T, s}$ over $\Delta$, $\Semiring$ and $\bar\Phi$,
-such that for all $t \in \Delta^*$, $A_{T, s}(t) = T(s, t)$.
-\end{proposition}
-%
-\begin{proof}
-Let $T = \< Q, \init, \bar{\wei}, \final >$,
-let 
-$\wei: Q \times \Sigma \times \Delta \times Q \to \Semiring$
-be the synthesized form of 
-$\bar{\wei} = \< \wei_{00}, \wei_{10}, \wei_{01}, \wei_{11}>$,
-and let $s = s_1 \ldots s_n$ with $s_1, \ldots, s_n \in \Sigma$.
-%$n = |s|$ be the length of $s$.
-%
-The state set of $A_{T, s}$ will be $Q' = [1..n] \times Q$.
-Its state entering weight function is defined by 
-$\init'(1, q) = \init(q)$ for all $q \in Q$, 
-and $\init'(i, q) = \zero$ for all $1 < i \leq n$, 
-and its state leaving weight function by 
-$\final'(n, q) = \final(q)$ or all $q \in Q$, 
-and $\final'(i, q) = \zero$ for all $1 \leq i <  n$.
-
-\noindent
-We define transition functions $\wei'_0$ and $\wei'_1$ for $A_{T, s}$, 
-together with their synthesized form $\wei'$ as follows: 
-for all $1 \leq i, j \leq n$, $q, q' \in Q$, %and $a \in \Sigma$, 
-and $b \in \Delta$, 
-\[
-\begin{array}{lclcl}
-\wei'_0\bigl(\< i, q>, \< i, q'>\bigr) & = & \wei_{00}(q, q')\\ 
-%
-\wei'_0\bigl(\< i, q>, \< i+1, q'>\bigr) & = & \phi(s_i) & 
-\mathrm{if} & i < n, \mathrm{where~} \phi = \wei_{10}(q, q'),\\
-%
-\wei'_0\bigl(\< i, q>, \< j, q'>\bigr) & = & \zero & 
-\mathrm{if} & j \neq i, i+1,\\ 
-%
-\wei'_1\bigl(\< i, q>, \< i, q'>\bigr) & = & \wei_{01}(q, q')\\
-%
-\wei'_1\bigl(\< i, q>, \< i+1, q'>\bigr) & = & \eta_{s_i} &
-\mathrm{if} & i < n, \mathrm{where~} \eta = \wei_{11}(q, q'),\\
-%
-\wei'_1\bigl(\< i, q>, \< j, q'>\bigr) & = & \zero & 
-\mathrm{if} & j \neq i, i+1.
-\end{array}      
-\]
-We use the assumption that the underlying label theory of $T$ is 
-is closed under partial applications (Section~\ref{section:symbols})
-in the above construction of $\wei'_1$.
-
-\noindent
-The $\SWA$ wanted is $A_{T, s} = \< Q', \init', {\bar\wei}', \final' >$.
-\qed
-\end{proof}
-%
-\noindent
-The construction time and size for $A_{T, s}$ are $O(\| T \| . | s |)$,
-where the size $\| T \|$ of $T$ is its number of states $|Q|$.
-
 % gen. to transducers
@@
 \]
-Moreover, we can show that
+Moreover, %we can show that
 \[
@@
 ... essentially the same as for \WA~\cite{Mohri02ijfcs}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SW-VPA
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{SW Visibly Pushdown Automata}
-\label{section:SWVPA}\label{sec:SWVPA}
-The following model generalizes Symbolic VPA~\cite{dAntonyAlur14SVPDA}
-from Boolean semirings to arbitrary semiring weight domains.
-
-\subsection{Definition}
-Let $\Omega$ be a countable alphabet 
-%finite (large) or infinite,
-that we assume partitioned into three subsets 
-$\Omegai$, $\Omegac$, $\Omegac$,
-whose elements are respectively called 
-\emph{internal}, \emph{call} and \emph{return} symbols.
-% \begin{itemize}
-% \item a set $\Omegai$ of \emph{internal symbols} denoted $a$,
-% \item a set $\Omegac$ of \emph{call symbols} denoted $\call{a}$,
-% \item a set $\Omegar$ of \emph{return symbols} denoted $\return{a}$.
-% \end{itemize}
-Let~$\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative and complete semiring and let  
-$\bar\Phi = \< \Phii, \Phic, \Phir, \Phici, \Phicc, \Phicr>$ 
-be a label theory over $\Semiring$
-%In order to simplify notations, %and following the definition of Section~\ref{section:transducer}, 
-%we shall write respectively 
-where $\Phii$, $\Phic$, $\Phir$ and~$\Phicx$ (with $\mathsf{x} \in \{ \mathsf{i}, \mathsf{c}, \mathsf{r}\}$) 
-stand respectively 
-for~$\Phi_\Omegai$, $\Phi_\Omegac$, $\Phi_\Omegar$ and~$\Phi_{\Omegac, \Omega_\mathsf{x}}$.
-%
-%Moreover, we extend this theory with a set $\Phii$ 
-%of unary functions in $\Omegai \to \Semiring$,
-%closed under $\oplus$ and $\otimes$.
-
-\begin{definition}
-A \emph{Symbolic Weighted Visibly Pushdown Automata} (\SWVPA) 
-over  $\Omega = \Omegai \uplus \Omegac \uplus \Omegar$, $\Semiring$ and $\bar\Phi$ 
-is a tuple $A = \< Q, P, \init, \weii, \weic, \weir, \weie, \final >$,
-where $Q$ is a finite set of states, 
-$P$ is a finite set of stack symbols, 
-$\mathsf{in} : Q \to \Semiring$, 
-respectively $\mathsf{out} : Q \to \Semiring,$
-are functions defining the weight for entering, 
-respectively leaving, a state, 
-and 
-$\weii : Q \times P \times Q \to \Phici$,  
-$\weic : Q \times P \times Q \times P \to \Phicc$,  
-$\weir : Q \times P \times Q \to \Phicr$,  
-$\weie : Q \times Q \to \Phir$,  
-are transition functions.
-\end{definition}
-%
-Similarly as in Section~\ref{section:transducer}, 
-we extend the above transition functions as follows
-for all $q, q' \in Q$, $p \in P$, 
-$a \in \Omegai$, 
-$\call{c} \in \Omegac$, 
-$\return{r} \in \Omegar$, 
-overloading their names: % for simplicity:
-\[
-\begin{array}{lll}
-\weii: Q \times \Omegac \times P \times \Omegai \times Q \to \Semiring & 
-\weii(q, c, p, a, q') = \eta_\mathsf{ci}(c, a) & 
-\mathrm{where~} \eta_\mathsf{ci} = \weii(q, p, q'),\\
-%
-\weic: Q \times \Omegac \times P \times  \Omegac \times Q \times P \to \Semiring & 
-\weic(q, c, p, \call{c'}, q', p') = \eta_\mathsf{cc}(c, \call{c'}) & 
-\mathrm{where~} \eta_\mathsf{cc} = \weic(q, p, q', p'),\\
-%
-\weir: Q \times \Omegac \times P \times \Omegar \times Q \to \Semiring & 
-\weir(q, {\call{c}},  p, {\return{r}}, q') = \eta_\mathsf{cr}({\call{c}},  {\return{r}}) & 
-\mathrm{where~} \eta_\mathsf{cr} = \weir(q, p, q'),\\
-%
-\weie: Q \times \Omegar \times Q \to \Semiring & 
-\weie(q, {\return{r}}, q') = \phi_\mathsf{e}({\return{r}}) &
-\mathrm{where~} \phi_\mathsf{e} = \weie(q, q').\\
-\end{array}      
-\]
-
-\noindent
-The intuition is the following for the above transitions.
-\begin{description}
-\item $\weii$ : read the input internal symbol $a$, change state to $q'$ (stack is untouched).
-\item $\weic$ : read the input call symbol $\call{c}$, push it to the stack along with $p$, change state to $q'$.
-\item $\weir$ : when the stack is not empty, 
-      read and pop from stack a pair made of $\call{c}$ and $p$, 
-      read the input return symbol $\return{r}$, change state to $q'$.
-      In this case, the weight function $\phi_\mathsf{r}$ 
-      computes a value of matching between the call and return symbols $c$ and $r$.
-      This value might be set to $\zero$ in order to express that the symbols do not match.
-\item $\weie$ : when the stack is empty, 
-      read the input symbol $\call{r}$, change state to $q'$.
-\end{description}
-
-We give now a formal definition of these transitions of the automaton~$A$
-in term of %a weight value computed by 
-an intermediate function $\weight_A$, like in Section~\ref{sec:SWT}.
-In the case of a pushdown automaton, a configuration is composed 
-of a state $q \in Q$ and a stack content $\gamma \in \Gamma^*$, where $\Gamma = \Omegac \times P$.
-Therefore, $\weight_A$ is a function from 
-$Q \times \Gamma^* \times \Omega^* \times Q \times \Gamma^*$ into $\Semiring$.
-\[
-\begin{array}{rcl}
-\weight_A\bigl(\config{q}{\gamma}, a\, u, \config{q'}{\gamma'}\bigr) & = & 
- {\displaystyle\bigoplus_{q'' \in Q}} \weii(q, a, q'') 
- \otimes \weight_A\bigl(\config{q''}{\gamma}, u, \config{q'}{\gamma'}\bigr)\\
-%
-\weight_A\bigl(\config{q}{\gamma}, {\call{c}}\, u, \config{q'}{\gamma'}\bigr) & = & 
- {\displaystyle\bigoplus_{\begin{array}{c}
-                          \scriptstyle q'' \in Q\\[-2pt]
-                          \scriptstyle p \in P
-                          \end{array}}}
- \weic\bigl(q, {\call{c}}, q'', p\bigr) 
- \otimes \weight_A\bigl(\config{q''}{{\call{c}}\, p\cdot \gamma}, u, \config{q'}{\gamma'}\bigr)\\[1mm]
-%
-\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot \gamma}, {\return{r}}\, u, \config{q'}{\gamma'}\bigr) & = & 
- {\displaystyle\bigoplus_{q'' \in Q}} 
- \weir\bigl(q, {\call{c}}, p, {\return{r}}, q''\bigr) 
- \otimes \weight_A\bigl(\config{q''}{\gamma}, u, \config{q'}{\gamma'}\bigr)\\
-%
-\weight_A\bigl(\config{q}{\bot}, {\return{r}}\, u, \config{q'}{\gamma'}\bigr) & = & 
- {\displaystyle\bigoplus_{q'' \in Q}} \weie(q, {\return{r}}, q'') 
- \otimes \weight_A\bigl(\config{q''}{\bot}, u, \config{q'}{\gamma'}\bigr)\\
-\end{array}
-\]
-where $\bot$ denotes the empty stack and ${\call{c}}\, p\cdot \gamma$ 
-denotes a stack where the pair made of ${\call{c}} \in \Omegac$ and $p \in P$ is the top symbol 
-and $\gamma$ is the rest of stack.
-
-\noindent
-The weight associated by $A$ to $s \in \Omega^*$
-is then defined as follows,
-following empty stack semantics: 
-\begin{equation}
-A(s)  = 
-{\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
-\mathsf{in}(q) \mathop{\otimes} 
-\weight_A\bigl(\config{q}{\bot}, s, \config{q'}{\bot}\bigr) 
-\mathop{\otimes} \mathsf{out}(q').
-\label{eq:weightA}
-\end{equation}
-
-\begin{example}
-structured words...
-intro language of music notation ?
-\end{example}
-
-\subsection{Properties}
-Like VPA and symbolic VPA, 
-the class of \SWVPA is closed under the binary operators of the underlying semiring.
-%
-\begin{proposition}
-Let $A_1$ and $A_2$ be two (\SWVPA)
-over the same $\Omega$ and $\Semiring$.
-There exists two $\SWVPA$ $A_1 \oplus A_2$ and $A_1 \otimes A_2$, 
-effectively constructible, 
-such that for all $s \in \Omega^*$, 
-$(A_1 \oplus A_2)(s) = A_1(s) \oplus A_2(s)$ and 
-$(A_1 \otimes A_2)(s) = A_1(s) \otimes A_2(s)$.
-\end{proposition}
-The construction is essentially the same 
-as in the case of the Boolean semiring~\cite{dAntonyAlur14SVPDA}.
-
-\subsection{Best-first Search} 
-\label{sec:best}\label{sec:search}
-**hypotheses**
-Let us assume that the semiring~$\Semiring$ is
-commutative, bounded, complete, and total.
-and assume an effective label theory.
-%s.t. for all $\phi \in \Phii$, 
-%$\psi \in \Phir$, 
-%and $\eta \in \Phicr$, 
-%$\displaystyle\bigoplus_{a \in \Omegai} \phi(a)$ 
-%$\displaystyle\bigoplus_{r \in \Omegar} \phi(r)$ and 
-%$\displaystyle\bigoplus_{{\call{c}} \in \Omegac} 
-%\displaystyle\bigoplus_{{\return{r}} \in \Omegar} \eta({\call{c}}, {\return{r}})$
-%are computable...
-
-% total ? 
-% monotonic and superior writ natural ordering
-%Regarding the infinite sum operator, note that
-%$\bigoplus_{x \in \Phi_\Omega} \phi(x)$, 
-%$\bigoplus_{y \in \Phi_\Delta} \psi(y)$, and  
-%... exist and in $\Semiring$.
-
-We propose a Dijkstra algorithm computing the minimal weight by $A$, 
-\wrt~$\leq_\oplus$, for a word in~$\Omega^*$.
-
-\noindent
-More precisely, 
-let $\top$ be a fresh stack symbol which does not belong to $\Gamma$,
-and for every two states $q, q' \in Q$ and $\sigma \in \{ \bot, \top \}$, 
-let 
-\[ 
-  d_0(q, \sigma, q') = \bigoplus_{s\in \Omega^*} 
-  \textstyle
-  \weight_A\bigl(\config{q}{\sigma}, s, \config{q'}{\sigma}\bigr). 
-\]
-Since $\Semiring$ is complete, this infinite sum is well defined, 
-and since $\leq_\oplus$ is assumed total, it is the minimum in $\Omega^*$
-of $s \mapsto \weight_A(\config{q}{\sigma}, s, \config{q'}{\sigma})$ \wrt this ordering.
-%
-When $\sigma = \bot$, $d_0(q, \sigma, q')$ is the central 
-expression in a term of the definition (\ref{eq:weightA}) of $A(s_0)$ for the minimum $s_0$
-(for the above function).
-When $\sigma = \bot$, intuitively, it is the minimum weight of a computation of $A$
-starting in state $q$ with a stack $\gamma \in \Gamma^*$ (possibly empty), 
-and ending in state $q$ with the same stack $\gamma$, such that moreover 
-the computation does not touch a symbol of $\gamma$. 
-That means that during the computation, 
-$A$ may apply the first case of in the definition of $\weight_A$ (internal symbol),
-as well as the second case, to can push call symbols on the top of $\gamma$, 
-and may pop these symbols with the third case (return symbol). 
-However, it cannot apply one of the two last cases (return symbol and empty stack)
-when the current stack is $\gamma$.
-%pop symbols in $\gamma$.
-% Note that having a stack reduced to such a symbol makes impossible the application of the 
-% two last cases in the definition of $\weight_A$ (return symbol and empty stack). 
-% However, it is possible to apply the two first cases 
-% (internal symbol or call symbol, with a push on the top of $\top$).
-
-The algorithm~\ref{algo:Dijkstra}
-constructs iteratively a marking $d : Q \times \{ \bot, \top \} \times Q \to \Semiring$
-%of the triplets $\<q, \sigma, q'>$ 
-%of states of $A$ by weight values in $\Semiring$, 
-that converges eventually to $d_0(q, \sigma, q')$. 
-%It uses for that purpose a priority queue $P$ containing triplets of 
-%$Q \times \{ \bot, \top \} \times Q$.
-
-\begin{algo}[1-best for \SWVPA] \label{algo:Dijkstra}
-\textbf{initially} let $P = Q \times \{ \bot, \top \} \times Q$, %contains all %triplets 
-%$\< q_1, \bot, q_2>$ and $\<q_1, \top, q_2>$ for $q_1, q_2 \in Q$, with 
-and 
-$d(q_1, \bot, q_2) = d(q_1, \top, q_2) = \one$ if $q_1 = q_2$ and
-$d(q_1, \bot, q_2) = d(q_1, \top, q_2) = \zero$ otherwise.
-
-\smallskip\noindent
-\textbf{while} $P$ is not empty
-
-\noindent
-\quad\textbf{extract} $\< q_1, \sigma, q_2>$ from $P$ 
- such that $d(q_1, \sigma, q_2)$ is minimal wrt $\leq_\oplus$.
-
-\noindent\quad \textbf{for all} $q_0, q_3 \in Q$ and $p \in P$ do
-
-\noindent
-\qquad
-\(\begin{array}{rrcl}
- & d(q_1, \sigma, q_3) & \opluseq &
-   d(q_1, \sigma, q_2) \otimes \displaystyle\bigoplus_{a \in \Omegai} \weii(q_2, a, q_3)\\
-%
-%  & d(q_0, \sigma, q_2) & \opluseq & 
-%    \displaystyle\bigoplus_{a \in \Omegai} \weii(q_0, a, q_1) \otimes d(q_1, \sigma, q_2)\\
-%
-\mathrm{if~} \sigma = \top & d(q_0, \top, q_3) & \opluseq &
-d(q_1, \sigma, q_2) \otimes
-%\displaystyle\bigoplus_{p \in P} 
-\displaystyle\bigoplus_{{\call{c}} \in \Omegac} 
-\displaystyle\bigoplus_{{\return{r}} \in \Omegar} \eta({\call{c}}, {\return{r}})\\
-\mathrm{and} &  d(q_0, \bot, q_3) & \opluseq &
-d(q_1, \sigma, q_2) \otimes
-%\displaystyle\bigoplus_{p \in P} 
-\displaystyle\bigoplus_{{\call{c}} \in \Omegac} 
-\displaystyle\bigoplus_{{\return{r}} \in \Omegar} \eta({\call{c}}, {\return{r}})\\
-& & & \mathrm{where~} \eta = \weic(q_0, q_1, p) \otimes \weir(q_2, p, q_3)\\
-%
-\mathrm{if~} \sigma = \bot & d(q_1, \bot, q_3) & \opluseq &
-  d(q_1, \sigma, q_2) \otimes \displaystyle\bigoplus_{\return{r} \in \Omegar} \weie(q_2, {\return{r}}, q_3)\\
-%
-%& d(q_0, \bot, q_2) & \opluseq & 
-%d(q_0, \bot, q_1) \otimes d(q_1, \sigma, q_2), \mathrm{if~} \< q_0, \bot, q_1> \notin P \\
-%
-%\mathrm{if~} \sigma = \top & d(q_0, \top, q_2) & \opluseq & 
-%d(q_0, \top, q_1) \otimes d(q_1, \sigma, q_2), \mathrm{if~} \< q_0, \top, q_1> \notin P \\
-%
-& d(q_1, \bot, q_3) & \opluseq & 
-   d(q_1, \sigma, q_2) \otimes d(q_2, \bot, q_3), \mathrm{if~} \< q_2, \bot, q_3> \notin P \\
-%
-\mathrm{if~} \sigma = \top & d(q_1, \top, q_3) & \opluseq & 
-    d(q_1, \sigma, q_2) \otimes d(q_2, \top, q_3), \mathrm{if~} \< q_2, \top, q_3> \notin P \\
-\end{array}
-\) 
-\end{algo}
-
-\noindent
-The infinite sums in the updates of $d$ in Algorithm~\ref{algo:Dijkstra} are well defined
-since~$\Semiring$ is complete.
-The algorithm performs $2.|Q|^2$ iterations until $P$ is empty, 
-and each iteration has a time complexity $O(|Q|^2 . |P|)$.
-This gives a time complexity $O(|Q|^4 . |P|)$. 
-It can be reduced by implementing $P$ as a priority queue, 
-prioritized by the value returned by $d$
-***complete***. %$|Q|^3.\log(|Q|^2)$
-
-The correctness of Algorithm~\ref{algo:Dijkstra} 
-is ensured by the invariant expressed in the following lemma.
-\begin{lemma}
-For all $\< q_1, \sigma, q_2> \notin P$, $d(q_1, \sigma, q_2) =  d_0(q_1, \sigma, q_2)$.
-\end{lemma}
-The proof is by contradiction, 
-assuming a counter-example minimal in the length of the witness word.
-
-% \begin{algorithm}
-% \caption{<your caption for this algorithm>}
-% \label{<your label for references later in your document>}
-% \begin{program}
-% \seq{|line| := `` '', i := 1};
-% \WHILE i \neq n+1 \DO
-%       |line| := |item|[i] \concat `` '' \concat |number|[i]; 
-%       i := i+1; 
-%       \WHILE i \neq n+1 \AND |item|[i] = |item|[i-1] \DO 
-%       |line| := |line| \concat ``, '' \concat |number|[i]);
-%       i := i+1 \OD ; 
-%       |write|(|line|) \OD 
-% \end{program}
-% \end{algorithm}
-
-% The weight of a transition acts as a guard: 
-% a transition is activated for a symbol $`a`$ iff its weight $`\phi(a) \neq 0`$ (or $`\psi(a, b) \neq 0`$), 
-% the absorbing element.
-
-%\begin{lemma}
-\noindent
-For computing the minimal weight of a computation of $A$, we use the fact that,
-at the termination of Algorithm~\ref{algo:Dijkstra}, %it holds that,
-%There exist $q_1, q_2 \in Q$
-\[
-  {\displaystyle \bigoplus_{s \in \Omega^*} A(s)} = 
-  {\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
-  \mathsf{in}(q) \mathop{\otimes} d(q, \bot, q') \mathop{\otimes} \mathsf{out}(q').
-\]
-%\end{lemma}
-
-%\medskip
-\noindent
-In order to obtain effectively a witness 
-(word of $\Semiring^*$ with computation of $A$ of minimal weight), 
-we require an additional property the of weight functions.
-%
-\begin{definition}
-Let $\Omega$ be an alphabet and $\Semiring$ a complete semiring.
-A function $\phi$ from $\Omega^n$ into $\Semiring$
-is called $k$-\emph{convex} for a natural number $k$ iff 
-$\mathit{card}\{ \mathbf{a} \in \Omega^n \mid \phi(\mathbf{a}) = 
-                 \displaystyle\bigoplus_{\mathbf{p} \in \Omega^n} \phi(\mathbf{p}) \} \leq k$.
-\end{definition}
-A label theory is $k$-convex if all its functions are $k$-convex.
-
-\begin{proposition}
-For a \SWVPA $A$ over a 
-commutative, idempotent, superior, total and complete semiring 
-and an alphabet $\Omega$ with a $k$-convex label theory,
-one can construct in PTIME a word $s \in \Omega^*$ 
-such that $A(s)$ is minimal \wrt the natural ordering for $\Semiring$. 
-%$\leg_\oplus$.
-\end{proposition}
-
-\subsection{Trees and Nested-Words}
-\label{sec:trees}
-The hierarchical structure of nested-words, defined the \emph{call} and \emph{return} markup symbols of 
-suggest a correspondence between these word and trees. 
-The lifting of this correspondence to languages, respective of tree automata and VPA have
-been discussed in~\cite{AlurMadhusudan09nested}, 
-see also~\cite{Caralp12VPAmult} for the weighted case.
-In this section, we discuss the correspondence between the symbolic-weighted extensions
-of tree automata and VPA.
-
-Let $\Omega$ be a countable ranked alphabet, such that 
-every symbol $a \in \Omega$ has a rank 
-$\rank(a) \in [0..M]$ where $M$ is a fixed natural number.
-We write $\Omega_k$ the subset of all symbols $a$ of $\Omega$
-with $\rank(a) = k$, where $0 \leq k \leq M$.
-%$\cup_{i = 0}^{i = M} \Omega_i$ 
-%
-$\T(\Omega)$ denotes the free $\Omega$-algebra of finite, ordered, 
-$\Omega$-labeled trees, 
-which is the smallest set such that  $\Omega_0 \subset \T(\Omega)$
-and for all $1 \leq k \leq M$, all $a \in \Omega_k$, 
-and all $t_1, \ldots, t_k \in \T(\Omega)$, $a(t_1, \ldots, t_k) \in \T(\Omega)$.
-%
-% tree = single node (leave) labeled with a symbol of $a \in \Omegai$
-% (such a tree is simply denoted by $a$)
-% or the composition, denoted by $b(t_1,\ldots, t_n$) of a node labeled with $b$
-% and $n$ subtrees $t_1$,\ldots, $t_n$.
-
-\renewcommand{\call}[1]{\ensuremath \langle_{#1}}
-\renewcommand{\return}[1]{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
-
-Let $\hat\Omega$ be the countable (unranked) alphabet obtained from 
-$\Omega$ as follows: 
-$\hat\Omega = \< \Omegai, \Omegac, \Omegar >$, with
-$\Omegai = \Omega_0$, 
-$\Omegac = \{ \; \call{a} \mid a \in \Omega_{>0} \}$,
-$\Omegac = \{ \; \return{a} \mid a \in \Omega_{>0} \}$,
-where $\Omega_{>0}$ denotes $\bigcup_{k = 1}^{M} \Omega_k$.
-
-\noindent
-We define a linearization of trees of $\T(\Omega)$ into 
-words of $\hat{\Omega}^*$ as follows:
-\begin{description}
-\item $\lin(a) = a$ for all $a \in \Omega_0$, 
-\item $\lin\bigl( b(t_1, \ldots, t_k)\bigr) = 
-       \call{b} \; \lin(t_1) \ldots \lin(t_k) \; \return{b}$ 
-       when $b \in \Omega_k$, $1 \leq k \leq M$.
-\end{description}
-
-% A \emph{regular tree grammar} over $\Omega$ 
-% is a triplet $G = \< N, q_\mathsf{i}, R>$ where
-% $N$ is a finite set of non-terminal symbols denoted $q$..., 
-% $q_\mathsf{i} \in N$ is the starting non-terminal, 
-% $R$ is a finite set of production rules of the form
-% $q_0 \to a(q_1\ldots q_k)$ where 
-% $q_0, q_1, \ldots, q_k \in N$
-% $a \in \Omega_k$.
-% A tree $t \in \T(\Omega)$ is in the language of $G$ 
-% if it can be generated from $q_\mathsf{i}$ by 
-% non terminal replacement following the rules of $R$.
-
-Let us assume a label theory $\Phi_{\Omega_k}$ 
-for each $k \in [0..M]$.
-%
-\begin{definition}  \label{def:SWTA}
-A \emph{symbolic-weighted tree automaton} (\SWTA)
-over the ranked input alphabet~$\Omega$
-and the commutative semiring $\Semiring$ 
-is a triplet $A = \< Q, \init, \bar{\wei} >$ where
-$Q$ is a finite set of states, 
-$\mathsf{in} : Q \to \Semiring$ is the starting weight function, 
-and $\bar{\wei}$ is a $M+2$-uplet of transition functions made of:
-$\wei_\epsilon$ from $Q \times Q$ into $\Semiring$, and, 
-for each $k \in [0..M]$, $\wei_{\Omega, k}$
-from $Q \times Q^{k}$ into $\Phi_{\Omega_k}$.
-\end{definition}
-%
-Like in Section~\ref{sec:SWAdef}, we define from $\bar{\wei}$ 
-a transition function~$\wei$, %from $\bar\wei$
-from 
-$Q \times (\Sigma \cup \{ \epsilon \}) \times \bigcup_{k=0}^{M} Q^k$
-into~$\Semiring$: %also called $\wei$ for simplicity, 
-%such that, for all $q, q' \in Q$, $a \in \Sigma$, and $b \in \Delta$, 
-\[
-\begin{array}{rcll}
-\wei(q_0, \epsilon, q_1) & = &  \wei_\epsilon(q_0, q_1),\\ %\phi_\epsilon\\
-\wei(q_0, a, q_1 \ldots q_k) & = & \phi_{\Omega, k}(a) &
-\quad\mathrm{where~} \phi_{\Omega, k} = \wei_{\Omega, k}(q_0, q_1\ldots q_k).\\
-\end{array}      
-\]
-Intuitively, $\wei(q_0, a, q_1 \ldots q_k)$ can be seen as
-the weight of a production rule $q_0 \to a(q_1, \ldots, q_k)$ 
-of a regular tree grammar~\cite{tata}, 
-that replaces the non-terminal symbol $q_0$, by $a(q_1, \ldots, q_k)$. 
-%in a step of tree building.
-%
-Such a grammar computes the weights of the derivation trees 
-of the Context-Free grammar obtained by forgetting the labeling symbols of $\Omega_{>0}$.
-
-\noindent
-The \SWTA of Definition~\ref{def:SWTA} defines a mapping 
-from trees of $\T(\Omega)$ into the weights of~$\Semiring$,
-based on the intermediate function $\weight_A$
-defined as follows for $q_0 \in Q$ and 
-$t = b(t_1,\ldots, t_k) \in \T(\Omega)$,
-åwith $0 \leq k \leq M$:
-\[
-\begin{array}{rccl}
-\weight_A(q_0, t) & = & 
-   \displaystyle\bigoplus_{q_1 \in Q} &
-   \wei(q, \epsilon, q_1) \otimes \weight_A(q_1, t)\\
- & \oplus & \displaystyle\bigoplus_{q_1 \ldots q_k \in Q^k} &
-              \wei(q_0, b, q_1 \ldots q_k ) 
-   \otimes \displaystyle\bigotimes_{i=1}^{k}
-           \weight_A(q_i, t_i).\\
-\end{array}
-\]
-
-\medskip\noindent
-The weight associated by $A$ to  $t \in \T(\Omega)$ is then
-\begin{equation}
-A(t)  = 
-\displaystyle\bigoplus_{q \in Q} \mathsf{in}(q) \mathop{\otimes} \weight_A(q, t).
-\label{eq:weightTA}
-\end{equation}
-
-\begin{lemma}  \label{lem:SWTA}
-For all \SWTA $A$ over~$\Omega$ and $\Semiring$, without $\epsilon$-transitions,
-there exists an effectively constructible \SWVPA $A'$ over 
-$\hat\Omega$ and $\Semiring$
-such that for all $t \in \T(\Omega)$, $A'\bigl(\lin(t)\bigr) = A(t)$.
-\end{lemma} 
-% 
-\begin{proof}
-Let $A = \< Q, \init, \bar{\wei} >$ where $\bar{\wei}$ is summarized as above by a function
-$\wei: Q \times (\Sigma \cup \{ \epsilon \}) \times \bigcup_{k=0}^{M} Q^k \to \Semiring$.
-
-We build 
-$A' = \< Q', P', \init', \weii, \weic, \weir, \weie, \final' >$,
-computing over $\hat\Omega = \< \Omegai, \Omegac, \Omegar >$,
-%
-%$\Omega = \Omegai \uplus \Omegac \uplus \Omegar$ and $\Semiring$ 
-where $Q' = \bigcup_{k=0}^{M} Q^k$ be  the set of sequences of state symbols of $A$, 
-of length at most $M$, including the empty sequence denoted by $\epsilon$, 
-and where $P' = Q'$.
-
-\[
-\begin{array}{ll}
-%\weii: Q \times \Omegai \times Q \to \Semiring & 
-\weii(\bar{q}, a, \bar{q} q') = \wei(q', a, \epsilon) & 
-\mathrm{for~all~} a \in \Omega_0,\\
-%
-%\weic: Q \times \Omegac \times Q \times P \to \Semiring & 
-\weic(\bar{q}, \call{b}, \epsilon, \bar{q}) = \one & %\wei(p', b, \bar{p}) & 
-\mathrm{for~all~} b \in \Omega_{>0},\\
-%
-%\weir: Q \times \Omegac \times P \times \Sigmar \times Q \to \Semiring & 
-\weir(\bar{q}, {\call{b}}, \bar{p}, {\return{b}}, \bar{p} q') = \wei(q', b, \bar{q}) & 
-\mathrm{for~all~}  b \in \Omega_{>0},\\
-%
-%\weie: Q \times \Omegar \times Q \to \Semiring & 
-\weie(\bar{p}, {\return{b}}, \bar{q}) = \zero &
-\mathrm{for~all~}  b \in \Omega_{>0}.\\
-\end{array}      
-\]
-
-In practice, it is sufficient to consider in $Q'$ only the prefixes of 
-sequences.
-\qed\end{proof}
-
-\section{Symbolic Weighted Parsing}
-\label{sec:parsing}
-Let us now use the models and results of the former sections in %order to define 
-an approach to the problem of parsing over infinite alphabet. %appropriate
-%
-Besides considering infinitely many possible of input symbols, 
-handled with suitable language formalisms, 
-this approach extends conventional parsing 
-by computing a derivation tree modulo 
-a generic distance between words,  
-defined by a SW transducer given in input.
-This enables considering finer word relationships than strict equality 
-as in the conventional parsing approach, 
-opening possibilities of quantitative analysis via this method.
-
-\subsection{Definition}
-
-Let $\Sigma$ be a countable input alphabet 
-and $\Omega = \Omegai \uplus \Omegac \uplus \Omegar$ an output countable alphabet,
-let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative, bounded, complete 
-and total semiring 
-and let  $(\Phi_\epsilon, \Phic, \Phir, \Phicr)$ be a label theory over $\Semiring$, 
-assumed computable and $k$-convex for some fixed $k$.
-
-\noindent
-Assuming given in input:
-\begin{itemize}
-\item a \SWT $T$ over $\Sigma$,  $\Omega$, and $\Semiring$,
-defining a measure %between words 
-$T: \Sigma^* \times \Omega^* \to \Semiring$,
-\item a \SWVPA $A$ over $\Omega$, and $\Semiring$, defining a series of nested words
-      $A : \Omega^* \to \Semiring$,
-\item an input word $s \in \Sigma^*$,
-\end{itemize}
-the problem of \emph{symbolic weighted parsing} is to find 
-a nested word $t \in \Omega$ minimizing
-\( T(s, t) \otimes A(t)\)
-\wrt $\leq_\oplus$, 
-\ie such that %Hence, it is the problem of finding 
-\( T(s, t) \otimes A(t) = 
-  \displaystyle\bigoplus_{t' \in \Omega^*} T(s, t') \otimes A(t') \).
-
-\noindent 
-Therefore, it is the problem of optimizing a measure 
-called the \emph{edit-distance between $s$ and $A$} in~\cite{Mohri03EDWA}.
-%
-The input language can also be expressed as a \SWTA, or, 
-as a particular case, as a weighted context-free grammar, 
-converted in turn into a \SWVPA following Lemma~\ref{lem:SWTA}.
-%
-In the case of finite alphabets, 
-the problem of searching, in a WTA language, 
-for the best parse tree for a given,  
-sometimes referred as~\emph{weighted parsing}
-(see \cite{Goodman99SemiringParsing,MorbitzVogler19weighted-parsing} **more general problems**)
-is a particular case of SW parsing.
-%
-Indeed, it corresponds to the case where $T$
-accepts only the pairs $\<s, t>$ such that 
-$s$ is the projection of $t$ on $\Omegai$. 
-This can be done with a single state $q$ and 
-with transition rules of the form:
-\begin{description}
-\item[] $\wei(q, \epsilon, a, q) = \one$ for all $a \in \Omegac \cup \Omegar$,
-\item[] $\wei(q, a, a, q) = \one$ for all $a \in \Omegai$,
-\item[] $\wei(q, a, b, q) = \zero$ for all $a, b \in \Omegai$, $a \neq b$.
-\end{description}
-
-\subsection{Computation}
-
-\begin{proposition}
-The problem of Symbolic Weighted  parsing 
-can be solved in PTIME in the size of the input \SWT, \SWVPA (or \SWTA) and input word, 
-and the computation time of the functions of the label theory.
-\end{proposition}
-%
-%\begin{proof}
-%Bar-Hillel construction
-%\qed  
-%\end{proof}
-
-%\input{transcription}
-
-\section*{Conclusion}
-% discussion
-
-\noindent
--- summary
-
-\noindent
--- other theoretical properties of SW models
-
-\noindent
--- room to improve complexity for best-search algorithm
-... modular approach with oracles ...
-
-and extention to $n$-best
-
-\noindent
--- offline algorithm, 
-semi-online implementation for AMT (bar-by-bar approach)
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% BIBLIO                                                                     %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\bibliographystyle{plain}
-%\bibliographystyle{plainurl} 
-\bibliographystyle{abbrv}
-%\bibliographystyle{splncs04}
-\bibliography{references}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% APPENDIX                                                                   %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\appendix
-
-\end{document}
-
-\section{Edit-Distance}
-
-%\subsection{Distance between words or languages}
-...algebraic definition of edit-distance of Mohri, in \cite{Mohri03EDWA}
-% Mehryar Mohri 
-% Edit-distance of weighted automata: General definitions and algorithms
-% International Journal of Foundations of Computer Science 14.06 (2003): 957-982.
-distance $d$ over $\Sigma^* \times \Sigma^*$ 
-into a semiring  $\Semiring = ( \Semiring, \oplus, \zero, \otimes, \one)$.
-
-%\noindent
-Let $\Omega = \Sigma \cup \{ \epsilon \} \times \Sigma \cup \{ \epsilon \} \setminus \{ (\epsilon, \epsilon) \}$,
-and let $h$ be the morphism from $\Omega^*$ into $\Sigma^* \times \Sigma^*$  
-defined over the concatenation of strings of $\Sigma^*$ (that removes the $\epsilon$'s).
-%
-\noindent
-An \emph{alignment} between 2 strings  $s, t \in \Sigma^*$ is an element $\omega \in \Omega^*$ 
-such that $h(\omega) = (s, t)$.
-%
-\noindent
-We assume a base cost function $\Omega$ : $\delta: \Omega \to S$, extended to $\Omega^*$ as follows  
-(for $\omega \in \Omega^*$): 
-\(
-\displaystyle\delta(\omega) = \bigotimes_{0 \leq i < |\omega|} \delta(\omega_i)
-\).
-
-\noindent
-\begin{definition}
-For  $s, t \in \Sigma^*$, the edit-distance between $s$ and $t$ is  
-\( 
-d(s, t) = \displaystyle\bigoplus_{\omega \in \Omega^*\, h(\omega) = (s, t)} \delta(\omega)
-\).
-\end{definition}
-
-e.g. Levenstein edit-distance: $S$ is min-plus and $\delta(a, b) = 1$ for all $(a, b) \in \Omega$.
-
-%\paragraph{Distance between a word and a regular language}

### 2021-05-21 main2.tex
applicaion partielle transducteur

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% prelin
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Preliminary Notions}
+\label{section:prelim}\label{sec:prelim}
+
+notations: for set $S$ : $S^*$ and $S^+$.
+interval $[i..j]$ of natural numbers
+
+\subsection{Semirings} 
+\label{section:semiring}\label{sec:semiring}
+\input{semiring}
+
+\subsection{Label Theory}
+\label{section:symbols}
+\input{labels}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %% SWT & SWA
@@
 
-\subsection{Semirings} 
-\label{section:semiring}
-\input{semiring}
-
-\subsection{Label Theory}
-\label{section:symbols}
-\input{labels}
-
-\subsection{Definitions} \label{sec:SWTdef}\label{sec:SWAdef}
+%\subsection{Definitions} \label{sec:SWTdef}\label{sec:SWAdef}
 %(SWT)
 
-Let $\Semiring$ be a commutative and complete semiring, 
+Let $\Semiring$ be a commutative semiring, 
 $\Sigma$ and $\Delta$ be alphabets called respectively \emph{input} and \emph{output}, %{alphabets}, 
@@
 where $Q$ is a finite set of states, 
-$\mathsf{in} : Q \to \Semiring$, 
-respectively $\mathsf{out} : Q \to \Semiring,$
+$\mathsf{in} : Q \to \Phi_{\Sigma, \Delta}$, 
+respectively $\mathsf{out} : Q \to \Phi_{\Sigma, \Delta}$,
 are functions defining the weight for entering, 
 respectively leaving, a state, 
-and $\bar{\wei}$ is a 4-uplet of transition functions 
-$\wei_{00}$, $\wei_{10}$, $\wei_{01}$, and $\wei_{11}$
+and $\bar{\wei}$ is a triplet of transition functions 
+$\wei_{10}$, $\wei_{01}$,  and $\wei_{11}$
 %$\wei_{i, j}: Q \times Q \to \Phi_{\Sigma, \Delta}$, with $i, j \in \{ 0, 1\}$.
@@
 
-A pair of states $\< q, q'>$ such that $\wei_{00}(q, q') \neq \zero$
-is called an \emph{$\epsilon$-transition}.
+%A pair of states $\< q, q'>$ such that $\wei_{00}(q, q') \neq \zero$
+%is called an \emph{$\epsilon$-transition}.
 %
 For convenience, when $\wei_{ij}(q, q') = \eta \in \Phi_{\Sigma, \Delta}$, 
-for $q, q' \in Q$ and $i, j \in \{ 0, 1\}$,
+for $q, q' \in Q$ and $i, j \in \{ 0, 1\} \setminus \{ \<0, 0> \}$,
+respectively $\mathsf{in}(q) = \eta$, $\mathsf{out}(q') = \eta$,
 we shall sometimes write, overloading the function names:
-$\wei_{ij}(q, a, b, q') = \eta(a, b)$ when  $a \in \Sigma$,  $b \in \Delta$.
-
-\noindent 
-The \SWT $T$ computes on pairs of words $\< s, t> \in \Sigma^* \times \Delta^*$,
-$s$, resp. $t$, being called input word, resp. output word.
+$\wei_{ij}(q, a, b, q') = \eta(a, b)$ for $a \in \Sigma$,  $b \in \Delta$,
+respectively $\mathsf{in}(q, a, b) = \eta(a, b)$,
+$\mathsf{out}(q', a, b) = \eta(a, b)$.
+
+\noindent 
+The \SWT $T$ computes on pairs of words $\< s, t> \in \Sigma^+ \times \Delta^+$,
+$s$ and $t$, being respectively called input and output word.
 %
@@
 $a \in \Sigma$, $u \in \Sigma^*$, $b\in \Delta$, $v\in \Delta^*$,
-\[
-\begin{array}{lccl}
-\weight_T(q, \epsilon, \epsilon, q)  & = & \multicolumn{2}{l}{\one}\\ % or \final(q)
-\weight_T(q, \epsilon, \epsilon, q') & = & \multicolumn{2}{l}{\zero \mathrm{~if~} q \neq q'}\\
+
+\begin{equation}
+\begin{array}[t]{lccl}
+%\weight_T(q, \epsilon, \epsilon, q)  & = & \multicolumn{2}{l}{\one}\\ % or \final(q)
+%\weight_T(q, \epsilon, \epsilon, q') & = & \multicolumn{2}{l}{\zero \mathrm{~if~} q \neq q'}\\
 \weight_T(q, au, bv, q') & = & 
-   \displaystyle\bigoplus_{q'' \in Q} &
-   \wei_{00}(q, a, b, q'') \otimes \weight_T(q'', au, bv, q')\\
-  & \oplus & \displaystyle\bigoplus_{q'' \in Q} &
+             \displaystyle\bigoplus_{q'' \in Q, u \neq \epsilon} &
     \wei_{10}(q, a, b, q'') \otimes \weight_T(q'', u, bv, q')\\
-  & \oplus & \displaystyle\bigoplus_{q'' \in Q} &
+  & \oplus & \displaystyle\bigoplus_{q'' \in Q, v \neq \epsilon} &
     \wei_{01}(q, a, b, q'') \otimes \weight_T(q'', au, v, q')\\
-  & \oplus & \displaystyle\bigoplus_{q'' \in Q} &
+  & \oplus & \displaystyle\bigoplus_{q'' \in Q, u,v \neq \epsilon} &
     \wei_{11}(q, a, b, q'') \otimes \weight_T(q'', u, v, q')\\
+  & \oplus & \displaystyle\bigoplus_{u,v = \epsilon} &
+    \wei_{11}(q, a, b, q')
 \end{array}
-\]
-%
-We recall that, by convention, an empty sum with $\bigoplus$ is $\zero$. 
-%
-The above sum may be infinite because of the $\epsilon$-transitions, % in the first and second lines.
-in this case, the infinite sum operators applies since $\Semiring$ is assumed complete.
-%As we shall see, under conditions, the $\epsilon$-transitions can be removed and hence the sum made finite.
-%** SUPPR. $\epsilon$-transition (with $\wei_{00}$) first. **
-%
-%*** rem. there is no case for $a, \epsilon$ or $\epsilon, b$.
-%   It can be simulated with $\wei_{10}$ and $\wei_{01}$ on the last letter. ***
-
-%** The application of $\oplus$ can be seen as a ND choice between transitions **
-% choose to advance head in input, in output or both (or none)
+\label{eq:SWT-weight}
+\end{equation}
+%
+We recall that, by convention (Section~\ref{sec:semiring}), 
+an empty sum with $\bigoplus$ is equal to~$\zero$. 
+%
 Intuitively, using a transition $\wei_{ij}(q, a, b, q')$ means for $T$:
-read $a$, resp. $b$, at the current position in the input word, resp. output word;
-if $i = 1$, resp. $j = 1$,
-increment the current position in the input word (otherwise, do not change it);
-change state from $q$ to $q'$.
-The input and output symbols are always read at current positions, 
-even during transition that do not increment the head's position.
-This is in contrast with the models of weighted transducers over finite alphabets~\cite{Mohri03ijfcs},
-which ignore the symbol at current position in this case.
-In our case, symbols in a infinite alphabet cannot be stored in the finite memory of the transducer, 
-whereas considering the current symbol may be useful to compute a transition weight .
-%(even when it does not change the head's position, like with $\epsilon$-transitions).
-%
-The cases $\weight_T(q, au, \epsilon, q')$ and $\weight_T(q, \epsilon, bv, q')$
-are missing in the definition of $\weight_T$.
-It means that $T$ must avoid configurations where it reached the end of 
-the output word and not of the input one, or vice-versa.
-This can be done by using $\wei_{10}$ and $\wei_{01}$
-before reaching the end of word, and using a special state for this purpose.
+when reading respectively $a$ and $b$ at the current positions in the input and output words, 
+increment the current position in the input word if if $i = 1$, 
+and in the output word if $j = 1$ (otherwise, do not change it),
+and change state from $q$ to $q'$.
+%
+In contrast with the models of weighted transducers over finite alphabets~\cite{Mohri03ijfcs},
+the input and output symbols at current positions are always read by transitions, 
+even when they do not change the reading position the head's position.
+This is an important feature in the case of an infinite alphabet in 
+order to compare input and output symbols.
+%which cannot be stored in the finite memory of the transducer.
 
@@
 will be ignored in the sum, meaning that there is no possible transition
-from state $q$ into state $q'$ while reading $a$ and writing $b$ 
-(including the case $a = \epsilon$ and $b = \epsilon$).
-%These functions $\phi$ act as guards for the transducer's transitions, 
-%preventing a transition when they return the absorbing $\zero$ of $\Semiring$.
-This is analogous to a transition's guard not satisfied by $\<a, b>$ in 
-the case of symbolic transducers.
+from state $q$ into state $q'$ while reading $a$ and $b$.
+This is analogous to the case of a transition's guard not satisfied by $\<a, b>$ for 
+symbolic transducers.
+
+%whereas considering the current symbol may be useful to compute a transition weight.
+%(even when it does not change the head's position, like with $\epsilon$-transitions).
+%
+%The cases $\weight_T(q, au, \epsilon, q')$ and $\weight_T(q, \epsilon, bv, q')$
+%are missing in the definition of $\weight_T$.
+%It means that $T$ must avoid configurations where it reached the end of 
+%the output word and not of the input one, or vice-versa.
+%This can be done by using $\wei_{10}$ and $\wei_{01}$
+%before reaching the end of word, and using a special state for this purpose.
 
@@
 \medskip\noindent
-The weight associated by $T$ to  $\< s, t> \in \Sigma^* \times \Delta^*$
+Let $\< s, t> \in \Sigma^+ \times \Delta^+$, with $s = s_1\ldots s_n$, 
+and $t = t_1\ldots t_m$. 
+The weight associated by $T$ to $\< s, t>$
 is defined as follows:
@@
 T(s, t)  = 
-\displaystyle\bigoplus_{q, q' \in Q} \mathsf{in}(q) 
-\mathop{\otimes} \weight_T(q, s, t, q') \mathop{\otimes} \mathsf{out}(q').
-\label{eq:weightT}
+\displaystyle\bigoplus_{q, q' \in Q} \mathsf{in}(q, s_1, t_1) 
+\mathop{\otimes} \weight_T(q, s, t, q') \mathop{\otimes} \mathsf{out}(q', s_n, t_m)
+\label{eq:SWT-value}
 \end{equation}
@@
  \wei_{10}(q, d, d', q) = |d' - d|$,
-and $\wei_{00}(q, d, d', q) = \zero$, for all $d, d' \in \mathbb{R}_+$.
+for all $d, d' \in \mathbb{R}_+$.
 % needs reading input/output symbols even by epsilon-transitions.
@@
 %
-Moreover, if $\<\mathsf{e}, d'_i>$ is followed by continuations 
+If $\<\mathsf{e}, d'_i>$ is followed by continuations 
 $\<\mathsf{-}, d'_{i+1}>$..., they are just skip with no cost (transitions $\wei_{01}$ or weight $\one$).
@@
 At last, we let $q_0$ be the only initial and final state, with
-$\init(q_0) = \one$, $\init(q_1) = \zero$, 
-$\final(q_0) = \one$, $\final(q_1) = \zero$, 
+$\init(q_0, d, b) = \final(q_0, d, b) = \one$, and 
+$\init(q_1, d, b) = \final(q_1, d, b) = \zero$, 
+for all $d \in \Sigma$ and $b \in \Delta$.
 \endex
@@
 is a tuple
-$A = \< Q, \init, \bar{\wei}, \final >$,
+$A = \< Q, \init, {\wei_1}, \final >$,
 where $Q$ is a finite set of states, 
-$\mathsf{in} : Q \to \Semiring$, 
-respectively $\mathsf{out} : Q \to \Semiring,$
+$\mathsf{in} : Q \to \Phi_\Sigma$, 
+respectively $\mathsf{out} : Q \to \Phi_\Sigma,$
 are functions defining the weight for entering, 
 respectively leaving, a state, 
-and $\bar{\wei}$ is a pair of transition functions 
-$\wei_{0}$, $\wei_{1}$
-%$\wei_{i, j}: Q \times Q \to \Phi_{\Sigma, \Delta}$, with $i, j \in \{ 0, 1\}$.
+and ${\wei_1}$ is a transition functions 
 from $Q \times Q$ into $\Phi_{\Sigma}$.
 \end{definition}
-A pair of states $\< q, q'>$ such that $\wei_{0}(q, q') \neq \zero$
-is called an \emph{$\epsilon$-transition}.
-
+%      
 \noindent
 As above in the case of \SWT, 
-the transition functions $\wei_0$ and $\wei_1$
-can be synthesized into a single function
-$\wei: Q \times \Sigma \times Q \to \Semiring$.
+when $\wei_1(q, q') = \phi \in \Phi_\Sigma$, 
+respectively $\mathsf{in}(q) = \phi$, $\mathsf{out}(q') = \phi$,
+we may write $\wei_1(q, a, q')$ for $\phi(a)$, 
+respectively $\mathsf{in}(q, a) = \phi(a)$, $\mathsf{out}(q', a) = \phi(a)$.
+%$\wei_1: Q \times \Sigma \times Q \to \Semiring$, 
+The computation of $A$ on words $s \in \Sigma^+$
+is defined with an intermediate function $\weight_A$, 
+defined as follows for $q, q' \in Q$, $a \in \Sigma$, $u \in \Sigma^*$,
+%
+\begin{equation}
+\begin{array}{lccl}
+%\weight_T(q, \epsilon, \epsilon, q)  & = & \multicolumn{2}{l}{\one}\\ % or \final(q)
+%\weight_T(q, \epsilon, \epsilon, q') & = & \multicolumn{2}{l}{\zero \mathrm{~if~} q \neq q'}\\
+\weight_A(q, au, q') & = & 
+            \displaystyle\bigoplus_{r \in Q, u \neq \epsilon} &
+    \wei_{1}(q, a, r) \otimes \weight_A(r, u, q')\\
+  & \oplus & \displaystyle\bigoplus_{u = \epsilon} & 
+    \wei_{1}(q, a, q')
+\end{array}
+\label{eq:SWA-weight}
+\end{equation}
+%
+\noindent
+and the weight value associated by $A$ to 
+$s = s_1\ldots s_n \in \Sigma^+$ is: % defined as follows:
+\begin{equation}
+A(s)  = 
+\displaystyle\bigoplus_{q, q' \in Q} \mathsf{in}(q, s_1) 
+\mathop{\otimes} \weight_A(q, s, q') \mathop{\otimes} \mathsf{out}(q', s_n)
+\label{eq:SWA-value}
+\end{equation}
+
 %
@@
 
-\subsection{Properties}
-The two following properties will be useful to our approach on 
-symbolic weighted parsing in Section~\ref{sec:parsing}.
-
-\begin{proposition}
+%\subsection{Properties}
+\noindent
+The following property will be useful to the approach on 
+symbolic weighted parsing presented in Section~\ref{sec:parsing}.
+
+\begin{proposition} \label{prop:epsilon}
 Given a \SWT $T$ over $\Sigma$, $\Delta$, $\Semiring$ and $\bar\Phi$,
-and $s \in \Sigma^*$, 
+and $s \in \Sigma^+$, 
 there exists an effectively constructible \SWA 
 $A_{T, s}$ over $\Delta$, $\Semiring$ and $\bar\Phi$,
-such that for all $t \in \Delta^*$, $A_{T, s}(t) = T(s, t)$.
+such that for all $t \in \Delta^+$, $A_{T, s}(t) = T(s, t)$.
 \end{proposition}
@@
 be the synthesized form of 
-$\bar{\wei} = \< \wei_{00}, \wei_{10}, \wei_{01}, \wei_{11}>$,
-and let $s = s_1 \ldots s_n$ with $s_1, \ldots, s_n \in \Sigma$.
+$\bar{\wei} = \< \wei_{10}, \wei_{01}, \wei_{11}>$,
+and let $s = s_1 \ldots s_n$ with $s_1, \ldots, s_n \in \Sigma$ ($n \geq 1$).
 %$n = |s|$ be the length of $s$.
-%
-The state set of $A_{T, s}$ will be $Q' = [1..n] \times Q$.
-Its state entering weight function is defined by 
-$\init'(1, q) = \init(q)$ for all $q \in Q$, 
-and $\init'(i, q) = \zero$ for all $1 < i \leq n$, 
-and its state leaving weight function by 
-$\final'(n, q) = \final(q)$ or all $q \in Q$, 
-and $\final'(i, q) = \zero$ for all $1 \leq i <  n$.
-
-\noindent
-We define transition functions $\wei'_0$ and $\wei'_1$ for $A_{T, s}$, 
-together with their synthesized form $\wei'$ as follows: 
-for all $1 \leq i, j \leq n$, $q, q' \in Q$, %and $a \in \Sigma$, 
-and $b \in \Delta$, 
+
+The state set of $A_{T, s}$ will be $Q' = [1..n+1] \times Q$.
+Its state entering weight function is defined by, 
+for all $q \in Q$ and $b \in \Delta$,  
+$\init'(\< 1, q>, b) = \init(q, s_1, b)$ 
+and $\init'(\< i, q>, b) = \zero$ for all $1 < i \leq n+1$. 
+Its state leaving weight function is defined by, for all $q \in Q$ and $b \in \Delta$,  
+$\final'(\< n+1, q>, b) = \final(q, s_n, b)$, 
+and $\final'(\< i, q>, b) = \zero$ for all $1 \leq i <  n+1$.
+
+Every non-null transition of $A_{T, s}$ will
+simulate a sequence of transitions of $T$ which advance in 
+the input word while staying immobile in the output word, 
+and then make one step in the output word (and advance in the input word or not).
+%
+The above initial sequence correspond to $\epsilon$-transitions of automata, 
+and its total weight is computed by the following intermediate function 
+$\wei'_0 : Q' \times Q' \to \Phi_{\Sigma}$, 
+defined for all $q, q' \in Q$: % all $1 \leq i < n$ and $1 \leq k \leq n-i$, 
+by:
 \[
-\begin{array}{lclcr}
-\wei'_0\bigl(\< i, q>, \< i, q'>\bigr) & = & \wei_{00}(q, q')\\ 
-%
-\wei'_0\bigl(\< i, q>, \< i+1, q'>\bigr) & = & \phi(s_i) & 
-\mathrm{if} & i < n, \mathrm{where~} \phi = \wei_{10}(q, q'),\\
-%
-\wei'_0\bigl(\< i, q>, \< j, q'>\bigr) & = & \zero & 
-\mathrm{if} & j \neq i, i+1,\\ 
-%
-\wei'_1\bigl(\< i, q>, \< i, q'>\bigr) & = & \wei_{01}(q, q')\\
-%
-\wei'_1\bigl(\< i, q>, \< i+1, q'>\bigr) & = & \eta_{s_i} &
-\mathrm{if} & i < n, \mathrm{where~} \eta = \wei_{11}(q, q'),\\
-%
-\wei'_1\bigl(\< i, q>, \< j, q'>\bigr) & = & \zero & 
-\mathrm{if} & j \neq i, i+1.
+\begin{array}{rcl}
+\wei'_0\bigl(\<i, q>, \< i, q>\bigr) & = & \one 
+\quad \mathrm{if~} 1 \leq i \leq n+1,\\
+%
+\wei'_0\bigl(\<i, q>, \< i, q'>\bigr) & = & \zero 
+\quad \mathrm{if~} 1 \leq i \leq n+1 \mathrm{~and~} q \neq q',\\
+%
+\wei'_0\bigl(\<i, q>, \< i+k, q'>\bigr) & = & 
+\displaystyle
+\bigoplus_{\begin{array}{c}
+           \scriptstyle q_0,\ldots, q_k \in Q\\[-2pt]
+           \scriptstyle q_0 = q, q_k = q'
+           \end{array}} 
+\bigotimes_{j=0}^{k-1} \bigl[\wei_{10}(q_j, q_{j+1})\bigr]_{s_{i+j}}
+\quad \begin{array}[t]{l}
+      \mathrm{if~} 1 \leq i < n\\
+      \mathrm{and~} 1 \leq k \leq n-i.
+      \end{array}
 \end{array}
 \]
-We use the assumption that the underlying label theory of $T$ is 
-is closed under partial applications (Section~\ref{section:symbols})
-in the above construction of $\wei'_1$.
-
-\noindent
-The $\SWA$ wanted is $A_{T, s} = \< Q', \init', {\bar\wei}', \final' >$.
+where $\bigl[\wei_{10}(q_j, q_{j+1})\bigr]_{s_{i+j}}$ 
+is the partial application $\eta_{s_{i+j}} \in \Phi_{\Delta}$
+for $\eta = \wei_{10}(q_j, q_{j+1}) \in \Phi_{\Sigma, \Delta}$ ($s_{i+j} \in \Sigma$).
+%
+The function $\wei'_0$ is defined thanks to the closure properties of the label theory $\bar\Phi$
+(Section~\ref{section:symbols}). 
+% closed under partial applications 
+The sum and product in its definition are finite, 
+we shall see below how to compute the first sum in polynomial time.
+%The effective computation of $\wei'_0$ does not require enumerating all the state sequences 
+%of length $k+1$ in the sum.
+%Indeed, it amounts to search short
+
+\noindent
+We define now the transition function 
+$\wei'_1: Q' \times Q' \to \Phi_\Sigma$
+of $A_{T, s}$ as follows,
+for $q, q' \in Q$,
+$1 \leq i \leq n$, and $0 \leq k \leq n-i$:
+\[
+\begin{array}{lcl}
+\wei'_1\bigl(\< i, q>, \< i, q'>\bigr) & = & \wei_{01}(q, q')\\
+%
+\wei'_1\bigl(\< i, q>, \< i+k, q'>\bigr) & = & 
+\displaystyle
+\bigoplus_{\begin{array}{c}
+           \scriptstyle q'' \in Q\\[-2pt]
+           \scriptstyle i+k < n
+           \end{array}} 
+\wei'_0\bigl(\<i, q>, \< i+k, q''>\bigr) \otimes \eta_{s_{i+k}}
+\quad \mathrm{where~} \eta = \wei_{01}(q'', q'),\\
+ & \oplus & \displaystyle
+\bigoplus_{q'' \in Q} \wei'_0\bigl(\<i, q>, \< i+k-1, q''>\bigr) \otimes \eta'_{s_{i+k}}
+\quad \mathrm{where~}\eta' = \wei_{11}(q'', q'),\\
+%
+\wei'_1\bigl(\< i, q>, \< j, q'>\bigr) & = & \zero  
+\quad \mathrm{if~} j < i.
+\end{array}      
+\]
+%
+
+\noindent
+We can show that the $\SWA$ $A_{T, s} = \< Q', \init', \wei'_1, \final' >$
+has the expected property: $\forall t \in \Delta^+, A_{T, s}(t) = T(s, t)$.
 \qed
@@
 
+\noindent
+On the quadratic computation of $\wei'_0$. **
+
+\noindent
 The construction time and size for $A_{T, s}$ are $O(\| T \| . | s |)$,
 where the size $\| T \|$ of $T$ is its number of states $|Q|$.
-
-% gen. to transducers
-\begin{proposition}
-When $\Semiring$ is commutative, bounded and complete, 
-given a \SWA $A$ 
-over $\Sigma$, $\Semiring$ and $\bar\Phi$
-there exists an effectively constructible \SWA $A'$ 
-without $\epsilon$-transitions 
-over $\Sigma$, $\Semiring$ and $\bar\Phi$,
-such that for all $s \in \Sigma^*$, $A'(s) = A(s)$.
-\end{proposition}
-%
-\begin{proof} Let $A = \< Q, \init, \<\wei_0, \wei_1>, \final >$.
-The automaton $A'$ is $\< Q, \init, \<\wei'_0, \wei'_1>, \final >$,
-where, for all $q, q' \in Q$,
-$\wei'_0(q, q') = \zero$  
-and 
-** Tb revised **   
-% \[
-% \begin{array}{rcccl}
-% \wei'_\Sigma(q, q') & = &  & \displaystyle\bigoplus_{q_1 \in Q} &
-%                \wei_\Sigma(q, q_1) \otimes  \wei_\epsilon(q_1, q')\\
-%                     &   & \oplus  & \displaystyle\bigoplus_{q_2 \in Q} &
-%                \wei_\epsilon(q, q_2) \otimes \wei_\Sigma(q_2, q')\\
-% \end{array}
-% \]
-\[
-\wei'_\Sigma(q, q') = 
-\displaystyle\bigoplus_{q'' \in Q} \wei_\epsilon(q, q'') \otimes \wei_\Sigma(q'', q')
-\oplus \displaystyle\bigoplus_{q_1 \in Q} \wei_\Sigma(q, q'') \otimes \wei_\epsilon(q'', q')
-\]
-In the above definition of $\wei'_\Sigma$ we use the operator
-of product of function of $\Phi_\Sigma$ by $\Semiring$ 
-described in Section~\ref{section:symbols}.
-%
-By definition of $\weight_{A}$ and distributivity of $\otimes$ on $\oplus$, 
-** NO. TBC see \cite{Mohri02ijfcs} **
-it holds that $\weight_{A}(q, s, q') = \weight_{A'}(q, s, q')$.
-\qed
-\end{proof}
 

### 2021-05-22 main2.tex
revision def. SWVPA

@@
 \def\Phir{{\Phi_\mathsf{r}}}
+\def\Phix{{\Phi_\mathsf{x}}}
 \def\Phici{{\Phi_\mathsf{ci}}}
@@
 \def\weie{{\wei_\mathsf{e}}}
+\def\weiei{{\wei^\mathsf{e}_\mathsf{i}}}
+\def\weiec{{\wei^\mathsf{e}_\mathsf{c}}}
+\def\weier{{\wei^\mathsf{e}_\mathsf{r}}}
+\def\weiex{{\wei^\mathsf{e}_\mathsf{x}}}
 \newcommand{\config}[2]{\ensuremath \genfrac{[}{]}{0pt}{}{#1}{#2}} 
@@
 On the quadratic computation of $\wei'_0$. **
+by a best path search in the graph with nodes $[i..i+k] \times Q$
+and edges labeled in $\Phi_\Sigma$...
+ordering = summary.
 
@@
 %finite (large) or infinite,
-that we assume partitioned into three subsets 
-$\Omegai$, $\Omegac$, $\Omegac$,
+that we assume partitioned into three 
+subsets~$\Omegai$, $\Omegac$, $\Omegac$,
 whose elements are respectively called 
 \emph{internal}, \emph{call} and \emph{return} symbols.
+% names are  coined by application to functional program verification
 % \begin{itemize}
@@
 over  $\Omega = \Omegai \uplus \Omegac \uplus \Omegar$, $\Semiring$ and $\bar\Phi$ 
-is a tuple $A = \< Q, P, \init, \weii, \weic, \weir, \weie, \final >$,
+is a tuple $A = \< Q, P, \init, \bar\wei, \final >$,
 where $Q$ is a finite set of states, 
@@
 respectively leaving, a state, 
-and 
+and $\bar\wei$ is a sextuplet composed of the transition functions:
 $\weii : Q \times P \times Q \to \Phici$,  
+$\weiei : Q \times Q \to \Phii$,  
 $\weic : Q \times P \times Q \times P \to \Phicc$,  
+$\weiec : Q \times P \times Q \to \Phic$,  
 $\weir : Q \times P \times Q \to \Phicr$,  
-$\weie : Q \times Q \to \Phir$,  
-are transition functions.
+$\weier : Q \times Q \to \Phir$.
+%and 
+%$\weiex : Q \times Q \to \Phix$ 
+%with $\mathsf{x} \in \{ \mathsf{i}, \mathsf{c}, \mathsf{r}\}$.
 \end{definition}
@@
 %
-\weic: Q \times \Omegac \times P \times  \Omegac \times Q \times P \to \Semiring & 
-\weic(q, c, p, \call{c'}, q', p') = \eta_\mathsf{cc}(c, \call{c'}) & 
-\mathrm{where~} \eta_\mathsf{cc} = \weic(q, p, q', p'),\\
+\weiei: Q \times \Omegai \times Q \to \Semiring & 
+\weiei(q, a, q') = \phi_\mathsf{i}(a) &
+\mathrm{where~} \phi_\mathsf{i} = \weiei(q, q').\\
+%
+\weic: Q \times \Omegac \times P \times  \Omegac \times P \times Q \to \Semiring & 
+\weic(q, c, p, \call{c'}, p', q') = \eta_\mathsf{cc}(c, \call{c'}) & 
+\mathrm{where~} \eta_\mathsf{cc} = \weic(q, p, p', q'),\\
+%
+\weiec: Q \times \Omegac \times P \times Q \to \Semiring & 
+\weiec(q, {\call{c}}, p, q') = \phi_\mathsf{c}({\call{c}}) &
+\mathrm{where~} \phi_\mathsf{c} = \weiec(q, p, q').\\
 %
@@
 %
-\weie: Q \times \Omegar \times Q \to \Semiring & 
-\weie(q, {\return{r}}, q') = \phi_\mathsf{e}({\return{r}}) &
-\mathrm{where~} \phi_\mathsf{e} = \weie(q, q').\\
+\weier: Q \times \Omegar \times Q \to \Semiring & 
+\weier(q, {\return{r}}, q') = \phi_\mathsf{r}({\return{r}}) &
+\mathrm{where~} \phi_\mathsf{r} = \weier(q, q').\\
 \end{array}      
@@
 The intuition is the following for the above transitions.
-\begin{description}
-\item $\weii$ : read the input internal symbol $a$, change state to $q'$ (stack is untouched).
-\item $\weic$ : read the input call symbol $\call{c}$, push it to the stack along with $p$, change state to $q'$.
-\item $\weir$ : when the stack is not empty, 
-      read and pop from stack a pair made of $\call{c}$ and $p$, 
-      read the input return symbol $\return{r}$, change state to $q'$.
-      In this case, the weight function $\phi_\mathsf{r}$ 
-      computes a value of matching between the call and return symbols $c$ and $r$.
-      This value might be set to $\zero$ in order to express that the symbols do not match.
-\item $\weie$ : when the stack is empty, 
-      read the input symbol $\call{r}$, change state to $q'$.
-\end{description}
-
-We give now a formal definition of these transitions of the automaton~$A$
+
+\noindent
+$\weii$ and $\weiei$ both read an input internal symbol $a$ and change state from $q$ to $q'$, 
+without changing the stack. 
+Moreover, $\weii$ reads 
+${\call{c}} \in \Omegac$ and $p \in P$ at the top of the stack 
+($c$ is compared to $a$ by the weight function $\eta_\mathsf{ci} \in \Phici$)
+and $\weiei$ applies when the stack is empty.
+
+\noindent
+$\weic$ and $\weiec$ read the input call symbol $\call{c}'$, 
+push it to the stack along with $p'$, and change state from $q$ to to $q'$.
+Moreover, $\weic$ reads ${\call{c}}$ and $p$ at the top of the stack 
+($c$ is compared to $c'$),
+and $\weiec$ applies when the stack is empty.
+
+\noindent
+$\weir$ and $\weier$ read the input return symbol $\return{r}$, and change state from $q$ to to $q'$.
+Moreover, $\weir$ reads and pop from stack a pair made of $\call{c}$ and $p$, 
+($\call{c}$ is compared to $\return{r}$),
+and $\weier$ applies when the stack is empty.
+%In this case, the weight function $\phi_\mathsf{r}$ 
+%computes a value of matching between the call and return symbols $c$ and $r$.
+%This value might be set to $\zero$ in order to express that the symbols do not match.
+
+Formally, the transitions of the automaton~$A$ are defined
 in term of %a weight value computed by 
@@
 Therefore, $\weight_A$ is a function from 
-$Q \times \Gamma^* \times \Omega^* \times Q \times \Gamma^*$ into $\Semiring$.
+$[Q \times \Gamma^*] \times \Omega^* \times [Q \times \Gamma^*]$ into~$\Semiring$.
 \[
 \begin{array}{rcl}
-\weight_A\bigl(\config{q}{\gamma}, a\, u, \config{q'}{\gamma'}\bigr) & = & 
- {\displaystyle\bigoplus_{q'' \in Q}} \weii(q, a, q'') 
+\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot\gamma}, a\, u, \config{q'}{\gamma'}\bigr) & = & 
+ {\displaystyle\bigoplus_{q'' \in Q}} \weii(q, c, p, a, q'') 
  \otimes \weight_A\bigl(\config{q''}{\gamma}, u, \config{q'}{\gamma'}\bigr)\\
 %
-\weight_A\bigl(\config{q}{\gamma}, {\call{c}}\, u, \config{q'}{\gamma'}\bigr) & = & 
+\weight_A\bigl(\config{q}{\bot}, a\, u, \config{q'}{\gamma'}\bigr) & = & 
+ {\displaystyle\bigoplus_{q'' \in Q}} \weiei(q, a, q'') 
+ \otimes \weight_A\bigl(\config{q''}{\bot}, u, \config{q'}{\gamma'}\bigr)\\
+%
+\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot\gamma}, {\call{c}'}\, u, \config{q'}{\gamma'}\bigr) & = & 
  {\displaystyle\bigoplus_{\begin{array}{c}
@@
                           \end{array}}}
- \weic\bigl(q, {\call{c}}, q'', p\bigr) 
- \otimes \weight_A\bigl(\config{q''}{{\call{c}}\, p\cdot \gamma}, u, \config{q'}{\gamma'}\bigr)\\[1mm]
+ \weic\bigl(q, {\call{c}}, p, {\call{c}'}, p', q''\bigr) 
+ \otimes \weight_A\bigl(\config{q''}{{\call{c}'}\, p' \cdot {\call{c}}\, p\cdot \gamma}, u, \config{q'}{\gamma'}\bigr)\\[1mm]
+%
+\weight_A\bigl(\config{q}{\bot}, {\call{c}}\, u, \config{q'}{\gamma'}\bigr) & = & 
+ {\displaystyle\bigoplus_{q'' \in Q}} \weiec(q, {\call{c}}, p, q'') 
+ \otimes \weight_A\bigl(\config{q''}{{\call{c}}\, p}, u, \config{q'}{\gamma'}\bigr)\\
 %
@@
 \weight_A\bigl(\config{q}{\bot}, {\return{r}}\, u, \config{q'}{\gamma'}\bigr) & = & 
- {\displaystyle\bigoplus_{q'' \in Q}} \weie(q, {\return{r}}, q'') 
+ {\displaystyle\bigoplus_{q'' \in Q}} \weier(q, {\return{r}}, q'') 
  \otimes \weight_A\bigl(\config{q''}{\bot}, u, \config{q'}{\gamma'}\bigr)\\
@@
 The weight associated by $A$ to $s \in \Omega^*$
-is then defined as follows,
-following empty stack semantics: 
+is defined according to empty stack semantics: 
+%
 \begin{equation}
@@
 
-\subsection{Best-first Search} 
+\subsection{Best Search} 
 \label{sec:best}\label{sec:search}
@@
 and since $\leq_\oplus$ is assumed total, it is the minimum in $\Omega^*$
-of $s \mapsto \weight_A(\config{q}{\sigma}, s, \config{q'}{\sigma})$ \wrt this ordering.
-%
-When $\sigma = \bot$, $d_0(q, \sigma, q')$ is the central 
-expression in a term of the definition (\ref{eq:weightA}) of $A(s_0)$ for the minimum $s_0$
-(for the above function).
-When $\sigma = \bot$, intuitively, it is the minimum weight of a computation of $A$
+of the fonction 
+$s \mapsto \weight_A(\config{q}{\sigma}, s, \config{q'}{\sigma})$ 
+\wrt this ordering.
+%
+When $\sigma = \bot$, $d_0(q, \sigma, q')$ is the central expression in %a term of 
+the definition (\ref{eq:weightA}) of $A(s_0)$ for the minimum $s_0$
+(for the function $\weight_A$).
+%
+When $\sigma = \top$, intuitively, it is the minimum weight of a computation of $A$
 starting in state $q$ with a stack $\gamma \in \Gamma^*$ (possibly empty), 
-and ending in state $q$ with the same stack $\gamma$, such that moreover 
+and ending in state $q'$ with the same stack $\gamma$, such that moreover 
 the computation does not touch a symbol of $\gamma$. 
@@
 $A$ may apply the first case of in the definition of $\weight_A$ (internal symbol),
-as well as the second case, to can push call symbols on the top of $\gamma$, 
+as well as the second case, to push call symbols on the top of $\gamma$, 
 and may pop these symbols with the third case (return symbol). 
@@
 % (internal symbol or call symbol, with a push on the top of $\top$).
-
-The algorithm~\ref{algo:Dijkstra}
+%
+Algorithm~\ref{algo:Dijkstra}
 constructs iteratively a marking $d : Q \times \{ \bot, \top \} \times Q \to \Semiring$
@@
 %of states of $A$ by weight values in $\Semiring$, 
-that converges eventually to $d_0(q, \sigma, q')$. 
+that converges eventually to $d_0$.
 %It uses for that purpose a priority queue $P$ containing triplets of 
@@
 
-\begin{algo}[1-best for \SWVPA] \label{algo:Dijkstra}
+\begin{algo}{best search for \SWVPA} \label{algo:Dijkstra}
 \textbf{initially} let $P = Q \times \{ \bot, \top \} \times Q$, %contains all %triplets 
@@
 
-\noindent\quad \textbf{for all} $q_0, q_3 \in Q$ and $p \in P$ do
-
-\noindent
-\qquad
-\(\begin{array}{rrcl}
+\noindent\quad \textbf{for all} $q_0, q_3 \in Q$ and $p \in P$, 
+update $d$ with $d(q_1, \sigma, q_2)$ (Figure~\ref{fig:best-update}).
+\end{algo}
+
+\begin{figure}[t]
+\[
+\begin{array}{rrcl}
+\multicolumn{4}{l}{\mathrm{For~all~} q_0, q_3 \in Q, p \in P,}\\
  & d(q_1, \sigma, q_3) & \opluseq &
@@
 \end{array}
-\) 
-\end{algo}
-
-\noindent
-The infinite sums in the updates of $d$ in Algorithm~\ref{algo:Dijkstra} are well defined
+\]
+\caption{Update $d$ with $d(q_1, \sigma, q_2)$.}
+\label{fig:best-update}
+\end{figure}
+
+\noindent
+The infinite sums in the updates of $d$ in Algorithm~\ref{algo:Dijkstra}, 
+Figure~\ref{fig:best-update}
+are well defined
 since~$\Semiring$ is complete.
+** effectively computable by hypothese that the label theory is effective**
 The algorithm performs $2.|Q|^2$ iterations until $P$ is empty, 
@@
 commutative, idempotent, superior, total and complete semiring 
-and an alphabet $\Omega$ with a $k$-convex label theory,
+and an alphabet $\Omega$ with a $k$-convex effective label theory,
 one can construct in PTIME a word $s \in \Omega^*$ 

### 2021-05-23 main2.tex
revision algo best search for new extended def. VPA

@@
 \weiei(q, a, q') = \phi_\mathsf{i}(a) &
-\mathrm{where~} \phi_\mathsf{i} = \weiei(q, q').\\
+\mathrm{where~} \phi_\mathsf{i} = \weiei(q, q').\\[2pt]
 %
@@
 \weiec(q, {\call{c}}, p, q') = \phi_\mathsf{c}({\call{c}}) &
-\mathrm{where~} \phi_\mathsf{c} = \weiec(q, p, q').\\
+\mathrm{where~} \phi_\mathsf{c} = \weiec(q, p, q').\\[2pt]
 %
@@
 without changing the stack. 
-Moreover, $\weii$ reads 
+Moreover, $\weii$ reads a pair made of 
 ${\call{c}} \in \Omegac$ and $p \in P$ at the top of the stack 
 ($c$ is compared to $a$ by the weight function $\eta_\mathsf{ci} \in \Phici$)
-and $\weiei$ applies when the stack is empty.
+and $\weiei$ applies if and only if the stack is empty.
 
@@
 ($c$ is compared to $c'$),
-and $\weiec$ applies when the stack is empty.
+and $\weiec$ applies iff the stack is empty.
 
@@
 ($\call{c}$ is compared to $\return{r}$),
-and $\weier$ applies when the stack is empty.
+and $\weier$ applies iff the stack is empty.
 %In this case, the weight function $\phi_\mathsf{r}$ 
@@
 \begin{example}
-structured words...
-intro language of music notation ?
+structured words with timed symbols...
+intro language of music notation? (markup = time division, leaves = events etc)
 \end{example}
@@
 Let $A_1$ and $A_2$ be two (\SWVPA)
-over the same $\Omega$ and $\Semiring$.
+over the same $\Omega$, $\Semiring$ and $\bar\Phi$.
 There exists two $\SWVPA$ $A_1 \oplus A_2$ and $A_1 \otimes A_2$, 
@@
 \label{sec:best}\label{sec:search}
-**hypotheses**
+%**hypotheses**
 Let us assume that the semiring~$\Semiring$ is
@@
 More precisely, 
-let $\top$ be a fresh stack symbol which does not belong to $\Gamma$,
-and for every two states $q, q' \in Q$ and $\sigma \in \{ \bot, \top \}$, 
-let 
-\[ 
-  d_0(q, \sigma, q') = \bigoplus_{s\in \Omega^*} 
+let $b_\bot : Q \times Q \to \Semiring$ be the function:
+%
+\begin{equation}
+  b_\bot(q, q') = \bigoplus_{s\in \Omega^*} 
   \textstyle
-  \weight_A\bigl(\config{q}{\sigma}, s, \config{q'}{\sigma}\bigr). 
-\]
+  \weight_A\bigl(\config{q}{\bot}, s, \config{q'}{\bot}\bigr)
+\end{equation}
+%
 Since $\Semiring$ is complete, this infinite sum is well defined, 
@@
 %
-When $\sigma = \bot$, $d_0(q, \sigma, q')$ is the central expression in %a term of 
+Let $\top$ be a fresh stack symbol which does not belong to $\Gamma$.
+and let $b_\top : Q \times P \times Q \to \Phic$ be such that,
+for every two states $q, q' \in Q$ 
+and stack symbol $p \in P$ %and $\sigma \in \{ \bot, \top \}$, 
+\begin{equation}
+  b_\top(q, p, q') : c \mapsto \bigoplus_{s\in \Omega^*} 
+  \textstyle
+  \weight_A\bigl(\config{q}{c\,p\cdot \top }, s, \config{q'}{c\, p\cdot \top}\bigr). 
+\end{equation}
+%
+When $\sigma = \bot$, $b_\bot(q, q')$ is the central expression in %a term of 
 the definition (\ref{eq:weightA}) of $A(s_0)$ for the minimum $s_0$
@@
 
-\begin{figure}[t]
+\begin{figure}
+For all $q_0, q_3 \in Q$, %$p \in P$,
 \[
-\begin{array}{rrcl}
-\multicolumn{4}{l}{\mathrm{For~all~} q_0, q_3 \in Q, p \in P,}\\
- & d(q_1, \sigma, q_3) & \opluseq &
-   d(q_1, \sigma, q_2) \otimes \displaystyle\bigoplus_{a \in \Omegai} \weii(q_2, a, q_3)\\
-%
-%  & d(q_0, \sigma, q_2) & \opluseq & 
-%    \displaystyle\bigoplus_{a \in \Omegai} \weii(q_0, a, q_1) \otimes d(q_1, \sigma, q_2)\\
-%
-\mathrm{if~} \sigma = \top & d(q_0, \top, q_3) & \opluseq &
-d(q_1, \sigma, q_2) \otimes
-%\displaystyle\bigoplus_{p \in P} 
+\begin{array}{lcl}
+%\multicolumn{3}{l}{\mathrm{For~all~} q_0, q_3 \in Q, p \in P,}\\
+d_\top(q_1, p, q_3) & \opluseq &
+  d_\top(q_1, p, q_2) \otimes 
+  \displaystyle\bigoplus_{a \in \Omegai} \weii(q_2, p, q_3)_a\\
+\multicolumn{3}{l}{\quad
+\mathrm{where~} \weii(q_2, p, q_3)_a \in \Phic
+\mathrm{~is~the~partial~application~}
+x_c \mapsto \weii(q_2, x_c, p, a, q_3)}\\
+%
+d_\bot(q_1, p, q_3) & \opluseq &
+  d_\bot(q_1, q_2) \otimes 
+  \displaystyle\bigoplus_{a \in \Omegai} \weie(q_2, a, q_3)\\
+%     
+d_\top(q_0, p, q_3) & \opluseq &
+  \displaystyle\bigoplus_{{\call{c}'} \in \Omegac} 
+  \displaystyle\bigoplus_{{\return{r}} \in \Omegar} 
+  \weic(q_0, p, p', q_1)_{c'} \otimes 
+  d_\top(q_1, p', q_2) \otimes
+  \weir(q_2, c', p', r, q_3)\\
+\multicolumn{3}{l}{\quad
+\mathrm{where~} \weic(q_0, p, p', q_1)_{c'} \in \Phic
+\mathrm{~is~the~partial~application~}
+x_c \mapsto \weic(q_0, p, x_c, c', p', q_1)}\\[2pt]
+%
+d_\bot(q_0, q_3) & \opluseq &
   \displaystyle\bigoplus_{{\call{c}} \in \Omegac} 
-\displaystyle\bigoplus_{{\return{r}} \in \Omegar} \eta({\call{c}}, {\return{r}})\\
-\mathrm{and} &  d(q_0, \bot, q_3) & \opluseq &
-d(q_1, \sigma, q_2) \otimes
-%\displaystyle\bigoplus_{p \in P} 
-\displaystyle\bigoplus_{{\call{c}} \in \Omegac} 
-\displaystyle\bigoplus_{{\return{r}} \in \Omegar} \eta({\call{c}}, {\return{r}})\\
-& & & \mathrm{where~} \eta = \weic(q_0, q_1, p) \otimes \weir(q_2, p, q_3)\\
-%
-\mathrm{if~} \sigma = \bot & d(q_1, \bot, q_3) & \opluseq &
-  d(q_1, \sigma, q_2) \otimes \displaystyle\bigoplus_{\return{r} \in \Omegar} \weie(q_2, {\return{r}}, q_3)\\
-%
-%& d(q_0, \bot, q_2) & \opluseq & 
-%d(q_0, \bot, q_1) \otimes d(q_1, \sigma, q_2), \mathrm{if~} \< q_0, \bot, q_1> \notin P \\
-%
-%\mathrm{if~} \sigma = \top & d(q_0, \top, q_2) & \opluseq & 
-%d(q_0, \top, q_1) \otimes d(q_1, \sigma, q_2), \mathrm{if~} \< q_0, \top, q_1> \notin P \\
-%
-& d(q_1, \bot, q_3) & \opluseq & 
-   d(q_1, \sigma, q_2) \otimes d(q_2, \bot, q_3), \mathrm{if~} \< q_2, \bot, q_3> \notin P \\
-%
-\mathrm{if~} \sigma = \top & d(q_1, \top, q_3) & \opluseq & 
-    d(q_1, \sigma, q_2) \otimes d(q_2, \top, q_3), \mathrm{if~} \< q_2, \top, q_3> \notin P \\
+  \displaystyle\bigoplus_{{\return{r}} \in \Omegar} 
+  \weiec(q_0, c, p, q_1) \otimes 
+   d_\top(q_1, p, q_2)(c) \otimes
+  \weir(q_2, c, p, r, q_3)\\
+%
+d_\bot(q_1, q_3) & \opluseq &
+  d_\bot(q_1, q_2) \otimes 
+  \displaystyle\bigoplus_{\return{r} \in \Omegar} \weier(q_2, {\return{r}}, q_3)\\
+%
+d_\top(q_1, p, q_3) & \opluseq & 
+  d_\top(q_1, p, q_2) \otimes d_\top(q_2, p, q_3), 
+  \mathrm{if~} \< q_2, \top, q_3> \notin P \\
+%
+d_\bot(q_1, q_3) & \opluseq & 
+  d_\bot(q_1, q_2) \otimes d_\bot(q_2, q_3), \mathrm{if~} \< q_2, \bot, q_3> \notin P \\
 \end{array}
 \]
-\caption{Update $d$ with $d(q_1, \sigma, q_2)$.}
+%$\weii(q_2, p, q_3)_a \in \Phic$, for $a \in \Omegai$,
+%is the partial application
+%$x_c \mapsto \weii(q_2, x_c, p, a, q_3)$.
+%
+%$\weic(q_0, p, p', q_1)_{c'} \in \Phic$, for $c' \in \Omegac$,
+%is the partial application
+%$x_c \mapsto \weic(q_0, p, x_c, c', p', q_1)$.
+\caption{Update 
+$d_\top$ with $d_\top(q_1, p, q_2)$ or $d_\bot$ with $d_\top(\bot, q_2)$.}
 \label{fig:best-update}

### 2021-05-24 main-V2.tex
rename V2 and revise algo best search

--- 2021-05-23 main2.tex
@@
 \newcommand{\E}{\mathbb{E}}
+%\newcommand{\P}{\mathcal{P}}
+\newcommand{\Q}{\mathcal{Q}}
+\newcommand{\R}{\mathcal{R}}
+\newcommand{\T}{\mathcal{T}}
 \newcommand{\W}{\mathbb{W}}
-\newcommand{\T}{\mathcal{T}}
-
 \newcommand{\Semiring}{\mathbb{S}}
@@
 of a state $q \in Q$ and a stack content $\gamma \in \Gamma^*$, where $\Gamma = \Omegac \times P$.
-Therefore, $\weight_A$ is a function from 
+Hence, $\weight_A$ is a function from 
 $[Q \times \Gamma^*] \times \Omega^* \times [Q \times \Gamma^*]$ into~$\Semiring$.
-\[
-\begin{array}{rcl}
+%
+\begin{equation}
+\begin{array}[t]{rcl}
 \weight_A\bigl(\config{q}{{\call{c}}\, p\cdot\gamma}, a\, u, \config{q'}{\gamma'}\bigr) & = & 
@@
  \otimes \weight_A\bigl(\config{q''}{\gamma}, u, \config{q'}{\gamma'}\bigr)\\
-%
 \weight_A\bigl(\config{q}{\bot}, a\, u, \config{q'}{\gamma'}\bigr) & = & 
@@
 \end{array}
-\]
+\label{eq:SWVPA-weight}
+\end{equation}
+%
 where $\bot$ denotes the empty stack and ${\call{c}}\, p\cdot \gamma$ 
@@
 Let us assume that the semiring~$\Semiring$ is
-commutative, bounded, complete, and total.
+commutative, bounded, total, and complete.
 and assume an effective label theory.
@@
 
-We propose a Dijkstra algorithm computing the minimal weight by $A$, 
-\wrt~$\leq_\oplus$, for a word in~$\Omega^*$.
+We propose a Dijkstra algorithm computing, for a $\SWVPA$ $A$
+over~$\Omega$, $\Semiring$ and~$\bar\Phi$, 
+the minimal weight \wrt~$\leq_\oplus$, for a word in~$\Omega^*$.
 
@@
 %
-\begin{equation}
+\begin{equation}\label{eq:bbot}
   b_\bot(q, q') = \bigoplus_{s\in \Omega^*} 
@@
 %
-Since $\Semiring$ is complete, this infinite sum is well defined, 
-and since $\leq_\oplus$ is assumed total, it is the minimum in $\Omega^*$
+Since $\Semiring$ is complete, the infinite sum in \eqref{eq:bbot} is well defined,
+and by totality of $\leq_\oplus$, it is the minimum in $\Omega^*$
 of the fonction 
@@
 %
-Let $\top$ be a fresh stack symbol which does not belong to $\Gamma$.
+The term $\config{q}{\bot}, s, \config{q'}{\bot}$ 
+of this sum is the central expression in 
+the definition \eqref{eq:weightA} of $A(s_0)$, for the minimum $s_0$
+of the function $\weight_A$.
+
+%
+\noindent
+Let $\top$ be a fresh stack symbol which does not belong to $\Gamma$,
 and let $b_\top : Q \times P \times Q \to \Phic$ be such that,
 for every two states $q, q' \in Q$ 
-and stack symbol $p \in P$ %and $\sigma \in \{ \bot, \top \}$, 
-\begin{equation}
+and stack symbol $p \in P$: %and $\sigma \in \{ \bot, \top \}$, 
+\begin{equation}\label{eq:btop}
   b_\top(q, p, q') : c \mapsto \bigoplus_{s\in \Omega^*} 
@@
 %
-When $\sigma = \bot$, $b_\bot(q, q')$ is the central expression in %a term of 
-the definition (\ref{eq:weightA}) of $A(s_0)$ for the minimum $s_0$
-(for the function $\weight_A$).
-%
-When $\sigma = \top$, intuitively, it is the minimum weight of a computation of $A$
-starting in state $q$ with a stack $\gamma \in \Gamma^*$ (possibly empty), 
-and ending in state $q'$ with the same stack $\gamma$, such that moreover 
-the computation does not touch a symbol of $\gamma$. 
-That means that during the computation, 
-$A$ may apply the first case of in the definition of $\weight_A$ (internal symbol),
-as well as the second case, to push call symbols on the top of $\gamma$, 
-and may pop these symbols with the third case (return symbol). 
-However, it cannot apply one of the two last cases (return symbol and empty stack)
-when the current stack is $\gamma$.
+Intuitively, the function defined in \eqref{eq:btop}
+associateds to $c \in \Omegac$ 
+the minimum weight of a computation of $A$
+starting in state $q$ with a stack 
+$c p \cdot \gamma \in \Gamma^+$ 
+and ending in state $q'$ with the same stack,
+such that the computation does pop 
+the pair made of $c$ and $p$ at the top of this stack,
+but may read these symbols.
+Moreover, $A$ may push another pair $\< c', p'>$ %call symbols 
+on the top of $c p \cdot \gamma$,
+following the the third case of 
+in the definition \eqref{eq:SWVPA-weight} of $\weight_A$,
+and may pop $\< c', p'>$ later, following the fifth case of \eqref{eq:SWVPA-weight} (return symbol). 
+%However, it cannot apply one of the two last cases (return symbol and empty stack)
+%when the current stack is $\gamma$.
 %pop symbols in $\gamma$.
@@
 % (internal symbol or call symbol, with a push on the top of $\top$).
-%
+
 Algorithm~\ref{algo:Dijkstra}
-constructs iteratively a marking $d : Q \times \{ \bot, \top \} \times Q \to \Semiring$
+constructs iteratively markings 
+$d_\bot : Q \times Q \to \Semiring$ and 
+$d_\top : Q \times P \times Q \to \Phic$
 %of the triplets $\<q, \sigma, q'>$ 
 %of states of $A$ by weight values in $\Semiring$, 
-that converges eventually to $d_0$.
+that converges eventually to $b_\top$ and $b_\bot$.
 %It uses for that purpose a priority queue $P$ containing triplets of 
@@
 
-\begin{algo}{best search for \SWVPA} \label{algo:Dijkstra}
-\textbf{initially} let $P = Q \times \{ \bot, \top \} \times Q$, %contains all %triplets 
-%$\< q_1, \bot, q_2>$ and $\<q_1, \top, q_2>$ for $q_1, q_2 \in Q$, with 
-and 
-$d(q_1, \bot, q_2) = d(q_1, \top, q_2) = \one$ if $q_1 = q_2$ and
-$d(q_1, \bot, q_2) = d(q_1, \top, q_2) = \zero$ otherwise.
+\begin{algo}[best search for \SWVPA]\label{algo:Dijkstra}
+\textbf{initially} 
+let $\Q = (Q \times Q) \cup (Q \times P \times Q)$, %contains all 
+and let 
+$d_\bot(q_1, q_2) = d_\top(q_1, p, q_2) = \one$ if $q_1 = q_2$ and
+$d_\bot(q_1, q_2) = d_\top(q_1, p, q_2) = \zero$ otherwise.
 
 \smallskip\noindent
-\textbf{while} $P$ is not empty
-
-\noindent
-\quad\textbf{extract} $\< q_1, \sigma, q_2>$ from $P$ 
- such that $d(q_1, \sigma, q_2)$ is minimal wrt $\leq_\oplus$.
-
-\noindent\quad \textbf{for all} $q_0, q_3 \in Q$ and $p \in P$, 
-update $d$ with $d(q_1, \sigma, q_2)$ (Figure~\ref{fig:best-update}).
+\textbf{while} $\Q$ is not empty
+
+\noindent\quad
+\textbf{extract} $\< q_1, q_2>$ or $\< q_1, p, q_2>$ from $\Q$
+
+\noindent\quad
+such that $d_\bot(q_1, q_2)$, resp. $\bigoplus_{c\in \Omegac} d_\top(q_1, p, q_2)(c)$,
+is minimal in $\Semiring$ wrt $\leq_\oplus$.
+
+\noindent\quad 
+%\textbf{for all} $q_0, q_3 \in Q$ 
+\textbf{update} $d_\bot$ with $\< q_1, q_2>$ or $d_\top$ with $\< q_1, p, q_2>$
+(Figure~\ref{fig:best-update}).
 \end{algo}
@@
 %$x_c \mapsto \weic(q_0, p, x_c, c', p', q_1)$.
-\caption{Update 
-$d_\top$ with $d_\top(q_1, p, q_2)$ or $d_\bot$ with $d_\top(\bot, q_2)$.}
+\caption{Update $d_\bot$ with $\<q_1, q_2>$ or $d_\top$ with $\< q_1, p, q_2>$.} 
 \label{fig:best-update}
@@
 is ensured by the invariant expressed in the following lemma.
-\begin{lemma}
-For all $\< q_1, \sigma, q_2> \notin P$, $d(q_1, \sigma, q_2) =  d_0(q_1, \sigma, q_2)$.
+\begin{lemma}\label{lem:bot}
+For all $\< q_1, q_2> \notin \Q$,
+$d_\bot(q_1, q_2) =  b_\bot(q_1, q_2)$/
 \end{lemma}
@@
 
-% \begin{algorithm}
-% \caption{<your caption for this algorithm>}
-% \label{<your label for references later in your document>}
-% \begin{program}
-% \seq{|line| := `` '', i := 1};
-% \WHILE i \neq n+1 \DO
-%       |line| := |item|[i] \concat `` '' \concat |number|[i]; 
-%       i := i+1; 
-%       \WHILE i \neq n+1 \AND |item|[i] = |item|[i-1] \DO 
-%       |line| := |line| \concat ``, '' \concat |number|[i]);
-%       i := i+1 \OD ; 
-%       |write|(|line|) \OD 
-% \end{program}
-% \end{algorithm}
-
-% The weight of a transition acts as a guard: 
-% a transition is activated for a symbol $`a`$ iff its weight $`\phi(a) \neq 0`$ (or $`\psi(a, b) \neq 0`$), 
-% the absorbing element.
-
-%\begin{lemma}
+\begin{lemma}\label{lem:top}
+For all $\< q_1, p, q_2> \notin \Q$, 
+$d_\top(q_1, p, q_2) =  b_\top(q_1, p, q_2)$,
+\end{lemma}
+
 \noindent
@@
   {\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
-  \mathsf{in}(q) \mathop{\otimes} d(q, \bot, q') \mathop{\otimes} \mathsf{out}(q').
+  \mathsf{in}(q) \mathop{\otimes} d_\bot(q, q') \mathop{\otimes} \mathsf{out}(q').
 \]
-%\end{lemma}
 
@@
 
-\begin{proposition}
-For a \SWVPA $A$ over a 
-commutative, idempotent, superior, total and complete semiring 
-and an alphabet $\Omega$ with a $k$-convex effective label theory,
+\begin{proposition}\label{th:best-search}
+For a \SWVPA $A$ 
+over $\Omega$, 
+$\Semiring$ commutative, bounded, total and complete, %semiring 
+and $\bar\Phi$ $k$-convex effective, % label theory,
 one can construct in PTIME a word $s \in \Omega^*$ 
 such that $A(s)$ is minimal \wrt the natural ordering for $\Semiring$. 
-%$\leg_\oplus$.
 \end{proposition}

### 2021-05-24 main-V2.tex
revise section tree-nested words

@@
 
-\subsection{Definition}
+\subsection{Definition}\label{se:SWVPA-def}
 Let $\Omega$ be a countable alphabet 
@@
 \label{sec:trees}
-The hierarchical structure of nested-words, defined the \emph{call} and \emph{return} markup symbols of 
-suggest a correspondence between these word and trees. 
-The lifting of this correspondence to languages, respective of tree automata and VPA have
-been discussed in~\cite{AlurMadhusudan09nested}, 
-see also~\cite{Caralp12VPAmult} for the weighted case.
-In this section, we discuss the correspondence between the symbolic-weighted extensions
+The hierarchical structure of nested-words, defined with the \emph{call} and \emph{return} markup symbols  
+suggest a correspondence with trees. 
+The lifting of this correspondence to languages, of tree automata and VPA,
+has been discussed in~\cite{AlurMadhusudan09nested}, 
+and~\cite{Caralp12VPAmult} for the weighted case.
+In this section, we describe a correspondence between the symbolic-weighted extensions
 of tree automata and VPA.
@@
 $\rank(a) \in [0..M]$ where $M$ is a fixed natural number.
-We write $\Omega_k$ the subset of all symbols $a$ of $\Omega$
-with $\rank(a) = k$, where $0 \leq k \leq M$.
-%$\cup_{i = 0}^{i = M} \Omega_i$ 
-%
-$\T(\Omega)$ denotes the free $\Omega$-algebra of finite, ordered, 
-$\Omega$-labeled trees, 
-which is the smallest set such that  $\Omega_0 \subset \T(\Omega)$
+We denote by $\Omega_k$ the subset of all symbols $a$ of $\Omega$
+with $\rank(a) = k$, where $0 \leq k \leq M$, 
+and $\Omega_{>0} = \Omega \setminus \Omega_0$.
+%
+\noindent 
+The free $\Omega$-algebra of finite, ordered, 
+$\Omega$-labeled trees is denoted by $\T(\Omega)$.
+It is the smallest set such that  $\Omega_0 \subset \T(\Omega)$
 and for all $1 \leq k \leq M$, all $a \in \Omega_k$, 
@@
 % and $n$ subtrees $t_1$,\ldots, $t_n$.
+%
+Let us assume a  commutative semiring $\Semiring$ 
+and a label theory $\bar{\Phi}$ over $\Semiring$ 
+containing one set $\Phi_{\Omega_k}$ for each $k \in [0..M]$.
 
@@
 
-Let $\hat\Omega$ be the countable (unranked) alphabet obtained from 
-$\Omega$ as follows: 
-$\hat\Omega = \< \Omegai, \Omegac, \Omegar >$, with
+Let $\hat\Omega$ be the countable (unranked) alphabet obtained
+from $\Omega$ by: 
+$\hat\Omega = \Omegai \uplus \Omegac \uplus \Omegar$, with
 $\Omegai = \Omega_0$, 
 $\Omegac = \{ \; \call{a} \mid a \in \Omega_{>0} \}$,
-$\Omegac = \{ \; \return{a} \mid a \in \Omega_{>0} \}$,
-where $\Omega_{>0}$ denotes $\bigcup_{k = 1}^{M} \Omega_k$.
+$\Omegar = \{ \; \return{a} \mid a \in \Omega_{>0} \}$.
+We associate to $\hat\Omega$
+a label theory $\hat{\Phi}$ 
+like in Section~\ref{se:SWVPA-def}.
 
@@
 % non terminal replacement following the rules of $R$.
-
-Let us assume a label theory $\Phi_{\Omega_k}$ 
-for each $k \in [0..M]$.
 %
@@
 A \emph{symbolic-weighted tree automaton} (\SWTA)
-over the ranked input alphabet~$\Omega$
-and the commutative semiring $\Semiring$ 
+over $\Omega$, $\Semiring$, and $\bar\Phi$
 is a triplet $A = \< Q, \init, \bar{\wei} >$ where
@@
 $\mathsf{in} : Q \to \Semiring$ is the starting weight function, 
-and $\bar{\wei}$ is a $M+2$-uplet of transition functions made of:
-$\wei_\epsilon$ from $Q \times Q$ into $\Semiring$, and, 
-for each $k \in [0..M]$, $\wei_{\Omega, k}$
-from $Q \times Q^{k}$ into $\Phi_{\Omega_k}$.
+and $\bar{\wei}$ is a tuplet of transition functions containing
+%$\wei_\epsilon$ from $Q \times Q$ into $\Semiring$, and, 
+$\wei_{k}: Q \times Q^{k} \to \Phi_{\Omega_k}$
+for each $k \in [0..M]$. 
 \end{definition}
 %
-Like in Section~\ref{sec:SWAdef}, we define from $\bar{\wei}$ 
+%Like in Section~\ref{sec:SWAdef}, 
+We define from $\bar{\wei}$ 
 a transition function~$\wei$, %from $\bar\wei$
 from 
-$Q \times (\Sigma \cup \{ \epsilon \}) \times \bigcup_{k=0}^{M} Q^k$
-into~$\Semiring$: %also called $\wei$ for simplicity, 
+$Q \times \Omega \times \bigcup_{k=0}^{M} Q^k$
+into~$\Semiring$ by: %also called $\wei$ for simplicity, 
 %such that, for all $q, q' \in Q$, $a \in \Sigma$, and $b \in \Delta$, 
@@
 \begin{array}{rcll}
-\wei(q_0, \epsilon, q_1) & = &  \wei_\epsilon(q_0, q_1),\\ %\phi_\epsilon\\
+%\wei(q_0, \epsilon, q_1) & = &  \wei_\epsilon(q_0, q_1),\\ %\phi_\epsilon\\
 \wei(q_0, a, q_1 \ldots q_k) & = & \phi_{\Omega, k}(a) &
-\quad\mathrm{where~} \phi_{\Omega, k} = \wei_{\Omega, k}(q_0, q_1\ldots q_k).\\
+\quad\mathrm{where~} \phi_{k} = \wei_{k}(q_0, q_1\ldots q_k).\\
 \end{array}      
@@
 of a regular tree grammar~\cite{tata}, 
-that replaces the non-terminal symbol $q_0$, by $a(q_1, \ldots, q_k)$. 
+that replaces the non-terminal symbol $q_0$ by $a(q_1, \ldots, q_k)$. 
 %in a step of tree building.
@@
 from trees of $\T(\Omega)$ into the weights of~$\Semiring$,
-based on the intermediate function $\weight_A$
+based on the intermediate function~$\weight_A$
 defined as follows for $q_0 \in Q$ and 
 $t = b(t_1,\ldots, t_k) \in \T(\Omega)$,
-åwith $0 \leq k \leq M$:
-\[
+with $0 \leq k \leq M$:
+\begin{equation}
 \begin{array}{rccl}
 \weight_A(q_0, t) & = & 
-   \displaystyle\bigoplus_{q_1 \in Q} &
-   \wei(q, \epsilon, q_1) \otimes \weight_A(q_1, t)\\
- & \oplus & \displaystyle\bigoplus_{q_1 \ldots q_k \in Q^k} &
+%   \displaystyle\bigoplus_{q_1 \in Q} &
+%   \wei(q, \epsilon, q_1) \otimes \weight_A(q_1, t)\\
+ \displaystyle\bigoplus_{q_1 \ldots q_k \in Q^k} &
               \wei(q_0, b, q_1 \ldots q_k ) 
    \otimes \displaystyle\bigotimes_{i=1}^{k}
-           \weight_A(q_i, t_i).\\
+           \weight_A(q_i, t_i)
+\end{array}
+\end{equation}
+
+\medskip\noindent
+The weight associated by $A$ to  $t \in \T(\Omega)$ is 
+\begin{equation}
+A(t)  = 
+\displaystyle\bigoplus_{q \in Q} \mathsf{in}(q) \mathop{\otimes} \weight_A(q, t)
+\label{eq:weightTA}
+\end{equation}
+
+\begin{lemma}  \label{lem:SWTA}
+For all \SWTA $A$ over~$\Omega$, $\Semiring$ commutative, and $\Phi$,
+there exists an effectively constructible \SWVPA $A'$ over 
+$\hat\Omega$, $\Semiring$ and $\hat\Phi$ 
+such that for all $t \in \T(\Omega)$, $A'\bigl(\lin(t)\bigr) = A(t)$.
+\end{lemma} 
+% 
+\begin{proof}
+Let $A = \< Q, \init, \bar{\wei} >$ where $\bar{\wei}$ is presented as above by a function
+    %
+We build 
+$A' = \< Q', P', \init', \bar{\wei}', \final' >$,
+%computing over $\hat\Omega = \Omegai \uplus \Omegac \uplus \Omegar$,
+where $Q' = \bigcup_{k=0}^{M} Q^k$ is the set of sequences of state symbols of $A$, 
+of length at most $M$, including the empty sequence denoted by $\epsilon$, 
+and where $P' = Q'$ and $\bar\wei$ is defined by:
+
+\[
+\begin{array}{ll}
+\weii(\bar{q}, {\call{b}}, \bar{p}, a, \bar{q} q') = \wei(q', a, \epsilon) & 
+\mathrm{for~all~} b \in \Omega_{>0}, \bar{p}\in P', a \in \Omega_0\\
+%
+\weiei(\bar{q}, a, \bar{q} q') = \wei(q', a, \epsilon) & 
+\mathrm{for~all~} a \in \Omega_0\\
+%
+\weic(\bar{q}, {\call{b}}, \bar{p}, \call{b}, \epsilon, \bar{q}) = \one & %\wei(p', b, \bar{p}) & 
+\mathrm{for~all~} b \in \Omega_{>0},  \bar{p}\in P', b \in \Omega_{>0}\\
+%
+\weiec(\bar{q}, \call{b}, \epsilon, \bar{q}) = \one & %\wei(p', b, \bar{p}) & 
+\mathrm{for~all~} b \in \Omega_{>0}\\
+%
+%\weir: Q \times \Omegac \times P \times \Sigmar \times Q \to \Semiring & 
+\weir(\bar{q}, {\call{b}}, \bar{p}, {\return{b}}, \bar{p} q') = \wei(q', b, \bar{q}) & 
+\mathrm{for~all~}  b \in \Omega_{>0}, \bar{p}\in P'\\
+%
+%\weie: Q \times \Omegar \times Q \to \Semiring & 
+\weier(\bar{p}, {\return{b}}, \bar{q}) = \zero &
+\mathrm{for~all~}  b \in \Omega_{>0}\\
 \end{array}      
 \]
-
-\medskip\noindent
-The weight associated by $A$ to  $t \in \T(\Omega)$ is then
-\begin{equation}
-A(t)  = 
-\displaystyle\bigoplus_{q \in Q} \mathsf{in}(q) \mathop{\otimes} \weight_A(q, t).
-\label{eq:weightTA}
-\end{equation}
-
-\begin{lemma}  \label{lem:SWTA}
-For all \SWTA $A$ over~$\Omega$ and $\Semiring$, without $\epsilon$-transitions,
-there exists an effectively constructible \SWVPA $A'$ over 
-$\hat\Omega$ and $\Semiring$
-such that for all $t \in \T(\Omega)$, $A'\bigl(\lin(t)\bigr) = A(t)$.
-\end{lemma} 
-% 
-\begin{proof}
-Let $A = \< Q, \init, \bar{\wei} >$ where $\bar{\wei}$ is summarized as above by a function
-$\wei: Q \times (\Sigma \cup \{ \epsilon \}) \times \bigcup_{k=0}^{M} Q^k \to \Semiring$.
-
-We build 
-$A' = \< Q', P', \init', \weii, \weic, \weir, \weie, \final' >$,
-computing over $\hat\Omega = \< \Omegai, \Omegac, \Omegar >$,
-%
-%$\Omega = \Omegai \uplus \Omegac \uplus \Omegar$ and $\Semiring$ 
-where $Q' = \bigcup_{k=0}^{M} Q^k$ be  the set of sequences of state symbols of $A$, 
-of length at most $M$, including the empty sequence denoted by $\epsilon$, 
-and where $P' = Q'$.
-
-\[
-\begin{array}{ll}
-%\weii: Q \times \Omegai \times Q \to \Semiring & 
-\weii(\bar{q}, a, \bar{q} q') = \wei(q', a, \epsilon) & 
-\mathrm{for~all~} a \in \Omega_0,\\
-%
-%\weic: Q \times \Omegac \times Q \times P \to \Semiring & 
-\weic(\bar{q}, \call{b}, \epsilon, \bar{q}) = \one & %\wei(p', b, \bar{p}) & 
-\mathrm{for~all~} b \in \Omega_{>0},\\
-%
-%\weir: Q \times \Omegac \times P \times \Sigmar \times Q \to \Semiring & 
-\weir(\bar{q}, {\call{b}}, \bar{p}, {\return{b}}, \bar{p} q') = \wei(q', b, \bar{q}) & 
-\mathrm{for~all~}  b \in \Omega_{>0},\\
-%
-%\weie: Q \times \Omegar \times Q \to \Semiring & 
-\weie(\bar{p}, {\return{b}}, \bar{q}) = \zero &
-\mathrm{for~all~}  b \in \Omega_{>0}.\\
-\end{array}      
-\]
-
-In practice, it is sufficient to consider in $Q'$ only the prefixes of 
-sequences.
+\noindent
+It is sufficient to consider in $Q'$ only the prefixes of 
+sequences in transition with a non-null weight.
 \qed\end{proof}

### 2021-05-25 main-V2.tex
V2 : section parsing : Def. complete & proof

@@
 $\Sigma$ and $\Delta$ be alphabets called respectively \emph{input} and \emph{output}, %{alphabets}, 
-and $\bar\Phi = \< \Phi_\Sigma, \Phi_\Delta, \Phi_{\Sigma, \Delta}>$ be an $\Semiring$-label theory.
+and $\bar\Phi = \< \Phi_\Sigma, \Phi_\Delta, \Phi_{\Sigma, \Delta}>$ be a label theory over $\Semiring$.
 
@@
 \label{section:SWVPA}\label{sec:SWVPA}
-The following model generalizes Symbolic VPA~\cite{dAntonyAlur14SVPDA}
+The model presented in this section generalizes Symbolic VPA~\cite{dAntonyAlur14SVPDA}
 from Boolean semirings to arbitrary semiring weight domains.
+It will compute on nested words over infinite alphabets, 
+associating to every such word a weight value. 
+Nest words are able to describe structures of labeled trees, 
+and in the context of parsing, they will be useful to 
+represent parse trees.
 
@@
 %
-\begin{proposition}
+\begin{proposition}\label{prop:SWVPA-product}
 Let $A_1$ and $A_2$ be two (\SWVPA)
@@
 
-\subsection{Trees and Nested-Words}
+\subsection{Nested-Words and Parse-Trees}
 \label{sec:trees}
@@
 %
-Let us assume a  commutative semiring $\Semiring$ 
-and a label theory $\bar{\Phi}$ over $\Semiring$ 
-containing one set $\Phi_{\Omega_k}$ for each $k \in [0..M]$.
+Let us assume a commutative semiring~$\Semiring$ 
+and a label theory~$\bar{\Phi}$ over~$\Semiring$ 
+containing one set~$\Phi_{\Omega_k}$ for each $k \in [0..M]$.
 
@@
 
-\begin{lemma}  \label{lem:SWTA}
+\begin{lemma}\label{lem:SWTA}
 For all \SWTA $A$ over~$\Omega$, $\Semiring$ commutative, and $\Phi$,
@@
 %
-\weic(\bar{q}, {\call{b}}, \bar{p}, \call{b}, \epsilon, \bar{q}) = \one & %\wei(p', b, \bar{p}) & 
-\mathrm{for~all~} b \in \Omega_{>0},  \bar{p}\in P', b \in \Omega_{>0}\\
+\weic(\bar{q}, {\call{b}}, \bar{p}, \call{b'}, \epsilon, \bar{q}) = \one & %\wei(p', b, \bar{p}) & 
+\mathrm{for~all~} b \in \Omega_{>0},  \bar{p}\in P', b, b' \in \Omega_{>0}\\
 %
@@
 \label{sec:parsing}
-Let us now use the models and results of the former sections in %order to define 
-an approach to the problem of parsing over infinite alphabet. %appropriate
+Let us now apply the models and results of the previous sections %in order to define 
+to the problem of parsing over infinite alphabet. %appropriate
 %
@@
 handled with suitable language formalisms, 
-this approach extends conventional parsing 
+this approach extends conventional parsing and weighted parsing
 by computing a derivation tree modulo 
@@
 Let $\Sigma$ be a countable input alphabet 
-and $\Omega = \Omegai \uplus \Omegac \uplus \Omegar$ an output countable alphabet,
-let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative, bounded, complete 
+and let $\Omega$ be a countable output ranked alphabet, 
+with maximal rank value $M$, 
+and $\hat\Omega = \Omegai \uplus \Omegac \uplus \Omegar$
+be the alphabet with nesting symbols, associated like Section~\ref{sec:trees}, 
+for the linearization of trees of $\T(\Omega)$
+(remember that $\Omegai = \Omega_0$).
+%
+Let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative, bounded, complete 
 and total semiring 
-and let  $(\Phi_\epsilon, \Phic, \Phir, \Phicr)$ be a label theory over $\Semiring$, 
-assumed computable and $k$-convex for some fixed $k$.
-
-\noindent
-Assuming given in input:
+and let $\bar\Phi$ be a label theory over $\Semiring$
+containing
+$\Phi_\Sigma$, $\Phi_{\Sigma, \Omegai}$, as well as
+$\Phii$, $\Phic$, $\Phir$, $\Phicr$
+(following the notations of Section~\ref{se:SWVPA-def}).
+It is moreover assumed computable and $k$-convex for some fixed $k$.
+
+\noindent
+Let us assume given the following input:
 \begin{itemize}
-\item a \SWT $T$ over $\Sigma$,  $\Omega$, and $\Semiring$,
+\item a \SWT $T$ over $\Sigma$, $\Omegai$, $\Semiring$, and $\bar\Phi$, 
 defining a measure %between words 
-$T: \Sigma^* \times \Omega^* \to \Semiring$,
-\item a \SWVPA $A$ over $\Omega$, and $\Semiring$, defining a series of nested words
-      $A : \Omega^* \to \Semiring$,
-\item an input word $s \in \Sigma^*$,
+$T: \Sigma^* \times \Omegai^* \to \Semiring$,
+
+\item a \SWTA $A$ over $\Omega$, $\Semiring$, and $\bar\Phi$,
+defining a measure $A: \T(\Omega) \to \Semiring$,
+%a \SWVPA $A$ over $\Omega$, and $\Semiring$, defining a series of nested words
+%      $A : \Omega^* \to \Semiring$,
+\item an input word $s \in \Sigma^*$.
 \end{itemize}
-the problem of \emph{symbolic weighted parsing} is to find 
-a nested word $t \in \Omega$ minimizing
-\( T(s, t) \otimes A(t)\)
+%
+As explained in Section~\ref{sec:trees}, 
+$A = \< Q, \init, \bar{\wei} >$ can be seen as a weighted regular tree grammar, 
+that generates (weighted) trees by replacement of a state symbol $q_0$ (non-terminal), 
+by a tree $a(q_1,\ldots, q_k)$, where $k = \rank(a)$.
+A replacement rule $q_0 \to a(q_1,\ldots, q_k)$, 
+of weight $\wei(q_0, a, q_1 \ldots q_k) \in \Semiring$ 
+according to Definition~\ref{def:SWTA}, 
+corresponds to the production rule $q_0 := a(q_1,\ldots, q_k)$ of a weighted CF grammar,
+with set non-terminal symbols $Q$ and set of terminal symbols $\Omega_0$.
+This actually is a slight generalization of CFG since 
+each such production rule is labelled by  a symbol of $\Omega_{>0}$, 
+hence parse trees %derivation trees
+are trees of $\T(\Omega)$.
+%
+Another (more original) generalization is that the set of terminal symbols 
+$\Omega_0$ may be infinite.
+
+\noindent 
+We extend the measure defined by $T$ to 
+$d : \Sigma^* \times \T(\Omega) \to \Semiring$ as follows.
+Given a word $w \in {\hat\Omega}^*$, the projection of $w$ onto $\Omegai$,
+denoted $w|_\Omegai$,
+is the word of $\Omegai^*$ obtained from $w$ by removing all symbols 
+in $\hat\Omega \setminus \Omegai$.
+Using this notation
+and the tree linearization operator defined in Section~\ref{sec:trees}, 
+$d$ is defined by:
+\begin{equation}
+d(s, t) = T\bigl(s, \lin(t)|_\Omegai \bigr)  \mathrm{~for~} s \in \Sigma^*, t\in \T(\Omega)
+\end{equation}
+
+\noindent 
+\emph{Symbolic weighted parsing} is the problem, 
+given the above input, 
+to find a tree $t \in \T(\Omega)$ %nested word $t \in \Omega$ 
+minimizing \( d(s, t) \otimes A(t)\)
 \wrt $\leq_\oplus$, 
-\ie such that %Hence, it is the problem of finding 
-\( T(s, t) \otimes A(t) = 
-  \displaystyle\bigoplus_{t' \in \Omega^*} T(s, t') \otimes A(t') \).
-
-\noindent 
-Therefore, it is the problem of optimizing a measure 
-called the \emph{edit-distance between $s$ and $A$} in~\cite{Mohri03EDWA}.
-%
-The input language can also be expressed as a \SWTA, or, 
-as a particular case, as a weighted context-free grammar, 
-converted in turn into a \SWVPA following Lemma~\ref{lem:SWTA}.
-%
-In the case of finite alphabets, 
-the problem of searching, in a WTA language, 
-for the best parse tree for a given,  
+\ie such that: %Hence, it is the problem of finding 
+%
+\begin{equation}\label{eq:distance-lang}
+d(s, t) \otimes A(t) = \displaystyle\bigoplus_{t' \in \T(\Omega)} d(s, t') \otimes A(t') 
+\end{equation}
+%
+The measure expressed in~\eqref{eq:distance-lang} is called
+called edit-distance between $s$ and $A$ in~\cite{Mohri03EDWA}.
+%
+%The input language can also be expressed as a \SWTA, or, 
+%as a particular case, as a weighted context-free grammar, 
+%converted in turn into a \SWVPA following Lemma~\ref{lem:SWTA}.
+%
+The problem of searching, in a WTA language, 
+the best parse tree matching a given input,  
 sometimes referred as~\emph{weighted parsing}
-(see \cite{Goodman99SemiringParsing,MorbitzVogler19weighted-parsing} **more general problems**)
-is a particular case of SW parsing.
-%
-Indeed, it corresponds to the case where $T$
-accepts only the pairs $\<s, t>$ such that 
-$s$ is the projection of $t$ on $\Omegai$. 
-This can be done with a single state $q$ and 
-with transition rules of the form:
-\begin{description}
-\item[] $\wei(q, \epsilon, a, q) = \one$ for all $a \in \Omegac \cup \Omegar$,
-\item[] $\wei(q, a, a, q) = \one$ for all $a \in \Omegai$,
-\item[] $\wei(q, a, b, q) = \zero$ for all $a, b \in \Omegai$, $a \neq b$.
-\end{description}
+corresponds to SW parsing in the case of finite alphabets
+and when the transducer $T$ characterizes identity
+see \eg~\cite{Goodman99SemiringParsing} 
+and~\cite{MorbitzVogler19weighted-parsing} for a more general 
+weighted parsing framework.
+%
+%Indeed, it corresponds to the case where $T$
+%accepts only the pairs $\<s, t>$ such that 
+%$s$ is the projection of $t$ on $\Omegai$. 
+%This can be done with a single state $q$ and 
+%with transition rules of the form:
+%\begin{description}
+%\item[] $\wei(q, \epsilon, a, q) = \one$ for all $a \in \Omegac \cup \Omegar$,
+%\item[] $\wei(q, a, a, q) = \one$ for all $a \in \Omegai$,
+%\item[] $\wei(q, a, b, q) = \zero$ for all $a, b \in \Omegai$, $a \neq b$.
+%\end{description}
 
@@
 The problem of Symbolic Weighted  parsing 
-can be solved in PTIME in the size of the input \SWT, \SWVPA (or \SWTA) and input word, 
+can be solved in PTIME in the size of the input \SWT $T$, \SWTA $A$ %\SWVPA (or ) 
+and input word $s$, 
 and the computation time of the functions of the label theory.
 \end{proposition}
-%
-%\begin{proof}
-%Bar-Hillel construction
-%\qed  
-%\end{proof}
-
+\begin{proof} (sketch)
+We follow a \emph{Bar-Hillel} construction, also called parsing by intersection.
+
+\noindent
+We first extend the \SWT $T$ over $\Sigma$, $\Omegai$, $\Semiring$, and $\bar\Phi$, 
+into a \SWT $T'$ over $\Sigma$ and $\hat\Omega$ (and the same semiring and label theory),
+such that for all $s \in \Sigma^*$, and $u \in {\hat\Omega}^*$, 
+$T'(s, u) = T(s, u|_{\Omegai})$. The transducer $T'$ simplify skips every symbol 
+$b \in {\hat\Omega} \setminus \Omegai$ in input, 
+with new transitions of the form $\wei_{01}(q, \epsilon, b, q')$.
+
+\noindent
+Then, given an input word $s \in \Sigma^*$, 
+we compute the \SWA $A_{T', s}$, using Proposition~\ref{prop:epsilon}.
+This automaton is such that for all $t \in \T(\Omega)$, 
+$A_{T', s}\bigl(\lin(t)\bigr) 
+ = T'\bigl(s, \lin(t)\bigr) 
+ = T'\bigl(s, \lin(t)|_{\Omegai}\bigr) 
+ = d(s, t)$.
+
+\noindent
+Next, we convert the input \SWTA $A$ over $\Omega$
+into a \SWVPA $A'$ over $\hat\Omega$, using Lemma~\ref{lem:SWTA}, 
+and we compute the \SWVPA $A_{T', s} \otimes A'$, 
+using Proposition~\ref{prop:SWVPA-product}.
+
+\noindent
+It remains to compute a best nested-word $w \in {\hat\Omega}^*$ 
+using the best-search procedure of Proposition~\ref{th:best-search},
+and convert it into a best tree in $\T(\Omega)$ in order to solve SW parsing
+for~$T$, $A$ and~$s$.
+\qed  
+\end{proof}
+
+\subsection{Application to Automated Music Transcription}
 %\input{transcription}

### 2021-05-25 main.tex
merge with V2

@@
+% fork of main.tex with more expressive models
+% SWT read both tapes (under 2 heads) at every transition (even epsilon or semi-epsilon)
+% SWA read tape (under head) at every transition (even epsilon)
+% SW-VPA read top of stack (when not empty) at every transition
+
 \documentclass[runningheads]{llncs}
@@
 \def\Phir{{\Phi_\mathsf{r}}}
+\def\Phix{{\Phi_\mathsf{x}}}
+\def\Phici{{\Phi_\mathsf{ci}}}
+\def\Phicc{{\Phi_\mathsf{cc}}}
 \def\Phicr{{\Phi_\mathsf{cr}}}
+\def\Phicx{{\Phi_\mathsf{cx}}}
 \def\weii{{\wei_\mathsf{i}}}
@@
 \input{introduction}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% prelin
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Preliminary Notions}
+\label{section:prelim}\label{sec:prelim}
+
+notations: for set $S$ : $S^*$ and $S^+$.
+interval $[i..j]$ of natural numbers
+
+\subsection{Semirings} 
+\label{section:semiring}\label{sec:semiring}
+\input{semiring}
+
+\subsection{Label Theory}
+\label{section:symbols}
+\input{labels}
 
@@
 
-\subsection{Semirings} 
-\label{section:semiring}
-\input{semiring}
-
-\subsection{Label Theory}
-\label{section:symbols}
-\input{labels}
-
-\subsection{Definitions} \label{sec:SWTdef}\label{sec:SWAdef}
+%\subsection{Definitions} \label{sec:SWTdef}\label{sec:SWAdef}
 %(SWT)
 
-Let $\Semiring$ be a commutative and complete semiring, 
+Let $\Semiring$ be a commutative semiring, 
 $\Sigma$ and $\Delta$ be alphabets called respectively \emph{input} and \emph{output}, %{alphabets}, 
-and $\bar\Phi = \< \Phi_\Sigma, \Phi_\Delta, \Phi_{\Sigma, \Delta}>$ be an $\Semiring$-label theory.
+and $\bar\Phi = \< \Phi_\Sigma, \Phi_\Delta, \Phi_{\Sigma, \Delta}>$ be a label theory over $\Semiring$.
 
@@
 where $Q$ is a finite set of states, 
-$\mathsf{in} : Q \to \Semiring$, 
-respectively $\mathsf{out} : Q \to \Semiring,$
+$\mathsf{in} : Q \to \Phi_{\Sigma, \Delta}$, 
+respectively $\mathsf{out} : Q \to \Phi_{\Sigma, \Delta}$,
 are functions defining the weight for entering, 
 respectively leaving, a state, 
-and $\bar{\wei}$ is a tuplet made of the 4 of transition functions 
-$\wei_\epsilon$, $\wei_\Sigma$, $\wei_\Delta$, and $\wei_{\Sigma, \Delta}$
-from $Q \times Q$ into respectively~$\Semiring$, %$\Phi_\epsilon$, 
-$\Phi_\Sigma$, $\Phi_\Delta$, and $\Phi_{\Sigma, \Delta}$.
+and $\bar{\wei}$ is a triplet of transition functions 
+$\wei_{10}$, $\wei_{01}$,  and $\wei_{11}$
+%$\wei_{i, j}: Q \times Q \to \Phi_{\Sigma, \Delta}$, with $i, j \in \{ 0, 1\}$.
+from $Q \times Q$ into $\Phi_{\Sigma, \Delta}$.
 \end{definition}
-%
-A pair of states $\< q, q'>$ such that $\wei_\epsilon(q, q') \neq \zero$
-is called an \emph{$\epsilon$-transition}.
-
-\noindent We synthesize the 4-uplet of transition functions into a unique 
-function~$\wei:
- Q \times (\Sigma \cup \{ \epsilon \}) \times (\Delta \cup \{ \epsilon \}) \times Q 
- \to \Semiring$, %also called $\wei$ for simplicity, 
-such that, for all $q, q' \in Q$, $a \in \Sigma$,  $b \in \Delta$, 
-%and with 
-%$\< \phi_\epsilon, \phi_\Sigma, \phi_\Delta, \phi_{\Sigma, \Delta}> = \wei(q, q')$, 
-\[
-\begin{array}{rcll}
-\wei(q, \epsilon, \epsilon, q') & = &  \wei_\epsilon(q, q'),\\ %\phi_\epsilon\\
-\wei(q, a, \epsilon, q') & = & \phi(a) & %\wei_\Sigma(q, q')(a)
-\quad\mathrm{where~} \phi = \wei_\Sigma(q, q') \in \Phi_\Sigma,\\
-\wei(q, \epsilon, b, q') & = & \psi(b) & 
-\quad\mathrm{where~} \psi = \wei_\Delta(q, q') \in \Phi_\Delta,\\
-\wei(q, a, b, q') & = & \eta(a, b) & 
-\quad\mathrm{where~} \eta = \wei_{\Sigma, \Delta}(q, q') \in \Phi_{\Sigma, \Delta}.\\
-\end{array}      
-\]
-%More precisely, $Q \times Q$, 
-%resp. $Q \times \Sigma \times Q$,
-%$Q \times \Delta \times Q$,
-%$Q \times \Sigma \times \Delta \times Q$,
-%into resp. $\Phi_0$, $\Phi_\Sigma$, $\Phi_\Delta$ $\Phi_{\Sigma, \Delta}$.
-
-\noindent
-The symbolic-weighted transducer~$T$ defines a mapping 
-from the pairs of strings of $\Sigma^* \times \Delta^*$ 
+
+%A pair of states $\< q, q'>$ such that $\wei_{00}(q, q') \neq \zero$
+%is called an \emph{$\epsilon$-transition}.
+%
+For convenience, when $\wei_{ij}(q, q') = \eta \in \Phi_{\Sigma, \Delta}$, 
+for $q, q' \in Q$ and $i, j \in \{ 0, 1\} \setminus \{ \<0, 0> \}$,
+respectively $\mathsf{in}(q) = \eta$, $\mathsf{out}(q') = \eta$,
+we shall sometimes write, overloading the function names:
+$\wei_{ij}(q, a, b, q') = \eta(a, b)$ for $a \in \Sigma$,  $b \in \Delta$,
+respectively $\mathsf{in}(q, a, b) = \eta(a, b)$,
+$\mathsf{out}(q', a, b) = \eta(a, b)$.
+
+\noindent 
+The \SWT $T$ computes on pairs of words $\< s, t> \in \Sigma^+ \times \Delta^+$,
+$s$ and $t$, being respectively called input and output word.
+%
+% \noindent
+More precisely, the symbolic-weighted transducer~$T$ defines a mapping 
+from the pairs of strings of $\Sigma^+ \times \Delta^+$ 
 into~$\Semiring$,
@@
 %$a \in \Sigma$, $b \in \Delta$ 
-for every strings of $s \in \Sigma^*$, $t \in \Delta^*$:
-\[
-\begin{array}{rccl}
-\weight_T(q, \epsilon, \epsilon, q') & = &              \wei(q, \epsilon, \epsilon, q')\\
-\weight_T(q, s, t, q') & = & 
-             \displaystyle\bigoplus_{q'' \in Q} &
-             \wei(q, \epsilon, \epsilon, q'') \otimes \weight_T(q'', s, t, q')\\
- & \oplus &  \displaystyle\bigoplus_{q'' \in Q} &
-             \wei(q'', \epsilon, \epsilon, q') \otimes \weight_T(q, s, t, q'')\\
- & \oplus &  \displaystyle\bigoplus_{\begin{array}{c}
-                                     \scriptstyle q'' \in Q\\[-2pt]
-                                     \scriptstyle s = au, a \in \Sigma
-                                     \end{array}} &
-   \wei(q, a, \epsilon, q'') \otimes \weight_T(q'', u, t, q')\\
- & \oplus & \displaystyle\bigoplus_{\begin{array}{c}
-                                    \scriptstyle q'' \in Q\\[-2pt]
-                                    \scriptstyle t = bv, b \in \Delta
-                                    \end{array}} &
-    \wei(q, \epsilon, b, q'') \otimes \weight_T(q'', s, v, q')\\ 
- & \oplus & \displaystyle\bigoplus_{\begin{array}{c}
-                                    \scriptstyle q'' \in Q\\[-2pt]
-                                    \scriptstyle s = au, a \in \Sigma\\[-2pt]
-                                    \scriptstyle t = bv, b \in \Delta
-                                    \end{array}} &
-   \wei(q, a, b, q'') \otimes \weight_T(q'', u, v, q').\\ 
+%for every strings of $s \in \Sigma^*$, $t \in \Delta^*$:
+$a \in \Sigma$, $u \in \Sigma^*$, $b\in \Delta$, $v\in \Delta^*$,
+
+\begin{equation}
+\begin{array}[t]{lccl}
+%\weight_T(q, \epsilon, \epsilon, q)  & = & \multicolumn{2}{l}{\one}\\ % or \final(q)
+%\weight_T(q, \epsilon, \epsilon, q') & = & \multicolumn{2}{l}{\zero \mathrm{~if~} q \neq q'}\\
+\weight_T(q, au, bv, q') & = & 
+             \displaystyle\bigoplus_{q'' \in Q, u \neq \epsilon} &
+    \wei_{10}(q, a, b, q'') \otimes \weight_T(q'', u, bv, q')\\
+  & \oplus & \displaystyle\bigoplus_{q'' \in Q, v \neq \epsilon} &
+    \wei_{01}(q, a, b, q'') \otimes \weight_T(q'', au, v, q')\\
+  & \oplus & \displaystyle\bigoplus_{q'' \in Q, u,v \neq \epsilon} &
+    \wei_{11}(q, a, b, q'') \otimes \weight_T(q'', u, v, q')\\
+  & \oplus & \displaystyle\bigoplus_{u,v = \epsilon} &
+    \wei_{11}(q, a, b, q')
 \end{array}
-\]
-%
-We recall that, by convention, an empty sum with $\bigoplus$ is $\zero$. 
-%
-The above sum may be infinite because of the $\epsilon$-transitions, % in the first and second lines.
-in this case, the infinite sum operators applies since $\Semiring$ is assumed complete.
-As we shall see, under conditions, the $\epsilon$-transitions can be removed and hence the sum made finite.
-%
-%** The application of $\oplus$ can be seen as a ND choice between transitions **
-% choose to advance head in input, in output or both (or none)
+\label{eq:SWT-weight}
+\end{equation}
+%
+We recall that, by convention (Section~\ref{sec:semiring}), 
+an empty sum with $\bigoplus$ is equal to~$\zero$. 
+%
+Intuitively, using a transition $\wei_{ij}(q, a, b, q')$ means for $T$:
+when reading respectively $a$ and $b$ at the current positions in the input and output words, 
+increment the current position in the input word if if $i = 1$, 
+and in the output word if $j = 1$ (otherwise, do not change it),
+and change state from $q$ to $q'$.
+%
+In contrast with the models of weighted transducers over finite alphabets~\cite{Mohri03ijfcs},
+the input and output symbols at current positions are always read by transitions, 
+even when they do not change the reading position the head's position.
+This is an important feature in the case of an infinite alphabet in 
+order to compare input and output symbols.
+%which cannot be stored in the finite memory of the transducer.
 
 Since $\zero$ is absorbing for~$\otimes$ in~$\Semiring$,
-one term $\wei(q, a, b, q'')$ equal to $\zero$ in the above expression 
+one term $\wei_{ij}(q, a, b, q'')$ equal to $\zero$ in the above expression 
 will be ignored in the sum, meaning that there is no possible transition
-from state $q$ into state $q'$ while reading $a$ and writing $b$ 
-(including the case $a = \epsilon$ and $b = \epsilon$).
-%These functions $\phi$ act as guards for the transducer's transitions, 
-%preventing a transition when they return the absorbing $\zero$ of $\Semiring$.
-This is analogous to a transition's guard not satisfied by $\<a, b>$ in 
-the case of symbolic transducers.
+from state $q$ into state $q'$ while reading $a$ and $b$.
+This is analogous to the case of a transition's guard not satisfied by $\<a, b>$ for 
+symbolic transducers.
+
+%whereas considering the current symbol may be useful to compute a transition weight.
+%(even when it does not change the head's position, like with $\epsilon$-transitions).
+%
+%The cases $\weight_T(q, au, \epsilon, q')$ and $\weight_T(q, \epsilon, bv, q')$
+%are missing in the definition of $\weight_T$.
+%It means that $T$ must avoid configurations where it reached the end of 
+%the output word and not of the input one, or vice-versa.
+%This can be done by using $\wei_{10}$ and $\wei_{01}$
+%before reaching the end of word, and using a special state for this purpose.
 
@@
 Intuitively, 
-$\wei(q, a, \epsilon, q'')$, in the third line, is the cost of 
+$\wei_{10}(q, a, b, r)$ is the cost of 
 the deletion of the symbol~$a \in \Sigma$ in~$s$, 
-$\wei(q, \epsilon, b, q'')$, in the fourth line, is the cost 
+$\wei{01}(q, a, b, r)$ is the cost 
 of the insertion of~$b \in \Delta$ in $t$, 
-and $\wei(q, a, b, q'')$, in the last line, is the cost 
+and $\wei_{11}(q, a, b, r)$ is the cost 
 of the substitution of  $a \in \Sigma$ by~$b \in \Delta$.
+%
 The cost of a sequence of such operations transforming $s$ into $t$, 
@@
 T(s, t)  = 
-\displaystyle\bigoplus_{q, q' \in Q} \mathsf{in}(q) 
-\mathop{\otimes} \weight_T(q, s, t, q') \mathop{\otimes} \mathsf{out}(q').
-\label{eq:weightT}
+\displaystyle\bigoplus_{q, q' \in Q} \mathsf{in}(q, s_1, t_1) 
+\mathop{\otimes} \weight_T(q, s, t, q') \mathop{\otimes} \mathsf{out}(q', s_n, t_m)
+\label{eq:SWT-value}
 \end{equation}
 
-% DTW for sequences of timestamped events 
-% impossible, needs reading input/output symbols even by epsilon-transitions.
+\begin{example}
+We can define with a \SWT the computation of a similarity measure 
+between timed sequences
+similar to dynamic time warping (DTW).
+
+Let $\Semiring$  be the tropical (\emph{min-plus}) semiring of Figure~\ref{fig:semirings} and 
+let $\Sigma = \Delta = \mathbb{R}_+$ be sets of timestamps.
+We consider a \SWT with one state $q$ and transitions
+$\wei_{11}(q, d, d', q) = 
+ \wei_{01}(q, d, d', q) = 
+ \wei_{10}(q, d, d', q) = |d' - d|$,
+for all $d, d' \in \mathbb{R}_+$.
+% needs reading input/output symbols even by epsilon-transitions.
+The recursive definition of $\weight_T$ correspond to the dynamic programming equations of DTW
+for the computation of an optimal match between words, 
+the matching cost for two symbols being the 
+the time distance between them.
+\endex
+\end{example}
 
@@
 \wei_{11}(q_1, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d|\\
-\wei_{01}(q_0, \epsilon, \< \mathsf{-}, d'>, q_0) & = & \one & &
-\wei_{01}(q_1, \epsilon, \< \mathsf{-}, d'>, q_0) & = & \one\\
-\wei_{10}(q_0, d, \epsilon, q_1) & = & \alpha\\
+\wei_{01}(q_0, d, \< \mathsf{-}, d'>, q_0) & = & \one & &
+\wei_{01}(q_1, d, \< \mathsf{-}, d'>, q_0) & = & \one\\
+\wei_{10}(q_0, d, b, q_1) & = & \alpha & & \multicolumn{3}{l}{\mathrm{for~all~} b \in \Delta}
 \end{array}
@@
 %
-Finally, it may happen that the performers plays an extra note accidentally, but only once in a row. 
+Moreover, it may happen that the performers plays an extra note accidentally, but only once in a row. 
 This is modelled by the transition $\wei_{10}$ with an arbitrary weight value $\alpha \in \Semiring$, 
-switch from state $q_0$ (normal) to $q_1$ (error).
+switching from state $q_0$ (normal) to $q_1$ (error).
 The transitions in the second column below switch back to the normal state $q_0$.
@@
 % between dates of input and output events.
+At last, we let $q_0$ be the only initial and final state, with
+$\init(q_0, d, b) = \final(q_0, d, b) = \one$, and 
+$\init(q_1, d, b) = \final(q_1, d, b) = \zero$, 
+for all $d \in \Sigma$ and $b \in \Delta$.
 \endex
@@
 is a tuple
-$A = \< Q, \init, \bar{\wei}, \final >$,
+$A = \< Q, \init, {\wei_1}, \final >$,
 where $Q$ is a finite set of states, 
-$\mathsf{in} : Q \to \Semiring$, 
-respectively $\mathsf{out} : Q \to \Semiring,$
+$\mathsf{in} : Q \to \Phi_\Sigma$, 
+respectively $\mathsf{out} : Q \to \Phi_\Sigma,$
 are functions defining the weight for entering, 
 respectively leaving, a state, 
-and $\bar{\wei}$ is a pair of transition functions 
-$\wei_\epsilon$ and $\wei_\Sigma$ from $Q \times Q$ into 
-respectively $\Semiring$ %$\Phi_\epsilon$ 
-and $\Phi_\Sigma$.
+and ${\wei_1}$ is a transition functions 
+from $Q \times Q$ into $\Phi_{\Sigma}$.
 \end{definition}
-
+%      
 \noindent
 As above in the case of \SWT, 
-the transition functions $\wei_\epsilon$ and $\wei_\Sigma$
-can be synthesized into a single function
-$\wei: Q \times (\Sigma \cup \{ \epsilon \}) \times Q \to \Semiring$.
+when $\wei_1(q, q') = \phi \in \Phi_\Sigma$, 
+respectively $\mathsf{in}(q) = \phi$, $\mathsf{out}(q') = \phi$,
+we may write $\wei_1(q, a, q')$ for $\phi(a)$, 
+respectively $\mathsf{in}(q, a) = \phi(a)$, $\mathsf{out}(q', a) = \phi(a)$.
+%$\wei_1: Q \times \Sigma \times Q \to \Semiring$, 
+The computation of $A$ on words $s \in \Sigma^+$
+is defined with an intermediate function $\weight_A$, 
+defined as follows for $q, q' \in Q$, $a \in \Sigma$, $u \in \Sigma^*$,
+%
+\begin{equation}
+\begin{array}{lccl}
+%\weight_T(q, \epsilon, \epsilon, q)  & = & \multicolumn{2}{l}{\one}\\ % or \final(q)
+%\weight_T(q, \epsilon, \epsilon, q') & = & \multicolumn{2}{l}{\zero \mathrm{~if~} q \neq q'}\\
+\weight_A(q, au, q') & = & 
+            \displaystyle\bigoplus_{r \in Q, u \neq \epsilon} &
+    \wei_{1}(q, a, r) \otimes \weight_A(r, u, q')\\
+  & \oplus & \displaystyle\bigoplus_{u = \epsilon} & 
+    \wei_{1}(q, a, q')
+\end{array}
+\label{eq:SWA-weight}
+\end{equation}
+%
+\noindent
+and the weight value associated by $A$ to 
+$s = s_1\ldots s_n \in \Sigma^+$ is: % defined as follows:
+\begin{equation}
+A(s)  = 
+\displaystyle\bigoplus_{q, q' \in Q} \mathsf{in}(q, s_1) 
+\mathop{\otimes} \weight_A(q, s, q') \mathop{\otimes} \mathsf{out}(q', s_n)
+\label{eq:SWA-value}
+\end{equation}
+
 %
@@
 
-\subsection{Properties}
-The two following properties will be useful to our approach on 
-symbolic weighted parsing in Section~\ref{sec:parsing}.
-
-\begin{proposition}
-Given a \SWT $T$ 
-over alphabet $\Sigma$, $\Delta$ and a commutative semiring~$\Semiring$, 
-and $s \in \Sigma^*$, 
+%\subsection{Properties}
+\noindent
+The following property will be useful to the approach on 
+symbolic weighted parsing presented in Section~\ref{sec:parsing}.
+
+\begin{proposition} \label{prop:epsilon}
+Given a \SWT $T$ over $\Sigma$, $\Delta$, $\Semiring$ and $\bar\Phi$,
+and $s \in \Sigma^+$, 
 there exists an effectively constructible \SWA 
-$A_{T, s}$ over $\Delta$ and $\Semiring$ 
-such that for all $t \in \Delta^*$, $A_{T, s}(t) = T(s, t)$.
+$A_{T, s}$ over $\Delta$, $\Semiring$ and $\bar\Phi$,
+such that for all $t \in \Delta^+$, $A_{T, s}(t) = T(s, t)$.
 \end{proposition}
@@
 let 
-$\wei: Q \times (\Sigma \cup \{ \epsilon \}) \times (\Delta \cup \{ \epsilon \}) \times Q 
- \to \Semiring$
+$\wei: Q \times \Sigma \times \Delta \times Q \to \Semiring$
 be the synthesized form of 
-$\bar{\wei} = \< \wei_\epsilon, \wei_\Sigma, \wei_\Delta, \wei_{\Sigma, \Delta}>$,
-and let $s = s_1 \ldots s_n$ with $s_1, \ldots, s_n \in \Sigma$.
+$\bar{\wei} = \< \wei_{10}, \wei_{01}, \wei_{11}>$,
+and let $s = s_1 \ldots s_n$ with $s_1, \ldots, s_n \in \Sigma$ ($n \geq 1$).
 %$n = |s|$ be the length of $s$.
-%
-The state set of $A_{T, s}$ will be $Q' = [1..n] \times Q$.
-Its state entering weight function is defined by 
-$\init'(1, q) = \init(q)$ for all $q \in Q$, 
-and $\init'(i, q) = \zero$ for all $1 < i \leq n$, 
-and its state leaving weight function by 
-$\final'(n, q) = \final(q)$ or all $q \in Q$, 
-and $\final'(i, q) = \zero$ for all $1 \leq i <  n$.
-
-\noindent
-We define transition functions $\wei'_\epsilon$ and $\wei'_\Sigma$ for $A_{T, s}$, 
-together with their synthesized form $\wei'$ as follows: 
-for all $1 \leq i, j \leq n$, $q, q' \in Q$, %and $a \in \Sigma$, 
-and $b \in \Delta$, 
+
+The state set of $A_{T, s}$ will be $Q' = [1..n+1] \times Q$.
+Its state entering weight function is defined by, 
+for all $q \in Q$ and $b \in \Delta$,  
+$\init'(\< 1, q>, b) = \init(q, s_1, b)$ 
+and $\init'(\< i, q>, b) = \zero$ for all $1 < i \leq n+1$. 
+Its state leaving weight function is defined by, for all $q \in Q$ and $b \in \Delta$,  
+$\final'(\< n+1, q>, b) = \final(q, s_n, b)$, 
+and $\final'(\< i, q>, b) = \zero$ for all $1 \leq i <  n+1$.
+
+Every non-null transition of $A_{T, s}$ will
+simulate a sequence of transitions of $T$ which advance in 
+the input word while staying immobile in the output word, 
+and then make one step in the output word (and advance in the input word or not).
+%
+The above initial sequence correspond to $\epsilon$-transitions of automata, 
+and its total weight is computed by the following intermediate function 
+$\wei'_0 : Q' \times Q' \to \Phi_{\Sigma}$, 
+defined for all $q, q' \in Q$: % all $1 \leq i < n$ and $1 \leq k \leq n-i$, 
+by:
 \[
-\begin{array}{lclcrcll}
-\wei'_\epsilon\bigl(\< i, q>, \< i, q'>\bigr) & = & \wei_\epsilon(q, q') & \mathrm{\ie} &
-\wei'\bigl(\< i, q>, \epsilon, \< i, q'>\bigr)   & = & \wei(q, \epsilon, \epsilon, q'), & \\ 
-%
-\wei'_\epsilon\bigl(\< i, q>, \< i+1, q'>\bigr) & = & \phi(s_i) & 
-\multicolumn{5}{l}{\mathrm{where~} \phi = \wei_\Sigma(q, q'),}\\
- & & & \mathrm{\ie} &
-\wei'\bigl(\< i, q>, \epsilon, \< i+1, q'>\bigr) & = & \wei(q, s_i, \epsilon, q') & 
-\mathrm{if~} i < n,\\ 
-%
-\wei'_\epsilon\bigl(\< i, q>, \< j, q'>\bigr) & = & \zero & 
-%\mathrm{\ie} & \wei'(\< i, q>, \epsilon, \< j, q'>)   & = & \zero & & 
-\multicolumn{5}{l}{\mathrm{if~} j \neq i, i+1,}\\ 
-%
-\wei'_\Sigma\bigl(\< i, q>, \< i, q'>\bigr) & = & \wei_\Delta(q, q') & \mathrm{\ie} &
-\wei'\bigl(\< i, q>, b, \< i, q'>\bigr)          & = & \wei(q, \epsilon, b, q'),\\ 
-%
-\wei'_\Sigma\bigl(\< i, q>, \< i+1, q'>\bigr) & = & \eta_{s_i} &
-\multicolumn{5}{l}{\mathrm{where~} \eta = \wei_{\Sigma, \Delta}(q, q'),}\\
- & & & \mathrm{\ie} & \wei'\bigl(\< i, q>, b, \< i+1, q'>\bigr) & = & \wei(q, s_i, b, q') & 
- \mathrm{if~} i < n,\\ 
-%
-\wei'_\Sigma\bigl(\< i, q>, \< j, q'>\bigr) & = & \zero & \mathrm{\ie} &
-\wei'\bigl(\< i, q>, b, \< j, q'>\bigr)     & = & \multicolumn{2}{l}{\zero\ \mathrm{~if~} j \neq i, i+1.}
+\begin{array}{rcl}
+\wei'_0\bigl(\<i, q>, \< i, q>\bigr) & = & \one 
+\quad \mathrm{if~} 1 \leq i \leq n+1,\\
+%
+\wei'_0\bigl(\<i, q>, \< i, q'>\bigr) & = & \zero 
+\quad \mathrm{if~} 1 \leq i \leq n+1 \mathrm{~and~} q \neq q',\\
+%
+\wei'_0\bigl(\<i, q>, \< i+k, q'>\bigr) & = & 
+\displaystyle
+\bigoplus_{\begin{array}{c}
+           \scriptstyle q_0,\ldots, q_k \in Q\\[-2pt]
+           \scriptstyle q_0 = q, q_k = q'
+           \end{array}} 
+\bigotimes_{j=0}^{k-1} \bigl[\wei_{10}(q_j, q_{j+1})\bigr]_{s_{i+j}}
+\quad \begin{array}[t]{l}
+      \mathrm{if~} 1 \leq i < n\\
+      \mathrm{and~} 1 \leq k \leq n-i.
+      \end{array}
 \end{array}
 \]
-We use the assumption that the underlying label theory of $T$ is 
-is closed under partial applications (Section~\ref{section:symbols})
-in the above construction of $\wei'_\Sigma$.
-
-\noindent
-The $\SWA$ wanted is $A_{T, s} = \< Q', \init', {\bar\wei}', \final' >$.
+where $\bigl[\wei_{10}(q_j, q_{j+1})\bigr]_{s_{i+j}}$ 
+is the partial application $\eta_{s_{i+j}} \in \Phi_{\Delta}$
+for $\eta = \wei_{10}(q_j, q_{j+1}) \in \Phi_{\Sigma, \Delta}$ ($s_{i+j} \in \Sigma$).
+%
+The function $\wei'_0$ is defined thanks to the closure properties of the label theory $\bar\Phi$
+(Section~\ref{section:symbols}). 
+% closed under partial applications 
+The sum and product in its definition are finite, 
+we shall see below how to compute the first sum in polynomial time.
+%The effective computation of $\wei'_0$ does not require enumerating all the state sequences 
+%of length $k+1$ in the sum.
+%Indeed, it amounts to search short
+
+\noindent
+We define now the transition function 
+$\wei'_1: Q' \times Q' \to \Phi_\Sigma$
+of $A_{T, s}$ as follows,
+for $q, q' \in Q$,
+$1 \leq i \leq n$, and $0 \leq k \leq n-i$:
+\[
+\begin{array}{lcl}
+\wei'_1\bigl(\< i, q>, \< i, q'>\bigr) & = & \wei_{01}(q, q')\\
+%
+\wei'_1\bigl(\< i, q>, \< i+k, q'>\bigr) & = & 
+\displaystyle
+\bigoplus_{\begin{array}{c}
+           \scriptstyle q'' \in Q\\[-2pt]
+           \scriptstyle i+k < n
+           \end{array}} 
+\wei'_0\bigl(\<i, q>, \< i+k, q''>\bigr) \otimes \eta_{s_{i+k}}
+\quad \mathrm{where~} \eta = \wei_{01}(q'', q'),\\
+ & \oplus & \displaystyle
+\bigoplus_{q'' \in Q} \wei'_0\bigl(\<i, q>, \< i+k-1, q''>\bigr) \otimes \eta'_{s_{i+k}}
+\quad \mathrm{where~}\eta' = \wei_{11}(q'', q'),\\
+%
+\wei'_1\bigl(\< i, q>, \< j, q'>\bigr) & = & \zero  
+\quad \mathrm{if~} j < i.
+\end{array}      
+\]
+%
+
+\noindent
+We can show that the $\SWA$ $A_{T, s} = \< Q', \init', \wei'_1, \final' >$
+has the expected property: $\forall t \in \Delta^+, A_{T, s}(t) = T(s, t)$.
 \qed
@@
 
+\noindent
+On the quadratic computation of $\wei'_0$. **
+by a best path search in the graph with nodes $[i..i+k] \times Q$
+and edges labeled in $\Phi_\Sigma$...
+ordering = summary.
+
+\noindent
 The construction time and size for $A_{T, s}$ are $O(\| T \| . | s |)$,
 where the size $\| T \|$ of $T$ is its number of states $|Q|$.
-
-% gen. to transducers
-\begin{proposition}
-Given a \SWA $A$ over alphabet $\Sigma$ 
-and a commutative and bounded semiring $\Semiring$, 
-there exists an effectively constructible \SWA $A'$ 
-without $\epsilon$-transitions 
-such that for all $s \in \Sigma^*$, $A'(s) = A(s)$.
-\end{proposition}
-%
-%\begin{proof}  ** Tb revised **   
-%Let $A = \< Q, \init, \<\wei_\epsilon, \wei_\Sigma>, \final >$.
-%The automaton $A'$ is $\< Q, \init, \<\wei'_\epsilon, \wei'_\Sigma>, \final >$,
-%where, for all $q, q' \in Q$,
-%$\wei'_\epsilon(q, q') = \zero$  
-%and 
-%% \[
-%% \begin{array}{rcccl}
-%% \wei'_\Sigma(q, q') & = &  & \displaystyle\bigoplus_{q_1 \in Q} &
-%%                \wei_\Sigma(q, q_1) \otimes  \wei_\epsilon(q_1, q')\\
-%%                     &   & \oplus  & \displaystyle\bigoplus_{q_2 \in Q} &
-%%                \wei_\epsilon(q, q_2) \otimes \wei_\Sigma(q_2, q')\\
-%% \end{array}
-%% \]
-%\[
-%\wei'_\Sigma(q, q') = 
-%\displaystyle\bigoplus_{q'' \in Q} \wei_\epsilon(q, q'') \otimes \wei_\Sigma(q'', q')
-%\oplus \displaystyle\bigoplus_{q_1 \in Q} \wei_\Sigma(q, q'') \otimes \wei_\epsilon(q'', q')
-%\]
-%In the above definition of $\wei'_\Sigma$ we use the operator
-%of product of function of $\Phi_\Sigma$ by $\Semiring$ 
-%described in Section~\ref{section:symbols}.
-%%
-%By definition of $\weight_{A}$ and distributivity of $\otimes$ on $\oplus$, 
-%** NO. TBC see \cite{Mohri02ijfcs} **
-%it holds that $\weight_{A}(q, s, q') = \weight_{A'}(q, s, q')$.
-%\qed
-%\end{proof}
 
@@
 \label{section:SWVPA}\label{sec:SWVPA}
-The following model generalizes Symbolic VPA~\cite{dAntonyAlur14SVPDA}
+The model presented in this section generalizes Symbolic VPA~\cite{dAntonyAlur14SVPDA}
 from Boolean semirings to arbitrary semiring weight domains.
-
-\subsection{Definition}
+It will compute on nested words over infinite alphabets, 
+associating to every such word a weight value. 
+Nest words are able to describe structures of labeled trees, 
+and in the context of parsing, they will be useful to 
+represent parse trees.
+
+\subsection{Definition}\label{se:SWVPA-def}
 Let $\Omega$ be a countable alphabet 
 %finite (large) or infinite,
-that we assume partitioned into three subsets 
-$\Omegai$, $\Omegac$, $\Omegac$,
+that we assume partitioned into three 
+subsets~$\Omegai$, $\Omegac$, $\Omegac$,
 whose elements are respectively called 
 \emph{internal}, \emph{call} and \emph{return} symbols.
+% names are  coined by application to functional program verification
 % \begin{itemize}
@@
 % \end{itemize}
-Let~$\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative semiring and  
-let  $(\Phi_\epsilon, \Phic, \Phir, \Phicr)$ be a label theory over $\Semiring$
+Let~$\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative and complete semiring and let  
+$\bar\Phi = \< \Phii, \Phic, \Phir, \Phici, \Phicc, \Phicr>$ 
+be a label theory over $\Semiring$
 %In order to simplify notations, %and following the definition of Section~\ref{section:transducer}, 
 %we shall write respectively 
-where $\Phic$, $\Phir$ and~$\Phicr$ stand respectively 
-for~$\Phi_\Omegac$, $\Phi_\Omegar$ and~$\Phi_{\Omegac, \Omegar}$.
-%
-Moreover, we extend this theory with a set $\Phii$ 
-of unary functions in $\Omegai \to \Semiring$,
-closed under $\oplus$ and $\otimes$.
+where $\Phii$, $\Phic$, $\Phir$ and~$\Phicx$ (with $\mathsf{x} \in \{ \mathsf{i}, \mathsf{c}, \mathsf{r}\}$) 
+stand respectively 
+for~$\Phi_\Omegai$, $\Phi_\Omegac$, $\Phi_\Omegar$ and~$\Phi_{\Omegac, \Omega_\mathsf{x}}$.
+%
+%Moreover, we extend this theory with a set $\Phii$ 
+%of unary functions in $\Omegai \to \Semiring$,
+%closed under $\oplus$ and $\otimes$.
 
@@
 A \emph{Symbolic Weighted Visibly Pushdown Automata} (\SWVPA) 
-over  $\Omega = \Omegai \uplus \Omegac \uplus \Omegar$ and $\Semiring$ is a tuple
-$A = \< Q, P, \init, \weii, \weic, \weir, \weie, \final >$,
+over  $\Omega = \Omegai \uplus \Omegac \uplus \Omegar$, $\Semiring$ and $\bar\Phi$ 
+is a tuple $A = \< Q, P, \init, \bar\wei, \final >$,
 where $Q$ is a finite set of states, 
@@
 respectively leaving, a state, 
-and 
-$\weii : Q \times Q \to \Phii$,  
-$\weic : Q \times Q \times P \to \Phic$,  
+and $\bar\wei$ is a sextuplet composed of the transition functions:
+$\weii : Q \times P \times Q \to \Phici$,  
+$\weiei : Q \times Q \to \Phii$,  
+$\weic : Q \times P \times Q \times P \to \Phicc$,  
+$\weiec : Q \times P \times Q \to \Phic$,  
 $\weir : Q \times P \times Q \to \Phicr$,  
-$\weie : Q \times Q \to \Phir$,  
-are transition functions.
+$\weier : Q \times Q \to \Phir$.
+%and 
+%$\weiex : Q \times Q \to \Phix$ 
+%with $\mathsf{x} \in \{ \mathsf{i}, \mathsf{c}, \mathsf{r}\}$.
 \end{definition}
@@
 \begin{array}{lll}
-\weii: Q \times \Omegai \times Q \to \Semiring & 
-\weii(q, a, q') = \phi_\mathsf{i}(a) & 
-\mathrm{where~} \phi_\mathsf{i} = \weii(q, q'),\\
-%
-\weic: Q \times \Omegac \times Q \times P \to \Semiring & 
-\weic(q, \call{c}, q', p) = \phi_\mathsf{c}(\call{c}) & 
-\mathrm{where~} \phi_\mathsf{c} = \weic(q, q', p),\\
+\weii: Q \times \Omegac \times P \times \Omegai \times Q \to \Semiring & 
+\weii(q, c, p, a, q') = \eta_\mathsf{ci}(c, a) & 
+\mathrm{where~} \eta_\mathsf{ci} = \weii(q, p, q'),\\
+%
+\weiei: Q \times \Omegai \times Q \to \Semiring & 
+\weiei(q, a, q') = \phi_\mathsf{i}(a) &
+\mathrm{where~} \phi_\mathsf{i} = \weiei(q, q').\\[2pt]
+%
+\weic: Q \times \Omegac \times P \times  \Omegac \times P \times Q \to \Semiring & 
+\weic(q, c, p, \call{c'}, p', q') = \eta_\mathsf{cc}(c, \call{c'}) & 
+\mathrm{where~} \eta_\mathsf{cc} = \weic(q, p, p', q'),\\
+%
+\weiec: Q \times \Omegac \times P \times Q \to \Semiring & 
+\weiec(q, {\call{c}}, p, q') = \phi_\mathsf{c}({\call{c}}) &
+\mathrm{where~} \phi_\mathsf{c} = \weiec(q, p, q').\\[2pt]
 %
 \weir: Q \times \Omegac \times P \times \Omegar \times Q \to \Semiring & 
-\weir(q, {\call{c}},  p, {\return{r}}, q') = \phi_\mathsf{r}({\call{c}},  {\return{r}}) & 
-\mathrm{where~} \phi_\mathsf{r} = \weir(q, p, q'),\\
-%
-\weie: Q \times \Omegar \times Q \to \Semiring & 
-\weie(q, {\return{r}}, q') = \phi_\mathsf{e}({\return{r}}) &
-\mathrm{where~} \phi_\mathsf{e} = \weie(q, q').\\
+\weir(q, {\call{c}},  p, {\return{r}}, q') = \eta_\mathsf{cr}({\call{c}},  {\return{r}}) & 
+\mathrm{where~} \eta_\mathsf{cr} = \weir(q, p, q'),\\
+%
+\weier: Q \times \Omegar \times Q \to \Semiring & 
+\weier(q, {\return{r}}, q') = \phi_\mathsf{r}({\return{r}}) &
+\mathrm{where~} \phi_\mathsf{r} = \weier(q, q').\\
 \end{array}      
@@
 The intuition is the following for the above transitions.
-\begin{description}
-\item $\weii$ : read the input internal symbol $a$, change state to $q'$ (stack is untouched).
-\item $\weic$ : read the input call symbol $\call{c}$, push it to the stack along with $p$, change state to $q'$.
-\item $\weir$ : when the stack is not empty, 
-      read and pop from stack a pair made of $\call{c}$ and $p$, 
-      read the input return symbol $\return{r}$, change state to $q'$.
-      In this case, the weight function $\phi_\mathsf{r}$ 
-      computes a value of matching between the call and return symbols $c$ and $r$.
-      This value might be set to $\zero$ in order to express that the symbols do not match.
-\item $\weie$ : when the stack is empty, 
-      read the input symbol $\call{r}$, change state to $q'$.
-\end{description}
-
-We give now a formal definition of these transitions of the automaton~$A$
+
+\noindent
+$\weii$ and $\weiei$ both read an input internal symbol $a$ and change state from $q$ to $q'$, 
+without changing the stack. 
+Moreover, $\weii$ reads a pair made of 
+${\call{c}} \in \Omegac$ and $p \in P$ at the top of the stack 
+($c$ is compared to $a$ by the weight function $\eta_\mathsf{ci} \in \Phici$)
+and $\weiei$ applies if and only if the stack is empty.
+
+\noindent
+$\weic$ and $\weiec$ read the input call symbol $\call{c}'$, 
+push it to the stack along with $p'$, and change state from $q$ to to $q'$.
+Moreover, $\weic$ reads ${\call{c}}$ and $p$ at the top of the stack 
+($c$ is compared to $c'$),
+and $\weiec$ applies iff the stack is empty.
+
+\noindent
+$\weir$ and $\weier$ read the input return symbol $\return{r}$, and change state from $q$ to to $q'$.
+Moreover, $\weir$ reads and pop from stack a pair made of $\call{c}$ and $p$, 
+($\call{c}$ is compared to $\return{r}$),
+and $\weier$ applies iff the stack is empty.
+%In this case, the weight function $\phi_\mathsf{r}$ 
+%computes a value of matching between the call and return symbols $c$ and $r$.
+%This value might be set to $\zero$ in order to express that the symbols do not match.
+
+Formally, the transitions of the automaton~$A$ are defined
 in term of %a weight value computed by 
@@
 of a state $q \in Q$ and a stack content $\gamma \in \Gamma^*$, where $\Gamma = \Omegac \times P$.
-Therefore, $\weight_A$ is a function from 
-$Q \times \Gamma^* \times \Omega^* \times Q \times \Gamma^*$ into $\Semiring$.
-\[
-\begin{array}{rcl}
-\weight_A\bigl(\config{q}{\gamma}, a\, u, \config{q'}{\gamma'}\bigr) & = & 
- {\displaystyle\bigoplus_{q'' \in Q}} \weii(q, a, q'') 
+Hence, $\weight_A$ is a function from 
+$[Q \times \Gamma^*] \times \Omega^* \times [Q \times \Gamma^*]$ into~$\Semiring$.
+%
+\begin{equation}
+\begin{array}[t]{rcl}
+\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot\gamma}, a\, u, \config{q'}{\gamma'}\bigr) & = & 
+ {\displaystyle\bigoplus_{q'' \in Q}} \weii(q, c, p, a, q'') 
  \otimes \weight_A\bigl(\config{q''}{\gamma}, u, \config{q'}{\gamma'}\bigr)\\
-%
-\weight_A\bigl(\config{q}{\gamma}, {\call{c}}\, u, \config{q'}{\gamma'}\bigr) & = & 
+\weight_A\bigl(\config{q}{\bot}, a\, u, \config{q'}{\gamma'}\bigr) & = & 
+ {\displaystyle\bigoplus_{q'' \in Q}} \weiei(q, a, q'') 
+ \otimes \weight_A\bigl(\config{q''}{\bot}, u, \config{q'}{\gamma'}\bigr)\\
+%
+\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot\gamma}, {\call{c}'}\, u, \config{q'}{\gamma'}\bigr) & = & 
  {\displaystyle\bigoplus_{\begin{array}{c}
@@
 \weight_A\bigl(\config{q}{\bot}, {\return{r}}\, u, \config{q'}{\gamma'}\bigr) & = & 
- {\displaystyle\bigoplus_{q'' \in Q}} \weie(q, {\return{r}}, q'') 
+ {\displaystyle\bigoplus_{q'' \in Q}} \weier(q, {\return{r}}, q'') 
  \otimes \weight_A\bigl(\config{q''}{\bot}, u, \config{q'}{\gamma'}\bigr)\\
 \end{array}
-\]
+\label{eq:SWVPA-weight}
+\end{equation}
+%
 where $\bot$ denotes the empty stack and ${\call{c}}\, p\cdot \gamma$ 
@@
 
-%\begin{example}
-%structured words...
-%intro language of music notation ?
-%\end{example}
+\begin{example}
+structured words with timed symbols...
+intro language of music notation? (markup = time division, leaves = events etc)
+\end{example}
 
@@
 %
-\begin{proposition}
+\begin{proposition}\label{prop:SWVPA-product}
 Let $A_1$ and $A_2$ be two (\SWVPA)
-over the same $\Omega$ and $\Semiring$.
+over the same $\Omega$, $\Semiring$ and $\bar\Phi$.
 There exists two $\SWVPA$ $A_1 \oplus A_2$ and $A_1 \otimes A_2$, 
@@
 More precisely, 
-let $\top$ be a fresh stack symbol which does not belong to $\Gamma$,
-and for every two states $q, q' \in Q$ and $\sigma \in \{ \bot, \top \}$, 
-let 
-\[ 
-  d_0(q, \sigma, q') = \bigoplus_{s\in \Omega^*} 
+let $b_\bot : Q \times Q \to \Semiring$ be the function:
+%
+\begin{equation}\label{eq:bbot}
+  b_\bot(q, q') = \bigoplus_{s\in \Omega^*} 
   \textstyle
-  \weight_A\bigl(\config{q}{\sigma}, s, \config{q'}{\sigma}\bigr). 
-\]
-Since $\Semiring$ is complete, this infinite sum is well defined, 
-and since $\leq_\oplus$ is assumed total, it is the minimum in $\Omega^*$
-of $s \mapsto \weight_A(\config{q}{\sigma}, s, \config{q'}{\sigma})$ \wrt this ordering.
-%
-When $\sigma = \bot$, $d_0(q, \sigma, q')$ is the central 
-expression in a term of the definition (\ref{eq:weightA}) of $A(s_0)$ for the minimum $s_0$
-(for the above function).
-When $\sigma = \bot$, intuitively, it is the minimum weight of a computation of $A$
-starting in state $q$ with a stack $\gamma \in \Gamma^*$ (possibly empty), 
-and ending in state $q$ with the same stack $\gamma$, such that moreover 
-the computation does not touch a symbol of $\gamma$. 
-That means that during the computation, 
-$A$ may apply the first case of in the definition of $\weight_A$ (internal symbol),
-as well as the second case, to can push call symbols on the top of $\gamma$, 
-and may pop these symbols with the third case (return symbol). 
-However, it cannot apply one of the two last cases (return symbol and empty stack)
-when the current stack is $\gamma$.
+  \weight_A\bigl(\config{q}{\bot}, s, \config{q'}{\bot}\bigr)
+\end{equation}
+%
+Since $\Semiring$ is complete, the infinite sum in \eqref{eq:bbot} is well defined,
+and by totality of $\leq_\oplus$, it is the minimum in $\Omega^*$
+of the fonction 
+$s \mapsto \weight_A(\config{q}{\sigma}, s, \config{q'}{\sigma})$ 
+\wrt this ordering.
+%
+The term $\config{q}{\bot}, s, \config{q'}{\bot}$ 
+of this sum is the central expression in 
+the definition \eqref{eq:weightA} of $A(s_0)$, for the minimum $s_0$
+of the function $\weight_A$.
+
+%
+\noindent
+Let $\top$ be a fresh stack symbol which does not belong to $\Gamma$,
+and let $b_\top : Q \times P \times Q \to \Phic$ be such that,
+for every two states $q, q' \in Q$ 
+and stack symbol $p \in P$: %and $\sigma \in \{ \bot, \top \}$, 
+\begin{equation}\label{eq:btop}
+  b_\top(q, p, q') : c \mapsto \bigoplus_{s\in \Omega^*} 
+  \textstyle
+  \weight_A\bigl(\config{q}{c\,p\cdot \top }, s, \config{q'}{c\, p\cdot \top}\bigr). 
+\end{equation}
+%
+Intuitively, the function defined in \eqref{eq:btop}
+associateds to $c \in \Omegac$ 
+the minimum weight of a computation of $A$
+starting in state $q$ with a stack 
+$c p \cdot \gamma \in \Gamma^+$ 
+and ending in state $q'$ with the same stack,
+such that the computation does pop 
+the pair made of $c$ and $p$ at the top of this stack,
+but may read these symbols.
+Moreover, $A$ may push another pair $\< c', p'>$ %call symbols 
+on the top of $c p \cdot \gamma$,
+following the the third case of 
+in the definition \eqref{eq:SWVPA-weight} of $\weight_A$,
+and may pop $\< c', p'>$ later, following the fifth case of \eqref{eq:SWVPA-weight} (return symbol). 
+%However, it cannot apply one of the two last cases (return symbol and empty stack)
+%when the current stack is $\gamma$.
 %pop symbols in $\gamma$.
@@
 
-The algorithm~\ref{algo:Dijkstra}
-constructs iteratively a marking $d : Q \times \{ \bot, \top \} \times Q \to \Semiring$
+Algorithm~\ref{algo:Dijkstra}
+constructs iteratively markings 
+$d_\bot : Q \times Q \to \Semiring$ and 
+$d_\top : Q \times P \times Q \to \Phic$
 %of the triplets $\<q, \sigma, q'>$ 
 %of states of $A$ by weight values in $\Semiring$, 
-that converges eventually to $d_0(q, \sigma, q')$. 
+that converges eventually to $b_\top$ and $b_\bot$.
 %It uses for that purpose a priority queue $P$ containing triplets of 
@@
 
-\begin{algo}[1-best for \SWVPA] \label{algo:Dijkstra}
-\textbf{initially} let $P = Q \times \{ \bot, \top \} \times Q$, %contains all %triplets 
-%$\< q_1, \bot, q_2>$ and $\<q_1, \top, q_2>$ for $q_1, q_2 \in Q$, with 
-and 
-$d(q_1, \bot, q_2) = d(q_1, \top, q_2) = \one$ if $q_1 = q_2$ and
-$d(q_1, \bot, q_2) = d(q_1, \top, q_2) = \zero$ otherwise.
+\begin{algo}[best search for \SWVPA]\label{algo:Dijkstra}
+\textbf{initially} 
+let $\Q = (Q \times Q) \cup (Q \times P \times Q)$, %contains all 
+and let 
+$d_\bot(q_1, q_2) = d_\top(q_1, p, q_2) = \one$ if $q_1 = q_2$ and
+$d_\bot(q_1, q_2) = d_\top(q_1, p, q_2) = \zero$ otherwise.
 
 \smallskip\noindent
-\textbf{while} $P$ is not empty
-
-\noindent
-\quad\textbf{extract} $\< q_1, \sigma, q_2>$ from $P$ 
- such that $d(q_1, \sigma, q_2)$ is minimal wrt $\leq_\oplus$.
-
-\noindent\quad \textbf{for all} $q_0, q_3 \in Q$ and $p \in P$ do
-
-\noindent
-\qquad
-\(\begin{array}{rrcl}
- & d(q_1, \sigma, q_3) & \opluseq &
-   d(q_1, \sigma, q_2) \otimes \displaystyle\bigoplus_{a \in \Omegai} \weii(q_2, a, q_3)\\
-%
-%  & d(q_0, \sigma, q_2) & \opluseq & 
-%    \displaystyle\bigoplus_{a \in \Omegai} \weii(q_0, a, q_1) \otimes d(q_1, \sigma, q_2)\\
-%
-\mathrm{if~} \sigma = \top & d(q_0, \top, q_3) & \opluseq &
-d(q_1, \sigma, q_2) \otimes
-%\displaystyle\bigoplus_{p \in P} 
+\textbf{while} $\Q$ is not empty
+
+\noindent\quad
+\textbf{extract} $\< q_1, q_2>$ or $\< q_1, p, q_2>$ from $\Q$
+
+\noindent\quad
+such that $d_\bot(q_1, q_2)$, resp. $\bigoplus_{c\in \Omegac} d_\top(q_1, p, q_2)(c)$,
+is minimal in $\Semiring$ wrt $\leq_\oplus$.
+
+\noindent\quad 
+%\textbf{for all} $q_0, q_3 \in Q$ 
+\textbf{update} $d_\bot$ with $\< q_1, q_2>$ or $d_\top$ with $\< q_1, p, q_2>$
+(Figure~\ref{fig:best-update}).
+\end{algo}
+
+\begin{figure}
+For all $q_0, q_3 \in Q$, %$p \in P$,
+\[
+\begin{array}{lcl}
+%\multicolumn{3}{l}{\mathrm{For~all~} q_0, q_3 \in Q, p \in P,}\\
+d_\top(q_1, p, q_3) & \opluseq &
+  d_\top(q_1, p, q_2) \otimes 
+  \displaystyle\bigoplus_{a \in \Omegai} \weii(q_2, p, q_3)_a\\
+\multicolumn{3}{l}{\quad
+\mathrm{where~} \weii(q_2, p, q_3)_a \in \Phic
+\mathrm{~is~the~partial~application~}
+x_c \mapsto \weii(q_2, x_c, p, a, q_3)}\\
+%
+d_\bot(q_1, p, q_3) & \opluseq &
+  d_\bot(q_1, q_2) \otimes 
+  \displaystyle\bigoplus_{a \in \Omegai} \weie(q_2, a, q_3)\\
+%     
+d_\top(q_0, p, q_3) & \opluseq &
+  \displaystyle\bigoplus_{{\call{c}'} \in \Omegac} 
+  \displaystyle\bigoplus_{{\return{r}} \in \Omegar} 
+  \weic(q_0, p, p', q_1)_{c'} \otimes 
+  d_\top(q_1, p', q_2) \otimes
+  \weir(q_2, c', p', r, q_3)\\
+\multicolumn{3}{l}{\quad
+\mathrm{where~} \weic(q_0, p, p', q_1)_{c'} \in \Phic
+\mathrm{~is~the~partial~application~}
+x_c \mapsto \weic(q_0, p, x_c, c', p', q_1)}\\[2pt]
+%
+d_\bot(q_0, q_3) & \opluseq &
   \displaystyle\bigoplus_{{\call{c}} \in \Omegac} 
-\displaystyle\bigoplus_{{\return{r}} \in \Omegar} \eta({\call{c}}, {\return{r}})\\
-\mathrm{and} &  d(q_0, \bot, q_3) & \opluseq &
-d(q_1, \sigma, q_2) \otimes
-%\displaystyle\bigoplus_{p \in P} 
-\displaystyle\bigoplus_{{\call{c}} \in \Omegac} 
-\displaystyle\bigoplus_{{\return{r}} \in \Omegar} \eta({\call{c}}, {\return{r}})\\
-& & & \mathrm{where~} \eta = \weic(q_0, q_1, p) \otimes \weir(q_2, p, q_3)\\
-%
-\mathrm{if~} \sigma = \bot & d(q_1, \bot, q_3) & \opluseq &
-  d(q_1, \sigma, q_2) \otimes \displaystyle\bigoplus_{\return{r} \in \Omegar} \weie(q_2, {\return{r}}, q_3)\\
-%
-%& d(q_0, \bot, q_2) & \opluseq & 
-%d(q_0, \bot, q_1) \otimes d(q_1, \sigma, q_2), \mathrm{if~} \< q_0, \bot, q_1> \notin P \\
-%
-%\mathrm{if~} \sigma = \top & d(q_0, \top, q_2) & \opluseq & 
-%d(q_0, \top, q_1) \otimes d(q_1, \sigma, q_2), \mathrm{if~} \< q_0, \top, q_1> \notin P \\
-%
-& d(q_1, \bot, q_3) & \opluseq & 
-   d(q_1, \sigma, q_2) \otimes d(q_2, \bot, q_3), \mathrm{if~} \< q_2, \bot, q_3> \notin P \\
-%
-\mathrm{if~} \sigma = \top & d(q_1, \top, q_3) & \opluseq & 
-    d(q_1, \sigma, q_2) \otimes d(q_2, \top, q_3), \mathrm{if~} \< q_2, \top, q_3> \notin P \\
+  \displaystyle\bigoplus_{{\return{r}} \in \Omegar} 
+  \weiec(q_0, c, p, q_1) \otimes 
+   d_\top(q_1, p, q_2)(c) \otimes
+  \weir(q_2, c, p, r, q_3)\\
+%
+d_\bot(q_1, q_3) & \opluseq &
+  d_\bot(q_1, q_2) \otimes 
+  \displaystyle\bigoplus_{\return{r} \in \Omegar} \weier(q_2, {\return{r}}, q_3)\\
+%
+d_\top(q_1, p, q_3) & \opluseq & 
+  d_\top(q_1, p, q_2) \otimes d_\top(q_2, p, q_3), 
+  \mathrm{if~} \< q_2, \top, q_3> \notin P \\
+%
+d_\bot(q_1, q_3) & \opluseq & 
+  d_\bot(q_1, q_2) \otimes d_\bot(q_2, q_3), \mathrm{if~} \< q_2, \bot, q_3> \notin P \\
 \end{array}
-\) 
-\end{algo}
-
-\noindent
-The infinite sums in the updates of $d$ in Algorithm~\ref{algo:Dijkstra} are well defined
+\]
+%$\weii(q_2, p, q_3)_a \in \Phic$, for $a \in \Omegai$,
+%is the partial application
+%$x_c \mapsto \weii(q_2, x_c, p, a, q_3)$.
+%
+%$\weic(q_0, p, p', q_1)_{c'} \in \Phic$, for $c' \in \Omegac$,
+%is the partial application
+%$x_c \mapsto \weic(q_0, p, x_c, c', p', q_1)$.
+\caption{Update $d_\bot$ with $\<q_1, q_2>$ or $d_\top$ with $\< q_1, p, q_2>$.} 
+\label{fig:best-update}
+\end{figure}
+
+\noindent
+The infinite sums in the updates of $d$ in Algorithm~\ref{algo:Dijkstra}, 
+Figure~\ref{fig:best-update}
+are well defined
 since~$\Semiring$ is complete.
+** effectively computable by hypothese that the label theory is effective**
 The algorithm performs $2.|Q|^2$ iterations until $P$ is empty, 
@@
 This gives a time complexity $O(|Q|^4 . |P|)$. 
-%
-%It can be reduced by implementing $P$ as a priority queue, 
-%prioritized by the value returned by $d$
-%***complete***. %$|Q|^3.\log(|Q|^2)$
+It can be reduced by implementing $P$ as a priority queue, 
+prioritized by the value returned by $d$
+***complete***. %$|Q|^3.\log(|Q|^2)$
 
@@
 
-\begin{proposition}
-For a \SWVPA $A$ over a 
-commutative, idempotent, superior, total and complete semiring 
-and an alphabet $\Omega$ with a $k$-convex label theory,
+\begin{proposition}\label{th:best-search}
+For a \SWVPA $A$ 
+over $\Omega$, 
+$\Semiring$ commutative, bounded, total and complete, %semiring 
+and $\bar\Phi$ $k$-convex effective, % label theory,
 one can construct in PTIME a word $s \in \Omega^*$ 
 such that $A(s)$ is minimal \wrt the natural ordering for $\Semiring$. 
-%$\leg_\oplus$.
 \end{proposition}
 
-\subsection{Trees and Nested-Words}
+\subsection{Nested-Words and Parse-Trees}
 \label{sec:trees}
-The hierarchical structure of nested-words, defined the \emph{call} and \emph{return} markup symbols of 
-suggest a correspondence between these word and trees. 
-The lifting of this correspondence to languages, respective of tree automata and VPA have
-been discussed in~\cite{AlurMadhusudan09nested}, 
-see also~\cite{Caralp12VPAmult} for the weighted case.
-In this section, we discuss the correspondence between the symbolic-weighted extensions
+The hierarchical structure of nested-words, defined with the \emph{call} and \emph{return} markup symbols  
+suggest a correspondence with trees. 
+The lifting of this correspondence to languages, of tree automata and VPA,
+has been discussed in~\cite{AlurMadhusudan09nested}, 
+and~\cite{Caralp12VPAmult} for the weighted case.
+In this section, we describe a correspondence between the symbolic-weighted extensions
 of tree automata and VPA.
@@
 % and $n$ subtrees $t_1$,\ldots, $t_n$.
+%
+Let us assume a commutative semiring~$\Semiring$ 
+and a label theory~$\bar{\Phi}$ over~$\Semiring$ 
+containing one set~$\Phi_{\Omega_k}$ for each $k \in [0..M]$.
 
@@
 from trees of $\T(\Omega)$ into the weights of~$\Semiring$,
-based on the intermediate function $\weight_A$
+based on the intermediate function~$\weight_A$
 defined as follows for $q_0 \in Q$ and 
 $t = b(t_1,\ldots, t_k) \in \T(\Omega)$,
-åwith $0 \leq k \leq M$:
-\[
+with $0 \leq k \leq M$:
+\begin{equation}
 \begin{array}{rccl}
 \weight_A(q_0, t) & = & 
-   \displaystyle\bigoplus_{q_1 \in Q} &
-   \wei(q, \epsilon, q_1) \otimes \weight_A(q_1, t)\\
- & \oplus & \displaystyle\bigoplus_{q_1 \ldots q_k \in Q^k} &
+%   \displaystyle\bigoplus_{q_1 \in Q} &
+%   \wei(q, \epsilon, q_1) \otimes \weight_A(q_1, t)\\
+ \displaystyle\bigoplus_{q_1 \ldots q_k \in Q^k} &
               \wei(q_0, b, q_1 \ldots q_k ) 
    \otimes \displaystyle\bigotimes_{i=1}^{k}
-           \weight_A(q_i, t_i).\\
+           \weight_A(q_i, t_i)
+\end{array}
+\end{equation}
+
+\medskip\noindent
+The weight associated by $A$ to  $t \in \T(\Omega)$ is 
+\begin{equation}
+A(t)  = 
+\displaystyle\bigoplus_{q \in Q} \mathsf{in}(q) \mathop{\otimes} \weight_A(q, t)
+\label{eq:weightTA}
+\end{equation}
+
+\begin{lemma}\label{lem:SWTA}
+For all \SWTA $A$ over~$\Omega$, $\Semiring$ commutative, and $\Phi$,
+there exists an effectively constructible \SWVPA $A'$ over 
+$\hat\Omega$, $\Semiring$ and $\hat\Phi$ 
+such that for all $t \in \T(\Omega)$, $A'\bigl(\lin(t)\bigr) = A(t)$.
+\end{lemma} 
+% 
+\begin{proof}
+Let $A = \< Q, \init, \bar{\wei} >$ where $\bar{\wei}$ is presented as above by a function
+    %
+We build 
+$A' = \< Q', P', \init', \bar{\wei}', \final' >$,
+%computing over $\hat\Omega = \Omegai \uplus \Omegac \uplus \Omegar$,
+where $Q' = \bigcup_{k=0}^{M} Q^k$ is the set of sequences of state symbols of $A$, 
+of length at most $M$, including the empty sequence denoted by $\epsilon$, 
+and where $P' = Q'$ and $\bar\wei$ is defined by:
+
+\[
+\begin{array}{ll}
+\weii(\bar{q}, {\call{b}}, \bar{p}, a, \bar{q} q') = \wei(q', a, \epsilon) & 
+\mathrm{for~all~} b \in \Omega_{>0}, \bar{p}\in P', a \in \Omega_0\\
+%
+\weiei(\bar{q}, a, \bar{q} q') = \wei(q', a, \epsilon) & 
+\mathrm{for~all~} a \in \Omega_0\\
+%
+\weic(\bar{q}, {\call{b}}, \bar{p}, \call{b'}, \epsilon, \bar{q}) = \one & %\wei(p', b, \bar{p}) & 
+\mathrm{for~all~} b \in \Omega_{>0},  \bar{p}\in P', b, b' \in \Omega_{>0}\\
+%
+\weiec(\bar{q}, \call{b}, \epsilon, \bar{q}) = \one & %\wei(p', b, \bar{p}) & 
+\mathrm{for~all~} b \in \Omega_{>0}\\
+%
+%\weir: Q \times \Omegac \times P \times \Sigmar \times Q \to \Semiring & 
+\weir(\bar{q}, {\call{b}}, \bar{p}, {\return{b}}, \bar{p} q') = \wei(q', b, \bar{q}) & 
+\mathrm{for~all~}  b \in \Omega_{>0}, \bar{p}\in P'\\
+%
+%\weie: Q \times \Omegar \times Q \to \Semiring & 
+\weier(\bar{p}, {\return{b}}, \bar{q}) = \zero &
+\mathrm{for~all~}  b \in \Omega_{>0}\\
 \end{array}      
 \]
-
-\medskip\noindent
-The weight associated by $A$ to  $t \in \T(\Omega)$ is then
-\begin{equation}
-A(t)  = 
-\displaystyle\bigoplus_{q \in Q} \mathsf{in}(q) \mathop{\otimes} \weight_A(q, t).
-\label{eq:weightTA}
-\end{equation}
-
-\begin{lemma}  \label{lem:SWTA}
-For all \SWTA $A$ over~$\Omega$ and $\Semiring$, without $\epsilon$-transitions,
-there exists an effectively constructible \SWVPA $A'$ over 
-$\hat\Omega$ and $\Semiring$
-such that for all $t \in \T(\Omega)$, $A'\bigl(\lin(t)\bigr) = A(t)$.
-\end{lemma} 
-% 
-\begin{proof}
-Let $A = \< Q, \init, \bar{\wei} >$ where $\bar{\wei}$ is summarized as above by a function
-$\wei: Q \times (\Sigma \cup \{ \epsilon \}) \times \bigcup_{k=0}^{M} Q^k \to \Semiring$.
-
-We build 
-$A' = \< Q', P', \init', \weii, \weic, \weir, \weie, \final' >$,
-computing over $\hat\Omega = \< \Omegai, \Omegac, \Omegar >$,
-%
-%$\Omega = \Omegai \uplus \Omegac \uplus \Omegar$ and $\Semiring$ 
-where $Q' = \bigcup_{k=0}^{M} Q^k$ be  the set of sequences of state symbols of $A$, 
-of length at most $M$, including the empty sequence denoted by $\epsilon$, 
-and where $P' = Q'$.
-
-\[
-\begin{array}{ll}
-%\weii: Q \times \Omegai \times Q \to \Semiring & 
-\weii(\bar{q}, a, \bar{q} q') = \wei(q', a, \epsilon) & 
-\mathrm{for~all~} a \in \Omega_0,\\
-%
-%\weic: Q \times \Omegac \times Q \times P \to \Semiring & 
-\weic(\bar{q}, \call{b}, \epsilon, \bar{q}) = \one & %\wei(p', b, \bar{p}) & 
-\mathrm{for~all~} b \in \Omega_{>0},\\
-%
-%\weir: Q \times \Omegac \times P \times \Sigmar \times Q \to \Semiring & 
-\weir(\bar{q}, {\call{b}}, \bar{p}, {\return{b}}, \bar{p} q') = \wei(q', b, \bar{q}) & 
-\mathrm{for~all~}  b \in \Omega_{>0},\\
-%
-%\weie: Q \times \Omegar \times Q \to \Semiring & 
-\weie(\bar{p}, {\return{b}}, \bar{q}) = \zero &
-\mathrm{for~all~}  b \in \Omega_{>0}.\\
-\end{array}      
-\]
-
-In practice, it is sufficient to consider in $Q'$ only the prefixes of 
-sequences.
+\noindent
+It is sufficient to consider in $Q'$ only the prefixes of 
+sequences in transition with a non-null weight.
 \qed\end{proof}

### 2021-05-25 main.tex
revise SWT & SWA : no epsilon, not read tape without move for T

@@
-% fork of main.tex with more expressive models
-% SWT read both tapes (under 2 heads) at every transition (even epsilon or semi-epsilon)
-% SWA read tape (under head) at every transition (even epsilon)
+% SWT without epsilon-transitions
+% SWT do not read the other tape in "skip" transitions 10 or 01
+%     (hence this version is simpler  than V2)
+% SWA without epsilon-transitions
 % SW-VPA read top of stack (when not empty) at every transition 
+%        (6 kinds of transitions).
 
@@
 \usepackage{amsmath} 
+\allowdisplaybreaks
 \usepackage{amssymb} 
@@
 
-%% for new macros
-\usepackage{xspace}
+%% allows for page break within arrays
+\usepackage{longtable}
 
@@
 
+%% for new macros
+\usepackage{xspace}
+
 %% Misc macros
-
 \def\ie{\textit{i.e.}\xspace}
@@
 where $Q$ is a finite set of states, 
-$\mathsf{in} : Q \to \Phi_{\Sigma, \Delta}$, 
-respectively $\mathsf{out} : Q \to \Phi_{\Sigma, \Delta}$,
+$\mathsf{in} : Q \to \Semiring$   %\Phi_{\Sigma, \Delta}
+respectively $\mathsf{out} : Q \to \Semiring$  %\Phi_{\Sigma, \Delta}
 are functions defining the weight for entering, 
@@
 and $\bar{\wei}$ is a triplet of transition functions 
-$\wei_{10}$, $\wei_{01}$,  and $\wei_{11}$
-%$\wei_{i, j}: Q \times Q \to \Phi_{\Sigma, \Delta}$, with $i, j \in \{ 0, 1\}$.
-from $Q \times Q$ into $\Phi_{\Sigma, \Delta}$.
+$\wei_{10}: Q \times Q \to \Phi_{\Sigma}$,
+$\wei_{01}: Q \times Q \to \Phi_{\Delta}$, and 
+$\wei_{11}: Q \times Q \to \Phi_{\Sigma, \Delta}$.
 \end{definition}
-
-%A pair of states $\< q, q'>$ such that $\wei_{00}(q, q') \neq \zero$
-%is called an \emph{$\epsilon$-transition}.
-%
-For convenience, when $\wei_{ij}(q, q') = \eta \in \Phi_{\Sigma, \Delta}$, 
-for $q, q' \in Q$ and $i, j \in \{ 0, 1\} \setminus \{ \<0, 0> \}$,
-respectively $\mathsf{in}(q) = \eta$, $\mathsf{out}(q') = \eta$,
-we shall sometimes write, overloading the function names:
-$\wei_{ij}(q, a, b, q') = \eta(a, b)$ for $a \in \Sigma$,  $b \in \Delta$,
-respectively $\mathsf{in}(q, a, b) = \eta(a, b)$,
-$\mathsf{out}(q', a, b) = \eta(a, b)$.
-
-\noindent 
-The \SWT $T$ computes on pairs of words $\< s, t> \in \Sigma^+ \times \Delta^+$,
+%
+\noindent
+For convenience, we shall the following presentation of transition functions 
+as functions of  
+$Q \times (\Sigma \cup \{ \epsilon \}) \times (\Delta \cup \{ \epsilon \}) \times Q \to \Semiring$,
+overloading the function names,
+such that, for all $q, q' \in Q$, $a \in \Sigma$,  $b \in \Delta$, 
+\[
+\begin{array}{rcll}
+\wei_{10}(q, a, \epsilon, q') & = & \phi(a) & %\wei_\Sigma(q, q')(a)
+\quad\mathrm{where~} \phi = \wei_{10}(q, q') \in \Phi_\Sigma,\\
+\wei_{01}(q, \epsilon, b, q') & = & \psi(b) & 
+\quad\mathrm{where~} \psi = \wei_{01}(q, q') \in \Phi_\Delta,\\
+\wei_{11}(q, a, b, q') & = & \eta(a, b) & 
+\quad\mathrm{where~} \eta = \wei_{11}(q, q') \in \Phi_{\Sigma, \Delta}.\\
+\end{array}      
+\]
+%
+\noindent 
+The \SWT $T$ computes on pairs of words $\< s, t> \in \Sigma^* \times \Delta^*$,
 $s$ and $t$, being respectively called input and output word.
@@
 More precisely, the symbolic-weighted transducer~$T$ defines a mapping 
-from the pairs of strings of $\Sigma^+ \times \Delta^+$ 
+from the pairs of strings of $\Sigma^* \times \Delta^*$ 
 into~$\Semiring$,
@@
 $a \in \Sigma$, $u \in \Sigma^*$, $b\in \Delta$, $v\in \Delta^*$, 
-
-\begin{equation}
-\begin{array}[t]{lccl}
-%\weight_T(q, \epsilon, \epsilon, q)  & = & \multicolumn{2}{l}{\one}\\ % or \final(q)
-%\weight_T(q, \epsilon, \epsilon, q') & = & \multicolumn{2}{l}{\zero \mathrm{~if~} q \neq q'}\\
-\weight_T(q, au, bv, q') & = & 
-             \displaystyle\bigoplus_{q'' \in Q, u \neq \epsilon} &
-    \wei_{10}(q, a, b, q'') \otimes \weight_T(q'', u, bv, q')\\
-  & \oplus & \displaystyle\bigoplus_{q'' \in Q, v \neq \epsilon} &
-    \wei_{01}(q, a, b, q'') \otimes \weight_T(q'', au, v, q')\\
-  & \oplus & \displaystyle\bigoplus_{q'' \in Q, u,v \neq \epsilon} &
-    \wei_{11}(q, a, b, q'') \otimes \weight_T(q'', u, v, q')\\
-  & \oplus & \displaystyle\bigoplus_{u,v = \epsilon} &
-    \wei_{11}(q, a, b, q')
-\end{array}
+%
+\begin{align}
+\weight_T(q, \epsilon, \epsilon, q)  & = \final(q)\\
+\weight_T(q, \epsilon, \epsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
+\weight_T(q, au, \epsilon, q') & = \displaystyle\bigoplus_{q'' \in Q} 
+    \wei_{10}(q, a, \epsilon, q'') \otimes \weight_T(q'', u, \epsilon, q')\nonumber\\
+\weight_T(q, \epsilon, bv, q') & = \displaystyle\bigoplus_{q'' \in Q} 
+    \wei_{01}(q, \epsilon, b, q'') \otimes \weight_T(q'', \epsilon, v, q')\nonumber\\
+\weight_T(q, au, bv, q') & = \displaystyle\bigoplus_{q'' \in Q} 
+    \wei_{10}(q, a, \epsilon, q'') \otimes \weight_T(q'', u, bv, q')\nonumber\\
+                    & \oplus \displaystyle\bigoplus_{q'' \in Q} 
+    \wei_{01}(q, \epsilon, b, q'') \otimes \weight_T(q'', au, v, q')\nonumber\\
+                    & \oplus \displaystyle\bigoplus_{q'' \in Q} 
+    \wei_{11}(q, a, b, q'') \otimes \weight_T(q'', u, v, q')\nonumber
 \label{eq:SWT-weight}
-\end{equation}
+\end{align}
 %
@@
 when reading respectively $a$ and $b$ at the current positions in the input and output words, 
-increment the current position in the input word if if $i = 1$, 
-and in the output word if $j = 1$ (otherwise, do not change it),
+increment the current position in the input word if and only if $i = 1$, 
+and in the output word iff $j = 1$ (otherwise, do not change it),
 and change state from $q$ to $q'$.
-%
-In contrast with the models of weighted transducers over finite alphabets~\cite{Mohri03ijfcs},
-the input and output symbols at current positions are always read by transitions, 
-even when they do not change the reading position the head's position.
-This is an important feature in the case of an infinite alphabet in 
-order to compare input and output symbols.
+When $a = \epsilon$ (resp. $b = \epsilon$), the current symbol 
+in the input (resp. output) is not read.
+%
+%In contrast with the models of weighted transducers over finite alphabets~\cite{Mohri03ijfcs},
+%the input and output symbols at current positions are always read by transitions, 
+%even when they do not change the reading position the head's position.
+%This is an important feature in the case of an infinite alphabet in 
+%order to compare input and output symbols.
 %which cannot be stored in the finite memory of the transducer.
-
+%
 Since $\zero$ is absorbing for~$\otimes$ in~$\Semiring$,
@@
 Intuitively, 
-$\wei_{10}(q, a, b, r)$ is the cost of 
+$\wei_{10}(q, a, \epsilon, r)$ is the cost of 
 the deletion of the symbol~$a \in \Sigma$ in~$s$, 
-$\wei{01}(q, a, b, r)$ is the cost 
+$\wei{01}(q, \epsilon, b, r)$ is the cost 
 of the insertion of~$b \in \Delta$ in $t$, 
@@
 \medskip\noindent
-Let $\< s, t> \in \Sigma^+ \times \Delta^+$, with $s = s_1\ldots s_n$, 
-and $t = t_1\ldots t_m$. 
-The weight associated by $T$ to $\< s, t>$
+%Let $\< s, t> \in \Sigma^* \times \Delta^*$, with $s = s_1\ldots s_n$, and $t = t_1\ldots t_m$. 
+The weight associated by $T$ to $\< s, t> \in \Sigma^* \times \Delta^*$
 is defined as follows:
@@
 T(s, t)  = 
-\displaystyle\bigoplus_{q, q' \in Q} \mathsf{in}(q, s_1, t_1) 
-\mathop{\otimes} \weight_T(q, s, t, q') \mathop{\otimes} \mathsf{out}(q', s_n, t_m)
+\displaystyle\bigoplus_{q, q' \in Q} \mathsf{in}(q) 
+\mathop{\otimes} \weight_T(q, s, t, q') %\mathop{\otimes} \mathsf{out}(q', s_n, t_m)
 \label{eq:SWT-value}
@@
 
-\begin{example}
-We can define with a \SWT the computation of a similarity measure 
-between timed sequences
-similar to dynamic time warping (DTW).
-
-Let $\Semiring$  be the tropical (\emph{min-plus}) semiring of Figure~\ref{fig:semirings} and 
-let $\Sigma = \Delta = \mathbb{R}_+$ be sets of timestamps.
-We consider a \SWT with one state $q$ and transitions
-$\wei_{11}(q, d, d', q) = 
- \wei_{01}(q, d, d', q) = 
- \wei_{10}(q, d, d', q) = |d' - d|$,
-for all $d, d' \in \mathbb{R}_+$.
-% needs reading input/output symbols even by epsilon-transitions.
-The recursive definition of $\weight_T$ correspond to the dynamic programming equations of DTW
-for the computation of an optimal match between words, 
-the matching cost for two symbols being the 
-the time distance between them.
-\endex
-\end{example}
+%\begin{example}
+%We can define with a \SWT the computation of a similarity measure 
+%between timed sequences
+%similar to dynamic time warping (DTW).
+%
+%Let $\Semiring$  be the tropical (\emph{min-plus}) semiring of Figure~\ref{fig:semirings} and 
+%let $\Sigma = \Delta = \mathbb{R}_+$ be sets of timestamps.
+%We consider a \SWT with one state $q$ and transitions
+%$\wei_{11}(q, d, d', q) = 
+% \wei_{01}(q, d, d', q) = 
+% \wei_{10}(q, d, d', q) = |d' - d|$,
+%for all $d, d' \in \mathbb{R}_+$.
+%% needs reading input/output symbols even by epsilon-transitions.
+%The recursive definition of $\weight_T$ correspond to the dynamic programming equations of DTW
+%for the computation of an optimal match between words, 
+%the matching cost for two symbols being the 
+%the time distance between them.
+%\endex
+%\end{example}
 
@@
 \wei_{11}(q_1, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d|\\
-\wei_{01}(q_0, d, \< \mathsf{-}, d'>, q_0) & = & \one & &
-\wei_{01}(q_1, d, \< \mathsf{-}, d'>, q_0) & = & \one\\
-\wei_{10}(q_0, d, b, q_1) & = & \alpha & & \multicolumn{3}{l}{\mathrm{for~all~} b \in \Delta}
+\wei_{01}(q_0, \epsilon, \< \mathsf{-}, d'>, q_0) & = & \one & &
+\wei_{01}(q_1, \epsilon, \< \mathsf{-}, d'>, q_0) & = & \one\\
+\wei_{10}(q_0, d, \epsilon, q_1) & = & \alpha & & %\multicolumn{3}{l}{\mathrm{for~all~} b \in \Delta}
 \end{array}
@@
 At last, we let $q_0$ be the only initial and final state, with
-$\init(q_0, d, b) = \final(q_0, d, b) = \one$, and 
-$\init(q_1, d, b) = \final(q_1, d, b) = \zero$, 
-for all $d \in \Sigma$ and $b \in \Delta$.
+$\init(q_0) = \final(q_0) = \one$, and 
+$\init(q_1) = \final(q_1) = \zero$. 
+%$\init(q_0, d, b) = \final(q_0, d, b) = \one$, and 
+%$\init(q_1, d, b) = \final(q_1, d, b) = \zero$, 
+%for all $d \in \Sigma$ and $b \in \Delta$.
 \endex
@@
 where $Q$ is a finite set of states, 
-$\mathsf{in} : Q \to \Phi_\Sigma$, 
-respectively $\mathsf{out} : Q \to \Phi_\Sigma,$
+$\mathsf{in} : Q \to \Semiring$, %\Phi_\Sigma$, 
+respectively $\mathsf{out} : Q \to \Semiring$  %\Phi_\Sigma,$
 are functions defining the weight for entering, 
@@
 when $\wei_1(q, q') = \phi \in \Phi_\Sigma$, 
-respectively $\mathsf{in}(q) = \phi$, $\mathsf{out}(q') = \phi$,
-we may write $\wei_1(q, a, q')$ for $\phi(a)$, 
-respectively $\mathsf{in}(q, a) = \phi(a)$, $\mathsf{out}(q', a) = \phi(a)$.
+%respectively $\mathsf{in}(q) = \phi$, $\mathsf{out}(q') = \phi$,
+we may write $\wei_1(q, a, q')$ for~$\phi(a)$. 
+%respectively $\mathsf{in}(q, a) = \phi(a)$, $\mathsf{out}(q', a) = \phi(a)$.
 %$\wei_1: Q \times \Sigma \times Q \to \Semiring$, 
-The computation of $A$ on words $s \in \Sigma^+$
+The computation of $A$ on words $s \in \Sigma^*$
 is defined with an intermediate function $\weight_A$, 
@@
 %
-\begin{equation}
-\begin{array}{lccl}
-%\weight_T(q, \epsilon, \epsilon, q)  & = & \multicolumn{2}{l}{\one}\\ % or \final(q)
-%\weight_T(q, \epsilon, \epsilon, q') & = & \multicolumn{2}{l}{\zero \mathrm{~if~} q \neq q'}\\
-\weight_A(q, au, q') & = & 
-            \displaystyle\bigoplus_{r \in Q, u \neq \epsilon} &
-    \wei_{1}(q, a, r) \otimes \weight_A(r, u, q')\\
-  & \oplus & \displaystyle\bigoplus_{u = \epsilon} & 
-    \wei_{1}(q, a, q')
-\end{array}
+\begin{align}
+\weight_A(q, \epsilon, q) & = \final(q)\\
+\weight_A(q, \epsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
+\weight_A(q, au, q') & =  \displaystyle\bigoplus_{q'' \in Q} 
+    \wei_{1}(q, a, q'') \otimes \weight_A(q'', u, q')\nonumber
 \label{eq:SWA-weight}
-\end{equation}
+\end{align}
 %
@@
 and the weight value associated by $A$ to 
-$s = s_1\ldots s_n \in \Sigma^+$ is: % defined as follows:
+$s \in \Sigma^*$ is defined as follows: %$s = s_1\ldots s_n \in \Sigma^+$ 
 \begin{equation}
 A(s)  = 
-\displaystyle\bigoplus_{q, q' \in Q} \mathsf{in}(q, s_1) 
-\mathop{\otimes} \weight_A(q, s, q') \mathop{\otimes} \mathsf{out}(q', s_n)
+\displaystyle\bigoplus_{q, q' \in Q} \mathsf{in}(q) 
+\mathop{\otimes} \weight_A(q, s, q') 
 \label{eq:SWA-value}
@@
 Let $T = \< Q, \init, \bar{\wei}, \final >$,
-let 
-$\wei: Q \times \Sigma \times \Delta \times Q \to \Semiring$
-be the synthesized form of 
-$\bar{\wei} = \< \wei_{10}, \wei_{01}, \wei_{11}>$,
-and let $s = s_1 \ldots s_n$ with $s_1, \ldots, s_n \in \Sigma$ ($n \geq 1$).
-%$n = |s|$ be the length of $s$.
+with $\bar{\wei} = \< \wei_{10}, \wei_{01}, \wei_{11}>$,
+%where the component functions of
+%be the synthesized form of 
+%$Q \times (\Sigma \cup \{ \epsilon \}) \times (\Delta \cup \{ \epsilon \}) \times Q \to \Semiring$,
+and let $s = s_1 \ldots s_n$ with $s_1, \ldots, s_n \in \Sigma$ ($n \geq 1$). %$n = |s|$ be the length of $s$.
 
@@
 Its state entering weight function is defined by, 
-for all $q \in Q$ and $b \in \Delta$,  
-$\init'(\< 1, q>, b) = \init(q, s_1, b)$ 
-and $\init'(\< i, q>, b) = \zero$ for all $1 < i \leq n+1$. 
-Its state leaving weight function is defined by, for all $q \in Q$ and $b \in \Delta$,  
-$\final'(\< n+1, q>, b) = \final(q, s_n, b)$, 
-and $\final'(\< i, q>, b) = \zero$ for all $1 \leq i <  n+1$.
+for all $q \in Q$ $\init'(1, q) = \init(q)$ 
+and $\init'(i, q) = \zero$ for all $1 < i \leq n+1$. 
+Its state leaving weight function is defined by, for all $q \in Q$ 
+$\final'(n+1, q) = \final(q)$, 
+and $\final'(i, q) = \zero$ for all $1 \leq i <  n+1$.
 
 Every non-null transition of $A_{T, s}$ will
-simulate a sequence of transitions of $T$ which advance in 
-the input word while staying immobile in the output word, 
+simulate a sequence of transitions of $T$ performing the following steps:
+advance in the input word while staying immobile in the output word, 
 and then make one step in the output word (and advance in the input word or not).
 %
-The above initial sequence correspond to $\epsilon$-transitions of automata, 
-and its total weight is computed by the following intermediate function 
-$\wei'_0 : Q' \times Q' \to \Phi_{\Sigma}$, 
+The first steps in the sequence correspond to $\epsilon$-transitions of automata, 
+and their total weight is computed by the following intermediate function 
+$\wei'_0 : Q' \times Q' \to \Semiring$, 
 defined for all $q, q' \in Q$: % all $1 \leq i < n$ and $1 \leq k \leq n-i$, 
@@
            \end{array}} 
-\bigotimes_{j=0}^{k-1} \bigl[\wei_{10}(q_j, q_{j+1})\bigr]_{s_{i+j}}
+\bigotimes_{j=0}^{k-1} \phi_{j, j+1}(s_{i+j})
 \quad \begin{array}[t]{l}
-      \mathrm{if~} 1 \leq i < n\\
-      \mathrm{and~} 1 \leq k \leq n-i.
+      \mathrm{if~} 1 \leq i < n,
+      \mathrm{~and~} 1 \leq k \leq n-i,\\
+      \mathrm{where~} \phi_{j, j+1} = \wei_{10}(q_j, q_{j+1})
       \end{array}
@@
 \]
-where $\bigl[\wei_{10}(q_j, q_{j+1})\bigr]_{s_{i+j}}$ 
-is the partial application $\eta_{s_{i+j}} \in \Phi_{\Delta}$
-for $\eta = \wei_{10}(q_j, q_{j+1}) \in \Phi_{\Sigma, \Delta}$ ($s_{i+j} \in \Sigma$).
+%where $\bigl[\wei_{10}(q_j, q_{j+1})\bigr]_{s_{i+j}}$ 
+%is the partial application $\eta_{s_{i+j}} \in \Phi_{\Delta}$
+%for $\eta = \wei_{10}(q_j, q_{j+1}) \in \Phi_{\Sigma, \Delta}$ ($s_{i+j} \in \Sigma$).
 %
@@
            \end{array}} 
-\wei'_0\bigl(\<i, q>, \< i+k, q''>\bigr) \otimes \eta_{s_{i+k}}
-\quad \mathrm{where~} \eta = \wei_{01}(q'', q'),\\
+\wei'_0\bigl(\<i, q>, \< i+k, q''>\bigr) \otimes \psi %\eta_{s_{i+k}}
+\quad \mathrm{where~} \psi = \wei_{01}(q'', q'),\\
  & \oplus & \displaystyle
-\bigoplus_{q'' \in Q} \wei'_0\bigl(\<i, q>, \< i+k-1, q''>\bigr) \otimes \eta'_{s_{i+k}}
-\quad \mathrm{where~}\eta' = \wei_{11}(q'', q'),\\
+\bigoplus_{q'' \in Q} \wei'_0\bigl(\<i, q>, \< i+k-1, q''>\bigr) \otimes \eta_{s_{i+k}}
+\quad \mathrm{where~}\eta = \wei_{11}(q'', q'),\\
 %
@@
 %
+In the second equation, 
+$\eta_{s_{i+j}} \in \Phi_{\Delta}$ is a partial application 
+for $\eta \in \Phi_{\Sigma, \Delta}$, $s_{i+j} \in \Sigma$, 
+defined by $\eta_{s_{i+j}}(y) = \eta(s_{i+j}, y)$ for all $y \in \Sigma$.
 

### 2021-05-27 main.tex
def complete label theory & prop

@@
 %(SWT)
-
+\noindent 
 Let $\Semiring$ be a commutative semiring, 
@@
 as functions of  
-$Q \times (\Sigma \cup \{ \epsilon \}) \times (\Delta \cup \{ \epsilon \}) \times Q \to \Semiring$,
+$Q \times (\Sigma \cup \{ \varepsilon \}) \times (\Delta \cup \{ \varepsilon \}) \times Q \to \Semiring$,
 overloading the function names,
@@
 \begin{array}{rcll}
-\wei_{10}(q, a, \epsilon, q') & = & \phi(a) & %\wei_\Sigma(q, q')(a)
+\wei_{10}(q, a, \varepsilon, q') & = & \phi(a) & %\wei_\Sigma(q, q')(a)
 \quad\mathrm{where~} \phi = \wei_{10}(q, q') \in \Phi_\Sigma,\\
-\wei_{01}(q, \epsilon, b, q') & = & \psi(b) & 
+\wei_{01}(q, \varepsilon, b, q') & = & \psi(b) & 
 \quad\mathrm{where~} \psi = \wei_{01}(q, q') \in \Phi_\Delta,\\
@@
 \begin{align}
-\weight_T(q, \epsilon, \epsilon, q)  & = \final(q)\\
-\weight_T(q, \epsilon, \epsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
-\weight_T(q, au, \epsilon, q') & = \displaystyle\bigoplus_{q'' \in Q} 
-    \wei_{10}(q, a, \epsilon, q'') \otimes \weight_T(q'', u, \epsilon, q')\nonumber\\
-\weight_T(q, \epsilon, bv, q') & = \displaystyle\bigoplus_{q'' \in Q} 
-    \wei_{01}(q, \epsilon, b, q'') \otimes \weight_T(q'', \epsilon, v, q')\nonumber\\
+\weight_T(q, \varepsilon, \varepsilon, q)  & = \final(q)\\
+\weight_T(q, \varepsilon, \varepsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
+\weight_T(q, au, \varepsilon, q') & = \displaystyle\bigoplus_{q'' \in Q} 
+    \wei_{10}(q, a, \varepsilon, q'') \otimes \weight_T(q'', u, \varepsilon, q')\nonumber\\
+\weight_T(q, \varepsilon, bv, q') & = \displaystyle\bigoplus_{q'' \in Q} 
+    \wei_{01}(q, \varepsilon, b, q'') \otimes \weight_T(q'', \varepsilon, v, q')\nonumber\\
 \weight_T(q, au, bv, q') & = \displaystyle\bigoplus_{q'' \in Q} 
-    \wei_{10}(q, a, \epsilon, q'') \otimes \weight_T(q'', u, bv, q')\nonumber\\
+    \wei_{10}(q, a, \varepsilon, q'') \otimes \weight_T(q'', u, bv, q')\nonumber\\
                     & \oplus \displaystyle\bigoplus_{q'' \in Q} 
-    \wei_{01}(q, \epsilon, b, q'') \otimes \weight_T(q'', au, v, q')\nonumber\\
+    \wei_{01}(q, \varepsilon, b, q'') \otimes \weight_T(q'', au, v, q')\nonumber\\
                     & \oplus \displaystyle\bigoplus_{q'' \in Q} 
@@
 and change state from $q$ to $q'$.
-When $a = \epsilon$ (resp. $b = \epsilon$), the current symbol 
+When $a = \varepsilon$ (resp. $b = \varepsilon$), the current symbol 
 in the input (resp. output) is not read.
@@
 %whereas considering the current symbol may be useful to compute a transition weight.
-%(even when it does not change the head's position, like with $\epsilon$-transitions).
-%
-%The cases $\weight_T(q, au, \epsilon, q')$ and $\weight_T(q, \epsilon, bv, q')$
+%(even when it does not change the head's position, like with $\varepsilon$-transitions).
+%
+%The cases $\weight_T(q, au, \varepsilon, q')$ and $\weight_T(q, \varepsilon, bv, q')$
 %are missing in the definition of $\weight_T$.
@@
 Intuitively, 
-$\wei_{10}(q, a, \epsilon, r)$ is the cost of 
+$\wei_{10}(q, a, \varepsilon, r)$ is the cost of 
 the deletion of the symbol~$a \in \Sigma$ in~$s$, 
-$\wei{01}(q, \epsilon, b, r)$ is the cost 
+$\wei{01}(q, \varepsilon, b, r)$ is the cost 
 of the insertion of~$b \in \Delta$ in $t$, 
@@
 \wei_{11}(q_1, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d|\\
-\wei_{01}(q_0, \epsilon, \< \mathsf{-}, d'>, q_0) & = & \one & &
-\wei_{01}(q_1, \epsilon, \< \mathsf{-}, d'>, q_0) & = & \one\\
-\wei_{10}(q_0, d, \epsilon, q_1) & = & \alpha & & %\multicolumn{3}{l}{\mathrm{for~all~} b \in \Delta}
+\wei_{01}(q_0, \varepsilon, \< \mathsf{-}, d'>, q_0) & = & \one & &
+\wei_{01}(q_1, \varepsilon, \< \mathsf{-}, d'>, q_0) & = & \one\\
+\wei_{10}(q_0, d, \varepsilon, q_1) & = & \alpha & & %\multicolumn{3}{l}{\mathrm{for~all~} b \in \Delta}
 \end{array}
@@
 \begin{align}
-\weight_A(q, \epsilon, q) & = \final(q)\\
-\weight_A(q, \epsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
+\weight_A(q, \varepsilon, q) & = \final(q)\\
+\weight_A(q, \varepsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
 \weight_A(q, au, q') & =  \displaystyle\bigoplus_{q'' \in Q} 
@@
 %
-%When $\wei_\epsilon(q, q') = \zero$ for all $q, q' \in Q$, 
-%the automaton~$A$ is called \emph{without $\epsilon$-transitions}.
+%When $\wei_\varepsilon(q, q') = \zero$ for all $q, q' \in Q$, 
+%the automaton~$A$ is called \emph{without $\varepsilon$-transitions}.
 
@@
 %be the synthesized form of 
-%$Q \times (\Sigma \cup \{ \epsilon \}) \times (\Delta \cup \{ \epsilon \}) \times Q \to \Semiring$,
+%$Q \times (\Sigma \cup \{ \varepsilon \}) \times (\Delta \cup \{ \varepsilon \}) \times Q \to \Semiring$,
 and let $s = s_1 \ldots s_n$ with $s_1, \ldots, s_n \in \Sigma$ ($n \geq 1$). %$n = |s|$ be the length of $s$.
@@
 %
-The first steps in the sequence correspond to $\epsilon$-transitions of automata, 
+The first steps in the sequence correspond to $\varepsilon$-transitions of automata, 
 and their total weight is computed by the following intermediate function 
@@
 %
-\begin{equation}
-\begin{array}[t]{rcl}
-\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot\gamma}, a\, u, \config{q'}{\gamma'}\bigr) & = & 
+\begin{align}
+%\begin{array}[t]{rcl}
+\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot\gamma}, a\, u, \config{q'}{\gamma'}\bigr) & =  
  {\displaystyle\bigoplus_{q'' \in Q}} \weii(q, c, p, a, q'') 
- \otimes \weight_A\bigl(\config{q''}{\gamma}, u, \config{q'}{\gamma'}\bigr)\\
-\weight_A\bigl(\config{q}{\bot}, a\, u, \config{q'}{\gamma'}\bigr) & = & 
+ \otimes \weight_A\bigl(\config{q''}{\gamma}, u, \config{q'}{\gamma'}\bigr)\nonumber\\
+\weight_A\bigl(\config{q}{\bot}, a\, u, \config{q'}{\gamma'}\bigr) & =  
  {\displaystyle\bigoplus_{q'' \in Q}} \weiei(q, a, q'') 
- \otimes \weight_A\bigl(\config{q''}{\bot}, u, \config{q'}{\gamma'}\bigr)\\
-%
-\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot\gamma}, {\call{c}'}\, u, \config{q'}{\gamma'}\bigr) & = & 
+ \otimes \weight_A\bigl(\config{q''}{\bot}, u, \config{q'}{\gamma'}\bigr)\nonumber\\
+%
+\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot\gamma}, {\call{c}'}\, u, \config{q'}{\gamma'}\bigr) & =  
  {\displaystyle\bigoplus_{\begin{array}{c}
@@
  \weic\bigl(q, {\call{c}}, p, {\call{c}'}, p', q''\bigr) 
- \otimes \weight_A\bigl(\config{q''}{{\call{c}'}\, p' \cdot {\call{c}}\, p\cdot \gamma}, u, \config{q'}{\gamma'}\bigr)\\[1mm]
-%
-\weight_A\bigl(\config{q}{\bot}, {\call{c}}\, u, \config{q'}{\gamma'}\bigr) & = & 
+ \otimes \weight_A\bigl(\config{q''}{{\call{c}'}\, p' \cdot {\call{c}}\, p\cdot \gamma}, u, \config{q'}{\gamma'}\bigr)\nonumber\\[1mm]
+%
+\weight_A\bigl(\config{q}{\bot}, {\call{c}}\, u, \config{q'}{\gamma'}\bigr) & =  
  {\displaystyle\bigoplus_{q'' \in Q}} \weiec(q, {\call{c}}, p, q'') 
- \otimes \weight_A\bigl(\config{q''}{{\call{c}}\, p}, u, \config{q'}{\gamma'}\bigr)\\
-%
-\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot \gamma}, {\return{r}}\, u, \config{q'}{\gamma'}\bigr) & = & 
+ \otimes \weight_A\bigl(\config{q''}{{\call{c}}\, p}, u, \config{q'}{\gamma'}\bigr)\nonumber\\
+%
+\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot \gamma}, {\return{r}}\, u, \config{q'}{\gamma'}\bigr) & =  
  {\displaystyle\bigoplus_{q'' \in Q}} 
  \weir\bigl(q, {\call{c}}, p, {\return{r}}, q''\bigr) 
- \otimes \weight_A\bigl(\config{q''}{\gamma}, u, \config{q'}{\gamma'}\bigr)\\
-%
-\weight_A\bigl(\config{q}{\bot}, {\return{r}}\, u, \config{q'}{\gamma'}\bigr) & = & 
+ \otimes \weight_A\bigl(\config{q''}{\gamma}, u, \config{q'}{\gamma'}\bigr)\nonumber\\
+%
+\weight_A\bigl(\config{q}{\bot}, {\return{r}}\, u, \config{q'}{\gamma'}\bigr) & =  
  {\displaystyle\bigoplus_{q'' \in Q}} \weier(q, {\return{r}}, q'') 
- \otimes \weight_A\bigl(\config{q''}{\bot}, u, \config{q'}{\gamma'}\bigr)\\
-\end{array}
+ \otimes \weight_A\bigl(\config{q''}{\bot}, u, \config{q'}{\gamma'}\bigr)
+%\end{array}
 \label{eq:SWVPA-weight}
-\end{equation}
+\end{align}
 %
@@
 and $\bar{\wei}$ is a tuplet of transition functions containing
-%$\wei_\epsilon$ from $Q \times Q$ into $\Semiring$, and, 
+%$\wei_\varepsilon$ from $Q \times Q$ into $\Semiring$, and, 
 $\wei_{k}: Q \times Q^{k} \to \Phi_{\Omega_k}$
@@
 \begin{array}{rcll}
-%\wei(q_0, \epsilon, q_1) & = &  \wei_\epsilon(q_0, q_1),\\ %\phi_\epsilon\\
+%\wei(q_0, \varepsilon, q_1) & = &  \wei_\varepsilon(q_0, q_1),\\ %\phi_\varepsilon\\
 \wei(q_0, a, q_1 \ldots q_k) & = & \phi_{\Omega, k}(a) &
@@
 %   \displaystyle\bigoplus_{q_1 \in Q} &
-%   \wei(q, \epsilon, q_1) \otimes \weight_A(q_1, t)\\
+%   \wei(q, \varepsilon, q_1) \otimes \weight_A(q_1, t)\\
  \displaystyle\bigoplus_{q_1 \ldots q_k \in Q^k} &
@@
 where $Q' = \bigcup_{k=0}^{M} Q^k$ is the set of sequences of state symbols of $A$, 
-of length at most $M$, including the empty sequence denoted by $\epsilon$, 
+of length at most $M$, including the empty sequence denoted by $\varepsilon$, 
 and where $P' = Q'$ and $\bar\wei$ is defined by:
@@
 \begin{array}{ll}
-\weii(\bar{q}, {\call{b}}, \bar{p}, a, \bar{q} q') = \wei(q', a, \epsilon) & 
+\weii(\bar{q}, {\call{b}}, \bar{p}, a, \bar{q} q') = \wei(q', a, \varepsilon) & 
 \mathrm{for~all~} b \in \Omega_{>0}, \bar{p}\in P', a \in \Omega_0\\
 %
-\weiei(\bar{q}, a, \bar{q} q') = \wei(q', a, \epsilon) & 
+\weiei(\bar{q}, a, \bar{q} q') = \wei(q', a, \varepsilon) & 
 \mathrm{for~all~} a \in \Omega_0\\
 %
-\weic(\bar{q}, {\call{b}}, \bar{p}, \call{b'}, \epsilon, \bar{q}) = \one & %\wei(p', b, \bar{p}) & 
+\weic(\bar{q}, {\call{b}}, \bar{p}, \call{b'}, \varepsilon, \bar{q}) = \one & %\wei(p', b, \bar{p}) & 
 \mathrm{for~all~} b \in \Omega_{>0},  \bar{p}\in P', b, b' \in \Omega_{>0}\\
 %
-\weiec(\bar{q}, \call{b}, \epsilon, \bar{q}) = \one & %\wei(p', b, \bar{p}) & 
+\weiec(\bar{q}, \call{b}, \varepsilon, \bar{q}) = \one & %\wei(p', b, \bar{p}) & 
 \mathrm{for~all~} b \in \Omega_{>0}\\
@@
 %\begin{description}
-%\item[] $\wei(q, \epsilon, a, q) = \one$ for all $a \in \Omegac \cup \Omegar$,
+%\item[] $\wei(q, \varepsilon, a, q) = \one$ for all $a \in \Omegac \cup \Omegar$,
 %\item[] $\wei(q, a, a, q) = \one$ for all $a \in \Omegai$,
@@
 $b \in {\hat\Omega} \setminus \Omegai$ in input, 
-with new transitions of the form $\wei_{01}(q, \epsilon, b, q')$.
+with new transitions of the form $\wei_{01}(q, \varepsilon, b, q')$.
 
@@
 %\noindent
-Let $\Omega = \Sigma \cup \{ \epsilon \} \times \Sigma \cup \{ \epsilon \} \setminus \{ (\epsilon, \epsilon) \}$,
+Let $\Omega = \Sigma \cup \{ \varepsilon \} \times \Sigma \cup \{ \varepsilon \} \setminus \{ (\varepsilon, \varepsilon) \}$,
 and let $h$ be the morphism from $\Omega^*$ into $\Sigma^* \times \Sigma^*$  
-defined over the concatenation of strings of $\Sigma^*$ (that removes the $\epsilon$'s).
+defined over the concatenation of strings of $\Sigma^*$ (that removes the $\varepsilon$'s).
 %

### 2021-05-27 main.tex
revise algo best w. notation operators complete label theory
convexity. macro stack

@@
 \newcommand{\pair}[2]{\langle{#1}, {#2}\rangle}
+
+%\newcommand{\config}[2]{\ensuremath \genfrac{[}{]}{0pt}{}{#1}{#2}} 
+\newcommand{\config}[2]{\ensuremath{#1}[{#2}]} 
+\newcommand{\configup}[2]{\ensuremath{#1}\left[\begin{array}{c} #2 \end{array}\right]}
+\def\stacksep{\cdot}
+\def\stackup{\\}
+
+\newcommand{\opluseq}{\ensuremath\mathrel{\oplus}=}
+
+% \bigominus
+\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
+\DeclareFontShape{U}{mathx}{m}{n}{
+<-6> mathx5 <6-7> mathx6 <7-8> matha7
+<8-9> mathx8 <9-10> mathx9
+<10-12> mathx10 <12-> mathx12
+}{}
+\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
+\DeclareMathSymbol{\bigominus}{\mathop}{mathx}{"C1}
+
 %\newcommand{\A}{\mathcal{A}}
@@
 \def\weiex{{\wei^\mathsf{e}_\mathsf{x}}}
-\newcommand{\config}[2]{\ensuremath \genfrac{[}{]}{0pt}{}{#1}{#2}} 
-\newcommand{\opluseq}{\ensuremath\mathrel{\oplus}=}
 \newcommand{\ioi}[1]{\mathsf{ioi}({#1})}
@@
 an intermediate function $\weight_A$, like in Section~\ref{sec:SWT}.
-In the case of a pushdown automaton, a configuration is composed 
-of a state $q \in Q$ and a stack content $\gamma \in \Gamma^*$, where $\Gamma = \Omegac \times P$.
+In the case of a pushdown automaton, a configuration, denoted by $q[\gamma]$ 
+is composed of a state $q \in Q$ 
+and a stack content $\gamma \in \Gamma^*$, 
+where $\Gamma = \Omegac \times P$.
 Hence, $\weight_A$ is a function from 
@@
 %\begin{array}[t]{rcl}
-\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot\gamma}, a\, u, \config{q'}{\gamma'}\bigr) & =  
+\weight_A\bigl(\configup{q}{\<{\call{c}}, p>\stackup\gamma}, a\, u, 
+               \config{q'}{\gamma'}\bigr) & =  
  {\displaystyle\bigoplus_{q'' \in Q}} \weii(q, c, p, a, q'') 
- \otimes \weight_A\bigl(\config{q''}{\gamma}, u, \config{q'}{\gamma'}\bigr)\nonumber\\
-\weight_A\bigl(\config{q}{\bot}, a\, u, \config{q'}{\gamma'}\bigr) & =  
+  \otimes \weight_A\bigl(\configup{q''}{\<{\call{c}}, p> \stackup \gamma}, u, 
+                         \config{q'}{\gamma'}\bigr)\nonumber\\
+ %
+\weight_A\bigl(\config{q}{\bot}, a\, u, 
+               \config{q'}{\gamma'}\bigr) & =  
   {\displaystyle\bigoplus_{q'' \in Q}} \weiei(q, a, q'') 
@@
 %
-\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot\gamma}, {\call{c}'}\, u, \config{q'}{\gamma'}\bigr) & =  
+\weight_A\bigl(\configup{q}{\<{\call{c}}, p>\stackup\gamma}, {\call{c}'}\, u, 
+               \config{q'}{\gamma'}\bigr) & =  
  {\displaystyle\bigoplus_{\begin{array}{c}
@@
  \weic\bigl(q, {\call{c}}, p, {\call{c}'}, p', q''\bigr) 
- \otimes \weight_A\bigl(\config{q''}{{\call{c}'}\, p' \cdot {\call{c}}\, p\cdot \gamma}, u, \config{q'}{\gamma'}\bigr)\nonumber\\[1mm]
-%
-\weight_A\bigl(\config{q}{\bot}, {\call{c}}\, u, \config{q'}{\gamma'}\bigr) & =  
+ \otimes \weight_A\bigl(\configup{q''}{\<{\call{c}'}, p'>\stackup \<{\call{c}}, p>\stackup \gamma}, u, 
+                        \config{q'}{\gamma'}\bigr)\nonumber\\[1mm]
+%
+\weight_A\bigl(\config{q}{\bot}, {\call{c}}\, u, 
+               \config{q'}{\gamma'}\bigr) & =  
  {\displaystyle\bigoplus_{q'' \in Q}} \weiec(q, {\call{c}}, p, q'') 
- \otimes \weight_A\bigl(\config{q''}{{\call{c}}\, p}, u, \config{q'}{\gamma'}\bigr)\nonumber\\
-%
-\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot \gamma}, {\return{r}}\, u, \config{q'}{\gamma'}\bigr) & =  
+  \otimes \weight_A\bigl(\config{q''}{{\call{c}}\, p}, u, 
+                         \config{q'}{\gamma'}\bigr)\nonumber\\
+%
+\weight_A\bigl(\configup{q}{\<{\call{c}}, p>\stackup \gamma}, {\return{r}}\, u, 
+               \config{q'}{\gamma'}\bigr) & =  
  {\displaystyle\bigoplus_{q'' \in Q}} 
   \weir\bigl(q, {\call{c}}, p, {\return{r}}, q''\bigr) 
- \otimes \weight_A\bigl(\config{q''}{\gamma}, u, \config{q'}{\gamma'}\bigr)\nonumber\\
-%
-\weight_A\bigl(\config{q}{\bot}, {\return{r}}\, u, \config{q'}{\gamma'}\bigr) & =  
+  \otimes \weight_A\bigl(\config{q''}{\gamma}, u, 
+                         \config{q'}{\gamma'}\bigr)\nonumber\\
+%
+\weight_A\bigl(\config{q}{\bot}, {\return{r}}\, u, 
+               \config{q'}{\gamma'}\bigr) & =  
  {\displaystyle\bigoplus_{q'' \in Q}} \weier(q, {\return{r}}, q'') 
- \otimes \weight_A\bigl(\config{q''}{\bot}, u, \config{q'}{\gamma'}\bigr)
+  \otimes \weight_A\bigl(\config{q''}{\bot}, u, 
+                         \config{q'}{\gamma'}\bigr)
 %\end{array}
@@
 %
-where $\bot$ denotes the empty stack and ${\call{c}}\, p\cdot \gamma$ 
-denotes a stack where the pair made of ${\call{c}} \in \Omegac$ and $p \in P$ is the top symbol 
-and $\gamma$ is the rest of stack.
+where $\bot$ denotes the empty stack.
+% and ${\call{c}}\, p\stacksep \gamma$ 
+%denotes a stack where the pair made of ${\call{c}} \in \Omegac$ and $p \in P$ is the top symbol 
+%and $\gamma$ is the rest of stack.
 
@@
   \textstyle
-  \weight_A\bigl(\config{q}{c\,p\cdot \top }, s, \config{q'}{c\, p\cdot \top}\bigr). 
+  \weight_A\bigl(\configup{q}{\< c, p> \stackup \top }, s, \configup{q'}{\<c, p> \stackup \top}\bigr). 
 \end{equation}
@@
   d_\top(q_1, p, q_2) \otimes 
-  \displaystyle\bigoplus_{a \in \Omegai} \weii(q_2, p, q_3)_a\\
-\multicolumn{3}{l}{\quad
-\mathrm{where~} \weii(q_2, p, q_3)_a \in \Phic
-\mathrm{~is~the~partial~application~}
-x_c \mapsto \weii(q_2, x_c, p, a, q_3)}\\
+  \displaystyle\bigoplus_{\Omegai} \weii(q_2, p, q_3)\\
+%\multicolumn{3}{l}{\quad
+%\mathrm{where~} \weii(q_2, p, q_3)_a \in \Phic
+%\mathrm{~is~the~partial~application~}
+%x_c \mapsto \weii(q_2, x_c, p, a, q_3)}\\
 %
@@
   d_\bot(q_1, q_2) \otimes 
-  \displaystyle\bigoplus_{a \in \Omegai} \weie(q_2, a, q_3)\\
+  \displaystyle\bigoplus_{\Omegai} \weiei(q_2, q_3)\\
 %     
 d_\top(q_0, p, q_3) & \opluseq &
-  \displaystyle\bigoplus_{{\call{c}'} \in \Omegac} 
-  \displaystyle\bigoplus_{{\return{r}} \in \Omegar} 
-  \weic(q_0, p, p', q_1)_{c'} \otimes 
-  d_\top(q_1, p', q_2) \otimes
-  \weir(q_2, c', p', r, q_3)\\
-\multicolumn{3}{l}{\quad
-\mathrm{where~} \weic(q_0, p, p', q_1)_{c'} \in \Phic
-\mathrm{~is~the~partial~application~}
-x_c \mapsto \weic(q_0, p, x_c, c', p', q_1)}\\[2pt]
+  {\displaystyle\bigoplus_{\Omegac}}^2 
+  \bigl[ \bigl( \weic(q_0, p, p', q_1) \otimes_2 
+  d_\top(q_1, p', q_2) \bigr) \otimes_2
+  {\displaystyle\bigoplus_{\Omegar}} \weir(q_2, p', q_3) \bigr]\\
+%\multicolumn{3}{l}{\quad
+%\mathrm{where~} \weic(q_0, p, p', q_1)_{c'} \in \Phic
+%\mathrm{~is~the~partial~application~}
+%x_c \mapsto \weic(q_0, p, x_c, c', p', q_1)}\\[2pt]
 %
 d_\bot(q_0, q_3) & \opluseq &
-  \displaystyle\bigoplus_{{\call{c}} \in \Omegac} 
-  \displaystyle\bigoplus_{{\return{r}} \in \Omegar} 
-  \weiec(q_0, c, p, q_1) \otimes 
-   d_\top(q_1, p, q_2)(c) \otimes
-  \weir(q_2, c, p, r, q_3)\\
+  {\displaystyle\bigoplus_{\Omegac}}
+  \bigl(
+  \weiec(q_0, p, q_1) \otimes 
+   d_\top(q_1, p, q_2) \otimes
+  \displaystyle\bigoplus_{\Omegar} \weir(q_2, p, q_3)\bigr)\\
 %
@@
   d_\bot(q_1, q_2) \otimes 
-  \displaystyle\bigoplus_{\return{r} \in \Omegar} \weier(q_2, {\return{r}}, q_3)\\
+  \displaystyle\bigoplus_{\Omegar} \weier(q_2, q_3)\\
 %
@@
 In order to obtain effectively a witness 
-(word of $\Semiring^*$ with computation of $A$ of minimal weight), 
-we require an additional property the of weight functions.
-%
-\begin{definition}
-Let $\Omega$ be an alphabet and $\Semiring$ a complete semiring.
-A function $\phi$ from $\Omega^n$ into $\Semiring$
-is called $k$-\emph{convex} for a natural number $k$ iff 
-$\mathit{card}\{ \mathbf{a} \in \Omega^n \mid \phi(\mathbf{a}) = 
-                 \displaystyle\bigoplus_{\mathbf{p} \in \Omega^n} \phi(\mathbf{p}) \} \leq k$.
-\end{definition}
-A label theory is $k$-convex if all its functions are $k$-convex.
+(word of $\Omega^*$ with a computation of $A$ of minimal weight), 
+we require the additional property of convexity of weight functions.
 
@@
 $\Semiring$ commutative, bounded, total and complete, %semiring 
-and $\bar\Phi$ $k$-convex effective, % label theory,
+and $\bar\Phi$ effective and $k$-convex, % label theory,
 one can construct in PTIME a word $s \in \Omega^*$ 

### 2021-05-28 introduction.tex
passe intro, semiring, label th.

@@
 of structuring a linear representation in input %typically
-(a finite word over an alphabet) according to a language model. % (a formal grammar).
+(a finite word) according to a language model. % (a formal grammar).
 % natural language, programming language, 
@@
 like constituency parsing or of programming languages compilers or interpreters.
-Considering large or infinite alphabets can however be of practical interest
-in other cases.
+Considering large or infinite alphabets can however be of practical interest. %in other cases.
 %
@@
 % (notion of monoid recognizability, based on registers, comparable to Bojancszik et al. data words)
-Regarding the latter case, we briefly mention
-at the end of the paper 
-a parse-based approach to automated music transcription~\cite{foscarin:hal-01988990}
-that motivated the present work. 
-In this problem, a symbolic music performance, 
-presented as a sequence of timed musical events, % timestamped
-% a symbolic representation of a music performance 
+The latter case is related to a case study that motivated the present work: 
+automated music transcription.
+In this problem, a symbolic representation of a music performance, % a symbolic representation of a music performance 
+in the form of a sequence of timed musical events, % timestamped
 is converted into a structured score in Common Western Music Notation.
-%generalizes weighted parsing: 
-%finding the best derivation for a weighted grammar.
+It can therefore be expressed as a parsing problem~\cite{foscarin:hal-01988990}, 
+over an infinite alphabet of timed events.
 
@@
 Other extensions of language models  %(automata and grammars) 
-help in dealing with non-determinism, by the computation of weight values. 
+help in dealing with non-determinism, with the computation of weight values. 
 With an ambiguous grammar, there may exists several derivations 
@@
 The association of one weight value %in semiring domain
-to each derivation permits to select a best one (or $n$ bests). % a fixed number of bests 
-% = ...a ranking of derivations 
+to each AST permits to select a best one (or $n$ bests). % a fixed number of bests 
+% = ...a ranking of ASTs 
 This is roughly the principle of \emph{weighted parsing}
@@
 and the rule's weights can be combined with a associative product operator~$\otimes$ into 
-the weight of a derivation.
+the weight of a AST.
 A second operator~$\oplus$, associative and commutative, 
@@
 In this paper, we present a uniform framework for weighted parsing over infinite input alphabets.
-It is based on weighted %finite states 
-language models generalizing 
+It is based on symbolic weighted  (SW) finite states language models generalizing 
 both~\SA, with functions in an arbitrary semiring instead of Boolean guards, 
@@
 % non-terminals and production rules).
-In their transition rules, input symbols appear as variables %(parameters)
+In the transition rules of SW models, input symbols appear as variables %(parameters)
 and the weight associated to a transition rule is a function of these variables.
@@
 based on properties of the SW models. 
-The first step is a Bar-Hillel construction where, 
-given a \SWT $T$, a \SWVPA $A$, and an input word $s$, 
+The first step is an intersection 
+(Bar-Hillel construction~\cite{GruneJacobs08parsing} \marginpar{chap. intersection in~\cite{GruneJacobs08parsing}}) 
+where, given a \SWT $T$, a \SWVPA $A$, and an input word $s$, 
 a \SWVPA $A_{T, s}$ is built, such that for all $t$, $A_{T, s}(t) = T(s, t) \otimes A(t)$.
@@
 ($iii.a$)~the \SWT-based definition of generic edit distances between input and output words,
-and ($iii.b$)~the use of nested words, and \SWVPA, 
+and ($iii.b$)~the use in this context of nested words, and \SWVPA, 
 instead of syntax trees and grammars. %(\S~\ref{sec:trees}).

### 2021-05-28 semiring.tex
passe intro, semiring, label th.

@@
 \begin{itemize}
-\item $\oplus$ is commutative;
+\item $\oplus$ is commutative: 
  $\< \Semiring, \oplus, \zero>$ is a commutative monoid 
@@
 (Figure~\ref{fig:semirings}).
+%
+\noindent
+An idempotent semiring $\Semiring$~is called \emph{total} if
+it~$\leq_\oplus$ is total
+\ie when for all $x, y \in \Semiring$, either $x \oplus y = x$ or $x \oplus y = y$.
 
@@
 
-In~\cite{Huang08advanceddynamic}, the property~$(i)$ 
-is called $\Semiring$ \emph{superior} \wrt the ordering~$\leq_\oplus$.
+In~\cite{Huang08advanceddynamic}, when the property~$(i)$ holds,  
+$\Semiring$ is called \emph{superior} \wrt the ordering~$\leq_\oplus$.
 We have seen in the proof of Lemma~\ref{lem:bounded} that it implies that 
@@
 %
-In~\cite{Mohri02semiring}, the property ($ii$) of Lemma~\ref{lem:bounded}  
-is called \emph{boundedness} of $\Semiring$ -- we shall use this term in the rest of the paper. 
+In~\cite{Mohri02semiring}, 
+$\Semiring$ with the property ($ii$) of Lemma~\ref{lem:bounded}  
+is called \emph{bounded} -- we shall use this term in the rest of the paper. 
 % \emph{negative boundedness} 
 It implies that, when looking for a best path in a graph whose edges
-are weighted by values of $\Semiring$, the loops can be safely avoided.
-
-%Following the terminology of~\cite{Mohri02semiring},
-%when $\forall x \in \dom(\Semiring), \one \oplus x = \one$,
-%the semiring $\Semiring$ is is called \emph{bounded}.
+are weighted by values of $\Semiring$, the loops can be safely avoided
+(because, for all $x \in \Semiring$ and $n \geq 1$, 
+ $x \oplus x^n = x \otimes (\one \oplus x^{n-1}) = x$).
 
@@
 \end{proof}
-
-\noindent
-An idempotent semiring $\Semiring$~is called \emph{total} if
-it~$\leq_\oplus$ is total
-\ie when for all $x, y \in \Semiring$, either $x \oplus y = x$ or $x \oplus y = y$.
 

### 2021-05-28 main.tex
passe intro, semiring, label th.
revise def. runs SWT and SWA

@@
 
-notations: for set $S$ : $S^*$ and $S^+$.
-interval $[i..j]$ of natural numbers
+%notations: for set $S$ : sets of sequences $S^*$ and $S^+$...
+%interval $[i..j]$ of natural numbers
 
@@
 $\Sigma$ and $\Delta$ be alphabets called respectively \emph{input} and \emph{output}, %{alphabets}, 
-and $\bar\Phi = \< \Phi_\Sigma, \Phi_\Delta, \Phi_{\Sigma, \Delta}>$ be a label theory over $\Semiring$.
+and $\bar\Phi$ be a label theory over $\Semiring$
+containing $\Phi_\Sigma$, $\Phi_\Delta$, $\Phi_{\Sigma, \Delta}$.
 
@@
 \noindent
-For convenience, we shall the following presentation of transition functions 
+For convenience, we shall sometimes present transition functions 
 as functions of  
@@
 \noindent 
+% The symbolic-weighted transducer
 The \SWT $T$ computes on pairs of words $\< s, t> \in \Sigma^* \times \Delta^*$,
 $s$ and $t$, being respectively called input and output word.
-%
 % \noindent
-More precisely, the symbolic-weighted transducer~$T$ defines a mapping 
-from the pairs of strings of $\Sigma^* \times \Delta^*$ 
-into~$\Semiring$,
-based on the following intermediate function $\weight_T$
-defined recursively for every $q, q' \in Q$, 
-%$a \in \Sigma$, $b \in \Delta$ 
-%for every strings of $s \in \Sigma^*$, $t \in \Delta^*$:
-$a \in \Sigma$, $u \in \Sigma^*$, $b\in \Delta$, $v\in \Delta^*$, 
+More precisely, $T$ defines a mapping 
+from $\Sigma^* \times \Delta^*$ into~$\Semiring$,
+based on an intermediate function $\weight_T$
+defined recursively for every states $q, q' \in Q$, 
+and every strings $\< s, t> \in \Sigma^* \times \Delta^* \setminus \{ \< \varepsilon, \varepsilon >\}$ by:
+%$a \in \Sigma$, $u \in \Sigma^*$, $b\in \Delta$, $v\in \Delta^*$, by:
 %
 \begin{align}
-\weight_T(q, \varepsilon, \varepsilon, q)  & = \final(q)\\
+\weight_T(q, \varepsilon, \varepsilon, q)  & = \one \label{eq:SWT-weight}\\ %\final(q)\\
 \weight_T(q, \varepsilon, \varepsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
-\weight_T(q, au, \varepsilon, q') & = \displaystyle\bigoplus_{q'' \in Q} 
-    \wei_{10}(q, a, \varepsilon, q'') \otimes \weight_T(q'', u, \varepsilon, q')\nonumber\\
-\weight_T(q, \varepsilon, bv, q') & = \displaystyle\bigoplus_{q'' \in Q} 
-    \wei_{01}(q, \varepsilon, b, q'') \otimes \weight_T(q'', \varepsilon, v, q')\nonumber\\
-\weight_T(q, au, bv, q') & = \displaystyle\bigoplus_{q'' \in Q} 
-    \wei_{10}(q, a, \varepsilon, q'') \otimes \weight_T(q'', u, bv, q')\nonumber\\
-                    & \oplus \displaystyle\bigoplus_{q'' \in Q} 
-    \wei_{01}(q, \varepsilon, b, q'') \otimes \weight_T(q'', au, v, q')\nonumber\\
-                    & \oplus \displaystyle\bigoplus_{q'' \in Q} 
+%\weight_T(q, au, \varepsilon, q') & = \displaystyle\bigoplus_{q'' \in Q} 
+%    \wei_{10}(q, a, \varepsilon, q'') \otimes \weight_T(q'', u, \varepsilon, q')\nonumber\\
+%\weight_T(q, \varepsilon, bv, q') & = \displaystyle\bigoplus_{q'' \in Q} 
+%    \wei_{01}(q, \varepsilon, b, q'') \otimes \weight_T(q'', \varepsilon, v, q')\nonumber\\
+\weight_T(q, s, t, q') & = \displaystyle\bigoplus_{\begin{array}{c}
+                                                   \scriptstyle q'' \in Q\\
+                                                   \scriptstyle s = au,\, a \in \Sigma
+                                                   \end{array}} 
+    \wei_{10}(q, a, \varepsilon, q'') \otimes \weight_T(q'', u, t, q')\nonumber\\
+                    & \oplus \displaystyle\bigoplus_{\begin{array}{c}
+                                                     \scriptstyle q'' \in Q\\ 
+                                                     \scriptstyle t = bv,\, b \in \Delta\\
+                                                     \end{array}} 
+    \wei_{01}(q, \varepsilon, b, q'') \otimes \weight_T(q'', s, v, q')\nonumber\\
+                    & \oplus \displaystyle\bigoplus_{\begin{array}{c}
+                                                     \scriptstyle q'' \in Q\\
+                                                     \scriptstyle s = au,\, t = bv\\
+                                                     \end{array}} 
     \wei_{11}(q, a, b, q'') \otimes \weight_T(q'', u, v, q')\nonumber
-\label{eq:SWT-weight}
 \end{align}
@@
 
-The expression of $\weight_T$ can be seen as a stateful definition of 
+The expression \eqref{eq:SWT-weight} 
+can be seen as a stateful definition of 
 an edit-distance between a word $s \in \Sigma^*$ and a word $t \in \Delta^*$,
@@
 And the distance between $s$ and $t$ is the sum, with $\oplus$,
-of all such product of costs.
+of all possible products.
 
@@
 %Let $\< s, t> \in \Sigma^* \times \Delta^*$, with $s = s_1\ldots s_n$, and $t = t_1\ldots t_m$. 
-The weight associated by $T$ to $\< s, t> \in \Sigma^* \times \Delta^*$
-is defined as follows:
+Formally, the weight associated by $T$ to $\< s, t> \in \Sigma^* \times \Delta^*$ is: 
+%defined as follows:
 \begin{equation}
 T(s, t)  = 
-\displaystyle\bigoplus_{q, q' \in Q} \mathsf{in}(q) 
-\mathop{\otimes} \weight_T(q, s, t, q') %\mathop{\otimes} \mathsf{out}(q', s_n, t_m)
+\displaystyle\bigoplus_{q, q' \in Q} \init(q) 
+\mathop{\otimes} \weight_T(q, s, t, q') \mathop{\otimes} \final(q')
 \label{eq:SWT-value}
@@
 \begin{example}
+In Common Western Music Notation~\cite{Gould11BB}, 
+several symbols may be used to represent a unique sounding event.
+For instance, several notes can be combined with a tie, 
+like in $\musQuarter\!\!\mathrel{\raisebox{-1.5mm}{$\smile$}}\!\!\musEighth$,  
+and a note can be augmented with a dot like in~\musQuarterDotted).
+We propose a small transducer that compares an input sequence of sounding 
+events (music "performance") 
+to an output sequence of written events (music "score").
 % simple pointwise distance between two sequences of timestamped events **
+
 Let us consider the tropical (\emph{min-plus}) semiring~$\Semiring$ 
@@
 %$\Sigma = \{ a, - \} \times \R_+$ and 
-This example of $\Delta$ is motivated by the case of music notation, 
-where several notated events (notes) can be tied together, 
-with a \emph{tie} or a \emph{dot}
-(like in $\musQuarter\!\!\mathrel{\raisebox{-1.5mm}{$\smile$}}\!\!\musEighth$ 
- or equivalently~\musQuarterDotted),
-meaning that they will be played as a unique sounding event.
+%This example of $\Delta$ is motivated by the case of music notation, 
+%where several notated events (notes) can be tied together, 
+%with a \emph{tie} or a \emph{dot}
+%meaning that they will be played as a unique sounding event.
 %The timestamp of $a \in \Sigma$, denoted by $\mathsf{t}(a)$, is expressed as a rational number.
@@
 \begin{align}
-\weight_A(q, \varepsilon, q) & = \final(q)\\
+\weight_A(q, \varepsilon, q) & = \one\\ %\final(q)\\
 \weight_A(q, \varepsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
@@
 A(s)  = 
-\displaystyle\bigoplus_{q, q' \in Q} \mathsf{in}(q) 
-\mathop{\otimes} \weight_A(q, s, q') 
+\displaystyle\bigoplus_{q, q' \in Q} \init(q) 
+\mathop{\otimes} \weight_A(q, s, q') \mathop{\otimes} \final(q') 
 \label{eq:SWA-value}

### 2021-06-03 image.tex
versions

--- 2021-05-27 main.tex
@@
-% SWT without epsilon-transitions
-% SWT do not read the other tape in "skip" transitions 10 or 01
-%     (hence this version is simpler  than V2)
-% SWA without epsilon-transitions
-% SW-VPA read top of stack (when not empty) at every transition 
-%        (6 kinds of transitions).
-
-\documentclass[runningheads]{llncs}
-%\documentclass[a4paper,11pt]{article}
-%\setcounter{page}{1}
-
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-
-\usepackage{hyperref}
-%\usepackage[bookmarks,bookmarksnumbered,naturalnames,plainpages=false]{hyperref}
-%usepackage{url}
-
-% for footnote ref
-\usepackage{refcount}
-
-% array and tabular
-\usepackage{array}
-\newcolumntype{C}[1]{>{\centering\arraybackslash\hspace{0pt}}p{#1}}
-
-% extension of enumerate env. (style for displaying counters)
-% \usepackage{enumerate}
-
-%% pictures
-% \usepackage{graphicx} 
-% \DeclareGraphicsExtensions{.pdf,.png,.jpg}
-% \graphicspath{fig/}
-
-%% PGF, Tikz
-\usepackage{tikz-cd}
-%% \usepackage{pgfplots}
-%% \usepgfplotslibrary{dateplot}
-%% \usepackage{pgf,pgfarrows,pgfnodes, pgfautomata}
-% \usepackage{tikz}
-% \usetikzlibrary{cd}
-%% \usetikzlibrary{arrows}
-%% \usetikzlibrary{calc}
-%% \usetikzlibrary{snakes}
-%% \usetikzlibrary{backgrounds}
-% \usetikzlibrary{trees}
-%% \usetikzlibrary{automata}
-%% \usetikzlibrary{positioning}
-%% \usetikzlibrary{matrix}
-%% \usetikzlibrary{patterns}
-%% \usetikzlibrary{shapes}
-
-% symbols
-\usepackage{amsmath} 
-\allowdisplaybreaks
-\usepackage{amssymb} 
-\usepackage{amsbsy}
-\usepackage{bbold}
-\usepackage{latexsym}
-%\usepackage{amsfonts}
-\usepackage{stmaryrd}
-%\usepackage{mathabx}
-%\usepackage{MnSymbol}
-\usepackage{harmony} % simple music fonts
-\usepackage{mathtools} % for arrows
-%\usepackage{mathptmx}
-
-%% theorem environments
-\usepackage{theorem}
-% \newtheorem{theorem}{Theorem} %[section]
-% \newtheorem{definition}[theorem]{Definition}
-% \newtheorem{lemma}[theorem]{Lemma}
-% \newtheorem{corollary}[theorem]{Corollary}
-% \newtheorem{proposition}[theorem]{Proposition}
-% \newenvironment{proof}{\vspace{-2ex}{\it Proof. }}{\hspace*{\fill} $\Box$\smallskip }
-% \theorembodyfont{\slshape}
-% \newtheorem{example}[theorem]{Example}
-% \newtheorem{remark}[theorem]{Remark}
-\def\endex{\hspace*{\fill} $\Diamond$\smallskip }
-{\theorembodyfont{\rmfamily} \theoremstyle{break} \newtheorem{algo}{Algorithm}}
-
-%% algorithms
-%\usepackage{algorithm} 
-%\usepackage{program}
-
-%% allows for page break within arrays
-\usepackage{longtable}
-
-%% arrows etc
-%\input{rewriting}
-
-%% music symbols
-% see http://tug.ctan.org/info/latex4musicians/latex4musicians.pdf
-\usepackage{musicography}
-
-%% for new macros
-\usepackage{xspace}
-
-%% Misc macros
-\def\ie{\textit{i.e.}\xspace}
-\def\eg{\textit{e.g.}\xspace}
-\def\wrt{\textit{wrt}\xspace}
-%\def\wlog{\textit{wlog}\xspace}
-\def\etc{\textit{etc}\xspace}
-
-\def\<#1>{\langle #1 \rangle}
-\newcommand{\pair}[2]{\langle{#1}, {#2}\rangle}
-%\newcommand{\A}{\mathcal{A}}
-%\newcommand{\B}{\mathcal{B}}
-\newcommand{\D}{\mathbb{D}}
-\newcommand{\E}{\mathbb{E}}
-%\newcommand{\P}{\mathcal{P}}
-\newcommand{\Q}{\mathcal{Q}}
-\newcommand{\R}{\mathcal{R}}
-\newcommand{\T}{\mathcal{T}}
-\newcommand{\W}{\mathbb{W}}
-\newcommand{\Semiring}{\mathbb{S}}
-\newcommand{\zero}{\mathbb{0}}
-\newcommand{\one}{\mathbb{1}}
-\newcommand{\dom}{\ensuremath{\mathit{dom}}}
-
-\def\SA{\textsf{sA}\xspace}
-\def\WA{\textsf{wA}\xspace}
-\def\SWT{\textsf{swT}\xspace}
-\def\SWA{\textsf{swA}\xspace}
-\def\SWTA{\textsf{swTA}\xspace}
-\def\SWVPA{\textsf{sw-VPA}\xspace}
-\def\weight{\mathsf{weight}}
-\def\wei{\mathsf{w}}
-\def\mei{\mathsf{m}}
-\def\init{\mathsf{in}}
-\def\final{\mathsf{out}}
-\newcommand{\call}[1]{\ensuremath #1} %{\ensuremath \langle_{#1}}
-\newcommand{\return}[1]{\ensuremath #1} %{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
-\def\Omegai{{\Omega_\mathsf{i}}}
-\def\Omegac{{\Omega_\mathsf{c}}}
-\def\Omegar{{\Omega_\mathsf{r}}}
-\def\Sigmai{{\Sigma_\mathsf{i}}}
-\def\Sigmac{{\Sigma_\mathsf{c}}}
-\def\Sigmar{{\Sigma_\mathsf{r}}}
-\def\Deltai{{\Delta_\mathsf{i}}}
-\def\Deltac{{\Delta_\mathsf{c}}}
-\def\Deltar{{\Delta_\mathsf{r}}}
-\def\Phii{{\Phi_\mathsf{i}}}
-\def\Phic{{\Phi_\mathsf{c}}}
-\def\Phir{{\Phi_\mathsf{r}}}
-\def\Phix{{\Phi_\mathsf{x}}}
-\def\Phici{{\Phi_\mathsf{ci}}}
-\def\Phicc{{\Phi_\mathsf{cc}}}
-\def\Phicr{{\Phi_\mathsf{cr}}}
-\def\Phicx{{\Phi_\mathsf{cx}}}
-\def\weii{{\wei_\mathsf{i}}}
-\def\weic{{\wei_\mathsf{c}}}
-\def\weir{{\wei_\mathsf{r}}}
-\def\weie{{\wei_\mathsf{e}}}
-\def\weiei{{\wei^\mathsf{e}_\mathsf{i}}}
-\def\weiec{{\wei^\mathsf{e}_\mathsf{c}}}
-\def\weier{{\wei^\mathsf{e}_\mathsf{r}}}
-\def\weiex{{\wei^\mathsf{e}_\mathsf{x}}}
-\newcommand{\config}[2]{\ensuremath \genfrac{[}{]}{0pt}{}{#1}{#2}} 
-\newcommand{\opluseq}{\ensuremath\mathrel{\oplus}=}
-\newcommand{\ioi}[1]{\mathsf{ioi}({#1})}
-\newcommand{\rank}{\mathsf{rk}}
-\newcommand{\lin}{\mathsf{lin}}
-
-%\sloppy
-
-% Parsing over infinite alphabet as optimal alignment computation 
-% as edit-distance between string and language
-% 
-%\title{Symbolic Weighted Parsing and Automated Music Transcription}
-%\title{Symbolic Weighted Language Models and Automated Music Transcription}
-\title{Symbolic Weighted Language Models and\\ Parsing over Infinite Alphabets}
-%\title{Weighted Visibly Pushdown Automata and Automated Music Transcription}
-\author{Florent Jacquemard}
-\institute{INRIA \& CNAM, Paris, France\\
-\email{florent.jacquemard@inria.fr}}
-
-%\titlerunning{WVPA \& AMT}
-%\titlerunning{Symbolic Weighted Parsing} % and Automated Music Transcription
-%\authorrunning{Florent Jacquemard}
-
-\date{\today}
-
-\begin{document}
-\thispagestyle{empty}
-\maketitle
-
-\begin{abstract}
-\input{abstract}
-\end{abstract}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% intro
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Introduction} \label{sec:intro}
-\input{introduction}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% prelin
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Preliminary Notions}
-\label{section:prelim}\label{sec:prelim}
-
-notations: for set $S$ : $S^*$ and $S^+$.
-interval $[i..j]$ of natural numbers
-
-\subsection{Semirings} 
-\label{section:semiring}\label{sec:semiring}
-\input{semiring}
-
-\subsection{Label Theory}
-\label{section:symbols}
-\input{labels}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SWT & SWA
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{SW Automata and Transducers}
-\label{section:transducer}\label{sec:transducer}
-\label{section:SWA}\label{sec:SWA}
-\label{section:SWT}\label{sec:SWT}
-
-We follow the approach of~\cite{Mohri03EDWA} for the computation of distances,
-between words and languages, using weighted transducers, 
-and extend it to infinite alphabets.
-% with models of symbolic weighted automata and transducers. 
-%
-The models introduced in this section generalize 
-weighted automata and transducers~\cite{Droste09handbook} 
-%over finite alphabets, see  e.g.~\cite{Mohri03EDWA}, 
-by labelling each transition with a weight function (instead of a simple weight value), 
-that takes the input and output symbols as parameters. 
-These functions are similar to the guards of symbolic automata~\cite{dAntoniVeanes17CAV,dAntoni21CACM},
-but they can return values in an generic semiring, 
-whereas the latter guards are restricted to the Boolean semiring.
-
-%\subsection{Definitions} \label{sec:SWTdef}\label{sec:SWAdef}
-%(SWT)
-\noindent 
-Let $\Semiring$ be a commutative semiring, 
-$\Sigma$ and $\Delta$ be alphabets called respectively \emph{input} and \emph{output}, %{alphabets}, 
-and $\bar\Phi = \< \Phi_\Sigma, \Phi_\Delta, \Phi_{\Sigma, \Delta}>$ be a label theory over $\Semiring$.
-
-\begin{definition}
-\label{def:transducer} \label{def:SWT}
-A \emph{symbolic-weighted transducer} (\SWT)
-over $\Sigma$, $\Delta$, $\Semiring$ and $\bar\Phi$
-%the input and output alphabets~$\Sigma$ and $\Delta$ with label theory $\bar\Phi$, and the semiring $\Semiring$ 
-is a tuple
-$T = \< Q, \init, \bar{\wei}, \final >$,
-where $Q$ is a finite set of states, 
-$\mathsf{in} : Q \to \Semiring$   %\Phi_{\Sigma, \Delta}
-respectively $\mathsf{out} : Q \to \Semiring$  %\Phi_{\Sigma, \Delta}
-are functions defining the weight for entering, 
-respectively leaving, a state, 
-and $\bar{\wei}$ is a triplet of transition functions 
-$\wei_{10}: Q \times Q \to \Phi_{\Sigma}$,
-$\wei_{01}: Q \times Q \to \Phi_{\Delta}$, and 
-$\wei_{11}: Q \times Q \to \Phi_{\Sigma, \Delta}$.
-\end{definition}
-%
-\noindent
-For convenience, we shall the following presentation of transition functions 
-as functions of  
-$Q \times (\Sigma \cup \{ \varepsilon \}) \times (\Delta \cup \{ \varepsilon \}) \times Q \to \Semiring$,
-overloading the function names,
-such that, for all $q, q' \in Q$, $a \in \Sigma$,  $b \in \Delta$, 
-\[
-\begin{array}{rcll}
-\wei_{10}(q, a, \varepsilon, q') & = & \phi(a) & %\wei_\Sigma(q, q')(a)
-\quad\mathrm{where~} \phi = \wei_{10}(q, q') \in \Phi_\Sigma,\\
-\wei_{01}(q, \varepsilon, b, q') & = & \psi(b) & 
-\quad\mathrm{where~} \psi = \wei_{01}(q, q') \in \Phi_\Delta,\\
-\wei_{11}(q, a, b, q') & = & \eta(a, b) & 
-\quad\mathrm{where~} \eta = \wei_{11}(q, q') \in \Phi_{\Sigma, \Delta}.\\
-\end{array}      
-\]
-%
-\noindent 
-The \SWT $T$ computes on pairs of words $\< s, t> \in \Sigma^* \times \Delta^*$,
-$s$ and $t$, being respectively called input and output word.
-%
-% \noindent
-More precisely, the symbolic-weighted transducer~$T$ defines a mapping 
-from the pairs of strings of $\Sigma^* \times \Delta^*$ 
-into~$\Semiring$,
-based on the following intermediate function $\weight_T$
-defined recursively for every $q, q' \in Q$, 
-%$a \in \Sigma$, $b \in \Delta$ 
-%for every strings of $s \in \Sigma^*$, $t \in \Delta^*$:
-$a \in \Sigma$, $u \in \Sigma^*$, $b\in \Delta$, $v\in \Delta^*$, 
-%
-\begin{align}
-\weight_T(q, \varepsilon, \varepsilon, q)  & = \final(q)\\
-\weight_T(q, \varepsilon, \varepsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
-\weight_T(q, au, \varepsilon, q') & = \displaystyle\bigoplus_{q'' \in Q} 
-    \wei_{10}(q, a, \varepsilon, q'') \otimes \weight_T(q'', u, \varepsilon, q')\nonumber\\
-\weight_T(q, \varepsilon, bv, q') & = \displaystyle\bigoplus_{q'' \in Q} 
-    \wei_{01}(q, \varepsilon, b, q'') \otimes \weight_T(q'', \varepsilon, v, q')\nonumber\\
-\weight_T(q, au, bv, q') & = \displaystyle\bigoplus_{q'' \in Q} 
-    \wei_{10}(q, a, \varepsilon, q'') \otimes \weight_T(q'', u, bv, q')\nonumber\\
-                    & \oplus \displaystyle\bigoplus_{q'' \in Q} 
-    \wei_{01}(q, \varepsilon, b, q'') \otimes \weight_T(q'', au, v, q')\nonumber\\
-                    & \oplus \displaystyle\bigoplus_{q'' \in Q} 
-    \wei_{11}(q, a, b, q'') \otimes \weight_T(q'', u, v, q')\nonumber
-\label{eq:SWT-weight}
-\end{align}
-%
-We recall that, by convention (Section~\ref{sec:semiring}), 
-an empty sum with $\bigoplus$ is equal to~$\zero$. 
-%
-Intuitively, using a transition $\wei_{ij}(q, a, b, q')$ means for $T$:
-when reading respectively $a$ and $b$ at the current positions in the input and output words, 
-increment the current position in the input word if and only if $i = 1$, 
-and in the output word iff $j = 1$ (otherwise, do not change it),
-and change state from $q$ to $q'$.
-When $a = \varepsilon$ (resp. $b = \varepsilon$), the current symbol 
-in the input (resp. output) is not read.
-%
-%In contrast with the models of weighted transducers over finite alphabets~\cite{Mohri03ijfcs},
-%the input and output symbols at current positions are always read by transitions, 
-%even when they do not change the reading position the head's position.
-%This is an important feature in the case of an infinite alphabet in 
-%order to compare input and output symbols.
-%which cannot be stored in the finite memory of the transducer.
-%
-Since $\zero$ is absorbing for~$\otimes$ in~$\Semiring$,
-one term $\wei_{ij}(q, a, b, q'')$ equal to $\zero$ in the above expression 
-will be ignored in the sum, meaning that there is no possible transition
-from state $q$ into state $q'$ while reading $a$ and $b$.
-This is analogous to the case of a transition's guard not satisfied by $\<a, b>$ for 
-symbolic transducers.
-
-%whereas considering the current symbol may be useful to compute a transition weight.
-%(even when it does not change the head's position, like with $\varepsilon$-transitions).
-%
-%The cases $\weight_T(q, au, \varepsilon, q')$ and $\weight_T(q, \varepsilon, bv, q')$
-%are missing in the definition of $\weight_T$.
-%It means that $T$ must avoid configurations where it reached the end of 
-%the output word and not of the input one, or vice-versa.
-%This can be done by using $\wei_{10}$ and $\wei_{01}$
-%before reaching the end of word, and using a special state for this purpose.
-
-The expression of $\weight_T$ can be seen as a stateful definition of 
-an edit-distance between a word $s \in \Sigma^*$ and a word $t \in \Delta^*$,
-see also~\cite{Mohri03ijfcs}.
-Intuitively, 
-$\wei_{10}(q, a, \varepsilon, r)$ is the cost of 
-the deletion of the symbol~$a \in \Sigma$ in~$s$, 
-$\wei{01}(q, \varepsilon, b, r)$ is the cost 
-of the insertion of~$b \in \Delta$ in $t$, 
-and $\wei_{11}(q, a, b, r)$ is the cost 
-of the substitution of  $a \in \Sigma$ by~$b \in \Delta$.
-%
-The cost of a sequence of such operations transforming $s$ into $t$, 
-is the product, with $\otimes$, of the individual costs of the operations involved;
-And the distance between $s$ and $t$ is the sum, with $\oplus$,
-of all such product of costs.
-
-\medskip\noindent
-%Let $\< s, t> \in \Sigma^* \times \Delta^*$, with $s = s_1\ldots s_n$, and $t = t_1\ldots t_m$. 
-The weight associated by $T$ to $\< s, t> \in \Sigma^* \times \Delta^*$
-is defined as follows:
-\begin{equation}
-T(s, t)  = 
-\displaystyle\bigoplus_{q, q' \in Q} \mathsf{in}(q) 
-\mathop{\otimes} \weight_T(q, s, t, q') %\mathop{\otimes} \mathsf{out}(q', s_n, t_m)
-\label{eq:SWT-value}
-\end{equation}
-
-%\begin{example}
-%We can define with a \SWT the computation of a similarity measure 
-%between timed sequences
-%similar to dynamic time warping (DTW).
-%
-%Let $\Semiring$  be the tropical (\emph{min-plus}) semiring of Figure~\ref{fig:semirings} and 
-%let $\Sigma = \Delta = \mathbb{R}_+$ be sets of timestamps.
-%We consider a \SWT with one state $q$ and transitions
-%$\wei_{11}(q, d, d', q) = 
-% \wei_{01}(q, d, d', q) = 
-% \wei_{10}(q, d, d', q) = |d' - d|$,
-%for all $d, d' \in \mathbb{R}_+$.
-%% needs reading input/output symbols even by epsilon-transitions.
-%The recursive definition of $\weight_T$ correspond to the dynamic programming equations of DTW
-%for the computation of an optimal match between words, 
-%the matching cost for two symbols being the 
-%the time distance between them.
-%\endex
-%\end{example}
-
-\begin{example}
-% simple pointwise distance between two sequences of timestamped events **
-Let us consider the tropical (\emph{min-plus}) semiring~$\Semiring$ 
-of Figure~\ref{fig:semirings} and 
-let $\Sigma = \mathbb{R}_+$ be an input alphabet of event dates
-and $\Delta = \{ \mathsf{e}, \mathsf{-} \} \times \mathbb{R}_+$ 
-be an output alphabet of symbols with timestamps. 
-A symbol $\< \mathsf{e}, d > \in \Delta$ represents an event starting at date $d$, 
-and $\< \mathsf{-}, d >$ is a continuation of the previous event.
-%More precisely, we let 
-%$\Sigma = \{ a, - \} \times \R_+$ and 
-This example of $\Delta$ is motivated by the case of music notation, 
-where several notated events (notes) can be tied together, 
-with a \emph{tie} or a \emph{dot}
-(like in $\musQuarter\!\!\mathrel{\raisebox{-1.5mm}{$\smile$}}\!\!\musEighth$ 
- or equivalently~\musQuarterDotted),
-meaning that they will be played as a unique sounding event.
-%The timestamp of $a \in \Sigma$, denoted by $\mathsf{t}(a)$, is expressed as a rational number.
-
-We consider a \SWT with two states $q_0$ and $q_1$ whose purpose 
-is to compare a recorded performance $s \in \Sigma^*$
-with notated music sheet $t \in \Delta^*$.
-One timestamp $d_i \in \Sigma$ may corresponds 
-to one notated event $\<\mathsf{e}, d'_i> \in \Sigma$, in which case 
-the weight value computed by the \SWT is the time distance between both
-(see transitions $\wei_{11}$ below).
-%
-If $\<\mathsf{e}, d'_i>$ is followed by continuations 
-$\<\mathsf{-}, d'_{i+1}>$..., they are just skip with no cost (transitions $\wei_{01}$ or weight $\one$).
-%transitions
-\[
-\begin{array}{rclcrcl}
-\wei_{11}(q_0, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d| & \quad &
-\wei_{11}(q_1, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d|\\
-\wei_{01}(q_0, \varepsilon, \< \mathsf{-}, d'>, q_0) & = & \one & &
-\wei_{01}(q_1, \varepsilon, \< \mathsf{-}, d'>, q_0) & = & \one\\
-\wei_{10}(q_0, d, \varepsilon, q_1) & = & \alpha & & %\multicolumn{3}{l}{\mathrm{for~all~} b \in \Delta}
-\end{array}
-\]
-%
-Moreover, it may happen that the performers plays an extra note accidentally, but only once in a row. 
-This is modelled by the transition $\wei_{10}$ with an arbitrary weight value $\alpha \in \Semiring$, 
-switching from state $q_0$ (normal) to $q_1$ (error).
-The transitions in the second column below switch back to the normal state $q_0$.
-% the metric computed by the \SWT is the smallest sum of point wise distances 
-% between dates of input and output events.
-At last, we let $q_0$ be the only initial and final state, with
-$\init(q_0) = \final(q_0) = \one$, and 
-$\init(q_1) = \final(q_1) = \zero$. 
-%$\init(q_0, d, b) = \final(q_0, d, b) = \one$, and 
-%$\init(q_1, d, b) = \final(q_1, d, b) = \zero$, 
-%for all $d \in \Sigma$ and $b \in \Delta$.
-\endex
-\end{example}
-
-\noindent
-The \emph{Symbolic Weighted Automata} %$A = \< Q, \init, \weight, \final >$
-%over $\Sigma$,  $\Semiring$ and $\bar\Phi$
-are defined similarly as the transducers of Definition~\ref{def:SWT}, 
-by simply omitting the output symbols.
-%
-In this case, the label theory $\bar\Phi$ can be reduced to a singleton $\< \Phi_\Sigma>$.
-%over $\Sigma$ is reduced to
-%a set $\Phi_\Sigma$ closed under~$\oplus$ and~$\otimes$.
-%
-\begin{definition} \label{def:SWA}
-A \emph{symbolic-weighted automaton} (\SWA)
-over $\Sigma$, $\Semiring$ and $\bar\Phi$
-%the input alphabet~$\Sigma$ and the commutative semiring $\Semiring$ 
-is a tuple
-$A = \< Q, \init, {\wei_1}, \final >$,
-where $Q$ is a finite set of states, 
-$\mathsf{in} : Q \to \Semiring$, %\Phi_\Sigma$, 
-respectively $\mathsf{out} : Q \to \Semiring$  %\Phi_\Sigma,$
-are functions defining the weight for entering, 
-respectively leaving, a state, 
-and ${\wei_1}$ is a transition functions 
-from $Q \times Q$ into $\Phi_{\Sigma}$.
-\end{definition}
-%      
-\noindent
-As above in the case of \SWT, 
-when $\wei_1(q, q') = \phi \in \Phi_\Sigma$, 
-%respectively $\mathsf{in}(q) = \phi$, $\mathsf{out}(q') = \phi$,
-we may write $\wei_1(q, a, q')$ for~$\phi(a)$. 
-%respectively $\mathsf{in}(q, a) = \phi(a)$, $\mathsf{out}(q', a) = \phi(a)$.
-%$\wei_1: Q \times \Sigma \times Q \to \Semiring$, 
-The computation of $A$ on words $s \in \Sigma^*$
-is defined with an intermediate function $\weight_A$, 
-defined as follows for $q, q' \in Q$, $a \in \Sigma$, $u \in \Sigma^*$,
-%
-\begin{align}
-\weight_A(q, \varepsilon, q) & = \final(q)\\
-\weight_A(q, \varepsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
-\weight_A(q, au, q') & =  \displaystyle\bigoplus_{q'' \in Q} 
-    \wei_{1}(q, a, q'') \otimes \weight_A(q'', u, q')\nonumber
-\label{eq:SWA-weight}
-\end{align}
-%
-\noindent
-and the weight value associated by $A$ to 
-$s \in \Sigma^*$ is defined as follows: %$s = s_1\ldots s_n \in \Sigma^+$ 
-\begin{equation}
-A(s)  = 
-\displaystyle\bigoplus_{q, q' \in Q} \mathsf{in}(q) 
-\mathop{\otimes} \weight_A(q, s, q') 
-\label{eq:SWA-value}
-\end{equation}
-
-%
-%When $\wei_\varepsilon(q, q') = \zero$ for all $q, q' \in Q$, 
-%the automaton~$A$ is called \emph{without $\varepsilon$-transitions}.
-
-%The \emph{summary} of a $\SWT$, resp. a $\SWA$, is ***
-
-%\subsection{Properties}
-\noindent
-The following property will be useful to the approach on 
-symbolic weighted parsing presented in Section~\ref{sec:parsing}.
-
 \begin{proposition} \label{prop:epsilon}
@@
 %
-The function $\wei'_0$ is defined thanks to the closure properties of the label theory $\bar\Phi$
+The function $\wei'_0$ is defined thanks to the closure properties of the label theory~$\bar\Phi$
 (Section~\ref{section:symbols}). 
@@
 where the size $\| T \|$ of $T$ is its number of states $|Q|$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SW-VPA
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{SW Visibly Pushdown Automata}
-\label{section:SWVPA}\label{sec:SWVPA}
-The model presented in this section generalizes Symbolic VPA~\cite{dAntonyAlur14SVPDA}
-from Boolean semirings to arbitrary semiring weight domains.
-It will compute on nested words over infinite alphabets, 
-associating to every such word a weight value. 
-Nest words are able to describe structures of labeled trees, 
-and in the context of parsing, they will be useful to 
-represent parse trees.
-
-\subsection{Definition}\label{se:SWVPA-def}
-Let $\Omega$ be a countable alphabet 
-%finite (large) or infinite,
-that we assume partitioned into three 
-subsets~$\Omegai$, $\Omegac$, $\Omegac$,
-whose elements are respectively called 
-\emph{internal}, \emph{call} and \emph{return} symbols.
-% names are  coined by application to functional program verification
-% \begin{itemize}
-% \item a set $\Omegai$ of \emph{internal symbols} denoted $a$,
-% \item a set $\Omegac$ of \emph{call symbols} denoted $\call{a}$,
-% \item a set $\Omegar$ of \emph{return symbols} denoted $\return{a}$.
-% \end{itemize}
-Let~$\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative and complete semiring and let  
-$\bar\Phi = \< \Phii, \Phic, \Phir, \Phici, \Phicc, \Phicr>$ 
-be a label theory over $\Semiring$
-%In order to simplify notations, %and following the definition of Section~\ref{section:transducer}, 
-%we shall write respectively 
-where $\Phii$, $\Phic$, $\Phir$ and~$\Phicx$ (with $\mathsf{x} \in \{ \mathsf{i}, \mathsf{c}, \mathsf{r}\}$) 
-stand respectively 
-for~$\Phi_\Omegai$, $\Phi_\Omegac$, $\Phi_\Omegar$ and~$\Phi_{\Omegac, \Omega_\mathsf{x}}$.
-%
-%Moreover, we extend this theory with a set $\Phii$ 
-%of unary functions in $\Omegai \to \Semiring$,
-%closed under $\oplus$ and $\otimes$.
-
-\begin{definition}
-A \emph{Symbolic Weighted Visibly Pushdown Automata} (\SWVPA) 
-over  $\Omega = \Omegai \uplus \Omegac \uplus \Omegar$, $\Semiring$ and $\bar\Phi$ 
-is a tuple $A = \< Q, P, \init, \bar\wei, \final >$,
-where $Q$ is a finite set of states, 
-$P$ is a finite set of stack symbols, 
-$\mathsf{in} : Q \to \Semiring$, 
-respectively $\mathsf{out} : Q \to \Semiring,$
-are functions defining the weight for entering, 
-respectively leaving, a state, 
-and $\bar\wei$ is a sextuplet composed of the transition functions:
-$\weii : Q \times P \times Q \to \Phici$,  
-$\weiei : Q \times Q \to \Phii$,  
-$\weic : Q \times P \times Q \times P \to \Phicc$,  
-$\weiec : Q \times P \times Q \to \Phic$,  
-$\weir : Q \times P \times Q \to \Phicr$,  
-$\weier : Q \times Q \to \Phir$.
-%and 
-%$\weiex : Q \times Q \to \Phix$ 
-%with $\mathsf{x} \in \{ \mathsf{i}, \mathsf{c}, \mathsf{r}\}$.
-\end{definition}
-%
-Similarly as in Section~\ref{section:transducer}, 
-we extend the above transition functions as follows
-for all $q, q' \in Q$, $p \in P$, 
-$a \in \Omegai$, 
-$\call{c} \in \Omegac$, 
-$\return{r} \in \Omegar$, 
-overloading their names: % for simplicity:
-\[
-\begin{array}{lll}
-\weii: Q \times \Omegac \times P \times \Omegai \times Q \to \Semiring & 
-\weii(q, c, p, a, q') = \eta_\mathsf{ci}(c, a) & 
-\mathrm{where~} \eta_\mathsf{ci} = \weii(q, p, q'),\\
-%
-\weiei: Q \times \Omegai \times Q \to \Semiring & 
-\weiei(q, a, q') = \phi_\mathsf{i}(a) &
-\mathrm{where~} \phi_\mathsf{i} = \weiei(q, q').\\[2pt]
-%
-\weic: Q \times \Omegac \times P \times  \Omegac \times P \times Q \to \Semiring & 
-\weic(q, c, p, \call{c'}, p', q') = \eta_\mathsf{cc}(c, \call{c'}) & 
-\mathrm{where~} \eta_\mathsf{cc} = \weic(q, p, p', q'),\\
-%
-\weiec: Q \times \Omegac \times P \times Q \to \Semiring & 
-\weiec(q, {\call{c}}, p, q') = \phi_\mathsf{c}({\call{c}}) &
-\mathrm{where~} \phi_\mathsf{c} = \weiec(q, p, q').\\[2pt]
-%
-\weir: Q \times \Omegac \times P \times \Omegar \times Q \to \Semiring & 
-\weir(q, {\call{c}},  p, {\return{r}}, q') = \eta_\mathsf{cr}({\call{c}},  {\return{r}}) & 
-\mathrm{where~} \eta_\mathsf{cr} = \weir(q, p, q'),\\
-%
-\weier: Q \times \Omegar \times Q \to \Semiring & 
-\weier(q, {\return{r}}, q') = \phi_\mathsf{r}({\return{r}}) &
-\mathrm{where~} \phi_\mathsf{r} = \weier(q, q').\\
-\end{array}      
-\]
-
-\noindent
-The intuition is the following for the above transitions.
-
-\noindent
-$\weii$ and $\weiei$ both read an input internal symbol $a$ and change state from $q$ to $q'$, 
-without changing the stack. 
-Moreover, $\weii$ reads a pair made of 
-${\call{c}} \in \Omegac$ and $p \in P$ at the top of the stack 
-($c$ is compared to $a$ by the weight function $\eta_\mathsf{ci} \in \Phici$)
-and $\weiei$ applies if and only if the stack is empty.
-
-\noindent
-$\weic$ and $\weiec$ read the input call symbol $\call{c}'$, 
-push it to the stack along with $p'$, and change state from $q$ to to $q'$.
-Moreover, $\weic$ reads ${\call{c}}$ and $p$ at the top of the stack 
-($c$ is compared to $c'$),
-and $\weiec$ applies iff the stack is empty.
-
-\noindent
-$\weir$ and $\weier$ read the input return symbol $\return{r}$, and change state from $q$ to to $q'$.
-Moreover, $\weir$ reads and pop from stack a pair made of $\call{c}$ and $p$, 
-($\call{c}$ is compared to $\return{r}$),
-and $\weier$ applies iff the stack is empty.
-%In this case, the weight function $\phi_\mathsf{r}$ 
-%computes a value of matching between the call and return symbols $c$ and $r$.
-%This value might be set to $\zero$ in order to express that the symbols do not match.
-
-Formally, the transitions of the automaton~$A$ are defined
-in term of %a weight value computed by 
-an intermediate function $\weight_A$, like in Section~\ref{sec:SWT}.
-In the case of a pushdown automaton, a configuration is composed 
-of a state $q \in Q$ and a stack content $\gamma \in \Gamma^*$, where $\Gamma = \Omegac \times P$.
-Hence, $\weight_A$ is a function from 
-$[Q \times \Gamma^*] \times \Omega^* \times [Q \times \Gamma^*]$ into~$\Semiring$.
-%
-\begin{align}
-%\begin{array}[t]{rcl}
-\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot\gamma}, a\, u, \config{q'}{\gamma'}\bigr) & =  
- {\displaystyle\bigoplus_{q'' \in Q}} \weii(q, c, p, a, q'') 
- \otimes \weight_A\bigl(\config{q''}{\gamma}, u, \config{q'}{\gamma'}\bigr)\nonumber\\
-\weight_A\bigl(\config{q}{\bot}, a\, u, \config{q'}{\gamma'}\bigr) & =  
- {\displaystyle\bigoplus_{q'' \in Q}} \weiei(q, a, q'') 
- \otimes \weight_A\bigl(\config{q''}{\bot}, u, \config{q'}{\gamma'}\bigr)\nonumber\\
-%
-\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot\gamma}, {\call{c}'}\, u, \config{q'}{\gamma'}\bigr) & =  
- {\displaystyle\bigoplus_{\begin{array}{c}
-                          \scriptstyle q'' \in Q\\[-2pt]
-                          \scriptstyle p \in P
-                          \end{array}}}
- \weic\bigl(q, {\call{c}}, p, {\call{c}'}, p', q''\bigr) 
- \otimes \weight_A\bigl(\config{q''}{{\call{c}'}\, p' \cdot {\call{c}}\, p\cdot \gamma}, u, \config{q'}{\gamma'}\bigr)\nonumber\\[1mm]
-%
-\weight_A\bigl(\config{q}{\bot}, {\call{c}}\, u, \config{q'}{\gamma'}\bigr) & =  
- {\displaystyle\bigoplus_{q'' \in Q}} \weiec(q, {\call{c}}, p, q'') 
- \otimes \weight_A\bigl(\config{q''}{{\call{c}}\, p}, u, \config{q'}{\gamma'}\bigr)\nonumber\\
-%
-\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot \gamma}, {\return{r}}\, u, \config{q'}{\gamma'}\bigr) & =  
- {\displaystyle\bigoplus_{q'' \in Q}} 
- \weir\bigl(q, {\call{c}}, p, {\return{r}}, q''\bigr) 
- \otimes \weight_A\bigl(\config{q''}{\gamma}, u, \config{q'}{\gamma'}\bigr)\nonumber\\
-%
-\weight_A\bigl(\config{q}{\bot}, {\return{r}}\, u, \config{q'}{\gamma'}\bigr) & =  
- {\displaystyle\bigoplus_{q'' \in Q}} \weier(q, {\return{r}}, q'') 
- \otimes \weight_A\bigl(\config{q''}{\bot}, u, \config{q'}{\gamma'}\bigr)
-%\end{array}
-\label{eq:SWVPA-weight}
-\end{align}
-%
-where $\bot$ denotes the empty stack and ${\call{c}}\, p\cdot \gamma$ 
-denotes a stack where the pair made of ${\call{c}} \in \Omegac$ and $p \in P$ is the top symbol 
-and $\gamma$ is the rest of stack.
-
-\noindent
-The weight associated by $A$ to $s \in \Omega^*$
-is defined according to empty stack semantics: 
-%
-\begin{equation}
-A(s)  = 
-{\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
-\mathsf{in}(q) \mathop{\otimes} 
-\weight_A\bigl(\config{q}{\bot}, s, \config{q'}{\bot}\bigr) 
-\mathop{\otimes} \mathsf{out}(q').
-\label{eq:weightA}
-\end{equation}
-
-\begin{example}
-structured words with timed symbols...
-intro language of music notation? (markup = time division, leaves = events etc)
-\end{example}
-
-\subsection{Properties}
-Like VPA and symbolic VPA, 
-the class of \SWVPA is closed under the binary operators of the underlying semiring.
-%
-\begin{proposition}\label{prop:SWVPA-product}
-Let $A_1$ and $A_2$ be two (\SWVPA)
-over the same $\Omega$, $\Semiring$ and $\bar\Phi$.
-There exists two $\SWVPA$ $A_1 \oplus A_2$ and $A_1 \otimes A_2$, 
-effectively constructible, 
-such that for all $s \in \Omega^*$, 
-$(A_1 \oplus A_2)(s) = A_1(s) \oplus A_2(s)$ and 
-$(A_1 \otimes A_2)(s) = A_1(s) \otimes A_2(s)$.
-\end{proposition}
-The construction is essentially the same 
-as in the case of the Boolean semiring~\cite{dAntonyAlur14SVPDA}.
-
-\subsection{Best Search} 
-\label{sec:best}\label{sec:search}
-%**hypotheses**
-Let us assume that the semiring~$\Semiring$ is
-commutative, bounded, total, and complete.
-and assume an effective label theory.
-%s.t. for all $\phi \in \Phii$, 
-%$\psi \in \Phir$, 
-%and $\eta \in \Phicr$, 
-%$\displaystyle\bigoplus_{a \in \Omegai} \phi(a)$ 
-%$\displaystyle\bigoplus_{r \in \Omegar} \phi(r)$ and 
-%$\displaystyle\bigoplus_{{\call{c}} \in \Omegac} 
-%\displaystyle\bigoplus_{{\return{r}} \in \Omegar} \eta({\call{c}}, {\return{r}})$
-%are computable...
-
-% total ? 
-% monotonic and superior writ natural ordering
-%Regarding the infinite sum operator, note that
-%$\bigoplus_{x \in \Phi_\Omega} \phi(x)$, 
-%$\bigoplus_{y \in \Phi_\Delta} \psi(y)$, and  
-%... exist and in $\Semiring$.
-
-We propose a Dijkstra algorithm computing, for a $\SWVPA$ $A$
-over~$\Omega$, $\Semiring$ and~$\bar\Phi$, 
-the minimal weight \wrt~$\leq_\oplus$, for a word in~$\Omega^*$.
-
-\noindent
-More precisely, 
-let $b_\bot : Q \times Q \to \Semiring$ be the function:
-%
-\begin{equation}\label{eq:bbot}
-  b_\bot(q, q') = \bigoplus_{s\in \Omega^*} 
-  \textstyle
-  \weight_A\bigl(\config{q}{\bot}, s, \config{q'}{\bot}\bigr)
-\end{equation}
-%
-Since $\Semiring$ is complete, the infinite sum in \eqref{eq:bbot} is well defined,
-and by totality of $\leq_\oplus$, it is the minimum in $\Omega^*$
-of the fonction 
-$s \mapsto \weight_A(\config{q}{\sigma}, s, \config{q'}{\sigma})$ 
-\wrt this ordering.
-%
-The term $\config{q}{\bot}, s, \config{q'}{\bot}$ 
-of this sum is the central expression in 
-the definition \eqref{eq:weightA} of $A(s_0)$, for the minimum $s_0$
-of the function $\weight_A$.
-
-%
-\noindent
-Let $\top$ be a fresh stack symbol which does not belong to $\Gamma$,
-and let $b_\top : Q \times P \times Q \to \Phic$ be such that,
-for every two states $q, q' \in Q$ 
-and stack symbol $p \in P$: %and $\sigma \in \{ \bot, \top \}$, 
-\begin{equation}\label{eq:btop}
-  b_\top(q, p, q') : c \mapsto \bigoplus_{s\in \Omega^*} 
-  \textstyle
-  \weight_A\bigl(\config{q}{c\,p\cdot \top }, s, \config{q'}{c\, p\cdot \top}\bigr). 
-\end{equation}
-%
-Intuitively, the function defined in \eqref{eq:btop}
-associateds to $c \in \Omegac$ 
-the minimum weight of a computation of $A$
-starting in state $q$ with a stack 
-$c p \cdot \gamma \in \Gamma^+$ 
-and ending in state $q'$ with the same stack,
-such that the computation does pop 
-the pair made of $c$ and $p$ at the top of this stack,
-but may read these symbols.
-Moreover, $A$ may push another pair $\< c', p'>$ %call symbols 
-on the top of $c p \cdot \gamma$,
-following the the third case of 
-in the definition \eqref{eq:SWVPA-weight} of $\weight_A$,
-and may pop $\< c', p'>$ later, following the fifth case of \eqref{eq:SWVPA-weight} (return symbol). 
-%However, it cannot apply one of the two last cases (return symbol and empty stack)
-%when the current stack is $\gamma$.
-%pop symbols in $\gamma$.
-% Note that having a stack reduced to such a symbol makes impossible the application of the 
-% two last cases in the definition of $\weight_A$ (return symbol and empty stack). 
-% However, it is possible to apply the two first cases 
-% (internal symbol or call symbol, with a push on the top of $\top$).
-
-Algorithm~\ref{algo:Dijkstra}
-constructs iteratively markings 
-$d_\bot : Q \times Q \to \Semiring$ and 
-$d_\top : Q \times P \times Q \to \Phic$
-%of the triplets $\<q, \sigma, q'>$ 
-%of states of $A$ by weight values in $\Semiring$, 
-that converges eventually to $b_\top$ and $b_\bot$.
-%It uses for that purpose a priority queue $P$ containing triplets of 
-%$Q \times \{ \bot, \top \} \times Q$.
-
-\begin{algo}[best search for \SWVPA]\label{algo:Dijkstra}
-\textbf{initially} 
-let $\Q = (Q \times Q) \cup (Q \times P \times Q)$, %contains all 
-and let 
-$d_\bot(q_1, q_2) = d_\top(q_1, p, q_2) = \one$ if $q_1 = q_2$ and
-$d_\bot(q_1, q_2) = d_\top(q_1, p, q_2) = \zero$ otherwise.
-
-\smallskip\noindent
-\textbf{while} $\Q$ is not empty
-
-\noindent\quad
-\textbf{extract} $\< q_1, q_2>$ or $\< q_1, p, q_2>$ from $\Q$
-
-\noindent\quad
-such that $d_\bot(q_1, q_2)$, resp. $\bigoplus_{c\in \Omegac} d_\top(q_1, p, q_2)(c)$,
-is minimal in $\Semiring$ wrt $\leq_\oplus$.
-
-\noindent\quad 
-%\textbf{for all} $q_0, q_3 \in Q$ 
-\textbf{update} $d_\bot$ with $\< q_1, q_2>$ or $d_\top$ with $\< q_1, p, q_2>$
-(Figure~\ref{fig:best-update}).
-\end{algo}
-
-\begin{figure}
-For all $q_0, q_3 \in Q$, %$p \in P$,
-\[
-\begin{array}{lcl}
-%\multicolumn{3}{l}{\mathrm{For~all~} q_0, q_3 \in Q, p \in P,}\\
-d_\top(q_1, p, q_3) & \opluseq &
-  d_\top(q_1, p, q_2) \otimes 
-  \displaystyle\bigoplus_{a \in \Omegai} \weii(q_2, p, q_3)_a\\
-\multicolumn{3}{l}{\quad
-\mathrm{where~} \weii(q_2, p, q_3)_a \in \Phic
-\mathrm{~is~the~partial~application~}
-x_c \mapsto \weii(q_2, x_c, p, a, q_3)}\\
-%
-d_\bot(q_1, p, q_3) & \opluseq &
-  d_\bot(q_1, q_2) \otimes 
-  \displaystyle\bigoplus_{a \in \Omegai} \weie(q_2, a, q_3)\\
-%     
-d_\top(q_0, p, q_3) & \opluseq &
-  \displaystyle\bigoplus_{{\call{c}'} \in \Omegac} 
-  \displaystyle\bigoplus_{{\return{r}} \in \Omegar} 
-  \weic(q_0, p, p', q_1)_{c'} \otimes 
-  d_\top(q_1, p', q_2) \otimes
-  \weir(q_2, c', p', r, q_3)\\
-\multicolumn{3}{l}{\quad
-\mathrm{where~} \weic(q_0, p, p', q_1)_{c'} \in \Phic
-\mathrm{~is~the~partial~application~}
-x_c \mapsto \weic(q_0, p, x_c, c', p', q_1)}\\[2pt]
-%
-d_\bot(q_0, q_3) & \opluseq &
-  \displaystyle\bigoplus_{{\call{c}} \in \Omegac} 
-  \displaystyle\bigoplus_{{\return{r}} \in \Omegar} 
-  \weiec(q_0, c, p, q_1) \otimes 
-   d_\top(q_1, p, q_2)(c) \otimes
-  \weir(q_2, c, p, r, q_3)\\
-%
-d_\bot(q_1, q_3) & \opluseq &
-  d_\bot(q_1, q_2) \otimes 
-  \displaystyle\bigoplus_{\return{r} \in \Omegar} \weier(q_2, {\return{r}}, q_3)\\
-%
-d_\top(q_1, p, q_3) & \opluseq & 
-  d_\top(q_1, p, q_2) \otimes d_\top(q_2, p, q_3), 
-  \mathrm{if~} \< q_2, \top, q_3> \notin P \\
-%
-d_\bot(q_1, q_3) & \opluseq & 
-  d_\bot(q_1, q_2) \otimes d_\bot(q_2, q_3), \mathrm{if~} \< q_2, \bot, q_3> \notin P \\
-\end{array}
-\]
-%$\weii(q_2, p, q_3)_a \in \Phic$, for $a \in \Omegai$,
-%is the partial application
-%$x_c \mapsto \weii(q_2, x_c, p, a, q_3)$.
-%
-%$\weic(q_0, p, p', q_1)_{c'} \in \Phic$, for $c' \in \Omegac$,
-%is the partial application
-%$x_c \mapsto \weic(q_0, p, x_c, c', p', q_1)$.
-\caption{Update $d_\bot$ with $\<q_1, q_2>$ or $d_\top$ with $\< q_1, p, q_2>$.} 
-\label{fig:best-update}
-\end{figure}
-
-\noindent
-The infinite sums in the updates of $d$ in Algorithm~\ref{algo:Dijkstra}, 
-Figure~\ref{fig:best-update}
-are well defined
-since~$\Semiring$ is complete.
-** effectively computable by hypothese that the label theory is effective**
-The algorithm performs $2.|Q|^2$ iterations until $P$ is empty, 
-and each iteration has a time complexity $O(|Q|^2 . |P|)$.
-This gives a time complexity $O(|Q|^4 . |P|)$. 
-It can be reduced by implementing $P$ as a priority queue, 
-prioritized by the value returned by $d$
-***complete***. %$|Q|^3.\log(|Q|^2)$
-
-The correctness of Algorithm~\ref{algo:Dijkstra} 
-is ensured by the invariant expressed in the following lemma.
-\begin{lemma}\label{lem:bot}
-For all $\< q_1, q_2> \notin \Q$,
-$d_\bot(q_1, q_2) =  b_\bot(q_1, q_2)$/
-\end{lemma}
-The proof is by contradiction, 
-assuming a counter-example minimal in the length of the witness word.
-
-\begin{lemma}\label{lem:top}
-For all $\< q_1, p, q_2> \notin \Q$, 
-$d_\top(q_1, p, q_2) =  b_\top(q_1, p, q_2)$,
-\end{lemma}
-
-\noindent
-For computing the minimal weight of a computation of $A$, we use the fact that,
-at the termination of Algorithm~\ref{algo:Dijkstra}, %it holds that,
-%There exist $q_1, q_2 \in Q$
-\[
-  {\displaystyle \bigoplus_{s \in \Omega^*} A(s)} = 
-  {\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
-  \mathsf{in}(q) \mathop{\otimes} d_\bot(q, q') \mathop{\otimes} \mathsf{out}(q').
-\]
-
-%\medskip
-\noindent
-In order to obtain effectively a witness 
-(word of $\Semiring^*$ with computation of $A$ of minimal weight), 
-we require an additional property the of weight functions.
-%
-\begin{definition}
-Let $\Omega$ be an alphabet and $\Semiring$ a complete semiring.
-A function $\phi$ from $\Omega^n$ into $\Semiring$
-is called $k$-\emph{convex} for a natural number $k$ iff 
-$\mathit{card}\{ \mathbf{a} \in \Omega^n \mid \phi(\mathbf{a}) = 
-                 \displaystyle\bigoplus_{\mathbf{p} \in \Omega^n} \phi(\mathbf{p}) \} \leq k$.
-\end{definition}
-A label theory is $k$-convex if all its functions are $k$-convex.
-
-\begin{proposition}\label{th:best-search}
-For a \SWVPA $A$ 
-over $\Omega$, 
-$\Semiring$ commutative, bounded, total and complete, %semiring 
-and $\bar\Phi$ $k$-convex effective, % label theory,
-one can construct in PTIME a word $s \in \Omega^*$ 
-such that $A(s)$ is minimal \wrt the natural ordering for $\Semiring$. 
-\end{proposition}
-
-\subsection{Nested-Words and Parse-Trees}
-\label{sec:trees}
-The hierarchical structure of nested-words, defined with the \emph{call} and \emph{return} markup symbols  
-suggest a correspondence with trees. 
-The lifting of this correspondence to languages, of tree automata and VPA,
-has been discussed in~\cite{AlurMadhusudan09nested}, 
-and~\cite{Caralp12VPAmult} for the weighted case.
-In this section, we describe a correspondence between the symbolic-weighted extensions
-of tree automata and VPA.
-
-Let $\Omega$ be a countable ranked alphabet, such that 
-every symbol $a \in \Omega$ has a rank 
-$\rank(a) \in [0..M]$ where $M$ is a fixed natural number.
-We denote by $\Omega_k$ the subset of all symbols $a$ of $\Omega$
-with $\rank(a) = k$, where $0 \leq k \leq M$, 
-and $\Omega_{>0} = \Omega \setminus \Omega_0$.
-%
-\noindent 
-The free $\Omega$-algebra of finite, ordered, 
-$\Omega$-labeled trees is denoted by $\T(\Omega)$.
-It is the smallest set such that  $\Omega_0 \subset \T(\Omega)$
-and for all $1 \leq k \leq M$, all $a \in \Omega_k$, 
-and all $t_1, \ldots, t_k \in \T(\Omega)$, $a(t_1, \ldots, t_k) \in \T(\Omega)$.
-%
-% tree = single node (leave) labeled with a symbol of $a \in \Omegai$
-% (such a tree is simply denoted by $a$)
-% or the composition, denoted by $b(t_1,\ldots, t_n$) of a node labeled with $b$
-% and $n$ subtrees $t_1$,\ldots, $t_n$.
-%
-Let us assume a commutative semiring~$\Semiring$ 
-and a label theory~$\bar{\Phi}$ over~$\Semiring$ 
-containing one set~$\Phi_{\Omega_k}$ for each $k \in [0..M]$.
-
-\renewcommand{\call}[1]{\ensuremath \langle_{#1}}
-\renewcommand{\return}[1]{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
-
-Let $\hat\Omega$ be the countable (unranked) alphabet obtained
-from $\Omega$ by: 
-$\hat\Omega = \Omegai \uplus \Omegac \uplus \Omegar$, with
-$\Omegai = \Omega_0$, 
-$\Omegac = \{ \; \call{a} \mid a \in \Omega_{>0} \}$,
-$\Omegar = \{ \; \return{a} \mid a \in \Omega_{>0} \}$.
-We associate to $\hat\Omega$
-a label theory $\hat{\Phi}$ 
-like in Section~\ref{se:SWVPA-def}.
-
-\noindent
-We define a linearization of trees of $\T(\Omega)$ into 
-words of $\hat{\Omega}^*$ as follows:
-\begin{description}
-\item $\lin(a) = a$ for all $a \in \Omega_0$, 
-\item $\lin\bigl( b(t_1, \ldots, t_k)\bigr) = 
-       \call{b} \; \lin(t_1) \ldots \lin(t_k) \; \return{b}$ 
-       when $b \in \Omega_k$, $1 \leq k \leq M$.
-\end{description}
-
-% A \emph{regular tree grammar} over $\Omega$ 
-% is a triplet $G = \< N, q_\mathsf{i}, R>$ where
-% $N$ is a finite set of non-terminal symbols denoted $q$..., 
-% $q_\mathsf{i} \in N$ is the starting non-terminal, 
-% $R$ is a finite set of production rules of the form
-% $q_0 \to a(q_1\ldots q_k)$ where 
-% $q_0, q_1, \ldots, q_k \in N$
-% $a \in \Omega_k$.
-% A tree $t \in \T(\Omega)$ is in the language of $G$ 
-% if it can be generated from $q_\mathsf{i}$ by 
-% non terminal replacement following the rules of $R$.
-%
-\begin{definition}  \label{def:SWTA}
-A \emph{symbolic-weighted tree automaton} (\SWTA)
-over $\Omega$, $\Semiring$, and $\bar\Phi$
-is a triplet $A = \< Q, \init, \bar{\wei} >$ where
-$Q$ is a finite set of states, 
-$\mathsf{in} : Q \to \Semiring$ is the starting weight function, 
-and $\bar{\wei}$ is a tuplet of transition functions containing
-%$\wei_\varepsilon$ from $Q \times Q$ into $\Semiring$, and, 
-$\wei_{k}: Q \times Q^{k} \to \Phi_{\Omega_k}$
-for each $k \in [0..M]$. 
-\end{definition}
-%
-%Like in Section~\ref{sec:SWAdef}, 
-We define from $\bar{\wei}$ 
-a transition function~$\wei$, %from $\bar\wei$
-from 
-$Q \times \Omega \times \bigcup_{k=0}^{M} Q^k$
-into~$\Semiring$ by: %also called $\wei$ for simplicity, 
-%such that, for all $q, q' \in Q$, $a \in \Sigma$, and $b \in \Delta$, 
-\[
-\begin{array}{rcll}
-%\wei(q_0, \varepsilon, q_1) & = &  \wei_\varepsilon(q_0, q_1),\\ %\phi_\varepsilon\\
-\wei(q_0, a, q_1 \ldots q_k) & = & \phi_{\Omega, k}(a) &
-\quad\mathrm{where~} \phi_{k} = \wei_{k}(q_0, q_1\ldots q_k).\\
-\end{array}      
-\]
-Intuitively, $\wei(q_0, a, q_1 \ldots q_k)$ can be seen as
-the weight of a production rule $q_0 \to a(q_1, \ldots, q_k)$ 
-of a regular tree grammar~\cite{tata}, 
-that replaces the non-terminal symbol $q_0$ by $a(q_1, \ldots, q_k)$. 
-%in a step of tree building.
-%
-Such a grammar computes the weights of the derivation trees 
-of the Context-Free grammar obtained by forgetting the labeling symbols of $\Omega_{>0}$.
-
-\noindent
-The \SWTA of Definition~\ref{def:SWTA} defines a mapping 
-from trees of $\T(\Omega)$ into the weights of~$\Semiring$,
-based on the intermediate function~$\weight_A$
-defined as follows for $q_0 \in Q$ and 
-$t = b(t_1,\ldots, t_k) \in \T(\Omega)$,
-with $0 \leq k \leq M$:
-\begin{equation}
-\begin{array}{rccl}
-\weight_A(q_0, t) & = & 
-%   \displaystyle\bigoplus_{q_1 \in Q} &
-%   \wei(q, \varepsilon, q_1) \otimes \weight_A(q_1, t)\\
- \displaystyle\bigoplus_{q_1 \ldots q_k \in Q^k} &
-              \wei(q_0, b, q_1 \ldots q_k ) 
-   \otimes \displaystyle\bigotimes_{i=1}^{k}
-           \weight_A(q_i, t_i)
-\end{array}
-\end{equation}
-
-\medskip\noindent
-The weight associated by $A$ to  $t \in \T(\Omega)$ is 
-\begin{equation}
-A(t)  = 
-\displaystyle\bigoplus_{q \in Q} \mathsf{in}(q) \mathop{\otimes} \weight_A(q, t)
-\label{eq:weightTA}
-\end{equation}
-
-\begin{lemma}\label{lem:SWTA}
-For all \SWTA $A$ over~$\Omega$, $\Semiring$ commutative, and $\Phi$,
-there exists an effectively constructible \SWVPA $A'$ over 
-$\hat\Omega$, $\Semiring$ and $\hat\Phi$ 
-such that for all $t \in \T(\Omega)$, $A'\bigl(\lin(t)\bigr) = A(t)$.
-\end{lemma} 
-% 
-\begin{proof}
-Let $A = \< Q, \init, \bar{\wei} >$ where $\bar{\wei}$ is presented as above by a function
-    %
-We build 
-$A' = \< Q', P', \init', \bar{\wei}', \final' >$,
-%computing over $\hat\Omega = \Omegai \uplus \Omegac \uplus \Omegar$,
-where $Q' = \bigcup_{k=0}^{M} Q^k$ is the set of sequences of state symbols of $A$, 
-of length at most $M$, including the empty sequence denoted by $\varepsilon$, 
-and where $P' = Q'$ and $\bar\wei$ is defined by:
-
-\[
-\begin{array}{ll}
-\weii(\bar{q}, {\call{b}}, \bar{p}, a, \bar{q} q') = \wei(q', a, \varepsilon) & 
-\mathrm{for~all~} b \in \Omega_{>0}, \bar{p}\in P', a \in \Omega_0\\
-%
-\weiei(\bar{q}, a, \bar{q} q') = \wei(q', a, \varepsilon) & 
-\mathrm{for~all~} a \in \Omega_0\\
-%
-\weic(\bar{q}, {\call{b}}, \bar{p}, \call{b'}, \varepsilon, \bar{q}) = \one & %\wei(p', b, \bar{p}) & 
-\mathrm{for~all~} b \in \Omega_{>0},  \bar{p}\in P', b, b' \in \Omega_{>0}\\
-%
-\weiec(\bar{q}, \call{b}, \varepsilon, \bar{q}) = \one & %\wei(p', b, \bar{p}) & 
-\mathrm{for~all~} b \in \Omega_{>0}\\
-%
-%\weir: Q \times \Omegac \times P \times \Sigmar \times Q \to \Semiring & 
-\weir(\bar{q}, {\call{b}}, \bar{p}, {\return{b}}, \bar{p} q') = \wei(q', b, \bar{q}) & 
-\mathrm{for~all~}  b \in \Omega_{>0}, \bar{p}\in P'\\
-%
-%\weie: Q \times \Omegar \times Q \to \Semiring & 
-\weier(\bar{p}, {\return{b}}, \bar{q}) = \zero &
-\mathrm{for~all~}  b \in \Omega_{>0}\\
-\end{array}      
-\]
-\noindent
-It is sufficient to consider in $Q'$ only the prefixes of 
-sequences in transition with a non-null weight.
-\qed\end{proof}
-
-\section{Symbolic Weighted Parsing}
-\label{sec:parsing}
-Let us now apply the models and results of the previous sections %in order to define 
-to the problem of parsing over infinite alphabet. %appropriate
-%
-Besides considering infinitely many possible of input symbols, 
-handled with suitable language formalisms, 
-this approach extends conventional parsing and weighted parsing
-by computing a derivation tree modulo 
-a generic distance between words,  
-defined by a SW transducer given in input.
-This enables considering finer word relationships than strict equality 
-as in the conventional parsing approach, 
-opening possibilities of quantitative analysis via this method.
-
-\subsection{Definition}
-
-Let $\Sigma$ be a countable input alphabet 
-and let $\Omega$ be a countable output ranked alphabet, 
-with maximal rank value $M$, 
-and $\hat\Omega = \Omegai \uplus \Omegac \uplus \Omegar$
-be the alphabet with nesting symbols, associated like Section~\ref{sec:trees}, 
-for the linearization of trees of $\T(\Omega)$
-(remember that $\Omegai = \Omega_0$).
-%
-Let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative, bounded, complete 
-and total semiring 
-and let $\bar\Phi$ be a label theory over $\Semiring$
-containing
-$\Phi_\Sigma$, $\Phi_{\Sigma, \Omegai}$, as well as
-$\Phii$, $\Phic$, $\Phir$, $\Phicr$
-(following the notations of Section~\ref{se:SWVPA-def}).
-It is moreover assumed computable and $k$-convex for some fixed $k$.
-
-\noindent
-Let us assume given the following input:
-\begin{itemize}
-\item a \SWT $T$ over $\Sigma$, $\Omegai$, $\Semiring$, and $\bar\Phi$, 
-defining a measure %between words 
-$T: \Sigma^* \times \Omegai^* \to \Semiring$,
-
-\item a \SWTA $A$ over $\Omega$, $\Semiring$, and $\bar\Phi$,
-defining a measure $A: \T(\Omega) \to \Semiring$,
-%a \SWVPA $A$ over $\Omega$, and $\Semiring$, defining a series of nested words
-%      $A : \Omega^* \to \Semiring$,
-\item an input word $s \in \Sigma^*$.
-\end{itemize}
-%
-As explained in Section~\ref{sec:trees}, 
-$A = \< Q, \init, \bar{\wei} >$ can be seen as a weighted regular tree grammar, 
-that generates (weighted) trees by replacement of a state symbol $q_0$ (non-terminal), 
-by a tree $a(q_1,\ldots, q_k)$, where $k = \rank(a)$.
-A replacement rule $q_0 \to a(q_1,\ldots, q_k)$, 
-of weight $\wei(q_0, a, q_1 \ldots q_k) \in \Semiring$ 
-according to Definition~\ref{def:SWTA}, 
-corresponds to the production rule $q_0 := a(q_1,\ldots, q_k)$ of a weighted CF grammar,
-with set non-terminal symbols $Q$ and set of terminal symbols $\Omega_0$.
-This actually is a slight generalization of CFG since 
-each such production rule is labelled by  a symbol of $\Omega_{>0}$, 
-hence parse trees %derivation trees
-are trees of $\T(\Omega)$.
-%
-Another (more original) generalization is that the set of terminal symbols 
-$\Omega_0$ may be infinite.
-
-\noindent 
-We extend the measure defined by $T$ to 
-$d : \Sigma^* \times \T(\Omega) \to \Semiring$ as follows.
-Given a word $w \in {\hat\Omega}^*$, the projection of $w$ onto $\Omegai$,
-denoted $w|_\Omegai$,
-is the word of $\Omegai^*$ obtained from $w$ by removing all symbols 
-in $\hat\Omega \setminus \Omegai$.
-Using this notation
-and the tree linearization operator defined in Section~\ref{sec:trees}, 
-$d$ is defined by:
-\begin{equation}
-d(s, t) = T\bigl(s, \lin(t)|_\Omegai \bigr)  \mathrm{~for~} s \in \Sigma^*, t\in \T(\Omega)
-\end{equation}
-
-\noindent 
-\emph{Symbolic weighted parsing} is the problem, 
-given the above input, 
-to find a tree $t \in \T(\Omega)$ %nested word $t \in \Omega$ 
-minimizing \( d(s, t) \otimes A(t)\)
-\wrt $\leq_\oplus$, 
-\ie such that: %Hence, it is the problem of finding 
-%
-\begin{equation}\label{eq:distance-lang}
-d(s, t) \otimes A(t) = \displaystyle\bigoplus_{t' \in \T(\Omega)} d(s, t') \otimes A(t') 
-\end{equation}
-%
-The measure expressed in~\eqref{eq:distance-lang} is called
-called edit-distance between $s$ and $A$ in~\cite{Mohri03EDWA}.
-%
-%The input language can also be expressed as a \SWTA, or, 
-%as a particular case, as a weighted context-free grammar, 
-%converted in turn into a \SWVPA following Lemma~\ref{lem:SWTA}.
-%
-The problem of searching, in a WTA language, 
-the best parse tree matching a given input,  
-sometimes referred as~\emph{weighted parsing}
-corresponds to SW parsing in the case of finite alphabets
-and when the transducer $T$ characterizes identity
-see \eg~\cite{Goodman99SemiringParsing} 
-and~\cite{MorbitzVogler19weighted-parsing} for a more general 
-weighted parsing framework.
-%
-%Indeed, it corresponds to the case where $T$
-%accepts only the pairs $\<s, t>$ such that 
-%$s$ is the projection of $t$ on $\Omegai$. 
-%This can be done with a single state $q$ and 
-%with transition rules of the form:
-%\begin{description}
-%\item[] $\wei(q, \varepsilon, a, q) = \one$ for all $a \in \Omegac \cup \Omegar$,
-%\item[] $\wei(q, a, a, q) = \one$ for all $a \in \Omegai$,
-%\item[] $\wei(q, a, b, q) = \zero$ for all $a, b \in \Omegai$, $a \neq b$.
-%\end{description}
-
-\subsection{Computation}
-
-\begin{proposition}
-The problem of Symbolic Weighted  parsing 
-can be solved in PTIME in the size of the input \SWT $T$, \SWTA $A$ %\SWVPA (or ) 
-and input word $s$, 
-and the computation time of the functions of the label theory.
-\end{proposition}
-\begin{proof} (sketch)
-We follow a \emph{Bar-Hillel} construction, also called parsing by intersection.
-
-\noindent
-We first extend the \SWT $T$ over $\Sigma$, $\Omegai$, $\Semiring$, and $\bar\Phi$, 
-into a \SWT $T'$ over $\Sigma$ and $\hat\Omega$ (and the same semiring and label theory),
-such that for all $s \in \Sigma^*$, and $u \in {\hat\Omega}^*$, 
-$T'(s, u) = T(s, u|_{\Omegai})$. The transducer $T'$ simplify skips every symbol 
-$b \in {\hat\Omega} \setminus \Omegai$ in input, 
-with new transitions of the form $\wei_{01}(q, \varepsilon, b, q')$.
-
-\noindent
-Then, given an input word $s \in \Sigma^*$, 
-we compute the \SWA $A_{T', s}$, using Proposition~\ref{prop:epsilon}.
-This automaton is such that for all $t \in \T(\Omega)$, 
-$A_{T', s}\bigl(\lin(t)\bigr) 
- = T'\bigl(s, \lin(t)\bigr) 
- = T'\bigl(s, \lin(t)|_{\Omegai}\bigr) 
- = d(s, t)$.
-
-\noindent
-Next, we convert the input \SWTA $A$ over $\Omega$
-into a \SWVPA $A'$ over $\hat\Omega$, using Lemma~\ref{lem:SWTA}, 
-and we compute the \SWVPA $A_{T', s} \otimes A'$, 
-using Proposition~\ref{prop:SWVPA-product}.
-
-\noindent
-It remains to compute a best nested-word $w \in {\hat\Omega}^*$ 
-using the best-search procedure of Proposition~\ref{th:best-search},
-and convert it into a best tree in $\T(\Omega)$ in order to solve SW parsing
-for~$T$, $A$ and~$s$.
-\qed  
-\end{proof}
-
-\subsection{Application to Automated Music Transcription}
-%\input{transcription}
-
-\section*{Conclusion}
-% discussion
-
-\noindent
--- summary
-
-\noindent
--- other theoretical properties of SW models
-
-\noindent
--- room to improve complexity for best-search algorithm
-... modular approach with oracles ...
-
-and extention to $n$-best
-
-\noindent
--- offline algorithm, 
-semi-online implementation for AMT (bar-by-bar approach)
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% BIBLIO                                                                     %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\bibliographystyle{plain}
-%\bibliographystyle{plainurl} 
-\bibliographystyle{abbrv}
-%\bibliographystyle{splncs04}
-\bibliography{references}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% APPENDIX                                                                   %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\appendix
-
-\end{document}
-
-\section{Edit-Distance}
-
-%\subsection{Distance between words or languages}
-...algebraic definition of edit-distance of Mohri, in \cite{Mohri03EDWA}
-% Mehryar Mohri 
-% Edit-distance of weighted automata: General definitions and algorithms
-% International Journal of Foundations of Computer Science 14.06 (2003): 957-982.
-distance $d$ over $\Sigma^* \times \Sigma^*$ 
-into a semiring  $\Semiring = ( \Semiring, \oplus, \zero, \otimes, \one)$.
-
-%\noindent
-Let $\Omega = \Sigma \cup \{ \varepsilon \} \times \Sigma \cup \{ \varepsilon \} \setminus \{ (\varepsilon, \varepsilon) \}$,
-and let $h$ be the morphism from $\Omega^*$ into $\Sigma^* \times \Sigma^*$  
-defined over the concatenation of strings of $\Sigma^*$ (that removes the $\varepsilon$'s).
-%
-\noindent
-An \emph{alignment} between 2 strings  $s, t \in \Sigma^*$ is an element $\omega \in \Omega^*$ 
-such that $h(\omega) = (s, t)$.
-%
-\noindent
-We assume a base cost function $\Omega$ : $\delta: \Omega \to S$, extended to $\Omega^*$ as follows  
-(for $\omega \in \Omega^*$): 
-\(
-\displaystyle\delta(\omega) = \bigotimes_{0 \leq i < |\omega|} \delta(\omega_i)
-\).
-
-\noindent
-\begin{definition}
-For  $s, t \in \Sigma^*$, the edit-distance between $s$ and $t$ is  
-\( 
-d(s, t) = \displaystyle\bigoplus_{\omega \in \Omega^*\, h(\omega) = (s, t)} \delta(\omega)
-\).
-\end{definition}
-
-e.g. Levenstein edit-distance: $S$ is min-plus and $\delta(a, b) = 1$ for all $(a, b) \in \Omega$.
-
-%\paragraph{Distance between a word and a regular language}

### 2021-06-03 main-eptcs.tex
versions

% !TEX root = main.tex

% http://www.eptcs.org
\documentclass[submission,copyright,creativecommons]{eptcs}
\providecommand{\event}{GandALF 2021} % Name of the event you are submitting to
\usepackage{breakurl}                 % Not needed if you use pdflatex only.
\usepackage{underscore}               % Only needed if you use pdflatex.

%% theorem environments
\usepackage{theorem}
\newtheorem{theorem}{Theorem} %[section]
\newtheorem{definition}[theorem]{Definition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newenvironment{proof}{\vspace{-2ex}{\it Proof. }}{\hspace*{\fill} $\Box$\smallskip }
\theorembodyfont{\slshape}
\newtheorem{example}[theorem]{Example}
\newtheorem{remark}[theorem]{Remark}
\def\endex{\hspace*{\fill} $\Diamond$\smallskip }
\def\qed{}
{\theorembodyfont{\rmfamily} \theoremstyle{break} \newtheorem{algo}{Algorithm}}

\author{Florent Jacquemard
\institute{INRIA \& CNAM, Paris, France\\
\email{florent.jacquemard@inria.fr}}
} % end author

\def\titlerunning{Symbolic Weighted Language Models and\\ Parsing over Infinite Alphabets}
\def\authorrunning{Florent Jacquemard}

### 2021-06-03 main.tex
revise def. runs SWT and SWA and SWVPA
versions

@@
 
-\documentclass[runningheads]{llncs}
-%\documentclass[a4paper,11pt]{article}
-%\setcounter{page}{1}
+% default style article.cls
+\input{main-article}
+
+% Springer llncs style
+% https://www.springer.com/gp/computer-science/lncs/conference-proceedings-guidelines
+%\input{main-llncs}
+
+% EPTCS style
+% http://www.eptcs.org
+%\input{main-eptcs}
 
@@
 %\usepackage{mathptmx}
-
-%% theorem environments
-\usepackage{theorem}
-% \newtheorem{theorem}{Theorem} %[section]
-% \newtheorem{definition}[theorem]{Definition}
-% \newtheorem{lemma}[theorem]{Lemma}
-% \newtheorem{corollary}[theorem]{Corollary}
-% \newtheorem{proposition}[theorem]{Proposition}
-% \newenvironment{proof}{\vspace{-2ex}{\it Proof. }}{\hspace*{\fill} $\Box$\smallskip }
-% \theorembodyfont{\slshape}
-% \newtheorem{example}[theorem]{Example}
-% \newtheorem{remark}[theorem]{Remark}
-\def\endex{\hspace*{\fill} $\Diamond$\smallskip }
-{\theorembodyfont{\rmfamily} \theoremstyle{break} \newtheorem{algo}{Algorithm}}
 
@@
 %\title{Weighted Visibly Pushdown Automata and Automated Music Transcription}
-\author{Florent Jacquemard}
-\institute{INRIA \& CNAM, Paris, France\\
-\email{florent.jacquemard@inria.fr}}
-
 %\titlerunning{WVPA \& AMT}
@@
 are functions defining the weight for entering, 
-respectively leaving, a state, 
+respectively leaving, computation in a state,  
 and $\bar{\wei}$ is a triplet of transition functions 
@@
 \begin{example}
-In Common Western Music Notation~\cite{Gould11BB}, 
+In Common Western Music Notation~\cite{Gould11Notation}, 
 several symbols may be used to represent a unique sounding event.
 For instance, several notes can be combined with a tie, 
-like in $\musQuarter\!\!\mathrel{\raisebox{-1.5mm}{$\smile$}}\!\!\musEighth$,  
-and a note can be augmented with a dot like in~\musQuarterDotted).
-We propose a small transducer that compares an input sequence of sounding 
+like in $\musQuarter\!\!\!\mathrel{\raisebox{-1.5mm}{$\smile$}}\!\!\!\musEighth$,  
+and one note can be augmented by half its duration with a dot like in~\musQuarterDotted{}
+
+\noindent 
+We propose a small transducer model that compares an input sequence of sounding 
 events (music "performance") 
@@
 % simple pointwise distance between two sequences of timestamped events **
-
+%
 Let us consider the tropical (\emph{min-plus}) semiring~$\Semiring$ 
@@
 are functions defining the weight for entering, 
-respectively leaving, a state, 
+respectively leaving, computation in a state, 
 and ${\wei_1}$ is a transition functions 
-from $Q \times Q$ into $\Phi_{\Sigma}$.
+from $Q \times Q$ into~$\Phi_{\Sigma}$.
 \end{definition}
@@
 \begin{proposition} \label{prop:epsilon}
-Given a \SWT $T$ over $\Sigma$, $\Delta$, $\Semiring$ and $\bar\Phi$,
+Given a \SWT $T$ over $\Sigma$, $\Delta$, 
+$\Semiring$ commutative, bounded and complete,
+and $\bar\Phi$ effective,
+and a $\SWTA$ $A$ over $\Sigma$, $\Semiring$ and $\bar\Phi$,
+there exists an effectively constructible \SWA 
+$B_{T, A}$ over $\Delta$, $\Semiring$ and $\bar\Phi$,
+such that for all $t \in \Delta^*$, 
+$B_{T, A}(t) = \displaystyle\bigoplus_{s\in \Sigma^*} A(s) \otimes T(s, t)$.
+\end{proposition}
+%
+\begin{proof}
+Let $T = \< Q, \init_T, \bar{\wei}, \final_T >$,
+with $\bar{\wei} = \< \wei_{10}, \wei_{01}, \wei_{11}>$,
+and $A = \< P, \init_A, \wei_1, \final_A >$.
+
+\noindent
+The state set of $B_{T, A}$ will be $Q' = P \times Q$.
+The entering, leaving and transition functions of $B_{T, A}$ will 
+simulate synchronized computations of $A$ and $T$, 
+while reading an output word of $\Delta^*$.
+%
+Its state entering functions is defined 
+for all $p \in P$, $q \in Q$ 
+by $\init'(p, q) = \init_A(p) \otimes \init_T(q)$.
+The transition function $\wei'_1$ will roughly perform product of 
+transitions defined by $\wei_1$, 
+$\wei_{01}$ ($T$ reading in output word and not in input word)
+and $\wei_{11}$ ($T$ reading in output word and input word).
+
+% Initialize state leaving functions i
+%and $\final'(p, q) = \final_A(p) \otimes \final_T(q)$.
+
+%Every transition of $B_{T, A}$ will
+%simulate a sequence of transitions of $T$ performing the following steps:
+%advance in the input word while staying immobile in the output word, 
+%and then make one step in the output word (and advance in the input word or not).
+
+\noindent
+Initially, let
+\(
+\wei'_1\bigl( \< p_1, q_1>, \< p_2, q_2>\bigr) = 
+\wei_1(p_1, p_2) \otimes
+\bigl[
+\wei_{01}(q_1, q_2) 
+\oplus
+\displaystyle\bigoplus_{\Sigma}
+\wei_{11}(q_1, q_2) 
+\bigr].
+\)
+
+\noindent
+For each $p_1\in P$ and $q_1, q_2 \in Q$, do let
+\[
+\wei'_1\bigl( \< p_1, q_1>, \< p_2, q_3>\bigr) \opluseq 
+\displaystyle\bigoplus_{\Sigma} \wei_{10}(q_1, q_2) 
+\otimes 
+\wei'_1\bigl( \< p_1, q_2>, \< p_2, q_3>\bigr)
+\]
+and
+\[
+\final'(p_1, q_1) \opluseq 
+\displaystyle\bigoplus_{\Sigma} \wei_{10}(q_1, q_2) 
+\otimes \final'(p_1, q_2)
+\]
+...
+\qed
+\end{proof}
+
+complexity of construction...
+
+\begin{corollary} \label{cor:epsilon}
+Given a \SWT $T$ over $\Sigma$, $\Delta$, 
+$\Semiring$ commutative, bounded and complete,
+and $\bar\Phi$ effective,
 and $s \in \Sigma^+$, 
 there exists an effectively constructible \SWA 
-$A_{T, s}$ over $\Delta$, $\Semiring$ and $\bar\Phi$,
-such that for all $t \in \Delta^+$, $A_{T, s}(t) = T(s, t)$.
-\end{proposition}
-%
-\begin{proof}
-Let $T = \< Q, \init, \bar{\wei}, \final >$,
-with $\bar{\wei} = \< \wei_{10}, \wei_{01}, \wei_{11}>$,
-%where the component functions of
-%be the synthesized form of 
-%$Q \times (\Sigma \cup \{ \varepsilon \}) \times (\Delta \cup \{ \varepsilon \}) \times Q \to \Semiring$,
-and let $s = s_1 \ldots s_n$ with $s_1, \ldots, s_n \in \Sigma$ ($n \geq 1$). %$n = |s|$ be the length of $s$.
-
-The state set of $A_{T, s}$ will be $Q' = [1..n+1] \times Q$.
-Its state entering weight function is defined by, 
-for all $q \in Q$ $\init'(1, q) = \init(q)$ 
-and $\init'(i, q) = \zero$ for all $1 < i \leq n+1$. 
-Its state leaving weight function is defined by, for all $q \in Q$ 
-$\final'(n+1, q) = \final(q)$, 
-and $\final'(i, q) = \zero$ for all $1 \leq i <  n+1$.
-
-Every non-null transition of $A_{T, s}$ will
-simulate a sequence of transitions of $T$ performing the following steps:
-advance in the input word while staying immobile in the output word, 
-and then make one step in the output word (and advance in the input word or not).
-%
-The first steps in the sequence correspond to $\varepsilon$-transitions of automata, 
-and their total weight is computed by the following intermediate function 
-$\wei'_0 : Q' \times Q' \to \Semiring$, 
-defined for all $q, q' \in Q$: % all $1 \leq i < n$ and $1 \leq k \leq n-i$, 
-by:
-\[
-\begin{array}{rcl}
-\wei'_0\bigl(\<i, q>, \< i, q>\bigr) & = & \one 
-\quad \mathrm{if~} 1 \leq i \leq n+1,\\
-%
-\wei'_0\bigl(\<i, q>, \< i, q'>\bigr) & = & \zero 
-\quad \mathrm{if~} 1 \leq i \leq n+1 \mathrm{~and~} q \neq q',\\
-%
-\wei'_0\bigl(\<i, q>, \< i+k, q'>\bigr) & = & 
-\displaystyle
-\bigoplus_{\begin{array}{c}
-           \scriptstyle q_0,\ldots, q_k \in Q\\[-2pt]
-           \scriptstyle q_0 = q, q_k = q'
-           \end{array}} 
-\bigotimes_{j=0}^{k-1} \phi_{j, j+1}(s_{i+j})
-\quad \begin{array}[t]{l}
-      \mathrm{if~} 1 \leq i < n,
-      \mathrm{~and~} 1 \leq k \leq n-i,\\
-      \mathrm{where~} \phi_{j, j+1} = \wei_{10}(q_j, q_{j+1})
-      \end{array}
-\end{array}
-\]
-%where $\bigl[\wei_{10}(q_j, q_{j+1})\bigr]_{s_{i+j}}$ 
-%is the partial application $\eta_{s_{i+j}} \in \Phi_{\Delta}$
-%for $\eta = \wei_{10}(q_j, q_{j+1}) \in \Phi_{\Sigma, \Delta}$ ($s_{i+j} \in \Sigma$).
-%
-The function $\wei'_0$ is defined thanks to the closure properties of the label theory $\bar\Phi$
-(Section~\ref{section:symbols}). 
-% closed under partial applications 
-The sum and product in its definition are finite, 
-we shall see below how to compute the first sum in polynomial time.
-%The effective computation of $\wei'_0$ does not require enumerating all the state sequences 
-%of length $k+1$ in the sum.
-%Indeed, it amounts to search short
-
-\noindent
-We define now the transition function 
-$\wei'_1: Q' \times Q' \to \Phi_\Sigma$
-of $A_{T, s}$ as follows,
-for $q, q' \in Q$,
-$1 \leq i \leq n$, and $0 \leq k \leq n-i$:
-\[
-\begin{array}{lcl}
-\wei'_1\bigl(\< i, q>, \< i, q'>\bigr) & = & \wei_{01}(q, q')\\
-%
-\wei'_1\bigl(\< i, q>, \< i+k, q'>\bigr) & = & 
-\displaystyle
-\bigoplus_{\begin{array}{c}
-           \scriptstyle q'' \in Q\\[-2pt]
-           \scriptstyle i+k < n
-           \end{array}} 
-\wei'_0\bigl(\<i, q>, \< i+k, q''>\bigr) \otimes \psi %\eta_{s_{i+k}}
-\quad \mathrm{where~} \psi = \wei_{01}(q'', q'),\\
- & \oplus & \displaystyle
-\bigoplus_{q'' \in Q} \wei'_0\bigl(\<i, q>, \< i+k-1, q''>\bigr) \otimes \eta_{s_{i+k}}
-\quad \mathrm{where~}\eta = \wei_{11}(q'', q'),\\
-%
-\wei'_1\bigl(\< i, q>, \< j, q'>\bigr) & = & \zero  
-\quad \mathrm{if~} j < i.
-\end{array}      
-\]
-%
-In the second equation, 
-$\eta_{s_{i+j}} \in \Phi_{\Delta}$ is a partial application 
-for $\eta \in \Phi_{\Sigma, \Delta}$, $s_{i+j} \in \Sigma$, 
-defined by $\eta_{s_{i+j}}(y) = \eta(s_{i+j}, y)$ for all $y \in \Sigma$.
-
-\noindent
-We can show that the $\SWA$ $A_{T, s} = \< Q', \init', \wei'_1, \final' >$
-has the expected property: $\forall t \in \Delta^+, A_{T, s}(t) = T(s, t)$.
-\qed
-\end{proof}
-
-\noindent
-On the quadratic computation of $\wei'_0$. **
-by a best path search in the graph with nodes $[i..i+k] \times Q$
-and edges labeled in $\Phi_\Sigma$...
-ordering = summary.
+$B_{T, s}$ over $\Delta$, $\Semiring$ and $\bar\Phi$,
+such that for all $t \in \Delta^*$, $B_{T, s}(t) = T(s, t)$.
+\end{corollary}
 
@@
 Hence, $\weight_A$ is a function from 
-$[Q \times \Gamma^*] \times \Omega^* \times [Q \times \Gamma^*]$ into~$\Semiring$.
+$[Q \times \Gamma^*] \times \Omega^* \times [Q \times \Gamma^*]$ into~$\Semiring$
+(the empty stack is denoted by $\bot$). 
 %
@@
 %\begin{array}[t]{rcl}
+\weight_A\bigl(\config{q}{\bot}, \varepsilon, \config{q}{\bot}) & = \one\label{eq:SWVPA-weight}\\
+\weight_A\bigl(\config{q}{\bot}, \varepsilon, \config{q'}{\gamma}) & = \zero 
+\mathrm{~if~} q \neq q'\nonumber\\
 \weight_A\bigl(\configup{q}{\<{\call{c}}, p>\stackup\gamma}, a\, u, 
@@
                           \scriptstyle q'' \in Q\\[-2pt]
-                          \scriptstyle p \in P
+                          \scriptstyle p' \in P
                           \end{array}}}
@@
                \config{q'}{\gamma'}\bigr) & =  
- {\displaystyle\bigoplus_{q'' \in Q}} \weiec(q, {\call{c}}, p, q'') 
+ {\displaystyle\bigoplus_{\begin{array}{c}
+                          \scriptstyle q'' \in Q\\[-2pt]
+                          \scriptstyle p \in P
+                          \end{array}}}
+  \weiec(q, {\call{c}}, p, q'') 
   \otimes \weight_A\bigl(\config{q''}{{\call{c}}\, p}, u, 
@@
   \otimes \weight_A\bigl(\config{q''}{\bot}, u, 
-                         \config{q'}{\gamma'}\bigr)
+                         \config{q'}{\gamma'}\bigr)\nonumber
 %\end{array}
-\label{eq:SWVPA-weight}
 \end{align}
 %
-where $\bot$ denotes the empty stack.
 % and ${\call{c}}\, p\stacksep \gamma$ 
@@
 sequences in transition with a non-null weight.
-\qed\end{proof}
+\qed
+\end{proof}
 
@@
 %\bibliographystyle{splncs04}
+%\bibliographystyle{eptcs}
+%\bibliography{generic}
+
 \bibliography{references}

### 2021-06-03 semiring.tex
versions

@@
 %\medskip
+\noindent
 We shall need below infinite sums with~$\oplus$.
@@
 $(x_i)_{i \in I}$ %$\{ x_i \mid i \in I \}$
-of elements of $\dom(\Semiring)$ over an index set $I \subset \mathbb{N}$, such that
-%is well-defined and in $\dom(\Semiring)$,
-the following holds: %properties
+of elements of $\dom(\Semiring)$ over an index set $I \subset \mathbb{N}$, such that:
+%the following holds: %properties
 \begin{description}

### 2021-06-04 main.tex
conclusion

@@
 Let $T = \< Q, \init_T, \bar{\wei}, \final_T >$,
-with $\bar{\wei} = \< \wei_{10}, \wei_{01}, \wei_{11}>$,
-and $A = \< P, \init_A, \wei_1, \final_A >$.
-
+where $\bar{\wei}$ contains $\wei_{10}$, $\wei_{01}$, and $\wei_{11}$,
+from $Q \times Q$ into respectively 
+$\Phi_{\Sigma}$, $\Phi_{\Delta}$, and $\Phi_{\Sigma, \Delta}$,
+and let $A = \< P, \init_A, \wei_1, \final_A >$
+with $\wei_1: Q \times Q \to \Phi_{\Sigma}$.
+%
 \noindent
@@
 by $\init'(p, q) = \init_A(p) \otimes \init_T(q)$.
-The transition function $\wei'_1$ will roughly perform product of 
-transitions defined by $\wei_1$, 
+The transition function $\wei'_1$ will roughly perform 
+a synchronized product of transitions defined by $\wei_1$, 
 $\wei_{01}$ ($T$ reading in output word and not in input word)
 and $\wei_{11}$ ($T$ reading in output word and input word).
-
+%
+Moreover, $\wei'_1$ also needs to simulate transitions 
+defined by $\wei_{10}$: $T$ reading in intput word and not in output word.
+Since $B_{T, A}$  will read only in the output word, such a transition corresponds
+to an $\varepsilon$-transition of $\SWA$, 
+but $\SWA$ have been defined without $\varepsilon$-transitions.
+Therefore, in order to take care of this case, we perform an on-the-fly
+suppression of $\varepsilon$-transition in the $\SWA$ in construction, 
+following the algorithm of~\cite{LombardySakarovitch12ciaa}. 
+%
 % Initialize state leaving functions i
@@
 \noindent
-Initially, let
-\(
+Initially, for all $p_1, p_2 \in P$, and $q_1, q_2 \in Q$, let
+\[
 \wei'_1\bigl( \< p_1, q_1>, \< p_2, q_2>\bigr) = 
@@
 \bigr].
-\)
-
-\noindent
-For each $p_1\in P$ and $q_1, q_2 \in Q$, do let
+\]
+
+\noindent
+Iterate the following for all $p_1\in P$ and $q_1, q_2 \in Q$:
+for all $p_2\in P$ and $q_3 \in Q$,
 \[
@@
 \] 
-...
+\marginpar{\tiny proof correctness}
 \qed
@@
 
-complexity of construction...
+%The complexity of construction
+\noindent
+The construction time and size for $B_{T, A}$ are $O(\| T \|^3 . \| A \|^2)$,
+where the sizes $\| T \|$ and $\| A \|$ are their number of states.
+%of~$T$ is its number of states $|Q|$.
 
@@
 
-\noindent
-The construction time and size for $A_{T, s}$ are $O(\| T \| . | s |)$,
-where the size $\| T \|$ of $T$ is its number of states $|Q|$.
+%\noindent
+%The construction time and size for $B_{T, s}$ are $O(\| T \|^3 . | s |^2)$,
+%where the size $\| T \|$ of~$T$ is its number of states $|Q|$.
 
@@
 
+\noindent 
+Every $\SWA$ $A = \< Q, \init, {\wei_1}, \final >$,
+over $\Sigma$, $\Semiring$ and $\bar\Phi$
+is a particular case of $\SWVPA$ 
+$\< Q, \emptyset, \init, \bar\wei, \final >$ 
+over $\Omega$, $\Semiring$ and $\bar\Phi$
+with $\Omegai = \Sigma$ and $\Omegac = \Omegar = \emptyset$,
+and computing with an always empty stack:
+$\weiei = \wei_1$ and all the other functions 
+of~$\bar\wei$ are the constant~$\zero$.
+
 \subsection{Properties}
@@
 \begin{proposition}\label{prop:SWVPA-product}
-Let $A_1$ and $A_2$ be two (\SWVPA)
+Let $A_1$ and $A_2$ be two \SWVPA
 over the same $\Omega$, $\Semiring$ and $\bar\Phi$.
-There exists two $\SWVPA$ $A_1 \oplus A_2$ and $A_1 \otimes A_2$, 
-effectively constructible, 
+There exists two effectively constructible $\SWVPA$ 
+$A_1 \oplus A_2$ and $A_1 \otimes A_2$,  
 such that for all $s \in \Omega^*$, 
@@
 \end{proposition}
+%
+\begin{proof}
 The construction is essentially the same 
 as in the case of the Boolean semiring~\cite{dAntonyAlur14SVPDA}.
+\end{proof}
 
@@
 Let us assume that the semiring~$\Semiring$ is
-commutative, bounded, total, and complete.
+commutative, bounded, and complete,  
+\marginpar{\tiny total?}
 and assume an effective label theory.
@@
 %are computable...
-
+% 
 % total ? 
@@
 %... exist and in $\Semiring$.
-
+%
 We propose a Dijkstra algorithm computing, for a $\SWVPA$ $A$
 over~$\Omega$, $\Semiring$ and~$\bar\Phi$, 
-the minimal weight \wrt~$\leq_\oplus$, for a word in~$\Omega^*$.
+the minimal weight %\wrt~$\leq_\oplus$, 
+for a word in~$\Omega^*$.
 
@@
 Since $\Semiring$ is complete, the infinite sum in \eqref{eq:bbot} is well defined,
-and by totality of $\leq_\oplus$, it is the minimum in $\Omega^*$
+and, roviding that $\Semiring$ is total, it is the minimum in $\Omega^*$,
+\wrt~$\leq_\oplus$, % this ordering.
 of the fonction 
-$s \mapsto \weight_A(\config{q}{\sigma}, s, \config{q'}{\sigma})$ 
-\wrt this ordering.
+$s \mapsto \weight_A(\config{q}{\sigma}, s, \config{q'}{\sigma})$.
 %
@@
 
-\subsection{Definition}
-
+%\subsection{Definition}
+%
 Let $\Sigma$ be a countable input alphabet 
@@
 \section*{Conclusion}
+% summary 
+We have introduced weighted language models (SW transducers and visibly pushdown automata)
+computing over infinite alphabets, 
+and applied them to the problem of parsing 
+with infinitely many possible input symbols (typically timed events).
+%handled with suitable language formalisms, 
+%
+This approach extends conventional parsing and weighted parsing
+by computing a derivation tree modulo 
+a generic distance between words,  
+defined by a SW transducer given in input.
+This enables considering finer word relationships than strict equality 
+as in the conventional parsing approach, 
+opening possibilities of quantitative analysis via this method.
+
 % discussion
-
-\noindent
--- summary
+Ongoing and future work include
 
@@
 -- other theoretical properties of SW models
+\eg extention to $n$-best, 
+to $k$-\emph{closed} semirings~\cite{Mohri02semiring}
+(instead of \emph{bounded}, which is equivalent to $0$-\emph{closed})
 
@@
 
-and extention to $n$-best
-
-\noindent
--- offline algorithm, 
+\noindent
+-- present here an offline algorithm for best search, 
 semi-online implementation for AMT (bar-by-bar approach)
+with an on-the-fly automata construction.
 

### 2021-06-04 main.tex
pass section trees and parsing

@@
 containing one set~$\Phi_{\Omega_k}$ for each $k \in [0..M]$.
-
+%
 \renewcommand{\call}[1]{\ensuremath \langle_{#1}}
 \renewcommand{\return}[1]{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
-
-Let $\hat\Omega$ be the countable (unranked) alphabet obtained
-from $\Omega$ by: 
-$\hat\Omega = \Omegai \uplus \Omegac \uplus \Omegar$, with
-$\Omegai = \Omega_0$, 
-$\Omegac = \{ \; \call{a} \mid a \in \Omega_{>0} \}$,
-$\Omegar = \{ \; \return{a} \mid a \in \Omega_{>0} \}$.
-We associate to $\hat\Omega$
-a label theory $\hat{\Phi}$ 
-like in Section~\ref{se:SWVPA-def}.
-
-\noindent
-We define a linearization of trees of $\T(\Omega)$ into 
-words of $\hat{\Omega}^*$ as follows:
-\begin{description}
-\item $\lin(a) = a$ for all $a \in \Omega_0$, 
-\item $\lin\bigl( b(t_1, \ldots, t_k)\bigr) = 
-       \call{b} \; \lin(t_1) \ldots \lin(t_k) \; \return{b}$ 
-       when $b \in \Omega_k$, $1 \leq k \leq M$.
-\end{description}
 
@@
 %$\wei_\varepsilon$ from $Q \times Q$ into $\Semiring$, and, 
-$\wei_{k}: Q \times Q^{k} \to \Phi_{\Omega_k}$
+one function $\wei_{k}: Q \times Q^{k} \to \Phi_{\Omega_k}$
 for each $k \in [0..M]$. 
@@
 %\wei(q_0, \varepsilon, q_1) & = &  \wei_\varepsilon(q_0, q_1),\\ %\phi_\varepsilon\\
-\wei(q_0, a, q_1 \ldots q_k) & = & \phi_{\Omega, k}(a) &
+\wei(q_0, a, q_1 \ldots q_k) & = & \phi_{k}(a) &
 \quad\mathrm{where~} \phi_{k} = \wei_{k}(q_0, q_1\ldots q_k).\\
@@
 \]
-Intuitively, $\wei(q_0, a, q_1 \ldots q_k)$ can be seen as
-the weight of a production rule $q_0 \to a(q_1, \ldots, q_k)$ 
-of a regular tree grammar~\cite{tata}, 
-that replaces the non-terminal symbol $q_0$ by $a(q_1, \ldots, q_k)$. 
-%in a step of tree building.
-%
-Such a grammar computes the weights of the derivation trees 
-of the Context-Free grammar obtained by forgetting the labeling symbols of $\Omega_{>0}$.
-
+%
 \noindent
 The \SWTA of Definition~\ref{def:SWTA} defines a mapping 
-from trees of $\T(\Omega)$ into the weights of~$\Semiring$,
+from trees of $\T(\Omega)$ into the weight values in~$\Semiring$,
 based on the intermediate function~$\weight_A$
@@
 
-\begin{lemma}\label{lem:SWTA}
-For all \SWTA $A$ over~$\Omega$, $\Semiring$ commutative, and $\Phi$,
+\noindent
+Intuitively, $\wei(q_0, a, q_1 \ldots q_k)$ can be seen as
+the weight of a production rule $q_0 \to a(q_1, \ldots, q_k)$ 
+of a regular tree grammar $G$~\cite{tata}, 
+that replaces the non-terminal symbol $q_0$ by $a(q_1, \ldots, q_k)$. 
+%in a step of tree building.
+%
+%Such a grammar computes the weights of the derivation trees 
+%of the Context-Free grammar obtained by forgetting the labeling symbols of $\Omega_{>0}$.
+And the above production rule can also be seen as 
+a rule of a weighted CF grammar, of the form
+$[a]\, q_0 := q_1 \ldots q_k$ if $k > 0$,
+and $q_0 := a$ if $k = 0$. 
+In the first case, $a$ is a label of the rule, 
+in the second case it is a terminal symbol.
+On can observe that the weight of a labeled derivation tree $t$
+of the weighted CF grammar associated to~$A$ as above, 
+is $\weight_A(q, t)$, 
+when $q$ is the start non-terminal.
+%
+We shall now establish a correspondence between such derivation tree $t$
+and some word describing a linearization of $t$, 
+in a way that $\weight_A(q, t)$ can be computed by a $\SWVPA$.
+
+Let $\hat\Omega$ be the countable (unranked) alphabet obtained
+from $\Omega$ by: 
+$\hat\Omega = \Omegai \uplus \Omegac \uplus \Omegar$, with
+$\Omegai = \Omega_0$, 
+$\Omegac = \{ \; \call{a} \mid a \in \Omega_{>0} \}$,
+$\Omegar = \{ \; \return{a} \mid a \in \Omega_{>0} \}$.
+
+\noindent
+We associate to $\hat\Omega$
+a label theory $\hat{\Phi}$ 
+like in Section~\ref{se:SWVPA-def}, 
+%
+\noindent
+and we define a linearization of trees of $\T(\Omega)$ into 
+words of $\hat{\Omega}^*$ as follows:
+\begin{description}
+\item $\lin(a) = a$ for all $a \in \Omega_0$, 
+\item $\lin\bigl( b(t_1, \ldots, t_k)\bigr) = 
+       \call{b} \; \lin(t_1) \ldots \lin(t_k) \; \return{b}$ 
+       when $b \in \Omega_k$ for $1 \leq k \leq M$.
+\end{description}
+
+\begin{proposition}\label{lem:SWTA}
+For all \SWTA $A$ over~$\Omega$, $\Semiring$ commutative, and $\bar\Phi$,
 there exists an effectively constructible \SWVPA $A'$ over 
@@
 such that for all $t \in \T(\Omega)$, $A'\bigl(\lin(t)\bigr) = A(t)$.
-\end{lemma} 
+\end{proposition} 
 % 
@@
 %
-Besides considering infinitely many possible of input symbols, 
-handled with suitable language formalisms, 
-this approach extends conventional parsing and weighted parsing
-by computing a derivation tree modulo 
-a generic distance between words,  
-defined by a SW transducer given in input.
-This enables considering finer word relationships than strict equality 
-as in the conventional parsing approach, 
-opening possibilities of quantitative analysis via this method.
-
 %\subsection{Definition}
 %
-Let $\Sigma$ be a countable input alphabet 
-and let $\Omega$ be a countable output ranked alphabet, 
-with maximal rank value $M$, 
-and $\hat\Omega = \Omegai \uplus \Omegac \uplus \Omegar$
-be the alphabet with nesting symbols, associated like Section~\ref{sec:trees}, 
-for the linearization of trees of $\T(\Omega)$
-(remember that $\Omegai = \Omega_0$).
-%
-Let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative, bounded, complete 
-and total semiring 
-and let $\bar\Phi$ be a label theory over $\Semiring$
-containing
-$\Phi_\Sigma$, $\Phi_{\Sigma, \Omegai}$, as well as
+Let~$\Sigma$ be a countable input alphabet 
+and~$\Omega$ be a countable output ranked alphabet, 
+with maximal rank value~$M$, 
+and let $\hat\Omega = \Omegai \uplus \Omegac \uplus \Omegar$
+be the alphabet with nesting symbols associated to $\Omega$ like in Section~\ref{sec:trees},
+for the linearization of trees of $\T(\Omega)$.
+%(remember that $\Omegai = \Omega_0$).
+%
+Let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a 
+commutative, bounded, and complete \marginpar{\tiny total?} semiring  
+and let $\bar\Phi$ be an effective label theory over $\Semiring$,
+containing $\Phi_\Sigma$, $\Phi_{\Sigma, \Omegai}$, as well as
 $\Phii$, $\Phic$, $\Phir$, $\Phicr$
 (following the notations of Section~\ref{se:SWVPA-def}).
-It is moreover assumed computable and $k$-convex for some fixed $k$.
-
-\noindent
-Let us assume given the following input:
-\begin{itemize}
-\item a \SWT $T$ over $\Sigma$, $\Omegai$, $\Semiring$, and $\bar\Phi$, 
+%It is moreover assumed computable and $k$-convex for some fixed~$k$.
+%
+\noindent
+We assume given the following input:
+\begin{description}
+\item[--] a \SWT $T$ over $\Sigma$, $\Omegai$, $\Semiring$, and $\bar\Phi$, 
 defining a measure %between words 
@@
 
-\item a \SWTA $A$ over $\Omega$, $\Semiring$, and $\bar\Phi$,
+\item[--] a \SWTA $A$ over $\Omega$, $\Semiring$, and $\bar\Phi$,
 defining a measure $A: \T(\Omega) \to \Semiring$,
@@
 %      $A : \Omega^* \to \Semiring$,
-\item an input word $s \in \Sigma^*$.
-\end{itemize}
+\item[--] an input word $s \in \Sigma^*$.
+\end{description}
 %
@@
 $A = \< Q, \init, \bar{\wei} >$ can be seen as a weighted regular tree grammar, 
-that generates (weighted) trees by replacement of a state symbol $q_0$ (non-terminal), 
-by a tree $a(q_1,\ldots, q_k)$, where $k = \rank(a)$.
-A replacement rule $q_0 \to a(q_1,\ldots, q_k)$, 
-of weight $\wei(q_0, a, q_1 \ldots q_k) \in \Semiring$ 
-according to Definition~\ref{def:SWTA}, 
-corresponds to the production rule $q_0 := a(q_1,\ldots, q_k)$ of a weighted CF grammar,
-with set non-terminal symbols $Q$ and set of terminal symbols $\Omega_0$.
-This actually is a slight generalization of CFG since 
-each such production rule is labelled by  a symbol of $\Omega_{>0}$, 
-hence parse trees %derivation trees
-are trees of $\T(\Omega)$.
-%
-Another (more original) generalization is that the set of terminal symbols 
-$\Omega_0$ may be infinite.
+or a labeled weighted CF grammar whose derivation trees are in $\T(\Omega)$, 
+with weight defined by $A$.
+%that generates (weighted) trees by replacement of a state symbol~$q_0$ (non-terminal), 
+%by a tree $a(q_1,\ldots, q_k)$, where $k = \rank(a)$.
+%A replacement rule $q_0 \to a(q_1,\ldots, q_k)$, 
+%of weight $\wei(q_0, a, q_1 \ldots q_k) \in \Semiring$ 
+%according to Definition~\ref{def:SWTA}, 
+%corresponds to the production rule $q_0 := a(q_1,\ldots, q_k)$ of a weighted CF grammar,
+%with set non-terminal symbols $Q$ and set of terminal symbols $\Omega_0$.
+%
+%This actually is a slight generalization of CFG since 
+%each such production rule is labelled by  a symbol of $\Omega_{>0}$, 
+%hence parse trees %derivation trees
+%are trees of $\T(\Omega)$.
+%
+%Another (more original) generalization is that the set of terminal symbols 
+%$\Omega_0$ may be infinite.
 
@@
 The measure expressed in~\eqref{eq:distance-lang} is called
-called edit-distance between $s$ and $A$ in~\cite{Mohri03EDWA}.
+called the edit-distance between~$s$ and~$A$ in~\cite{Mohri03EDWA}.
 %
@@
 the best parse tree matching a given input,  
-sometimes referred as~\emph{weighted parsing}
+sometimes referred as~\emph{weighted parsing},
 corresponds to SW parsing in the case of finite alphabets
@@
 
-\subsection{Computation}
-
+%\subsection{Computation}
+%
 \begin{proposition}
@@
 
-\subsection{Application to Automated Music Transcription}
+\paragraph{Application to Automated Music Transcription.}
 %\input{transcription}
+...
 
@@
 defined by a SW transducer given in input.
-This enables considering finer word relationships than strict equality 
-as in the conventional parsing approach, 
+This enables to consider finer word relationships than strict equality, 
+%as in the conventional parsing approach, 
 opening possibilities of quantitative analysis via this method.
@@
 % discussion
+\noindent
 Ongoing and future work include
@@
 \noindent
--- other theoretical properties of SW models
-\eg extention to $n$-best, 
-to $k$-\emph{closed} semirings~\cite{Mohri02semiring}
-(instead of \emph{bounded}, which is equivalent to $0$-\emph{closed})
-
-\noindent
--- room to improve complexity for best-search algorithm
+-- The study of other theoretical properties of SW models, 
+such as the extension of the best search algorithm from $1$-best to $n$-best~\cite{Huang05kbest}, 
+and to $k$-\emph{closed} semirings~\cite{Mohri02semiring}
+(instead of \emph{bounded}, which corresponds to $0$-\emph{closed}).
+
+\noindent
+-- ...there is room to improve the complexity bounds for the algorithms
 ... modular approach with oracles ...

### 2021-06-08 main.tex
rev. def. swTA (check parent node) and conversiion VPA

@@
 \def\weir{{\wei_\mathsf{r}}}
-\def\weie{{\wei_\mathsf{e}}}
-\def\weiei{{\wei^\mathsf{e}_\mathsf{i}}}
-\def\weiec{{\wei^\mathsf{e}_\mathsf{c}}}
-\def\weier{{\wei^\mathsf{e}_\mathsf{r}}}
-\def\weiex{{\wei^\mathsf{e}_\mathsf{x}}}
+\newcommand{\weie[1]}{\wei_{#1}^\mathsf{e}}
+\def\weiei{\weie[\mathsf{i}]}
+\def\weiec{\weie[\mathsf{c}]}
+\def\weier{\weie[\mathsf{r}]}
+\def\weiex{\weie[\mathsf{x}]}
 \newcommand{\ioi}[1]{\mathsf{ioi}({#1})}
@@
 A \emph{symbolic-weighted tree automaton} (\SWTA)
-over $\Omega$, $\Semiring$, and $\bar\Phi$
+over $\Omega$, $\Semiring$, and~$\bar\Phi$
 is a triplet $A = \< Q, \init, \bar{\wei} >$ where
 $Q$ is a finite set of states, 
-$\mathsf{in} : Q \to \Semiring$ is the starting weight function, 
-and $\bar{\wei}$ is a tuplet of transition functions containing
+$\mathsf{in} : Q \to \Phi_\Omega$ is the starting weight function, 
+and $\bar{\wei}$ is a tuplet of transition functions containing, 
+for each $k \in [0..M]$, 
 %$\wei_\varepsilon$ from $Q \times Q$ into $\Semiring$, and, 
-one function $\wei_{k}: Q \times Q^{k} \to \Phi_{\Omega_k}$
-for each $k \in [0..M]$. 
+the functions $\wei_{k}: Q \times Q^{k} \to \Phi_{\Omega_{>0},\Omega_k}$
+and $\weie[k]: Q \times Q^{k} \to \Phi_{\Omega_k}$.
 \end{definition}
@@
 %Like in Section~\ref{sec:SWAdef}, 
-We define from $\bar{\wei}$ 
-a transition function~$\wei$, %from $\bar\wei$
-from 
-$Q \times \Omega \times \bigcup_{k=0}^{M} Q^k$
-into~$\Semiring$ by: %also called $\wei$ for simplicity, 
+We define %from $\bar{\wei}$ 
+a transition 
+function~$\wei: Q \times (\Omega_{> 0} \cup \{ \varepsilon \}) \times \Omega \times \bigcup_{k=0}^{M} Q^k 
+  \to \Semiring$
+by: %also called $\wei$ for simplicity, 
 %such that, for all $q, q' \in Q$, $a \in \Sigma$, and $b \in \Delta$, 
@@
 %\wei(q_0, \varepsilon, q_1) & = &  \wei_\varepsilon(q_0, q_1),\\ %\phi_\varepsilon\\
-\wei(q_0, a, q_1 \ldots q_k) & = & \phi_{k}(a) &
-\quad\mathrm{where~} \phi_{k} = \wei_{k}(q_0, q_1\ldots q_k).\\
+\wei(q_0, a, b, q_1 \ldots q_k) & = & \eta(a, b) &
+\quad\mathrm{where~} \eta = \wei_{k}(q_0, q_1\ldots q_k)\\
+\wei(q_0, \varepsilon, b, q_1 \ldots q_k) & = & \phi(b) &
+\quad\mathrm{where~} \phi = \weie[k](q_0, q_1\ldots q_k).
 \end{array}      
@@
 \noindent
-The \SWTA of Definition~\ref{def:SWTA} defines a mapping 
+The first case deals with a strict subtree, with a parent node labeled by $a$, 
+and the second case with a root tree.
+%
+Every \SWTA %of Definition~\ref{def:SWTA} 
+defines a mapping 
 from trees of $\T(\Omega)$ into the weight values in~$\Semiring$,
-based on the intermediate function~$\weight_A$
-defined as follows for $q_0 \in Q$ and 
-$t = b(t_1,\ldots, t_k) \in \T(\Omega)$,
-with $0 \leq k \leq M$:
+based on an intermediate function
+$\weight_A: Q \times (\Omega \cup \{ \varepsilon \}) \times \T(\Omega) \to \Semiring$
+defined as follows 
 \begin{equation}
 \begin{array}{rccl}
-\weight_A(q_0, t) & = & 
+\weight_A(q_0, a, t) & = & 
 %   \displaystyle\bigoplus_{q_1 \in Q} &
@@
  \displaystyle\bigoplus_{q_1 \ldots q_k \in Q^k} &
-              \wei(q_0, b, q_1 \ldots q_k ) 
+              \wei(q_0, a, b, q_1 \ldots q_k ) 
    \otimes \displaystyle\bigotimes_{i=1}^{k}
-           \weight_A(q_i, t_i)
+           \weight_A(q_i, b, t_i)
 \end{array}
 \end{equation}
+where $q_0 \in Q$, $a \in \Omega_{>0} \cup \{ \varepsilon \}$ and 
+$t = b(t_1,\ldots, t_k) \in \T(\Omega)$,
+$0 \leq k \leq M$.
 
 \medskip\noindent
-The weight associated by $A$ to  $t \in \T(\Omega)$ is 
+Finally, the weight associated by $A$ to  $t \in \T(\Omega)$ is 
 \begin{equation}
 A(t)  = 
-\displaystyle\bigoplus_{q \in Q} \mathsf{in}(q) \mathop{\otimes} \weight_A(q, t)
+\displaystyle\bigoplus_{q \in Q} \mathsf{in}(q) \mathop{\otimes} \weight_A(q, \varepsilon, t)
 \label{eq:weightTA}
@@
 \noindent
-Intuitively, $\wei(q_0, a, q_1 \ldots q_k)$ can be seen as
-the weight of a production rule $q_0 \to a(q_1, \ldots, q_k)$ 
-of a regular tree grammar $G$~\cite{tata}, 
-that replaces the non-terminal symbol $q_0$ by $a(q_1, \ldots, q_k)$. 
+Intuitively, $\wei(q_0, a, b, q_1 \ldots q_k)$ can be seen as
+the weight of a production rule $q_0 \to b(q_1, \ldots, q_k)$ 
+of a regular tree grammar~\cite{tata}, 
+that replaces the non-terminal symbol $q_0$ by $b(q_1, \ldots, q_k)$, 
+provided that the parent of $q_0$ is labeled by $a$
+(or $q_0$ is the root node if $a = \varepsilon$).
 %in a step of tree building.
@@
 %of the Context-Free grammar obtained by forgetting the labeling symbols of $\Omega_{>0}$.
-And the above production rule can also be seen as 
+The above production rule can also be seen as 
 a rule of a weighted CF grammar, of the form
-$[a]\, q_0 := q_1 \ldots q_k$ if $k > 0$,
-and $q_0 := a$ if $k = 0$. 
-In the first case, $a$ is a label of the rule, 
-in the second case it is a terminal symbol.
-On can observe that the weight of a labeled derivation tree $t$
+$[a, b]\, q_0 := q_1 \ldots q_k$ if $k > 0$,
+and $[a]\, q_0 := b$ if $k = 0$. 
+In the first case, $b$ is a label of the rule, 
+and in the second case, it is a terminal symbol.
+And in both cases, $a$ is a constraint on the label of the 
+parent node in the derivation tree.
+This features of observation of parent's label  
+are useful in the case of infinite alphabet, 
+where it is not possible to memorize a label with the states.
+%
+%One can observe that 
+The weight of a labeled derivation tree $t$
 of the weighted CF grammar associated to~$A$ as above, 
@@
 \begin{array}{ll}
-\weii(\bar{q}, {\call{b}}, \bar{p}, a, \bar{q} q') = \wei(q', a, \varepsilon) & 
-\mathrm{for~all~} b \in \Omega_{>0}, \bar{p}\in P', a \in \Omega_0\\
-%
-\weiei(\bar{q}, a, \bar{q} q') = \wei(q', a, \varepsilon) & 
+\weii(\bar{q}, {\call{c}}, \bar{p}, a, \bar{q} q') = \wei(q', c, a, \varepsilon) & 
+\mathrm{for~all~} c \in \Omega_{>0}, \bar{p}\in P', a \in \Omega_0\\
+%
+\weiei(\bar{q}, a, \bar{q} q') = \wei(q', \varepsilon, a, \varepsilon) & 
 \mathrm{for~all~} a \in \Omega_0\\
 %
-\weic(\bar{q}, {\call{b}}, \bar{p}, \call{b'}, \varepsilon, \bar{q}) = \one & %\wei(p', b, \bar{p}) & 
-\mathrm{for~all~} b \in \Omega_{>0},  \bar{p}\in P', b, b' \in \Omega_{>0}\\
-%
-\weiec(\bar{q}, \call{b}, \varepsilon, \bar{q}) = \one & %\wei(p', b, \bar{p}) & 
-\mathrm{for~all~} b \in \Omega_{>0}\\
+\weic(\bar{q}, {\call{c}}, \bar{p}, \call{c'}, \varepsilon, \bar{q}) = \wei(q', c, c', \bar{q}) &  
+\mathrm{for~all~} c, c' \in \Omega_{>0},  \bar{p}\in P'\\
+%
+\weiec(\bar{q}, \call{c}, \varepsilon, \bar{q}) = \wei(q', \varepsilon, c, \bar{q}) & 
+\mathrm{for~all~} c \in \Omega_{>0}\\
 %
 %\weir: Q \times \Omegac \times P \times \Sigmar \times Q \to \Semiring & 
-\weir(\bar{q}, {\call{b}}, \bar{p}, {\return{b}}, \bar{p} q') = \wei(q', b, \bar{q}) & 
+\weir(\bar{q}, {\call{b}}, \bar{p}, {\return{b}}, \bar{p} q') = \one & 
 \mathrm{for~all~}  b \in \Omega_{>0}, \bar{p}\in P'\\

### 2021-06-08 main.tex
revise conversion swTA to swVPA

@@
 \noindent
+where $q_1\ldots q_k$ is $\varepsilon$ if $k = 0$.
 The first case deals with a strict subtree, with a parent node labeled by $a$, 
-and the second case with a root tree.
-%
+and the second case is for a root tree.
+
+\noindent
 Every \SWTA %of Definition~\ref{def:SWTA} 
 defines a mapping 
-from trees of $\T(\Omega)$ into the weight values in~$\Semiring$,
-based on an intermediate function
+from trees of $\T(\Omega)$ into~$\Semiring$, %the weight values in~$\Semiring$,
+based on the following intermediate function
 $\weight_A: Q \times (\Omega \cup \{ \varepsilon \}) \times \T(\Omega) \to \Semiring$ 
-defined as follows 
 \begin{equation}
-\begin{array}{rccl}
-\weight_A(q_0, a, t) & = & 
+%\begin{array}{rccl}
+\weight_A(q_0, a, t) =  % & = &
 %   \displaystyle\bigoplus_{q_1 \in Q} &
 %   \wei(q, \varepsilon, q_1) \otimes \weight_A(q_1, t)\\
- \displaystyle\bigoplus_{q_1 \ldots q_k \in Q^k} &
+ \displaystyle\bigoplus_{q_1 \ldots q_k \in Q^k} % &
               \wei(q_0, a, b, q_1 \ldots q_k ) 
@@
            \weight_A(q_i, b, t_i)
-\end{array}
+%\end{array}
 \end{equation}
@@
 and in the second case, it is a terminal symbol.
-And in both cases, $a$ is a constraint on the label of the 
-parent node in the derivation tree.
-This features of observation of parent's label  
+And in both cases, $a$ is a constraint on the label of rule applied 
+on the parent node in the derivation tree.
+This features of observing the parent's label  
 are useful in the case of infinite alphabet, 
@@
 %One can observe that 
-The weight of a labeled derivation tree $t$
+\noindent The weight of a labeled derivation tree $t$
 of the weighted CF grammar associated to~$A$ as above, 
@@
 where $Q' = \bigcup_{k=0}^{M} Q^k$ is the set of sequences of state symbols of $A$, 
-of length at most $M$, including the empty sequence denoted by $\varepsilon$, 
+of length at most $M$, including the empty sequence denoted by~$\varepsilon$, 
 and where $P' = Q'$ and $\bar\wei$ is defined by:
@@
 \[
-\begin{array}{ll}
-\weii(\bar{q}, {\call{c}}, \bar{p}, a, \bar{q} q') = \wei(q', c, a, \varepsilon) & 
-\mathrm{for~all~} c \in \Omega_{>0}, \bar{p}\in P', a \in \Omega_0\\
-%
-\weiei(\bar{q}, a, \bar{q} q') = \wei(q', \varepsilon, a, \varepsilon) & 
+\begin{array}{lcll}
+\weii(q_0\, \bar{u}, {\call{c}}, \bar{p}, a, \bar{u}) & = & \wei(q_0, c, a, \varepsilon) & 
+\mathrm{for~all~} c \in \Omega_{>0}, a \in \Omega_0\\ %\bar{p}\in P', 
+%
+\weiei(q_0\,\bar{u}, a, \bar{u}) & = & \wei(q_0, \varepsilon, a, \varepsilon) & 
 \mathrm{for~all~} a \in \Omega_0\\
 %
-\weic(\bar{q}, {\call{c}}, \bar{p}, \call{c'}, \varepsilon, \bar{q}) = \wei(q', c, c', \bar{q}) &  
-\mathrm{for~all~} c, c' \in \Omega_{>0},  \bar{p}\in P'\\
-%
-\weiec(\bar{q}, \call{c}, \varepsilon, \bar{q}) = \wei(q', \varepsilon, c, \bar{q}) & 
+\weic(q_0\,\bar{u}, {\call{c}}, \bar{p}, \call{d}, \bar{u}, \bar{q}) & = & \wei(q_0, c, d, \bar{q}) &  
+\mathrm{for~all~} c, d \in \Omega_{>0}\\ % \bar{p}\in P'
+%
+\weiec(q_0\,\bar{u}, \call{c}, \bar{u}, \bar{q}) & = & \wei(q_0, \varepsilon, c, \bar{q}) & 
 \mathrm{for~all~} c \in \Omega_{>0}\\
@@
 %\weir: Q \times \Omegac \times P \times \Sigmar \times Q \to \Semiring & 
-\weir(\bar{q}, {\call{b}}, \bar{p}, {\return{b}}, \bar{p} q') = \one & 
-\mathrm{for~all~}  b \in \Omega_{>0}, \bar{p}\in P'\\
+\weir(\varepsilon, {\call{c}}, \bar{p}, {\return{c}}, \bar{p}) & = & \one & 
+\mathrm{for~all~}  c \in \Omega_{>0}\\ % \bar{p}\in P'
 %
 %\weie: Q \times \Omegar \times Q \to \Semiring & 
-\weier(\bar{p}, {\return{b}}, \bar{q}) = \zero &
-\mathrm{for~all~}  b \in \Omega_{>0}\\
+\weier(\bar{u}, {\return{c}}, \bar{q}) & = & \zero &
+\mathrm{for~all~}  c \in \Omega_{>0}
 \end{array}      
@@
 \noindent
-It is sufficient to consider in $Q'$ only the prefixes of 
-sequences in transition with a non-null weight.
+All cases not matched by one of the above equations have a weight $\zero$, 
+for instance  %This is in particular the case of
+$\weir(\bar{u}, {\call{c}}, \bar{p}, {\return{d}}, \bar{q}) = \zero$
+if $c \neq d$
+or $\bar{u} \neq \varepsilon$
+or $\bar{q} \neq \bar{p}$.
+%
+%\noindent
+%It is sufficient to consider in $Q'$ only the prefixes of 
+%sequences in transition with a non-null weight.
 \qed

### 2021-06-08 main.tex
1 pass section parsing

@@
 with weight defined by $A$.
+The purpose of the transducer $T$ is to measure a distance between input and output words, 
+\ie between the word $s$ and the sequence leaves of a derivation tree, in $\Omegai^*$.
 %that generates (weighted) trees by replacement of a state symbol~$q_0$ (non-terminal), 
@@
 such that for all $s \in \Sigma^*$, and $u \in {\hat\Omega}^*$, 
-$T'(s, u) = T(s, u|_{\Omegai})$. The transducer $T'$ simplify skips every symbol 
-$b \in {\hat\Omega} \setminus \Omegai$ in input, 
-with new transitions of the form $\wei_{01}(q, \varepsilon, b, q')$.
+$T'(s, u) = T(s, u|_{\Omegai})$. The transducer $T'$ simply skips every symbol 
+$b \in {\hat\Omega} \setminus \Omegai$, 
+by the addition to the transition of $T$,
+of new transitions of the form $\wei_{01}(q, \varepsilon, b, q')$.
 
@@
 This automaton is such that for all $t \in \T(\Omega)$, 
-$A_{T', s}\bigl(\lin(t)\bigr) 
+\[ 
+   A_{T', s}\bigl(\lin(t)\bigr) 
  = T'\bigl(s, \lin(t)\bigr) 
  = T'\bigl(s, \lin(t)|_{\Omegai}\bigr) 
- = d(s, t)$.
+ = d(s, t).
+\]
 

### 2021-06-09 main.tex
format LIPIcs

@@
 % default style article.cls
-\input{main-article}
+%\input{main-article}
 
@@
 
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
+% LIPICS v.2021 style
+% https://submission.dagstuhl.de/documentation/authors
+\input{main-lipics}
 
@@
 %\usepackage{program} 
+\usepackage[ruled,vlined]{algorithm2e}
 
@@
 and
-\[
+\(
 \final'(p_1, q_1) \opluseq 
@@
 \otimes \final'(p_1, q_2)
-\] 
+\) 
 \marginpar{\tiny proof correctness}
-\qed
+%\qed
 \end{proof}
@@
 
-\begin{algo}[best search for \SWVPA]\label{algo:Dijkstra}
-\textbf{initially} 
-let $\Q = (Q \times Q) \cup (Q \times P \times Q)$, %contains all 
-and let 
-$d_\bot(q_1, q_2) = d_\top(q_1, p, q_2) = \one$ if $q_1 = q_2$ and
-$d_\bot(q_1, q_2) = d_\top(q_1, p, q_2) = \zero$ otherwise.
+\begin{algorithm}
+\SetKw{Extract}{extract}
+\SetKw{Update}{update}
+
+\textbf{initially} let $\Q = (Q \times Q) \cup (Q \times P \times Q)$, %contains all 
+and let $d_\bot(q_1, q_2) = d_\top(q_1, p, q_2) = \one$ if $q_1 = q_2$ and $d_\bot(q_1, q_2) = d_\top(q_1, p, q_2) = \zero$ otherwise$\;$
 
 \smallskip\noindent
-\textbf{while} $\Q$ is not empty
-
-\noindent\quad
-\textbf{extract} $\< q_1, q_2>$ or $\< q_1, p, q_2>$ from $\Q$
-
-\noindent\quad
-such that $d_\bot(q_1, q_2)$, resp. $\bigoplus_{c\in \Omegac} d_\top(q_1, p, q_2)(c)$,
-is minimal in $\Semiring$ wrt $\leq_\oplus$.
-
-\noindent\quad 
+\While{$\Q \neq \emptyset$}{
+
+%\noindent\quad
+\Extract $\< q_1, q_2>$ or $\< q_1, p, q_2>$ from $\Q$
+%
+%\noindent\quad
+such that $d_\bot(q_1, q_2)$, resp. $\bigoplus_{c\in \Omegac} d_\top(q_1, p, q_2)(c)$, is minimal in $\Semiring$ wrt $\leq_\oplus$$\;$
+
+%\noindent\quad 
 %\textbf{for all} $q_0, q_3 \in Q$ 
-\textbf{update} $d_\bot$ with $\< q_1, q_2>$ or $d_\top$ with $\< q_1, p, q_2>$
-(Figure~\ref{fig:best-update}).
-\end{algo}
+\Update $d_\bot$ with $\< q_1, q_2>$ or $d_\top$ with $\< q_1, p, q_2>$ (Figure~\ref{fig:best-update}).$\;$
+}
+\caption{Best search for \SWVPA}
+\label{algo:Dijkstra}
+\end{algorithm}
 
@@
 Next, we convert the input \SWTA $A$ over $\Omega$
-into a \SWVPA $A'$ over $\hat\Omega$, using Lemma~\ref{lem:SWTA}, 
+into a \SWVPA $A'$ over $\hat\Omega$, using Proposition~\ref{lem:SWTA}, 
 and we compute the \SWVPA $A_{T', s} \otimes A'$, 
@@
 %\bibliographystyle{plain}
-%\bibliographystyle{plainurl} 
-\bibliographystyle{abbrv}
-%\bibliographystyle{splncs04}
-%\bibliographystyle{eptcs}
+%\bibliographystyle{abbrv}
+\bibliographystyle{plainurl}  % LIPICS
+%\bibliographystyle{splncs04} % LLNCS
+%\bibliographystyle{eptcs}    % EPTCS
 %\bibliography{generic}

### 2021-06-09 introduction.tex
versions
conclusion
format LIPIcs
pass for new format

@@
 % (notion of monoid recognizability, based on registers, comparable to Bojancszik et al. data words)
-The latter case is related to a case study that motivated the present work: 
+The latter is related to a case study that motivated the present work: 
 automated music transcription.
@@
 in the form of a sequence of timed musical events, % timestamped
-is converted into a structured score in Common Western Music Notation.
-It can therefore be expressed as a parsing problem~\cite{foscarin:hal-01988990}, 
+is converted into a structured score in Common Western Music Notation~\cite{Gould11Notation}.
+It can therefore be stated as a parsing problem~\cite{foscarin:hal-01988990}, 
 over an infinite alphabet of timed events.
@@
 (see~\cite{Segoufin06csl} for a survey),
+\marginpar{\tiny register: skip refs and details, add Mikolaj}
 with pebbles for marking positions~\cite{NevenSchwentickVianu04FSMinfinite}, 
@@
 %Intuitively,~$\oplus$ selects, or ranking, the syntax trees.
-The weight domain, equipped with these two operators is assumed, at minima, 
-to form a \emph{semiring} %$\Semiring$ 
+The weight domain, equipped with these two operators shall be, at minima, 
+a \emph{semiring} %$\Semiring$ 
 where $\oplus$ can be extended to infinite sums, 
@@
 In this paper, we present a uniform framework for weighted parsing over infinite input alphabets.
-It is based on symbolic weighted  (SW) finite states language models generalizing 
-both~\SA, with functions in an arbitrary semiring instead of Boolean guards, 
-and~\WA, by handling infinite alphabets -- Figure~\ref{fig:hierarchy}.
+It is based on \emph{symbolic weighted} finite states language models (\swM), generalizing 
+both~\SA, with functions into an arbitrary semiring instead of Boolean guards, 
+and~\WA, by handling infinite alphabets, see Figure~\ref{fig:hierarchy}.
 %***
@@
 % non-terminals and production rules).
-In the transition rules of SW models, input symbols appear as variables %(parameters)
+In the transition rules of \swM models, input symbols appear as variables %(parameters)
 and the weight associated to a transition rule is a function of these variables.
@@
   \begin{array}{c} 
-  \mathsf{NFA} : \Sigma_{\mathsf{fin}}^* \to \mathbb{B}\\
+  \mathsf{FA} : \Sigma_{\mathsf{fin}}^* \to \mathbb{B}\\
   q \xrightarrow{a} q' \quad  a \in \Sigma_\mathsf{fin}
@@
 %
-\node (WA) at (-2,2) {%
+\node (WA) at (-2.5,2) {%
   \(
@@
 %
-\node (SA) at (2,2) {%
+\node (SA) at (2.5,2) {%
   \(
@@
 $\mathbb{B}$ is the Boolean algebra, 
-$\mathbb{S}$ is an arbitrary commutative semiring, 
-$q \xrightarrow{\dots} q'$ represents the form of a transition between states $q$ and $q'$.}
+$\mathbb{S}$ is a commutative semiring, 
+$q \xrightarrow{\dots} q'$ is a transition between states $q$ and $q'$.}
 \label{fig:hierarchy}  
@@
 corresponding to a linearization of trees.
-In the context of parsing, they are used here to represent (weighted) AST of CF grammars.
+In the context of parsing, they can represent (weighted) AST of CF grammars.
 More precisely, a \SWVPA $A$ associates a weight value $A(t)$ % \in \Semiring$ 
@@
 a \SWT is used to define a distance $T(s, t)$ between two finite words $s$ and $t$
-over an infinite alphabet, following~\cite{Mohri03EDWA}.
+over infinite alphabets. % following~\cite{Mohri03EDWA}.
 Then, the \emph{SW-parsing} problem aims at %computing 
 finding $t$ minimizing 
-$T(s, t) \otimes A(t)$ (\wrt the ranking defined by $\oplus$)
--- this value is called the distance between $s$ and $A$ in~\cite{Mohri03EDWA}.
+$T(s, t) \otimes A(t)$ (\wrt the ranking defined by $\oplus$).
+The latter value is called the distance between $s$ and $A$ in~\cite{Mohri03EDWA}.
 %
-Similarly to weighted-parsing 
+Like weighted-parsing 
 methods~\cite{Goodman99SemiringParsing,Nederhof03weightedParsing,MorbitzVogler19weighted-parsing}, 
 our approach proceeds in two steps, 
-based on properties of the SW models. 
+based on properties of the \swM. 
 The first step is an intersection 
-(Bar-Hillel construction~\cite{GruneJacobs08parsing} \marginpar{chap. intersection in~\cite{GruneJacobs08parsing}}) 
+(Bar-Hillel construction~\cite{GruneJacobs08parsing}) 
+\marginpar{\tiny chap. intersection in~\cite{GruneJacobs08parsing}} 
 where, given a \SWT $T$, a \SWVPA $A$, and an input word $s$, 
@@
 the keys to which are 
-($iii.a$)~the \SWT-based definition of generic edit distances between input and output words,
+($iii.a$)~the \SWT-based definition of generic edit distances between input and output (yield) words,
 and ($iii.b$)~the use in this context of nested words, and \SWVPA, 

### 2021-06-09 main.tex
pass for new format

@@
 \def\WA{\textsf{wA}\xspace}
+\def\SW{\textsf{sw}\xspace}
+\def\swM{\textsf{swM}\xspace}
 \def\SWT{\textsf{swT}\xspace}
@@
 \noindent
-For convenience, we shall sometimes present transition functions 
+We call \emph{number of transitions} of $T$ the number of pairs of states 
+$q, q' \in Q$ such that $\wei_{10}$ or $\wei_{01}$ or $\wei_{11}$
+is not the constant $\zero$.
+%
+\noindent
+For convenience, we shall sometimes present transitions 
 as functions of  
@@
 The \SWT $T$ computes on pairs of words $\< s, t> \in \Sigma^* \times \Delta^*$,
-$s$ and $t$, being respectively called input and output word.
+$s$ and $t$, being respectively called \emph{input} and \emph{output} word.
 % \noindent
@@
 based on an intermediate function $\weight_T$
-defined recursively for every states $q, q' \in Q$, 
-and every strings $\< s, t> \in \Sigma^* \times \Delta^* \setminus \{ \< \varepsilon, \varepsilon >\}$ by:
+defined recursively, for every states $q, q' \in Q$, 
+and every strings $\< s, t> \in \Sigma^* \times \Delta^* \setminus \{ \< \varepsilon, \varepsilon >\}$, by:
 %$a \in \Sigma$, $u \in \Sigma^*$, $b\in \Delta$, $v\in \Delta^*$, by:
@@
 \begin{align}
-\weight_T(q, \varepsilon, \varepsilon, q)  & = \one \label{eq:SWT-weight}\\ %\final(q)\\
-\weight_T(q, \varepsilon, \varepsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
-%\weight_T(q, au, \varepsilon, q') & = \displaystyle\bigoplus_{q'' \in Q} 
-%    \wei_{10}(q, a, \varepsilon, q'') \otimes \weight_T(q'', u, \varepsilon, q')\nonumber\\
-%\weight_T(q, \varepsilon, bv, q') & = \displaystyle\bigoplus_{q'' \in Q} 
-%    \wei_{01}(q, \varepsilon, b, q'') \otimes \weight_T(q'', \varepsilon, v, q')\nonumber\\
+%\weight_T(q, \varepsilon, \varepsilon, q)  & = \one \label{eq:SWT-weight}\\ %\final(q)\\
+%\weight_T(q, \varepsilon, \varepsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
+\weight_T(q, \varepsilon, \varepsilon, q')  & = \one 
+ \quad \mathrm{if~} q = q' \mathrm{~and~} \zero \mathrm{~otherwise} \label{eq:SWT-weight}\\ 
 \weight_T(q, s, t, q') & = \displaystyle\bigoplus_{\begin{array}{c}
-                                                   \scriptstyle q'' \in Q\\
+                                                   \scriptstyle q'' \in Q\\[-1ex]
                                                    \scriptstyle s = au,\, a \in \Sigma
@@
                     & \oplus \displaystyle\bigoplus_{\begin{array}{c}
-                                                     \scriptstyle q'' \in Q\\ 
+                                                     \scriptstyle q'' \in Q\\[-1ex]
                                                      \scriptstyle t = bv,\, b \in \Delta\\
@@
                     & \oplus \displaystyle\bigoplus_{\begin{array}{c}
-                                                     \scriptstyle q'' \in Q\\
+                                                     \scriptstyle q'' \in Q\\[-1ex]
                                                      \scriptstyle s = au,\, t = bv\\
@@
 increment the current position in the input word if and only if $i = 1$, 
-and in the output word iff $j = 1$ (otherwise, do not change it),
+and in the output word iff $j = 1$, %(otherwise, do not change it),
 and change state from $q$ to $q'$.
@@
 %
-Since $\zero$ is absorbing for~$\otimes$ in~$\Semiring$,
+Since~$\zero$ is absorbing for~$\otimes$ in~$\Semiring$,
 one term $\wei_{ij}(q, a, b, q'')$ equal to $\zero$ in the above expression 
@@
 of all possible products.
-
+%
 \medskip\noindent
@@
 %of~$T$ is its number of states $|Q|$.
+\marginpar{\tiny revise with nb of tr. and states}
 
@@
 \weiei(q, a, q') = \phi_\mathsf{i}(a) &
-\mathrm{where~} \phi_\mathsf{i} = \weiei(q, q').\\[2pt]
+\mathrm{where~} \phi_\mathsf{i} = \weiei(q, q').\\
 %
@@
 \weiec(q, {\call{c}}, p, q') = \phi_\mathsf{c}({\call{c}}) &
-\mathrm{where~} \phi_\mathsf{c} = \weiec(q, p, q').\\[2pt]
+\mathrm{where~} \phi_\mathsf{c} = \weiec(q, p, q').\\
 %
@@
 The intuition is the following for the above transitions.
-
+%
 \noindent
@@
 Moreover, $\weii$ reads a pair made of 
-${\call{c}} \in \Omegac$ and $p \in P$ at the top of the stack 
+${\call{c}} \in \Omegac$ and $p \in P$ on the top of the stack 
 ($c$ is compared to $a$ by the weight function $\eta_\mathsf{ci} \in \Phici$)
-and $\weiei$ applies if and only if the stack is empty.
-
+and $\weiei$, $\weiec$, $\weier$ apply if and only if the stack is empty.
+%
 \noindent
@@
 Moreover, $\weic$ reads ${\call{c}}$ and $p$ at the top of the stack 
-($c$ is compared to $c'$),
-and $\weiec$ applies iff the stack is empty.
-
+($c$ is compared to $c'$).
+%and $\weiec$ applies iff the stack is empty.
+%
 \noindent
@@
 Moreover, $\weir$ reads and pop from stack a pair made of $\call{c}$ and $p$, 
-($\call{c}$ is compared to $\return{r}$),
-and $\weier$ applies iff the stack is empty.
+($\call{c}$ is compared to $\return{r}$).
+%and $\weier$ applies iff the stack is empty.
 %In this case, the weight function $\phi_\mathsf{r}$ 
@@
 an intermediate function $\weight_A$, like in Section~\ref{sec:SWT}.
-In the case of a pushdown automaton, a configuration, denoted by $q[\gamma]$ 
-is composed of a state $q \in Q$ 
+%In the case of a pushdown automaton, 
+A configuration, denoted by $q[\gamma]$, 
+is here composed of a state $q \in Q$ 
 and a stack content $\gamma \in \Gamma^*$, 
@@
 \begin{align}
-%\begin{array}[t]{rcl}
-\weight_A\bigl(\config{q}{\bot}, \varepsilon, \config{q}{\bot}) & = \one\label{eq:SWVPA-weight}\\
-\weight_A\bigl(\config{q}{\bot}, \varepsilon, \config{q'}{\gamma}) & = \zero 
-\mathrm{~if~} q \neq q'\nonumber\\
+%\weight_A\bigl(\config{q}{\bot}, \varepsilon, \config{q}{\bot}) & = \one\label{eq:SWVPA-weight}\\
+%\weight_A\bigl(\config{q}{\bot}, \varepsilon, \config{q'}{\gamma}) & = \zero 
+%\mathrm{~if~} q \neq q'\nonumber\\
+\weight_A\bigl(\config{q}{\bot}, \varepsilon, \config{q'}{\bot}) & = \one
+\mathrm{~if~} q = q' \mathrm{~and~} \zero \mathrm{~otherwise}\label{eq:SWVPA-weight}\\
 \weight_A\bigl(\configup{q}{\<{\call{c}}, p>\stackup\gamma}, a\, u, 
@@
  {\displaystyle\bigoplus_{\begin{array}{c}
-                          \scriptstyle q'' \in Q\\[-2pt]
+                          \scriptstyle q'' \in Q\\[-1ex]
                           \scriptstyle p' \in P
@@
  {\displaystyle\bigoplus_{\begin{array}{c}
-                          \scriptstyle q'' \in Q\\[-2pt]
+                          \scriptstyle q'' \in Q\\[-1ex]
                           \scriptstyle p \in P
@@
                          \config{q'}{\gamma'}\bigr)\nonumber
-%\end{array}
 \end{align}
@@
 structured words with timed symbols...
+\marginpar{\tiny todo}
 intro language of music notation? (markup = time division, leaves = events etc)
@@
 
+\begin{algorithm}
+\SetKw{Extract}{extract}
+\SetKw{Update}{update}
+
+\textbf{initially} let $\Q = (Q \times Q) \cup (Q \times P \times Q)$, %contains all 
+and let $d_\bot(q_1, q_2) = d_\top(q_1, p, q_2) = \one$ if $q_1 = q_2$ and $d_\bot(q_1, q_2) = d_\top(q_1, p, q_2) = \zero$ otherwise$\;$
+
+\smallskip\noindent
+\While{$\Q \neq \emptyset$}{
+
+%\noindent\quad
+\Extract $\< q_1, q_2>$ or $\< q_1, p, q_2>$ from $\Q$
+%
+%\noindent\quad
+such that $d_\bot(q_1, q_2)$, resp. $\bigoplus_{c\in \Omegac} d_\top(q_1, p, q_2)(c)$, is minimal in $\Semiring$ wrt $\leq_\oplus$$\;$
+
+%\noindent\quad 
+%\textbf{for all} $q_0, q_3 \in Q$ 
+\Update $d_\bot$ with $\< q_1, q_2>$ or $d_\top$ with $\< q_1, p, q_2>$ (Figure~\ref{fig:best-update}).$\;$
+}
+\caption{Best search for \SWVPA}
+\label{algo:Dijkstra}
+\end{algorithm}
+
 Algorithm~\ref{algo:Dijkstra}
@@
 %$Q \times \{ \bot, \top \} \times Q$.
-
-\begin{algorithm}
-\SetKw{Extract}{extract}
-\SetKw{Update}{update}
-
-\textbf{initially} let $\Q = (Q \times Q) \cup (Q \times P \times Q)$, %contains all 
-and let $d_\bot(q_1, q_2) = d_\top(q_1, p, q_2) = \one$ if $q_1 = q_2$ and $d_\bot(q_1, q_2) = d_\top(q_1, p, q_2) = \zero$ otherwise$\;$
-
-\smallskip\noindent
-\While{$\Q \neq \emptyset$}{
-
-%\noindent\quad
-\Extract $\< q_1, q_2>$ or $\< q_1, p, q_2>$ from $\Q$
-%
-%\noindent\quad
-such that $d_\bot(q_1, q_2)$, resp. $\bigoplus_{c\in \Omegac} d_\top(q_1, p, q_2)(c)$, is minimal in $\Semiring$ wrt $\leq_\oplus$$\;$
-
-%\noindent\quad 
-%\textbf{for all} $q_0, q_3 \in Q$ 
-\Update $d_\bot$ with $\< q_1, q_2>$ or $d_\top$ with $\< q_1, p, q_2>$ (Figure~\ref{fig:best-update}).$\;$
-}
-\caption{Best search for \SWVPA}
-\label{algo:Dijkstra}
-\end{algorithm}
 
@@
 since~$\Semiring$ is complete.
+\marginpar{\tiny complete} 
 ** effectively computable by hypothese that the label theory is effective**
@@
 It can be reduced by implementing $P$ as a priority queue, 
-prioritized by the value returned by $d$
-***complete***. %$|Q|^3.\log(|Q|^2)$
+prioritized by the value returned by $d$.
+%$|Q|^3.\log(|Q|^2)$
 
@@
 $d$ is defined by:
-\begin{equation}
+%\begin{equation}
+\(
 d(s, t) = T\bigl(s, \lin(t)|_\Omegai \bigr)  \mathrm{~for~} s \in \Sigma^*, t\in \T(\Omega)
-\end{equation}
+\).
+%\end{equation}
 
@@
 for~$T$, $A$ and~$s$.
-\qed  
+%\qed  
 \end{proof}
 
-\paragraph{Application to Automated Music Transcription.}
+%\paragraph{Application to Automated Music Transcription.}
 %\input{transcription}
-...
+%...
 

### 2021-06-10 introduction.tex
revise section parsing

@@
 ($iii.a$)~the \SWT-based definition of generic edit distances between input and output (yield) words,
-and ($iii.b$)~the use in this context of nested words, and \SWVPA, 
+and ($iii.b$)~the use, convenient in this context, 
+of nested words, and \SWVPA, 
 instead of syntax trees and grammars. %(\S~\ref{sec:trees}).

### 2021-06-10 main.tex
revise section parsing

@@
 \def\final{\mathsf{out}}
-\newcommand{\call}[1]{\ensuremath #1} %{\ensuremath \langle_{#1}}
-\newcommand{\return}[1]{\ensuremath #1} %{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
 \def\Omegai{{\Omega_\mathsf{i}}}
@@
 
+\newcommand{\call}[1]{\ensuremath #1} %{\ensuremath \langle_{#1}}
+\newcommand{\return}[1]{\ensuremath #1} %{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
+\newcommand{\ccall}[1]{\ensuremath \langle_{#1}}
+\newcommand{\creturn}[1]{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
+
 %\sloppy
@@
 
-\subsection{Semirings} 
+\paragraph*{Semirings} 
 \label{section:semiring}\label{sec:semiring}
@@
 
-\subsection{Label Theory}
+\paragraph*{Label Theory}
 \label{section:symbols}
@@
 whereas the latter guards are restricted to the Boolean semiring.
+%
 
@@
 associating to every such word a weight value. 
-Nest words are able to describe structures of labeled trees, 
+Nested words are able to describe structures of labeled trees, 
 and in the context of parsing, they will be useful to 
-represent parse trees.
-
-\subsection{Definition}\label{se:SWVPA-def}
+represent AST. % parse trees.
+
+%\subsection{Definition}
+\noindent
+\label{sec:SWVPA-def}
 Let $\Omega$ be a countable alphabet 
@@
 
-\subsection{Properties}
+%\subsection{Properties}
+\medskip\noindent
 Like VPA and symbolic VPA, 
@@
 
-\subsection{Best Search} 
+%\subsection{Best Search} 
 \label{sec:best}\label{sec:search}
+\noindent
 %**hypotheses**
@@
 for a word in~$\Omega^*$.
-
-\noindent
-More precisely, 
-let $b_\bot : Q \times Q \to \Semiring$ be the function:
+%
+\noindent
+%More precisely, 
+Let $b_\bot : Q \times Q \to \Semiring$ be such that: %the function:
 %
@@
 Since $\Semiring$ is complete, the infinite sum in \eqref{eq:bbot} is well defined,
-and, roviding that $\Semiring$ is total, it is the minimum in $\Omega^*$,
+and, providing that $\Semiring$ is total, it is the minimum in $\Omega^*$,
 \wrt~$\leq_\oplus$, % this ordering.
@@
   \textstyle
-  \weight_A\bigl(\configup{q}{\< c, p> \stackup \top }, s, \configup{q'}{\<c, p> \stackup \top}\bigr). 
+  \weight_A\bigl(\configup{q}{\< c, p> \stackup \top }, s, \configup{q'}{\<c, p> \stackup \top}\bigr)
 \end{equation}
@@
 Intuitively, the function defined in \eqref{eq:btop}
-associateds to $c \in \Omegac$ 
+associates to $c \in \Omegac$ 
 the minimum weight of a computation of $A$
 starting in state $q$ with a stack 
-$c p \cdot \gamma \in \Gamma^+$ 
+$\<c, p> \cdot \gamma \in \Gamma^+$ 
 and ending in state $q'$ with the same stack,
-such that the computation does pop 
+such that the computation does not pop 
 the pair made of $c$ and $p$ at the top of this stack,
@@
 Moreover, $A$ may push another pair $\< c', p'>$ %call symbols 
-on the top of $c p \cdot \gamma$,
-following the the third case of 
+on the top of $\<c, p> \cdot \gamma$,
+following the third case of 
 in the definition \eqref{eq:SWVPA-weight} of $\weight_A$,
@@
 \textbf{initially} let $\Q = (Q \times Q) \cup (Q \times P \times Q)$, %contains all 
-and let $d_\bot(q_1, q_2) = d_\top(q_1, p, q_2) = \one$ if $q_1 = q_2$ and $d_\bot(q_1, q_2) = d_\top(q_1, p, q_2) = \zero$ otherwise$\;$
+and let $d_\bot(q_1, q_2) = d_\top(q_1, p, q_2) = \one$ 
+if $q_1 = q_2$ and $d_\bot(q_1, q_2) = d_\top(q_1, p, q_2) = \zero$ otherwise$\;$
 
@@
 and each iteration has a time complexity $O(|Q|^2 . |P|)$.
-This gives a time complexity $O(|Q|^4 . |P|)$. 
+\marginpar{\tiny detail with nb tr. and states}
+That gives a time complexity $O(|Q|^4 . |P|)$. 
 It can be reduced by implementing $P$ as a priority queue, 
@@
 %There exist $q_1, q_2 \in Q$
-\[
+\(
   {\displaystyle \bigoplus_{s \in \Omega^*} A(s)} = 
@@
   \mathsf{in}(q) \mathop{\otimes} d_\bot(q, q') \mathop{\otimes} \mathsf{out}(q').
-\]
+\)
 
@@
 $\Semiring$ commutative, bounded, total and complete, %semiring 
-and $\bar\Phi$ effective and $k$-convex, % label theory,
-one can construct in PTIME a word $s \in \Omega^*$ 
-such that $A(s)$ is minimal \wrt the natural ordering for $\Semiring$. 
+and $\bar\Phi$ effective, %and $k$-convex, % label theory,
+one can construct in PTIME a word $t \in \Omega^*$ 
+such that $A(t)$ is minimal \wrt the natural ordering for $\Semiring$. 
 \end{proposition}
 
-\subsection{Nested-Words and Parse-Trees}
+\section{Symbolic Weighted Parsing}
+\label{sec:parsing}
+%% see parse-tree.tex for a definition with parse tree (instead of nested words)
+
+Let us now apply the models and results of the previous sections %in order to define 
+to the problem of parsing over infinite alphabet. %appropriate
+%
+%\subsection{Definition}
+%
+Let~$\Sigma$ be a countable input alphabet, 
+and~$\Omega = \Omegai \uplus \Omegac \uplus \Omegar$ be a countable output alphabet.
+%
+Let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a 
+commutative, bounded, and complete \marginpar{\tiny total?} semiring  
+and let $\bar\Phi$ be an effective label theory over $\Semiring$,
+containing $\Phi_\Sigma$, $\Phi_{\Sigma, \Omegai}$, as well as
+$\Phii$, $\Phic$, $\Phir$, $\Phicr$
+(following the notations of Section~\ref{sec:SWVPA-def}).
+%
+\noindent
+We assume given the following input:
+\begin{description}
+\item[--] a \SWT $T$ over $\Sigma$, $\Omegai$, $\Semiring$, and $\bar\Phi$, 
+defining a measure %between words 
+$T: \Sigma^* \times \Omegai^* \to \Semiring$,
+
+\item[--] a \SWVPA $A$ over $\Omega$, $\Semiring$, and $\bar\Phi$,
+defining a measure $A: \Omega^* \to \Semiring$,
+%a \SWVPA $A$ over $\Omega$, and $\Semiring$, defining a series of nested words
+%      $A : \Omega^* \to \Semiring$,
+\item[--] an input word $s \in \Sigma^*$.
+\end{description}
+%
+For all $s \in \Sigma^*$ and $t \in \Omega^*$, let 
+\(d(s, t) = T\bigl(s, t|_\Omegai \bigr)\), 
+where $t|_\Omegai \in \Omegai^*$ is the projection of $t$ onto~$\Omegai$,
+obtained from $t$ by removing all symbols in $\Omega \setminus \Omegai$.
+%\end{equation}
+%
+\noindent 
+\emph{Symbolic weighted parsing} is the problem, 
+given the above input, 
+to find $t \in \Omega^*$ %nested word $t \in \Omega$ 
+minimizing \( d(s, t) \otimes A(t)\)
+\wrt $\leq_\oplus$, 
+\ie s.t. %such that: %Hence, it is the problem of finding 
+%
+\begin{equation}\label{eq:distance-lang}
+d(s, t) \otimes A(t) = \displaystyle\bigoplus_{t' \in \T(\Omega)} d(s, t') \otimes A(t') 
+\end{equation}
+%
+Following the terminology of~\cite{Mohri03EDWA}, 
+$\SW$-parsing is the problem of computing  
+the distance~\eqref{eq:distance-lang} between the input~$s$ and the output weighted language of~$A$,
+and returning a witness $t$. 
+%
+Every labeled tree $t$ can be linearized into a nested word $\lin(t) \in \Omega^*$, 
+assuming \eg that $\Omegai$ contain the symbols labelling the leaves (symbols of rank $0$) 
+and $\Omegac$ and $\Omegar$ contain respectively one left and right parenthesis
+$\ccall{b}$ and $\creturn{b}$ for each symbol $b$ labelling inner nodes (symbols of rank $>0$). 
+%
+With this representation, the projection $\lin(t)|_\Omegai$ is then the 
+sequence of leaves of $t$, enumerated in a \emph{dfs}-traversal. 
+%
+We show in Appendix~\ref{sec:trees} how to convert a ($\SW$) tree automaton~$A$
+into a $\SWVPA$ computing $A(\lin(t))$ for every tree $t$. 
+That also holds, for the set of ASTs of a weighted CF-grammar.
+%The purpose of the transducer $T$ is to measure a distance between input and output words, 
+%\ie between the word $s$ and the sequence leaves of a derivation tree, in $\Omegai^*$.
+%
+%that generates (weighted) trees by replacement of a state symbol~$q_0$ (non-terminal), 
+%by a tree $a(q_1,\ldots, q_k)$, where $k = \rank(a)$.
+%A replacement rule $q_0 \to a(q_1,\ldots, q_k)$, 
+%of weight $\wei(q_0, a, q_1 \ldots q_k) \in \Semiring$ 
+%according to Definition~\ref{def:SWTA}, 
+%corresponds to the production rule $q_0 := a(q_1,\ldots, q_k)$ of a weighted CF grammar,
+%with set non-terminal symbols $Q$ and set of terminal symbols $\Omega_0$.
+%
+%This actually is a slight generalization of CFG since 
+%each such production rule is labelled by  a symbol of $\Omega_{>0}$, 
+%hence parse trees %derivation trees
+%are trees of $\T(\Omega)$.
+%
+%Another (more original) generalization is that the set of terminal symbols 
+%$\Omega_0$ may be infinite.
+%
+%
+%The input language can also be expressed as a \SWTA, or, 
+%as a particular case, as a weighted context-free grammar, 
+%converted in turn into a \SWVPA following Lemma~\ref{lem:SWTA}.
+%
+Therefore, $\SW$-parsing generalizes 
+the problem of searching, the best derivation of a weighted CF-grammar  
+that yields a given input, sometimes referred as~\emph{weighted parsing},
+%corresponds to $\SW$-parsing in the case of finite alphabets
+%and when the transducer $T$ characterizes identity
+see \eg~\cite{Goodman99SemiringParsing} 
+and~\cite{MorbitzVogler19weighted-parsing} for a more general weighted parsing framework.
+%
+%weighted parsing and convention parsing 
+The latter indeed corresponds to the 
+particular case where the alphabet is finite, 
+$T$ computes identity and $A$ is obtained from the weighted CF grammar.
+%Indeed, it corresponds to the case where $T$
+%accepts only the pairs $\<s, t>$ such that 
+%$s$ is the projection of $t$ on $\Omegai$. 
+%This can be done with a single state $q$ and 
+%with transition rules of the form:
+%\begin{description}
+%\item[] $\wei(q, \varepsilon, a, q) = \one$ for all $a \in \Omegac \cup \Omegar$,
+%\item[] $\wei(q, a, a, q) = \one$ for all $a \in \Omegai$,
+%\item[] $\wei(q, a, b, q) = \zero$ for all $a, b \in \Omegai$, $a \neq b$.
+%\end{description}
+
+%\subsection{Computation}
+%
+\begin{proposition}
+The problem of Symbolic Weighted  parsing 
+can be solved in PTIME in the size of the input \SWT $T$, \SWVPA $A$
+and input word $s$, 
+and the computation time of the functions of the label theory.
+\end{proposition}
+\begin{proof} (sketch)
+We follow a \emph{Bar-Hillel} construction, also called parsing by intersection.
+
+\noindent
+We first extend the \SWT $T$ over $\Sigma$, $\Omegai$, $\Semiring$, and $\bar\Phi$, 
+into a \SWT $T'$ over $\Sigma$ and $\Omega$ (and the same semiring and label theory),
+such that for all $s \in \Sigma^*$, and $u \in {\Omega}^*$, 
+$T'(s, u) = T(s, u|_{\Omegai})$. The transducer $T'$ simply skips every symbol 
+$b \in {\Omega} \setminus \Omegai$, 
+by the addition to the transition of $T$,
+of new transitions of the form $\wei_{01}(q, \varepsilon, b, q')$.
+%
+\noindent
+Then, given an input word $s \in \Sigma^*$, 
+using Corolary~\ref{cor:epsilon}, 
+we compute the \SWA $B_{T', s}$, 
+such that for all $t \in \Omega^*$, $B_{T', s}(t) = d(s, t)$.
+%
+%This automaton is such that for all $t \in \T(\Omega)$, 
+%\[ 
+%   A_{T', s}\bigl(\lin(t)\bigr) 
+% = T'\bigl(s, \lin(t)\bigr) 
+% = T'\bigl(s, \lin(t)|_{\Omegai}\bigr) 
+% = d(s, t).
+%\]
+
+\noindent
+Next, %we convert the input \SWTA $A$ over $\Omega$
+%into a \SWVPA $A'$ over $\hat\Omega$, using Proposition~\ref{lem:SWTA}, and 
+we compute the \SWVPA $B_{T', s} \otimes A$, 
+using Proposition~\ref{prop:SWVPA-product}.
+%
+\noindent
+It remains to compute a best nested-word $w \in {\Omega}^*$ 
+using the best-search procedure of Proposition~\ref{th:best-search}.
+%and convert it into a best tree in $\T(\Omega)$ in order to solve SW parsing
+%for~$T$, $A$ and~$s$.
+%\qed  
+\end{proof}
+
+%\paragraph{Application to Automated Music Transcription.}
+%\input{transcription}
+%...
+
+\section*{Conclusion}
+% summary 
+We have introduced weighted language models (SW transducers and visibly pushdown automata)
+computing over infinite alphabets, 
+and applied them to the problem of parsing 
+with infinitely many possible input symbols (typically timed events).
+%handled with suitable language formalisms, 
+%
+This approach extends conventional parsing and weighted parsing
+by computing a derivation tree modulo 
+a generic distance between words,  
+defined by a SW transducer given in input.
+This enables to consider finer word relationships than strict equality, 
+%as in the conventional parsing approach, 
+opening possibilities of quantitative analysis via this method.
+
+% discussion
+\noindent
+Ongoing and future work include
+
+\noindent
+-- The study of other theoretical properties of SW models, 
+such as the extension of the best search algorithm from $1$-best to $n$-best~\cite{Huang05kbest}, 
+and to $k$-\emph{closed} semirings~\cite{Mohri02semiring}
+(instead of \emph{bounded}, which corresponds to $0$-\emph{closed}).
+
+\noindent
+-- ...there is room to improve the complexity bounds for the algorithms
+... modular approach with oracles ...
+
+\noindent
+-- present here an offline algorithm for best search, 
+semi-online implementation for AMT (bar-by-bar approach)
+with an on-the-fly automata construction.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% BIBLIO                                                                     %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\bibliographystyle{plain}
+%\bibliographystyle{abbrv}
+\bibliographystyle{plainurl}  % LIPICS
+%\bibliographystyle{splncs04} % LLNCS
+%\bibliographystyle{eptcs}    % EPTCS
+%\bibliography{generic}
+
+\bibliography{references}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% APPENDIX                                                                   %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newpage
+\appendix
+
+\section{Nested-Words and Parse-Trees}
 \label{sec:trees}
@@
 a label theory $\hat{\Phi}$ 
-like in Section~\ref{se:SWVPA-def}, 
+like in Section~\ref{sec:SWVPA-def}, 
 %
@@
 %sequences in transition with a non-null weight.
-\qed
-\end{proof}
-
-\section{Symbolic Weighted Parsing}
-\label{sec:parsing}
-Let us now apply the models and results of the previous sections %in order to define 
-to the problem of parsing over infinite alphabet. %appropriate
-%
-%\subsection{Definition}
-%
-Let~$\Sigma$ be a countable input alphabet 
-and~$\Omega$ be a countable output ranked alphabet, 
-with maximal rank value~$M$, 
-and let $\hat\Omega = \Omegai \uplus \Omegac \uplus \Omegar$
-be the alphabet with nesting symbols associated to $\Omega$ like in Section~\ref{sec:trees},
-for the linearization of trees of $\T(\Omega)$.
-%(remember that $\Omegai = \Omega_0$).
-%
-Let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a 
-commutative, bounded, and complete \marginpar{\tiny total?} semiring  
-and let $\bar\Phi$ be an effective label theory over $\Semiring$,
-containing $\Phi_\Sigma$, $\Phi_{\Sigma, \Omegai}$, as well as
-$\Phii$, $\Phic$, $\Phir$, $\Phicr$
-(following the notations of Section~\ref{se:SWVPA-def}).
-%It is moreover assumed computable and $k$-convex for some fixed~$k$.
-%
-\noindent
-We assume given the following input:
-\begin{description}
-\item[--] a \SWT $T$ over $\Sigma$, $\Omegai$, $\Semiring$, and $\bar\Phi$, 
-defining a measure %between words 
-$T: \Sigma^* \times \Omegai^* \to \Semiring$,
-
-\item[--] a \SWTA $A$ over $\Omega$, $\Semiring$, and $\bar\Phi$,
-defining a measure $A: \T(\Omega) \to \Semiring$,
-%a \SWVPA $A$ over $\Omega$, and $\Semiring$, defining a series of nested words
-%      $A : \Omega^* \to \Semiring$,
-\item[--] an input word $s \in \Sigma^*$.
-\end{description}
-%
-As explained in Section~\ref{sec:trees}, 
-$A = \< Q, \init, \bar{\wei} >$ can be seen as a weighted regular tree grammar, 
-or a labeled weighted CF grammar whose derivation trees are in $\T(\Omega)$, 
-with weight defined by $A$.
-The purpose of the transducer $T$ is to measure a distance between input and output words, 
-\ie between the word $s$ and the sequence leaves of a derivation tree, in $\Omegai^*$.
-%that generates (weighted) trees by replacement of a state symbol~$q_0$ (non-terminal), 
-%by a tree $a(q_1,\ldots, q_k)$, where $k = \rank(a)$.
-%A replacement rule $q_0 \to a(q_1,\ldots, q_k)$, 
-%of weight $\wei(q_0, a, q_1 \ldots q_k) \in \Semiring$ 
-%according to Definition~\ref{def:SWTA}, 
-%corresponds to the production rule $q_0 := a(q_1,\ldots, q_k)$ of a weighted CF grammar,
-%with set non-terminal symbols $Q$ and set of terminal symbols $\Omega_0$.
-%
-%This actually is a slight generalization of CFG since 
-%each such production rule is labelled by  a symbol of $\Omega_{>0}$, 
-%hence parse trees %derivation trees
-%are trees of $\T(\Omega)$.
-%
-%Another (more original) generalization is that the set of terminal symbols 
-%$\Omega_0$ may be infinite.
-
-\noindent 
-We extend the measure defined by $T$ to 
-$d : \Sigma^* \times \T(\Omega) \to \Semiring$ as follows.
-Given a word $w \in {\hat\Omega}^*$, the projection of $w$ onto $\Omegai$,
-denoted $w|_\Omegai$,
-is the word of $\Omegai^*$ obtained from $w$ by removing all symbols 
-in $\hat\Omega \setminus \Omegai$.
-Using this notation
-and the tree linearization operator defined in Section~\ref{sec:trees}, 
-$d$ is defined by:
-%\begin{equation}
-\(
-d(s, t) = T\bigl(s, \lin(t)|_\Omegai \bigr)  \mathrm{~for~} s \in \Sigma^*, t\in \T(\Omega)
-\).
-%\end{equation}
-
-\noindent 
-\emph{Symbolic weighted parsing} is the problem, 
-given the above input, 
-to find a tree $t \in \T(\Omega)$ %nested word $t \in \Omega$ 
-minimizing \( d(s, t) \otimes A(t)\)
-\wrt $\leq_\oplus$, 
-\ie such that: %Hence, it is the problem of finding 
-%
-\begin{equation}\label{eq:distance-lang}
-d(s, t) \otimes A(t) = \displaystyle\bigoplus_{t' \in \T(\Omega)} d(s, t') \otimes A(t') 
-\end{equation}
-%
-The measure expressed in~\eqref{eq:distance-lang} is called
-called the edit-distance between~$s$ and~$A$ in~\cite{Mohri03EDWA}.
-%
-%The input language can also be expressed as a \SWTA, or, 
-%as a particular case, as a weighted context-free grammar, 
-%converted in turn into a \SWVPA following Lemma~\ref{lem:SWTA}.
-%
-The problem of searching, in a WTA language, 
-the best parse tree matching a given input,  
-sometimes referred as~\emph{weighted parsing},
-corresponds to SW parsing in the case of finite alphabets
-and when the transducer $T$ characterizes identity
-see \eg~\cite{Goodman99SemiringParsing} 
-and~\cite{MorbitzVogler19weighted-parsing} for a more general 
-weighted parsing framework.
-%
-%Indeed, it corresponds to the case where $T$
-%accepts only the pairs $\<s, t>$ such that 
-%$s$ is the projection of $t$ on $\Omegai$. 
-%This can be done with a single state $q$ and 
-%with transition rules of the form:
-%\begin{description}
-%\item[] $\wei(q, \varepsilon, a, q) = \one$ for all $a \in \Omegac \cup \Omegar$,
-%\item[] $\wei(q, a, a, q) = \one$ for all $a \in \Omegai$,
-%\item[] $\wei(q, a, b, q) = \zero$ for all $a, b \in \Omegai$, $a \neq b$.
-%\end{description}
-
-%\subsection{Computation}
-%
-\begin{proposition}
-The problem of Symbolic Weighted  parsing 
-can be solved in PTIME in the size of the input \SWT $T$, \SWTA $A$ %\SWVPA (or ) 
-and input word $s$, 
-and the computation time of the functions of the label theory.
-\end{proposition}
-\begin{proof} (sketch)
-We follow a \emph{Bar-Hillel} construction, also called parsing by intersection.
-
-\noindent
-We first extend the \SWT $T$ over $\Sigma$, $\Omegai$, $\Semiring$, and $\bar\Phi$, 
-into a \SWT $T'$ over $\Sigma$ and $\hat\Omega$ (and the same semiring and label theory),
-such that for all $s \in \Sigma^*$, and $u \in {\hat\Omega}^*$, 
-$T'(s, u) = T(s, u|_{\Omegai})$. The transducer $T'$ simply skips every symbol 
-$b \in {\hat\Omega} \setminus \Omegai$, 
-by the addition to the transition of $T$,
-of new transitions of the form $\wei_{01}(q, \varepsilon, b, q')$.
-
-\noindent
-Then, given an input word $s \in \Sigma^*$, 
-we compute the \SWA $A_{T', s}$, using Proposition~\ref{prop:epsilon}.
-This automaton is such that for all $t \in \T(\Omega)$, 
-\[ 
-   A_{T', s}\bigl(\lin(t)\bigr) 
- = T'\bigl(s, \lin(t)\bigr) 
- = T'\bigl(s, \lin(t)|_{\Omegai}\bigr) 
- = d(s, t).
-\]
-
-\noindent
-Next, we convert the input \SWTA $A$ over $\Omega$
-into a \SWVPA $A'$ over $\hat\Omega$, using Proposition~\ref{lem:SWTA}, 
-and we compute the \SWVPA $A_{T', s} \otimes A'$, 
-using Proposition~\ref{prop:SWVPA-product}.
-
-\noindent
-It remains to compute a best nested-word $w \in {\hat\Omega}^*$ 
-using the best-search procedure of Proposition~\ref{th:best-search},
-and convert it into a best tree in $\T(\Omega)$ in order to solve SW parsing
-for~$T$, $A$ and~$s$.
 %\qed
 \end{proof}
-
-%\paragraph{Application to Automated Music Transcription.}
-%\input{transcription}
-%...
-
-\section*{Conclusion}
-% summary 
-We have introduced weighted language models (SW transducers and visibly pushdown automata)
-computing over infinite alphabets, 
-and applied them to the problem of parsing 
-with infinitely many possible input symbols (typically timed events).
-%handled with suitable language formalisms, 
-%
-This approach extends conventional parsing and weighted parsing
-by computing a derivation tree modulo 
-a generic distance between words,  
-defined by a SW transducer given in input.
-This enables to consider finer word relationships than strict equality, 
-%as in the conventional parsing approach, 
-opening possibilities of quantitative analysis via this method.
-
-% discussion
-\noindent
-Ongoing and future work include
-
-\noindent
--- The study of other theoretical properties of SW models, 
-such as the extension of the best search algorithm from $1$-best to $n$-best~\cite{Huang05kbest}, 
-and to $k$-\emph{closed} semirings~\cite{Mohri02semiring}
-(instead of \emph{bounded}, which corresponds to $0$-\emph{closed}).
-
-\noindent
--- ...there is room to improve the complexity bounds for the algorithms
-... modular approach with oracles ...
-
-\noindent
--- present here an offline algorithm for best search, 
-semi-online implementation for AMT (bar-by-bar approach)
-with an on-the-fly automata construction.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% BIBLIO                                                                     %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\bibliographystyle{plain}
-%\bibliographystyle{abbrv}
-\bibliographystyle{plainurl}  % LIPICS
-%\bibliographystyle{splncs04} % LLNCS
-%\bibliographystyle{eptcs}    % EPTCS
-%\bibliography{generic}
-
-\bibliography{references}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% APPENDIX                                                                   %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\appendix
 

### 2021-06-10 parse-tree.tex
revise section parsing

--- 2021-05-24 main-V2.tex
@@
-% fork of main.tex with more expressive models
-% SWT read both tapes (under 2 heads) at every transition (even epsilon or semi-epsilon)
-% SWA read tape (under head) at every transition (even epsilon)
-% SW-VPA read top of stack (when not empty) at every transition
+Let $\Omega$ be a ranked alphabet
+with maximal rank value~$M$, 
+and let $\hat\Omega$
+be the alphabet with nesting symbols associated to $\Omega$ like in Section~\ref{sec:trees},
+for the linearization of trees of $\T(\Omega)$.
+%(remember that $\Omegai = \Omega_0$).
+%...
+%As explained in Section~\ref{sec:trees}, 
+%$A = \< Q, \init, \bar{\wei} >$ can be seen as a weighted regular tree grammar, 
+%or a labeled weighted CF grammar whose derivation trees are in $\T(\Omega)$, 
+%with weight defined by $A$.
+%...
 
-\documentclass[runningheads]{llncs}
-%\documentclass[a4paper,11pt]{article}
-%\setcounter{page}{1}
-
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-
-\usepackage{hyperref}
-%\usepackage[bookmarks,bookmarksnumbered,naturalnames,plainpages=false]{hyperref}
-%usepackage{url}
-
-% for footnote ref
-\usepackage{refcount}
-
-% array and tabular
-\usepackage{array}
-\newcolumntype{C}[1]{>{\centering\arraybackslash\hspace{0pt}}p{#1}}
-
-% extension of enumerate env. (style for displaying counters)
-% \usepackage{enumerate}
-
-%% pictures
-% \usepackage{graphicx} 
-% \DeclareGraphicsExtensions{.pdf,.png,.jpg}
-% \graphicspath{fig/}
-
-%% PGF, Tikz
-\usepackage{tikz-cd}
-%% \usepackage{pgfplots}
-%% \usepgfplotslibrary{dateplot}
-%% \usepackage{pgf,pgfarrows,pgfnodes, pgfautomata}
-% \usepackage{tikz}
-% \usetikzlibrary{cd}
-%% \usetikzlibrary{arrows}
-%% \usetikzlibrary{calc}
-%% \usetikzlibrary{snakes}
-%% \usetikzlibrary{backgrounds}
-% \usetikzlibrary{trees}
-%% \usetikzlibrary{automata}
-%% \usetikzlibrary{positioning}
-%% \usetikzlibrary{matrix}
-%% \usetikzlibrary{patterns}
-%% \usetikzlibrary{shapes}
-
-% symbols
-\usepackage{amsmath} 
-\usepackage{amssymb} 
-\usepackage{amsbsy}
-\usepackage{bbold}
-\usepackage{latexsym}
-%\usepackage{amsfonts}
-\usepackage{stmaryrd}
-%\usepackage{mathabx}
-%\usepackage{MnSymbol}
-\usepackage{harmony} % simple music fonts
-\usepackage{mathtools} % for arrows
-%\usepackage{mathptmx}
-
-%% theorem environments
-\usepackage{theorem}
-% \newtheorem{theorem}{Theorem} %[section]
-% \newtheorem{definition}[theorem]{Definition}
-% \newtheorem{lemma}[theorem]{Lemma}
-% \newtheorem{corollary}[theorem]{Corollary}
-% \newtheorem{proposition}[theorem]{Proposition}
-% \newenvironment{proof}{\vspace{-2ex}{\it Proof. }}{\hspace*{\fill} $\Box$\smallskip }
-% \theorembodyfont{\slshape}
-% \newtheorem{example}[theorem]{Example}
-% \newtheorem{remark}[theorem]{Remark}
-\def\endex{\hspace*{\fill} $\Diamond$\smallskip }
-{\theorembodyfont{\rmfamily} \theoremstyle{break} \newtheorem{algo}{Algorithm}}
-
-%% algorithms
-%\usepackage{algorithm} 
-%\usepackage{program}
-
-%% for new macros
-\usepackage{xspace}
-
-%% arrows etc
-%\input{rewriting}
-
-%% music symbols
-% see http://tug.ctan.org/info/latex4musicians/latex4musicians.pdf
-\usepackage{musicography}
-
-%% Misc macros
-
-\def\ie{\textit{i.e.}\xspace}
-\def\eg{\textit{e.g.}\xspace}
-\def\wrt{\textit{wrt}\xspace}
-%\def\wlog{\textit{wlog}\xspace}
-\def\etc{\textit{etc}\xspace}
-
-\def\<#1>{\langle #1 \rangle}
-\newcommand{\pair}[2]{\langle{#1}, {#2}\rangle}
-%\newcommand{\A}{\mathcal{A}}
-%\newcommand{\B}{\mathcal{B}}
-\newcommand{\D}{\mathbb{D}}
-\newcommand{\E}{\mathbb{E}}
-%\newcommand{\P}{\mathcal{P}}
-\newcommand{\Q}{\mathcal{Q}}
-\newcommand{\R}{\mathcal{R}}
-\newcommand{\T}{\mathcal{T}}
-\newcommand{\W}{\mathbb{W}}
-\newcommand{\Semiring}{\mathbb{S}}
-\newcommand{\zero}{\mathbb{0}}
-\newcommand{\one}{\mathbb{1}}
-\newcommand{\dom}{\ensuremath{\mathit{dom}}}
-
-\def\SA{\textsf{sA}\xspace}
-\def\WA{\textsf{wA}\xspace}
-\def\SWT{\textsf{swT}\xspace}
-\def\SWA{\textsf{swA}\xspace}
-\def\SWTA{\textsf{swTA}\xspace}
-\def\SWVPA{\textsf{sw-VPA}\xspace}
-\def\weight{\mathsf{weight}}
-\def\wei{\mathsf{w}}
-\def\mei{\mathsf{m}}
-\def\init{\mathsf{in}}
-\def\final{\mathsf{out}}
-\newcommand{\call}[1]{\ensuremath #1} %{\ensuremath \langle_{#1}}
-\newcommand{\return}[1]{\ensuremath #1} %{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
-\def\Omegai{{\Omega_\mathsf{i}}}
-\def\Omegac{{\Omega_\mathsf{c}}}
-\def\Omegar{{\Omega_\mathsf{r}}}
-\def\Sigmai{{\Sigma_\mathsf{i}}}
-\def\Sigmac{{\Sigma_\mathsf{c}}}
-\def\Sigmar{{\Sigma_\mathsf{r}}}
-\def\Deltai{{\Delta_\mathsf{i}}}
-\def\Deltac{{\Delta_\mathsf{c}}}
-\def\Deltar{{\Delta_\mathsf{r}}}
-\def\Phii{{\Phi_\mathsf{i}}}
-\def\Phic{{\Phi_\mathsf{c}}}
-\def\Phir{{\Phi_\mathsf{r}}}
-\def\Phix{{\Phi_\mathsf{x}}}
-\def\Phici{{\Phi_\mathsf{ci}}}
-\def\Phicc{{\Phi_\mathsf{cc}}}
-\def\Phicr{{\Phi_\mathsf{cr}}}
-\def\Phicx{{\Phi_\mathsf{cx}}}
-\def\weii{{\wei_\mathsf{i}}}
-\def\weic{{\wei_\mathsf{c}}}
-\def\weir{{\wei_\mathsf{r}}}
-\def\weie{{\wei_\mathsf{e}}}
-\def\weiei{{\wei^\mathsf{e}_\mathsf{i}}}
-\def\weiec{{\wei^\mathsf{e}_\mathsf{c}}}
-\def\weier{{\wei^\mathsf{e}_\mathsf{r}}}
-\def\weiex{{\wei^\mathsf{e}_\mathsf{x}}}
-\newcommand{\config}[2]{\ensuremath \genfrac{[}{]}{0pt}{}{#1}{#2}} 
-\newcommand{\opluseq}{\ensuremath\mathrel{\oplus}=}
-\newcommand{\ioi}[1]{\mathsf{ioi}({#1})}
-\newcommand{\rank}{\mathsf{rk}}
-\newcommand{\lin}{\mathsf{lin}}
-
-%\sloppy
-
-% Parsing over infinite alphabet as optimal alignment computation 
-% as edit-distance between string and language
-% 
-%\title{Symbolic Weighted Parsing and Automated Music Transcription}
-%\title{Symbolic Weighted Language Models and Automated Music Transcription}
-\title{Symbolic Weighted Language Models and\\ Parsing over Infinite Alphabets}
-%\title{Weighted Visibly Pushdown Automata and Automated Music Transcription}
-\author{Florent Jacquemard}
-\institute{INRIA \& CNAM, Paris, France\\
-\email{florent.jacquemard@inria.fr}}
-
-%\titlerunning{WVPA \& AMT}
-%\titlerunning{Symbolic Weighted Parsing} % and Automated Music Transcription
-%\authorrunning{Florent Jacquemard}
-
-\date{\today}
-
-\begin{document}
-\thispagestyle{empty}
-\maketitle
-
-\begin{abstract}
-\input{abstract}
-\end{abstract}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% intro
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Introduction} \label{sec:intro}
-\input{introduction}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% prelin
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Preliminary Notions}
-\label{section:prelim}\label{sec:prelim}
-
-notations: for set $S$ : $S^*$ and $S^+$.
-interval $[i..j]$ of natural numbers
-
-\subsection{Semirings} 
-\label{section:semiring}\label{sec:semiring}
-\input{semiring}
-
-\subsection{Label Theory}
-\label{section:symbols}
-\input{labels}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SWT & SWA
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{SW Automata and Transducers}
-\label{section:transducer}\label{sec:transducer}
-\label{section:SWA}\label{sec:SWA}
-\label{section:SWT}\label{sec:SWT}
-
-We follow the approach of~\cite{Mohri03EDWA} for the computation of distances,
-between words and languages, using weighted transducers, 
-and extend it to infinite alphabets.
-% with models of symbolic weighted automata and transducers. 
-%
-The models introduced in this section generalize 
-weighted automata and transducers~\cite{Droste09handbook} 
-%over finite alphabets, see  e.g.~\cite{Mohri03EDWA}, 
-by labelling each transition with a weight function (instead of a simple weight value), 
-that takes the input and output symbols as parameters. 
-These functions are similar to the guards of symbolic automata~\cite{dAntoniVeanes17CAV,dAntoni21CACM},
-but they can return values in an generic semiring, 
-whereas the latter guards are restricted to the Boolean semiring.
-
-%\subsection{Definitions} \label{sec:SWTdef}\label{sec:SWAdef}
-%(SWT)
-
-Let $\Semiring$ be a commutative semiring, 
-$\Sigma$ and $\Delta$ be alphabets called respectively \emph{input} and \emph{output}, %{alphabets}, 
-and $\bar\Phi = \< \Phi_\Sigma, \Phi_\Delta, \Phi_{\Sigma, \Delta}>$ be an $\Semiring$-label theory.
-
-\begin{definition}
-\label{def:transducer} \label{def:SWT}
-A \emph{symbolic-weighted transducer} (\SWT)
-over $\Sigma$, $\Delta$, $\Semiring$ and $\bar\Phi$
-%the input and output alphabets~$\Sigma$ and $\Delta$ with label theory $\bar\Phi$, and the semiring $\Semiring$ 
-is a tuple
-$T = \< Q, \init, \bar{\wei}, \final >$,
-where $Q$ is a finite set of states, 
-$\mathsf{in} : Q \to \Phi_{\Sigma, \Delta}$, 
-respectively $\mathsf{out} : Q \to \Phi_{\Sigma, \Delta}$,
-are functions defining the weight for entering, 
-respectively leaving, a state, 
-and $\bar{\wei}$ is a triplet of transition functions 
-$\wei_{10}$, $\wei_{01}$,  and $\wei_{11}$
-%$\wei_{i, j}: Q \times Q \to \Phi_{\Sigma, \Delta}$, with $i, j \in \{ 0, 1\}$.
-from $Q \times Q$ into $\Phi_{\Sigma, \Delta}$.
-\end{definition}
-
-%A pair of states $\< q, q'>$ such that $\wei_{00}(q, q') \neq \zero$
-%is called an \emph{$\epsilon$-transition}.
-%
-For convenience, when $\wei_{ij}(q, q') = \eta \in \Phi_{\Sigma, \Delta}$, 
-for $q, q' \in Q$ and $i, j \in \{ 0, 1\} \setminus \{ \<0, 0> \}$,
-respectively $\mathsf{in}(q) = \eta$, $\mathsf{out}(q') = \eta$,
-we shall sometimes write, overloading the function names:
-$\wei_{ij}(q, a, b, q') = \eta(a, b)$ for $a \in \Sigma$,  $b \in \Delta$,
-respectively $\mathsf{in}(q, a, b) = \eta(a, b)$,
-$\mathsf{out}(q', a, b) = \eta(a, b)$.
+we assume in input
+(...)
+-- a \SWTA $A$ over $\Omega$, $\Semiring$, and $\bar\Phi$,
+defining a measure $A: T(\Omega) \to \Semiring$,
+(...)
 
 \noindent 
-The \SWT $T$ computes on pairs of words $\< s, t> \in \Sigma^+ \times \Delta^+$,
-$s$ and $t$, being respectively called input and output word.
+We extend the measure defined by $T$ to 
+$d : \Sigma^* \times \T(\Omega) \to \Semiring$ as follows.
+Given a word $w \in {\hat\Omega}^*$, the projection of $w$ onto $\Omegai$,
+denoted $w|_\Omegai$,
+is the word of $\Omegai^*$ obtained from $w$ by removing all symbols 
+in $\hat\Omega \setminus \Omegai$.
+Using this notation
+and the tree linearization operator defined in Section~\ref{sec:trees}, 
+$d$ is defined by:
+%\begin{equation}
+\(
+d(s, t) = T\bigl(s, \lin(t)|_\Omegai \bigr)  \mathrm{~for~} s \in \Sigma^*, t\in \T(\Omega)
+\).
+%\end{equation}
+
+\noindent 
+\emph{Symbolic weighted parsing} is the problem, 
+given the above input, 
+to find a tree $t \in \T(\Omega)$ %nested word $t \in \Omega$ 
+minimizing \( d(s, t) \otimes A(t)\)
+\wrt $\leq_\oplus$, 
+\ie such that: %Hence, it is the problem of finding 
 %
-% \noindent
-More precisely, the symbolic-weighted transducer~$T$ defines a mapping 
-from the pairs of strings of $\Sigma^+ \times \Delta^+$ 
-into~$\Semiring$,
-based on the following intermediate function $\weight_T$
-defined recursively for every $q, q' \in Q$, 
-%$a \in \Sigma$, $b \in \Delta$ 
-%for every strings of $s \in \Sigma^*$, $t \in \Delta^*$:
-$a \in \Sigma$, $u \in \Sigma^*$, $b\in \Delta$, $v\in \Delta^*$,
-
-\begin{equation}
-\begin{array}[t]{lccl}
-%\weight_T(q, \epsilon, \epsilon, q)  & = & \multicolumn{2}{l}{\one}\\ % or \final(q)
-%\weight_T(q, \epsilon, \epsilon, q') & = & \multicolumn{2}{l}{\zero \mathrm{~if~} q \neq q'}\\
-\weight_T(q, au, bv, q') & = & 
-             \displaystyle\bigoplus_{q'' \in Q, u \neq \epsilon} &
-    \wei_{10}(q, a, b, q'') \otimes \weight_T(q'', u, bv, q')\\
-  & \oplus & \displaystyle\bigoplus_{q'' \in Q, v \neq \epsilon} &
-    \wei_{01}(q, a, b, q'') \otimes \weight_T(q'', au, v, q')\\
-  & \oplus & \displaystyle\bigoplus_{q'' \in Q, u,v \neq \epsilon} &
-    \wei_{11}(q, a, b, q'') \otimes \weight_T(q'', u, v, q')\\
-  & \oplus & \displaystyle\bigoplus_{u,v = \epsilon} &
-    \wei_{11}(q, a, b, q')
-\end{array}
-\label{eq:SWT-weight}
+\begin{equation}\label{eq:distance-lang}
+d(s, t) \otimes A(t) = \displaystyle\bigoplus_{t' \in \T(\Omega)} d(s, t') \otimes A(t') 
 \end{equation}
 %
-We recall that, by convention (Section~\ref{sec:semiring}), 
-an empty sum with $\bigoplus$ is equal to~$\zero$. 
-%
-Intuitively, using a transition $\wei_{ij}(q, a, b, q')$ means for $T$:
-when reading respectively $a$ and $b$ at the current positions in the input and output words, 
-increment the current position in the input word if if $i = 1$, 
-and in the output word if $j = 1$ (otherwise, do not change it),
-and change state from $q$ to $q'$.
-%
-In contrast with the models of weighted transducers over finite alphabets~\cite{Mohri03ijfcs},
-the input and output symbols at current positions are always read by transitions, 
-even when they do not change the reading position the head's position.
-This is an important feature in the case of an infinite alphabet in 
-order to compare input and output symbols.
-%which cannot be stored in the finite memory of the transducer.
-
-Since $\zero$ is absorbing for~$\otimes$ in~$\Semiring$,
-one term $\wei_{ij}(q, a, b, q'')$ equal to $\zero$ in the above expression 
-will be ignored in the sum, meaning that there is no possible transition
-from state $q$ into state $q'$ while reading $a$ and $b$.
-This is analogous to the case of a transition's guard not satisfied by $\<a, b>$ for 
-symbolic transducers.
-
-%whereas considering the current symbol may be useful to compute a transition weight.
-%(even when it does not change the head's position, like with $\epsilon$-transitions).
-%
-%The cases $\weight_T(q, au, \epsilon, q')$ and $\weight_T(q, \epsilon, bv, q')$
-%are missing in the definition of $\weight_T$.
-%It means that $T$ must avoid configurations where it reached the end of 
-%the output word and not of the input one, or vice-versa.
-%This can be done by using $\wei_{10}$ and $\wei_{01}$
-%before reaching the end of word, and using a special state for this purpose.
-
-The expression of $\weight_T$ can be seen as a stateful definition of 
-an edit-distance between a word $s \in \Sigma^*$ and a word $t \in \Delta^*$,
-see also~\cite{Mohri03ijfcs}.
-Intuitively, 
-$\wei_{10}(q, a, b, r)$ is the cost of 
-the deletion of the symbol~$a \in \Sigma$ in~$s$, 
-$\wei{01}(q, a, b, r)$ is the cost 
-of the insertion of~$b \in \Delta$ in $t$, 
-and $\wei_{11}(q, a, b, r)$ is the cost 
-of the substitution of  $a \in \Sigma$ by~$b \in \Delta$.
-%
-The cost of a sequence of such operations transforming $s$ into $t$, 
-is the product, with $\otimes$, of the individual costs of the operations involved;
-And the distance between $s$ and $t$ is the sum, with $\oplus$,
-of all such product of costs.
-
-\medskip\noindent
-Let $\< s, t> \in \Sigma^+ \times \Delta^+$, with $s = s_1\ldots s_n$, 
-and $t = t_1\ldots t_m$. 
-The weight associated by $T$ to $\< s, t>$
-is defined as follows:
-\begin{equation}
-T(s, t)  = 
-\displaystyle\bigoplus_{q, q' \in Q} \mathsf{in}(q, s_1, t_1) 
-\mathop{\otimes} \weight_T(q, s, t, q') \mathop{\otimes} \mathsf{out}(q', s_n, t_m)
-\label{eq:SWT-value}
-\end{equation}
-
-\begin{example}
-We can define with a \SWT the computation of a similarity measure 
-between timed sequences
-similar to dynamic time warping (DTW).
-
-Let $\Semiring$  be the tropical (\emph{min-plus}) semiring of Figure~\ref{fig:semirings} and 
-let $\Sigma = \Delta = \mathbb{R}_+$ be sets of timestamps.
-We consider a \SWT with one state $q$ and transitions
-$\wei_{11}(q, d, d', q) = 
- \wei_{01}(q, d, d', q) = 
- \wei_{10}(q, d, d', q) = |d' - d|$,
-for all $d, d' \in \mathbb{R}_+$.
-% needs reading input/output symbols even by epsilon-transitions.
-The recursive definition of $\weight_T$ correspond to the dynamic programming equations of DTW
-for the computation of an optimal match between words, 
-the matching cost for two symbols being the 
-the time distance between them.
-\endex
-\end{example}
-
-\begin{example}
-% simple pointwise distance between two sequences of timestamped events **
-Let us consider the tropical (\emph{min-plus}) semiring~$\Semiring$ 
-of Figure~\ref{fig:semirings} and 
-let $\Sigma = \mathbb{R}_+$ be an input alphabet of event dates
-and $\Delta = \{ \mathsf{e}, \mathsf{-} \} \times \mathbb{R}_+$ 
-be an output alphabet of symbols with timestamps. 
-A symbol $\< \mathsf{e}, d > \in \Delta$ represents an event starting at date $d$, 
-and $\< \mathsf{-}, d >$ is a continuation of the previous event.
-%More precisely, we let 
-%$\Sigma = \{ a, - \} \times \R_+$ and 
-This example of $\Delta$ is motivated by the case of music notation, 
-where several notated events (notes) can be tied together, 
-with a \emph{tie} or a \emph{dot}
-(like in $\musQuarter\!\!\mathrel{\raisebox{-1.5mm}{$\smile$}}\!\!\musEighth$ 
- or equivalently~\musQuarterDotted),
-meaning that they will be played as a unique sounding event.
-%The timestamp of $a \in \Sigma$, denoted by $\mathsf{t}(a)$, is expressed as a rational number.
-
-We consider a \SWT with two states $q_0$ and $q_1$ whose purpose 
-is to compare a recorded performance $s \in \Sigma^*$
-with notated music sheet $t \in \Delta^*$.
-One timestamp $d_i \in \Sigma$ may corresponds 
-to one notated event $\<\mathsf{e}, d'_i> \in \Sigma$, in which case 
-the weight value computed by the \SWT is the time distance between both
-(see transitions $\wei_{11}$ below).
-%
-If $\<\mathsf{e}, d'_i>$ is followed by continuations 
-$\<\mathsf{-}, d'_{i+1}>$..., they are just skip with no cost (transitions $\wei_{01}$ or weight $\one$).
-%transitions
-\[
-\begin{array}{rclcrcl}
-\wei_{11}(q_0, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d| & \quad &
-\wei_{11}(q_1, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d|\\
-\wei_{01}(q_0, d, \< \mathsf{-}, d'>, q_0) & = & \one & &
-\wei_{01}(q_1, d, \< \mathsf{-}, d'>, q_0) & = & \one\\
-\wei_{10}(q_0, d, b, q_1) & = & \alpha & & \multicolumn{3}{l}{\mathrm{for~all~} b \in \Delta}
-\end{array}
-\]
-%
-Moreover, it may happen that the performers plays an extra note accidentally, but only once in a row. 
-This is modelled by the transition $\wei_{10}$ with an arbitrary weight value $\alpha \in \Semiring$, 
-switching from state $q_0$ (normal) to $q_1$ (error).
-The transitions in the second column below switch back to the normal state $q_0$.
-% the metric computed by the \SWT is the smallest sum of point wise distances 
-% between dates of input and output events.
-At last, we let $q_0$ be the only initial and final state, with
-$\init(q_0, d, b) = \final(q_0, d, b) = \one$, and 
-$\init(q_1, d, b) = \final(q_1, d, b) = \zero$, 
-for all $d \in \Sigma$ and $b \in \Delta$.
-\endex
-\end{example}
-
-\noindent
-The \emph{Symbolic Weighted Automata} %$A = \< Q, \init, \weight, \final >$
-%over $\Sigma$,  $\Semiring$ and $\bar\Phi$
-are defined similarly as the transducers of Definition~\ref{def:SWT}, 
-by simply omitting the output symbols.
-%
-In this case, the label theory $\bar\Phi$ can be reduced to a singleton $\< \Phi_\Sigma>$.
-%over $\Sigma$ is reduced to
-%a set $\Phi_\Sigma$ closed under~$\oplus$ and~$\otimes$.
-%
-\begin{definition} \label{def:SWA}
-A \emph{symbolic-weighted automaton} (\SWA)
-over $\Sigma$, $\Semiring$ and $\bar\Phi$
-%the input alphabet~$\Sigma$ and the commutative semiring $\Semiring$ 
-is a tuple
-$A = \< Q, \init, {\wei_1}, \final >$,
-where $Q$ is a finite set of states, 
-$\mathsf{in} : Q \to \Phi_\Sigma$, 
-respectively $\mathsf{out} : Q \to \Phi_\Sigma,$
-are functions defining the weight for entering, 
-respectively leaving, a state, 
-and ${\wei_1}$ is a transition functions 
-from $Q \times Q$ into $\Phi_{\Sigma}$.
-\end{definition}
-%      
-\noindent
-As above in the case of \SWT, 
-when $\wei_1(q, q') = \phi \in \Phi_\Sigma$, 
-respectively $\mathsf{in}(q) = \phi$, $\mathsf{out}(q') = \phi$,
-we may write $\wei_1(q, a, q')$ for $\phi(a)$, 
-respectively $\mathsf{in}(q, a) = \phi(a)$, $\mathsf{out}(q', a) = \phi(a)$.
-%$\wei_1: Q \times \Sigma \times Q \to \Semiring$, 
-The computation of $A$ on words $s \in \Sigma^+$
-is defined with an intermediate function $\weight_A$, 
-defined as follows for $q, q' \in Q$, $a \in \Sigma$, $u \in \Sigma^*$,
-%
-\begin{equation}
-\begin{array}{lccl}
-%\weight_T(q, \epsilon, \epsilon, q)  & = & \multicolumn{2}{l}{\one}\\ % or \final(q)
-%\weight_T(q, \epsilon, \epsilon, q') & = & \multicolumn{2}{l}{\zero \mathrm{~if~} q \neq q'}\\
-\weight_A(q, au, q') & = & 
-            \displaystyle\bigoplus_{r \in Q, u \neq \epsilon} &
-    \wei_{1}(q, a, r) \otimes \weight_A(r, u, q')\\
-  & \oplus & \displaystyle\bigoplus_{u = \epsilon} & 
-    \wei_{1}(q, a, q')
-\end{array}
-\label{eq:SWA-weight}
-\end{equation}
-%
-\noindent
-and the weight value associated by $A$ to 
-$s = s_1\ldots s_n \in \Sigma^+$ is: % defined as follows:
-\begin{equation}
-A(s)  = 
-\displaystyle\bigoplus_{q, q' \in Q} \mathsf{in}(q, s_1) 
-\mathop{\otimes} \weight_A(q, s, q') \mathop{\otimes} \mathsf{out}(q', s_n)
-\label{eq:SWA-value}
-\end{equation}
-
-%
-%When $\wei_\epsilon(q, q') = \zero$ for all $q, q' \in Q$, 
-%the automaton~$A$ is called \emph{without $\epsilon$-transitions}.
-
-%The \emph{summary} of a $\SWT$, resp. a $\SWA$, is ***
-
-%\subsection{Properties}
-\noindent
-The following property will be useful to the approach on 
-symbolic weighted parsing presented in Section~\ref{sec:parsing}.
-
-\begin{proposition} \label{prop:epsilon}
-Given a \SWT $T$ over $\Sigma$, $\Delta$, $\Semiring$ and $\bar\Phi$,
-and $s \in \Sigma^+$, 
-there exists an effectively constructible \SWA 
-$A_{T, s}$ over $\Delta$, $\Semiring$ and $\bar\Phi$,
-such that for all $t \in \Delta^+$, $A_{T, s}(t) = T(s, t)$.
-\end{proposition}
-%
-\begin{proof}
-Let $T = \< Q, \init, \bar{\wei}, \final >$,
-let 
-$\wei: Q \times \Sigma \times \Delta \times Q \to \Semiring$
-be the synthesized form of 
-$\bar{\wei} = \< \wei_{10}, \wei_{01}, \wei_{11}>$,
-and let $s = s_1 \ldots s_n$ with $s_1, \ldots, s_n \in \Sigma$ ($n \geq 1$).
-%$n = |s|$ be the length of $s$.
-
-The state set of $A_{T, s}$ will be $Q' = [1..n+1] \times Q$.
-Its state entering weight function is defined by, 
-for all $q \in Q$ and $b \in \Delta$,  
-$\init'(\< 1, q>, b) = \init(q, s_1, b)$ 
-and $\init'(\< i, q>, b) = \zero$ for all $1 < i \leq n+1$. 
-Its state leaving weight function is defined by, for all $q \in Q$ and $b \in \Delta$,  
-$\final'(\< n+1, q>, b) = \final(q, s_n, b)$, 
-and $\final'(\< i, q>, b) = \zero$ for all $1 \leq i <  n+1$.
-
-Every non-null transition of $A_{T, s}$ will
-simulate a sequence of transitions of $T$ which advance in 
-the input word while staying immobile in the output word, 
-and then make one step in the output word (and advance in the input word or not).
-%
-The above initial sequence correspond to $\epsilon$-transitions of automata, 
-and its total weight is computed by the following intermediate function 
-$\wei'_0 : Q' \times Q' \to \Phi_{\Sigma}$, 
-defined for all $q, q' \in Q$: % all $1 \leq i < n$ and $1 \leq k \leq n-i$, 
-by:
-\[
-\begin{array}{rcl}
-\wei'_0\bigl(\<i, q>, \< i, q>\bigr) & = & \one 
-\quad \mathrm{if~} 1 \leq i \leq n+1,\\
-%
-\wei'_0\bigl(\<i, q>, \< i, q'>\bigr) & = & \zero 
-\quad \mathrm{if~} 1 \leq i \leq n+1 \mathrm{~and~} q \neq q',\\
-%
-\wei'_0\bigl(\<i, q>, \< i+k, q'>\bigr) & = & 
-\displaystyle
-\bigoplus_{\begin{array}{c}
-           \scriptstyle q_0,\ldots, q_k \in Q\\[-2pt]
-           \scriptstyle q_0 = q, q_k = q'
-           \end{array}} 
-\bigotimes_{j=0}^{k-1} \bigl[\wei_{10}(q_j, q_{j+1})\bigr]_{s_{i+j}}
-\quad \begin{array}[t]{l}
-      \mathrm{if~} 1 \leq i < n\\
-      \mathrm{and~} 1 \leq k \leq n-i.
-      \end{array}
-\end{array}
-\]
-where $\bigl[\wei_{10}(q_j, q_{j+1})\bigr]_{s_{i+j}}$ 
-is the partial application $\eta_{s_{i+j}} \in \Phi_{\Delta}$
-for $\eta = \wei_{10}(q_j, q_{j+1}) \in \Phi_{\Sigma, \Delta}$ ($s_{i+j} \in \Sigma$).
-%
-The function $\wei'_0$ is defined thanks to the closure properties of the label theory $\bar\Phi$
-(Section~\ref{section:symbols}). 
-% closed under partial applications 
-The sum and product in its definition are finite, 
-we shall see below how to compute the first sum in polynomial time.
-%The effective computation of $\wei'_0$ does not require enumerating all the state sequences 
-%of length $k+1$ in the sum.
-%Indeed, it amounts to search short
-
-\noindent
-We define now the transition function 
-$\wei'_1: Q' \times Q' \to \Phi_\Sigma$
-of $A_{T, s}$ as follows,
-for $q, q' \in Q$,
-$1 \leq i \leq n$, and $0 \leq k \leq n-i$:
-\[
-\begin{array}{lcl}
-\wei'_1\bigl(\< i, q>, \< i, q'>\bigr) & = & \wei_{01}(q, q')\\
-%
-\wei'_1\bigl(\< i, q>, \< i+k, q'>\bigr) & = & 
-\displaystyle
-\bigoplus_{\begin{array}{c}
-           \scriptstyle q'' \in Q\\[-2pt]
-           \scriptstyle i+k < n
-           \end{array}} 
-\wei'_0\bigl(\<i, q>, \< i+k, q''>\bigr) \otimes \eta_{s_{i+k}}
-\quad \mathrm{where~} \eta = \wei_{01}(q'', q'),\\
- & \oplus & \displaystyle
-\bigoplus_{q'' \in Q} \wei'_0\bigl(\<i, q>, \< i+k-1, q''>\bigr) \otimes \eta'_{s_{i+k}}
-\quad \mathrm{where~}\eta' = \wei_{11}(q'', q'),\\
-%
-\wei'_1\bigl(\< i, q>, \< j, q'>\bigr) & = & \zero  
-\quad \mathrm{if~} j < i.
-\end{array}      
-\]
-%
-
-\noindent
-We can show that the $\SWA$ $A_{T, s} = \< Q', \init', \wei'_1, \final' >$
-has the expected property: $\forall t \in \Delta^+, A_{T, s}(t) = T(s, t)$.
-\qed
-\end{proof}
-
-\noindent
-On the quadratic computation of $\wei'_0$. **
-by a best path search in the graph with nodes $[i..i+k] \times Q$
-and edges labeled in $\Phi_\Sigma$...
-ordering = summary.
-
-\noindent
-The construction time and size for $A_{T, s}$ are $O(\| T \| . | s |)$,
-where the size $\| T \|$ of $T$ is its number of states $|Q|$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SW-VPA
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{SW Visibly Pushdown Automata}
-\label{section:SWVPA}\label{sec:SWVPA}
-The following model generalizes Symbolic VPA~\cite{dAntonyAlur14SVPDA}
-from Boolean semirings to arbitrary semiring weight domains.
-
-\subsection{Definition}
-Let $\Omega$ be a countable alphabet 
-%finite (large) or infinite,
-that we assume partitioned into three 
-subsets~$\Omegai$, $\Omegac$, $\Omegac$,
-whose elements are respectively called 
-\emph{internal}, \emph{call} and \emph{return} symbols.
-% names are  coined by application to functional program verification
-% \begin{itemize}
-% \item a set $\Omegai$ of \emph{internal symbols} denoted $a$,
-% \item a set $\Omegac$ of \emph{call symbols} denoted $\call{a}$,
-% \item a set $\Omegar$ of \emph{return symbols} denoted $\return{a}$.
-% \end{itemize}
-Let~$\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative and complete semiring and let  
-$\bar\Phi = \< \Phii, \Phic, \Phir, \Phici, \Phicc, \Phicr>$ 
-be a label theory over $\Semiring$
-%In order to simplify notations, %and following the definition of Section~\ref{section:transducer}, 
-%we shall write respectively 
-where $\Phii$, $\Phic$, $\Phir$ and~$\Phicx$ (with $\mathsf{x} \in \{ \mathsf{i}, \mathsf{c}, \mathsf{r}\}$) 
-stand respectively 
-for~$\Phi_\Omegai$, $\Phi_\Omegac$, $\Phi_\Omegar$ and~$\Phi_{\Omegac, \Omega_\mathsf{x}}$.
-%
-%Moreover, we extend this theory with a set $\Phii$ 
-%of unary functions in $\Omegai \to \Semiring$,
-%closed under $\oplus$ and $\otimes$.
-
-\begin{definition}
-A \emph{Symbolic Weighted Visibly Pushdown Automata} (\SWVPA) 
-over  $\Omega = \Omegai \uplus \Omegac \uplus \Omegar$, $\Semiring$ and $\bar\Phi$ 
-is a tuple $A = \< Q, P, \init, \bar\wei, \final >$,
-where $Q$ is a finite set of states, 
-$P$ is a finite set of stack symbols, 
-$\mathsf{in} : Q \to \Semiring$, 
-respectively $\mathsf{out} : Q \to \Semiring,$
-are functions defining the weight for entering, 
-respectively leaving, a state, 
-and $\bar\wei$ is a sextuplet composed of the transition functions:
-$\weii : Q \times P \times Q \to \Phici$,  
-$\weiei : Q \times Q \to \Phii$,  
-$\weic : Q \times P \times Q \times P \to \Phicc$,  
-$\weiec : Q \times P \times Q \to \Phic$,  
-$\weir : Q \times P \times Q \to \Phicr$,  
-$\weier : Q \times Q \to \Phir$.
-%and 
-%$\weiex : Q \times Q \to \Phix$ 
-%with $\mathsf{x} \in \{ \mathsf{i}, \mathsf{c}, \mathsf{r}\}$.
-\end{definition}
-%
-Similarly as in Section~\ref{section:transducer}, 
-we extend the above transition functions as follows
-for all $q, q' \in Q$, $p \in P$, 
-$a \in \Omegai$, 
-$\call{c} \in \Omegac$, 
-$\return{r} \in \Omegar$, 
-overloading their names: % for simplicity:
-\[
-\begin{array}{lll}
-\weii: Q \times \Omegac \times P \times \Omegai \times Q \to \Semiring & 
-\weii(q, c, p, a, q') = \eta_\mathsf{ci}(c, a) & 
-\mathrm{where~} \eta_\mathsf{ci} = \weii(q, p, q'),\\
-%
-\weiei: Q \times \Omegai \times Q \to \Semiring & 
-\weiei(q, a, q') = \phi_\mathsf{i}(a) &
-\mathrm{where~} \phi_\mathsf{i} = \weiei(q, q').\\[2pt]
-%
-\weic: Q \times \Omegac \times P \times  \Omegac \times P \times Q \to \Semiring & 
-\weic(q, c, p, \call{c'}, p', q') = \eta_\mathsf{cc}(c, \call{c'}) & 
-\mathrm{where~} \eta_\mathsf{cc} = \weic(q, p, p', q'),\\
-%
-\weiec: Q \times \Omegac \times P \times Q \to \Semiring & 
-\weiec(q, {\call{c}}, p, q') = \phi_\mathsf{c}({\call{c}}) &
-\mathrm{where~} \phi_\mathsf{c} = \weiec(q, p, q').\\[2pt]
-%
-\weir: Q \times \Omegac \times P \times \Omegar \times Q \to \Semiring & 
-\weir(q, {\call{c}},  p, {\return{r}}, q') = \eta_\mathsf{cr}({\call{c}},  {\return{r}}) & 
-\mathrm{where~} \eta_\mathsf{cr} = \weir(q, p, q'),\\
-%
-\weier: Q \times \Omegar \times Q \to \Semiring & 
-\weier(q, {\return{r}}, q') = \phi_\mathsf{r}({\return{r}}) &
-\mathrm{where~} \phi_\mathsf{r} = \weier(q, q').\\
-\end{array}      
-\]
-
-\noindent
-The intuition is the following for the above transitions.
-
-\noindent
-$\weii$ and $\weiei$ both read an input internal symbol $a$ and change state from $q$ to $q'$, 
-without changing the stack. 
-Moreover, $\weii$ reads a pair made of 
-${\call{c}} \in \Omegac$ and $p \in P$ at the top of the stack 
-($c$ is compared to $a$ by the weight function $\eta_\mathsf{ci} \in \Phici$)
-and $\weiei$ applies if and only if the stack is empty.
-
-\noindent
-$\weic$ and $\weiec$ read the input call symbol $\call{c}'$, 
-push it to the stack along with $p'$, and change state from $q$ to to $q'$.
-Moreover, $\weic$ reads ${\call{c}}$ and $p$ at the top of the stack 
-($c$ is compared to $c'$),
-and $\weiec$ applies iff the stack is empty.
-
-\noindent
-$\weir$ and $\weier$ read the input return symbol $\return{r}$, and change state from $q$ to to $q'$.
-Moreover, $\weir$ reads and pop from stack a pair made of $\call{c}$ and $p$, 
-($\call{c}$ is compared to $\return{r}$),
-and $\weier$ applies iff the stack is empty.
-%In this case, the weight function $\phi_\mathsf{r}$ 
-%computes a value of matching between the call and return symbols $c$ and $r$.
-%This value might be set to $\zero$ in order to express that the symbols do not match.
-
-Formally, the transitions of the automaton~$A$ are defined
-in term of %a weight value computed by 
-an intermediate function $\weight_A$, like in Section~\ref{sec:SWT}.
-In the case of a pushdown automaton, a configuration is composed 
-of a state $q \in Q$ and a stack content $\gamma \in \Gamma^*$, where $\Gamma = \Omegac \times P$.
-Hence, $\weight_A$ is a function from 
-$[Q \times \Gamma^*] \times \Omega^* \times [Q \times \Gamma^*]$ into~$\Semiring$.
-%
-\begin{equation}
-\begin{array}[t]{rcl}
-\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot\gamma}, a\, u, \config{q'}{\gamma'}\bigr) & = & 
- {\displaystyle\bigoplus_{q'' \in Q}} \weii(q, c, p, a, q'') 
- \otimes \weight_A\bigl(\config{q''}{\gamma}, u, \config{q'}{\gamma'}\bigr)\\
-\weight_A\bigl(\config{q}{\bot}, a\, u, \config{q'}{\gamma'}\bigr) & = & 
- {\displaystyle\bigoplus_{q'' \in Q}} \weiei(q, a, q'') 
- \otimes \weight_A\bigl(\config{q''}{\bot}, u, \config{q'}{\gamma'}\bigr)\\
-%
-\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot\gamma}, {\call{c}'}\, u, \config{q'}{\gamma'}\bigr) & = & 
- {\displaystyle\bigoplus_{\begin{array}{c}
-                          \scriptstyle q'' \in Q\\[-2pt]
-                          \scriptstyle p \in P
-                          \end{array}}}
- \weic\bigl(q, {\call{c}}, p, {\call{c}'}, p', q''\bigr) 
- \otimes \weight_A\bigl(\config{q''}{{\call{c}'}\, p' \cdot {\call{c}}\, p\cdot \gamma}, u, \config{q'}{\gamma'}\bigr)\\[1mm]
-%
-\weight_A\bigl(\config{q}{\bot}, {\call{c}}\, u, \config{q'}{\gamma'}\bigr) & = & 
- {\displaystyle\bigoplus_{q'' \in Q}} \weiec(q, {\call{c}}, p, q'') 
- \otimes \weight_A\bigl(\config{q''}{{\call{c}}\, p}, u, \config{q'}{\gamma'}\bigr)\\
-%
-\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot \gamma}, {\return{r}}\, u, \config{q'}{\gamma'}\bigr) & = & 
- {\displaystyle\bigoplus_{q'' \in Q}} 
- \weir\bigl(q, {\call{c}}, p, {\return{r}}, q''\bigr) 
- \otimes \weight_A\bigl(\config{q''}{\gamma}, u, \config{q'}{\gamma'}\bigr)\\
-%
-\weight_A\bigl(\config{q}{\bot}, {\return{r}}\, u, \config{q'}{\gamma'}\bigr) & = & 
- {\displaystyle\bigoplus_{q'' \in Q}} \weier(q, {\return{r}}, q'') 
- \otimes \weight_A\bigl(\config{q''}{\bot}, u, \config{q'}{\gamma'}\bigr)\\
-\end{array}
-\label{eq:SWVPA-weight}
-\end{equation}
-%
-where $\bot$ denotes the empty stack and ${\call{c}}\, p\cdot \gamma$ 
-denotes a stack where the pair made of ${\call{c}} \in \Omegac$ and $p \in P$ is the top symbol 
-and $\gamma$ is the rest of stack.
-
-\noindent
-The weight associated by $A$ to $s \in \Omega^*$
-is defined according to empty stack semantics: 
-%
-\begin{equation}
-A(s)  = 
-{\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
-\mathsf{in}(q) \mathop{\otimes} 
-\weight_A\bigl(\config{q}{\bot}, s, \config{q'}{\bot}\bigr) 
-\mathop{\otimes} \mathsf{out}(q').
-\label{eq:weightA}
-\end{equation}
-
-\begin{example}
-structured words with timed symbols...
-intro language of music notation? (markup = time division, leaves = events etc)
-\end{example}
-
-\subsection{Properties}
-Like VPA and symbolic VPA, 
-the class of \SWVPA is closed under the binary operators of the underlying semiring.
-%
-\begin{proposition}
-Let $A_1$ and $A_2$ be two (\SWVPA)
-over the same $\Omega$, $\Semiring$ and $\bar\Phi$.
-There exists two $\SWVPA$ $A_1 \oplus A_2$ and $A_1 \otimes A_2$, 
-effectively constructible, 
-such that for all $s \in \Omega^*$, 
-$(A_1 \oplus A_2)(s) = A_1(s) \oplus A_2(s)$ and 
-$(A_1 \otimes A_2)(s) = A_1(s) \otimes A_2(s)$.
-\end{proposition}
-The construction is essentially the same 
-as in the case of the Boolean semiring~\cite{dAntonyAlur14SVPDA}.
-
-\subsection{Best Search} 
-\label{sec:best}\label{sec:search}
-%**hypotheses**
-Let us assume that the semiring~$\Semiring$ is
-commutative, bounded, total, and complete.
-and assume an effective label theory.
-%s.t. for all $\phi \in \Phii$, 
-%$\psi \in \Phir$, 
-%and $\eta \in \Phicr$, 
-%$\displaystyle\bigoplus_{a \in \Omegai} \phi(a)$ 
-%$\displaystyle\bigoplus_{r \in \Omegar} \phi(r)$ and 
-%$\displaystyle\bigoplus_{{\call{c}} \in \Omegac} 
-%\displaystyle\bigoplus_{{\return{r}} \in \Omegar} \eta({\call{c}}, {\return{r}})$
-%are computable...
-
-% total ? 
-% monotonic and superior writ natural ordering
-%Regarding the infinite sum operator, note that
-%$\bigoplus_{x \in \Phi_\Omega} \phi(x)$, 
-%$\bigoplus_{y \in \Phi_\Delta} \psi(y)$, and  
-%... exist and in $\Semiring$.
-
-We propose a Dijkstra algorithm computing, for a $\SWVPA$ $A$
-over~$\Omega$, $\Semiring$ and~$\bar\Phi$, 
-the minimal weight \wrt~$\leq_\oplus$, for a word in~$\Omega^*$.
-
-\noindent
-More precisely, 
-let $b_\bot : Q \times Q \to \Semiring$ be the function:
-%
-\begin{equation}\label{eq:bbot}
-  b_\bot(q, q') = \bigoplus_{s\in \Omega^*} 
-  \textstyle
-  \weight_A\bigl(\config{q}{\bot}, s, \config{q'}{\bot}\bigr)
-\end{equation}
-%
-Since $\Semiring$ is complete, the infinite sum in \eqref{eq:bbot} is well defined,
-and by totality of $\leq_\oplus$, it is the minimum in $\Omega^*$
-of the fonction 
-$s \mapsto \weight_A(\config{q}{\sigma}, s, \config{q'}{\sigma})$ 
-\wrt this ordering.
-%
-The term $\config{q}{\bot}, s, \config{q'}{\bot}$ 
-of this sum is the central expression in 
-the definition \eqref{eq:weightA} of $A(s_0)$, for the minimum $s_0$
-of the function $\weight_A$.
-
-%
-\noindent
-Let $\top$ be a fresh stack symbol which does not belong to $\Gamma$,
-and let $b_\top : Q \times P \times Q \to \Phic$ be such that,
-for every two states $q, q' \in Q$ 
-and stack symbol $p \in P$: %and $\sigma \in \{ \bot, \top \}$, 
-\begin{equation}\label{eq:btop}
-  b_\top(q, p, q') : c \mapsto \bigoplus_{s\in \Omega^*} 
-  \textstyle
-  \weight_A\bigl(\config{q}{c\,p\cdot \top }, s, \config{q'}{c\, p\cdot \top}\bigr). 
-\end{equation}
-%
-Intuitively, the function defined in \eqref{eq:btop}
-associateds to $c \in \Omegac$ 
-the minimum weight of a computation of $A$
-starting in state $q$ with a stack 
-$c p \cdot \gamma \in \Gamma^+$ 
-and ending in state $q'$ with the same stack,
-such that the computation does pop 
-the pair made of $c$ and $p$ at the top of this stack,
-but may read these symbols.
-Moreover, $A$ may push another pair $\< c', p'>$ %call symbols 
-on the top of $c p \cdot \gamma$,
-following the the third case of 
-in the definition \eqref{eq:SWVPA-weight} of $\weight_A$,
-and may pop $\< c', p'>$ later, following the fifth case of \eqref{eq:SWVPA-weight} (return symbol). 
-%However, it cannot apply one of the two last cases (return symbol and empty stack)
-%when the current stack is $\gamma$.
-%pop symbols in $\gamma$.
-% Note that having a stack reduced to such a symbol makes impossible the application of the 
-% two last cases in the definition of $\weight_A$ (return symbol and empty stack). 
-% However, it is possible to apply the two first cases 
-% (internal symbol or call symbol, with a push on the top of $\top$).
-
-Algorithm~\ref{algo:Dijkstra}
-constructs iteratively markings 
-$d_\bot : Q \times Q \to \Semiring$ and 
-$d_\top : Q \times P \times Q \to \Phic$
-%of the triplets $\<q, \sigma, q'>$ 
-%of states of $A$ by weight values in $\Semiring$, 
-that converges eventually to $b_\top$ and $b_\bot$.
-%It uses for that purpose a priority queue $P$ containing triplets of 
-%$Q \times \{ \bot, \top \} \times Q$.
-
-\begin{algo}[best search for \SWVPA]\label{algo:Dijkstra}
-\textbf{initially} 
-let $\Q = (Q \times Q) \cup (Q \times P \times Q)$, %contains all 
-and let 
-$d_\bot(q_1, q_2) = d_\top(q_1, p, q_2) = \one$ if $q_1 = q_2$ and
-$d_\bot(q_1, q_2) = d_\top(q_1, p, q_2) = \zero$ otherwise.
-
-\smallskip\noindent
-\textbf{while} $\Q$ is not empty
-
-\noindent\quad
-\textbf{extract} $\< q_1, q_2>$ or $\< q_1, p, q_2>$ from $\Q$
-
-\noindent\quad
-such that $d_\bot(q_1, q_2)$, resp. $\bigoplus_{c\in \Omegac} d_\top(q_1, p, q_2)(c)$,
-is minimal in $\Semiring$ wrt $\leq_\oplus$.
-
-\noindent\quad 
-%\textbf{for all} $q_0, q_3 \in Q$ 
-\textbf{update} $d_\bot$ with $\< q_1, q_2>$ or $d_\top$ with $\< q_1, p, q_2>$
-(Figure~\ref{fig:best-update}).
-\end{algo}
-
-\begin{figure}
-For all $q_0, q_3 \in Q$, %$p \in P$,
-\[
-\begin{array}{lcl}
-%\multicolumn{3}{l}{\mathrm{For~all~} q_0, q_3 \in Q, p \in P,}\\
-d_\top(q_1, p, q_3) & \opluseq &
-  d_\top(q_1, p, q_2) \otimes 
-  \displaystyle\bigoplus_{a \in \Omegai} \weii(q_2, p, q_3)_a\\
-\multicolumn{3}{l}{\quad
-\mathrm{where~} \weii(q_2, p, q_3)_a \in \Phic
-\mathrm{~is~the~partial~application~}
-x_c \mapsto \weii(q_2, x_c, p, a, q_3)}\\
-%
-d_\bot(q_1, p, q_3) & \opluseq &
-  d_\bot(q_1, q_2) \otimes 
-  \displaystyle\bigoplus_{a \in \Omegai} \weie(q_2, a, q_3)\\
-%     
-d_\top(q_0, p, q_3) & \opluseq &
-  \displaystyle\bigoplus_{{\call{c}'} \in \Omegac} 
-  \displaystyle\bigoplus_{{\return{r}} \in \Omegar} 
-  \weic(q_0, p, p', q_1)_{c'} \otimes 
-  d_\top(q_1, p', q_2) \otimes
-  \weir(q_2, c', p', r, q_3)\\
-\multicolumn{3}{l}{\quad
-\mathrm{where~} \weic(q_0, p, p', q_1)_{c'} \in \Phic
-\mathrm{~is~the~partial~application~}
-x_c \mapsto \weic(q_0, p, x_c, c', p', q_1)}\\[2pt]
-%
-d_\bot(q_0, q_3) & \opluseq &
-  \displaystyle\bigoplus_{{\call{c}} \in \Omegac} 
-  \displaystyle\bigoplus_{{\return{r}} \in \Omegar} 
-  \weiec(q_0, c, p, q_1) \otimes 
-   d_\top(q_1, p, q_2)(c) \otimes
-  \weir(q_2, c, p, r, q_3)\\
-%
-d_\bot(q_1, q_3) & \opluseq &
-  d_\bot(q_1, q_2) \otimes 
-  \displaystyle\bigoplus_{\return{r} \in \Omegar} \weier(q_2, {\return{r}}, q_3)\\
-%
-d_\top(q_1, p, q_3) & \opluseq & 
-  d_\top(q_1, p, q_2) \otimes d_\top(q_2, p, q_3), 
-  \mathrm{if~} \< q_2, \top, q_3> \notin P \\
-%
-d_\bot(q_1, q_3) & \opluseq & 
-  d_\bot(q_1, q_2) \otimes d_\bot(q_2, q_3), \mathrm{if~} \< q_2, \bot, q_3> \notin P \\
-\end{array}
-\]
-%$\weii(q_2, p, q_3)_a \in \Phic$, for $a \in \Omegai$,
-%is the partial application
-%$x_c \mapsto \weii(q_2, x_c, p, a, q_3)$.
-%
-%$\weic(q_0, p, p', q_1)_{c'} \in \Phic$, for $c' \in \Omegac$,
-%is the partial application
-%$x_c \mapsto \weic(q_0, p, x_c, c', p', q_1)$.
-\caption{Update $d_\bot$ with $\<q_1, q_2>$ or $d_\top$ with $\< q_1, p, q_2>$.} 
-\label{fig:best-update}
-\end{figure}
-
-\noindent
-The infinite sums in the updates of $d$ in Algorithm~\ref{algo:Dijkstra}, 
-Figure~\ref{fig:best-update}
-are well defined
-since~$\Semiring$ is complete.
-** effectively computable by hypothese that the label theory is effective**
-The algorithm performs $2.|Q|^2$ iterations until $P$ is empty, 
-and each iteration has a time complexity $O(|Q|^2 . |P|)$.
-This gives a time complexity $O(|Q|^4 . |P|)$. 
-It can be reduced by implementing $P$ as a priority queue, 
-prioritized by the value returned by $d$
-***complete***. %$|Q|^3.\log(|Q|^2)$
-
-The correctness of Algorithm~\ref{algo:Dijkstra} 
-is ensured by the invariant expressed in the following lemma.
-\begin{lemma}\label{lem:bot}
-For all $\< q_1, q_2> \notin \Q$,
-$d_\bot(q_1, q_2) =  b_\bot(q_1, q_2)$/
-\end{lemma}
-The proof is by contradiction, 
-assuming a counter-example minimal in the length of the witness word.
-
-\begin{lemma}\label{lem:top}
-For all $\< q_1, p, q_2> \notin \Q$, 
-$d_\top(q_1, p, q_2) =  b_\top(q_1, p, q_2)$,
-\end{lemma}
-
-\noindent
-For computing the minimal weight of a computation of $A$, we use the fact that,
-at the termination of Algorithm~\ref{algo:Dijkstra}, %it holds that,
-%There exist $q_1, q_2 \in Q$
-\[
-  {\displaystyle \bigoplus_{s \in \Omega^*} A(s)} = 
-  {\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
-  \mathsf{in}(q) \mathop{\otimes} d_\bot(q, q') \mathop{\otimes} \mathsf{out}(q').
-\]
-
-%\medskip
-\noindent
-In order to obtain effectively a witness 
-(word of $\Semiring^*$ with computation of $A$ of minimal weight), 
-we require an additional property the of weight functions.
-%
-\begin{definition}
-Let $\Omega$ be an alphabet and $\Semiring$ a complete semiring.
-A function $\phi$ from $\Omega^n$ into $\Semiring$
-is called $k$-\emph{convex} for a natural number $k$ iff 
-$\mathit{card}\{ \mathbf{a} \in \Omega^n \mid \phi(\mathbf{a}) = 
-                 \displaystyle\bigoplus_{\mathbf{p} \in \Omega^n} \phi(\mathbf{p}) \} \leq k$.
-\end{definition}
-A label theory is $k$-convex if all its functions are $k$-convex.
-
-\begin{proposition}\label{th:best-search}
-For a \SWVPA $A$ 
-over $\Omega$, 
-$\Semiring$ commutative, bounded, total and complete, %semiring 
-and $\bar\Phi$ $k$-convex effective, % label theory,
-one can construct in PTIME a word $s \in \Omega^*$ 
-such that $A(s)$ is minimal \wrt the natural ordering for $\Semiring$. 
-\end{proposition}
-
-\subsection{Trees and Nested-Words}
-\label{sec:trees}
-The hierarchical structure of nested-words, defined the \emph{call} and \emph{return} markup symbols of 
-suggest a correspondence between these word and trees. 
-The lifting of this correspondence to languages, respective of tree automata and VPA have
-been discussed in~\cite{AlurMadhusudan09nested}, 
-see also~\cite{Caralp12VPAmult} for the weighted case.
-In this section, we discuss the correspondence between the symbolic-weighted extensions
-of tree automata and VPA.
-
-Let $\Omega$ be a countable ranked alphabet, such that 
-every symbol $a \in \Omega$ has a rank 
-$\rank(a) \in [0..M]$ where $M$ is a fixed natural number.
-We write $\Omega_k$ the subset of all symbols $a$ of $\Omega$
-with $\rank(a) = k$, where $0 \leq k \leq M$.
-%$\cup_{i = 0}^{i = M} \Omega_i$ 
-%
-$\T(\Omega)$ denotes the free $\Omega$-algebra of finite, ordered, 
-$\Omega$-labeled trees, 
-which is the smallest set such that  $\Omega_0 \subset \T(\Omega)$
-and for all $1 \leq k \leq M$, all $a \in \Omega_k$, 
-and all $t_1, \ldots, t_k \in \T(\Omega)$, $a(t_1, \ldots, t_k) \in \T(\Omega)$.
-%
-% tree = single node (leave) labeled with a symbol of $a \in \Omegai$
-% (such a tree is simply denoted by $a$)
-% or the composition, denoted by $b(t_1,\ldots, t_n$) of a node labeled with $b$
-% and $n$ subtrees $t_1$,\ldots, $t_n$.
-
-\renewcommand{\call}[1]{\ensuremath \langle_{#1}}
-\renewcommand{\return}[1]{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
-
-Let $\hat\Omega$ be the countable (unranked) alphabet obtained from 
-$\Omega$ as follows: 
-$\hat\Omega = \< \Omegai, \Omegac, \Omegar >$, with
-$\Omegai = \Omega_0$, 
-$\Omegac = \{ \; \call{a} \mid a \in \Omega_{>0} \}$,
-$\Omegac = \{ \; \return{a} \mid a \in \Omega_{>0} \}$,
-where $\Omega_{>0}$ denotes $\bigcup_{k = 1}^{M} \Omega_k$.
-
-\noindent
-We define a linearization of trees of $\T(\Omega)$ into 
-words of $\hat{\Omega}^*$ as follows:
-\begin{description}
-\item $\lin(a) = a$ for all $a \in \Omega_0$, 
-\item $\lin\bigl( b(t_1, \ldots, t_k)\bigr) = 
-       \call{b} \; \lin(t_1) \ldots \lin(t_k) \; \return{b}$ 
-       when $b \in \Omega_k$, $1 \leq k \leq M$.
-\end{description}
-
-% A \emph{regular tree grammar} over $\Omega$ 
-% is a triplet $G = \< N, q_\mathsf{i}, R>$ where
-% $N$ is a finite set of non-terminal symbols denoted $q$..., 
-% $q_\mathsf{i} \in N$ is the starting non-terminal, 
-% $R$ is a finite set of production rules of the form
-% $q_0 \to a(q_1\ldots q_k)$ where 
-% $q_0, q_1, \ldots, q_k \in N$
-% $a \in \Omega_k$.
-% A tree $t \in \T(\Omega)$ is in the language of $G$ 
-% if it can be generated from $q_\mathsf{i}$ by 
-% non terminal replacement following the rules of $R$.
-
-Let us assume a label theory $\Phi_{\Omega_k}$ 
-for each $k \in [0..M]$.
-%
-\begin{definition}  \label{def:SWTA}
-A \emph{symbolic-weighted tree automaton} (\SWTA)
-over the ranked input alphabet~$\Omega$
-and the commutative semiring $\Semiring$ 
-is a triplet $A = \< Q, \init, \bar{\wei} >$ where
-$Q$ is a finite set of states, 
-$\mathsf{in} : Q \to \Semiring$ is the starting weight function, 
-and $\bar{\wei}$ is a $M+2$-uplet of transition functions made of:
-$\wei_\epsilon$ from $Q \times Q$ into $\Semiring$, and, 
-for each $k \in [0..M]$, $\wei_{\Omega, k}$
-from $Q \times Q^{k}$ into $\Phi_{\Omega_k}$.
-\end{definition}
-%
-Like in Section~\ref{sec:SWAdef}, we define from $\bar{\wei}$ 
-a transition function~$\wei$, %from $\bar\wei$
-from 
-$Q \times (\Sigma \cup \{ \epsilon \}) \times \bigcup_{k=0}^{M} Q^k$
-into~$\Semiring$: %also called $\wei$ for simplicity, 
-%such that, for all $q, q' \in Q$, $a \in \Sigma$, and $b \in \Delta$, 
-\[
-\begin{array}{rcll}
-\wei(q_0, \epsilon, q_1) & = &  \wei_\epsilon(q_0, q_1),\\ %\phi_\epsilon\\
-\wei(q_0, a, q_1 \ldots q_k) & = & \phi_{\Omega, k}(a) &
-\quad\mathrm{where~} \phi_{\Omega, k} = \wei_{\Omega, k}(q_0, q_1\ldots q_k).\\
-\end{array}      
-\]
-Intuitively, $\wei(q_0, a, q_1 \ldots q_k)$ can be seen as
-the weight of a production rule $q_0 \to a(q_1, \ldots, q_k)$ 
-of a regular tree grammar~\cite{tata}, 
-that replaces the non-terminal symbol $q_0$, by $a(q_1, \ldots, q_k)$. 
-%in a step of tree building.
-%
-Such a grammar computes the weights of the derivation trees 
-of the Context-Free grammar obtained by forgetting the labeling symbols of $\Omega_{>0}$.
-
-\noindent
-The \SWTA of Definition~\ref{def:SWTA} defines a mapping 
-from trees of $\T(\Omega)$ into the weights of~$\Semiring$,
-based on the intermediate function $\weight_A$
-defined as follows for $q_0 \in Q$ and 
-$t = b(t_1,\ldots, t_k) \in \T(\Omega)$,
-åwith $0 \leq k \leq M$:
-\[
-\begin{array}{rccl}
-\weight_A(q_0, t) & = & 
-   \displaystyle\bigoplus_{q_1 \in Q} &
-   \wei(q, \epsilon, q_1) \otimes \weight_A(q_1, t)\\
- & \oplus & \displaystyle\bigoplus_{q_1 \ldots q_k \in Q^k} &
-              \wei(q_0, b, q_1 \ldots q_k ) 
-   \otimes \displaystyle\bigotimes_{i=1}^{k}
-           \weight_A(q_i, t_i).\\
-\end{array}
-\]
-
-\medskip\noindent
-The weight associated by $A$ to  $t \in \T(\Omega)$ is then
-\begin{equation}
-A(t)  = 
-\displaystyle\bigoplus_{q \in Q} \mathsf{in}(q) \mathop{\otimes} \weight_A(q, t).
-\label{eq:weightTA}
-\end{equation}
-
-\begin{lemma}  \label{lem:SWTA}
-For all \SWTA $A$ over~$\Omega$ and $\Semiring$, without $\epsilon$-transitions,
-there exists an effectively constructible \SWVPA $A'$ over 
-$\hat\Omega$ and $\Semiring$
-such that for all $t \in \T(\Omega)$, $A'\bigl(\lin(t)\bigr) = A(t)$.
-\end{lemma} 
-% 
-\begin{proof}
-Let $A = \< Q, \init, \bar{\wei} >$ where $\bar{\wei}$ is summarized as above by a function
-$\wei: Q \times (\Sigma \cup \{ \epsilon \}) \times \bigcup_{k=0}^{M} Q^k \to \Semiring$.
-
-We build 
-$A' = \< Q', P', \init', \weii, \weic, \weir, \weie, \final' >$,
-computing over $\hat\Omega = \< \Omegai, \Omegac, \Omegar >$,
-%
-%$\Omega = \Omegai \uplus \Omegac \uplus \Omegar$ and $\Semiring$ 
-where $Q' = \bigcup_{k=0}^{M} Q^k$ be  the set of sequences of state symbols of $A$, 
-of length at most $M$, including the empty sequence denoted by $\epsilon$, 
-and where $P' = Q'$.
-
-\[
-\begin{array}{ll}
-%\weii: Q \times \Omegai \times Q \to \Semiring & 
-\weii(\bar{q}, a, \bar{q} q') = \wei(q', a, \epsilon) & 
-\mathrm{for~all~} a \in \Omega_0,\\
-%
-%\weic: Q \times \Omegac \times Q \times P \to \Semiring & 
-\weic(\bar{q}, \call{b}, \epsilon, \bar{q}) = \one & %\wei(p', b, \bar{p}) & 
-\mathrm{for~all~} b \in \Omega_{>0},\\
-%
-%\weir: Q \times \Omegac \times P \times \Sigmar \times Q \to \Semiring & 
-\weir(\bar{q}, {\call{b}}, \bar{p}, {\return{b}}, \bar{p} q') = \wei(q', b, \bar{q}) & 
-\mathrm{for~all~}  b \in \Omega_{>0},\\
-%
-%\weie: Q \times \Omegar \times Q \to \Semiring & 
-\weie(\bar{p}, {\return{b}}, \bar{q}) = \zero &
-\mathrm{for~all~}  b \in \Omega_{>0}.\\
-\end{array}      
-\]
-
-In practice, it is sufficient to consider in $Q'$ only the prefixes of 
-sequences.
-\qed\end{proof}
-
-\section{Symbolic Weighted Parsing}
-\label{sec:parsing}
-Let us now use the models and results of the former sections in %order to define 
-an approach to the problem of parsing over infinite alphabet. %appropriate
-%
-Besides considering infinitely many possible of input symbols, 
-handled with suitable language formalisms, 
-this approach extends conventional parsing 
-by computing a derivation tree modulo 
-a generic distance between words,  
-defined by a SW transducer given in input.
-This enables considering finer word relationships than strict equality 
-as in the conventional parsing approach, 
-opening possibilities of quantitative analysis via this method.
-
-\subsection{Definition}
-
-Let $\Sigma$ be a countable input alphabet 
-and $\Omega = \Omegai \uplus \Omegac \uplus \Omegar$ an output countable alphabet,
-let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative, bounded, complete 
-and total semiring 
-and let  $(\Phi_\epsilon, \Phic, \Phir, \Phicr)$ be a label theory over $\Semiring$, 
-assumed computable and $k$-convex for some fixed $k$.
-
-\noindent
-Assuming given in input:
-\begin{itemize}
-\item a \SWT $T$ over $\Sigma$,  $\Omega$, and $\Semiring$,
-defining a measure %between words 
-$T: \Sigma^* \times \Omega^* \to \Semiring$,
-\item a \SWVPA $A$ over $\Omega$, and $\Semiring$, defining a series of nested words
-      $A : \Omega^* \to \Semiring$,
-\item an input word $s \in \Sigma^*$,
-\end{itemize}
-the problem of \emph{symbolic weighted parsing} is to find 
-a nested word $t \in \Omega$ minimizing
-\( T(s, t) \otimes A(t)\)
-\wrt $\leq_\oplus$, 
-\ie such that %Hence, it is the problem of finding 
-\( T(s, t) \otimes A(t) = 
-  \displaystyle\bigoplus_{t' \in \Omega^*} T(s, t') \otimes A(t') \).
-
-\noindent 
-Therefore, it is the problem of optimizing a measure 
-called the \emph{edit-distance between $s$ and $A$} in~\cite{Mohri03EDWA}.
-%
-The input language can also be expressed as a \SWTA, or, 
-as a particular case, as a weighted context-free grammar, 
-converted in turn into a \SWVPA following Lemma~\ref{lem:SWTA}.
-%
-In the case of finite alphabets, 
-the problem of searching, in a WTA language, 
-for the best parse tree for a given,  
-sometimes referred as~\emph{weighted parsing}
-(see \cite{Goodman99SemiringParsing,MorbitzVogler19weighted-parsing} **more general problems**)
-is a particular case of SW parsing.
-%
-Indeed, it corresponds to the case where $T$
-accepts only the pairs $\<s, t>$ such that 
-$s$ is the projection of $t$ on $\Omegai$. 
-This can be done with a single state $q$ and 
-with transition rules of the form:
-\begin{description}
-\item[] $\wei(q, \epsilon, a, q) = \one$ for all $a \in \Omegac \cup \Omegar$,
-\item[] $\wei(q, a, a, q) = \one$ for all $a \in \Omegai$,
-\item[] $\wei(q, a, b, q) = \zero$ for all $a, b \in \Omegai$, $a \neq b$.
-\end{description}
-
-\subsection{Computation}
-
-\begin{proposition}
-The problem of Symbolic Weighted  parsing 
-can be solved in PTIME in the size of the input \SWT, \SWVPA (or \SWTA) and input word, 
-and the computation time of the functions of the label theory.
-\end{proposition}
-%
-%\begin{proof}
-%Bar-Hillel construction
-%\qed  
-%\end{proof}
-
-%\input{transcription}
-
-\section*{Conclusion}
-% discussion
-
-\noindent
--- summary
-
-\noindent
--- other theoretical properties of SW models
-
-\noindent
--- room to improve complexity for best-search algorithm
-... modular approach with oracles ...
-
-and extention to $n$-best
-
-\noindent
--- offline algorithm, 
-semi-online implementation for AMT (bar-by-bar approach)
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% BIBLIO                                                                     %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\bibliographystyle{plain}
-%\bibliographystyle{plainurl} 
-\bibliographystyle{abbrv}
-%\bibliographystyle{splncs04}
-\bibliography{references}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% APPENDIX                                                                   %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\appendix
-
-\end{document}
-
-\section{Edit-Distance}
-
-%\subsection{Distance between words or languages}
-...algebraic definition of edit-distance of Mohri, in \cite{Mohri03EDWA}
-% Mehryar Mohri 
-% Edit-distance of weighted automata: General definitions and algorithms
-% International Journal of Foundations of Computer Science 14.06 (2003): 957-982.
-distance $d$ over $\Sigma^* \times \Sigma^*$ 
-into a semiring  $\Semiring = ( \Semiring, \oplus, \zero, \otimes, \one)$.
-
-%\noindent
-Let $\Omega = \Sigma \cup \{ \epsilon \} \times \Sigma \cup \{ \epsilon \} \setminus \{ (\epsilon, \epsilon) \}$,
-and let $h$ be the morphism from $\Omega^*$ into $\Sigma^* \times \Sigma^*$  
-defined over the concatenation of strings of $\Sigma^*$ (that removes the $\epsilon$'s).
-%
-\noindent
-An \emph{alignment} between 2 strings  $s, t \in \Sigma^*$ is an element $\omega \in \Omega^*$ 
-such that $h(\omega) = (s, t)$.
-%
-\noindent
-We assume a base cost function $\Omega$ : $\delta: \Omega \to S$, extended to $\Omega^*$ as follows  
-(for $\omega \in \Omega^*$): 
-\(
-\displaystyle\delta(\omega) = \bigotimes_{0 \leq i < |\omega|} \delta(\omega_i)
-\).
-
-\noindent
-\begin{definition}
-For  $s, t \in \Sigma^*$, the edit-distance between $s$ and $t$ is  
-\( 
-d(s, t) = \displaystyle\bigoplus_{\omega \in \Omega^*\, h(\omega) = (s, t)} \delta(\omega)
-\).
-\end{definition}
-
-e.g. Levenstein edit-distance: $S$ is min-plus and $\delta(a, b) = 1$ for all $(a, b) \in \Omega$.
-
-%\paragraph{Distance between a word and a regular language}
+The measure expressed in~\eqref{eq:distance-lang} is called
+called the edit-distance between~$s$ and~$A$ in~\cite{Mohri03EDWA}.

### 2021-06-16 distance.tex
comments Masahiko

--- 2021-06-09 main.tex
@@
-% SWT without epsilon-transitions
-% SWT do not read the other tape in "skip" transitions 10 or 01
-%     (hence this version is simpler  than V2)
-% SWA without epsilon-transitions
-% SW-VPA read top of stack (when not empty) at every transition 
-%        (6 kinds of transitions).
-
-% default style article.cls
-%\input{main-article}
-
-% Springer llncs style
-% https://www.springer.com/gp/computer-science/lncs/conference-proceedings-guidelines
-%\input{main-llncs}
-
-% EPTCS style
-% http://www.eptcs.org
-%\input{main-eptcs}
-
-% LIPICS v.2021 style
-% https://submission.dagstuhl.de/documentation/authors
-\input{main-lipics}
-
-\usepackage{hyperref}
-%\usepackage[bookmarks,bookmarksnumbered,naturalnames,plainpages=false]{hyperref}
-%usepackage{url}
-
-% for footnote ref
-\usepackage{refcount}
-
-% array and tabular
-\usepackage{array}
-\newcolumntype{C}[1]{>{\centering\arraybackslash\hspace{0pt}}p{#1}}
-
-% extension of enumerate env. (style for displaying counters)
-% \usepackage{enumerate}
-
-%% pictures
-% \usepackage{graphicx} 
-% \DeclareGraphicsExtensions{.pdf,.png,.jpg}
-% \graphicspath{fig/}
-
-%% PGF, Tikz
-\usepackage{tikz-cd}
-%% \usepackage{pgfplots}
-%% \usepgfplotslibrary{dateplot}
-%% \usepackage{pgf,pgfarrows,pgfnodes, pgfautomata}
-% \usepackage{tikz}
-% \usetikzlibrary{cd}
-%% \usetikzlibrary{arrows}
-%% \usetikzlibrary{calc}
-%% \usetikzlibrary{snakes}
-%% \usetikzlibrary{backgrounds}
-% \usetikzlibrary{trees}
-%% \usetikzlibrary{automata}
-%% \usetikzlibrary{positioning}
-%% \usetikzlibrary{matrix}
-%% \usetikzlibrary{patterns}
-%% \usetikzlibrary{shapes}
-
-% symbols
-\usepackage{amsmath} 
-\allowdisplaybreaks
-\usepackage{amssymb} 
-\usepackage{amsbsy}
-\usepackage{bbold}
-\usepackage{latexsym}
-%\usepackage{amsfonts}
-\usepackage{stmaryrd}
-%\usepackage{mathabx}
-%\usepackage{MnSymbol}
-\usepackage{harmony} % simple music fonts
-\usepackage{mathtools} % for arrows
-%\usepackage{mathptmx}
-
-%% algorithms
-%\usepackage{algorithm} 
-%\usepackage{program} 
-\usepackage[ruled,vlined]{algorithm2e}
-
-%% allows for page break within arrays
-\usepackage{longtable}
-
-%% arrows etc
-%\input{rewriting}
-
-%% music symbols
-% see http://tug.ctan.org/info/latex4musicians/latex4musicians.pdf
-\usepackage{musicography}
-
-%% for new macros
-\usepackage{xspace}
-
-%% Misc macros
-\def\ie{\textit{i.e.}\xspace}
-\def\eg{\textit{e.g.}\xspace}
-\def\wrt{\textit{wrt}\xspace}
-%\def\wlog{\textit{wlog}\xspace}
-\def\etc{\textit{etc}\xspace}
-
-\def\<#1>{\langle #1 \rangle}
-\newcommand{\pair}[2]{\langle{#1}, {#2}\rangle}
-
-%\newcommand{\config}[2]{\ensuremath \genfrac{[}{]}{0pt}{}{#1}{#2}} 
-\newcommand{\config}[2]{\ensuremath{#1}[{#2}]} 
-\newcommand{\configup}[2]{\ensuremath{#1}\left[\begin{array}{c} #2 \end{array}\right]}
-\def\stacksep{\cdot}
-\def\stackup{\\}
-
-\newcommand{\opluseq}{\ensuremath\mathrel{\oplus}=}
-
-% \bigominus
-\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
-\DeclareFontShape{U}{mathx}{m}{n}{
-<-6> mathx5 <6-7> mathx6 <7-8> matha7
-<8-9> mathx8 <9-10> mathx9
-<10-12> mathx10 <12-> mathx12
-}{}
-\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
-\DeclareMathSymbol{\bigominus}{\mathop}{mathx}{"C1}
-
-%\newcommand{\A}{\mathcal{A}}
-%\newcommand{\B}{\mathcal{B}}
-\newcommand{\D}{\mathbb{D}}
-\newcommand{\E}{\mathbb{E}}
-%\newcommand{\P}{\mathcal{P}}
-\newcommand{\Q}{\mathcal{Q}}
-\newcommand{\R}{\mathcal{R}}
-\newcommand{\T}{\mathcal{T}}
-\newcommand{\W}{\mathbb{W}}
-\newcommand{\Semiring}{\mathbb{S}}
-\newcommand{\zero}{\mathbb{0}}
-\newcommand{\one}{\mathbb{1}}
-\newcommand{\dom}{\ensuremath{\mathit{dom}}}
-
-\def\SA{\textsf{sA}\xspace}
-\def\WA{\textsf{wA}\xspace}
-\def\SWT{\textsf{swT}\xspace}
-\def\SWA{\textsf{swA}\xspace}
-\def\SWTA{\textsf{swTA}\xspace}
-\def\SWVPA{\textsf{sw-VPA}\xspace}
-\def\weight{\mathsf{weight}}
-\def\wei{\mathsf{w}}
-\def\mei{\mathsf{m}}
-\def\init{\mathsf{in}}
-\def\final{\mathsf{out}}
-\newcommand{\call}[1]{\ensuremath #1} %{\ensuremath \langle_{#1}}
-\newcommand{\return}[1]{\ensuremath #1} %{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
-\def\Omegai{{\Omega_\mathsf{i}}}
-\def\Omegac{{\Omega_\mathsf{c}}}
-\def\Omegar{{\Omega_\mathsf{r}}}
-\def\Sigmai{{\Sigma_\mathsf{i}}}
-\def\Sigmac{{\Sigma_\mathsf{c}}}
-\def\Sigmar{{\Sigma_\mathsf{r}}}
-\def\Deltai{{\Delta_\mathsf{i}}}
-\def\Deltac{{\Delta_\mathsf{c}}}
-\def\Deltar{{\Delta_\mathsf{r}}}
-\def\Phii{{\Phi_\mathsf{i}}}
-\def\Phic{{\Phi_\mathsf{c}}}
-\def\Phir{{\Phi_\mathsf{r}}}
-\def\Phix{{\Phi_\mathsf{x}}}
-\def\Phici{{\Phi_\mathsf{ci}}}
-\def\Phicc{{\Phi_\mathsf{cc}}}
-\def\Phicr{{\Phi_\mathsf{cr}}}
-\def\Phicx{{\Phi_\mathsf{cx}}}
-\def\weii{{\wei_\mathsf{i}}}
-\def\weic{{\wei_\mathsf{c}}}
-\def\weir{{\wei_\mathsf{r}}}
-\newcommand{\weie[1]}{\wei_{#1}^\mathsf{e}}
-\def\weiei{\weie[\mathsf{i}]}
-\def\weiec{\weie[\mathsf{c}]}
-\def\weier{\weie[\mathsf{r}]}
-\def\weiex{\weie[\mathsf{x}]}
-\newcommand{\ioi}[1]{\mathsf{ioi}({#1})}
-\newcommand{\rank}{\mathsf{rk}}
-\newcommand{\lin}{\mathsf{lin}}
-
-%\sloppy
-
-% Parsing over infinite alphabet as optimal alignment computation 
-% as edit-distance between string and language
-% 
-%\title{Symbolic Weighted Parsing and Automated Music Transcription}
-%\title{Symbolic Weighted Language Models and Automated Music Transcription}
-\title{Symbolic Weighted Language Models and\\ Parsing over Infinite Alphabets}
-%\title{Weighted Visibly Pushdown Automata and Automated Music Transcription}
-%\titlerunning{WVPA \& AMT}
-%\titlerunning{Symbolic Weighted Parsing} % and Automated Music Transcription
-%\authorrunning{Florent Jacquemard}
-
-\date{\today}
-
-\begin{document}
-\thispagestyle{empty}
-\maketitle
-
-\begin{abstract}
-\input{abstract}
-\end{abstract}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% intro
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Introduction} \label{sec:intro}
-\input{introduction}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% prelin
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Preliminary Notions}
-\label{section:prelim}\label{sec:prelim}
-
-%notations: for set $S$ : sets of sequences $S^*$ and $S^+$...
-%interval $[i..j]$ of natural numbers
-
-\subsection{Semirings} 
-\label{section:semiring}\label{sec:semiring}
-\input{semiring}
-
-\subsection{Label Theory}
-\label{section:symbols}
-\input{labels}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SWT & SWA
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{SW Automata and Transducers}
-\label{section:transducer}\label{sec:transducer}
-\label{section:SWA}\label{sec:SWA}
-\label{section:SWT}\label{sec:SWT}
-
-We follow the approach of~\cite{Mohri03EDWA} for the computation of distances,
-between words and languages, using weighted transducers, 
-and extend it to infinite alphabets.
-% with models of symbolic weighted automata and transducers. 
-%
-The models introduced in this section generalize 
-weighted automata and transducers~\cite{Droste09handbook} 
-%over finite alphabets, see  e.g.~\cite{Mohri03EDWA}, 
-by labelling each transition with a weight function (instead of a simple weight value), 
-that takes the input and output symbols as parameters. 
-These functions are similar to the guards of symbolic automata~\cite{dAntoniVeanes17CAV,dAntoni21CACM},
-but they can return values in an generic semiring, 
-whereas the latter guards are restricted to the Boolean semiring.
-
-%\subsection{Definitions} \label{sec:SWTdef}\label{sec:SWAdef}
-%(SWT)
-\noindent 
-Let $\Semiring$ be a commutative semiring, 
-$\Sigma$ and $\Delta$ be alphabets called respectively \emph{input} and \emph{output}, %{alphabets}, 
-and $\bar\Phi$ be a label theory over $\Semiring$
-containing $\Phi_\Sigma$, $\Phi_\Delta$, $\Phi_{\Sigma, \Delta}$.
-
-\begin{definition}
-\label{def:transducer} \label{def:SWT}
-A \emph{symbolic-weighted transducer} (\SWT)
-over $\Sigma$, $\Delta$, $\Semiring$ and $\bar\Phi$
-%the input and output alphabets~$\Sigma$ and $\Delta$ with label theory $\bar\Phi$, and the semiring $\Semiring$ 
-is a tuple
-$T = \< Q, \init, \bar{\wei}, \final >$,
-where $Q$ is a finite set of states, 
-$\mathsf{in} : Q \to \Semiring$   %\Phi_{\Sigma, \Delta}
-respectively $\mathsf{out} : Q \to \Semiring$  %\Phi_{\Sigma, \Delta}
-are functions defining the weight for entering, 
-respectively leaving, computation in a state,  
-and $\bar{\wei}$ is a triplet of transition functions 
-$\wei_{10}: Q \times Q \to \Phi_{\Sigma}$,
-$\wei_{01}: Q \times Q \to \Phi_{\Delta}$, and 
-$\wei_{11}: Q \times Q \to \Phi_{\Sigma, \Delta}$.
-\end{definition}
-%
-\noindent
-For convenience, we shall sometimes present transition functions 
-as functions of  
-$Q \times (\Sigma \cup \{ \varepsilon \}) \times (\Delta \cup \{ \varepsilon \}) \times Q \to \Semiring$,
-overloading the function names,
-such that, for all $q, q' \in Q$, $a \in \Sigma$,  $b \in \Delta$, 
-\[
-\begin{array}{rcll}
-\wei_{10}(q, a, \varepsilon, q') & = & \phi(a) & %\wei_\Sigma(q, q')(a)
-\quad\mathrm{where~} \phi = \wei_{10}(q, q') \in \Phi_\Sigma,\\
-\wei_{01}(q, \varepsilon, b, q') & = & \psi(b) & 
-\quad\mathrm{where~} \psi = \wei_{01}(q, q') \in \Phi_\Delta,\\
-\wei_{11}(q, a, b, q') & = & \eta(a, b) & 
-\quad\mathrm{where~} \eta = \wei_{11}(q, q') \in \Phi_{\Sigma, \Delta}.\\
-\end{array}      
-\]
-%
-\noindent 
-% The symbolic-weighted transducer
-The \SWT $T$ computes on pairs of words $\< s, t> \in \Sigma^* \times \Delta^*$,
-$s$ and $t$, being respectively called input and output word.
-% \noindent
-More precisely, $T$ defines a mapping 
-from $\Sigma^* \times \Delta^*$ into~$\Semiring$,
-based on an intermediate function $\weight_T$
-defined recursively for every states $q, q' \in Q$, 
-and every strings $\< s, t> \in \Sigma^* \times \Delta^* \setminus \{ \< \varepsilon, \varepsilon >\}$ by:
-%$a \in \Sigma$, $u \in \Sigma^*$, $b\in \Delta$, $v\in \Delta^*$, by:
-%
-\begin{align}
-\weight_T(q, \varepsilon, \varepsilon, q)  & = \one \label{eq:SWT-weight}\\ %\final(q)\\
-\weight_T(q, \varepsilon, \varepsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
-%\weight_T(q, au, \varepsilon, q') & = \displaystyle\bigoplus_{q'' \in Q} 
-%    \wei_{10}(q, a, \varepsilon, q'') \otimes \weight_T(q'', u, \varepsilon, q')\nonumber\\
-%\weight_T(q, \varepsilon, bv, q') & = \displaystyle\bigoplus_{q'' \in Q} 
-%    \wei_{01}(q, \varepsilon, b, q'') \otimes \weight_T(q'', \varepsilon, v, q')\nonumber\\
-\weight_T(q, s, t, q') & = \displaystyle\bigoplus_{\begin{array}{c}
-                                                   \scriptstyle q'' \in Q\\
-                                                   \scriptstyle s = au,\, a \in \Sigma
-                                                   \end{array}} 
-    \wei_{10}(q, a, \varepsilon, q'') \otimes \weight_T(q'', u, t, q')\nonumber\\
-                    & \oplus \displaystyle\bigoplus_{\begin{array}{c}
-                                                     \scriptstyle q'' \in Q\\ 
-                                                     \scriptstyle t = bv,\, b \in \Delta\\
-                                                     \end{array}} 
-    \wei_{01}(q, \varepsilon, b, q'') \otimes \weight_T(q'', s, v, q')\nonumber\\
-                    & \oplus \displaystyle\bigoplus_{\begin{array}{c}
-                                                     \scriptstyle q'' \in Q\\
-                                                     \scriptstyle s = au,\, t = bv\\
-                                                     \end{array}} 
-    \wei_{11}(q, a, b, q'') \otimes \weight_T(q'', u, v, q')\nonumber
-\end{align}
-%
-We recall that, by convention (Section~\ref{sec:semiring}), 
-an empty sum with $\bigoplus$ is equal to~$\zero$. 
-%
-Intuitively, using a transition $\wei_{ij}(q, a, b, q')$ means for $T$:
-when reading respectively $a$ and $b$ at the current positions in the input and output words, 
-increment the current position in the input word if and only if $i = 1$, 
-and in the output word iff $j = 1$ (otherwise, do not change it),
-and change state from $q$ to $q'$.
-When $a = \varepsilon$ (resp. $b = \varepsilon$), the current symbol 
-in the input (resp. output) is not read.
-%
-%In contrast with the models of weighted transducers over finite alphabets~\cite{Mohri03ijfcs},
-%the input and output symbols at current positions are always read by transitions, 
-%even when they do not change the reading position the head's position.
-%This is an important feature in the case of an infinite alphabet in 
-%order to compare input and output symbols.
-%which cannot be stored in the finite memory of the transducer.
-%
-Since $\zero$ is absorbing for~$\otimes$ in~$\Semiring$,
-one term $\wei_{ij}(q, a, b, q'')$ equal to $\zero$ in the above expression 
-will be ignored in the sum, meaning that there is no possible transition
-from state $q$ into state $q'$ while reading $a$ and $b$.
-This is analogous to the case of a transition's guard not satisfied by $\<a, b>$ for 
-symbolic transducers.
-
-%whereas considering the current symbol may be useful to compute a transition weight.
-%(even when it does not change the head's position, like with $\varepsilon$-transitions).
-%
-%The cases $\weight_T(q, au, \varepsilon, q')$ and $\weight_T(q, \varepsilon, bv, q')$
-%are missing in the definition of $\weight_T$.
-%It means that $T$ must avoid configurations where it reached the end of 
-%the output word and not of the input one, or vice-versa.
-%This can be done by using $\wei_{10}$ and $\wei_{01}$
-%before reaching the end of word, and using a special state for this purpose.
-
-The expression \eqref{eq:SWT-weight} 
-can be seen as a stateful definition of 
-an edit-distance between a word $s \in \Sigma^*$ and a word $t \in \Delta^*$,
-see also~\cite{Mohri03ijfcs}.
-Intuitively, 
-$\wei_{10}(q, a, \varepsilon, r)$ is the cost of 
-the deletion of the symbol~$a \in \Sigma$ in~$s$, 
-$\wei{01}(q, \varepsilon, b, r)$ is the cost 
-of the insertion of~$b \in \Delta$ in $t$, 
-and $\wei_{11}(q, a, b, r)$ is the cost 
-of the substitution of  $a \in \Sigma$ by~$b \in \Delta$.
-%
-The cost of a sequence of such operations transforming $s$ into $t$, 
-is the product, with $\otimes$, of the individual costs of the operations involved;
-And the distance between $s$ and $t$ is the sum, with $\oplus$,
-of all possible products.
-
-\medskip\noindent
-%Let $\< s, t> \in \Sigma^* \times \Delta^*$, with $s = s_1\ldots s_n$, and $t = t_1\ldots t_m$. 
-Formally, the weight associated by $T$ to $\< s, t> \in \Sigma^* \times \Delta^*$ is: 
-%defined as follows:
-\begin{equation}
-T(s, t)  = 
-\displaystyle\bigoplus_{q, q' \in Q} \init(q) 
-\mathop{\otimes} \weight_T(q, s, t, q') \mathop{\otimes} \final(q')
-\label{eq:SWT-value}
-\end{equation}
-
-%\begin{example}
-%We can define with a \SWT the computation of a similarity measure 
-%between timed sequences
-%similar to dynamic time warping (DTW).
-%
-%Let $\Semiring$  be the tropical (\emph{min-plus}) semiring of Figure~\ref{fig:semirings} and 
-%let $\Sigma = \Delta = \mathbb{R}_+$ be sets of timestamps.
-%We consider a \SWT with one state $q$ and transitions
-%$\wei_{11}(q, d, d', q) = 
-% \wei_{01}(q, d, d', q) = 
-% \wei_{10}(q, d, d', q) = |d' - d|$,
-%for all $d, d' \in \mathbb{R}_+$.
-%% needs reading input/output symbols even by epsilon-transitions.
-%The recursive definition of $\weight_T$ correspond to the dynamic programming equations of DTW
-%for the computation of an optimal match between words, 
-%the matching cost for two symbols being the 
-%the time distance between them.
-%\endex
-%\end{example}
-
-\begin{example}
-In Common Western Music Notation~\cite{Gould11Notation}, 
-several symbols may be used to represent a unique sounding event.
-For instance, several notes can be combined with a tie, 
-like in $\musQuarter\!\!\!\mathrel{\raisebox{-1.5mm}{$\smile$}}\!\!\!\musEighth$,  
-and one note can be augmented by half its duration with a dot like in~\musQuarterDotted{}
-
-\noindent 
-We propose a small transducer model that compares an input sequence of sounding 
-events (music "performance") 
-to an output sequence of written events (music "score").
-% simple pointwise distance between two sequences of timestamped events **
-%
-Let us consider the tropical (\emph{min-plus}) semiring~$\Semiring$ 
-of Figure~\ref{fig:semirings} and 
-let $\Sigma = \mathbb{R}_+$ be an input alphabet of event dates
-and $\Delta = \{ \mathsf{e}, \mathsf{-} \} \times \mathbb{R}_+$ 
-be an output alphabet of symbols with timestamps. 
-A symbol $\< \mathsf{e}, d > \in \Delta$ represents an event starting at date $d$, 
-and $\< \mathsf{-}, d >$ is a continuation of the previous event.
-%More precisely, we let 
-%$\Sigma = \{ a, - \} \times \R_+$ and 
-%This example of $\Delta$ is motivated by the case of music notation, 
-%where several notated events (notes) can be tied together, 
-%with a \emph{tie} or a \emph{dot}
-%meaning that they will be played as a unique sounding event.
-%The timestamp of $a \in \Sigma$, denoted by $\mathsf{t}(a)$, is expressed as a rational number.
-
-We consider a \SWT with two states $q_0$ and $q_1$ whose purpose 
-is to compare a recorded performance $s \in \Sigma^*$
-with notated music sheet $t \in \Delta^*$.
-One timestamp $d_i \in \Sigma$ may corresponds 
-to one notated event $\<\mathsf{e}, d'_i> \in \Sigma$, in which case 
-the weight value computed by the \SWT is the time distance between both
-(see transitions $\wei_{11}$ below).
-%
-If $\<\mathsf{e}, d'_i>$ is followed by continuations 
-$\<\mathsf{-}, d'_{i+1}>$..., they are just skip with no cost (transitions $\wei_{01}$ or weight $\one$).
-%transitions
-\[
-\begin{array}{rclcrcl}
-\wei_{11}(q_0, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d| & \quad &
-\wei_{11}(q_1, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d|\\
-\wei_{01}(q_0, \varepsilon, \< \mathsf{-}, d'>, q_0) & = & \one & &
-\wei_{01}(q_1, \varepsilon, \< \mathsf{-}, d'>, q_0) & = & \one\\
-\wei_{10}(q_0, d, \varepsilon, q_1) & = & \alpha & & %\multicolumn{3}{l}{\mathrm{for~all~} b \in \Delta}
-\end{array}
-\]
-%
-Moreover, it may happen that the performers plays an extra note accidentally, but only once in a row. 
-This is modelled by the transition $\wei_{10}$ with an arbitrary weight value $\alpha \in \Semiring$, 
-switching from state $q_0$ (normal) to $q_1$ (error).
-The transitions in the second column below switch back to the normal state $q_0$.
-% the metric computed by the \SWT is the smallest sum of point wise distances 
-% between dates of input and output events.
-At last, we let $q_0$ be the only initial and final state, with
-$\init(q_0) = \final(q_0) = \one$, and 
-$\init(q_1) = \final(q_1) = \zero$. 
-%$\init(q_0, d, b) = \final(q_0, d, b) = \one$, and 
-%$\init(q_1, d, b) = \final(q_1, d, b) = \zero$, 
-%for all $d \in \Sigma$ and $b \in \Delta$.
-\endex
-\end{example}
-
-\noindent
-The \emph{Symbolic Weighted Automata} %$A = \< Q, \init, \weight, \final >$
-%over $\Sigma$,  $\Semiring$ and $\bar\Phi$
-are defined similarly as the transducers of Definition~\ref{def:SWT}, 
-by simply omitting the output symbols.
-%
-In this case, the label theory $\bar\Phi$ can be reduced to a singleton $\< \Phi_\Sigma>$.
-%over $\Sigma$ is reduced to
-%a set $\Phi_\Sigma$ closed under~$\oplus$ and~$\otimes$.
-%
-\begin{definition} \label{def:SWA}
-A \emph{symbolic-weighted automaton} (\SWA)
-over $\Sigma$, $\Semiring$ and $\bar\Phi$
-%the input alphabet~$\Sigma$ and the commutative semiring $\Semiring$ 
-is a tuple
-$A = \< Q, \init, {\wei_1}, \final >$,
-where $Q$ is a finite set of states, 
-$\mathsf{in} : Q \to \Semiring$, %\Phi_\Sigma$, 
-respectively $\mathsf{out} : Q \to \Semiring$  %\Phi_\Sigma,$
-are functions defining the weight for entering, 
-respectively leaving, computation in a state, 
-and ${\wei_1}$ is a transition functions 
-from $Q \times Q$ into~$\Phi_{\Sigma}$.
-\end{definition}
-%      
-\noindent
-As above in the case of \SWT, 
-when $\wei_1(q, q') = \phi \in \Phi_\Sigma$, 
-%respectively $\mathsf{in}(q) = \phi$, $\mathsf{out}(q') = \phi$,
-we may write $\wei_1(q, a, q')$ for~$\phi(a)$. 
-%respectively $\mathsf{in}(q, a) = \phi(a)$, $\mathsf{out}(q', a) = \phi(a)$.
-%$\wei_1: Q \times \Sigma \times Q \to \Semiring$, 
-The computation of $A$ on words $s \in \Sigma^*$
-is defined with an intermediate function $\weight_A$, 
-defined as follows for $q, q' \in Q$, $a \in \Sigma$, $u \in \Sigma^*$,
-%
-\begin{align}
-\weight_A(q, \varepsilon, q) & = \one\\ %\final(q)\\
-\weight_A(q, \varepsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
-\weight_A(q, au, q') & =  \displaystyle\bigoplus_{q'' \in Q} 
-    \wei_{1}(q, a, q'') \otimes \weight_A(q'', u, q')\nonumber
-\label{eq:SWA-weight}
-\end{align}
-%
-\noindent
-and the weight value associated by $A$ to 
-$s \in \Sigma^*$ is defined as follows: %$s = s_1\ldots s_n \in \Sigma^+$ 
-\begin{equation}
-A(s)  = 
-\displaystyle\bigoplus_{q, q' \in Q} \init(q) 
-\mathop{\otimes} \weight_A(q, s, q') \mathop{\otimes} \final(q') 
-\label{eq:SWA-value}
-\end{equation}
-
-%
-%When $\wei_\varepsilon(q, q') = \zero$ for all $q, q' \in Q$, 
-%the automaton~$A$ is called \emph{without $\varepsilon$-transitions}.
-
-%The \emph{summary} of a $\SWT$, resp. a $\SWA$, is ***
-
-%\subsection{Properties}
-\noindent
-The following property will be useful to the approach on 
-symbolic weighted parsing presented in Section~\ref{sec:parsing}.
-
-\begin{proposition} \label{prop:epsilon}
-Given a \SWT $T$ over $\Sigma$, $\Delta$, 
-$\Semiring$ commutative, bounded and complete,
-and $\bar\Phi$ effective,
-and a $\SWTA$ $A$ over $\Sigma$, $\Semiring$ and $\bar\Phi$,
-there exists an effectively constructible \SWA 
-$B_{T, A}$ over $\Delta$, $\Semiring$ and $\bar\Phi$,
-such that for all $t \in \Delta^*$, 
-$B_{T, A}(t) = \displaystyle\bigoplus_{s\in \Sigma^*} A(s) \otimes T(s, t)$.
-\end{proposition}
-%
-\begin{proof}
-Let $T = \< Q, \init_T, \bar{\wei}, \final_T >$,
-where $\bar{\wei}$ contains $\wei_{10}$, $\wei_{01}$, and $\wei_{11}$,
-from $Q \times Q$ into respectively 
-$\Phi_{\Sigma}$, $\Phi_{\Delta}$, and $\Phi_{\Sigma, \Delta}$,
-and let $A = \< P, \init_A, \wei_1, \final_A >$
-with $\wei_1: Q \times Q \to \Phi_{\Sigma}$.
-%
-\noindent
-The state set of $B_{T, A}$ will be $Q' = P \times Q$.
-The entering, leaving and transition functions of $B_{T, A}$ will 
-simulate synchronized computations of $A$ and $T$, 
-while reading an output word of $\Delta^*$.
-%
-Its state entering functions is defined 
-for all $p \in P$, $q \in Q$ 
-by $\init'(p, q) = \init_A(p) \otimes \init_T(q)$.
-The transition function $\wei'_1$ will roughly perform 
-a synchronized product of transitions defined by $\wei_1$, 
-$\wei_{01}$ ($T$ reading in output word and not in input word)
-and $\wei_{11}$ ($T$ reading in output word and input word).
-%
-Moreover, $\wei'_1$ also needs to simulate transitions 
-defined by $\wei_{10}$: $T$ reading in intput word and not in output word.
-Since $B_{T, A}$  will read only in the output word, such a transition corresponds
-to an $\varepsilon$-transition of $\SWA$, 
-but $\SWA$ have been defined without $\varepsilon$-transitions.
-Therefore, in order to take care of this case, we perform an on-the-fly
-suppression of $\varepsilon$-transition in the $\SWA$ in construction, 
-following the algorithm of~\cite{LombardySakarovitch12ciaa}. 
-%
-% Initialize state leaving functions i
-%and $\final'(p, q) = \final_A(p) \otimes \final_T(q)$.
-
-%Every transition of $B_{T, A}$ will
-%simulate a sequence of transitions of $T$ performing the following steps:
-%advance in the input word while staying immobile in the output word, 
-%and then make one step in the output word (and advance in the input word or not).
-
-\noindent
-Initially, for all $p_1, p_2 \in P$, and $q_1, q_2 \in Q$, let
-\[
-\wei'_1\bigl( \< p_1, q_1>, \< p_2, q_2>\bigr) = 
-\wei_1(p_1, p_2) \otimes
-\bigl[
-\wei_{01}(q_1, q_2) 
-\oplus
-\displaystyle\bigoplus_{\Sigma}
-\wei_{11}(q_1, q_2) 
-\bigr].
-\]
-
-\noindent
-Iterate the following for all $p_1\in P$ and $q_1, q_2 \in Q$:
-for all $p_2\in P$ and $q_3 \in Q$,
-\[
-\wei'_1\bigl( \< p_1, q_1>, \< p_2, q_3>\bigr) \opluseq 
-\displaystyle\bigoplus_{\Sigma} \wei_{10}(q_1, q_2) 
-\otimes 
-\wei'_1\bigl( \< p_1, q_2>, \< p_2, q_3>\bigr)
-\]
-and
-\(
-\final'(p_1, q_1) \opluseq 
-\displaystyle\bigoplus_{\Sigma} \wei_{10}(q_1, q_2) 
-\otimes \final'(p_1, q_2)
-\) 
-\marginpar{\tiny proof correctness}
-%\qed
-\end{proof}
-
-%The complexity of construction
-\noindent
-The construction time and size for $B_{T, A}$ are $O(\| T \|^3 . \| A \|^2)$,
-where the sizes $\| T \|$ and $\| A \|$ are their number of states.
-%of~$T$ is its number of states $|Q|$.
-
-\begin{corollary} \label{cor:epsilon}
-Given a \SWT $T$ over $\Sigma$, $\Delta$, 
-$\Semiring$ commutative, bounded and complete,
-and $\bar\Phi$ effective,
-and $s \in \Sigma^+$, 
-there exists an effectively constructible \SWA 
-$B_{T, s}$ over $\Delta$, $\Semiring$ and $\bar\Phi$,
-such that for all $t \in \Delta^*$, $B_{T, s}(t) = T(s, t)$.
-\end{corollary}
-
-%\noindent
-%The construction time and size for $B_{T, s}$ are $O(\| T \|^3 . | s |^2)$,
-%where the size $\| T \|$ of~$T$ is its number of states $|Q|$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SW-VPA
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{SW Visibly Pushdown Automata}
-\label{section:SWVPA}\label{sec:SWVPA}
-The model presented in this section generalizes Symbolic VPA~\cite{dAntonyAlur14SVPDA}
-from Boolean semirings to arbitrary semiring weight domains.
-It will compute on nested words over infinite alphabets, 
-associating to every such word a weight value. 
-Nest words are able to describe structures of labeled trees, 
-and in the context of parsing, they will be useful to 
-represent parse trees.
-
-\subsection{Definition}\label{se:SWVPA-def}
-Let $\Omega$ be a countable alphabet 
-%finite (large) or infinite,
-that we assume partitioned into three 
-subsets~$\Omegai$, $\Omegac$, $\Omegac$,
-whose elements are respectively called 
-\emph{internal}, \emph{call} and \emph{return} symbols.
-% names are  coined by application to functional program verification
-% \begin{itemize}
-% \item a set $\Omegai$ of \emph{internal symbols} denoted $a$,
-% \item a set $\Omegac$ of \emph{call symbols} denoted $\call{a}$,
-% \item a set $\Omegar$ of \emph{return symbols} denoted $\return{a}$.
-% \end{itemize}
-Let~$\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative and complete semiring and let  
-$\bar\Phi = \< \Phii, \Phic, \Phir, \Phici, \Phicc, \Phicr>$ 
-be a label theory over $\Semiring$
-%In order to simplify notations, %and following the definition of Section~\ref{section:transducer}, 
-%we shall write respectively 
-where $\Phii$, $\Phic$, $\Phir$ and~$\Phicx$ (with $\mathsf{x} \in \{ \mathsf{i}, \mathsf{c}, \mathsf{r}\}$) 
-stand respectively 
-for~$\Phi_\Omegai$, $\Phi_\Omegac$, $\Phi_\Omegar$ and~$\Phi_{\Omegac, \Omega_\mathsf{x}}$.
-%
-%Moreover, we extend this theory with a set $\Phii$ 
-%of unary functions in $\Omegai \to \Semiring$,
-%closed under $\oplus$ and $\otimes$.
-
-\begin{definition}
-A \emph{Symbolic Weighted Visibly Pushdown Automata} (\SWVPA) 
-over  $\Omega = \Omegai \uplus \Omegac \uplus \Omegar$, $\Semiring$ and $\bar\Phi$ 
-is a tuple $A = \< Q, P, \init, \bar\wei, \final >$,
-where $Q$ is a finite set of states, 
-$P$ is a finite set of stack symbols, 
-$\mathsf{in} : Q \to \Semiring$, 
-respectively $\mathsf{out} : Q \to \Semiring,$
-are functions defining the weight for entering, 
-respectively leaving, a state, 
-and $\bar\wei$ is a sextuplet composed of the transition functions:
-$\weii : Q \times P \times Q \to \Phici$,  
-$\weiei : Q \times Q \to \Phii$,  
-$\weic : Q \times P \times Q \times P \to \Phicc$,  
-$\weiec : Q \times P \times Q \to \Phic$,  
-$\weir : Q \times P \times Q \to \Phicr$,  
-$\weier : Q \times Q \to \Phir$.
-%and 
-%$\weiex : Q \times Q \to \Phix$ 
-%with $\mathsf{x} \in \{ \mathsf{i}, \mathsf{c}, \mathsf{r}\}$.
-\end{definition}
-%
-Similarly as in Section~\ref{section:transducer}, 
-we extend the above transition functions as follows
-for all $q, q' \in Q$, $p \in P$, 
-$a \in \Omegai$, 
-$\call{c} \in \Omegac$, 
-$\return{r} \in \Omegar$, 
-overloading their names: % for simplicity:
-\[
-\begin{array}{lll}
-\weii: Q \times \Omegac \times P \times \Omegai \times Q \to \Semiring & 
-\weii(q, c, p, a, q') = \eta_\mathsf{ci}(c, a) & 
-\mathrm{where~} \eta_\mathsf{ci} = \weii(q, p, q'),\\
-%
-\weiei: Q \times \Omegai \times Q \to \Semiring & 
-\weiei(q, a, q') = \phi_\mathsf{i}(a) &
-\mathrm{where~} \phi_\mathsf{i} = \weiei(q, q').\\[2pt]
-%
-\weic: Q \times \Omegac \times P \times  \Omegac \times P \times Q \to \Semiring & 
-\weic(q, c, p, \call{c'}, p', q') = \eta_\mathsf{cc}(c, \call{c'}) & 
-\mathrm{where~} \eta_\mathsf{cc} = \weic(q, p, p', q'),\\
-%
-\weiec: Q \times \Omegac \times P \times Q \to \Semiring & 
-\weiec(q, {\call{c}}, p, q') = \phi_\mathsf{c}({\call{c}}) &
-\mathrm{where~} \phi_\mathsf{c} = \weiec(q, p, q').\\[2pt]
-%
-\weir: Q \times \Omegac \times P \times \Omegar \times Q \to \Semiring & 
-\weir(q, {\call{c}},  p, {\return{r}}, q') = \eta_\mathsf{cr}({\call{c}},  {\return{r}}) & 
-\mathrm{where~} \eta_\mathsf{cr} = \weir(q, p, q'),\\
-%
-\weier: Q \times \Omegar \times Q \to \Semiring & 
-\weier(q, {\return{r}}, q') = \phi_\mathsf{r}({\return{r}}) &
-\mathrm{where~} \phi_\mathsf{r} = \weier(q, q').\\
-\end{array}      
-\]
-
-\noindent
-The intuition is the following for the above transitions.
-
-\noindent
-$\weii$ and $\weiei$ both read an input internal symbol $a$ and change state from $q$ to $q'$, 
-without changing the stack. 
-Moreover, $\weii$ reads a pair made of 
-${\call{c}} \in \Omegac$ and $p \in P$ at the top of the stack 
-($c$ is compared to $a$ by the weight function $\eta_\mathsf{ci} \in \Phici$)
-and $\weiei$ applies if and only if the stack is empty.
-
-\noindent
-$\weic$ and $\weiec$ read the input call symbol $\call{c}'$, 
-push it to the stack along with $p'$, and change state from $q$ to to $q'$.
-Moreover, $\weic$ reads ${\call{c}}$ and $p$ at the top of the stack 
-($c$ is compared to $c'$),
-and $\weiec$ applies iff the stack is empty.
-
-\noindent
-$\weir$ and $\weier$ read the input return symbol $\return{r}$, and change state from $q$ to to $q'$.
-Moreover, $\weir$ reads and pop from stack a pair made of $\call{c}$ and $p$, 
-($\call{c}$ is compared to $\return{r}$),
-and $\weier$ applies iff the stack is empty.
-%In this case, the weight function $\phi_\mathsf{r}$ 
-%computes a value of matching between the call and return symbols $c$ and $r$.
-%This value might be set to $\zero$ in order to express that the symbols do not match.
-
-Formally, the transitions of the automaton~$A$ are defined
-in term of %a weight value computed by 
-an intermediate function $\weight_A$, like in Section~\ref{sec:SWT}.
-In the case of a pushdown automaton, a configuration, denoted by $q[\gamma]$ 
-is composed of a state $q \in Q$ 
-and a stack content $\gamma \in \Gamma^*$, 
-where $\Gamma = \Omegac \times P$.
-Hence, $\weight_A$ is a function from 
-$[Q \times \Gamma^*] \times \Omega^* \times [Q \times \Gamma^*]$ into~$\Semiring$
-(the empty stack is denoted by $\bot$). 
-%
-\begin{align}
-%\begin{array}[t]{rcl}
-\weight_A\bigl(\config{q}{\bot}, \varepsilon, \config{q}{\bot}) & = \one\label{eq:SWVPA-weight}\\
-\weight_A\bigl(\config{q}{\bot}, \varepsilon, \config{q'}{\gamma}) & = \zero 
-\mathrm{~if~} q \neq q'\nonumber\\
-\weight_A\bigl(\configup{q}{\<{\call{c}}, p>\stackup\gamma}, a\, u, 
-               \config{q'}{\gamma'}\bigr) & =  
- {\displaystyle\bigoplus_{q'' \in Q}} \weii(q, c, p, a, q'') 
-  \otimes \weight_A\bigl(\configup{q''}{\<{\call{c}}, p> \stackup \gamma}, u, 
-                         \config{q'}{\gamma'}\bigr)\nonumber\\
- %
-\weight_A\bigl(\config{q}{\bot}, a\, u, 
-               \config{q'}{\gamma'}\bigr) & =  
-  {\displaystyle\bigoplus_{q'' \in Q}} \weiei(q, a, q'') 
-   \otimes \weight_A\bigl(\config{q''}{\bot}, u, \config{q'}{\gamma'}\bigr)\nonumber\\
-%
-\weight_A\bigl(\configup{q}{\<{\call{c}}, p>\stackup\gamma}, {\call{c}'}\, u, 
-               \config{q'}{\gamma'}\bigr) & =  
- {\displaystyle\bigoplus_{\begin{array}{c}
-                          \scriptstyle q'' \in Q\\[-2pt]
-                          \scriptstyle p' \in P
-                          \end{array}}}
- \weic\bigl(q, {\call{c}}, p, {\call{c}'}, p', q''\bigr) 
- \otimes \weight_A\bigl(\configup{q''}{\<{\call{c}'}, p'>\stackup \<{\call{c}}, p>\stackup \gamma}, u, 
-                        \config{q'}{\gamma'}\bigr)\nonumber\\[1mm]
-%
-\weight_A\bigl(\config{q}{\bot}, {\call{c}}\, u, 
-               \config{q'}{\gamma'}\bigr) & =  
- {\displaystyle\bigoplus_{\begin{array}{c}
-                          \scriptstyle q'' \in Q\\[-2pt]
-                          \scriptstyle p \in P
-                          \end{array}}}
-  \weiec(q, {\call{c}}, p, q'') 
-  \otimes \weight_A\bigl(\config{q''}{{\call{c}}\, p}, u, 
-                         \config{q'}{\gamma'}\bigr)\nonumber\\
-%
-\weight_A\bigl(\configup{q}{\<{\call{c}}, p>\stackup \gamma}, {\return{r}}\, u, 
-               \config{q'}{\gamma'}\bigr) & =  
- {\displaystyle\bigoplus_{q'' \in Q}} 
-  \weir\bigl(q, {\call{c}}, p, {\return{r}}, q''\bigr) 
-  \otimes \weight_A\bigl(\config{q''}{\gamma}, u, 
-                         \config{q'}{\gamma'}\bigr)\nonumber\\
-%
-\weight_A\bigl(\config{q}{\bot}, {\return{r}}\, u, 
-               \config{q'}{\gamma'}\bigr) & =  
- {\displaystyle\bigoplus_{q'' \in Q}} \weier(q, {\return{r}}, q'') 
-  \otimes \weight_A\bigl(\config{q''}{\bot}, u, 
-                         \config{q'}{\gamma'}\bigr)\nonumber
-%\end{array}
-\end{align}
-%
-% and ${\call{c}}\, p\stacksep \gamma$ 
-%denotes a stack where the pair made of ${\call{c}} \in \Omegac$ and $p \in P$ is the top symbol 
-%and $\gamma$ is the rest of stack.
-
-\noindent
-The weight associated by $A$ to $s \in \Omega^*$
-is defined according to empty stack semantics: 
-%
-\begin{equation}
-A(s)  = 
-{\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
-\mathsf{in}(q) \mathop{\otimes} 
-\weight_A\bigl(\config{q}{\bot}, s, \config{q'}{\bot}\bigr) 
-\mathop{\otimes} \mathsf{out}(q').
-\label{eq:weightA}
-\end{equation}
-
-\begin{example}
-structured words with timed symbols...
-intro language of music notation? (markup = time division, leaves = events etc)
-\end{example}
-
-\noindent 
-Every $\SWA$ $A = \< Q, \init, {\wei_1}, \final >$,
-over $\Sigma$, $\Semiring$ and $\bar\Phi$
-is a particular case of $\SWVPA$ 
-$\< Q, \emptyset, \init, \bar\wei, \final >$ 
-over $\Omega$, $\Semiring$ and $\bar\Phi$
-with $\Omegai = \Sigma$ and $\Omegac = \Omegar = \emptyset$,
-and computing with an always empty stack:
-$\weiei = \wei_1$ and all the other functions 
-of~$\bar\wei$ are the constant~$\zero$.
-
-\subsection{Properties}
-Like VPA and symbolic VPA, 
-the class of \SWVPA is closed under the binary operators of the underlying semiring.
-%
-\begin{proposition}\label{prop:SWVPA-product}
-Let $A_1$ and $A_2$ be two \SWVPA
-over the same $\Omega$, $\Semiring$ and $\bar\Phi$.
-There exists two effectively constructible $\SWVPA$ 
-$A_1 \oplus A_2$ and $A_1 \otimes A_2$,  
-such that for all $s \in \Omega^*$, 
-$(A_1 \oplus A_2)(s) = A_1(s) \oplus A_2(s)$ and 
-$(A_1 \otimes A_2)(s) = A_1(s) \otimes A_2(s)$.
-\end{proposition}
-%
-\begin{proof}
-The construction is essentially the same 
-as in the case of the Boolean semiring~\cite{dAntonyAlur14SVPDA}.
-\end{proof}
-
-\subsection{Best Search} 
-\label{sec:best}\label{sec:search}
-%**hypotheses**
-Let us assume that the semiring~$\Semiring$ is
-commutative, bounded, and complete,  
-\marginpar{\tiny total?}
-and assume an effective label theory.
-%s.t. for all $\phi \in \Phii$, 
-%$\psi \in \Phir$, 
-%and $\eta \in \Phicr$, 
-%$\displaystyle\bigoplus_{a \in \Omegai} \phi(a)$ 
-%$\displaystyle\bigoplus_{r \in \Omegar} \phi(r)$ and 
-%$\displaystyle\bigoplus_{{\call{c}} \in \Omegac} 
-%\displaystyle\bigoplus_{{\return{r}} \in \Omegar} \eta({\call{c}}, {\return{r}})$
-%are computable...
-% 
-% total ? 
-% monotonic and superior writ natural ordering
-%Regarding the infinite sum operator, note that
-%$\bigoplus_{x \in \Phi_\Omega} \phi(x)$, 
-%$\bigoplus_{y \in \Phi_\Delta} \psi(y)$, and  
-%... exist and in $\Semiring$.
-%
-We propose a Dijkstra algorithm computing, for a $\SWVPA$ $A$
-over~$\Omega$, $\Semiring$ and~$\bar\Phi$, 
-the minimal weight %\wrt~$\leq_\oplus$, 
-for a word in~$\Omega^*$.
-
-\noindent
-More precisely, 
-let $b_\bot : Q \times Q \to \Semiring$ be the function:
-%
-\begin{equation}\label{eq:bbot}
-  b_\bot(q, q') = \bigoplus_{s\in \Omega^*} 
-  \textstyle
-  \weight_A\bigl(\config{q}{\bot}, s, \config{q'}{\bot}\bigr)
-\end{equation}
-%
-Since $\Semiring$ is complete, the infinite sum in \eqref{eq:bbot} is well defined,
-and, roviding that $\Semiring$ is total, it is the minimum in $\Omega^*$,
-\wrt~$\leq_\oplus$, % this ordering.
-of the fonction 
-$s \mapsto \weight_A(\config{q}{\sigma}, s, \config{q'}{\sigma})$.
-%
-The term $\config{q}{\bot}, s, \config{q'}{\bot}$ 
-of this sum is the central expression in 
-the definition \eqref{eq:weightA} of $A(s_0)$, for the minimum $s_0$
-of the function $\weight_A$.
-
-%
-\noindent
-Let $\top$ be a fresh stack symbol which does not belong to $\Gamma$,
-and let $b_\top : Q \times P \times Q \to \Phic$ be such that,
-for every two states $q, q' \in Q$ 
-and stack symbol $p \in P$: %and $\sigma \in \{ \bot, \top \}$, 
-\begin{equation}\label{eq:btop}
-  b_\top(q, p, q') : c \mapsto \bigoplus_{s\in \Omega^*} 
-  \textstyle
-  \weight_A\bigl(\configup{q}{\< c, p> \stackup \top }, s, \configup{q'}{\<c, p> \stackup \top}\bigr). 
-\end{equation}
-%
-Intuitively, the function defined in \eqref{eq:btop}
-associateds to $c \in \Omegac$ 
-the minimum weight of a computation of $A$
-starting in state $q$ with a stack 
-$c p \cdot \gamma \in \Gamma^+$ 
-and ending in state $q'$ with the same stack,
-such that the computation does pop 
-the pair made of $c$ and $p$ at the top of this stack,
-but may read these symbols.
-Moreover, $A$ may push another pair $\< c', p'>$ %call symbols 
-on the top of $c p \cdot \gamma$,
-following the the third case of 
-in the definition \eqref{eq:SWVPA-weight} of $\weight_A$,
-and may pop $\< c', p'>$ later, following the fifth case of \eqref{eq:SWVPA-weight} (return symbol). 
-%However, it cannot apply one of the two last cases (return symbol and empty stack)
-%when the current stack is $\gamma$.
-%pop symbols in $\gamma$.
-% Note that having a stack reduced to such a symbol makes impossible the application of the 
-% two last cases in the definition of $\weight_A$ (return symbol and empty stack). 
-% However, it is possible to apply the two first cases 
-% (internal symbol or call symbol, with a push on the top of $\top$).
-
-Algorithm~\ref{algo:Dijkstra}
-constructs iteratively markings 
-$d_\bot : Q \times Q \to \Semiring$ and 
-$d_\top : Q \times P \times Q \to \Phic$
-%of the triplets $\<q, \sigma, q'>$ 
-%of states of $A$ by weight values in $\Semiring$, 
-that converges eventually to $b_\top$ and $b_\bot$.
-%It uses for that purpose a priority queue $P$ containing triplets of 
-%$Q \times \{ \bot, \top \} \times Q$.
-
-\begin{algorithm}
-\SetKw{Extract}{extract}
-\SetKw{Update}{update}
-
-\textbf{initially} let $\Q = (Q \times Q) \cup (Q \times P \times Q)$, %contains all 
-and let $d_\bot(q_1, q_2) = d_\top(q_1, p, q_2) = \one$ if $q_1 = q_2$ and $d_\bot(q_1, q_2) = d_\top(q_1, p, q_2) = \zero$ otherwise$\;$
-
-\smallskip\noindent
-\While{$\Q \neq \emptyset$}{
-
-%\noindent\quad
-\Extract $\< q_1, q_2>$ or $\< q_1, p, q_2>$ from $\Q$
-%
-%\noindent\quad
-such that $d_\bot(q_1, q_2)$, resp. $\bigoplus_{c\in \Omegac} d_\top(q_1, p, q_2)(c)$, is minimal in $\Semiring$ wrt $\leq_\oplus$$\;$
-
-%\noindent\quad 
-%\textbf{for all} $q_0, q_3 \in Q$ 
-\Update $d_\bot$ with $\< q_1, q_2>$ or $d_\top$ with $\< q_1, p, q_2>$ (Figure~\ref{fig:best-update}).$\;$
-}
-\caption{Best search for \SWVPA}
-\label{algo:Dijkstra}
-\end{algorithm}
-
-\begin{figure}
-For all $q_0, q_3 \in Q$, %$p \in P$,
-\[
-\begin{array}{lcl}
-%\multicolumn{3}{l}{\mathrm{For~all~} q_0, q_3 \in Q, p \in P,}\\
-d_\top(q_1, p, q_3) & \opluseq &
-  d_\top(q_1, p, q_2) \otimes 
-  \displaystyle\bigoplus_{\Omegai} \weii(q_2, p, q_3)\\
-%\multicolumn{3}{l}{\quad
-%\mathrm{where~} \weii(q_2, p, q_3)_a \in \Phic
-%\mathrm{~is~the~partial~application~}
-%x_c \mapsto \weii(q_2, x_c, p, a, q_3)}\\
-%
-d_\bot(q_1, p, q_3) & \opluseq &
-  d_\bot(q_1, q_2) \otimes 
-  \displaystyle\bigoplus_{\Omegai} \weiei(q_2, q_3)\\
-%     
-d_\top(q_0, p, q_3) & \opluseq &
-  {\displaystyle\bigoplus_{\Omegac}}^2 
-  \bigl[ \bigl( \weic(q_0, p, p', q_1) \otimes_2 
-  d_\top(q_1, p', q_2) \bigr) \otimes_2
-  {\displaystyle\bigoplus_{\Omegar}} \weir(q_2, p', q_3) \bigr]\\
-%\multicolumn{3}{l}{\quad
-%\mathrm{where~} \weic(q_0, p, p', q_1)_{c'} \in \Phic
-%\mathrm{~is~the~partial~application~}
-%x_c \mapsto \weic(q_0, p, x_c, c', p', q_1)}\\[2pt]
-%
-d_\bot(q_0, q_3) & \opluseq &
-  {\displaystyle\bigoplus_{\Omegac}}
-  \bigl(
-  \weiec(q_0, p, q_1) \otimes 
-   d_\top(q_1, p, q_2) \otimes
-  \displaystyle\bigoplus_{\Omegar} \weir(q_2, p, q_3)\bigr)\\
-%
-d_\bot(q_1, q_3) & \opluseq &
-  d_\bot(q_1, q_2) \otimes 
-  \displaystyle\bigoplus_{\Omegar} \weier(q_2, q_3)\\
-%
-d_\top(q_1, p, q_3) & \opluseq & 
-  d_\top(q_1, p, q_2) \otimes d_\top(q_2, p, q_3), 
-  \mathrm{if~} \< q_2, \top, q_3> \notin P \\
-%
-d_\bot(q_1, q_3) & \opluseq & 
-  d_\bot(q_1, q_2) \otimes d_\bot(q_2, q_3), \mathrm{if~} \< q_2, \bot, q_3> \notin P \\
-\end{array}
-\]
-%$\weii(q_2, p, q_3)_a \in \Phic$, for $a \in \Omegai$,
-%is the partial application
-%$x_c \mapsto \weii(q_2, x_c, p, a, q_3)$.
-%
-%$\weic(q_0, p, p', q_1)_{c'} \in \Phic$, for $c' \in \Omegac$,
-%is the partial application
-%$x_c \mapsto \weic(q_0, p, x_c, c', p', q_1)$.
-\caption{Update $d_\bot$ with $\<q_1, q_2>$ or $d_\top$ with $\< q_1, p, q_2>$.} 
-\label{fig:best-update}
-\end{figure}
-
-\noindent
-The infinite sums in the updates of $d$ in Algorithm~\ref{algo:Dijkstra}, 
-Figure~\ref{fig:best-update}
-are well defined
-since~$\Semiring$ is complete.
-** effectively computable by hypothese that the label theory is effective**
-The algorithm performs $2.|Q|^2$ iterations until $P$ is empty, 
-and each iteration has a time complexity $O(|Q|^2 . |P|)$.
-This gives a time complexity $O(|Q|^4 . |P|)$. 
-It can be reduced by implementing $P$ as a priority queue, 
-prioritized by the value returned by $d$
-***complete***. %$|Q|^3.\log(|Q|^2)$
-
-The correctness of Algorithm~\ref{algo:Dijkstra} 
-is ensured by the invariant expressed in the following lemma.
-\begin{lemma}\label{lem:bot}
-For all $\< q_1, q_2> \notin \Q$,
-$d_\bot(q_1, q_2) =  b_\bot(q_1, q_2)$/
-\end{lemma}
-The proof is by contradiction, 
-assuming a counter-example minimal in the length of the witness word.
-
-\begin{lemma}\label{lem:top}
-For all $\< q_1, p, q_2> \notin \Q$, 
-$d_\top(q_1, p, q_2) =  b_\top(q_1, p, q_2)$,
-\end{lemma}
-
-\noindent
-For computing the minimal weight of a computation of $A$, we use the fact that,
-at the termination of Algorithm~\ref{algo:Dijkstra}, %it holds that,
-%There exist $q_1, q_2 \in Q$
-\[
-  {\displaystyle \bigoplus_{s \in \Omega^*} A(s)} = 
-  {\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
-  \mathsf{in}(q) \mathop{\otimes} d_\bot(q, q') \mathop{\otimes} \mathsf{out}(q').
-\]
-
-%\medskip
-\noindent
-In order to obtain effectively a witness 
-(word of $\Omega^*$ with a computation of $A$ of minimal weight), 
-we require the additional property of convexity of weight functions.
-
-\begin{proposition}\label{th:best-search}
-For a \SWVPA $A$ 
-over $\Omega$, 
-$\Semiring$ commutative, bounded, total and complete, %semiring 
-and $\bar\Phi$ effective and $k$-convex, % label theory,
-one can construct in PTIME a word $s \in \Omega^*$ 
-such that $A(s)$ is minimal \wrt the natural ordering for $\Semiring$. 
-\end{proposition}
-
-\subsection{Nested-Words and Parse-Trees}
-\label{sec:trees}
-The hierarchical structure of nested-words, defined with the \emph{call} and \emph{return} markup symbols  
-suggest a correspondence with trees. 
-The lifting of this correspondence to languages, of tree automata and VPA,
-has been discussed in~\cite{AlurMadhusudan09nested}, 
-and~\cite{Caralp12VPAmult} for the weighted case.
-In this section, we describe a correspondence between the symbolic-weighted extensions
-of tree automata and VPA.
-
-Let $\Omega$ be a countable ranked alphabet, such that 
-every symbol $a \in \Omega$ has a rank 
-$\rank(a) \in [0..M]$ where $M$ is a fixed natural number.
-We denote by $\Omega_k$ the subset of all symbols $a$ of $\Omega$
-with $\rank(a) = k$, where $0 \leq k \leq M$, 
-and $\Omega_{>0} = \Omega \setminus \Omega_0$.
-%
-\noindent 
-The free $\Omega$-algebra of finite, ordered, 
-$\Omega$-labeled trees is denoted by $\T(\Omega)$.
-It is the smallest set such that  $\Omega_0 \subset \T(\Omega)$
-and for all $1 \leq k \leq M$, all $a \in \Omega_k$, 
-and all $t_1, \ldots, t_k \in \T(\Omega)$, $a(t_1, \ldots, t_k) \in \T(\Omega)$.
-%
-% tree = single node (leave) labeled with a symbol of $a \in \Omegai$
-% (such a tree is simply denoted by $a$)
-% or the composition, denoted by $b(t_1,\ldots, t_n$) of a node labeled with $b$
-% and $n$ subtrees $t_1$,\ldots, $t_n$.
-%
-Let us assume a commutative semiring~$\Semiring$ 
-and a label theory~$\bar{\Phi}$ over~$\Semiring$ 
-containing one set~$\Phi_{\Omega_k}$ for each $k \in [0..M]$.
-%
-\renewcommand{\call}[1]{\ensuremath \langle_{#1}}
-\renewcommand{\return}[1]{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
-
-% A \emph{regular tree grammar} over $\Omega$ 
-% is a triplet $G = \< N, q_\mathsf{i}, R>$ where
-% $N$ is a finite set of non-terminal symbols denoted $q$..., 
-% $q_\mathsf{i} \in N$ is the starting non-terminal, 
-% $R$ is a finite set of production rules of the form
-% $q_0 \to a(q_1\ldots q_k)$ where 
-% $q_0, q_1, \ldots, q_k \in N$
-% $a \in \Omega_k$.
-% A tree $t \in \T(\Omega)$ is in the language of $G$ 
-% if it can be generated from $q_\mathsf{i}$ by 
-% non terminal replacement following the rules of $R$.
-%
-\begin{definition}  \label{def:SWTA}
-A \emph{symbolic-weighted tree automaton} (\SWTA)
-over $\Omega$, $\Semiring$, and~$\bar\Phi$
-is a triplet $A = \< Q, \init, \bar{\wei} >$ where
-$Q$ is a finite set of states, 
-$\mathsf{in} : Q \to \Phi_\Omega$ is the starting weight function, 
-and $\bar{\wei}$ is a tuplet of transition functions containing, 
-for each $k \in [0..M]$, 
-%$\wei_\varepsilon$ from $Q \times Q$ into $\Semiring$, and, 
-the functions $\wei_{k}: Q \times Q^{k} \to \Phi_{\Omega_{>0},\Omega_k}$
-and $\weie[k]: Q \times Q^{k} \to \Phi_{\Omega_k}$.
-\end{definition}
-%
-%Like in Section~\ref{sec:SWAdef}, 
-We define %from $\bar{\wei}$ 
-a transition 
-function~$\wei: Q \times (\Omega_{> 0} \cup \{ \varepsilon \}) \times \Omega \times \bigcup_{k=0}^{M} Q^k 
-  \to \Semiring$
-by: %also called $\wei$ for simplicity, 
-%such that, for all $q, q' \in Q$, $a \in \Sigma$, and $b \in \Delta$, 
-\[
-\begin{array}{rcll}
-%\wei(q_0, \varepsilon, q_1) & = &  \wei_\varepsilon(q_0, q_1),\\ %\phi_\varepsilon\\
-\wei(q_0, a, b, q_1 \ldots q_k) & = & \eta(a, b) &
-\quad\mathrm{where~} \eta = \wei_{k}(q_0, q_1\ldots q_k)\\
-\wei(q_0, \varepsilon, b, q_1 \ldots q_k) & = & \phi(b) &
-\quad\mathrm{where~} \phi = \weie[k](q_0, q_1\ldots q_k).
-\end{array}      
-\]
-%
-\noindent
-where $q_1\ldots q_k$ is $\varepsilon$ if $k = 0$.
-The first case deals with a strict subtree, with a parent node labeled by $a$, 
-and the second case is for a root tree.
-
-\noindent
-Every \SWTA %of Definition~\ref{def:SWTA} 
-defines a mapping 
-from trees of $\T(\Omega)$ into~$\Semiring$, %the weight values in~$\Semiring$,
-based on the following intermediate function
-$\weight_A: Q \times (\Omega \cup \{ \varepsilon \}) \times \T(\Omega) \to \Semiring$ 
-\begin{equation}
-%\begin{array}{rccl}
-\weight_A(q_0, a, t) =  % & = &
-%   \displaystyle\bigoplus_{q_1 \in Q} &
-%   \wei(q, \varepsilon, q_1) \otimes \weight_A(q_1, t)\\
- \displaystyle\bigoplus_{q_1 \ldots q_k \in Q^k} % &
-              \wei(q_0, a, b, q_1 \ldots q_k ) 
-   \otimes \displaystyle\bigotimes_{i=1}^{k}
-           \weight_A(q_i, b, t_i)
-%\end{array}
-\end{equation}
-where $q_0 \in Q$, $a \in \Omega_{>0} \cup \{ \varepsilon \}$ and 
-$t = b(t_1,\ldots, t_k) \in \T(\Omega)$,
-$0 \leq k \leq M$.
-
-\medskip\noindent
-Finally, the weight associated by $A$ to  $t \in \T(\Omega)$ is 
-\begin{equation}
-A(t)  = 
-\displaystyle\bigoplus_{q \in Q} \mathsf{in}(q) \mathop{\otimes} \weight_A(q, \varepsilon, t)
-\label{eq:weightTA}
-\end{equation}
-
-\noindent
-Intuitively, $\wei(q_0, a, b, q_1 \ldots q_k)$ can be seen as
-the weight of a production rule $q_0 \to b(q_1, \ldots, q_k)$ 
-of a regular tree grammar~\cite{tata}, 
-that replaces the non-terminal symbol $q_0$ by $b(q_1, \ldots, q_k)$, 
-provided that the parent of $q_0$ is labeled by $a$
-(or $q_0$ is the root node if $a = \varepsilon$).
-%in a step of tree building.
-%
-%Such a grammar computes the weights of the derivation trees 
-%of the Context-Free grammar obtained by forgetting the labeling symbols of $\Omega_{>0}$.
-The above production rule can also be seen as 
-a rule of a weighted CF grammar, of the form
-$[a, b]\, q_0 := q_1 \ldots q_k$ if $k > 0$,
-and $[a]\, q_0 := b$ if $k = 0$. 
-In the first case, $b$ is a label of the rule, 
-and in the second case, it is a terminal symbol.
-And in both cases, $a$ is a constraint on the label of rule applied 
-on the parent node in the derivation tree.
-This features of observing the parent's label  
-are useful in the case of infinite alphabet, 
-where it is not possible to memorize a label with the states.
-%
-%One can observe that 
-\noindent The weight of a labeled derivation tree $t$
-of the weighted CF grammar associated to~$A$ as above, 
-is $\weight_A(q, t)$, 
-when $q$ is the start non-terminal.
-%
-We shall now establish a correspondence between such derivation tree $t$
-and some word describing a linearization of $t$, 
-in a way that $\weight_A(q, t)$ can be computed by a $\SWVPA$.
-
-Let $\hat\Omega$ be the countable (unranked) alphabet obtained
-from $\Omega$ by: 
-$\hat\Omega = \Omegai \uplus \Omegac \uplus \Omegar$, with
-$\Omegai = \Omega_0$, 
-$\Omegac = \{ \; \call{a} \mid a \in \Omega_{>0} \}$,
-$\Omegar = \{ \; \return{a} \mid a \in \Omega_{>0} \}$.
-
-\noindent
-We associate to $\hat\Omega$
-a label theory $\hat{\Phi}$ 
-like in Section~\ref{se:SWVPA-def}, 
-%
-\noindent
-and we define a linearization of trees of $\T(\Omega)$ into 
-words of $\hat{\Omega}^*$ as follows:
-\begin{description}
-\item $\lin(a) = a$ for all $a \in \Omega_0$, 
-\item $\lin\bigl( b(t_1, \ldots, t_k)\bigr) = 
-       \call{b} \; \lin(t_1) \ldots \lin(t_k) \; \return{b}$ 
-       when $b \in \Omega_k$ for $1 \leq k \leq M$.
-\end{description}
-
-\begin{proposition}\label{lem:SWTA}
-For all \SWTA $A$ over~$\Omega$, $\Semiring$ commutative, and $\bar\Phi$,
-there exists an effectively constructible \SWVPA $A'$ over 
-$\hat\Omega$, $\Semiring$ and $\hat\Phi$ 
-such that for all $t \in \T(\Omega)$, $A'\bigl(\lin(t)\bigr) = A(t)$.
-\end{proposition} 
-% 
-\begin{proof}
-Let $A = \< Q, \init, \bar{\wei} >$ where $\bar{\wei}$ is presented as above by a function
-    %
-We build 
-$A' = \< Q', P', \init', \bar{\wei}', \final' >$,
-%computing over $\hat\Omega = \Omegai \uplus \Omegac \uplus \Omegar$,
-where $Q' = \bigcup_{k=0}^{M} Q^k$ is the set of sequences of state symbols of $A$, 
-of length at most $M$, including the empty sequence denoted by~$\varepsilon$, 
-and where $P' = Q'$ and $\bar\wei$ is defined by:
-
-\[
-\begin{array}{lcll}
-\weii(q_0\, \bar{u}, {\call{c}}, \bar{p}, a, \bar{u}) & = & \wei(q_0, c, a, \varepsilon) & 
-\mathrm{for~all~} c \in \Omega_{>0}, a \in \Omega_0\\ %\bar{p}\in P', 
-%
-\weiei(q_0\,\bar{u}, a, \bar{u}) & = & \wei(q_0, \varepsilon, a, \varepsilon) & 
-\mathrm{for~all~} a \in \Omega_0\\
-%
-\weic(q_0\,\bar{u}, {\call{c}}, \bar{p}, \call{d}, \bar{u}, \bar{q}) & = & \wei(q_0, c, d, \bar{q}) &  
-\mathrm{for~all~} c, d \in \Omega_{>0}\\ % \bar{p}\in P'
-%
-\weiec(q_0\,\bar{u}, \call{c}, \bar{u}, \bar{q}) & = & \wei(q_0, \varepsilon, c, \bar{q}) & 
-\mathrm{for~all~} c \in \Omega_{>0}\\
-%
-%\weir: Q \times \Omegac \times P \times \Sigmar \times Q \to \Semiring & 
-\weir(\varepsilon, {\call{c}}, \bar{p}, {\return{c}}, \bar{p}) & = & \one & 
-\mathrm{for~all~}  c \in \Omega_{>0}\\ % \bar{p}\in P'
-%
-%\weie: Q \times \Omegar \times Q \to \Semiring & 
-\weier(\bar{u}, {\return{c}}, \bar{q}) & = & \zero &
-\mathrm{for~all~}  c \in \Omega_{>0}
-\end{array}      
-\]
-\noindent
-All cases not matched by one of the above equations have a weight $\zero$, 
-for instance  %This is in particular the case of
-$\weir(\bar{u}, {\call{c}}, \bar{p}, {\return{d}}, \bar{q}) = \zero$
-if $c \neq d$
-or $\bar{u} \neq \varepsilon$
-or $\bar{q} \neq \bar{p}$.
-%
-%\noindent
-%It is sufficient to consider in $Q'$ only the prefixes of 
-%sequences in transition with a non-null weight.
-\qed
-\end{proof}
-
-\section{Symbolic Weighted Parsing}
-\label{sec:parsing}
-Let us now apply the models and results of the previous sections %in order to define 
-to the problem of parsing over infinite alphabet. %appropriate
-%
-%\subsection{Definition}
-%
-Let~$\Sigma$ be a countable input alphabet 
-and~$\Omega$ be a countable output ranked alphabet, 
-with maximal rank value~$M$, 
-and let $\hat\Omega = \Omegai \uplus \Omegac \uplus \Omegar$
-be the alphabet with nesting symbols associated to $\Omega$ like in Section~\ref{sec:trees},
-for the linearization of trees of $\T(\Omega)$.
-%(remember that $\Omegai = \Omega_0$).
-%
-Let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a 
-commutative, bounded, and complete \marginpar{\tiny total?} semiring  
-and let $\bar\Phi$ be an effective label theory over $\Semiring$,
-containing $\Phi_\Sigma$, $\Phi_{\Sigma, \Omegai}$, as well as
-$\Phii$, $\Phic$, $\Phir$, $\Phicr$
-(following the notations of Section~\ref{se:SWVPA-def}).
-%It is moreover assumed computable and $k$-convex for some fixed~$k$.
-%
-\noindent
-We assume given the following input:
-\begin{description}
-\item[--] a \SWT $T$ over $\Sigma$, $\Omegai$, $\Semiring$, and $\bar\Phi$, 
-defining a measure %between words 
-$T: \Sigma^* \times \Omegai^* \to \Semiring$,
-
-\item[--] a \SWTA $A$ over $\Omega$, $\Semiring$, and $\bar\Phi$,
-defining a measure $A: \T(\Omega) \to \Semiring$,
-%a \SWVPA $A$ over $\Omega$, and $\Semiring$, defining a series of nested words
-%      $A : \Omega^* \to \Semiring$,
-\item[--] an input word $s \in \Sigma^*$.
-\end{description}
-%
-As explained in Section~\ref{sec:trees}, 
-$A = \< Q, \init, \bar{\wei} >$ can be seen as a weighted regular tree grammar, 
-or a labeled weighted CF grammar whose derivation trees are in $\T(\Omega)$, 
-with weight defined by $A$.
-The purpose of the transducer $T$ is to measure a distance between input and output words, 
-\ie between the word $s$ and the sequence leaves of a derivation tree, in $\Omegai^*$.
-%that generates (weighted) trees by replacement of a state symbol~$q_0$ (non-terminal), 
-%by a tree $a(q_1,\ldots, q_k)$, where $k = \rank(a)$.
-%A replacement rule $q_0 \to a(q_1,\ldots, q_k)$, 
-%of weight $\wei(q_0, a, q_1 \ldots q_k) \in \Semiring$ 
-%according to Definition~\ref{def:SWTA}, 
-%corresponds to the production rule $q_0 := a(q_1,\ldots, q_k)$ of a weighted CF grammar,
-%with set non-terminal symbols $Q$ and set of terminal symbols $\Omega_0$.
-%
-%This actually is a slight generalization of CFG since 
-%each such production rule is labelled by  a symbol of $\Omega_{>0}$, 
-%hence parse trees %derivation trees
-%are trees of $\T(\Omega)$.
-%
-%Another (more original) generalization is that the set of terminal symbols 
-%$\Omega_0$ may be infinite.
-
-\noindent 
-We extend the measure defined by $T$ to 
-$d : \Sigma^* \times \T(\Omega) \to \Semiring$ as follows.
-Given a word $w \in {\hat\Omega}^*$, the projection of $w$ onto $\Omegai$,
-denoted $w|_\Omegai$,
-is the word of $\Omegai^*$ obtained from $w$ by removing all symbols 
-in $\hat\Omega \setminus \Omegai$.
-Using this notation
-and the tree linearization operator defined in Section~\ref{sec:trees}, 
-$d$ is defined by:
-\begin{equation}
-d(s, t) = T\bigl(s, \lin(t)|_\Omegai \bigr)  \mathrm{~for~} s \in \Sigma^*, t\in \T(\Omega)
-\end{equation}
-
-\noindent 
-\emph{Symbolic weighted parsing} is the problem, 
-given the above input, 
-to find a tree $t \in \T(\Omega)$ %nested word $t \in \Omega$ 
-minimizing \( d(s, t) \otimes A(t)\)
-\wrt $\leq_\oplus$, 
-\ie such that: %Hence, it is the problem of finding 
-%
-\begin{equation}\label{eq:distance-lang}
-d(s, t) \otimes A(t) = \displaystyle\bigoplus_{t' \in \T(\Omega)} d(s, t') \otimes A(t') 
-\end{equation}
-%
-The measure expressed in~\eqref{eq:distance-lang} is called
-called the edit-distance between~$s$ and~$A$ in~\cite{Mohri03EDWA}.
-%
-%The input language can also be expressed as a \SWTA, or, 
-%as a particular case, as a weighted context-free grammar, 
-%converted in turn into a \SWVPA following Lemma~\ref{lem:SWTA}.
-%
-The problem of searching, in a WTA language, 
-the best parse tree matching a given input,  
-sometimes referred as~\emph{weighted parsing},
-corresponds to SW parsing in the case of finite alphabets
-and when the transducer $T$ characterizes identity
-see \eg~\cite{Goodman99SemiringParsing} 
-and~\cite{MorbitzVogler19weighted-parsing} for a more general 
-weighted parsing framework.
-%
-%Indeed, it corresponds to the case where $T$
-%accepts only the pairs $\<s, t>$ such that 
-%$s$ is the projection of $t$ on $\Omegai$. 
-%This can be done with a single state $q$ and 
-%with transition rules of the form:
-%\begin{description}
-%\item[] $\wei(q, \varepsilon, a, q) = \one$ for all $a \in \Omegac \cup \Omegar$,
-%\item[] $\wei(q, a, a, q) = \one$ for all $a \in \Omegai$,
-%\item[] $\wei(q, a, b, q) = \zero$ for all $a, b \in \Omegai$, $a \neq b$.
-%\end{description}
-
-%\subsection{Computation}
-%
-\begin{proposition}
-The problem of Symbolic Weighted  parsing 
-can be solved in PTIME in the size of the input \SWT $T$, \SWTA $A$ %\SWVPA (or ) 
-and input word $s$, 
-and the computation time of the functions of the label theory.
-\end{proposition}
-\begin{proof} (sketch)
-We follow a \emph{Bar-Hillel} construction, also called parsing by intersection.
-
-\noindent
-We first extend the \SWT $T$ over $\Sigma$, $\Omegai$, $\Semiring$, and $\bar\Phi$, 
-into a \SWT $T'$ over $\Sigma$ and $\hat\Omega$ (and the same semiring and label theory),
-such that for all $s \in \Sigma^*$, and $u \in {\hat\Omega}^*$, 
-$T'(s, u) = T(s, u|_{\Omegai})$. The transducer $T'$ simply skips every symbol 
-$b \in {\hat\Omega} \setminus \Omegai$, 
-by the addition to the transition of $T$,
-of new transitions of the form $\wei_{01}(q, \varepsilon, b, q')$.
-
-\noindent
-Then, given an input word $s \in \Sigma^*$, 
-we compute the \SWA $A_{T', s}$, using Proposition~\ref{prop:epsilon}.
-This automaton is such that for all $t \in \T(\Omega)$, 
-\[ 
-   A_{T', s}\bigl(\lin(t)\bigr) 
- = T'\bigl(s, \lin(t)\bigr) 
- = T'\bigl(s, \lin(t)|_{\Omegai}\bigr) 
- = d(s, t).
-\]
-
-\noindent
-Next, we convert the input \SWTA $A$ over $\Omega$
-into a \SWVPA $A'$ over $\hat\Omega$, using Proposition~\ref{lem:SWTA}, 
-and we compute the \SWVPA $A_{T', s} \otimes A'$, 
-using Proposition~\ref{prop:SWVPA-product}.
-
-\noindent
-It remains to compute a best nested-word $w \in {\hat\Omega}^*$ 
-using the best-search procedure of Proposition~\ref{th:best-search},
-and convert it into a best tree in $\T(\Omega)$ in order to solve SW parsing
-for~$T$, $A$ and~$s$.
-\qed  
-\end{proof}
-
-\paragraph{Application to Automated Music Transcription.}
-%\input{transcription}
-...
-
-\section*{Conclusion}
-% summary 
-We have introduced weighted language models (SW transducers and visibly pushdown automata)
-computing over infinite alphabets, 
-and applied them to the problem of parsing 
-with infinitely many possible input symbols (typically timed events).
-%handled with suitable language formalisms, 
-%
-This approach extends conventional parsing and weighted parsing
-by computing a derivation tree modulo 
-a generic distance between words,  
-defined by a SW transducer given in input.
-This enables to consider finer word relationships than strict equality, 
-%as in the conventional parsing approach, 
-opening possibilities of quantitative analysis via this method.
-
-% discussion
-\noindent
-Ongoing and future work include
-
-\noindent
--- The study of other theoretical properties of SW models, 
-such as the extension of the best search algorithm from $1$-best to $n$-best~\cite{Huang05kbest}, 
-and to $k$-\emph{closed} semirings~\cite{Mohri02semiring}
-(instead of \emph{bounded}, which corresponds to $0$-\emph{closed}).
-
-\noindent
--- ...there is room to improve the complexity bounds for the algorithms
-... modular approach with oracles ...
-
-\noindent
--- present here an offline algorithm for best search, 
-semi-online implementation for AMT (bar-by-bar approach)
-with an on-the-fly automata construction.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% BIBLIO                                                                     %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\bibliographystyle{plain}
-%\bibliographystyle{abbrv}
-\bibliographystyle{plainurl}  % LIPICS
-%\bibliographystyle{splncs04} % LLNCS
-%\bibliographystyle{eptcs}    % EPTCS
-%\bibliography{generic}
-
-\bibliography{references}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% APPENDIX                                                                   %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\appendix
-
-\end{document}
-
 \section{Edit-Distance}

### 2021-06-16 introduction.tex
comments Masahiko

@@
 % !TEX root = main.tex
-%
+%\todo
 % Introduction
@@
 (see~\cite{Segoufin06csl} for a survey),
-\marginpar{\tiny register: skip refs and details, add Mikolaj}
+\mynote{register: skip refs and details, add Mikolaj}
 with pebbles for marking positions~\cite{NevenSchwentickVianu04FSMinfinite}, 
@@
 (Bar-Hillel construction~\cite{GruneJacobs08parsing}) 
-\marginpar{\tiny chap. intersection in~\cite{GruneJacobs08parsing}} 
+\mynote{chap. intersection in~\cite{GruneJacobs08parsing}} 
 where, given a \SWT $T$, a \SWVPA $A$, and an input word $s$, 

### 2021-06-16 main.tex
comments Masahiko

@@
 \input{main-lipics}
+
+% comments in margin with \todo command
+\usepackage[colorinlistoftodos]{todonotes} % option disable to hide todos
+% command todo : options
+% - noline: no line connecting the note with the place in the text where the note occurs in the latex code
+% - fancyline : curved arrow from note to inssertion point
+% - inline: place a todonote inside the text instead of in the margin
+% - author=name
+\newcommand{\mynote}[1]{\todo[noline,size=\tiny,color=yellow!40]{#1}}
 
@@
 $\mathsf{in} : Q \to \Semiring$   %\Phi_{\Sigma, \Delta}
-respectively $\mathsf{out} : Q \to \Semiring$  %\Phi_{\Sigma, \Delta}
-are functions defining the weight for entering, 
-respectively leaving, computation in a state,  
+(respectively $\mathsf{out} : Q \to \Semiring$)  %\Phi_{\Sigma, \Delta}
+are functions defining the weight for entering
+(respectively leaving) computation in a state,  
 and $\bar{\wei}$ is a triplet of transition functions 
@@
 One timestamp $d_i \in \Sigma$ may corresponds 
-to one notated event $\<\mathsf{e}, d'_i> \in \Sigma$, in which case 
+to one notated event $\<\mathsf{e}, d'_i> \in \Delta$, in which case 
 the weight value computed by the \SWT is the time distance between both
@@
 %
-In this case, the label theory $\bar\Phi$ can be reduced to a singleton $\< \Phi_\Sigma>$.
-%over $\Sigma$ is reduced to
-%a set $\Phi_\Sigma$ closed under~$\oplus$ and~$\otimes$.
+%In this case, the label theory $\bar\Phi$ can be reduced to a singleton $\< \Phi_\Sigma>$.
 %
@@
 where $Q$ is a finite set of states, 
-$\mathsf{in} : Q \to \Semiring$, %\Phi_\Sigma$, 
-respectively $\mathsf{out} : Q \to \Semiring$  %\Phi_\Sigma,$
-are functions defining the weight for entering, 
-respectively leaving, computation in a state, 
+$\mathsf{in} : Q \to \Semiring$ %\Phi_\Sigma$, 
+(respectively $\mathsf{out} : Q \to \Semiring$)  %\Phi_\Sigma,$
+are functions defining the weight for entering
+(respectively leaving) computation in a state, 
 and ${\wei_1}$ is a transition functions 
@@
 and $\bar\Phi$ effective,
-and a $\SWTA$ $A$ over $\Sigma$, $\Semiring$ and $\bar\Phi$,
+and a $\SWA$ $A$ over $\Sigma$, $\Semiring$ and $\bar\Phi$,
 there exists an effectively constructible \SWA 
@@
 \) 
-\marginpar{\tiny proof correctness}
+\mynote{proof correctness}
 %\qed
@@
 %of~$T$ is its number of states $|Q|$.
-\marginpar{\tiny revise with nb of tr. and states}
+\mynote{revise with nb of tr. and states}
 
@@
 $P$ is a finite set of stack symbols, 
-$\mathsf{in} : Q \to \Semiring$, 
-respectively $\mathsf{out} : Q \to \Semiring,$
-are functions defining the weight for entering, 
-respectively leaving, a state, 
+$\mathsf{in} : Q \to \Semiring$ 
+(respectively $\mathsf{out} : Q \to \Semiring$)
+are functions defining the weight for entering
+(respectively leaving) a state, 
 and $\bar\wei$ is a sextuplet composed of the transition functions:
@@
 structured words with timed symbols...
-\marginpar{\tiny todo}
+\mynote{todo example VPA}
 intro language of music notation? (markup = time division, leaves = events etc)
@@
 commutative, bounded, and complete,  
-\marginpar{\tiny total?}
-and assume an effective label theory.
+\mynote{total?}
+and that $\bar\Phi$ is an effective label theory.
 %s.t. for all $\phi \in \Phii$, 
@@
 since~$\Semiring$ is complete.
-\marginpar{\tiny complete} 
+\mynote{complete **} 
 ** effectively computable by hypothese that the label theory is effective**
@@
 and each iteration has a time complexity $O(|Q|^2 . |P|)$.
-\marginpar{\tiny detail with nb tr. and states}
+\mynote{detail with nb tr. and states}
 That gives a time complexity $O(|Q|^4 . |P|)$. 
@@
 Let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a 
-commutative, bounded, and complete \marginpar{\tiny total?} semiring  
+commutative, bounded, and complete \mynote{total?} semiring  
 and let $\bar\Phi$ be an effective label theory over $\Semiring$,
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% TODO                                                                       %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newpage
+\listoftodos
+
 \end{document}
-
-\section{Edit-Distance}
-
-%\subsection{Distance between words or languages}
-...algebraic definition of edit-distance of Mohri, in \cite{Mohri03EDWA}
-% Mehryar Mohri 
-% Edit-distance of weighted automata: General definitions and algorithms
-% International Journal of Foundations of Computer Science 14.06 (2003): 957-982.
-distance $d$ over $\Sigma^* \times \Sigma^*$ 
-into a semiring  $\Semiring = ( \Semiring, \oplus, \zero, \otimes, \one)$.
-
-%\noindent
-Let $\Omega = \Sigma \cup \{ \varepsilon \} \times \Sigma \cup \{ \varepsilon \} \setminus \{ (\varepsilon, \varepsilon) \}$,
-and let $h$ be the morphism from $\Omega^*$ into $\Sigma^* \times \Sigma^*$  
-defined over the concatenation of strings of $\Sigma^*$ (that removes the $\varepsilon$'s).
-%
-\noindent
-An \emph{alignment} between 2 strings  $s, t \in \Sigma^*$ is an element $\omega \in \Omega^*$ 
-such that $h(\omega) = (s, t)$.
-%
-\noindent
-We assume a base cost function $\Omega$ : $\delta: \Omega \to S$, extended to $\Omega^*$ as follows  
-(for $\omega \in \Omega^*$): 
-\(
-\displaystyle\delta(\omega) = \bigotimes_{0 \leq i < |\omega|} \delta(\omega_i)
-\).
-
-\noindent
-\begin{definition}
-For  $s, t \in \Sigma^*$, the edit-distance between $s$ and $t$ is  
-\( 
-d(s, t) = \displaystyle\bigoplus_{\omega \in \Omega^*\, h(\omega) = (s, t)} \delta(\omega)
-\).
-\end{definition}
-
-e.g. Levenstein edit-distance: $S$ is min-plus and $\delta(a, b) = 1$ for all $(a, b) \in \Omega$.
-
-%\paragraph{Distance between a word and a regular language}

### 2021-07-01 introduction.tex
1 passe intro legere

@@
 is the problem %process 
-of structuring a linear representation in input %typically
-(a finite word) according to a language model. % (a formal grammar).
+of structuring a linear representation on input %typically
+(a finite word), according to a language model. % (a formal grammar).
 % natural language, programming language, 
@@
 assume a finite and reasonably small input alphabet. %models and algorithms 
-Such a restriction makes perfectly sense in the context of NLP tasks 
-like constituency parsing or of programming languages compilers or interpreters.
-Considering large or infinite alphabets can however be of practical interest. %in other cases.
+Such a restriction makes perfect sense in the context of NLP tasks such as constituency parsing,
+or of programming languages compilers or interpreters.
+Considering large or infinite alphabets can however be of practical interest, %in other cases.
 %
-For instance, when dealing with large characters encodings such as UTF-16, % processing strings in 
+for instance, when dealing with large characters encodings such as UTF-16, % processing strings in 
 \eg for vulnerability detection in Web-applications~\cite{dAntoni21CACM}, 
@@
 of data streams or serialization of structured documents 
-(which may contain textual or numerical attributes)~\cite{Segoufin06csl}, 
+(with textual or numerical attributes)~\cite{Segoufin06csl}, 
 or for processing timed execution traces~\cite{Bouyer03algebraic}.
@@
 % (notion of monoid recognizability, based on registers, comparable to Bojancszik et al. data words)
-The latter is related to a case study that motivated the present work: 
+%
+The latter case is related to a study that motivated the present work: 
 automated music transcription.
-In this problem, a symbolic representation of a music performance, % a symbolic representation of a music performance 
-in the form of a sequence of timed musical events, % timestamped
-is converted into a structured score in Common Western Music Notation~\cite{Gould11Notation}.
+In this problem, a music performance, % a symbolic representation of a music performance 
+represented symbolically in the form of a sequence of timed musical events, % timestamped
+is converted into a score in Common Western Music Notation~\cite{Gould11Notation},
+structured according to nested grouping and metric strength of events.
+%(the structure represents hierarchical  grouping of events, and their relative temporal positions)
 It can therefore be stated as a parsing problem~\cite{foscarin:hal-01988990}, 
@@
 (see~\cite{Segoufin06csl} for a survey),
-\mynote{register: skip refs and details, add Mikolaj}
+\mynote{register: skip refs and details, add Mikolaj recent}
 with pebbles for marking positions~\cite{NevenSchwentickVianu04FSMinfinite}, 
@@
 %...For the representation of  in model checking, verification and 
+%
 The automata at the core of model checkers
@@
 the transitions are guarded by predicates over infinite alphabet domains.
-With closure conditions on the sets of such predicates, % (alphabet theories), 
+With appropriate closure conditions on the sets of such predicates, % (alphabet theories), 
 all the good properties enjoyed by automata over finite alphabets are preserved.
@@
 Other extensions of language models  %(automata and grammars) 
-help in dealing with non-determinism, with the computation of weight values. 
-With an ambiguous grammar, there may exists several derivations 
-(\emph{abstract syntax trees} -- AST), % \ie the result of an analyze.
+help in dealing with non-determinism, by the computation of weight values. 
+With an ambiguous grammar, there may exist several derivations 
+(\emph{abstract syntax trees} -- AST) % \ie the result of an analyze.
 yielding one input word. % structuring a word, 
@@
 and the rule's weights can be combined with a associative product operator~$\otimes$ into 
-the weight of a AST.
+the weight of an AST.
 A second operator~$\oplus$, associative and commutative, 
-is moreover used to handle ambiguity of the model, 
+is moreover used to handle the ambiguity of the model, 
 by summing the weights of the possibly several (in general exponentially many) AST %syntax trees 
@@
 In this paper, we present a uniform framework for weighted parsing over infinite input alphabets.
-It is based on \emph{symbolic weighted} finite states language models (\swM), generalizing 
-both~\SA, with functions into an arbitrary semiring instead of Boolean guards, 
+It is based on \emph{symbolic weighted} finite states language models (\swM), 
+generalizing~\SA, with functions into an arbitrary semiring instead of Boolean guards, 
 and~\WA, by handling infinite alphabets, see Figure~\ref{fig:hierarchy}.
@@
 % non-terminals and production rules).
-In the transition rules of \swM models, input symbols appear as variables %(parameters)
+In the transition rules of \swM models, input symbols appear as variables, %(parameters)
 and the weight associated to a transition rule is a function of these variables.
@@
 The models presented here are finite automata called symbolic-weighted (\SWA),
-transducers (\SWT) and pushdown automata,
+transducers (\SWT), and pushdown automata
 with a visibly restriction~\cite{AlurMadhusudan09nested} (\SWVPA).
@@
 On the other hand, 
-a \SWT is used to define a distance $T(s, t)$ between two finite words $s$ and $t$
+a \SWT can define a distance $T(s, t)$ between finite words $s$ and $t$
 over infinite alphabets. % following~\cite{Mohri03EDWA}.
@@
 In the second step, a best AST $t$ is found by applying to $A_{T, s}$ 
-a best search algorithm similar to shortest distance in graphs~\cite{Mohri02semiring,Huang05kbest}.
+a best search algorithm similar to the shortest distance 
+in graphs~\cite{Mohri02semiring,Huang05kbest}.
 %In expressiveness, they are equivalent to weighted CFG. 
@@
 % value in a semiring...
+\mynote{expressiveness: VPA have restricted equality test. 
+        comparable to pebble automata? $\to$ conclusion}
 

### 2021-07-02 introduction.tex
modifs Lydia

@@
 (see~\cite{Segoufin06csl} for a survey),
-\mynote{register: skip refs and details, add Mikolaj recent}
+\florent{register: skip refs and details, add Mikolaj recent}
 with pebbles for marking positions~\cite{NevenSchwentickVianu04FSMinfinite}, 
@@
 (Bar-Hillel construction~\cite{GruneJacobs08parsing}) 
-\mynote{chap. intersection in~\cite{GruneJacobs08parsing}} 
+\florent{chap. intersection in~\cite{GruneJacobs08parsing}} 
 where, given a \SWT $T$, a \SWVPA $A$, and an input word $s$, 
@@
 % value in a semiring...
-\mynote{expressiveness: VPA have restricted equality test. 
+\florent{expressiveness: VPA have restricted equality test. 
         comparable to pebble automata? $\to$ conclusion}

### 2021-07-02 main.tex
macros commentaires
modifs Lydia

@@
 % - author=name
-\newcommand{\mynote}[1]{\todo[noline,size=\tiny,color=yellow!40]{#1}}
+\newcommand{\florent}[1]{\todo[noline,size=\tiny,color=yellow!40]{#1}}
+\newcommand{\lydia}[1]{\todo[noline,size=\tiny,color=red!30]{#1}}
+\newcommand{\philippe}[1]{\todo[noline,size=\tiny,color=green!30]{#1}}
 
@@
 %over finite alphabets, see  e.g.~\cite{Mohri03EDWA}, 
-by labelling each transition with a weight function (instead of a simple weight value), 
+by labeling each transition with a weight function (instead of a simple weight value), 
 that takes the input and output symbols as parameters. 
 These functions are similar to the guards of symbolic automata~\cite{dAntoniVeanes17CAV,dAntoni21CACM},
-but they can return values in an generic semiring, 
+but they can return values in a generic semiring, 
 whereas the latter guards are restricted to the Boolean semiring.
@@
 defined recursively, for every states $q, q' \in Q$, 
-and every strings $\< s, t> \in \Sigma^* \times \Delta^* \setminus \{ \< \varepsilon, \varepsilon >\}$, by:
+and every strings $\< s, t> \in \Sigma^* \times \Delta^* \setminus \{ \< \varepsilon, \varepsilon >\}$ considered as concatenation
+of the symbol $a \in \Sigma$ (resp. $b \in \Delta$) with a word $u \in \Sigma^*$ (resp. $v in \Delta^*$), by :
+\lydia{added $u$ and $v$}
+
 %$a \in \Sigma$, $u \in \Sigma^*$, $b\in \Delta$, $v\in \Delta^*$, by:
@@
 \end{align}
+
 %
@@
 the deletion of the symbol~$a \in \Sigma$ in~$s$, 
-$\wei{01}(q, \varepsilon, b, r)$ is the cost 
+$\wei_{01}(q, \varepsilon, b, r)$ is the cost 
 of the insertion of~$b \in \Delta$ in $t$, 
@@
 is the product, with $\otimes$, of the individual costs of the operations involved;
-And the distance between $s$ and $t$ is the sum, with $\oplus$,
+and the distance between $s$ and $t$ is the sum, with $\oplus$,
 of all possible products.
@@
 In Common Western Music Notation~\cite{Gould11Notation}, 
-several symbols may be used to represent a unique sounding event.
+several symbols may be used to represent one single sounding event. 
+\lydia{unique $\to$ similar}
+\florent{similar $\to$ single}
 For instance, several notes can be combined with a tie, 
 like in $\musQuarter\!\!\!\mathrel{\raisebox{-1.5mm}{$\smile$}}\!\!\!\musEighth$,  
-and one note can be augmented by half its duration with a dot like in~\musQuarterDotted{}
-
-\noindent 
-We propose a small transducer model that compares an input sequence of sounding 
+and one note can be augmented by half its duration with a dot like in~\musQuarterDotted{}.
+\lydia{modif.}
+These notations are perceived equivalent when played, as their duration is equal, yet the notation is different.
+We thus want to be able to recognize all the possible notations of equivalent durations of notes,
+when comparing a written score and a music performance.
+
+\noindent 
+We propose a small weighted transducer model that computes a distance between an input sequence of sounding 
 events (music "performance") 
@@
 is to compare a recorded performance $s \in \Sigma^*$
-with notated music sheet $t \in \Delta^*$.
-One timestamp $d_i \in \Sigma$ may corresponds 
+with a notated music sheet $t \in \Delta^*$.
+One timestamp $d_i \in \Sigma$ may correspond 
 to one notated event $\<\mathsf{e}, d'_i> \in \Delta$, in which case 
@@
 If $\<\mathsf{e}, d'_i>$ is followed by continuations 
-$\<\mathsf{-}, d'_{i+1}>$..., they are just skip with no cost (transitions $\wei_{01}$ or weight $\one$).
+$\<\mathsf{-}, d'_{i+1}>$..., they are just skipped with no cost (transitions $\wei_{01}$ or weight $\one$).
 %transitions
@@
 %
-Moreover, it may happen that the performers plays an extra note accidentally, but only once in a row. 
+We also must able to take performing errors into account, while still being able to compare with the score, 
+since a performer could, for example, play an unwritten extra note. 
+\lydia{modif}
 This is modelled by the transition $\wei_{10}$ with an arbitrary weight value $\alpha \in \Semiring$, 
@@
 Moreover, $\wei'_1$ also needs to simulate transitions 
-defined by $\wei_{10}$: $T$ reading in intput word and not in output word.
+defined by $\wei_{10}$: $T$ reading in input word and not in output word.
 Since $B_{T, A}$  will read only in the output word, such a transition corresponds
@@
 \) 
-\mynote{proof correctness}
+\florent{proof correctness}
 %\qed
@@
 %of~$T$ is its number of states $|Q|$.
-\mynote{revise with nb of tr. and states}
+\florent{revise with nb of tr. and states}
 
@@
 that we assume partitioned into three 
-subsets~$\Omegai$, $\Omegac$, $\Omegac$,
+subsets~$\Omegai$, $\Omegac$, $\Omegar$,
 whose elements are respectively called 
@@
 (respectively leaving) a state, 
-and $\bar\wei$ is a sextuplet composed of the transition functions:
+and $\bar\wei$ is a sextuplet composed of the transition functions :
 $\weii : Q \times P \times Q \to \Phici$,  
@@
 The intuition is the following for the above transitions. 
+$\weiei$, $\weiec$, $\weier$ describe the cases where the stack is empty. 
+\lydia{moved this to the beginning}
 %
@@
 ${\call{c}} \in \Omegac$ and $p \in P$ on the top of the stack 
-($c$ is compared to $a$ by the weight function $\eta_\mathsf{ci} \in \Phici$)
-and $\weiei$, $\weiec$, $\weier$ apply if and only if the stack is empty.
+($c$ is compared to $a$ by the weight function $\eta_\mathsf{ci} \in \Phici$).
 %
@@
 Hence, $\weight_A$ is a function from 
-$[Q \times \Gamma^*] \times \Omega^* \times [Q \times \Gamma^*]$ into~$\Semiring$
-(the empty stack is denoted by $\bot$). 
+$[Q \times \Gamma^*] \times \Omega^* \times [Q \times \Gamma^*]$ into~$\Semiring$.
+The empty stack is denoted by $\bot$, and the upmost symbol is the last pushed content. 
+The following functions illustrate each of the possible cases, being : reading $a \in \Omegai$, 
+\lydia{intro to func}
+or $\call{c} \in \Omegac$, or $\return{r} \in \Omegar$ for each possible state of the stack (empty or not), 
+to add to $u \in {\Omega}^*$.
+
 %
@@
 structured words with timed symbols...
-\mynote{todo example VPA}
+\florent{todo example VPA}
 intro language of music notation? (markup = time division, leaves = events etc)
@@
 commutative, bounded, and complete,  
-\mynote{total?}
+\florent{total?}
 and that $\bar\Phi$ is an effective label theory.
@@
 %More precisely, 
-Let $b_\bot : Q \times Q \to \Semiring$ be such that: %the function:
+We distinguish two cases: when the stack is empty, and when it is not. 
+\lydia{2 cases of $b$}
+In the case of an empty stack, let $b_\bot : Q \times Q \to \Semiring$ be such that: %the function:
 %
@@
   \textstyle
-  \weight_A\bigl(\config{q}{\bot}, s, \config{q'}{\bot}\bigr)
+  \weight_A\bigl(\config{q}{\bot}, s, \config{q'}{\bot}\bigr).
 \end{equation}
@@
 \noindent
-Let $\top$ be a fresh stack symbol which does not belong to $\Gamma$,
+If the stack is not empty, let $\top$ be a fresh stack symbol which does not belong to $\Gamma$,
 and let $b_\top : Q \times P \times Q \to \Phic$ be such that,
@@
 and ending in state $q'$ with the same stack,
-such that the computation does not pop 
+such that the computation can not pop 
 the pair made of $c$ and $p$ at the top of this stack,
-but may read these symbols.
+but may only read these symbols.
 Moreover, $A$ may push another pair $\< c', p'>$ %call symbols 
@@
 %\noindent\quad
-such that $d_\bot(q_1, q_2)$, resp. $\bigoplus_{c\in \Omegac} d_\top(q_1, p, q_2)(c)$, is minimal in $\Semiring$ wrt $\leq_\oplus$$\;$
+such that $d_\bot(q_1, q_2)$, resp. $\bigoplus_{c\in \Omegac} d_\top(q_1, p, q_2)(c)$, is minimal in $\Semiring$ \wrt $\leq_\oplus$$\;$
 
@@
 since~$\Semiring$ is complete.
-\mynote{complete **} 
-** effectively computable by hypothese that the label theory is effective**
+\florent{complete **} 
+** effectively computable by hypothesis that the label theory is effective**
 The algorithm performs $2.|Q|^2$ iterations until $P$ is empty, 
 and each iteration has a time complexity $O(|Q|^2 . |P|)$.
-\mynote{detail with nb tr. and states}
+\florent{detail with nb tr. and states}
 That gives a time complexity $O(|Q|^4 . |P|)$. 
@@
 Let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a 
-commutative, bounded, and complete \mynote{total?} semiring  
+commutative, bounded, and complete \florent{total?} semiring  
 and let $\bar\Phi$ be an effective label theory over $\Semiring$,

### 2021-07-02 main.tex
commentaires florent

@@
 
+\florent{notation $cp$ for $\< c, p>$ ?}
 %
@@
 If the stack is not empty, let $\top$ be a fresh stack symbol which does not belong to $\Gamma$,
+\florent{$b_\top$ : mot bien parenthèsé $c$/$r$}
 and let $b_\top : Q \times P \times Q \to \Phic$ be such that,
@@
 \noindent
+\florent{explication Fig.~\ref{fig:best-update} suivant cas de \eqref{eq:SWVPA-weight}}
 The infinite sums in the updates of $d$ in Algorithm~\ref{algo:Dijkstra}, 
@@
 %...
+\florent{2 lines Application to Automated Music Transcription: 
+implementation $\neq$ but same principle, on-the-fly automata construction
+during best search, for efficiency.}
 
@@
 Ongoing and future work include 
+\florent{TODO future work}
 

### 2021-07-02 main.tex
modifs Lydia

@@
 of the symbol $a \in \Sigma$ (resp. $b \in \Delta$) with a word $u \in \Sigma^*$ (resp. $v in \Delta^*$), by :
-\lydia{added $u$ and $v$}
+\lydia{added u and v def}
 
@@
 These notations are perceived equivalent when played, as their duration is equal, yet the notation is different.
-We thus want to be able to recognize all the possible notations of equivalent durations of notes,
-when comparing a written score and a music performance.
-
-\noindent 
-We propose a small weighted transducer model that computes a distance between an input sequence of sounding 
+We thus want to be able to compare a music score with music played by a performer. 
+\lydia{changed end}
+%
+\noindent 
+We propose a small weighted transducer model that calculates the distance bewteen an input sequence of sounding 
 events (music "performance") 
@@
 since a performer could, for example, play an unwritten extra note. 
-\lydia{modif}
+\lydia{reformulated this sentence} 
+%
 This is modelled by the transition $\wei_{10}$ with an arbitrary weight value $\alpha \in \Semiring$, 
@@
 $\init(q_1) = \final(q_1) = \zero$.
+
+\lydia{ccl to the ex}
+That way, an \SWT is capable of evaluating the differences between a score and a performance, 
+all the while ensuring that performance errors are plausible.
+
 %$\init(q_0, d, b) = \final(q_0, d, b) = \one$, and 
@@
 The intuition is the following for the above transitions. 
-$\weiei$, $\weiec$, $\weier$ describe the cases where the stack is empty. 
+$\weiei$, $\weiec$, and $\weier$ describe the cases where the stack is empty. 
 \lydia{moved this to the beginning}
@@
 to add to $u \in {\Omega}^*$.
-
+\lydia{introduced the 6 cases}
 \florent{notation $cp$ for $\< c, p>$ ?}
-%
+
 \begin{align}
@@
 %\mathrm{~if~} q \neq q'\nonumber\\
+%
 \weight_A\bigl(\config{q}{\bot}, \varepsilon, \config{q'}{\bot}) & = \one
@@
   \weiec(q, {\call{c}}, p, q'') 
-  \otimes \weight_A\bigl(\config{q''}{{\call{c}}\, p}, u, 
+  \otimes \weight_A\bigl(\config{q''}{\<{\call{c}}, p>}, u, 
                          \config{q'}{\gamma'}\bigr)\nonumber\\
 %
-\weight_A\bigl(\configup{q}{\<{\call{c}}, p>\stackup \gamma}, {\return{r}}\, u, 
+\weight_A\bigl(\configup{q}{\<{\call{c}}, p>\stackup\gamma}, {\return{r}}\, u, 
                \config{q'}{\gamma'}\bigr) & =  
@@
 \end{align}
+\lydia{c p to <c, p>} 
 %
@@
 %More precisely, 
-We distinguish two cases: when the stack is empty, and when it is not. 
-\lydia{2 cases of $b$}
-In the case of an empty stack, let $b_\bot : Q \times Q \to \Semiring$ be such that: %the function:
+We distinguish two cases : when the stack is empty, and when it is not. 
+\lydia{introduced 2 cases for b}
+%
+In the case of an empty stack, let $b_\bot : Q \times Q \to \Semiring$ be such that : %the function:
 %
@@
 of the function $\weight_A$.
+\lydia{so ?}
 

### 2021-07-05 introduction.tex
Commentaires Philippe

--- 2021-07-01 introduction.tex
@@
 automated music transcription.
+\mynote{I think that we can make the case for a larger set of situations: given a linear 
+music notation input, for instance elements in an XML file, we we want to structure
+the input according to a hierarchical rhythmic space. I can elaborate...}
 In this problem, a music performance, % a symbolic representation of a music performance 

### 2021-07-05 introduction.tex

--- 2021-07-02 introduction.tex
@@
 automated music transcription.
+\philippe{I think that we can make the case for a larger set of situations: given a linear 
+music notation input, for instance elements in an XML file, we we want to structure
+the input according to a hierarchical rhythmic space. I can elaborate...}
 In this problem, a music performance, % a symbolic representation of a music performance 

### 2021-07-05 semiring.tex
pass for new format
revise section parsing
commentaires florent

@@
 We shall consider semirings for the weight values of our language models.
+\florent{The results are established for a general class of semirings. They can be instantiated for concrete cases}
 %
@@
 by definition of $\leq_\oplus$, $\one \oplus x = \one$.
-\qed
+%\qed
 \end{proof}
@@
 both sides by $x$ and distributing. 
-\qed
+%\qed
 \end{proof}
@@
 
-\begin{example}
-Figure~\ref{fig:semirings} presents examples of semirings interesting in practice 
-and enjoying the above properties.
-\end{example}
+%\begin{example}
+%Figure~\ref{fig:semirings} presents examples of semirings interesting in practice 
+%and enjoying the above properties.
+%\end{example}
 
@@
 \end{figure}
+\florent{results of this paper: for semirings commutative, bounded, total and complete}
 

### 2021-07-05 introduction.tex
Philippe sectons 1-2

@@
 and the weight associated to a transition rule is a function of these variables.
+\philippe{This sentence (symbols as variables) is not immediatly clear to me. Maybe a short example or intuition?}
 %
@@
 %\wrt an edit distance
-The latter model of automata computes on \emph{nested words}~\cite{AlurMadhusudan09nested}, 
+The latter model of automata operates on \emph{nested words}~\cite{AlurMadhusudan09nested}, 
 a structured form of words parenthesized with markup symbols, 
@@
 More precisely, a \SWVPA $A$ associates a weight value $A(t)$ % \in \Semiring$ 
-to a given a nested word $t$, which is the linearization of an AST. %representing a parse tree.
+to a given nested word $t$, which is the linearization of an AST. %representing a parse tree.
+\philippe{The weight is a label on edges in the derivation tree, right? Not sure I understand 
+the sentence}
 %
@@
 finding $t$ minimizing 
-$T(s, t) \otimes A(t)$ (\wrt the ranking defined by $\oplus$).
+$T(s, t) \otimes A(t)$ (\wrt the ranking defined by $\oplus$), given an input word $s$.
 The latter value is called the distance between $s$ and $A$ in~\cite{Mohri03EDWA}.
+\philippe{You mean the distance between $s$ and $t$? "The latter value" is a bit ambiguous}
 %
@@
 a \SWVPA $A_{T, s}$ is built, such that for all $t$, $A_{T, s}(t) = T(s, t) \otimes A(t)$.
+\philippe{The notation  $A_{T, s}$ has  not been introduced so far. It is not clear why $T$ is a parameter there}
 In the second step, a best AST $t$ is found by applying to $A_{T, s}$ 
@@
 in graphs~\cite{Mohri02semiring,Huang05kbest}.
+\philippe{OK, à quelques détails (pour moi), tout ça est très bien expliqué}
 %In expressiveness, they are equivalent to weighted CFG. 

### 2021-07-05 semiring.tex
Philippe sectons 1-2

@@
 %
+\philippe{There is sometimes a confusion in the text between the struture and the domain $\Semiring$. Not
+essential}.
 A \emph{semiring} $\< \Semiring, \oplus, \zero, \otimes, \one>$ 
@@
 \noindent
+\philippe{Here the difference between $\Semiring$ as a structure and as a domain is blurred.}
 We shall need below infinite sums with~$\oplus$.
@@
 of elements of $\dom(\Semiring)$ over an index set $I \subset \mathbb{N}$, such that:
+\philippe{$j\in \mathbb{N}$: $j$ is en element of $\mathbb{N}$, not the same s $j\subset \mathbb{N}$}
 %the following holds: %properties

### 2021-07-06 main.tex
Fin première lecture P.

@@
 as functions of  
+\philippe{I missed sth: what is this $\varepsilon$? Intuitively clear but not defined?}
 $Q \times (\Sigma \cup \{ \varepsilon \}) \times (\Delta \cup \{ \varepsilon \}) \times Q \to \Semiring$,
@@
 such that, for all $q, q' \in Q$, $a \in \Sigma$,  $b \in \Delta$,
+
 \[
@@
 defined recursively, for every states $q, q' \in Q$, 
-and every strings $\< s, t> \in \Sigma^* \times \Delta^* \setminus \{ \< \varepsilon, \varepsilon >\}$ considered as concatenation
-of the symbol $a \in \Sigma$ (resp. $b \in \Delta$) with a word $u \in \Sigma^*$ (resp. $v in \Delta^*$), by :
+and every pairs of strings $\< s, t> \in \Sigma^* \times \Delta^* \setminus \{ \< \varepsilon, \varepsilon >\}$,
+each  considered as the concatenation
+of the symbol $a \in \Sigma$ (resp. $b \in \Delta$) with a word $u \in \Sigma^*$ (resp. $v \in \Delta^*$), by :
 \lydia{added u and v def}
@@
 %
+\philippe{OK tout ça se lit bien :-)}
 Intuitively, using a transition $\wei_{ij}(q, a, b, q')$ means for $T$:
@@
 
+\philippe{Je crois qu'il faudrait numéroter les exemples indépendamment des définitions. Cet exemple est le premier
+qui donne des détails sur l'application visée. Il arrive peut-être un peu tard et est long. On pourrait introduire
+la motivation dans l'intro, et développer des petits exemples au fur et à mesure.}
 \begin{example}
@@
 %
-We also must able to take performing errors into account, while still being able to compare with the score, 
+We also must be able to take performing errors into account, while still being able to compare with the score, 
 since a performer could, for example, play an unwritten extra note. 
@@
 (respectively leaving) computation in a state, 
-and ${\wei_1}$ is a transition functions 
+and ${\wei_1}$ is a transition function 
 from $Q \times Q$ into~$\Phi_{\Sigma}$.
@@
 a synchronized product of transitions defined by $\wei_1$, 
-$\wei_{01}$ ($T$ reading in output word and not in input word)
-and $\wei_{11}$ ($T$ reading in output word and input word).
+$\wei_{01}$ ($T$ reading in output word and not an input word)
+and $\wei_{11}$ ($T$ reading both an input word and an output word).
 %
 Moreover, $\wei'_1$ also needs to simulate transitions 
-defined by $\wei_{10}$: $T$ reading in input word and not in output word.
+defined by $\wei_{10}$: $T$ reading in input word and not an output word.
 Since $B_{T, A}$  will read only in the output word, such a transition corresponds
@@
 \emph{internal}, \emph{call} and \emph{return} symbols.
+\philippe{Là je crois qu'il faudrait expliquer ces Omega, je commence à fatiguer et je suis un
+peu largué par toutes ces définitions. J'intuite qu'il s'agit des symboles, parenthèses ouvrantes
+et fermantes? Pourquoi il faut un alphabet pour les parenthèses?}
 % names are  coined by application to functional program verification
@@
 where $Q$ is a finite set of states, 
+\philippe{Est-ce que tout le monde sait ce qu'est un pushdown automata? Je suppose que c'est
+lié à la pile.}
 $P$ is a finite set of stack symbols, 
@@
 and $\Omegac$ and $\Omegar$ contain respectively one left and right parenthesis
+\philippe{Ah oui, ça aurait pu être dit avant.}
 $\ccall{b}$ and $\creturn{b}$ for each symbol $b$ labelling inner nodes (symbols of rank $>0$). 

### 2021-07-06 main.tex
comments
short pass section 2 (esp. labels)

@@
 %\title{Symbolic Weighted Language Models and Automated Music Transcription}
-\title{Symbolic Weighted Language Models and\\ Parsing over Infinite Alphabets}
+\title{Symbolic Weighted Language Models and\\ Quantitative Parsing over Infinite Alphabets}
 %\title{Weighted Visibly Pushdown Automata and Automated Music Transcription}
@@
 defined recursively, for every states $q, q' \in Q$, 
-and every pairs of strings $\< s, t> \in \Sigma^* \times \Delta^* \setminus \{ \< \varepsilon, \varepsilon >\}$,
-each  considered as the concatenation
-of the symbol $a \in \Sigma$ (resp. $b \in \Delta$) with a word $u \in \Sigma^*$ (resp. $v \in \Delta^*$), by :
-\lydia{added u and v def}
-
+and every pairs of strings $\< s, t> \in \Sigma^* \times \Delta^*$,
+% \setminus \{ \< \varepsilon, \varepsilon >\}$,
+where $au$, and $bv$, denote the concatenation
+of the symbol $a \in \Sigma$ (resp. $b \in \Delta$) 
+with a word $u \in \Sigma^*$ (resp. $v \in \Delta^*$).
+\lydia{added $u$ and $v$ def}
+%
 %$a \in \Sigma$, $u \in \Sigma^*$, $b\in \Delta$, $v\in \Delta^*$, by:

### 2021-07-06 semiring.tex
short pass section 2 (esp. labels)

@@
 A semiring $\Semiring$ is \emph{commutative} if $\otimes$ is commutative.
-It is \emph{idempotent} if for each $x \in \dom(\Semiring)$, $x \oplus x = x$.
+It is \emph{idempotent} if for all $x \in \Semiring$, $x \oplus x = x$.
 %
@@
 called the \emph{natural ordering} of~$\Semiring$~\cite{Mohri02semiring} 
-and defined,  by: 
+defined,  by: 
 %implicitly defined by the semiring $\Semiring$ 
-for all $x$ and $y$,
+for all $x, y\in \Semiring$,
 $x \leq_\oplus y \;\mbox{iff}\; x \oplus y = x$.
@@
 \ie when for all $x, y \in \Semiring$, either $x \oplus y = x$ or $x \oplus y = y$.
+\florent{is total necessary?}
 
@@
 \end{lemma}   
-When the property of Lemma~\ref{lem:monotonic} holds, 
+%When  holds, 
 %We say in this case that 
-$\Semiring$ is called  \emph{monotonic}. % \wrt $\leq_\oplus$.
+To express the property of Lemma~\ref{lem:monotonic}, we call
+$\Semiring$ \emph{monotonic} \wrt $\leq_\oplus$.
 %A semiring $\Semiring$ 
@@
 It implies that, when looking for a best path in a graph whose edges
-are weighted by values of $\Semiring$, the loops can be safely avoided
-(because, for all $x \in \Semiring$ and $n \geq 1$, 
- $x \oplus x^n = x \otimes (\one \oplus x^{n-1}) = x$).
+are weighted by values of $\Semiring$, the loops can be safely avoided,
+because, for all $x \in \Semiring$ and $n \geq 1$, 
+ $x \oplus x^n = x \otimes (\one \oplus x^{n-1}) = x$.
 

### 2021-07-07 main.tex
passe section finale parsing

@@
 there exists an effectively constructible \SWA 
-$B_{T, A}$ over $\Delta$, $\Semiring$ and $\bar\Phi$,
+$B_{A, T}$ over $\Delta$, $\Semiring$ and $\bar\Phi$,
 such that for all $t \in \Delta^*$, 
-$B_{T, A}(t) = \displaystyle\bigoplus_{s\in \Sigma^*} A(s) \otimes T(s, t)$.
+$B_{A, T}(t) = \displaystyle\bigoplus_{s\in \Sigma^*} A(s) \otimes T(s, t)$.
 \end{proposition}
@@
 \noindent
-The state set of $B_{T, A}$ will be $Q' = P \times Q$.
-The entering, leaving and transition functions of $B_{T, A}$ will 
+The state set of $B_{A, T}$ will be $Q' = P \times Q$.
+The entering, leaving and transition functions of $B_{A, T}$ will 
 simulate synchronized computations of $A$ and $T$, 
@@
 defined by $\wei_{10}$: $T$ reading in input word and not an output word.
-Since $B_{T, A}$  will read only in the output word, such a transition corresponds
+Since $B_{A, T}$  will read only in the output word, such a transition corresponds
 to an $\varepsilon$-transition of $\SWA$, 
@@
 
-%Every transition of $B_{T, A}$ will
+%Every transition of $B_{A, T}$ will
 %simulate a sequence of transitions of $T$ performing the following steps:
@@
 \noindent
-The construction time and size for $B_{T, A}$ are $O(\| T \|^3 . \| A \|^2)$,
+The construction time and size for $B_{A, T}$ are $O(\| T \|^3 . \| A \|^2)$,
 where the sizes $\| T \|$ and $\| A \|$ are their number of states.
@@
 there exists an effectively constructible \SWA 
-$B_{T, s}$ over $\Delta$, $\Semiring$ and $\bar\Phi$,
-such that for all $t \in \Delta^*$, $B_{T, s}(t) = T(s, t)$.
+$B_{s, T}$ over $\Delta$, $\Semiring$ and $\bar\Phi$,
+such that for all $t \in \Delta^*$, $B_{s, T}(t) = T(s, t)$.
 \end{corollary}
@@
 Let us now apply the models and results of the previous sections %in order to define 
-to the problem of parsing over infinite alphabet. %appropriate
+to the problem of parsing over an infinite alphabet. %appropriate
 %
@@
 %
-Let~$\Sigma$ be a countable input alphabet, 
-and~$\Omega = \Omegai \uplus \Omegac \uplus \Omegar$ be a countable output alphabet.
-%
-Let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a 
+Let~$\Sigma$ 
+and~$\Omega = \Omegai \uplus \Omegac \uplus \Omegar$ 
+be countable input and output alphabets,
+let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a 
 commutative, bounded, and complete \florent{total?} semiring  
@@
 %
-For all $s \in \Sigma^*$ and $t \in \Omega^*$, let 
-\(d(s, t) = T\bigl(s, t|_\Omegai \bigr)\), 
+For all $u \in \Sigma^*$ and $t \in \Omega^*$, let 
+\(d(u, t) = T\bigl(u, t|_\Omegai \bigr)\), 
 where $t|_\Omegai \in \Omegai^*$ is the projection of $t$ onto~$\Omegai$,
@@
 %
-Every labeled tree $t$ can be linearized into a nested word $\lin(t) \in \Omega^*$, 
-assuming \eg that $\Omegai$ contain the symbols labelling the leaves (symbols of rank $0$) 
-and $\Omegac$ and $\Omegar$ contain respectively one left and right parenthesis
-\philippe{Ah oui, ça aurait pu être dit avant.}
-$\ccall{b}$ and $\creturn{b}$ for each symbol $b$ labelling inner nodes (symbols of rank $>0$). 
-%
-With this representation, the projection $\lin(t)|_\Omegai$ is then the 
-sequence of leaves of $t$, enumerated in a \emph{dfs}-traversal. 
-%
-We show in Appendix~\ref{sec:trees} how to convert a ($\SW$) tree automaton~$A$
-into a $\SWVPA$ computing $A(\lin(t))$ for every tree $t$. 
-That also holds, for the set of ASTs of a weighted CF-grammar.
-%The purpose of the transducer $T$ is to measure a distance between input and output words, 
-%\ie between the word $s$ and the sequence leaves of a derivation tree, in $\Omegai^*$.
-%
-%that generates (weighted) trees by replacement of a state symbol~$q_0$ (non-terminal), 
-%by a tree $a(q_1,\ldots, q_k)$, where $k = \rank(a)$.
-%A replacement rule $q_0 \to a(q_1,\ldots, q_k)$, 
-%of weight $\wei(q_0, a, q_1 \ldots q_k) \in \Semiring$ 
-%according to Definition~\ref{def:SWTA}, 
-%corresponds to the production rule $q_0 := a(q_1,\ldots, q_k)$ of a weighted CF grammar,
-%with set non-terminal symbols $Q$ and set of terminal symbols $\Omega_0$.
-%
-%This actually is a slight generalization of CFG since 
-%each such production rule is labelled by  a symbol of $\Omega_{>0}$, 
-%hence parse trees %derivation trees
-%are trees of $\T(\Omega)$.
-%
-%Another (more original) generalization is that the set of terminal symbols 
-%$\Omega_0$ may be infinite.
-%
-%
-%The input language can also be expressed as a \SWTA, or, 
-%as a particular case, as a weighted context-free grammar, 
-%converted in turn into a \SWVPA following Lemma~\ref{lem:SWTA}.
-%
-Therefore, $\SW$-parsing generalizes 
-the problem of searching, the best derivation of a weighted CF-grammar  
-that yields a given input, sometimes referred as~\emph{weighted parsing},
-%corresponds to $\SW$-parsing in the case of finite alphabets
-%and when the transducer $T$ characterizes identity
-see \eg~\cite{Goodman99SemiringParsing} 
-and~\cite{MorbitzVogler19weighted-parsing} for a more general weighted parsing framework.
-%
-%weighted parsing and convention parsing 
-The latter indeed corresponds to the 
-particular case where the alphabet is finite, 
-$T$ computes identity and $A$ is obtained from the weighted CF grammar.
+%\subsection{Computation}
+%
+\begin{proposition}
+The problem of Symbolic Weighted  parsing 
+can be solved in PTIME in the size of the input \SWT $T$, \SWVPA $A$
+and input word $s$, 
+and the computation time of the functions and operators of the label theory.
+\end{proposition}
+%
+\begin{proof} (sketch)
+We follow a \emph{Bar-Hillel} construction, for parsing by intersection.
+%
+\noindent
+Let us first extend the \SWT $T$ over $\Sigma$, $\Omegai$ 
+into a \SWT $T'$ over $\Sigma$ and $\Omega$ 
+(and the same semiring and label theory $\Semiring$ and $\bar\Phi$),
+such that for all $u \in \Sigma^*$, and $t \in {\Omega}^*$, 
+$T'(u, u) = T(u, t|_{\Omegai})$. 
+%
+The transducer $T'$ simply skips every symbol 
+$b \in {\Omega} \setminus \Omegai$, 
+by the addition to $T$,
+of new transitions of the form $\wei_{01}(q, \varepsilon, b, q')$.
+%
+\noindent
+Then, using Corolary~\ref{cor:epsilon}, 
+we construct from the input word $s \in \Sigma^*$ and $T'$
+a \SWA $B_{s, T'}$, 
+such that for all $t \in \Omega^*$, $B_{s, T'}(t) = d(s, t)$.
+%
+%This automaton is such that for all $t \in \T(\Omega)$, 
+%\[ 
+%   A_{T', s}\bigl(\lin(t)\bigr) 
+% = T'\bigl(s, \lin(t)\bigr) 
+% = T'\bigl(s, \lin(t)|_{\Omegai}\bigr) 
+% = d(s, t).
+%\]
+%
+\noindent
+Next, %we convert the input \SWTA $A$ over $\Omega$
+%into a \SWVPA $A'$ over $\hat\Omega$, using Proposition~\ref{lem:SWTA}, and 
+we compute the \SWVPA $B_{s, T'} \otimes A$, 
+using Proposition~\ref{prop:SWVPA-product}.
+%
+\noindent
+It remains to compute a best nested-word $t \in {\Omega}^*$ 
+using the best-search procedure of Proposition~\ref{th:best-search}.
+%and convert it into a best tree in $\T(\Omega)$ in order to solve SW parsing
+%for~$T$, $A$ and~$s$.
+%\qed  
+\end{proof}
+The $\SW$-parsing generalizes 
+the problem of searching the best derivation (AST) of a weighted CF-grammar  
+that yields a given input word.
+The latter problem, sometimes called~\emph{weighted parsing},
+(see \eg~\cite{Goodman99SemiringParsing} 
+ and~\cite{MorbitzVogler19weighted-parsing} for general weighted parsing frameworks)
+corresponds to $\SW$-parsing in the case of finite alphabets,
+a transducer $T$ computing the identity and some \SWVPA~$A$ 
+obtained from the weighted CF grammar.
+%
 %Indeed, it corresponds to the case where $T$
@@
 %\end{description}
-
-%\subsection{Computation}
-%
-\begin{proposition}
-The problem of Symbolic Weighted  parsing 
-can be solved in PTIME in the size of the input \SWT $T$, \SWVPA $A$
-and input word $s$, 
-and the computation time of the functions of the label theory.
-\end{proposition}
-\begin{proof} (sketch)
-We follow a \emph{Bar-Hillel} construction, also called parsing by intersection.
-
-\noindent
-We first extend the \SWT $T$ over $\Sigma$, $\Omegai$, $\Semiring$, and $\bar\Phi$, 
-into a \SWT $T'$ over $\Sigma$ and $\Omega$ (and the same semiring and label theory),
-such that for all $s \in \Sigma^*$, and $u \in {\Omega}^*$, 
-$T'(s, u) = T(s, u|_{\Omegai})$. The transducer $T'$ simply skips every symbol 
-$b \in {\Omega} \setminus \Omegai$, 
-by the addition to the transition of $T$,
-of new transitions of the form $\wei_{01}(q, \varepsilon, b, q')$.
-%
-\noindent
-Then, given an input word $s \in \Sigma^*$, 
-using Corolary~\ref{cor:epsilon}, 
-we compute the \SWA $B_{T', s}$, 
-such that for all $t \in \Omega^*$, $B_{T', s}(t) = d(s, t)$.
-%
-%This automaton is such that for all $t \in \T(\Omega)$, 
-%\[ 
-%   A_{T', s}\bigl(\lin(t)\bigr) 
-% = T'\bigl(s, \lin(t)\bigr) 
-% = T'\bigl(s, \lin(t)|_{\Omegai}\bigr) 
-% = d(s, t).
-%\]
-
-\noindent
-Next, %we convert the input \SWTA $A$ over $\Omega$
-%into a \SWVPA $A'$ over $\hat\Omega$, using Proposition~\ref{lem:SWTA}, and 
-we compute the \SWVPA $B_{T', s} \otimes A$, 
-using Proposition~\ref{prop:SWVPA-product}.
-%
-\noindent
-It remains to compute a best nested-word $w \in {\Omega}^*$ 
-using the best-search procedure of Proposition~\ref{th:best-search}.
-%and convert it into a best tree in $\T(\Omega)$ in order to solve SW parsing
-%for~$T$, $A$ and~$s$.
-%\qed  
-\end{proof}
+%
+Indeed, the \emph{depth-first} traversal of an AST $\tau$ 
+yields a well-parenthesised word $\lin(\tau)$ over an alphabet 
+$\Omega = \Omegai \uplus \Omegac \uplus \Omegar$,  
+assuming \eg that $\Omegai$ contain the symbols labelling the leaves of $\tau$ (symbols of rank $0$) 
+and $\Omegac$ and $\Omegar$ contain respectively one left and right parenthesis
+\philippe{Ah oui, ça aurait pu être dit avant.}
+$\ccall{b}$ and $\creturn{b}$ for each symbol $b$ labelling inner nodes of $\tau$ (symbols of rank $>0$). 
+%
+With this representation, the projection $\lin(t)|_\Omegai$ is then the 
+sequence of leaves of $\tau$. %, enumerated in a \emph{dfs}-traversal. 
+%
+We show in Appendix~\ref{sec:trees} how to convert a ($\SW$) tree automaton~$A$ 
+%(in particular a TA computing on the AST of a given CF-grammar)
+into a $\SWVPA$ computing $A(\lin(\tau))$ for every tree $\tau$. 
+That also holds for the set of ASTs of a weighted CF-grammar.
+%The purpose of the transducer $T$ is to measure a distance between input and output words, 
+%\ie between the word $s$ and the sequence leaves of a derivation tree, in $\Omegai^*$.
+%
+%that generates (weighted) trees by replacement of a state symbol~$q_0$ (non-terminal), 
+%by a tree $a(q_1,\ldots, q_k)$, where $k = \rank(a)$.
+%A replacement rule $q_0 \to a(q_1,\ldots, q_k)$, 
+%of weight $\wei(q_0, a, q_1 \ldots q_k) \in \Semiring$ 
+%according to Definition~\ref{def:SWTA}, 
+%corresponds to the production rule $q_0 := a(q_1,\ldots, q_k)$ of a weighted CF grammar,
+%with set non-terminal symbols $Q$ and set of terminal symbols $\Omega_0$.
+%
+%This actually is a slight generalization of CFG since 
+%each such production rule is labelled by  a symbol of $\Omega_{>0}$, 
+%hence parse trees %derivation trees
+%are trees of $\T(\Omega)$.
+%
+%Another (more original) generalization is that the set of terminal symbols 
+%$\Omega_0$ may be infinite.
+%
+%
+%The input language can also be expressed as a \SWTA, or, 
+%as a particular case, as a weighted context-free grammar, 
+%converted in turn into a \SWVPA following Lemma~\ref{lem:SWTA}.
+%
 

### 2021-07-08 abstract.tex
Intro Philippe

@@
 %
-It is based on language models called Symbolic Weighted Automata (SWA) 
+It is based on language models called Symbolic Weighted Automata (\SWA)
 at the joint %intersection
-between Symbolic Automata (SA) and Weighted Automata (WA), 
-as well as Transducers (SWT) and Visibly Pushdown (SWPDA) variants.
+between Symbolic Automata (\SA) and Weighted Automata (\WA),
+as well as Transducers (\SWT) and Visibly Pushdown (\SWVPA) variants.
 %
-Like SA, SWA deal with large or infinite input alphabets, 
-and like WA, they output a weight value in a semiring domain.
-The transitions of WA are labeled by functions from an infinite alphabet into the weight domain. 
-This is unlike WA whose transitions are guarded by boolean predicates 
+Like \SA, \SWA deal with large or infinite input alphabets,
+and like \WA, they output a weight value in a semiring domain.
+The transitions of \SWA are labeled by functions from an infinite alphabet into the weight domain.
+This is unlike \SA whose transitions are guarded by boolean predicates
 overs symbols in an infinite alphabet
-and also unlike WA whose transitions are labeled by constant weight values, 
+and also unlike \WA whose transitions are labeled by constant weight values,
 and who deal only with finite automata.
 %
-We present some properties of WA, WT and SWPDA models, 
+We present some properties of \SWA, \SWT and \SWVPA models,
 %and show how they can be used
@@
 %
-We also briefly describe the application that motivated the introduction of these models: 
+We illustrate the models with examples taken from
+a motivating application, namely
 a parse-based approach to automated music transcription.

### 2021-07-08 introduction.tex
Intro Philippe

--- 2021-07-05 introduction.tex
@@
 assume a finite and reasonably small input alphabet. %models and algorithms
-Such a restriction makes perfect sense in the context of NLP tasks such as constituency parsing,
+Such a restriction makes perfect sense in the context of
+NLP tasks such as constituency parsing,
 or of programming languages compilers or interpreters.
-Considering large or infinite alphabets can however be of practical interest, %in other cases.
+Considering large or infinite alphabets can however be of
+practical interest, %in other cases.
 %
@@
 %
-for the analyse (\eg validation or filtering) 
+for the analysis (\eg validation or filtering)
 of data streams or serialization of structured documents
@@
 %
+
 The latter case is related to a study that motivated the present work:
-automated music transcription.
-\philippe{I think that we can make the case for a larger set of situations: given a linear 
-music notation input, for instance elements in an XML file, we we want to structure
-the input according to a hierarchical rhythmic space. I can elaborate...}
-In this problem, a music performance, % a symbolic representation of a music performance 
-represented symbolically in the form of a sequence of timed musical events, % timestamped
-is converted into a score in Common Western Music Notation~\cite{Gould11Notation},
-structured according to nested grouping and metric strength of events.
-%(the structure represents hierarchical  grouping of events, and their relative temporal positions)
+automated music transcription. Most representations of music are
+essentially linear. This is true for audio files,
+but also for widely used symbolic representations like MIDI.
+Such representations ignore the hierarchical structures that frame the
+conception of music, at least in the western area. These structures, on the other hand,
+are present, either explicitly  or implicitly,
+in music notation~\cite{Gould11Notation}: music scores are partitioned in measures, measures
+in beats, and so on so forth. Music events do not occur at arbitrary timestamps,
+but respect a discrete division of the  timeline incurred by
+these recursive divisions.
+ The \emph{transcription problem} takes
+a linear input (audio or MIDI) and aims at re-constructing these structures
+by mapping input events to this hierarchical rhythmic space.
 It can therefore be stated as a parsing problem~\cite{foscarin:hal-01988990},
@@
 a weight value is associated to each transition rule, % production rule
-and the rule's weights can be combined with a associative product operator~$\otimes$ into 
-the weight of an AST.
+and the rule's weights can be combined with an
+associative product operator~$\otimes$ into the weight of an AST.
 A second operator~$\oplus$, associative and commutative,
-is moreover used to handle the ambiguity of the model, 
-by summing the weights of the possibly several (in general exponentially many) AST %syntax trees 
+is moreover used to resolve the ambiguity raised by the existence
+of several (in general exponentially many) AST
 associated to a given input word.
@@
 such as the Viterbi semiring and the tropical min-plus algebra, see Figure~\ref{fig:semirings}.
+\philippe{La figure 2 est citée avant la figure 1 mais apparait longtemps après. A corriger.}
 %of domain $\mathbb{R}^+ \cup \{ +\infty\}$,
@@
 
-In this paper, we present a uniform framework for weighted parsing over infinite input alphabets.
+In this paper, we present a framework for weighted parsing over infinite input alphabets.
+\philippe{Tu fais une différence entre model et automata?}
 It is based on \emph{symbolic weighted} finite states language models (\swM),
-generalizing~\SA, with functions into an arbitrary semiring instead of Boolean guards, 
+generalizing both~\SA, with functions into an arbitrary semiring instead of Boolean guards,
 and~\WA, by handling infinite alphabets, see Figure~\ref{fig:hierarchy}.
@@
 %
+\philippe{Tu veux dire: les modèles formels que tu combines?}
 The models presented here are finite automata called symbolic-weighted (\SWA),
@@
 %\wrt an edit distance
-The latter model of automata computes on \emph{nested words}~\cite{AlurMadhusudan09nested}, 
+The latter model of automata operates on \emph{nested words}~\cite{AlurMadhusudan09nested},
 a structured form of words parenthesized with markup symbols,
@@
 More precisely, a \SWVPA $A$ associates a weight value $A(t)$ % \in \Semiring$
-to a given a nested word $t$, which is the linearization of an AST. %representing a parse tree.
+to a given nested word $t$, which is the linearization of an AST. %representing a parse tree.
 %
@@
 finding $t$ minimizing
-$T(s, t) \otimes A(t)$ (\wrt the ranking defined by $\oplus$).
+$T(s, t) \otimes A(t)$ (\wrt the ranking defined by $\oplus$), given an input word $s$.
 The latter value is called the distance between $s$ and $A$ in~\cite{Mohri03EDWA}.
 %
+
 Like weighted-parsing
@@
 a \SWVPA $A_{T, s}$ is built, such that for all $t$, $A_{T, s}(t) = T(s, t) \otimes A(t)$.
+\philippe{The notation  $A_{T, s}$ has  not been introduced so far. It is not clear why $T$ is a parameter there}
 In the second step, a best AST $t$ is found by applying to $A_{T, s}$
@@
         comparable to pebble automata? $\to$ conclusion}
-
 The main contributions of the paper are:

### 2021-07-08 introduction.tex

@@
 
-In this paper, we present a framework for weighted parsing over infinite input alphabets.
+In this paper, we present a uniform framework for weighted parsing over infinite input alphabets.
 \philippe{Tu fais une différence entre model et automata?}
 It is based on \emph{symbolic weighted} finite states language models (\swM),
-generalizing both~\SA, with functions into an arbitrary semiring instead of Boolean guards,
-and~\WA, by handling infinite alphabets, see Figure~\ref{fig:hierarchy}.
+generalizing the Boolean guards of~\SA %in transitions
+into functions into an arbitrary semiring,
+and generalizing also~\WA, by handling infinite alphabets, see Figure~\ref{fig:hierarchy}.
+
 %***
@@
 % non-terminals and production rules).
-In the transition rules of \swM models, input symbols appear as variables, %(parameters)
-and the weight associated to a transition rule is a function of these variables.
+
+In short, a transition rule $q \xrightarrow{\phi} q'$ from state $q$ to $q'$ of a \swM,
+%input symbols are variables, %(parameters)
+is labeled by a function $\phi$ associating to every input symbol $a$ a weight value $\phi(a)$
+in a semiring domain.
+%and the weight associated to a transition rule is a function of these variables.
+\philippe{This sentence (symbols as variables) is not immediately clear to me. Maybe a short example or intuition?}
+\florent{modified}
 %

### 2021-07-08 introduction.tex
Intro revisee Ph.

@@
 in music notation~\cite{Gould11Notation}: music scores are partitioned in measures, measures
-in beats, and so on so forth. Music events do not occur at arbitrary timestamps,
+in beats, and beats can be further recursively divided.
+It follows that music events do not occur at arbitrary timestamps,
 but respect a discrete division of the  timeline incurred by
@@
 The \emph{transcription problem} takes
-a linear input (audio or MIDI) and aims at re-constructing these structures
+as input a linear repreentation (audio or MIDI) and aims at re-constructing
+these structures
 by mapping input events to this hierarchical rhythmic space.
 It can therefore be stated as a parsing problem~\cite{foscarin:hal-01988990},
-over an infinite alphabet of timed events.
+over an infinite alphabet of timed events. Throughout the paper we will
+illustrate our framework with examples taken for this motivating application.
 

### 2021-07-08 semiring.tex
Intro revisee Ph.

@@
 We shall consider semirings for the weight values of our language models.
-\florent{The results are established for a general class of semirings. They can be instantiated for concrete cases}
-%
-\philippe{There is sometimes a confusion in the text between the struture and the domain $\Semiring$. Not
-essential}.
+%\florent{The results are established for a general class of semirings. They can be instantiated for concrete cases}
 A \emph{semiring} $\< \Semiring, \oplus, \zero, \otimes, \one>$
@@
 in order to handle non-determinism,
-and $\otimes$ combines two values into a single value, 
-in a chaining of transitions.
+and $\otimes$ combines two values into a single value.
 %and let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a {semiring},
@@
 % \emph{negative boundedness}
+\philippe{Ca j'ai pas compris}
 It implies that, when looking for a best path in a graph whose edges

### 2021-07-08 introduction.tex
Ajout exemples

@@
 The \emph{transcription problem} takes
-as input a linear repreentation (audio or MIDI) and aims at re-constructing
+as input a linear representation (audio or MIDI) and aims at re-constructing
 these structures
@@
 It can therefore be stated as a parsing problem~\cite{foscarin:hal-01988990},
-over an infinite alphabet of timed events. Throughout the paper we will
-illustrate our framework with examples taken for this motivating application.
+over an infinite alphabet of timed events.
 
@@
 instead of syntax trees and grammars. %(\S~\ref{sec:trees}).
-%
-%Finally, Section~\ref{sec:transcription} describes
-%the implemented application %of this approach
-%to automated music transcription
-%that motivated this study.
 
-%based these models and on generic def. of distances;
-%the use of VPA for parsing, word instead of trees, use comparison input-output.
-%and best-search
-
-%In Section~\ref{section:SWA} we introduce \SWA and \SWT.
-%Then \SWVPA are defined in Section~\ref{sec:SWVPA},
-%where a polynomial 1-best algorithm is described that can be use to solve
-%Symbolic Weighted Parsing, as described Section~\ref{sec:parsing}.
-%Finally, in Section~\ref{sec:transcription}, we present an application
-%of this approach to automated music transcription that has been implemented.
+\begin{example}[Running example]
+Throughout the paper we
+illustrate our framework with music transcription examples: Given
+a \emph{timeline} of musical events with arbitrary timestamps as input, parse it
+into a structured music score. In our example, input events
+are pairs $\< \eta, \tau>$ made of
+a symbol $\eta \in \Sigma$, where $\Sigma$ stands for the
+set of MIDI message symbols~\cite{?}
+and  $\tau \in \mathbb{Q}$ is a timestamp. The output of parsing
+is a representation of the sequence in
+Common Western Music Notation (CWMN)~\cite{Gould11Notation}
+where event symbols belong to the domain $\Delta$
+of \emph{pitches} (e.g., A4, G5, etc.), temporal
+information is encoded as \emph{durations} (whole \musWhole$,
+$quarter, $\musQuarter$, eight $\musEighth$, etc), and notes are grouped in
+high-level structures (beams, measures, tuplets). The following inputs
+will be used:
+\begin{enumerate}
+  \item $I_1  = [ <e_1, 0.07>,<e_2, 0.72>,<e_3, 0.91>]$, over interval $[0,1[$
+  \item $I_2  = [ <e_3, 1.05>,<e_4, 1.36>,<e_5, 1.71>]$, over interval $[1,2[$
+\end{enumerate}
+There exists many possible parsings of $I_1 \cup I_2$ in music notation, among
+which
+\includegraphics[scale=0.20]{pictures/score5.png}
+and \includegraphics[scale=0.20]{pictures/score4.png}.  Wheighted
+parsing associates a cost to each solution, and our framework
+aims at selecting the best one with respect to this cost.
+\endex
+\end{example}

### 2021-07-08 semiring.tex
Philippe semirings + labels

@@
 
+\begin{example}
+The recursive subdivision of time that leads to hierarchichal
+structures of music notation can be modeled as
+production rules. Since there exists several possible division,
+rules can be weighted 
+in the tropical semiring whose domain $\mathbb{R}_+ \cup \{ +\infty\}$,
+$\oplus$ is min, $\zero = +\infty$, $\otimes$ is sum, and $\one = 0$.
+For instance, the following production rules define two possible divisions
+of a bounded time interval into respectively a duplet and a triplet.
+\[
+\rho_1: q_0 \xrightarrow{0.06} \< q_{1}, q_{2}>,\
+\rho_2: q_0 \xrightarrow{0.12} \< q_{1}, q_{2}, q_{2}>.
+\]
+\endex
+\end{example}
 %%% semiring properties used in paper

### 2021-07-08 main.tex
description algo search

@@
 
-%\newcommand{\A}{\mathcal{A}}
-%\newcommand{\B}{\mathcal{B}}
-\newcommand{\D}{\mathbb{D}}
+\newcommand{\A}{\mathcal{A}} % automata, VPA
+\newcommand{\B}{\mathcal{B}} % automata (in constructions)
+\newcommand{\C}{\mathcal{C}} % automata (in constructions)
+\newcommand{\D}{\mathcal{D}}  % transducer = distance
 \newcommand{\E}{\mathbb{E}}
@@
 \def\SWVPA{\textsf{sw-VPA}\xspace}
+\def\VPA{\textsf{VPA}\xspace}
+\def\SVPA{\textsf{sVPA}\xspace}
 \def\weight{\mathsf{weight}}
@@
 \label{section:SWVPA}\label{sec:SWVPA}
-The model presented in this section generalizes Symbolic VPA~\cite{dAntonyAlur14SVPDA}
+The model presented in this section generalizes symbolic \VPA
+(\SVPA~\cite{dAntonyAlur14SVPDA}, 
+ generalizing themselves \VPA~\cite{AlurMadhusudan09nested} to infinite alphabets)
 from Boolean semirings to arbitrary semiring weight domains.
@@
 represent AST. % parse trees.
+\florent{see \S 5 and App.A}
 
@@
 whose elements are respectively called 
-\emph{internal}, \emph{call} and \emph{return} symbols.
+\emph{internal}, \emph{call} and \emph{return} symbols~\cite{AlurMadhusudan09nested}.
 \philippe{Là je crois qu'il faudrait expliquer ces Omega, je commence à fatiguer et je suis un
@@
 \begin{array}{lll}
-\weii: Q \times \Omegac \times P \times \Omegai \times Q \to \Semiring & 
+\weii: Q \times [\Omegac \times P] \times \Omegai \times Q \to \Semiring & 
 \weii(q, c, p, a, q') = \eta_\mathsf{ci}(c, a) & 
@@
 %
-\weic: Q \times \Omegac \times P \times  \Omegac \times P \times Q \to \Semiring & 
+\weic: Q \times [\Omegac \times P] \times  [\Omegac \times P] \times Q \to \Semiring & 
 \weic(q, c, p, \call{c'}, p', q') = \eta_\mathsf{cc}(c, \call{c'}) & 
@@
 %
-\weiec: Q \times \Omegac \times P \times Q \to \Semiring & 
+\weiec: Q \times [\Omegac \times P] \times Q \to \Semiring & 
 \weiec(q, {\call{c}}, p, q') = \phi_\mathsf{c}({\call{c}}) &
@@
 %
-\weir: Q \times \Omegac \times P \times \Omegar \times Q \to \Semiring & 
+\weir: Q \times [\Omegac \times P] \times \Omegar \times Q \to \Semiring & 
 \weir(q, {\call{c}},  p, {\return{r}}, q') = \eta_\mathsf{cr}({\call{c}},  {\return{r}}) & 
@@
 \mathop{\otimes} \mathsf{out}(q').
-\label{eq:weightA}
+\label{eq:SWVPA-value}
 \end{equation}
@@
 \medskip\noindent
-Like VPA and symbolic VPA, 
+Similarly to \VPA~\cite{AlurMadhusudan09nested} 
+and \SVPA~\cite{dAntonyAlur14SVPDA}, 
 the class of \SWVPA is closed under the binary operators of the underlying semiring.
@@
 as in the case of the Boolean semiring~\cite{dAntonyAlur14SVPDA}.
+\florent{complete proof}
 \end{proof}
@@
 \noindent
-%**hypotheses**
-Let us assume that the semiring~$\Semiring$ is
-commutative, bounded, and complete,  
+We shall now present a procedure for searching, for a~$\SWVPA$ $A$, 
+a word %$w \in \Omega^*$ 
+of minimal weight for~$A$, 
+as stated in the following proposition. %\wrt~$\leq_\oplus$, 
+%
+\begin{proposition}\label{th:best-search}
+For a \SWVPA $A$ 
+over $\Omega$, 
+$\Semiring$ commutative, bounded, total and complete, %semiring 
+and $\bar\Phi$ effective, %and $k$-convex, % label theory,
+one can construct in PTIME a word $t \in \Omega^*$ 
+such that $A(t)$ is minimal \wrt the natural ordering for $\Semiring$. 
+\end{proposition} 
 \florent{total?}
-and that $\bar\Phi$ is an effective label theory.
-%s.t. for all $\phi \in \Phii$, 
-%$\psi \in \Phir$, 
-%and $\eta \in \Phicr$, 
-%$\displaystyle\bigoplus_{a \in \Omegai} \phi(a)$ 
-%$\displaystyle\bigoplus_{r \in \Omegar} \phi(r)$ and 
-%$\displaystyle\bigoplus_{{\call{c}} \in \Omegac} 
-%\displaystyle\bigoplus_{{\return{r}} \in \Omegar} \eta({\call{c}}, {\return{r}})$
-%are computable...
-% 
-% total ? 
-% monotonic and superior writ natural ordering
-%Regarding the infinite sum operator, note that
-%$\bigoplus_{x \in \Phi_\Omega} \phi(x)$, 
-%$\bigoplus_{y \in \Phi_\Delta} \psi(y)$, and  
-%... exist and in $\Semiring$.
-%
-We propose a Dijkstra algorithm computing, for a $\SWVPA$ $A$
+%
+Let $A = \< Q, P, \init, \bar\wei, \final >$.
+%
+We propose a Dijkstra algorithm computing,
+for every $q, q' \in Q$, 
+the minimum, \wrt $\leq_\oplus$, of the function 
+$\beta_{q, q'} : t \mapsto \weight_A(\config{q}{\bot}, t, \config{q'}{\bot})$.
+Let us denote by $b_\bot(q, q')$ this minimum. 
+By definition of $\leq_\oplus$ it holds that:
+%
+\begin{equation}\label{eq:bbot}
+  b_\bot(q, q') = \bigoplus_{t\in \Omega^*} 
+  \textstyle
+  \weight_A\bigl(\config{q}{\bot}, t, \config{q'}{\bot}\bigr).
+\end{equation}
+Hence, following~\eqref{eq:SWVPA-value}, and the associativity and commutativity 
+and distributivity for~$\otimes$ and~$\oplus$, the minimum of~$A(t)$ is 
+\(
+{\displaystyle\bigoplus_{t\in \Omega^*}}
+{\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
+\mathsf{in}(q) \mathop{\otimes} 
+\beta_{q, q'}(t)
+\mathop{\otimes} \mathsf{out}(q')
+= 
+{\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
+\mathsf{in}(q) \mathop{\otimes} 
+b_\bot(q, q')
+\mathop{\otimes} \mathsf{out}(q')
+\).
+
+\noindent
+In order to compute
+
+****
+
 over~$\Omega$, $\Semiring$ and~$\bar\Phi$, 
@@
 %
+
 \noindent
@@
 
-%\medskip
 \noindent
@@
 we require the additional property of convexity of weight functions.
-
-\begin{proposition}\label{th:best-search}
-For a \SWVPA $A$ 
-over $\Omega$, 
-$\Semiring$ commutative, bounded, total and complete, %semiring 
-and $\bar\Phi$ effective, %and $k$-convex, % label theory,
-one can construct in PTIME a word $t \in \Omega^*$ 
-such that $A(t)$ is minimal \wrt the natural ordering for $\Semiring$. 
-\end{proposition}
 
@@
 The $\SW$-parsing generalizes 
-the problem of searching the best derivation (AST) of a weighted CF-grammar  
-that yields a given input word.
+the problem of searching the best derivation (AST) of a weighted CF-grammar $G$ 
+that yields a given input word $w$.
 The latter problem, sometimes called~\emph{weighted parsing},
@@
 a transducer $T$ computing the identity and some \SWVPA~$A$ 
-obtained from the weighted CF grammar.
+obtained from $G$. %the weighted CF grammar.
 %
@@
 $\Omega = \Omegai \uplus \Omegac \uplus \Omegar$,  
-assuming \eg that $\Omegai$ contain the symbols labelling the leaves of $\tau$ (symbols of rank $0$) 
+assuming \eg that $\Omegai$ contains the symbols labelling the leaves of $\tau$ (symbols of rank $0$), 
 and $\Omegac$ and $\Omegar$ contain respectively one left and right parenthesis
@@
 %
-With this representation, the projection $\lin(t)|_\Omegai$ is then the 
-sequence of leaves of $\tau$. %, enumerated in a \emph{dfs}-traversal. 
-%
-We show in Appendix~\ref{sec:trees} how to convert a ($\SW$) tree automaton~$A$ 
-%(in particular a TA computing on the AST of a given CF-grammar)
-into a $\SWVPA$ computing $A(\lin(\tau))$ for every tree $\tau$. 
-That also holds for the set of ASTs of a weighted CF-grammar.
+%With this representation, the projection $\lin(t)|_\Omegai$ is the 
+%sequence of leaves of $\tau$. %, enumerated in a \emph{dfs}-traversal. 
+%
+We show in Appendix~\ref{sec:trees} how to construct a 
+%($\SW$) tree automaton~$A$ %(in particular the TA computing on the AST of a given CF-grammar)
+a $\SWVPA$ $A$ such that $A(\lin(\tau))$ is the weight the AST $\tau$ of $G$.
+
+%That also holds for the set of ASTs of a weighted CF-grammar.
 %The purpose of the transducer $T$ is to measure a distance between input and output words, 
@@
 %
+
+%In practice, 
+%the functions in transition might be represented by diagram such as
+%Algebraic Decision Diagrams~\cite{Bahar97ADD}
 

### 2021-07-08 main.tex
Modifs P.

--- 2021-07-07 main.tex
@@
 %\usepackage{mathptmx}
+\usepackage{comment}
 
@@
 as functions of
-\philippe{I missed sth: what is this $\varepsilon$? Intuitively clear but not defined?}
 $Q \times (\Sigma \cup \{ \varepsilon \}) \times (\Delta \cup \{ \varepsilon \}) \times Q \to \Semiring$,
@@
 %
-\philippe{OK tout ça se lit bien :-)}
 Intuitively, using a transition $\wei_{ij}(q, a, b, q')$ means for $T$:
@@
 
-\philippe{Je crois qu'il faudrait numéroter les exemples indépendamment des définitions. Cet exemple est le premier
-qui donne des détails sur l'application visée. Il arrive peut-être un peu tard et est long. On pourrait introduire
-la motivation dans l'intro, et développer des petits exemples au fur et à mesure.}
+\begin{example}
+Let us develop the example of comparison between music played by a performer,
+represented as a sequence $s \in \Sigma^*$ of events in the MIDI alphabet $\Sigma$,
+and  a music score represented as a sequence $t \in \Delta^*$
+in the CWMN alphabet $\Delta$. We build a small weighted transducer model
+ with two states $q_0$ and $q_1$ that
+calculates the distance between $s$ and $t$.
+
+If one performed event $s_i$  corresponds
+to one notated event $t_1$ (for instance MIDI code 61 and pitch A4),
+the weight value computed by the \SWT is the time distance between both,
+as in Example~\ref{distance-time}, and is modeled by
+transitions $\wei_{11}$ below.
+%
+If we meet the music notation symbol '-' that
+represents continuation (such as instance in \emph{ties}
+$\musQuarter\!\!\!\mathrel{\raisebox{-1.5mm}{$\smile$}}\!\!\!\musEighth$,
+or \emph{dots} \musQuarterDotted{}), it is  skipped with no cost (transitions $\wei_{01}$ or weight $\one$).
+%transitions
+\[
+\begin{array}{rclcrcl}
+\wei_{11}(q_0, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d| & \quad &
+\wei_{11}(q_1, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d|\\
+\wei_{01}(q_0, \varepsilon, \< \mathsf{-}, d'>, q_0) & = & \one & &
+\wei_{01}(q_1, \varepsilon, \< \mathsf{-}, d'>, q_0) & = & \one\\
+\wei_{10}(q_0, d, \varepsilon, q_1) & = & \alpha & & %\multicolumn{3}{l}{\mathrm{for~all~} b \in \Delta}
+\end{array}
+\]
+%
+We also must be able to take performing errors into account, while still being able to compare with the score,
+since a performer could, for example, play an unwritten extra note.
+\lydia{reformulated this sentence}
+%
+This is modelled by the transition $\wei_{10}$ with an arbitrary weight value $\alpha \in \Semiring$,
+switching from state $q_0$ (normal) to $q_1$ (error).
+\philippe{Comprends pas cette phrase}
+The transitions in the second column below switch back to the normal state $q_0$.
+% the metric computed by the \SWT is the smallest sum of point wise distances
+% between dates of input and output events.
+At last, we let $q_0$ be the only initial and final state, with
+$\init(q_0) = \final(q_0) = \one$, and
+$\init(q_1) = \final(q_1) = \zero$.
+
+\lydia{ccl to the ex}
+That way, an \SWT is capable of evaluating the differences between a score and a performance,
+all the while ensuring that performance errors are plausible.
+%$\init(q_0, d, b) = \final(q_0, d, b) = \one$, and
+%$\init(q_1, d, b) = \final(q_1, d, b) = \zero$,
+%for all $d \in \Sigma$ and $b \in \Delta$.
+\endex
+\end{example}
+
+\begin{comment}
 \begin{example}
@@
 \noindent
-We propose a small weighted transducer model that calculates the distance bewteen an input sequence of sounding 
+We propose a small weighted transducer model with two states $q_0$ and $q_1$
+that calculates the distance bewteen an input sequence of sounding
 events (music "performance")
@@
 \end{example}
+\end{comment}
 
@@
 \noindent
+\philippe{Il me manque une explication:  on construit un automate
+qui, étant donnée une partition $t$, renvoie la distance minimale
+avec n'importe quelle performance (distance donnée par un transducer)?
+Quel est le rôle de A(s)?
+}
 The following property will be useful to the approach on
@@
 \emph{internal}, \emph{call} and \emph{return} symbols.
-\philippe{Là je crois qu'il faudrait expliquer ces Omega, je commence à fatiguer et je suis un
-peu largué par toutes ces définitions. J'intuite qu'il s'agit des symboles, parenthèses ouvrantes
-et fermantes? Pourquoi il faut un alphabet pour les parenthèses?}
 % names are  coined by application to functional program verification
@@
 where $Q$ is a finite set of states,
-\philippe{Est-ce que tout le monde sait ce qu'est un pushdown automata? Je suppose que c'est
 lié à la pile.}

### 2021-07-08 introduction.tex

--- 2021-07-08 introduction.tex
@@
 assume a finite and reasonably small input alphabet. %models and algorithms 
-Such a restriction makes perfect sense in the context of
-NLP tasks such as constituency parsing,
+Such a restriction makes perfect sense in the context of NLP tasks such as constituency parsing,
 or of programming languages compilers or interpreters.
-Considering large or infinite alphabets can however be of
-practical interest, %in other cases.
+Considering large or infinite alphabets can however be of practical interest, %in other cases.
 %
@@
 %
-for the analysis (\eg validation or filtering)
+for the analyse (\eg validation or filtering) 
 of data streams or serialization of structured documents 
@@
 %
-
 The latter case is related to a study that motivated the present work: 
-automated music transcription. Most representations of music are
-essentially linear. This is true for audio files,
-but also for widely used symbolic representations like MIDI.
-Such representations ignore the hierarchical structures that frame the
-conception of music, at least in the western area. These structures, on the other hand,
-are present, either explicitly  or implicitly,
-in music notation~\cite{Gould11Notation}: music scores are partitioned in measures, measures
-in beats, and so on so forth. Music events do not occur at arbitrary timestamps,
-but respect a discrete division of the  timeline incurred by
-these recursive divisions.
- The \emph{transcription problem} takes
-a linear input (audio or MIDI) and aims at re-constructing these structures
-by mapping input events to this hierarchical rhythmic space.
+automated music transcription.
+\philippe{I think that we can make the case for a larger set of situations: given a linear 
+music notation input, for instance elements in an XML file, we we want to structure
+the input according to a hierarchical rhythmic space. I can elaborate...}
+In this problem, a music performance, % a symbolic representation of a music performance 
+represented symbolically in the form of a sequence of timed musical events, % timestamped
+is converted into a score in Common Western Music Notation~\cite{Gould11Notation},
+structured according to nested grouping and metric strength of events.
+%(the structure represents hierarchical  grouping of events, and their relative temporal positions)
 It can therefore be stated as a parsing problem~\cite{foscarin:hal-01988990}, 
@@
 a weight value is associated to each transition rule, % production rule
-and the rule's weights can be combined with an
-associative product operator~$\otimes$ into the weight of an AST.
+and the rule's weights can be combined with a associative product operator~$\otimes$ into 
+the weight of an AST.
 A second operator~$\oplus$, associative and commutative, 
-is moreover used to resolve the ambiguity raised by the existence
-of several (in general exponentially many) AST
+is moreover used to handle the ambiguity of the model, 
+by summing the weights of the possibly several (in general exponentially many) AST %syntax trees 
 associated to a given input word.
@@
 such as the Viterbi semiring and the tropical min-plus algebra, see Figure~\ref{fig:semirings}. 
-\philippe{La figure 2 est citée avant la figure 1 mais apparait longtemps après. A corriger.}
 %of domain $\mathbb{R}^+ \cup \{ +\infty\}$, 
@@
 In this paper, we present a uniform framework for weighted parsing over infinite input alphabets.
-\philippe{Tu fais une différence entre model et automata?}
 It is based on \emph{symbolic weighted} finite states language models (\swM), 
@@
 and generalizing also~\WA, by handling infinite alphabets, see Figure~\ref{fig:hierarchy}.
-
 %***
@@
 % non-terminals and production rules).
-
 In short, a transition rule $q \xrightarrow{\phi} q'$ from state $q$ to $q'$ of a \swM, 
@@
 %
-\philippe{Tu veux dire: les modèles formels que tu combines?}
 The models presented here are finite automata called symbolic-weighted (\SWA),
@@
 to a given nested word $t$, which is the linearization of an AST. %representing a parse tree.
+\philippe{The weight is a label on edges in the derivation tree, right? Not sure I understand 
+the sentence}
 %
@@
 The latter value is called the distance between $s$ and $A$ in~\cite{Mohri03EDWA}.
+\philippe{You mean the distance between $s$ and $t$? "The latter value" is a bit ambiguous}
 %
-
 Like weighted-parsing 
@@
         comparable to pebble automata? $\to$ conclusion}
+
 The main contributions of the paper are: 

### 2021-07-08 main.tex

--- 2021-07-08 main.tex
@@
 whose elements are respectively called
-\emph{internal}, \emph{call} and \emph{return} symbols~\cite{AlurMadhusudan09nested}.
-\philippe{Là je crois qu'il faudrait expliquer ces Omega, je commence à fatiguer et je suis un
-peu largué par toutes ces définitions. J'intuite qu'il s'agit des symboles, parenthèses ouvrantes
-et fermantes? Pourquoi il faut un alphabet pour les parenthèses?}
+\emph{internal}, \emph{call} and \emph{return} symbols.
 % names are  coined by application to functional program verification
@@
 where $Q$ is a finite set of states,
-\philippe{Est-ce que tout le monde sait ce qu'est un pushdown automata? Je suppose que c'est
-lié à la pile.}
 $P$ is a finite set of stack symbols,
@@
 
+\begin{proposition}\label{th:best-search}
+For a \SWVPA $A$
+over $\Omega$,
+$\Semiring$ commutative, bounded, total and complete, %semiring
+and $\bar\Phi$ effective, %and $k$-convex, % label theory,
+one can construct in PTIME a word $t \in \Omega^*$
+such that $A(t)$ is minimal \wrt the natural ordering for $\Semiring$.
+\end{proposition}
+
 \section{Symbolic Weighted Parsing}
@@
 and $\Omegac$ and $\Omegar$ contain respectively one left and right parenthesis
-\philippe{Ah oui, ça aurait pu être dit avant.}
 $\ccall{b}$ and $\creturn{b}$ for each symbol $b$ labelling inner nodes of $\tau$ (symbols of rank $>0$).

### 2021-07-08 introduction.tex
recovery intro Philippe
Petite modif intro

--- 2021-07-08 introduction.tex
@@
 It follows that music events do not occur at arbitrary timestamps,
-but respect a discrete division of the  timeline incurred by
+but respect a discrete partitioning of the  timeline incurred by
 these recursive divisions.
@@
 %\cite{Eilenberg74automata}
-such as the Viterbi semiring and the tropical min-plus algebra, see Figure~\ref{fig:semirings}.
-\philippe{La figure 2 est citée avant la figure 1 mais apparait longtemps après. A corriger.}
+such as the Viterbi semiring and the tropical min-plus algebra% % see Figure~\ref{fig:semirings}.
 %of domain $\mathbb{R}^+ \cup \{ +\infty\}$,

### 2021-07-08 introduction.tex
Petite modif intro (2)

@@
 In this paper, we present a uniform framework for weighted parsing over infinite input alphabets.
-\philippe{Tu fais une différence entre model et automata?}
 It is based on \emph{symbolic weighted} finite states language models (\swM),
@@
 %and the weight associated to a transition rule is a function of these variables.
-\philippe{This sentence (symbols as variables) is not immediately clear to me. Maybe a short example or intuition?}
-\florent{modified}
 %
@@
 %
-\philippe{Tu veux dire: les modèles formels que tu combines?}
-The models presented here are finite automata called symbolic-weighted (\SWA),
+The framework relies on several language models:
+finite automata called symbolic-weighted (\SWA),
 transducers (\SWT), and pushdown automata
@@
 a \SWVPA $A_{T, s}$ is built, such that for all $t$, $A_{T, s}(t) = T(s, t) \otimes A(t)$.
-\philippe{The notation  $A_{T, s}$ has  not been introduced so far. It is not clear why $T$ is a parameter there}
 In the second step, a best AST $t$ is found by applying to $A_{T, s}$

### 2021-07-08 main.tex
recovery intro Philippe
expl. algo best-search (en cours)

@@
 \end{proposition}
-\florent{total?}
+%\florent{total?}
 %
@@
 Let us denote by $b_\bot(q, q')$ this minimum.
-By definition of $\leq_\oplus$ it holds that:
+By definition of $\leq_\oplus$, and since $\Semiring$ is total,
+it holds that:
 %
@@
 \end{equation}
+Since $\Semiring$ is complete, the infinite sum in \eqref{eq:bbot} is well defined,
+and,  it is the minimum in $\Omega^*$,
+\wrt~$\leq_\oplus$, % this ordering.
+of the fonction
+$s \mapsto \weight_A(\config{q}{\sigma}, s, \config{q'}{\sigma})$.
+%
 Hence, following~\eqref{eq:SWVPA-value}, and the associativity and commutativity
@@
 \noindent
-In order to compute
-
-****
-
-over~$\Omega$, $\Semiring$ and~$\bar\Phi$,
-the minimal weight %\wrt~$\leq_\oplus$,
-for a word in~$\Omega^*$.
-%
-
-\noindent
-%More precisely,
-We distinguish two cases : when the stack is empty, and when it is not.
-\lydia{introduced 2 cases for b}
-%
-In the case of an empty stack, let $b_\bot : Q \times Q \to \Semiring$ be such that : %the function:
-%
-\begin{equation}\label{eq:bbot}
-  b_\bot(q, q') = \bigoplus_{s\in \Omega^*}
-  \textstyle
-  \weight_A\bigl(\config{q}{\bot}, s, \config{q'}{\bot}\bigr).
-\end{equation}
-%
-Since $\Semiring$ is complete, the infinite sum in \eqref{eq:bbot} is well defined,
-and, providing that $\Semiring$ is total, it is the minimum in $\Omega^*$,
-\wrt~$\leq_\oplus$, % this ordering.
-of the fonction
-$s \mapsto \weight_A(\config{q}{\sigma}, s, \config{q'}{\sigma})$.
-%
-The term $\config{q}{\bot}, s, \config{q'}{\bot}$
-of this sum is the central expression in
-the definition \eqref{eq:weightA} of $A(s_0)$, for the minimum $s_0$
-of the function $\weight_A$.
-\lydia{so ?}
-
-%
-\noindent
-If the stack is not empty, let $\top$ be a fresh stack symbol which does not belong to $\Gamma$,
-\florent{$b_\top$ : mot bien parenthèsé $c$/$r$}
-and let $b_\top : Q \times P \times Q \to \Phic$ be such that,
-for every two states $q, q' \in Q$
-and stack symbol $p \in P$: %and $\sigma \in \{ \bot, \top \}$,
-\begin{equation}\label{eq:btop}
-  b_\top(q, p, q') : c \mapsto \bigoplus_{s\in \Omega^*}
-  \textstyle
-  \weight_A\bigl(\configup{q}{\< c, p> \stackup \top }, s, \configup{q'}{\<c, p> \stackup \top}\bigr)
-\end{equation}
+In order to compute the above function $b_\bot : Q \times Q \to \Semiring$, 
+we shall use the auxiliary function $b_\top :  Q \times P \times Q \to \Phic$.
 %
@@
 
+over~$\Omega$, $\Semiring$ and~$\bar\Phi$,
+the minimal weight %\wrt~$\leq_\oplus$,
+for a word in~$\Omega^*$.
+%
+
+\noindent
+%More precisely,
+We distinguish two cases : when the stack is empty, and when it is not.
+\lydia{introduced 2 cases for b}
+%
+In the case of an empty stack, let $b_\bot : Q \times Q \to \Semiring$ be such that : %the function:
+%
+EQBBOT
+%
+The term $\config{q}{\bot}, s, \config{q'}{\bot}$
+of this sum is the central expression in
+the definition \eqref{eq:weightA} of $A(s_0)$, for the minimum $s_0$
+of the function $\weight_A$.
+\lydia{so ?}
+
+%
+\noindent
+If the stack is not empty, let $\top$ be a fresh stack symbol which does not belong to $\Gamma$,
+\florent{$b_\top$ : mot bien parenthèsé $c$/$r$}
+and let $b_\top : Q \times P \times Q \to \Phic$ be such that,
+for every two states $q, q' \in Q$
+and stack symbol $p \in P$: %and $\sigma \in \{ \bot, \top \}$,
+\begin{equation}\label{eq:btop}
+  b_\top(q, p, q') : c \mapsto \bigoplus_{s\in \Omega^*}
+  \textstyle
+  \weight_A\bigl(\configup{q}{\< c, p> \stackup \top }, s, \configup{q'}{\<c, p> \stackup \top}\bigr)
+\end{equation}
+%
+
 \begin{algorithm}

### 2021-07-08 main.tex

@@
 \label{sec:SWVPA-def}
-Let $\Omega$ be a countable alphabet
+Let $\Delta$ be a countable alphabet
 %finite (large) or infinite,
 that we assume partitioned into three
-subsets~$\Omegai$, $\Omegac$, $\Omegar$,
+subsets~$\Deltai$, $\Deltac$, $\Deltar$,
 whose elements are respectively called
@@
 % \begin{itemize}
-% \item a set $\Omegai$ of \emph{internal symbols} denoted $a$,
-% \item a set $\Omegac$ of \emph{call symbols} denoted $\call{a}$,
-% \item a set $\Omegar$ of \emph{return symbols} denoted $\return{a}$.
+% \item a set $\Deltai$ of \emph{internal symbols} denoted $a$,
+% \item a set $\Deltac$ of \emph{call symbols} denoted $\call{a}$,
+% \item a set $\Deltar$ of \emph{return symbols} denoted $\return{a}$.
 % \end{itemize}
@@
 stand respectively
-for~$\Phi_\Omegai$, $\Phi_\Omegac$, $\Phi_\Omegar$ and~$\Phi_{\Omegac, \Omega_\mathsf{x}}$.
+for~$\Phi_\Deltai$, $\Phi_\Deltac$, $\Phi_\Deltar$ and~$\Phi_{\Deltac, \Delta_\mathsf{x}}$.
 %
 %Moreover, we extend this theory with a set $\Phii$
-%of unary functions in $\Omegai \to \Semiring$,
+%of unary functions in $\Deltai \to \Semiring$,
 %closed under $\oplus$ and $\otimes$.
@@
 A \emph{Symbolic Weighted Visibly Pushdown Automata} (\SWVPA)
-over  $\Omega = \Omegai \uplus \Omegac \uplus \Omegar$, $\Semiring$ and $\bar\Phi$
+over  $\Delta = \Deltai \uplus \Deltac \uplus \Deltar$, $\Semiring$ and $\bar\Phi$
 is a tuple $A = \< Q, P, \init, \bar\wei, \final >$,
@@
 for all $q, q' \in Q$, $p \in P$,
-$a \in \Omegai$,
-$\call{c} \in \Omegac$,
-$\return{r} \in \Omegar$,
+$a \in \Deltai$,
+$\call{c} \in \Deltac$,
+$\return{r} \in \Deltar$,
 overloading their names: % for simplicity:
@@
 \begin{array}{lll}
-\weii: Q \times [\Omegac \times P] \times \Omegai \times Q \to \Semiring &
+\weii: Q \times [\Deltac \times P] \times \Deltai \times Q \to \Semiring &
 \weii(q, c, p, a, q') = \eta_\mathsf{ci}(c, a) &
@@
 %
-\weiei: Q \times \Omegai \times Q \to \Semiring &
+\weiei: Q \times \Deltai \times Q \to \Semiring &
 \weiei(q, a, q') = \phi_\mathsf{i}(a) &
@@
 %
-\weic: Q \times [\Omegac \times P] \times  [\Omegac \times P] \times Q \to \Semiring &
+\weic: Q \times [\Deltac \times P] \times  [\Deltac \times P] \times Q \to \Semiring &
 \weic(q, c, p, \call{c'}, p', q') = \eta_\mathsf{cc}(c, \call{c'}) &
@@
 %
-\weiec: Q \times [\Omegac \times P] \times Q \to \Semiring &
+\weiec: Q \times [\Deltac \times P] \times Q \to \Semiring &
 \weiec(q, {\call{c}}, p, q') = \phi_\mathsf{c}({\call{c}}) &
@@
 %
-\weir: Q \times [\Omegac \times P] \times \Omegar \times Q \to \Semiring &
+\weir: Q \times [\Deltac \times P] \times \Deltar \times Q \to \Semiring &
 \weir(q, {\call{c}},  p, {\return{r}}, q') = \eta_\mathsf{cr}({\call{c}},  {\return{r}}) &
@@
 %
-\weier: Q \times \Omegar \times Q \to \Semiring &
+\weier: Q \times \Deltar \times Q \to \Semiring &
 \weier(q, {\return{r}}, q') = \phi_\mathsf{r}({\return{r}}) &
@@
 Moreover, $\weii$ reads a pair made of
-${\call{c}} \in \Omegac$ and $p \in P$ on the top of the stack
+${\call{c}} \in \Deltac$ and $p \in P$ on the top of the stack
 ($c$ is compared to $a$ by the weight function $\eta_\mathsf{ci} \in \Phici$).
@@
 and a stack content $\gamma \in \Gamma^*$,
-where $\Gamma = \Omegac \times P$.
+where $\Gamma = \Deltac \times P$.
 Hence, $\weight_A$ is a function from
-$[Q \times \Gamma^*] \times \Omega^* \times [Q \times \Gamma^*]$ into~$\Semiring$.
+$[Q \times \Gamma^*] \times \Delta^* \times [Q \times \Gamma^*]$ into~$\Semiring$.
 The empty stack is denoted by $\bot$, and the upmost symbol is the last pushed content.
-The following functions illustrate each of the possible cases, being : reading $a \in \Omegai$,
+The following functions illustrate each of the possible cases, being : reading $a \in \Deltai$,
 \lydia{intro to func}
-or $\call{c} \in \Omegac$, or $\return{r} \in \Omegar$ for each possible state of the stack (empty or not),
-to add to $u \in {\Omega}^*$.
+or $\call{c} \in \Deltac$, or $\return{r} \in \Deltar$ for each possible state of the stack (empty or not),
+to add to $u \in {\Delta}^*$.
 \lydia{introduced the 6 cases}
@@
 % and ${\call{c}}\, p\stacksep \gamma$
-%denotes a stack where the pair made of ${\call{c}} \in \Omegac$ and $p \in P$ is the top symbol
+%denotes a stack where the pair made of ${\call{c}} \in \Deltac$ and $p \in P$ is the top symbol
 %and $\gamma$ is the rest of stack.
@@
 \noindent
-The weight associated by $A$ to $s \in \Omega^*$
+The weight associated by $A$ to $s \in \Delta^*$
 is defined according to empty stack semantics:
@@
 $\< Q, \emptyset, \init, \bar\wei, \final >$
-over $\Omega$, $\Semiring$ and $\bar\Phi$
-with $\Omegai = \Sigma$ and $\Omegac = \Omegar = \emptyset$,
+over $\Delta$, $\Semiring$ and $\bar\Phi$
+with $\Deltai = \Sigma$ and $\Deltac = \Deltar = \emptyset$,
 and computing with an always empty stack:
@@
 Let $A_1$ and $A_2$ be two \SWVPA
-over the same $\Omega$, $\Semiring$ and $\bar\Phi$.
+over the same $\Delta$, $\Semiring$ and $\bar\Phi$.
 There exists two effectively constructible $\SWVPA$
 $A_1 \oplus A_2$ and $A_1 \otimes A_2$,
-such that for all $s \in \Omega^*$,
+such that for all $s \in \Delta^*$,
 $(A_1 \oplus A_2)(s) = A_1(s) \oplus A_2(s)$ and
@@
 We shall now present a procedure for searching, for a~$\SWVPA$ $A$,
-a word %$w \in \Omega^*$
+a word %$w \in \Delta^*$
 of minimal weight for~$A$,
@@
 For a \SWVPA $A$
-over $\Omega$,
+over $\Delta$,
 $\Semiring$ commutative, bounded, total and complete, %semiring
 and $\bar\Phi$ effective, %and $k$-convex, % label theory,
-one can construct in PTIME a word $t \in \Omega^*$
+one can construct in PTIME a word $t \in \Delta^*$
 such that $A(t)$ is minimal \wrt the natural ordering for $\Semiring$.
@@
 \begin{equation}\label{eq:bbot}
-  b_\bot(q, q') = \bigoplus_{t\in \Omega^*}
+  b_\bot(q, q') = \bigoplus_{t\in \Delta^*}
   \textstyle
@@
 Since $\Semiring$ is complete, the infinite sum in \eqref{eq:bbot} is well defined,
-and,  it is the minimum in $\Omega^*$,
+and,  it is the minimum in $\Delta^*$,
 \wrt~$\leq_\oplus$, % this ordering.
@@
 \(
-{\displaystyle\bigoplus_{t\in \Omega^*}}
+{\displaystyle\bigoplus_{t\in \Delta^*}}
 {\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
@@
 Intuitively, the function defined in \eqref{eq:btop}
-associates to $c \in \Omegac$ 
+associates to $c \in \Deltac$ 
 the minimum weight of a computation of $A$
@@
 
-over~$\Omega$, $\Semiring$ and~$\bar\Phi$,
+over~$\Delta$, $\Semiring$ and~$\bar\Phi$,
 the minimal weight %\wrt~$\leq_\oplus$,
-for a word in~$\Omega^*$.
+for a word in~$\Delta^*$.
 %
@@
 \begin{equation}\label{eq:btop}
-  b_\top(q, p, q') : c \mapsto \bigoplus_{s\in \Omega^*}
+  b_\top(q, p, q') : c \mapsto \bigoplus_{s\in \Delta^*}
   \textstyle
@@
 %\noindent\quad
-such that $d_\bot(q_1, q_2)$, resp. $\bigoplus_{c\in \Omegac} d_\top(q_1, p, q_2)(c)$, is minimal in $\Semiring$ \wrt $\leq_\oplus$$\;$
+such that $d_\bot(q_1, q_2)$, resp. $\bigoplus_{c\in \Deltac} d_\top(q_1, p, q_2)(c)$, is minimal in $\Semiring$ \wrt $\leq_\oplus$$\;$
 
@@
   d_\top(q_1, p, q_2) \otimes
-  \displaystyle\bigoplus_{\Omegai} \weii(q_2, p, q_3)\\
+  \displaystyle\bigoplus_{\Deltai} \weii(q_2, p, q_3)\\
 %\multicolumn{3}{l}{\quad
@@
   d_\bot(q_1, q_2) \otimes
-  \displaystyle\bigoplus_{\Omegai} \weiei(q_2, q_3)\\
+  \displaystyle\bigoplus_{\Deltai} \weiei(q_2, q_3)\\
 %
 d_\top(q_0, p, q_3) & \opluseq &
-  {\displaystyle\bigoplus_{\Omegac}}^2
+  {\displaystyle\bigoplus_{\Deltac}}^2
   \bigl[ \bigl( \weic(q_0, p, p', q_1) \otimes_2
   d_\top(q_1, p', q_2) \bigr) \otimes_2
-  {\displaystyle\bigoplus_{\Omegar}} \weir(q_2, p', q_3) \bigr]\\
+  {\displaystyle\bigoplus_{\Deltar}} \weir(q_2, p', q_3) \bigr]\\
 %\multicolumn{3}{l}{\quad
@@
 d_\bot(q_0, q_3) & \opluseq &
-  {\displaystyle\bigoplus_{\Omegac}}
+  {\displaystyle\bigoplus_{\Deltac}}
   \bigl(
@@
    d_\top(q_1, p, q_2) \otimes
-  \displaystyle\bigoplus_{\Omegar} \weir(q_2, p, q_3)\bigr)\\
+  \displaystyle\bigoplus_{\Deltar} \weir(q_2, p, q_3)\bigr)\\
 %
@@
   d_\bot(q_1, q_2) \otimes
-  \displaystyle\bigoplus_{\Omegar} \weier(q_2, q_3)\\
+  \displaystyle\bigoplus_{\Deltar} \weier(q_2, q_3)\\
 %
@@
 \]
-%$\weii(q_2, p, q_3)_a \in \Phic$, for $a \in \Omegai$,
+%$\weii(q_2, p, q_3)_a \in \Phic$, for $a \in \Deltai$,
 %is the partial application
@@
 %
-%$\weic(q_0, p, p', q_1)_{c'} \in \Phic$, for $c' \in \Omegac$,
+%$\weic(q_0, p, p', q_1)_{c'} \in \Phic$, for $c' \in \Deltac$,
 %is the partial application
@@
 \(
-  {\displaystyle \bigoplus_{s \in \Omega^*} A(s)} =
+  {\displaystyle \bigoplus_{s \in \Delta^*} A(s)} =
   {\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
@@
 In order to obtain effectively a witness
-(word of $\Omega^*$ with a computation of $A$ of minimal weight),
+(word of $\Delta^*$ with a computation of $A$ of minimal weight),
 we require the additional property of convexity of weight functions.
@@
 Let~$\Sigma$
-and~$\Omega = \Omegai \uplus \Omegac \uplus \Omegar$
+and~$\Delta = \Deltai \uplus \Deltac \uplus \Deltar$
 be countable input and output alphabets,
@@
 and let $\bar\Phi$ be an effective label theory over $\Semiring$,
-containing $\Phi_\Sigma$, $\Phi_{\Sigma, \Omegai}$, as well as
+containing $\Phi_\Sigma$, $\Phi_{\Sigma, \Deltai}$, as well as
 $\Phii$, $\Phic$, $\Phir$, $\Phicr$
@@
 \begin{description}
-\item[--] a \SWT $T$ over $\Sigma$, $\Omegai$, $\Semiring$, and $\bar\Phi$,
+\item[--] a \SWT $T$ over $\Sigma$, $\Deltai$, $\Semiring$, and $\bar\Phi$,
 defining a measure %between words
-$T: \Sigma^* \times \Omegai^* \to \Semiring$,
-
-\item[--] a \SWVPA $A$ over $\Omega$, $\Semiring$, and $\bar\Phi$,
-defining a measure $A: \Omega^* \to \Semiring$,
-%a \SWVPA $A$ over $\Omega$, and $\Semiring$, defining a series of nested words
-%      $A : \Omega^* \to \Semiring$,
+$T: \Sigma^* \times \Deltai^* \to \Semiring$,
+
+\item[--] a \SWVPA $A$ over $\Delta$, $\Semiring$, and $\bar\Phi$,
+defining a measure $A: \Delta^* \to \Semiring$,
+%a \SWVPA $A$ over $\Delta$, and $\Semiring$, defining a series of nested words
+%      $A : \Delta^* \to \Semiring$,
 \item[--] an input word $s \in \Sigma^*$.
@@
 %
-For all $u \in \Sigma^*$ and $t \in \Omega^*$, let
-\(d(u, t) = T\bigl(u, t|_\Omegai \bigr)\),
-where $t|_\Omegai \in \Omegai^*$ is the projection of $t$ onto~$\Omegai$,
-obtained from $t$ by removing all symbols in $\Omega \setminus \Omegai$.
+For all $u \in \Sigma^*$ and $t \in \Delta^*$, let
+\(d(u, t) = T\bigl(u, t|_\Deltai \bigr)\),
+where $t|_\Deltai \in \Deltai^*$ is the projection of $t$ onto~$\Deltai$,
+obtained from $t$ by removing all symbols in $\Delta \setminus \Deltai$.
 %\end{equation}
@@
 given the above input,
-to find $t \in \Omega^*$ %nested word $t \in \Omega$
+to find $t \in \Delta^*$ %nested word $t \in \Delta$
 minimizing \( d(s, t) \otimes A(t)\)
@@
 \begin{equation}\label{eq:distance-lang}
-d(s, t) \otimes A(t) = \displaystyle\bigoplus_{t' \in \T(\Omega)} d(s, t') \otimes A(t')
+d(s, t) \otimes A(t) = \displaystyle\bigoplus_{t' \in \Delta^*} d(s, t') \otimes A(t')
 \end{equation}
@@
 \noindent
-Let us first extend the \SWT $T$ over $\Sigma$, $\Omegai$
-into a \SWT $T'$ over $\Sigma$ and $\Omega$
+Let us first extend the \SWT $T$ over $\Sigma$, $\Deltai$
+into a \SWT $T'$ over $\Sigma$ and $\Delta$
 (and the same semiring and label theory $\Semiring$ and $\bar\Phi$),
-such that for all $u \in \Sigma^*$, and $t \in {\Omega}^*$,
-$T'(u, u) = T(u, t|_{\Omegai})$.
+such that for all $u \in \Sigma^*$, and $t \in {\Delta}^*$,
+$T'(u, u) = T(u, t|_{\Deltai})$.
 %
 The transducer $T'$ simply skips every symbol
-$b \in {\Omega} \setminus \Omegai$,
+$b \in {\Delta} \setminus \Deltai$,
 by the addition to $T$,
@@
 a \SWA $B_{s, T'}$,
-such that for all $t \in \Omega^*$, $B_{s, T'}(t) = d(s, t)$.
-%
-%This automaton is such that for all $t \in \T(\Omega)$,
+such that for all $t \in \Delta^*$, $B_{s, T'}(t) = d(s, t)$.
+%
+%This automaton is such that for all $t \in \Delta^*$,
 %\[
@@
 % = T'\bigl(s, \lin(t)\bigr)
-% = T'\bigl(s, \lin(t)|_{\Omegai}\bigr)
+% = T'\bigl(s, \lin(t)|_{\Deltai}\bigr)
 % = d(s, t).
@@
 \noindent
-Next, %we convert the input \SWTA $A$ over $\Omega$
-%into a \SWVPA $A'$ over $\hat\Omega$, using Proposition~\ref{lem:SWTA}, and
+Next, %we convert the input \SWTA $A$ over $\Delta$
+%into a \SWVPA $A'$ over $\hat\Delta$, using Proposition~\ref{lem:SWTA}, and
 we compute the \SWVPA $B_{s, T'} \otimes A$,
@@
 \noindent
-It remains to compute a best nested-word $t \in {\Omega}^*$
+It remains to compute a best nested-word $t \in {\Delta}^*$
 using the best-search procedure of Proposition~\ref{th:best-search}.
-%and convert it into a best tree in $\T(\Omega)$ in order to solve SW parsing
+%and convert it into a best tree in $\T_\Delta$ in order to solve SW parsing
 %for~$T$, $A$ and~$s$.
@@
 %accepts only the pairs $\<s, t>$ such that
-%$s$ is the projection of $t$ on $\Omegai$.
+%$s$ is the projection of $t$ on $\Deltai$.
 %This can be done with a single state $q$ and
@@
 %\begin{description}
-%\item[] $\wei(q, \varepsilon, a, q) = \one$ for all $a \in \Omegac \cup \Omegar$,
-%\item[] $\wei(q, a, a, q) = \one$ for all $a \in \Omegai$,
-%\item[] $\wei(q, a, b, q) = \zero$ for all $a, b \in \Omegai$, $a \neq b$.
+%\item[] $\wei(q, \varepsilon, a, q) = \one$ for all $a \in \Deltac \cup \Deltar$,
+%\item[] $\wei(q, a, a, q) = \one$ for all $a \in \Deltai$,
+%\item[] $\wei(q, a, b, q) = \zero$ for all $a, b \in \Deltai$, $a \neq b$.
 %\end{description}
@@
 yields a well-parenthesised word $\lin(\tau)$ over an alphabet
-$\Omega = \Omegai \uplus \Omegac \uplus \Omegar$,
-assuming \eg that $\Omegai$ contains the symbols labelling the leaves of $\tau$ (symbols of rank $0$),
-and $\Omegac$ and $\Omegar$ contain respectively one left and right parenthesis
+$\Delta = \Deltai \uplus \Deltac \uplus \Deltar$,
+assuming \eg that $\Deltai$ contains the symbols labelling the leaves of $\tau$ (symbols of rank $0$),
+and $\Deltac$ and $\Deltar$ contain respectively one left and right parenthesis
 $\ccall{b}$ and $\creturn{b}$ for each symbol $b$ labelling inner nodes of $\tau$ (symbols of rank $>0$).
 %
-%With this representation, the projection $\lin(t)|_\Omegai$ is the
+%With this representation, the projection $\lin(t)|_\Deltai$ is the
 %sequence of leaves of $\tau$. %, enumerated in a \emph{dfs}-traversal.
 %
-We show in Appendix~\ref{sec:trees} how to construct a
+We show in Appendix~\ref{sec:trees} how to construct 
 %($\SW$) tree automaton~$A$ %(in particular the TA computing on the AST of a given CF-grammar)
@@
 %The purpose of the transducer $T$ is to measure a distance between input and output words,
-%\ie between the word $s$ and the sequence leaves of a derivation tree, in $\Omegai^*$.
+%\ie between the word $s$ and the sequence leaves of a derivation tree, in $\Deltai^*$.
 %
@@
 %corresponds to the production rule $q_0 := a(q_1,\ldots, q_k)$ of a weighted CF grammar,
-%with set non-terminal symbols $Q$ and set of terminal symbols $\Omega_0$.
+%with set non-terminal symbols $Q$ and set of terminal symbols $\Delta_0$.
 %
 %This actually is a slight generalization of CFG since
-%each such production rule is labelled by  a symbol of $\Omega_{>0}$,
+%each such production rule is labelled by  a symbol of $\Delta_{>0}$,
 %hence parse trees %derivation trees
-%are trees of $\T(\Omega)$.
+%are trees of $\T_\Delta)$.
 %
 %Another (more original) generalization is that the set of terminal symbols
-%$\Omega_0$ may be infinite.
+%$\Delta_0$ may be infinite.
 %
@@
 The free $\Omega$-algebra of finite, ordered,
-$\Omega$-labeled trees is denoted by $\T(\Omega)$.
-It is the smallest set such that  $\Omega_0 \subset \T(\Omega)$
+$\Omega$-labeled trees is denoted by $\T_\Omega$.
+It is the smallest set such that  $\Omega_0 \subset \T_\Omega$
 and for all $1 \leq k \leq M$, all $a \in \Omega_k$,
-and all $t_1, \ldots, t_k \in \T(\Omega)$, $a(t_1, \ldots, t_k) \in \T(\Omega)$.
+and all $t_1, \ldots, t_k \in \T_\Omega$, $a(t_1, \ldots, t_k) \in \T_\Omega$.
 %
@@
 % $a \in \Omega_k$.
-% A tree $t \in \T(\Omega)$ is in the language of $G$
+% A tree $t \in \T_\Omega$ is in the language of $G$
 % if it can be generated from $q_\mathsf{i}$ by
@@
 defines a mapping
-from trees of $\T(\Omega)$ into~$\Semiring$, %the weight values in~$\Semiring$,
+from trees of $\T_\Omega$ into~$\Semiring$, %the weight values in~$\Semiring$,
 based on the following intermediate function
-$\weight_A: Q \times (\Omega \cup \{ \varepsilon \}) \times \T(\Omega) \to \Semiring$
+$\weight_A: Q \times (\Omega \cup \{ \varepsilon \}) \times \T_\Omega \to \Semiring$
 \begin{equation}
@@
 where $q_0 \in Q$, $a \in \Omega_{>0} \cup \{ \varepsilon \}$ and
-$t = b(t_1,\ldots, t_k) \in \T(\Omega)$,
+$t = b(t_1,\ldots, t_k) \in \T_\Omega$,
 $0 \leq k \leq M$.
@@
 \medskip\noindent
-Finally, the weight associated by $A$ to  $t \in \T(\Omega)$ is
+Finally, the weight associated by $A$ to  $t \in \T_\Omega$ is
 \begin{equation}
@@
 %
-We shall now establish a correspondence between such derivation tree $t$
+We shall now establish a correspondence between such a derivation tree $t$
 and some word describing a linearization of $t$,
@@
 
-Let $\hat\Omega$ be the countable (unranked) alphabet obtained
+Let $\Delta$ be the countable (unranked) alphabet obtained
 from $\Omega$ by:
-$\hat\Omega = \Omegai \uplus \Omegac \uplus \Omegar$, with
-$\Omegai = \Omega_0$,
-$\Omegac = \{ \; \call{a} \mid a \in \Omega_{>0} \}$,
-$\Omegar = \{ \; \return{a} \mid a \in \Omega_{>0} \}$.
-
-\noindent
-We associate to $\hat\Omega$
+$\Delta = \Deltai \uplus \Deltac \uplus \Deltar$, with
+$\Deltai = \Delta_0$,
+$\Deltac = \{ \; \call{a} \mid a \in \Omega_{>0} \}$,
+$\Deltar = \{ \; \return{a} \mid a \in \Omega_{>0} \}$.
+
+\noindent
+We associate to $\Delta$
 a label theory $\hat{\Phi}$
@@
 \noindent
-and we define a linearization of trees of $\T(\Omega)$ into
-words of $\hat{\Omega}^*$ as follows:
+and we define a linearization of trees of $\T_\Omega$ into
+words of $\Delta^*$ as follows:
 \begin{description}
@@
 there exists an effectively constructible \SWVPA $A'$ over
-$\hat\Omega$, $\Semiring$ and $\hat\Phi$
-such that for all $t \in \T(\Omega)$, $A'\bigl(\lin(t)\bigr) = A(t)$.
+$\Delta$, $\Semiring$ and $\hat\Phi$
+such that for all $t \in \T_\Omega$, $A'\bigl(\lin(t)\bigr) = A(t)$.
 \end{proposition}
@@
 $A' = \< Q', P', \init', \bar{\wei}', \final' >$,
-%computing over $\hat\Omega = \Omegai \uplus \Omegac \uplus \Omegar$,
+%computing over $\Delta = \Omegai \uplus \Omegac \uplus \Omegar$,
 where $Q' = \bigcup_{k=0}^{M} Q^k$ is the set of sequences of state symbols of $A$,

### 2021-07-08 introduction.tex

@@
 
-\begin{example}[Running example]
+\begin{example}[Running example]\label{ex:running}
 Throughout the paper we

### 2021-07-08 main.tex

@@
 
-Let $\Delta$ be the countable (unranked) alphabet obtained
+Let $\hat\Omega$ be the countable (unranked) alphabet obtained
 from $\Omega$ by:
-$\Delta = \Deltai \uplus \Deltac \uplus \Deltar$, with
-$\Deltai = \Delta_0$,
+$\hat\Omega = \Deltai \uplus \Deltac \uplus \Deltar$, with
+$\Deltai = \Omega_0$,
 $\Deltac = \{ \; \call{a} \mid a \in \Omega_{>0} \}$,
@@
 \noindent
-We associate to $\Delta$
+We associate to $\hat\Omega$
 a label theory $\hat{\Phi}$
@@
 and we define a linearization of trees of $\T_\Omega$ into
-words of $\Delta^*$ as follows:
+words of ${\hat\Omega}^*$ as follows:
 \begin{description}
@@
 there exists an effectively constructible \SWVPA $A'$ over
-$\Delta$, $\Semiring$ and $\hat\Phi$
+$\hat\Omega$, $\Semiring$ and $\hat\Phi$
 such that for all $t \in \T_\Omega$, $A'\bigl(\lin(t)\bigr) = A(t)$.

### 2021-07-08 introduction.tex
intro: symbol A_T,s

--- 2021-07-08 introduction.tex
@@
 approaches~\cite{Goodman99SemiringParsing,Nederhof03weightedParsing,MorbitzVogler19weighted-parsing}.
+\florent{\cite{MorbitzVogler19weighted-parsing} much is more general}
 In \emph{weighted language models},
@@
 where, given a \SWT $T$, a \SWVPA $A$, and an input word $s$,
-a \SWVPA $A_{T, s}$ is built, such that for all $t$, $A_{T, s}(t) = T(s, t) \otimes A(t)$.
-In the second step, a best AST $t$ is found by applying to $A_{T, s}$
+a \SWVPA $B$ %$A_{T, s}$ 
+is built, such that for all~$t$,  $B(t) = T(s, t) \otimes A(t)$.% $A_{T, s}(t) = T(s, t) \otimes A(t)$.
+In the second step, a best AST $t$ is found by applying to $B$ %$A_{T, s}$
 a best search algorithm similar to the shortest distance
@@
 \begin{enumerate}
-  \item $I_1  = [ <e_1, 0.07>,<e_2, 0.72>,<e_3, 0.91>]$, over interval $[0,1[$
-  \item $I_2  = [ <e_3, 1.05>,<e_4, 1.36>,<e_5, 1.71>]$, over interval $[1,2[$
+  \item $I_1  = [ \<e_1, 0.07>,\<e_2, 0.72>, \<e_3, 0.91>]$, over interval $[0,1[$
+  \item $I_2  = [ \<e_3, 1.05>,\<e_4, 1.36>, \<e_5, 1.71>]$, over interval $[1,2[$
 \end{enumerate}
@@
 \includegraphics[scale=0.20]{pictures/score5.png}
-and \includegraphics[scale=0.20]{pictures/score4.png}.  Wheighted
-parsing associates a cost to each solution, and our framework
+and \includegraphics[scale=0.20]{pictures/score4.png}.  
+SW-parsing associates a cost 
+\florent{weight value}
+to each solution, and our framework
 aims at selecting the best one with respect to this cost.

### 2021-07-08 semiring.tex

@@
 
-\begin{example}
+\begin{example}\label{rules-division}
 The recursive subdivision of time that leads to hierarchichal
@@
 \]
+Further binary divisions of time sub-intervals %represented respectively by~$q_2$ and~$q_3$
+are possible with:
+
+\[
+\rho_3:\,q_2 \xrightarrow{0.1} \< q_{3}, q_{3}>, \
+\rho_4:\,q_3 \xrightarrow{0.11} \< q_{4}, q_{4}>. \
+\]
+
 \endex

### 2021-07-08 main.tex
Philippe ajout deux exemples

@@
 
-\begin{example}
+\begin{example}\label{ex:SWT}
 Let us develop the example of comparison between music played by a performer,
@@
 
+\begin{example}
+Consider once more the scores of Exemple~\ref{ex:running}.
+$\Deltai$ represents the set of music events of the form $<n, t>$
+where $n$ is either a note (e.g. A4), the continuation symbol $-$, or a
+rest symbol R. $\Deltac$ and  $\Deltar$ are used to encode the score structure:
+Let $\Deltac = \{[, <]\}$ and  $\Deltar = \{], >]\}$ where $[$ (rep. $]$)
+and $<$ (resp. $>$) correspond to start/end of measures, and start/end of
+beams. Then  \includegraphics[scale=0.20]{pictures/score5.png}
+is encoded with the nested word $[0, \frac{3}{4}, \frac{7}{8}>][<1, \frac{4}{3}, \frac{5}{3}]$
+\endex
+\end{example}
+
 \begin{definition}
@@
 
-\begin{example}
-structured words with timed symbols...
-\florent{todo example VPA}
-intro language of music notation? (markup = time division, leaves = events etc)
-\end{example}
-
 \noindent
@@
 and returning a witness $t$.
+
+\begin{example}[Symbolic Weighted Parsing and the transcription problem]
+Applied to the music transcription problem, the above formalism
+is interpreted as follows:
+\begin{enumerate}
+
+  \item The \SWT $T$ evaluates a ``fitness measure''  that expresses
+    a correspondance between a performance and its notation. See Example~\ref{ex:SWT}.
+\item The \SWVPA $A$ expresses a cost related to the music notation. One possibility
+ is to relate this cost to the structural complexity: given a set of equivalent
+ representations, we aim at choosing the simpler one. For instance
+ \includegraphics[scale=0.20]{pictures/score4.png} should be favored
+ on \includegraphics[scale=0.20]{pictures/score5.png}, although both yield
+ a same linearization. Costs  can be expressed
+ in the grammar with weight associated to each production rule. See Example~\ref{rules-division}.
+
+ \end{enumerate}
+Therefore, the framework, applied to the transcription problem, allows to find an optimal
+solution that considers both the fitness of the result to the input, and the complexity of the former.
+\end{example}
 %

### 2021-07-08 main.tex
explication algo best-search & debut de preuve

--- 2021-07-08 main.tex
@@
 \noindent
-The weight associated by $A$ to $s \in \Delta^*$
+The weight associated by $A$ to $t \in \Delta^*$
 is defined according to empty stack semantics:
@@
 \begin{equation}
-A(s)  =
+A(t)  =
 {\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
 \mathsf{in}(q) \mathop{\otimes}
-\weight_A\bigl(\config{q}{\bot}, s, \config{q'}{\bot}\bigr)
+\weight_A\bigl(\config{q}{\bot}, t, \config{q'}{\bot}\bigr)
 \mathop{\otimes} \mathsf{out}(q').
@@
 \end{equation}
-Since $\Semiring$ is complete, the infinite sum in \eqref{eq:bbot} is well defined,
-and,  it is the minimum in $\Delta^*$,
-\wrt~$\leq_\oplus$, % this ordering.
-of the fonction
-$s \mapsto \weight_A(\config{q}{\sigma}, s, \config{q'}{\sigma})$.
-%
-Hence, following~\eqref{eq:SWVPA-value}, and the associativity and commutativity
-and distributivity for~$\otimes$ and~$\oplus$, the minimum of~$A(t)$ is
-\(
+Since $\Semiring$ is complete, the infinite sum in~\eqref{eq:bbot} is well defined.
+%
+Following~\eqref{eq:SWVPA-value}, and the associativity and commutativity
+and distributivity for~$\otimes$ and~$\oplus$, the minimum of~$A(t)$ is:
+\begin{equation}\label{eq:min}
+{\displaystyle\bigoplus_{t\in \Delta^*}} A(t)
+= 
 {\displaystyle\bigoplus_{t\in \Delta^*}}
@@
 \mathop{\otimes} \mathsf{out}(q')
-\).
+\end{equation}
 
@@
 In order to compute the above function $b_\bot : Q \times Q \to \Semiring$, 
-we shall use the auxiliary function $b_\top :  Q \times P \times Q \to \Phic$.
-%
-Intuitively, the function defined in \eqref{eq:btop}
-associates to $c \in \Deltac$ 
-the minimum weight of a computation of $A$
-starting in state $q$ with a stack 
-$\<c, p> \cdot \gamma \in \Gamma^+$ 
-and ending in state $q'$ with the same stack,
-such that the computation can not pop 
-the pair made of $c$ and $p$ at the top of this stack,
-but may only read these symbols.
-Moreover, $A$ may push another pair $\< c', p'>$ %call symbols 
-on the top of $\<c, p> \cdot \gamma$,
+we shall consider an auxiliary function $b_\top :  Q \times P \times Q \to \Phic$.
+%
+Intuitively, when $b_\top(q, p, q') = \phi \in \Phic$, 
+for all $c \in \Deltac$,  $\phi(c)$ is 
+the minimum weight of a computation of~$A$
+starting in state $q$ with a non-empty stack 
+$\gamma' = \<c, p> \, \gamma \in \Gamma^+$ 
+and ending in state $q'$ with the same stack $\gamma'$,
+such that moreover, 
+the computation does not pop the pair $\<c, p>$ at the top of $\gamma'$
+(\ie $\gamma'$ is left untouched during the computation).
+However, the computation can read $\<c, p>$ at the top of $\gamma'$,
+and can also push another pair $\< c', p'> \in \Gamma$ on $\gamma'$,
 following the third case of 
-in the definition \eqref{eq:SWVPA-weight} of $\weight_A$,
-and may pop $\< c', p'>$ later, following the fifth case of \eqref{eq:SWVPA-weight} (return symbol). 
+in the definition \eqref{eq:SWVPA-weight} of $\weight_A$ (call symbol).
+The pair $\< c', p'>$ can be pop later, during the computation from $q$ to $q'$,  
+following the fifth case of \eqref{eq:SWVPA-weight} (return symbol). 
 %However, it cannot apply one of the two last cases (return symbol and empty stack)
@@
 %pop symbols in $\gamma$.
+%
+Formally, in order to define $b_\top$, we consider 
+a fresh stack symbol $\top \notin \Gamma$,   %which does not belong to $\Gamma$,
+representing the above untouched stack, and let:
+%
+\begin{equation}\label{eq:btop}
+  b_\top(q, p, q') : c \mapsto \bigoplus_{s\in \Delta^*}
+  \textstyle
+  \weight_A\bigl(\configup{q}{\< c, p> \stackup \top }, s, \configup{q'}{\<c, p> \stackup \top}\bigr)
+\quad  
+\mathrm{~for~all~} c \in \Deltac  
+\end{equation}
+%
+By definition of $\weight_A$ in \eqref{eq:SWVPA-weight}, 
+using the symbol $\top$ for the part of the stack below $\<c, p>$
+(\ie the substack $\gamma$ in the above $\gamma' = \<c, p> \, \gamma$)
+ensures that this part $\gamma'$ is not touched during the computation.
+% there cannot be a pop of $\<c, p>$ followed by a push of $\<c, p>$
+% because it is not possible to push over $\top$ (push read the top of stack).
+%
+This ensures in particular that the subword 
+read during the computation is well parenthesized
+(every symbol in $\Deltac$ has a matching symbol in $\Deltar$).
+
 % Note that having a stack reduced to such a symbol makes impossible the application of the 
@@
 
-over~$\Delta$, $\Semiring$ and~$\bar\Phi$,
-the minimal weight %\wrt~$\leq_\oplus$,
-for a word in~$\Delta^*$.
-%
-
-\noindent
-%More precisely,
-We distinguish two cases : when the stack is empty, and when it is not.
-\lydia{introduced 2 cases for b}
-%
-In the case of an empty stack, let $b_\bot : Q \times Q \to \Semiring$ be such that : %the function:
-%
-EQBBOT
-%
-The term $\config{q}{\bot}, s, \config{q'}{\bot}$
-of this sum is the central expression in
-the definition \eqref{eq:weightA} of $A(s_0)$, for the minimum $s_0$
-of the function $\weight_A$.
-\lydia{so ?}
-
-%
-\noindent
-If the stack is not empty, let $\top$ be a fresh stack symbol which does not belong to $\Gamma$,
-\florent{$b_\top$ : mot bien parenthèsé $c$/$r$}
-and let $b_\top : Q \times P \times Q \to \Phic$ be such that,
-for every two states $q, q' \in Q$
-and stack symbol $p \in P$: %and $\sigma \in \{ \bot, \top \}$,
-\begin{equation}\label{eq:btop}
-  b_\top(q, p, q') : c \mapsto \bigoplus_{s\in \Delta^*}
-  \textstyle
-  \weight_A\bigl(\configup{q}{\< c, p> \stackup \top }, s, \configup{q'}{\<c, p> \stackup \top}\bigr)
-\end{equation}
-%
+%The term $\config{q}{\bot}, s, \config{q'}{\bot}$
+%of this sum is the central expression in
+%the definition \eqref{eq:weightA} of $A(s_0)$, for the minimum $s_0$
+%of the function $\weight_A$.
 
@@
 
-\textbf{initially} let $\Q = (Q \times Q) \cup (Q \times P \times Q)$, %contains all
+\textbf{initially} let $\Q = (Q \times Q) \cup (Q \times P \times Q)$, 
 and let $d_\bot(q_1, q_2) = d_\top(q_1, p, q_2) = \one$
@@
 Algorithm~\ref{algo:Dijkstra}
-constructs iteratively markings
+constructs iteratively two markings
 $d_\bot : Q \times Q \to \Semiring$ and
-$d_\top : Q \times P \times Q \to \Phic$
+$d_\top : Q \times P \times Q \to \Phic$,
 %of the triplets $\<q, \sigma, q'>$
@@
 %$Q \times \{ \bot, \top \} \times Q$.
+%
+\noindent
+\florent{explication Fig.~\ref{fig:best-update} suivant cas de \eqref{eq:SWVPA-weight}}
+The infinite sums in the updates of $d$ in Algorithm~\ref{algo:Dijkstra},
+Figure~\ref{fig:best-update}
+are well defined
+since~$\Semiring$ is complete.
 
@@
 
-\noindent
-\florent{explication Fig.~\ref{fig:best-update} suivant cas de \eqref{eq:SWVPA-weight}}
-The infinite sums in the updates of $d$ in Algorithm~\ref{algo:Dijkstra},
-Figure~\ref{fig:best-update}
-are well defined
-since~$\Semiring$ is complete.
 \florent{complete **}
 ** effectively computable by hypothesis that the label theory is effective**
-The algorithm performs $2.|Q|^2$ iterations until $P$ is empty,
+termination: The algorithm performs $2.|Q|^2$ iterations until $P$ is empty,
 and each iteration has a time complexity $O(|Q|^2 . |P|)$.
@@
 %$|Q|^3.\log(|Q|^2)$
-
-The correctness of Algorithm~\ref{algo:Dijkstra}
-is ensured by the invariant expressed in the following lemma.
-\begin{lemma}\label{lem:bot}
-For all $\< q_1, q_2> \notin \Q$,
-$d_\bot(q_1, q_2) =  b_\bot(q_1, q_2)$/
-\end{lemma}
-The proof is by contradiction,
-assuming a counter-example minimal in the length of the witness word.
-
-\begin{lemma}\label{lem:top}
-For all $\< q_1, p, q_2> \notin \Q$,
-$d_\top(q_1, p, q_2) =  b_\top(q_1, p, q_2)$,
-\end{lemma}
-
-\noindent
-For computing the minimal weight of a computation of $A$, we use the fact that,
-at the termination of Algorithm~\ref{algo:Dijkstra}, %it holds that,
-%There exist $q_1, q_2 \in Q$
-\(
-  {\displaystyle \bigoplus_{s \in \Delta^*} A(s)} =
-  {\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
-  \mathsf{in}(q) \mathop{\otimes} d_\bot(q, q') \mathop{\otimes} \mathsf{out}(q').
-\)
+%
+At termination, $d_\bot = b_\bot$ and $d_\top = b_\top$
+(see Appendix~\ref{sec:bestsearch}).
+With~\eqref{eq:min}, this ensures
+the correctness of Algorithm~\ref{algo:Dijkstra}.
 
@@
 (word of $\Delta^*$ with a computation of $A$ of minimal weight),
-we require the additional property of convexity of weight functions.
+we require the additional property of convexity of weight functions.***
+\florent{completer}
 
@@
 
-\section{Nested-Words and Parse-Trees}
-\label{sec:trees}
+\section{Nested-Words and Parse-Trees} \label{sec:trees}
 The hierarchical structure of nested-words, defined with the \emph{call} and \emph{return} markup symbols
@@
 
+\section{Correctness of the Best-Search Algorithm} \label{sec:bestsearch}
+\input{app-bestsearch}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2021-07-08 app-trees.tex
appendices et commentaires

--- 2021-07-08 main.tex
@@
-% SWT without epsilon-transitions
-% SWT do not read the other tape in "skip" transitions 10 or 01
-%     (hence this version is simpler  than V2)
-% SWA without epsilon-transitions
-% SW-VPA read top of stack (when not empty) at every transition
-%        (6 kinds of transitions).
+% !TEX root = main.tex
+%
+% Correspondance between nested-words and trees
 
-% default style article.cls
-%\input{main-article}
-
-% Springer llncs style
-% https://www.springer.com/gp/computer-science/lncs/conference-proceedings-guidelines
-%\input{main-llncs}
-
-% EPTCS style
-% http://www.eptcs.org
-%\input{main-eptcs}
-
-% LIPICS v.2021 style
-% https://submission.dagstuhl.de/documentation/authors
-\input{main-lipics}
-
-% comments in margin with \todo command
-\usepackage[colorinlistoftodos]{todonotes} % option disable to hide todos
-% command todo : options
-% - noline: no line connecting the note with the place in the text where the note occurs in the latex code
-% - fancyline : curved arrow from note to inssertion point
-% - inline: place a todonote inside the text instead of in the margin
-% - author=name
-\newcommand{\florent}[1]{\todo[noline,size=\tiny,color=yellow!40]{#1}}
-\newcommand{\lydia}[1]{\todo[noline,size=\tiny,color=red!30]{#1}}
-\newcommand{\philippe}[1]{\todo[noline,size=\tiny,color=green!30]{#1}}
-
-\usepackage{hyperref}
-%\usepackage[bookmarks,bookmarksnumbered,naturalnames,plainpages=false]{hyperref}
-%usepackage{url}
-
-% for footnote ref
-\usepackage{refcount}
-
-% array and tabular
-\usepackage{array}
-\newcolumntype{C}[1]{>{\centering\arraybackslash\hspace{0pt}}p{#1}}
-
-% extension of enumerate env. (style for displaying counters)
-% \usepackage{enumerate}
-
-%% pictures
-% \usepackage{graphicx}
-% \DeclareGraphicsExtensions{.pdf,.png,.jpg}
-% \graphicspath{fig/}
-
-%% PGF, Tikz
-\usepackage{tikz-cd}
-%% \usepackage{pgfplots}
-%% \usepgfplotslibrary{dateplot}
-%% \usepackage{pgf,pgfarrows,pgfnodes, pgfautomata}
-% \usepackage{tikz}
-% \usetikzlibrary{cd}
-%% \usetikzlibrary{arrows}
-%% \usetikzlibrary{calc}
-%% \usetikzlibrary{snakes}
-%% \usetikzlibrary{backgrounds}
-% \usetikzlibrary{trees}
-%% \usetikzlibrary{automata}
-%% \usetikzlibrary{positioning}
-%% \usetikzlibrary{matrix}
-%% \usetikzlibrary{patterns}
-%% \usetikzlibrary{shapes}
-
-% symbols
-\usepackage{amsmath}
-\allowdisplaybreaks
-\usepackage{amssymb}
-\usepackage{amsbsy}
-\usepackage{bbold}
-\usepackage{latexsym}
-%\usepackage{amsfonts}
-\usepackage{stmaryrd}
-%\usepackage{mathabx}
-%\usepackage{MnSymbol}
-\usepackage{harmony} % simple music fonts
-\usepackage{mathtools} % for arrows
-%\usepackage{mathptmx}
-\usepackage{comment}
-
-%% algorithms
-%\usepackage{algorithm}
-%\usepackage{program}
-\usepackage[ruled,vlined]{algorithm2e}
-
-%% allows for page break within arrays
-\usepackage{longtable}
-
-%% arrows etc
-%\input{rewriting}
-
-%% music symbols
-% see http://tug.ctan.org/info/latex4musicians/latex4musicians.pdf
-\usepackage{musicography}
-
-%% for new macros
-\usepackage{xspace}
-
-%% Misc macros
-\def\ie{\textit{i.e.}\xspace}
-\def\eg{\textit{e.g.}\xspace}
-\def\wrt{\textit{wrt}\xspace}
-%\def\wlog{\textit{wlog}\xspace}
-\def\etc{\textit{etc}\xspace}
-
-\def\<#1>{\langle #1 \rangle}
-\newcommand{\pair}[2]{\langle{#1}, {#2}\rangle}
-
-%\newcommand{\config}[2]{\ensuremath \genfrac{[}{]}{0pt}{}{#1}{#2}}
-\newcommand{\config}[2]{\ensuremath{#1}[{#2}]}
-\newcommand{\configup}[2]{\ensuremath{#1}\left[\begin{array}{c} #2 \end{array}\right]}
-\def\stacksep{\cdot}
-\def\stackup{\\}
-
-\newcommand{\opluseq}{\ensuremath\mathrel{\oplus}=}
-
-% \bigominus
-\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
-\DeclareFontShape{U}{mathx}{m}{n}{
-<-6> mathx5 <6-7> mathx6 <7-8> matha7
-<8-9> mathx8 <9-10> mathx9
-<10-12> mathx10 <12-> mathx12
-}{}
-\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
-\DeclareMathSymbol{\bigominus}{\mathop}{mathx}{"C1}
-
-\newcommand{\A}{\mathcal{A}} % automata, VPA
-\newcommand{\B}{\mathcal{B}} % automata (in constructions)
-\newcommand{\C}{\mathcal{C}} % automata (in constructions)
-\newcommand{\D}{\mathcal{D}}  % transducer = distance
-\newcommand{\E}{\mathbb{E}}
-%\newcommand{\P}{\mathcal{P}}
-\newcommand{\Q}{\mathcal{Q}}
-\newcommand{\R}{\mathcal{R}}
-\newcommand{\T}{\mathcal{T}}
-\newcommand{\W}{\mathbb{W}}
-\newcommand{\Semiring}{\mathbb{S}}
-\newcommand{\zero}{\mathbb{0}}
-\newcommand{\one}{\mathbb{1}}
-\newcommand{\dom}{\ensuremath{\mathit{dom}}}
-
-\def\SA{\textsf{sA}\xspace}
-\def\WA{\textsf{wA}\xspace}
-\def\SW{\textsf{sw}\xspace}
-\def\swM{\textsf{swM}\xspace}
-\def\SWT{\textsf{swT}\xspace}
-\def\SWA{\textsf{swA}\xspace}
-\def\SWTA{\textsf{swTA}\xspace}
-\def\SWVPA{\textsf{sw-VPA}\xspace}
-\def\VPA{\textsf{VPA}\xspace}
-\def\SVPA{\textsf{sVPA}\xspace}
-\def\weight{\mathsf{weight}}
-\def\wei{\mathsf{w}}
-\def\mei{\mathsf{m}}
-\def\init{\mathsf{in}}
-\def\final{\mathsf{out}}
-\def\Omegai{{\Omega_\mathsf{i}}}
-\def\Omegac{{\Omega_\mathsf{c}}}
-\def\Omegar{{\Omega_\mathsf{r}}}
-\def\Sigmai{{\Sigma_\mathsf{i}}}
-\def\Sigmac{{\Sigma_\mathsf{c}}}
-\def\Sigmar{{\Sigma_\mathsf{r}}}
-\def\Deltai{{\Delta_\mathsf{i}}}
-\def\Deltac{{\Delta_\mathsf{c}}}
-\def\Deltar{{\Delta_\mathsf{r}}}
-\def\Phii{{\Phi_\mathsf{i}}}
-\def\Phic{{\Phi_\mathsf{c}}}
-\def\Phir{{\Phi_\mathsf{r}}}
-\def\Phix{{\Phi_\mathsf{x}}}
-\def\Phici{{\Phi_\mathsf{ci}}}
-\def\Phicc{{\Phi_\mathsf{cc}}}
-\def\Phicr{{\Phi_\mathsf{cr}}}
-\def\Phicx{{\Phi_\mathsf{cx}}}
-\def\weii{{\wei_\mathsf{i}}}
-\def\weic{{\wei_\mathsf{c}}}
-\def\weir{{\wei_\mathsf{r}}}
-\newcommand{\weie[1]}{\wei_{#1}^\mathsf{e}}
-\def\weiei{\weie[\mathsf{i}]}
-\def\weiec{\weie[\mathsf{c}]}
-\def\weier{\weie[\mathsf{r}]}
-\def\weiex{\weie[\mathsf{x}]}
-\newcommand{\ioi}[1]{\mathsf{ioi}({#1})}
-\newcommand{\rank}{\mathsf{rk}}
-\newcommand{\lin}{\mathsf{lin}}
-
-\newcommand{\call}[1]{\ensuremath #1} %{\ensuremath \langle_{#1}}
-\newcommand{\return}[1]{\ensuremath #1} %{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
-\newcommand{\ccall}[1]{\ensuremath \langle_{#1}}
-\newcommand{\creturn}[1]{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
-
-%\sloppy
-
-% Parsing over infinite alphabet as optimal alignment computation
-% as edit-distance between string and language
-%
-%\title{Symbolic Weighted Parsing and Automated Music Transcription}
-%\title{Symbolic Weighted Language Models and Automated Music Transcription}
-\title{Symbolic Weighted Language Models and\\ Quantitative Parsing over Infinite Alphabets}
-%\title{Weighted Visibly Pushdown Automata and Automated Music Transcription}
-%\titlerunning{WVPA \& AMT}
-%\titlerunning{Symbolic Weighted Parsing} % and Automated Music Transcription
-%\authorrunning{Florent Jacquemard}
-
-\date{\today}
-
-\begin{document}
-\thispagestyle{empty}
-\maketitle
-
-\begin{abstract}
-\input{abstract}
-\end{abstract}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% intro
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Introduction} \label{sec:intro}
-\input{introduction}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% prelin
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Preliminary Notions}
-\label{section:prelim}\label{sec:prelim}
-
-%notations: for set $S$ : sets of sequences $S^*$ and $S^+$...
-%interval $[i..j]$ of natural numbers
-
-\paragraph*{Semirings}
-\label{section:semiring}\label{sec:semiring}
-\input{semiring}
-
-\paragraph*{Label Theory}
-\label{section:symbols}
-\input{labels}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SWT & SWA
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{SW Automata and Transducers}
-\label{section:transducer}\label{sec:transducer}
-\label{section:SWA}\label{sec:SWA}
-\label{section:SWT}\label{sec:SWT}
-
-We follow the approach of~\cite{Mohri03EDWA} for the computation of distances,
-between words and languages, using weighted transducers,
-and extend it to infinite alphabets.
-% with models of symbolic weighted automata and transducers.
-%
-The models introduced in this section generalize
-weighted automata and transducers~\cite{Droste09handbook}
-%over finite alphabets, see  e.g.~\cite{Mohri03EDWA},
-by labeling each transition with a weight function (instead of a simple weight value),
-that takes the input and output symbols as parameters.
-These functions are similar to the guards of symbolic automata~\cite{dAntoniVeanes17CAV,dAntoni21CACM},
-but they can return values in a generic semiring,
-whereas the latter guards are restricted to the Boolean semiring.
-%
-
-%\subsection{Definitions} \label{sec:SWTdef}\label{sec:SWAdef}
-%(SWT)
-\noindent
-Let $\Semiring$ be a commutative semiring,
-$\Sigma$ and $\Delta$ be alphabets called respectively \emph{input} and \emph{output}, %{alphabets},
-and $\bar\Phi$ be a label theory over $\Semiring$
-containing $\Phi_\Sigma$, $\Phi_\Delta$, $\Phi_{\Sigma, \Delta}$.
-
-\begin{definition}
-\label{def:transducer} \label{def:SWT}
-A \emph{symbolic-weighted transducer} (\SWT)
-over $\Sigma$, $\Delta$, $\Semiring$ and $\bar\Phi$
-%the input and output alphabets~$\Sigma$ and $\Delta$ with label theory $\bar\Phi$, and the semiring $\Semiring$
-is a tuple
-$T = \< Q, \init, \bar{\wei}, \final >$,
-where $Q$ is a finite set of states,
-$\mathsf{in} : Q \to \Semiring$   %\Phi_{\Sigma, \Delta}
-(respectively $\mathsf{out} : Q \to \Semiring$)  %\Phi_{\Sigma, \Delta}
-are functions defining the weight for entering
-(respectively leaving) computation in a state,
-and $\bar{\wei}$ is a triplet of transition functions
-$\wei_{10}: Q \times Q \to \Phi_{\Sigma}$,
-$\wei_{01}: Q \times Q \to \Phi_{\Delta}$, and
-$\wei_{11}: Q \times Q \to \Phi_{\Sigma, \Delta}$.
-\end{definition}
-%
-\noindent
-We call \emph{number of transitions} of $T$ the number of pairs of states
-$q, q' \in Q$ such that $\wei_{10}$ or $\wei_{01}$ or $\wei_{11}$
-is not the constant $\zero$.
-%
-\noindent
-For convenience, we shall sometimes present transitions
-as functions of
-$Q \times (\Sigma \cup \{ \varepsilon \}) \times (\Delta \cup \{ \varepsilon \}) \times Q \to \Semiring$,
-overloading the function names,
-such that, for all $q, q' \in Q$, $a \in \Sigma$,  $b \in \Delta$,
-
-\[
-\begin{array}{rcll}
-\wei_{10}(q, a, \varepsilon, q') & = & \phi(a) & %\wei_\Sigma(q, q')(a)
-\quad\mathrm{where~} \phi = \wei_{10}(q, q') \in \Phi_\Sigma,\\
-\wei_{01}(q, \varepsilon, b, q') & = & \psi(b) &
-\quad\mathrm{where~} \psi = \wei_{01}(q, q') \in \Phi_\Delta,\\
-\wei_{11}(q, a, b, q') & = & \eta(a, b) &
-\quad\mathrm{where~} \eta = \wei_{11}(q, q') \in \Phi_{\Sigma, \Delta}.\\
-\end{array}
-\]
-%
-\noindent
-% The symbolic-weighted transducer
-The \SWT $T$ computes on pairs of words $\< s, t> \in \Sigma^* \times \Delta^*$,
-$s$ and $t$, being respectively called \emph{input} and \emph{output} word.
-% \noindent
-More precisely, $T$ defines a mapping
-from $\Sigma^* \times \Delta^*$ into~$\Semiring$,
-based on an intermediate function $\weight_T$
-defined recursively, for every states $q, q' \in Q$,
-and every pairs of strings $\< s, t> \in \Sigma^* \times \Delta^*$,
-% \setminus \{ \< \varepsilon, \varepsilon >\}$,
-where $au$, and $bv$, denote the concatenation
-of the symbol $a \in \Sigma$ (resp. $b \in \Delta$)
-with a word $u \in \Sigma^*$ (resp. $v \in \Delta^*$).
-\lydia{added $u$ and $v$ def}
-%
-%$a \in \Sigma$, $u \in \Sigma^*$, $b\in \Delta$, $v\in \Delta^*$, by:
-%
-\begin{align}
-%\weight_T(q, \varepsilon, \varepsilon, q)  & = \one \label{eq:SWT-weight}\\ %\final(q)\\
-%\weight_T(q, \varepsilon, \varepsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
-\weight_T(q, \varepsilon, \varepsilon, q')  & = \one
- \quad \mathrm{if~} q = q' \mathrm{~and~} \zero \mathrm{~otherwise} \label{eq:SWT-weight}\\
-\weight_T(q, s, t, q') & = \displaystyle\bigoplus_{\begin{array}{c}
-                                                   \scriptstyle q'' \in Q\\[-1ex]
-                                                   \scriptstyle s = au,\, a \in \Sigma
-                                                   \end{array}}
-    \wei_{10}(q, a, \varepsilon, q'') \otimes \weight_T(q'', u, t, q')\nonumber\\
-                    & \oplus \displaystyle\bigoplus_{\begin{array}{c}
-                                                     \scriptstyle q'' \in Q\\[-1ex]
-                                                     \scriptstyle t = bv,\, b \in \Delta\\
-                                                     \end{array}}
-    \wei_{01}(q, \varepsilon, b, q'') \otimes \weight_T(q'', s, v, q')\nonumber\\
-                    & \oplus \displaystyle\bigoplus_{\begin{array}{c}
-                                                     \scriptstyle q'' \in Q\\[-1ex]
-                                                     \scriptstyle s = au,\, t = bv\\
-                                                     \end{array}}
-    \wei_{11}(q, a, b, q'') \otimes \weight_T(q'', u, v, q')\nonumber
-\end{align}
-
-%
-We recall that, by convention (Section~\ref{sec:semiring}),
-an empty sum with $\bigoplus$ is equal to~$\zero$.
-%
-Intuitively, using a transition $\wei_{ij}(q, a, b, q')$ means for $T$:
-when reading respectively $a$ and $b$ at the current positions in the input and output words,
-increment the current position in the input word if and only if $i = 1$,
-and in the output word iff $j = 1$, %(otherwise, do not change it),
-and change state from $q$ to $q'$.
-When $a = \varepsilon$ (resp. $b = \varepsilon$), the current symbol
-in the input (resp. output) is not read.
-%
-%In contrast with the models of weighted transducers over finite alphabets~\cite{Mohri03ijfcs},
-%the input and output symbols at current positions are always read by transitions,
-%even when they do not change the reading position the head's position.
-%This is an important feature in the case of an infinite alphabet in
-%order to compare input and output symbols.
-%which cannot be stored in the finite memory of the transducer.
-%
-Since~$\zero$ is absorbing for~$\otimes$ in~$\Semiring$,
-one term $\wei_{ij}(q, a, b, q'')$ equal to $\zero$ in the above expression
-will be ignored in the sum, meaning that there is no possible transition
-from state $q$ into state $q'$ while reading $a$ and $b$.
-This is analogous to the case of a transition's guard not satisfied by $\<a, b>$ for
-symbolic transducers.
-
-%whereas considering the current symbol may be useful to compute a transition weight.
-%(even when it does not change the head's position, like with $\varepsilon$-transitions).
-%
-%The cases $\weight_T(q, au, \varepsilon, q')$ and $\weight_T(q, \varepsilon, bv, q')$
-%are missing in the definition of $\weight_T$.
-%It means that $T$ must avoid configurations where it reached the end of
-%the output word and not of the input one, or vice-versa.
-%This can be done by using $\wei_{10}$ and $\wei_{01}$
-%before reaching the end of word, and using a special state for this purpose.
-
-The expression \eqref{eq:SWT-weight}
-can be seen as a stateful definition of
-an edit-distance between a word $s \in \Sigma^*$ and a word $t \in \Delta^*$,
-see also~\cite{Mohri03ijfcs}.
-Intuitively,
-$\wei_{10}(q, a, \varepsilon, r)$ is the cost of
-the deletion of the symbol~$a \in \Sigma$ in~$s$,
-$\wei_{01}(q, \varepsilon, b, r)$ is the cost
-of the insertion of~$b \in \Delta$ in $t$,
-and $\wei_{11}(q, a, b, r)$ is the cost
-of the substitution of  $a \in \Sigma$ by~$b \in \Delta$.
-%
-The cost of a sequence of such operations transforming $s$ into $t$,
-is the product, with $\otimes$, of the individual costs of the operations involved;
-and the distance between $s$ and $t$ is the sum, with $\oplus$,
-of all possible products.
-%
-\medskip\noindent
-%Let $\< s, t> \in \Sigma^* \times \Delta^*$, with $s = s_1\ldots s_n$, and $t = t_1\ldots t_m$.
-Formally, the weight associated by $T$ to $\< s, t> \in \Sigma^* \times \Delta^*$ is:
-%defined as follows:
-\begin{equation}
-T(s, t)  =
-\displaystyle\bigoplus_{q, q' \in Q} \init(q)
-\mathop{\otimes} \weight_T(q, s, t, q') \mathop{\otimes} \final(q')
-\label{eq:SWT-value}
-\end{equation}
-
-%\begin{example}
-%We can define with a \SWT the computation of a similarity measure
-%between timed sequences
-%similar to dynamic time warping (DTW).
-%
-%Let $\Semiring$  be the tropical (\emph{min-plus}) semiring of Figure~\ref{fig:semirings} and
-%let $\Sigma = \Delta = \mathbb{R}_+$ be sets of timestamps.
-%We consider a \SWT with one state $q$ and transitions
-%$\wei_{11}(q, d, d', q) =
-% \wei_{01}(q, d, d', q) =
-% \wei_{10}(q, d, d', q) = |d' - d|$,
-%for all $d, d' \in \mathbb{R}_+$.
-%% needs reading input/output symbols even by epsilon-transitions.
-%The recursive definition of $\weight_T$ correspond to the dynamic programming equations of DTW
-%for the computation of an optimal match between words,
-%the matching cost for two symbols being the
-%the time distance between them.
-%\endex
-%\end{example}
-
-\begin{example}
-Let us develop the example of comparison between music played by a performer,
-represented as a sequence $s \in \Sigma^*$ of events in the MIDI alphabet $\Sigma$,
-and  a music score represented as a sequence $t \in \Delta^*$
-in the CWMN alphabet $\Delta$. We build a small weighted transducer model
- with two states $q_0$ and $q_1$ that
-calculates the distance between $s$ and $t$.
-
-If one performed event $s_i$  corresponds
-to one notated event $t_1$ (for instance MIDI code 61 and pitch A4),
-the weight value computed by the \SWT is the time distance between both,
-as in Example~\ref{distance-time}, and is modeled by
-transitions $\wei_{11}$ below.
-%
-If we meet the music notation symbol '-' that
-represents continuation (such as instance in \emph{ties}
-$\musQuarter\!\!\!\mathrel{\raisebox{-1.5mm}{$\smile$}}\!\!\!\musEighth$,
-or \emph{dots} \musQuarterDotted{}), it is  skipped with no cost (transitions $\wei_{01}$ or weight $\one$).
-%transitions
-\[
-\begin{array}{rclcrcl}
-\wei_{11}(q_0, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d| & \quad &
-\wei_{11}(q_1, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d|\\
-\wei_{01}(q_0, \varepsilon, \< \mathsf{-}, d'>, q_0) & = & \one & &
-\wei_{01}(q_1, \varepsilon, \< \mathsf{-}, d'>, q_0) & = & \one\\
-\wei_{10}(q_0, d, \varepsilon, q_1) & = & \alpha & & %\multicolumn{3}{l}{\mathrm{for~all~} b \in \Delta}
-\end{array}
-\]
-%
-We also must be able to take performing errors into account, while still being able to compare with the score,
-since a performer could, for example, play an unwritten extra note.
-\lydia{reformulated this sentence}
-%
-This is modelled by the transition $\wei_{10}$ with an arbitrary weight value $\alpha \in \Semiring$,
-switching from state $q_0$ (normal) to $q_1$ (error).
-\philippe{Comprends pas cette phrase}
-The transitions in the second column below switch back to the normal state $q_0$.
-% the metric computed by the \SWT is the smallest sum of point wise distances
-% between dates of input and output events.
-At last, we let $q_0$ be the only initial and final state, with
-$\init(q_0) = \final(q_0) = \one$, and
-$\init(q_1) = \final(q_1) = \zero$.
-
-\lydia{ccl to the ex}
-That way, an \SWT is capable of evaluating the differences between a score and a performance,
-all the while ensuring that performance errors are plausible.
-%$\init(q_0, d, b) = \final(q_0, d, b) = \one$, and
-%$\init(q_1, d, b) = \final(q_1, d, b) = \zero$,
-%for all $d \in \Sigma$ and $b \in \Delta$.
-\endex
-\end{example}
-
-\begin{comment}
-\begin{example}
-In Common Western Music Notation~\cite{Gould11Notation},
-several symbols may be used to represent one single sounding event.
-\lydia{unique $\to$ similar}
-\florent{similar $\to$ single}
-For instance, several notes can be combined with a tie,
-like in $\musQuarter\!\!\!\mathrel{\raisebox{-1.5mm}{$\smile$}}\!\!\!\musEighth$,
-and one note can be augmented by half its duration with a dot like in~\musQuarterDotted{}.
-\lydia{modif.}
-These notations are perceived equivalent when played, as their duration is equal, yet the notation is different.
-We thus want to be able to compare a music score with music played by a performer.
-\lydia{changed end}
-%
-\noindent
-We propose a small weighted transducer model with two states $q_0$ and $q_1$
-that calculates the distance bewteen an input sequence of sounding
-events (music "performance")
-to an output sequence of written events (music "score").
-% simple pointwise distance between two sequences of timestamped events **
-%
-Let us consider the tropical (\emph{min-plus}) semiring~$\Semiring$
-of Figure~\ref{fig:semirings} and
-let $\Sigma = \mathbb{R}_+$ be an input alphabet of event dates
-and $\Delta = \{ \mathsf{e}, \mathsf{-} \} \times \mathbb{R}_+$
-be an output alphabet of symbols with timestamps.
-A symbol $\< \mathsf{e}, d > \in \Delta$ represents an event starting at date $d$,
-and $\< \mathsf{-}, d >$ is a continuation of the previous event.
-%More precisely, we let
-%$\Sigma = \{ a, - \} \times \R_+$ and
-%This example of $\Delta$ is motivated by the case of music notation,
-%where several notated events (notes) can be tied together,
-%with a \emph{tie} or a \emph{dot}
-%meaning that they will be played as a unique sounding event.
-%The timestamp of $a \in \Sigma$, denoted by $\mathsf{t}(a)$, is expressed as a rational number.
-
-We consider a \SWT with two states $q_0$ and $q_1$ whose purpose
-is to compare a recorded performance $s \in \Sigma^*$
-with a notated music sheet $t \in \Delta^*$.
-One timestamp $d_i \in \Sigma$ may correspond
-to one notated event $\<\mathsf{e}, d'_i> \in \Delta$, in which case
-the weight value computed by the \SWT is the time distance between both
-(see transitions $\wei_{11}$ below).
-%
-If $\<\mathsf{e}, d'_i>$ is followed by continuations
-$\<\mathsf{-}, d'_{i+1}>$..., they are just skipped with no cost (transitions $\wei_{01}$ or weight $\one$).
-%transitions
-\[
-\begin{array}{rclcrcl}
-\wei_{11}(q_0, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d| & \quad &
-\wei_{11}(q_1, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d|\\
-\wei_{01}(q_0, \varepsilon, \< \mathsf{-}, d'>, q_0) & = & \one & &
-\wei_{01}(q_1, \varepsilon, \< \mathsf{-}, d'>, q_0) & = & \one\\
-\wei_{10}(q_0, d, \varepsilon, q_1) & = & \alpha & & %\multicolumn{3}{l}{\mathrm{for~all~} b \in \Delta}
-\end{array}
-\]
-%
-We also must be able to take performing errors into account, while still being able to compare with the score,
-since a performer could, for example, play an unwritten extra note.
-\lydia{reformulated this sentence}
-%
-This is modelled by the transition $\wei_{10}$ with an arbitrary weight value $\alpha \in \Semiring$,
-switching from state $q_0$ (normal) to $q_1$ (error).
-The transitions in the second column below switch back to the normal state $q_0$.
-% the metric computed by the \SWT is the smallest sum of point wise distances
-% between dates of input and output events.
-At last, we let $q_0$ be the only initial and final state, with
-$\init(q_0) = \final(q_0) = \one$, and
-$\init(q_1) = \final(q_1) = \zero$.
-
-\lydia{ccl to the ex}
-That way, an \SWT is capable of evaluating the differences between a score and a performance,
-all the while ensuring that performance errors are plausible.
-
-%$\init(q_0, d, b) = \final(q_0, d, b) = \one$, and
-%$\init(q_1, d, b) = \final(q_1, d, b) = \zero$,
-%for all $d \in \Sigma$ and $b \in \Delta$.
-\endex
-\end{example}
-\end{comment}
-
-\noindent
-The \emph{Symbolic Weighted Automata} %$A = \< Q, \init, \weight, \final >$
-%over $\Sigma$,  $\Semiring$ and $\bar\Phi$
-are defined similarly as the transducers of Definition~\ref{def:SWT},
-by simply omitting the output symbols.
-%
-%In this case, the label theory $\bar\Phi$ can be reduced to a singleton $\< \Phi_\Sigma>$.
-%
-\begin{definition} \label{def:SWA}
-A \emph{symbolic-weighted automaton} (\SWA)
-over $\Sigma$, $\Semiring$ and $\bar\Phi$
-%the input alphabet~$\Sigma$ and the commutative semiring $\Semiring$
-is a tuple
-$A = \< Q, \init, {\wei_1}, \final >$,
-where $Q$ is a finite set of states,
-$\mathsf{in} : Q \to \Semiring$ %\Phi_\Sigma$,
-(respectively $\mathsf{out} : Q \to \Semiring$)  %\Phi_\Sigma,$
-are functions defining the weight for entering
-(respectively leaving) computation in a state,
-and ${\wei_1}$ is a transition function
-from $Q \times Q$ into~$\Phi_{\Sigma}$.
-\end{definition}
-%
-\noindent
-As above in the case of \SWT,
-when $\wei_1(q, q') = \phi \in \Phi_\Sigma$,
-%respectively $\mathsf{in}(q) = \phi$, $\mathsf{out}(q') = \phi$,
-we may write $\wei_1(q, a, q')$ for~$\phi(a)$.
-%respectively $\mathsf{in}(q, a) = \phi(a)$, $\mathsf{out}(q', a) = \phi(a)$.
-%$\wei_1: Q \times \Sigma \times Q \to \Semiring$,
-The computation of $A$ on words $s \in \Sigma^*$
-is defined with an intermediate function $\weight_A$,
-defined as follows for $q, q' \in Q$, $a \in \Sigma$, $u \in \Sigma^*$,
-%
-\begin{align}
-\weight_A(q, \varepsilon, q) & = \one\\ %\final(q)\\
-\weight_A(q, \varepsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
-\weight_A(q, au, q') & =  \displaystyle\bigoplus_{q'' \in Q}
-    \wei_{1}(q, a, q'') \otimes \weight_A(q'', u, q')\nonumber
-\label{eq:SWA-weight}
-\end{align}
-%
-\noindent
-and the weight value associated by $A$ to
-$s \in \Sigma^*$ is defined as follows: %$s = s_1\ldots s_n \in \Sigma^+$
-\begin{equation}
-A(s)  =
-\displaystyle\bigoplus_{q, q' \in Q} \init(q)
-\mathop{\otimes} \weight_A(q, s, q') \mathop{\otimes} \final(q')
-\label{eq:SWA-value}
-\end{equation}
-
-%
-%When $\wei_\varepsilon(q, q') = \zero$ for all $q, q' \in Q$,
-%the automaton~$A$ is called \emph{without $\varepsilon$-transitions}.
-
-%The \emph{summary} of a $\SWT$, resp. a $\SWA$, is ***
-
-%\subsection{Properties}
-\noindent
-\philippe{Il me manque une explication:  on construit un automate
-qui, étant donnée une partition $t$, renvoie la distance minimale
-avec n'importe quelle performance (distance donnée par un transducer)?
-Quel est le rôle de A(s)?
-}
-The following property will be useful to the approach on
-symbolic weighted parsing presented in Section~\ref{sec:parsing}.
-
-\begin{proposition} \label{prop:epsilon}
-Given a \SWT $T$ over $\Sigma$, $\Delta$,
-$\Semiring$ commutative, bounded and complete,
-and $\bar\Phi$ effective,
-and a $\SWA$ $A$ over $\Sigma$, $\Semiring$ and $\bar\Phi$,
-there exists an effectively constructible \SWA
-$B_{A, T}$ over $\Delta$, $\Semiring$ and $\bar\Phi$,
-such that for all $t \in \Delta^*$,
-$B_{A, T}(t) = \displaystyle\bigoplus_{s\in \Sigma^*} A(s) \otimes T(s, t)$.
-\end{proposition}
-%
-\begin{proof}
-Let $T = \< Q, \init_T, \bar{\wei}, \final_T >$,
-where $\bar{\wei}$ contains $\wei_{10}$, $\wei_{01}$, and $\wei_{11}$,
-from $Q \times Q$ into respectively
-$\Phi_{\Sigma}$, $\Phi_{\Delta}$, and $\Phi_{\Sigma, \Delta}$,
-and let $A = \< P, \init_A, \wei_1, \final_A >$
-with $\wei_1: Q \times Q \to \Phi_{\Sigma}$.
-%
-\noindent
-The state set of $B_{A, T}$ will be $Q' = P \times Q$.
-The entering, leaving and transition functions of $B_{A, T}$ will
-simulate synchronized computations of $A$ and $T$,
-while reading an output word of $\Delta^*$.
-%
-Its state entering functions is defined
-for all $p \in P$, $q \in Q$
-by $\init'(p, q) = \init_A(p) \otimes \init_T(q)$.
-The transition function $\wei'_1$ will roughly perform
-a synchronized product of transitions defined by $\wei_1$,
-$\wei_{01}$ ($T$ reading in output word and not an input word)
-and $\wei_{11}$ ($T$ reading both an input word and an output word).
-%
-Moreover, $\wei'_1$ also needs to simulate transitions
-defined by $\wei_{10}$: $T$ reading in input word and not an output word.
-Since $B_{A, T}$  will read only in the output word, such a transition corresponds
-to an $\varepsilon$-transition of $\SWA$,
-but $\SWA$ have been defined without $\varepsilon$-transitions.
-Therefore, in order to take care of this case, we perform an on-the-fly
-suppression of $\varepsilon$-transition in the $\SWA$ in construction,
-following the algorithm of~\cite{LombardySakarovitch12ciaa}.
-%
-% Initialize state leaving functions i
-%and $\final'(p, q) = \final_A(p) \otimes \final_T(q)$.
-
-%Every transition of $B_{A, T}$ will
-%simulate a sequence of transitions of $T$ performing the following steps:
-%advance in the input word while staying immobile in the output word,
-%and then make one step in the output word (and advance in the input word or not).
-
-\noindent
-Initially, for all $p_1, p_2 \in P$, and $q_1, q_2 \in Q$, let
-\[
-\wei'_1\bigl( \< p_1, q_1>, \< p_2, q_2>\bigr) =
-\wei_1(p_1, p_2) \otimes
-\bigl[
-\wei_{01}(q_1, q_2)
-\oplus
-\displaystyle\bigoplus_{\Sigma}
-\wei_{11}(q_1, q_2)
-\bigr].
-\]
-
-\noindent
-Iterate the following for all $p_1\in P$ and $q_1, q_2 \in Q$:
-for all $p_2\in P$ and $q_3 \in Q$,
-\[
-\wei'_1\bigl( \< p_1, q_1>, \< p_2, q_3>\bigr) \opluseq
-\displaystyle\bigoplus_{\Sigma} \wei_{10}(q_1, q_2)
-\otimes
-\wei'_1\bigl( \< p_1, q_2>, \< p_2, q_3>\bigr)
-\]
-and
-\(
-\final'(p_1, q_1) \opluseq
-\displaystyle\bigoplus_{\Sigma} \wei_{10}(q_1, q_2)
-\otimes \final'(p_1, q_2)
-\)
-\florent{proof correctness}
-%\qed
-\end{proof}
-
-%The complexity of construction
-\noindent
-The construction time and size for $B_{A, T}$ are $O(\| T \|^3 . \| A \|^2)$,
-where the sizes $\| T \|$ and $\| A \|$ are their number of states.
-%of~$T$ is its number of states $|Q|$.
-\florent{revise with nb of tr. and states}
-
-\begin{corollary} \label{cor:epsilon}
-Given a \SWT $T$ over $\Sigma$, $\Delta$,
-$\Semiring$ commutative, bounded and complete,
-and $\bar\Phi$ effective,
-and $s \in \Sigma^+$,
-there exists an effectively constructible \SWA
-$B_{s, T}$ over $\Delta$, $\Semiring$ and $\bar\Phi$,
-such that for all $t \in \Delta^*$, $B_{s, T}(t) = T(s, t)$.
-\end{corollary}
-
-%\noindent
-%The construction time and size for $B_{T, s}$ are $O(\| T \|^3 . | s |^2)$,
-%where the size $\| T \|$ of~$T$ is its number of states $|Q|$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SW-VPA
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{SW Visibly Pushdown Automata}
-\label{section:SWVPA}\label{sec:SWVPA}
-The model presented in this section generalizes symbolic \VPA
-(\SVPA~\cite{dAntonyAlur14SVPDA},
- generalizing themselves \VPA~\cite{AlurMadhusudan09nested} to infinite alphabets)
-from Boolean semirings to arbitrary semiring weight domains.
-It will compute on nested words over infinite alphabets,
-associating to every such word a weight value.
-Nested words are able to describe structures of labeled trees,
-and in the context of parsing, they will be useful to
-represent AST. % parse trees.
-\florent{see \S 5 and App.A}
-
-%\subsection{Definition}
-\noindent
-\label{sec:SWVPA-def}
-Let $\Delta$ be a countable alphabet
-%finite (large) or infinite,
-that we assume partitioned into three
-subsets~$\Deltai$, $\Deltac$, $\Deltar$,
-whose elements are respectively called
-\emph{internal}, \emph{call} and \emph{return} symbols.
-% names are  coined by application to functional program verification
-% \begin{itemize}
-% \item a set $\Deltai$ of \emph{internal symbols} denoted $a$,
-% \item a set $\Deltac$ of \emph{call symbols} denoted $\call{a}$,
-% \item a set $\Deltar$ of \emph{return symbols} denoted $\return{a}$.
-% \end{itemize}
-Let~$\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative and complete semiring and let
-$\bar\Phi = \< \Phii, \Phic, \Phir, \Phici, \Phicc, \Phicr>$
-be a label theory over $\Semiring$
-%In order to simplify notations, %and following the definition of Section~\ref{section:transducer},
-%we shall write respectively
-where $\Phii$, $\Phic$, $\Phir$ and~$\Phicx$ (with $\mathsf{x} \in \{ \mathsf{i}, \mathsf{c}, \mathsf{r}\}$)
-stand respectively
-for~$\Phi_\Deltai$, $\Phi_\Deltac$, $\Phi_\Deltar$ and~$\Phi_{\Deltac, \Delta_\mathsf{x}}$.
-%
-%Moreover, we extend this theory with a set $\Phii$
-%of unary functions in $\Deltai \to \Semiring$,
-%closed under $\oplus$ and $\otimes$.
-
-\begin{definition}
-A \emph{Symbolic Weighted Visibly Pushdown Automata} (\SWVPA)
-over  $\Delta = \Deltai \uplus \Deltac \uplus \Deltar$, $\Semiring$ and $\bar\Phi$
-is a tuple $A = \< Q, P, \init, \bar\wei, \final >$,
-where $Q$ is a finite set of states,
-$P$ is a finite set of stack symbols,
-$\mathsf{in} : Q \to \Semiring$
-(respectively $\mathsf{out} : Q \to \Semiring$)
-are functions defining the weight for entering
-(respectively leaving) a state,
-and $\bar\wei$ is a sextuplet composed of the transition functions :
-$\weii : Q \times P \times Q \to \Phici$,
-$\weiei : Q \times Q \to \Phii$,
-$\weic : Q \times P \times Q \times P \to \Phicc$,
-$\weiec : Q \times P \times Q \to \Phic$,
-$\weir : Q \times P \times Q \to \Phicr$,
-$\weier : Q \times Q \to \Phir$.
-%and
-%$\weiex : Q \times Q \to \Phix$
-%with $\mathsf{x} \in \{ \mathsf{i}, \mathsf{c}, \mathsf{r}\}$.
-\end{definition}
-%
-Similarly as in Section~\ref{section:transducer},
-we extend the above transition functions as follows
-for all $q, q' \in Q$, $p \in P$,
-$a \in \Deltai$,
-$\call{c} \in \Deltac$,
-$\return{r} \in \Deltar$,
-overloading their names: % for simplicity:
-\[
-\begin{array}{lll}
-\weii: Q \times [\Deltac \times P] \times \Deltai \times Q \to \Semiring &
-\weii(q, c, p, a, q') = \eta_\mathsf{ci}(c, a) &
-\mathrm{where~} \eta_\mathsf{ci} = \weii(q, p, q'),\\
-%
-\weiei: Q \times \Deltai \times Q \to \Semiring &
-\weiei(q, a, q') = \phi_\mathsf{i}(a) &
-\mathrm{where~} \phi_\mathsf{i} = \weiei(q, q').\\
-%
-\weic: Q \times [\Deltac \times P] \times  [\Deltac \times P] \times Q \to \Semiring &
-\weic(q, c, p, \call{c'}, p', q') = \eta_\mathsf{cc}(c, \call{c'}) &
-\mathrm{where~} \eta_\mathsf{cc} = \weic(q, p, p', q'),\\
-%
-\weiec: Q \times [\Deltac \times P] \times Q \to \Semiring &
-\weiec(q, {\call{c}}, p, q') = \phi_\mathsf{c}({\call{c}}) &
-\mathrm{where~} \phi_\mathsf{c} = \weiec(q, p, q').\\
-%
-\weir: Q \times [\Deltac \times P] \times \Deltar \times Q \to \Semiring &
-\weir(q, {\call{c}},  p, {\return{r}}, q') = \eta_\mathsf{cr}({\call{c}},  {\return{r}}) &
-\mathrm{where~} \eta_\mathsf{cr} = \weir(q, p, q'),\\
-%
-\weier: Q \times \Deltar \times Q \to \Semiring &
-\weier(q, {\return{r}}, q') = \phi_\mathsf{r}({\return{r}}) &
-\mathrm{where~} \phi_\mathsf{r} = \weier(q, q').\\
-\end{array}
-\]
-
-\noindent
-The intuition is the following for the above transitions.
-$\weiei$, $\weiec$, and $\weier$ describe the cases where the stack is empty.
-\lydia{moved this to the beginning}
-%
-\noindent
-$\weii$ and $\weiei$ both read an input internal symbol $a$ and change state from $q$ to $q'$,
-without changing the stack.
-Moreover, $\weii$ reads a pair made of
-${\call{c}} \in \Deltac$ and $p \in P$ on the top of the stack
-($c$ is compared to $a$ by the weight function $\eta_\mathsf{ci} \in \Phici$).
-%
-\noindent
-$\weic$ and $\weiec$ read the input call symbol $\call{c}'$,
-push it to the stack along with $p'$, and change state from $q$ to to $q'$.
-Moreover, $\weic$ reads ${\call{c}}$ and $p$ at the top of the stack
-($c$ is compared to $c'$).
-%and $\weiec$ applies iff the stack is empty.
-%
-\noindent
-$\weir$ and $\weier$ read the input return symbol $\return{r}$, and change state from $q$ to to $q'$.
-Moreover, $\weir$ reads and pop from stack a pair made of $\call{c}$ and $p$,
-($\call{c}$ is compared to $\return{r}$).
-%and $\weier$ applies iff the stack is empty.
-%In this case, the weight function $\phi_\mathsf{r}$
-%computes a value of matching between the call and return symbols $c$ and $r$.
-%This value might be set to $\zero$ in order to express that the symbols do not match.
-
-Formally, the transitions of the automaton~$A$ are defined
-in term of %a weight value computed by
-an intermediate function $\weight_A$, like in Section~\ref{sec:SWT}.
-%In the case of a pushdown automaton,
-A configuration, denoted by $q[\gamma]$,
-is here composed of a state $q \in Q$
-and a stack content $\gamma \in \Gamma^*$,
-where $\Gamma = \Deltac \times P$.
-Hence, $\weight_A$ is a function from
-$[Q \times \Gamma^*] \times \Delta^* \times [Q \times \Gamma^*]$ into~$\Semiring$.
-The empty stack is denoted by $\bot$, and the upmost symbol is the last pushed content.
-The following functions illustrate each of the possible cases, being : reading $a \in \Deltai$,
-\lydia{intro to func}
-or $\call{c} \in \Deltac$, or $\return{r} \in \Deltar$ for each possible state of the stack (empty or not),
-to add to $u \in {\Delta}^*$.
-\lydia{introduced the 6 cases}
-\florent{notation $cp$ for $\< c, p>$ ?}
-
-\begin{align}
-%\weight_A\bigl(\config{q}{\bot}, \varepsilon, \config{q}{\bot}) & = \one\label{eq:SWVPA-weight}\\
-%\weight_A\bigl(\config{q}{\bot}, \varepsilon, \config{q'}{\gamma}) & = \zero
-%\mathrm{~if~} q \neq q'\nonumber\\
-%
-\weight_A\bigl(\config{q}{\bot}, \varepsilon, \config{q'}{\bot}) & = \one
-\mathrm{~if~} q = q' \mathrm{~and~} \zero \mathrm{~otherwise}\label{eq:SWVPA-weight}\\
-\weight_A\bigl(\configup{q}{\<{\call{c}}, p>\stackup\gamma}, a\, u,
-               \config{q'}{\gamma'}\bigr) & =
- {\displaystyle\bigoplus_{q'' \in Q}} \weii(q, c, p, a, q'')
-  \otimes \weight_A\bigl(\configup{q''}{\<{\call{c}}, p> \stackup \gamma}, u,
-                         \config{q'}{\gamma'}\bigr)\nonumber\\
-%
-\weight_A\bigl(\config{q}{\bot}, a\, u,
-               \config{q'}{\gamma'}\bigr) & =
-  {\displaystyle\bigoplus_{q'' \in Q}} \weiei(q, a, q'')
-   \otimes \weight_A\bigl(\config{q''}{\bot}, u, \config{q'}{\gamma'}\bigr)\nonumber\\
-%
-\weight_A\bigl(\configup{q}{\<{\call{c}}, p>\stackup\gamma}, {\call{c}'}\, u,
-               \config{q'}{\gamma'}\bigr) & =
- {\displaystyle\bigoplus_{\begin{array}{c}
-                          \scriptstyle q'' \in Q\\[-1ex]
-                          \scriptstyle p' \in P
-                          \end{array}}}
- \weic\bigl(q, {\call{c}}, p, {\call{c}'}, p', q''\bigr)
- \otimes \weight_A\bigl(\configup{q''}{\<{\call{c}'}, p'>\stackup \<{\call{c}}, p>\stackup \gamma}, u,
-                        \config{q'}{\gamma'}\bigr)\nonumber\\[1mm]
-%
-\weight_A\bigl(\config{q}{\bot}, {\call{c}}\, u,
-               \config{q'}{\gamma'}\bigr) & =
- {\displaystyle\bigoplus_{\begin{array}{c}
-                          \scriptstyle q'' \in Q\\[-1ex]
-                          \scriptstyle p \in P
-                          \end{array}}}
-  \weiec(q, {\call{c}}, p, q'')
-  \otimes \weight_A\bigl(\config{q''}{\<{\call{c}}, p>}, u,
-                         \config{q'}{\gamma'}\bigr)\nonumber\\
-%
-\weight_A\bigl(\configup{q}{\<{\call{c}}, p>\stackup\gamma}, {\return{r}}\, u,
-               \config{q'}{\gamma'}\bigr) & =
- {\displaystyle\bigoplus_{q'' \in Q}}
-  \weir\bigl(q, {\call{c}}, p, {\return{r}}, q''\bigr)
-  \otimes \weight_A\bigl(\config{q''}{\gamma}, u,
-                         \config{q'}{\gamma'}\bigr)\nonumber\\
-%
-\weight_A\bigl(\config{q}{\bot}, {\return{r}}\, u,
-               \config{q'}{\gamma'}\bigr) & =
- {\displaystyle\bigoplus_{q'' \in Q}} \weier(q, {\return{r}}, q'')
-  \otimes \weight_A\bigl(\config{q''}{\bot}, u,
-                         \config{q'}{\gamma'}\bigr)\nonumber
-\end{align}
-\lydia{c p to <c, p>}
-%
-% and ${\call{c}}\, p\stacksep \gamma$
-%denotes a stack where the pair made of ${\call{c}} \in \Deltac$ and $p \in P$ is the top symbol
-%and $\gamma$ is the rest of stack.
-
-\noindent
-The weight associated by $A$ to $t \in \Delta^*$
-is defined according to empty stack semantics:
-%
-\begin{equation}
-A(t)  =
-{\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
-\mathsf{in}(q) \mathop{\otimes}
-\weight_A\bigl(\config{q}{\bot}, t, \config{q'}{\bot}\bigr)
-\mathop{\otimes} \mathsf{out}(q').
-\label{eq:SWVPA-value}
-\end{equation}
-
-\begin{example}
-structured words with timed symbols...
-\florent{todo example VPA}
-intro language of music notation? (markup = time division, leaves = events etc)
-\end{example}
-
-\noindent
-Every $\SWA$ $A = \< Q, \init, {\wei_1}, \final >$,
-over $\Sigma$, $\Semiring$ and $\bar\Phi$
-is a particular case of $\SWVPA$
-$\< Q, \emptyset, \init, \bar\wei, \final >$
-over $\Delta$, $\Semiring$ and $\bar\Phi$
-with $\Deltai = \Sigma$ and $\Deltac = \Deltar = \emptyset$,
-and computing with an always empty stack:
-$\weiei = \wei_1$ and all the other functions
-of~$\bar\wei$ are the constant~$\zero$.
-
-%\subsection{Properties}
-\medskip\noindent
-Similarly to \VPA~\cite{AlurMadhusudan09nested}
-and \SVPA~\cite{dAntonyAlur14SVPDA},
-the class of \SWVPA is closed under the binary operators of the underlying semiring.
-%
-\begin{proposition}\label{prop:SWVPA-product}
-Let $A_1$ and $A_2$ be two \SWVPA
-over the same $\Delta$, $\Semiring$ and $\bar\Phi$.
-There exists two effectively constructible $\SWVPA$
-$A_1 \oplus A_2$ and $A_1 \otimes A_2$,
-such that for all $s \in \Delta^*$,
-$(A_1 \oplus A_2)(s) = A_1(s) \oplus A_2(s)$ and
-$(A_1 \otimes A_2)(s) = A_1(s) \otimes A_2(s)$.
-\end{proposition}
-%
-\begin{proof}
-The construction is essentially the same
-as in the case of the Boolean semiring~\cite{dAntonyAlur14SVPDA}.
-\florent{complete proof}
-\end{proof}
-
-%\subsection{Best Search}
-\label{sec:best}\label{sec:search}
-\noindent
-We shall now present a procedure for searching, for a~$\SWVPA$ $A$,
-a word %$w \in \Delta^*$
-of minimal weight for~$A$,
-as stated in the following proposition. %\wrt~$\leq_\oplus$,
-%
-\begin{proposition}\label{th:best-search}
-For a \SWVPA $A$
-over $\Delta$,
-$\Semiring$ commutative, bounded, total and complete, %semiring
-and $\bar\Phi$ effective, %and $k$-convex, % label theory,
-one can construct in PTIME a word $t \in \Delta^*$
-such that $A(t)$ is minimal \wrt the natural ordering for $\Semiring$.
-\end{proposition}
-%\florent{total?}
-%
-Let $A = \< Q, P, \init, \bar\wei, \final >$.
-%
-We propose a Dijkstra algorithm computing,
-for every $q, q' \in Q$,
-the minimum, \wrt $\leq_\oplus$, of the function
-$\beta_{q, q'} : t \mapsto \weight_A(\config{q}{\bot}, t, \config{q'}{\bot})$.
-Let us denote by $b_\bot(q, q')$ this minimum.
-By definition of $\leq_\oplus$, and since $\Semiring$ is total,
-it holds that:
-%
-\begin{equation}\label{eq:bbot}
-  b_\bot(q, q') = \bigoplus_{t\in \Delta^*}
-  \textstyle
-  \weight_A\bigl(\config{q}{\bot}, t, \config{q'}{\bot}\bigr).
-\end{equation}
-Since $\Semiring$ is complete, the infinite sum in~\eqref{eq:bbot} is well defined.
-%
-Following~\eqref{eq:SWVPA-value}, and the associativity and commutativity
-and distributivity for~$\otimes$ and~$\oplus$, the minimum of~$A(t)$ is:
-\begin{equation}\label{eq:min}
-{\displaystyle\bigoplus_{t\in \Delta^*}} A(t)
-= 
-{\displaystyle\bigoplus_{t\in \Delta^*}}
-{\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
-\mathsf{in}(q) \mathop{\otimes}
-\beta_{q, q'}(t)
-\mathop{\otimes} \mathsf{out}(q')
-=
-{\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
-\mathsf{in}(q) \mathop{\otimes}
-b_\bot(q, q')
-\mathop{\otimes} \mathsf{out}(q')
-\end{equation}
-
-\noindent
-In order to compute the above function $b_\bot : Q \times Q \to \Semiring$, 
-we shall consider an auxiliary function $b_\top :  Q \times P \times Q \to \Phic$.
-%
-Intuitively, when $b_\top(q, p, q') = \phi \in \Phic$, 
-for all $c \in \Deltac$,  $\phi(c)$ is 
-the minimum weight of a computation of~$A$
-starting in state $q$ with a non-empty stack 
-$\gamma' = \<c, p> \, \gamma \in \Gamma^+$ 
-and ending in state $q'$ with the same stack $\gamma'$,
-such that moreover, 
-the computation does not pop the pair $\<c, p>$ at the top of $\gamma'$
-(\ie $\gamma'$ is left untouched during the computation).
-However, the computation can read $\<c, p>$ at the top of $\gamma'$,
-and can also push another pair $\< c', p'> \in \Gamma$ on $\gamma'$,
-following the third case of 
-in the definition \eqref{eq:SWVPA-weight} of $\weight_A$ (call symbol).
-The pair $\< c', p'>$ can be pop later, during the computation from $q$ to $q'$,  
-following the fifth case of \eqref{eq:SWVPA-weight} (return symbol). 
-%However, it cannot apply one of the two last cases (return symbol and empty stack)
-%when the current stack is $\gamma$.
-%pop symbols in $\gamma$.
-%
-Formally, in order to define $b_\top$, we consider 
-a fresh stack symbol $\top \notin \Gamma$,   %which does not belong to $\Gamma$,
-representing the above untouched stack, and let:
-%
-\begin{equation}\label{eq:btop}
-  b_\top(q, p, q') : c \mapsto \bigoplus_{s\in \Delta^*}
-  \textstyle
-  \weight_A\bigl(\configup{q}{\< c, p> \stackup \top }, s, \configup{q'}{\<c, p> \stackup \top}\bigr)
-\quad  
-\mathrm{~for~all~} c \in \Deltac  
-\end{equation}
-%
-By definition of $\weight_A$ in \eqref{eq:SWVPA-weight}, 
-using the symbol $\top$ for the part of the stack below $\<c, p>$
-(\ie the substack $\gamma$ in the above $\gamma' = \<c, p> \, \gamma$)
-ensures that this part $\gamma'$ is not touched during the computation.
-% there cannot be a pop of $\<c, p>$ followed by a push of $\<c, p>$
-% because it is not possible to push over $\top$ (push read the top of stack).
-%
-This ensures in particular that the subword 
-read during the computation is well parenthesized
-(every symbol in $\Deltac$ has a matching symbol in $\Deltar$).
-
-% Note that having a stack reduced to such a symbol makes impossible the application of the 
-% two last cases in the definition of $\weight_A$ (return symbol and empty stack). 
-% However, it is possible to apply the two first cases 
-% (internal symbol or call symbol, with a push on the top of $\top$).
-
-%The term $\config{q}{\bot}, s, \config{q'}{\bot}$
-%of this sum is the central expression in
-%the definition \eqref{eq:weightA} of $A(s_0)$, for the minimum $s_0$
-%of the function $\weight_A$.
-
-\begin{algorithm}
-\SetKw{Extract}{extract}
-\SetKw{Update}{update}
-
-\textbf{initially} let $\Q = (Q \times Q) \cup (Q \times P \times Q)$, 
-and let $d_\bot(q_1, q_2) = d_\top(q_1, p, q_2) = \one$
-if $q_1 = q_2$ and $d_\bot(q_1, q_2) = d_\top(q_1, p, q_2) = \zero$ otherwise$\;$
-
-\smallskip\noindent
-\While{$\Q \neq \emptyset$}{
-
-%\noindent\quad
-\Extract $\< q_1, q_2>$ or $\< q_1, p, q_2>$ from $\Q$
-%
-%\noindent\quad
-such that $d_\bot(q_1, q_2)$, resp. $\bigoplus_{c\in \Deltac} d_\top(q_1, p, q_2)(c)$, is minimal in $\Semiring$ \wrt $\leq_\oplus$$\;$
-
-%\noindent\quad
-%\textbf{for all} $q_0, q_3 \in Q$
-\Update $d_\bot$ with $\< q_1, q_2>$ or $d_\top$ with $\< q_1, p, q_2>$ (Figure~\ref{fig:best-update}).$\;$
-}
-\caption{Best search for \SWVPA}
-\label{algo:Dijkstra}
-\end{algorithm}
-
-Algorithm~\ref{algo:Dijkstra}
-constructs iteratively two markings
-$d_\bot : Q \times Q \to \Semiring$ and
-$d_\top : Q \times P \times Q \to \Phic$,
-%of the triplets $\<q, \sigma, q'>$
-%of states of $A$ by weight values in $\Semiring$,
-that converges eventually to $b_\top$ and $b_\bot$.
-%It uses for that purpose a priority queue $P$ containing triplets of
-%$Q \times \{ \bot, \top \} \times Q$.
-%
-\noindent
-\florent{explication Fig.~\ref{fig:best-update} suivant cas de \eqref{eq:SWVPA-weight}}
-The infinite sums in the updates of $d$ in Algorithm~\ref{algo:Dijkstra},
-Figure~\ref{fig:best-update}
-are well defined
-since~$\Semiring$ is complete.
-
-\begin{figure}
-For all $q_0, q_3 \in Q$, %$p \in P$,
-\[
-\begin{array}{lcl}
-%\multicolumn{3}{l}{\mathrm{For~all~} q_0, q_3 \in Q, p \in P,}\\
-d_\top(q_1, p, q_3) & \opluseq &
-  d_\top(q_1, p, q_2) \otimes
-  \displaystyle\bigoplus_{\Deltai} \weii(q_2, p, q_3)\\
-%\multicolumn{3}{l}{\quad
-%\mathrm{where~} \weii(q_2, p, q_3)_a \in \Phic
-%\mathrm{~is~the~partial~application~}
-%x_c \mapsto \weii(q_2, x_c, p, a, q_3)}\\
-%
-d_\bot(q_1, p, q_3) & \opluseq &
-  d_\bot(q_1, q_2) \otimes
-  \displaystyle\bigoplus_{\Deltai} \weiei(q_2, q_3)\\
-%
-d_\top(q_0, p, q_3) & \opluseq &
-  {\displaystyle\bigoplus_{\Deltac}}^2
-  \bigl[ \bigl( \weic(q_0, p, p', q_1) \otimes_2
-  d_\top(q_1, p', q_2) \bigr) \otimes_2
-  {\displaystyle\bigoplus_{\Deltar}} \weir(q_2, p', q_3) \bigr]\\
-%\multicolumn{3}{l}{\quad
-%\mathrm{where~} \weic(q_0, p, p', q_1)_{c'} \in \Phic
-%\mathrm{~is~the~partial~application~}
-%x_c \mapsto \weic(q_0, p, x_c, c', p', q_1)}\\[2pt]
-%
-d_\bot(q_0, q_3) & \opluseq &
-  {\displaystyle\bigoplus_{\Deltac}}
-  \bigl(
-  \weiec(q_0, p, q_1) \otimes
-   d_\top(q_1, p, q_2) \otimes
-  \displaystyle\bigoplus_{\Deltar} \weir(q_2, p, q_3)\bigr)\\
-%
-d_\bot(q_1, q_3) & \opluseq &
-  d_\bot(q_1, q_2) \otimes
-  \displaystyle\bigoplus_{\Deltar} \weier(q_2, q_3)\\
-%
-d_\top(q_1, p, q_3) & \opluseq &
-  d_\top(q_1, p, q_2) \otimes d_\top(q_2, p, q_3),
-  \mathrm{if~} \< q_2, \top, q_3> \notin P \\
-%
-d_\bot(q_1, q_3) & \opluseq &
-  d_\bot(q_1, q_2) \otimes d_\bot(q_2, q_3), \mathrm{if~} \< q_2, \bot, q_3> \notin P \\
-\end{array}
-\]
-%$\weii(q_2, p, q_3)_a \in \Phic$, for $a \in \Deltai$,
-%is the partial application
-%$x_c \mapsto \weii(q_2, x_c, p, a, q_3)$.
-%
-%$\weic(q_0, p, p', q_1)_{c'} \in \Phic$, for $c' \in \Deltac$,
-%is the partial application
-%$x_c \mapsto \weic(q_0, p, x_c, c', p', q_1)$.
-\caption{Update $d_\bot$ with $\<q_1, q_2>$ or $d_\top$ with $\< q_1, p, q_2>$.}
-\label{fig:best-update}
-\end{figure}
-
-\florent{complete **}
-** effectively computable by hypothesis that the label theory is effective**
-termination: The algorithm performs $2.|Q|^2$ iterations until $P$ is empty,
-and each iteration has a time complexity $O(|Q|^2 . |P|)$.
-\florent{detail with nb tr. and states}
-That gives a time complexity $O(|Q|^4 . |P|)$.
-It can be reduced by implementing $P$ as a priority queue,
-prioritized by the value returned by $d$.
-%$|Q|^3.\log(|Q|^2)$
-%
-At termination, $d_\bot = b_\bot$ and $d_\top = b_\top$
-(see Appendix~\ref{sec:bestsearch}).
-With~\eqref{eq:min}, this ensures
-the correctness of Algorithm~\ref{algo:Dijkstra}.
-
-\noindent
-In order to obtain effectively a witness
-(word of $\Delta^*$ with a computation of $A$ of minimal weight),
-we require the additional property of convexity of weight functions.***
-\florent{completer}
-
-\begin{proposition}\label{th:best-search}
-For a \SWVPA $A$
-over $\Delta$,
-$\Semiring$ commutative, bounded, total and complete, %semiring
-and $\bar\Phi$ effective, %and $k$-convex, % label theory,
-one can construct in PTIME a word $t \in \Delta^*$
-such that $A(t)$ is minimal \wrt the natural ordering for $\Semiring$.
-\end{proposition}
-
-\section{Symbolic Weighted Parsing}
-\label{sec:parsing}
-%% see parse-tree.tex for a definition with parse tree (instead of nested words)
-
-Let us now apply the models and results of the previous sections %in order to define
-to the problem of parsing over an infinite alphabet. %appropriate
-%
-%\subsection{Definition}
-%
-Let~$\Sigma$
-and~$\Delta = \Deltai \uplus \Deltac \uplus \Deltar$
-be countable input and output alphabets,
-let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a
-commutative, bounded, and complete \florent{total?} semiring
-and let $\bar\Phi$ be an effective label theory over $\Semiring$,
-containing $\Phi_\Sigma$, $\Phi_{\Sigma, \Deltai}$, as well as
-$\Phii$, $\Phic$, $\Phir$, $\Phicr$
-(following the notations of Section~\ref{sec:SWVPA-def}).
-%
-\noindent
-We assume given the following input:
-\begin{description}
-\item[--] a \SWT $T$ over $\Sigma$, $\Deltai$, $\Semiring$, and $\bar\Phi$,
-defining a measure %between words
-$T: \Sigma^* \times \Deltai^* \to \Semiring$,
-
-\item[--] a \SWVPA $A$ over $\Delta$, $\Semiring$, and $\bar\Phi$,
-defining a measure $A: \Delta^* \to \Semiring$,
-%a \SWVPA $A$ over $\Delta$, and $\Semiring$, defining a series of nested words
-%      $A : \Delta^* \to \Semiring$,
-\item[--] an input word $s \in \Sigma^*$.
-\end{description}
-%
-For all $u \in \Sigma^*$ and $t \in \Delta^*$, let
-\(d(u, t) = T\bigl(u, t|_\Deltai \bigr)\),
-where $t|_\Deltai \in \Deltai^*$ is the projection of $t$ onto~$\Deltai$,
-obtained from $t$ by removing all symbols in $\Delta \setminus \Deltai$.
-%\end{equation}
-%
-\noindent
-\emph{Symbolic weighted parsing} is the problem,
-given the above input,
-to find $t \in \Delta^*$ %nested word $t \in \Delta$
-minimizing \( d(s, t) \otimes A(t)\)
-\wrt $\leq_\oplus$,
-\ie s.t. %such that: %Hence, it is the problem of finding
-%
-\begin{equation}\label{eq:distance-lang}
-d(s, t) \otimes A(t) = \displaystyle\bigoplus_{t' \in \Delta^*} d(s, t') \otimes A(t')
-\end{equation}
-%
-Following the terminology of~\cite{Mohri03EDWA},
-$\SW$-parsing is the problem of computing
-the distance~\eqref{eq:distance-lang} between the input~$s$ and the output weighted language of~$A$,
-and returning a witness $t$.
-%
-%\subsection{Computation}
-%
-\begin{proposition}
-The problem of Symbolic Weighted  parsing
-can be solved in PTIME in the size of the input \SWT $T$, \SWVPA $A$
-and input word $s$,
-and the computation time of the functions and operators of the label theory.
-\end{proposition}
-%
-\begin{proof} (sketch)
-We follow a \emph{Bar-Hillel} construction, for parsing by intersection.
-%
-\noindent
-Let us first extend the \SWT $T$ over $\Sigma$, $\Deltai$
-into a \SWT $T'$ over $\Sigma$ and $\Delta$
-(and the same semiring and label theory $\Semiring$ and $\bar\Phi$),
-such that for all $u \in \Sigma^*$, and $t \in {\Delta}^*$,
-$T'(u, u) = T(u, t|_{\Deltai})$.
-%
-The transducer $T'$ simply skips every symbol
-$b \in {\Delta} \setminus \Deltai$,
-by the addition to $T$,
-of new transitions of the form $\wei_{01}(q, \varepsilon, b, q')$.
-%
-\noindent
-Then, using Corolary~\ref{cor:epsilon},
-we construct from the input word $s \in \Sigma^*$ and $T'$
-a \SWA $B_{s, T'}$,
-such that for all $t \in \Delta^*$, $B_{s, T'}(t) = d(s, t)$.
-%
-%This automaton is such that for all $t \in \Delta^*$,
-%\[
-%   A_{T', s}\bigl(\lin(t)\bigr)
-% = T'\bigl(s, \lin(t)\bigr)
-% = T'\bigl(s, \lin(t)|_{\Deltai}\bigr)
-% = d(s, t).
-%\]
-%
-\noindent
-Next, %we convert the input \SWTA $A$ over $\Delta$
-%into a \SWVPA $A'$ over $\hat\Delta$, using Proposition~\ref{lem:SWTA}, and
-we compute the \SWVPA $B_{s, T'} \otimes A$,
-using Proposition~\ref{prop:SWVPA-product}.
-%
-\noindent
-It remains to compute a best nested-word $t \in {\Delta}^*$
-using the best-search procedure of Proposition~\ref{th:best-search}.
-%and convert it into a best tree in $\T_\Delta$ in order to solve SW parsing
-%for~$T$, $A$ and~$s$.
-%\qed
-\end{proof}
-The $\SW$-parsing generalizes
-the problem of searching the best derivation (AST) of a weighted CF-grammar $G$
-that yields a given input word $w$.
-The latter problem, sometimes called~\emph{weighted parsing},
-(see \eg~\cite{Goodman99SemiringParsing}
- and~\cite{MorbitzVogler19weighted-parsing} for general weighted parsing frameworks)
-corresponds to $\SW$-parsing in the case of finite alphabets,
-a transducer $T$ computing the identity and some \SWVPA~$A$
-obtained from $G$. %the weighted CF grammar.
-%
-%Indeed, it corresponds to the case where $T$
-%accepts only the pairs $\<s, t>$ such that
-%$s$ is the projection of $t$ on $\Deltai$.
-%This can be done with a single state $q$ and
-%with transition rules of the form:
-%\begin{description}
-%\item[] $\wei(q, \varepsilon, a, q) = \one$ for all $a \in \Deltac \cup \Deltar$,
-%\item[] $\wei(q, a, a, q) = \one$ for all $a \in \Deltai$,
-%\item[] $\wei(q, a, b, q) = \zero$ for all $a, b \in \Deltai$, $a \neq b$.
-%\end{description}
-%
-Indeed, the \emph{depth-first} traversal of an AST $\tau$
-yields a well-parenthesised word $\lin(\tau)$ over an alphabet
-$\Delta = \Deltai \uplus \Deltac \uplus \Deltar$,
-assuming \eg that $\Deltai$ contains the symbols labelling the leaves of $\tau$ (symbols of rank $0$),
-and $\Deltac$ and $\Deltar$ contain respectively one left and right parenthesis
-$\ccall{b}$ and $\creturn{b}$ for each symbol $b$ labelling inner nodes of $\tau$ (symbols of rank $>0$).
-%
-%With this representation, the projection $\lin(t)|_\Deltai$ is the
-%sequence of leaves of $\tau$. %, enumerated in a \emph{dfs}-traversal.
-%
-We show in Appendix~\ref{sec:trees} how to construct 
-%($\SW$) tree automaton~$A$ %(in particular the TA computing on the AST of a given CF-grammar)
-a $\SWVPA$ $A$ such that $A(\lin(\tau))$ is the weight the AST $\tau$ of $G$.
-
-%That also holds for the set of ASTs of a weighted CF-grammar.
-%The purpose of the transducer $T$ is to measure a distance between input and output words,
-%\ie between the word $s$ and the sequence leaves of a derivation tree, in $\Deltai^*$.
-%
-%that generates (weighted) trees by replacement of a state symbol~$q_0$ (non-terminal),
-%by a tree $a(q_1,\ldots, q_k)$, where $k = \rank(a)$.
-%A replacement rule $q_0 \to a(q_1,\ldots, q_k)$,
-%of weight $\wei(q_0, a, q_1 \ldots q_k) \in \Semiring$
-%according to Definition~\ref{def:SWTA},
-%corresponds to the production rule $q_0 := a(q_1,\ldots, q_k)$ of a weighted CF grammar,
-%with set non-terminal symbols $Q$ and set of terminal symbols $\Delta_0$.
-%
-%This actually is a slight generalization of CFG since
-%each such production rule is labelled by  a symbol of $\Delta_{>0}$,
-%hence parse trees %derivation trees
-%are trees of $\T_\Delta)$.
-%
-%Another (more original) generalization is that the set of terminal symbols
-%$\Delta_0$ may be infinite.
-%
-%
-%The input language can also be expressed as a \SWTA, or,
-%as a particular case, as a weighted context-free grammar,
-%converted in turn into a \SWVPA following Lemma~\ref{lem:SWTA}.
-%
-
-%In practice,
-%the functions in transition might be represented by diagram such as
-%Algebraic Decision Diagrams~\cite{Bahar97ADD}
-
-%\paragraph{Application to Automated Music Transcription.}
-%\input{transcription}
-%...
-\florent{2 lines Application to Automated Music Transcription:
-implementation $\neq$ but same principle, on-the-fly automata construction
-during best search, for efficiency.}
-
-\section*{Conclusion}
-% summary
-We have introduced weighted language models (SW transducers and visibly pushdown automata)
-computing over infinite alphabets,
-and applied them to the problem of parsing
-with infinitely many possible input symbols (typically timed events).
-%handled with suitable language formalisms,
-%
-This approach extends conventional parsing and weighted parsing
-by computing a derivation tree modulo
-a generic distance between words,
-defined by a SW transducer given in input.
-This enables to consider finer word relationships than strict equality,
-%as in the conventional parsing approach,
-opening possibilities of quantitative analysis via this method.
-
-% discussion
-\noindent
-Ongoing and future work include
-\florent{TODO future work}
-
-\noindent
--- The study of other theoretical properties of SW models,
-such as the extension of the best search algorithm from $1$-best to $n$-best~\cite{Huang05kbest},
-and to $k$-\emph{closed} semirings~\cite{Mohri02semiring}
-(instead of \emph{bounded}, which corresponds to $0$-\emph{closed}).
-
-\noindent
--- ...there is room to improve the complexity bounds for the algorithms
-... modular approach with oracles ...
-
-\noindent
--- present here an offline algorithm for best search,
-semi-online implementation for AMT (bar-by-bar approach)
-with an on-the-fly automata construction.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% BIBLIO                                                                     %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\bibliographystyle{plain}
-%\bibliographystyle{abbrv}
-\bibliographystyle{plainurl}  % LIPICS
-%\bibliographystyle{splncs04} % LLNCS
-%\bibliographystyle{eptcs}    % EPTCS
-%\bibliography{generic}
-
-\bibliography{references}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% APPENDIX                                                                   %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\appendix
-
-\section{Nested-Words and Parse-Trees} \label{sec:trees}
-The hierarchical structure of nested-words, defined with the \emph{call} and \emph{return} markup symbols
+The hierarchical structure of nested-words, defined with the \emph{call} 
+and \emph{return} markup symbols
 suggest a correspondence with trees.
@@
 \end{proof}
-
-\section{Correctness of the Best-Search Algorithm} \label{sec:bestsearch}
-\input{app-bestsearch}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% TODO                                                                       %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\listoftodos
-
-\end{document}

### 2021-07-08 main.tex
appendices et commentaires

@@
 
+\florent{indep. counters Def. Prop. Ex...}
+
 \section{Introduction} \label{sec:intro}
@@
 overloading the function names,
-such that, for all $q, q' \in Q$, $a \in \Sigma$,  $b \in \Delta$,
+such that, for all $q, q' \in Q$, $a \in \Sigma$,  $b \in \Delta$:
 
@@
 $s$ and $t$, being respectively called \emph{input} and \emph{output} word.
+\florent{It compute asynchronously: advance in $s$ and not in $t$ ($\wei_{10}$), 
+or the opposite ($\wei_{01}$),
+or advance in both $s$ and $t$ ($\wei_{11}$).}
 % \noindent
@@
 such that for all $u \in \Sigma^*$, and $t \in {\Delta}^*$,
-$T'(u, u) = T(u, t|_{\Deltai})$.
+$T'(u, t) = T(u, t|_{\Deltai})$.
 %
@@
 
-%In practice,
-%the functions in transition might be represented by diagram such as
-%Algebraic Decision Diagrams~\cite{Bahar97ADD}
+\florent{In practice,
+the functions in transition might be represented by diagram such as
+Algebraic Decision Diagrams~\cite{Bahar97ADD}.}
 
@@
 \section{Nested-Words and Parse-Trees} \label{sec:trees}
-The hierarchical structure of nested-words, defined with the \emph{call} and \emph{return} markup symbols
-suggest a correspondence with trees.
-The lifting of this correspondence to languages, of tree automata and VPA,
-has been discussed in~\cite{AlurMadhusudan09nested},
-and~\cite{Caralp12VPAmult} for the weighted case.
-In this section, we describe a correspondence between the symbolic-weighted extensions
-of tree automata and VPA.
-
-Let $\Omega$ be a countable ranked alphabet, such that
-every symbol $a \in \Omega$ has a rank
-$\rank(a) \in [0..M]$ where $M$ is a fixed natural number.
-We denote by $\Omega_k$ the subset of all symbols $a$ of $\Omega$
-with $\rank(a) = k$, where $0 \leq k \leq M$,
-and $\Omega_{>0} = \Omega \setminus \Omega_0$.
-%
-\noindent
-The free $\Omega$-algebra of finite, ordered,
-$\Omega$-labeled trees is denoted by $\T_\Omega$.
-It is the smallest set such that  $\Omega_0 \subset \T_\Omega$
-and for all $1 \leq k \leq M$, all $a \in \Omega_k$,
-and all $t_1, \ldots, t_k \in \T_\Omega$, $a(t_1, \ldots, t_k) \in \T_\Omega$.
-%
-% tree = single node (leave) labeled with a symbol of $a \in \Omegai$
-% (such a tree is simply denoted by $a$)
-% or the composition, denoted by $b(t_1,\ldots, t_n$) of a node labeled with $b$
-% and $n$ subtrees $t_1$,\ldots, $t_n$.
-%
-Let us assume a commutative semiring~$\Semiring$
-and a label theory~$\bar{\Phi}$ over~$\Semiring$
-containing one set~$\Phi_{\Omega_k}$ for each $k \in [0..M]$.
-%
-\renewcommand{\call}[1]{\ensuremath \langle_{#1}}
-\renewcommand{\return}[1]{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
-
-% A \emph{regular tree grammar} over $\Omega$
-% is a triplet $G = \< N, q_\mathsf{i}, R>$ where
-% $N$ is a finite set of non-terminal symbols denoted $q$...,
-% $q_\mathsf{i} \in N$ is the starting non-terminal,
-% $R$ is a finite set of production rules of the form
-% $q_0 \to a(q_1\ldots q_k)$ where
-% $q_0, q_1, \ldots, q_k \in N$
-% $a \in \Omega_k$.
-% A tree $t \in \T_\Omega$ is in the language of $G$
-% if it can be generated from $q_\mathsf{i}$ by
-% non terminal replacement following the rules of $R$.
-%
-\begin{definition}  \label{def:SWTA}
-A \emph{symbolic-weighted tree automaton} (\SWTA)
-over $\Omega$, $\Semiring$, and~$\bar\Phi$
-is a triplet $A = \< Q, \init, \bar{\wei} >$ where
-$Q$ is a finite set of states,
-$\mathsf{in} : Q \to \Phi_\Omega$ is the starting weight function,
-and $\bar{\wei}$ is a tuplet of transition functions containing,
-for each $k \in [0..M]$,
-%$\wei_\varepsilon$ from $Q \times Q$ into $\Semiring$, and,
-the functions $\wei_{k}: Q \times Q^{k} \to \Phi_{\Omega_{>0},\Omega_k}$
-and $\weie[k]: Q \times Q^{k} \to \Phi_{\Omega_k}$.
-\end{definition}
-%
-%Like in Section~\ref{sec:SWAdef},
-We define %from $\bar{\wei}$
-a transition
-function~$\wei: Q \times (\Omega_{> 0} \cup \{ \varepsilon \}) \times \Omega \times \bigcup_{k=0}^{M} Q^k
-  \to \Semiring$
-by: %also called $\wei$ for simplicity,
-%such that, for all $q, q' \in Q$, $a \in \Sigma$, and $b \in \Delta$,
-\[
-\begin{array}{rcll}
-%\wei(q_0, \varepsilon, q_1) & = &  \wei_\varepsilon(q_0, q_1),\\ %\phi_\varepsilon\\
-\wei(q_0, a, b, q_1 \ldots q_k) & = & \eta(a, b) &
-\quad\mathrm{where~} \eta = \wei_{k}(q_0, q_1\ldots q_k)\\
-\wei(q_0, \varepsilon, b, q_1 \ldots q_k) & = & \phi(b) &
-\quad\mathrm{where~} \phi = \weie[k](q_0, q_1\ldots q_k).
-\end{array}
-\]
-%
-\noindent
-where $q_1\ldots q_k$ is $\varepsilon$ if $k = 0$.
-The first case deals with a strict subtree, with a parent node labeled by $a$,
-and the second case is for a root tree.
-
-\noindent
-Every \SWTA %of Definition~\ref{def:SWTA}
-defines a mapping
-from trees of $\T_\Omega$ into~$\Semiring$, %the weight values in~$\Semiring$,
-based on the following intermediate function
-$\weight_A: Q \times (\Omega \cup \{ \varepsilon \}) \times \T_\Omega \to \Semiring$
-\begin{equation}
-%\begin{array}{rccl}
-\weight_A(q_0, a, t) =  % & = &
-%   \displaystyle\bigoplus_{q_1 \in Q} &
-%   \wei(q, \varepsilon, q_1) \otimes \weight_A(q_1, t)\\
- \displaystyle\bigoplus_{q_1 \ldots q_k \in Q^k} % &
-              \wei(q_0, a, b, q_1 \ldots q_k )
-   \otimes \displaystyle\bigotimes_{i=1}^{k}
-           \weight_A(q_i, b, t_i)
-%\end{array}
-\end{equation}
-where $q_0 \in Q$, $a \in \Omega_{>0} \cup \{ \varepsilon \}$ and
-$t = b(t_1,\ldots, t_k) \in \T_\Omega$,
-$0 \leq k \leq M$.
-
-\medskip\noindent
-Finally, the weight associated by $A$ to  $t \in \T_\Omega$ is
-\begin{equation}
-A(t)  =
-\displaystyle\bigoplus_{q \in Q} \mathsf{in}(q) \mathop{\otimes} \weight_A(q, \varepsilon, t)
-\label{eq:weightTA}
-\end{equation}
-
-\noindent
-Intuitively, $\wei(q_0, a, b, q_1 \ldots q_k)$ can be seen as
-the weight of a production rule $q_0 \to b(q_1, \ldots, q_k)$
-of a regular tree grammar~\cite{tata},
-that replaces the non-terminal symbol $q_0$ by $b(q_1, \ldots, q_k)$,
-provided that the parent of $q_0$ is labeled by $a$
-(or $q_0$ is the root node if $a = \varepsilon$).
-%in a step of tree building.
-%
-%Such a grammar computes the weights of the derivation trees
-%of the Context-Free grammar obtained by forgetting the labeling symbols of $\Omega_{>0}$.
-The above production rule can also be seen as
-a rule of a weighted CF grammar, of the form
-$[a, b]\, q_0 := q_1 \ldots q_k$ if $k > 0$,
-and $[a]\, q_0 := b$ if $k = 0$.
-In the first case, $b$ is a label of the rule,
-and in the second case, it is a terminal symbol.
-And in both cases, $a$ is a constraint on the label of rule applied
-on the parent node in the derivation tree.
-This features of observing the parent's label
-are useful in the case of infinite alphabet,
-where it is not possible to memorize a label with the states.
-%
-%One can observe that
-\noindent The weight of a labeled derivation tree $t$
-of the weighted CF grammar associated to~$A$ as above,
-is $\weight_A(q, t)$,
-when $q$ is the start non-terminal.
-%
-We shall now establish a correspondence between such a derivation tree $t$
-and some word describing a linearization of $t$,
-in a way that $\weight_A(q, t)$ can be computed by a $\SWVPA$.
-
-Let $\hat\Omega$ be the countable (unranked) alphabet obtained
-from $\Omega$ by:
-$\hat\Omega = \Deltai \uplus \Deltac \uplus \Deltar$, with
-$\Deltai = \Omega_0$,
-$\Deltac = \{ \; \call{a} \mid a \in \Omega_{>0} \}$,
-$\Deltar = \{ \; \return{a} \mid a \in \Omega_{>0} \}$.
-
-\noindent
-We associate to $\hat\Omega$
-a label theory $\hat{\Phi}$
-like in Section~\ref{sec:SWVPA-def},
-%
-\noindent
-and we define a linearization of trees of $\T_\Omega$ into
-words of ${\hat\Omega}^*$ as follows:
-\begin{description}
-\item $\lin(a) = a$ for all $a \in \Omega_0$,
-\item $\lin\bigl( b(t_1, \ldots, t_k)\bigr) =
-       \call{b} \; \lin(t_1) \ldots \lin(t_k) \; \return{b}$
-       when $b \in \Omega_k$ for $1 \leq k \leq M$.
-\end{description}
-
-\begin{proposition}\label{lem:SWTA}
-For all \SWTA $A$ over~$\Omega$, $\Semiring$ commutative, and $\bar\Phi$,
-there exists an effectively constructible \SWVPA $A'$ over
-$\hat\Omega$, $\Semiring$ and $\hat\Phi$
-such that for all $t \in \T_\Omega$, $A'\bigl(\lin(t)\bigr) = A(t)$.
-\end{proposition}
-%
-\begin{proof}
-Let $A = \< Q, \init, \bar{\wei} >$ where $\bar{\wei}$ is presented as above by a function
-    %
-We build
-$A' = \< Q', P', \init', \bar{\wei}', \final' >$,
-%computing over $\Delta = \Omegai \uplus \Omegac \uplus \Omegar$,
-where $Q' = \bigcup_{k=0}^{M} Q^k$ is the set of sequences of state symbols of $A$,
-of length at most $M$, including the empty sequence denoted by~$\varepsilon$,
-and where $P' = Q'$ and $\bar\wei$ is defined by:
-
-\[
-\begin{array}{lcll}
-\weii(q_0\, \bar{u}, {\call{c}}, \bar{p}, a, \bar{u}) & = & \wei(q_0, c, a, \varepsilon) &
-\mathrm{for~all~} c \in \Omega_{>0}, a \in \Omega_0\\ %\bar{p}\in P',
-%
-\weiei(q_0\,\bar{u}, a, \bar{u}) & = & \wei(q_0, \varepsilon, a, \varepsilon) &
-\mathrm{for~all~} a \in \Omega_0\\
-%
-\weic(q_0\,\bar{u}, {\call{c}}, \bar{p}, \call{d}, \bar{u}, \bar{q}) & = & \wei(q_0, c, d, \bar{q}) &
-\mathrm{for~all~} c, d \in \Omega_{>0}\\ % \bar{p}\in P'
-%
-\weiec(q_0\,\bar{u}, \call{c}, \bar{u}, \bar{q}) & = & \wei(q_0, \varepsilon, c, \bar{q}) &
-\mathrm{for~all~} c \in \Omega_{>0}\\
-%
-%\weir: Q \times \Omegac \times P \times \Sigmar \times Q \to \Semiring &
-\weir(\varepsilon, {\call{c}}, \bar{p}, {\return{c}}, \bar{p}) & = & \one &
-\mathrm{for~all~}  c \in \Omega_{>0}\\ % \bar{p}\in P'
-%
-%\weie: Q \times \Omegar \times Q \to \Semiring &
-\weier(\bar{u}, {\return{c}}, \bar{q}) & = & \zero &
-\mathrm{for~all~}  c \in \Omega_{>0}
-\end{array}
-\]
-\noindent
-All cases not matched by one of the above equations have a weight $\zero$,
-for instance  %This is in particular the case of
-$\weir(\bar{u}, {\call{c}}, \bar{p}, {\return{d}}, \bar{q}) = \zero$
-if $c \neq d$
-or $\bar{u} \neq \varepsilon$
-or $\bar{q} \neq \bar{p}$.
-%
-%\noindent
-%It is sufficient to consider in $Q'$ only the prefixes of
-%sequences in transition with a non-null weight.
-%\qed
-\end{proof}
+\input{app-trees}
 

### 2021-07-09 app-bestsearch.tex

% !TEX root = main.tex
%
% Proof of best-search algo

The correctness of Algorithm~\ref{algo:Dijkstra}
is stated by the following lemma.

\begin{lemma}\label{lem:bot}
At the termination of Algorithm~\ref{algo:Dijkstra}, 
for all $\< q_1, q_2> \notin \Q$,
$d_\bot(q_1, q_2) =  b_\bot(q_1, q_2)$.
\end{lemma}

The proof is by contradiction,
assuming a counter-example minimal in the length of the witness word.

is ensured by the invariant... expressed in the following lemma.

\begin{lemma}\label{lem:top}
At every step of Algorithm~\ref{algo:Dijkstra}, 
for all $\< q_1, p, q_2> \notin \Q$,
$d_\top(q_1, p, q_2) = b_\top(q_1, p, q_2)$.
\end{lemma}

\noindent
For computing the minimal weight of a computation of $A$, we use the equality~\eqref{eq:min}, 
which, with Lemma~\ref{lem:bot} implies that
at the termination of Algorithm~\ref{algo:Dijkstra}, %it holds that,
%There exist $q_1, q_2 \in Q$
\(
  {\displaystyle \bigoplus_{s \in \Delta^*} A(s)} =
  {\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
  \mathsf{in}(q) \mathop{\otimes} d_\bot(q, q') \mathop{\otimes} \mathsf{out}(q').
\)

### 2021-07-09 introduction.tex
explication algo best-search & debut de preuve

@@
 % !TEX root = main.tex
-%\todo
+%
 % Introduction
@@
 approaches~\cite{Goodman99SemiringParsing,Nederhof03weightedParsing,MorbitzVogler19weighted-parsing}.
-\florent{\cite{MorbitzVogler19weighted-parsing} much is more general}
+\florent{WARNING: \cite{MorbitzVogler19weighted-parsing} much is more general that weighted parsing}
 In \emph{weighted language models},
@@
 
-In short, a transition rule $q \xrightarrow{\phi} q'$ from state $q$ to $q'$ of a \swM,
+In short, a transition rule $q \xrightarrow{\phi} q'$ from state $q$ to $q'$ of a \SWA,
 %input symbols are variables, %(parameters)
@@
 %\wrt an edit distance
-The latter model of automata operates on \emph{nested words}~\cite{AlurMadhusudan09nested},
-a structured form of words parenthesized with markup symbols,
-corresponding to a linearization of trees.
+The latter model of automata operates sequentially on \emph{nested words}~\cite{AlurMadhusudan09nested},
+structured with markup symbols (parentheses) and able to describe linearizations of trees.
+%
 In the context of parsing, they can represent (weighted) AST of CF grammars.
@@
 $T(s, t) \otimes A(t)$ (\wrt the ranking defined by $\oplus$), given an input word $s$.
-The latter value is called the distance between $s$ and $A$ in~\cite{Mohri03EDWA}.
+The latter product is called the distance between $s$ and $A$ in~\cite{Mohri03EDWA}.
 %
-
 Like weighted-parsing
 methods~\cite{Goodman99SemiringParsing,Nederhof03weightedParsing,MorbitzVogler19weighted-parsing},
+\florent{WARNING: \cite{MorbitzVogler19weighted-parsing} much is more general that weighted parsing}
 our approach proceeds in two steps,
@@
 of \emph{pitches} (e.g., A4, G5, etc.), temporal
-information is encoded as \emph{durations} (whole \musWhole$,
-$quarter, $\musQuarter$, eight $\musEighth$, etc), and notes are grouped in
+information is encoded as \emph{durations} (whole $\musWhole$,
+quarter $\musQuarter$, eight $\musEighth$, etc), and notes are grouped in
 high-level structures (beams, measures, tuplets). The following inputs

### 2021-07-09 main.tex

@@
 %
-Intuitively, when $b_\top(q, p, q') = \phi \in \Phic$, 
-for all $c \in \Deltac$,  $\phi(c)$ is 
+Intuitively, $b_\top(q, p, q')$ is a function of $\Phic$, 
+mapping every $c \in \Deltac$ to 
 the minimum weight of a computation of~$A$
 starting in state $q$ with a non-empty stack 
-$\gamma' = \<c, p> \, \gamma \in \Gamma^+$ 
+$\gamma' = \<c, p> \, \gamma \in \Gamma^+$, 
 and ending in state $q'$ with the same stack $\gamma'$,
@@
 %
+% Note that having a stack reduced to such a symbol makes impossible the application of the
+% two last cases in the definition of $\weight_A$ (return symbol and empty stack).
+% However, it is possible to apply the two first cases
+% (internal symbol or call symbol, with a push on the top of $\top$).
+%
 Formally, in order to define $b_\top$, we consider 

### 2021-07-09 main.tex
, expl. Corr. section SWT & expl. def. SWVPA

@@
 Since $B_{A, T}$  will read only in the output word, such a transition corresponds
-to an $\varepsilon$-transition of $\SWA$,
-but $\SWA$ have been defined without $\varepsilon$-transitions.
+to an $\varepsilon$-transition of $\SWA$.
+But $\SWA$ have been defined without $\varepsilon$-transitions.
 Therefore, in order to take care of this case, we perform an on-the-fly
@@
 \]
+\florent{init. $\final'$}
 
@@
 \)
-\florent{proof correctness}
+\florent{proof correctness (app)}
 %\qed
@@
 %of~$T$ is its number of states $|Q|$.
-\florent{revise with nb of tr. and states}
-
+\florent{revise complexity with nb. of tr. and states}
+%
+The particular case of Proposition~\ref{prop:epsilon} for a singleton $A$,
+\ie $A(s) = \one$ for a given $s$ and $A(s') = \zero$ for all $s' \neq s$, 
+corresponds to a construction of a \SWA for the partial application of a \SWT, 
+fixing the first argument $s$.
 \begin{corollary} \label{cor:epsilon}
@@
 whose elements are respectively called
-\emph{internal}, \emph{call} and \emph{return} symbols.
+\emph{internal}, \emph{call} and \emph{return} symbols, 
+following~\cite{AlurMadhusudan09nested}.
 % names are  coined by application to functional program verification
@@
 $[Q \times \Gamma^*] \times \Delta^* \times [Q \times \Gamma^*]$ into~$\Semiring$.
-The empty stack is denoted by $\bot$, and the upmost symbol is the last pushed content.
-The following functions illustrate each of the possible cases, being : reading $a \in \Deltai$,
-\lydia{intro to func}
-or $\call{c} \in \Deltac$, or $\return{r} \in \Deltar$ for each possible state of the stack (empty or not),
-to add to $u \in {\Delta}^*$.
-\lydia{introduced the 6 cases}
-\florent{notation $cp$ for $\< c, p>$ ?}
+The empty stack is denoted by $\bot$, and the first %upmost 
+symbol is the last pushed content.
+%
+The recursive definition of $\weight_A$
+enumerates each of the six possible cases: 
+reading $a \in \Deltai$,
+%\lydia{intro to func}
+or $\call{c} \in \Deltac$, or $\return{r} \in \Deltar$,
+for each possible state of the stack (empty or not).
+%to add to $u \in {\Delta}^*$.
+%\lydia{introduced the 6 cases}
+\florent{shorter notation $cp$ for $\< c, p>$ ?}
 
@@
 as in the case of the Boolean semiring~\cite{dAntonyAlur14SVPDA}.
-\florent{complete proof}
+\florent{complete proof (app)}
 \end{proof}
@@
 %...
+
 \florent{2 lines Application to Automated Music Transcription:

### 2021-07-09 main.tex
petite passe (preuves) 2 premieres section)

@@
 }
+\florent{ajout d'une explication avant le Corrolaire}
 The following property will be useful to the approach on
@@
 \[
-\wei'_1\bigl( \< p_1, q_1>, \< p_2, q_2>\bigr) =
+\begin{array}{rcl}
+\wei'_1\bigl( \< p_1, q_1>, \< p_2, q_2>\bigr) & = & 
 \wei_1(p_1, p_2) \otimes
@@
 \wei_{11}(q_1, q_2)
-\bigr].
+\bigr]\\
+\final'(p_1, q_1) & = & \final_A(p_1) \otimes \final_T(q_1)
+\end{array}
 \]
-\florent{init. $\final'$}
-
-\noindent
-Iterate the following for all $p_1\in P$ and $q_1, q_2 \in Q$:
+
+\noindent
+Then, we iterate the following for all $p_1\in P$ and $q_1, q_2 \in Q$:
 for all $p_2\in P$ and $q_3 \in Q$,
 \[
-\wei'_1\bigl( \< p_1, q_1>, \< p_2, q_3>\bigr) \opluseq
+\begin{array}{rcl}
+\wei'_1\bigl( \< p_1, q_1>, \< p_2, q_3>\bigr) & \opluseq & 
 \displaystyle\bigoplus_{\Sigma} \wei_{10}(q_1, q_2)
 \otimes
-\wei'_1\bigl( \< p_1, q_2>, \< p_2, q_3>\bigr)
-\]
-and
-\(
-\final'(p_1, q_1) \opluseq
+\wei'_1\bigl( \< p_1, q_2>, \< p_2, q_3>\bigr)\\
+\final'(p_1, q_1) & \opluseq & 
 \displaystyle\bigoplus_{\Sigma} \wei_{10}(q_1, q_2)
 \otimes \final'(p_1, q_2)
-\)
+\end{array}
+\]
 \florent{proof correctness (app)}
@@
 \noindent
-\florent{explication Fig.~\ref{fig:best-update} suivant cas de \eqref{eq:SWVPA-weight}}
+\florent{explication Fig.~\ref{fig:best-update} (update) suivant cas de \eqref{eq:SWVPA-weight}?}
 The infinite sums in the updates of $d$ in Algorithm~\ref{algo:Dijkstra},
@@
 
-\florent{complete **}
-** effectively computable by hypothesis that the label theory is effective**
-termination: The algorithm performs $2.|Q|^2$ iterations until $P$ is empty,
+Termination: the algorithm performs $2.|Q|^2$ iterations until $P$ is empty,
 and each iteration has a time complexity $O(|Q|^2 . |P|)$.
-\florent{detail with nb tr. and states}
+\florent{complexity: detail with nb tr. and states?}
 That gives a time complexity $O(|Q|^4 . |P|)$.
@@
 \noindent
+** effectively computable by hypothesis that the label theory is effective**
+\florent{REVOIR}
 In order to obtain effectively a witness
@@
 we require the additional property of convexity of weight functions.***
-\florent{completer}
 

### 2021-07-09 introduction.tex

@@
 %
-Parsing %variant of membership problem for formal languages
-%- given a language model and an input word $s$, compute a derivation of the model that yields $s$.
-is the problem %process
-of structuring a linear representation on input %typically
-(a finite word), according to a language model. % (a formal grammar).
+Parsing is the problem
+of structuring a linear representation
+(a finite word) according to a language model. % (a formal grammar).
 % natural language, programming language,
@@
 Considering large or infinite alphabets can however be of
-practical interest, %in other cases.
+practical interest, when dealing with large characters encodings such as UTF-16,
+% processing strings in \eg for vulnerability detection in Web-applications~\cite{dAntoni21CACM},
 %
-for instance, when dealing with large characters encodings such as UTF-16, % processing strings in
-\eg for vulnerability detection in Web-applications~\cite{dAntoni21CACM},
-%
-for the analysis (\eg validation or filtering)
-of data streams or serialization of structured documents
-(with textual or numerical attributes)~\cite{Segoufin06csl},
-or for processing timed execution traces~\cite{Bouyer03algebraic}.
+analysis of data streams or serialization of structured documents~\cite{Segoufin06csl},
+or processing timed execution traces~\cite{Bouyer03algebraic}.
 % algebraic definition of a class of data languages
@@
 
-The latter case is related to a study that motivated the present work:
-automated music transcription. Most representations of music are
-essentially linear. This is true for audio files,
-but also for widely used symbolic representations like MIDI.
+The latter case is related to a motivation  of  the present work:
+automated music transcription. Representations that capture  music performances
+are
+essentially linear:   audio files, or the widely used MIDI format~\cite{?}.
 Such representations ignore the hierarchical structures that frame the
@@
 % = ...a ranking of ASTs
-This is roughly the principle of \emph{weighted parsing}
-approaches~\cite{Goodman99SemiringParsing,Nederhof03weightedParsing,MorbitzVogler19weighted-parsing}.
+%This is roughly the principle of \emph{weighted parsing} approaches
 \florent{WARNING: \cite{MorbitzVogler19weighted-parsing} much is more general that weighted parsing}
-In \emph{weighted language models},
+In \emph{weighted language models}~\cite{Goodman99SemiringParsing,Nederhof03weightedParsing,MorbitzVogler19weighted-parsing},
 like \eg probabilistic context-free grammars % (CFG),
@@
 generalizing the Boolean guards of~\SA %in transitions
-into functions into an arbitrary semiring,
+to functions into an arbitrary semiring,
 and generalizing also~\WA, by handling infinite alphabets, see Figure~\ref{fig:hierarchy}.
@@
 \caption{Classes of Symbolic/Weighted Automata.
-$\Sigma_\mathsf{fin}$ is a finite alphabet,
-$\Sigma_\mathsf{inf}$ is a countable alphabet,
-$\mathbb{B}$ is the Boolean algebra,
-$\mathbb{S}$ is a commutative semiring,
+$\Sigma_\mathsf{fin}$ and $\Sigma_\mathsf{inf}$ denote finite/countable alphabets,
+$\mathbb{B}$  the Boolean algebra,
+$\mathbb{S}$ a commutative semiring.
 $q \xrightarrow{\dots} q'$ is a transition between states $q$ and $q'$.}
@@
 a \SWVPA $B$ %$A_{T, s}$
-is built, such that for all~$t$,  $B(t) = T(s, t) \otimes A(t)$.% $A_{T, s}(t) = T(s, t) \otimes A(t)$.
+is built, such that for all~$t$,
+$B(t) = T(s, t) \otimes A(t)$.% $A_{T, s}(t) = T(s, t) \otimes A(t)$.
 In the second step, a best AST $t$ is found by applying to $B$ %$A_{T, s}$

### 2021-07-09 semiring.tex
Sections labels et semirings

@@
 % Semiring basics
-%
-We shall consider semirings for the weight values of our language models.
+%%
+%We shall consider semirings for the weight values of our language models.
 %\florent{The results are established for a general class of semirings. They can be instantiated for concrete cases}
@@
 in order to handle non-determinism,
-and $\otimes$ combines two values into a single value.
+and $\otimes$ combines two values into a single one.
 %and let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a {semiring},
 
-\medskip%\noindent
 A semiring $\Semiring$ is \emph{commutative} if $\otimes$ is commutative.
@@
 %When  holds,
-%We say in this case that
-To express the property of Lemma~\ref{lem:monotonic}, we call
-$\Semiring$ \emph{monotonic} \wrt $\leq_\oplus$.
+
+We then say the
+$\Semiring$ is \emph{monotonic} \wrt $\leq_\oplus$.
 %A semiring $\Semiring$
@@
 \noindent
-\philippe{Here the difference between $\Semiring$ as a structure and as a domain is blurred.}
-We shall need below infinite sums with~$\oplus$.
+We  need infinite sums with~$\oplus$.
 A semiring~$\Semiring$ is called \emph{complete}~\cite{Droste09handbook}
@@
 of elements of $\dom(\Semiring)$ over an index set $I \subset \mathbb{N}$, such that:
-\philippe{$j\in \mathbb{N}$: $j$ is en element of $\mathbb{N}$, not the same s $j\subset \mathbb{N}$}
-%the following holds: %properties
 \begin{description}
@@
 production rules. Since there exists several possible division,
-rules can be weighted
-in the tropical semiring whose domain $\mathbb{R}_+ \cup \{ +\infty\}$,
-$\oplus$ is min, $\zero = +\infty$, $\otimes$ is sum, and $\one = 0$.
+rules can be weighted in the tropical semiring.
 For instance, the following production rules define two possible divisions
-of a bounded time interval into respectively a duplet and a triplet.
+of a time interval into respectively a duplet and a triplet.
 \[
@@
 are possible with:
-
 \[
@@
 \]
-
-\endex
 \end{example}

### 2021-07-09 main.tex

@@
 \noindent
+\florent{mv formal def. weight for \SWA to app.?}
 As above in the case of \SWT,
@@
 \begin{proof}
+\florent{proof: sketch and mv to app.?}
 Let $T = \< Q, \init_T, \bar{\wei}, \final_T >$,

### 2021-07-09 main.tex
effectiveness

@@
 \end{equation}
-Since $\Semiring$ is complete, the infinite sum in~\eqref{eq:bbot} is well defined.
+The infinite sum in~\eqref{eq:bbot} is well defined since $\Semiring$ is complete.
 %
@@
 %$Q \times \{ \bot, \top \} \times Q$.
-%
-\noindent
 \florent{explication Fig.~\ref{fig:best-update} (update) suivant cas de \eqref{eq:SWVPA-weight}?}
+
+\noindent
 The infinite sums in the updates of $d$ in Algorithm~\ref{algo:Dijkstra},
@@
 we require the additional property of convexity of weight functions.***
-
-\begin{proposition}\label{th:best-search}
-For a \SWVPA $A$
-over $\Delta$,
-$\Semiring$ commutative, bounded, total and complete, %semiring
-and $\bar\Phi$ effective, %and $k$-convex, % label theory,
-one can construct in PTIME a word $t \in \Delta^*$
-such that $A(t)$ is minimal \wrt the natural ordering for $\Semiring$.
-\end{proposition}
 

### 2021-07-09 main.tex
Fin relecture (rapide) P.

--- 2021-07-09 main.tex
@@
 These functions are similar to the guards of symbolic automata~\cite{dAntoniVeanes17CAV,dAntoni21CACM},
-but they can return values in a generic semiring,
-whereas the latter guards are restricted to the Boolean semiring.
+but the latter guards are restricted to the Boolean semiring.
 %
@@
 such that, for all $q, q' \in Q$, $a \in \Sigma$,  $b \in \Delta$:
-
 \[
@@
 \begin{example}\label{ex:SWT}
-Let us develop the example of comparison between music played by a performer,
+We develop the example of comparison between music played by a performer,
 represented as a sequence $s \in \Sigma^*$ of events in the MIDI alphabet $\Sigma$,
@@
 If one performed event $s_i$  corresponds
-to one notated event $t_1$ (for instance MIDI code 61 and pitch A4),
+to one notated event $t_j$ (for instance MIDI code 61 and pitch A4),
 the weight value computed by the \SWT is the time distance between both,
@@
 %
-We also must be able to take performing errors into account, while still being able to compare with the score,
+We also want to take performing errors into account, while still being able to compare with the score,
 since a performer could, for example, play an unwritten extra note.
@@
 $\init(q_1) = \final(q_1) = \zero$.
-
-\lydia{ccl to the ex}
-That way, an \SWT is capable of evaluating the differences between a score and a performance,
+This \SWT is capable of evaluating the differences between a score and a performance,
 all the while ensuring that performance errors are plausible.
@@
 \noindent
-\philippe{Il me manque une explication:  on construit un automate
-qui, étant donnée une partition $t$, renvoie la distance minimale
-avec n'importe quelle performance (distance donnée par un transducer)?
-Quel est le rôle de A(s)?
-}
-\florent{ajout d'une explication avant le Corrolaire}
 The following property will be useful to the approach on
@@
 \begin{example}
-Consider once more the scores of Exemple~\ref{ex:running}.
+Consider once more the scores of Example~\ref{ex:running}.
 $\Deltai$ represents the set of music events of the form $<n, t>$
@@
 \begin{proof}
-The construction is essentially the same
-as in the case of the Boolean semiring~\cite{dAntonyAlur14SVPDA}.
+The construction is similar
+to that of the Boolean semiring~\cite{dAntonyAlur14SVPDA}.
 \florent{complete proof (app)}
 \end{proof}
-
-%\subsection{Best Search}
-\label{sec:best}\label{sec:search}
 \noindent
@@
 Therefore, the framework, applied to the transcription problem, allows to find an optimal
-solution that considers both the fitness of the result to the input, and the complexity of the former.
+solution that considers both the fitness of the result, and its structural complexity.
 \end{example}

### 2021-07-09 main.tex
revise proof prop. closure and mv def. effectiveness

@@
 $\Semiring$ commutative, bounded and complete,
-and $\bar\Phi$ effective,
+and $\bar\Phi$, %effective,
 and a $\SWA$ $A$ over $\Sigma$, $\Semiring$ and $\bar\Phi$,
@@
 \end{proposition}
+\florent{AV: not needed $\bar\Phi$ effective}
 %
@@
 \florent{proof: sketch and mv to app.?}
-Let $T = \< Q, \init_T, \bar{\wei}, \final_T >$,
-where $\bar{\wei}$ contains $\wei_{10}$, $\wei_{01}$, and $\wei_{11}$,
-from $Q \times Q$ into respectively
-$\Phi_{\Sigma}$, $\Phi_{\Delta}$, and $\Phi_{\Sigma, \Delta}$,
-and let $A = \< P, \init_A, \wei_1, \final_A >$
-with $\wei_1: Q \times Q \to \Phi_{\Sigma}$.
-%
-\noindent
-The state set of $B_{A, T}$ will be $Q' = P \times Q$.
-The entering, leaving and transition functions of $B_{A, T}$ will
-simulate synchronized computations of $A$ and $T$,
-while reading an output word of $\Delta^*$.
-%
-Its state entering functions is defined
-for all $p \in P$, $q \in Q$
-by $\init'(p, q) = \init_A(p) \otimes \init_T(q)$.
-The transition function $\wei'_1$ will roughly perform
-a synchronized product of transitions defined by $\wei_1$,
-$\wei_{01}$ ($T$ reading in output word and not an input word)
-and $\wei_{11}$ ($T$ reading both an input word and an output word).
-%
-Moreover, $\wei'_1$ also needs to simulate transitions
-defined by $\wei_{10}$: $T$ reading in input word and not an output word.
-Since $B_{A, T}$  will read only in the output word, such a transition corresponds
-to an $\varepsilon$-transition of $\SWA$.
-But $\SWA$ have been defined without $\varepsilon$-transitions.
-Therefore, in order to take care of this case, we perform an on-the-fly
-suppression of $\varepsilon$-transition in the $\SWA$ in construction,
-following the algorithm of~\cite{LombardySakarovitch12ciaa}.
-%
-% Initialize state leaving functions i
-%and $\final'(p, q) = \final_A(p) \otimes \final_T(q)$.
-
-%Every transition of $B_{A, T}$ will
-%simulate a sequence of transitions of $T$ performing the following steps:
-%advance in the input word while staying immobile in the output word,
-%and then make one step in the output word (and advance in the input word or not).
-
-\noindent
-Initially, for all $p_1, p_2 \in P$, and $q_1, q_2 \in Q$, let
-\[
-\begin{array}{rcl}
-\wei'_1\bigl( \< p_1, q_1>, \< p_2, q_2>\bigr) & = &
-\wei_1(p_1, p_2) \otimes
-\bigl[
-\wei_{01}(q_1, q_2)
-\oplus
-\displaystyle\bigoplus_{\Sigma}
-\wei_{11}(q_1, q_2)
-\bigr]\\
-\final'(p_1, q_1) & = & \final_A(p_1) \otimes \final_T(q_1)
-\end{array}
-\]
-
-\noindent
-Then, we iterate the following for all $p_1\in P$ and $q_1, q_2 \in Q$:
-for all $p_2\in P$ and $q_3 \in Q$,
-\[
-\begin{array}{rcl}
-\wei'_1\bigl( \< p_1, q_1>, \< p_2, q_3>\bigr) & \opluseq &
-\displaystyle\bigoplus_{\Sigma} \wei_{10}(q_1, q_2)
-\otimes
-\wei'_1\bigl( \< p_1, q_2>, \< p_2, q_3>\bigr)\\
-\final'(p_1, q_1) & \opluseq &
-\displaystyle\bigoplus_{\Sigma} \wei_{10}(q_1, q_2)
-\otimes \final'(p_1, q_2)
-\end{array}
-\]
-\florent{proof correctness (app)}
+\input{app-closure}
 %\qed
@@
 Formally, the transitions of the automaton~$A$ are defined
-in term of %a weight value computed by
-an intermediate function $\weight_A$, like in Section~\ref{sec:SWT}.
-%In the case of a pushdown automaton,
-A configuration, denoted by $q[\gamma]$,
-is here composed of a state $q \in Q$
+with an intermediate function $\weight_A$, like in Section~\ref{sec:SWT}.
+%
+A configuration $q[\gamma]$
+is composed of a state $q \in Q$
 and a stack content $\gamma \in \Gamma^*$,
@@
 or $\call{c} \in \Deltac$, or $\return{r} \in \Deltar$,
-for each possible state of the stack (empty or not).
+and for each possible state of the stack (empty or not).
 %to add to $u \in {\Delta}^*$.
@@
 \end{align}
-\lydia{c p to <c, p>}
+%\lydia{c p to <c, p>}
 %
@@
 a word %$w \in \Delta^*$
-of minimal weight for~$A$,
-as stated in the following proposition. %\wrt~$\leq_\oplus$,
+of minimal weight for~$A$.
+%as stated in the following proposition. %\wrt~$\leq_\oplus$,
+It requires the following property of label theories.
+%
+\begin{definition}\label{def:effective}
+A label theory is called \emph{effective} when
+for all $\phi \in \Phi_\Sigma$ and $\eta \in \Phi_{\Sigma, \Delta}$,
+$\bigoplus_{\Sigma} \phi$,
+$\bigoplus_{\Delta}\bigoplus_{\Sigma} \eta$, and
+$\bigoplus_{\Sigma}\bigoplus_{\Delta} \eta$
+can be effectively computed from $\phi$ and $\eta$, 
+and moreover, the number of symbols reaching these bounds is finite 
+and can be effectively computed.
+\end{definition}
 %
@@
 prioritized by the value returned by $d$.
+** where  $T$ is the worst time complexity for an oracle computing the values
+of Definition~\ref{def:effective}.
 %$|Q|^3.\log(|Q|^2)$
@@
 In order to obtain effectively a witness
-(word of $\Delta^*$ with a computation of $A$ of minimal weight),
-we require the additional property of convexity of weight functions.***
+(word of $\Delta^*$ with a computation of $A$ of minimal weight),...
+%we require the additional property of convexity of weight functions.***
+using the second requirement in Definition~\ref{def:effective}.
 

### 2021-07-10 main.tex
Révision exemple phase 1

@@
 \begin{example}\label{ex:SWT}
-We develop the example of comparison between music played by a performer,
-represented as a sequence $s \in \Sigma^*$ of events in the MIDI alphabet $\Sigma$,
-and  a music score represented as a sequence $t \in \Delta^*$
-in the CWMN alphabet $\Delta$. We build a small weighted transducer model
+We build a small weighted transducer model
 with two states $q_0$ and $q_1$ that
-calculates the distance between $s$ and $t$.
-
-If one performed event $s_i$  corresponds
-to one notated event $t_j$ (for instance MIDI code 61 and pitch A4),
+calculates the distance between the performance $I$ and
+the score $O$ of Example~\ref{ex:running}.
+
+If one performed event $\mu_i$  corresponds
+to one notated event $\nu_j$ (for instance MIDI code 61 and pitch A4),
 the weight value computed by the \SWT is the time distance between both,
-as in Example~\ref{distance-time}, and is modeled by
+(see Example~\ref{distance-time}), and is modeled by
 transitions $\wei_{11}$ below.
@@
 If we meet the music notation symbol '-' that
-represents continuation (such as instance in \emph{ties}
+represents continuation (this occurs for instance in \emph{ties}
 $\musQuarter\!\!\!\mathrel{\raisebox{-1.5mm}{$\smile$}}\!\!\!\musEighth$,
@@
 \begin{array}{rclcrcl}
-\wei_{11}(q_0, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d| & \quad &
-\wei_{11}(q_1, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d|\\
-\wei_{01}(q_0, \varepsilon, \< \mathsf{-}, d'>, q_0) & = & \one & &
-\wei_{01}(q_1, \varepsilon, \< \mathsf{-}, d'>, q_0) & = & \one\\
-\wei_{10}(q_0, d, \varepsilon, q_1) & = & \alpha & & %\multicolumn{3}{l}{\mathrm{for~all~} b \in \Delta}
+\wei_{11}(q_0, \<\mu, \tau>, \<\nu, \tau'>, q_0) & = & |\tau' - \tau| & \quad &
+\wei_{11}(q_1, \<\mu, \tau>, \<\nu, \tau'>, q_0) & = & |\tau' - \tau|\\
+\wei_{01}(q_0, \varepsilon, \< \mathsf{-}, \tau'>, q_0) & = & \one & &
+\wei_{01}(q_1, \varepsilon, \< \mathsf{-}, \tau'>, q_0) & = & \one\\
+\wei_{10}(q_0, \<\mu, \tau>, \varepsilon, q_1) & = & \alpha & & %\multicolumn{3}{l}{\mathrm{for~all~} b \in \Delta}
 \end{array}
@@
 \begin{example}
+
 Consider once more the scores of Example~\ref{ex:running}.
-$\Deltai$ represents the set of music events of the form $<n, t>$
-where $n$ is either a note (e.g. A4), the continuation symbol $-$, or a
+$\Deltai$ represents the set of music events of the form $<\nu, \tau>$
+where $\nu$ is either a note (e.g. A4), the continuation symbol $-$, or a
 rest symbol R. $\Deltac$ and  $\Deltar$ are used to encode the score structure:
-Let $\Deltac = \{[, <]\}$ and  $\Deltar = \{], >]\}$ where $[$ (rep. $]$)
-and $<$ (resp. $>$) correspond to start/end of measures, and start/end of
-beams. Then  \includegraphics[scale=0.20]{pictures/score5.png}
-is encoded with the nested word $[0, \frac{3}{4}, \frac{7}{8}>][<1, \frac{4}{3}, \frac{5}{3}]$
+Let $\Deltac = \{\langle, \prec\}$ and  $\Deltar = \{\rangle, \succ\}$ where $\langle$
+(resp. $\rangle$)
+and $\prec$ (resp. $\succ$) correspond to start/end of measures, and start/end of
+tuplets. Then  \includegraphics[scale=0.20]{pictures/score5.png}
+is encoded with the nested word $\langle \prec 0, \frac{3}{4}, \frac{7}{8} \succ \rangle \langle \prec 1, \frac{4}{3}, \frac{5}{3}  \succ \rangle$
 \endex
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Symbolic Weighted Parsing}
@@
 \end{example}
+
+\begin{example}\label{rules-division}
+The recursive subdivision of time that leads to hierarchichal
+structures of music notation can be modeled as
+production rules. Since there exists several possible division,
+rules can be weighted in the tropical semiring.
+For instance, the following production rules define two possible divisions
+of a time interval into respectively a duplet and a triplet.
+\[
+\rho_1: q_0 \xrightarrow{0.06} \< q_{1}, q_{2}>,\
+\rho_2: q_0 \xrightarrow{0.12} \< q_{1}, q_{2}, q_{2}>.
+\]
+Further binary divisions of time sub-intervals %represented respectively by~$q_2$ and~$q_3$
+are possible with:
+\[
+\rho_3:\,q_2 \xrightarrow{0.1} \< q_{3}, q_{3}>, \
+\rho_4:\,q_3 \xrightarrow{0.11} \< q_{4}, q_{4}>. \
+\]
+
+The latter shows a subdivision
+based on the rules from Example~\ref{rules-division}. The first measure ($I_1$)
+results from the successive applications of  rules $\rho_1$ (division in two)
+and rules $\rho_3$ (division of the second half) in two. The second measure
+is a division in three obtained by rule $\rho_3$. It follows that the notation
+defines the sequence of timestamps $0, \frac{3}{4}, \frac{7}{8}, 1, \frac{4}{3}, \frac{5}{3}$.
+\end{example}
+
 %

### 2021-07-10 main.tex
revise proof prop. closure and mv def. effectiveness

--- 2021-07-09 main.tex
@@
 %
-The particular case of Proposition~\ref{prop:epsilon} for a singleton $A$,
-\ie $A(s) = \one$ for a given $s$ and $A(s') = \zero$ for all $s' \neq s$,
-corresponds to a construction of a \SWA for the partial application of a \SWT,
+The particular case of Proposition~\ref{prop:epsilon} with a singleton $A$,
+\ie such that $A(s) = \one$ for a given $s \in \Sigma^*$ 
+and $A(s') = \zero$ for all $s' \neq s$,
+corresponds to a construction of a \SWA for the partial application of the \SWT $T$,
 fixing the first argument $s$.

### 2021-07-10 introduction.tex

@@
 illustrate our framework with music transcription examples: Given
-a \emph{timeline} of musical events with arbitrary timestamps as input, parse it
+a \emph{timeline} of musical events with arbitrary timestamps from an infinite
+domain as input, parse it
 into a structured music score. In our example, input events
-are pairs $\< \eta, \tau>$ made of
-a symbol $\eta \in \Sigma$, where $\Sigma$ stands for the
+are pairs $\< \mu, \tau>$ made of
+a symbol $\mu \in \Sigma$, where $\Sigma$ stands for the
 set of MIDI message symbols~\cite{?}
-and  $\tau \in \mathbb{Q}$ is a timestamp. The output of parsing
-is a representation of the sequence in
+and  $\tau \in \mathbb{Q}$ is a timestamp. Such inputs typically correspond
+to the recording of a live performance. The output of parsing
+is a sequence of events (or \emph{notes}) $\< \nu, \tau'>$   in
 Common Western Music Notation (CWMN)~\cite{Gould11Notation}
-where event symbols belong to the domain $\Delta$
-of \emph{pitches} (e.g., A4, G5, etc.), temporal
-information is encoded as \emph{durations} (whole $\musWhole$,
-quarter $\musQuarter$, eight $\musEighth$, etc), and notes are grouped in
-high-level structures (beams, measures, tuplets). The following inputs
+where event symbols $\nu$ belong to the domain $\Delta$
+of \emph{pitches} (e.g., A4, G5, etc.), and timestamp $\tau'$
+belongs to a rhythmic ``grid'' obtained from recursive divisions:
+whole notes ($\musWhole$) splitted in halves ($\musHalf$), halves
+in quarters ($\musQuarter$), eights ($\musEighth$), etc.
+
+The following concrete sequences
 will be used:
 \begin{enumerate}
-  \item $I_1  = [ \<e_1, 0.07>,\<e_2, 0.72>, \<e_3, 0.91>]$, over interval $[0,1[$
-  \item $I_2  = [ \<e_3, 1.05>,\<e_4, 1.36>, \<e_5, 1.71>]$, over interval $[1,2[$
+  \item A performance  $I = [ \<\mu_1, 0.07>, \<\mu_2, 0.72>, \<\mu_3, 0.91>, \<\mu_4, 1.05>, \<\mu_5, 1.36>, \<\mu_6, 1.71>]$,
+     over interval $[0,2[$,
+  \item A score $O =$ \includegraphics[scale=0.20]{pictures/score5.png}, corresponding to
+      a hierarchical structures (measures, tuplets) that can be linearized as a sequence of events
+      $[\<\nu_1,0>, \<\nu_2,\frac{3}{4}>, \<\nu_3,\frac{7}{8}>, \<\nu_4,1>, \<\nu_5,\frac{4}{3}>, \<\nu_6\frac{5}{3}>]$
 \end{enumerate}
-There exists many possible parsings of $I_1 \cup I_2$ in music notation, among
-which
-\includegraphics[scale=0.20]{pictures/score5.png}
-and \includegraphics[scale=0.20]{pictures/score4.png}.
-SW-parsing associates a cost
-\florent{weight value}
+\smallskip
+
+We will show that $O$  can result for the
+parsing of $I$. Note that several other parsings (e.g.,
+\includegraphics[scale=0.20]{pictures/score4.png}) are possible.
+SW-parsing associates a weight value
 to each solution, and our framework
-aims at selecting the best one with respect to this cost.
+aims at selecting the best one with respect to this weight.
 \endex

### 2021-07-10 semiring.tex

@@
 
-\begin{example}\label{rules-division}
-The recursive subdivision of time that leads to hierarchichal
-structures of music notation can be modeled as
-production rules. Since there exists several possible division,
-rules can be weighted in the tropical semiring.
-For instance, the following production rules define two possible divisions
-of a time interval into respectively a duplet and a triplet.
-\[
-\rho_1: q_0 \xrightarrow{0.06} \< q_{1}, q_{2}>,\
-\rho_2: q_0 \xrightarrow{0.12} \< q_{1}, q_{2}, q_{2}>.
-\]
-Further binary divisions of time sub-intervals %represented respectively by~$q_2$ and~$q_3$
-are possible with:
-\[
-\rho_3:\,q_2 \xrightarrow{0.1} \< q_{3}, q_{3}>, \
-\rho_4:\,q_3 \xrightarrow{0.11} \< q_{4}, q_{4}>. \
-\]
-\end{example}
 %%% semiring properties used in paper

### 2021-07-10 main.tex
Gain de place (1)

--- 2021-07-10 main.tex
@@
 
-\begin{example}
-
+\begin{example}\label{ex:nested-word}
 Consider once more the scores of Example~\ref{ex:running}.
@@
 
-\noindent
 The infinite sums in the updates of $d$ in Algorithm~\ref{algo:Dijkstra},
@@
 since~$\Semiring$ is complete.
+
+Termination: the algorithm performs $2.|Q|^2$ iterations until $P$ is empty,
+and each iteration has a time complexity $O(|Q|^2 . |P|)$.
+\florent{complexity: detail with nb tr. and states?}
+That gives a time complexity $O(|Q|^4 . |P|)$.
+It can be reduced by implementing $P$ as a priority queue,
+prioritized by the value returned by $d$.
+** where  $T$ is the worst time complexity for an oracle computing the values
+of Definition~\ref{def:effective}.
+%$|Q|^3.\log(|Q|^2)$
+%
+At termination, $d_\bot = b_\bot$ and $d_\top = b_\top$
+(see Appendix~\ref{sec:bestsearch}).
+With~\eqref{eq:min}, this ensures
+the correctness of Algorithm~\ref{algo:Dijkstra}.
 
@@
 
-Termination: the algorithm performs $2.|Q|^2$ iterations until $P$ is empty,
-and each iteration has a time complexity $O(|Q|^2 . |P|)$.
-\florent{complexity: detail with nb tr. and states?}
-That gives a time complexity $O(|Q|^4 . |P|)$.
-It can be reduced by implementing $P$ as a priority queue,
-prioritized by the value returned by $d$.
-** where  $T$ is the worst time complexity for an oracle computing the values
-of Definition~\ref{def:effective}.
-%$|Q|^3.\log(|Q|^2)$
-%
-At termination, $d_\bot = b_\bot$ and $d_\top = b_\top$
-(see Appendix~\ref{sec:bestsearch}).
-With~\eqref{eq:min}, this ensures
-the correctness of Algorithm~\ref{algo:Dijkstra}.
-
 \noindent
@@
  representations, we aim at choosing the simpler one. For instance
- \includegraphics[scale=0.20]{pictures/score4.png} should be favored
- on \includegraphics[scale=0.20]{pictures/score5.png}, although both yield
- a same linearization. Costs  can be expressed
- in the grammar with weight associated to each production rule. See Example~\ref{rules-division}.
-
- \end{enumerate}
-Therefore, the framework, applied to the transcription problem, allows to find an optimal
-solution that considers both the fitness of the result, and its structural complexity.
-\end{example}
-
-\begin{example}\label{rules-division}
-The recursive subdivision of time that leads to hierarchichal
-structures of music notation can be modeled as
-production rules. Since there exists several possible division,
-rules can be weighted in the tropical semiring.
+ \includegraphics[scale=0.20]{pictures/score5.png} should be favored
+ on \includegraphics[scale=0.20]{pictures/score4.png}. Costs  can be expressed
+ in the grammar as weights associated to each production rule.
  For instance, the following production rules define two possible divisions
-of a time interval into respectively a duplet and a triplet.
+ of a time interval into respectively a duplet and a triplet. Each comes with
+ a specific cost.
  \[
@@
 
-The latter shows a subdivision
-based on the rules from Example~\ref{rules-division}. The first measure ($I_1$)
+Score \includegraphics[scale=0.20]{pictures/score5.png} is obtained as follows:
+the first measure
  results from the successive applications of  rules $\rho_1$ (division in two)
  and rules $\rho_3$ (division of the second half) in two. The second measure
-is a division in three obtained by rule $\rho_3$. It follows that the notation
-defines the sequence of timestamps $0, \frac{3}{4}, \frac{7}{8}, 1, \frac{4}{3}, \frac{5}{3}$.
+ is a division in three obtained by rule $\rho_3$. One obtains a parse tree
+ that can be linearized in the nested word of Example~\ref{ex:nested-word}.
+ \end{enumerate}
+Therefore, the framework, applied to the transcription problem, allows to find an optimal
+solution that considers both the fitness of the result, and its structural complexity.
 \end{example}

### 2021-07-10 abstract.tex
description algo search
Petits changements abstract

@@
 The transitions of \SWA are labeled by functions from an infinite alphabet into the weight domain.
-This is unlike \SA whose transitions are guarded by boolean predicates
+This is unlike \SA whose transitions are guarded by Boolean predicates
 overs symbols in an infinite alphabet
@@
 %
-We illustrate the models with examples taken from
-a motivating application, namely
-a parse-based approach to automated music transcription.
+We illustrate the model with a motivating application, namely
+automated music transcription.

### 2021-07-10 introduction.tex
Révision intro

@@
 NLP tasks such as constituency parsing,
-or of programming languages compilers or interpreters.
+programming languages compilers or interpreters.
 Considering large or infinite alphabets can however be of
@@
 %
-analysis of data streams or serialization of structured documents~\cite{Segoufin06csl},
+analysis of data streams, serialization of structured documents~\cite{Segoufin06csl},
 or processing timed execution traces~\cite{Bouyer03algebraic}.
@@
 by mapping input events to this hierarchical rhythmic space.
-It can therefore be stated as a parsing problem~\cite{foscarin:hal-01988990},
+It can therefore be stated as a parsing problem~\cite{foscarin:hal-01988990}
 over an infinite alphabet of timed events.
@@
 %words carrying data values in an infinite domain (e.g. integers) e.g. data processing
-For instance, some automata with memory extensions
+Some automata with memory extensions
 allow restricted storage and comparison of input symbols,
@@
 (sets of assignments of Boolean variables) %propositional variables)
-and in practice,  %implementation
+and, in practice,  %implementation
 each transition accepts a set of such symbols (instead of an individual symbol),
@@
 in symbolic automata (\SA)~\cite{dAntoniVeanes17CAV,dAntoni21CACM},
-the transitions are guarded by predicates over infinite alphabet domains.
+transitions are guarded by predicates over infinite  domains.
 With appropriate closure conditions on the sets of such predicates, % (alphabet theories),
-all the good properties enjoyed by automata over finite alphabets are preserved.
+all the good properties enjoyed over finite alphabets are preserved.
 
 Other extensions of language models  %(automata and grammars)
-help in dealing with non-determinism, by the computation of weight values.
+help in dealing with non-determinism by computing weight values.
 With an ambiguous grammar, there may exist several derivations
@@
 and weighted automata (\WA) \cite{Droste09handbook},
-a weight value is associated to each transition rule, % production rule
+a weight is associated to each transition rule, % production rule
 and the rule's weights can be combined with an
-associative product operator~$\otimes$ into the weight of an AST.
-A second operator~$\oplus$, associative and commutative,
+associative product operator~$\otimes$ to yield the weight of an AST.
+A second operator~$\oplus$
 is moreover used to resolve the ambiguity raised by the existence
@@
 associated to a given input word.
-Typically, $\oplus$ will select the best of two weight values.
+Typically, $\oplus$  selects the best of two weight values.
 %Intuitively,~$\oplus$ selects, or ranking, the syntax trees.
-The weight domain, equipped with these two operators shall be, at minima,
+The weight domain, equipped with these two operators is, at minima,
 a \emph{semiring} %$\Semiring$
@@
 to functions into an arbitrary semiring,
-and generalizing also~\WA, by handling infinite alphabets, see Figure~\ref{fig:hierarchy}.
-
-%***
-%some weighted language models
-%model of weighted CFG
-%computing on words over infinite input alphabets %sets of terminal symbol
-%- but with finite sets of states and transitions rules.
-% non-terminals and production rules).
+and generalizing also~\WA, by handling infinite alphabets (see Figure~\ref{fig:hierarchy}).
 
 In short, a transition rule $q \xrightarrow{\phi} q'$ from state $q$ to $q'$ of a \SWA,
+\philippe{Je ne comprends pas très ce que cette première phrase fait là. Je commencerais
+directement le paragraphe avec "This framework relies..."}
 %input symbols are variables, %(parameters)
@@
 with a visibly restriction~\cite{AlurMadhusudan09nested} (\SWVPA).
+A \SWT defines a distance $T(s, t)$ between finite words $s$ and $t$
+over infinite alphabets. % following~\cite{Mohri03EDWA}.
 %\wrt an edit distance
-The latter model of automata operates sequentially on \emph{nested words}~\cite{AlurMadhusudan09nested},
-structured with markup symbols (parentheses) and able to describe linearizations of trees.
+A  \SWVPA operates sequentially on \emph{nested words}~\cite{AlurMadhusudan09nested},
+structured with markup symbols (parentheses), and apt at describing linearizations of trees.
+In a parsing context, a \SWVPA $A$ associates a weight value $A(t)$ % \in \Semiring$
+to a given nested word $t$, which is itsef the linearization of a weighted AST. %representing a parse tree.
 %
-In the context of parsing, they can represent (weighted) AST of CF grammars.
-More precisely, a \SWVPA $A$ associates a weight value $A(t)$ % \in \Semiring$
-to a given nested word $t$, which is the linearization of an AST. %representing a parse tree.
-%
-On the other hand,
-a \SWT can define a distance $T(s, t)$ between finite words $s$ and $t$
-over infinite alphabets. % following~\cite{Mohri03EDWA}.
-Then, the \emph{SW-parsing} problem aims at %computing
+Then, given an input word $s$, the \emph{SW-parsing} problem aims at
 finding $t$ minimizing
-$T(s, t) \otimes A(t)$ (\wrt the ranking defined by $\oplus$), given an input word $s$.
-The latter product is called the distance between $s$ and $A$ in~\cite{Mohri03EDWA}.
-%
+$T(s, t) \otimes A(t)$, called the distance between $s$ and $A$ in~\cite{Mohri03EDWA},
+\wrt the ranking defined by $\oplus$,
 Like weighted-parsing
@@
 \florent{WARNING: \cite{MorbitzVogler19weighted-parsing} much is more general that weighted parsing}
-our approach proceeds in two steps,
-based on properties of the \swM.
+our approach proceeds in two steps.
 The first step is an intersection
@@
 \begin{example}[Running example]\label{ex:running}
-Throughout the paper we
-illustrate our framework with music transcription examples: Given
-a \emph{timeline} of musical events with arbitrary timestamps from an infinite
+We
+illustrate our framework with music transcription: given
+a \emph{timeline} of musical events with timestamps from an infinite
 domain as input, parse it
-into a structured music score. In our example, input events
-are pairs $\< \mu, \tau>$ made of
-a symbol $\mu \in \Sigma$, where $\Sigma$ stands for the
+into a structured music score. Input events
+are pairs $\< \mu, \tau>$, with $\mu \in \Sigma$, where $\Sigma$ stands for the
 set of MIDI message symbols~\cite{?}
@@
 in quarters ($\musQuarter$), eights ($\musEighth$), etc.
-
 The following concrete sequences
@@
 
-We will show that $O$  can result for the
+We will show that $O$ is a soution for the
 parsing of $I$. Note that several other parsings (e.g.,

### 2021-07-10 main.tex
Petits changements abstract
work on proofs in appendix

@@
 $\Semiring$ commutative, bounded and complete,
-and $\bar\Phi$, %effective,
+and $\bar\Phi$ effective,
 and a $\SWA$ $A$ over $\Sigma$, $\Semiring$ and $\bar\Phi$,
@@
 \end{proposition}
-\florent{AV: not needed $\bar\Phi$ effective}
 %
 \begin{proof}
-\florent{proof: sketch and mv to app.?}
-\input{app-closure}
+(sketch, see Appendix~\ref{sec:closure} for details).
+The state set of $B_{A, T}$ is the Cartesian product of 
+the state sets of $A$ and $T$ and its transitions simulate,
+while reading an output word $t \in \Delta^*$, the
+synchronized behaviour of $A$ and $T$ on $t$ 
+and some input word $s \in \Sigma^*$. 
+The weight for reading of the input is summarized using the operator
+$\bigoplus^1_\Sigma$.
+%
+The main difficulty comes from the transitions of $T$
+of the form $\wei_{10}$, who read in input and ignore the output.
+Such transition shall be simulated by $\varepsilon$-transitions in $B_{A, T}$,
+but $\varepsilon$-transitions are not defined for $\SWA$.
+Therefore, the $\varepsilon$-transitions
+are eliminated on-the-fly during the construction of~$B_{A, T}$,
+following a procedure of~\cite{LombardySakarovitch12ciaa}.
 %\qed
@@
 The construction is similar
-to that of the Boolean semiring~\cite{dAntonyAlur14SVPDA}.
+to the case of the Boolean semiring~\cite{dAntonyAlur14SVPDA}.
 \florent{complete proof (app)}
@@
 %as stated in the following proposition. %\wrt~$\leq_\oplus$,
-It requires the following property of label theories.
-%
-\begin{definition}\label{def:effective}
-A label theory is called \emph{effective} when
-for all $\phi \in \Phi_\Sigma$ and $\eta \in \Phi_{\Sigma, \Delta}$,
-$\bigoplus_{\Sigma} \phi$,
-$\bigoplus_{\Delta}\bigoplus_{\Sigma} \eta$, and
-$\bigoplus_{\Sigma}\bigoplus_{\Delta} \eta$
-can be effectively computed from $\phi$ and $\eta$,
-and moreover, the number of symbols reaching these bounds is finite
-and can be effectively computed.
-\end{definition}
+%It requires the following property of label theories.
 %
@@
 
-\begin{figure}
+\begin{figure}[ht]
 For all $q_0, q_3 \in Q$, %$p \in P$,
@@
 
+\section{Properties of Label Theory Operators} \label{sec:closure}
+\input{app-labels}
+
+\section{Proof of Proposition~\ref{prop:epsilon}} \label{sec:closure}
+\input{app-closure}
+
+\section{Correctness of the Best-Search Algorithm} \label{sec:bestsearch}
+\input{app-bestsearch}
+
 \section{Nested-Words and Parse-Trees} \label{sec:trees}
 \input{app-trees}
-
-\section{Correctness of the Best-Search Algorithm} \label{sec:bestsearch}
-\input{app-bestsearch}
 

### 2021-07-10 introduction.tex
1 pass intro

@@
 automated music transcription. Representations that capture  music performances
-are
-essentially linear:   audio files, or the widely used MIDI format~\cite{?}.
+are essentially linear:   
+audio files, or the widely used MIDI file format~\cite{?}.
 Such representations ignore the hierarchical structures that frame the
@@
 and generalizing also~\WA, by handling infinite alphabets (see Figure~\ref{fig:hierarchy}).
-
-In short, a transition rule $q \xrightarrow{\phi} q'$ from state $q$ to $q'$ of a \SWA,
+%
+In short, a transition rule $q \xrightarrow{\phi} q'$, from state $q$ to $q'$,
 \philippe{Je ne comprends pas très ce que cette première phrase fait là. Je commencerais
 directement le paragraphe avec "This framework relies..."}
+\florent{c'est un résumé de la Fig. \ref{fig:hierarchy}}
 %input symbols are variables, %(parameters)
-is labeled by a function $\phi$ associating to every input symbol $a$ a weight value $\phi(a)$
-in a semiring domain.
+is labeled by a function~$\phi$ associating to every input symbol $a$ a weight value $\phi(a)$
+in a semiring domain $\Semiring$.
 %and the weight associated to a transition rule is a function of these variables.
@@
 %but can be an arbitrary commutative semiring (assuming some restrictions).
-%
+
 The framework relies on several language models:

### 2021-07-11 introduction.tex
ex.1 introduction

@@
 are essentially linear:   
-audio files, or the widely used MIDI file format~\cite{?}.
+audio files, or the widely used MIDI file format~\cite{MIDIfile}.
 Such representations ignore the hierarchical structures that frame the
@@
 are present, either explicitly  or implicitly,
-in music notation~\cite{Gould11Notation}: music scores are partitioned in measures, measures
-in beats, and beats can be further recursively divided.
+in Common Western Music Notation~\cite{Gould11Notation}
+%in music notation~\cite{Gould11Notation}: 
+music scores are partitioned in measures, 
+measures in beats, and beats can be further recursively divided.
 It follows that music events do not occur at arbitrary timestamps,
@@
 \begin{example}[Running example]\label{ex:running}
-We
-illustrate our framework with music transcription: given
-a \emph{timeline} of musical events with timestamps from an infinite
-domain as input, parse it
-into a structured music score. Input events
-are pairs $\< \mu, \tau>$, with $\mu \in \Sigma$, where $\Sigma$ stands for the
-set of MIDI message symbols~\cite{?}
+We illustrate our framework with music transcription: given
+a \emph{timeline} of timestamped musical events from an infinite
+alphabet $\Sigma$ as input, 
+parse it into a structured music score. 
+Input events of $\Sigma$
+are pairs $\< \mu, \tau>$, were $\mu$ is a MIDI message symbol~\cite{MIDIfile}
 and $\tau \in \mathbb{Q}$ is a timestamp. Such inputs typically correspond
 to the recording of a live performance. The output of parsing
-is a sequence of events (or \emph{notes}) $\< \nu, \tau'>$   in
-Common Western Music Notation (CWMN)~\cite{Gould11Notation}
-where event symbols $\nu$ belong to the domain $\Delta$
-of \emph{pitches} (e.g., A4, G5, etc.), and timestamp $\tau'$
-belongs to a rhythmic ``grid'' obtained from recursive divisions:
+is a sequence of events (or \emph{notes}) $\< \nu, \tau'>$ 
+in an alphabet $\Delta$
+%in Common Western Music Notation (CWMN)~\cite{Gould11Notation}
+where event symbols $\nu$ are \emph{pitch names} (e.g., A4, G5, \etc.), 
+and timestamp $\tau'$ belongs to a rhythmic ``grid'' obtained from recursive divisions:
 whole notes ($\musWhole$) splitted in halves ($\musHalf$), halves
@@
 \begin{enumerate}
-  \item A performance  $I = [ \<\mu_1, 0.07>, \<\mu_2, 0.72>, \<\mu_3, 0.91>, \<\mu_4, 1.05>, \<\mu_5, 1.36>, \<\mu_6, 1.71>]$,
-     over interval $[0,2[$,
-  \item A score $O =$ \includegraphics[scale=0.20]{pictures/score5.png}, corresponding to
-      a hierarchical structures (measures, tuplets) that can be linearized as a sequence of events
-      $[\<\nu_1,0>, \<\nu_2,\frac{3}{4}>, \<\nu_3,\frac{7}{8}>, \<\nu_4,1>, \<\nu_5,\frac{4}{3}>, \<\nu_6\frac{5}{3}>]$
+  \item A performance  $I = [ \<49, 0.07>, \<51, 0.72>, \<52, 0.91>, \<54, 1.05>, \<56, 1.36>, \<57, 1.71>]$, % over interval $[0,2[$,
+  \item A score $O =$ 
+  \includegraphics[scale=0.35,trim=0 5mm 0 0]{pictures/ex1.pdf},
+  %\includegraphics[scale=0.20]{pictures/score5.png}, 
+  corresponding to a hierarchical structures (measures, tuplets) 
+  that can be linearized as a sequence of events
+      $[\<\mathrm{A}4, 0>, 
+        \<\mathrm{B}4, \frac{3}{4}>, 
+        \<\mathrm{C}5, \frac{7}{8}>, 
+        \<\mathrm{D}5, 1>, 
+        \<\mathrm{E}5,\frac{4}{3}>, 
+        \<\mathrm{F}5\frac{5}{3}>]$
 \end{enumerate}
@@
 parsing of $I$. Note that several other parsings (e.g.,
-\includegraphics[scale=0.20]{pictures/score4.png}) are possible.
+\includegraphics[scale=0.35,trim=0 5mm 0 0]{pictures/ex2.pdf}) are possible.
 SW-parsing associates a weight value

### 2021-07-14 introduction.tex
revisit exemples (ex. sw-VPA pas termine)

@@
 are essentially linear:   
-audio files, or the widely used MIDI file format~\cite{MIDIfile}.
+audio files, or the widely used 
+MIDI format~\cite{Selfridge-Field97beyondMIDI}. %\cite{MIDIfile}.
 Such representations ignore the hierarchical structures that frame the
@@
 
-\begin{example}[Running example]\label{ex:running}
-We illustrate our framework with music transcription: given
-a \emph{timeline} of timestamped musical events from an infinite
-alphabet $\Sigma$ as input, 
-parse it into a structured music score. 
+\begin{example}%[Running example]
+\label{ex:running}
+We illustrate our framework with a very simplified %toy
+running example of \emph{music transcription}: 
+a given \emph{timeline} of timestamped musical events 
+from an infinite alphabet $\Sigma$ as input, 
+is parsed into a structured music score. 
 Input events of $\Sigma$
-are pairs $\< \mu, \tau>$, were $\mu$ is a MIDI message symbol~\cite{MIDIfile}
-and $\tau \in \mathbb{Q}$ is a timestamp. Such inputs typically correspond
-to the recording of a live performance. The output of parsing
-is a sequence of events (or \emph{notes}) $\< \nu, \tau'>$ 
-in an alphabet $\Delta$
+are pairs $\mu \ts{\tau}$, where $\mu$ is a 
+MIDI pitch~\cite{Selfridge-Field97beyondMIDI}, %\cite{MIDIfile}
+and $\tau \in \mathbb{Q}$ is a timestamp in seconds. 
+Such inputs typically correspond
+to the recording of a live performance, \eg
+$I = 69\ts{0.07}, 
+     71\ts{0.72}, 
+     72\ts{0.91}, 
+     74\ts{1.05}, 
+     76\ts{1.36}, 
+     77\ts{1.71}$. % over interval $[0,2[$,
+
+The output of parsing is a sequence of 
+timed symbols $\nu\ts{\tau'}$ in an alphabet $\Delta$,
+%events (or \emph{notes}) 
 %in Common Western Music Notation (CWMN)~\cite{Gould11Notation}
-where event symbols $\nu$ are \emph{pitch names} (e.g., A4, G5, \etc.), 
-and timestamp $\tau'$ belongs to a rhythmic ``grid'' obtained from recursive divisions:
-whole notes ($\musWhole$) splitted in halves ($\musHalf$), halves
-in quarters ($\musQuarter$), eights ($\musEighth$), etc.
-The following concrete sequences
-will be used:
-\begin{enumerate}
-  \item A performance  $I = [ \<49, 0.07>, \<51, 0.72>, \<52, 0.91>, \<54, 1.05>, \<56, 1.36>, \<57, 1.71>]$, % over interval $[0,2[$,
-  \item A score $O =$ 
+where every symbol $\nu$ may represent 
+an \emph{event} (or \emph{note}), 
+specified by its \emph{pitch name} 
+(\eg, $\mbox{\footnotesize A4}$, $\mbox{\footnotesize G5}$, \etc.), 
+an event \emph{continuation} (symbol `$-$`, see Ex.~\ref{ex:SWT}),
+or a \emph{markup}. %(\emph{parenthesis}),
+The timestamp $\tau'$, expressed in number of \emph{measures},
+belongs to a rhythmic ``grid'' obtained from recursive divisions:
+whole notes ($\musWhole$) split in halves ($\musHalf$), halves
+in quarters ($\musQuarter$), eights ($\musEighth$), \etc.
+%
+For instance, the output score 
 \includegraphics[scale=0.35,trim=0 5mm 0 0]{pictures/ex1.pdf},
   %\includegraphics[scale=0.20]{pictures/score5.png}, 
-  corresponding to a hierarchical structures (measures, tuplets) 
-  that can be linearized as a sequence of events
-      $[\<\mathrm{A}4, 0>, 
-        \<\mathrm{B}4, \frac{3}{4}>, 
-        \<\mathrm{C}5, \frac{7}{8}>, 
-        \<\mathrm{D}5, 1>, 
-        \<\mathrm{E}5,\frac{4}{3}>, 
-        \<\mathrm{F}5\frac{5}{3}>]$
-\end{enumerate}
+corresponds to a hierarchical structure 
+(measures, tuplets..., see Figure~\ref{fig:score-tree}) 
+that can be linearized as the sequence
+$O =$ 
+$\ccall{\mathsf{m}}\ts{0}$,
+$\ccall{2}\ts{0}$,
+$\mbox{\footnotesize A4}\ts{0}$, 
+$\ccall{2}\ts{\frac{1}{2}}$,
+$-\ts{\frac{1}{2}}$,
+$\ccall{2}\ts{\frac{3}{4}}$,
+$\mbox{\footnotesize B4}\ts{\frac{3}{4}}$, 
+$\mbox{\footnotesize C$\sharp$5}\ts{\frac{7}{8}}$, 
+$\creturn{2}\ts{1}$,
+$\creturn{2}\ts{1}$,
+$\creturn{2}\ts{1}$,
+$\ccall{\mathsf{m}}\ts{1}$,
+$\ccall{3}\ts{1}$,
+$\mbox{\footnotesize D5}\ts{1}$, 
+$\mbox{\footnotesize E5}\ts{\frac{4}{3}}$, 
+$\mbox{\footnotesize F5}\ts{\frac{5}{3}}$,
+$\creturn{3}\ts{2}$,
+$\creturn{\mathsf{m}}\ts{2}$,
+$\creturn{\mathsf{m}}\ts{2}$.
+%\end{enumerate}
+The opening markups $\ccall{\mathsf{m}}$ %and $\creturn{\mathsf{m}}$
+delimit \emph{measures}, 
+which are time intervals of duration 1 in this example, 
+while the subsequences of $O$ between markups 
+$\ccall{d}$ and $\creturn{d}$,
+for some natural number $d$, 
+represent a division of the current 
+time interval into $d$ sub-intervals of equal length.
+
 \smallskip
-
-We will show that $O$ is a soution for the
+We will show that $O$ is a solution for the
 parsing of $I$. Note that several other parsings (e.g.,

### 2021-07-14 main.tex
ex.1 introduction
revisit exemples (ex. sw-VPA pas termine)

@@
 \newcommand{\lin}{\mathsf{lin}}
+\newcommand{\ts}[1]{\mathop{:}{#1}} % timestamp marker
+%\newcommand{\ts}[1]{^{#1}} % timestamp marker
 
@@
 \newcommand{\ccall}[1]{\ensuremath \langle_{#1}}
-\newcommand{\creturn}[1]{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
+\newcommand{\creturn}[1]{\ensuremath \rangle_{#1}} 
+%\newcommand{\creturn}[1]{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
 
@@
 \begin{example}\label{ex:SWT}
-We build a small weighted transducer model
+We build a small \SWT over the alphabets $\Sigma$ and $\Deltai$ of 
+Examples~\ref{ex:running} and \ref{distance-time},
+%weighted transducer model
 with two states $q_0$ and $q_1$ that
-calculates the distance between the performance $I$ and
-the score $O$ of Example~\ref{ex:running}.
-
-If one performed event $\mu_i$  corresponds
-to one notated event $\nu_j$ (for instance MIDI code 61 and pitch A4),
+calculates the temporal distance between an input performance in~$\Sigma^*$
+and the subsequence 
+of the events of $\Deltai$ in a score.
+%the performance $I$ and
+%the score $O$ of Example~\ref{ex:running}.
+%
+When reading one performed event $\mu$ 
+and the corresponding notated event $\nu$ 
+(for instance MIDI pitch 69 and note name $\mbox{\footnotesize A4}$),
 the weight value computed by the \SWT is the time distance between both,
-(see Example~\ref{distance-time}), and is modeled by
-transitions $\wei_{11}$ below.
-%
-If we meet the music notation symbol '-' that
-represents continuation (this occurs for instance in \emph{ties}
+(see Example~\ref{distance-time}), as modeled by transitions $\wei_{11}$ below.
+%
+If we meet the continuation symbol $'-'$ 
+(this occurs for instance in \emph{ties}
 $\musQuarter\!\!\!\mathrel{\raisebox{-1.5mm}{$\smile$}}\!\!\!\musEighth$,
-or \emph{dots} \musQuarterDotted{}), it is  skipped with no cost (transitions $\wei_{01}$ or weight $\one$).
-%transitions
+or \emph{dots} \musQuarterDotted{}),
+%that represents continuation 
+it is  skipped with no cost (transitions $\wei_{01}$ or weight $\one$).
+%Likewise, the markups are ignored.
 \[
-\begin{array}{rclcrcl}
-\wei_{11}(q_0, \<\mu, \tau>, \<\nu, \tau'>, q_0) & = & |\tau' - \tau| & \quad &
-\wei_{11}(q_1, \<\mu, \tau>, \<\nu, \tau'>, q_0) & = & |\tau' - \tau|\\
-\wei_{01}(q_0, \varepsilon, \< \mathsf{-}, \tau'>, q_0) & = & \one & &
-\wei_{01}(q_1, \varepsilon, \< \mathsf{-}, \tau'>, q_0) & = & \one\\
-\wei_{10}(q_0, \<\mu, \tau>, \varepsilon, q_1) & = & \alpha & & %\multicolumn{3}{l}{\mathrm{for~all~} b \in \Delta}
+\begin{array}{lclclcll}
+\wei_{11}(q_0, \mu\ts{\tau}, \nu\ts{\tau'}, q_0) & = & 
+     \delta(\mu\ts{\tau}, \nu\ts{\tau'}) & \quad & %|\tau' - \tau| & \quad &
+\wei_{11}(q_1, \mu\ts{\tau}, \nu\ts{\tau'}, q_0) & = & 
+     \delta(\mu\ts{\tau}, \nu\ts{\tau'}) & 
+     \mbox{if~} \nu \neq - \\ %|\tau' - \tau| & \quad &
+\wei_{01}(q_0, \varepsilon, \mathsf{-}\ts{\tau'}, q_0) & = & \one & &
+\wei_{01}(q_1, \varepsilon, \mathsf{-}\ts{\tau'}, q_0) & = & \one\\
+\wei_{10}(q_0, \mu\ts{\tau}, \varepsilon, q_1) & = & \alpha & & %\multicolumn{3}{l}{\mathrm{for~all~} b \in \Delta}
 \end{array}
@@
 %
-We also want to take performing errors into account, while still being able to compare with the score,
+We also want to take performing errors into account, 
+while still being able to compare with the score,
 since a performer could, for example, play an unwritten extra note.
-\lydia{reformulated this sentence}
-%
-This is modelled by the transition $\wei_{10}$ with an arbitrary weight value $\alpha \in \Semiring$,
-switching from state $q_0$ (normal) to $q_1$ (error).
+%\lydia{reformulated this sentence}
+%
+%Reading such a note is modelled by 
+The transition $\wei_{10}$, 
+with an fixed weight value $\alpha \in \Semiring$,
+switches from state $q_0$ (normal) to $q_1$ (error)
+when reading an extra note $\mu$.
 \philippe{Comprends pas cette phrase}
+\florent{$\to$ reformulation}
 The transitions in the second column below switch back to the normal state $q_0$.
@@
 $\init(q_1) = \final(q_1) = \zero$.
+\florent{dernière phrase utile?}
 This \SWT is capable of evaluating the differences between a score and a performance,
@@
 \begin{example}\label{ex:nested-word}
-Consider once more the scores of Example~\ref{ex:running}.
-$\Deltai$ represents the set of music events of the form $<\nu, \tau>$
-where $\nu$ is either a note (e.g. A4), the continuation symbol $-$, or a
-rest symbol R. $\Deltac$ and  $\Deltar$ are used to encode the score structure:
-Let $\Deltac = \{\langle, \prec\}$ and  $\Deltar = \{\rangle, \succ\}$ where $\langle$
-(resp. $\rangle$)
-and $\prec$ (resp. $\succ$) correspond to start/end of measures, and start/end of
-tuplets. Then  \includegraphics[scale=0.20]{pictures/score5.png}
-is encoded with the nested word $\langle \prec 0, \frac{3}{4}, \frac{7}{8} \succ \rangle \langle \prec 1, \frac{4}{3}, \frac{5}{3}  \succ \rangle$
+Consider once more the linearized score of Example~\ref{ex:running},
+with $\Deltai$ representing timed notes and continuations.
+%the set of music events of the form $<\nu, \tau>$
+%where $\nu$ is either a note (e.g. A4), the continuation symbol $-$, or a
+%rest symbol R. 
+The subset $\Delta \setminus \Deltai$ is partitioned into 
+$\Deltac$ and $\Deltar$, containing respectively opening and closing parentheses.
+%are used to encode the score structure:
+%Let $\Deltac = \{\langle, \prec\}$ and  $\Deltar = \{\rangle, \succ\}$ where $\langle$
+%(resp. $\rangle$)
+%and $\prec$ (resp. $\succ$) correspond to start/end of measures, and start/end of
+%tuplets. Then  \includegraphics[scale=0.20]{pictures/score5.png}
+%is encoded with the nested word $\langle \prec 0, \frac{3}{4}, \frac{7}{8} \succ \rangle %\langle \prec 1, \frac{4}{3}, \frac{5}{3}  \succ \rangle$
+The score representation $O$ in Example~\ref{ex:running}
+% score: \includegraphics[scale=0.35,trim=0 5mm 0 0]{pictures/ex1.pdf},
+is a nested-word of~$\Delta^*$. 
+%
+Another example is the second candidate 
+\includegraphics[scale=0.35,trim=0 5mm 0 0]{pictures/ex2.pdf}
+of transcription of~$I$,
+linearized into
+$O' =$ 
+$\ccall{\mathsf{m}}\ts{0}$,
+$\ccall{2}\ts{0}$,
+$\mbox{\footnotesize A4}\ts{0}$, 
+$\ccall{2}\ts{\frac{1}{2}}$,
+$-\ts{\frac{1}{2}}$,
+$\mbox{\footnotesize B4}\ts{\frac{3}{4}}$, 
+$\creturn{2}\ts{1}$,
+$\creturn{2}\ts{1}$,
+$\ccall{\mathsf{m}}\ts{1}$,
+$\ccall{3}\ts{1}$,
+$\mbox{\footnotesize `C$\sharp$5`}\ts{1}$, 
+$\mbox{\footnotesize D5}\ts{1}$, 
+$\mbox{\footnotesize E5}\ts{\frac{4}{3}}$, 
+$\mbox{\footnotesize F5}\ts{\frac{5}{3}}$,
+$\creturn{3}\ts{2}$,
+$\creturn{\mathsf{m}}\ts{2}$,
+$\creturn{\mathsf{m}}\ts{2}$
+(see also Figure~\ref{fig:score-tree}).
+The symbol between quotes $\mbox{\footnotesize `C$\sharp$5`}$
+represents an \emph{appogiatura}, %(grace-note),
+that is an ornemental note with theoretical duration zero.
 \endex
@@
 \weiec: Q \times [\Deltac \times P] \times Q \to \Semiring &
-\weiec(q, {\call{c}}, p, q') = \phi_\mathsf{c}({\call{c}}) &
+\weiec(q, {\call{c}}, p', q') = \phi_\mathsf{c}({\call{c}}) &
 \mathrm{where~} \phi_\mathsf{c} = \weiec(q, p, q').\\
@@
 of~$\bar\wei$ are the constant~$\zero$.
+
+\begin{example}
+... \SWVPA over the alphabet of Example~\ref{ex:nested-word}
+that expresses a weight related to the music notation,
+more precisely to its structural complexity: 
+given a set of equivalent representations, we aim at choosing the simpler one. 
+...
+
+*** SUPPR ***
+For instance
+\includegraphics[scale=0.20]{pictures/score5.png} will be favored
+on \includegraphics[scale=0.20]{pictures/score4.png}. 
+%
+Costs  can be expressed
+in the grammar as weights associated to each production rule.
+For instance, the following production rules define two possible divisions
+of a time interval into respectively a duplet and a triplet. Each comes with
+a specific cost.
+\[
+\rho_1: q_0 \xrightarrow{0.06} \< q_{1}, q_{2}>,\
+\rho_2: q_0 \xrightarrow{0.12} \< q_{1}, q_{2}, q_{2}>.
+\]
+Further binary divisions of time sub-intervals %represented respectively by~$q_2$ and~$q_3$
+are possible with:
+\[
+\rho_3:\,q_2 \xrightarrow{0.1} \< q_{3}, q_{3}>, \
+\rho_4:\,q_3 \xrightarrow{0.11} \< q_{4}, q_{4}>. \
+\]
+
+Score \includegraphics[scale=0.20]{pictures/score5.png} is obtained as follows:
+the first measure
+results from the successive applications of  rules $\rho_1$ (division in two)
+and rules $\rho_3$ (division of the second half) in two. The second measure
+is a division in three obtained by rule $\rho_3$. One obtains a parse tree
+that can be linearized in the nested word of Example~\ref{ex:nested-word}.
+...
+\end{example}
 
@@
 is interpreted as follows:
-\begin{enumerate}
-
+%
+\begin{itemize}
+\item The input word is $I$ of Example~\ref{ex:running}.
 \item The \SWT $T$ evaluates a ``fitness measure''  that expresses
-    a correspondance between a performance and its notation. See Example~\ref{ex:SWT}.
-\item The \SWVPA $A$ expresses a cost related to the music notation. One possibility
- is to relate this cost to the structural complexity: given a set of equivalent
- representations, we aim at choosing the simpler one. For instance
- \includegraphics[scale=0.20]{pictures/score5.png} should be favored
+    a correspondance between a performance and notation events. See Example~\ref{ex:SWT}.
+\item The \SWVPA $A$ expresses a cost related to the music notation. 
+\end{itemize}
+One possibility is to relate this cost to the structural complexity: 
+given a set of equivalent representations, 
+we aim at choosing the simpler one. 
+% 
+For instance
+\includegraphics[scale=0.20]{pictures/score5.png} will be favored
 on \includegraphics[scale=0.20]{pictures/score4.png}. Costs  can be expressed
@@
 that can be linearized in the nested word of Example~\ref{ex:nested-word}.
- \end{enumerate}
+
 Therefore, the framework, applied to the transcription problem, allows to find an optimal

### 2021-07-15 examples.tex
backup examples in commit 422547a

% backup of examples
% commit 422547a 10/07/2021 by Philippe
% https://github.com/cedric-cnam/WSVPA/commit/422547aa703906e8234eddc51968cfd618400c4f

% introduction.tex

\begin{example}[Running example]\label{ex:running}
We
illustrate our framework with music transcription: given
a \emph{timeline} of musical events with timestamps from an infinite
domain as input, parse it
into a structured music score. Input events
are pairs $\< \mu, \tau>$, with $\mu \in \Sigma$, where $\Sigma$ stands for the
set of MIDI message symbols~\cite{?}
and  $\tau \in \mathbb{Q}$ is a timestamp. Such inputs typically correspond
to the recording of a live performance. The output of parsing
is a sequence of events (or \emph{notes}) $\< \nu, \tau'>$   in
Common Western Music Notation (CWMN)~\cite{Gould11Notation}
where event symbols $\nu$ belong to the domain $\Delta$
of \emph{pitches} (e.g., A4, G5, etc.), and timestamp $\tau'$
belongs to a rhythmic ``grid'' obtained from recursive divisions:
whole notes ($\musWhole$) splitted in halves ($\musHalf$), halves
in quarters ($\musQuarter$), eights ($\musEighth$), etc.
The following concrete sequences
will be used:
\begin{enumerate}
  \item A performance  $I = [ \<\mu_1, 0.07>, \<\mu_2, 0.72>, \<\mu_3, 0.91>, \<\mu_4, 1.05>, \<\mu_5, 1.36>, \<\mu_6, 1.71>]$,
     over interval $[0,2[$,
  \item A score $O =$ \includegraphics[scale=0.20]{pictures/score5.png}, corresponding to
      a hierarchical structures (measures, tuplets) that can be linearized as a sequence of events
      $[\<\nu_1,0>, \<\nu_2,\frac{3}{4}>, \<\nu_3,\frac{7}{8}>, \<\nu_4,1>, \<\nu_5,\frac{4}{3}>, \<\nu_6\frac{5}{3}>]$
\end{enumerate}
\smallskip

We will show that $O$ is a soution for the
parsing of $I$. Note that several other parsings (e.g.,
\includegraphics[scale=0.20]{pictures/score4.png}) are possible.
SW-parsing associates a weight value
to each solution, and our framework
aims at selecting the best one with respect to this weight.
\endex
\end{example}

% labels.tex

\begin{example}\label{distance-time}
We return to the music transcription problem (Exemple~\ref{ex:running}). In order to align the
input with a music score, we must take into consideration
the expressive timing of human performance that
results in small time shifts between an input event and the corresponding
notation event.  These shifts can be weighted as the time distance between both,
computed in the tropical semiring as follows.
$$\delta\bigl(\<\mu, \tau>, \<\nu, \tau'>\bigr) =
\left\{
\begin{array}{ll}
   | \tau_1 - \tau_2 | & \mathrm{if}\  \mu \rm{\ corresponds\  to\ } \nu \\
   \zero  & \mathrm{otherwise}
\end{array}
\right.$$
The performance is
$I = [ \<\mu_1, 0.07>, \<\mu_2, 0.72>, \<\mu_3, 0.91>, \<\mu_4, 1.05>, \<\mu_5, 1.36>, \<\mu_6, 1.71>]$,
and the (linearized) score is
$[\<\nu_1,0>, \<\nu_2,\frac{3}{4}>, \<\nu_3,\frac{7}{8}>, \<\nu_4,1>, \<\nu_5,\frac{4}{3}>, \<\nu_6\frac{5}{3}>]$
Assuming the pairwise correspondence of MIDI symbols
$\mu_i$ and notation symbols $\nu_i$, for $i \in [1, 6]$,
the distance between $I$ and $O$ is the  aggregation with $\otimes$
of the pairwise differences between the
timestamps. In the tropical semiring, this yields
$|0.07 - 0| + |0.72 - \frac{3}{4}| + |0.91- \frac{7}{8} | +
|1.05-1| + |1.36-\frac{4}{3}| + |1.71-\frac{5}{3}|= 0.255$.
\endex
\end{example}

% main.tex

\begin{example}\label{ex:SWT}
We build a small weighted transducer model
with two states $q_0$ and $q_1$ that
calculates the distance between the performance $I$ and
the score $O$ of Example~\ref{ex:running}.

If one performed event $\mu_i$  corresponds
to one notated event $\nu_j$ (for instance MIDI code 61 and pitch A4),
the weight value computed by the \SWT is the time distance between both,
(see Example~\ref{distance-time}), and is modeled by
transitions $\wei_{11}$ below.
%
If we meet the music notation symbol '-' that
represents continuation (this occurs for instance in \emph{ties}
$\musQuarter\!\!\!\mathrel{\raisebox{-1.5mm}{$\smile$}}\!\!\!\musEighth$,
or \emph{dots} \musQuarterDotted{}), it is  skipped with no cost (transitions $\wei_{01}$ or weight $\one$).
%transitions
\[
\begin{array}{rclcrcl}
\wei_{11}(q_0, \<\mu, \tau>, \<\nu, \tau'>, q_0) & = & |\tau' - \tau| & \quad &
\wei_{11}(q_1, \<\mu, \tau>, \<\nu, \tau'>, q_0) & = & |\tau' - \tau|\\
\wei_{01}(q_0, \varepsilon, \< \mathsf{-}, \tau'>, q_0) & = & \one & &
\wei_{01}(q_1, \varepsilon, \< \mathsf{-}, \tau'>, q_0) & = & \one\\
\wei_{10}(q_0, \<\mu, \tau>, \varepsilon, q_1) & = & \alpha & & %\multicolumn{3}{l}{\mathrm{for~all~} b \in \Delta}
\end{array}
\]
%
We also want to take performing errors into account, while still being able to compare with the score,
since a performer could, for example, play an unwritten extra note.
\lydia{reformulated this sentence}
%
This is modelled by the transition $\wei_{10}$ with an arbitrary weight value $\alpha \in \Semiring$,
switching from state $q_0$ (normal) to $q_1$ (error).
\philippe{Comprends pas cette phrase}
The transitions in the second column below switch back to the normal state $q_0$.
% the metric computed by the \SWT is the smallest sum of point wise distances
% between dates of input and output events.
At last, we let $q_0$ be the only initial and final state, with
$\init(q_0) = \final(q_0) = \one$, and
$\init(q_1) = \final(q_1) = \zero$.
This \SWT is capable of evaluating the differences between a score and a performance,
all the while ensuring that performance errors are plausible.
%$\init(q_0, d, b) = \final(q_0, d, b) = \one$, and
%$\init(q_1, d, b) = \final(q_1, d, b) = \zero$,
%for all $d \in \Sigma$ and $b \in \Delta$.
\endex
\end{example}

\begin{example}\label{ex:nested-word}
Consider once more the scores of Example~\ref{ex:running}.
$\Deltai$ represents the set of music events of the form $<\nu, \tau>$
where $\nu$ is either a note (e.g. A4), the continuation symbol $-$, or a
rest symbol R. $\Deltac$ and  $\Deltar$ are used to encode the score structure:
Let $\Deltac = \{\langle, \prec\}$ and  $\Deltar = \{\rangle, \succ\}$ where $\langle$
(resp. $\rangle$)
and $\prec$ (resp. $\succ$) correspond to start/end of measures, and start/end of
tuplets. Then  \includegraphics[scale=0.20]{pictures/score5.png}
is encoded with the nested word $\langle \prec 0, \frac{3}{4}, \frac{7}{8} \succ \rangle \langle \prec 1, \frac{4}{3}, \frac{5}{3}  \succ \rangle$
\endex
\end{example}

\begin{example}[Symbolic Weighted Parsing and the transcription problem]
Applied to the music transcription problem, the above formalism
is interpreted as follows:
\begin{enumerate}
\item The \SWT $T$ evaluates a ``fitness measure''  that expresses
 a correspondance between a performance and its notation. See Example~\ref{ex:SWT}.
\item The \SWVPA $A$ expresses a cost related to the music notation. One possibility
 is to relate this cost to the structural complexity: given a set of equivalent
 representations, we aim at choosing the simpler one. For instance
 \includegraphics[scale=0.20]{pictures/score5.png} should be favored
 on \includegraphics[scale=0.20]{pictures/score4.png}. Costs  can be expressed
 in the grammar as weights associated to each production rule.
 For instance, the following production rules define two possible divisions
 of a time interval into respectively a duplet and a triplet. Each comes with
 a specific cost.
 \[
 \rho_1: q_0 \xrightarrow{0.06} \< q_{1}, q_{2}>,\
 \rho_2: q_0 \xrightarrow{0.12} \< q_{1}, q_{2}, q_{2}>.
 \]
 Further binary divisions of time sub-intervals %represented respectively by~$q_2$ and~$q_3$
 are possible with:
 \[
 \rho_3:\,q_2 \xrightarrow{0.1} \< q_{3}, q_{3}>, \
 \rho_4:\,q_3 \xrightarrow{0.11} \< q_{4}, q_{4}>. \
 \]

Score \includegraphics[scale=0.20]{pictures/score5.png} is obtained as follows:
the first measure
 results from the successive applications of  rules $\rho_1$ (division in two)
 and rules $\rho_3$ (division of the second half) in two. The second measure
 is a division in three obtained by rule $\rho_3$. One obtains a parse tree
 that can be linearized in the nested word of Example~\ref{ex:nested-word}.
 \end{enumerate}
Therefore, the framework, applied to the transcription problem, allows to find an optimal
solution that considers both the fitness of the result, and its structural complexity.
\end{example}

\begin{example}[Symbolic Weighted Parsing and the transcription problem]
Applied to the music transcription problem, the above formalism
is interpreted as follows:
\begin{enumerate}

  \item The \SWT $T$ evaluates a ``fitness measure''  that expresses
    a correspondance between a performance and its notation. See Example~\ref{ex:SWT}.
\item The \SWVPA $A$ expresses a cost related to the music notation. One possibility
 is to relate this cost to the structural complexity: given a set of equivalent
 representations, we aim at choosing the simpler one. For instance
 \includegraphics[scale=0.20]{pictures/score5.png} should be favored
 on \includegraphics[scale=0.20]{pictures/score4.png}. Costs  can be expressed
 in the grammar as weights associated to each production rule.
 For instance, the following production rules define two possible divisions
 of a time interval into respectively a duplet and a triplet. Each comes with
 a specific cost.
 \[
 \rho_1: q_0 \xrightarrow{0.06} \< q_{1}, q_{2}>,\
 \rho_2: q_0 \xrightarrow{0.12} \< q_{1}, q_{2}, q_{2}>.
 \]
 Further binary divisions of time sub-intervals %represented respectively by~$q_2$ and~$q_3$
 are possible with:
 \[
 \rho_3:\,q_2 \xrightarrow{0.1} \< q_{3}, q_{3}>, \
 \rho_4:\,q_3 \xrightarrow{0.11} \< q_{4}, q_{4}>. \
 \]

Score \includegraphics[scale=0.20]{pictures/score5.png} is obtained as follows:
the first measure
 results from the successive applications of  rules $\rho_1$ (division in two)
 and rules $\rho_3$ (division of the second half) in two. The second measure
 is a division in three obtained by rule $\rho_3$. One obtains a parse tree
 that can be linearized in the nested word of Example~\ref{ex:nested-word}.
 \end{enumerate}
Therefore, the framework, applied to the transcription problem, allows to find an optimal
solution that considers both the fitness of the result, and its structural complexity.
\end{example}

### 2021-07-15 introduction.tex
2 last examples (VPA and parsing)

@@
      71\ts{0.72}, 
-     72\ts{0.91}, 
+     73\ts{0.91}, 
      74\ts{1.05}, 

### 2021-07-15 main.tex
2 last examples (VPA and parsing)

@@
 
-\begin{example}
-... \SWVPA over the alphabet of Example~\ref{ex:nested-word}
+\begin{example} \label{ex:SWVPA}
+We consider a \SWVPA over the alphabet of Example~\ref{ex:nested-word}
 that expresses a weight related to the music notation,
@@
 given a set of equivalent representations, we aim at choosing the simpler one. 
-...
-
-*** SUPPR ***
-For instance
-\includegraphics[scale=0.20]{pictures/score5.png} will be favored
-on \includegraphics[scale=0.20]{pictures/score4.png}. 
-%
-Costs  can be expressed
-in the grammar as weights associated to each production rule.
-For instance, the following production rules define two possible divisions
-of a time interval into respectively a duplet and a triplet. Each comes with
-a specific cost.
-\[
-\rho_1: q_0 \xrightarrow{0.06} \< q_{1}, q_{2}>,\
-\rho_2: q_0 \xrightarrow{0.12} \< q_{1}, q_{2}, q_{2}>.
-\]
-Further binary divisions of time sub-intervals %represented respectively by~$q_2$ and~$q_3$
-are possible with:
-\[
-\rho_3:\,q_2 \xrightarrow{0.1} \< q_{3}, q_{3}>, \
-\rho_4:\,q_3 \xrightarrow{0.11} \< q_{4}, q_{4}>. \
-\]
-
-Score \includegraphics[scale=0.20]{pictures/score5.png} is obtained as follows:
-the first measure
-results from the successive applications of  rules $\rho_1$ (division in two)
-and rules $\rho_3$ (division of the second half) in two. The second measure
-is a division in three obtained by rule $\rho_3$. One obtains a parse tree
-that can be linearized in the nested word of Example~\ref{ex:nested-word}.
-...
+This si specified by the weights associated to the transition.
+
+Let us assume a third component in the symbols of $\Deltac$ 
+which is the duration of the associated time interval.
+It is used to check the timestamps in the symbols.
+%
+For instance, the following call transition starts in state
+$q_{{i}/{c}}$, meaning that the current interval is number~$i$ 
+amongst $c$ sub-intervals of duration~$\ell$ (as indicated by the stack top).
+It reads a new time-division symbol~$\ccall{d}$.
+The date for this symbol has to be the time $\tau$ 
+for the previous time-division symbol read~$\ccall{c}$ (found on the stack)
+plus the total duration of intervals read so far, $i \ell$.
+The new current interval duration is~$\frac{\ell}{d}$.
+
+\(
+\weic\bigl(q_{{i}/{c}}, 
+           [\ccall{c}\ts{\tau}\ts{\ell}], q, 
+           [\ccall{d}\ts{\tau + i\ell}\ts{\frac{\ell}{d}}], q_{{i+1}/{c}},
+           q_{{0}/{d}} \bigr) = \alpha_d
+\).
+
+\noindent
+The weight of this transition is $\alpha_d$. 
+We can penalize triplets compared to duplets with $\alpha_2 < \alpha_3$.
+%
+\noindent 
+Along with $\ccall{d}$, 
+the above transition pushes
+the state $q_{{i+1}/{c}}$ on the stack,  
+to start the next sub-interval after reading $\creturn{d}$, 
+with:
+\(
+\weir\bigl(q_{{d}/{d}}, 
+           [\ccall{d}\ts{\tau}\ts{\ell}], q_{{i}/{c}}, 
+           [\,\creturn{d}\ts{\tau + d\ell}], q_{{i}/{c}} \bigr) = \one
+\).
+
+\noindent
+Reading an event is done with:
+\(
+\weii\bigl(q_{{i}/{d}}, 
+           [\ccall{d}\ts{\tau}\ts{\ell}], q_{{j}/{c}}, 
+           [\mu\ts{\tau + i\ell}], q_{{i+1}/{d}} \bigr) = \alpha_\mu
+\).
+% cas des barres de mesure?
+\endex
 \end{example}
@@
 \end{itemize}
-One possibility is to relate this cost to the structural complexity: 
-given a set of equivalent representations, 
-we aim at choosing the simpler one. 
-% 
-For instance
-\includegraphics[scale=0.20]{pictures/score5.png} will be favored
-on \includegraphics[scale=0.20]{pictures/score4.png}. Costs  can be expressed
-in the grammar as weights associated to each production rule.
-For instance, the following production rules define two possible divisions
-of a time interval into respectively a duplet and a triplet. Each comes with
-a specific cost.
-\[
-\rho_1: q_0 \xrightarrow{0.06} \< q_{1}, q_{2}>,\
-\rho_2: q_0 \xrightarrow{0.12} \< q_{1}, q_{2}, q_{2}>.
-\]
-Further binary divisions of time sub-intervals %represented respectively by~$q_2$ and~$q_3$
-are possible with:
-\[
-\rho_3:\,q_2 \xrightarrow{0.1} \< q_{3}, q_{3}>, \
-\rho_4:\,q_3 \xrightarrow{0.11} \< q_{4}, q_{4}>. \
-\]
-
-Score \includegraphics[scale=0.20]{pictures/score5.png} is obtained as follows:
-the first measure
-results from the successive applications of  rules $\rho_1$ (division in two)
-and rules $\rho_3$ (division of the second half) in two. The second measure
-is a division in three obtained by rule $\rho_3$. One obtains a parse tree
-that can be linearized in the nested word of Example~\ref{ex:nested-word}.
-
-Therefore, the framework, applied to the transcription problem, allows to find an optimal
-solution that considers both the fitness of the result, and its structural complexity.
+%One possibility is to relate this cost to the structural complexity: 
+%given a set of equivalent representations, 
+%we aim at choosing the simpler one. 
+% 
+As seen in Example~\ref{ex:SWVPA}, 
+\includegraphics[scale=0.35,trim=0 5mm 0 0]{pictures/ex1.pdf},
+%\includegraphics[scale=0.20]{pictures/score5.png} 
+will be favored on 
+\includegraphics[scale=0.35,trim=0 5mm 0 0]{pictures/ex2.pdf}
+%\includegraphics[scale=0.20]{pictures/score4.png}. 
+when the weight a second time division with $\ccall{2}$ is less
+than the difference of weight between 
+$\mbox{`\footnotesize C$\sharp$5`}$
+and $\mbox{\footnotesize C$\sharp$5}$.
+
+\noindent
+The SW-parsing framework, applied to the transcription problem, 
+allows to find an optimal solution 
+that considers both the fitness of the result, 
+and its structural complexity.
 \end{example}

### 2021-07-16 main.tex
example VPA: measure

@@
 % command todo : options
-% - noline: no line connecting the note with the place in the text where the note occurs in the latex code
+% - noline: no line connecting the note with the place in the text 
+%   where the note occurs in the latex code
 % - fancyline : curved arrow from note to inssertion point
@@
 For instance, the following call transition starts in state
-$q_{{i}/{c}}$, meaning that the current interval is number~$i$ 
+$q_{{i}/{c}}$, meaning that the current interval has number~$i$ 
 amongst $c$ sub-intervals of duration~$\ell$ (as indicated by the stack top).
@@
 The new current interval duration is~$\frac{\ell}{d}$.
-
+%
 \(
 \weic\bigl(q_{{i}/{c}}, 
-           [\ccall{c}\ts{\tau}\ts{\ell}], q, 
-           [\ccall{d}\ts{\tau + i\ell}\ts{\frac{\ell}{d}}], q_{{i+1}/{c}},
+           \ccall{c}\ts{\tau}\ts{\ell}, q, 
+           \ccall{d}\ts{\tau + i\ell}\ts{\frac{\ell}{d}}, q_{{i+1}/{c}},
            q_{{0}/{d}} \bigr) = \alpha_d
@@
 % cas des barres de mesure?
+** case of new measure symbol: $q_0$ is a special (initial) state.
+first measure:
+\(
+\weiec\bigl(q_0, 
+           \ccall{\mathsf{m}}\ts{0}\ts{1}, q_0, 
+           q_0 \bigr) = \alpha_d
+\).
+new measure:
+\(
+\weic\bigl(q_{{c}/{c}}, 
+           \ccall{\mathsf{m}}\ts{\tau}\ts{1}, q_0, 
+           \ccall{\mathsf{m}}\ts{\tau + 1}\ts{1}, q_0,
+           q_0 \bigr) = \alpha_d
+\).
 \endex

### 2021-07-16 main.tex
example VPA: new measure

@@
 for the previous time-division symbol read~$\ccall{c}$ (found on the stack)
-plus the total duration of intervals read so far, $i \ell$.
-The new current interval duration is~$\frac{\ell}{d}$.
+plus the total duration of intervals read so far, $i \ell$, and
+the new current interval duration is~$\frac{\ell}{d}$:
 %
@@
 The weight of this transition is $\alpha_d$. 
-We can penalize triplets compared to duplets with $\alpha_2 < \alpha_3$.
+We can penalize \eg triplets compared to duplets with $\alpha_2 < \alpha_3$.
 %
@@
 \weir\bigl(q_{{d}/{d}}, 
-           [\ccall{d}\ts{\tau}\ts{\ell}], q_{{i}/{c}}, 
-           [\,\creturn{d}\ts{\tau + d\ell}], q_{{i}/{c}} \bigr) = \one
+           \ccall{d}\ts{\tau}\ts{\ell}, q_{{i}/{c}}, 
+           \,\creturn{d}\ts{\tau + d\ell}, q_{{i}/{c}} \bigr) = \one
 \).
@@
 \weii\bigl(q_{{i}/{d}}, 
-           [\ccall{d}\ts{\tau}\ts{\ell}], q_{{j}/{c}}, 
-           [\mu\ts{\tau + i\ell}], q_{{i+1}/{d}} \bigr) = \alpha_\mu
+           \ccall{d}\ts{\tau}\ts{\ell}, q_{{j}/{c}}, 
+           \mu\ts{\tau + i\ell}, q_{{i+1}/{d}} \bigr) = \alpha_\mu
 \).
 % cas des barres de mesure?
-** case of new measure symbol: $q_0$ is a special (initial) state.
-first measure:
+%For the case of a new measure symbol,
+%$q_{1/1}$ is a special (initial) state.
+The transition for starting to read the first measure is:
 \(
-\weiec\bigl(q_0, 
-           \ccall{\mathsf{m}}\ts{0}\ts{1}, q_0, 
-           q_0 \bigr) = \alpha_d
-\).
-new measure:
+\weiec\bigl(q_{1/1}, 
+           \ccall{\mathsf{m}}\ts{0}\ts{1}, q_{1/1}, 
+           q_{1/1} \bigr) = \one
+\),
+and for a new measure:
 \(
-\weic\bigl(q_{{c}/{c}}, 
-           \ccall{\mathsf{m}}\ts{\tau}\ts{1}, q_0, 
-           \ccall{\mathsf{m}}\ts{\tau + 1}\ts{1}, q_0,
-           q_0 \bigr) = \alpha_d
+\weic\bigl(q_{1/1}, 
+           \ccall{\mathsf{m}}\ts{\tau}\ts{1}, q_{1/1}, 
+           \ccall{\mathsf{m}}\ts{\tau + 1}\ts{1}, q_{1/1},
+           q_{1/1} \bigr) = \one
 \).

### 2021-07-17 app-trees.tex

@@
 \end{description}
+%
+\begin{example}
+The trees in Figure~\ref{fig:score-tree}
+represent the two scores in Examples~\ref{ex:running},\ref{ex:nested-word}, 
+and their linearization are respectively~$O$ and~$O'$ inn the same examples.
+\end{example}
 
@@
 \end{proof}
+
+\begin{figure}
+\begin{center}
+\begin{tikzpicture} [-,thick]
+\tikzstyle{level 1}=[level distance=9mm,sibling distance=24mm]
+\tikzstyle{level 2}=[level distance=8mm,sibling distance=6mm]
+\tikzstyle{level 3}=[level distance=8mm,sibling distance=6mm]
+\node {$\mathsf{m}$}
+  child { node {$2$}
+    child { node {$\mbox{\footnotesize A4}$} }
+    child { node {$2$}
+      child { node {$-$} }
+      child { node {$2$}
+        child { node {$\mbox{\footnotesize B4}$} }
+        child { node {$\mbox{\footnotesize C$\sharp$5}$} }}}}
+  child { node {$\mathsf{m}$}
+    child { node {$3$}
+      child { node {$\mbox{\footnotesize D5}$} }
+      child { node {$\mbox{\footnotesize E5}$} }
+      child { node {$\mbox{\footnotesize F5}$} }}
+    child { node {$\varepsilon$} }};        
+\end{tikzpicture}
+%
+\quad 
+%
+\begin{tikzpicture} [-,thick]
+\tikzstyle{level 1}=[level distance=9mm,sibling distance=30mm]
+\tikzstyle{level 2}=[level distance=8mm,sibling distance=8mm]
+\tikzstyle{level 3}=[level distance=8mm,sibling distance=8mm]
+\node {$\mathsf{m}$}
+  child { node {$2$}
+    child { node {$\mbox{\footnotesize A4}$} }
+    child { node {$2$}
+      child { node {$-$} }
+      child { node {$\mbox{\footnotesize B4}$} } } }
+  child { node {$\mathsf{m}$}
+    child { node {$3$}
+      child { node {$`\mbox{\footnotesize C$\sharp$5}`\,\mbox{\footnotesize D5}$} }
+%                   {$\begin{array}{c}
+%                     \mbox{\footnotesize C5}\\
+%                     \mbox{\footnotesize D5}
+%                     \end{array}$} }
+      child { node {$\mbox{\footnotesize E5}$} }
+      child { node {$\mbox{\footnotesize F5}$} } }
+    child { node {$\varepsilon$} } };        
+\end{tikzpicture}
+\end{center}
+\caption{Tree representation of scores 
+of Examples~\ref{ex:running},\ref{ex:nested-word}, 
+linearized respectively into~$O$ and~$O'$.}
+\label{fig:score-tree}
+\end{figure}

### 2021-07-17 introduction.tex

@@
 which are time intervals of duration 1 in this example, 
-while the subsequences of $O$ between markups 
-$\ccall{d}$ and $\creturn{d}$,
-for some natural number $d$, 
+while the subsequences of $O$ between markups~$\ccall{d}$ and~$\creturn{d}$,
+for some natural number~$d$, 
 represent a division of the current 
@@
 We will show that $O$ is a solution for the
-parsing of $I$. Note that several other parsings (e.g.,
+parsing of $I$. Note that several other parsings (\eg,
 \includegraphics[scale=0.35,trim=0 5mm 0 0]{pictures/ex2.pdf}) are possible.

### 2021-07-17 introduction.tex
Réduction espaces Phil

@@
 are present, either explicitly  or implicitly,
-in Common Western Music Notation~\cite{Gould11Notation}
+in Common Western Music Notation~\cite{Gould11Notation}.
 %in music notation~\cite{Gould11Notation}:
@@
 
-In this paper, we present a uniform framework for weighted parsing over infinite input alphabets.
+In this paper, we present a framework for weighted parsing over infinite input alphabets.
 It is based on \emph{symbolic weighted} finite states language models (\swM),
@@
 to functions into an arbitrary semiring,
-and generalizing also~\WA, by handling infinite alphabets (see Figure~\ref{fig:hierarchy}).
+and generalizing also~\WA, by handling infinite alphabets (Fig.~\ref{fig:hierarchy}).
 %
 In short, a transition rule $q \xrightarrow{\phi} q'$, from state $q$ to $q'$,
-\philippe{Je ne comprends pas très ce que cette première phrase fait là. Je commencerais
-directement le paragraphe avec "This framework relies..."}
-\florent{c'est un résumé de la Fig. \ref{fig:hierarchy}}
+%\florent{c'est un résumé de la Fig. \ref{fig:hierarchy}}
 %input symbols are variables, %(parameters)
 is labeled by a function~$\phi$ associating to every input symbol $a$ a weight value $\phi(a)$
-in a semiring domain $\Semiring$.
-%and the weight associated to a transition rule is a function of these variables.
-%
+in a semiring $\Semiring$.
+
 \begin{figure}
@@
 The framework relies on several language models:
-finite automata called symbolic-weighted (\SWA),
+symbolic-weighted automata (\SWA),
 transducers (\SWT), and pushdown automata
-with a visibly restriction~\cite{AlurMadhusudan09nested} (\SWVPA).
+with a visibility restriction~\cite{AlurMadhusudan09nested} (\SWVPA).
 A \SWT defines a distance $T(s, t)$ between finite words $s$ and $t$
@@
 A  \SWVPA operates sequentially on \emph{nested words}~\cite{AlurMadhusudan09nested},
-structured with markup symbols (parentheses), and apt at describing linearizations of trees.
-In a parsing context, a \SWVPA $A$ associates a weight value $A(t)$ % \in \Semiring$
+structured with markup symbols (parentheses), and describing linearizations of trees.
+A \SWVPA $A$ associates a weight value $A(t)$ % \in \Semiring$
 to a given nested word $t$, which is itsef the linearization of a weighted AST. %representing a parse tree.
@@
 $T(s, t) \otimes A(t)$, called the distance between $s$ and $A$ in~\cite{Mohri03EDWA},
-\wrt the ranking defined by $\oplus$,
-Like weighted-parsing
+\wrt the ranking defined by $\oplus$Like weighted-parsing
 methods~\cite{Goodman99SemiringParsing,Nederhof03weightedParsing,MorbitzVogler19weighted-parsing},
@@
 
-\smallskip
 We will show that $O$ is a solution for the

### 2021-07-17 main.tex
Réduction espaces Phil

@@
 
-\paragraph*{Semirings}
-\label{section:semiring}\label{sec:semiring}
+%\paragraph*{Semirings}
+%\label{section:semiring}\label{sec:semiring}
 \input{semiring}
 
-\paragraph*{Label Theory}
-\label{section:symbols}
+%\paragraph*{Label Theory}
+%\label{section:symbols}
 \input{labels}
@@
 
-We follow the approach of~\cite{Mohri03EDWA} for the computation of distances,
-between words and languages, using weighted transducers,
+We follow the approach of~\cite{Mohri03EDWA} for the computation of distances
+between words and languages
 and extend it to infinite alphabets.
@@
 %over finite alphabets, see  e.g.~\cite{Mohri03EDWA},
-by labeling each transition with a weight function (instead of a simple weight value),
+by labeling each transition with a weight function 
 that takes the input and output symbols as parameters.
@@
 % The symbolic-weighted transducer
-The \SWT $T$ computes on pairs of words $\< s, t> \in \Sigma^* \times \Delta^*$,
+The \SWT $T$ computes on pairs $\< s, t> \in \Sigma^* \times \Delta^*$,
 $s$ and $t$, being respectively called \emph{input} and \emph{output} word.
@@
 % \noindent
-More precisely, $T$ defines a mapping
-from $\Sigma^* \times \Delta^*$ into~$\Semiring$,
+$T$ is
 based on an intermediate function $\weight_T$
@@
 %
-We recall that, by convention (Section~\ref{sec:semiring}),
+We recall that, by convention (Section~\ref{sec:prelim}),
 an empty sum with $\bigoplus$ is equal to~$\zero$.
 %
-Intuitively, using a transition $\wei_{ij}(q, a, b, q')$ means for $T$:
-when reading respectively $a$ and $b$ at the current positions in the input and output words,
+Intuitively,  a transition $\wei_{ij}(q, a, b, q')$ is interpreted as follows:
+when reading  $a$ and $b$ the input and output words,
 increment the current position in the input word if and only if $i = 1$,
@@
 We also want to take performing errors into account,
-while still being able to compare with the score,
 since a performer could, for example, play an unwritten extra note.
@@
 $\init(q_1) = \final(q_1) = \zero$.
-\florent{dernière phrase utile?}
-This \SWT is capable of evaluating the differences between a score and a performance,
-all the while ensuring that performance errors are plausible.
+%\florent{dernière phrase utile?}
+%This \SWT is capable of evaluating the differences between a score and a performance,
+%all the while ensuring that performance errors are plausible.
 %$\init(q_0, d, b) = \final(q_0, d, b) = \one$, and
@@
 \noindent
-The \emph{Symbolic Weighted Automata} %$A = \< Q, \init, \weight, \final >$
+ \emph{Symbolic Weighted Automata} %$A = \< Q, \init, \weight, \final >$
 %over $\Sigma$,  $\Semiring$ and $\bar\Phi$
-are defined similarly as the transducers of Definition~\ref{def:SWT},
+are defined  as the transducers of Definition~\ref{def:SWT},
 by simply omitting the output symbols.
@@
 \noindent
-The following property will be useful to the approach on
-symbolic weighted parsing presented in Section~\ref{sec:parsing}.
+The following property will be useful for
+symbolic weighted parsing (Section~\ref{sec:parsing}).
 
@@
 and some input word $s \in \Sigma^*$.
-The weight for reading of the input is summarized using the operator
+The weight for reading of the input is obtained with
 $\bigoplus^1_\Sigma$.
 %
-The main difficulty comes from the transitions of $T$
-of the form $\wei_{10}$, who read in input and ignore the output.
+The main difficulty comes from the transitions
+of the form $\wei_{10}$, which read in input and ignore the output.
 Such transition shall be simulated by $\varepsilon$-transitions in $B_{A, T}$,
@@
 given a set of equivalent representations, we aim at choosing the simpler one.
-This si specified by the weights associated to the transition.
 
@@
 \noindent
-We shall now present a procedure for searching, for a~$\SWVPA$ $A$,
+We now present a procedure for searching, for a~$\SWVPA$ $A$,
 a word %$w \in \Delta^*$
@@
 \florent{explication Fig.~\ref{fig:best-update} (update) suivant cas de \eqref{eq:SWVPA-weight}?}
-
 The infinite sums in the updates of $d$ in Algorithm~\ref{algo:Dijkstra},
@@
 since~$\Semiring$ is complete.
-
-Termination: the algorithm performs $2.|Q|^2$ iterations until $P$ is empty,
+The algorithm performs $2.|Q|^2$ iterations until $P$ is empty,
 and each iteration has a time complexity $O(|Q|^2 . |P|)$.
@@
 \begin{proof} (sketch)
-We follow a \emph{Bar-Hillel} construction, for parsing by intersection.
-%
-\noindent
-Let us first extend the \SWT $T$ over $\Sigma$, $\Deltai$
+We follow a \emph{Bar-Hillel} construction for parsing by intersection.
+%
+\noindent
+We first extend the \SWT $T$ over $\Sigma$, $\Deltai$
 into a \SWT $T'$ over $\Sigma$ and $\Delta$
@@
 %
-The transducer $T'$ simply skips every symbol
-$b \in {\Delta} \setminus \Deltai$,
+$T'$ simply skips every symbol
+$b \in {\Delta} \setminus \Deltai$
 by the addition to $T$,
@@
 Then, using Corolary~\ref{cor:epsilon},
-we construct from the input word $s \in \Sigma^*$ and $T'$
+we construct from $s \in \Sigma^*$ and $T'$
 a \SWA $B_{s, T'}$,
@@
 It remains to compute a best nested-word $t \in {\Delta}^*$
-using the best-search procedure of Proposition~\ref{th:best-search}.
+using the procedure of Proposition~\ref{th:best-search}.
 %and convert it into a best tree in $\T_\Delta$ in order to solve SW parsing
@@
 \end{proof}
+
 The $\SW$-parsing generalizes
 the problem of searching the best derivation (AST) of a weighted CF-grammar $G$
-that yields a given input word $w$.
-The latter problem, sometimes called~\emph{weighted parsing},
+that yields a given input word $w$,   called~\emph{weighted parsing},
 (see \eg~\cite{Goodman99SemiringParsing}
- and~\cite{MorbitzVogler19weighted-parsing} for general weighted parsing frameworks)
+ and~\cite{MorbitzVogler19weighted-parsing}).
+Weighted parsing
 corresponds to $\SW$-parsing in the case of finite alphabets,
@@
 %
-
-\florent{In practice,
-the functions in transition might be represented by diagram such as
-Algebraic Decision Diagrams~\cite{Bahar97ADD}.}
-
+%\florent{In practice,
+%the functions in transition might be represented by diagram such as
+%Algebraic Decision Diagrams~\cite{Bahar97ADD}.}
 %\paragraph{Application to Automated Music Transcription.}
@@
 %...
-
-\florent{2 lines Application to Automated Music Transcription:
-implementation $\neq$ but same principle, on-the-fly automata construction
-during best search, for efficiency.}
+%\florent{2 lines Application to Automated Music Transcription:
+%implementation $\neq$ but same principle, on-the-fly automata construction
+%during best search, for efficiency.}
 

### 2021-07-17 semiring.tex
Réduction espaces Phil

@@
 %\florent{The results are established for a general class of semirings. They can be instantiated for concrete cases}
-A \emph{semiring} $\< \Semiring, \oplus, \zero, \otimes, \one>$
+\noindent
+\textbf{Semirings}. A \emph{semiring} $\< \Semiring, \oplus, \zero, \otimes, \one>$
 is a structure with a domain~$\Semiring$,
@@
 binary operators~$\oplus$ and $\otimes$,
-with respective neutral elements $\zero$ and $\one$, and such that:
+with respective neutral elements $\zero$ and $\one$, such that:
 %$\< \mathbb{S}, \oplus, \zero>$ is a commutative monoid
@@
 \emph{non-negative weights} condition in shortest-path algorithms~\cite{Dijkstra59anote}.
-Intuitively, it means that combining elements with $\otimes$ always increase their weight.
+Intuitively, it means that combining elements with~$\otimes$ always increase their weight.
 Formally, it is defined as the property ($i$) below. %of the following lemma.
@@
 %
-Hence, in a superior semiring, % \wrt~$\leq$,
-it holds that %$\one \leq \zero$
+Hence, in a superior semiring,
 for all $x \in \Semiring$, $\one \leq_\oplus x \leq_\oplus \zero$.
@@
 % \emph{negative boundedness}
-\philippe{Ca j'ai pas compris}
 It implies that, when looking for a best path in a graph whose edges
@@
 \end{figure}
-\florent{results of this paper: for semirings commutative, bounded, total and complete}
-
+%\florent{results of this paper: for semirings commutative, bounded, total and complete}
 %%% semiring properties used in paper

### 2021-07-18 introduction.tex
Conclusion + 12 pages

@@
 running example of \emph{music transcription}:
-a given \emph{timeline} of timestamped musical events
+a given \emph{timeline} of musical events
 from an infinite alphabet $\Sigma$ as input,
@@
 %in Common Western Music Notation (CWMN)~\cite{Gould11Notation}
-where every symbol $\nu$ may represent
+where $\nu$  represents
 an \emph{event} (or \emph{note}),
@@
 or a \emph{markup}. %(\emph{parenthesis}),
-The timestamp $\tau'$, expressed in number of \emph{measures},
+The timestamp $\tau'$
 belongs to a rhythmic ``grid'' obtained from recursive divisions:
@@
 corresponds to a hierarchical structure
-(measures, tuplets..., see Figure~\ref{fig:score-tree})
 that can be linearized as the sequence

### 2021-07-18 main.tex
proof prop. product SW-VPA
Conclusion + 12 pages

@@
 \def\weight{\mathsf{weight}}
-\def\wei{\mathsf{w}}
 \def\mei{\mathsf{m}}
@@
 \def\Phicx{{\Phi_\mathsf{cx}}}
+\def\wei{\mathsf{w}}
 \def\weii{{\wei_\mathsf{i}}}
@@
 \noindent
-The intuition is the following for the above transitions.
+%The intuition is the following for the above transitions.
 $\weiei$, $\weiec$, and $\weier$ describe the cases where the stack is empty.
@@
 \begin{proof}
-The construction is similar
-to the case of the Boolean semiring~\cite{dAntonyAlur14SVPDA}.
-\florent{complete proof (app)}
+We do a classical product construction, see Appendix~\ref{sec:product} for details.
+%similar to the case of the Boolean semiring~\cite{dAntonyAlur14SVPDA}.
 \end{proof}
@@
 \section*{Conclusion}
-% summary
-We have introduced weighted language models (SW transducers and visibly pushdown automata)
-computing over infinite alphabets,
+
+We  introduced weighted language models over infinite alphabets,
 and applied them to the problem of parsing
 with infinitely many possible input symbols (typically timed events).
-%handled with suitable language formalisms,
-%
-This approach extends conventional parsing and weighted parsing
+Our approach extends conventional parsing and weighted parsing
 by computing a derivation tree modulo
-a generic distance between words,
+a distance between words
 defined by a SW transducer given in input.
-This enables to consider finer word relationships than strict equality,
-%as in the conventional parsing approach,
-opening possibilities of quantitative analysis via this method.
-
-% discussion
-\noindent
-Ongoing and future work include
-\florent{TODO future work}
-
-\noindent
--- The study of other theoretical properties of SW models,
-such as the extension of the best search algorithm from $1$-best to $n$-best~\cite{Huang05kbest},
+This allows to consider finer word relationships than strict equality.
+
+This work can be extended in several directions. First, the
+the best search algorithm could bbe generalized from $1$-best to $n$-best~\cite{Huang05kbest},
 and to $k$-\emph{closed} semirings~\cite{Mohri02semiring}
-(instead of \emph{bounded}, which corresponds to $0$-\emph{closed}).
-
-\noindent
--- ...there is room to improve the complexity bounds for the algorithms
-... modular approach with oracles ...
-
-\noindent
--- present here an offline algorithm for best search,
-semi-online implementation for AMT (bar-by-bar approach)
-with an on-the-fly automata construction.
+(instead of \emph{bounded}, which corresponds to $0$-\emph{closed}). Second,
+we believe that
+the complexity bounds of the algorithms can be more precisely characterized.
+Finally, the  best search algorithm presented here offline, whereas
+an on-the-fly automata construction would allow for online parsing,
+a suitable feature in the context of applications such as, e.g., automatic music transcription.
 
@@
 
+\section{Proof of Proposition~\ref{prop:SWVPA-product}} \label{sec:product}
+\input{app-product}
+
 \section{Correctness of the Best-Search Algorithm} \label{sec:bestsearch}

### 2021-07-18 semiring.tex
Conclusion + 12 pages

@@
 %
-Intuitively, in the models presented in this paper,
+In the models presented in this paper,
 $\oplus$ selects an optimal value from two given values,
@@
 %and let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a {semiring},
-
 A semiring $\Semiring$ is \emph{commutative} if $\otimes$ is commutative.
@@
 
-In~\cite{Huang08advanceddynamic}, when the property~$(i)$ holds,
-$\Semiring$ is called \emph{superior} \wrt the ordering~$\leq_\oplus$.
-We have seen in the proof of Lemma~\ref{lem:bounded} that it implies that
+In~\cite{Huang08advanceddynamic}, when  property~$(i)$ holds,
+$\Semiring$ is called \emph{superior} \wrt~$\leq_\oplus$.
+It implies (proof of Lemma~\ref{lem:bounded})
+that
 $\one \leq_\oplus x$ for all $x \in \Semiring$.
@@
 %
-Intuitively, from an optimization point of view,
+From an optimization point of view,
 it means that $\one$ is the best value, and $\zero$ the worst.

### 2021-07-18 main.tex
Révision Philippe sur les exemples

@@
 \begin{example}\label{ex:SWT}
-We build a small \SWT over the alphabets $\Sigma$ and $\Deltai$ of
-Examples~\ref{ex:running} and \ref{distance-time},
+We build a small \SWT over  alphabets $\Sigma$ and $\Deltai$
+(Ex.~\ref{ex:running} and \ref{distance-time}),
 %weighted transducer model
-with two states $q_0$ and $q_1$ that
+with two states $q_0$ and $q_1$, that
 calculates the temporal distance between an input performance in~$\Sigma^*$
 and the subsequence
-of the events of $\Deltai$ in a score.
+of $\Deltai$ events in a score.
 %the performance $I$ and
@@
 %
-When reading one performed event $\mu$
+Given  a performed event $\mu$
 and the corresponding notated event $\nu$
-(for instance MIDI pitch 69 and note name $\mbox{\footnotesize A4}$),
-the weight value computed by the \SWT is the time distance between both,
-(see Example~\ref{distance-time}), as modeled by transitions $\wei_{11}$ below.
-%
-If we meet the continuation symbol $'-'$
-(this occurs for instance in \emph{ties}
+(e.g., MIDI pitch 69 and note $\mbox{\footnotesize A4}$),
+the weight  computed by the \SWT is the time distance between both,
+as modeled by transitions $\wei_{11}$ below.
+%
+The continuation symbol $'-'$
+(met for instance in \emph{ties}
 $\musQuarter\!\!\!\mathrel{\raisebox{-1.5mm}{$\smile$}}\!\!\!\musEighth$,
@@
 %that represents continuation
-it is  skipped with no cost (transitions $\wei_{01}$ or weight $\one$).
+is  skipped with no cost (transitions $\wei_{01}$).
 %Likewise, the markups are ignored.
@@
 \label{section:SWVPA}\label{sec:SWVPA}
-The model presented in this section generalizes symbolic \VPA
+The model presented now generalizes symbolic \VPA
 (\SVPA~\cite{dAntonyAlur14SVPDA},
  generalizing themselves \VPA~\cite{AlurMadhusudan09nested} to infinite alphabets)
-from Boolean semirings to arbitrary semiring weight domains.
-It will compute on nested words over infinite alphabets,
-associating to every such word a weight value.
-Nested words are able to describe structures of labeled trees,
-and in the context of parsing, they will be useful to
+from Boolean semirings to arbitrary semiring domains.
+It associates to every nested words over infinite alphabets a weight value.
+Nested words can describe structures of labeled trees.
+In the context of parsing, they will be useful to
 represent AST. % parse trees.
@@
 %finite (large) or infinite,
-that we assume partitioned into three
+artitioned into three
 subsets~$\Deltai$, $\Deltac$, $\Deltar$,
 whose elements are respectively called
-\emph{internal}, \emph{call} and \emph{return} symbols,
-following~\cite{AlurMadhusudan09nested}.
+\emph{internal}, \emph{call} and \emph{return} symbols~\cite{AlurMadhusudan09nested}.
 % names are  coined by application to functional program verification
@@
 \begin{example}\label{ex:nested-word}
-Consider once more the linearized score of Example~\ref{ex:running},
-with $\Deltai$ representing timed notes and continuations.
+The score representation $O$ in Example~\ref{ex:running}
+is a nested-word of~$\Delta^*$.
+$\Deltai$ corresponds to timed notes and continuations.
 %the set of music events of the form $<\nu, \tau>$
@@
 %rest symbol R.
-The subset $\Delta \setminus \Deltai$ is partitioned into
-$\Deltac$ and $\Deltar$, containing respectively opening and closing parentheses.
+$\Deltac$ and $\Deltar$ contain respectively opening and closing parentheses.
 %are used to encode the score structure:
@@
 %is encoded with the nested word $\langle \prec 0, \frac{3}{4}, \frac{7}{8} \succ \rangle %\langle \prec 1, \frac{4}{3}, \frac{5}{3}  \succ \rangle$
-The score representation $O$ in Example~\ref{ex:running}
-% score: \includegraphics[scale=0.35,trim=0 5mm 0 0]{pictures/ex1.pdf},
-is a nested-word of~$\Delta^*$.
-%
 Another example is the second candidate
@@
 $\creturn{\mathsf{m}}\ts{2}$
-(see also Figure~\ref{fig:score-tree}).
+(see also Fig.~\ref{fig:score-tree}).
 The symbol between quotes $\mbox{\footnotesize `C$\sharp$5`}$
 represents an \emph{appogiatura}, %(grace-note),
-that is an ornemental note with theoretical duration zero.
+i.e., an ornemental note with theoretical duration zero.
 \endex
@@
 %
-Similarly as in Section~\ref{section:transducer},
+As in Section~\ref{section:transducer},
 we extend the above transition functions as follows
@@
 \begin{example} \label{ex:SWVPA}
-We consider a \SWVPA over the alphabet of Example~\ref{ex:nested-word}
+We consider a \SWVPA over the alphabet of Ex.~\ref{ex:nested-word}
 that expresses a weight related to the music notation,
-more precisely to its structural complexity:
-given a set of equivalent representations, we aim at choosing the simpler one.
+more precisely to its structural complexity.
+Given a set of equivalent representations, we aim at choosing the simpler one.
 
 Let us assume a third component in the symbols of $\Deltac$
+\philippe{Je trouve l'exemple un peu confus. Je ne sais pas
+quel est "l'associate time interval" d'un symbole. Je ne sais
+pas à quoi fait référence "the following call transition". Je
+n'ose pas trop toucher mais je crois qu'il faut revoir.}
 which is the duration of the associated time interval.

### 2021-07-19 main.tex
Ajout du newpage pour marquer les 12 pages
proof property closure SWA byb SWT

--- 2021-07-17 main.tex
@@
 \newcommand{\dom}{\ensuremath{\mathit{dom}}}
+\newcommand{\seq}{\ensuremath{\mathit{seq}}}
+\newcommand{\word}{\mathit{word}}
 
@@
 %\paragraph*{Semirings}
-%\label{section:semiring}\label{sec:semiring}
+\label{section:semiring}\label{sec:semiring}
 \input{semiring}
@@
 %\paragraph*{Label Theory}
-%\label{section:symbols}
+\label{section:symbols}
 \input{labels}
@@
 with a word $u \in \Sigma^*$ (resp. $v \in \Delta^*$).
-\lydia{added $u$ and $v$ def}
+%\lydia{added $u$ and $v$ def}
 %
@@
 when reading an extra note $\mu$.
-\philippe{Comprends pas cette phrase}
-\florent{$\to$ reformulation}
+%\philippe{Comprends pas cette phrase}
+%\florent{$\to$ reformulation}
 The transitions in the second column below switch back to the normal state $q_0$.
@@
 $B_{A, T}$ over $\Delta$, $\Semiring$ and $\bar\Phi$,
-such that for all $t \in \Delta^*$,
+such that for all $t \in \Delta^+$,
 $B_{A, T}(t) = \displaystyle\bigoplus_{s\in \Sigma^*} A(s) \otimes T(s, t)$.
@@
 and some input word $s \in \Sigma^*$.
-The weight for reading of the input is obtained with
+The weight for reading the input $s$ is obtained with
 $\bigoplus^1_\Sigma$.
@@
 %of~$T$ is its number of states $|Q|$.
-\florent{revise complexity with nb. of tr. and states}
+\florent{revise complexity with nb. of tr. and states? 
+         or skip details and just say polynomial?}
 %
@@
 $B_{s, T}$ over $\Delta$, $\Semiring$ and $\bar\Phi$,
-such that for all $t \in \Delta^*$, $B_{s, T}(t) = T(s, t)$.
+such that for all $t \in \Delta^+$, $B_{s, T}(t) = T(s, t)$.
 \end{corollary}
@@
 and in the context of parsing, they will be useful to
-represent AST. % parse trees.
-\florent{see \S 5 and App.A}
+represent AST 
+(see Section~\ref{sec:parsing} and Appendix~\ref{app:trees}). % parse trees. 
+%\florent{see \S 5 and App.A}
 
@@
 
-\section{Nested-Words and Parse-Trees} \label{sec:trees}
+\section{Nested-Words and Parse-Trees} \label{sec:trees}\label{app:trees}
 \input{app-trees}

### 2021-07-19 main.tex

--- 2021-07-18 main.tex
@@
 \noindent
-\florent{mv formal def. weight for \SWA to app.?}
+%\florent{mv formal def. weight for \SWA to app.?}
 As above in the case of \SWT,
@@
 The computation of $A$ on words $s \in \Sigma^*$
-is defined with an intermediate function $\weight_A$,
-defined as follows for $q, q' \in Q$, $a \in \Sigma$, $u \in \Sigma^*$,
+is based with an intermediate function $\weight_A$,
+defined as follows for $q, q' \in Q$, $a \in \Sigma$, $u \in \Sigma^*$
 %
@@
 and a $\SWA$ $A$ over $\Sigma$, $\Semiring$ and $\bar\Phi$,
-there exists an effectively constructible \SWA
-$B_{A, T}$ over $\Delta$, $\Semiring$ and $\bar\Phi$,
-such that for all $t \in \Delta^*$,
+there exists a \SWA $B_{A, T}$ over $\Delta$, $\Semiring$ and $\bar\Phi$,
+effectively constructible in PTIME, 
+such that for all $t \in \Delta^+$,
 $B_{A, T}(t) = \displaystyle\bigoplus_{s\in \Sigma^*} A(s) \otimes T(s, t)$.
@@
 %The complexity of construction
-\noindent
-The construction time and size for $B_{A, T}$ are $O(\| T \|^3 . \| A \|^2)$,
-where the sizes $\| T \|$ and $\| A \|$ are their number of states.
-%of~$T$ is its number of states $|Q|$.
-\florent{revise complexity with nb. of tr. and states}
+%\noindent
+%The construction time and size for $B_{A, T}$ are $O(\| T \|^3 . \| A \|^2)$,
+%where the sizes $\| T \|$ and $\| A \|$ are their number of states.
+%%of~$T$ is its number of states $|Q|$.
+%\florent{revise complexity with nb. of tr. and states? 
+%         or skip details and just say polynomial?}
 %
@@
 In the context of parsing, they will be useful to
-represent AST. % parse trees.
-\florent{see \S 5 and App.A}
+represent AST % parse trees.
+(see Section~\ref{sec:parsing} and Appendix~\ref{app:trees}). % parse trees. 
+%\florent{see \S 5 and App.A}
 
@@
 $\weiei$, $\weiec$, and $\weier$ describe the cases where the stack is empty.
-\lydia{moved this to the beginning}
+%\lydia{moved this to the beginning}
 %
@@
 since~$\Semiring$ is complete.
-The algorithm performs $2.|Q|^2$ iterations until $P$ is empty,
-and each iteration has a time complexity $O(|Q|^2 . |P|)$.
+%
+% COMPLEXITY
+%The algorithm performs $2.|Q|^2$ iterations until $P$ is empty,
+%and each iteration has a time complexity $O(|Q|^2 . |P|)$.
 \florent{complexity: detail with nb tr. and states?}
-That gives a time complexity $O(|Q|^4 . |P|)$.
-It can be reduced by implementing $P$ as a priority queue,
-prioritized by the value returned by $d$.
-** where  $T$ is the worst time complexity for an oracle computing the values
-of Definition~\ref{def:effective}.
+%That gives a time complexity $O(|Q|^4 . |P|)$.
+%It can be reduced by implementing $P$ as a priority queue,
+%prioritized by the value returned by $d$.
+%** where  $T$ is the worst time complexity for an oracle computing the values
+%of Definition~\ref{def:effective}.
 %$|Q|^3.\log(|Q|^2)$
 %
-At termination, $d_\bot = b_\bot$ and $d_\top = b_\top$
-(see Appendix~\ref{sec:bestsearch}).
+Algorithm~\ref{algo:Dijkstra}
+terminates in PTIME and at termination, $d_\bot = b_\bot$ and $d_\top = b_\top$.
+\florent{(see Appendix~\ref{sec:bestsearch}).}
 With~\eqref{eq:min}, this ensures
-the correctness of Algorithm~\ref{algo:Dijkstra}.
-
+its correctness. %of Algorithm~\ref{algo:Dijkstra}.
+%
 \begin{figure}[ht]
@@
 \end{figure}
-
-\noindent
-** effectively computable by hypothesis that the label theory is effective**
-\florent{REVOIR}
-In order to obtain effectively a witness
-(word of $\Delta^*$ with a computation of $A$ of minimal weight),...
+%
+%\noindent
+%** effectively computable by hypothesis that the label theory is effective**
+%\florent{REVOIR}
+Thanks to the hypothesis of effectiveness of $\bar\Phi$, 
+it is possible to construct suring the iteration a witness word
+for Proposition~\ref{th:best-search}, 
+\ie a word of $\Delta^*$ with a minimal weight \wrt $A$.
+
 %we require the additional property of convexity of weight functions.***
-using the second requirement in Definition~\ref{def:effective}.
+%using the second requirement in Definition~\ref{def:effective}.
 
@@
 
+\newpage
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2021-07-19 introduction.tex
revise examples. 1 passe

@@
 an event \emph{continuation} (symbol `$-$`, see Ex.~\ref{ex:SWT}),
-or a \emph{markup}. %(\emph{parenthesis}),
-The timestamp $\tau'$
-belongs to a rhythmic ``grid'' obtained from recursive divisions:
+or a \emph{markup} (opening or closing parenthesis). %(\emph{parenthesis}),
+The temporal information $\tau'$ 
+is either a time interval, for the opening parentheses
+(representing the duration between the parenthesis and the mathing markup), 
+or a timestamp, for the other symbols.
+The time points in $\tau'$ belong to a rhythmic ``grid'' obtained from recursive divisions:
 whole notes ($\musWhole$) split in halves ($\musHalf$), halves
@@
 $O =$
-$\ccall{\mathsf{m}}\ts{0}$,
-$\ccall{2}\ts{0}$,
+$\ccall{\mathsf{m}}\ts{[0,1]}$,
+$\ccall{2}\ts{[0,1]}$,
 $\mbox{\footnotesize A4}\ts{0}$,
-$\ccall{2}\ts{\frac{1}{2}}$,
+$\ccall{2}\ts{[\frac{1}{2},1]}$,
 $-\ts{\frac{1}{2}}$,
-$\ccall{2}\ts{\frac{3}{4}}$,
+$\ccall{2}\ts{[\frac{3}{4},1]}$,
 $\mbox{\footnotesize B4}\ts{\frac{3}{4}}$,
@@
 $\creturn{2}\ts{1}$,
-$\ccall{\mathsf{m}}\ts{1}$,
-$\ccall{3}\ts{1}$,
+$\ccall{\mathsf{m}}\ts{[1,2]}$,
+$\ccall{3}\ts{[1,2]}$,
 $\mbox{\footnotesize D5}\ts{1}$,
@@
 represent a division of the current
-time interval into $d$ sub-intervals of equal length.
+time interval into $d$ sub-intervals of equal duration
+$\frac{\ell}{d}$ where $\ell$ is the length of the time interval attached to $\ccall{d}$.
 
 We will show that $O$ is a solution for the
-parsing of $I$. Note that several other parsings (\eg,
-\includegraphics[scale=0.35,trim=0 5mm 0 0]{pictures/ex2.pdf}) are possible.
+parsing of $I$. Note that several other parsings are possible
+like \eg \includegraphics[scale=0.35,trim=0 5mm 0 0]{pictures/ex2.pdf}.
 SW-parsing associates a weight value

### 2021-07-19 main.tex
revise examples. 1 passe

@@
 $O' =$
-$\ccall{\mathsf{m}}\ts{0}$,
-$\ccall{2}\ts{0}$,
+$\ccall{\mathsf{m}}\ts{[0,1]}$,
+$\ccall{2}\ts{[0,1]}$,
 $\mbox{\footnotesize A4}\ts{0}$,
-$\ccall{2}\ts{\frac{1}{2}}$,
+$\ccall{2}\ts{[\frac{1}{2},1]}$,
 $-\ts{\frac{1}{2}}$,
@@
 $\creturn{2}\ts{1}$,
-$\ccall{\mathsf{m}}\ts{1}$,
-$\ccall{3}\ts{1}$,
+$\ccall{\mathsf{m}}\ts{[1,2]}$,
+$\ccall{3}\ts{[1,2]}$,
 $\mbox{\footnotesize `C$\sharp$5`}\ts{1}$,
@@
 represents an \emph{appogiatura}, %(grace-note),
-i.e., an ornemental note with theoretical duration zero.
+i.e., an ornemental note with theoretical duration~0.
 \endex
@@
 
-Let us assume a third component in the symbols of $\Deltac$
-\philippe{Je trouve l'exemple un peu confus. Je ne sais pas
-quel est "l'associate time interval" d'un symbole. Je ne sais
-pas à quoi fait référence "the following call transition". Je
-n'ose pas trop toucher mais je crois qu'il faut revoir.}
-which is the duration of the associated time interval.
-It is used to check the timestamps in the symbols.
+%Let us assume a third component in the symbols of $\Deltac$
+%\philippe{Je trouve l'exemple un peu confus. Je ne sais pas
+%quel est "l'associate time interval" d'un symbole. Je ne sais
+%pas à quoi fait référence "the following call transition". Je
+%n'ose pas trop toucher mais je crois qu'il faut revoir.}
+%which is the duration of the associated time interval.
+%It is used to check the timestamps in the symbols.
 %
@@
 amongst $c$ sub-intervals of duration~$\ell$ (as indicated by the stack top).
-It reads a new time-division symbol~$\ccall{d}$.
-The date for this symbol has to be the time $\tau$
-for the previous time-division symbol read~$\ccall{c}$ (found on the stack)
-plus the total duration of intervals read so far, $i \ell$, and
-the new current interval duration is~$\frac{\ell}{d}$:
-%
+It reads a new time-division symbol~$\ccall{d}$:
 \(
 \weic\bigl(q_{{i}/{c}},
-           \ccall{c}\ts{\tau}\ts{\ell}, q,
-           \ccall{d}\ts{\tau + i\ell}\ts{\frac{\ell}{d}}, q_{{i+1}/{c}},
+           \ccall{c}\ts{[\tau, \tau+\ell]}, q,
+           \ccall{d}\ts{\iota}, q_{{i+1}/{c}},
            q_{{0}/{d}} \bigr) = \alpha_d
 \).
-
-\noindent
-The weight of this transition is $\alpha_d$.
+The time interval $\iota$ attached to the~$\ccall{d}$ read 
+must have a starting time $\tau + \frac{i\ell}{c}$,
+where~$\tau$ and~$\ell$ are respectively
+the starting time and duration of the interval attached to
+the previous time-division symbol read $\ccall{c}$ (found on the stack).
+%plus the total duration of intervals read so far, $i \ell$, and
+%the new current interval duration is~$\frac{\ell}{d}$.
+%
+And it must have a duration $\frac{\ell}{d}$.
+
+\noindent
+The weight of the above transition is $\alpha_d$.
 We can penalize \eg triplets compared to duplets with $\alpha_2 < \alpha_3$.
@@
 the state $q_{{i+1}/{c}}$ on the stack,
-to start the next sub-interval after reading $\creturn{d}$,
-with:
+in order to start the next sub-interval after reading $\creturn{d}$,
+with the transition:
 \(
 \weir\bigl(q_{{d}/{d}},
-           \ccall{d}\ts{\tau}\ts{\ell}, q_{{i}/{c}},
-           \,\creturn{d}\ts{\tau + d\ell}, q_{{i}/{c}} \bigr) = \one
+           \ccall{d}\ts{[\tau, \tau+\ell]}, q_{{i+1}/{c}},
+           \,\creturn{d}\ts{\tau + \ell}, q_{{i+1}/{c}} \bigr) = \one
 \).
@@
 \weii\bigl(q_{{i}/{d}},
-           \ccall{d}\ts{\tau}\ts{\ell}, q_{{j}/{c}},
-           \mu\ts{\tau + i\ell}, q_{{i+1}/{d}} \bigr) = \alpha_\mu
+           \ccall{d}\ts{[\tau, \tau+\ell]}, q_{{j}/{c}},
+           \mu\ts{\tau + \frac{i\ell}{d}}, q_{{i+1}/{d}} \bigr) = \alpha_\mu
 \).
@@
 %$q_{1/1}$ is a special (initial) state.
-The transition for starting to read the first measure is:
+The transition to start reading the first measure is:
 \(
 \weiec\bigl(q_{1/1},
-           \ccall{\mathsf{m}}\ts{0}\ts{1}, q_{1/1},
+           \ccall{\mathsf{m}}\ts{[0, 1]}, q_{1/1},
            q_{1/1} \bigr) = \one
@@
 \weic\bigl(q_{1/1},
-           \ccall{\mathsf{m}}\ts{\tau}\ts{1}, q_{1/1},
-           \ccall{\mathsf{m}}\ts{\tau + 1}\ts{1}, q_{1/1},
+          \ccall{\mathsf{m}}\ts{[\tau-1, \tau]}, q_{1/1},
+          \ccall{\mathsf{m}}\ts{[\tau, \tau+1]}, q_{1/1},
           q_{1/1} \bigr) = \one
@@
 Let $A_1$ and $A_2$ be two \SWVPA
-over the same $\Delta$, $\Semiring$ and $\bar\Phi$.
+over the same $\Delta$, commutative $\Semiring$ and effective $\bar\Phi$.
 There exists two effectively constructible $\SWVPA$
@@
 one can construct in PTIME a word $t \in \Delta^*$
-such that $A(t)$ is minimal \wrt the natural ordering for $\Semiring$.
+such that $A(t)$ is minimal \wrt the natural ordering $\leq_\oplus$ for $\Semiring$.
 \end{proposition}
@@
 %and each iteration has a time complexity $O(|Q|^2 . |P|)$.
-\florent{complexity: detail with nb tr. and states?}
+\florent{complexity: detail with      and states?}
 %That gives a time complexity $O(|Q|^4 . |P|)$.
@@
 let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a
-commutative, bounded, and complete \florent{total?} semiring
-and let $\bar\Phi$ be an effective label theory over $\Semiring$,
+commutative, bounded, total and complete %\florent{total?} 
+semiring and let $\bar\Phi$ be an effective label theory over $\Semiring$,
 containing $\Phi_\Sigma$, $\Phi_{\Sigma, \Deltai}$, as well as
@@
 a suitable feature in the context of applications such as, e.g., automatic music transcription.
+\florent{compare expressiveness with register etc automata}
 
@@
 
-\section{Correctness of the Best-Search Algorithm} \label{sec:bestsearch}
-\input{app-bestsearch}
+%\section{Correctness of the Best-Search Algorithm} \label{sec:bestsearch}
+%\input{app-bestsearch}
 

### 2021-07-19 abstract.tex
subbmission FST-TCS

@@
 and also unlike \WA whose transitions are labeled by constant weight values,
-and who deal only with finite automata.
+and who deal only with finite alphabets.
 %
@@
 %
-We illustrate the model with a motivating application, namely
+We illustrate the model with a motivating application to
 automated music transcription.

### 2021-07-19 main.tex
subbmission FST-TCS

@@
 \newcommand{\florent}[1]{\todo[noline,size=\tiny,color=yellow!40]{#1}}
+\renewcommand{\florent}[1]{} % uncomment for final
 \newcommand{\lydia}[1]{\todo[noline,size=\tiny,color=red!30]{#1}}
+\renewcommand{\lydia}[1]{} % uncomment for final
 \newcommand{\philippe}[1]{\todo[noline,size=\tiny,color=green!30]{#1}}
+\renewcommand{\philippe}[1]{} % uncomment for final
 
@@
 
-\florent{indep. counters Def. Prop. Ex...}
-
 \section{Introduction} \label{sec:intro}
@@
 weighted automata and transducers~\cite{Droste09handbook}
-%over finite alphabets, see  e.g.~\cite{Mohri03EDWA},
+%over finite alphabets, see \eg~\cite{Mohri03EDWA},
 by labeling each transition with a weight function
@@
 and the corresponding notated event $\nu$
-(e.g., MIDI pitch 69 and note $\mbox{\footnotesize A4}$),
+(\eg MIDI pitch 69 and note $\mbox{\footnotesize A4}$),
 the weight  computed by the \SWT is the time distance between both,
@@
 \begin{align}
-\weight_A(q, \varepsilon, q) & = \one\\ %\final(q)\\
-\weight_A(q, \varepsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
+\weight_A(q, \varepsilon, q') & = \one\ \mathrm{if}\  q = q',\ 
+          \mathrm{or}\  \zero\  \mathrm{otherwise}\\
+%\weight_A(q, \varepsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
 \weight_A(q, au, q') & =  \displaystyle\bigoplus_{q'' \in Q}
@@
 represents an \emph{appogiatura}, %(grace-note),
-i.e., an ornemental note with theoretical duration~0.
+\ie an ornemental note with theoretical duration~0.
 \endex
@@
 and its structural complexity.
+\endex
 \end{example}
+The application to music transcription suggested briefly in the examples
+has been implemented in a C++ tool~\cite{qparse}, %\footnote{https://qparse.gitlabpages.inria.fr}.
+following the principles of the present SW-parsing framework, 
+although it differs in several points.
+In particular, the automata constructions are performed on the on-the-fly
+during the search of a best AST, for efficiency reasons.
 
@@
 (see \eg~\cite{Goodman99SemiringParsing}
- and~\cite{MorbitzVogler19weighted-parsing}).
-Weighted parsing
-corresponds to $\SW$-parsing in the case of finite alphabets,
-a transducer $T$ computing the identity and some \SWVPA~$A$
-obtained from $G$. %the weighted CF grammar.
+ and the more general framework~\cite{MorbitzVogler19weighted-parsing}),
+with infinite input alphabet instead of a finite one and
+transducer-defined distances instead of equality. 
+%
+See Appendix~\ref{sec:trees} for more details on the correspondence between 
+nested words $t \in \Delta^*$ and AST and CF grammars and \SWVPA.
+
 %
@@
 %
-Indeed, the \emph{depth-first} traversal of an AST $\tau$
-yields a well-parenthesised word $\lin(\tau)$ over an alphabet
-$\Delta = \Deltai \uplus \Deltac \uplus \Deltar$,
-assuming \eg that $\Deltai$ contains the symbols labelling the leaves of $\tau$ (symbols of rank $0$),
-and $\Deltac$ and $\Deltar$ contain respectively one left and right parenthesis
-$\ccall{b}$ and $\creturn{b}$ for each symbol $b$ labelling inner nodes of $\tau$ (symbols of rank $>0$).
+%Weighted parsing
+%corresponds to $\SW$-parsing in the case of finite alphabets,
+%a transducer $T$ computing the identity and some \SWVPA~$A$
+%obtained from $G$. %the weighted CF grammar.
+%Indeed, the \emph{depth-first} traversal of an AST $\tau$
+%yields a well-parenthesised word $\lin(\tau)$ over an alphabet
+%$\Delta = \Deltai \uplus \Deltac \uplus \Deltar$,
+%assuming \eg that $\Deltai$ contains the symbols labelling the leaves of $\tau$ 
+%(symbols of rank $0$),
+%and $\Deltac$ and $\Deltar$ contain respectively one left and right parenthesis
+%$\ccall{b}$ and $\creturn{b}$ for each symbol $b$ labelling inner nodes of $\tau$ 
+%(symbols of rank $>0$).
 %
@@
 %
-We show in Appendix~\ref{sec:trees} how to construct
-%($\SW$) tree automaton~$A$ %(in particular the TA computing on the AST of a given CF-grammar)
-a $\SWVPA$ $A$ such that $A(\lin(\tau))$ is the weight the AST $\tau$ of $G$.
+%We show in Appendix~\ref{sec:trees} how to construct
+%%($\SW$) tree automaton~$A$ %(in particular the TA computing on the AST of a given CF-grammar)
+%a $\SWVPA$ $A$ such that $A(\lin(\tau))$ is the weight the AST $\tau$ of $G$.
 
@@
 %...
-%\florent{2 lines Application to Automated Music Transcription:
-%implementation $\neq$ but same principle, on-the-fly automata construction
-%during best search, for efficiency.}
 
@@
 
-We  introduced weighted language models over infinite alphabets,
+We  introduced Symbolic Weighted language models %over infinite alphabets,
 and applied them to the problem of parsing
@@
 This allows to consider finer word relationships than strict equality.
+%
 
 This work can be extended in several directions. First, the
-the best search algorithm could bbe generalized from $1$-best to $n$-best~\cite{Huang05kbest},
+the best search algorithm could be generalized from $1$-best to $n$-best~\cite{Huang05kbest},
 and to $k$-\emph{closed} semirings~\cite{Mohri02semiring}
 (instead of \emph{bounded}, which corresponds to $0$-\emph{closed}). Second,
-we believe that
-the complexity bounds of the algorithms can be more precisely characterized.
+%we believe that
+the complexity bounds of the algorithms could be more precisely characterized,
+as well as expressiveness of \swM compared to the automata of 
+\eg~\cite{Segoufin06csl,KaminskiFrancez94,NevenSchwentickVianu04FSMinfinite,Bojanczyk11FO2}.
 Finally, the  best search algorithm presented here offline, whereas
 an on-the-fly automata construction would allow for online parsing,
-a suitable feature in the context of applications such as, e.g., automatic music transcription.
-\florent{compare expressiveness with register etc automata}
+a suitable feature in the context of applications such as, 
+\eg automatic music transcription.
+
+\florent{comparison of expressiveness with the automata 
+of\cite{Segoufin06csl,KaminskiFrancez94,NevenSchwentickVianu04FSMinfinite,Bojanczyk11FO2}.}
 

### 2021-07-20 main.tex
passe section 4 swVPA

@@
 from Boolean semirings to arbitrary semiring domains.
-It associates to every nested words over infinite alphabets a weight value.
+It associates to every nested word over an infinite alphabet a weight value
+in a semiring.
 Nested words can describe structures of labeled trees.
@@
 \begin{example}\label{ex:nested-word}
-The score representation $O$ in Example~\ref{ex:running}
-is a nested-word of~$\Delta^*$.
-$\Deltai$ corresponds to timed notes and continuations.
+In the nested score representation $O \in \Delta^*$ in Ex.~\ref{ex:running},
+$\Deltai$ corresponds to timed notes and continuations,
 %the set of music events of the form $<\nu, \tau>$
@@
 %rest symbol R.
-$\Deltac$ and $\Deltar$ contain respectively opening and closing parentheses.
+and $\Deltac$ and $\Deltar$ contain respectively opening and closing parentheses.
 %are used to encode the score structure:
@@
 %is encoded with the nested word $\langle \prec 0, \frac{3}{4}, \frac{7}{8} \succ \rangle %\langle \prec 1, \frac{4}{3}, \frac{5}{3}  \succ \rangle$
-Another example is the second candidate
+Another example is the other candidate
 \includegraphics[scale=0.35,trim=0 5mm 0 0]{pictures/ex2.pdf}
@@
 $[Q \times \Gamma^*] \times \Delta^* \times [Q \times \Gamma^*]$ into~$\Semiring$.
-The empty stack is denoted by $\bot$, and the first %upmost
+The empty stack is denoted by $\bot$, and the topupmost
 symbol is the last pushed content.
@@
 or $\call{c} \in \Deltac$, or $\return{r} \in \Deltar$,
-and for each possible state of the stack (empty or not).
+for each possible state of the stack (empty or not).
 %to add to $u \in {\Delta}^*$.
@@
 \begin{example} \label{ex:SWVPA}
-We consider a \SWVPA over the alphabet of Ex.~\ref{ex:nested-word}
+We consider a \SWVPA over the alphabet of Example~\ref{ex:nested-word}
 that expresses a weight related to the music notation,
-more precisely to its structural complexity.
+or more precisely to its structural complexity.
 Given a set of equivalent representations, we aim at choosing the simpler one.
@@
 For instance, the following call transition starts in state
-$q_{{i}/{c}}$, meaning that the current interval has number~$i$
+$q_{{i}/{c}}$, meaning that the current interval has number~$i < c$
 amongst $c$ sub-intervals of duration~$\ell$ (as indicated by the stack top).
@@
 the state $q_{{i+1}/{c}}$ on the stack,
-in order to start the next sub-interval after reading $\creturn{d}$,
+in order to start the next sub-interval after reading 
+a return symbol $\creturn{d}$,
 with the transition:
@@
 \noindent
-Reading an event is done with:
+Reading a musical event $\mu$ is done with:
 \(
@@
 \noindent
-We now present a procedure for searching, for a~$\SWVPA$ $A$,
+We present now a procedure for searching, for a~$\SWVPA$ $A$,
 a word %$w \in \Delta^*$
@@
 %
-Following~\eqref{eq:SWVPA-value}, and the associativity and commutativity
+Following~\eqref{eq:SWVPA-value}, and the associativity, commutativity
 and distributivity for~$\otimes$ and~$\oplus$, the minimum of~$A(t)$ is:
@@
 (\ie the substack $\gamma$ in the above $\gamma' = \<c, p> \, \gamma$)
-ensures that this part $\gamma'$ is not touched during the computation.
+ensures that this part is not touched during the computation.
 % there cannot be a pop of $\<c, p>$ followed by a push of $\<c, p>$
@@
 %\noindent\quad
-such that $d_\bot(q_1, q_2)$, resp. $\bigoplus_{c\in \Deltac} d_\top(q_1, p, q_2)(c)$, is minimal in $\Semiring$ \wrt $\leq_\oplus$$\;$
+such that $d_\bot(q_1, q_2)$, 
+resp. $\bigoplus_{\Deltac} d_\top(q_1, p, q_2)$, is minimal in $\Semiring$ \wrt $\leq_\oplus$$\;$
 
@@
 %\textbf{for all} $q_0, q_3 \in Q$
-\Update $d_\bot$ with $\< q_1, q_2>$ or $d_\top$ with $\< q_1, p, q_2>$ (Figure~\ref{fig:best-update}).$\;$
+\Update $d_\bot$ with $\< q_1, q_2>$ or $d_\top$ with $\< q_1, p, q_2>$ 
+(Figure~\ref{fig:best-update}).$\;$
 }
@@
 Algorithm~\ref{algo:Dijkstra}
-constructs iteratively two markings
+constructs iteratively,
+using a priority queue $\Q$, 
+two markings
 $d_\bot : Q \times Q \to \Semiring$ and
@@
 that converges eventually to $b_\top$ and $b_\bot$. 
-%It uses for that purpose a priority queue $P$ containing triplets of
-%$Q \times \{ \bot, \top \} \times Q$.
-\florent{explication Fig.~\ref{fig:best-update} (update) suivant cas de \eqref{eq:SWVPA-weight}?}
-The infinite sums in the updates of $d$ in Algorithm~\ref{algo:Dijkstra},
-Figure~\ref{fig:best-update}
-are well defined
-since~$\Semiring$ is complete.
+%\florent{explication Fig.~\ref{fig:best-update} (update) suivant cas de \eqref{eq:SWVPA-weight}?}
+%\florent{The infinite sums in the updates of $d$ in Algorithm~\ref{algo:Dijkstra},
+%Figure~\ref{fig:best-update} are well defined since~$\Semiring$ is complete.}
 %
@@
 %and each iteration has a time complexity $O(|Q|^2 . |P|)$.
-\florent{complexity: detail with      and states?}
+%\florent{complexity: detail with and states?}
 %That gives a time complexity $O(|Q|^4 . |P|)$.
@@
 %
-Algorithm~\ref{algo:Dijkstra}
-terminates in PTIME and at termination, $d_\bot = b_\bot$ and $d_\top = b_\top$.
+%Algorithm~\ref{algo:Dijkstra}
+It terminates in PTIME and at termination (when $\Q$ is empty), 
+and its correctness is ensured by~\eqref{eq:min} and  
+the following invariants: 
+$\< q_1, q_2> \notin \Q$ iff $d_\bot(q_1, q_2) = b_\bot(q_1, q_2)$, 
+and $\< q_1, p, q_2> \notin \Q$ iff 
+$\bigoplus_{\Deltac} d_\top(q_1, p, q_2) = \bigoplus_{\Deltac} b_\top(q_1, p, q_2)$.
 \florent{(see Appendix~\ref{sec:bestsearch}).}
-With~\eqref{eq:min}, this ensures
-its correctness. %of Algorithm~\ref{algo:Dijkstra}.
-%
+%of Algorithm~\ref{algo:Dijkstra}.
+
 \begin{figure}[ht]
@@
 %\florent{REVOIR}
-Thanks to the hypothesis of effectiveness of $\bar\Phi$, 
-it is possible to construct suring the iteration a witness word
+Thanks to the hypothesis that $\bar\Phi$ is effective, 
+it is possible to construct during the iteration a witness word
 for Proposition~\ref{th:best-search}, 
-\ie a word of $\Delta^*$ with a minimal weight \wrt $A$.
+\ie a word $t \in \Delta^*$ with a minimal weight $A(t)$ \wrt~$\leq_\oplus$.
 
@@
 \noindent
-It remains to compute a best nested-word $t \in {\Delta}^*$
+It remains to compute a best nested word $t \in {\Delta}^*$
 using the procedure of Proposition~\ref{th:best-search}.
@@
 
-\section{Nested-Words and Parse-Trees} \label{sec:trees}\label{app:trees}
+\section{Nested Words and Parse Trees} \label{sec:trees}\label{app:trees}
 \input{app-trees}

### 2021-07-20 abstract.tex
passe abstract, introduction

@@
 It is based on language models called Symbolic Weighted Automata (\SWA)
-at the joint %intersection
-between Symbolic Automata (\SA) and Weighted Automata (\WA),
+at the joint between Symbolic Automata (\SA) and Weighted Automata (\WA),
 as well as Transducers (\SWT) and Visibly Pushdown (\SWVPA) variants.
@@
 The transitions of \SWA are labeled by functions from an infinite alphabet into the weight domain.
-This is unlike \SA whose transitions are guarded by Boolean predicates
-overs symbols in an infinite alphabet
-and also unlike \WA whose transitions are labeled by constant weight values,
+This generalizes \SA, whose transitions are guarded by Boolean predicates
+overs symbols in an infinite alphabet,
+and also \WA, whose transitions are labeled by constant weight values,
 and who deal only with finite alphabets.
@@
 We present some properties of \SWA, \SWT and \SWVPA models,
-%and show how they can be used
 that we use to define and solve a variant of parsing

### 2021-07-20 introduction.tex
passe abstract, introduction

@@
 Considering large or infinite alphabets can however be of
-practical interest, when dealing with large characters encodings such as UTF-16,
+practical interest, when dealing with large characters encodings such as UTF-16~\cite{dAntoni21CACM},
 % processing strings in \eg for vulnerability detection in Web-applications~\cite{dAntoni21CACM},
 %
-analysis of data streams, serialization of structured documents~\cite{Segoufin06csl},
+analysis of data streams, serialization of structured documents~\cite{Segoufin06csl,NevenSchwentickVianu04FSMinfinite},
 or processing timed execution traces~\cite{Bouyer03algebraic}.
@@
 are present, either explicitly  or implicitly,
-in Common Western Music Notation~\cite{Gould11Notation}.
+in Common Western Music Notation~\cite{Gould11Notation}:
 %in music notation~\cite{Gould11Notation}:
-music scores are partitioned in measures,
+Music scores are partitioned in measures,
 measures in beats, and beats can be further recursively divided.
@@
 to functions into an arbitrary semiring,
-and generalizing also~\WA, by handling infinite alphabets (Fig.~\ref{fig:hierarchy}).
+and generalizing also~\WA, by handling infinite alphabets (Figure~\ref{fig:hierarchy}).
 %
@@
 (\eg, $\mbox{\footnotesize A4}$, $\mbox{\footnotesize G5}$, \etc.),
-an event \emph{continuation} (symbol `$-$`, see Ex.~\ref{ex:SWT}),
+an event \emph{continuation} (symbol `$-$`, see Example~\ref{ex:SWT}),
 or a \emph{markup} (opening or closing parenthesis). %(\emph{parenthesis}),
@@
 is either a time interval, for the opening parentheses
-(representing the duration between the parenthesis and the mathing markup), 
+(representing the duration between the parenthesis 
+ and the matching closing one), 
 or a timestamp, for the other symbols.

### 2021-07-20 main.tex
passe abstract, introduction

@@
 \begin{equation}\label{eq:distance-lang}
-d(s, t) \otimes A(t) = \displaystyle\bigoplus_{t' \in \Delta^*} d(s, t') \otimes A(t')
+d(s, t) \otimes A(t) = \displaystyle\bigoplus_{u \in \Delta^*} d(s, u) \otimes A(u)
 \end{equation}
@@
 \item The \SWT $T$ evaluates a ``fitness measure''  that expresses
-    a correspondance between a performance and notation events. See Example~\ref{ex:SWT}.
+    a correspondance between a performance and a nested representation of a music score.
+    See Example~\ref{ex:SWT}.
 \item The \SWVPA $A$ expresses a cost related to the music notation.
@@
 \begin{proposition}
-The problem of Symbolic Weighted  parsing
+The problem of Symbolic Weighted Parsing
 can be solved in PTIME in the size of the input \SWT $T$, \SWVPA $A$
@@
 that yields a given input word $w$, called~\emph{weighted parsing},
-(see \eg~\cite{Goodman99SemiringParsing}
+(see~\cite{Goodman99SemiringParsing}
  and the more general framework~\cite{MorbitzVogler19weighted-parsing}),

### 2021-07-20 main.tex
passe section 3: swT, swA. resubmit.

@@
 defined recursively, for every states $q, q' \in Q$,
-and every pairs of strings $\< s, t> \in \Sigma^* \times \Delta^*$,
+and every pairs of strings $\< s, t> \in \Sigma^* \times \Delta^*$.
 % \setminus \{ \< \varepsilon, \varepsilon >\}$,
-where $au$, and $bv$, denote the concatenation
-of the symbol $a \in \Sigma$ (resp. $b \in \Delta$)
-with a word $u \in \Sigma^*$ (resp. $v \in \Delta^*$).
-%\lydia{added $u$ and $v$ def}
 %
@@
 Intuitively,  a transition $\wei_{ij}(q, a, b, q')$ is interpreted as follows:
-when reading  $a$ and $b$ the input and output words,
+when reading $a$ and $b$ in the input and output words,
 increment the current position in the input word if and only if $i = 1$,
@@
 We propose a small weighted transducer model with two states $q_0$ and $q_1$
-that calculates the distance bewteen an input sequence of sounding
+that calculates the distance between an input sequence of sounding
 events (music "performance")

### 2021-08-31 FST-TCS/rebuttal/rebuttal.md
rebuttal TSF TCS

Thank you for your submission to FSTTCS 2021. The FSTTCS 2021
rebuttal period is now open, and it will close at 23:59, September 1, AoE.

During this time, you will have access to the current state of your
reviews and have the opportunity to submit a response. 
Please keep in mind the following during this process:

* The response should mainly focus on any factual errors in the reviews and
any questions posed by the reviewers. It must not provide new
research results or reformulate the presentation.  There is no limit on the length of the response, but please try to be concise and to the point.

* The rebuttal period is an opportunity to react to the reviews, but
not a requirement to do so. Thus, if you feel the reviews are
accurate and the reviewers have not asked any questions, then you do
not have to respond.

* The reviews are as submitted by the PC members, without any
coordination between them. Thus, there may be inconsistencies.
Furthermore, these are not the final versions of the reviews. The
reviews can later be updated to take into account the discussions at
the program committee meeting, and we may find it necessary to
solicit other outside reviews after the rebuttal period.

* The program committee will read your responses carefully and take
this information into account during the discussions. On the other
hand, the program committee will not directly respond to your
responses, either before the program committee meeting or in the
final versions of the reviews.

* Your response will be seen by all PC members who have access to the
discussion of your paper, so please try to be polite and constructive.

The reviews on your paper are attached to this letter. To submit your
response you should log on the EasyChair Web page for FSTTCS 2021 and
select your submission on the menu.

Best wishes,
Chandra (Track A) and Mikolaj (Track B)

---
## REVIEW 1
SUBMISSION: 89
TITLE: Symbolic Weighted Language Models and Quantitative Parsing over Infinite Alphabets
AUTHORS: Florent Jacquemard, Philippe Rigaux and Lydia Rodrigez de la Nava

----------- Overall evaluation -----------
The paper introduces symbolic weighted automata (swA) as a generalization of
finite automata in two directions: first, like symbolic automata, they allow
infinite input alphabets and second, like weighted automata, they map each
input word to a value in a semiring .  In the same spirit, symbolic weighted
transducers (swT) and symbolic visibly pushdown automata (sw-VPA) are
introduced.

Based on these concepts, the paper introduces a variety of constructions.
The first one is a Bar-Hillel Perles Shamir construction for a given swT and
a given sWA.  The second one performs the closure of two sw-VPA under the
semiring operations.  The third one computes, for a given sw-VPA, the best
word with respect to the natural order of the underlying semiring.

Finally, the paper introduces the framework of Symbolic Weighted Parsing,
which, given
- a word from some input alphabet,
- an sw-VPA over some target alphabet, and
- an swT mapping words from the input alphabet to the internal symbols of the
  sw-VPA's alphabet,
aims to compute a nested word which minimizes both the weight assigned to it
by the sw-VPA and the distance to the input word (which is computed by the
swT).  The authors show that this problem can be solved in polynomial time by
combining the three constructions introduced earlier.  Moreover, they
motivate their work by the example of the automatic transcription of music
recordings into music scores.

The formalism of symbolic weighted automata is not original to the paper.  It
was already introduced by Herrmann and Vogler (2016) and later revised by
Herrmann (2020) under the name of weighted symbolic automata with data
storage.  This literature subsumes swA (by choosing the trivial storage),
swT (since transducers can be understood as particular weighted automata),
and sw-VPA, for which the data storage VP was defined.  Moreover, the earlier
work is more general, since it allows unital valuation monoids (rather than
just semirings) as weight structures.  They authors of the present paper
should refer to the earlier publications and clearly distinguish between
adopted concepts and their own work.

Moreover, the authors claim that their framework is a generalization of
previous approaches to weighted parsing, in the sense that it allows infinite
input alphabets.  However, some of the previous approaches allowed more
general weight structures than just semirings, so their are incomparable to
the present paper.

The paper is very well structured and it was a pleasure to see all three
automaton models and constructions fit together in the final section.
However, the original content seems to have been squeezed into the final
pages of the paper, while in the beginning and in the center many pages were
used to repeat the concepts of Herrmann and Vogler (2016).  Moreover, the
appendix contains interesting remarks regarding the relation between nested
words and parse trees.  The are several occations where the paper refers to
the appendix, which should be avoided, since the appendix is not a part of
the published conference paper.  I enjoyed the running example of automatic
music transcription which serves to illustrate every introduced concept and
brings a different domain into consideration.  The overall impression of the
paper is spoiled by many careless mistakes, such as formulas sticking out
into the page margin (see below for a list).

Overall, I think that the paper cannot be published in its present form because
the introduction of symbolic weighted automata without reference to the
original literature could be seen as plagiarism.  I think, however, that the
original work presented in the paper can be of interest to the scientific
community.  This is why I opt to weakly reject the paper.  In order for the
paper to be published, its focus should be shifted from the theory of symbolic
weighted automata to the framework of symbolic weighted parsing.  In
particular, the authors could spend fewer pages repeating already established
concepts from the literature and integrate Appendix D into the main paper.

List of mistakes:
- 17: who -> which; missing comma after swT
- 18: superfluous comma
- 26: compilers or interpreters for programming languages
- 27: superfluous comma
- 32: superfluous comma
- 33: Western
- 43: skip commas around parentheses; `with' is not an optimal conjuction
- Figure 1: do no start sentences with symbols; add an `and' before S
- 47: Boolean Formulas
- 51--53: the jump from automata to grammars is uneasy and not really
  necessary, as the paper mainly deals with automata
- 57: weighted automata don't have ASTs, but runs
- 60: other approaches use multioperator monoids or valuation monoids, so `at
  minima a semiring' is wrong
- 70: the second comma is superfluous
- 74: missing period and space after \oplus
- 75: maybe cite Parsing as intersection by Nederhof and Satta (2003)
- 77: missing space after period
- 84: convenient in this context -> for convenience?
- 87: superfluous comma
- 116: \oplus `combines two values into a single one' as well (in fact, every
  binary operation does)
- 118--119: I would say: … \le_\oplus which is called the natural ordering of
  S and is, for every …, defined by …
  Also, you should be consistent with using `for all' or `for every' in
  quantifications (I would prefer for every)
- 122: if it IS; I'd also add a comma after total
- 123: Lemma 2 is Lemma 2 in [21], you should refer to it
- 125: you write `then', but the following sentence seems to be disconnected
  from the previous lemma
- 126: you should cite Knuth's generalization of Dijkstra's algorithm for
  superiority, he defines the notion for functions of arbitrary arity, so not
  just semirings (but multioperator monoids); you should give an example for
  superiority in the semiring notation (with \otimes)
  Moreover, the `corresponds' is wrong, since superiority generalizes
  non-negative edge weights
  PS: Knuth's definition of superiority includes monotonicity
- 128: increaseS
- 136: what implies?
- 143: Lemma 4 is Lemma 3 in [21]
- 146: need infinite sums with \oplus -> need to extend \oplus to infinitely
  many operands
- 147: dom is undefined; I can also be ℕ
- 151: partitionS
- 152: I'd use plural, i.e. products and sums
- 153: quantify x and the family (y_i) for i \in ℕ
- 155--156: alphabets are usually finite and non-empty sets
- 156: the sentence ends after … which is difficult to spot
- 159: denoted by; also it seems like \bar\Phi is a (\Sigma \cup \Sigma
  \times \Delta)-indexed family, why don't you introduce it like this
- 161--171: use an itemize environment
- 162--165: I think these should also hold for \Delta, why don't you write
  this explicitly?
- 186: calculability -> computability?
- 188: the definition of \bigoplus_\Delta can be seen to be implicit.  Please
  fully quantify the alphabets of a label theory
- 195: references: there is earlier research on symbolic automata, e.g.
  Veanes (2013)
- 196: put respectively after output
- 199: use \colon rather than : after function names (this also applies to
  several other places in the paper)
- 203: THE number
- 208: What does T compute?
- 216: by convention -> by definition
- 219: resp.\@
- 220--221: you mean that a term CONTAINING w_{ij}(…) equal to 0 is ignored
- 228: three superfluous commas
- 228--229: with \oplus/\otimes -> in terms of \oplus/\otimes
- 232: over THE alphabets
- 237: superfluous commas
- 238: text in the margin, also it is not clear whether the if refers to the
  entire line or just the right definition
- 247: I wouldn't define out just inside parentheses
- 251: based ON
- 252: for EVERY
- 253: replace comma by and
- 281: Partitioned
- 284: x is a variable, so it should be typeset in math mode
- 290: appogGiatura; ornAmental; also, I'd give this explanation already in
  Example 1 where the appoggiatura is first used
- 291: I'd at least say once that \Delta_i, \Delta_c, and \Delta_r are
  pairwise disjoint (or explain the symbol)
- 297: the extension here is not as in Section 3, there, the different
  function hat at least similar types
- 299: every line expands into the margin; moreover, in the rightmost
  equation of line three, in the application of w_c, p' and q' are swapped
- 310: topupmost -> topmost
- 316: equation extends into margin
- 330--331: as indicated by the stack top: better explain that stack symbols
  are time intervals
- 331: I think the application of w_c is wrong (same with w_r and w_i later),
  shouldn't the : be a comma and what is q in the middle?
- 336: how are states (which are in Q) pushed onto the stack if \Gamma =
  \Delta_c \times P?
- 339: what are j and c, do you mean i and d?
- 341: shouldn't there be a transition for returning from a measure?
- 345: exist (singular)
- 348: for?  do you mean accepted by?  but the language of an swA isn't
  defined either
- 352: Dijkstra-like algorithm?
- 354: readers may have an easier time seeing this argument if you say that S
  is extremal.  You should say in the preliminaries that S is extremal if and
  only if \le_\oplus is total.
- 366: be popPED
- 371: How is it ensured that the symbol <c,p> isn't popped and pushed again
  later during the computation?
- 375: converge (singular)
- 376: I do not understand the and at termination, and its correctness part.
  What do you want to express?
- 379: Almost all update equations need to compute infinite sums.  Don't you
  think that the assumption of \bar\Phi being effective is pretty strong and
  gives your the result `for free'?  I think you should at least elaborate on
  this and say to what degree your effective label theory is a generalization
  of the decidable label theory of unweighted symbolic automata.
- 386: We assume the following input to be given
- 394: the notation sw-parsing is undefined
- 416: word order: put `to T' in the end
- 420: `the sw-parsing problem' or `sw-parsing', but not `the sw-parsing'
- 424: the triple use and does not work well
- 431: double the
- 435: there is something missing between `here' and `offline'
- 629: typical references for regular tree grammars are Brainerd (1969) and
  Gecseg and Steinby (1984)
- 647: inn -> in
- 650: missing period after function
- 652: I think \bar w should be \bar w'
- 654: an intuition for the \bar q would have made understanding the
  construction far easier

Original literature:
- Herrmann and Vogler (2016): L. Herrmann & H. Vogler (2016): Weighted Symbolic Automata with Data Storage. In: Developments in Language Theory, Springer, pp. 203–215, doi:10.1007/978-3-662-53132-7 17.
- Herrmann (2020): L. Herrmann (2016): Weighted Automata with Storage. PhD thesis. Technische Universität Dresden, 2020, urn:nbn:de:bsz:14-qucosa2-740685

---
## REVIEW 2
SUBMISSION: 89
TITLE: Symbolic Weighted Language Models and Quantitative Parsing over Infinite Alphabets
AUTHORS: Florent Jacquemard, Philippe Rigaux and Lydia Rodrigez de la Nava

----------- Overall evaluation -----------
Transducers and visibly pushdown automata are used for parsing input words into abstract syntax trees. When the input words carry information from an infinite domain, such as time stamps, the transducers and automata need to be extended to handle these, which is the topic of this paper. The transducers are extended in one direction to handle inputs from the infinite domain, using the concepts from symbolic automata. Transducers are extended in another direction to give multiple outputs with associated weights. These can be used to produce multiple options for intermediate representations before computing the abstract syntax trees. The multiple intermediate representations can be compared and the one with the minimum weight can be picked up for further processing by visibly pushdown automata to construct the abstract syntax tree. The visibly pushdown automata are again extended to handle symbolic inputs and ouptut weights for different possible runs, which represent different!
  possible abstract syntax trees.

The contributions of this paper are extending transducers and automata to handle symbolic inputs and to output weights in a semiring instead of Boolean values. The paper also gives procedures for computing minimum weights for given inputs and to compute witnesses for these minimum weights, which will be helpful to construct syntax trees. The definitions have been extended from the respective symbolic and weighted models. The technical details are mostly well written and complemented with intuitive explanations using a running example. However, for now, I am still giving weak accept instead of accept, since I am having difficulty understading a few points. See the last two comments below.

Line 209:  "T is based on an intermediate function ...": should this be "The semantics of T is based on an intermediate function ..."?

Line 281: "artitioned into three" -> "partitioned into three"

Last line of algorithm 1: What exactly is meant by "update d_\bot with <q_1,q_2>"? Does it mean update the image of <q_1,q_2> under d_\bot with the new value as computed in Figure 3? Write this clearly.

Figure 3: This figure is supposed to compute <q_1,q_2>. First of all, it is not clear what is meant by "compute <q_1,q_2>"; I am guessing, as I wrote in the comment above that the computed quantity is the image of <q_1,q_2> under d_\bot. However, even this doesn't make sense; there are some series of equantions written, for d_\top(q_1,p,q_3) (not for d_\top(q_1,p,q_2). Even in this equation, what is q_2? Do we take the minimum (wrt \oplus) over all q_2? The second line says d_\bot(q_1,p,q_3), which doesn't make sense because d_\bot takes two arguments, not three. There are two equations for d_\top(q_1,p,q_3) and one for d_top(q_0,p,q_3). How are these related? It is very confusing what is being computed here and how. Please update this to make it clear.

## REVIEW 3
SUBMISSION: 89
TITLE: Symbolic Weighted Language Models and Quantitative Parsing over Infinite Alphabets
AUTHORS: Florent Jacquemard, Philippe Rigaux and Lydia Rodrigez de la Nava

----------- Overall evaluation -----------
**Review - "Symbolic Weighted Language Models and Parsing over Infinite Alphabets"**

Overall the paper is good, though somewhat in need of editing, spelling correction etc.

It describes how to usefully incorporate weights into symbolic automata, and how to leverage existing algorithms over symbolic automata and transducers to achieve a form of parsing of strings over infinite alphabet. In particular, the running example can take into account both linear fit along a continuous interval, as well as the complexity of the inferred syntax tree.

The method is described as empirically useful, and theoretically novel.

The descriptions of the various definitions and proofs are mostly clear and correct, apart from typos, and combines ideas from several previous works into a set of formalisms that are both independently interesting, and combine to form an efficient weighted parsing algorithm for VPL over infinite alphabets.

Comments of a more editorial character follows:

Introduction:

 The literature review should also include work by Hermann and Vogler, who bring in weights into symbolic automata as well, as well as stacks and other types of data storage, though their focus is more on exploring theoretical properties than empirical implementation, and make no complexity claims, as far as I can see. (Herrmann, Luisa, and Heiko Vogler. "Weighted symbolic automata with data storage." International Conference on Developments in Language Theory. Springer, Berlin, Heidelberg, 2016.)

 On lines 72+, the authors refer to Mohri in saying that "[...]given an input word $s$, the _SW-parsing_ problem aims at finding $t$ minimizing $T(s,t) \times A(t)$, called the distance between $s$ and $A$[...]". It is not entirely clear to me what part of this expression they consider to be the distance as described by Mohri. Is it the distance, as defined by $T$, from the $s$ to the distribution over ASTs $t$ defined by $A$? If so, it may be useful to structure the sentence somewhat differently, e.g. "[...] finding the linearized AST $t$ minimizing $d = T(s,t) \times A(t)$, where $d$ denotes the distance between $s$ and the distribution over linearized ASTs described by $A$, using the distance measure defined by $T$."

 Another possible interpretation would be that the distance is between the two distributions over linearized ASTs defined by, on the one hand, $T$ partially applied to $s$, and on the other, $A$, but this does not seem to fit with the importance of the specific AST $t$, at least not with my understanding of Mohri's definition of distances between distributions.

 I think that, for me, the confusion stems partly from Mohri exclusively talking about distances between two strings (languages,distributions) over the _same_ alphabet, while $s$, crucially, comes from a _different_ alphabet than the one over which $A$ defines a distribution.

Preliminary Notions:

 The introduction of Label theories (l 154++) does not depend on its first paragraph containing the definitions of words and alphabets, as far as I can tell, so I would let this first paragraph sit under its own title of e.g. "Alphabets and words". It is also, separately, not clear to me what the intuition and motivation is for using a label theory, at least at this point in the paper. In particular, I am unsure why any of the specific closures in ll 161+ are considered necessary. It may be a better use of space to define label theories rather simply, here, and put the specifics of the closures into the appendix, together with the various properties already present there.

SW Automata and Transducers

 A brief explanation of what you consider Mohri's approach to the computation of distances between words and languages may be helpful.

 The paragraph starting at l 224 is a great summary of the way you consider edit-distances in this paper.

 In Example 8, you disallow the last note being an error - is there some specific justification for that, or is it just for simplicity?

SW Visibly Pushdown Automata

 Example 14 is quite limited in detail in comparison to the earlier examples, while being more complex in scope. This makes the specifics of the example hard to double check, and to properly understand. It is also somewhat difficult to separate the symbols that form part of the language from those of the meta-language in the specifications of the various $w$ function calls. Additionally, you write "the above transition pushes the state $q_{i + 1/c}$ on[to] the stack" - I would encourage you to choose different letters and terminology than "state $q$" for stack symbols, to reduce the risk for the reader confusing the stack symbols and states. In fact, it may be a good idea to show a fuller schematic of a few steps of the sw-VPA computation, with relevant connections between various variables highlighted, what equalities need to hold etc.

 On line 366 of Proposition 16, the "can be" should be "must be", surely, since what you are describing is a computation that should leave the original stack (exactly) untouched?

 It seems to me that there is a consistent lack of quantification over $P$ in the section starting with the definition of $b_\top$. This is not hugely problematic for the proof, since the algorithm is still defined (mostly) properly, and the complexity should only be impacted by a factor of something like $|P|$.

Minor comments:

 l 87. I would front "input" in the sentence, yielding "a given input _timeline_ of [...] alphabet $\Sigma$, is parsed into[...]"

 l 90. Please parenthesize the pitch-timecode pairs for readability. As an aside: these are strictly the onsets? And strictly monophonic?

 l 100. Shouldn't the first measure end before the second starts? This would make $O$ a forest rather than a tree - easily fixed with a "piece" root element.

 l 156. The ellipsis between $\Delta$ and $\Sigma^*$ confuses the sentence boundaries for the reader. Better to go "The set of finite sequences (_words_) over $\Sigma$ is denoted $\Sigma^*$, while $\varepsilon$ denotes the empty word[...]".

 l 220. The sentence starting here is somewhat confusing (and possibly wrong, depending on if $q'$ and $q''$ are intended to be distinct). I suppose that the point is that one can quit the recursion "early" whenever $w_{ij}$ returns $0$, but the language could be clearer, something like "Since $0$ is absorbing for $\times$, and the identity element for $+$ in $S$, if $w_{ij}(q,a,b,q'')$ is $0$ (meaning there is no transition from $q$ to $q'$ reading $a$ and $b$) that entire term can be ignored in the sum."

 l 240: an fixed weight -> a fixed weight

 l 251: based with an intermediate -> based on an intermediate

 l 276: generalizing themselves -> themselves generalizing

 l 281: artitioned -> partitioned

 l 288: same as on l 100, measure 1 should maybe end before measure 2

 l 290: ornemental -> ornamental

 l 306: reads and pop from stack -> reads and pops from the stack

 l 307: I would use "computations" or something like that rather than "transitions" here

 l 310: topupmost -> topmost, though I'm not sure if it's necessary to describe the basic workings of a stack. If you choose to do so, then also

 l 311: content -> pair

 l 323: particular -> special

 l 365: "case of in the" -> either "case in the" or "case of the"

 l 366: pop -> popped

 The second case in Figure 3 should start with $d_\top$, not $d_\bot$

 l 395: I would omit "output". If you want to draw attention to the different alphabets involved, I would say "between the string $s$ over the input alphabet, and the weighted language over the output alphabet defined by $A$"

 l 399: correspondance -> correspondence

 l 402: "weight a second time division" -> something like "weight assigned to an additional second time division"

 l 416: extra commen between "$T$" and "of new transition"

 l 417: Corolary -> Corollary

 l 420: This whole paragraph should be shaped up, writing wise.

 l 434: $swM$??

 l 435: "offline" -> "works offline"

 l 437: extraneous comma after "such as"

 References:
- [ ]  Ref [22] and [23] both seem to refer to the same work by Mohri

------------------------------------------------------

### 2021-08-31 FST-TCS/rebuttal/rebuttal.md
rebbuttal: reviews

@@
+# rebuttal FST TCS
+
+> date: 30 August 2021
+>
+> from: fsttcs2021@easychair.org
+
 Thank you for your submission to FSTTCS 2021. The FSTTCS 2021
@@
 ----------- Overall evaluation -----------
-The paper introduces symbolic weighted automata (swA) as a generalization of
-finite automata in two directions: first, like symbolic automata, they allow
-infinite input alphabets and second, like weighted automata, they map each
-input word to a value in a semiring .  In the same spirit, symbolic weighted
-transducers (swT) and symbolic visibly pushdown automata (sw-VPA) are
-introduced.
+The paper introduces symbolic weighted automata (swA) as a generalization of finite automata in two directions: first, like symbolic automata, they allow infinite input alphabets and second, like weighted automata, they map each input word to a value in a semiring .
+In the same spirit, symbolic weighted transducers (swT) and symbolic visibly pushdown automata (sw-VPA) are introduced.
 
 Based on these concepts, the paper introduces a variety of constructions.
-The first one is a Bar-Hillel Perles Shamir construction for a given swT and
-a given sWA.  The second one performs the closure of two sw-VPA under the
-semiring operations.  The third one computes, for a given sw-VPA, the best
-word with respect to the natural order of the underlying semiring.
+The first one is a Bar-Hillel Perles Shamir construction for a given swT and a given sWA.
+The second one performs the closure of two sw-VPA under the semiring operations.
+The third one computes, for a given sw-VPA, the best word with respect to the natural order of the underlying semiring.
 
-Finally, the paper introduces the framework of Symbolic Weighted Parsing,
-which, given
+Finally, the paper introduces the framework of Symbolic Weighted Parsing, which, given
 - a word from some input alphabet,
 - an sw-VPA over some target alphabet, and
-- an swT mapping words from the input alphabet to the internal symbols of the
-  sw-VPA's alphabet,
-aims to compute a nested word which minimizes both the weight assigned to it
-by the sw-VPA and the distance to the input word (which is computed by the
-swT).  The authors show that this problem can be solved in polynomial time by
-combining the three constructions introduced earlier.  Moreover, they
-motivate their work by the example of the automatic transcription of music
-recordings into music scores.
+- an swT mapping words from the input alphabet to the internal symbols of the sw-VPA's alphabet,
+  aims to compute a nested word which minimizes both the weight assigned to it by the sw-VPA and the distance to the input word (which is computed by the swT).
+The authors show that this problem can be solved in polynomial time by combining the three constructions introduced earlier.
+Moreover, they
+  motivate their work by the example of the automatic transcription of music recordings into music scores.
 
-The formalism of symbolic weighted automata is not original to the paper.  It
-was already introduced by Herrmann and Vogler (2016) and later revised by
-Herrmann (2020) under the name of weighted symbolic automata with data
-storage.  This literature subsumes swA (by choosing the trivial storage),
-swT (since transducers can be understood as particular weighted automata),
-and sw-VPA, for which the data storage VP was defined.  Moreover, the earlier
-work is more general, since it allows unital valuation monoids (rather than
-just semirings) as weight structures.  They authors of the present paper
-should refer to the earlier publications and clearly distinguish between
-adopted concepts and their own work.
+The formalism of symbolic weighted automata is not original to the paper.
+It was already introduced by Herrmann and Vogler (2016) and later revised by Herrmann (2020) under the name of weighted symbolic automata with data storage.
+This literature subsumes swA (by choosing the trivial storage), swT (since transducers can be understood as particular weighted automata), and sw-VPA, for which the data storage VP was defined.
+Moreover, the earlier work is more general, since it allows unital valuation monoids (rather than just semirings) as weight structures.
+They authors of the present paper should refer to the earlier publications and clearly distinguish between adopted concepts and their own work.
 
-Moreover, the authors claim that their framework is a generalization of
-previous approaches to weighted parsing, in the sense that it allows infinite
-input alphabets.  However, some of the previous approaches allowed more
-general weight structures than just semirings, so their are incomparable to
-the present paper.
+Moreover, the authors claim that their framework is a generalization of previous approaches to weighted parsing, in the sense that it allows infinite input alphabets.
+However, some of the previous approaches allowed more general weight structures than just semirings, so their are incomparable to the present paper.
 
-The paper is very well structured and it was a pleasure to see all three
-automaton models and constructions fit together in the final section.
-However, the original content seems to have been squeezed into the final
-pages of the paper, while in the beginning and in the center many pages were
-used to repeat the concepts of Herrmann and Vogler (2016).  Moreover, the
-appendix contains interesting remarks regarding the relation between nested
-words and parse trees.  The are several occations where the paper refers to
-the appendix, which should be avoided, since the appendix is not a part of
-the published conference paper.  I enjoyed the running example of automatic
-music transcription which serves to illustrate every introduced concept and
-brings a different domain into consideration.  The overall impression of the
-paper is spoiled by many careless mistakes, such as formulas sticking out
-into the page margin (see below for a list).
+The paper is very well structured and it was a pleasure to see all three automaton models and constructions fit together in the final section.
 
-Overall, I think that the paper cannot be published in its present form because
-the introduction of symbolic weighted automata without reference to the
-original literature could be seen as plagiarism.  I think, however, that the
-original work presented in the paper can be of interest to the scientific
-community.  This is why I opt to weakly reject the paper.  In order for the
-paper to be published, its focus should be shifted from the theory of symbolic
-weighted automata to the framework of symbolic weighted parsing.  In
-particular, the authors could spend fewer pages repeating already established
+However, the original content seems to have been squeezed into the final pages of the paper, while in the beginning and in the center many pages were used to repeat the concepts of Herrmann and Vogler (2016).
+
+Moreover, the appendix contains interesting remarks regarding the relation between nested words and parse trees.
+The are several occations where the paper refers to the appendix, which should be avoided, since the appendix is not a part of the published conference paper.
+
+I enjoyed the running example of automatic music transcription which serves to illustrate every introduced concept and brings a different domain into consideration.
+The overall impression of the paper is spoiled by many careless mistakes, such as formulas sticking out into the page margin (see below for a list).
+
+Overall, I think that the paper cannot be published in its present form because the introduction of symbolic weighted automata without reference to the original literature could be seen as plagiarism.
+I think, however, that the original work presented in the paper can be of interest to the scientific community.
+This is why I opt to weakly reject the paper.
+In order for the paper to be published, its focus should be shifted from the theory of symbolic weighted automata to the framework of symbolic weighted parsing.
+
+In particular, the authors could spend fewer pages repeating already established
 concepts from the literature and integrate Appendix D into the main paper.
@@
 List of mistakes:
-- 17: who -> which; missing comma after swT
-- 18: superfluous comma
-- 26: compilers or interpreters for programming languages
-- 27: superfluous comma
-- 32: superfluous comma
-- 33: Western
-- 43: skip commas around parentheses; `with' is not an optimal conjuction
-- Figure 1: do no start sentences with symbols; add an `and' before S
-- 47: Boolean Formulas
-- 51--53: the jump from automata to grammars is uneasy and not really
-  necessary, as the paper mainly deals with automata
-- 57: weighted automata don't have ASTs, but runs
-- 60: other approaches use multioperator monoids or valuation monoids, so `at
-  minima a semiring' is wrong
-- 70: the second comma is superfluous
-- 74: missing period and space after \oplus
-- 75: maybe cite Parsing as intersection by Nederhof and Satta (2003)
-- 77: missing space after period
-- 84: convenient in this context -> for convenience?
-- 87: superfluous comma
-- 116: \oplus `combines two values into a single one' as well (in fact, every
-  binary operation does)
-- 118--119: I would say: … \le_\oplus which is called the natural ordering of
-  S and is, for every …, defined by …
-  Also, you should be consistent with using `for all' or `for every' in
-  quantifications (I would prefer for every)
-- 122: if it IS; I'd also add a comma after total
-- 123: Lemma 2 is Lemma 2 in [21], you should refer to it
-- 125: you write `then', but the following sentence seems to be disconnected
-  from the previous lemma
-- 126: you should cite Knuth's generalization of Dijkstra's algorithm for
-  superiority, he defines the notion for functions of arbitrary arity, so not
-  just semirings (but multioperator monoids); you should give an example for
-  superiority in the semiring notation (with \otimes)
-  Moreover, the `corresponds' is wrong, since superiority generalizes
-  non-negative edge weights
-  PS: Knuth's definition of superiority includes monotonicity
-- 128: increaseS
-- 136: what implies?
-- 143: Lemma 4 is Lemma 3 in [21]
-- 146: need infinite sums with \oplus -> need to extend \oplus to infinitely
-  many operands
-- 147: dom is undefined; I can also be ℕ
-- 151: partitionS
-- 152: I'd use plural, i.e. products and sums
-- 153: quantify x and the family (y_i) for i \in ℕ
-- 155--156: alphabets are usually finite and non-empty sets
-- 156: the sentence ends after … which is difficult to spot
-- 159: denoted by; also it seems like \bar\Phi is a (\Sigma \cup \Sigma
-  \times \Delta)-indexed family, why don't you introduce it like this
-- 161--171: use an itemize environment
-- 162--165: I think these should also hold for \Delta, why don't you write
-  this explicitly?
-- 186: calculability -> computability?
-- 188: the definition of \bigoplus_\Delta can be seen to be implicit.  Please
-  fully quantify the alphabets of a label theory
-- 195: references: there is earlier research on symbolic automata, e.g.
-  Veanes (2013)
-- 196: put respectively after output
-- 199: use \colon rather than : after function names (this also applies to
-  several other places in the paper)
-- 203: THE number
-- 208: What does T compute?
-- 216: by convention -> by definition
-- 219: resp.\@
-- 220--221: you mean that a term CONTAINING w_{ij}(…) equal to 0 is ignored
-- 228: three superfluous commas
-- 228--229: with \oplus/\otimes -> in terms of \oplus/\otimes
-- 232: over THE alphabets
-- 237: superfluous commas
-- 238: text in the margin, also it is not clear whether the if refers to the
-  entire line or just the right definition
-- 247: I wouldn't define out just inside parentheses
-- 251: based ON
-- 252: for EVERY
-- 253: replace comma by and
-- 281: Partitioned
-- 284: x is a variable, so it should be typeset in math mode
-- 290: appogGiatura; ornAmental; also, I'd give this explanation already in
-  Example 1 where the appoggiatura is first used
-- 291: I'd at least say once that \Delta_i, \Delta_c, and \Delta_r are
-  pairwise disjoint (or explain the symbol)
-- 297: the extension here is not as in Section 3, there, the different
-  function hat at least similar types
-- 299: every line expands into the margin; moreover, in the rightmost
-  equation of line three, in the application of w_c, p' and q' are swapped
-- 310: topupmost -> topmost
-- 316: equation extends into margin
-- 330--331: as indicated by the stack top: better explain that stack symbols
-  are time intervals
-- 331: I think the application of w_c is wrong (same with w_r and w_i later),
-  shouldn't the : be a comma and what is q in the middle?
-- 336: how are states (which are in Q) pushed onto the stack if \Gamma =
-  \Delta_c \times P?
-- 339: what are j and c, do you mean i and d?
-- 341: shouldn't there be a transition for returning from a measure?
-- 345: exist (singular)
-- 348: for?  do you mean accepted by?  but the language of an swA isn't
-  defined either
-- 352: Dijkstra-like algorithm?
-- 354: readers may have an easier time seeing this argument if you say that S
-  is extremal.  You should say in the preliminaries that S is extremal if and
-  only if \le_\oplus is total.
-- 366: be popPED
-- 371: How is it ensured that the symbol <c,p> isn't popped and pushed again
-  later during the computation?
-- 375: converge (singular)
-- 376: I do not understand the and at termination, and its correctness part.
-  What do you want to express?
-- 379: Almost all update equations need to compute infinite sums.  Don't you
-  think that the assumption of \bar\Phi being effective is pretty strong and
-  gives your the result `for free'?  I think you should at least elaborate on
-  this and say to what degree your effective label theory is a generalization
-  of the decidable label theory of unweighted symbolic automata.
-- 386: We assume the following input to be given
-- 394: the notation sw-parsing is undefined
-- 416: word order: put `to T' in the end
-- 420: `the sw-parsing problem' or `sw-parsing', but not `the sw-parsing'
-- 424: the triple use and does not work well
-- 431: double the
-- 435: there is something missing between `here' and `offline'
-- 629: typical references for regular tree grammars are Brainerd (1969) and
-  Gecseg and Steinby (1984)
-- 647: inn -> in
-- 650: missing period after function
-- 652: I think \bar w should be \bar w'
-- 654: an intuition for the \bar q would have made understanding the
-  construction far easier
+- [ ] 17: who -> which; missing comma after swT
+- [ ] 18: superfluous comma
+- [ ] 26: compilers or interpreters for programming languages
+- [ ] 27: superfluous comma
+- [ ] 32: superfluous comma
+- [ ] 33: Western
+- [ ] 43: skip commas around parentheses; `with' is not an optimal conjuction 
+- [ ] Figure 1: do no start sentences with symbols; add an `and' before $S$
+- [ ] 47: Boolean Formulas
+- [ ] 51--53: the jump from automata to grammars is uneasy and not really necessary, as the paper mainly deals with automata
+- [ ] 57: weighted automata don't have ASTs, but runs
+- [ ] 60: other approaches use multioperator monoids or valuation monoids, so `at minima a semiring' is wrong
+- [ ] 70: the second comma is superfluous
+- [ ] 74: missing period and space after $\oplus$
+- [ ] 75: maybe cite Parsing as intersection by Nederhof and Satta (2003)
+- [ ] 77: missing space after period
+- [ ] 84: convenient in this context -> for convenience?
+- [ ] 87: superfluous comma
+- [ ] 116: \oplus `combines two values into a single one' as well (in fact, every   binary operation does)
+- [ ] 118--119: I would say: … $\leq_\oplus$ which is called the natural ordering of $S$ and is, for every …, defined by …
+- [ ] Also, you should be consistent with using  'for all' or `for every' in quantifications (I would prefer for every)
+
+- [ ] 122: if it IS; I'd also add a comma after total
+- [ ] 123: Lemma 2 is Lemma 2 in [21], you should refer to it
+- [ ] 125: you write `then', but the following sentence seems to be disconnected from the previous lemma
+- [ ] 126: you should cite Knuth's generalization of Dijkstra's algorithm for superiority, he defines the notion for functions of arbitrary arity, so not just semirings (but multioperator monoids); you should give an example for superiority in the semiring notation (with $\otimes$)
+  Moreover, the `corresponds' is wrong, since superiority generalizes non-negative edge weights
+- [ ] PS: Knuth's definition of superiority includes monotonicity
+
+- [ ] 128: increaseS
+- [ ] 136: what implies?
+- [ ] 143: Lemma 4 is Lemma 3 in [21]
+- [ ] 146: need infinite sums with \oplus -> need to extend \oplus to infinitely many operands
+- [ ] 147: dom is undefined; I can also be ℕ
+- [ ] 151: partitionS
+- [ ] 152: I'd use plural, i.e. products and sums
+- [ ] 153: quantify $x$ and the family ($y_i$) for $i ∈ ℕ$
+- [ ] 155--156: alphabets are usually finite and non-empty sets
+- [ ] 156: the sentence ends after … which is difficult to spot
+- [ ] 159: denoted by; also it seems like $\bar\Phi$ is a ($\Sigma \cup \Sigma \times \Delta$)-indexed family, why don't you introduce it like this
+- [ ] 161--171: use an itemize environment
+- [ ] 162--165: I think these should also hold for $\Delta$, why don't you write this explicitly?
+- [ ] 186: calculability -> computability?
+- [ ] 188: the definition of $\bigoplus_\Delta$ can be seen to be implicit.  Please fully quantify the alphabets of a label theory
+- [ ] 195: references: there is earlier research on symbolic automata, e.g. Veanes (2013)
+- [ ] 196: put respectively after output
+- [ ] 199: use \colon rather than : after function names (this also applies to several other places in the paper)
+- [ ] 203: THE number
+- [ ] 208: What does $T$ compute?
+- [ ] 216: by convention -> by definition
+- [ ] 219: resp.\@
+- [ ] 220--221: you mean that a term CONTAINING $w_{ij}(…)$ equal to 0 is ignored
+- [ ] 228: three superfluous commas
+- [ ] 228--229: with $\oplus/\otimes$ -> in terms of $\oplus/\otimes$
+- [ ] 232: over THE alphabets
+- [ ] 237: superfluous commas
+- [ ] 238: text in the margin, also it is not clear whether the if refers to the entire line or just the right definition
+- [ ] 247: I wouldn't define out just inside parentheses
+- [ ] 251: based ON
+- [ ] 252: for EVERY
+- [ ] 253: replace comma by and
+- [ ] 281: Partitioned
+- [ ] 284: $x$ is a variable, so it should be typeset in math mode
+- [ ] 290: appogGiatura; ornAmental; also, I'd give this explanation already in Example 1 where the appoggiatura is first used
+- [ ] 291: I'd at least say once that $\Delta_i$, $\Delta_c$, and $\Delta_r$ are pairwise disjoint (or explain the symbol)
+- [ ] 297: the extension here is not as in Section 3, there, the different function hat at least similar types
+- [ ] 299: every line expands into the margin; moreover, in the rightmost equation of line three, in the application of $w_c$, $p'$ and $q'$ are swapped
+- [ ] 310: topupmost -> topmost
+- [ ] 316: equation extends into margin
+- [ ] 330--331: as indicated by the stack top: better explain that stack symbols are time intervals
+- [ ] 331: I think the application of $w_c$ is wrong (same with $w_r$ and $w_i$ later), shouldn't the : be a comma and what is q in the middle?
+- [ ] 336: how are states (which are in $Q$) pushed onto the stack if $\Gamma = \Delta_c \times P$?
+- [ ] 339: what are $j$ and $c$, do you mean $i$ and $d$?
+- [ ] 341: shouldn't there be a transition for returning from a measure?
+- [ ] 345: exist (singular)
+- [ ] 348: for?  do you mean accepted by?  but the language of an swA isn't defined either
+- [ ] 352: Dijkstra-like algorithm?
+- [ ] 354: readers may have an easier time seeing this argument if you say that $S$ is extremal.  You should say in the preliminaries that S is extremal if and only if $\leq_\oplus$ is total.
+- [ ] 366: be popPED
+- [ ] 371: How is it ensured that the symbol <c,p> isn't popped and pushed again later during the computation?
+- [ ] 375: converge (singular)
+- [ ] 376: I do not understand the and at termination, and its correctness part.  What do you want to express?
+- [ ] 379: Almost all update equations need to compute infinite sums.
+Don't you think that the assumption of $\bar\Phi$ being effective is pretty strong and gives your the result `for free'?
+I think you should at least elaborate on this and say to what degree your effective label theory is a generalization of the decidable label theory of unweighted symbolic automata.
+- [ ] 386: We assume the following input to be given
+- [ ] 394: the notation sw-parsing is undefined
+- [ ] 416: word order: put `to T' in the end
+- [ ] 420: "the sw-parsing problem" or "sw-parsing", but not "the sw-parsing"
+- [ ] 424: the triple use and does not work well
+- [ ] 431: double the
+- [ ] 435: there is something missing between "here" and "offline"
+- [ ] 629: typical references for regular tree grammars are Brainerd (1969) and  Gecseg and Steinby (1984)
+- [ ] 647: inn -> in
+- [ ] 650: missing period after function
+- [ ] 652: I think $\bar w$ should be $\bar w'$
+- [ ] 654: an intuition for the $\bar q$ would have made understanding the construction far easier
 
@@
 
-Line 209:  "T is based on an intermediate function ...": should this be "The semantics of T is based on an intermediate function ..."?
+- [ ] Line 209:  "T is based on an intermediate function ...": should this be "The semantics of $T$ is based on an intermediate function ..."?
 
-Line 281: "artitioned into three" -> "partitioned into three"
+- [ ] Line 281: "artitioned into three" -> "partitioned into three"
 
-Last line of algorithm 1: What exactly is meant by "update d_\bot with <q_1,q_2>"?
-Does it mean update the image of <q_1,q_2> under d_\bot with the new value as computed in Figure 3?
+- [ ] Last line of algorithm 1: What exactly is meant by "update $d_⊥$ with $<q_1,q_2>$"?
+Does it mean update the image of $<q_1,q_2>$ under $d_⊥$ with the new value as computed in Figure 3?
 Write this clearly.
 
-Figure 3: This figure is supposed to compute <q_1,q_2>.
-First of all, it is not clear what is meant by "compute <q_1,q_2>"; I am guessing, as I wrote in the comment above that the computed quantity is the image of <q_1,q_2> under d_\bot.
-However, even this doesn't make sense; there are some series of equantions written, for d_\top(q_1,p,q_3) (not for d_\top(q_1,p,q_2).
-Even in this equation, what is q_2?
-Do we take the minimum (wrt \oplus) over all q_2?
-The second line says d_\bot(q_1,p,q_3), which doesn't make sense because d_\bot takes two arguments, not three.
-There are two equations for d_\top(q_1,p,q_3) and one for d_top(q_0,p,q_3).
+Figure 3: This figure is supposed to compute $<q_1,q_2>$. 
+- [ ] First of all, it is not clear what is meant by "compute $<q_1,q_2>$"; I am guessing, as I wrote in the comment above that the computed quantity is the image of $<q_1,q_2>$ under $d_⊥$.
+However, even this doesn't make sense; there are some series of equantions written, for $d_\top(q_1,p,q_3)$ (not for $d_\top(q_1,p,q_2)$.
+- [ ] Even in this equation, what is $q_2$? Do we take the minimum (wrt $\oplus$) over all $q_2$? 
+- [ ] The second line says $d_\bot(q_1,p,q_3)$, which doesn't make sense because $d_⊥$ takes two arguments, not three. 
+- [ ] There are two equations for $d_\top(q_1,p,q_3)$ and one for $d_\top(q_0,p,q_3)$.
 How are these related?
@@
 
-Introduction:
+**Introduction**:
 
- The literature review should also include work by Hermann and Vogler, who bring in weights into symbolic automata as well, as well as stacks and other types of data storage, though their focus is more on exploring theoretical properties than empirical implementation, and make no complexity claims, as far as I can see.
+- [ ] The literature review should also include work by Hermann and Vogler, who bring in weights into symbolic automata as well, as well as stacks and other types of data storage, though their focus is more on exploring theoretical properties than empirical implementation, and make no complexity claims, as far as I can see.
 (Herrmann, Luisa, and Heiko Vogler.
@@
 On lines 72+, the authors refer to Mohri in saying that "[...]given an input word $s$, the _SW-parsing_ problem aims at finding $t$ minimizing $T(s,t) \times A(t)$, called the distance between $s$ and $A$[...]".
-It is not entirely clear to me what part of this expression they consider to be the distance as described by Mohri.
+
+- [ ] It is not entirely clear to me what part of this expression they consider to be the distance as described by Mohri.
 Is it the distance, as defined by $T$, from the $s$ to the distribution over ASTs $t$ defined by $A$?
@@
 
-Preliminary Notions:
+**Preliminary Notions**:
 
- The introduction of Label theories (l 154++) does not depend on its first paragraph containing the definitions of words and alphabets, as far as I can tell, so I would let this first paragraph sit under its own title of e.g.
+- [ ] The introduction of Label theories (l.
+154++) does not depend on its first paragraph containing the definitions of words and alphabets, as far as I can tell, so I would let this first paragraph sit under its own title of e.g.
 "Alphabets and words".
-It is also, separately, not clear to me what the intuition and motivation is for using a label theory, at least at this point in the paper.
+- [ ] It is also, separately, not clear to me what the intuition and motivation is for using a label theory, at least at this point in the paper.
 In particular, I am unsure why any of the specific closures in ll 161+ are considered necessary.
@@
 
-SW Automata and Transducers
+**SW Automata and Transducers**
 
- A brief explanation of what you consider Mohri's approach to the computation of distances between words and languages may be helpful.
+- [ ] A brief explanation of what you consider Mohri's approach to the computation of distances between words and languages may be helpful.
 
- The paragraph starting at l 224 is a great summary of the way you consider edit-distances in this paper.
+The paragraph starting at line 224 is a great summary of the way you consider edit-distances in this paper.
 
- In Example 8, you disallow the last note being an error - is there some specific justification for that, or is it just for simplicity?
+- [ ] In Example 8, you disallow the last note being an error - is there some specific justification for that, or is it just for simplicity?
 
-SW Visibly Pushdown Automata
+**SW Visibly Pushdown Automata**
 
@@
 This makes the specifics of the example hard to double check, and to properly understand.
-It is also somewhat difficult to separate the symbols that form part of the language from those of the meta-language in the specifications of the various $w$ function calls.
-Additionally, you write "the above transition pushes the state $q_{i + 1/c}$ on[to] the stack" - I would encourage you to choose different letters and terminology than "state $q$" for stack symbols, to reduce the risk for the reader confusing the stack symbols and states.
-In fact, it may be a good idea to show a fuller schematic of a few steps of the sw-VPA computation, with relevant connections between various variables highlighted, what equalities need to hold etc.
 
- On line 366 of Proposition 16, the "can be" should be "must be", surely, since what you are describing is a computation that should leave the original stack (exactly) untouched?
+- [ ] It is also somewhat difficult to separate the symbols that form part of the language from those of the meta-language in the specifications of the various $w$ function calls. 
+- [ ] Additionally, you write "the above transition pushes the state $q_{i + 1/c}$ on[to] the stack" - I would encourage you to choose different letters and terminology than "state $q$" for stack symbols, to reduce the risk for the reader confusing the stack symbols and states.
+- [ ] In fact, it may be a good idea to show a fuller schematic of a few steps of the sw-VPA computation, with relevant connections between various variables highlighted, what equalities need to hold etc.
 
- It seems to me that there is a consistent lack of quantification over $P$ in the section starting with the definition of $b_\top$.
+- [ ] On line 366 of Proposition 16, the "can be" should be "must be", surely, since what you are describing is a computation that should leave the original stack (exactly) untouched?
+- [ ]  It seems to me that there is a consistent lack of quantification over $P$ in the section starting with the definition of $b_\top$.
 This is not hugely problematic for the proof, since the algorithm is still defined (mostly) properly, and the complexity should only be impacted by a factor of something like $|P|$.
 
-Minor comments:
+**Minor comments**:
 
- l 87. I would front "input" in the sentence, yielding "a given input _timeline_ of [...] alphabet $\Sigma$, is parsed into[...]"
+- [ ] l. 87. I would front "input" in the sentence, yielding "a given input _timeline_ of [...] alphabet $\Sigma$, is parsed into[...]"
 
- l 90. Please parenthesize the pitch-timecode pairs for readability. As an aside: these are strictly the onsets? And strictly monophonic?
+- [ ] l. 90. Please parenthesize the pitch-timecode pairs for readability. As an aside: these are strictly the onsets? And strictly monophonic?
 
- l 100. Shouldn't the first measure end before the second starts? This would make $O$ a forest rather than a tree - easily fixed with a "piece" root element.
+- [ ] l. 100. Shouldn't the first measure end before the second starts? This would make $O$ a forest rather than a tree - easily fixed with a "piece" root element.
 
- l 156.
+- [ ] l.
+156.
 The ellipsis between $\Delta$ and $\Sigma^*$ confuses the sentence boundaries for the reader.
@@
 
- l 220.
+- [ ] l.
+220.
 The sentence starting here is somewhat confusing (and possibly wrong, depending on if $q'$ and $q''$ are intended to be distinct).
@@
 
- l 240: an fixed weight -> a fixed weight
+- [ ] l. 240: an fixed weight -> a fixed weight
 
- l 251: based with an intermediate -> based on an intermediate
+- [ ] l. 251: based with an intermediate -> based on an intermediate
 
- l 276: generalizing themselves -> themselves generalizing
+- [ ] l. 276: generalizing themselves -> themselves generalizing
 
- l 281: artitioned -> partitioned
+- [ ] l. 281: artitioned -> partitioned
 
- l 288: same as on l 100, measure 1 should maybe end before measure 2
+- [ ] l. 288: same as on line 100, measure 1 should maybe end before measure 2
 
- l 290: ornemental -> ornamental
+- [ ] l. 290: ornemental -> ornamental
 
- l 306: reads and pop from stack -> reads and pops from the stack
+- [ ] l. 306: reads and pop from stack -> reads and pops from the stack
 
- l 307: I would use "computations" or something like that rather than "transitions" here
+- [ ] l. 307: I would use "computations" or something like that rather than "transitions" here
 
- l 310: topupmost -> topmost, though I'm not sure if it's necessary to describe the basic workings of a stack. If you choose to do so, then also
+- [ ] l. 310: topupmost -> topmost, though I'm not sure if it's necessary to describe the basic workings of a stack. If you choose to do so, then also
 
- l 311: content -> pair
+- [ ] l. 311: content -> pair
 
- l 323: particular -> special
+- [ ] l. 323: particular -> special
 
- l 365: "case of in the" -> either "case in the" or "case of the"
+- [ ] l. 365: "case of in the" -> either "case in the" or "case of the"
 
- l 366: pop -> popped
+- [ ] l. 366: pop -> popped
 
@@
 
- l 395: I would omit "output".
+- [ ] l.
+395: I would omit "output".
 If you want to draw attention to the different alphabets involved, I would say "between the string $s$ over the input alphabet, and the weighted language over the output alphabet defined by $A$"
 
- l 399: correspondance -> correspondence
+- [ ] l. 399: correspondance -> correspondence
 
- l 402: "weight a second time division" -> something like "weight assigned to an additional second time division"
+- [ ] l. 402: "weight a second time division" -> something like "weight assigned to an additional second time division"
 
- l 416: extra commen between "$T$" and "of new transition"
+- [ ] l. 416: extra commen between "$T$" and "of new transition"
 
- l 417: Corolary -> Corollary
+- [ ] l. 417: Corolary -> Corollary
 
- l 420: This whole paragraph should be shaped up, writing wise.
+- [ ] l. 420: This whole paragraph should be shaped up, writing wise.
 
- l 434: $swM$??
+- [ ] l. 434: $swM$??
 
- l 435: "offline" -> "works offline"
+- [ ] l. 435: "offline" -> "works offline"
 
- l 437: extraneous comma after "such as"
+- [ ] l. 437: extraneous comma after "such as"
 

### 2021-08-31 FST-TCS/rebuttal/rebuttal.md
update rebuttal notes

@@
 > date: 30 August 2021
->
 > from: fsttcs2021@easychair.org
@@
 - [ ] 87: superfluous comma
-- [ ] 116: \oplus `combines two values into a single one' as well (in fact, every   binary operation does)
+- [ ] 116: $\oplus$ `combines two values into a single one' as well (in fact, every   binary operation does)
 - [ ] 118--119: I would say: … $\leq_\oplus$ which is called the natural ordering of $S$ and is, for every …, defined by …
@@
 - [ ] 366: be popPED
-- [ ] 371: How is it ensured that the symbol <c,p> isn't popped and pushed again later during the computation?
+- [ ] 371: How is it ensured that the symbol $<c,p>$ isn't popped and pushed again later during the computation?
 - [ ] 375: converge (singular)
@@
 - [ ] 386: We assume the following input to be given
-- [ ] 394: the notation sw-parsing is undefined
-- [ ] 416: word order: put `to T' in the end
+- [ ] 394: the notation "sw-parsing" is undefined
+- [ ] 416: word order: put "to $T$" in the end
 - [ ] 420: "the sw-parsing problem" or "sw-parsing", but not "the sw-parsing"
@@
 
-On lines 72+, the authors refer to Mohri in saying that "[...]given an input word $s$, the _SW-parsing_ problem aims at finding $t$ minimizing $T(s,t) \times A(t)$, called the distance between $s$ and $A$[...]".
+On lines 72+, the authors refer to Mohri in saying that "[...] given an input word $s$, the _SW-parsing_ problem aims at finding $t$ minimizing $T(s,t) \times A(t)$, called the distance between $s$ and $A$[...]".
 
@@
 - [ ]  Ref [22] and [23] both seem to refer to the same work by Mohri
-
-------------------------------------------------------

### 2021-08-31 FST-TCS/rebuttal/response.txt
1st draft response after discussion

We would like to thank the reviewers for their valuable work, 
and their comments and references/suggestion.

In particular, we are grateful to them for pointing us to the work of L. Herrmann & Heiko... in DLT 2016 & PhD 2020.  
Indeed, the models that we are using in the paper are particular (restricted) cases of the Weighted Symbolic Automata with Storage defined in DLT 2016 & PhD 2020.
We agree that it should definitely be explained at the beginning of the paper, and the 2 references must be cited as well.

shift focus to the 2 main contributions of the paper are actually the following:
- a best-search algorithm
  (as far as we know no such algorithm has been proposed for swVPA)
- a framework for parsing over infinite alphabet and with tree-automata. 
  in particular the use of weighted transducers for def. of distance and of VPA for computing on linearization of parse trees.

About the parsing : rem. reviewer1 on "previous approaches" which are more general.
Our work is also oriented toward the formalisation of an empirical problem, the Automated Music Transcription (in parallel to the development of a tool).

Regarding the best-search algorithm, 
answer to reviewer 2:... (DETAILS)

questions of reviewer 3 about the distance and the work of Mohri.

other rem. review 1:
the relationship between VPA and TA has been discussed in several places, 
**we did [or not?]** find an explicit construction of VPA from TA.
could be move to paper if you think that it is original.
[ ] check in paper [5]

### 2021-09-01 FST-TCS/rebuttal/response.txt
1st draft response after discussion

@@
 
-shift focus to the 2 main contributions of the paper are actually the following:
+the 2 main contributions of the paper are the following:
 - a best-search algorithm
@@
 
-About the parsing : rem. reviewer1 on "previous approaches" which are more general.
-Our work is also oriented toward the formalisation of an empirical problem, the Automated Music Transcription (in parallel to the development of a tool).
+Although this only mentioned quickly in the paper, our main motivation / purpose is a formal study of an empirical (challenging) problem : Automated Music Transcription, based on consistent theoretical framework made of formal language models / formalisms.
+in parallel to developpement of a transcription tool
+so far results show that theoretical… improve …
+
+About the weighted parsing : rem. reviewer1 on "previous approaches" which are more general.
 

### 2021-09-01 FST-TCS/rebuttal/response.txt
draft response

@@
 We agree that it should definitely be explained at the beginning of the paper, and the 2 references must be cited as well. 
+however, It seems difficult to completely skip definitions of the models, for self-completeness reasons.
 
-the 2 main contributions of the paper are the following:
+also agree to focus the paper on its 2 main contributions, which are the following:
 - a best-search algorithm
@@
 in parallel to developpement of a transcription tool
-so far results show that theoretical… improve …
-
-About the weighted parsing : rem. reviewer1 on "previous approaches" which are more general.
+so far results have shown that theoretical foundations improve accuracy…
 
@@
 
+About the weighted parsing : rem. reviewer1 on "previous approaches" which are more general.
+
 other rem. review 1:

### 2021-09-01 FST-TCS/rebuttal/response.txt
draft response

--- 2021-08-31 FST-TCS/rebuttal/response.txt
@@
 
-shift focus to the 2 main contributions of the paper are actually the following:
+However, It seems difficult to completely skip definitions of the models, for self-completeness reasons.
+definition of models less general sufficient for our purpose
+not clear whether Prop. 10 (in center pages) is a direct consequence of Hermann V. 2016...
+
+also agree to focus the paper on its 2 main contributions, which are the following:
 - a best-search algorithm
-  (as far as we know no such algorithm has been proposed for swVPA)
+  (as far as we know this is a new result
+  no such algorithm has been proposed for swVPA)
 - a framework for parsing over infinite alphabet and with tree-automata. 
@@
 
-About the parsing : rem. reviewer1 on "previous approaches" which are more general.
-Our work is also oriented toward the formalisation of an empirical problem, the Automated Music Transcription (in parallel to the development of a tool).
+second is new application 
+Moreover, 
+Although this only mentioned quickly in the paper, our main motivation / purpose is a formal study of an empirical (challenging) problem : Automated Music Transcription, based on consistent theoretical framework made of formal language models / formalisms.
+in parallel to developpement of a transcription tool
+so far results have shown that theoretical foundations improve accuracy…
 

### 2021-09-01 FST-TCS/rebuttal/response.txt
typoes in algorithm + 1 ref. Lemma 4

@@
-We would like to thank the reviewers for their valuable work, 
-and their comments and references/suggestion.
+We would like to thank the reviewers for their hard work, and all their comments and references/suggestions which are very useful to us.
 
 In particular, we are grateful to them for pointing us to the work of L. Herrmann & Heiko... in DLT 2016 & PhD 2020.  
-Indeed, the models that we are using in the paper are particular (restricted) cases of the Weighted Symbolic Automata with Storage defined in DLT 2016 & PhD 2020.
+Indeed, the models that we are using in the paper are particular (and strictly restricted) cases of the Weighted Symbolic Automata with Storage defined in DLT 2016 & PhD 2020. 
 We agree that it should definitely be explained at the beginning of the paper, and the 2 references must be cited as well.
@@
 answer to reviewer 2:... (DETAILS)
+update: intuitively, we assume that values of d_bot and d_top known for every tuplets extracted from the queue Q, 
+and use these values to update the other values of the functions d_bot and d_top .
 

### 2021-09-01 main.tex
typos algo best
passe appendix
typoes in algorithm + 1 ref. Lemma 4

@@
 %\textbf{for all} $q_0, q_3 \in Q$
-\Update $d_\bot$ with $\< q_1, q_2>$ or $d_\top$ with $\< q_1, p, q_2>$ 
-(Figure~\ref{fig:best-update}).$\;$
+\Update $d_\bot$ if $\< q_1, q_2>$ was extracted, or $d_\top$ 
+if it was $\< q_1, p, q_2>$, 
+as in Figure~\ref{fig:best-update}.$\;$
 }
@@
 \begin{figure}[ht]
-For all $q_0, q_3 \in Q$, %$p \in P$,
+For all $q_0, q_3 \in Q$, for all $p' \in P$,
 \[
@@
 %
-d_\bot(q_1, p, q_3) & \opluseq &
+d_\bot(q_1, q_3) & \opluseq &
   d_\bot(q_1, q_2) \otimes
@@
 %
-d_\top(q_0, p, q_3) & \opluseq &
+d_\top(q_0, p', q_3) & \opluseq &
   {\displaystyle\bigoplus_{\Deltac}}^2
-  \bigl[ \bigl( \weic(q_0, p, p', q_1) \otimes_2
-  d_\top(q_1, p', q_2) \bigr) \otimes_2
-  {\displaystyle\bigoplus_{\Deltar}} \weir(q_2, p', q_3) \bigr]\\
+  \bigl[ \bigl( \weic(q_0, p', q_1, p) \otimes_2
+  d_\top(q_1, p, q_2) \bigr) \otimes_2
+  {\displaystyle\bigoplus_{\Deltar}} \weir(q_2, p, q_3) \bigr]\\
 %\multicolumn{3}{l}{\quad
@@
   d_\top(q_1, p, q_2) \otimes d_\top(q_2, p, q_3),
-  \mathrm{if~} \< q_2, \top, q_3> \notin P \\
+  \mathrm{if~} \< q_2, p, q_3> \notin \Q \\
 %
 d_\bot(q_1, q_3) & \opluseq &
-  d_\bot(q_1, q_2) \otimes d_\bot(q_2, q_3), \mathrm{if~} \< q_2, \bot, q_3> \notin P \\
+  d_\bot(q_1, q_2) \otimes d_\bot(q_2, q_3), \mathrm{if~} \< q_2, q_3> \notin \Q \\
 \end{array}
@@
 %$x_c \mapsto \weic(q_0, p, x_c, c', p', q_1)$.
-\caption{Update $d_\bot$ with $\<q_1, q_2>$ or $d_\top$ with $\< q_1, p, q_2>$.}
+%\caption{Update $d_\bot$ with $\<q_1, q_2>$ or $d_\top$ with $\< q_1, p, q_2>$.}
+\caption{Update $d_\bot$ or $d_\top$.}
 \label{fig:best-update}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\listoftodos
+%\newpage
+%\listoftodos
 

### 2021-09-01 semiring.tex
proof property closure SWA byb SWT
revise examples. 1 passe
passe section prelim
typoes in algorithm + 1 ref. Lemma 4

@@
    and $\< \Semiring, \otimes, \one>$ a monoid,
-\item $\otimes$ distributes over~$\oplus$:  $\forall x, y, z \in \mathbb{S}$,
+\item $\otimes$ distributes over~$\oplus$: \\ $\forall x, y, z \in \mathbb{S}$,
 $x \otimes (y \oplus z) = (x \otimes y) \oplus (x \otimes z)$,
@@
 %and let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a {semiring},
-A semiring $\Semiring$ is \emph{commutative} if $\otimes$ is commutative.
+A semiring~$\Semiring$ is \emph{commutative} if $\otimes$ is commutative.
 It is \emph{idempotent} if for all $x \in \Semiring$, $x \oplus x = x$.
@@
 \ie when for all $x, y \in \Semiring$, either $x \oplus y = x$ or $x \oplus y = y$.
-\florent{is total necessary?}
+%\florent{is total necessary?}
 
@@
 
-\begin{lemma}
+\begin{lemma}[\cite{Mohri02semiring}, Lemma~3]\label{lem:idempotent}
 Every bounded semiring is idempotent.

### 2021-09-01 FST-TCS/rebuttal/response.txt
response cont.

@@
 
-In particular, we are grateful to them for pointing us to the work of L. Herrmann & Heiko... in DLT 2016 & PhD 2020.  
+In particular, we are grateful to them for pointing us to the work of Luisa Herrmann & Heiko Vögler in DLT 2016 & PhD 2020.  
 Indeed, the models that we are using in the paper are particular (and strictly restricted) cases of the Weighted Symbolic Automata with Storage defined in DLT 2016 & PhD 2020. 
@@
 
-also agree to focus the paper on its 2 main contributions, which are the following:
+we plan to rephrase the presentation at the begining of the paper, in order to focus on the 2 following contributions:
 - a best-search algorithm
@@
   in particular the use of weighted transducers for def. of distance and of VPA for computing on linearization of parse trees.
+  to our knowledge, this second contribution is new application of symbolic and weighted models.
 
-second is new application 
-Moreover, 
 Although this only mentioned quickly in the paper, our main motivation / purpose is a formal study of an empirical (challenging) problem : Automated Music Transcription, based on consistent theoretical framework made of formal language models / formalisms.

### 2021-09-01 FST-TCS/rebuttal/rebuttal.md
review typos

@@
 List of mistakes:
-- [ ] 17: who -> which; missing comma after swT
-- [ ] 18: superfluous comma
-- [ ] 26: compilers or interpreters for programming languages
-- [ ] 27: superfluous comma
-- [ ] 32: superfluous comma
-- [ ] 33: Western
-- [ ] 43: skip commas around parentheses; `with' is not an optimal conjuction 
-- [ ] Figure 1: do no start sentences with symbols; add an `and' before $S$
-- [ ] 47: Boolean Formulas
+- [x] 17: who -> which; missing comma after swT
+- [x] 18: superfluous comma
+- [x] 26: compilers or interpreters for programming languages
+- [x] 27: superfluous comma
+- [x] 32: superfluous comma
+- [x] 33: Western
+- [x] 43: skip commas around parentheses; `with' is not an optimal conjuction 
+- [x] Figure 1: do no start sentences with symbols; add an `and' before $S$
+- [x] 47: Boolean Formulas
 - [ ] 51--53: the jump from automata to grammars is uneasy and not really necessary, as the paper mainly deals with automata
@@
 - [ ] 60: other approaches use multioperator monoids or valuation monoids, so `at minima a semiring' is wrong
-- [ ] 70: the second comma is superfluous
-- [ ] 74: missing period and space after $\oplus$
+- [x] 70: the second comma is superfluous
+- [x] 74: missing period and space after $\oplus$
 - [ ] 75: maybe cite Parsing as intersection by Nederhof and Satta (2003)
-- [ ] 77: missing space after period
-- [ ] 84: convenient in this context -> for convenience?
-- [ ] 87: superfluous comma
+- [x] 77: missing space after period
+- [x] 84: convenient in this context -> for convenience?
+- [x] 87: superfluous comma
 - [ ] 116: $\oplus$ `combines two values into a single one' as well (in fact, every   binary operation does)
-- [ ] 118--119: I would say: … $\leq_\oplus$ which is called the natural ordering of $S$ and is, for every …, defined by …
-- [ ] Also, you should be consistent with using  'for all' or `for every' in quantifications (I would prefer for every)
+- [x] 118--119: I would say: … $\leq_\oplus$ which is called the natural ordering of $S$ and is, for every …, defined by …
+- [x] Also, you should be consistent with using  'for all' or `for every' in quantifications (I would prefer for every)
 
-- [ ] 122: if it IS; I'd also add a comma after total
-- [ ] 123: Lemma 2 is Lemma 2 in [21], you should refer to it
-- [ ] 125: you write `then', but the following sentence seems to be disconnected from the previous lemma
+- [x] 122: if it IS; I'd also add a comma after total
+- [x] 123: Lemma 2 is Lemma 2 in [21], you should refer to it
+- [x] 125: you write `then', but the following sentence seems to be disconnected from the previous lemma
 - [ ] 126: you should cite Knuth's generalization of Dijkstra's algorithm for superiority, he defines the notion for functions of arbitrary arity, so not just semirings (but multioperator monoids); you should give an example for superiority in the semiring notation (with $\otimes$)
@@
 
-- [ ] 128: increaseS
+- [x] 128: increaseS
 - [ ] 136: what implies?
 - [ ] 143: Lemma 4 is Lemma 3 in [21]
-- [ ] 146: need infinite sums with \oplus -> need to extend \oplus to infinitely many operands
-- [ ] 147: dom is undefined; I can also be ℕ
-- [ ] 151: partitionS
-- [ ] 152: I'd use plural, i.e. products and sums
+- [x] 146: need infinite sums with \oplus -> need to extend \oplus to infinitely many operands
+- [x] 147: dom is undefined; I can also be ℕ
+- [x] 151: partitionS
+- [x] 152: I'd use plural, i.e. products and sums
 - [ ] 153: quantify $x$ and the family ($y_i$) for $i ∈ ℕ$
 - [ ] 155--156: alphabets are usually finite and non-empty sets
-- [ ] 156: the sentence ends after … which is difficult to spot
+- [x] 156: the sentence ends after … which is difficult to spot
 - [ ] 159: denoted by; also it seems like $\bar\Phi$ is a ($\Sigma \cup \Sigma \times \Delta$)-indexed family, why don't you introduce it like this
@@
 - [ ] 162--165: I think these should also hold for $\Delta$, why don't you write this explicitly?
-- [ ] 186: calculability -> computability?
+- [x] 186: calculability -> computability?
 - [ ] 188: the definition of $\bigoplus_\Delta$ can be seen to be implicit.  Please fully quantify the alphabets of a label theory
 - [ ] 195: references: there is earlier research on symbolic automata, e.g. Veanes (2013)
-- [ ] 196: put respectively after output
-- [ ] 199: use \colon rather than : after function names (this also applies to several other places in the paper)
-- [ ] 203: THE number
+- [x] 196: put respectively after output
+- [x] 199: use \colon rather than : after function names (this also applies to several other places in the paper)
+- [x] 203: THE number
 - [ ] 208: What does $T$ compute?
-- [ ] 216: by convention -> by definition
+- [x] 216: by convention -> by definition
 - [ ] 219: resp.\@
 - [ ] 220--221: you mean that a term CONTAINING $w_{ij}(…)$ equal to 0 is ignored
-- [ ] 228: three superfluous commas
-- [ ] 228--229: with $\oplus/\otimes$ -> in terms of $\oplus/\otimes$
-- [ ] 232: over THE alphabets
-- [ ] 237: superfluous commas
+- [x] 228: three superfluous commas
+- [x] 228--229: with $\oplus/\otimes$ -> in terms of $\oplus/\otimes$
+- [x] 232: over THE alphabets
+- [x] 237: superfluous commas
 - [ ] 238: text in the margin, also it is not clear whether the if refers to the entire line or just the right definition
-- [ ] 247: I wouldn't define out just inside parentheses
-- [ ] 251: based ON
-- [ ] 252: for EVERY
-- [ ] 253: replace comma by and
-- [ ] 281: Partitioned
+- [x] 247: I wouldn't define out just inside parentheses
+- [x] 251: based ON
+- [x] 252: for EVERY
+- [x] 253: replace comma by and
+- [x] 281: Partitioned
 - [ ] 284: $x$ is a variable, so it should be typeset in math mode
-- [ ] 290: appogGiatura; ornAmental; also, I'd give this explanation already in Example 1 where the appoggiatura is first used
+- [x] 290: appogGiatura; ornAmental; also, I'd give this explanation already in Example 1 where the appoggiatura is first used
 - [ ] 291: I'd at least say once that $\Delta_i$, $\Delta_c$, and $\Delta_r$ are pairwise disjoint (or explain the symbol)

### 2021-09-01 introduction.tex
review typos

@@
 NLP tasks such as constituency parsing, 
-programming languages compilers or interpreters.
+compilers or interpreters for programming languages.
 Considering large or infinite alphabets can however be of
-practical interest, when dealing with large characters encodings such as UTF-16~\cite{dAntoni21CACM},
+practical interest when dealing with large characters encodings such as UTF-16~\cite{dAntoni21CACM},
 % processing strings in \eg for vulnerability detection in Web-applications~\cite{dAntoni21CACM},
@@
 are essentially linear:
-audio files, or the widely used
+audio files or the widely used
 MIDI format~\cite{Selfridge-Field97beyondMIDI}. %\cite{MIDIfile}.
 Such representations ignore the hierarchical structures that frame the
-conception of music, at least in the western area. These structures, on the other hand,
+conception of music, at least in the Western area. These structures, on the other hand,
 are present, either explicitly  or implicitly,
@@
 Some automata with memory extensions
-allow restricted storage and comparison of input symbols,
+allow restricted storage and comparison of input symbols
 %and correspond with logics,
-(see~\cite{Segoufin06csl} for a survey),
+(see~\cite{Segoufin06csl} for a survey)
 \florent{register: skip refs and details, add Mikolaj recent}
-with pebbles for marking positions~\cite{NevenSchwentickVianu04FSMinfinite},
+using pebbles for marking positions~\cite{NevenSchwentickVianu04FSMinfinite},
 registers~\cite{KaminskiFrancez94},
@@
 each transition accepts a set of such symbols (instead of an individual symbol),
-represented by Boolean formula or Binary Decision Diagrams.
+represented by Boolean Formulas or Binary Decision Diagrams.
 %
@@
 \caption{Classes of Symbolic/Weighted Automata.
-$\Sigma_\mathsf{fin}$ and $\Sigma_\mathsf{inf}$ denote finite/countable alphabets,
+Here, $\Sigma_\mathsf{fin}$ and $\Sigma_\mathsf{inf}$ denote finite/countable alphabets,
 $\mathbb{B}$  the Boolean algebra,
@@
 A  \SWVPA operates sequentially on \emph{nested words}~\cite{AlurMadhusudan09nested},
-structured with markup symbols (parentheses), and describing linearizations of trees.
+structured with markup symbols (parentheses) and describing linearizations of trees.
 A \SWVPA $A$ associates a weight value $A(t)$ % \in \Semiring$
@@
 $T(s, t) \otimes A(t)$, called the distance between $s$ and $A$ in~\cite{Mohri03EDWA},
-\wrt the ranking defined by $\oplus$Like weighted-parsing
+\wrt the ranking defined by $\oplus$. Like weighted-parsing
 methods~\cite{Goodman99SemiringParsing,Nederhof03weightedParsing,MorbitzVogler19weighted-parsing},
@@
 is built, such that for all~$t$,
-$B(t) = T(s, t) \otimes A(t)$.% $A_{T, s}(t) = T(s, t) \otimes A(t)$.
+$B(t) = T(s, t) \otimes A(t)$. % $A_{T, s}(t) = T(s, t) \otimes A(t)$.
 In the second step, a best AST $t$ is found by applying to $B$ %$A_{T, s}$
@@
 ($iii.a$)~the \SWT-based definition of generic edit distances between input and output (yield) words,
-and ($iii.b$)~the use, convenient in this context,
+and ($iii.b$)~the use, for convenience in this context,
 of nested words, and \SWVPA,
@@
 a given \emph{timeline} of musical events
-from an infinite alphabet $\Sigma$ as input,
+from an infinite alphabet $\Sigma$ as input
 is parsed into a structured music score.

### 2021-09-01 main.tex
review typos

@@
 Let $\Semiring$ be a commutative semiring,
-$\Sigma$ and $\Delta$ be alphabets called respectively \emph{input} and \emph{output}, %{alphabets},
+$\Sigma$ and $\Delta$ be alphabets called \emph{input} and \emph{output} respectively, %{alphabets},
 and $\bar\Phi$ be a label theory over $\Semiring$
@@
 where $Q$ is a finite set of states,
-$\mathsf{in} : Q \to \Semiring$   %\Phi_{\Sigma, \Delta}
-(respectively $\mathsf{out} : Q \to \Semiring$)  %\Phi_{\Sigma, \Delta}
+$\mathsf{in} \colon Q \to \Semiring$   %\Phi_{\Sigma, \Delta}
+(respectively $\mathsf{out} \colon Q \to \Semiring$)  %\Phi_{\Sigma, \Delta}
 are functions defining the weight for entering
@@
 \noindent
-We call \emph{number of transitions} of $T$ the number of pairs of states
+We call the \emph{number of transitions} of $T$ the number of pairs of states
 $q, q' \in Q$ such that $\wei_{10}$ or $\wei_{01}$ or $\wei_{11}$
@@
 overloading the function names,
-such that, for all $q, q' \in Q$, $a \in \Sigma$,  $b \in \Delta$:
+such that, for every $q, q' \in Q$, $a \in \Sigma$,  $b \in \Delta$:
 \[
@@
 %
-We recall that, by convention (Section~\ref{sec:prelim}),
+We recall that, by definition (Section~\ref{sec:prelim}),
 an empty sum with $\bigoplus$ is equal to~$\zero$.
@@
 %
-The cost of a sequence of such operations transforming $s$ into $t$,
-is the product, with $\otimes$, of the individual costs of the operations involved;
-and the distance between $s$ and $t$ is the sum, with $\oplus$,
+The cost of a sequence of such operations transforming $s$ into $t$
+is the product in terms of $\otimes$ of the individual costs of the operations involved;
+and the distance between $s$ and $t$ is the sum in terms of $\oplus$
 of all possible products.
@@
 \begin{example}\label{ex:SWT}
-We build a small \SWT over  alphabets $\Sigma$ and $\Deltai$
+We build a small \SWT over the alphabets $\Sigma$ and $\Deltai$
 (Ex.~\ref{ex:running} and \ref{distance-time}),
@@
 $\musQuarter\!\!\!\mathrel{\raisebox{-1.5mm}{$\smile$}}\!\!\!\musEighth$,
-or \emph{dots} \musQuarterDotted{}),
+or \emph{dots} \musQuarterDotted{})
 %that represents continuation
@@
 $\mathsf{in} : Q \to \Semiring$ %\Phi_\Sigma$,
-(respectively $\mathsf{out} : Q \to \Semiring$)  %\Phi_\Sigma,$
+, respectively $\mathsf{out} : Q \to \Semiring$,  %\Phi_\Sigma,$
 are functions defining the weight for entering
@@
 The computation of $A$ on words $s \in \Sigma^*$
-is based with an intermediate function $\weight_A$,
+is based on with an intermediate function $\weight_A$,
 defined as follows for $q, q' \in Q$, $a \in \Sigma$, $u \in \Sigma^*$
@@
 \begin{align}
-\weight_A(q, \varepsilon, q') & = \one\ \mathrm{if}\  q = q',\ 
-          \mathrm{or}\  \zero\  \mathrm{otherwise}\\
+\weight_A(q, \varepsilon, q') & = \one\ \mathrm{if}\  q = q'\ 
+          \mathrm{and}\  \zero\  \mathrm{otherwise}\\
 %\weight_A(q, \varepsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
@@
 effectively constructible in PTIME, 
-such that for all $t \in \Delta^+$,
+such that for every $t \in \Delta^+$,
 $B_{A, T}(t) = \displaystyle\bigoplus_{s\in \Sigma^*} A(s) \otimes T(s, t)$.
@@
 \ie such that $A(s) = \one$ for a given $s \in \Sigma^*$
-and $A(s') = \zero$ for all $s' \neq s$,
+and $A(s') = \zero$ for every $s' \neq s$,
 corresponds to a construction of a \SWA for the partial application of the \SWT $T$,
@@
 $B_{s, T}$ over $\Delta$, $\Semiring$ and $\bar\Phi$,
-such that for all $t \in \Delta^+$, $B_{s, T}(t) = T(s, t)$.
+such that for every $t \in \Delta^+$, $B_{s, T}(t) = T(s, t)$.
 \end{corollary}
@@
 %finite (large) or infinite,
-artitioned into three
+partitioned into three
 subsets~$\Deltai$, $\Deltac$, $\Deltar$,
@@
 represents an \emph{appogiatura}, %(grace-note),
-\ie an ornemental note with theoretical duration~0.
+\ie an ornamental note with theoretical duration~0.
 \endex
@@
 we extend the above transition functions as follows
-for all $q, q' \in Q$, $p \in P$,
+for every $q, q' \in Q$, $p \in P$,
 $a \in \Deltai$,
@@
 $A_1 \oplus A_2$ and $A_1 \otimes A_2$,
-such that for all $s \in \Delta^*$,
+such that for every $s \in \Delta^*$,
 $(A_1 \oplus A_2)(s) = A_1(s) \oplus A_2(s)$ and
@@
 \begin{figure}[ht]
-For all $q_0, q_3 \in Q$, for all $p' \in P$,
+For every $q_0, q_3 \in Q$, for every $p' \in P$,
 \[
@@
 %
-For all $u \in \Sigma^*$ and $t \in \Delta^*$, let
+For every $u \in \Sigma^*$ and $t \in \Delta^*$, let
 \(d(u, t) = T\bigl(u, t|_\Deltai \bigr)\),
@@
 (and the same semiring and label theory $\Semiring$ and $\bar\Phi$),
-such that for all $u \in \Sigma^*$, and $t \in {\Delta}^*$,
+such that for every $u \in \Sigma^*$, and $t \in {\Delta}^*$,
 $T'(u, t) = T(u, t|_{\Deltai})$.
@@
 a \SWA $B_{s, T'}$,
-such that for all $t \in \Delta^*$, $B_{s, T'}(t) = d(s, t)$.
+such that for every $t \in \Delta^*$, $B_{s, T'}(t) = d(s, t)$.
 %

### 2021-09-01 semiring.tex
review typos

@@
 A semiring~$\Semiring$ is \emph{commutative} if $\otimes$ is commutative.
-It is \emph{idempotent} if for all $x \in \Semiring$, $x \oplus x = x$.
+It is \emph{idempotent} if for every $x \in \Semiring$, $x \oplus x = x$.
 %
@@
 a partial ordering~$\leq_\oplus$
-called the \emph{natural ordering} of~$\Semiring$~\cite{Mohri02semiring}
-defined,  by:
+which is called the \emph{natural ordering} of~$\Semiring$~\cite{Mohri02semiring},
+and is defined by:
 %implicitly defined by the semiring $\Semiring$
-for all $x, y\in \Semiring$,
+for every $x, y\in \Semiring$,
 $x \leq_\oplus y \;\mbox{iff}\; x \oplus y = x$.
@@
 An idempotent semiring $\Semiring$~is called \emph{total} if
-it~$\leq_\oplus$ is total
-\ie when for all $x, y \in \Semiring$, either $x \oplus y = x$ or $x \oplus y = y$.
+~$\leq_\oplus$ is total,
+\ie when for every $x, y \in \Semiring$, either $x \oplus y = x$ or $x \oplus y = y$.
 %\florent{is total necessary?}
@@
 Let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be an idempotent semiring.
-For all $x, y, z  \in \Semiring$,
+For every $x, y, z  \in \Semiring$,
 if $x \leq_\oplus y$ then
@@
 
-We then say the
+We thus say the
 $\Semiring$ is \emph{monotonic} \wrt $\leq_\oplus$.
@@
 \emph{non-negative weights} condition in shortest-path algorithms~\cite{Dijkstra59anote}.
-Intuitively, it means that combining elements with~$\otimes$ always increase their weight.
+Intuitively, it means that combining elements with~$\otimes$ always increases their weight.
 Formally, it is defined as the property ($i$) below. %of the following lemma.
@@
 \begin{itemize}
-\item [$i.$] for all $x, y \in \Semiring$,
+\item [$i.$] for every $x, y \in \Semiring$,
 $x \leq_\oplus x \otimes y$ and
 $y \leq_\oplus x \otimes y$
-\item[$ii.$] for all $x \in \Semiring$,  $\one \oplus x = \one$.
+\item[$ii.$] for every $x \in \Semiring$,  $\one \oplus x = \one$.
 \end{itemize}
@@
 that
-$\one \leq_\oplus x$ for all $x \in \Semiring$.
+$\one \leq_\oplus x$ for every $x \in \Semiring$.
 Similarly, by the first inequality of ($i$) with $y = \zero$,
@@
 Hence, in a superior semiring,
-for all $x \in \Semiring$, $\one \leq_\oplus x \leq_\oplus \zero$.
+for every $x \in \Semiring$, $\one \leq_\oplus x \leq_\oplus \zero$.
 %
@@
 are weighted by values of $\Semiring$, the loops can be safely avoided,
-because, for all $x \in \Semiring$ and $n \geq 1$,
+because, for every $x \in \Semiring$ and $n \geq 1$,
  $x \oplus x^n = x \otimes (\one \oplus x^{n-1}) = x$.
@@
 \noindent
-We  need infinite sums with~$\oplus$.
+We need to extend ~$\oplus to infinitely many operands$.
 A semiring~$\Semiring$ is called \emph{complete}~\cite{Droste09handbook}
@@
 $(x_i)_{i \in I}$ %$\{ x_i \mid i \in I \}$
-of elements of $\dom(\Semiring)$ over an index set $I \subset \mathbb{N}$, such that:
+of elements of $\dom(\Semiring)$ over an index set $I \subseteq \mathbb{N}$, such that:
 \begin{description}
@@
 \emph{associativity and commutativity:}\\
-for all $I \subseteq \mathbb{N}$
+for evert $I \subseteq \mathbb{N}$
 and all partition $(I_{j})_{j \in J}$ of $I$, %\subseteq \mathbb{N}$,
@@
 \emph{distributivity of product over infinite sum:}\\
-for all $I \subseteq \mathbb{N}$,
+for every $I \subseteq \mathbb{N}$,
 \(

### 2021-09-01 FST-TCS/rebuttal/rebuttal.md
corrected review typos

@@
 
+REMAINS TO DO :
+references :
+- [ ] 75: maybe cite Parsing as intersection by Nederhof and Satta (2003)
+- [ ] 126: you should cite Knuth's generalization of Dijkstra's algorithm for superiority, he defines the notion for functions of arbitrary arity, so not just semirings (but multioperator monoids); you should give an example for superiority in the semiring notation (with $\otimes$)
+Moreover, the `corresponds' is wrong, since superiority generalizes non-negative edge weights
+- [ ] PS: Knuth's definition of superiority includes monotonicity
+- [ ] 143: Lemma 4 is Lemma 3 in [21]
+- [ ] 195: references: there is earlier research on symbolic automata, e.g. Veanes (2013)
+- [ ]  Ref [22] and [23] both seem to refer to the same work by Mohri
+
+- [ ] 51--53: the jump from automata to grammars is uneasy and not really necessary, as the paper mainly deals with automata
+- [ ] 57: weighted automata don't have ASTs, but runs
+- [ ] 60: other approaches use multioperator monoids or valuation monoids, so `at minima a semiring' is wrong
+- [ ] 116: $\oplus$ `combines two values into a single one' as well (in fact, every   binary operation does)
+- [ ] 136: what implies?
+- [ ] 153: quantify $x$ and the family ($y_i$) for $i ∈ ℕ$
+- [ ] 155--156: alphabets are usually finite and non-empty sets
+- [ ] 159: denoted by; also it seems like $\bar\Phi$ is a ($\Sigma \cup \Sigma \times \Delta$)-indexed family, why don't you introduce it like this
+- [ ] 161--171: use an itemize environment 
+- [ ] 162--165: I think these should also hold for $\Delta$, why don't you write this explicitly?
+- [ ] 188: the definition of $\bigoplus_\Delta$ can be seen to be implicit.  Please fully quantify the alphabets of a label theory
+- [ ] 208: What does $T$ compute?
+- [ ] 219: resp.\@
+- [ ] 220--221: you mean that a term CONTAINING $w_{ij}(…)$ equal to 0 is ignored
+- [ ] 238: text in the margin, also it is not clear whether the if refers to the entire line or just the right definition
+- [ ] 284: $x$ is a variable, so it should be typeset in math mode
+- [ ] 291: I'd at least say once that $\Delta_i$, $\Delta_c$, and $\Delta_r$ are pairwise disjoint (or explain the symbol)
+- [ ] 297: the extension here is not as in Section 3, there, the different function hat at least similar types
+- [ ] 299: every line expands into the margin; moreover, in the rightmost equation of line three, in the application of $w_c$, $p'$ and $q'$ are swapped
+- [ ] 316: equation extends into margin
+- [ ] 331: I think the application of $w_c$ is wrong (same with $w_r$ and $w_i$ later), shouldn't the : be a comma and what is q in the middle?
+- [ ] 336: how are states (which are in $Q$) pushed onto the stack if $\Gamma = \Delta_c \times P$?
+- [ ] 339: what are $j$ and $c$, do you mean $i$ and $d$?
+- [ ] 341: shouldn't there be a transition for returning from a measure?
+- [ ] 348: for?  do you mean accepted by?  but the language of an swA isn't defined either
+- [ ] 354: readers may have an easier time seeing this argument if you say that $S$ is extremal.  You should say in the preliminaries that S is extremal if and only if $\leq_\oplus$ is total.
+- [ ] 371: How is it ensured that the symbol $<c,p>$ isn't popped and pushed again later during the computation?
+- [ ] 376: I do not understand the and at termination, and its correctness part.  What do you want to express?
+- [ ] 379: Almost all update equations need to compute infinite sums.
+Don't you think that the assumption of $\bar\Phi$ being effective is pretty strong and gives your the result `for free'?
+I think you should at least elaborate on this and say to what degree your effective label theory is a generalization of the decidable label theory of unweighted symbolic automata.
+- [ ] 629: typical references for regular tree grammars are Brainerd (1969) and  Gecseg and Steinby (1984)
+- [ ] 654: an intuition for the $\bar q$ would have made understanding the construction far easier
+
+R3 :
+
+- [ ] l. 100. Shouldn't the first measure end before the second starts? This would make $O$ a forest rather than a tree - easily fixed with a "piece" root element.
+- [ ] l.
+220.
+The sentence starting here is somewhat confusing (and possibly wrong, depending on if $q'$ and $q''$ are intended to be distinct).
+I suppose that the point is that one can quit the recursion "early" whenever $w_{ij}$ returns $0$, but the language could be clearer, something like "Since $0$ is absorbing for $\times$, and the identity element for $+$ in $S$, if $w_{ij}(q,a,b,q'')$ is $0$ (meaning there is no transition from $q$ to $q'$ reading $a$ and $b$) that entire term can be ignored in the sum."
+- [ ] l. 288: same as on line 100, measure 1 should maybe end before measure 2
+- [ ] l. 434: $swM$??
+
 List of mistakes:
@@
 - [x] Also, you should be consistent with using  'for all' or `for every' in quantifications (I would prefer for every)
-
 - [x] 122: if it IS; I'd also add a comma after total
@@
 - [ ] PS: Knuth's definition of superiority includes monotonicity
-
 - [x] 128: increaseS
@@
 - [ ] 299: every line expands into the margin; moreover, in the rightmost equation of line three, in the application of $w_c$, $p'$ and $q'$ are swapped
-- [ ] 310: topupmost -> topmost
+- [x] 310: topupmost -> topmost
 - [ ] 316: equation extends into margin
-- [ ] 330--331: as indicated by the stack top: better explain that stack symbols are time intervals
+- [x] 330--331: as indicated by the stack top: better explain that stack symbols are time intervals
 - [ ] 331: I think the application of $w_c$ is wrong (same with $w_r$ and $w_i$ later), shouldn't the : be a comma and what is q in the middle?
@@
 - [ ] 341: shouldn't there be a transition for returning from a measure?
-- [ ] 345: exist (singular)
+- [x] 345: exist (singular)
 - [ ] 348: for?  do you mean accepted by?  but the language of an swA isn't defined either
-- [ ] 352: Dijkstra-like algorithm?
+- [x] 352: Dijkstra-like algorithm?
 - [ ] 354: readers may have an easier time seeing this argument if you say that $S$ is extremal.  You should say in the preliminaries that S is extremal if and only if $\leq_\oplus$ is total.
-- [ ] 366: be popPED
+- [x] 366: be popPED
 - [ ] 371: How is it ensured that the symbol $<c,p>$ isn't popped and pushed again later during the computation?
-- [ ] 375: converge (singular)
+- [x] 375: converge (singular)
 - [ ] 376: I do not understand the and at termination, and its correctness part.  What do you want to express?
@@
 I think you should at least elaborate on this and say to what degree your effective label theory is a generalization of the decidable label theory of unweighted symbolic automata.
-- [ ] 386: We assume the following input to be given
-- [ ] 394: the notation "sw-parsing" is undefined
-- [ ] 416: word order: put "to $T$" in the end
-- [ ] 420: "the sw-parsing problem" or "sw-parsing", but not "the sw-parsing"
-- [ ] 424: the triple use and does not work well
-- [ ] 431: double the
-- [ ] 435: there is something missing between "here" and "offline"
+- [x] 386: We assume the following input to be given
+- [x] 394: the notation "sw-parsing" is undefined
+- [x] 416: word order: put "to $T$" in the end
+- [x] 420: "the sw-parsing problem" or "sw-parsing", but not "the sw-parsing"
+- [x] 424: the triple use and does not work well
+- [x] 431: double the
+- [x] 435: there is something missing between "here" and "offline"
 - [ ] 629: typical references for regular tree grammars are Brainerd (1969) and  Gecseg and Steinby (1984)
-- [ ] 647: inn -> in
-- [ ] 650: missing period after function
-- [ ] 652: I think $\bar w$ should be $\bar w'$
+- [x] 647: inn -> in
+- [x] 650: missing period after function
+- [x] 652: I think $\bar w$ should be $\bar w'$
 - [ ] 654: an intuition for the $\bar q$ would have made understanding the construction far easier
@@
 
-- [ ] Line 281: "artitioned into three" -> "partitioned into three"
+- [x] Line 281: "artitioned into three" -> "partitioned into three"
 
@@
 
-- [ ] l. 87. I would front "input" in the sentence, yielding "a given input _timeline_ of [...] alphabet $\Sigma$, is parsed into[...]"
+- [x] l. 87. I would front "input" in the sentence, yielding "a given input _timeline_ of [...] alphabet $\Sigma$, is parsed into[...]"
 
-- [ ] l. 90. Please parenthesize the pitch-timecode pairs for readability. As an aside: these are strictly the onsets? And strictly monophonic?
+- [x] l. 90. Please parenthesize the pitch-timecode pairs for readability. As an aside: these are strictly the onsets? And strictly monophonic?
 
@@
 
-- [ ] l.
+- [x] l.
 156.
@@
 
-- [ ] l. 240: an fixed weight -> a fixed weight
+- [x] l. 240: an fixed weight -> a fixed weight
 
-- [ ] l. 251: based with an intermediate -> based on an intermediate
+- [x] l. 251: based with an intermediate -> based on an intermediate
 
-- [ ] l. 276: generalizing themselves -> themselves generalizing
+- [x] l. 276: generalizing themselves -> themselves generalizing
 
-- [ ] l. 281: artitioned -> partitioned
+- [x] l. 281: artitioned -> partitioned
 
@@
 
-- [ ] l. 290: ornemental -> ornamental
+- [x] l. 290: ornemental -> ornamental
 
-- [ ] l. 306: reads and pop from stack -> reads and pops from the stack
+- [x] l. 306: reads and pop from stack -> reads and pops from the stack
 
-- [ ] l. 307: I would use "computations" or something like that rather than "transitions" here
+- [x] l. 307: I would use "computations" or something like that rather than "transitions" here
 
-- [ ] l. 310: topupmost -> topmost, though I'm not sure if it's necessary to describe the basic workings of a stack. If you choose to do so, then also
+- [x] l. 310: topupmost -> topmost, though I'm not sure if it's necessary to describe the basic workings of a stack. If you choose to do so, then also
 
-- [ ] l. 311: content -> pair
+- [x] l. 311: content -> pair
 
-- [ ] l. 323: particular -> special
+- [x] l. 323: particular -> special
 
-- [ ] l. 365: "case of in the" -> either "case in the" or "case of the"
+- [x] l. 365: "case of in the" -> either "case in the" or "case of the"
 
-- [ ] l. 366: pop -> popped
+- [x] l. 366: pop -> popped
 
- The second case in Figure 3 should start with $d_\top$, not $d_\bot$
+ - [x] The second case in Figure 3 should start with $d_\top$, not $d_\bot$
 
-- [ ] l.
+- [x] l.
 395: I would omit "output".
@@
 
-- [ ] l. 399: correspondance -> correspondence
+- [x] l. 399: correspondance -> correspondence
 
-- [ ] l. 402: "weight a second time division" -> something like "weight assigned to an additional second time division"
+- [x] l. 402: "weight a second time division" -> something like "weight assigned to an additional second time division"
 
-- [ ] l. 416: extra commen between "$T$" and "of new transition"
+- [x] l. 416: extra commen between "$T$" and "of new transition"
 
-- [ ] l. 417: Corolary -> Corollary
+- [x] l. 417: Corolary -> Corollary
 
-- [ ] l. 420: This whole paragraph should be shaped up, writing wise.
+- [x] l. 420: This whole paragraph should be shaped up, writing wise.
 
@@
 
-- [ ] l. 435: "offline" -> "works offline"
+- [x] l. 435: "offline" -> "works offline"
 
-- [ ] l. 437: extraneous comma after "such as"
+- [x] l. 437: extraneous comma after "such as"
 

### 2021-09-01 app-trees.tex
passe section 4 swVPA
corrected review typos

@@
 
-The hierarchical structure of nested-words, defined with the \emph{call} 
+The hierarchical structure of nested words, defined with the \emph{call} 
 and \emph{return} markup symbols
@@
 represent the two scores in Examples~\ref{ex:running},\ref{ex:nested-word}, 
-and their linearization are respectively~$O$ and~$O'$ inn the same examples.
+and their linearization are respectively~$O$ and~$O'$ in the same examples.
 \end{example}
@@
 \begin{proof}
-Let $A = \< Q, \init, \bar{\wei} >$ where $\bar{\wei}$ is presented as above by a function
+Let $A = \< Q, \init, \bar{\wei} >$ where $\bar{\wei}$ is presented as above by a function.
     %
@@
 of length at most $M$, including the empty sequence denoted by~$\varepsilon$,
-and where $P' = Q'$ and $\bar\wei$ is defined by:
+and where $P' = Q'$ and $\bar{\wei}'$ is defined by:
 

### 2021-09-01 main.tex
corrected review typos

@@
 The transition $\wei_{10}$,
-with an fixed weight value $\alpha \in \Semiring$,
+with a fixed weight value $\alpha \in \Semiring$,
 switches from state $q_0$ (normal) to $q_1$ (error)
@@
 The computation of $A$ on words $s \in \Sigma^*$
-is based on with an intermediate function $\weight_A$,
+is based on an intermediate function $\weight_A$,
 defined as follows for $q, q' \in Q$, $a \in \Sigma$, $u \in \Sigma^*$
@@
 (\SVPA~\cite{dAntonyAlur14SVPDA},
- generalizing themselves \VPA~\cite{AlurMadhusudan09nested} to infinite alphabets)
+ themselves generalizing \VPA~\cite{AlurMadhusudan09nested} to infinite alphabets)
 from Boolean semirings to arbitrary semiring domains.
@@
 $\weir$ and $\weier$ read the input return symbol $\return{r}$, and change state from $q$ to to $q'$.
-Moreover, $\weir$ reads and pop from stack a pair made of $\call{c}$ and $p$,
+Moreover, $\weir$ reads and pops from stack a pair made of $\call{c}$ and $p$,
 ($\call{c}$ is compared to $\return{r}$).
@@
 $[Q \times \Gamma^*] \times \Delta^* \times [Q \times \Gamma^*]$ into~$\Semiring$.
-The empty stack is denoted by $\bot$, and the topupmost
-symbol is the last pushed content.
+The empty stack is denoted by $\bot$, and the topmost
+symbol is the last pushed pair.
 %
@@
 over $\Sigma$, $\Semiring$ and $\bar\Phi$
-is a particular case of $\SWVPA$
+is a special case of $\SWVPA$
 $\< Q, \emptyset, \init, \bar\wei, \final >$
@@
 $q_{{i}/{c}}$, meaning that the current interval has number~$i < c$
-amongst $c$ sub-intervals of duration~$\ell$ (as indicated by the stack top).
+amongst $c$ sub-intervals of duration~$\ell$ (as indicated by the interval at the stack's top).
 It reads a new time-division symbol~$\ccall{d}$:
@@
 over the same $\Delta$, commutative $\Semiring$ and effective $\bar\Phi$.
-There exists two effectively constructible $\SWVPA$
+There exist two effectively constructible $\SWVPA$
 $A_1 \oplus A_2$ and $A_1 \otimes A_2$,
@@
 %
-We propose a Dijkstra algorithm computing,
+We propose a Dijkstra-like algorithm computing,
 for every $q, q' \in Q$,
@@
 and can also push another pair $\< c', p'> \in \Gamma$ on $\gamma'$,
-following the third case of
+following the third case 
 in the definition \eqref{eq:SWVPA-weight} of $\weight_A$ (call symbol).
-The pair $\< c', p'>$ can be pop later, during the computation from $q$ to $q'$,
+The pair $\< c', p'>$ can be popped later, during the computation from $q$ to $q'$,
 following the fifth case of \eqref{eq:SWVPA-weight} (return symbol).
@@
 %of states of $A$ by weight values in $\Semiring$,
-that converges eventually to $b_\top$ and $b_\bot$. 
+that converge eventually to $b_\top$ and $b_\bot$. 
 %\florent{explication Fig.~\ref{fig:best-update} (update) suivant cas de \eqref{eq:SWVPA-weight}?}
@@
 \noindent
-We assume given the following input:
+We assume the following input to be given:
 \begin{description}
@@
 $\SW$-parsing is the problem of computing
-the distance~\eqref{eq:distance-lang} between the input~$s$ and the output weighted language of~$A$,
+the distance~\eqref{eq:distance-lang} between the input string ~$s$ and the weighted language over the output alphabet defined by ~$A$,
 and returning a witness $t$.
@@
 \item The \SWT $T$ evaluates a ``fitness measure''  that expresses
-    a correspondance between a performance and a nested representation of a music score.
+    a correspondence between a performance and a nested representation of a music score.
     See Example~\ref{ex:SWT}.
@@
 %\includegraphics[scale=0.20]{pictures/score4.png}.
-when the weight a second time division with $\ccall{2}$ is less
+when the weight assigned to an additional second time division with $\ccall{2}$ is less
 than the difference of weight between
@@
 $b \in {\Delta} \setminus \Deltai$
-by the addition to $T$,
-of new transitions of the form $\wei_{01}(q, \varepsilon, b, q')$.
-%
-\noindent
-Then, using Corolary~\ref{cor:epsilon},
+by the addition 
+of new transitions of the form $\wei_{01}(q, \varepsilon, b, q')$ to $T$.
+%
+\noindent
+Then, using Corollary~\ref{cor:epsilon},
 we construct from $s \in \Sigma^*$ and $T'$
@@
 
-The $\SW$-parsing generalizes
-the problem of searching the best derivation (AST) of a weighted CF-grammar $G$
+The $\SW$-parsing problem generalizes
+the problem of searching for the best derivation (AST) of a weighted CF-grammar $G$
 that yields a given input word $w$, called~\emph{weighted parsing},
-(see~\cite{Goodman99SemiringParsing}
- and the more general framework~\cite{MorbitzVogler19weighted-parsing}),
-with infinite input alphabet instead of a finite one and
+(see~\cite{Goodman99SemiringParsing},
+ also the more general framework~\cite{MorbitzVogler19weighted-parsing}),
+with an infinite input alphabet instead of a finite one and
 transducer-defined distances instead of equality. 
@@
 See Appendix~\ref{sec:trees} for more details on the correspondence between 
-nested words $t \in \Delta^*$ and AST and CF grammars and \SWVPA.
+nested words $t \in \Delta^*$, AST, CF grammars and \SWVPA.
 
@@
 
-This work can be extended in several directions. First, the
+This work can be extended in several directions. First, 
 the best search algorithm could be generalized from $1$-best to $n$-best~\cite{Huang05kbest},
@@
 \eg~\cite{Segoufin06csl,KaminskiFrancez94,NevenSchwentickVianu04FSMinfinite,Bojanczyk11FO2}.
-Finally, the  best search algorithm presented here offline, whereas
+Finally, the  best search algorithm presented here works offline, whereas
 an on-the-fly automata construction would allow for online parsing,
-a suitable feature in the context of applications such as, 
+a suitable feature in the context of applications such as 
 \eg automatic music transcription.

### 2021-09-01 semiring.tex
corrected review typos

@@
 for evert $I \subseteq \mathbb{N}$
-and all partition $(I_{j})_{j \in J}$ of $I$, %\subseteq \mathbb{N}$,
+and all partitions $(I_{j})_{j \in J}$ of $I$, %\subseteq \mathbb{N}$,
 \(
@@
 \item[$iii.$]
-\emph{distributivity of product over infinite sum:}\\
+\emph{distributivity of products over infinite sums:}\\
 for every $I \subseteq \mathbb{N}$,

### 2021-09-01 FST-TCS/rebuttal/response.txt
response rebutal

--- 2021-09-01 FST-TCS/rebuttal/response.txt
@@
-We would like to thank the reviewers for their valuable work, 
-and their comments and references/suggestion.
+We first warmly thank the reviewers for their hard work, and the impressive list of comments and references/suggestions which will be very useful to improve the text, whatever the final decision is.
 
-In particular, we are grateful to them for pointing us to the work of L. Herrmann & Heiko... in DLT 2016 & PhD 2020.  
-Indeed, the models that we are using in the paper are particular (restricted) cases of the Weighted Symbolic Automata with Storage defined in DLT 2016 & PhD 2020.
-We agree that it should definitely be explained at the beginning of the paper, and the 2 references must be cited as well.
+Position and contributions (mostly review 1)
+--------------------------------------------
 
-However, It seems difficult to completely skip definitions of the models, for self-completeness reasons.
-definition of models less general sufficient for our purpose
-not clear whether Prop. 10 (in center pages) is a direct consequence of Hermann V. 2016...
+In particular, we are grateful to them for pointing out the work of Luisa Herrmann & Heiko Vögler in DLT 2016 & PhD 2020.
+We fully agree that the models presented in our paper are particular (and strictly restricted) cases of the Weighted Symbolic Automata with Storage defined in DLT 2016 & PhD 2020.
+This must definitely be explained at the beginning of the paper, and both references must be cited.
 
-also agree to focus the paper on its 2 main contributions, which are the following:
-- a best-search algorithm
-  (as far as we know this is a new result
-  no such algorithm has been proposed for swVPA)
-- a framework for parsing over infinite alphabet and with tree-automata. 
-  in particular the use of weighted transducers for def. of distance and of VPA for computing on linearization of parse trees.
+We believe that our work still contains significant and original contributions.
+However, given these new references and the reviewer’s comment, we do realize that the focus of the paper should be more clearly stated.
+In short, the purpose of developing a theoretical framework combining several language models is not to study their formal properties as such (in this respect our work does not directly compare with the study conducted in PhD 2020) but  to address a specific class of problems via weighted parsing.
+In the current text, we agree that it is fully revealed in the last section, but we will rephrase the beginning of the paper by shifting the list of contributions as follows:
 
-second is new application 
-Moreover, 
-Although this only mentioned quickly in the paper, our main motivation / purpose is a formal study of an empirical (challenging) problem : Automated Music Transcription, based on consistent theoretical framework made of formal language models / formalisms.
-in parallel to developpement of a transcription tool
-so far results have shown that theoretical foundations improve accuracy…
+- Our framework is intended to address the optimisation problem of symbolic weighted parsing over an infinite alphabet and with tree-automata.
+It combines weighted transducers for defining distances, and VPA for linearizing parse trees.
+To our knowledge, this contribution is a new application of symbolic and weighted models.
+- Our work features a best-search algorithm (as far as we know this is a new result no such algorithm has been proposed for swVPA)
+- Finally, in order to highlight the central motivation of the paper which is the definition of a formal approach yielding a tractable algorithm to solve a challenging parsing problem, we will mention the challenging problem of Automatic Music Transcription as a main empirical target.
 
-Regarding the best-search algorithm, 
-answer to reviewer 2:... (DETAILS)
+We hope that it would reinforce and clarify the purpose of the theoretical contributions developed in the text. An additional change to reflect this focus shift is to change the title to: 
+Quantitative Parsing over Infinite Alphabets with Symbolic Weighted Language Models.
 
-questions of reviewer 3 about the distance and the work of Mohri.
+We propose this rephrasing, along with the improved SOA, as a means to more accurately position our work, and let the reviewers decide whether this is satisfactory regarding the objections raised in the first review.
 
-About the weighted parsing : rem. reviewer1 on "previous approaches" which are more general.
+Regarding the suggestion of the first reviewer to remove that part of our text to avoid rephrasing, we believe that it would be difficult to completely skip the definitions of the models without impacting the readability of the text.
+We will of course devote our best efforts to improve the formal material by pointing to related work when it does not hinder self-completeness.
 
-other rem. review 1:
-the relationship between VPA and TA has been discussed in several places, 
-**we did [or not?]** find an explicit construction of VPA from TA.
-could be move to paper if you think that it is original.
-[ ] check in paper [5]
+Algorithm (mostly review 2)
+-------------------------------------
+
+Regarding the best-search algorithm, in particular Figure 3, we agree that it could be better explained (we do also apologize for the several unfortunate typos…). 
+For the updates defined in Figure 3, intuitively, we assume that the values of d_bot and d_top are definitely fixed for every tuplet which has been extracted from the queue \Q, and we use these values in order to update the functions d_bot and d_top on the other tuplets (those  which are still in the queue).
+The typos In Figure 3 are the following:
+- At the beginning of Fig. 3, the quantification is “for all q_0, q_3 in Q and p’ in P”
+- the second line should be of the form:  d_\bot(q_1, q_3) += … (as pointed by reviewer 2)
+- In line 3, p and p’ should be swapped.
+
+Misc.
+------
+
+Some reviewers raised questions about the notion of label theory, in particular the condition of effectiveness (Definition 6).
+This assumption is indeed strong, but in the context of solving symbolic weighted parsing, it is not unrealistic.
+In fact, for an instance of the problem, as stated in Section 5, we have only a finite number of fixed weight functions to consider (the functions in the automata’s transitions) and we may assume that they are effective (i.e.
+their global minimum are known).
+Then, the other functions to consider are obtained by combination with the operators of Section 2, and these operators preserve effectiveness, by monotony of the semiring.
+In practice, the combinations may be represented by structures like Algebraic Decision Diagrams (Bahar et al 1997).
+
+Reviewer 3 has questions about the relationship between our notion of distance and the definitions of Mohri in [22].
+Our interpretation is actually the second one mentioned in the review: using the notation of [22], Def.
+6, we consider the distance d({ s }, A) between a singleton distribution {s} (all words other than s have weight 0) and the distribution of A, wrt a partial application of T (on a subset of the alphabet).

### 2021-09-01 abstract.tex

@@
 and also \WA, whose transitions are labeled by constant weight values,
-and who deal only with finite alphabets.
+and which deal only with finite alphabets.
 %
-We present some properties of \SWA, \SWT and \SWVPA models,
+We present some properties of \SWA, \SWT, and \SWVPA models
 that we use to define and solve a variant of parsing

### 2021-09-01 introduction.tex

@@
 running example of \emph{music transcription}:
-a given \emph{timeline} of musical events
-from an infinite alphabet $\Sigma$ as input
+a given input \emph{timeline} of musical events
+from an infinite alphabet $\Sigma$,
 is parsed into a structured music score.
@@
 to the recording of a live performance, \eg
-$I = 69\ts{0.07},
-     71\ts{0.72},
-     73\ts{0.91},
-     74\ts{1.05},
-     76\ts{1.36},
-     77\ts{1.71}$. % over interval $[0,2[$,
+$I = (69\ts{0.07}),
+     (71\ts{0.72}),
+     (73\ts{0.91}),
+     (74\ts{1.05}),
+     (76\ts{1.36}),
+     (77\ts{1.71})$. % over interval $[0,2[$,
 

### 2021-09-02 FST-TCS/rebuttal/rebuttal-reviews.md
1 passe modifs suivant certaines remarques des reviwers

--- 2021-09-01 FST-TCS/rebuttal/rebuttal.md
@@
 
-The formalism of symbolic weighted automata is not original to the paper.
+- [ ] The formalism of symbolic weighted automata is not original to the paper.
 It was already introduced by Herrmann and Vogler (2016) and later revised by Herrmann (2020) under the name of weighted symbolic automata with data storage.
@@
 
-Moreover, the authors claim that their framework is a generalization of previous approaches to weighted parsing, in the sense that it allows infinite input alphabets.
-However, some of the previous approaches allowed more general weight structures than just semirings, so their are incomparable to the present paper.
+- [ ] Moreover, the authors claim that their framework is a generalization of previous approaches to weighted parsing, in the sense that it allows infinite input alphabets.
+However, some of the previous approaches allowed more general weight structures than just semirings, so their are **incomparable** to the present paper.
 
@@
 - [ ] 195: references: there is earlier research on symbolic automata, e.g. Veanes (2013)
-- [ ]  Ref [22] and [23] both seem to refer to the same work by Mohri
+- [x]  Ref [22] and [23] both seem to refer to the same work by Mohri
 
@@
 - [ ] 57: weighted automata don't have ASTs, but runs
-- [ ] 60: other approaches use multioperator monoids or valuation monoids, so `at minima a semiring' is wrong
+- [x] 60: other approaches use multioperator monoids or valuation monoids, so `at minima a semiring' is wrong
 - [ ] 116: $\oplus$ `combines two values into a single one' as well (in fact, every   binary operation does)
 - [ ] 136: what implies?
-- [ ] 153: quantify $x$ and the family ($y_i$) for $i ∈ ℕ$
+- [x] 153: quantify $x$ and the family ($y_i$) for $i ∈ ℕ$
 - [ ] 155--156: alphabets are usually finite and non-empty sets
@@
 - [ ] 371: How is it ensured that the symbol $<c,p>$ isn't popped and pushed again later during the computation?
-- [ ] 376: I do not understand the and at termination, and its correctness part.  What do you want to express?
+- [x] 376: I do not understand the and at termination, and its correctness part.  What do you want to express?
 - [ ] 379: Almost all update equations need to compute infinite sums.
@@
 I think you should at least elaborate on this and say to what degree your effective label theory is a generalization of the decidable label theory of unweighted symbolic automata.
-- [ ] 629: typical references for regular tree grammars are Brainerd (1969) and  Gecseg and Steinby (1984)
+- [ ] 629: typical references for regular tree grammars are Brainerd (1969) 
+- [ ] ​         and  Gecseg and Steinby (1984)
 - [ ] 654: an intuition for the $\bar q$ would have made understanding the construction far easier
@@
 - [ ] l. 100. Shouldn't the first measure end before the second starts? This would make $O$ a forest rather than a tree - easily fixed with a "piece" root element.
+
 - [ ] l.
@@
 I suppose that the point is that one can quit the recursion "early" whenever $w_{ij}$ returns $0$, but the language could be clearer, something like "Since $0$ is absorbing for $\times$, and the identity element for $+$ in $S$, if $w_{ij}(q,a,b,q'')$ is $0$ (meaning there is no transition from $q$ to $q'$ reading $a$ and $b$) that entire term can be ignored in the sum."
+
 - [ ] l. 288: same as on line 100, measure 1 should maybe end before measure 2
@@
 - [x] 375: converge (singular)
-- [ ] 376: I do not understand the and at termination, and its correctness part.  What do you want to express?
+- [x] 376: I do not understand the and at termination, and its correctness part.  What do you want to express?
 - [ ] 379: Almost all update equations need to compute infinite sums.
@@
 Original literature:
-- Herrmann and Vogler (2016): L.
+
+- [ ] Herrmann and Vogler (2016): L.
 Herrmann & H.
@@
 203–215, doi:10.1007/978-3-662-53132-7 17.
-- Herrmann (2020): L. Herrmann (2016): Weighted Automata with Storage. PhD thesis. Technische Universität Dresden, 2020, urn:nbn:de:bsz:14-qucosa2-740685
+
+- [ ] Herrmann (2020): L. Herrmann (2016): Weighted Automata with Storage. PhD thesis. Technische Universität Dresden, 2020, urn:nbn:de:bsz:14-qucosa2-740685
 
@@
 
-- [ ] Last line of algorithm 1: What exactly is meant by "update $d_⊥$ with $<q_1,q_2>$"?
+- [x] Last line of algorithm 1: What exactly is meant by "update $d_⊥$ with $<q_1,q_2>$"?
 Does it mean update the image of $<q_1,q_2>$ under $d_⊥$ with the new value as computed in Figure 3?
@@
 Figure 3: This figure is supposed to compute $<q_1,q_2>$. 
-- [ ] First of all, it is not clear what is meant by "compute $<q_1,q_2>$"; I am guessing, as I wrote in the comment above that the computed quantity is the image of $<q_1,q_2>$ under $d_⊥$.
+- [x] First of all, it is not clear what is meant by "compute $<q_1,q_2>$"; I am guessing, as I wrote in the comment above that the computed quantity is the image of $<q_1,q_2>$ under $d_⊥$.
 However, even this doesn't make sense; there are some series of equantions written, for $d_\top(q_1,p,q_3)$ (not for $d_\top(q_1,p,q_2)$.
-- [ ] Even in this equation, what is $q_2$? Do we take the minimum (wrt $\oplus$) over all $q_2$? 
-- [ ] The second line says $d_\bot(q_1,p,q_3)$, which doesn't make sense because $d_⊥$ takes two arguments, not three. 
-- [ ] There are two equations for $d_\top(q_1,p,q_3)$ and one for $d_\top(q_0,p,q_3)$.
+- [x] Even in this equation, what is $q_2$? Do we take the minimum (wrt $\oplus$) over all $q_2$? 
+- [x] The second line says $d_\bot(q_1,p,q_3)$, which doesn't make sense because $d_⊥$ takes two arguments, not three. 
+- [x] There are two equations for $d_\top(q_1,p,q_3)$ and one for $d_\top(q_0,p,q_3)$.
 How are these related?
@@
  References:
-- [ ]  Ref [22] and [23] both seem to refer to the same work by Mohri
+- [x]  Ref [22] and [23] both seem to refer to the same work by Mohri

### 2021-09-02 introduction.tex
1 passe modifs suivant certaines remarques des reviwers

@@
 %Intuitively,~$\oplus$ selects, or ranking, the syntax trees.
-The weight domain, equipped with these two operators is, at minima,
+The weight domain, equipped with these two operators is typically 
 a \emph{semiring} %$\Semiring$
@@
 A \SWT defines a distance $T(s, t)$ between finite words $s$ and $t$
-over infinite alphabets. % following~\cite{Mohri03EDWA}.
+over infinite alphabets. % following~\cite{Mohri03ijfcs}.
 %\wrt an edit distance
@@
 finding $t$ minimizing
-$T(s, t) \otimes A(t)$, called the distance between $s$ and $A$ in~\cite{Mohri03EDWA},
+$T(s, t) \otimes A(t)$, called the distance between $s$ and $A$ in~\cite{Mohri03ijfcs},
 \wrt the ranking defined by $\oplus$. Like weighted-parsing
@@
 %the distance $d(s, w) \otimes A(w)$ between $s$ and $A$,
-%as defined in~\cite{Mohri03EDWA}.
+%as defined in~\cite{Mohri03ijfcs}.
 

### 2021-09-02 main-V0.tex
comment versions
1 passe modifs suivant certaines remarques des reviwers

@@
-% old version
+%
+% initial version 
+% 
+%
 
@@
 a \SWT is used to define a distance $T(s, t)$ between two finite words $s$ and $t$
-over an infinite alphabet, following~\cite{Mohri03EDWA}.
+over an infinite alphabet, following~\cite{Mohri03ijfcs}.
 Then, the \emph{SW-parsing} problem aims at %computing 
@@
 $T(s, t) \otimes A(t)$ (\wrt the ranking defined by $\oplus$)
--- this value is called the distance between $s$ and $A$ in~\cite{Mohri03EDWA}.
+-- this value is called the distance between $s$ and $A$ in~\cite{Mohri03ijfcs}.
 %
@@
 %the distance $d(s, w) \otimes A(w)$ between $s$ and $A$, 
-%as defined in~\cite{Mohri03EDWA}.
+%as defined in~\cite{Mohri03ijfcs}.
 
@@
 
-We follow the approach of~\cite{Mohri03EDWA} for the computation of distances
+We follow the approach of~\cite{Mohri03ijfcs} for the computation of distances
 between words and languages, with weighted transducers, 
@@
 weighted automata and transducers~\cite{Droste09handbook} 
-%over finite alphabets, see  e.g.~\cite{Mohri03EDWA}, 
+%over finite alphabets, see  e.g.~\cite{Mohri03ijfcs}, 
 by labeling each transition with a weight function that takes the 
@@
 Therefore, it is the problem of optimizing a measure 
-called the \emph{edit-distance between $s$ and $A$} in~\cite{Mohri03EDWA}.
+called the \emph{edit-distance between $s$ and $A$} in~\cite{Mohri03ijfcs}.
 %
@@
 %\subsection{Distance between words or languages}
-...algebraic definition of edit-distance of Mohri, in \cite{Mohri03EDWA}
+...algebraic definition of edit-distance of Mohri, in \cite{Mohri03ijfcs}
 % Mehryar Mohri 

### 2021-09-02 main-V1.tex
comment versions
V1: file descr. in comment
1 passe modifs suivant certaines remarques des reviwers

--- 2021-05-17 main.tex
@@
+%
+% transducers and automata with epsilon transitions
+%
+
 \documentclass[runningheads]{llncs}
@@
 
-We follow the approach of~\cite{Mohri03EDWA} for the computation of distances,
+We follow the approach of~\cite{Mohri03ijfcs} for the computation of distances,
 between words and languages, using weighted transducers, 
@@
 weighted automata and transducers~\cite{Droste09handbook} 
-%over finite alphabets, see  e.g.~\cite{Mohri03EDWA}, 
+%over finite alphabets, see  e.g.~\cite{Mohri03ijfcs}, 
 by labelling each transition with a weight function (instead of a simple weight value), 

### 2021-09-02 main-V2.tex
revise algo best w. notation operators complete label theory
1 passe modifs suivant certaines remarques des reviwers

@@
 \mathrm{~is~the~partial~application~}
-x_c \mapsto \weic(q_0, p, x_c, c', p', q_1)}\\[2pt]
+x_c \mapsto \weic(q_0, x_c, p, c', p', q_1)}\\[2pt]
 %
@@
 The measure expressed in~\eqref{eq:distance-lang} is called
-called edit-distance between $s$ and $A$ in~\cite{Mohri03EDWA}.
+called edit-distance between $s$ and $A$ in~\cite{Mohri03ijfcs}.
 %

### 2021-09-02 main.tex
1 passe modifs suivant certaines remarques des reviwers

@@
 \newcommand{\florent}[1]{\todo[noline,size=\tiny,color=yellow!40]{#1}}
-\renewcommand{\florent}[1]{} % uncomment for final
+%\renewcommand{\florent}[1]{} % uncomment for final
 \newcommand{\lydia}[1]{\todo[noline,size=\tiny,color=red!30]{#1}}
-\renewcommand{\lydia}[1]{} % uncomment for final
+%\renewcommand{\lydia}[1]{} % uncomment for final
 \newcommand{\philippe}[1]{\todo[noline,size=\tiny,color=green!30]{#1}}
-\renewcommand{\philippe}[1]{} % uncomment for final
+%\renewcommand{\philippe}[1]{} % uncomment for final
 
@@
 
-We follow the approach of~\cite{Mohri03EDWA} for the computation of distances
+We follow the approach of~\cite{Mohri03ijfcs} for the computation of distances
 between words and languages
@@
 weighted automata and transducers~\cite{Droste09handbook}
-%over finite alphabets, see \eg~\cite{Mohri03EDWA},
+%over finite alphabets, see \eg~\cite{Mohri03ijfcs},
 by labeling each transition with a weight function
@@
 
-Algorithm~\ref{algo:Dijkstra}
-constructs iteratively,
-using a priority queue $\Q$, 
+Algorithm~\ref{algo:Dijkstra} constructs iteratively, using a priority queue $\Q$, 
 two markings
@@
 that converge eventually to $b_\top$ and $b_\bot$. 
-%\florent{explication Fig.~\ref{fig:best-update} (update) suivant cas de \eqref{eq:SWVPA-weight}?}
-%\florent{The infinite sums in the updates of $d$ in Algorithm~\ref{algo:Dijkstra},
-%Figure~\ref{fig:best-update} are well defined since~$\Semiring$ is complete.}
+%
+\florent{intuition sur update en commentaire}
+%Intuitively, in the update defined in Figure~\ref{fig:best-update},
+%we assume that the values of $d_\bot$ and $d_\top$ are definitely 
+%fixed for every tuplet which has been extracted from the queue $\Q$ 
+%(like in particular $\<q_1, q_2>$ or $\< q_1, p, q_2>$ when the algorithm calls this update), 
+%and we use these values 
+%(and the functions $w$ of transitions) 
+%in order to update the values of $d_\bot$ and $d_\top$ 
+%on the other tuplets (those which are still in the queue $\Q$).
+%
+\florent{explication Fig.~\ref{fig:best-update} (update) suivant cas de \eqref{eq:SWVPA-weight}?}
+\florent{The infinite sums in the updates of $d$ in Algorithm~\ref{algo:Dijkstra},
+ Figure~\ref{fig:best-update} are well defined since~$\Semiring$ is complete.}
 %
@@
 %Algorithm~\ref{algo:Dijkstra}
-It terminates in PTIME and at termination (when $\Q$ is empty), 
+It terminates in PTIME (when $\Q$ is empty), 
 and its correctness is ensured by~\eqref{eq:min} and  
@@
 \begin{figure}[ht]
-For every $q_0, q_3 \in Q$, for every $p' \in P$,
+For every $q_0, q_3 \in Q$ and $p' \in P$, if $\< q_1, q_2> \notin \Q$ 
 \[
@@
 %\multicolumn{3}{l}{\mathrm{For~all~} q_0, q_3 \in Q, p \in P,}\\
+%
+d_\bot(q_1, q_3) & \opluseq &
+  d_\bot(q_1, q_2) \otimes
+  \displaystyle\bigoplus_{\Deltai} \weiei(q_2, q_3)
+  \mathrm{if~} \< q_1, q_2> \notin \Q \\
+%
+d_\bot(q_0, q_3) & \opluseq &
+  {\displaystyle\bigoplus_{\Deltac}}
+  \bigl(
+  \weiec(q_0, p, q_1) \otimes
+   d_\top(q_1, p, q_2) \otimes
+  \displaystyle\bigoplus_{\Deltar} \weir(q_2, p, q_3)\bigr)\\
+%
+d_\bot(q_1, q_3) & \opluseq &
+  d_\bot(q_1, q_2) \otimes
+  \displaystyle\bigoplus_{\Deltar} \weier(q_2, q_3)\\
+%
+d_\bot(q_1, q_3) & \opluseq &
+  d_\bot(q_1, q_2) \otimes d_\bot(q_2, q_3)\  \mathrm{if~} \< q_2, q_3> \notin \Q \\
+\end{array}
+\]
+%
+For every $q_0, q_3 \in Q$ and $p' \in P$, if $\< q_1, p, q_2> \notin \Q$
+\[
+\begin{array}{lcl}
 d_\top(q_1, p, q_3) & \opluseq &
   d_\top(q_1, p, q_2) \otimes
-  \displaystyle\bigoplus_{\Deltai} \weii(q_2, p, q_3)\\
+  \displaystyle\bigoplus_{\Deltai} \weii(q_2, p, q_3)
+  \mathrm{if~} \< q_1, p, q_2> \notin \Q \\
 %\multicolumn{3}{l}{\quad
@@
 %x_c \mapsto \weii(q_2, x_c, p, a, q_3)}\\
-%
-d_\bot(q_1, q_3) & \opluseq &
-  d_\bot(q_1, q_2) \otimes
-  \displaystyle\bigoplus_{\Deltai} \weiei(q_2, q_3)\\
-%
 d_\top(q_0, p', q_3) & \opluseq &
@@
 %
-d_\bot(q_0, q_3) & \opluseq &
-  {\displaystyle\bigoplus_{\Deltac}}
-  \bigl(
-  \weiec(q_0, p, q_1) \otimes
-   d_\top(q_1, p, q_2) \otimes
-  \displaystyle\bigoplus_{\Deltar} \weir(q_2, p, q_3)\bigr)\\
-%
-d_\bot(q_1, q_3) & \opluseq &
-  d_\bot(q_1, q_2) \otimes
-  \displaystyle\bigoplus_{\Deltar} \weier(q_2, q_3)\\
-%
 d_\top(q_1, p, q_3) & \opluseq &
-  d_\top(q_1, p, q_2) \otimes d_\top(q_2, p, q_3),
+  d_\top(q_1, p, q_2) \otimes d_\top(q_2, p, q_3)\ 
   \mathrm{if~} \< q_2, p, q_3> \notin \Q \\
-%
-d_\bot(q_1, q_3) & \opluseq &
-  d_\bot(q_1, q_2) \otimes d_\bot(q_2, q_3), \mathrm{if~} \< q_2, q_3> \notin \Q \\
 \end{array}
@@
 %\caption{Update $d_\bot$ with $\<q_1, q_2>$ or $d_\top$ with $\< q_1, p, q_2>$.}
-\caption{Update $d_\bot$ or $d_\top$.}
+\caption{The update of $d_\bot$ when $\< q_1, q_2> \notin \Q$, 
+         and the update $d_\top$ when $\< q_1, q_2> \notin \Q$.}
 \label{fig:best-update}
@@
 %
-Following the terminology of~\cite{Mohri03EDWA},
+Following the terminology of~\cite{Mohri03ijfcs},
 $\SW$-parsing is the problem of computing

### 2021-09-02 semiring.tex
1 passe modifs suivant certaines remarques des reviwers

--- 2021-09-01 semiring.tex
@@
 \noindent
-We need to extend ~$\oplus to infinitely many operands$.
+We need to extend $\oplus$ to infinitely many operands.
 A semiring~$\Semiring$ is called \emph{complete}~\cite{Droste09handbook}
@@
 \emph{associativity and commutativity:}\\
-for evert $I \subseteq \mathbb{N}$
-and all partition $(I_{j})_{j \in J}$ of $I$, %\subseteq \mathbb{N}$,
+for all $I \subseteq \mathbb{N}$
+and all partitions $(I_{j})_{j \in J}$ of $I$, %\subseteq \mathbb{N}$,
 \(
@@
 \item[$iii.$]
-\emph{distributivity of product over infinite sum:}\\
-for every $I \subseteq \mathbb{N}$,
+\emph{distributivity of products over infinite sums:}\\
+for all $I \subseteq \mathbb{N}$, 
+%and all $x, y \in \Semiring$,
 \(
 \displaystyle
-\bigoplus_{i \in I} (x \otimes y_i) = x \otimes \bigoplus_{i\in I} y_i\), and
+\forall x \in \Semiring, \bigoplus_{i \in I} (x \otimes y_i) = x \otimes \bigoplus_{i\in I} y_i\), and
 \(
 \displaystyle
-\bigoplus_{i \in I} (x_i \otimes y) = (\bigoplus_{i \in I} x_i ) \otimes y\).
+\forall y \in \Semiring, \bigoplus_{i \in I} (x_i \otimes y) = (\bigoplus_{i \in I} x_i ) \otimes y\).
 \end{description}

### 2021-09-02 FST-TCS/rebuttal/response.txt
reponse finale rebutal

@@
 
-In particular, we are grateful to them for pointing out the work of Luisa Herrmann & Heiko Vögler in DLT 2016 & PhD 2020.
-We fully agree that the models presented in our paper are particular (and strictly restricted) cases of the Weighted Symbolic Automata with Storage defined in DLT 2016 & PhD 2020.
+In particular, we are grateful to them for pointing out the work of Luisa Herrmann & Heiko Vögler in DLT 2016 & Ph.D.
+2020.
+We fully agree that the models presented in our paper are particular (and strictly restricted) cases of the Weighted Symbolic Automata with Storage defined in DLT 2016 & Ph.D.
+2020.
 This must definitely be explained at the beginning of the paper, and both references must be cited.
@@
 However, given these new references and the reviewer’s comment, we do realize that the focus of the paper should be more clearly stated.
-In short, the purpose of developing a theoretical framework combining several language models is not to study their formal properties as such (in this respect our work does not directly compare with the study conducted in PhD 2020) but  to address a specific class of problems via weighted parsing.
+In short, the purpose of developing a theoretical framework combining several language models is not to study their formal properties as such (in this respect our work does not directly compare with the study conducted in Ph.D.
+2020) but to address a specific class of problems via weighted parsing.
 In the current text, we agree that it is fully revealed in the last section, but we will rephrase the beginning of the paper by shifting the list of contributions as follows:
 
-- Our framework is intended to address the optimisation problem of symbolic weighted parsing over an infinite alphabet and with tree-automata.
-It combines weighted transducers for defining distances, and VPA for linearizing parse trees.
+- Our framework is intended to address the optimization problem of symbolic weighted parsing over an infinite alphabet and with tree-automata.
+It combines weighted transducers for defining distances and VPA for linearizing parse trees.
 To our knowledge, this contribution is a new application of symbolic and weighted models.
-- Our work features a best-search algorithm (as far as we know this is a new result no such algorithm has been proposed for swVPA)
+- Our work features a best-search algorithm (as far as we know this is a new result, no such algorithm has been proposed for swVPA)
 - Finally, in order to highlight the central motivation of the paper which is the definition of a formal approach yielding a tractable algorithm to solve a challenging parsing problem, we will mention the challenging problem of Automatic Music Transcription as a main empirical target.
@@
 
-Regarding the best-search algorithm, in particular Figure 3, we agree that it could be better explained (we do also apologize for the several unfortunate typos…). 
-For the updates defined in Figure 3, intuitively, we assume that the values of d_bot and d_top are definitely fixed for every tuplet which has been extracted from the queue \Q, and we use these values in order to update the functions d_bot and d_top on the other tuplets (those  which are still in the queue).
-The typos In Figure 3 are the following:
+Regarding the best-search algorithm, in particular Figure 3, we agree that it could be better explained. We do also apologize for the several unfortunate typos… 
+For the updates defined in Figure 3, intuitively, we assume that the values of d_bot and d_top are definitely fixed for every tuplet which has been extracted from the queue \Q (in particular <q_1, q_2> and <q_1, p, q_2>), and we use these values (and the functions w of transitions) in order to update the values of d_bot and d_top on the other tuplets (those which are still in the queue).
+
+The typos in Figure 3 are the following:
 - At the beginning of Fig. 3, the quantification is “for all q_0, q_3 in Q and p’ in P”
@@
 This assumption is indeed strong, but in the context of solving symbolic weighted parsing, it is not unrealistic.
-In fact, for an instance of the problem, as stated in Section 5, we have only a finite number of fixed weight functions to consider (the functions in the automata’s transitions) and we may assume that they are effective (i.e.
+In fact, for an instance of the problem, as stated in Section 5, we have first only a finite number of fixed weight functions to consider (the functions in the automata’s transitions) and we may assume that they are effective (i.e.
 their global minimum are known).
@@
 Our interpretation is actually the second one mentioned in the review: using the notation of [22], Def.
-6, we consider the distance d({ s }, A) between a singleton distribution {s} (all words other than s have weight 0) and the distribution of A, wrt a partial application of T (on a subset of the alphabet).
+6, we consider the distance d({ s }, A) between a singleton distribution {s} (all words other than s have weight 0) and the distribution of A, with respect to a partial application of T (on a subset of the alphabet).
+
+The other comments raised by the reviewer are relevant, but for the sake of conciseness, we do not describe in this response how we will address them in the forthcoming version of our text. 
+Thanks again for these very helpful reviews!

### 2021-09-22 FST-TCS/rebuttal/rebuttal-reviews.md
cleaned mistake list

@@
 concepts from the literature and integrate Appendix D into the main paper.
-
-REMAINS TO DO :
-references :
-- [ ] 75: maybe cite Parsing as intersection by Nederhof and Satta (2003)
-- [ ] 126: you should cite Knuth's generalization of Dijkstra's algorithm for superiority, he defines the notion for functions of arbitrary arity, so not just semirings (but multioperator monoids); you should give an example for superiority in the semiring notation (with $\otimes$)
-Moreover, the `corresponds' is wrong, since superiority generalizes non-negative edge weights
-- [ ] PS: Knuth's definition of superiority includes monotonicity
-- [ ] 143: Lemma 4 is Lemma 3 in [21]
-- [ ] 195: references: there is earlier research on symbolic automata, e.g. Veanes (2013)
-- [x]  Ref [22] and [23] both seem to refer to the same work by Mohri
-
-- [ ] 51--53: the jump from automata to grammars is uneasy and not really necessary, as the paper mainly deals with automata
-- [ ] 57: weighted automata don't have ASTs, but runs
-- [x] 60: other approaches use multioperator monoids or valuation monoids, so `at minima a semiring' is wrong
-- [ ] 116: $\oplus$ `combines two values into a single one' as well (in fact, every   binary operation does)
-- [ ] 136: what implies?
-- [x] 153: quantify $x$ and the family ($y_i$) for $i ∈ ℕ$
-- [ ] 155--156: alphabets are usually finite and non-empty sets
-- [ ] 159: denoted by; also it seems like $\bar\Phi$ is a ($\Sigma \cup \Sigma \times \Delta$)-indexed family, why don't you introduce it like this
-- [ ] 161--171: use an itemize environment 
-- [ ] 162--165: I think these should also hold for $\Delta$, why don't you write this explicitly?
-- [ ] 188: the definition of $\bigoplus_\Delta$ can be seen to be implicit.  Please fully quantify the alphabets of a label theory
-- [ ] 208: What does $T$ compute?
-- [ ] 219: resp.\@
-- [ ] 220--221: you mean that a term CONTAINING $w_{ij}(…)$ equal to 0 is ignored
-- [ ] 238: text in the margin, also it is not clear whether the if refers to the entire line or just the right definition
-- [ ] 284: $x$ is a variable, so it should be typeset in math mode
-- [ ] 291: I'd at least say once that $\Delta_i$, $\Delta_c$, and $\Delta_r$ are pairwise disjoint (or explain the symbol)
-- [ ] 297: the extension here is not as in Section 3, there, the different function hat at least similar types
-- [ ] 299: every line expands into the margin; moreover, in the rightmost equation of line three, in the application of $w_c$, $p'$ and $q'$ are swapped
-- [ ] 316: equation extends into margin
-- [ ] 331: I think the application of $w_c$ is wrong (same with $w_r$ and $w_i$ later), shouldn't the : be a comma and what is q in the middle?
-- [ ] 336: how are states (which are in $Q$) pushed onto the stack if $\Gamma = \Delta_c \times P$?
-- [ ] 339: what are $j$ and $c$, do you mean $i$ and $d$?
-- [ ] 341: shouldn't there be a transition for returning from a measure?
-- [ ] 348: for?  do you mean accepted by?  but the language of an swA isn't defined either
-- [ ] 354: readers may have an easier time seeing this argument if you say that $S$ is extremal.  You should say in the preliminaries that S is extremal if and only if $\leq_\oplus$ is total.
-- [ ] 371: How is it ensured that the symbol $<c,p>$ isn't popped and pushed again later during the computation?
-- [x] 376: I do not understand the and at termination, and its correctness part.  What do you want to express?
-- [ ] 379: Almost all update equations need to compute infinite sums.
-Don't you think that the assumption of $\bar\Phi$ being effective is pretty strong and gives your the result `for free'?
-I think you should at least elaborate on this and say to what degree your effective label theory is a generalization of the decidable label theory of unweighted symbolic automata.
-- [ ] 629: typical references for regular tree grammars are Brainerd (1969) 
-- [ ] ​         and  Gecseg and Steinby (1984)
-- [ ] 654: an intuition for the $\bar q$ would have made understanding the construction far easier
-
-R3 :
-
-- [ ] l. 100. Shouldn't the first measure end before the second starts? This would make $O$ a forest rather than a tree - easily fixed with a "piece" root element.
-
-- [ ] l.
-220.
-The sentence starting here is somewhat confusing (and possibly wrong, depending on if $q'$ and $q''$ are intended to be distinct).
-I suppose that the point is that one can quit the recursion "early" whenever $w_{ij}$ returns $0$, but the language could be clearer, something like "Since $0$ is absorbing for $\times$, and the identity element for $+$ in $S$, if $w_{ij}(q,a,b,q'')$ is $0$ (meaning there is no transition from $q$ to $q'$ reading $a$ and $b$) that entire term can be ignored in the sum."
-
-- [ ] l. 288: same as on line 100, measure 1 should maybe end before measure 2
-- [ ] l. 434: $swM$??
 

### 2021-09-23 FST-TCS/rebuttal/rebuttal-reviews.md
reviews fsttcs

--- 2021-09-02 FST-TCS/rebuttal/rebuttal-reviews.md
@@
 - an sw-VPA over some target alphabet, and
-- an swT mapping words from the input alphabet to the internal symbols of the sw-VPA's alphabet,
-  aims to compute a nested word which minimizes both the weight assigned to it by the sw-VPA and the distance to the input word (which is computed by the swT).
+- an swT mapping words from the input alphabet to the internal symbols of the sw-VPA's alphabet,  aims to compute a nested word which minimizes both the weight assigned to it by the sw-VPA and the distance to the input word (which is computed by the swT).
+
 The authors show that this problem can be solved in polynomial time by combining the three constructions introduced earlier.
-Moreover, they
-  motivate their work by the example of the automatic transcription of music recordings into music scores.
+Moreover, they motivate their work by the example of the automatic transcription of music recordings into music scores.
 
@@
 
+---
 ## REVIEW 3

### 2021-09-23 introduction.tex
minor modifs

@@
 for some natural number~$d$,
-represent a division of the current
-time interval into $d$ sub-intervals of equal duration
-$\frac{\ell}{d}$ where $\ell$ is the length of the time interval attached to $\ccall{d}$.
-
+represent a division of the time interval attached to $\ccall{d}$,
+of duration $\ell$,
+into $d$ sub-intervals of equal duration $\frac{\ell}{d}$.
+%
 We will show that $O$ is a solution for the

### 2021-09-23 introduction.tex
remarks reviewer 1 fixed or in margin (blue)

@@
 %words carrying data values in an infinite domain (e.g. integers) e.g. data processing
-Some automata with memory extensions
+Some automata 
+\reviews{1) and grammar with storage?}
+with memory extensions
 allow restricted storage and comparison of input symbols
@@
 help in dealing with non-determinism by computing weight values.
-With an ambiguous grammar, there may exist several derivations
+With an ambiguous grammar, 
+\reviews{1) jump from automata to grammars is uneasy and not really necessary}
+there may exist several derivations
 (\emph{abstract syntax trees} -- AST) % \ie the result of an analyze.
@@
 %This is roughly the principle of \emph{weighted parsing} approaches
-\florent{WARNING: \cite{MorbitzVogler19weighted-parsing} much is more general that weighted parsing}
+\florent{introduce pb "weighted parsing" and ref.\cite{MorbitzVogler19weighted-parsing}.} 
 In \emph{weighted language models}~\cite{Goodman99SemiringParsing,Nederhof03weightedParsing,MorbitzVogler19weighted-parsing},
@@
 associative product operator~$\otimes$ to yield the weight of an AST.
+\reviews{1) AST for grammar, run for WA}
 A second operator~$\oplus$
@@
 and generalizing also~\WA, by handling infinite alphabets (Figure~\ref{fig:hierarchy}).
+\reviews{cite \cite{Herrmann16dlt,Herrmann20phd}}
 %
@@
 methods~\cite{Goodman99SemiringParsing,Nederhof03weightedParsing,MorbitzVogler19weighted-parsing},
-\florent{WARNING: \cite{MorbitzVogler19weighted-parsing} much is more general that weighted parsing}
+\reviews{1) WARNING: \cite{MorbitzVogler19weighted-parsing} incomparable to our def. weighted parsing}
 our approach proceeds in two steps.
@@
 (Bar-Hillel construction~\cite{GruneJacobs08parsing})
+\reviews{1) also \cite{NederhofSatta03ParsingIntersection} pasing as intersection}
 \florent{chap. intersection in~\cite{GruneJacobs08parsing}}
@@
 The main contributions of the paper are:
+\reviews{1) not $i$! see response.txt}
 ($i$)~the introduction of automata, \SWA, transducers, \SWT (Section~\ref{section:SWA}),

### 2021-09-23 main.tex
minor modifs
remarks reviewer 1 fixed or in margin (blue)

@@
 %\renewcommand{\philippe}[1]{} % uncomment for final
+\newcommand{\reviews}[1]{\todo[noline,size=\tiny,color=blue!30]{#1}}
+%\renewcommand{\reviews}[1]{} % uncomment for final
 
@@
 
+\begin{comment}
 %\begin{example}
@@
 %\end{example}
+\end{comment}
 
@@
 \label{sec:SWVPA-def}
-Let $\Delta$ be a countable alphabet
+\reviews{3) pb countable}
+Let $\Delta$ be a countable alphabet,
 %finite (large) or infinite,
 partitioned into three
-subsets~$\Deltai$, $\Deltac$, $\Deltar$,
+pairwise disjoint subsets~$\Deltai$, $\Deltac$, $\Deltar$,
 whose elements are respectively called
@@
 %we shall write respectively
-where $\Phii$, $\Phic$, $\Phir$ and~$\Phicx$ (with $\mathsf{x} \in \{ \mathsf{i}, \mathsf{c}, \mathsf{r}\}$)
+where $\Phii$, $\Phic$, $\Phir$ and~$\Phicx$ 
+(where $\mathsf{x}$ is either $\mathsf{i}$, $\mathsf{c}$, or $\mathsf{r}$)
 stand respectively
@@
 \begin{example} \label{ex:SWVPA}
+\reviews{revision ex. needed. see rev. 3) and app. $\weic$, $\weir$, $\weii$ wrong.}
 We consider a \SWVPA over the alphabet of Example~\ref{ex:nested-word}
@@
 \).
-The time interval $\iota$ attached to the~$\ccall{d}$ read 
+The time interval $\iota$ attached to~$\ccall{d}$ 
 must have a starting time $\tau + \frac{i\ell}{c}$,
@@
 By definition of $\leq_\oplus$, and since $\Semiring$ is total,
+\reviews{1) $\Semiring$ total $\to$ $\Semiring$ extremal}
 it holds that:

### 2021-09-23 semiring.tex
remarks reviewer 1 fixed or in margin (blue)

@@
 is {superiority}~\cite{Huang08advanceddynamic},
-which corresponds to the
+\reviews{1) also \cite{Knuth77}: "he defines the notion for functions of arbitrary arity, so not just semirings (but multioperator monoids"}
+which generalizes to the
 \emph{non-negative weights} condition in shortest-path algorithms~\cite{Dijkstra59anote}.
@@
 
+\reviews{1) example superiority}
 In~\cite{Huang08advanceddynamic}, when  property~$(i)$ holds,
@@
 % \emph{negative boundedness}
-It implies that, when looking for a best path in a graph whose edges
+Boundedeness implies that, when looking for a best path in a graph whose edges
 are weighted by values of $\Semiring$, the loops can be safely avoided,

### 2021-10-05 introduction.tex
macro \tstamp

@@
 Input events of $\Sigma$
-are pairs $\mu \ts{\tau}$, where $\mu$ is a
-MIDI pitch~\cite{Selfridge-Field97beyondMIDI}, %\cite{MIDIfile}
+are pairs $\mu \tstamp{\tau}$, where $\mu$ is a
+MIDI key number~\cite{Selfridge-Field97beyondMIDI}, %\cite{MIDIfile}
 and $\tau \in \mathbb{Q}$ is a timestamp in seconds.
@@
 to the recording of a live performance, \eg
-$I = (69\ts{0.07}),
-     (71\ts{0.72}),
-     (73\ts{0.91}),
-     (74\ts{1.05}),
-     (76\ts{1.36}),
-     (77\ts{1.71})$. % over interval $[0,2[$,
+$I = (69\tstamp{0.07}),
+     (71\tstamp{0.72}),
+     (73\tstamp{0.91}),
+     (74\tstamp{1.05}),
+     (76\tstamp{1.36}),
+     (77\tstamp{1.71})$. % over interval $[0,2[$,
 
 The output of parsing is a sequence of
-timed symbols $\nu\ts{\tau'}$ in an alphabet $\Delta$,
+timed symbols $\nu\tstamp{\tau'}$ in an alphabet $\Delta$,
 %events (or \emph{notes})
@@
 $O =$
-$\ccall{\mathsf{m}}\ts{[0,1]}$,
-$\ccall{2}\ts{[0,1]}$,
-$\mbox{\footnotesize A4}\ts{0}$,
-$\ccall{2}\ts{[\frac{1}{2},1]}$,
-$-\ts{\frac{1}{2}}$,
-$\ccall{2}\ts{[\frac{3}{4},1]}$,
-$\mbox{\footnotesize B4}\ts{\frac{3}{4}}$,
-$\mbox{\footnotesize C$\sharp$5}\ts{\frac{7}{8}}$,
-$\creturn{2}\ts{1}$,
-$\creturn{2}\ts{1}$,
-$\creturn{2}\ts{1}$,
-$\ccall{\mathsf{m}}\ts{[1,2]}$,
-$\ccall{3}\ts{[1,2]}$,
-$\mbox{\footnotesize D5}\ts{1}$,
-$\mbox{\footnotesize E5}\ts{\frac{4}{3}}$,
-$\mbox{\footnotesize F5}\ts{\frac{5}{3}}$,
-$\creturn{3}\ts{2}$,
-$\creturn{\mathsf{m}}\ts{2}$,
-$\creturn{\mathsf{m}}\ts{2}$.
+$\ccall{\mathsf{m}}\tstamp{[0,1]}$,
+$\ccall{2}\tstamp{[0,1]}$,
+$\mbox{\footnotesize A4}\tstamp{0}$,
+$\ccall{2}\tstamp{[\frac{1}{2},1]}$,
+$-\tstamp{\frac{1}{2}}$,
+$\ccall{2}\tstamp{[\frac{3}{4},1]}$,
+$\mbox{\footnotesize B4}\tstamp{\frac{3}{4}}$,
+$\mbox{\footnotesize C$\sharp$5}\tstamp{\frac{7}{8}}$,
+$\creturn{2}\tstamp{1}$,
+$\creturn{2}\tstamp{1}$,
+$\creturn{2}\tstamp{1}$,
+$\ccall{\mathsf{m}}\tstamp{[1,2]}$,
+$\ccall{3}\tstamp{[1,2]}$,
+$\mbox{\footnotesize D5}\tstamp{1}$,
+$\mbox{\footnotesize E5}\tstamp{\frac{4}{3}}$,
+$\mbox{\footnotesize F5}\tstamp{\frac{5}{3}}$,
+$\creturn{3}\tstamp{2}$,
+$\creturn{\mathsf{m}}\tstamp{2}$,
+$\creturn{\mathsf{m}}\tstamp{2}$.
 %\end{enumerate}

### 2021-10-05 main.tex
format llncs
macro \tstamp

@@
 % https://www.springer.com/gp/computer-science/lncs/conference-proceedings-guidelines
-%\input{main-llncs}
+\input{main-llncs}
 
@@
 % https://submission.dagstuhl.de/documentation/authors
-\input{main-lipics}
+%\input{main-lipics}
 
@@
 \newcommand{\lin}{\mathsf{lin}}
-\newcommand{\ts}[1]{\mathop{:}{#1}} % timestamp marker
+\newcommand{\tstamp}[1]{\mathpunct{:}{\scriptstyle #1}} % timestamp marker
 %\newcommand{\ts}[1]{^{#1}} % timestamp marker
@@
 \begin{array}{lclclcll}
-\wei_{11}(q_0, \mu\ts{\tau}, \nu\ts{\tau'}, q_0) & = &
-     \delta(\mu\ts{\tau}, \nu\ts{\tau'}) & \quad & %|\tau' - \tau| & \quad &
-\wei_{11}(q_1, \mu\ts{\tau}, \nu\ts{\tau'}, q_0) & = &
-     \delta(\mu\ts{\tau}, \nu\ts{\tau'}) &
+\wei_{11}(q_0, \mu\tstamp{\tau}, \nu\tstamp{\tau'}, q_0) & = &
+     \delta(\mu\tstamp{\tau}, \nu\tstamp{\tau'}) & \quad & %|\tau' - \tau| & \quad &
+\wei_{11}(q_1, \mu\tstamp{\tau}, \nu\tstamp{\tau'}, q_0) & = &
+     \delta(\mu\tstamp{\tau}, \nu\tstamp{\tau'}) &
      \mbox{if~} \nu \neq - \\ %|\tau' - \tau| & \quad &
-\wei_{01}(q_0, \varepsilon, \mathsf{-}\ts{\tau'}, q_0) & = & \one & &
-\wei_{01}(q_1, \varepsilon, \mathsf{-}\ts{\tau'}, q_0) & = & \one\\
-\wei_{10}(q_0, \mu\ts{\tau}, \varepsilon, q_1) & = & \alpha & & %\multicolumn{3}{l}{\mathrm{for~all~} b \in \Delta}
+\wei_{01}(q_0, \varepsilon, \mathsf{-}\tstamp{\tau'}, q_0) & = & \one & &
+\wei_{01}(q_1, \varepsilon, \mathsf{-}\tstamp{\tau'}, q_0) & = & \one\\
+\wei_{10}(q_0, \mu\tstamp{\tau}, \varepsilon, q_1) & = & \alpha & & %\multicolumn{3}{l}{\mathrm{for~all~} b \in \Delta}
 \end{array}
@@
 $O' =$
-$\ccall{\mathsf{m}}\ts{[0,1]}$,
-$\ccall{2}\ts{[0,1]}$,
-$\mbox{\footnotesize A4}\ts{0}$,
-$\ccall{2}\ts{[\frac{1}{2},1]}$,
-$-\ts{\frac{1}{2}}$,
-$\mbox{\footnotesize B4}\ts{\frac{3}{4}}$,
-$\creturn{2}\ts{1}$,
-$\creturn{2}\ts{1}$,
-$\ccall{\mathsf{m}}\ts{[1,2]}$,
-$\ccall{3}\ts{[1,2]}$,
-$\mbox{\footnotesize `C$\sharp$5`}\ts{1}$,
-$\mbox{\footnotesize D5}\ts{1}$,
-$\mbox{\footnotesize E5}\ts{\frac{4}{3}}$,
-$\mbox{\footnotesize F5}\ts{\frac{5}{3}}$,
-$\creturn{3}\ts{2}$,
-$\creturn{\mathsf{m}}\ts{2}$,
-$\creturn{\mathsf{m}}\ts{2}$
+$\ccall{\mathsf{m}}\tstamp{[0,1]}$,
+$\ccall{2}\tstamp{[0,1]}$,
+$\mbox{\footnotesize A4}\tstamp{0}$,
+$\ccall{2}\tstamp{[\frac{1}{2},1]}$,
+$-\tstamp{\frac{1}{2}}$,
+$\mbox{\footnotesize B4}\tstamp{\frac{3}{4}}$,
+$\creturn{2}\tstamp{1}$,
+$\creturn{2}\tstamp{1}$,
+$\ccall{\mathsf{m}}\tstamp{[1,2]}$,
+$\ccall{3}\tstamp{[1,2]}$,
+$\mbox{\footnotesize `C$\sharp$5`}\tstamp{1}$,
+$\mbox{\footnotesize D5}\tstamp{1}$,
+$\mbox{\footnotesize E5}\tstamp{\frac{4}{3}}$,
+$\mbox{\footnotesize F5}\tstamp{\frac{5}{3}}$,
+$\creturn{3}\tstamp{2}$,
+$\creturn{\mathsf{m}}\tstamp{2}$,
+$\creturn{\mathsf{m}}\tstamp{2}$
 (see also Fig.~\ref{fig:score-tree}).
@@
 \weic\bigl(q_{{i}/{c}},
-           \ccall{c}\ts{[\tau, \tau+\ell]}, q,
-           \ccall{d}\ts{\iota}, q_{{i+1}/{c}},
+           \ccall{c}\tstamp{[\tau, \tau+\ell]}, q,
+           \ccall{d}\tstamp{\iota}, q_{{i+1}/{c}},
            q_{{0}/{d}} \bigr) = \alpha_d
@@
 \weir\bigl(q_{{d}/{d}},
-           \ccall{d}\ts{[\tau, \tau+\ell]}, q_{{i+1}/{c}},
-           \,\creturn{d}\ts{\tau + \ell}, q_{{i+1}/{c}} \bigr) = \one
+           \ccall{d}\tstamp{[\tau, \tau+\ell]}, q_{{i+1}/{c}},
+           \,\creturn{d}\tstamp{\tau + \ell}, q_{{i+1}/{c}} \bigr) = \one
 \).
@@
 \weii\bigl(q_{{i}/{d}},
-           \ccall{d}\ts{[\tau, \tau+\ell]}, q_{{j}/{c}},
-           \mu\ts{\tau + \frac{i\ell}{d}}, q_{{i+1}/{d}} \bigr) = \alpha_\mu
+           \ccall{d}\tstamp{[\tau, \tau+\ell]}, q_{{j}/{c}},
+           \mu\tstamp{\tau + \frac{i\ell}{d}}, q_{{i+1}/{d}} \bigr) = \alpha_\mu
 \).
@@
 \weiec\bigl(q_{1/1},
-           \ccall{\mathsf{m}}\ts{[0, 1]}, q_{1/1},
+           \ccall{\mathsf{m}}\tstamp{[0, 1]}, q_{1/1},
            q_{1/1} \bigr) = \one
@@
 \weic\bigl(q_{1/1},
-          \ccall{\mathsf{m}}\ts{[\tau-1, \tau]}, q_{1/1},
-          \ccall{\mathsf{m}}\ts{[\tau, \tau+1]}, q_{1/1},
+          \ccall{\mathsf{m}}\tstamp{[\tau-1, \tau]}, q_{1/1},
+          \ccall{\mathsf{m}}\tstamp{[\tau, \tau+1]}, q_{1/1},
           q_{1/1} \bigr) = \one

### 2021-10-05 main.tex
proof draft + figure 3 separated

--- 2021-09-23 main.tex
@@
 \newcommand{\lin}{\mathsf{lin}}
-\newcommand{\ts}[1]{\mathop{:}{#1}} % timestamp marker
+%\newcommand{\ts}[1]{\mathop{:}{#1}} % timestamp marker
 %\newcommand{\ts}[1]{^{#1}} % timestamp marker
@@
 \noindent
-We present now a procedure for searching, for a~$\SWVPA$ $A$,
-a word %$w \in \Delta^*$
-of minimal weight for~$A$.
+%We present now a procedure for searching, for a~$\SWVPA$ $A$,
+%a word %$w \in \Delta^*$
+%of minimal weight for~$A$.
+We present now a procedure for searching a word of minimal weight for a~$\SWVPA$ $A$.
 %as stated in the following proposition. %\wrt~$\leq_\oplus$,
@@
 we shall consider an auxiliary function $b_\top :  Q \times P \times Q \to \Phic$.
+%%%%%% lydaddition %%%%%%
+The function $b_\bot$ will allow to compute $b_\top$ recursively in the case where the stack in not empty.
+%%%%%% lydaddition %%%%%%
 %
@@
 and ending in state $q'$ with the same stack $\gamma'$,
-such that moreover,
+such that
 the computation does not pop the pair $\<c, p>$ at the top of $\gamma'$
@@
 However, the computation can read $\<c, p>$ at the top of $\gamma'$,
-and can also push another pair $\< c', p'> \in \Gamma$ on $\gamma'$,
+and can also push another pair $\< c', p'> \in \Gamma$ on top of $\gamma'$,
 following the third case 
@@
 %
-By definition of $\weight_A$ in \eqref{eq:SWVPA-weight},
-using the symbol $\top$ for the part of the stack below $\<c, p>$
-(\ie the substack $\gamma$ in the above $\gamma' = \<c, p> \, \gamma$)
-ensures that this part is not touched during the computation.
+%By definition of $\weight_A$ in \eqref{eq:SWVPA-weight},
+%using the symbol $\top$ for the part of the stack below $\<c, p>$
+%(\ie the substack $\gamma$ in the above $\gamma' = \<c, p> \, \gamma$)
+%ensures that this part is not touched during the computation.
 % there cannot be a pop of $\<c, p>$ followed by a push of $\<c, p>$
@@
 %\textbf{for all} $q_0, q_3 \in Q$
-\Update $d_\bot$ if $\< q_1, q_2>$ was extracted, or $d_\top$ 
-if it was $\< q_1, p, q_2>$, 
-as in Figure~\ref{fig:best-update}.$\;$
+\Update $d_\bot$ if $\< q_1, q_2>$ was extracted (see Figure~\ref{fig:best-update-bottom}), or $d_\top$ 
+if it was $\< q_1, p, q_2>$
+(see Figure~\ref{fig:best-update-top}).$\;$
 }
@@
 \end{algorithm}
+
+\input{proof}
 
@@
   d_\bot(q_1, q_2) \otimes
-  \displaystyle\bigoplus_{\Deltai} \weiei(q_2, q_3)
-  \mathrm{if~} \< q_1, q_2> \notin \Q \\
+  \displaystyle\bigoplus_{\Deltai} \weiei(q_2, q_3)\\
 %
@@
   \bigl(
-  \weiec(q_0, p, q_1) \otimes
-   d_\top(q_1, p, q_2) \otimes
-  \displaystyle\bigoplus_{\Deltar} \weir(q_2, p, q_3)\bigr)\\
+  \weiec(q_0, p', q_1) \otimes
+   d_\top(q_1, p', q_2) \otimes
+  \displaystyle\bigoplus_{\Deltar} \weir(q_2, p', q_3)\bigr)\\
 %
@@
 \]
-%
+\caption{The update $d_\top$ when $\< q_1, q_2> \notin \Q$.}
+\label{fig:best-update-bottom}
+\end{figure}
+%
+
+\begin{figure}[ht]
 For every $q_0, q_3 \in Q$ and $p' \in P$, if $\< q_1, p, q_2> \notin \Q$
@@
   d_\top(q_1, p, q_2) \otimes
-  \displaystyle\bigoplus_{\Deltai} \weii(q_2, p, q_3)
-  \mathrm{if~} \< q_1, p, q_2> \notin \Q \\
+  \displaystyle\bigoplus_{\Deltai} \weii(q_2, p, q_3)\\
 %\multicolumn{3}{l}{\quad
@@
   d_\top(q_1, p, q_2) \otimes d_\top(q_2, p, q_3)\ 
-  \mathrm{if~} \< q_2, p, q_3> \notin \Q \\
+\\
 \end{array}
@@
 %\caption{Update $d_\bot$ with $\<q_1, q_2>$ or $d_\top$ with $\< q_1, p, q_2>$.}
-\caption{The update of $d_\bot$ when $\< q_1, q_2> \notin \Q$, 
-         and the update $d_\top$ when $\< q_1, q_2> \notin \Q$.}
-\label{fig:best-update}
+\caption{The update of $d_\bot$ when $\< q_1, q_2> \notin \Q$.}
+\label{fig:best-update-top}
 \end{figure}

### 2021-10-05 proof.tex
proof draft + figure 3 separated

\textbf{Proof of correctness}

To prove that the algorithm does compute the word with the best weight, we use the following invariant :
$$
<q_1, q_2> \notin Q \iff d_{bot} (q_1, q_2) = b_{bot} (q_1, q_2)
$$
In other words, a tuple $<q_1, q_2>$ is removed from the queue during an iteration only when $d_{bot}(q1, q2)$ has its minimal value.

Initialization : Before the while loop, no tuple has been extracted from the queue, so the invariant is trivially true.

Recursion :

Let's show by contradiction that at each iteration of the while loop, for the tuple $<q, q'>$ extracted from Q we have:
$$
d_{bot}(q, q') = b_{bot}(q, q'), 
$$

Let's consider the states $q_0$, $q_1$, $q_2$ and $q_3$ $\in Q$ such as they form a path $q$ ~...~>$q_0$ -> $q_1$ -> $q_2$ -> $q_3$ ~..~> $q'$.

We take $<q_0, q_1>$ as the first tuple to be removed from the queue. Then when $<q_0, q_1>$ is extracted from the queue, the algorithms updates $d_{bot}(q_0, q_2)$ and $d_{bot}(q_0, q_3)$. Thus, we have:
$$
d_{bot}(q_0, q_2) \geq b_{bot}(q_0, q_1) \\
d_{bot}(q_0, q_3) \geq b_{bot}(q_0, q_1)
$$
Let's show that it is true for each update possible, shown in Figure 3. TODO

Let $<q_0, q_3>$ be the first tuple for which the invariant is false:
$$
For <q_0, q_3> \notin Q, d_{bot} (q_0, q_3) \neq b_{bot} (q_0, q_3)
$$
Since $d_{bot}(q_0, q_2)$ has to be computed to compute $d_{bot}(q_0, q3)$, and because the invariant is true for  $d_{bot}(q_0, q_2)$ we can say:
$$
d_{bot}(q_0, q_2) = b_{bot}(q_0, q_2) \\
d_{bot}(q_0, q_2) \leq b_{bot}(q_0, q_3) \\
\\d_{bot}(q_0, q_2) \leq d_{bot}(q_0, q_3) \\
$$
 However, when $<q_0, q_3>$ was extracted, $<q_0, q_2>$ was still in $Q$. Thus, by definition 
$$
d_{bot}(q_0, q_2) \geq d_{bot}(q_0, q_3) \\
$$
Then the two inequalities are actually equal : 
$$
d_{bot}(q_0, q_2) = b_{bot}(q_0, q_2) = d_{bot}(q_0, q_3) = b_{bot}(q_0, q_3) \\
$$
We find that $d_{bot}(q_0, q_3) = b_{bot}(q_0, q_3)$, so by contradiction, we find that indeed $d_{bot}(q_0, q_3) = b_{bot}(q_0, q_3) iff <q_0, q_3> \notin Q$.

Termination: TODO

### 2021-10-05 main.tex

@@
 \newcommand{\lin}{\mathsf{lin}}
-%\newcommand{\ts}[1]{\mathop{:}{#1}} % timestamp marker
+\newcommand{\tstamp}[1]{\mathpunct{:}{\scriptstyle #1}} % timestamp marker
 %\newcommand{\ts}[1]{^{#1}} % timestamp marker

### 2021-10-05 main.tex
revision presentation algo best-search

@@
 %\textbf{for all} $q_0, q_3 \in Q$
-\Update $d_\bot$ if $\< q_1, q_2>$ was extracted (see Figure~\ref{fig:best-update-bottom}), or $d_\top$ 
-if it was $\< q_1, p, q_2>$
-(see Figure~\ref{fig:best-update-top}).$\;$
+\Update $d_\bot$ if $\< q_1, q_2>$ was extracted (Figure~\ref{fig:best-update-bottom}), or \\
+\Update $d_\top$ if $\< q_1, p, q_2>$ was extracted (Figure~\ref{fig:best-update-top}).$\;$
 }
@@
 \begin{figure}[ht]
-For every $q_0, q_3 \in Q$ and $p' \in P$, if $\< q_1, q_2> \notin \Q$ 
-\[
+\(
 \begin{array}{lcl}
-%\multicolumn{3}{l}{\mathrm{For~all~} q_0, q_3 \in Q, p \in P,}\\
-%
-d_\bot(q_1, q_3) & \opluseq &
-  d_\bot(q_1, q_2) \otimes
-  \displaystyle\bigoplus_{\Deltai} \weiei(q_2, q_3)\\
+\multicolumn{3}{l}{%
+ \mbox{For~every}\  q_0, q_3 \in Q,}\\[1mm] % if $\< q_1, p, q_2> \notin \Q$
 %
 d_\bot(q_0, q_3) & \opluseq &
-  {\displaystyle\bigoplus_{\Deltac}}
+  d_\bot(q_0, q_2) \otimes
+  \bigoplus_{\Deltai} \weiei(q_2, q_3)\ 
+  \mathrm{if~} \< q_0, q_2> \notin \Q \\
+%
+d_\bot(q_0, q_3) & \opluseq &
+  {\bigoplus_{\Deltac}}
   \bigl(
-  \weiec(q_0, p', q_1) \otimes
-   d_\top(q_1, p', q_2) \otimes
-  \displaystyle\bigoplus_{\Deltar} \weir(q_2, p', q_3)\bigr)\\
-%
-d_\bot(q_1, q_3) & \opluseq &
-  d_\bot(q_1, q_2) \otimes
-  \displaystyle\bigoplus_{\Deltar} \weier(q_2, q_3)\\
-%
-d_\bot(q_1, q_3) & \opluseq &
-  d_\bot(q_1, q_2) \otimes d_\bot(q_2, q_3)\  \mathrm{if~} \< q_2, q_3> \notin \Q \\
+  \weiec(q_0, p, q_1) \otimes
+   d_\top(q_1, p, q_2) \otimes
+  \bigoplus_{\Deltar} \weir(q_2, p, q_3)\bigr)\ 
+  \mathrm{if~} \< q_1, p, q_2> \notin \Q \\
+%
+d_\bot(q_0, q_3) & \opluseq &
+  d_\bot(q_0, q_2) \otimes
+  \bigoplus_{\Deltar} \weier(q_2, q_3)\ 
+  \mathrm{if~} \< q_0, q_2> \notin \Q \\ 
+%
+d_\bot(q_0, q_3) & \opluseq &
+  d_\bot(q_0, q_2) \otimes d_\bot(q_2, q_3)\  
+  \mathrm{if~} \< q_0, q_2>\notin \Q, \< q_2, q_3> \notin \Q \\[-1mm]
 \end{array}
-\]
-\caption{The update $d_\top$ when $\< q_1, q_2> \notin \Q$.}
+\)
+\caption{\textbf{update} $d_\bot$ (Algorithm~\ref{algo:Dijkstra}).}
 \label{fig:best-update-bottom}
@@
 \begin{figure}[ht]
-For every $q_0, q_3 \in Q$ and $p' \in P$, if $\< q_1, p, q_2> \notin \Q$
-\[
+\(
 \begin{array}{lcl}
-d_\top(q_1, p, q_3) & \opluseq &
-  d_\top(q_1, p, q_2) \otimes
-  \displaystyle\bigoplus_{\Deltai} \weii(q_2, p, q_3)\\
+\multicolumn{3}{l}{%
+ \mbox{For~every}\  q_0, q_3 \in Q\ \mbox{and}\ p \in P,}\\[1mm] % if $\< q_1, p, q_2> \notin \Q$
+d_\top(q_0, p, q_3) & \opluseq &
+  d_\top(q_0, p, q_2) \otimes
+  \bigoplus_{\Deltai} \weii(q_2, p, q_3)\
+  \mathrm{if~} \< q_0, p, q_2> \notin \Q \\
 %\multicolumn{3}{l}{\quad
@@
 %x_c \mapsto \weii(q_2, x_c, p, a, q_3)}\\
-d_\top(q_0, p', q_3) & \opluseq &
-  {\displaystyle\bigoplus_{\Deltac}}^2
-  \bigl[ \bigl( \weic(q_0, p', q_1, p) \otimes_2
-  d_\top(q_1, p, q_2) \bigr) \otimes_2
-  {\displaystyle\bigoplus_{\Deltar}} \weir(q_2, p, q_3) \bigr]\\
+d_\top(q_0, p, q_3) & \opluseq &
+  {\bigoplus^2_{\Deltac}}
+  \bigl[ \bigl( \weic(q_0, p, q_1, p) \otimes_2
+  d_\top(q_1, p', q_2) \bigr) \otimes_2
+  {\bigoplus_{\Deltar}} \weir(q_2, p', q_3) \bigr]\
+  \mathrm{if~} \< q_1, p', q_2> \notin \Q \\
 %\multicolumn{3}{l}{\quad
@@
 %
-d_\top(q_1, p, q_3) & \opluseq &
-  d_\top(q_1, p, q_2) \otimes d_\top(q_2, p, q_3)\ 
-\\
+d_\top(q_0, p, q_3) & \opluseq &
+  d_\top(q_0, p, q_2) \otimes d_\top(q_2, p, q_3)\ 
+  \mathrm{if~} \< q_0, p, q_2> \notin \Q, < q_2, p, q_3> \notin \Q \\[-1mm]
 \end{array}
-\]
+\)
 %$\weii(q_2, p, q_3)_a \in \Phic$, for $a \in \Deltai$,
@@
 %\caption{Update $d_\bot$ with $\<q_1, q_2>$ or $d_\top$ with $\< q_1, p, q_2>$.}
-\caption{The update of $d_\bot$ when $\< q_1, q_2> \notin \Q$.}
+\caption{\textbf{update} $d_\top$ (Algorithm~\ref{algo:Dijkstra}).}
 \label{fig:best-update-top}

### 2021-10-07 main.tex
example 5 makeover

@@
 \reviews{revision ex. needed. see rev. 3) and app. $\weic$, $\weir$, $\weii$ wrong.}
-We consider a \SWVPA over the alphabet of Example~\ref{ex:nested-word}
+We consider a \SWVPA over the alphabet of Example~\ref{ex:nested-word} and with $P = Q$,
 that expresses a weight related to the music notation,
 or more precisely to its structural complexity.
-Given a set of equivalent representations, we aim at choosing the simpler one.
-
-%Let us assume a third component in the symbols of $\Deltac$
-%\philippe{Je trouve l'exemple un peu confus. Je ne sais pas
-%quel est "l'associate time interval" d'un symbole. Je ne sais
-%pas à quoi fait référence "the following call transition". Je
-%n'ose pas trop toucher mais je crois qu'il faut revoir.}
-%which is the duration of the associated time interval.
-%It is used to check the timestamps in the symbols.
-%
-For instance, the following call transition starts in state
-$q_{{i}/{c}}$, meaning that the current interval has number~$i < c$
-amongst $c$ sub-intervals of duration~$\ell$ (as indicated by the interval at the stack's top).
-It reads a new time-division symbol~$\ccall{d}$:
+Given a set of equivalent representations, we aim at choosing the simpler one, \ie the one with the smallest weight.
+
+\lydia{We should find a better variable name than j... it does not render well in fractions}
+Let's say the top of the stack is a tuple of the call~$\ccall{n}\tstamp{[\tau, \tau+\ell]}$ and a state $q$.
+Let's now consider a new call transition starting in state $q_{\frac{j}{n}}$, that is the $j$-th state of the $n$ sub-intervals of duration~$\frac{\ell}{n}$.
+We thus read a new time-division symbol ~$\ccall{d}$ and compute the weight of the transition from $q_{\frac{j}{n}}$ to $q_{\frac{1}{d}}$ (the first state after the new ~$\ccall{d}$), reading the top of the stack $<\ccall{n}\tstamp{[\tau, \tau+\ell]}, q>$, and pushing $<\ccall{d}\tstamp{\iota}, q_{\frac{j+1}{c}}>$ on top, or:
 \(
-\weic\bigl(q_{{i}/{c}},
-           \ccall{c}\tstamp{[\tau, \tau+\ell]}, q,
-           \ccall{d}\tstamp{\iota}, q_{{i+1}/{c}},
-           q_{{0}/{d}} \bigr) = \alpha_d
+\weic\bigl(q_{\frac{j}{n}},
+           \ccall{n}\tstamp{[\tau, \tau+\ell]}, q,
+           \ccall{d}\tstamp{\iota}, q_{\frac{j+1}{c}},
+           q_{\frac{1}{d}} \bigr) = \alpha_d
 \).
-The time interval $\iota$ attached to~$\ccall{d}$ 
-must have a starting time $\tau + \frac{i\ell}{c}$,
-where~$\tau$ and~$\ell$ are respectively
-the starting time and duration of the interval attached to
-the previous time-division symbol read $\ccall{c}$ (found on the stack).
-%plus the total duration of intervals read so far, $i \ell$, and
-%the new current interval duration is~$\frac{\ell}{d}$.
-%
-And it must have a duration $\frac{\ell}{d}$.
-
-\noindent
-The weight of the above transition is $\alpha_d$.
-We can penalize \eg triplets compared to duplets with $\alpha_2 < \alpha_3$.
-%
-\noindent
-Along with $\ccall{d}$,
-the above transition pushes
-the state $q_{{i+1}/{c}}$ on the stack,
-in order to start the next sub-interval after reading 
-a return symbol $\creturn{d}$,
-with the transition:
+\lydia{is it ok to explain precisely this computation and not the others ? is it redundant with the explaniation before equation (5) ?}
+\noindent
+Reading the $k$-th musical event $\mu$ in this current sub-interval is computed with:
 \(
-\weir\bigl(q_{{d}/{d}},
-           \ccall{d}\tstamp{[\tau, \tau+\ell]}, q_{{i+1}/{c}},
-           \,\creturn{d}\tstamp{\tau + \ell}, q_{{i+1}/{c}} \bigr) = \one
+\weii\bigl(q_{{k}/{d}},
+           \ccall{d}\tstamp{[\tau, \tau+\ell]}, q_{\frac{j}{n}},
+           \mu\tstamp{\tau + \frac{j\ell}{d}}, q_{\frac{k+1}{d}} \bigr) = \alpha_\mu
 \).
-
-\noindent
-Reading a musical event $\mu$ is done with:
+Finally, this sub-interval will end after reading a return symbol $\creturn{d}$ and compute:
 \(
-\weii\bigl(q_{{i}/{d}},
-           \ccall{d}\tstamp{[\tau, \tau+\ell]}, q_{{j}/{c}},
-           \mu\tstamp{\tau + \frac{i\ell}{d}}, q_{{i+1}/{d}} \bigr) = \alpha_\mu
+\weir\bigl(q_{\frac{d}{d}},
+           \ccall{d}\tstamp{[\tau, \tau+\ell]}, q_{\frac{i+1}{c}},
+           \,\creturn{d}\tstamp{\tau + \ell}, q_{\frac{i+1}{c}} \bigr) = \one
 \).
-% cas des barres de mesure?
-%For the case of a new measure symbol,
-%$q_{1/1}$ is a special (initial) state.
-The transition to start reading the first measure is:
+\noindent
+We described earlier the special cases where the stack is empty, which applies for example when reading the first measure for which we only push $<\ccall{\mathsf{m}}\tstamp{[0, 1]}, q_{1/1}>$ in the stack:
 \(
@@
            q_{1/1} \bigr) = \one
-\),
-and for a new measure:
+\).
+In comparison, any other new measure would compute instead:
 \(
@@
 \).
+
+Each transition thus has a weight computing an overall weight for one representation of music notation. 
+When setting the weights, we decide which notation is preferred if possible. 
+For example, let's say we want to prioritize tuplets over triplets, then we set $\alpha_2$ and $\alpha_3$ such as $\alpha_2 > \alpha_3$.
+In conclusion, a \SWVPA is capable of computing several representations of the same piece of music, allowing to choose the best one.
 \endex
@@
 obtained from $t$ by removing all symbols in $\Delta \setminus \Deltai$.
-%\end{equation}
 %

### 2021-10-10 main.tex
old example 5 in comment
minor

@@
 \(
-\weii\bigl(q_{{k}/{d}},
+\weii\bigl(q_{\frac{k}{d}},
            \ccall{d}\tstamp{[\tau, \tau+\ell]}, q_{\frac{j}{n}},
@@
 \noindent
-We described earlier the special cases where the stack is empty, which applies for example when reading the first measure for which we only push $<\ccall{\mathsf{m}}\tstamp{[0, 1]}, q_{1/1}>$ in the stack:
+We described earlier the special cases where the stack is empty, which applies for example when reading the first measure for which we only push $<\ccall{\mathsf{m}}\tstamp{[0, 1]}, q_{\frac{1}{1}}>$ in the stack:
 \(
-\weiec\bigl(q_{1/1},
+\weiec\bigl(q_{\frac{1}{1}},
            \ccall{\mathsf{m}}\tstamp{[0, 1]}, q_{1/1},
@@
 \(
-\weic\bigl(q_{1/1},
-          \ccall{\mathsf{m}}\tstamp{[\tau-1, \tau]}, q_{1/1},
-          \ccall{\mathsf{m}}\tstamp{[\tau, \tau+1]}, q_{1/1},
-          q_{1/1} \bigr) = \one
+\weic\bigl(q_{\frac{1}{1}},
+          \ccall{\mathsf{m}}\tstamp{[\tau-1, \tau]}, q_{\frac{1}{1}},
+          \ccall{\mathsf{m}}\tstamp{[\tau, \tau+1]}, q_{\frac{1}{1}},
+          q_{\frac{1}{1}} \bigr) = \one
 \).
@@
 \end{example}
+
+%\begin{example} \label{ex:SWVPA}
+%\reviews{revision ex. needed. see rev. 3) and app. $\weic$, $\weir$, $\weii$ wrong.}
+%We consider a \SWVPA over the alphabet of Example~\ref{ex:nested-word} and with $P = Q$
+%that expresses a weight related to the music notation,
+%or more precisely to its structural complexity.
+%Given a set of equivalent representations, we aim at choosing the simpler one.
+%
+%Let's say the top of the stack is a tuple of the call~$\ccall{n}\tstamp{[\tau, \tau+\ell] and a state $q$.
+Let's now consider a new call transition starting in state $q_{\frac{j}{n}}$, that is the $j$-th state of the $n$ sub-intervals of duration~$\frac{\ell}{n}$.
+%We thus read a new time-division symbol ~$\ccall{d}$ and compute the weight of the transition from q_{\frac{j}{n}} to q_{\frac{1}{d} (the first state after the new ~$\ccall{d}$), reading the top of the stack <\ccall{n}\tstamp{[\tau, \tau+\ell]}, q>, and pushing <\ccall{d}\tstamp{\iota}, q_{\frac{j+1}{c}}> on top, or:
+%\(
+%\weic\bigl(q_{\frac{j}{n}},
+           %\ccall{n}\tstamp{[\tau, \tau+\ell]}, q,
+           %\ccall{d}\tstamp{\iota}, q_{\frac{j+1}{c}},
+           %q_{\frac{1}{d}} \bigr) = \alpha_d
+%%\).
+%
+%\noindent
+%Reading the $k$-th musical event $\mu$ in this current sub-interval is computed with:
+%\(
+%\weii\bigl(q_{{k}/{d}},
+           %\ccall{d}\tstamp{[\tau, \tau+\ell]}, q_{\frac{j}{n}},
+           %\mu\tstamp{\tau + \frac{j\ell}{d}}, q_{\frac{k+1}{d}} \bigr) = \alpha_\mu
+%\).
+%Finally, this sub-interval will end after reading a return symbol \creturn{d} which will compute the transition:
+%\(
+%\weir\bigl(q_{\frac{d}{d}},
+           %\ccall{d}\tstamp{[\tau, \tau+\ell]}, q_{\frac{i+1}{c}},
+           %\,\creturn{d}\tstamp{\tau + \ell}, q_{\frac{i+1}{c}} \bigr) = \one
+%\).
+%
+%\noindent
+%We described earlier the special cases where the stack is empty, which applies for example when reading the first measure for which we push <\ccall{\mathsf{m}}\tstamp{[0, 1]}, q_{1/1}>, but don't read anything on the stack:
+%\(
+%\weiec\bigl(q_{1/1},
+ %          \ccall{\mathsf{m}}\tstamp{[0, 1]}, q_{1/1},
+%           q_{1/1} \bigr) = \one
+%\).
+%In comparison, any other new measure would compute instead:
+%\(
+%\weic\bigl(q_{1/1},
+%          \ccall{\mathsf{m}}\tstamp{[\tau-1, \tau]}, q_{1/1},
+%          \ccall{\mathsf{m}}\tstamp{[\tau, \tau+1]}, q_{1/1},
+%          q_{1/1} \bigr) = \one
+%\).
+%\endex
+%\end{example}
 
@@
 the minimum weight of a computation of~$A$
-starting in state $q$ with a non-empty stack
-$\gamma' = \<c, p> \, \gamma \in \Gamma^+$,
+starting in state $q$, 
+with a non-empty stack $\gamma' = \<c, p> \, \gamma \in \Gamma^+$,
 and ending in state $q'$ with the same stack $\gamma'$,
@@
    d_\top(q_1, p, q_2) \otimes
-  \bigoplus_{\Deltar} \weir(q_2, p, q_3)\bigr)\ 
+  \bigoplus^2_{\Deltar} \weir(q_2, p, q_3)\bigr)\ 
   \mathrm{if~} \< q_1, p, q_2> \notin \Q \\
@@
   {\bigoplus^2_{\Deltac}}
-  \bigl[ \bigl( \weic(q_0, p, q_1, p) \otimes_2
+  \bigl[ \bigl( \weic(q_0, p, p', q_1) \otimes_2
   d_\top(q_1, p', q_2) \bigr) \otimes_2
@@
   d_\top(q_0, p, q_2) \otimes d_\top(q_2, p, q_3)\ 
-  \mathrm{if~} \< q_0, p, q_2> \notin \Q, < q_2, p, q_3> \notin \Q \\[-1mm]
+  \mathrm{if~} \< q_0, p, q_2> \notin \Q, \< q_2, p, q_3> \notin \Q \\[-1mm]
 \end{array}

### 2021-10-10 proof.tex
minor

@@
 $$
-We find that $d_{bot}(q_0, q_3) = b_{bot}(q_0, q_3)$, so by contradiction, we find that indeed $d_{bot}(q_0, q_3) = b_{bot}(q_0, q_3) iff <q_0, q_3> \notin Q$.
+We find that $d_{bot}(q_0, q_3) = b_{bot}(q_0, q_3)$, so by contradiction, we find that indeed $d_{bot}(q_0, q_3) = b_{bot}(q_0, q_3)$ iff 
+$\<q_0, q_3> \notin Q$.
 

### 2021-10-11 proof.tex
alternative algo

@@
 $$
-<q_1, q_2> \notin Q \iff d_{bot} (q_1, q_2) = b_{bot} (q_1, q_2)
+\<q_1, q_2> \notin Q \iff d_{\bot}(q_1, q_2) = b_{\bot}(q_1, q_2)
 $$
-In other words, a tuple $<q_1, q_2>$ is removed from the queue during an iteration only when $d_{bot}(q1, q2)$ has its minimal value.
+In other words, a tuple $\<q_1, q_2>$ is removed from the queue during an iteration only when $d_{\bot}(q1, q2)$ has its minimal value.
 

### 2021-10-11 main.tex
alternative algo
add def. of transitions

@@
 \newcommand{\E}{\mathbb{E}}
+\newcommand{\F}{\mathcal{F}}
+\newcommand{\G}{\mathcal{G}}
 %\newcommand{\P}{\mathcal{P}}
@@
 %\newcommand{\ts}[1]{^{#1}} % timestamp marker
-
 \newcommand{\call}[1]{\ensuremath #1} %{\ensuremath \langle_{#1}}
@@
 %\newcommand{\creturn}[1]{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
+\newcommand{\src}{\mathsf{src}}
+\newcommand{\trg}{\mathsf{trg}}
 
@@
 \noindent
-We call the \emph{number of transitions} of $T$ the number of pairs of states
-$q, q' \in Q$ such that $\wei_{10}$ or $\wei_{01}$ or $\wei_{11}$
-is not the constant $\zero$.
-%
-\noindent
-For convenience, we shall sometimes present transitions
-as functions of
-$Q \times (\Sigma \cup \{ \varepsilon \}) \times (\Delta \cup \{ \varepsilon \}) \times Q \to \Semiring$,
-overloading the function names,
-such that, for every $q, q' \in Q$, $a \in \Sigma$,  $b \in \Delta$:
+%We call the \emph{number of transitions} of $T$ the number of pairs of states
+%$q, q' \in Q$ such that $\wei_{10}$ or $\wei_{01}$ or $\wei_{11}$
+%is not the constant $\zero$.
+%
+\noindent
+A tuple of $Q \times (\Sigma \cup \{ \varepsilon \}) \times (\Delta \cup \{ \varepsilon \}) \times Q \to \Semiring$
+is called a \emph{transition} of $T$.
+For convenience, we shall sometimes present the above $\wei_{10}$, $\wei_{01}$, $\wei_{11}$
+%(overloading the function names),
+as mappings associating to every transition a weight value in $\Semiring$ as follows,
+for every $q, q' \in Q$, $a \in \Sigma$,  $b \in \Delta$:
 \[
@@
 %\florent{mv formal def. weight for \SWA to app.?}
-As above in the case of \SWT,
-when $\wei_1(q, q') = \phi \in \Phi_\Sigma$,
+A \SWA \emph{transition} is a triplet of $Q \times \Sigma \times Q$, 
+and, as in the case of \SWT,
+we may write $\wei_1(q, a, q')$ for~$\phi(a)$
+when $\wei_1(q, q') = \phi \in \Phi_\Sigma$.
 %respectively $\mathsf{in}(q) = \phi$, $\mathsf{out}(q') = \phi$,
-we may write $\wei_1(q, a, q')$ for~$\phi(a)$.
 %respectively $\mathsf{in}(q, a) = \phi(a)$, $\mathsf{out}(q', a) = \phi(a)$.
@@
 \]
-
+The tuples in the definition domains of the above functions 
+are also called \emph{transitions} of the \SWVPA.
+%
 \noindent
 %The intuition is the following for the above transitions.
-$\weiei$, $\weiec$, and $\weier$ describe the cases where the stack is empty.
-%\lydia{moved this to the beginning}
-%
-\noindent
-$\weii$ and $\weiei$ both read an input internal symbol $a$ and change state from $q$ to $q'$,
-without changing the stack.
+Intuitively, $\weiei$, $\weiec$, and $\weier$ describe the cases where the stack is empty.
+The transitions of $\weii$ and $\weiei$ both read an internal symbol~$a$ 
+and change state from $q$ to $q'$, without changing the stack.
 Moreover, $\weii$ reads a pair made of
@@
 \noindent
-$\weic$ and $\weiec$ read the input call symbol $\call{c}'$,
+The transitions of $\weic$ and $\weiec$ read the call symbol $\call{c}'$,
 push it to the stack along with $p'$, and change state from $q$ to to $q'$.
@@
 \noindent
-$\weir$ and $\weier$ read the input return symbol $\return{r}$, and change state from $q$ to to $q'$.
-Moreover, $\weir$ reads and pops from stack a pair made of $\call{c}$ and $p$,
+Finally, $\weir$ and $\weier$ read the return symbol $\return{r}$, and change state from $q$ to to $q'$.
+Moreover, $\weir$ reads and pops from stack a pair made of $\call{c}$ and $p$
 ($\call{c}$ is compared to $\return{r}$).
@@
 
-Formally, the transitions of the automaton~$A$ are defined
+Formally, the computations of the automaton~$A$ are defined
 with an intermediate function $\weight_A$, like in Section~\ref{sec:SWT}.
@@
 %of minimal weight for~$A$.
-We present now a procedure for searching a word of minimal weight for a~$\SWVPA$ $A$.
+We present now a procedure for searching a word of minimal weight for a~$\SWVPA$~$A$.
 %as stated in the following proposition. %\wrt~$\leq_\oplus$,
@@
 %of the function $\weight_A$.
+
+\input{altalgo}
 

### 2021-10-11 algo-best-alt.tex
include proofs in text. rename files.

% !TEX root = main.tex
%

\subsection*{Alternative Algorithm for best-search}
We associate to the \SWVPA $A$ a $\Semiring$-labeled graph
$\G_A = \< V_A, E_A, \eta_A >$
with set of vertices $V_A = (Q \times Q) \cup (Q \times P \times Q)$,
set of edges $E_A = V_A \times V_A$, and with an edge labelling  
$\eta_A : E_A \to \Semiring$ defined as follows,
for $q_0, q_1, q_2, q_3 \in Q$, $p, p' \in P$:
%
\[
\begin{array}{rclcl}
\< q_0, q_1> & \to & \< q_0, q_2> & : &
\bigoplus_{\Deltai} \weiei(q_1, q_2)\\
%
\< q_1, p, q_2> & \to & \<q_0, q_3> & : &
\bigoplus_{\Deltac} \bigl( \weiec(q_0, p, q_1) \otimes \bigoplus^2_{\Deltar} \weir(q_2, p, q_3)\bigr)\\ 
%
\< q_0, q_2> & \to & \< q_0, q_3> & : & 
\bigoplus_{\Deltar} \weier(q_2, q_3)\\
% 
\< q_0, p, q_1> & \to & \< q_0, p, q_2> & : &
\bigoplus_{\Deltai} \weii(q_1, p, q_2)\\[1mm]
%
\< q_1, p', q_2> & \to & \< q_0, p, q_3> & : &
\bigoplus^2_{\Deltac} \bigl[ \weic(q_0, p, p', q_1) \otimes_2 {\bigoplus_{\Deltar}} \weir(q_2, p', q_3) \bigr]\\
\end{array}
\]
Removing from $\G_A$ every edge $e \in E_A$ such that $\eta(e) = \zero$ 
is preserving the results below.
%
A \emph{path} of $\G_A$ is a sequence $e_1,\ldots, e_n \in V^*_A$ such that 
for all $1 \leq i < n$, the second component of $e_i$ is the first component of $e_{i+1}$.
We write $\src(\pi) = e_1$ and $\trg(\pi) = e_n$.
The weight of the path is $\weight(\pi) = \bigotimes_{i=1}^{n} \eta(e_i)$.

%Intuitively, 
\begin{lemma}[Correctness]
For all $q_0, q_1, q_2 \in Q$, $p\in P$, 
and all path $\pi$ of $\G_A$ with $\src(\pi) = \<q_0, q_0>$,
($i$) if $\trg(\pi) = \< q_1, q_2>$, 
then there exists $u \in \Delta^*$ such that 
$\weight_A(q_1[\bot], u, q_1[\bot]) = \weight(\pi)$,
%
($ii$) if $\trg(\pi) = \< q_1, p, q_2>$,
then there exists $v \in \Delta^*$ such that 
${\displaystyle\bigoplus_{c \in \Deltac}}
 \weight_A\bigl(\configup{q_1}{\< c, p> \stackup \top }, v, \configup{q_2}{\<c, p> \stackup \top}\bigr)
 = \weight(\pi)$.
\end{lemma}
%
\begin{proof}
We prove ($i$) and ($ii$) simultaneously by induction on the length of the path $\pi$.
...\qed
\end{proof}

\begin{lemma}[Completeness]
For all $q_1, q_2 \in Q$, $p\in P$, 
($i$) there exists a path $\pi$ of $\G_A$ 
such that $\src(\pi) = \<q_0, q_0>$ for some $q_0 \in Q$, 
$\trg(\pi) = \< q_1, q_2>$,
and $\weight(\pi) = b_\bot(q_1, q_2)$, 
($ii$) there exists a path $\pi'$ of $\G_A$ 
such that $\src(\pi) = \<q_0, q_0>$ for some $q_0 \in Q$, 
$\trg(\pi) = \< q_1, p, q_2>$,
and $\weight(\pi) = b_\top(q_1, q_2)$.
\end{lemma}
%
\begin{proof}
Since $\Semiring$ is  total, there exists $t \in \Delta^*$ such that  
$b_\bot(q_1, q_2) = \weight_A\bigl(\config{q_1}{\bot}, t, \config{q_2}{\bot}\bigr)$.
We represent a computation of $A$ on $t$ by a sequence of transitions...

We show that we can construct a path $\pi$ from $t$.
...\qed
\end{proof}

### 2021-10-11 main.tex
include proofs in text. rename files.

@@
 \begin{proof}
-We do a classical product construction, see Appendix~\ref{sec:product} for details.
+We do a classical product construction. % see Appendix~\ref{sec:product} for details.
 %similar to the case of the Boolean semiring~\cite{dAntonyAlur14SVPDA}.
+\input{app-product}
 \end{proof}
@@
 
-\input{altalgo}
-
-\begin{algorithm}
-\SetKw{Extract}{extract}
-\SetKw{Update}{update}
-
-\textbf{initially} let $\Q = (Q \times Q) \cup (Q \times P \times Q)$,
-and let $d_\bot(q_1, q_2) = d_\top(q_1, p, q_2) = \one$
-if $q_1 = q_2$ and $d_\bot(q_1, q_2) = d_\top(q_1, p, q_2) = \zero$ otherwise$\;$
-
-\smallskip\noindent
-\While{$\Q \neq \emptyset$}{
-
-%\noindent\quad
-\Extract $\< q_1, q_2>$ or $\< q_1, p, q_2>$ from $\Q$
-%
-%\noindent\quad
-such that $d_\bot(q_1, q_2)$, 
-resp. $\bigoplus_{\Deltac} d_\top(q_1, p, q_2)$, is minimal in $\Semiring$ \wrt $\leq_\oplus$$\;$
-
-%\noindent\quad
-%\textbf{for all} $q_0, q_3 \in Q$
-\Update $d_\bot$ if $\< q_1, q_2>$ was extracted (Figure~\ref{fig:best-update-bottom}), or \\
-\Update $d_\top$ if $\< q_1, p, q_2>$ was extracted (Figure~\ref{fig:best-update-top}).$\;$
-}
-\caption{Best search for \SWVPA}
-\label{algo:Dijkstra}
-\end{algorithm}
-
-\input{proof}
-
-Algorithm~\ref{algo:Dijkstra} constructs iteratively, using a priority queue $\Q$, 
-two markings
-$d_\bot : Q \times Q \to \Semiring$ and
-$d_\top : Q \times P \times Q \to \Phic$,
-%of the triplets $\<q, \sigma, q'>$
-%of states of $A$ by weight values in $\Semiring$,
-that converge eventually to $b_\top$ and $b_\bot$. 
-%
-\florent{intuition sur update en commentaire}
-%Intuitively, in the update defined in Figure~\ref{fig:best-update},
-%we assume that the values of $d_\bot$ and $d_\top$ are definitely 
-%fixed for every tuplet which has been extracted from the queue $\Q$ 
-%(like in particular $\<q_1, q_2>$ or $\< q_1, p, q_2>$ when the algorithm calls this update), 
-%and we use these values 
-%(and the functions $w$ of transitions) 
-%in order to update the values of $d_\bot$ and $d_\top$ 
-%on the other tuplets (those which are still in the queue $\Q$).
-%
-\florent{explication Fig.~\ref{fig:best-update} (update) suivant cas de \eqref{eq:SWVPA-weight}?}
-\florent{The infinite sums in the updates of $d$ in Algorithm~\ref{algo:Dijkstra},
- Figure~\ref{fig:best-update} are well defined since~$\Semiring$ is complete.}
-%
-% COMPLEXITY
-%The algorithm performs $2.|Q|^2$ iterations until $P$ is empty,
-%and each iteration has a time complexity $O(|Q|^2 . |P|)$.
-%\florent{complexity: detail with and states?}
-%That gives a time complexity $O(|Q|^4 . |P|)$.
-%It can be reduced by implementing $P$ as a priority queue,
-%prioritized by the value returned by $d$.
-%** where  $T$ is the worst time complexity for an oracle computing the values
-%of Definition~\ref{def:effective}.
-%$|Q|^3.\log(|Q|^2)$
-%
-%Algorithm~\ref{algo:Dijkstra}
-It terminates in PTIME (when $\Q$ is empty), 
-and its correctness is ensured by~\eqref{eq:min} and  
-the following invariants: 
-$\< q_1, q_2> \notin \Q$ iff $d_\bot(q_1, q_2) = b_\bot(q_1, q_2)$, 
-and $\< q_1, p, q_2> \notin \Q$ iff 
-$\bigoplus_{\Deltac} d_\top(q_1, p, q_2) = \bigoplus_{\Deltac} b_\top(q_1, p, q_2)$.
-\florent{(see Appendix~\ref{sec:bestsearch}).}
-%of Algorithm~\ref{algo:Dijkstra}.
-
-\begin{figure}[ht]
-\(
-\begin{array}{lcl}
-\multicolumn{3}{l}{%
- \mbox{For~every}\  q_0, q_3 \in Q,}\\[1mm] % if $\< q_1, p, q_2> \notin \Q$
-%
-d_\bot(q_0, q_3) & \opluseq &
-  d_\bot(q_0, q_2) \otimes
-  \bigoplus_{\Deltai} \weiei(q_2, q_3)\ 
-  \mathrm{if~} \< q_0, q_2> \notin \Q \\
-%
-d_\bot(q_0, q_3) & \opluseq &
-  {\bigoplus_{\Deltac}}
-  \bigl(
-  \weiec(q_0, p, q_1) \otimes
-   d_\top(q_1, p, q_2) \otimes
-  \bigoplus^2_{\Deltar} \weir(q_2, p, q_3)\bigr)\ 
-  \mathrm{if~} \< q_1, p, q_2> \notin \Q \\
-%
-d_\bot(q_0, q_3) & \opluseq &
-  d_\bot(q_0, q_2) \otimes
-  \bigoplus_{\Deltar} \weier(q_2, q_3)\ 
-  \mathrm{if~} \< q_0, q_2> \notin \Q \\ 
-%
-d_\bot(q_0, q_3) & \opluseq &
-  d_\bot(q_0, q_2) \otimes d_\bot(q_2, q_3)\  
-  \mathrm{if~} \< q_0, q_2>\notin \Q, \< q_2, q_3> \notin \Q \\[-1mm]
-\end{array}
-\)
-\caption{\textbf{update} $d_\bot$ (Algorithm~\ref{algo:Dijkstra}).}
-\label{fig:best-update-bottom}
-\end{figure}
-%
-
-\begin{figure}[ht]
-\(
-\begin{array}{lcl}
-\multicolumn{3}{l}{%
- \mbox{For~every}\  q_0, q_3 \in Q\ \mbox{and}\ p \in P,}\\[1mm] % if $\< q_1, p, q_2> \notin \Q$
-d_\top(q_0, p, q_3) & \opluseq &
-  d_\top(q_0, p, q_2) \otimes
-  \bigoplus_{\Deltai} \weii(q_2, p, q_3)\
-  \mathrm{if~} \< q_0, p, q_2> \notin \Q \\
-%\multicolumn{3}{l}{\quad
-%\mathrm{where~} \weii(q_2, p, q_3)_a \in \Phic
-%\mathrm{~is~the~partial~application~}
-%x_c \mapsto \weii(q_2, x_c, p, a, q_3)}\\
-d_\top(q_0, p, q_3) & \opluseq &
-  {\bigoplus^2_{\Deltac}}
-  \bigl[ \bigl( \weic(q_0, p, p', q_1) \otimes_2
-  d_\top(q_1, p', q_2) \bigr) \otimes_2
-  {\bigoplus_{\Deltar}} \weir(q_2, p', q_3) \bigr]\
-  \mathrm{if~} \< q_1, p', q_2> \notin \Q \\
-%\multicolumn{3}{l}{\quad
-%\mathrm{where~} \weic(q_0, p, p', q_1)_{c'} \in \Phic
-%\mathrm{~is~the~partial~application~}
-%x_c \mapsto \weic(q_0, p, x_c, c', p', q_1)}\\[2pt]
-%
-d_\top(q_0, p, q_3) & \opluseq &
-  d_\top(q_0, p, q_2) \otimes d_\top(q_2, p, q_3)\ 
-  \mathrm{if~} \< q_0, p, q_2> \notin \Q, \< q_2, p, q_3> \notin \Q \\[-1mm]
-\end{array}
-\)
-%$\weii(q_2, p, q_3)_a \in \Phic$, for $a \in \Deltai$,
-%is the partial application
-%$x_c \mapsto \weii(q_2, x_c, p, a, q_3)$.
-%
-%$\weic(q_0, p, p', q_1)_{c'} \in \Phic$, for $c' \in \Deltac$,
-%is the partial application
-%$x_c \mapsto \weic(q_0, p, x_c, c', p', q_1)$.
-%\caption{Update $d_\bot$ with $\<q_1, q_2>$ or $d_\top$ with $\< q_1, p, q_2>$.}
-\caption{\textbf{update} $d_\top$ (Algorithm~\ref{algo:Dijkstra}).}
-\label{fig:best-update-top}
-\end{figure}
-%
-%\noindent
-%** effectively computable by hypothesis that the label theory is effective**
-%\florent{REVOIR}
-Thanks to the hypothesis that $\bar\Phi$ is effective, 
-it is possible to construct during the iteration a witness word
-for Proposition~\ref{th:best-search}, 
-\ie a word $t \in \Delta^*$ with a minimal weight $A(t)$ \wrt~$\leq_\oplus$.
-
-%we require the additional property of convexity of weight functions.***
-%using the second requirement in Definition~\ref{def:effective}.
+%\input{algo-best}
+\input{algo-best-alt}
 
@@
 
-\section{Properties of Label Theory Operators} \label{sec:closure}
-\input{app-labels}
+%\section{Properties of Label Theory Operators} \label{sec:closure}
+%\input{app-labels}
 
@@
 
-\section{Proof of Proposition~\ref{prop:SWVPA-product}} \label{sec:product}
-\input{app-product}
+%\section{Proof of Proposition~\ref{prop:SWVPA-product}} \label{sec:product}
+%\input{app-product}
 
@@
 
-\section{Nested Words and Parse Trees} \label{sec:trees}\label{app:trees}
-\input{app-trees}
+%\section{Nested Words and Parse Trees} \label{sec:trees}\label{app:trees}
+%\input{app-trees}
 

### 2021-10-11 proof-best.tex
include proofs in text. rename files.

\textbf{Proof of correctness}

To prove that the algorithm does compute the word with the best weight, we use the following invariant :
$$
\<q_1, q_2> \notin Q \iff d_{\bot}(q_1, q_2) = b_{\bot}(q_1, q_2)
$$
In other words, a tuple $\<q_1, q_2>$ is removed from the queue during an iteration only when $d_{\bot}(q1, q2)$ has its minimal value.

Initialization : Before the while loop, no tuple has been extracted from the queue, so the invariant is trivially true.

Recursion :

Let's show by contradiction that at each iteration of the while loop, for the tuple $\<q, q'>$ extracted from Q we have:
$$
d_{\bot}(q, q') = b_{\bot}(q, q'), 
$$

Let's consider the states $q_0$, $q_1$, $q_2$ and $q_3$ $\in Q$ such as they form a path $q$ ~...~>$q_0$ -> $q_1$ -> $q_2$ -> $q_3$ ~..~> $q'$.

We take $\<q_0, q_1>$ as the first tuple to be removed from the queue. Then when $\<q_0, q_1>$ is extracted from the queue, the algorithms updates $d_{\bot}(q_0, q_2)$ and $d_{\bot}(q_0, q_3)$. Thus, we have:
$$
d_{\bot}(q_0, q_2) \geq b_{\bot}(q_0, q_1) \\
d_{\bot}(q_0, q_3) \geq b_{\bot}(q_0, q_1)
$$
Let's show that it is true for each update possible, shown in Figure 3. TODO

Let $\<q_0, q_3>$ be the first tuple for which the invariant is false:
$$
For \<q_0, q_3> \notin Q, d_{\bot} (q_0, q_3) \neq b_{\bot} (q_0, q_3)
$$
Since $d_{\bot}(q_0, q_2)$ has to be computed to compute $d_{\bot}(q_0, q3)$, and because the invariant is true for  $d_{\bot}(q_0, q_2)$ we can say:
$$
d_{\bot}(q_0, q_2) = b_{\bot}(q_0, q_2) \\
d_{\bot}(q_0, q_2) \leq b_{\bot}(q_0, q_3) \\
\\d_{\bot}(q_0, q_2) \leq d_{\bot}(q_0, q_3) \\
$$
 However, when $\<q_0, q_3>$ was extracted, $\<q_0, q_2>$ was still in $Q$. Thus, by definition 
$$
d_{\bot}(q_0, q_2) \geq d_{\bot}(q_0, q_3) \\
$$
Then the two inequalities are actually equal : 
$$
d_{\bot}(q_0, q_2) = b_{\bot}(q_0, q_2) = d_{\bot}(q_0, q_3) = b_{\bot}(q_0, q_3) \\
$$
We find that $d_{\bot}(q_0, q_3) = b_{\bot}(q_0, q_3)$, so by contradiction, we find that indeed $d_{\bot}(q_0, q_3) = b_{\bot}(q_0, q_3)$ iff 
$\<q_0, q_3> \notin Q$.

Termination: TODO

### 2021-10-11 introduction.tex
minor
alternative algo
include proofs in text. rename files.
remarques modifs intro

@@
 Parsing is the problem
+\florent{mv}
 of structuring a linear representation
@@
 
-In this paper, we present a framework for weighted parsing over infinite input alphabets.
+\florent{Both WA and SA generalized \cite{Herrmann16dlt,Herrmann20phd}...
+         we present results...}
+In this paper, we present 
+a framework for weighted parsing over infinite input alphabets.
 It is based on \emph{symbolic weighted} finite states language models (\swM),
@@
 \begin{tikzpicture}
-\node (NFA) at (0,4) {%
+\node (SWADS) at (0,5.0) {%
   \(
   \begin{array}{c}
-  \mathsf{FA} : \Sigma_{\mathsf{fin}}^* \to \mathbb{B}\\
-  q \xrightarrow{a} q' \quad  a \in \Sigma_\mathsf{fin}
+  \mathsf{SWADS}\mbox{\cite{Herrmann16dlt,Herrmann20phd}}: \Sigma_{\mathsf{inf}}^* \to \mathbb{S}
   \end{array}
@@
 %
-\node (WA) at (-2.5,2) {%
+\node (SWA) at (0,3.4) {%
   \(
   \begin{array}{c}
-  \mathsf{WA} : \Sigma_{\mathsf{fin}}^* \to \mathbb{S}\\
-  q \xrightarrow{a, w} q' \quad  a \in \Sigma_\mathsf{fin}, w \in \mathbb{S}
+  \mathsf{SWA}\mbox{\cite{Herrmann16dlt,Herrmann20phd}}: \Sigma_{\mathsf{inf}}^* \to \mathbb{S}\\
+  q \xrightarrow{\phi} q', \phi : \Sigma_\mathsf{inf} \to \mathbb{S}
   \end{array}
@@
 %
-\node (SA) at (2.5,2) {%
+\node (WA) at (-1.8,1.7) {%
   \(
   \begin{array}{c}
-  \mathsf{SA} : \Sigma_{\mathsf{inf}}^* \to \mathbb{B}\\
-  q \xrightarrow{\phi} q' \quad \phi : \Sigma_\mathsf{inf} \to \mathbb{B}
+  \mathsf{WA}\mbox{\cite{Droste09handbook}}: \Sigma_{\mathsf{fin}}^* \to \mathbb{S}\\
+  q \xrightarrow{a, w} q', a \in \Sigma_\mathsf{fin}, w \in \mathbb{S}
   \end{array}
@@
 %
-\node (SWA) at (0,0) {%
+\node (SA) at (1.8,1.7) {%8
   \(
   \begin{array}{c}
-  \mathsf{SWA} : \Sigma_{\mathsf{inf}}^* \to \mathbb{S}\\
-  q \xrightarrow{\phi} q' \quad \phi : \Sigma_\mathsf{inf} \to \mathbb{S}
+  \mathsf{SA}\mbox{\cite{dAntoni21CACM}}: \Sigma_{\mathsf{inf}}^* \to \mathbb{B}\\
+  q \xrightarrow{\phi} q', \phi : \Sigma_\mathsf{inf} \to \mathbb{B}
+  \end{array}
+  \)
+};
+%
+\node (NFA) at (0,0) {%
+  \(
+  \begin{array}{c}
+  \mathsf{FA} : \Sigma_{\mathsf{fin}}^* \to \mathbb{B}\\
+  q \xrightarrow{a} q', a \in \Sigma_\mathsf{fin}
   \end{array}
@@
 \draw[->] (SA)--(SWA);
+\draw[->] (SWA)--(SWADS);
 %\begin{array}{c} \mathsf{NFA} : \Sigma^* \to \mathbb{B} \end{array}
@@
         comparable to pebble automata? $\to$ conclusion}
-The main contributions of the paper are:
+The main contributions 
+\florent{main results = 3 Prop.}
+of the paper are:
 \reviews{1) not $i$! see response.txt}
@@
 
+\florent{Moreover, application parsing. believe results appliable to other decision pb.}
+
 \begin{example}%[Running example]
@@
 Such inputs typically correspond
-to the recording of a live performance, \eg
-$I = (69\tstamp{0.07}),
-     (71\tstamp{0.72}),
-     (73\tstamp{0.91}),
-     (74\tstamp{1.05}),
-     (76\tstamp{1.36}),
-     (77\tstamp{1.71})$. % over interval $[0,2[$,
+to the recording of a live performance, 
+\eg~$I = 69\tstamp{0.07},
+	     71\tstamp{0.72},
+    	 73\tstamp{0.91},
+	     74\tstamp{1.05},
+	     76\tstamp{1.36},
+	     77\tstamp{1.71}$. % over interval $[0,2[$,
 

### 2021-10-11 algo-best-alt.tex
cont. proof algo graph constr.

@@
 
-\subsection*{Alternative Algorithm for best-search}
+\subsection*{(Alternative Algorithm for best-search)}
 We associate to the \SWVPA $A$ a $\Semiring$-labeled graph
@@
 with set of vertices $V_A = (Q \times Q) \cup (Q \times P \times Q)$,
-set of edges $E_A = V_A \times V_A$, and with an edge labelling  
+set of edges $E_A = V_A \times V_A$
+(an edge $\< v_1, v_2>$, with $v_1, v_2 \in V_A$ is denoted by $v_1 \to v_2$),
+and with an edge labelling  
 $\eta_A : E_A \to \Semiring$ defined as follows,
@@
 %
-A \emph{path} of $\G_A$ is a sequence $e_1,\ldots, e_n \in V^*_A$ such that 
-for all $1 \leq i < n$, the second component of $e_i$ is the first component of $e_{i+1}$.
-We write $\src(\pi) = e_1$ and $\trg(\pi) = e_n$.
-The weight of the path is $\weight(\pi) = \bigotimes_{i=1}^{n} \eta(e_i)$.
+A \emph{path} of $\G_A$ is a sequence $v_0,\ldots, v_n \in V^*_A$,
+%for all $0 \leq i < n$. %the second component of $e_i$ is the first component of $e_{i+1}$.
+%We write $\src(\pi) = e_1$ and $\trg(\pi) = e_n$.
+it is assigned a weight value $\weight(\pi) = \bigotimes_{i=1}^{n} \eta(v_{i-1} \to v_i)$.
 
@@
 \begin{proof}
-Since $\Semiring$ is  total, there exists $t \in \Delta^*$ such that  
+By~\eqref{eq:bbot}, 
+and since $\Semiring$ is total, for all $q_1, q_2 \in Q$, 
+there exists one or several $t \in \Delta^*$ such that  
 $b_\bot(q_1, q_2) = \weight_A\bigl(\config{q_1}{\bot}, t, \config{q_2}{\bot}\bigr)$.
-We represent a computation of $A$ on $t$ by a sequence of transitions...
+Let $T(q_1, q_2) \subseteq \Delta^*$ denotes the set of such words $t$. % of minimal length.
+%
+Moreover, following~\eqref{eq:SWVPA-weight}, and by distributivity and commutativity in $\Semiring$, 
+$b_\bot(q_1, q_2)$ is equal to a finite product $\bigotimes_{i=0}^n w_i(\tau_i)$, 
+where every $\tau_i$ is a transition of $A$, every $w_i$ is one of the functions of $\bar\wei$,
+the sequence of input symbols in $\tau_0,\ldots, \tau_n$ belongs to $T(q_1, q_2)$, 
+$\src(\tau_0) = q_1$, and $\snd(\tau_n) = q_2$, where 
+$\src(\tau)$ (resp. $\snd(\tau)$) is the first (resp. second) state component of a transition~$\tau$.
+There may be several such sequences of $\tau_i$ and $w_i$, we choose an arbitrarily one of minimal length $n$.
 
-We show that we can construct a path $\pi$ from $t$.
-...\qed
+%We represent a computation of $A$ on $t$ by a sequence $\rho$ of its transitions,
+
+We show now by induction on $n$ that we can construct a path $\pi$ of weight $b_\bot(q_1, q_2)$
+from $\tau_0,\ldots, \tau_n$ and $w_0,\ldots, w_n$. ...\qed
 \end{proof}

### 2021-10-11 main.tex
cont. proof algo graph constr.

@@
 %\newcommand{\creturn}[1]{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
-\newcommand{\src}{\mathsf{src}}
-\newcommand{\trg}{\mathsf{trg}}
+\newcommand{\src}{\mathit{src}}
+\newcommand{\trg}{\mathit{trg}}
+\newcommand{\fst}{\mathit{fst}}
+\newcommand{\snd}{\mathit{snd}}
 
@@
 %
-We propose a Dijkstra-like algorithm computing,
+We propose a method for %Dijkstra-like algorithm 
+computing,
 for every $q, q' \in Q$,
@@
 Let us denote by $b_\bot(q, q')$ this minimum.
+\florent{it exists by effectiveness of label th. ?}
+%
 By definition of $\leq_\oplus$, and since $\Semiring$ is total,

### 2021-10-11 main.tex
parentheses in exemples VPA (linearization trees)

@@
 \newcommand{\return}[1]{\ensuremath #1} %{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
-\newcommand{\ccall}[1]{\ensuremath \langle_{#1}}
-\newcommand{\creturn}[1]{\ensuremath \rangle_{#1}}
+\newcommand{\ccall}[1]{\ensuremath \lceil_{#1}}   % \langle
+\newcommand{\creturn}[1]{\ensuremath \rceil_{#1}}
 %\newcommand{\creturn}[1]{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
@@
 Let's now consider a new call transition starting in state $q_{\frac{j}{n}}$, that is the $j$-th state of the $n$ sub-intervals of duration~$\frac{\ell}{n}$.
-We thus read a new time-division symbol ~$\ccall{d}$ and compute the weight of the transition from $q_{\frac{j}{n}}$ to $q_{\frac{1}{d}}$ (the first state after the new ~$\ccall{d}$), reading the top of the stack $<\ccall{n}\tstamp{[\tau, \tau+\ell]}, q>$, and pushing $<\ccall{d}\tstamp{\iota}, q_{\frac{j+1}{c}}>$ on top, or:
+We thus read a new time-division symbol ~$\ccall{d}$ 
+and compute the weight of the transition from $q_{\frac{j}{n}}$ to $q_{\frac{1}{d}}$, 
+the first state after the new~$\ccall{d}$, reading on the top of the stack the pair
+$\<\ccall{n}\tstamp{[\tau, \tau+\ell]}, q>$, 
+and pushing $\<\ccall{d}\tstamp{\iota}, q_{\frac{j+1}{c}}>$ on top, or:
 \(
@@
 \noindent
-We described earlier the special cases where the stack is empty, which applies for example when reading the first measure for which we only push $<\ccall{\mathsf{m}}\tstamp{[0, 1]}, q_{\frac{1}{1}}>$ in the stack:
+We described earlier the special cases where the stack is empty, which applies for example when reading the first measure 
+for which we only push $\<\ccall{\mathsf{m}}\tstamp{[0, 1]}, q_{\frac{1}{1}}>$ in the stack:
 \(

### 2021-10-11 algo-best-alt.tex
parentheses in exemples VPA (linearization trees)

@@
 \subsection*{(Alternative Algorithm for best-search)}
-We associate to the \SWVPA $A$ a $\Semiring$-labeled graph
-$\G_A = \< V_A, E_A, \eta_A >$
+We shall compute the values of the functions $b_\bot$ and $b_\top$
+by searching shortests paths in a $\Semiring$-labeled graph $\G_A$
+associated to the \SWVPA $A$. 
+This graph is defined by $\G_A = \< V_A, E_A, \eta_A >$, 
 with set of vertices $V_A = (Q \times Q) \cup (Q \times P \times Q)$,
@@
 
-We show now by induction on $n$ that we can construct a path $\pi$ of weight $b_\bot(q_1, q_2)$
-from $\tau_0,\ldots, \tau_n$ and $w_0,\ldots, w_n$. ...\qed
+We show now the existence of a path $\pi$ as expected, by induction on $n$.
+%that we can construct a path $\pi$ of weight $b_\bot(q_1, q_2)$
+%from $\tau_0,\ldots, \tau_n$ and $w_0,\ldots, w_n$. ...\qed
 \end{proof}

### 2021-10-11 main.tex
parentheses in exemples VPA (linearization trees)

@@
 \begin{proof}
-We do a classical product construction. % see Appendix~\ref{sec:product} for details.
+%We do a classical product construction, see Appendix~\ref{sec:product} for details.
 %similar to the case of the Boolean semiring~\cite{dAntonyAlur14SVPDA}.
@@
 \reviews{1) $\Semiring$ total $\to$ $\Semiring$ extremal}
-it holds that:
+it holds that
+(the infinite sum in~\eqref{eq:bbot} is well defined since $\Semiring$ is complete):
 %
@@
 \end{equation}
-The infinite sum in~\eqref{eq:bbot} is well defined since $\Semiring$ is complete.
 %

### 2021-10-11 algo-best-alt.tex
proof algo cont.

@@
 \begin{lemma}[Completeness]
-For all $q_1, q_2 \in Q$, $p\in P$, 
-($i$) there exists a path $\pi$ of $\G_A$ 
+For all $q, q' \in Q$, $p\in P$, 
+($i$) there exists a path~$\pi$ of~$\G_A$ 
 such that $\src(\pi) = \<q_0, q_0>$ for some $q_0 \in Q$, 
-$\trg(\pi) = \< q_1, q_2>$,
-and $\weight(\pi) = b_\bot(q_1, q_2)$, 
+$\trg(\pi) = \< q, q'>$,
+and $\weight(\pi) = b_\bot(q, q')$, 
 ($ii$) there exists a path $\pi'$ of $\G_A$ 
@@
 $\trg(\pi) = \< q_1, p, q_2>$,
-and $\weight(\pi) = b_\top(q_1, q_2)$.
+and $\weight(\pi) = \bigoplus_\Deltac b_\top(q_1, p, q_2)$.
 \end{lemma}
@@
 \begin{proof}
-By~\eqref{eq:bbot}, 
-and since $\Semiring$ is total, for all $q_1, q_2 \in Q$, 
-there exists one or several $t \in \Delta^*$ such that  
-$b_\bot(q_1, q_2) = \weight_A\bigl(\config{q_1}{\bot}, t, \config{q_2}{\bot}\bigr)$.
-Let $T(q_1, q_2) \subseteq \Delta^*$ denotes the set of such words $t$. % of minimal length.
+By associativity, commutativity and distributivity for $\Semiring$,
+\eqref{eq:bbot} can be rewritten into the form, unfolding~\eqref{eq:SWVPA-weight}:
+\[
+b_\bot(q, q') = 
+\bigoplus_{t\in \Delta^*} 
+\bigoplus_{\begin{array}{c}
+           \scriptstyle q_0,\ldots, q_n \in Q\\
+           \scriptstyle p_{i},\ldots, p_{k} \in P
+           \end{array}}
+\bigotimes_{i=i}^n w_i(\tau_i)           
+\]
+where $n$ is the length of $t$, $k \leq n$,
+$q_0 = q$, $q_n = q'$, 
+for all $1 \leq i < n$, 
+$w_i$ is one of the functions of $\bar\wei$,
+$\tau_i$ is a transition of $A$ and
+$\src(\tau_i) = q_{i-1}$, $\snd(\tau_i) = q_i$.
 %
-Moreover, following~\eqref{eq:SWVPA-weight}, and by distributivity and commutativity in $\Semiring$, 
-$b_\bot(q_1, q_2)$ is equal to a finite product $\bigotimes_{i=0}^n w_i(\tau_i)$, 
-where every $\tau_i$ is a transition of $A$, every $w_i$ is one of the functions of $\bar\wei$,
-the sequence of input symbols in $\tau_0,\ldots, \tau_n$ belongs to $T(q_1, q_2)$, 
-$\src(\tau_0) = q_1$, and $\snd(\tau_n) = q_2$, where 
-$\src(\tau)$ (resp. $\snd(\tau)$) is the first (resp. second) state component of a transition~$\tau$.
-There may be several such sequences of $\tau_i$ and $w_i$, we choose an arbitrarily one of minimal length $n$.
+Since $\Semiring$ is total, there exists finite sequences as above such that
+\( b_\bot(q, q') = \bigotimes_{i=i}^n w_i(\tau_i)\).
+%
+There might exists, for $q$ and $q'$,
+several finite sequences~$\bar\tau$ and~$\bar{w}$ as above, 
+let us choose arbitrarily one of minimal length~$n$. 
+This integer $n$ will be denoted $n_{q, q'}$ in the following.
+%We represent a computation of $A$ on $t$ by a sequence $\rho$ of its transitions,
 
-%We represent a computation of $A$ on $t$ by a sequence $\rho$ of its transitions,
+Similarly, following~\eqref{eq:btop} and~\eqref{eq:SWVPA-weight}, 
+$\bigoplus_\Deltac b_\top(q_1, q_2)$ can be put in the form:
+\[
+{\textstyle\bigoplus_\Deltac} b_\top(q_1, q_2) = 
+{\textstyle\bigoplus_\Deltac} \bigoplus_{t\in \Delta^*} 
+\bigoplus_{\begin{array}{c}
+           \scriptstyle q_0,\ldots, q_n \in Q\\
+           \scriptstyle p_{i},\ldots, p_{k} \in P
+           \end{array}}
+\bigotimes_{i=i}^n w_i(\tau_i)           
+\]
+and hence 
+\( {\textstyle\bigoplus_\Deltac} b_\top(q_1, q_2) = \bigotimes_{i=i}^n w_i(\tau_i) \),
+using $c \in \Deltac$ that minimize...
+We denote the smallest $n$ as above by $n_{q, p, q'}$.
 

### 2021-10-12 algo-best-alt.tex
proof algo cont.

@@
 \begin{lemma}[Correctness]
-For all $q_0, q_1, q_2 \in Q$, $p\in P$, 
+For all $q_0, q, q' \in Q$, $p\in P$, 
 and all path $\pi$ of $\G_A$ with $\src(\pi) = \<q_0, q_0>$,
-($i$) if $\trg(\pi) = \< q_1, q_2>$, 
+($i$) if $\trg(\pi) = \< q, q'>$, 
 then there exists $u \in \Delta^*$ such that 
-$\weight_A(q_1[\bot], u, q_1[\bot]) = \weight(\pi)$,
+$\weight_A(q[\bot], u, q'[\bot]) = \weight(\pi)$,
 %
-($ii$) if $\trg(\pi) = \< q_1, p, q_2>$,
+($ii$) if $\trg(\pi) = \< q, p, q'>$,
 then there exists $v \in \Delta^*$ such that 
 ${\displaystyle\bigoplus_{c \in \Deltac}}
- \weight_A\bigl(\configup{q_1}{\< c, p> \stackup \top }, v, \configup{q_2}{\<c, p> \stackup \top}\bigr)
+ \weight_A\bigl(\configup{q}{\< c, p> \stackup \top }, v, \configup{q'}{\<c, p> \stackup \top}\bigr)
  = \weight(\pi)$.
@@
 
-\begin{lemma}[Completeness]
+\begin{lemma}[Completeness] \label{lem:algo-completeness}
 For all $q, q' \in Q$, $p\in P$, 
@@
 \eqref{eq:bbot} can be rewritten into the form, unfolding~\eqref{eq:SWVPA-weight}:
-\[
+%
+\begin{equation}\label{eq:bbot-unfold}
 b_\bot(q, q') = 
@@
            \end{array}}
-\bigotimes_{i=i}^n w_i(\tau_i)           
-\]
+\bigotimes_{i=1}^n w_i(\tau_i)           
+\end{equation}
+%
 where $n$ is the length of $t$, $k \leq n$,
@@
 Similarly, following~\eqref{eq:btop} and~\eqref{eq:SWVPA-weight}, 
-$\bigoplus_\Deltac b_\top(q_1, q_2)$ can be put in the form:
-\[
-{\textstyle\bigoplus_\Deltac} b_\top(q_1, q_2) = 
+$\bigoplus_\Deltac b_\top(q, p, q')$ can be put in the form:
+%
+\begin{equation}\label{eq:btop-unfold}
+{\textstyle\bigoplus_\Deltac} b_\top(q, p, q') = 
 {\textstyle\bigoplus_\Deltac} \bigoplus_{t\in \Delta^*} 
@@
            \end{array}}
-\bigotimes_{i=i}^n w_i(\tau_i)           
-\]
+\bigotimes_{i=1}^n w_i(\tau_i)           
+\end{equation}
+%
 and hence 
-\( {\textstyle\bigoplus_\Deltac} b_\top(q_1, q_2) = \bigotimes_{i=i}^n w_i(\tau_i) \),
-using $c \in \Deltac$ that minimize...
+\( {\textstyle\bigoplus_\Deltac} b_\top(q, p, q') = \bigotimes_{i=i}^n w_i(\tau_i) \),
+using $c \in \Deltac$ that minimize the function in rhs of~\eqref{eq:btop-unfold}.
 We denote the smallest $n$ as above by $n_{q, p, q'}$.
 
-We show now the existence of a path $\pi$ as expected, by induction on $n$.
-%that we can construct a path $\pi$ of weight $b_\bot(q_1, q_2)$
-%from $\tau_0,\ldots, \tau_n$ and $w_0,\ldots, w_n$. ...\qed
+\noindent
+We show now the existence of a path $\pi$ of Lemma~\ref{lem:algo-completeness}, 
+by simultaneous induction on $n_{q, q'}$ and $n_{q, p, q'}$.
+
+The base case, $n = 0$ corresponds to $t = \varepsilon$. 
+In this case, by~\eqref{eq:SWVPA-weight}, $b_\top(q, p, q') = \zero$,
+$b_\bot(q, q) = \one$ and $b_\bot(q, q') = \zero$ if $q \neq q'$.
+
+\noindent
+For $n = n_{q, q'} > 0$, let $\bar\tau$ and~$\bar{w}$
+be the sequences associated to $q$ and $q'$ as above.
+Let us do a case analysis of $w_n$.
+
 \end{proof}

### 2021-10-12 introduction.tex
intro: remarque org.

@@
 Parsing is the problem
-\florent{mv}
+\florent{deplacer plus loin (application)}
 of structuring a linear representation
@@
 
-The latter case is related to a motivation  of  the present work:
+The latter case is related to a motivation  of  the present work: \florent{idem}
 automated music transcription. Representations that capture  music performances
@@
 help in dealing with non-determinism by computing weight values.
+\florent{quantitative reasoning about programs}
 With an ambiguous grammar, 
@@
 
-\florent{Both WA and SA generalized \cite{Herrmann16dlt,Herrmann20phd}...
+\florent{Both WA and SA generalized by \cite{Herrmann16dlt,Herrmann20phd}...
          we present results...}
@@
 
-\florent{Moreover, application parsing. believe results appliable to other decision pb.}
+\florent{Application to parsing and tanscription. these results are appliable to other decision pb.}
 

### 2021-10-12 algo-best-alt.tex
algo def. weight path

@@
 %
-A \emph{path} of $\G_A$ is a sequence $v_0,\ldots, v_n \in V^*_A$,
-%for all $0 \leq i < n$. %the second component of $e_i$ is the first component of $e_{i+1}$.
-%We write $\src(\pi) = e_1$ and $\trg(\pi) = e_n$.
-it is assigned a weight value $\weight(\pi) = \bigotimes_{i=1}^{n} \eta(v_{i-1} \to v_i)$.
+A \emph{path} of $\G_A$ is a sequence $\pi = v_0,\ldots, v_n \in V^*_A$,
+it is assigned a weight value $\weight(\pi)$ which is the product by $\otimes$ of the weights of the 
+edges involved in the path.
+More precisely, for all $q, q' \in Q$ and $p \in P$, we let 
+$\weight(\< q, q' > ) = \one$ if $q = q'$, and $\zero$ otherwise, 
+$\weight(\< q, p, q' > ) = \zero$ and 
+for all $n \geq 1$, 
+$\weight(v_0,\ldots, v_n) = \weight(v_0,\ldots, v_{n-1}) \otimes \eta_A(v_{n-1} \to v_{n})$.
+%$\weight(\pi) = \bigotimes_{i=1}^{n} \eta(v_{i-1} \to v_i)$.
 

### 2021-10-12 algo-best-alt.tex
revision def. graph

@@
 \begin{array}{rclcl}
+%
 \< q_0, q_1> & \to & \< q_0, q_2> & : &
-\bigoplus_{\Deltai} \weiei(q_1, q_2)\\
+\bigoplus_{\Deltai} \weiei(q_1, q_2) \oplus \bigoplus_{\Deltar} \weier(q_1, q_2)\\
 %
 \< q_1, p, q_2> & \to & \<q_0, q_3> & : &
-\bigoplus_{\Deltac} \bigl( \weiec(q_0, p, q_1) \otimes \bigoplus^2_{\Deltar} \weir(q_2, p, q_3)\bigr)\\ 
+\bigoplus_{\Deltac} \bigl[ \weiec(q_0, p, q_1) \otimes \bigoplus^2_{\Deltar} \weir(q_2, p, q_3)\bigr]\\ 
 %
-\< q_0, q_2> & \to & \< q_0, q_3> & : & 
-\bigoplus_{\Deltar} \weier(q_2, q_3)\\
-% 
-\< q_0, p, q_1> & \to & \< q_0, p, q_2> & : &
-\bigoplus_{\Deltai} \weii(q_1, p, q_2)\\[1mm]
+%\< q_0, p, q_1> & \to & \< q_0, p, q_2> & : &
+%\bigoplus_{\Deltai} \weii(q_1, p, q_2)\\
 %
 \< q_1, p', q_2> & \to & \< q_0, p, q_3> & : &
-\bigoplus^2_{\Deltac} \bigl[ \weic(q_0, p, p', q_1) \otimes_2 {\bigoplus_{\Deltar}} \weir(q_2, p', q_3) \bigr]\\
+{\displaystyle\bigoplus_{q_0=q_1}}
+{\displaystyle\bigoplus_{p=p'}}
+ \bigoplus_{\Deltai} \weii(q_2, p, q_3)\\
+ & & & & 
+\qquad\quad 
+\oplus \bigoplus^2_{\Deltac} \bigl[ \weic(q_0, p, p', q_1) \otimes_2 {\bigoplus_{\Deltar}} \weir(q_2, p', q_3) \bigr]\\
+%
 \end{array}
@@
 A \emph{path} of $\G_A$ is a sequence $\pi = v_0,\ldots, v_n \in V^*_A$,
-it is assigned a weight value $\weight(\pi)$ which is the product by $\otimes$ of the weights of the 
-edges involved in the path.
+it is called \emph{safe} when $v_0$ has the form $\< q, q>$ for some $q \in Q$.
+A path $\pi$ is assigned a weight value $\weight(\pi)$ which is the product by $\otimes$ 
+of the weights of the edges involved in the path.
 More precisely, for all $q, q' \in Q$ and $p \in P$, we let 
@@
 \begin{lemma}[Correctness]
-For all $q_0, q, q' \in Q$, $p\in P$, 
-and all path $\pi$ of $\G_A$ with $\src(\pi) = \<q_0, q_0>$,
-($i$) if $\trg(\pi) = \< q, q'>$, 
+For all safe path~$\pi$ of~$\G_A$ %with $\src(\pi) = \<q_0, q_0>$,
+with last element $v$, \\
+($i$) if $v = \< q, q'>$, with $q, q' \in Q$, 
 then there exists $u \in \Delta^*$ such that 
@@
 %
-($ii$) if $\trg(\pi) = \< q, p, q'>$,
+($ii$) if $v = \< q, p, q'>$, with $q, q' \in Q$, $p\in P$, 
 then there exists $v \in \Delta^*$ such that 
@@
 \begin{proof}
-We prove ($i$) and ($ii$) simultaneously by induction on the length of the path $\pi$.
-...\qed
+We prove ($i$) and ($ii$) simultaneously by induction on the length $|\pi|$ of the path $\pi$.
+
+\noindent
+The base case $|pi| = 1$ is a direct consequence of the definition of weight of paths of length 1
+and the first line of \eqref{eq:SWVPA-weight}.
+
+\noindent
+If $\pi = v_0,\ldots, v_n$ with $n \geq 1$, 
+let us do a case analysis on the edge $v_{n-1} \to v_n$.
+...
+\qed
 \end{proof}

### 2021-10-12 algo-best-alt.tex
revision lemmes algos et debut revision def. weight Sw-VPA (inversion)

@@
 A \emph{path} of $\G_A$ is a sequence $\pi = v_0,\ldots, v_n \in V^*_A$,
-it is called \emph{safe} when $v_0$ has the form $\< q, q>$ for some $q \in Q$.
+it is called \emph{safe} when $v_0$ has the form $\< q, q>$ for some $q \in Q$, 
+and $n_v$ is called $\last(\pi)$.
+%
 A path $\pi$ is assigned a weight value $\weight(\pi)$ which is the product by $\otimes$ 
@@
 \begin{lemma}[Correctness]
-For all safe path~$\pi$ of~$\G_A$ %with $\src(\pi) = \<q_0, q_0>$,
-with last element $v$, \\
-($i$) if $v = \< q, q'>$, with $q, q' \in Q$, 
+For all safe path~$\pi$ of~$\G_A$,  %with $\src(\pi) = \<q_0, q_0>$,
+%with last element $v$, \\
+($i$) if $\last(\pi) = \< q, q'>$, %with $q, q' \in Q$, 
 then there exists $u \in \Delta^*$ such that 
-$\weight_A(q[\bot], u, q'[\bot]) = \weight(\pi)$,
+$\weight_A(q[\bot], u, q'[\bot]) = \weight(\pi)$,\\
 %
-($ii$) if $v = \< q, p, q'>$, with $q, q' \in Q$, $p\in P$, 
-then there exists $v \in \Delta^*$ such that 
+($ii$) if $\last(\pi) = \< q, p, q'>$, %with $q, q' \in Q$, $p\in P$, 
+then there exists $v \in \Delta^*$ such that \\
 ${\displaystyle\bigoplus_{c \in \Deltac}}
@@
 For all $q, q' \in Q$, $p\in P$, 
-($i$) there exists a path~$\pi$ of~$\G_A$ 
-such that $\src(\pi) = \<q_0, q_0>$ for some $q_0 \in Q$, 
-$\trg(\pi) = \< q, q'>$,
-and $\weight(\pi) = b_\bot(q, q')$, 
-($ii$) there exists a path $\pi'$ of $\G_A$ 
-such that $\src(\pi) = \<q_0, q_0>$ for some $q_0 \in Q$, 
-$\trg(\pi) = \< q_1, p, q_2>$,
-and $\weight(\pi) = \bigoplus_\Deltac b_\top(q_1, p, q_2)$.
+($i$) there exists a safe path~$\pi$ of~$\G_A$ such that
+%such that $\src(\pi) = \<q_0, q_0>$ for some $q_0 \in Q$, 
+$\last(\pi) = \< q, q'>$,
+and $\weight(\pi) = b_\bot(q, q')$, \\
+($ii$) there exists a safe path $\pi'$ of $\G_A$ 
+such that $\last(\pi') = \< q, p, q'>$,
+and $\weight(\pi') = \bigoplus_\Deltac b_\top(q, p, q')$.
 \end{lemma}

### 2021-10-12 algo-best-alt.tex
inversion dans les 3 def. weight

@@
 \noindent
-The base case $|pi| = 1$ is a direct consequence of the definition of weight of paths of length 1
+The base case $|\pi| = 1$ is a direct consequence of the definition of weight of paths of length 1
 and the first line of \eqref{eq:SWVPA-weight}.
@@
 $\last(\pi) = \< q, q'>$,
-and $\weight(\pi) = b_\bot(q, q')$, \\
+and $\weight(\pi) = b_\bot(q, q')$, 
 ($ii$) there exists a safe path $\pi'$ of $\G_A$ 

### 2021-10-12 introduction.tex
figure intro

@@
   \begin{array}{c}
-  \mathsf{SWADS}\mbox{\cite{Herrmann16dlt,Herrmann20phd}}: \Sigma_{\mathsf{inf}}^* \to \mathbb{S}
+  \mathsf{SWADS}\,\mbox{\cite{Herrmann16dlt,Herrmann20phd}}: \Sigma_{\mathsf{inf}}^* \to \mathbb{S}
   \end{array}
@@
   \begin{array}{c}
-  \mathsf{SWA}\mbox{\cite{Herrmann16dlt,Herrmann20phd}}: \Sigma_{\mathsf{inf}}^* \to \mathbb{S}\\
+  \mathsf{SWA}: \Sigma_{\mathsf{inf}}^* \to \mathbb{S}\\
   q \xrightarrow{\phi} q', \phi : \Sigma_\mathsf{inf} \to \mathbb{S}
@@
   \begin{array}{c}
-  \mathsf{WA}\mbox{\cite{Droste09handbook}}: \Sigma_{\mathsf{fin}}^* \to \mathbb{S}\\
+  \mathsf{WA}\,\mbox{\cite{Droste09handbook}}: \Sigma_{\mathsf{fin}}^* \to \mathbb{S}\\
   q \xrightarrow{a, w} q', a \in \Sigma_\mathsf{fin}, w \in \mathbb{S}
@@
   \begin{array}{c}
-  \mathsf{SA}\mbox{\cite{dAntoni21CACM}}: \Sigma_{\mathsf{inf}}^* \to \mathbb{B}\\
+  \mathsf{SA}\,\mbox{\cite{dAntoni21CACM}}: \Sigma_{\mathsf{inf}}^* \to \mathbb{B}\\
   q \xrightarrow{\phi} q', \phi : \Sigma_\mathsf{inf} \to \mathbb{B}
@@
 };
+%
+\node (VPA) at (5.6,0) {%
+  \(
+  \mathsf{VPA}\,\cite{AlurMadhusudan09nested}
+  \)
+};
+%
+\node (WVPA) at (4.6,1.7) {%
+  \(
+  \mathsf{WVPA}\,\cite{Caralp12VPAmult}
+  \)
+};
+%
+\node (SVPA) at (6.6,1.7) {%
+  \(
+  \mathsf{SVPA}\,\cite{dAntonyAlur14SVPDA}
+  \)
+};
+%
+\node (SWVPA) at (5.6,3.4) {%
+  \(
+  \mathsf{SWVPA}
+  \)
+};
+%
 \draw[->] (NFA)--(WA);
@@
 \draw[->] (SWA)--(SWADS);
+\draw[->] (VPA)--(WVPA);
+\draw[->] (VPA)--(SVPA);
+\draw[->] (WVPA)--(SWVPA);
+\draw[->] (SVPA)--(SWVPA);
+\draw[->] (SWVPA)--(SWADS);
 %\begin{array}{c} \mathsf{NFA} : \Sigma^* \to \mathbb{B} \end{array}

### 2021-10-13 algo-best-alt.tex
1 passe label theories

@@
 
-\subsection*{(Alternative Algorithm for best-search)}
+\florent{New Algorithm for best-search}
 We shall compute the values of the functions $b_\bot$ and $b_\top$
-by searching shortests paths in a $\Semiring$-labeled graph $\G_A$
+by search of shortest paths in a $\Semiring$-labeled graph $\G_A$
 associated to the \SWVPA $A$. 
@@
 and with an edge labelling  
-$\eta_A : E_A \to \Semiring$ defined as follows,
+$\eta_A : E_A \to \Semiring$ defined by, %as follows,
 for $q_0, q_1, q_2, q_3 \in Q$, $p, p' \in P$:
@@
 %
-\< q_0, q_1> & \to & \< q_0, q_2> & : &
+\< q_0, q_1> & \to & \< q_0, q_2> & \mapsto &
 \bigoplus_{\Deltai} \weiei(q_1, q_2) \oplus \bigoplus_{\Deltar} \weier(q_1, q_2)\\
 %
-\< q_1, p, q_2> & \to & \<q_0, q_3> & : &
+\< q_1, p, q_2> & \to & \<q_0, q_3> & \mapsto &
 \bigoplus_{\Deltac} \bigl[ \weiec(q_0, p, q_1) \otimes \bigoplus^2_{\Deltar} \weir(q_2, p, q_3)\bigr]\\ 
@@
 %
-\< q_1, p', q_2> & \to & \< q_0, p, q_3> & : &
+\< q_1, p', q_2> & \to & \< q_0, p, q_3> & \mapsto &
 {\displaystyle\bigoplus_{q_0=q_1}}
@@
 \]
-Removing from $\G_A$ every edge $e \in E_A$ such that $\eta(e) = \zero$ 
-is preserving the results below.
+%Removing from $\G_A$ every edge $e \in E_A$ such that $\eta(e) = \zero$ 
+%is preserving the results below.
 %
@@
 %
-A path $\pi$ is assigned a weight value $\weight(\pi)$ which is the product by $\otimes$ 
+A path $\pi$ is assigned a weight value $\weight(\pi) \in \Semiring$ which is the product by $\otimes$ 
 of the weights of the edges involved in the path.

### 2021-10-13 main.tex
proof algo cont.
revision lemmes algos et debut revision def. weight Sw-VPA (inversion)
inversion dans les 3 def. weight
1 passe label theories

@@
 \newcommand{\snd}{\mathit{snd}}
+\newcommand{\last}{\mathit{last}}
 
@@
                                                    \scriptstyle q'' \in Q\\[-1ex]
-                                                   \scriptstyle s = au,\, a \in \Sigma
+                                                   \scriptstyle s = ua,\, a \in \Sigma
                                                    \end{array}}
-    \wei_{10}(q, a, \varepsilon, q'') \otimes \weight_T(q'', u, t, q')\nonumber\\
+    \weight_T(q, u, t, q'') \otimes \wei_{10}(q'', a, \varepsilon, q')\nonumber\\
                     & \oplus \displaystyle\bigoplus_{\begin{array}{c}
                                                      \scriptstyle q'' \in Q\\[-1ex]
-                                                     \scriptstyle t = bv,\, b \in \Delta\\
+                                                     \scriptstyle t = vb,\, b \in \Delta\\
                                                      \end{array}}
-    \wei_{01}(q, \varepsilon, b, q'') \otimes \weight_T(q'', s, v, q')\nonumber\\
+    \weight_T(q, s, v, q'') \otimes \wei_{01}(q'', \varepsilon, b, q')\nonumber\\
                     & \oplus \displaystyle\bigoplus_{\begin{array}{c}
                                                      \scriptstyle q'' \in Q\\[-1ex]
-                                                     \scriptstyle s = au,\, t = bv\\
+                                                     \scriptstyle s = ua,\, t = vb\\
                                                      \end{array}}
-    \wei_{11}(q, a, b, q'') \otimes \weight_T(q'', u, v, q')\nonumber
+    \weight_T(q, u, v, q'') \otimes \wei_{11}(q'', a, b, q')\nonumber
 \end{align}
@@
 %\weight_A(q, \varepsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
-\weight_A(q, au, q') & =  \displaystyle\bigoplus_{q'' \in Q}
-    \wei_{1}(q, a, q'') \otimes \weight_A(q'', u, q')\nonumber
+\weight_A(q, ua, q') & =  \displaystyle\bigoplus_{q'' \in Q}
+\weight_A(q, u, q'') \otimes  \wei_{1}(q'', a, q')\nonumber
 \label{eq:SWA-weight}
@@
 are also called \emph{transitions} of the \SWVPA.
+We denote by $\src(\tau)$ (respectively $\snd(\tau)$) the first (respectively second) 
+state component of a transition~$\tau$.
 %
@@
 %
-\weight_A\bigl(\config{q}{\bot}, \varepsilon, \config{q'}{\bot}) & = \one
-\mathrm{~if~} q = q' \mathrm{~and~} \zero \mathrm{~otherwise}\label{eq:SWVPA-weight}\\
-\weight_A\bigl(\configup{q}{\<{\call{c}}, p>\stackup\gamma}, a\, u,
-               \config{q'}{\gamma'}\bigr) & =
- {\displaystyle\bigoplus_{q'' \in Q}} \weii(q, c, p, a, q'')
-  \otimes \weight_A\bigl(\configup{q''}{\<{\call{c}}, p> \stackup \gamma}, u,
-                         \config{q'}{\gamma'}\bigr)\nonumber\\
-%
-\weight_A\bigl(\config{q}{\bot}, a\, u,
-               \config{q'}{\gamma'}\bigr) & =
-  {\displaystyle\bigoplus_{q'' \in Q}} \weiei(q, a, q'')
-   \otimes \weight_A\bigl(\config{q''}{\bot}, u, \config{q'}{\gamma'}\bigr)\nonumber\\
-%
-\weight_A\bigl(\configup{q}{\<{\call{c}}, p>\stackup\gamma}, {\call{c}'}\, u,
-               \config{q'}{\gamma'}\bigr) & =
+\weight_A\bigl(\config{q}{\gamma}, \varepsilon, \config{q'}{\gamma'}) & = \one
+\mathrm{~if~} q = q', \gamma =\gamma' = \bot
+\mathrm{~and~} \zero \mathrm{~otherwise}\label{eq:SWVPA-weight}\\
+%
+\weight_A\bigl(\config{q}{\gamma}, u\, a,
+               \configup{q'}{\<{\call{c}}, p>\stackup\gamma'}\bigr) & =
+ {\displaystyle\bigoplus_{q'' \in Q}} 
+  \weight_A\bigl(\config{q}{\gamma}, u,
+                 \configup{q''}{\<{\call{c}}, p> \stackup \gamma'}\bigr) 
+  \otimes \weii(q'', c, p, a, q') \nonumber\\
+%
+\weight_A\bigl(\config{q}{\gamma}, u\, a,
+               \config{q'}{\bot}\bigr) & =
+  {\displaystyle\bigoplus_{q'' \in Q}} 
+   \weight_A\bigl(\config{q}{\gamma}, u, \config{q''}{\bot}\bigr)
+   \otimes\weiei(q'', a, q')\nonumber\\
+%
+\weight_A\bigl(\config{q}{\gamma}, u\,{\call{c}'},
+               \configup{q'}{\<{\call{c}'}, p'>\stackup \<{\call{c}}, p>\stackup \gamma'}\bigr) & =
  {\displaystyle\bigoplus_{\begin{array}{c}
@@
                           \end{array}}}
- \weic\bigl(q, {\call{c}}, p, {\call{c}'}, p', q''\bigr)
- \otimes \weight_A\bigl(\configup{q''}{\<{\call{c}'}, p'>\stackup \<{\call{c}}, p>\stackup \gamma}, u,
-                        \config{q'}{\gamma'}\bigr)\nonumber\\[1mm]
-%
-\weight_A\bigl(\config{q}{\bot}, {\call{c}}\, u,
-               \config{q'}{\gamma'}\bigr) & =
+ \weight_A\bigl(\config{q}{\gamma}, u,
+                \configup{q''}{\<{\call{c}}, p>\stackup\gamma'}\bigr)
+ \otimes \weic\bigl(q'', {\call{c}}, p, {\call{c}'}, p', q'\bigr)
+\nonumber\\[1mm]
+%
+\weight_A\bigl(\config{q}{\gamma}, u\, {\call{c}},
+               \config{q'}{\<{\call{c}}, p>}\bigr) & =
  {\displaystyle\bigoplus_{\begin{array}{c}
@@
                           \end{array}}}
-  \weiec(q, {\call{c}}, p, q'')
-  \otimes \weight_A\bigl(\config{q''}{\<{\call{c}}, p>}, u,
-                         \config{q'}{\gamma'}\bigr)\nonumber\\
-%
-\weight_A\bigl(\configup{q}{\<{\call{c}}, p>\stackup\gamma}, {\return{r}}\, u,
+ \weight_A\bigl(\config{q}{\gamma}, u,
+                \config{q''}{\bot}\bigr)
+ \otimes \weiec(q'', {\call{c}}, p, q')
+\nonumber\\
+%
+\weight_A\bigl(\config{q}{\gamma}, u\,{\return{r}},
                \config{q'}{\gamma'}\bigr) & =
  {\displaystyle\bigoplus_{q'' \in Q}}
-  \weir\bigl(q, {\call{c}}, p, {\return{r}}, q''\bigr)
-  \otimes \weight_A\bigl(\config{q''}{\gamma}, u,
-                         \config{q'}{\gamma'}\bigr)\nonumber\\
-%
-\weight_A\bigl(\config{q}{\bot}, {\return{r}}\, u,
-               \config{q'}{\gamma'}\bigr) & =
- {\displaystyle\bigoplus_{q'' \in Q}} \weier(q, {\return{r}}, q'')
-  \otimes \weight_A\bigl(\config{q''}{\bot}, u,
-                         \config{q'}{\gamma'}\bigr)\nonumber
+  \weight_A\bigl(\config{q}{\gamma}, u,
+                 \configup{q''}{\<{\call{c}}, p>\stackup\gamma'}\bigr)
+  \otimes \weir\bigl(q'', {\call{c}}, p, {\return{r}}, q'\bigr)\nonumber\\
+%
+\weight_A\bigl(\config{q}{\gamma}, u\,{\return{r}},
+               \config{q'}{\bot}\bigr) & =
+ {\displaystyle\bigoplus_{q'' \in Q}} 
+          \weight_A\bigl(\config{q}{\gamma}, u,
+                         \config{q''}{\bot}\bigr)
+  \otimes \weier(q'', {\return{r}}, q')\nonumber
 \end{align}
@@
 \weight_A\bigl(\config{q}{\bot}, t, \config{q'}{\bot}\bigr)
-\mathop{\otimes} \mathsf{out}(q').
+\mathop{\otimes} \mathsf{out}(q')
 \label{eq:SWVPA-value}
@@
 Let $A_1$ and $A_2$ be two \SWVPA
-over the same $\Delta$, commutative $\Semiring$ and effective $\bar\Phi$.
+over the same~$\Delta$, commutative~$\Semiring$ and effective~$\bar\Phi$.
 There exist two effectively constructible $\SWVPA$
@@
 %similar to the case of the Boolean semiring~\cite{dAntonyAlur14SVPDA}.
-\input{app-product}
+\input{app-product}\qed
 \end{proof}
@@
 %%%%%% lydaddition %%%%%%
-The function $b_\bot$ will allow to compute $b_\top$ recursively in the case where the stack in not empty.
+%The function $b_\top$ will allow to compute~$b_\bot$ recursively in the case where the stack in not empty.
 %%%%%% lydaddition %%%%%%
 %
-Intuitively, $b_\top(q, p, q')$ is a function of $\Phic$,
+Intuitively, $b_\top(q, p, q')$ is a function of~$\Phic$,
 mapping every $c \in \Deltac$ to
 the minimum weight of a computation of~$A$
-starting in state $q$, 
+starting in state~$q$, 
 with a non-empty stack $\gamma' = \<c, p> \, \gamma \in \Gamma^+$,
-and ending in state $q'$ with the same stack $\gamma'$,
+and ending in state~$q'$ with the same stack~$\gamma'$,
 such that
@@
 %
-Formally, in order to define $b_\top$, we consider
+Formally, in order to define~$b_\top$, we consider
 a fresh stack symbol $\top \notin \Gamma$,   %which does not belong to $\Gamma$,
@@
 
-%\input{algo-best}
+\medskip
 \input{algo-best-alt}
+%\input{algo-best} % former version
 

### 2021-10-13 semiring.tex
alternative algo
figure intro
1 passe section semirings (remarks reviews)

@@
 %\florent{The results are established for a general class of semirings. They can be instantiated for concrete cases}
-\noindent
-\textbf{Semirings}. A \emph{semiring} $\< \Semiring, \oplus, \zero, \otimes, \one>$
+\subsection{Semirings}
+%\noindent
+A \emph{semiring} $\< \Semiring, \oplus, \zero, \otimes, \one>$
 is a structure with a domain~$\Semiring$,
@@
 %
-In the models presented in this paper,
-$\oplus$ selects an optimal value from two given values,
-in order to handle non-determinism,
-and $\otimes$ combines two values into a single one.
+%In the models presented in this paper,
+%$\oplus$ selects an optimal value from two given values,
+%in order to handle non-determinism,
+%and $\otimes$ combines two values into a single one. 
+%\florent{revise (rev. 1)}
+
 %and let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a {semiring},
@@
 Another important semiring property in the context of optimization
-is {superiority}~\cite{Huang08advanceddynamic},
-\reviews{1) also \cite{Knuth77}: "he defines the notion for functions of arbitrary arity, so not just semirings (but multioperator monoids"}
-which generalizes to the
+is {superiority}~\cite{Knuth77,Huang08advanceddynamic},
+%\reviews{1) also \cite{Knuth77}: "he defines the notion for functions of arbitrary arity, so not just semirings (but multioperator monoids"}
+which generalizes the
 \emph{non-negative weights} condition in shortest-path algorithms~\cite{Dijkstra59anote}.
 Intuitively, it means that combining elements with~$\otimes$ always increases their weight.
-Formally, it is defined as the property ($i$) below. %of the following lemma.
+Formally, %In~\cite{Huang08advanceddynamic}, 
+$\Semiring$ is called \emph{superior} \wrt~$\leq_\oplus$
+when property~$(i)$ of Lemma~\ref{lem:bounded} below holds.
 
@@
 \reviews{1) example superiority}
-In~\cite{Huang08advanceddynamic}, when  property~$(i)$ holds,
-$\Semiring$ is called \emph{superior} \wrt~$\leq_\oplus$.
-It implies (proof of Lemma~\ref{lem:bounded})
-that
-$\one \leq_\oplus x$ for every $x \in \Semiring$.
+The property~$(i)$ of superiority implies %(proof of Lemma~\ref{lem:bounded})
+that $\one \leq_\oplus x$ for every $x \in \Semiring$.
 Similarly, by the first inequality of ($i$) with $y = \zero$,

### 2021-10-13 algo-best-alt.tex
fin preuve best search

@@
 A \emph{path} of $\G_A$ is a sequence $\pi = v_0,\ldots, v_n \in V^*_A$,
-it is called \emph{safe} when $v_0$ has the form $\< q, q>$ for some $q \in Q$, 
-and $n_v$ is called $\last(\pi)$.
+where $v_0$ and $v_n$ are respectively called $\fst(\pi)$ and $\last(\pi)$.
+The path $\pi$ is called \emph{safe} when $\fst(\pi)$ has the form $\< q, q>$ for some $q \in Q$, 
 %
@@
 %$\weight(\pi) = \bigotimes_{i=1}^{n} \eta(v_{i-1} \to v_i)$.
+%
+%The following results establish that $b_\bot(q, q')$ by be computed from a shortest path in $\G_A$.
+For $v, v' \in V_A$, let $\mathit{short}(v, v')$ be the weight of the shortest path from $v$ to $v'$, 
+more precisely, 
+\(\mathit{short}(v, v') = \displaystyle\bigoplus_{\begin{array}{c}
+                                                  \scriptstyle\pi\,\mathrm{s.t.}\,
+										          \scriptstyle\fst(\pi) = v,
+	                                              \scriptstyle\last(\pi) = v'
+											      \end{array}} \weight(\pi)$.
+
+\begin{proposition}\label{algo-shortest}
+%Let  and let $\pi$ be the shortest safe path with $\last(\pi) = \< q, q'>$, i.e. 
+For all $q, q' \in Q$, 
+$b_\bot(q, q') = \displaystyle\bigoplus_{q_0 \in Q} \mathit{short}\bigl(\< q_0, q_0>, \< q, q'>\bigr)$.
+\end{proposition}
+This proposition is a consequence of Lemmata~\ref{algo-correct} and~\ref{algo-complete} below.
 
 %Intuitively, 
-\begin{lemma}[Correctness]
+\begin{lemma}[Correctness]\label{algo-correct}
 For all safe path~$\pi$ of~$\G_A$,  %with $\src(\pi) = \<q_0, q_0>$,
@@
 
-\begin{lemma}[Completeness] \label{lem:algo-completeness}
+\begin{lemma}[Completeness] \label{algo-complete}
 For all $q, q' \in Q$, $p\in P$, 
@@
 Let us do a case analysis of $w_n$.
+... \qed
+\end{proof}
 
-\end{proof}
+\medskip\noindent
+Following~\eqref{eq:min} and Proposition~\ref{algo-shortest}, 
+in order to compute the minimum of $t \mapsto A(t)$ \wrt $\leq_\oplus$, 
+it is sufficient to compute one shortest path in $\G_{A}$ of source
+$\<q_0, q_0>$ and target $\<q, q'>$ for all $q_0, q, q' \in Q$.
+This can be done with~~\cite{Mohri02semiring}~Theorem~?? \florent{completer}
+with a complexity ***. Altogether, the complexity for computing the minimum of $A$
+is in $O(...)$.
+Moreover, a witness $t \in \Delta^*$ for this minimum can be associated to the appropriate
+shortest path, like in the proof of Lemma~\ref{algo-correct}

### 2021-10-13 introduction.tex
fin preuve best search

@@
 \begin{tikzpicture}
-\node (SWADS) at (0,5.0) {%
+\node (SWADS) at (3.0,5.0) {%
   \(
   \begin{array}{c}
-  \mathsf{SWADS}\,\mbox{\cite{Herrmann16dlt,Herrmann20phd}}: \Sigma_{\mathsf{inf}}^* \to \mathbb{S}
+  \mathsf{SWADS}\,\mbox{\cite{Herrmann16dlt,Herrmann20phd}}
   \end{array}
@@
   \(
-  \mathsf{VPA}\,\cite{AlurMadhusudan09nested}
+  \mathsf{VPA}\,\mbox{\cite{AlurMadhusudan09nested}}
   \)
@@
   \(
-  \mathsf{WVPA}\,\cite{Caralp12VPAmult}
+  \mathsf{WVPA}\,\mbox{\cite{Caralp12VPAmult}}
   \)
@@
   \(
-  \mathsf{SVPA}\,\cite{dAntonyAlur14SVPDA}
+  \mathsf{SVPA}\,\mbox{\cite{dAntonyAlur14SVPDA}}
   \)
@@
 $\creturn{2}\tstamp{1}$,
+$\creturn{\mathsf{m}}\tstamp{1}$,
 $\ccall{\mathsf{m}}\tstamp{[1,2]}$,
@@
 $\creturn{3}\tstamp{2}$,
-$\creturn{\mathsf{m}}\tstamp{2}$,
 $\creturn{\mathsf{m}}\tstamp{2}$.
@@
 delimit \emph{measures},
-which are time intervals of duration 1 in this example,
+which are time intervals of duration~1 in this example,
 while the subsequences of $O$ between markups~$\ccall{d}$ and~$\creturn{d}$,

### 2021-10-13 main.tex
fin preuve best search

@@
 \newcommand{\snd}{\mathit{snd}}
+\newcommand{\first}{\mathit{first}}
 \newcommand{\last}{\mathit{last}}
@@
 %
-We propose a method for %Dijkstra-like algorithm 
+We propose first a method for %Dijkstra-like algorithm 
 computing,
@@
 \noindent
-In order to compute the above function $b_\bot : Q \times Q \to \Semiring$,
-we shall consider an auxiliary function $b_\top :  Q \times P \times Q \to \Phic$.
-%%%%%% lydaddition %%%%%%
-%The function $b_\top$ will allow to compute~$b_\bot$ recursively in the case where the stack in not empty.
-%%%%%% lydaddition %%%%%%
+Hence, in order to prove Proposition~\ref{th:best-search}, 
+it is sufficient to show that we can compute $b_\bot(q, q')$ for all $q, q' \in Q$. 
+For this purpose, we shall consider an auxiliary function $b_\top :  Q \times P \times Q \to \Phic$
+that corresponds to the cases where the stack of $A$ is not empty.
 %
@@
 the computation does not pop the pair $\<c, p>$ at the top of $\gamma'$
-(\ie $\gamma'$ is left untouched during the computation).
+(\ie~$\gamma'$ is left untouched during the computation).
 However, the computation can read $\<c, p>$ at the top of $\gamma'$,

### 2021-10-13 semiring.tex
fin preuve best search

@@
 \item[$i.$]
-\emph{infinite sums extend finite sums:}\\
-$\displaystyle\bigoplus_{i \in \emptyset} x_i = \zero$,\quad
-      $\forall j\in \mathbb{N}, \displaystyle\bigoplus_{i \in \{ j \}} x_i = x_j$,
-      $\forall j, k\in \mathbb{N}, j\neq k,
-      \displaystyle\bigoplus_{i \in \{ j, k \}} x_i = x_j \oplus x_k$,
+\emph{infinite sums extend finite sums:} $\forall j, k\in \mathbb{N}, j\neq k$,\\
+$\displaystyle\bigoplus_{i \in \emptyset} x_i = \zero$,\ 
+%$\forall j\in \mathbb{N}$, 
+$\displaystyle\bigoplus_{i \in \{ j \}} x_i = x_j$,\ 
+%$\forall j, k\in \mathbb{N}, j\neq k$,
+$\displaystyle\bigoplus_{i \in \{ j, k \}} x_i = x_j \oplus x_k$,
 %
 \item[$ii.$]
-\emph{associativity and commutativity:}\\
-for all $I \subseteq \mathbb{N}$
-and all partitions $(I_{j})_{j \in J}$ of $I$, %\subseteq \mathbb{N}$,
+\emph{associativity and commutativity:}
+%for all $I \subseteq \mathbb{N}$
+for all partitions $(I_{j})_{j \in J}$ of $I$,\\ %\subseteq \mathbb{N}$,
 \(
 \displaystyle
-\bigoplus_{j \in J}\bigoplus_{i \in I_j} x_i =
-\bigoplus_{i \in I} x_i
+\bigoplus_{j \in J}\bigoplus_{i \in I_j} x_i = \bigoplus_{i \in I} x_i
 \),
@@
 \item[$iii.$]
-\emph{distributivity of products over infinite sums:}\\
-for all $I \subseteq \mathbb{N}$, 
+\emph{distributivity of products over infinite sums:} 
+for all $I \subseteq \mathbb{N}$, $\forall x, y \in \Semiring$,\\
 %and all $x, y \in \Semiring$,
@@
 \displaystyle
-\forall x \in \Semiring, \bigoplus_{i \in I} (x \otimes y_i) = x \otimes \bigoplus_{i\in I} y_i\), and
+%\forall x \in \Semiring, 
+\bigoplus_{i \in I} (x \otimes y_i) = x \otimes \bigoplus_{i\in I} y_i\), and
 \(
 \displaystyle
-\forall y \in \Semiring, \bigoplus_{i \in I} (x_i \otimes y) = (\bigoplus_{i \in I} x_i ) \otimes y\).
+%\forall y \in \Semiring, 
+\bigoplus_{i \in I} (x_i \otimes y) = (\bigoplus_{i \in I} x_i ) \otimes y\).
 \end{description}

### 2021-10-13 algo-best-alt.tex
fin preuve best search

@@
 %The following results establish that $b_\bot(q, q')$ by be computed from a shortest path in $\G_A$.
-For $v, v' \in V_A$, let $\mathit{short}(v, v')$ be the weight of the shortest path from $v$ to $v'$, 
+For $v, v' \in V_A$,
+let $\Pi_{v, v'}$ be the set of pathes $\pi$ such that 
+$\fst(\pi) = v$ and $\last(\pi) = v'$, 
+and let $\mathit{short}(v, v')$ be the weight of the shortest path from $v$ to $v'$, 
 more precisely, 
-\(\mathit{short}(v, v') = \displaystyle\bigoplus_{\begin{array}{c}
-                                                  \scriptstyle\pi\,\mathrm{s.t.}\,
-										          \scriptstyle\fst(\pi) = v,
-	                                              \scriptstyle\last(\pi) = v'
-											      \end{array}} \weight(\pi)$.
+\(\mathit{short}(v, v') = \displaystyle\bigoplus_{\pi\in \Pi_{v, v'}} \weight(\pi)\).
 
@@
 in order to compute the minimum of $t \mapsto A(t)$ \wrt $\leq_\oplus$, 
-it is sufficient to compute one shortest path in $\G_{A}$ of source
+it is sufficient to compute one shortest pathes in $\G_{A}$ of source
 $\<q_0, q_0>$ and target $\<q, q'>$ for all $q_0, q, q' \in Q$.
@@
 Moreover, a witness $t \in \Delta^*$ for this minimum can be associated to the appropriate
-shortest path, like in the proof of Lemma~\ref{algo-correct}
+shortest path, with nno additional cost, like in the proof of Lemma~\ref{algo-correct}

### 2021-10-13 algo-best-alt.tex
petite revision section label th.

@@
 \noindent
-We show now the existence of a path $\pi$ of Lemma~\ref{lem:algo-completeness}, 
+We show now the existence of a path $\pi$ as expected, %in Lemma~\ref{lem:algo-complete}, 
 by simultaneous induction on $n_{q, q'}$ and $n_{q, p, q'}$.
 
+\noindent
 The base case, $n = 0$ corresponds to $t = \varepsilon$. 

### 2021-10-13 main.tex
biblio

@@
 %
-The models introduced in this section generalize
-weighted automata and transducers~\cite{Droste09handbook}
+The models presented in this section generalize weighted automata and transducers~\cite{Droste09handbook}
 %over finite alphabets, see \eg~\cite{Mohri03ijfcs},
@@
 that takes the input and output symbols as parameters.
-These functions are similar to the guards of symbolic automata~\cite{dAntoniVeanes17CAV,dAntoni21CACM},
-but the latter guards are restricted to the Boolean semiring.
-%
+These functions generalize the guards 
+of symbolic automata~\cite{Veanes12symbolic,dAntoniVeanes17CAV,dAntoni21CACM},
+from Boolean domains to commutative semirings.
+%
+They are particular cases of the very general model of 
+Weighted Symbolic Automata with Data Storage~\cite{Herrmann16dlt}.
 
@@
 %
-We recall that, by definition (Section~\ref{sec:prelim}),
+We recall that, by convention (Section~\ref{sec:prelim}),
 an empty sum with $\bigoplus$ is equal to~$\zero$.
@@
 This is analogous to the case of a transition's guard not satisfied by $\<a, b>$ for
-symbolic transducers.
+symbolic transducers~\cite{Veanes12transducers}.
 
@@
 
-\section{SW Visibly Pushdown Automata}
+\section{SW Visibly Pushdown Automata and Best-Search}
 \label{section:SWVPA}\label{sec:SWVPA}
@@
 
-%\section{Nested Words and Parse Trees} \label{sec:trees}\label{app:trees}
-%\input{app-trees}
+\section{Nested Words and Parse Trees} \label{sec:trees}\label{app:trees}
+\input{app-trees}
 

### 2021-10-13 main.tex
quelques remarques reviews

@@
 % \noindent
-$T$ is
+The semantics of $T$ is
 based on an intermediate function $\weight_T$
-defined recursively, for every states $q, q' \in Q$,
+defined recursively as follows, for every states $q, q' \in Q$,
 and every pairs of strings $\< s, t> \in \Sigma^* \times \Delta^*$.
@@
 \label{section:SWVPA}\label{sec:SWVPA}
-The model presented now generalizes symbolic \VPA
+We consider now a language model generalizing symbolic \VPA
 (\SVPA~\cite{dAntonyAlur14SVPDA},
@@
 from Boolean semirings to arbitrary semiring domains.
+It is also a particular case of 
+Weighted Symbolic Automata with Data Storage~\cite{Herrmann16dlt} for a special 
+type of data storage.
 It associates to every nested word over an infinite alphabet a weight value

### 2021-10-13 abstract.tex
premier redécouopage de l'intro

@@
 %
+\florent{revise following intro}
 It is based on language models called Symbolic Weighted Automata (\SWA)

### 2021-10-13 introduction.tex
biblio
premier redécouopage de l'intro

@@
 %
-Parsing is the problem
-\florent{deplacer plus loin (application)}
-of structuring a linear representation
-(a finite word) according to a language model. % (a formal grammar).
-% natural language, programming language,
-%
-Most of the context-free parsing approaches~\cite{GruneJacobs08parsing}
-assume a finite and reasonably small input alphabet. %models and algorithms
-Such a restriction makes perfect sense in the context of
-NLP tasks such as constituency parsing, 
-compilers or interpreters for programming languages.
-Considering large or infinite alphabets can however be of
-practical interest when dealing with large characters encodings such as UTF-16~\cite{dAntoni21CACM},
-% processing strings in \eg for vulnerability detection in Web-applications~\cite{dAntoni21CACM},
-%
-analysis of data streams, serialization of structured documents~\cite{Segoufin06csl,NevenSchwentickVianu04FSMinfinite},
-or processing timed execution traces~\cite{Bouyer03algebraic}.
-% algebraic definition of a class of data languages
-% (notion of monoid recognizability, based on registers, comparable to Bojancszik et al. data words)
-%
-
-The latter case is related to a motivation  of  the present work: \florent{idem}
-automated music transcription. Representations that capture  music performances
-are essentially linear:
-audio files or the widely used
-MIDI format~\cite{Selfridge-Field97beyondMIDI}. %\cite{MIDIfile}.
-Such representations ignore the hierarchical structures that frame the
-conception of music, at least in the Western area. These structures, on the other hand,
-are present, either explicitly  or implicitly,
-in Common Western Music Notation~\cite{Gould11Notation}:
-%in music notation~\cite{Gould11Notation}:
-Music scores are partitioned in measures,
-measures in beats, and beats can be further recursively divided.
-It follows that music events do not occur at arbitrary timestamps,
-but respect a discrete partitioning of the  timeline incurred by
-these recursive divisions.
-The \emph{transcription problem} takes
-as input a linear representation (audio or MIDI) and aims at re-constructing
-these structures
-by mapping input events to this hierarchical rhythmic space.
-It can therefore be stated as a parsing problem~\cite{foscarin:hal-01988990}
-over an infinite alphabet of timed events.
-
-Various extensions of language models for handling infinite alphabets have been studied.
+Various extensions of language models for handling infinite alphabets have been proposed.
 %words carrying data values in an infinite domain (e.g. integers) e.g. data processing
@@
 Other extensions of language models  %(automata and grammars)
-help in dealing with non-determinism by computing weight values.
-\florent{quantitative reasoning about programs}
-With an ambiguous grammar, 
-\reviews{1) jump from automata to grammars is uneasy and not really necessary}
-there may exist several derivations
-(\emph{abstract syntax trees} -- AST) % \ie the result of an analyze.
-yielding one input word. % structuring a word,
-The association of one weight value %in semiring domain
-to each AST permits to select a best one (or $n$ bests). % a fixed number of bests
-% = ...a ranking of ASTs
+help in dealing with non-determinism by computing weight values... associated to input..
+*** In the context of parsing for instance, associating one weight to every grammar's derivation
+permits to rank \emph{abstract syntax trees} -- AST, 
+hence to select a best one (or $n$ bests) when using am ambiguous grammar, 
 %This is roughly the principle of \emph{weighted parsing} approaches
 \florent{introduce pb "weighted parsing" and ref.\cite{MorbitzVogler19weighted-parsing}.} 
+%\reviews{1) jump from automata to grammars is uneasy and not really necessary}
+% for which there may exist several derivations yielding one input word.
+\florent{revise}
+*** For verification, ... quantitative reasoning about programs
+%
 In \emph{weighted language models}~\cite{Goodman99SemiringParsing,Nederhof03weightedParsing,MorbitzVogler19weighted-parsing},
@@
 
-\florent{Both WA and SA generalized by \cite{Herrmann16dlt,Herrmann20phd}...
-         we present results...}
-In this paper, we present 
-a framework for weighted parsing over infinite input alphabets.
-It is based on \emph{symbolic weighted} finite states language models (\swM),
+In this paper, we present results for 
+\emph{symbolic weighted} finite states language models (\swM) extensions
 generalizing the Boolean guards of~\SA %in transitions
@@
 and generalizing also~\WA, by handling infinite alphabets (Figure~\ref{fig:hierarchy}).
+...
 \reviews{cite \cite{Herrmann16dlt,Herrmann20phd}}
@@
 
-The framework relies on several language models:
+%The framework relies on several language models:
+The models studied are: 
 symbolic-weighted automata (\SWA),
@@
 to a given nested word $t$, which is itsef the linearization of a weighted AST. %representing a parse tree.
+
+\SWA and \SWVPA are both particular cases of \cite{Herrmann16dlt,Herrmann20phd}, 
+with special data storage types...
+\florent{Both WA and SA generalized by \cite{Herrmann16dlt,Herrmann20phd}...}
+
+THE RESULTS/CONTRIBUTIONS of this paper are:
+\florent{revise: main results = 3 theorems}
+%\reviews{1) not $i$! see response.txt}
+%($i$)~the introduction of automata, \SWA, transducers, \SWT (Section~\ref{section:SWA}),
+%and visibly pushdown automata \SWVPA (Section~\ref{sec:SWVPA}),
+%generalizing the corresponding classes of symbolic and weighted models,
+%for the weighted computation on (nested) words over infinite alphabets;
+($i$)~a construction %Bar-Hillel Perles Shamir 
+for computing a \SWT-defined distance between an \SWA language and a word (Section~\ref{section:SWA}), 
+($ii$)~closure results for \SWVPA (Section~\ref{sec:SWVPA}), 
+($iii$)~a polynomial best-search algorithm for \SWVPA, %(Section~\ref{sec:best})
+% a framework for parsing over infinite alphabets,
 %
-Then, given an input word $s$, the \emph{SW-parsing} problem aims at
+Moreover, we detail 
+($iv$)~an application to the problem of weighted parsing over infinite input alphabets 
+(\emph{SW-parsing}, Section~\ref{sec:parsing}), 
+based on the above models integrated in a uniform framework, with in particular:
+($iv.a$)~the \SWT-based definition of generic edit distances between input and output (yield) words,
+and ($iv.b$)~the use, for technical convenience, of nested words, and \SWVPA,
+instead of syntax trees and grammars. %(\S~\ref{sec:trees}).
+
+... These results should also apply in the context of the quantitive verification of programs...
+
+***
+Parsing is the problem of structuring a linear representation
+(a finite word) according to a language model. % (a formal grammar). % natural language, programming language,
+%
+Most of the context-free parsing approaches~\cite{GruneJacobs08parsing}
+assume a finite and reasonably small input alphabet. %models and algorithms
+Such a restriction makes perfect sense in the context of
+NLP tasks such as constituency parsing, 
+compilers or interpreters for programming languages.
+Considering large or infinite alphabets can however be of
+practical interest when dealing with large characters encodings such as UTF-16~\cite{dAntoni21CACM},
+\florent{applis à la Fossacs chez Veanes et al?}
+% processing strings in \eg for vulnerability detection in Web-applications~\cite{dAntoni21CACM},
+%
+analysis of data streams, serialization of structured documents~\cite{Segoufin06csl,NevenSchwentickVianu04FSMinfinite},
+or processing timed execution traces~\cite{Bouyer03algebraic}.
+% algebraic definition of a class of data languages
+% (notion of monoid recognizability, based on registers, comparable to Bojancszik et al. data words)
+
+The latter case is related to a motivation  of  the present work: \florent{idem}
+automated music transcription. Representations that capture  music performances
+are essentially linear:
+audio files or the widely used
+MIDI format~\cite{Selfridge-Field97beyondMIDI}. %\cite{MIDIfile}.
+Such representations ignore the hierarchical structures that frame the
+conception of music, at least in the Western area. These structures, on the other hand,
+are present, either explicitly  or implicitly,
+in Common Western Music Notation~\cite{Gould11Notation}:
+%in music notation~\cite{Gould11Notation}:
+Music scores are partitioned in measures,
+measures in beats, and beats can be further recursively divided.
+It follows that music events do not occur at arbitrary timestamps,
+but respect a discrete partitioning of the  timeline incurred by
+these recursive divisions.
+The \emph{transcription problem} takes
+as input a linear representation (audio or MIDI) and aims at re-constructing
+these structures
+by mapping input events to this hierarchical rhythmic space.
+It can therefore be stated as a parsing problem~\cite{foscarin:hal-01988990}
+over an infinite alphabet of timed events.
+
+Given an input word $s$, the \emph{SW-parsing} problem aims at
 finding $t$ minimizing
@@
 The first step is an intersection
-(Bar-Hillel construction~\cite{GruneJacobs08parsing})
+(Bar-Hillel construction~\cite{NederhofSatta03ParsingIntersection,GruneJacobs08parsing})
 \reviews{1) also \cite{NederhofSatta03ParsingIntersection} pasing as intersection}
@@
 % value in a semiring...
+
 \florent{expressiveness: VPA have restricted equality test.
         comparable to pebble automata? $\to$ conclusion}
-The main contributions 
-\florent{main results = 3 Prop.}
-of the paper are:
-\reviews{1) not $i$! see response.txt}
-($i$)~the introduction of automata, \SWA, transducers, \SWT (Section~\ref{section:SWA}),
-and visibly pushdown automata \SWVPA (Section~\ref{sec:SWVPA}),
-generalizing the corresponding classes of symbolic and weighted models,
-%for the weighted computation on (nested) words over infinite alphabets;
-($ii$)~a polynomial best-search algorithm for \SWVPA, %(Section~\ref{sec:best})
-% a framework for parsing over infinite alphabets,
-and ($iii$)~a uniform framework (Section~\ref{sec:parsing}) for parsing over infinite alphabets,
-the keys to which are
-($iii.a$)~the \SWT-based definition of generic edit distances between input and output (yield) words,
-and ($iii.b$)~the use, for convenience in this context,
-of nested words, and \SWVPA,
-instead of syntax trees and grammars. %(\S~\ref{sec:trees}).
-
-\florent{Application to parsing and tanscription. these results are appliable to other decision pb.}
 

### 2021-10-14 introduction.tex
inclusion contruction Prop. Bar-Hillel Perles Shamir et preuve correction en appendix

@@
 Following a similar idea, % and generalizing,
-in symbolic automata (\SA)~\cite{dAntoniVeanes17CAV,dAntoni21CACM},
+in symbolic automata (\SA)~\cite{Veanes12symbolic,dAntoniVeanes17CAV,dAntoni21CACM},
 transitions are guarded by predicates over infinite  domains.
@@
   \begin{array}{c}
-  \mathsf{SA}\,\mbox{\cite{dAntoni21CACM}}: \Sigma_{\mathsf{inf}}^* \to \mathbb{B}\\
+  \mathsf{SA}\,\mbox{\cite{Veanes12symbolic,dAntoni21CACM}}: \Sigma_{\mathsf{inf}}^* \to \mathbb{B}\\
   q \xrightarrow{\phi} q', \phi : \Sigma_\mathsf{inf} \to \mathbb{B}
@@
   \(
-  \mathsf{WVPA}\,\mbox{\cite{Caralp12VPAmult}}
+  \mathsf{WVPA}\,\mbox{\cite{Mathissen08weighted,Caralp12VPAmult}}
   \)
@@
 SW-parsing associates a weight value
-to each solution, and our framework
+to each solution, and our approach
 aims at selecting the best one with respect to this weight.

### 2021-10-14 main.tex
inclusion contruction Prop. Bar-Hillel Perles Shamir et preuve correction en appendix

@@
 \newcommand{\W}{\mathbb{W}}
+\newcommand{\Rho}{\mathit{R}}
 \newcommand{\Semiring}{\mathbb{S}}
@@
 We build a small \SWT over the alphabets $\Sigma$ and $\Deltai$
-(Ex.~\ref{ex:running} and \ref{distance-time}),
+of Ex.~\ref{ex:running} and~\ref{distance-time},
 %weighted transducer model
@@
 \begin{proof}
-(sketch, see Appendix~\ref{sec:closure} for details).
-The state set of $B_{A, T}$ is the Cartesian product of
-the state sets of $A$ and $T$ and its transitions simulate,
-while reading an output word $t \in \Delta^*$, the
-synchronized behaviour of $A$ and $T$ on $t$
-and some input word $s \in \Sigma^*$.
-The weight for reading the input $s$ is obtained with
-$\bigoplus^1_\Sigma$.
-%
-The main difficulty comes from the transitions
-of the form $\wei_{10}$, which read in input and ignore the output.
-Such transition shall be simulated by $\varepsilon$-transitions in $B_{A, T}$,
-but $\varepsilon$-transitions are not defined for $\SWA$.
-Therefore, the $\varepsilon$-transitions
-are eliminated on-the-fly during the construction of~$B_{A, T}$,
-following a procedure of~\cite{LombardySakarovitch12ciaa}.
-%\qed
+%(sketch, see Appendix~\ref{sec:closure} for details).
+Let $T = \< Q, \init_T, \bar{\wei}, \final_T >$,
+where $\bar{\wei}$ contains $\wei_{10}$, $\wei_{01}$, and $\wei_{11}$,
+from $Q \times Q$ into respectively
+$\Phi_{\Sigma}$, $\Phi_{\Delta}$, and $\Phi_{\Sigma, \Delta}$,
+and let $A = \< P, \init_A, \wei_1, \final_A >$
+with $\wei_1: P \times P \to \Phi_{\Sigma}$.
+The state set of $B_{A, T}$ will be %the Cartesian product of the state sets of $A$ and $T$ 
+$Q' = P \times Q$,
+and its entering, leaving and transition functions will
+simulate, while reading an output word $t \in \Delta^*$, 
+the synchronized computations of $A$ and $T$,
+when they read both the output word $t$
+and an input word $s \in \Sigma^*$.
+%The weight for reading the input $s$ is obtained with $\bigoplus^1_\Sigma$.
+
+\noindent
+The main difficulty comes from the transitions of $T$
+of the form $\wei_{10}$, which read in input $s$ and ignore the output $t$.
+Since the automaton $B_{A, T}$ only reads the output word $t$, 
+such transitions would correspond to $\varepsilon$-transitions in $B_{A, T}$.
+But $\varepsilon$-transitions are not defined for $\SWA$.
+Therefore, we shall perform on-the-fly elimination of the $\varepsilon$-transitions
+during the construction of~$B_{A, T}$,
+following an approach of~\cite{LombardySakarovitch12ciaa}, Algorithm~1.
+
+\noindent
+The state entering function of $B_{A, T}$ is defined
+for all $\<p_1, q_1> \in Q'$, %$p_1 \in P$, $q_1 \in Q$
+by:
+%
+\begin{equation} \label{eq:initprime}
+\init'\bigl(\< p_1, q_1>\bigr) = \init_A(p_1) \otimes \init_T(q_1).
+\end{equation}
+%
+The transition function $\wei'_1$ of $B_{A, T}$
+will roughly perform a synchronized product of transitions defined by $\wei_1$
+($A$ reading in input word $s$),
+$\wei_{01}$ ($T$ reading in output word $t$ and not in input word $s$),
+and $\wei_{11}$ ($T$ reading both in input word $s$ and in output word $t$).
+%
+Moreover, the $\varepsilon$-transitions to eliminate come from the simulation, 
+by $\wei'_1$, of the transitions defined by $\wei_1$ and $\wei_{10}$ 
+($A$ and $T$ reading in input word $s$ and not in output word $t$).
+%
+% Initialize state leaving functions i
+%and $\final'(p, q) = \final_A(p) \otimes \final_T(q)$.
+%
+%Every transition of $B_{A, T}$ will
+%simulate a sequence of transitions of $T$ performing the following steps:
+%advance in the input word while staying immobile in the output word,
+%and then make one step in the output word (and advance in the input word or not).
+%
+\noindent
+Let us construct the transition function $\wei'_1$ iteratively.
+%
+\noindent
+Initially, for all $p_1, p_2 \in P$, and $q_1, q_2 \in Q$, let
+%
+\begin{align}
+%\begin{array}{rcl}
+\wei'_1\bigl( \< p_1, q_1>, \< p_2, q_2>\bigr) & = 
+\bigl({\displaystyle\bigoplus_{p_1 = p_2}} \wei_{01}(q_1, q_2)\;\bigr)
+\oplus
+{\bigoplus}_\Sigma^1 \bigl(\wei_1(p_1, p_2) \otimes_1 \wei_{11}(q_1, q_2)\bigr)
+\label{eq:init-wei}\\
+%
+\final'\bigl(\<p_1, q_1>\bigr) & =  \final_A(p_1) \otimes \final_T(q_1)
+\label{eq:init-final}
+%\end{array}
+\end{align}
+We recall that by convention, 
+${\displaystyle\bigoplus_{p_1 = p_2}} \wei_{01}(q_1, q_2)$
+is equal to $\zero$ if $p_1 \neq p_2$.
+
+\noindent
+Then, we iterate the following updates for all $p_1, p_2, p_3\in P$ 
+and $q_1, q_2, q_3 \in Q$:
+%
+\begin{align}
+%\begin{array}{rcl}
+\wei'_1\bigl( \< p_1, q_1>, \< p_3, q_3>\bigr) & \opluseq 
+%\displaystyle\bigoplus_{\Sigma} 
+{\textstyle\bigoplus_\Sigma} \bigl(\wei_{1}(p_1, p_2) \otimes \wei_{10}(q_1, q_2)\bigr)
+\otimes \wei'_1\bigl( \< p_2, q_2>, \< p_3, q_3>\bigr) \label{eq:iter-wei}\\
+%
+\final'\bigl(\<p_2, q_2>\bigr) & \opluseq 
+%\displaystyle\bigoplus_{\Sigma} 
+{\textstyle\bigoplus_\Sigma} \bigl(\wei_{1}(p_1, p_2)\otimes \wei_{10}(q_1, q_2)\bigr)
+\otimes \final'(p_1, q_1) \label{eq:iter-final}
+%\end{array}
+\end{align}
+%
+In both cases of updates of $\wei'_1$~\eqref{eq:iter-wei}
+and $\final'$~\eqref{eq:iter-final} during the iteration, 
+$\wei_{1}(p_1, p_2) \otimes \wei_{10}(q_1, q_2)$ 
+is the weight of an $\varepsilon$-transition.
+It corresponds to the reading, by~$A$ and~$T$, 
+of a symbol $a$ in the input word $s$ without moving in the output word,
+\ie the synchronization of 
+a transition $\wei_{1}(p_1, a, p_2)$ of $A$ and 
+a transition $\wei_{10}(q_1, a, \varepsilon, q_2)$ of~$T$.
+%
+The iteration stops if it does not change the value of $\wei'_1$ and $\final'$.
+By hypothesis and Lemma~\ref{lem:idempotent}, 
+$\Semiring$ is idempotent. 
+Therefore, the construction of $B_{A, T}$ will stop after 
+at most $|P|^2 . |Q|^2$ iterations.
+%
+The correctness of this construction of $B_{A, T}$
+is proved in Appendix~\ref{sec:closure}.
+\qed
 \end{proof}
@@
 \noindent
-We assume the following input to be given:
+We assume to be given the following input:
 \begin{description}
@@
 $\SW$-parsing is the problem of computing
-the distance~\eqref{eq:distance-lang} between the input string ~$s$ and the weighted language over the output alphabet defined by ~$A$,
-and returning a witness $t$.
+the distance~\eqref{eq:distance-lang} between the input string~$s$ 
+and the weighted language over the output alphabet defined by~$A$,
+and returning a witness~$t$.
 
@@
 \item The input word is $I$ of Example~\ref{ex:running}.
-\item The \SWT $T$ evaluates a ``fitness measure''  that expresses
-    a correspondence between a performance and a nested representation of a music score.
-    See Example~\ref{ex:SWT}.
-\item The \SWVPA $A$ expresses a cost related to the music notation.
+\item The \SWT $T$ evaluates a ``fitness measure'' expressing
+    a correspondence between a performance and a nested representation of a music score
+    - Ex.~\ref{ex:SWT}.
+\item The \SWVPA $A$ expresses a cost related to the complexity of music notation.
 \end{itemize}
@@
 when the weight assigned to an additional second time division with $\ccall{2}$ is less
-than the difference of weight between
-$\mbox{`\footnotesize C$\sharp$5`}$
-and $\mbox{\footnotesize C$\sharp$5}$.
+than the difference of weight between\
+the appogiatura $\mbox{`\footnotesize C$\sharp$5`}$
+and the standard note $\mbox{\footnotesize C$\sharp$5}$.
 
@@
 \noindent
-It remains to compute a best nested word $t \in {\Delta}^*$
+It remains to compute a best nested word $t \in {\Delta}^*$ for this \SWVPA,
 using the procedure of Proposition~\ref{th:best-search}.
@@
 %for~$T$, $A$ and~$s$.
-%\qed
+\qed
 \end{proof}
@@
 
-\section{Proof of Proposition~\ref{prop:epsilon}} \label{sec:closure}
-\input{app-closure}
+\section{End of proof of Proposition~\ref{prop:epsilon}} \label{sec:closure}
+\input{app-Bar-Hillel-Perles-Shamir}
 

### 2021-10-14 algo-best-alt.tex
proof correctness

@@
 let us do a case analysis on the edge $v_{n-1} \to v_n$.
-...
-\qed
+%
+We suppose Lemma ~\ref{algo-correct} is true for a safe path $\pi' = v_0,\ldots, v_{n-1}$ and a word $u' \in \Delta^*$.
+Firstly, let's assume $v_{n-1} = \< q, q''>$ and $last(\pi) = \< q, q'>$.
+By definition, the edge $\< q, q''> \to \< q, q'>$ is possible only if an internal symbol or a return symbol is read, in order to keep the stack empty:
+
+\begin{itemize}
+\item if we read $a \in \Delta_i$ and have an empty stack, then the new $\weight_A$ is computed with the third case in equation~\eqref{eq:SWVPA-weight}.
+We have
+\begin{align*}
+ \weight_A\bigl(\config{q}{\gamma}, u\, a,
+               \config{q'}{\bot}\bigr) &=
+   \weight_A\bigl(\config{q}{\gamma}, u, \config{q''}{\bot}\bigr)
+   \otimes\weiei(q'', a, q')  \\
+ &= \weight(\pi')\otimes\weiei(q'', a, q')\\
+ &= \weight(\pi)\nonumber 
+\end{align*}
+with $\pi = (\pi', v_n)$, since we know that $\weight(v_0,\ldots, v_{n-1}) \otimes \eta_A(v_{n-1} \to v_{n}) = \weight(v_0,\ldots, v_n)$ and $\weiei(q'', a, q')$ is indeed a possible edge label $\eta_A$ for the edge $\< q, q''> \to \< q, q'>$ (first line of label definition).
+Hence in this case the Lemma is true.
+
+\item Following the same reasoning as before, we find that the Lemma is also true when $a \in \Delta_r$ by using the last equation in \eqref{eq:SWVPA-weight}.
+%
+\end{itemize}
+\noindent
+Let's now assume that we have a safe path $\pi' = (v_0,\ldots, v_{n-1})$ such that $v_{n-1} = \< q_1, p, q_2>$ and $last(\pi) = \< q_0, q_3>$.
+The edge $\< q_1, p, q_2> \to \< q_0, q_3>$ exists if we read a return then try to match with the call on top of the stack which then leaves the stack empty.
+With the 6th equation in~\eqref{eq:SWVPA-weight}, we can compute the $\weight_A$ for $\<q_0, q_3>$ and then replace the first term with the 5th equation in~\eqref{eq:SWVPA-weight}, such as:
+\begin{align*}
+\weight_A\bigl(\config{q_0}{\bot}, u\,{\return{r}},
+               \config{q_3}{\bot}\bigr) &=
+  \weight_A\bigl(\config{q_0}{\bot}, u,
+                 \configup{q_2}{\<{\call{c}}, p>\stackup\bot'}\bigr)
+  \otimes \weir\bigl(q_2, {\call{c}}, p, {\return{r}}, q_3\bigr)\\
+  &=
+   \weight_A(\config{q_1}{\bot}, u, \config{q_2}{\bot}\bigr)
+ \otimes \weiec(q_0, {\call{c}}, p, q_1)
+  \otimes \weir\bigl(q_2, {\call{c}}, p, {\return{r}}, q_3\bigr)\\
+  &= \weight_A(\pi')\otimes \weiec(q_0, {\call{c}}, p, q_1)
+  \otimes \weir\bigl(q_2, {\call{c}}, p, {\return{r}}, q_3\bigr)\\
+  &= \weight_A(\pi)
+\end{align*}
+for $\pi = (\pi', v_n)$. Here the Lemma holds since $\weir\bigl(q_2, {\call{c}}, p, {\return{r}}, q_3\bigr)$ is also part of the edge labels (second line).
+
+Finally, let's assume that $v_{n-1} = \<q_1, p, q_2>$ and $last(\pi) = \<q_0, p, q_3>$ for $p, p' \in Q$.
+Following the same reasoning as the two previous cases, we can similarly prove that the Lemma is also true.
+
 \end{proof}
@@
 Since $\Semiring$ is total, there exists finite sequences as above such that
-\( b_\bot(q, q') = \bigotimes_{i=i}^n w_i(\tau_i)\).
+\( b_\bot(q, q') = \bigotimes_{i=1}^n w_i(\tau_i)\).
 %
@@
 and hence 
-\( {\textstyle\bigoplus_\Deltac} b_\top(q, p, q') = \bigotimes_{i=i}^n w_i(\tau_i) \),
+\( {\textstyle\bigoplus_\Deltac} b_\top(q, p, q') = \bigotimes_{i=1}^n w_i(\tau_i) \),
 using $c \in \Deltac$ that minimize the function in rhs of~\eqref{eq:btop-unfold}.
@@
 Moreover, a witness $t \in \Delta^*$ for this minimum can be associated to the appropriate
-shortest path, with nno additional cost, like in the proof of Lemma~\ref{algo-correct}
+shortest path, with no additional cost, like in the proof of Lemma~\ref{algo-correct}.

### 2021-10-14 algo-best-alt.tex
corr
lemma to app

--- 2021-10-13 algo-best-alt.tex
@@
 \end{proposition}
-This proposition is a consequence of Lemmata~\ref{algo-correct} and~\ref{algo-complete} below.
-
-%Intuitively, 
-\begin{lemma}[Correctness]\label{algo-correct}
-For all safe path~$\pi$ of~$\G_A$,  %with $\src(\pi) = \<q_0, q_0>$,
-%with last element $v$, \\
-($i$) if $\last(\pi) = \< q, q'>$, %with $q, q' \in Q$, 
-then there exists $u \in \Delta^*$ such that 
-$\weight_A(q[\bot], u, q'[\bot]) = \weight(\pi)$,\\
-%
-($ii$) if $\last(\pi) = \< q, p, q'>$, %with $q, q' \in Q$, $p\in P$, 
-then there exists $v \in \Delta^*$ such that \\
-${\displaystyle\bigoplus_{c \in \Deltac}}
- \weight_A\bigl(\configup{q}{\< c, p> \stackup \top }, v, \configup{q'}{\<c, p> \stackup \top}\bigr)
- = \weight(\pi)$.
-\end{lemma}
-%
-\begin{proof}
-We prove ($i$) and ($ii$) simultaneously by induction on the length $|\pi|$ of the path $\pi$.
-
-\noindent
-The base case $|\pi| = 1$ is a direct consequence of the definition of weight of paths of length 1
-and the first line of \eqref{eq:SWVPA-weight}.
-
-\noindent
-If $\pi = v_0,\ldots, v_n$ with $n \geq 1$, 
-let us do a case analysis on the edge $v_{n-1} \to v_n$.
-...
-\qed
-\end{proof}
-
-\begin{lemma}[Completeness] \label{algo-complete}
-For all $q, q' \in Q$, $p\in P$, 
-($i$) there exists a safe path~$\pi$ of~$\G_A$ such that
-%such that $\src(\pi) = \<q_0, q_0>$ for some $q_0 \in Q$, 
-$\last(\pi) = \< q, q'>$,
-and $\weight(\pi) = b_\bot(q, q')$, 
-($ii$) there exists a safe path $\pi'$ of $\G_A$ 
-such that $\last(\pi') = \< q, p, q'>$,
-and $\weight(\pi') = \bigoplus_\Deltac b_\top(q, p, q')$.
-\end{lemma}
-%
-\begin{proof}
-By associativity, commutativity and distributivity for $\Semiring$,
-\eqref{eq:bbot} can be rewritten into the form, unfolding~\eqref{eq:SWVPA-weight}:
-%
-\begin{equation}\label{eq:bbot-unfold}
-b_\bot(q, q') = 
-\bigoplus_{t\in \Delta^*} 
-\bigoplus_{\begin{array}{c}
-           \scriptstyle q_0,\ldots, q_n \in Q\\
-           \scriptstyle p_{i},\ldots, p_{k} \in P
-           \end{array}}
-\bigotimes_{i=1}^n w_i(\tau_i)           
-\end{equation}
-%
-where $n$ is the length of $t$, $k \leq n$,
-$q_0 = q$, $q_n = q'$, 
-for all $1 \leq i < n$, 
-$w_i$ is one of the functions of $\bar\wei$,
-$\tau_i$ is a transition of $A$ and
-$\src(\tau_i) = q_{i-1}$, $\snd(\tau_i) = q_i$.
-%
-Since $\Semiring$ is total, there exists finite sequences as above such that
-\( b_\bot(q, q') = \bigotimes_{i=i}^n w_i(\tau_i)\).
-%
-There might exists, for $q$ and $q'$,
-several finite sequences~$\bar\tau$ and~$\bar{w}$ as above, 
-let us choose arbitrarily one of minimal length~$n$. 
-This integer $n$ will be denoted $n_{q, q'}$ in the following.
-%We represent a computation of $A$ on $t$ by a sequence $\rho$ of its transitions,
-
-Similarly, following~\eqref{eq:btop} and~\eqref{eq:SWVPA-weight}, 
-$\bigoplus_\Deltac b_\top(q, p, q')$ can be put in the form:
-%
-\begin{equation}\label{eq:btop-unfold}
-{\textstyle\bigoplus_\Deltac} b_\top(q, p, q') = 
-{\textstyle\bigoplus_\Deltac} \bigoplus_{t\in \Delta^*} 
-\bigoplus_{\begin{array}{c}
-           \scriptstyle q_0,\ldots, q_n \in Q\\
-           \scriptstyle p_{i},\ldots, p_{k} \in P
-           \end{array}}
-\bigotimes_{i=1}^n w_i(\tau_i)           
-\end{equation}
-%
-and hence 
-\( {\textstyle\bigoplus_\Deltac} b_\top(q, p, q') = \bigotimes_{i=i}^n w_i(\tau_i) \),
-using $c \in \Deltac$ that minimize the function in rhs of~\eqref{eq:btop-unfold}.
-We denote the smallest $n$ as above by $n_{q, p, q'}$.
-
-\noindent
-We show now the existence of a path $\pi$ of Lemma~\ref{lem:algo-completeness}, 
-by simultaneous induction on $n_{q, q'}$ and $n_{q, p, q'}$.
-
-The base case, $n = 0$ corresponds to $t = \varepsilon$. 
-In this case, by~\eqref{eq:SWVPA-weight}, $b_\top(q, p, q') = \zero$,
-$b_\bot(q, q) = \one$ and $b_\bot(q, q') = \zero$ if $q \neq q'$.
-
-\noindent
-For $n = n_{q, q'} > 0$, let $\bar\tau$ and~$\bar{w}$
-be the sequences associated to $q$ and $q'$ as above.
-Let us do a case analysis of $w_n$.
-... \qed
-\end{proof}
+This proposition is a consequence of Lemmata~\ref{algo-correct} and~\ref{algo-complete} in~Appendix \ref{app:shortpath}.
 

### 2021-10-14 abstract.tex
debut intro

--- 2021-09-01 abstract.tex
@@
 %
-We propose a framework for weighted parsing over infinite alphabets.
-%
-It is based on language models called Symbolic Weighted Automata (\SWA)
+\florent{revise following intro}
+We study properties and relationship between 3 class of
+quantitative language models computing over infinite input alphabets:
+%propose a framework for weighted parsing over infinite alphabets.
+Symbolic Weighted Automata (\SWA)
 at the joint between Symbolic Automata (\SA) and Weighted Automata (\WA),

### 2021-10-14 introduction.tex
debut intro

@@
 %
-Various extensions of language models for handling infinite alphabets have been proposed.
+Various extensions of language models have been proposed for handling infinite alphabets.
 %words carrying data values in an infinite domain (e.g. integers) e.g. data processing
 Some automata 
-\reviews{1) and grammar with storage?}
+%\reviews{1) and grammar with storage?}
 with memory extensions
@@
 %and correspond with logics,
-(see~\cite{Segoufin06csl} for a survey)
+(see~\cite{Segoufin06csl} for a survey).
 \florent{register: skip refs and details, add Mikolaj recent}
-using pebbles for marking positions~\cite{NevenSchwentickVianu04FSMinfinite},
+They use pebbles for marking positions~\cite{NevenSchwentickVianu04FSMinfinite},
 registers~\cite{KaminskiFrancez94},
 or %computations in 2 steps
-the possibility to compute on subsequences
+separate computation on subsequences
 with the same attribute values~\cite{Bojanczyk11FO2}. % data words automata.
@@
 With appropriate closure conditions on the sets of such predicates, % (alphabet theories),
-all the good properties enjoyed over finite alphabets are preserved.
+all the good properties enjoyed by automata over finite alphabets are preserved.
+The ability to compare input symbols of \SA is quite restricted compared to the 
+former automata with memory. 
+It can however be extended with the addition of a stack~\cite{dAntonyAlur14SVPDA}.
 
 Other extensions of language models  %(automata and grammars)
-help in dealing with non-determinism by computing weight values... associated to input..
-*** In the context of parsing for instance, associating one weight to every grammar's derivation
-permits to rank \emph{abstract syntax trees} -- AST, 
-hence to select a best one (or $n$ bests) when using am ambiguous grammar, 
+are able to assign a weight value to their input.
+%help in dealing with non-determinism
+This is useful in order for quantitative modelling 
+\eg of probabilistic or stochastic programs (recursive for pushdown models),
+quantitative database queries, or semi-structured data (in infinite domains), 
+as well as for the verification of quantitative properties about 
+\florent{quantitative reasoning/verification on programs}
+quality measures, resource-consumption, distance metrics, probabilistic guarantees, \etc
+%
+In the context of parsing, for grammars computing a weight value, 
+it is possible to rank derivations (and \emph{abstract syntax trees})
+in order to select a best one 
+(or $n$ bests)~\cite{Goodman99SemiringParsing,Nederhof03weightedParsing,MorbitzVogler19weighted-parsing-parsing}. 
+%\florent{= "weighted parsing"}
+This is useful in order to deal with ambiguity of grammars.
 %This is roughly the principle of \emph{weighted parsing} approaches
-\florent{introduce pb "weighted parsing" and ref.\cite{MorbitzVogler19weighted-parsing}.} 
 %\reviews{1) jump from automata to grammars is uneasy and not really necessary}
 % for which there may exist several derivations yielding one input word.
-\florent{revise}
-*** For verification, ... quantitative reasoning about programs
 %
-In \emph{weighted language models}~\cite{Goodman99SemiringParsing,Nederhof03weightedParsing,MorbitzVogler19weighted-parsing},
+In \emph{weighted language models}
 like \eg probabilistic context-free grammars % (CFG),
-and weighted automata (\WA) \cite{Droste09handbook},
+and weighted automata (\WA)~\cite{Droste09handbook},
 a weight is associated to each transition rule, % production rule
 and the rule's weights can be combined with an
-associative product operator~$\otimes$ to yield the weight of an AST.
-\reviews{1) AST for grammar, run for WA}
+associative product operator~$\otimes$ to yield the weight of a whole computation.
+%\reviews{1) AST for grammar, run for WA}
 A second operator~$\oplus$
 is moreover used to resolve the ambiguity raised by the existence
-of several (in general exponentially many) AST
-associated to a given input word.
+of several (in general exponentially many) computations on a given input.
 Typically, $\oplus$ selects the best of two weight values.
@@
 
-In this paper, we present results for 
-\emph{symbolic weighted} finite states language models (\swM) extensions
+In this paper, we present some results for 
+\emph{symbolic weighted} finite states language models %(\swM) 
 generalizing the Boolean guards of~\SA %in transitions
 to functions into an arbitrary semiring, 
-and generalizing also~\WA, by handling infinite alphabets (Figure~\ref{fig:hierarchy}).
-...
-\reviews{cite \cite{Herrmann16dlt,Herrmann20phd}}
+and extending also~\WA, by handling infinite alphabets (Figure~\ref{fig:hierarchy}).
 %
-In short, a transition rule $q \xrightarrow{\phi} q'$, from state $q$ to $q'$,
+The models considered in this paper are also 
+%\SWA and \SWVPA are both 
+particular cases of the very general class of 
+Weighted Symbolic Automata with Data Storage defined in~\cite{Herrmann16dlt,Herrmann20phd}. 
+%with special data storage types...
+%\florent{Both WA and SA generalized by \cite{Herrmann16dlt,Herrmann20phd}...}
+%\reviews{cite \cite{Herrmann16dlt,Herrmann20phd}}
+
+In short, a transition rule $q \xrightarrow{\phi} q'$ of a SW automaton, 
+from state $q$ to $q'$,
 %\florent{c'est un résumé de la Fig. \ref{fig:hierarchy}}
@@
 is labeled by a function~$\phi$ associating to every input symbol $a$ a weight value $\phi(a)$
-in a semiring $\Semiring$.
+in a semiring~$\Semiring$.
 
@@
 to a given nested word $t$, which is itsef the linearization of a weighted AST. %representing a parse tree.
-
-\SWA and \SWVPA are both particular cases of \cite{Herrmann16dlt,Herrmann20phd}, 
-with special data storage types...
-\florent{Both WA and SA generalized by \cite{Herrmann16dlt,Herrmann20phd}...}
 

### 2021-10-14 main.tex
onte on related work/models
lemma to app
debut intro

@@
 \label{sec:SWVPA-def}
-\reviews{3) pb countable}
+%\reviews{3) pb countable}
 Let $\Delta$ be a countable alphabet,
@@
 
-We  introduced Symbolic Weighted language models %over infinite alphabets,
+We studied 3 properties of Symbolic Weighted language models, %over infinite alphabets,
 and applied them to the problem of parsing
@@
 by computing a derivation tree modulo
-a distance between words
-defined by a SW transducer given in input.
+a distance between words defined by a SW transducer given in input.
 This allows to consider finer word relationships than strict equality.
-%
-
-This work can be extended in several directions. First, 
-the best search algorithm could be generalized from $1$-best to $n$-best~\cite{Huang05kbest},
+
+This work can be extended in several directions. 
+First,  the best search algorithm could be generalized from $1$-best to $n$-best~\cite{Huang05kbest},
 and to $k$-\emph{closed} semirings~\cite{Mohri02semiring}

### 2021-10-14 algo-best-lemma.tex

% !TEX root = main.tex
%

%Intuitively, 
\begin{lemma}[Correctness]\label{algo-correct}
For all safe path~$\pi$ of~$\G_A$,  %with $\src(\pi) = \<q_0, q_0>$,
%with last element $v$, \\
($i$) if $\last(\pi) = \< q, q'>$, %with $q, q' \in Q$, 
then there exists $u \in \Delta^*$ such that 
$\weight_A(q[\bot], u, q'[\bot]) = \weight(\pi)$,\\
%
($ii$) if $\last(\pi) = \< q, p, q'>$, %with $q, q' \in Q$, $p\in P$, 
then there exists $v \in \Delta^*$ such that \\
${\displaystyle\bigoplus_{c \in \Deltac}}
 \weight_A\bigl(\configup{q}{\< c, p> \stackup \top }, v, \configup{q'}{\<c, p> \stackup \top}\bigr)
 = \weight(\pi)$.
\end{lemma}
%
\begin{proof}
We prove ($i$) and ($ii$) simultaneously by induction on the length $|\pi|$ of the path $\pi$.

\noindent
The base case $|\pi| = 1$ is a direct consequence of the definition of weight of paths of length 1
and the first line of \eqref{eq:SWVPA-weight}.

\noindent
If $\pi = v_0,\ldots, v_n$ with $n \geq 1$, 
let us do a case analysis on the edge $v_{n-1} \to v_n$.
%
We suppose Lemma ~\ref{algo-correct} is true for a safe path $\pi' = v_0,\ldots, v_{n-1}$ and a word $u' \in \Delta^*$. Firstly, let's assume $v_{n-1} = \< q, q''>$ and $last(\pi) = \< q, q'>$. By definition, the edge $\< q, q''> \to \< q, q'>$ is possible only if an internal symbol or a return symbol is read, in order to keep the stack empty:

\begin{itemize}
\item if we read $a \in \Delta_i$ and have an empty stack, then the new $\weight_A$ is computed with the third case in equation~\eqref{eq:SWVPA-weight}.
We have
\begin{align*}
 \weight_A\bigl(\config{q}{\gamma}, u\, a,
               \config{q'}{\bot}\bigr) &=
   \weight_A\bigl(\config{q}{\gamma}, u, \config{q''}{\bot}\bigr)
   \otimes\weiei(q'', a, q')  \\
 &= \weight(\pi')\otimes\weiei(q'', a, q')\\
 &= \weight(\pi)\nonumber 
\end{align*}
with $\pi = (\pi', v_n)$, since we know that $\weight(v_0,\ldots, v_{n-1}) \otimes \eta_A(v_{n-1} \to v_{n}) = \weight(v_0,\ldots, v_n)$ and $\weiei(q'', a, q')$ is indeed a possible edge label $\eta_A$ for the edge $\< q, q''> \to \< q, q'>$ (first line of label definition). Hence in this case the Lemma is true.

\item Following the same reasoning as before, we find that the Lemma is also true when $a \in \Delta_r$ by using the last equation in \eqref{eq:SWVPA-weight}.
%
\end{itemize}
\noindent
Let's now assume that we have a safe path $\pi' = (v_0,\ldots, v_{n-1})$ such that $v_{n-1} = \< q_1, p, q_2>$ and $last(\pi) = \< q_0, q_3>$. The edge $\< q_1, p, q_2> \to \< q_0, q_3>$ exists if we read a return then try to match with the call on top of the stack which then leaves the stack empty. With the 6th equation in~\eqref{eq:SWVPA-weight}, we can compute the $\weight_A$ for $\<q_0, q_3>$ and then replace the first term with the 5th equation in~\eqref{eq:SWVPA-weight}, such as:
\begin{align*}
\weight_A\bigl(\config{q_0}{\bot}, u\,{\return{r}},
               \config{q_3}{\bot}\bigr) &=
  \weight_A\bigl(\config{q_0}{\bot}, u,
                 \configup{q_2}{\<{\call{c}}, p>\stackup\bot}\bigr)
  \otimes \weir\bigl(q_2, {\call{c}}, p, {\return{r}}, q_3\bigr)\\
  &=
   \weight_A(\config{q_1}{\bot}, u, \config{q_2}{\bot}\bigr)
 \otimes \weiec(q_0, {\call{c}}, p, q_1)
  \otimes \weir\bigl(q_2, {\call{c}}, p, {\return{r}}, q_3\bigr)\\
  &= \weight_A(\pi')\otimes \weiec(q_0, {\call{c}}, p, q_1)
  \otimes \weir\bigl(q_2, {\call{c}}, p, {\return{r}}, q_3\bigr)\\
  &= \weight_A(\pi)
\end{align*}
for $\pi = (\pi', v_n)$. Here the Lemma holds since $\weir\bigl(q_2, {\call{c}}, p, {\return{r}}, q_3\bigr)$ is also part of the edge labels (second line).

Finally, let's assume that $v_{n-1} = \<q_1, p, q_2>$ and $last(\pi) = \<q_0, p, q_3>$ for $p, p' \in Q$. Following the same reasoning as the two previous cases, we can similarly prove that the Lemma is also true.

\end{proof}

\begin{lemma}[Completeness] \label{algo-complete}
For all $q, q' \in Q$, $p\in P$, 
($i$) there exists a safe path~$\pi$ of~$\G_A$ such that
%such that $\src(\pi) = \<q_0, q_0>$ for some $q_0 \in Q$, 
$\last(\pi) = \< q, q'>$,
and $\weight(\pi) = b_\bot(q, q')$, 
($ii$) there exists a safe path $\pi'$ of $\G_A$ 
such that $\last(\pi') = \< q, p, q'>$,
and $\weight(\pi') = \bigoplus_\Deltac b_\top(q, p, q')$.
\end{lemma}
%
\begin{proof}
By associativity, commutativity and distributivity for $\Semiring$,
\eqref{eq:bbot} can be rewritten into the form, unfolding~\eqref{eq:SWVPA-weight}:
%
\begin{equation}\label{eq:bbot-unfold}
b_\bot(q, q') = 
\bigoplus_{t\in \Delta^*} 
\bigoplus_{\begin{array}{c}
           \scriptstyle q_0,\ldots, q_n \in Q\\
           \scriptstyle p_{i},\ldots, p_{k} \in P
           \end{array}}
\bigotimes_{i=1}^n w_i(\tau_i)           
\end{equation}
%
where $n$ is the length of $t$, $k \leq n$,
$q_0 = q$, $q_n = q'$, 
for all $1 \leq i < n$, 
$w_i$ is one of the functions of $\bar\wei$,
$\tau_i$ is a transition of $A$ and
$\src(\tau_i) = q_{i-1}$, $\snd(\tau_i) = q_i$.
%
Since $\Semiring$ is total, there exists finite sequences as above such that
\( b_\bot(q, q') = \bigotimes_{i=1}^n w_i(\tau_i)\).
%
There might exists, for $q$ and $q'$,
several finite sequences~$\bar\tau$ and~$\bar{w}$ as above, 
let us choose arbitrarily one of minimal length~$n$. 
This integer $n$ will be denoted $n_{q, q'}$ in the following.
%We represent a computation of $A$ on $t$ by a sequence $\rho$ of its transitions,

Similarly, following~\eqref{eq:btop} and~\eqref{eq:SWVPA-weight}, 
$\bigoplus_\Deltac b_\top(q, p, q')$ can be put in the form:
%
\begin{equation}\label{eq:btop-unfold}
{\textstyle\bigoplus_\Deltac} b_\top(q, p, q') = 
{\textstyle\bigoplus_\Deltac} \bigoplus_{t\in \Delta^*} 
\bigoplus_{\begin{array}{c}
           \scriptstyle q_0,\ldots, q_n \in Q\\
           \scriptstyle p_{i},\ldots, p_{k} \in P
           \end{array}}
\bigotimes_{i=1}^n w_i(\tau_i)           
\end{equation}
%
and hence 
\( {\textstyle\bigoplus_\Deltac} b_\top(q, p, q') = \bigotimes_{i=1}^n w_i(\tau_i) \),
using $c \in \Deltac$ that minimize the function in rhs of~\eqref{eq:btop-unfold}.
We denote the smallest $n$ as above by $n_{q, p, q'}$.

\noindent
We show now the existence of a path $\pi$ as expected, %in Lemma~\ref{lem:algo-complete}, 
by simultaneous induction on $n_{q, q'}$ and $n_{q, p, q'}$.

\noindent
The base case, $n = 0$ corresponds to $t = \varepsilon$. 
In this case, by~\eqref{eq:SWVPA-weight}, $b_\top(q, p, q') = \zero$,
$b_\bot(q, q) = \one$ and $b_\bot(q, q') = \zero$ if $q \neq q'$.

\noindent
For $n = n_{q, q'} > 0$, let $\bar\tau$ and~$\bar{w}$
be the sequences associated to $q$ and $q'$ as above.
Let us do a case analysis of $w_n$.
... \qed
\end{proof}

### 2021-10-14 algo-best-lemma.tex
proof completeness

@@
 
-\item Following the same reasoning as before, we find that the Lemma is also true when $a \in \Delta_r$ by using the last equation in \eqref{eq:SWVPA-weight}.
+\item Following the same reasoning as before, we find that the Lemma is also true when $a \in \Delta_r$ by using the last equation in \eqref{eq:SWVPA-weight} which illustrates the case when a return symbol is unmatched.
 %
@@
 \noindent
-Let's now assume that we have a safe path $\pi' = (v_0,\ldots, v_{n-1})$ such that $v_{n-1} = \< q_1, p, q_2>$ and $last(\pi) = \< q_0, q_3>$.
+Let's now assume that $v_{n-1} = \< q_1, p, q_2>$ and $last(\pi) = \< q_0, q_3>$.
 The edge $\< q_1, p, q_2> \to \< q_0, q_3>$ exists if we read a return then try to match with the call on top of the stack which then leaves the stack empty.
@@
 \begin{align*}
-\weight_A\bigl(\config{q_0}{\bot}, u\,{\return{r}},
-               \config{q_3}{\bot}\bigr) &=
+&\weight_A\bigl(\config{q_0}{\bot}, u\,{\return{r}},
+               \config{q_3}{\bot}\bigr) =
   \weight_A\bigl(\config{q_0}{\bot}, u,
@@
   &=
-   \weight_A(\config{q_1}{\bot}, u, \config{q_2}{\bot}\bigr)
+   \weight_A(\configup{q_1}{\<{\call{c}}, p>\stackup\bot}, u, \configup{q_2}{\<{\call{c}}, p>\stackup\bot}\bigr)
  \otimes \weiec(q_0, {\call{c}}, p, q_1)
@@
 
-Finally, let's assume that $v_{n-1} = \<q_1, p, q_2>$ and $last(\pi) = \<q_0, p, q_3>$ for $p, p' \in Q$.
+Finally, let's assume that $v_{n-1} = \<q_1, p, q_2>$ and $last(\pi) = \<q_0, p', q_3>$ for $p, p' \in Q$.
 Following the same reasoning as the two previous cases, we can similarly prove that the Lemma is also true.
-
+\qed
 \end{proof}
@@
 Let us do a case analysis of $w_n$.
-... \qed
+Computing $b_\bot(q, q')$ comes down to running $A$ on a well-matched word $u \in \Delta^*$ for which $|u| = n$.
+Let's take the case where $w_n  = \weiei$.
+Then using Equation~\ref{eq:SWVPA-weight}, we decompose $b_\bot(q, q')$ as follows:
+%\config{q}{\gamma}, u\,{\call{c}'},
+%\configup{q'}{\<{\call{c}'}, p'>\stackup \<{\call{c}}, p>\stackup \gamma'}\bigr)
+\begin{align*}
+b_\bot(q, q') &= \weight_A(\config{q}{\bot}, u', \config{q''}{\bot}) \otimes\weiei(q'', a, q') \\
+&= b_\bot(q, q'')\otimes\weiei(q'', a, q')
+\end{align*}
+Since $u=u'a$ is minimal, then $u'$ is minimal, and $\weight_A(\config{q}{\bot}, u', \config{q''}{\bot}) =b_\bot(q, q'')$.
+Consequently, for this case the Lemma is true.
+The other cases can be proven similarly,  but only for the $w_n$ that can form a well-matched word.
+For example, $w_n$ cannot be $\weic$ since this call will never have a matching return.
+\qed
 \end{proof}

### 2021-10-14 abstract.tex
intro & abstract

@@
 %
-We present some properties of \SWA, \SWT, and \SWVPA models
-that we use to define and solve a variant of parsing
-over infinite alphabets.
+We present a Bar-Hillel Perles Shamir construction of a \SWA
+computing a \SWT-defined distance between a \SWA input language and a word, 
+some closure results 
+and a polynomial best-search algorithm for \SWVPA. 
+These results are applied to solve a variant of parsing over infinite alphabets.
 %
-We illustrate the model with a motivating application to
-automated music transcription.
+%We illustrate the model with a motivating application to
+%automated music transcription.

### 2021-10-14 introduction.tex
intro & abstract

@@
 %
-The automata at the core of model checkers
+Moreover, automata at the core of model checkers
 compute on input symbols represented by large bitvectors~\cite{Vardi07ciaa} %\cite{BaierKatoen08MC}
-(sets of assignments of Boolean variables) %propositional variables)
+%(sets of assignments of Boolean variables) %propositional variables)
 and, in practice,  %implementation
@@
 %
-Following a similar idea, % and generalizing,
+Following this idea, % and generalizing,
 in symbolic automata (\SA)~\cite{Veanes12symbolic,dAntoniVeanes17CAV,dAntoni21CACM},
-transitions are guarded by predicates over infinite  domains.
+transitions are guarded by predicates over large or infinite domains.
 With appropriate closure conditions on the sets of such predicates, % (alphabet theories),
@@
 Other extensions of language models  %(automata and grammars)
-are able to assign a weight value to their input.
+assign one weight value to every input~\cite{Droste09handbook}.
 %help in dealing with non-determinism
-This is useful in order for quantitative modelling 
-\eg of probabilistic or stochastic programs (recursive for pushdown models),
-quantitative database queries, or semi-structured data (in infinite domains), 
+This is useful for the quantitative modelling of
+\eg probabilistic or stochastic recursive programs, %( for pushdown models),
+quantitative database queries, or semi-structured data, % (in infinite domains), 
 as well as for the verification of quantitative properties about 
@@
 %
-In the context of parsing, for grammars computing a weight value, 
+In the context of parsing, when grammars return weight values, 
 it is possible to rank derivations (and \emph{abstract syntax trees})
-in order to select a best one 
-(or $n$ bests)~\cite{Goodman99SemiringParsing,Nederhof03weightedParsing,MorbitzVogler19weighted-parsing-parsing}. 
+in order to select a best one (or $n$ bests),
 %\florent{= "weighted parsing"}
-This is useful in order to deal with ambiguity of grammars.
+\eg in case of ambiguity~\cite{Goodman99SemiringParsing,Nederhof03weightedParsing,MorbitzVogler19weighted-parsing}.
 %This is roughly the principle of \emph{weighted parsing} approaches
@@
 a weight is associated to each transition rule, % production rule
-and the rule's weights can be combined with an
-associative product operator~$\otimes$ to yield the weight of a whole computation.
+and the weights of the rules involved in a computation are combined with an
+associative product operator~$\otimes$. % to yield the weight of a whole computation.
 %\reviews{1) AST for grammar, run for WA}
@@
 In this paper, we present some results for 
-\emph{symbolic weighted} finite states language models %(\swM) 
+\emph{Symbolic Weighted} finite states language models %(\swM) 
 generalizing the Boolean guards of~\SA %in transitions
 to functions into an arbitrary semiring, 
-and extending also~\WA, by handling infinite alphabets (Figure~\ref{fig:hierarchy}).
+and also the~\WA, by handling infinite alphabets (Figure~\ref{fig:hierarchy}).
 %
-The models considered in this paper are also 
+The models considered are also 
 %\SWA and \SWVPA are both 
@@
 %\reviews{cite \cite{Herrmann16dlt,Herrmann20phd}}
-
-In short, a transition rule $q \xrightarrow{\phi} q'$ of a SW automaton, 
-from state $q$ to $q'$,
-%\florent{c'est un résumé de la Fig. \ref{fig:hierarchy}}
-%input symbols are variables, %(parameters)
-is labeled by a function~$\phi$ associating to every input symbol $a$ a weight value $\phi(a)$
-in a semiring~$\Semiring$.
 
@@
 
-%The framework relies on several language models:
+\noindent
+In short, a transition rule $q \xrightarrow{\phi} q'$ of a SW automaton, 
+from state $q$ to $q'$,
+%\florent{c'est un résumé de la Fig. \ref{fig:hierarchy}}
+%input symbols are variables, %(parameters)
+is labeled by a function~$\phi$ associating to every input symbol~$a$, a weight value~$\phi(a)$
+in a semiring~$\Semiring$.
+%
 The models studied are: 
 symbolic-weighted automata (\SWA),
-transducers (\SWT), and pushdown automata
-with a visibility restriction~\cite{AlurMadhusudan09nested} (\SWVPA).
-A \SWT defines a distance $T(s, t)$ between finite words $s$ and $t$
-over infinite alphabets. % following~\cite{Mohri03ijfcs}.
-%\wrt an edit distance
-A  \SWVPA operates sequentially on \emph{nested words}~\cite{AlurMadhusudan09nested},
-structured with markup symbols (parentheses) and describing linearizations of trees.
-A \SWVPA $A$ associates a weight value $A(t)$ % \in \Semiring$
-to a given nested word $t$, which is itsef the linearization of a weighted AST. %representing a parse tree.
-
-THE RESULTS/CONTRIBUTIONS of this paper are:
-\florent{revise: main results = 3 theorems}
+defining series over infinite alphabets, 
+transducers (\SWT), 
+defining distances between finite words over infinite alphabets, % following~\cite{Mohri03ijfcs}.
+and pushdown automata with a visibility restriction~\cite{AlurMadhusudan09nested} (\SWVPA),
+operating sequentially on %\emph{nested words}~\cite{AlurMadhusudan09nested},
+words structured with markup symbols (parentheses) and describing linearizations of trees.
+%A \SWVPA $A$ associates a weight value $A(t)$ % \in \Semiring$
+%to a given nested word $t$, which is itsef the linearization of a weighted AST. %representing a parse tree.
+%
+The main contributions of this paper are:
 %\reviews{1) not $i$! see response.txt}
@@
 %for the weighted computation on (nested) words over infinite alphabets;
-($i$)~a construction %Bar-Hillel Perles Shamir 
-for computing a \SWT-defined distance between an \SWA language and a word (Section~\ref{section:SWA}), 
-($ii$)~closure results for \SWVPA (Section~\ref{sec:SWVPA}), 
-($iii$)~a polynomial best-search algorithm for \SWVPA, %(Section~\ref{sec:best})
+($i$)~the construction à la Bar-Hillel Perles Shamir of a \SWA
+     computing a \SWT-defined distance between a \SWA input language and a word (Section~\ref{section:SWA}), 
+($ii$)~closure results for \SWVPA, 
+($iii$)~a polynomial best-search algorithm for \SWVPA (Section~\ref{sec:SWVPA}). %(Section~\ref{sec:best})
 % a framework for parsing over infinite alphabets,
+Moreover, we present in Section~\ref{sec:parsing} an 
+($iv$) application to the problem of weighted parsing over infinite input alphabets, 
+called \emph{SW-parsing}. 
 %
-Moreover, we detail 
-($iv$)~an application to the problem of weighted parsing over infinite input alphabets 
-(\emph{SW-parsing}, Section~\ref{sec:parsing}), 
-based on the above models integrated in a uniform framework, with in particular:
-($iv.a$)~the \SWT-based definition of generic edit distances between input and output (yield) words,
-and ($iv.b$)~the use, for technical convenience, of nested words, and \SWVPA,
-instead of syntax trees and grammars. %(\S~\ref{sec:trees}).
+The goal of this problem is, given an input word~$s$, 
+to find~$t$ minimizing the distance, in the sense of~\cite{Mohri03ijfcs}, 
+$T(s, t) \otimes A(t)$, where $T$ is a \SWT and $A$ a \SWVPA.
+The notion of transducer-based distances allows to consider 
+different infinite alphabets for the input $s$ and output $t$.
+Moreover, the use of \SWVPA permits to consider output $t$ in the form of a (nested) word, instead of a tree.
+%technical convenience
+%the SW models are integrated in a uniform framework.
+\emph{SW-parsing} is solved with the Bar-Hillel construction~\cite{NederhofSatta03ParsingIntersection} 
+%\cite{GruneJacobs08parsing}
+of a \SWVPA $B$ such that, for all~$t$, $B(t) = T(s, t) \otimes A(t)$
+and the application of a best-search procedure to this automaton $B$.
+% reduction to the shortest distance in graphs~\cite{Mohri02semiring,Huang05kbest}.
 
-... These results should also apply in the context of the quantitive verification of programs...
+%Like weighted-parsing methods~\cite{Goodman99SemiringParsing,Nederhof03weightedParsing,MorbitzVogler19weighted-parsing},
+%\reviews{1) WARNING: \cite{MorbitzVogler19weighted-parsing} incomparable to our def. weighted parsing}
 
-***
-Parsing is the problem of structuring a linear representation
-(a finite word) according to a language model. % (a formal grammar). % natural language, programming language,
-%
-Most of the context-free parsing approaches~\cite{GruneJacobs08parsing}
-assume a finite and reasonably small input alphabet. %models and algorithms
-Such a restriction makes perfect sense in the context of
-NLP tasks such as constituency parsing, 
-compilers or interpreters for programming languages.
+%Parsing is the problem of structuring a linear representation
+%(a finite word) according to a language model. % (a formal grammar). % natural language, programming language,
+
+Context-free parsing approaches~\cite{GruneJacobs08parsing}
+generally assume a finite and reasonably small input alphabet. %models and algorithms
+%Such a restriction makes perfect sense in the context of
+%NLP tasks such as constituency parsing, 
+%compilers or interpreters for programming languages.
 Considering large or infinite alphabets can however be of
-practical interest when dealing with large characters encodings such as UTF-16~\cite{dAntoni21CACM},
+practical interest when dealing with large characters encodings such as UTF-16~\cite{dAntoni21CACM}.
 \florent{applis à la Fossacs chez Veanes et al?}
 % processing strings in \eg for vulnerability detection in Web-applications~\cite{dAntoni21CACM},
-%
-analysis of data streams, serialization of structured documents~\cite{Segoufin06csl,NevenSchwentickVianu04FSMinfinite},
-or processing timed execution traces~\cite{Bouyer03algebraic}.
+%We believe that these results could also be useful 
+It is also true in the context of automata-based quantitative verification,  %of systems processing %analysis of 
+\eg when dealing with data streams, 
+serialization of structured documents~\cite{Segoufin06csl,NevenSchwentickVianu04FSMinfinite},
+or timed execution traces~\cite{Bouyer03algebraic}.
+
 % algebraic definition of a class of data languages
@@
 The latter case is related to a motivation  of  the present work: \florent{idem}
-automated music transcription. Representations that capture  music performances
-are essentially linear:
-audio files or the widely used
-MIDI format~\cite{Selfridge-Field97beyondMIDI}. %\cite{MIDIfile}.
-Such representations ignore the hierarchical structures that frame the
-conception of music, at least in the Western area. These structures, on the other hand,
-are present, either explicitly  or implicitly,
+automated music transcription. Representations capturing music performances
+are essentially linear~\cite{Selfridge-Field97beyondMIDI};
+%audio files or the widely used MIDI format %\cite{MIDIfile}.
+they ignore the hierarchical structures that frame the
+conception of music, at least in the Western area. 
+These structures, on the other hand, are present, either explicitly  or implicitly,
 in Common Western Music Notation~\cite{Gould11Notation}:
@@
 measures in beats, and beats can be further recursively divided.
-It follows that music events do not occur at arbitrary timestamps,
+It follows that written music events do not occur at arbitrary timestamps,
 but respect a discrete partitioning of the timeline incurred by
@@
 The \emph{transcription problem} takes
-as input a linear representation (audio or MIDI) and aims at re-constructing
-these structures
+as input a linear performance (audio or MIDI) and aims at re-constructing
+structured notation, 
 by mapping input events to this hierarchical rhythmic space.
-It can therefore be stated as a parsing problem~\cite{foscarin:hal-01988990}
-over an infinite alphabet of timed events.
+It can therefore be stated as a parsing problem
+over an infinite alphabet of timed events~\cite{foscarin:hal-01988990}.
 
-Given an input word $s$, the \emph{SW-parsing} problem aims at
-finding $t$ minimizing
-$T(s, t) \otimes A(t)$, called the distance between $s$ and $A$ in~\cite{Mohri03ijfcs},
-\wrt the ranking defined by $\oplus$. Like weighted-parsing
-methods~\cite{Goodman99SemiringParsing,Nederhof03weightedParsing,MorbitzVogler19weighted-parsing},
-\reviews{1) WARNING: \cite{MorbitzVogler19weighted-parsing} incomparable to our def. weighted parsing}
-our approach proceeds in two steps.
-The first step is an intersection
-(Bar-Hillel construction~\cite{NederhofSatta03ParsingIntersection,GruneJacobs08parsing})
-\reviews{1) also \cite{NederhofSatta03ParsingIntersection} pasing as intersection}
-\florent{chap. intersection in~\cite{GruneJacobs08parsing}}
-where, given a \SWT $T$, a \SWVPA $A$, and an input word $s$,
-a \SWVPA $B$ %$A_{T, s}$
-is built, such that for all~$t$,
-$B(t) = T(s, t) \otimes A(t)$. % $A_{T, s}(t) = T(s, t) \otimes A(t)$.
-In the second step, a best AST $t$ is found by applying to $B$ %$A_{T, s}$
-a best search algorithm similar to the shortest distance
-in graphs~\cite{Mohri02semiring,Huang05kbest}.
 %In expressiveness, they are equivalent to weighted CFG.
@@
 \label{ex:running}
-We illustrate our framework with a very simplified %toy
+We illustrate our approach with a very simplified %toy
 running example of \emph{music transcription}:

### 2021-10-15 main.tex
intro & abstract
conclusion

@@
 Weighted Symbolic Automata with Data Storage~\cite{Herrmann16dlt}. 
+\florent{review 3: = classical transducers}
 
@@
 \noindent
-Let us construct the transition function $\wei'_1$ iteratively.
+Let us construct the transition function~$\wei'_1$ iteratively.
 %
@@
 
-We studied 3 properties of Symbolic Weighted language models, %over infinite alphabets,
+We studied properties of three Symbolic Weighted language models, %over infinite alphabets,
 and applied them to the problem of parsing
 with infinitely many possible input symbols (typically timed events).
-Our approach extends conventional parsing and weighted parsing
-by computing a derivation tree modulo
-a distance between words defined by a SW transducer given in input.
+One originality of this approach 
+%extends conventional parsing and weighted parsing
+is to compute derivation trees modulo
+a distance between words defined by a given SW transducer.
 This allows to consider finer word relationships than strict equality.
@@
 This work can be extended in several directions. 
-First,  the best search algorithm could be generalized from $1$-best to $n$-best~\cite{Huang05kbest},
+The best search algorithm could be generalized from $1$-best to $n$-best~\cite{Huang05kbest},
 and to $k$-\emph{closed} semirings~\cite{Mohri02semiring}
-(instead of \emph{bounded}, which corresponds to $0$-\emph{closed}). Second,
-%we believe that
-the complexity bounds of the algorithms could be more precisely characterized,
-as well as expressiveness of \swM compared to the automata of 
-\eg~\cite{Segoufin06csl,KaminskiFrancez94,NevenSchwentickVianu04FSMinfinite,Bojanczyk11FO2}.
-Finally, the  best search algorithm presented here works offline, whereas
-an on-the-fly automata construction would allow for online parsing,
-a suitable feature in the context of applications such as 
-\eg automatic music transcription.
-
-\florent{comparison of expressiveness with the automata 
-of\cite{Segoufin06csl,KaminskiFrancez94,NevenSchwentickVianu04FSMinfinite,Bojanczyk11FO2}.}
+(instead of \emph{bounded}, which corresponds to $0$-\emph{closed}). 
+The complexity bounds of the best-search algorithms 
+could also certainly be improved.
+%as well as expressiveness of \swM compared to the automata of 
+%\eg~\cite{Segoufin06csl,KaminskiFrancez94,NevenSchwentickVianu04FSMinfinite,Bojanczyk11FO2}.
+Another possible study could be the existence of best-search algorithms
+for general models like~\cite{Herrmann16dlt}.
+Finally, the  best search algorithm presented here works offline, 
+whereas an on-the-fly approach coupling automata construction and best-search 
+would be interesting \eg for online XML validation or filtering.
+%a suitable feature in the context of applications such as 
+%\eg automatic music transcription.
+
+%\florent{comparison of expressiveness with the automata 
+%of\cite{Segoufin06csl,KaminskiFrancez94,NevenSchwentickVianu04FSMinfinite,Bojanczyk11FO2}.}
 
@@
 
+\section{Proof of Proposition~\ref{algo-shortest}}\label{app:shortpath} % Shortest path
+\input{algo-best-lemma.tex}
+
 \section{Nested Words and Parse Trees} \label{sec:trees}\label{app:trees}

### 2021-10-15 algo-best-alt.tex
intro & abstract
preliminary submission

@@
 
-\florent{New Algorithm for best-search}
+%\florent{New Algorithm for best-search}
 We shall compute the values of the functions $b_\bot$ and $b_\top$
@@
 \end{proposition}
-This proposition is a consequence of Lemmata~\ref{algo-correct} and~\ref{algo-complete} in~Appendix \ref{app:shortpath}.
+This proposition is a consequence of Lemmata~\ref{algo-correct} and~\ref{algo-complete} 
+found in~Appendix \ref{app:shortpath}.
 
@@
 in order to compute the minimum of $t \mapsto A(t)$ \wrt $\leq_\oplus$, 
-it is sufficient to compute one shortest pathes in $\G_{A}$ of source
+it is sufficient to compute one shortest paths in $\G_{A}$ of source
 $\<q_0, q_0>$ and target $\<q, q'>$ for all $q_0, q, q' \in Q$.
-This can be done with~~\cite{Mohri02semiring}~Theorem~?? \florent{completer}
-with a complexity ***. Altogether, the complexity for computing the minimum of $A$
-is in $O(...)$.
+This can be done with~\cite{Mohri02semiring} \florent{completer, Theorem~??}
+with a worst case time complexity in $O\bigl(|Q|^2.|P|)^3\bigr)$. 
+%Altogether, the complexity for computing the minimum of $A$ is in 
+%worst case $O(...)$.
 Moreover, a witness $t \in \Delta^*$ for this minimum can be associated to the appropriate

### 2021-10-15 introduction.tex
conclusion
preliminary submission

@@
 particular cases of the very general class of 
-Weighted Symbolic Automata with Data Storage defined in~\cite{Herrmann16dlt,Herrmann20phd}. 
+Weighted Symbolic Automata with Data Storage~\cite{Herrmann16dlt,Herrmann20phd}, 
+using appropriate storage types.
 %with special data storage types...
@@
 \begin{tikzpicture}
-\node (SWADS) at (3.0,5.0) {%
+\node (SWADS) at (2.8,5.0) {%
   \(
@@
 %(a finite word) according to a language model. % (a formal grammar). % natural language, programming language,
-
 Context-free parsing approaches~\cite{GruneJacobs08parsing}
@@
 
-The latter case is related to a motivation  of  the present work: \florent{idem}
-automated music transcription. Representations capturing music performances
+The latter case is related to a motivation study for the present work:
+automated music transcription, 
+that we shall use as a running example
+thorough the paper, in order to illustrate our results.
+%
+Representations capturing music performances
 are essentially linear~\cite{Selfridge-Field97beyondMIDI};
@@
 The \emph{transcription problem} takes
-as input a linear performance (audio or MIDI) and aims at re-constructing
-structured notation, 
+as input a linear performance (in audio or MIDI format) 
+and aims at re-constructing structured notation, 
 by mapping input events to this hierarchical rhythmic space.
@@
 \label{ex:running}
-We illustrate our approach with a very simplified %toy
-running example of \emph{music transcription}:
+We consider a very simplified %toy
+example of \emph{music transcription}:
 a given input \emph{timeline} of musical events
@@
 Such inputs typically correspond
-to the recording of a live performance, 
-\eg~$I = 69\tstamp{0.07},
+to the recording of a live performance, \eg 
+\[I = 69\tstamp{0.07},
       71\tstamp{0.72},
@@
 	  76\tstamp{1.36},
-	     77\tstamp{1.71}$. % over interval $[0,2[$,
+	  77\tstamp{1.71}.\] % over interval $[0,2[$,
 

### 2021-10-15 main.tex
preliminary submission

@@
 %\title{Symbolic Weighted Language Models and Automated Music Transcription}
-\title{Symbolic Weighted Language Models and\\ Quantitative Parsing over Infinite Alphabets}
 %\title{Weighted Visibly Pushdown Automata and Automated Music Transcription}
+\title{Symbolic Weighted Language Models,
+       Quantitative Parsing and Verification\\ over Infinite Alphabets}
 %\titlerunning{WVPA \& AMT}
@@
 Another possible study could be the existence of best-search algorithms
-for general models like~\cite{Herrmann16dlt}.
+for the more general models of~\cite{Herrmann16dlt}.
 Finally, the  best search algorithm presented here works offline, 

### 2021-10-15 algo-best-lemma.tex
intro & abstract
submission FoSSaCS

@@
 % !TEX root = main.tex
-%
+Proposition~\ref{algo-shortest} states the correctness of the construction of the graph~$G_A$ for the computation of $b_\bot$
+and, by extension, the minimum of $A$.
+Its proof is a consequence of the two following Lemmata~\ref{algo-correct} and~\ref{algo-complete}.
 
@@
 If $\pi = v_0,\ldots, v_n$ with $n \geq 1$, 
-let us do a case analysis on the edge $v_{n-1} \to v_n$.
-%
-We suppose Lemma ~\ref{algo-correct} is true for a safe path $\pi' = v_0,\ldots, v_{n-1}$ and a word $u' \in \Delta^*$.
-Firstly, let's assume $v_{n-1} = \< q, q''>$ and $last(\pi) = \< q, q'>$.
-By definition, the edge $\< q, q''> \to \< q, q'>$ is possible only if an internal symbol or a return symbol is read, in order to keep the stack empty:
+we assume that Lemma~\ref{algo-correct} holds for a safe path $\pi' = v_0,\ldots, v_{n-1}$ and a word~$u' \in \Delta^*$, 
+and do a case analysis on the edge~$v_{n-1} \to v_n$.
+
+\smallskip\noindent
+Firstly, let us consider the case where $v_{n-1} = \< q, q''>$ and $v_{n} = \< q, q'>$. 
+By definition, the edge $\< q, q''> \to \< q, q'>$ is possible only if an internal symbol or a return symbol is read, 
+keeping the stack empty:
 
 \begin{itemize}
-\item if we read $a \in \Delta_i$ and have an empty stack, then the new $\weight_A$ is computed with the third case in equation~\eqref{eq:SWVPA-weight}.
+\item if we read $a \in \Delta_i$ and have an empty stack, 
+then the new $\weight_A$ is computed 
+with the third case in equation~\eqref{eq:SWVPA-weight}.
 We have
 \begin{align*}
- \weight_A\bigl(\config{q}{\gamma}, u\, a,
+ \weight_A\bigl(\config{q}{\gamma}, u'\, a,
                \config{q'}{\bot}\bigr) &=
-   \weight_A\bigl(\config{q}{\gamma}, u, \config{q''}{\bot}\bigr)
+   \weight_A\bigl(\config{q}{\gamma}, u', \config{q''}{\bot}\bigr)
    \otimes\weiei(q'', a, q')  \\
@@
 \end{align*}
-with $\pi = (\pi', v_n)$, since we know that $\weight(v_0,\ldots, v_{n-1}) \otimes \eta_A(v_{n-1} \to v_{n}) = \weight(v_0,\ldots, v_n)$ and $\weiei(q'', a, q')$ is indeed a possible edge label $\eta_A$ for the edge $\< q, q''> \to \< q, q'>$ (first line of label definition).
-Hence in this case the Lemma is true.
+since we know that $\weight(v_0,\ldots, v_{n-1}) \otimes \eta_A(v_{n-1} \to v_{n}) = \weight(v_0,\ldots, v_n)$ and $\weiei(q'', a, q')$ is indeed a possible label $\eta_A$ for the edge $\< q, q''> \to \< q, q'>$ (first line of label definition).
+%Hence in this case the Lemma is true.
 
-\item Following the same reasoning as before, we find that the Lemma is also true when $a \in \Delta_r$ by using the last equation in \eqref{eq:SWVPA-weight} which illustrates the case when a return symbol is unmatched.
+\item Following the same reasoning as above, we find that the Lemma is also true when $a \in \Delta_r$ by using the last equation in \eqref{eq:SWVPA-weight} which corresponds to the case of an unmatched return symbol.
 %
@@
 \noindent
-Let's now assume that $v_{n-1} = \< q_1, p, q_2>$ and $last(\pi) = \< q_0, q_3>$.
-The edge $\< q_1, p, q_2> \to \< q_0, q_3>$ exists if we read a return then try to match with the call on top of the stack which then leaves the stack empty.
-With the 6th equation in~\eqref{eq:SWVPA-weight}, we can compute the $\weight_A$ for $\<q_0, q_3>$ and then replace the first term with the 5th equation in~\eqref{eq:SWVPA-weight}, such as:
+let us now consider the case where $v_{n-1} = \< q_1, p, q_2>$ and $v_{n} = \< q_0, q_3>$. 
+The edge $\< q_1, p, q_2> \to \< q_0, q_3>$ exists if we have a return symbol 
+matching a call symbol on the top of the stack, 
+which is pop, leaving the stack empty. 
+With the 6th equation in~\eqref{eq:SWVPA-weight}, 
+we can compute the $\weight_A$ for $\<q_0, q_3>$ 
+and then replace the first term with the 5th equation in~\eqref{eq:SWVPA-weight}, such as:
 \begin{align*}
@@
 \end{align*}
-for $\pi = (\pi', v_n)$. Here the Lemma holds since $\weir\bigl(q_2, {\call{c}}, p, {\return{r}}, q_3\bigr)$ is also part of the edge labels (second line).
+%
+Here the Lemma holds since $\weir\bigl(q_2, {\call{c}}, p, {\return{r}}, q_3\bigr)$ 
+is also part of the edge labels (second line).
 
-Finally, let's assume that $v_{n-1} = \<q_1, p, q_2>$ and $last(\pi) = \<q_0, p', q_3>$ for $p, p' \in Q$.
-Following the same reasoning as the two previous cases, we can similarly prove that the Lemma is also true.
+Finally, for the case where $v_{n-1} = \<q_1, p, q_2>$ and $v_n = \<q_0, p', q_3>$ for $p, p' \in Q$, 
+we can following the same reasoning as the two previous cases, and complete the proof of Lemma~{algo-correct}.
 \qed
@@
 be the sequences associated to $q$ and $q'$ as above.
-Let us do a case analysis of $w_n$.
-Computing $b_\bot(q, q')$ comes down to running $A$ on a well-matched word $u \in \Delta^*$ for which $|u| = n$.
-Let's take the case where $w_n  = \weiei$.
-Then using Equation~\ref{eq:SWVPA-weight}, we decompose $b_\bot(q, q')$ as follows:
+%
+We can perform a case analysis of $w_n$.
+Computing $b_\bot(q, q')$ comes down to running $A$ on a word $u \in \Delta^*$, 
+such that $|u| = n$, and $u$ is well-matched except for unmatched return symbols
+read on empty stack.
+%
+Let us consider the case where $w_n  = \weiei$. 
+Then using Equation~\ref{eq:SWVPA-weight}, we can decompose $b_\bot(q, q')$ as follows:
 %\config{q}{\gamma}, u\,{\call{c}'},
@@
 \end{align*}
-Since $u=u'a$ is minimal, then $u'$ is minimal, and $\weight_A(\config{q}{\bot}, u', \config{q''}{\bot}) =b_\bot(q, q'')$.
-Consequently, for this case the Lemma is true.
+Since $u = u'a$ is minimal in length, then so is $u'$, 
+and $\weight_A(\config{q}{\bot}, u', \config{q''}{\bot}) = b_\bot(q, q'')$. 
+Consequently, for this case the Lemma holds. 
 The other cases can be proven similarly,  but only for the $w_n$ that can form a well-matched word. 
-For example, $w_n$ cannot be $\weic$ since this call will never have a matching return.
+For example, $w_n$ cannot be $\weic$ since this call would not have a matching return.
 \qed

### 2021-10-15 introduction.tex
submission FoSSaCS

@@
 all the good properties enjoyed by automata over finite alphabets are preserved.
-The ability to compare input symbols of \SA is quite restricted compared to the 
+The ability of \SA to compare input symbols is restricted compared to the 
 former automata with memory. 
@@
 quantitative database queries, or semi-structured data, % (in infinite domains), 
-as well as for the verification of quantitative properties about 
-\florent{quantitative reasoning/verification on programs}
+as well as for the verification of quantitative properties of systems
+related to %\florent{quantitative reasoning/verification on programs}
 quality measures, resource-consumption, distance metrics, probabilistic guarantees, \etc
@@
 %
-The models considered are also 
+In short, a transition rule $q \xrightarrow{\phi} q'$ of a Symbolic Weighted Automaton, 
+from state $q$ to $q'$,
+%\florent{c'est un résumé de la Fig. \ref{fig:hierarchy}}
+%input symbols are variables, %(parameters)
+is labeled by a function~$\phi$ associating to every input symbol~$a$, a weight value~$\phi(a)$
+in a semiring~$\Semiring$.
+%
+These models are also 
 %\SWA and \SWVPA are both 
@@
 Weighted Symbolic Automata with Data Storage~\cite{Herrmann16dlt,Herrmann20phd}, 
-using appropriate storage types.
+for appropriate storage types.
 %with special data storage types...
@@
 \noindent
-In short, a transition rule $q \xrightarrow{\phi} q'$ of a SW automaton, 
-from state $q$ to $q'$,
-%\florent{c'est un résumé de la Fig. \ref{fig:hierarchy}}
-%input symbols are variables, %(parameters)
-is labeled by a function~$\phi$ associating to every input symbol~$a$, a weight value~$\phi(a)$
-in a semiring~$\Semiring$.
-%
 The models studied are: 
@@
 %for the weighted computation on (nested) words over infinite alphabets;
-($i$)~the construction à la Bar-Hillel Perles Shamir of a \SWA
-     computing a \SWT-defined distance between a \SWA input language and a word (Section~\ref{section:SWA}), 
-($ii$)~closure results for \SWVPA, 
-($iii$)~a polynomial best-search algorithm for \SWVPA (Section~\ref{sec:SWVPA}). %(Section~\ref{sec:best})
+\begin{description}
+\item[($i$)] a construction à la Bar-Hillel Perles Shamir of a \SWA
+     computing a \SWT-defined distance between a \SWA input language and a word (Proposition~\ref{prop:epsilon}), 
+\item[($ii$)] closure results for \SWVPA (Proposition~\ref{prop:SWVPA-product}), 
+\item[($iii$)] a polynomial best-search algorithm for \SWVPA (Proposition~{th:best-search}. %(Section~\ref{sec:best})
+\end{description}
+%
 % a framework for parsing over infinite alphabets,
 Moreover, we present in Section~\ref{sec:parsing} an 
-($iv$) application to the problem of weighted parsing over infinite input alphabets, 
+\begin{description}
+\item[($iv$)] application to the problem of weighted parsing over infinite input alphabets, 
 called \emph{SW-parsing}. 
+\end{description}
 %
-The goal of this problem is, given an input word~$s$, 
+The goal of the latter problem is, given an input word~$s$, 
 to find~$t$ minimizing the distance, in the sense of~\cite{Mohri03ijfcs}, 
@@
 different infinite alphabets for the input $s$ and output $t$.
-Moreover, the use of \SWVPA permits to consider output $t$ in the form of a (nested) word, instead of a tree.
+Moreover, the use of \SWVPA permits to search for an output $t$ in the form of a (nested) word, 
+instead of a tree.
 %technical convenience
 %the SW models are integrated in a uniform framework.
-\emph{SW-parsing} is solved with the Bar-Hillel construction~\cite{NederhofSatta03ParsingIntersection} 
+\emph{SW-parsing} is solved with a Bar-Hillel construction~\cite{NederhofSatta03ParsingIntersection} 
 %\cite{GruneJacobs08parsing}
@@
 practical interest when dealing with large characters encodings such as UTF-16~\cite{dAntoni21CACM}.
-\florent{applis à la Fossacs chez Veanes et al?}
 % processing strings in \eg for vulnerability detection in Web-applications~\cite{dAntoni21CACM},
 %We believe that these results could also be useful 
-It is also true in the context of automata-based quantitative verification,  %of systems processing %analysis of 
-\eg when dealing with data streams, 
+It is also true in the context of automata-based quantitative verification techniques~\cite{Kwiatkowska07quantitativeVerif},  %of systems processing %analysis of 
+in particular when dealing with data streams, 
 serialization of structured documents~\cite{Segoufin06csl,NevenSchwentickVianu04FSMinfinite},
@@
 
-The latter case is related to a motivation study for the present work:
+The latter case of timed traces 
+is related to a problem that motivated the present work:
 automated music transcription, 
@@
 conception of music, at least in the Western area. 
-These structures, on the other hand, are present, either explicitly  or implicitly,
+Hierarchical structures, on the other hand, are present, either explicitly  or implicitly,
 in Common Western Music Notation~\cite{Gould11Notation}:

### 2021-10-15 main.tex
submission FoSSaCS

@@
 \newcommand{\florent}[1]{\todo[noline,size=\tiny,color=yellow!40]{#1}}
-%\renewcommand{\florent}[1]{} % uncomment for final
+\renewcommand{\florent}[1]{} % uncomment for final
 \newcommand{\lydia}[1]{\todo[noline,size=\tiny,color=red!30]{#1}}
-%\renewcommand{\lydia}[1]{} % uncomment for final
+\renewcommand{\lydia}[1]{} % uncomment for final
 \newcommand{\philippe}[1]{\todo[noline,size=\tiny,color=green!30]{#1}}
-%\renewcommand{\philippe}[1]{} % uncomment for final
+\renewcommand{\philippe}[1]{} % uncomment for final
 \newcommand{\reviews}[1]{\todo[noline,size=\tiny,color=blue!30]{#1}}
-%\renewcommand{\reviews}[1]{} % uncomment for final
+\renewcommand{\reviews}[1]{} % uncomment for final
 
@@
 %
-The models presented in this section generalize weighted automata and transducers~\cite{Droste09handbook}
+The models presented in this section are weighted automata 
+and transducers~\cite{Droste09handbook}
 %over finite alphabets, see \eg~\cite{Mohri03ijfcs},
-by labeling each transition with a weight function
-that takes the input and output symbols as parameters.
+with transitions labeled by weight functions
+that take the input and output symbols as parameters.
 These functions generalize the guards 
@@
 %
-They are particular cases of the very general model of 
+These models are also particular cases of the very general model of 
 Weighted Symbolic Automata with Data Storage~\cite{Herrmann16dlt}. 
-\florent{review 3: = classical transducers}
+%\florent{review 3: = classical transducers}
 
@@
 
-\begin{definition}
-\label{def:transducer} \label{def:SWT}
+\begin{definition}\label{def:transducer} \label{def:SWT}
 A \emph{symbolic-weighted transducer} (\SWT)
@@
 %is not the constant $\zero$.
+%
+Note that the SW-transducers of Definition~\ref{def:transducer} %do not form a new model, 
+are actually classical weighted transducers where the alphabets are permitted to be infinite.
 %
@@
 %
-Since~$\zero$ is absorbing for~$\otimes$ in~$\Semiring$,
-one term $\wei_{ij}(q, a, b, q'')$ equal to $\zero$ in the above expression
-will be ignored in the sum, meaning that there is no possible transition
-from state $q$ into state $q'$ while reading $a$ and $b$.
+Since~$\zero$ is absorbing for~$\otimes$, 
+and neutral for $\oplus$ in~$\Semiring$,
+if $\wei_{ij}(q, a, b, q'')$ is equal to $\zero$
+(meaning that there is no possible transition
+from state $q$ into state $q'$ while reading $a$ and $b$), 
+then the entire term containing this expression can be ignored in the sum.
 This is analogous to the case of a transition's guard not satisfied by $\<a, b>$ for
@@
      \delta(\mu\tstamp{\tau}, \nu\tstamp{\tau'}) & \quad & %|\tau' - \tau| & \quad &
+     \mbox{if~} \nu \neq - \\ %|\tau' - \tau| & \quad &
 \wei_{11}(q_1, \mu\tstamp{\tau}, \nu\tstamp{\tau'}, q_0) & = &
-     \delta(\mu\tstamp{\tau}, \nu\tstamp{\tau'}) &
+     \delta(\mu\tstamp{\tau}, \nu\tstamp{\tau'}) & &
      \mbox{if~} \nu \neq - \\ %|\tau' - \tau| & \quad &
@@
 %\subsection{Properties}
-\noindent
-The following property will be useful for
-symbolic weighted parsing (Section~\ref{sec:parsing}).
+%\noindent
+%The following property will be useful for
+%symbolic weighted parsing (Section~\ref{sec:parsing}).
 
@@
 Weighted Symbolic Automata with Data Storage~\cite{Herrmann16dlt} for a special 
-type of data storage.
-It associates to every nested word over an infinite alphabet a weight value
+type of data storage, and 
+associates to every nested word over an infinite alphabet a weight value
 in a semiring.
@@
 $\creturn{3}\tstamp{2}$,
-$\creturn{\mathsf{m}}\tstamp{2}$,
 $\creturn{\mathsf{m}}\tstamp{2}$
@@
 %
-As in Section~\ref{section:transducer},
-we extend the above transition functions as follows
+%As in Section~\ref{section:transducer},
+We extend the above transition functions as follows
 for every $q, q' \in Q$, $p \in P$,
@@
 \[
-\begin{array}{lll}
-\weii: Q \times [\Deltac \times P] \times \Deltai \times Q \to \Semiring &
-\weii(q, c, p, a, q') = \eta_\mathsf{ci}(c, a) &
-\mathrm{where~} \eta_\mathsf{ci} = \weii(q, p, q'),\\
-%
-\weiei: Q \times \Deltai \times Q \to \Semiring &
-\weiei(q, a, q') = \phi_\mathsf{i}(a) &
-\mathrm{where~} \phi_\mathsf{i} = \weiei(q, q').\\
-%
-\weic: Q \times [\Deltac \times P] \times  [\Deltac \times P] \times Q \to \Semiring &
-\weic(q, c, p, \call{c'}, p', q') = \eta_\mathsf{cc}(c, \call{c'}) &
-\mathrm{where~} \eta_\mathsf{cc} = \weic(q, p, p', q'),\\
-%
-\weiec: Q \times [\Deltac \times P] \times Q \to \Semiring &
-\weiec(q, {\call{c}}, p', q') = \phi_\mathsf{c}({\call{c}}) &
-\mathrm{where~} \phi_\mathsf{c} = \weiec(q, p, q').\\
-%
-\weir: Q \times [\Deltac \times P] \times \Deltar \times Q \to \Semiring &
-\weir(q, {\call{c}},  p, {\return{r}}, q') = \eta_\mathsf{cr}({\call{c}},  {\return{r}}) &
-\mathrm{where~} \eta_\mathsf{cr} = \weir(q, p, q'),\\
-%
-\weier: Q \times \Deltar \times Q \to \Semiring &
-\weier(q, {\return{r}}, q') = \phi_\mathsf{r}({\return{r}}) &
-\mathrm{where~} \phi_\mathsf{r} = \weier(q, q').\\
+\begin{array}{lcl}
+\weii: Q \times [\Deltac \times P] \times \Deltai \times Q \to \Semiring & &
+\weii(q, c, p, a, q') = \eta_\mathsf{ci}(c, a)\\ 
+\multicolumn{3}{r}{\mathrm{where~} \eta_\mathsf{ci} = \weii(q, p, q'),}\\
+%
+\weiei: Q \times \Deltai \times Q \to \Semiring & &
+\weiei(q, a, q') = \phi_\mathsf{i}(a)\\
+\multicolumn{3}{r}{\mathrm{where~} \phi_\mathsf{i} = \weiei(q, q'),}\\
+%
+\weic: Q \times [\Deltac \times P] \times  [\Deltac \times P] \times Q \to \Semiring \  & &
+\weic(q, c, p, \call{c'}, p', q') = \eta_\mathsf{cc}(c, \call{c'})\qquad\\
+\multicolumn{3}{r}{\mathrm{where~} \eta_\mathsf{cc} = \weic(q, p, p', q'),}\\
+%
+\weiec: Q \times [\Deltac \times P] \times Q \to \Semiring & &
+\weiec(q, {\call{c}}, p', q') = \phi_\mathsf{c}({\call{c}})\\
+\multicolumn{3}{r}{\mathrm{where~} \phi_\mathsf{c} = \weiec(q, p, q'),}\\
+%
+\weir: Q \times [\Deltac \times P] \times \Deltar \times Q \to \Semiring & &
+\weir(q, {\call{c}},  p, {\return{r}}, q') = \eta_\mathsf{cr}({\call{c}},  {\return{r}})\\
+\multicolumn{3}{r}{\mathrm{where~} \eta_\mathsf{cr} = \weir(q, p, q'),}\\
+%
+\weier: Q \times \Deltar \times Q \to \Semiring & &
+\weier(q, {\return{r}}, q') = \phi_\mathsf{r}({\return{r}})\\
+\multicolumn{3}{r}{\mathrm{where~} \phi_\mathsf{r} = \weier(q, q')}.\\
 \end{array}
@@
 or more precisely to its structural complexity.
-Given a set of equivalent representations, we aim at choosing the simpler one, \ie the one with the smallest weight.
+Given a set of equivalent representations, 
+we aim at choosing the simplest one, \ie the one with the smallest weight.
 
 \lydia{We should find a better variable name than j... it does not render well in fractions}
-Let's say the top of the stack is a tuple of the call~$\ccall{n}\tstamp{[\tau, \tau+\ell]}$ and a state $q$.
-Let's now consider a new call transition starting in state $q_{\frac{j}{n}}$, that is the $j$-th state of the $n$ sub-intervals of duration~$\frac{\ell}{n}$.
+Let us assume that the top of the stack is a tuple 
+made of the call symbol~$\ccall{n}\tstamp{[\tau, \tau+\ell]}$ 
+and the state~$q$. 
+Let us now consider a new call transition starting in state $q_{\frac{j}{n}}$, that is the $j$-th state of the $n$ sub-intervals of duration~$\frac{\ell}{n}$.
 We thus read a new time-division symbol ~$\ccall{d}$ 
@@
 \(
-\weic\bigl(q_{\frac{j}{n}},
+{\weic\bigl(q_{\frac{j}{n}},
            \ccall{n}\tstamp{[\tau, \tau+\ell]}, q,
            \ccall{d}\tstamp{\iota}, q_{\frac{j+1}{c}},
-           q_{\frac{1}{d}} \bigr) = \alpha_d
+           q_{\frac{1}{d}} \bigr) = \alpha_d}
 \).
@@
 \(
-\weir\bigl(q_{\frac{d}{d}},
+{\weir\bigl(q_{\frac{d}{d}},
            \ccall{d}\tstamp{[\tau, \tau+\ell]}, q_{\frac{i+1}{c}},
-           \,\creturn{d}\tstamp{\tau + \ell}, q_{\frac{i+1}{c}} \bigr) = \one
+           \,\creturn{d}\tstamp{\tau + \ell}, q_{\frac{i+1}{c}} \bigr) = \one}
 \).
@@
 (\ie~$\gamma'$ is left untouched during the computation).
-However, the computation can read $\<c, p>$ at the top of $\gamma'$,
+However, the computation can read~$\<c, p>$ at the top of~$\gamma'$,
 and can also push another pair $\< c', p'> \in \Gamma$ on top of $\gamma'$,
@@
 in the definition \eqref{eq:SWVPA-weight} of $\weight_A$ (call symbol).
-The pair $\< c', p'>$ can be popped later, during the computation from $q$ to $q'$,
+The pair $\< c', p'>$ must be popped later, during the computation from $q$ to $q'$,
 following the fifth case of \eqref{eq:SWVPA-weight} (return symbol).
@@
 %
-\begin{proof} (sketch)
+\begin{proof} %(sketch)
 We follow a \emph{Bar-Hillel} construction for parsing by intersection.
@@
 the problem of searching for the best derivation (AST) of a weighted CF-grammar $G$
-that yields a given input word $w$, called~\emph{weighted parsing},
-(see~\cite{Goodman99SemiringParsing},
- also the more general framework~\cite{MorbitzVogler19weighted-parsing}),
+that yields a given input word $w$, 
 with an infinite input alphabet instead of a finite one and
-transducer-defined distances instead of equality. 
-%
-See Appendix~\ref{sec:trees} for more details on the correspondence between 
+transducer-defined distances instead of equality
+(it is however uncomparable to 
+ related problem called~\emph{semiring parsing},
+ \cite{Goodman99SemiringParsing},
+  called~\emph{weighted parsing}~\cite{MorbitzVogler19weighted-parsing}).
+%
+The interested reader might find in Appendix~\ref{sec:trees} 
+more details on the correspondence between 
 nested words $t \in \Delta^*$, AST, CF grammars and \SWVPA.
-
 %
@@
 
-We studied properties of three Symbolic Weighted language models, %over infinite alphabets,
+We presented three properties of three Symbolic Weighted language models
+(automata, transducers and visibly pushdown automata), %over infinite alphabets,
 and applied them to the problem of parsing
@@
 %extends conventional parsing and weighted parsing
-is to compute derivation trees modulo
+is the comparison of words using 
 a distance between words defined by a given SW transducer.
@@
 This work can be extended in several directions. 
-The best search algorithm could be generalized from $1$-best to $n$-best~\cite{Huang05kbest},
+The best-search algorithm for \SWVPA
+could be generalized from $1$-best to $n$-best~\cite{Huang05kbest},
 and to $k$-\emph{closed} semirings~\cite{Mohri02semiring}
 (instead of \emph{bounded}, which corresponds to $0$-\emph{closed}). 
-The complexity bounds of the best-search algorithms 
-could also certainly be improved.
+It complexity upper bound could also certainly be improved.
 %as well as expressiveness of \swM compared to the automata of 
@@
 for the more general models of~\cite{Herrmann16dlt}.
-Finally, the  best search algorithm presented here works offline, 
+
+\noindent
+Finally, the  best-search algorithm presented here works offline, 
 whereas an on-the-fly approach coupling automata construction and best-search 
-would be interesting \eg for online XML validation or filtering.
+would be interesting \eg for online XML validation or filtering, 
+or program monitoring~\cite{dAntonyAlur14SVPDA}.
 %a suitable feature in the context of applications such as 
@@
 
-\newpage
+%\newpage
 

### 2021-12-07 FoSSaCS/rebuttal.md
rebuttal fossacs

Thank you for your submission to FoSSaCS 2022. The rebuttal period will be between now and December 9, 2021, 23:59 AoE.

During this time, you will have access to the current state of your reviews and have the opportunity to submit a response. Please try to limit your response to 600 words or less and keep in mind the following during this process:

* The response must focus on any factual errors in the reviews and any questions posed by the reviewers. It must not provide new research results or reformulate the presentation. Try to be as concise and to the point as possible.

* The rebuttal period is an opportunity to react to the reviews, but not a requirement to do so. Thus, if you feel the reviews are accurate and the reviewers have not  asked any questions, then you do not have to respond.

* The reviews are as submitted by the PC members, without any coordination between them. Thus, there may be inconsistencies. Furthermore, these are not the final versions of the reviews. The reviews can later be updated to take into account the discussions at the program committee meeting, and we may find it necessary to solicit other outside reviews after the rebuttal period.

* The program committee will read your responses carefully and take this information into account during the discussions. On the other hand, the program committee will not directly respond to your responses, either before the program committee discussion or in the final versions of the reviews.

* Your response will be seen by all PC members who have access to the discussion of your paper, so please be polite and constructive.

The reviews on your paper are attached to this letter. To submit your response you should log on the EasyChair Web page for FoSSaCS 2022 and select your submission on the menu.

----------------------- REVIEW 1 ---------------------
SUBMISSION: 55
TITLE: Symbolic Weighted Language Models, Quantitative Parsing and Verification over Infinite Alphabets
AUTHORS: Florent Jacquemard, Lydia Rodrigez de la Nava and Philippe Rigaux

----------- Overall evaluation -----------
SCORE: -2 (reject)
----- TEXT:
The goal of the paper is to set up relevant models and constructions to solve efficiently 
the weighted parsing problem over infinite alphabets.

I especially liked the motivation example used throughout the paper: the music transcription problem.  The input is a sequence of timed events (MIDI format). Timestamps being rational numbers explains the need to consider infinite alphabets.  The goal is to produce a score which best corresponds to the input sequence.  The score has a hierarchical nested representation defining time points (grid) where notes of the score could occur.  So the timestamps of the input sequence have to be aligned with the above grid.  Several alignments are possible and a weighted transducer (wT) is used to measure the distance between the input sequence and the alignment.  The hierarchical division of the timeline defining the grid used for alignments is naturally represented as a nested word.  Several such hierarchical divisions are possible and a weighted visibly pushdown automaton (wVPA) is used to measure the complexity of the resulting score.  Finally, the best score is obtained by finding a witness of the minimization problem for a synchronized product of the wT and the wVPA. Since the input alphabet is infinite, one needs symbolic representations for wT and wVPA.

The paper introduces and study symbolic weighted automata (swA), transducers (swT) and  visibly pushdonw automata (swVPA). Some basic properties of these models are established  (Propositions 1 and 2). The notations are quite heavy and technical, but the intuition is  clear and the techniques are classic. The most interesting problem is the minimisation of  a swVPA (Proposition 3) in order to find a best parsing solution. Unfortunately, as  explained below in the detailed comments, I think the proof is wrong/incomplete and the  problems are not simple typos. Due to that I cannot recommend to accept the paper.

Major problem in the proof of Prop. 3:

On page 15, your definition of the weight function $\eta_A$ has two main problems.

1. It is not defined for edges from $Q\times Q$ to $Q\times P\times Q$.
2. On edges from $Q\times P\times Q$ to $Q\times P\times Q$, the weight you define is not 
   an element of the semiring $S$ as it is supposed to be. You are summing $w_i$, $w_r$ and 
   $w_c$ over the second argument $\bigoplus^{2}$ and the first argument remains free. So 
   you are defining a function from $\Delta_c$ to the semiring and not an element of $S$.

Other detailed comments:

Fig.2: the counting semiring is not idempotent, bounded, complete.

Lemma 4 (ii) 'and similarly for $\oplus$': This is not clear.
It is probably wrong if the semiring is not idempotent (complete does not imply 
idempotent) and even when idempotent this is not clear. You need that $\alpha=\bigoplus_i 
\alpha$ and without some continuity asumption this may not hold.
Same remark for (v) and (vi).

Lemma 4 (iii) second equality: I don't see why this should hold. This would imply 
$(a+b)\times(a'+b')=aa'+bb'$. Why should this hold?
Same remark for (iv)

Lemma 4 (v) and (vi): exponents ^{2} or ^{1} are missing on some \bigoplus

Def 1: How do you represent $\phi$ or $\eta$? What do you mean by 'effectively computed'? 
Turing machine?

Proposition 1: the PTIME claim is strange. This seems to assume that all operations on 
functions from the label theory are done in constant time, but def 1 only says they are 
effectively computable, without complexity bounds.
Same remark for Prop. 3

Proof of Prop 1:

- I think it would be better (easier to understand) to construct a product with
  $\epsilon$-transitions, and to separately state and prove that $\epsilon$-transitions can
  be removed from swA.
- in input -> an input (many occurrences)

page 12:

- p -> p' in w_c^{e}(q,p,q')
- what is the meaning/usefulness of $w_r^{e}$? Why should you (do you need to) 
  allow a return with an empty stack when dealing with VPAs?

page 13:

- Due to the base case, all $weight_A(q[\gamma],...)$ equal 0 when $\gamma\neq\bot$.
  So why do you add $\gamma$ to the definitions?
- Remove $p'\in P$ and $p\in P$ form the \bigoplus of the w_c cases
- Add c,p to the \bigoplus of the w_r case
- Remove the w_r^{e} case.
  If you choose to keep the w_r^{e} case then your definition is not coherent as you may have $\gamma'=\bot$ in the w_r case.

Example 5:

- I don't understand starting from the last 4 lines of page 13.
  $q_{\frac{j+1}{c}}$? $\lceil_d[\tau,\tau+\ell]$?
- line 3, page 14: measeures are not nested?

Proof of Prop 2:

- the case $\oplus$ is similar: No, this is simply a disjoint union of the  automata
- line 4 of the proof: $in_1 ∈\to in$

page 15, equation (14): the value is 0 according to your def.

Section 5, first para: $\phi_{cc}$ and $\phi_{ci}$ are missing
----------- Reviewer's confidence -----------
SCORE: 5 ((expert))

----------------------- REVIEW 2 ---------------------
SUBMISSION: 55
TITLE: Symbolic Weighted Language Models, Quantitative Parsing and Verification over Infinite Alphabets
AUTHORS: Florent Jacquemard, Lydia Rodrigez de la Nava and Philippe Rigaux

----------- Overall evaluation -----------
SCORE: 1 (weak accept)
----- TEXT:
The authors study symbolic variants of weighted automata (sWA), weighted transducers (sWT), and weighted visibly pushdown automata (sWVPA). The term "symbolic" here signifies that the input alphabet can be large or even infinite. For this reason, the automata transitions cannot enumerate the alphabet symbols. Instead, each transition carries a function from the alphabet to the weight semiring.

The idea of extending automata from small finite alphabets to large or infinite alphabets using predicates is rather straightforward and does not give rise to significant complications for many (but not all) problems about them. In fact, this idea can be found in papers much earlier than the cited works [9, 10, 36], for example, in (Noord & Gerdemann, 2001) and guarded automata for Kleene Algebra with Tests (Kozen, 2001).

>  Finite State Transducers with Predicates and Identities, by G. van Noord & D. Gerdemann (2001)

> Automata on Guarded Strings and Applications, by D. Kozen (2001)

The authors claim the following contributions for the symbolic weighted machines that they investigate:
(1) **Proposition 1**: A construction for an sWA that computes distance of a string, where the notion of distance is given by a sWT, from a weighted language given by a sWA.
(2) **Proposition 2**: The transductions that are computed by sWVPA are closed under semiring addition and semiring multiplication.
(3) **Proposition 3**: An algorithm for finding an input nested word with minimal weight, where the word/weight association is given by a sWVPA.
(4) **Proposition 5**: An application of the previous results to a variant of weighted parsing.

The paper is well-written and the technical ideas are communicated clearly. Several related works are discussed in the introduction. In particular, the authors mention some works on register automata and other quantitative automata models. There is also a relevant and large line of work on **Cost Register Automata** and related notions of regular expressions and domain-specific languages for quantitative properties. Some representative references:

>  Streamable Regular Transductions, by R. Alur et al. (2020) [TCS]

> StreamQRE: Modular Specification and Efficient Evaluation of Quantitative Queries over Streaming Data, by K. Mamouras et al. (2017) [PLDI]

> Regular Functions and Cost Register Automata, by R. Alur et al. (2013) [LICS]

While I find the content of the paper interesting and relevant, I have some reservations regarding how novel the main contributions are. For example, I am wondering whether Proposition 1 is essentially the composition of two well-known constructions. First, the product automaton (A x T) can be constructed using essentially the same idea as the composition of weighted transducers (see Mohri's papers). One would need to allow epsilon-transitions, but these can be eliminated as a second step, given that the authors consider a class of semirings that allows epsilon-elimination. Similarly, Proposition 2 seems to rely on a straightforward adaptation of product constructions.

It is nice that the authors consider a motivating parsing application, namely music transcription that introduces hierchical structure into the musical notation. This application seems a bit niche, so it would be interesting to have a discussion of other more impactful applications as well.

To summarize, this is overall an interesting paper, but I am only marginally leaning towards acceptance because of questions regarding the novelty of the technical results.

More remarks:

The authors seem to be referencing a result by Bar-Hillel, Perles and Shamir, but no citation is given. Please provide a citation and clarify the relationship with the reference.

It would help the reader if you restructure the end of Section 4 so that you can give an explicit proof for Proposition 3, instead of integrating the proof in the prose of the section.

----------- Reviewer's confidence -----------
SCORE: 3 ((medium))

----------------------- REVIEW 3 ---------------------
SUBMISSION: 55
TITLE: Symbolic Weighted Language Models, Quantitative Parsing and Verification over Infinite Alphabets
AUTHORS: Florent Jacquemard, Lydia Rodrigez de la Nava and Philippe Rigaux

----------- Overall evaluation -----------
SCORE: 2 (accept)
----- TEXT:
The paper "Symbolic Weighted Language Models, Quantitative Parsing and Verification over Infinite Alphabets" by Jacquemard et al studies quantitative language models over infinite input alphabets:

- Symbolic Weighted Automata (swA)
- Transducers (swT)
- Visibly Pushdown (sw-VPA)

This is achieved by generalizing the Boolean guards of symbolic automata to functions into a commutative semiring.

The main contribution is to show how these automata can be used to parse strings with infinitely many possible input symbols (such as timed events). The running example is that of transcribing music as a parsing problem. This example is developed to illustrate all aspects of the work, including the symbolic weighted parsing algorithm in Section 5.

The authors advertise their parsing algorithm saying that "The SW-parsing framework, applied to the transcription problem, allows to find an optimal solution that considers both the fitness of the result, and its structural complexity." There is an implementation at https://qparse.gitlabpages.inria.fr/ but I didn't have time to test it.

I very much like this work. It brings together the classical topic of parsing with cutting edge quantitative methods and the practical application of transcribing music from midi to scores.

Typos:

"(Proposition th:best-search."

"languages models"

"constructions are performed on the on-the-fly"

----------- Reviewer's confidence -----------
SCORE: 3 ((medium))

------------------------------------------------------

Best wishes,

Patricia and Lutz

### 2021-12-09 FoSSaCS/response.md
response (draft)

about the two first remarks in review 1, presented as "major problems in the proof of Prop.3":

1. indeed, there are no edges from $Q\times Q$ to $Q\times P\times Q$.
   It means that the weight, by $\eta_A$, of every such edge is 0 (of the semiring).
   It is generally the convention for missing edges in weighted graphs or missing transition of weighted automata. However, we could have staten it explicitely.

2. the weight the edges from $Q\times P\times Q$ to $Q\times P\times Q$ is actually an element of the semiring $S$.

The typing is as follows:   
on the one hand,

- $w_i(...) \in \Phi_{ci}$
- $\oplus_{\Delta_i} w_i \in \Phi_{c}$

on the other hand,

- $w_c(...) \in \Phi_{cc}$   
- $w_r(...) \in \Phi_{cr}$   
- $\oplus_{\Delta_r} w_r \in \Phi_{c}$
- $w_c(...) \otimes_2 \oplus_{\Delta_r} w_r \in \Phi_{c}$

Hence $\oplus_{\Delta_r}$ of the sum of the two above terms (i.e. $\eta_A$ of the edge) is in $S$.

### 2021-12-09 FoSSaCS/rebuttal.md
rev. parenthese Prop.3

@@
 1. It is not defined for edges from $Q\times Q$ to $Q\times P\times Q$.
-2. On edges from $Q\times P\times Q$ to $Q\times P\times Q$, the weight you define is not 
-   an element of the semiring $S$ as it is supposed to be. You are summing $w_i$, $w_r$ and 
-   $w_c$ over the second argument $\bigoplus^{2}$ and the first argument remains free. So 
-   you are defining a function from $\Delta_c$ to the semiring and not an element of $S$.
+2.
+On edges from $Q\times P\times Q$ to $Q\times P\times Q$, the weight you define is not an element of the semiring $S$ as it is supposed to be.
+You are summing $w_i$, $w_r$ and $w_c$ over the second argument $\bigoplus^{2}$ and the first argument remains free.
+So you are defining a function from $\Delta_c$ to the semiring and not an element of $S$.
 
@@
 
-Proposition 1: the PTIME claim is strange. This seems to assume that all operations on 
-functions from the label theory are done in constant time, but def 1 only says they are 
-effectively computable, without complexity bounds.
+Proposition 1: the PTIME claim is strange.
+This seems to assume that all operations on functions from the label theory are done in constant time, but def 1 only says they are effectively computable, without complexity bounds.
 Same remark for Prop. 3
@@
 Proof of Prop 1: 
-
 - I think it would be better (easier to understand) to construct a product with
-  $\epsilon$-transitions, and to separately state and prove that $\epsilon$-transitions can
-  be removed from swA.
+  $\epsilon$-transitions, and to separately state and prove that $\epsilon$-transitions can be removed from swA.
 - in input -> an input (many occurrences)

### 2021-12-09 FoSSaCS/response.md
reponse

@@
-about the two first remarks in review 1, presented as "major problems in the proof of Prop.3":
+We would like to thank the reviewers for their careful reading of the paper and their remarks, 
+and in particular the suggestion of reviewer 2 to consider the related model of Cost Register Automata.
+
+Regarding the two first comments in review 1, presented as "major problems in the proof of Prop.3", they are actually simple typos:
 
@@
    It means that the weight, by $\eta_A$, of every such edge is 0 (of the semiring).
-   It is generally the convention for missing edges in weighted graphs or missing transition of weighted automata. However, we could have staten it explicitely.
+   It is generally the convention for missing edges in weighted graphs or missing transitions in weighted automata. However, we could have staten it explicitly...
 
-2. the weight the edges from $Q\times P\times Q$ to $Q\times P\times Q$ is actually an element of the semiring $S$.
+2. there is a misplaced parenthesis in the expression for the weight of the edges from $Q\times P\times Q$ to $Q\times P\times Q$:
+   It should be: $\bigoplus_{\Delta_c} ( \bigoplus_{\Delta_i} w_i(...) \oplus w_c(...) \otimes_2 \bigoplus_{\Delta_r} w_r(...))$
+   (which is an element of the semiring), 
+   and not $\bigoplus_{\Delta_i} w_i(...) \oplus  \bigoplus_{\Delta_c} ( w_c(...) \otimes_2 \bigoplus_{\Delta_r} w_r(...))$.
+   The proofs of Prop.3 (in appendix) actually use the first (correct) formulation.
+
+We agree with reviewer 1 that the statements iii and iv of Lemma 4 are incorrect.
+This lemma is only used in an informal remark following Def. 1, and is not used in the proofs of the main results of the paper. Hence it could be removed without harming these results.
+
+Regarding other comments :
+
+- w^e_r (page 12) allows to deal with unmatched return symbols, following the original definitions of visibly PDA and symbolic VPDA [8], where the input words are not assumed to be well parenthesised.
+- also, following the original definitions of symbolic automata and VPDA, the functions of the label theory are supposed to be oracles, and the PTIME complexity assumes a constant time computation for these oracles.
+
+Considering the above problems identified by reviewers, we prefer to withdraw our submission.
+However, we would be grateful if you could transmit this answer to the reviewers.
 

### 2021-12-09 FoSSaCS/response.md
reponse

@@
    It means that the weight, by $\eta_A$, of every such edge is 0 (of the semiring).
-   It is generally the convention for missing edges in weighted graphs or missing transitions in weighted automata. However, we could have staten it explicitly...
+   It is generally the convention for missing edges in weighted graphs or missing transitions in weighted automata. However, we could have stated it explicitly...
 
@@
 
-We agree with reviewer 1 that the statements iii and iv of Lemma 4 are incorrect.
+We agree with reviewer 1 that statements iii and iv of Lemma 4 are incorrect.
 This lemma is only used in an informal remark following Def. 1, and is not used in the proofs of the main results of the paper. Hence it could be removed without harming these results.
@@
 
-- w^e_r (page 12) allows to deal with unmatched return symbols, following the original definitions of visibly PDA and symbolic VPDA [8], where the input words are not assumed to be well parenthesised.
+- w^e_r (page 12) allows dealing with unmatched return symbols, following the original definitions of visibly PDA and symbolic VPDA [8], where the input words are not assumed to be well parenthesized.
 - also, following the original definitions of symbolic automata and VPDA, the functions of the label theory are supposed to be oracles, and the PTIME complexity assumes a constant time computation for these oracles.
@@
 However, we would be grateful if you could transmit this answer to the reviewers.
+
+Thank you
+Best Regards
 

### 2021-12-09 FoSSaCS/response.md
reponse

@@
-We would like to thank the reviewers for their careful reading of the paper and their remarks, 
-and in particular the suggestion of reviewer 2 to consider the related model of Cost Register Automata.
+We would like to thank the reviewers for their careful reading of the paper and for their remarks, 
+in particular the suggestion of reviewer 2 to consider the related model of Cost Register Automata.
 
@@
 
-Considering the above problems identified by reviewers, we prefer to withdraw our submission.
-However, we would be grateful if you could transmit this answer to the reviewers.
+Although the the above problems identified by reviewers are small mistakes, they are several (hence too many!) mistakes, and we prefer to withdraw our submission.
+Nevertheless, we would be grateful if you could transmit this answer to the reviewers.
 
